### 2016-05-20 thesis/methods/methods.tex
added glossary file

\chapter{Methods (draft)}
\label{ch:methods}

\section{\gls{bmi}}
\label{sec:bmi}

\subsection{\gls{bmi} Calculation}
\label{subsec:bmicalc}

\gls{bmi} was calculated from the height and weight data of the samples using the following equation:

\begin{equation}
	\label{eq:bmicalc}
	\gls{bmi} = \frac{Weight (kg)}{Height^2(m^2)}\\
\end{equation}

\subsection{\gls{bmi} Classification}
\label{subsec:bmiclassification}

Samples were classified based on the \gls{who} definition, as shown in Table~\ref{tab:whobmiclass}.
\begin{table}[hb]
	\caption{\gls{who} defined \gls{bmi} classification}
	\label{tab:whobmiclass}
	\begin{center}
		\begin{tabular}{lc}
			\bfseries {Classification} & \bfseries {\gls{bmi} Value}\\
			\hline
			\rule{0pt}{2.25ex}Underweight & \textless{} 20.0\\
			Normal weight/lean & 20.0$\sim$24.9\\
			Overweight & 25.0$\sim$29.9\\
			Obese & \textgreater{} 30.0\\
		\end{tabular}
	\end{center}
\end{table}

\section{Publicly available cancer data}
\label{sec:data}

Data related to multiple different cancer types were downloaded from free and publicly available sources such as \gls{tcga} and the \gls{icgc}.

\subsection{\gls{tcga}}
\label{subsec:tcga}

The clinical data for all cancer types (33 types in total) were downloaded from \gls{tcga} database and were checked for the height and weight data for each sample, for each cancer type.
Any cancer type with no weight and/or height data of the samples were excluded from the project, as no \gls{bmi} information can be obtained without these data.

This gave a total of eight cancer types (\gls{blca}, \gls{cesc}, \gls{coad}, \gls{kirp}, \gls{lihc}, \gls{read}, \gls{skcm}, \gls{ucec}) to be used for further analyses.

\subsection{\gls{icgc}}
\label{subsec:icgc}

RNA-seq data for the eight cancer types with sample \gls{bmi} information available were downloaded from the \gls{icgc} database.
The RNA-seq data for each of the eight cancer types were imported into R.

\section{Validation of methodology used by \citet{Creighton2012}}
\label{sec:methodvalidation}

\subsection{Raw data from the \citet{Creighton2012} study}
\label{subsec:rawdatacr}

The raw microarray gene expression data files of all the samples used by \citet{Creighton2012} were downloaded from the \gls{geo} database.
Raw data were imported into R/RStudio for further analyses.

(add data/sample composition -- ethinicity, grade, etc -- maybe in table form)

\subsection{Normalisation of raw \citet{Creighton2012} data}
\label{subsec:normcrdata}

The raw microarray data for all 103 samples in the study were normalised using \gls{rma} method from the \textit{affy} package in R.
The data was then standardised so that each of the genes had a mean of 0 and a standard deviation of 1.
(mention scaling anything over 3 were made to be 3 for heatmap scaling)

\subsection{Obesity associated gene probes from the \citet{Creighton2012} study}
\label{subsec:crobsgene}

799 obesity associated gene probes identified by \citet{Creighton2012} were downloaded from (where? - i think the same place as microarray data).
The list of gene probes were imported into R for further analyses and validation of methodology.

\subsection{Differential gene expression analysis}
\label{subsec:deg}

All 103 samples from the \citet{Creighton2012} study were split into two groups (obese and non-obese) based on the sample \gls{bmi} status as defined by the \gls{who} classification (see Table~\ref{tab:whobmiclass}).
The difference in the gene expression between these two sample groups were analysed by fitting a linear model to the data, using the \textit{limma} package in R.
The top 799 differentially expressed gene probes that had p-value \textless{} 0.01 were identified.

The process above was repeated on the same data, with the same group (obese and non-obese) after controlling for all other clinical variables that were not related to the sample \gls{bmi} (age, sex, ethinicity, tumour grade, \gls{ln} status, \gls{pr} status, \gls{er} status, \gls{her2} status).
the top 799 differentially expressed gene probes that had p-value \textless{} 0.01 between the two groups were identified.

To completely remove the effect of ethinicity on the analysis, the analysis was repeated using only the Caucasian samples.
This reduced the sample size from 103 samples to 77 samples.
The data was controlled for all other clinical variables except the variables that were related to \gls{bmi} (age, sex, tumour grade, \gls{ln} status, \gls{pr} status, \gls{er} status, and \gls{her2} status).
Again, the top 799 differentially expressed genes between the two groups with p-value \textless{} 0.01 were identified.

\subsection{Comparison of \citet{Creighton2012} gene probes and identified gene probes}
\label{subsec:cfgeneprobes}

The obesity associated gene probes identified by \citet{Creighton2012} were compared with each of the three sets of gene probes identified from the differential gene expression analysis (see \ref{subsec:deg} \nameref{subsec:deg}).
The presence of any of the \citet{Creighton2012} genes were checked for in each of the gene sets identified at a p-value \textless{} 0.05, as well as the top 799 gene probes at p-Value \textless{} 0.01.

\subsection{\gls{svd}}
\label{subsec:svd}

(need svd derivation and why I'm using it)

\subsection{Correlation between obesity associated metagene and sample \gls{bmi}}
\label{subsec:metavsbmi}

In order to check whether the metagenes were actually specific to the \gls{bmi} status of the samples, the metagenes were plotted against both the \gls{bmi} statuses and the \gls{bmi} values of the samples.
The metagenes were ranked for each sample, and a box plot and a scatter plot were plotted for sample \gls{bmi} statuses and \gls{bmi} values, respectively.

\section{Applying metagene to other cancer types}
\label{sec:metagene}

(How you could transfer the metagene to other cancer data)

\subsection{Transformation matrix}
\label{subsec:transmat}

(why transformation matrix is used)
(refer back to eqiations in the  svd section)

\section{Pathway Enrichment Analysis}
\label{sec:pathenrich}

### 2016-05-20 thesis/misc/glossary.tex
added glossary file

\newacronym{blca}{BLCA}{bladder urothelial carcinoma}
\newacronym{cesc}{CESC}{cervical squamous cell carcinoma and endocervical adenocarcinoma}
\newacronym{coad}{COAD}{colon adenocarcinoma}
\newacronym{kirp}{KIRP}{kidney renal clear cell carcinoma}
\newacronym{lihc}{LIHC}{liver hepatocellular carcinoma}
\newacronym{read}{READ}{rectum adenocarcinoma}
\newacronym{skcm}{SKCM}{skin cutaneous melanoma}
\newacronym{ucec}{UCEC}{uterine corpus endometrial carcinoma}
\newacronym{geo}{GEO}{Gene Expression Omnibus}
\newacronym{tcga}{TCGA}{The Cancer Genome Atlas}
\newacronym{icgc}{ICGC}{International Cancer Genome Consortium}
\newacronym{bmi}{BMI}{Body Mass Index}
\newacronym{who}{WHO}{World Health Organization}
\newacronym{ln}{LN}{Lymph node}
\newacronym{pr}{PR}{Progesterone receptor}
\newacronym{er}{ER}{Estrogen receptor}
\newacronym{her2}{HER2}{Human epidermal growth factor receptor 2}
\newacronym{svd}{SVD}{Singular Value Decomposition}
\newacronym{rma}{RMA}{Robust Multi Array}

### 2016-05-20 thesis/thesis.tex
added glossary file

\documentclass[a4paper, 12pt, oneside]{report}

% Preamble:
\usepackage[T1]{fontenc}                     % package for specifying output font encoding
\usepackage[utf8]{inputenc}                  % package for specifying input font encoding
\usepackage{natbib}                          % package for reference management
\usepackage{graphicx}                        % package for graphics handling (i.e. figure images)
\usepackage{mathptmx}                        % package for Times New Roman font
\usepackage{chngcntr}                        % package for making continuous figure numbering, instead of section-based numbering
\usepackage{setspace}                        % package for setting line space (used in title page)
\usepackage[hidelinks, colorlinks]{hyperref} % package for hyperlinking references - hidelinks option removes the ugly boxes around the huperlinks
\usepackage[md]{titlesec}                    % package for tweaking the chapter/section/subsection titles format
\usepackage[acronym, toc]{glossaries}          % package for making glossaries

% Set up the hyperlinks - hange the colour of the links to blue:
\hypersetup{allcolors=blue}

% change figure numbering from section-based to continuous:
\counterwithout{figure}{chapter}
\counterwithout{equation}{chapter}
\counterwithout{table}{chapter}

% If you want to compile specific files, uncomment the line below and add the filenames in the {} :
%\includeonly{}

% Change the title of bibliography to references:
\renewcommand{\bibname}{References}

% To generate glossaries:
\include{./misc/glossary}
\makeglossaries

\begin{document}

\include{./misc/title}

\pagenumbering{roman} % set page numbering to roman numeral

\include{./misc/abstract}

\include{./misc/acknowledgement}

\tableofcontents

\addcontentsline{toc}{section}{\contentsname}

\clearpage

\addcontentsline{toc}{section}{\listfigurename}

\listoffigures

\clearpage

\addcontentsline{toc}{section}{\listtablename}

\listoftables

\clearpage

\printglossaries

\clearpage

% \listoftables

% \clearpage

\pagenumbering{arabic} % revert the page numbering to normal numbering
\glsresetall

\include{./intro/intro}

\include{./methods/methods}

\include{./results/results}

\include{./results/results2}

\include{./discussion/discussion}

% Use the reference style specified in BiocRefStyle.bst:
\bibliographystyle{BiocRefStyle}

% Print out the bibliography:
\bibliography{MSc}

\end{document}

### 2016-05-24 thesis/methods/methods.tex
tmp commit

@@
 
-Samples were classified based on the \gls{who} definition, as shown in Table~\ref{tab:whobmiclass}.
+Samples were classified based on the \gls{who} definition, as shown in Table~\cref{tab:whobmiclass}.
 \begin{table}[hb]
@@
 
-The clinical data for all cancer types (33 types in total) were downloaded from \gls{tcga} database and were checked for the height and weight data for each sample, for each cancer type.
+The clinical data for all available cancer types (33 types in total) were downloaded from \gls{tcga} database and were checked for the height and weight data for each sample, for each cancer type.
 Any cancer type with no weight and/or height data of the samples were excluded from the project, as no \gls{bmi} information can be obtained without these data.
 
-This gave a total of eight cancer types (\gls{blca}, \gls{cesc}, \gls{coad}, \gls{kirp}, \gls{lihc}, \gls{read}, \gls{skcm}, \gls{ucec}) to be used for further analyses.
+This gave a total of eight cancer types: \gls{blca}, \gls{cesc}, \gls{coad}, \gls{kirp}, \gls{lihc}, \gls{read}, \gls{skcm}, \gls{ucec}.
+All eight cancer types were used for further analyses.
 
@@
 
-RNA-seq data for the eight cancer types with sample \gls{bmi} information available were downloaded from the \gls{icgc} database.
-The RNA-seq data for each of the eight cancer types were imported into R.
+\gls{rnaseq} data for the eight cancer types with sample \gls{bmi} information available were downloaded from the \gls{icgc} database.
+The \gls{rnaseq} data for each of the eight cancer types were imported into \gls{r}.
 
-\section{Validation of methodology used by \citet{Creighton2012}}
-\label{sec:methodvalidation}
+%TODO: change the title name?
+\section{\gls{icgc} sample exclusion criteria}
+\label{sec:sampleexclusion}
+
+Samples from the downloaded data were removed using the following criteria.
+
+\subsection{Sample BMI availability}
+\label{subsec:samplebmiavailability}
+
+Some of the samples did not have either height or weight data, or both.
+These samples were removed from the analyses.
+
+\subsection{Sample ID conversion and matching}
+\label{subsec:sampleidconversion}
+
+%TODO: check the actual ID stripping process
+%TODO: check the WGCNA function and check if it uses median/mean/whatever for choosing the sample
+
+\gls{icgc} sample ID in the \gls{rnaseq} data were stripped so that the ID from the \gls{rnaseq} data matched with the \gls{tcga} ID in the clinical data.
+After stripping the ID, samples were checked for any overlaps.
+Where there was an overlap in the sample ID in the \gls{rnaseq} data, a single sample was chosen to represent that particular ID by using the \texttt{collapseRows} function in the \textit{WGCNA} package in \gls{r}.
+
+In each of the eight cancer types, the \gls{tcga} sample ID in the clinical data were cross-checked with the \gls{tcga} sample ID in the \gls{rnaseq} data, and vice versa.
+Any sample that did not have either clinical or \gls{rnaseq} data were removed, so both clinical and \gls{rnaseq} data were available for all of the samples that were included in the analyses.
+See \cref{tab:samplesize} for a summary of the number of samples included in the analyses for each cancer type.
+
+\begin{table}[h]
+	\caption{Summary of the total number of samples included in the analyses for each cancer type}
+	\label{tab:samplesize}
+	\begin{center}
+		\begin{tabular}{cc}
+			\textbf{Cancer Type}   & \textbf{Total} \\
+			\hline
+			\rule{0pt}{2.25ex}BLCA & 261   \\
+			CESC                   & 224   \\
+			COAD                   & 226   \\
+			KIRP                   & 124   \\
+			LIHC                   & 264   \\
+			READ                   & 73    \\
+			SKCM                   & 218   \\
+			UCEC                   & 482   \\
+		\end{tabular}
+	\end{center}
+\end{table}
+
+\section{Data processing -- \citet{Creighton2012} data}
+\label{sec:dataprocessingcr}
 
@@
 
-The raw microarray gene expression data files of all the samples used by \citet{Creighton2012} were downloaded from the \gls{geo} database.
-Raw data were imported into R/RStudio for further analyses.
+The raw microarray gene expression data and clinical data files of all *** samples from the \citet{Creighton2012} study were downloaded from the \gls{geo} database (include website name/GEO ID here).
+Raw data were imported into \gls{r}/RStudio for subsequent analyses.
 
-(add data/sample composition -- ethinicity, grade, etc -- maybe in table form)
+%TODO: add data/sample composition -- ethinicity, grade, etc -- maybe in table form
 
@@
 
-The raw microarray data for all 103 samples in the study were normalised using \gls{rma} method from the \textit{affy} package in R.
-The data was then standardised so that each of the genes had a mean of 0 and a standard deviation of 1.
-(mention scaling anything over 3 were made to be 3 for heatmap scaling)
+The raw microarray data for all 103 samples from the study were normalised using \gls{rma} method from the \textit{affy} package in \gls{r}.
+The data was then standardised so that each of the genes had a mean ($\mu$) of 0 and a standard deviation ($\sigma$) of 1.
+This ensured that all of the genes were on the same scale as one another, allowing for comparison.
 
@@
 
-799 obesity associated gene probes identified by \citet{Creighton2012} were downloaded from (where? - i think the same place as microarray data).
-The list of gene probes were imported into R for further analyses and validation of methodology.
+799 obesity associated gene probes identified by \citet{Creighton2012} were downloaded from (supplementary data??).
+The list of gene probes were imported into \gls{r}.
+The gene probe IDs were converted into gene symbols using the \textit{hgu133a.db} package in \gls{r}.
+Any duplicates and any gene symbols that were not present in both the Creighton data and the \gls{icgc} data were removed, which left *** unique obesity associated genes.
 
-\subsection{Differential gene expression analysis}
-\label{subsec:deg}
+\section{Validation of results -- \citet{Creighton2012}}
+\label{sec:valresultscr}
 
-All 103 samples from the \citet{Creighton2012} study were split into two groups (obese and non-obese) based on the sample \gls{bmi} status as defined by the \gls{who} classification (see Table~\ref{tab:whobmiclass}).
-The difference in the gene expression between these two sample groups were analysed by fitting a linear model to the data, using the \textit{limma} package in R.
+The results presented by \citet{Creighton2012} were validated using the same raw data they have used in their study.
+See  \cref{sec:dataprocessingcr} for details on how the data was processed before the analyses.
+
+\subsection{Metagene and sample gene expression pattern}
+\label{subsec:crmetage}
+
+The gene probe IDs were converted into gene symbols using the \textit{hgu133a.db} package in \gls{r}.
+Where there was a duplicated gene in the data, appropriate gene data was chosen to represent that particular gene by using the \texttt{collapseRows} function in the \textit{WGCNA} package in \gls{r}.
+All of the *** obesity associated genes from  \cref{subsec:crobsgene} were selected from the data.
+
+\gls{svd} was applied to this data to obtain the decomposed matrices of the data.
+%TODO: how do I explain how I got the \gls{metagene} scores?
+The \gls{metagene} scores of the samples were obtained by taking the first principal component of the (eigenvector...?).
+The \gls{metagene} scores were ranked based on the number of samples in the data (103 samples in Creighton data), and the values of the scores were between 0 and 1.
+
+These \gls{metagene} scores were plotted together with the gene expression of the obesity associated gene probes for each sample on a heatmap.
+The samples in the heatmap were ordered based on their ranked \gls{metagene} scores.
+To have a better colour scales in the heatmap, any expression values that were greater than 3 or less than -3 were assigned the value of 3 and -3, respectively.
+
+\subsection{Metagene and sample \gls{bmi}}
+\label{subsec:crmetabmi}
+
+The same \gls{metagene} scores as in the previous section (\cref{subsec:crmetage}) were used to compare the \gls{metagene} scores with the samples' \gls{bmi} statuses and values.
+Boxplot was plotted to show the association of the \gls{metagene} scores with the sample \gls{bmi} status, and scatter plot was plotted to show the association with the sample \gls{bmi} value.
+
+%TODO: check the formal definition of these test/values
+For the boxplot, the p-value was obtained from the ANOVA test.
+For the scatter plot, the regression line was plotted and the adjusted $R^2$ value was calculated for the regression line.
+
+\section{Creighton metagene in \gls{icgc} cancer data}
+\label{sec:crmetaicgc}
+
+\subsection{\gls{icgc} data processing}
+\label{subsec:datprocicgc}
+
+The obesity associated genes from \cref{subsec:crobsgene} were selected in each of the eight cancer types.
+All cancer data had 1 added to the count data, so no infinite values were introduced when the data was logged.
+All cancer data were logged to the base 10.
+
+All of the data were then standardised so that each of the genes had a mean ($\mu$) of 0 and a standard deviation ($\sigma$) of 1.
+Any ``NaN'' values in the data introduced from this were assigned a value of 0.
+
+\subsection{Transformation matrix}
+\label{subsec:transmatcr}
+
+Transformation matrix was formed from the decomposed matrices that resulted from the application of \gls{svd} to the Creighton data (see \cref{subsec:crmetage}).
+%TODO: add the derivation of transformation matrix??
+
+\subsection{Transformed \gls{icgc} data and metagene}
+\label{subsec:icgctranscr}
+
+Each of the eight cancer data from \cref{subsec:datprocicgc} were transformed using the transformation matrix from \cref{subsec:transmatcr}.
+Metagene scores of the samples in each cancer types were obtained by taking the first column of the cancer matrix that resulted from the transformation.
+The \gls{metagene} scores were ranked based on the number of samples in the data, and the scores ranged from 0 to 1.
+
+\subsection{Creighton metagene and sample gene expression in \gls{icgc} data}
+\label{subsec:crmetaicgcge}
+
+\Gls{metagene} scores from \cref{subsec:icgctranscr} were plotted together with the gene expression of the obesity associated gene probes for each sample on a heatmap, for each cancer type.
+The samples in the heatmap were ordered based on their ranked \gls{metagene} scores.
+To have a better colour scales in the heatmap, any expression values that were greater than 3 or less than -3 were assigned the value of 3 and -3, respectively.
+
+\subsection{Creighton metagene and sample \gls{bmi} in \gls{icgc} data}
+\label{subsec:crmetaicgcge}
+
+The same \gls{metagene} scores as in the previous section were used to compare the \gls{metagene} scores with the samples' \gls{bmi} statuses and values for each cancer type.
+Boxplot was plotted to show the association of the \gls{metagene} scores with the sample \gls{bmi} status, and scatter plot was plotted to show the association with the sample \gls{bmi} value.
+
+%TODO: check the formal definition of these test/values
+For the boxplot, the p-value was obtained from the ANOVA test.
+For the scatter plot, the regression line was plotted and the adjusted $R^2$ value was calculated for the regression line.
+
+%TODO: Not sure if I should cut this section to "see previous section for details"
+\section{Data processing -- \citet{Fuentes-Mattei2014} data}
+\label{sec:dataprocessingfm}
+
+\subsection{Raw data from the \citet{Fuentes-Mattei2014} study}
+\label{subsec:rawdatafm}
+
+The raw microarray gene expression data files of all *** samples from the  \citet{Fuentes-Mattei2014} study were downloaded from the \gls{geo} database (include website name/GEO ID here).
+Raw data were imported into \gls{r}/RStudio for subsequent analyses.
+
+%TODO: add data/sample composition -- ethinicity, grade, etc -- maybe in table form
+
+\subsection{Normalisation of raw \citet{Fuentes-Mattei2014} data}
+\label{subsec:normfmdata}
+
+The raw microarray data for all *** samples from the study were normalised using \gls{rma} method from the \textit{affy} package in \gls{r}.
+The data was then standardised so that each of the genes had a mean ($\mu$) of 0 and a standard deviation ($\sigma$) of 1.
+This ensured that all of the genes were on the same scale as one another, allowing for comparison.
+
+\subsection{Obesity associated gene probes from the \citet{Fuentes-Mattei2014} study}
+\label{subsec:fmobsgene}
+
+130 obesity associated gene probes identified by \citet{Fuentes-Mattei2014} were downloaded from (supplementary data??).
+The list of gene probes were imported into \gls{r}.
+Any gene symbols that were not present in the Fuentes-Mattei data, the Creighton data and the \gls{icgc} data were removed, which left *** unique obesity associated genes.
+
+\section{Validation of results -- \citet{Fuentes-Mattei2014}}
+\label{sec:valresultsfm}
+
+The results presented by \citet{Fuentes-Mattei2014} were validated using the \citet{Creighton2012} data, as the clinical data of the samples were not available.
+See \cref{subsec:rawdatacr,subsec:normcrdata} for details on how the data was processed before the analyses.
+
+\subsection{Fuentes-Mattei transformation matrix}
+\label{subsec:transmatfm}
+
+\gls{svd} was applied to Fuentes-Mattei data to obtain the decomposed matrices of the data, and this was used to make the transformation matrix.
+%TODO: add the derivation of transformation matrix??
+
+\subsection{Fuentes-Mattei metagene and gene expression pattern in Creighton data}
+\label{subsec:fmmetage}
+
+The gene probe IDs in Creighton data were converted into gene symbols using the \textit{hgu133a.db} package in \gls{r}.
+Where there was a duplicated gene, appropriate gene data was chosen to represent that particular gene by using the \texttt{collapseRows} function in the \textit{WGCNA} package in \gls{r}.
+All of the *** obesity associated genes from  \cref{subsec:fmobsgene} were selected from the data.
+
+%TODO: how do I explain how I got the \gls{metagene} scores?
+The \gls{metagene} scores of the samples were obtained by taking the first principal component of the transformed matrix.
+The \gls{metagene} scores were ranked based on the number of samples in the data (*** samples in \citet{Creighton2012} data), and the scores ranged from 0 to 1.
+
+These \gls{metagene} scores were plotted together with the gene expression of the obesity associated gene probes for each sample on a heatmap.
+The samples in the heatmap were ordered based on their ranked \gls{metagene} scores.
+To have a better colour scales in the heatmap, any expression values that were greater than 3 or less than -3 were assigned the value of 3 and -3, respectively.
+
+\subsection{Metagene and sample \gls{bmi}}
+\label{subsec:fmmetabmi}
+
+The same \gls{metagene} scores from the previous section were used to compare the \gls{metagene} scores with the samples' \gls{bmi} statuses and values.
+Boxplot was plotted to show the association of the \gls{metagene} scores with the sample \gls{bmi} status, and scatter plot was plotted to show the association with the sample \gls{bmi} value.
+
+%TODO: check the formal definition of these test/values
+For the boxplot, the p-value was obtained from the ANOVA test.
+For the scatter plot, the regression line was plotted and the adjusted $R^2$ value was calculated for the regression line.
+
+\section{Fuentes-Mattei metagene in \gls{icgc} cancer data}
+\label{sec:fmmetaicgc}
+
+\section{Differential gene expression analysis}
+\label{sec:deg}
+
+All 103 samples from the \citet{Creighton2012} study were split into two groups (obese and non-obese) based on the sample \gls{bmi} status as defined by the \gls{who} classification (see \cref{tab:whobmiclass}).
+The difference in the gene expression between these two sample groups were analysed by fitting a linear model to the data, using the \textit{limma} package in \gls{r}.
 The top 799 differentially expressed gene probes that had p-value \textless{} 0.01 were identified.
@@
 
-The obesity associated gene probes identified by \citet{Creighton2012} were compared with each of the three sets of gene probes identified from the differential gene expression analysis (see \ref{subsec:deg} \nameref{subsec:deg}).
-The presence of any of the \citet{Creighton2012} genes were checked for in each of the gene sets identified at a p-value \textless{} 0.05, as well as the top 799 gene probes at p-Value \textless{} 0.01.
+The obesity associated gene probes identified by \citet{Creighton2012} were compared with each of the three sets of gene probes identified from the differential gene expression analysis (see \cref{sec:deg} \nameref{sec:deg}).
+The presence of any of the \citet{Creighton2012} genes were checked for in each of the gene sets identified at a p-value \textless{} 0.05, as well as the top 799 gene probes at p-value \textless{} 0.01.
 

### 2016-05-28 thesis/methods/methods.tex
tmp commit

@@
 The \gls{rnaseq} data for each of the eight cancer types were imported into \gls{r}.
+%TODO: add in how I processed the ICGC data format into genes by samples.
 
@@
 
-\section{Data processing -- \citet{Creighton2012} data}
+\section{Data processing -- Creighton data}
 \label{sec:dataprocessingcr}
@@
 
-The raw microarray gene expression data and clinical data files of all *** samples from the \citet{Creighton2012} study were downloaded from the \gls{geo} database (include website name/GEO ID here).
+The raw microarray gene expression data files of all *** samples from the \citet{Creighton2012} study were downloaded from the \gls{geo} database (\gls{geo} accession ID: GSE24185).
+Clinical data of the samples were obtained from the supplementary table 1 from their paper.
 Raw data were imported into \gls{r}/RStudio for subsequent analyses.
@@
 
-\subsection{Normalisation of raw \citet{Creighton2012} data}
+\subsection{Normalisation of raw Creighton data}
 \label{subsec:normcrdata}
@@
 
-799 obesity associated gene probes identified by \citet{Creighton2012} were downloaded from (supplementary data??).
+799 obesity associated gene probes identified by \citet{Creighton2012} were obtained from the supplementary data file 1 from their paper.
 The list of gene probes were imported into \gls{r}.
@@
 The results presented by \citet{Creighton2012} were validated using the same raw data they have used in their study.
-See  \cref{sec:dataprocessingcr} for details on how the data was processed before the analyses.
+See  \cref{sec:dataprocessingcr} for details on how the raw data was accessed and processed.
 
@@
 %TODO: how do I explain how I got the \gls{metagene} scores?
-The \gls{metagene} scores of the samples were obtained by taking the first principal component of the (eigenvector...?).
+The \gls{metagene} scores of the samples were obtained by taking the first eigenvector of the (orthogonal matrix...?).
 The \gls{metagene} scores were ranked based on the number of samples in the data (103 samples in Creighton data), and the values of the scores were between 0 and 1.
@@
 
-\Gls{metagene} scores from \cref{subsec:icgctranscr} were plotted together with the gene expression of the obesity associated gene probes for each sample on a heatmap, for each cancer type.
+\Gls{metagene} scores from \cref{subsec:icgctranscr} were plotted together with the gene expression of the obesity associated gene probes for each sample on a heatmap for each cancer type.
+The data from \cref{subsec:datprocicgc} were used for the heatmap.
 The samples in the heatmap were ordered based on their ranked \gls{metagene} scores.
@@
 %TODO: Not sure if I should cut this section to "see previous section for details"
-\section{Data processing -- \citet{Fuentes-Mattei2014} data}
+\section{Data processing -- Fuentes-Mattei data}
 \label{sec:dataprocessingfm}
@@
 The raw microarray gene expression data files of all *** samples from the  \citet{Fuentes-Mattei2014} study were downloaded from the \gls{geo} database (include website name/GEO ID here).
+Clinical data for the samples in this study was not available.
 Raw data were imported into \gls{r}/RStudio for subsequent analyses.
@@
 
-\subsection{Normalisation of raw \citet{Fuentes-Mattei2014} data}
+\subsection{Normalisation of raw Fuentes-Mattei data}
 \label{subsec:normfmdata}
@@
 The data was then standardised so that each of the genes had a mean ($\mu$) of 0 and a standard deviation ($\sigma$) of 1.
-This ensured that all of the genes were on the same scale as one another, allowing for comparison.
 
@@
 
-The results presented by \citet{Fuentes-Mattei2014} were validated using the \citet{Creighton2012} data, as the clinical data of the samples were not available.
-See \cref{subsec:rawdatacr,subsec:normcrdata} for details on how the data was processed before the analyses.
+The results presented by \citet{Fuentes-Mattei2014} were validated using the Creighton data, as the clinical data of the samples were not available.
+See \cref{subsec:rawdatacr,subsec:normcrdata} for details on how the data was processed.
 
@@
 %TODO: how do I explain how I got the \gls{metagene} scores?
-The \gls{metagene} scores of the samples were obtained by taking the first principal component of the transformed matrix.
-The \gls{metagene} scores were ranked based on the number of samples in the data (*** samples in \citet{Creighton2012} data), and the scores ranged from 0 to 1.
+The \gls{metagene} scores of the samples were obtained by taking the first eigenvector of the transformed Creighton data.
+The \gls{metagene} scores were ranked based on the number of samples in the data (*** samples in Creighton data), and the scores ranged from 0 to 1.
 
@@
 The same \gls{metagene} scores from the previous section were used to compare the \gls{metagene} scores with the samples' \gls{bmi} statuses and values.
-Boxplot was plotted to show the association of the \gls{metagene} scores with the sample \gls{bmi} status, and scatter plot was plotted to show the association with the sample \gls{bmi} value.
-
-%TODO: check the formal definition of these test/values
-For the boxplot, the p-value was obtained from the ANOVA test.
-For the scatter plot, the regression line was plotted and the adjusted $R^2$ value was calculated for the regression line.
+Box and scatter plots, and any statistical values associated with these plots were calculated as in \cref{subsec:crmetaicgcge}.
 
@@
 
+\subsection{\gls{icgc} data processing}
+\label{subsec:datprocicgcfm}
+
+All of the \gls{icgc} data were processed in a similar fashion as in \cref{subsec:datprocicgc}, but the obesity associated genes from \cref{subsec:fmobsgene} were used instead.
+
+\subsection{Transformation matrix}
+\label{subsec:icgctransmatfm}
+
+The same transformation matrix from \cref{subsec:transmatfm} was used.
+
+\subsection{Transformed \gls{icgc} data and metagene}
+\label{subsec:icgctransfm}
+
+Each of the eight cancer data from \cref{subsec:datprocicgcfm} were transformed using the transformation matrix from \cref{subsec:transmatfm}.
+Metagene scores were ranked as in \cref{subsec:icgctranscr}.
+
+\subsection{Fuentes-Mattei metagene and sample gene expression in \gls{icgc} data}
+\label{subsec:fmmetaicgcge}
+
+\Gls{metagene} scores from \cref{subsec:icgctransfm} were plotted together with the gene expression of the obesity associated gene probes for each sample on a heatmap for each cancer type.
+The data from \cref{subsec:datprocicgcfm} were used for the heatmap.
+The samples in the heatmap were ordered based on their ranked \gls{metagene} scores.
+To have a better colour scales in the heatmap, any expression values that were greater than 3 or less than -3 were assigned the value of 3 and -3, respectively.
+
+\subsection{Creighton metagene and sample \gls{bmi} in \gls{icgc} data}
+\label{subsec:fmmetaicgcge}
+
+The same \gls{metagene} scores as in the previous section were used to compare the \gls{metagene} scores with the samples' \gls{bmi} statuses and values for each cancer type.
+Box and scatter plots, and any statistical values associated with these plots were calculated as in \cref{subsec:crmetaicgcge}.
+
 \section{Differential gene expression analysis}
@@
 
-All 103 samples from the \citet{Creighton2012} study were split into two groups (obese and non-obese) based on the sample \gls{bmi} status as defined by the \gls{who} classification (see \cref{tab:whobmiclass}).
+\subsection{Data processing and subsetting}
+\label{subsec:subdatadeg}
+
+Raw data from \cref{subsec:rawdatacr,subsec:normcrdata} were used in the analysis.
+Together with the raw data, few other data sets were created based on the raw data: residual data, Caucasian-only raw data, and Caucasian-only residual data.
+
+Residual data was created from the raw data after it has been controlled for all clinical variables, except those related to \gls{bmi}.
+These clinical variables included age, sex, ethinicity, tumour grade, \gls{pr}, \gls{er}, \gls{her2} statuses, and \gls{ln} status.
+
+Caucasian-only data was created by taking only the Caucasian samples from the raw data, which reduced the sample size from 103 samples to 77 samples.
+Caucasian-only residual data was created using the same approach as with the residual data: the data that remained after all the non-\gls{bmi} clinical variables were controlled in the Caucasian-only raw data.
+
+\subsection{Group classification}
+\label{subsec:crdegclass}
+
+In each of the datasets from \cref{subsec:subdatadeg}, all of the samples were split into two groups: obese and non-obese groups.
+These two groups were classified based on the sample \gls{bmi}, where the obese groups were defined as samples with \gls{bmi}  \textgreater{} 30, and non-obese group were defined as samples with \gls{bmi}  \textless{} 30 (see \cref{tab:whobmiclass}).
+
+\subsection{Gene expression analysis}
+\label{subsec:crdeg}
+
+The gene probe IDs were converted into gene symbols using the \textit{hgu133a.db} package in \gls{r}.
+Where there was a duplicated gene in the data, appropriate gene data was chosen to represent that particular gene by using the \texttt{collapseRows} function in the \textit{WGCNA} package in \gls{r}.
+
 The difference in the gene expression between these two sample groups were analysed by fitting a linear model to the data, using the \textit{limma} package in \gls{r}.
 The top 799 differentially expressed gene probes that had p-value \textless{} 0.01 were identified.
+
+\newpage
 

### 2016-05-28 thesis/methods/methods.tex
tmp commit

@@
 
+The difference in the gene expression between the obese and non-obese sample groups in each of the four data sets described in \cref{subsec:subdatadeg} were analysed by fitting a linear model to the data, using the \textit{limma} package in \gls{r}.
+From the top significantly expressed genes produced by the \texttt{topTable} function, all of the gene probes that had a p-value \textless{} 0.01 were selected.
+
+When there were more than 799 gene probes in the list of significant genes, the top 799 gene probes were selected from the list and used as the gene probe set for the metagene analysis, for that dataset.
+
+For each gene probe set, the gene probes that were common with the obesity associated genes identified by \citet{Creighton2012} were selected and used as a separate gene probe set.
+There were eight metagenes in total after this process: four gene probe sets from the datasets from \cref{subsec:subdatadeg}, and four gene probe sets that were in common with the obesity associated genes from \citet{Creighton2012} study.
+
+\subsection{Gene probe set conversion}
+\label{subsec:gpconv}
+
+\subsection{BMI metagenes in Creighton data}
+\label{subsec:bmidegmetacr}
+
+\subsection{BMI metagenes in \gls{icgc} data}
+\label{subsec:bmidegmetaicgc}
+
 The gene probe IDs were converted into gene symbols using the \textit{hgu133a.db} package in \gls{r}.
 Where there was a duplicated gene in the data, appropriate gene data was chosen to represent that particular gene by using the \texttt{collapseRows} function in the \textit{WGCNA} package in \gls{r}.
-
-The difference in the gene expression between these two sample groups were analysed by fitting a linear model to the data, using the \textit{limma} package in \gls{r}.
-The top 799 differentially expressed gene probes that had p-value \textless{} 0.01 were identified.
 

### 2016-05-30 thesis/methods/methods.tex
tmp commit

@@
 %TODO: change the title name?
-\section{\gls{icgc} sample exclusion criteria}
-\label{sec:sampleexclusion}
+\subsection{\gls{icgc} sample exclusion criteria}
+\label{subsec:sampleexclusion}
 
@@
 
-\subsection{Sample BMI availability}
-\label{subsec:samplebmiavailability}
+\subsubsection{\textbf{Sample BMI availability}}
+\label{subsubsec:samplebmiavailability}
 
@@
 
-\subsection{Sample ID conversion and matching}
-\label{subsec:sampleidconversion}
+\subsubsection{\textbf{Sample ID conversion and matching}}
+\label{subsubsec:sampleidconversion}
 
@@
 
-\subsection{Metagene and sample gene expression pattern}
-\label{subsec:crmetage}
+\subsection{Creighton metagene generation}
+\label{subsec:crmetagen}
 
@@
 
-These \gls{metagene} scores were plotted together with the gene expression of the obesity associated gene probes for each sample on a heatmap.
+\subsection{Metagene and sample gene expression pattern}
+\label{subsec:crmetage}
+
+\Gls{metagene} scores from \cref{subsec:crmetagen} were plotted together with the gene expression of the obesity associated gene probes for each sample on a heatmap.
 The samples in the heatmap were ordered based on their ranked \gls{metagene} scores.
@@
 Each of the eight cancer data from \cref{subsec:datprocicgc} were transformed using the transformation matrix from \cref{subsec:transmatcr}.
-Metagene scores of the samples in each cancer types were obtained by taking the first column of the cancer matrix that resulted from the transformation.
+Metagene scores of the samples in each cancer types were obtained by taking the first eigenvector of the cancer matrix that resulted from the transformation.
 The \gls{metagene} scores were ranked based on the number of samples in the data, and the scores ranged from 0 to 1.
 
-\subsection{Creighton metagene and sample gene expression in \gls{icgc} data}
+\subsection{Creighton metagene and sample gene expression}
 \label{subsec:crmetaicgcge}
 
-\Gls{metagene} scores from \cref{subsec:icgctranscr} were plotted together with the gene expression of the obesity associated gene probes for each sample on a heatmap for each cancer type.
+\Gls{metagene} scores from \cref{subsec:icgctranscr} were plotted together with the gene expression of the obesity associated gene probes on a heatmap for each cancer type.
 The data from \cref{subsec:datprocicgc} were used for the heatmap.
@@
 
-\subsection{Creighton metagene and sample \gls{bmi} in \gls{icgc} data}
+\subsection{Creighton metagene and sample \gls{bmi}}
 \label{subsec:crmetaicgcge}
@@
 
-\subsection{Fuentes-Mattei metagene and gene expression pattern in Creighton data}
-\label{subsec:fmmetage}
+\subsection{Fuentes-Mattei metagene generation}
+\label{subsec:fmmetagen}
 
@@
 
+\subsection{Fuentes-Mattei metagene and gene expression in Creighton data}
+\label{subsec:fmmetage}
+
 These \gls{metagene} scores were plotted together with the gene expression of the obesity associated gene probes for each sample on a heatmap.
@@
 
-\subsection{Metagene and sample \gls{bmi}}
+\subsection{Fuentes-Mattei metagene and sample \gls{bmi} in Creighton data}
 \label{subsec:fmmetabmi}
@@
 
-\subsection{Fuentes-Mattei metagene and sample gene expression in \gls{icgc} data}
+\subsection{Fuentes-Mattei metagene and sample gene expression}
 \label{subsec:fmmetaicgcge}
@@
 
-\subsection{Creighton metagene and sample \gls{bmi} in \gls{icgc} data}
+\subsection{Creighton metagene and sample \gls{bmi}}
 \label{subsec:fmmetaicgcge}
@@
 The difference in the gene expression between the obese and non-obese sample groups in each of the four data sets described in \cref{subsec:subdatadeg} were analysed by fitting a linear model to the data, using the \textit{limma} package in \gls{r}.
-From the top significantly expressed genes produced by the \texttt{topTable} function, all of the gene probes that had a p-value \textless{} 0.01 were selected.
+From the list of significantly expressed (genes produced by the \texttt{topTable} function), all of the gene probes that had a p-value \textless{} 0.01 were selected.
 
-When there were more than 799 gene probes in the list of significant genes, the top 799 gene probes were selected from the list and used as the gene probe set for the metagene analysis, for that dataset.
+When there was more than 799 gene probes in the list of significant genes, the top 799 gene probes were selected from the list.
+This list of 799 gene probes represented the obesity associated metagene for that dataset.
 
 For each gene probe set, the gene probes that were common with the obesity associated genes identified by \citet{Creighton2012} were selected and used as a separate gene probe set.
-There were eight metagenes in total after this process: four gene probe sets from the datasets from \cref{subsec:subdatadeg}, and four gene probe sets that were in common with the obesity associated genes from \citet{Creighton2012} study.
+There were eight metagenes in total after this process: four gene probe sets from the datasets from \cref{subsec:subdatadeg}, and four gene probe sets that were in common with the obesity associated genes from the \citet{Creighton2012} study.
 
-\subsection{Gene probe set conversion}
+\subsection{Gene probe set conversion and selection}
 \label{subsec:gpconv}
 
-\subsection{BMI metagenes in Creighton data}
-\label{subsec:bmidegmetacr}
+For each of the eight gene probe sets identified in \cref{subsec:crdeg}, the gene probe IDs were converted into gene symbols using the \textit{hgu133a.db} package in \gls{r}.
+Any duplicates and any gene symbols that were not present in both the Creighton data and the ICGC data were removed from the gene probe set.
+
+\section{BMI metagenes in Creighton data}
+\label{sec:bmidegmetacr}
+
+\subsection{BMI metagene generation}
+\label{subsec:bmimetagen}
+
+\subsection{BMI metagenes and sample gene expression in Creighton data}
+\label{subsec:gpconv}
+
+The gene probe IDs in the Creighton data were converted into gene symbols using the \textit{hgu133a.db} package in \gls{r}.
+Where there was a duplicated gene in the data, appropriate gene data was chosen to represent that particular gene by using the \texttt{collapseRows} function in the \textit{WGCNA} package in \gls{r}.
+
+For each BMI gene probe set identified in \cref{sec:deg}, relevant genes were taken from the Creighton data.
+The data was then standardised so all of the genes had a mean ($\mu$) of 0 and standard deviation ($\sigma$) of 1.
 
@@
 \label{subsec:bmidegmetaicgc}
-
-The gene probe IDs were converted into gene symbols using the \textit{hgu133a.db} package in \gls{r}.
-Where there was a duplicated gene in the data, appropriate gene data was chosen to represent that particular gene by using the \texttt{collapseRows} function in the \textit{WGCNA} package in \gls{r}.
 

### 2016-06-01 thesis/methods/methods.tex
tmp commit

@@
 All of the *** obesity associated genes from  \cref{subsec:fmobsgene} were selected from the data.
+The data was then standardised so that each of the genes had a mean ($\mu$) of 0 and a standard deviation ($\sigma$) of 1.
 
@@
 When there was more than 799 gene probes in the list of significant genes, the top 799 gene probes were selected from the list.
-This list of 799 gene probes represented the obesity associated metagene for that dataset.
+This list of 799 gene probes represented the gene probes that were associated with obesity  for that dataset.
+
+\subsection{Common gene probes from the \citet{Creighton2012} study}
+\label{subsec:crol}
 
 For each gene probe set, the gene probes that were common with the obesity associated genes identified by \citet{Creighton2012} were selected and used as a separate gene probe set.
-There were eight metagenes in total after this process: four gene probe sets from the datasets from \cref{subsec:subdatadeg}, and four gene probe sets that were in common with the obesity associated genes from the \citet{Creighton2012} study.
+There were eight gene probe sets in total after this process: four gene probe sets from the datasets from \cref{subsec:subdatadeg}, and four gene probe sets that were in common with the obesity associated genes from the \citet{Creighton2012} study.
+
+%TODO: add a table summarizing how many genes were in each gene probe set
 
@@
 
-For each of the eight gene probe sets identified in \cref{subsec:crdeg}, the gene probe IDs were converted into gene symbols using the \textit{hgu133a.db} package in \gls{r}.
+For each of the eight gene probe sets identified in \cref{subsec:crdeg,subsec:crol}, the gene probe IDs were converted into gene symbols using the \textit{hgu133a.db} package in \gls{r}.
 Any duplicates and any gene symbols that were not present in both the Creighton data and the ICGC data were removed from the gene probe set.
@@
 
+The gene probe IDs in Creighton data were converted into gene symbols using the \textit{hgu133a.db} package in \gls{r}.
+Where there was a duplicated gene, appropriate gene data was chosen to represent that particular gene by using the \texttt{collapseRows} function in the \textit{WGCNA} package in \gls{r}.
+For each gene probe sets identified in \cref{subsec:crdeg,subsec:crol}, the gene probes were selected from this data.
+The data was then standardised so that each of the genes had a mean ($\mu$) of 0 and a standard deviation ($\sigma$) of 1.
+
+SVD was applied to the standardised data, and the first eigenvector of the decomposed matrix was used as the sample metagene score for that gene probe set.
+The \gls{metagene} scores were ranked based on the number of samples in the data (*** samples in Creighton data), and the scores ranged from 0 to 1.
+
 \subsection{BMI metagenes and sample gene expression in Creighton data}
@@
 
-The gene probe IDs in the Creighton data were converted into gene symbols using the \textit{hgu133a.db} package in \gls{r}.
-Where there was a duplicated gene in the data, appropriate gene data was chosen to represent that particular gene by using the \texttt{collapseRows} function in the \textit{WGCNA} package in \gls{r}.
-
-For each BMI gene probe set identified in \cref{sec:deg}, relevant genes were taken from the Creighton data.
-The data was then standardised so all of the genes had a mean ($\mu$) of 0 and standard deviation ($\sigma$) of 1.
+For each \gls{metagene} scores in \cref{subsec:bmimetagen} were plotted together with the gene expression of the associated gene probes for each sample on a heatmap.
+The samples in the heatmap were ordered based on their ranked \gls{metagene} scores.
+To have a better colour scales in the heatmap, any expression values that were greater than 3 or less than -3 were assigned the value of 3 and -3, respectively.
 
@@
 \label{subsec:bmidegmetaicgc}
+
+The same \gls{metagene} scores from the previous section were used to compare the \gls{metagene} scores with the samples' \gls{bmi} statuses and values.
+Box and scatter plots, and any statistical values associated with these plots were calculated as in \cref{subsec:crmetaicgcge}.
 

### 2016-06-14 thesis/methods/methods.tex
re-writing methods section (it was too clunky)

--- 2016-05-28 thesis/methods/methods.tex
@@
 
+The original raw data from the \citet{Creighton2012} and \citet{Fuentes-Mattei2014} studies were downloaded from the \gls{geo} website.
 Data related to multiple different cancer types were downloaded from free and publicly available sources such as \gls{tcga} and the \gls{icgc}.
+
+\subsection{Raw data from the \citet{Creighton2012} study}
+\label{subsec:crrawdat}
+
+The raw microarray gene expression data files of all *** samples from the \citet{Creighton2012} study were downloaded from the \gls{geo} database (\gls{geo} accession ID: GSE24185).
+Clinical data of the samples were obtained from the supplementary table 1 from their paper.
+%TODO: add data/sample composition -- ethinicity, grade, etc -- maybe in table form
+799 obesity associated gene probes identified in the \citet{Creighton2012} study were obtained from the supplementary data file 1 from their paper.
+The raw expression data, clinical data, and the list of gene probes were imported into \gls{r}/RStudio.
+
+\subsection{Raw data from the \citet{Fuentes-Mattei2014} study}
+\label{subsec:fmrawdat}
+
+The raw microarray gene expression data files of all *** samples from the  \citet{Fuentes-Mattei2014} study were downloaded from the \gls{geo} database (include website name/GEO ID here).
+Clinical data for the samples in this study was not available.
+130 obesity associated gene probes identified by \citet{Fuentes-Mattei2014} were downloaded from (supplementary data??).
+The raw expression data and the list of gene probes were imported into \gls{r}/RStudio.
 
@@
 
-The clinical data for all available cancer types (33 types in total) were downloaded from \gls{tcga} database and were checked for the height and weight data for each sample, for each cancer type.
+The clinical data for all available cancer types (33 types in total) were downloaded from \gls{tcga} database and were checked for the height and weight data for each sample.
 Any cancer type with no weight and/or height data of the samples were excluded from the project, as no \gls{bmi} information can be obtained without these data.
@@
 
-\gls{rnaseq} data for the eight cancer types with sample \gls{bmi} information available were downloaded from the \gls{icgc} database.
+\gls{rnaseq} data for the eight cancer types identified in \cref{subsec:tcga} were downloaded from the \gls{icgc} database.
 The \gls{rnaseq} data for each of the eight cancer types were imported into \gls{r}.
+
+\section{Data processing}
+\label{sec:datproc}
+
+\subsection{Creighton data}
+\label{subsec:crdatproc}
+
+% \subsubsection{\textbf{Data processing}}
+
+\paragraph{\textit{Raw data normalisation}}
+
+The raw microarray data for all 103 samples from \cref{subsec:crrawdat} were normalised using the \gls{rma} method from the \textit{affy} package in \gls{r}.
+\gls{rma} normalisation was used to account for the inter-array differences that may have arisen during the data collection process.
+
+\paragraph{\textit{ Raw data standardisation }}
+
+The data was standardised so that each of the genes had a mean ($\mu$) of 0 and a standard deviation ($\sigma$) of 1.
+This ensured that all of the genes were on the same scale as one another, allowing for comparison across all the genes.
+
+\paragraph{\textit{ Raw data standardisation }}
+
+\subsection{Fuentes-Mattei data}
+\label{subsec:fmdatproc}
+
+% \subsubsection{\textbf{Data processing}}
+
+\paragraph{\textit{Raw data normalisation}}
+
+The raw microarray data for all *** samples from \cref{subsec:fmrawdat} were normalised using the \gls{rma} method from the \textit{affy} package in \gls{r}.
+\gls{rma} normalisation was used to account for the inter-array differences that may have arisen during the data collection process.
+
+\paragraph{\textit{ Raw data standardisation }}
+
+The data was standardised so that each of the genes had a mean ($\mu$) of 0 and a standard deviation ($\sigma$) of 1.
+This ensured that all of the genes were on the same scale as one another, allowing for comparison across all the genes.
+
+\subsection{\gls{tcga}}
+\label{subsec:datproctcga}
+
+\subsection{\gls{icgc}}
+\label{subsec:datprocicgc}
+
+\subsubsection{\textbf{Reformatting the raw data}}
+\label{subsec:icgcreformat}
+
 %TODO: add in how I processed the ICGC data format into genes by samples.
+TODO: add in how I processed the ICGC data format into genes by samples.
 
 %TODO: change the title name?
-\section{\gls{icgc} sample exclusion criteria}
-\label{sec:sampleexclusion}
+\subsubsection{\textbf{\gls{icgc} sample exclusion criteria}}
+\label{subsec:icgcsampleexclusion}
 
-Samples from the downloaded data were removed using the following criteria.
-
-\subsection{Sample BMI availability}
-\label{subsec:samplebmiavailability}
-
+\paragraph{\textit{Sample BMI availability}}
 Some of the samples did not have either height or weight data, or both.
@@
 
-\subsection{Sample ID conversion and matching}
-\label{subsec:sampleidconversion}
+\paragraph{\textit{Sample ID conversion and matching}}
 
@@
 
-\section{Data processing -- Creighton data}
-\label{sec:dataprocessingcr}
-
-\subsection{Raw data from the \citet{Creighton2012} study}
-\label{subsec:rawdatacr}
-
-The raw microarray gene expression data files of all *** samples from the \citet{Creighton2012} study were downloaded from the \gls{geo} database (\gls{geo} accession ID: GSE24185).
-Clinical data of the samples were obtained from the supplementary table 1 from their paper.
-Raw data were imported into \gls{r}/RStudio for subsequent analyses.
-
-%TODO: add data/sample composition -- ethinicity, grade, etc -- maybe in table form
-
 \subsection{Normalisation of raw Creighton data}
 \label{subsec:normcrdata}
-
-The raw microarray data for all 103 samples from the study were normalised using \gls{rma} method from the \textit{affy} package in \gls{r}.
-The data was then standardised so that each of the genes had a mean ($\mu$) of 0 and a standard deviation ($\sigma$) of 1.
-This ensured that all of the genes were on the same scale as one another, allowing for comparison.
 
@@
 
-799 obesity associated gene probes identified by \citet{Creighton2012} were obtained from the supplementary data file 1 from their paper.
-The list of gene probes were imported into \gls{r}.
 The gene probe IDs were converted into gene symbols using the \textit{hgu133a.db} package in \gls{r}.
 Any duplicates and any gene symbols that were not present in both the Creighton data and the \gls{icgc} data were removed, which left *** unique obesity associated genes.
-
-\section{Validation of results -- \citet{Creighton2012}}
-\label{sec:valresultscr}
-
-The results presented by \citet{Creighton2012} were validated using the same raw data they have used in their study.
-See  \cref{sec:dataprocessingcr} for details on how the raw data was accessed and processed.
-
-\subsection{Metagene and sample gene expression pattern}
-\label{subsec:crmetage}
-
-The gene probe IDs were converted into gene symbols using the \textit{hgu133a.db} package in \gls{r}.
-Where there was a duplicated gene in the data, appropriate gene data was chosen to represent that particular gene by using the \texttt{collapseRows} function in the \textit{WGCNA} package in \gls{r}.
-All of the *** obesity associated genes from  \cref{subsec:crobsgene} were selected from the data.
-
-\gls{svd} was applied to this data to obtain the decomposed matrices of the data.
-%TODO: how do I explain how I got the \gls{metagene} scores?
-The \gls{metagene} scores of the samples were obtained by taking the first eigenvector of the (orthogonal matrix...?).
-The \gls{metagene} scores were ranked based on the number of samples in the data (103 samples in Creighton data), and the values of the scores were between 0 and 1.
-
-These \gls{metagene} scores were plotted together with the gene expression of the obesity associated gene probes for each sample on a heatmap.
-The samples in the heatmap were ordered based on their ranked \gls{metagene} scores.
-To have a better colour scales in the heatmap, any expression values that were greater than 3 or less than -3 were assigned the value of 3 and -3, respectively.
-
-\subsection{Metagene and sample \gls{bmi}}
-\label{subsec:crmetabmi}
-
-The same \gls{metagene} scores as in the previous section (\cref{subsec:crmetage}) were used to compare the \gls{metagene} scores with the samples' \gls{bmi} statuses and values.
-Boxplot was plotted to show the association of the \gls{metagene} scores with the sample \gls{bmi} status, and scatter plot was plotted to show the association with the sample \gls{bmi} value.
-
-%TODO: check the formal definition of these test/values
-For the boxplot, the p-value was obtained from the ANOVA test.
-For the scatter plot, the regression line was plotted and the adjusted $R^2$ value was calculated for the regression line.
-
-\section{Creighton metagene in \gls{icgc} cancer data}
-\label{sec:crmetaicgc}
-
-\subsection{\gls{icgc} data processing}
-\label{subsec:datprocicgc}
-
-The obesity associated genes from \cref{subsec:crobsgene} were selected in each of the eight cancer types.
-All cancer data had 1 added to the count data, so no infinite values were introduced when the data was logged.
-All cancer data were logged to the base 10.
-
-All of the data were then standardised so that each of the genes had a mean ($\mu$) of 0 and a standard deviation ($\sigma$) of 1.
-Any ``NaN'' values in the data introduced from this were assigned a value of 0.
-
-\subsection{Transformation matrix}
-\label{subsec:transmatcr}
-
-Transformation matrix was formed from the decomposed matrices that resulted from the application of \gls{svd} to the Creighton data (see \cref{subsec:crmetage}).
-%TODO: add the derivation of transformation matrix??
-
-\subsection{Transformed \gls{icgc} data and metagene}
-\label{subsec:icgctranscr}
-
-Each of the eight cancer data from \cref{subsec:datprocicgc} were transformed using the transformation matrix from \cref{subsec:transmatcr}.
-Metagene scores of the samples in each cancer types were obtained by taking the first column of the cancer matrix that resulted from the transformation.
-The \gls{metagene} scores were ranked based on the number of samples in the data, and the scores ranged from 0 to 1.
-
-\subsection{Creighton metagene and sample gene expression in \gls{icgc} data}
-\label{subsec:crmetaicgcge}
-
-\Gls{metagene} scores from \cref{subsec:icgctranscr} were plotted together with the gene expression of the obesity associated gene probes for each sample on a heatmap for each cancer type.
-The data from \cref{subsec:datprocicgc} were used for the heatmap.
-The samples in the heatmap were ordered based on their ranked \gls{metagene} scores.
-To have a better colour scales in the heatmap, any expression values that were greater than 3 or less than -3 were assigned the value of 3 and -3, respectively.
-
-\subsection{Creighton metagene and sample \gls{bmi} in \gls{icgc} data}
-\label{subsec:crmetaicgcge}
-
-The same \gls{metagene} scores as in the previous section were used to compare the \gls{metagene} scores with the samples' \gls{bmi} statuses and values for each cancer type.
-Boxplot was plotted to show the association of the \gls{metagene} scores with the sample \gls{bmi} status, and scatter plot was plotted to show the association with the sample \gls{bmi} value.
-
-%TODO: check the formal definition of these test/values
-For the boxplot, the p-value was obtained from the ANOVA test.
-For the scatter plot, the regression line was plotted and the adjusted $R^2$ value was calculated for the regression line.
-
-%TODO: Not sure if I should cut this section to "see previous section for details"
-\section{Data processing -- Fuentes-Mattei data}
-\label{sec:dataprocessingfm}
-
-\subsection{Raw data from the \citet{Fuentes-Mattei2014} study}
-\label{subsec:rawdatafm}
-
-The raw microarray gene expression data files of all *** samples from the  \citet{Fuentes-Mattei2014} study were downloaded from the \gls{geo} database (include website name/GEO ID here).
-Clinical data for the samples in this study was not available.
-Raw data were imported into \gls{r}/RStudio for subsequent analyses.
-
-%TODO: add data/sample composition -- ethinicity, grade, etc -- maybe in table form
-
-\subsection{Normalisation of raw Fuentes-Mattei data}
-\label{subsec:normfmdata}
-
-The raw microarray data for all *** samples from the study were normalised using \gls{rma} method from the \textit{affy} package in \gls{r}.
-The data was then standardised so that each of the genes had a mean ($\mu$) of 0 and a standard deviation ($\sigma$) of 1.
-
-\subsection{Obesity associated gene probes from the \citet{Fuentes-Mattei2014} study}
-\label{subsec:fmobsgene}
-
-130 obesity associated gene probes identified by \citet{Fuentes-Mattei2014} were downloaded from (supplementary data??).
-The list of gene probes were imported into \gls{r}.
-Any gene symbols that were not present in the Fuentes-Mattei data, the Creighton data and the \gls{icgc} data were removed, which left *** unique obesity associated genes.
-
-\section{Validation of results -- \citet{Fuentes-Mattei2014}}
-\label{sec:valresultsfm}
-
-The results presented by \citet{Fuentes-Mattei2014} were validated using the Creighton data, as the clinical data of the samples were not available.
-See \cref{subsec:rawdatacr,subsec:normcrdata} for details on how the data was processed.
-
-\subsection{Fuentes-Mattei transformation matrix}
-\label{subsec:transmatfm}
-
-\gls{svd} was applied to Fuentes-Mattei data to obtain the decomposed matrices of the data, and this was used to make the transformation matrix.
-%TODO: add the derivation of transformation matrix??
-
-\subsection{Fuentes-Mattei metagene and gene expression pattern in Creighton data}
-\label{subsec:fmmetage}
-
-The gene probe IDs in Creighton data were converted into gene symbols using the \textit{hgu133a.db} package in \gls{r}.
-Where there was a duplicated gene, appropriate gene data was chosen to represent that particular gene by using the \texttt{collapseRows} function in the \textit{WGCNA} package in \gls{r}.
-All of the *** obesity associated genes from  \cref{subsec:fmobsgene} were selected from the data.
-
-%TODO: how do I explain how I got the \gls{metagene} scores?
-The \gls{metagene} scores of the samples were obtained by taking the first eigenvector of the transformed Creighton data.
-The \gls{metagene} scores were ranked based on the number of samples in the data (*** samples in Creighton data), and the scores ranged from 0 to 1.
-
-These \gls{metagene} scores were plotted together with the gene expression of the obesity associated gene probes for each sample on a heatmap.
-The samples in the heatmap were ordered based on their ranked \gls{metagene} scores.
-To have a better colour scales in the heatmap, any expression values that were greater than 3 or less than -3 were assigned the value of 3 and -3, respectively.
-
-\subsection{Metagene and sample \gls{bmi}}
-\label{subsec:fmmetabmi}
-
-The same \gls{metagene} scores from the previous section were used to compare the \gls{metagene} scores with the samples' \gls{bmi} statuses and values.
-Box and scatter plots, and any statistical values associated with these plots were calculated as in \cref{subsec:crmetaicgcge}.
-
-\section{Fuentes-Mattei metagene in \gls{icgc} cancer data}
-\label{sec:fmmetaicgc}
-
-\subsection{\gls{icgc} data processing}
-\label{subsec:datprocicgcfm}
-
-All of the \gls{icgc} data were processed in a similar fashion as in \cref{subsec:datprocicgc}, but the obesity associated genes from \cref{subsec:fmobsgene} were used instead.
-
-\subsection{Transformation matrix}
-\label{subsec:icgctransmatfm}
-
-The same transformation matrix from \cref{subsec:transmatfm} was used.
-
-\subsection{Transformed \gls{icgc} data and metagene}
-\label{subsec:icgctransfm}
-
-Each of the eight cancer data from \cref{subsec:datprocicgcfm} were transformed using the transformation matrix from \cref{subsec:transmatfm}.
-Metagene scores were ranked as in \cref{subsec:icgctranscr}.
-
-\subsection{Fuentes-Mattei metagene and sample gene expression in \gls{icgc} data}
-\label{subsec:fmmetaicgcge}
-
-\Gls{metagene} scores from \cref{subsec:icgctransfm} were plotted together with the gene expression of the obesity associated gene probes for each sample on a heatmap for each cancer type.
-The data from \cref{subsec:datprocicgcfm} were used for the heatmap.
-The samples in the heatmap were ordered based on their ranked \gls{metagene} scores.
-To have a better colour scales in the heatmap, any expression values that were greater than 3 or less than -3 were assigned the value of 3 and -3, respectively.
-
-\subsection{Creighton metagene and sample \gls{bmi} in \gls{icgc} data}
-\label{subsec:fmmetaicgcge}
-
-The same \gls{metagene} scores as in the previous section were used to compare the \gls{metagene} scores with the samples' \gls{bmi} statuses and values for each cancer type.
-Box and scatter plots, and any statistical values associated with these plots were calculated as in \cref{subsec:crmetaicgcge}.
-
-\section{Differential gene expression analysis}
-\label{sec:deg}
-
-\subsection{Data processing and subsetting}
-\label{subsec:subdatadeg}
-
-Raw data from \cref{subsec:rawdatacr,subsec:normcrdata} were used in the analysis.
-Together with the raw data, few other data sets were created based on the raw data: residual data, Caucasian-only raw data, and Caucasian-only residual data.
-
-Residual data was created from the raw data after it has been controlled for all clinical variables, except those related to \gls{bmi}.
-These clinical variables included age, sex, ethinicity, tumour grade, \gls{pr}, \gls{er}, \gls{her2} statuses, and \gls{ln} status.
-
-Caucasian-only data was created by taking only the Caucasian samples from the raw data, which reduced the sample size from 103 samples to 77 samples.
-Caucasian-only residual data was created using the same approach as with the residual data: the data that remained after all the non-\gls{bmi} clinical variables were controlled in the Caucasian-only raw data.
-
-\subsection{Group classification}
-\label{subsec:crdegclass}
-
-In each of the datasets from \cref{subsec:subdatadeg}, all of the samples were split into two groups: obese and non-obese groups.
-These two groups were classified based on the sample \gls{bmi}, where the obese groups were defined as samples with \gls{bmi}  \textgreater{} 30, and non-obese group were defined as samples with \gls{bmi}  \textless{} 30 (see \cref{tab:whobmiclass}).
-
-\subsection{Gene expression analysis}
-\label{subsec:crdeg}
-
-The difference in the gene expression between the obese and non-obese sample groups in each of the four data sets described in \cref{subsec:subdatadeg} were analysed by fitting a linear model to the data, using the \textit{limma} package in \gls{r}.
-From the top significantly expressed genes produced by the \texttt{topTable} function, all of the gene probes that had a p-value \textless{} 0.01 were selected.
-
-When there were more than 799 gene probes in the list of significant genes, the top 799 gene probes were selected from the list and used as the gene probe set for the metagene analysis, for that dataset.
-
-For each gene probe set, the gene probes that were common with the obesity associated genes identified by \citet{Creighton2012} were selected and used as a separate gene probe set.
-There were eight metagenes in total after this process: four gene probe sets from the datasets from \cref{subsec:subdatadeg}, and four gene probe sets that were in common with the obesity associated genes from \citet{Creighton2012} study.
-
-\subsection{Gene probe set conversion}
-\label{subsec:gpconv}
-
-\subsection{BMI metagenes in Creighton data}
-\label{subsec:bmidegmetacr}
-
-\subsection{BMI metagenes in \gls{icgc} data}
-\label{subsec:bmidegmetaicgc}
-
-The gene probe IDs were converted into gene symbols using the \textit{hgu133a.db} package in \gls{r}.
-Where there was a duplicated gene in the data, appropriate gene data was chosen to represent that particular gene by using the \texttt{collapseRows} function in the \textit{WGCNA} package in \gls{r}.
-
-\newpage
-
-The process above was repeated on the same data, with the same group (obese and non-obese) after controlling for all other clinical variables that were not related to the sample \gls{bmi} (age, sex, ethinicity, tumour grade, \gls{ln} status, \gls{pr} status, \gls{er} status, \gls{her2} status).
-the top 799 differentially expressed gene probes that had p-value \textless{} 0.01 between the two groups were identified.
-
-To completely remove the effect of ethinicity on the analysis, the analysis was repeated using only the Caucasian samples.
-This reduced the sample size from 103 samples to 77 samples.
-The data was controlled for all other clinical variables except the variables that were related to \gls{bmi} (age, sex, tumour grade, \gls{ln} status, \gls{pr} status, \gls{er} status, and \gls{her2} status).
-Again, the top 799 differentially expressed genes between the two groups with p-value \textless{} 0.01 were identified.
-
-\subsection{Comparison of \citet{Creighton2012} gene probes and identified gene probes}
-\label{subsec:cfgeneprobes}
-
-The obesity associated gene probes identified by \citet{Creighton2012} were compared with each of the three sets of gene probes identified from the differential gene expression analysis (see \cref{sec:deg} \nameref{sec:deg}).
-The presence of any of the \citet{Creighton2012} genes were checked for in each of the gene sets identified at a p-value \textless{} 0.05, as well as the top 799 gene probes at p-value \textless{} 0.01.
-
-\subsection{\gls{svd}}
-\label{subsec:svd}
-
-(need svd derivation and why I'm using it)
-
-\subsection{Correlation between obesity associated metagene and sample \gls{bmi}}
-\label{subsec:metavsbmi}
-
-In order to check whether the metagenes were actually specific to the \gls{bmi} status of the samples, the metagenes were plotted against both the \gls{bmi} statuses and the \gls{bmi} values of the samples.
-The metagenes were ranked for each sample, and a box plot and a scatter plot were plotted for sample \gls{bmi} statuses and \gls{bmi} values, respectively.
-
-\section{Applying metagene to other cancer types}
-\label{sec:metagene}
-
-(How you could transfer the metagene to other cancer data)
-
-\subsection{Transformation matrix}
-\label{subsec:transmat}
-
-(why transformation matrix is used)
-(refer back to eqiations in the  svd section)
-
-\section{Pathway Enrichment Analysis}
-\label{sec:pathenrich}

### 2016-06-14 thesis/methods/methods.tmp.tex
re-writing methods section (it was too clunky)

\chapter{Methods (draft)}
\label{ch:methods}

\section{\gls{bmi}}
\label{sec:bmi}

\subsection{\gls{bmi} Calculation}
\label{subsec:bmicalc}

\gls{bmi} was calculated from the height and weight data of the samples using the following equation:

\begin{equation}
	\label{eq:bmicalc}
	\gls{bmi} = \frac{Weight (kg)}{Height^2(m^2)}\\
\end{equation}

\subsection{\gls{bmi} Classification}
\label{subsec:bmiclassification}

Samples were classified based on the \gls{who} definition, as shown in Table~\cref{tab:whobmiclass}.
\begin{table}[hb]
	\caption{\gls{who} defined \gls{bmi} classification}
	\label{tab:whobmiclass}
	\begin{center}
		\begin{tabular}{lc}
			\bfseries {Classification} & \bfseries {\gls{bmi} Value}\\
			\hline
			\rule{0pt}{2.25ex}Underweight & \textless{} 20.0\\
			Normal weight/lean & 20.0$\sim$24.9\\
			Overweight & 25.0$\sim$29.9\\
			Obese & \textgreater{} 30.0\\
		\end{tabular}
	\end{center}
\end{table}

\section{Publicly available cancer data}
\label{sec:data}

Data related to multiple different cancer types were downloaded from free and publicly available sources such as \gls{tcga} and the \gls{icgc}.

\subsection{\gls{tcga}}
\label{subsec:tcga}

The clinical data for all available cancer types (33 types in total) were downloaded from \gls{tcga} database and were checked for the height and weight data for each sample, for each cancer type.
Any cancer type with no weight and/or height data of the samples were excluded from the project, as no \gls{bmi} information can be obtained without these data.

This gave a total of eight cancer types: \gls{blca}, \gls{cesc}, \gls{coad}, \gls{kirp}, \gls{lihc}, \gls{read}, \gls{skcm}, \gls{ucec}.
All eight cancer types were used for further analyses.

\subsection{\gls{icgc}}
\label{subsec:icgc}

\gls{rnaseq} data for the eight cancer types with sample \gls{bmi} information available were downloaded from the \gls{icgc} database.
The \gls{rnaseq} data for each of the eight cancer types were imported into \gls{r}.
%TODO: add in how I processed the ICGC data format into genes by samples.

%TODO: change the title name?
\subsection{\gls{icgc} sample exclusion criteria}
\label{subsec:sampleexclusion}

Samples from the downloaded data were removed using the following criteria.

\subsubsection{\textbf{Sample BMI availability}}
\label{subsubsec:samplebmiavailability}

Some of the samples did not have either height or weight data, or both.
These samples were removed from the analyses.

\subsubsection{\textbf{Sample ID conversion and matching}}
\label{subsubsec:sampleidconversion}

%TODO: check the actual ID stripping process
%TODO: check the WGCNA function and check if it uses median/mean/whatever for choosing the sample

\gls{icgc} sample ID in the \gls{rnaseq} data were stripped so that the ID from the \gls{rnaseq} data matched with the \gls{tcga} ID in the clinical data.
After stripping the ID, samples were checked for any overlaps.
Where there was an overlap in the sample ID in the \gls{rnaseq} data, a single sample was chosen to represent that particular ID by using the \texttt{collapseRows} function in the \textit{WGCNA} package in \gls{r}.

In each of the eight cancer types, the \gls{tcga} sample ID in the clinical data were cross-checked with the \gls{tcga} sample ID in the \gls{rnaseq} data, and vice versa.
Any sample that did not have either clinical or \gls{rnaseq} data were removed, so both clinical and \gls{rnaseq} data were available for all of the samples that were included in the analyses.
See \cref{tab:samplesize} for a summary of the number of samples included in the analyses for each cancer type.

\begin{table}[h]
	\caption{Summary of the total number of samples included in the analyses for each cancer type}
	\label{tab:samplesize}
	\begin{center}
		\begin{tabular}{cc}
			\textbf{Cancer Type}   & \textbf{Total} \\
			\hline
			\rule{0pt}{2.25ex}BLCA & 261   \\
			CESC                   & 224   \\
			COAD                   & 226   \\
			KIRP                   & 124   \\
			LIHC                   & 264   \\
			READ                   & 73    \\
			SKCM                   & 218   \\
			UCEC                   & 482   \\
		\end{tabular}
	\end{center}
\end{table}

\section{Data processing -- Creighton data}
\label{sec:dataprocessingcr}

\subsection{Raw data from the \citet{Creighton2012} study}
\label{subsec:rawdatacr}

The raw microarray gene expression data files of all *** samples from the \citet{Creighton2012} study were downloaded from the \gls{geo} database (\gls{geo} accession ID: GSE24185).
Clinical data of the samples were obtained from the supplementary table 1 from their paper.
Raw data were imported into \gls{r}/RStudio for subsequent analyses.

%TODO: add data/sample composition -- ethinicity, grade, etc -- maybe in table form

\subsection{Normalisation of raw Creighton data}
\label{subsec:normcrdata}

The raw microarray data for all 103 samples from the study were normalised using \gls{rma} method from the \textit{affy} package in \gls{r}.
The data was then standardised so that each of the genes had a mean ($\mu$) of 0 and a standard deviation ($\sigma$) of 1.
This ensured that all of the genes were on the same scale as one another, allowing for comparison.

\subsection{Obesity associated gene probes from the \citet{Creighton2012} study}
\label{subsec:crobsgene}

799 obesity associated gene probes identified by \citet{Creighton2012} were obtained from the supplementary data file 1 from their paper.
The list of gene probes were imported into \gls{r}.
The gene probe IDs were converted into gene symbols using the \textit{hgu133a.db} package in \gls{r}.
Any duplicates and any gene symbols that were not present in both the Creighton data and the \gls{icgc} data were removed, which left *** unique obesity associated genes.

\section{Validation of results -- \citet{Creighton2012}}
\label{sec:valresultscr}

The results presented by \citet{Creighton2012} were validated using the same raw data they have used in their study.
See  \cref{sec:dataprocessingcr} for details on how the raw data was accessed and processed.

\subsection{Creighton metagene generation}
\label{subsec:crmetagen}

The gene probe IDs were converted into gene symbols using the \textit{hgu133a.db} package in \gls{r}.
Where there was a duplicated gene in the data, appropriate gene data was chosen to represent that particular gene by using the \texttt{collapseRows} function in the \textit{WGCNA} package in \gls{r}.
All of the *** obesity associated genes from  \cref{subsec:crobsgene} were selected from the data.

\gls{svd} was applied to this data to obtain the decomposed matrices of the data.
%TODO: how do I explain how I got the \gls{metagene} scores?
The \gls{metagene} scores of the samples were obtained by taking the first eigenvector of the (orthogonal matrix...?).
The \gls{metagene} scores were ranked based on the number of samples in the data (103 samples in the Creighton data), and the values of the scores were between 0 and 1.

\subsection{Metagene and sample gene expression pattern}
\label{subsec:crmetage}

\Gls{metagene} scores from \cref{subsec:crmetagen} were plotted together with the gene expression of the obesity associated gene probes for each sample on a heatmap.
The samples in the heatmap were ordered based on their ranked \gls{metagene} scores.
To have a better colour scales in the heatmap, any expression values that were greater than 3 or less than -3 were assigned the value of 3 and -3, respectively.

\subsection{Metagene and sample \gls{bmi}}
\label{subsec:crmetabmi}

The same \gls{metagene} scores as in the previous section (\cref{subsec:crmetage}) were used to compare the \gls{metagene} scores with the samples' \gls{bmi} statuses and values.
Boxplot was plotted to show the association of the \gls{metagene} scores with the sample \gls{bmi} status, and scatter plot was plotted to show the association with the sample \gls{bmi} value.

%TODO: check the formal definition of these test/values
For the boxplot, the p-value was obtained from the ANOVA test.
For the scatter plot, the regression line was plotted and the adjusted $R^2$ value was calculated for the regression line.

\section{Creighton metagene in \gls{icgc} cancer data}
\label{sec:crmetaicgc}

\subsection{\gls{icgc} data processing}
\label{subsec:datprocicgc}

The obesity associated genes from \cref{subsec:crobsgene} were selected in each of the eight cancer types.
All cancer data had 1 added to the count data, so no infinite values were introduced when the data was logged.
All cancer data were logged to the base 10.

All of the data were then standardised so that each of the genes had a mean ($\mu$) of 0 and a standard deviation ($\sigma$) of 1.
Any ``NaN'' values in the data introduced from this were assigned a value of 0.

\subsection{Transformation matrix}
\label{subsec:transmatcr}

Transformation matrix was formed from the decomposed matrices that resulted from the application of \gls{svd} to the Creighton data (see \cref{subsec:crmetage}).
%TODO: add the derivation of transformation matrix??

\subsection{Transformed \gls{icgc} data and metagene}
\label{subsec:icgctranscr}

Each of the eight cancer data from \cref{subsec:datprocicgc} were transformed using the transformation matrix from \cref{subsec:transmatcr}.
Metagene scores of the samples in each cancer types were obtained by taking the first eigenvector of the cancer matrix that resulted from the transformation.
The \gls{metagene} scores were ranked based on the number of samples in the data, and the scores ranged from 0 to 1.

\subsection{Creighton metagene and sample gene expression}
\label{subsec:crmetaicgcge}

\Gls{metagene} scores from \cref{subsec:icgctranscr} were plotted together with the gene expression of the obesity associated gene probes on a heatmap for each cancer type.
The data from \cref{subsec:datprocicgc} were used for the heatmap.
The samples in the heatmap were ordered based on their ranked \gls{metagene} scores.
To have a better colour scales in the heatmap, any expression values that were greater than 3 or less than -3 were assigned the value of 3 and -3, respectively.

\subsection{Creighton metagene and sample \gls{bmi}}
\label{subsec:crmetaicgcge}

The same \gls{metagene} scores as in the previous section were used to compare the \gls{metagene} scores with the samples' \gls{bmi} statuses and values for each cancer type.
Boxplot was plotted to show the association of the \gls{metagene} scores with the sample \gls{bmi} status, and scatter plot was plotted to show the association with the sample \gls{bmi} value.

%TODO: check the formal definition of these test/values
For the boxplot, the p-value was obtained from the ANOVA test.
For the scatter plot, the regression line was plotted and the adjusted $R^2$ value was calculated for the regression line.

%TODO: Not sure if I should cut this section to "see previous section for details"
\section{Data processing -- Fuentes-Mattei data}
\label{sec:dataprocessingfm}

\subsection{Raw data from the \citet{Fuentes-Mattei2014} study}
\label{subsec:rawdatafm}

%TODO: add data/sample composition -- ethinicity, grade, etc -- maybe in table form

\subsection{Normalisation of raw Fuentes-Mattei data}
\label{subsec:normfmdata}

The raw microarray data for all *** samples from the study were normalised using \gls{rma} method from the \textit{affy} package in \gls{r}.
The data was then standardised so that each of the genes had a mean ($\mu$) of 0 and a standard deviation ($\sigma$) of 1.

\subsection{Obesity associated gene probes from the \citet{Fuentes-Mattei2014} study}
\label{subsec:fmobsgene}

Any gene symbols that were not present in the Fuentes-Mattei data, the Creighton data and the \gls{icgc} data were removed, which left *** unique obesity associated genes.

\section{Validation of results -- \citet{Fuentes-Mattei2014}}
\label{sec:valresultsfm}

The results presented by \citet{Fuentes-Mattei2014} were validated using the Creighton data, as the clinical data of the samples were not available.
See \cref{subsec:rawdatacr,subsec:normcrdata} for details on how the data was processed.

\subsection{Fuentes-Mattei transformation matrix}
\label{subsec:transmatfm}

\gls{svd} was applied to Fuentes-Mattei data to obtain the decomposed matrices of the data, and this was used to make the transformation matrix.
%TODO: add the derivation of transformation matrix??

\subsection{Fuentes-Mattei metagene generation}
\label{subsec:fmmetagen}

The gene probe IDs in the Creighton data were converted into gene symbols using the \textit{hgu133a.db} package in \gls{r}.
Where there was a duplicated gene, appropriate gene data was chosen to represent that particular gene by using the \texttt{collapseRows} function in the \textit{WGCNA} package in \gls{r}.
All of the *** obesity associated genes from  \cref{subsec:fmobsgene} were selected from the data.
The data was then standardised so that each of the genes had a mean ($\mu$) of 0 and a standard deviation ($\sigma$) of 1.

%TODO: how do I explain how I got the \gls{metagene} scores?
The \gls{metagene} scores of the samples were obtained by taking the first eigenvector of the transformed Creighton data.
The \gls{metagene} scores were ranked based on the number of samples in the data (*** samples in the Creighton data), and the scores ranged from 0 to 1.

\subsection{Fuentes-Mattei metagene and gene expression in the Creighton data}
\label{subsec:fmmetage}

These \gls{metagene} scores were plotted together with the gene expression of the obesity associated gene probes for each sample on a heatmap.
The samples in the heatmap were ordered based on their ranked \gls{metagene} scores.
To have a better colour scales in the heatmap, any expression values that were greater than 3 or less than -3 were assigned the value of 3 and -3, respectively.

\subsection{Fuentes-Mattei metagene and sample \gls{bmi} in the Creighton data}
\label{subsec:fmmetabmi}

The same \gls{metagene} scores from the previous section were used to compare the \gls{metagene} scores with the samples' \gls{bmi} statuses and values.
Box and scatter plots, and any statistical values associated with these plots were calculated as in \cref{subsec:crmetaicgcge}.

\section{Fuentes-Mattei metagene in \gls{icgc} cancer data}
\label{sec:fmmetaicgc}

\subsection{\gls{icgc} data processing}
\label{subsec:datprocicgcfm}

All of the \gls{icgc} data were processed in a similar fashion as in \cref{subsec:datprocicgc}, but the obesity associated genes from \cref{subsec:fmobsgene} were used instead.

\subsection{Transformation matrix}
\label{subsec:icgctransmatfm}

The same transformation matrix from \cref{subsec:transmatfm} was used.

\subsection{Transformed \gls{icgc} data and metagene}
\label{subsec:icgctransfm}

Each of the eight cancer data from \cref{subsec:datprocicgcfm} were transformed using the transformation matrix from \cref{subsec:transmatfm}.
Metagene scores were ranked as in \cref{subsec:icgctranscr}.

\subsection{Fuentes-Mattei metagene and sample gene expression}
\label{subsec:fmmetaicgcge}

\Gls{metagene} scores from \cref{subsec:icgctransfm} were plotted together with the gene expression of the obesity associated gene probes for each sample on a heatmap for each cancer type.
The data from \cref{subsec:datprocicgcfm} were used for the heatmap.
The samples in the heatmap were ordered based on their ranked \gls{metagene} scores.
To have a better colour scales in the heatmap, any expression values that were greater than 3 or less than -3 were assigned the value of 3 and -3, respectively.

\subsection{Creighton metagene and sample \gls{bmi}}
\label{subsec:fmmetaicgcge}

The same \gls{metagene} scores as in the previous section were used to compare the \gls{metagene} scores with the samples' \gls{bmi} statuses and values for each cancer type.
Box and scatter plots, and any statistical values associated with these plots were calculated as in \cref{subsec:crmetaicgcge}.

\section{Differential gene expression analysis}
\label{sec:deg}

\subsection{Data processing and subsetting}
\label{subsec:subdatadeg}

Raw data from \cref{subsec:rawdatacr,subsec:normcrdata} were used in the analysis.
Together with the raw data, few other data sets were created based on the raw data: residual data, Caucasian-only raw data, and Caucasian-only residual data.

Residual data was created from the raw data after it has been controlled for all clinical variables, except those related to \gls{bmi}.
These clinical variables included age, sex, ethinicity, tumour grade, \gls{pr}, \gls{er}, \gls{her2} statuses, and \gls{ln} status.

Caucasian-only data was created by taking only the Caucasian samples from the raw data, which reduced the sample size from 103 samples to 77 samples.
Caucasian-only residual data was created using the same approach as with the residual data: the data that remained after all the non-\gls{bmi} clinical variables were controlled in the Caucasian-only raw data.

\subsection{Group classification}
\label{subsec:crdegclass}

In each of the datasets from \cref{subsec:subdatadeg}, all of the samples were split into two groups: obese and non-obese groups.
These two groups were classified based on the sample \gls{bmi}, where the obese groups were defined as samples with \gls{bmi}  \textgreater{} 30, and non-obese group were defined as samples with \gls{bmi}  \textless{} 30 (see \cref{tab:whobmiclass}).

\subsection{Gene expression analysis}
\label{subsec:crdeg}

The difference in the gene expression between the obese and non-obese sample groups in each of the four data sets described in \cref{subsec:subdatadeg} were analysed by fitting a linear model to the data, using the \textit{limma} package in \gls{r}.
From the list of significantly expressed (genes produced by the \texttt{topTable} function), all of the gene probes that had a p-value \textless{} 0.01 were selected.

When there was more than 799 gene probes in the list of significant genes, the top 799 gene probes were selected from the list.
This list of 799 gene probes represented the gene probes that were associated with obesity  for that dataset.

\subsection{Common gene probes from the \citet{Creighton2012} study}
\label{subsec:crol}

For each gene probe set, the gene probes that were common with the obesity associated genes identified by \citet{Creighton2012} were selected and used as a separate gene probe set.
There were eight gene probe sets in total after this process: four gene probe sets from the datasets from \cref{subsec:subdatadeg}, and four gene probe sets that were in common with the obesity associated genes from the \citet{Creighton2012} study.

%TODO: add a table summarizing how many genes were in each gene probe set

\subsection{Gene probe set conversion and selection}
\label{subsec:gpconv}

For each of the eight gene probe sets identified in \cref{subsec:crdeg,subsec:crol}, the gene probe IDs were converted into gene symbols using the \textit{hgu133a.db} package in \gls{r}.
Any duplicates and any gene symbols that were not present in both the Creighton data and the ICGC data were removed from the gene probe set.

\section{BMI metagenes in the Creighton data}
\label{sec:bmidegmetacr}

\subsection{BMI metagene generation}
\label{subsec:bmimetagen}

The gene probe IDs in the Creighton data were converted into gene symbols using the \textit{hgu133a.db} package in \gls{r}.
Where there was a duplicated gene, appropriate gene data was chosen to represent that particular gene by using the \texttt{collapseRows} function in the \textit{WGCNA} package in \gls{r}.
For each gene probe sets identified in \cref{subsec:crdeg,subsec:crol}, the gene probes were selected from the Creighton data.
The data was then standardised so that each of the genes had a mean ($\mu$) of 0 and a standard deviation ($\sigma$) of 1.

SVD was applied to the standardised data, and the first eigenvector of the decomposed matrix was used as the sample metagene score for that gene probe set.
The \gls{metagene} scores were ranked based on the number of samples in the data (103 samples in the Creighton data), and the scores ranged from 0 to 1.

\subsection{BMI metagenes and sample gene expression}
\label{subsec:bmimetavscr}

For each \gls{metagene} scores in \cref{subsec:bmimetagen} were plotted together with the gene expression of the associated gene probes for each sample on a heatmap.
The samples in the heatmap were ordered based on their ranked \gls{metagene} scores.
To have a better colour scales in the heatmap, any expression values that were greater than 3 or less than -3 were assigned the value of 3 and -3, respectively.

\subsection{BMI metagenes and sample \gls{bmi}}
\label{subsec:bmidegmetacr}

The same \gls{metagene} scores from the previous section were used to compare the \gls{metagene} scores with the samples' \gls{bmi} statuses and values.
Box and scatter plots, and any statistical values associated with these plots were calculated as in \cref{subsec:crmetaicgcge}.

\section{BMI metagenes in the \gls{icgc} data}
\label{sec:bmidegmetaicgc}

\subsection{\gls{icgc} data processing}

All of the \gls{icgc} data were processed in a similar fashion as in \cref{subsec:datprocicgc}, but each of the eight BMI metagene gene probes from \cref{subsec:bmimetagen} were used instead.

\subsection{BMI metagene transformation matrix}
\label{subsec:bmimetagenicgc}

Transformation matrix for each BMI metagene gene probe sets were created using the decomposed matrices of the Creighton data that resulted from the application of SVD in \ref{subsec:bmimetagen}.

\subsection{BMI metagenes and sample gene expression in the \gls{icgc} data}
\label{subsec:gpconv}

For each \gls{metagene} scores in \cref{subsec:bmimetagen} were plotted together with the gene expression of the associated gene probes for each sample on a heatmap for each cancer type.
The samples in the heatmap were ordered based on their ranked \gls{metagene} scores.
To have a better colour scales in the heatmap, any expression values that were greater than 3 or less than -3 were assigned the value of 3 and -3, respectively.

\subsection{BMI metagenes in \gls{icgc} data}
\label{subsec:bmidegmetaicgc}

The same \gls{metagene} scores from the previous section were used to compare the \gls{metagene} scores with the samples' \gls{bmi} statuses and values.
Box and scatter plots, and any statistical values associated with these plots were calculated as in \cref{subsec:crmetaicgcge}.

\newpage

%TODO: need svd derivation and why I'm using it

%TODO: How you could transfer the metagene to other cancer data

%TODO: why transformation matrix is used
%TODO: refer back to eqiations in the  svd section

\section{Gene expression analysis in \gls{icgc} data}
\label{sec:degicgc}

\section{Pathway enrichment analysis in \gls{icgc} data}
\label{sec:pathenrich}

### 2016-06-14 thesis/intro/intro.tex
tmp commit, added .swp in .gitignore

\chapter{Introduction (draft)}
\label{ch:intro}

% TODO: also have a look at the masters talk write up from 2015

% This is taken from the committee report (I think it's a good intro to the intro)
% TODO: past tense...??
It is now commonly understood that cancers are caused by dysregulation of various mechanisms or pathways (known as the ``Hallmarks of Cancer'') that allow tumour cells to prolilferate, survive and migrate \citet{Hanahan2011}.
There have been many reports on the association between obesity and worsened cancer prognosis, suggesting that there may be distinct changes that underlie tumorigenesis and/or cancer progression that are specific to patients with higher \gls{bmi}.
The exact mechanism in which obesity accelerates cancer progression is unknown, but obesity is thought to disrupt the levels of various hormones (such as insulin, leptin, adeponectin, and other adipokines) that eventually activate tumorigenic pathways (reference the overview on obesity and tumour resistance).

% TODO: in the aims section, or here?
This research aims to determine whether gene expression signatures exist that are specific to obesity across multiple cancer types, and to investigate whether there aare any common pathways being dysregulated in cancers based on these genetic signatures.
Better understanding of the pathways being dysregulated in cancer cells in obese patients may lead to improved clinical decisions, and could thus contribute towards personalised treatment in the future.

\section{Obesity}
\label{sec:obesity}

General intro to obesity.
Implications of obesity -- disease associated with it and/or cancer.

\section{Cancer}
\label{sec:cancer}

What is cancer.

\subsection{Hallmarks of Cancer}
\label{subsec:cancerhallmarks}

What is it caused by (10 hallmarks of cancer).

\subsection{Relationship between obesity}
\label{subsec:obsbackground}

Why is it important in obesity context.

\section{Obesity associated genetic signatures}
\label{sec:obsgene}

Literature review, focussing on \citet{Creighton2012} and \citet{Fuentes-Mattei2014}.

\section{Aim of the project}
\label{sec:aim}

### 2016-06-14 thesis/methods/methods.tex
tmp commit, added .swp in .gitignore

@@
 
-\paragraph{\textit{Raw data normalisation}}
+\subsubsection{\textbf{Raw data normalisation}}
 
@@
 
-\paragraph{\textit{ Raw data standardisation }}
+\subsubsection{\textbf{Raw data standardisation}}
 
-The data was standardised so that each of the genes had a mean ($\mu$) of 0 and a standard deviation ($\sigma$) of 1.
+The data was standardised so that each gene had a mean ($\mu$) of 0 and a standard deviation ($\sigma$) of 1.
 This ensured that all of the genes were on the same scale as one another, allowing for comparison across all the genes.
 
-\paragraph{\textit{ Raw data standardisation }}
+\subsubsection{\textbf{Gene probe ID conversion}}
+
+Where appropriate, the gene probe IDs in the raw data were converted into their corresponding gene symbols using the \textit{hgu133a.db} package in \gls{r}.
+Where there was an overlap in the gene symbols, a single data was chosen to represent that particular gene symbol by using the \texttt{collapseRows} function in the \textit{WGCNA} package in \gls{r}.
+Likewise, the obesity associated gene probes were also converted into gene symbols where required.
+
+\subsubsection{\textbf{Sample \gls{bmi} calculation}}
+
+\gls{bmi} of the samples were calculated from the clinical data using \cref{eq:bmicalc}.
 
@@
 
-\paragraph{\textit{ Raw data standardisation }}
+\subsubsection{\textbf{Raw data standardisation}}
 
-The data was standardised so that each of the genes had a mean ($\mu$) of 0 and a standard deviation ($\sigma$) of 1.
+The data was standardised so that each gene had a mean ($\mu$) of 0 and a standard deviation ($\sigma$) of 1.
 This ensured that all of the genes were on the same scale as one another, allowing for comparison across all the genes.
+
+\subsubsection{\textbf{Gene probe ID conversion}}
+
+Where appropriate, the gene probe IDs in the raw data were converted into their corresponding gene symbols using the \textit{hgu133a.db} package in \gls{r}.
+Where there was an overlap in the gene symbols, a single data was chosen to represent that particular gene symbol by using the \texttt{collapseRows} function in the \textit{WGCNA} package in \gls{r}.
 
@@
 
+\subsubsection{\textbf{Raw data normalisation}}
+
+The raw \gls{rnaseq} data for each cancer types from \cref{subsec:tcga} had 1 added, and then logged to the power of 10.
+
+\subsubsection{\textbf{Raw data standardisation}}
+
+The normalised data was standardised so that each gene had a mean ($\mu$) of 0 and a standard deviation ($\sigma$) of 1.
+This ensured that all of the genes were on the same scale as one another, allowing for comparison across all the genes.
+
 \subsection{\gls{icgc}}
@@
 
-\subsubsection{\textbf{Reformatting the raw data}}
-\label{subsec:icgcreformat}
+\subsubsection{\textbf{Raw data format conversion}}
+%TODO: add in how I processed the ICGC data format into genes by samples.
+The raw \gls{rnaseq} data for each cancer types were formatted so that the rows corresponded to the genes and the columns to the samples.
+% TODO: find out what functions I used to reformat the icgc data
+This was done using the \textit{dplyr} package in \gls{r}.
 
-%TODO: add in how I processed the ICGC data format into genes by samples.
-TODO: add in how I processed the ICGC data format into genes by samples.
+\subsubsection{\textbf{Sample ID conversion}}
+%TODO: add details on how I stripped the ICGC ID into TCGA ID
+%TODO: check the actual ID stripping process
+%TODO: check the WGCNA function and check if it uses median/mean/whatever for choosing the sample
+\gls{icgc} sample ID in the \gls{rnaseq} data were stripped so that the ID matched with the \gls{tcga} ID in the clinical data.
+After the IDs were stripped, samples were checked for any overlaps.
+Where there was an overlap in the sample ID in the \gls{rnaseq} data, a single sample was chosen to represent that particular ID by using the \texttt{collapseRows} function in the \textit{WGCNA} package in \gls{r}.
 
-%TODO: change the title name?
 \subsubsection{\textbf{\gls{icgc} sample exclusion criteria}}
@@
 
-\paragraph{\textit{Sample BMI availability}}
+% \subsubsection{\textbf{Sample BMI availability}}
+
 Some of the samples did not have either height or weight data, or both.
 These samples were removed from the analyses.
-
-\paragraph{\textit{Sample ID conversion and matching}}
-
-%TODO: check the actual ID stripping process
-%TODO: check the WGCNA function and check if it uses median/mean/whatever for choosing the sample
-
-\gls{icgc} sample ID in the \gls{rnaseq} data were stripped so that the ID from the \gls{rnaseq} data matched with the \gls{tcga} ID in the clinical data.
-After stripping the ID, samples were checked for any overlaps.
-Where there was an overlap in the sample ID in the \gls{rnaseq} data, a single sample was chosen to represent that particular ID by using the \texttt{collapseRows} function in the \textit{WGCNA} package in \gls{r}.
-
 In each of the eight cancer types, the \gls{tcga} sample ID in the clinical data were cross-checked with the \gls{tcga} sample ID in the \gls{rnaseq} data, and vice versa.
-Any sample that did not have either clinical or \gls{rnaseq} data were removed, so both clinical and \gls{rnaseq} data were available for all of the samples that were included in the analyses.
+Any sample that did not have either clinical or \gls{rnaseq} data were removed.
+This ensured that both clinical and \gls{rnaseq} data were available for all of the samples that were included in the analyses.
 See \cref{tab:samplesize} for a summary of the number of samples included in the analyses for each cancer type.
 
+%TODO: May need more details on the sample (bmi status, etc.)
 \begin{table}[h]
@@
 
-\subsection{Normalisation of raw Creighton data}
-\label{subsec:normcrdata}
+\section{Plot creation}
+\label{sec:plot_creation}
 
-\subsection{Obesity associated gene probes from the \citet{Creighton2012} study}
-\label{subsec:crobsgene}
+\subsection{Heatmaps}
+\label{sub:heatmaps}
 
-The gene probe IDs were converted into gene symbols using the \textit{hgu133a.db} package in \gls{r}.
-Any duplicates and any gene symbols that were not present in both the Creighton data and the \gls{icgc} data were removed, which left *** unique obesity associated genes.
+\subsection{Box plots}
+\label{sub:box_plots}
+
+\subsection{Scatter plots}
+\label{sub:scatter_plots}
+
+\section{Gene expression analysis}
+\label{sec:gene_expression_analysis}
+
+\section{Metagene analysis}
+\label{sec:metagene_analysis}
+
+\section{Pathway enrichment analysis}
+\label{sec:pathway_enrichment_analysis}

### 2016-06-15 thesis/intro/intro.tex
added some stuff in Hallmarks of Cancer

@@
 % TODO: past tense...??
-It is now commonly understood that cancers are caused by dysregulation of various mechanisms or pathways (known as the ``Hallmarks of Cancer'') that allow tumour cells to prolilferate, survive and migrate \citet{Hanahan2011}.
+It is now commonly understood that cancers are caused by dysregulation of various mechanisms or pathways that allow tumour cells to prolilferate, survive and migrate.
 There have been many reports on the association between obesity and worsened cancer prognosis, suggesting that there may be distinct changes that underlie tumorigenesis and/or cancer progression that are specific to patients with higher \gls{bmi}.
-The exact mechanism in which obesity accelerates cancer progression is unknown, but obesity is thought to disrupt the levels of various hormones (such as insulin, leptin, adeponectin, and other adipokines) that eventually activate tumorigenic pathways (reference the overview on obesity and tumour resistance).
 
 % TODO: in the aims section, or here?
-This research aims to determine whether gene expression signatures exist that are specific to obesity across multiple cancer types, and to investigate whether there aare any common pathways being dysregulated in cancers based on these genetic signatures.
+In brief, this research aims to determine whether gene expression signatures exist that are specific to obesity across multiple cancer types, and to investigate whether there aare any common pathways being dysregulated in cancers based on these genetic signatures.
 Better understanding of the pathways being dysregulated in cancer cells in obese patients may lead to improved clinical decisions, and could thus contribute towards personalised treatment in the future.
@@
 
-What is it caused by (10 hallmarks of cancer).
+It is now commonly understood that cancers are caused by dysregulation of various mechanisms or pathways that allow tumour cells to prolilferate, survive and migrate.
+Many of these dysregulated pathways frequently occur in cancer cells, and these characteristic traits, commonly known as the ``Hallmarks of Cancer'', were first described by \citet{Hanahan2000}.
+
+These hallmarks were initially presented in 2000 to provide a common framework for scientists to better understand cancer biology, and also to direct the focus of future research.
+After a decade of advances in cancer research, the ``Hallmarks of Cancer'' were revisited by Hanahan and Weinberg in 2011, and four more hallmarks were added on top of the six presented in the original paper in 2000 \citep{Hanahan2011}.
+
+Together, these ``Hallmarks of Cancer'' (fig 1) have provided a foundation of the general biological knowledge of cancer and its characteristics that allowed the researchers to better understand the underlying cancer biology.
+
+Although much reseach has been conducted on the ``Hallmarks of Cancer'', the molecular processes underlying these hallmarks have still not been fully explained \citep{Hanahan2011}.
+This is due to the
 
@@
 Why is it important in obesity context.
+
+The exact mechanism in which obesity accelerates cancer progression is unknown, but obesity is thought to disrupt the levels of various hormones (such as insulin, leptin, adeponectin, and other adipokines) that eventually activate tumorigenic pathways (reference the overview on obesity and tumour resistance).
 
@@
 \label{sec:aim}
+
+\newpage
+
+CONSISTENCY OF VERB TENSE helps ensure smooth expression in your writing.
+The practice of the discipline for which you write typically determines which verb tenses to use in various parts of a scientific document.
+In general, however, the following guidelines may help you know when to use past and present tense.
+If you have questions about tense or other writing concerns specific to your discipline, check with your adviser.\\
+\\
+USE PAST TENSE. . .\\
+\\
+To describe your methodology and report your results.\\
+\\
+At the time you are writing your report, thesis, dissertation or article, you have already completed your study, so you should use past tense in your methodology section to record what you did, and in your results section to report what you found.\\
+\\
+       We hypothesized that adults would remember more items than children.\\
+\\
+       We extracted tannins from the leaves by bringing them to a boil in 50% methanol.\\
+\\
+       In experiment 2, response varied.\\
+\\
+When referring to the work of previous researchers.\\
+\\
+When citing previous research in your article, use past tense.
+Whatever a previous researcher said, did or wrote happened at some specific, definite time in the past and is not still being done.
+Results that were relevant only in the past or to a particular study and have not yet been generally accepted as fact also should be expressed in past tense:\\
+\\
+Smith (2008) reported that adult respondents in his study remembered 30 percent more than children. (Smith's study was completed in the past and his finding was specific to that particular study.)\\
+\\
+Previous research showed that children confuse the source of their memories more often than adults (Lindsey et al., 1991).
+(The research was conducted in the past, but the finding is now a widely accepted fact.)\\
+\\
+To describe a fact, law or finding that is no longer considered valid and relevant.\\
+\\
+Nineteenth-century physicians held that women got migraines because they were "the weaker sex," but current research shows that the causes of migraine are unrelated to gender.
+(Note the shift here from past tense [discredited belief] to present [current belief].)\\
+\\
+USE PRESENT TENSE. . .\\
+\\
+To express findings that continue to be true.\\
+\\
+Use present tense to express general truths or facts or conclusions supported by research results that are unlikely to change - in other words, something that is believed to be always true:\\
+\\
+       Genetic information is encoded in the sequence of nucleotides on DNA.\\
+\\
+       Galileo asserted that the earth revolves the sun.(The asserting took place in the past, but the\\
+       earth is still revolving around the sun. Note also that no source citation is needed here since it is a\\
+       widely known and well-accepted fact that Galileo made this assertion.)\\
+\\
+       Sexual dimorphism in body size is common among butterflies  (Singer1982).(Note how this\\
+       statement differs from one in which you refer to the researcher's work in the sentence: "Singer\\
+       (1982) stated that sexual dimorphism in body size is common among butterflies." Here you use past\\
+       tense to indicate what Singer reported, but present tense to indicate a research result that is\\
+       unlikely to change.)\\
+\\
+       We chose Vietnam for this study because it has a long coastline. (Use past tense to indicate what\\
+       you did [chose Vietnam], but present tense to indicate you assume that the length of Vietnam's\\
+       coastline is unlikely to change.)\\
+\\
+       We used cornmeal to feed the fingerlings because it provides high nutritional content at a\\
+       relatively low cost. (Past tense reflects what you did [used cornmeal], but present tense indicates\\
+       that neither the nutritional content nor the cost of corn meal is likely to change.)\\
+\\
+To refer to the article, thesis or dissertation itself.\\
+\\
+Use the present tense in reference to the thesis or dissertation itself and what it contains, shows, etc. For example:\\
+\\
+       Table 3 shows that the main cause of weight increase was nutritional value of the feed. (Table 3\\
+       will always show this; it is now a fact that is unlikely to change, and will be true whenever anyone\\
+       reads this sentence, so use present tense.)\\
+\\
+To discuss your findings and present your conclusions. Also use present tense to discuss your results and their implications.\\
+\\
+       Weight increased as the nutritional value of feed increased. These results suggest that feeds\\
+       higher in nutritional value contribute to greater weight gain in livestock. (Use past tense to\\
+       indicate what you found [weight increased], but use present tense to suggest what the result\\
+       implies.)\\

### 2016-06-17 thesis/intro/intro.tex
tmp commit

@@
 
+\subsection{Prevalence of obesity}
+\label{sub:prevalence_of_obesity}
+
+\subsection{Obesity and cancer risk}
+\label{sub:obesity_and_cancer_risk}
+
 \section{Cancer}
@@
 
+% TODO: edit the starting sentence so it's not a copy from the beginning
 It is now commonly understood that cancers are caused by dysregulation of various mechanisms or pathways that allow tumour cells to prolilferate, survive and migrate.
@@
 Although much reseach has been conducted on the ``Hallmarks of Cancer'', the molecular processes underlying these hallmarks have still not been fully explained \citep{Hanahan2011}.
-This is due to the
+This is due to the fact that cancer cells develop progressively under selective pressures, and are also affected by many environmental factors \citep{Gatza2011}, which makes it difficult to identify causal factors of cancer development.
+Furthermore, these external factors affect how the cancer cells respond to the environment, and may also cause cancer progression and/or drug resistance.
+
+\subsection{Tumour heterogeneity}
+\label{sub:tumour_heterogeneity}
 
@@
 
-\section{Obesity associated genetic signatures}
-\label{sec:obsgene}
+\section{Genetic signatures}
+\label{sec:genetic_signatures}
 
-Literature review, focussing on \citet{Creighton2012} and \citet{Fuentes-Mattei2014}.
+Due to advances in sequencing technology over the last few years, many researchers have started to focus on bioinformatics based analyses of cancer cells to better understand the biology of cancer.
+Advantages of this strategy are that very specific gene expression and/or mutational signatures for certain cancer subtypes can be identified from the analyses, which the researchers and clinicians can study in much greater depth and details.
+
+% TODO: have a leading sentence that introduces the gene signatures
+
+\subsection{Obesity associated genetic signatures}
+\label{subsec:obesity_associated_genetic_signatures}
+
+\subsubsection{ \textbf{\citet{Creighton2012} study}}
+\label{ssub:creighton_study}
+
+\subsubsection{ \textbf{\citet{Fuentes-Mattei2014} study}}
+\label{ssub:fuentes_mattei_study}
+
+\subsection{Pathway associated genetic signatures}
+\label{sub:pathway_associated_genetic_signatures}
+
+Talk about \citet{Gatza2010a}
 
@@
 If you have questions about tense or other writing concerns specific to your discipline, check with your adviser.\\
-\\
+
 USE PAST TENSE. . .\\
-\\
+
 To describe your methodology and report your results.\\
-\\
+
 At the time you are writing your report, thesis, dissertation or article, you have already completed your study, so you should use past tense in your methodology section to record what you did, and in your results section to report what you found.\\
-\\
+
        We hypothesized that adults would remember more items than children.\\
-\\
+
        We extracted tannins from the leaves by bringing them to a boil in 50% methanol.\\
-\\
+
        In experiment 2, response varied.\\
-\\
+
 When referring to the work of previous researchers.\\
-\\
+
 When citing previous research in your article, use past tense.
@@
 Results that were relevant only in the past or to a particular study and have not yet been generally accepted as fact also should be expressed in past tense:\\
-\\
+
 Smith (2008) reported that adult respondents in his study remembered 30 percent more than children. (Smith's study was completed in the past and his finding was specific to that particular study.)\\
-\\
+
 Previous research showed that children confuse the source of their memories more often than adults (Lindsey et al., 1991).
 (The research was conducted in the past, but the finding is now a widely accepted fact.)\\
-\\
+
 To describe a fact, law or finding that is no longer considered valid and relevant.\\
-\\
+
 Nineteenth-century physicians held that women got migraines because they were "the weaker sex," but current research shows that the causes of migraine are unrelated to gender.
 (Note the shift here from past tense [discredited belief] to present [current belief].)\\
-\\
+
 USE PRESENT TENSE. . .\\
-\\
+
 To express findings that continue to be true.\\
-\\
+
 Use present tense to express general truths or facts or conclusions supported by research results that are unlikely to change - in other words, something that is believed to be always true:\\
-\\
+
        Genetic information is encoded in the sequence of nucleotides on DNA.\\
-\\
+
        Galileo asserted that the earth revolves the sun.(The asserting took place in the past, but the\\
@@
        widely known and well-accepted fact that Galileo made this assertion.)\\
-\\
+
        Sexual dimorphism in body size is common among butterflies  (Singer1982).(Note how this\\
@@
        unlikely to change.)\\
-\\
+
        We chose Vietnam for this study because it has a long coastline. (Use past tense to indicate what\\
@@
        coastline is unlikely to change.)\\
-\\
+
        We used cornmeal to feed the fingerlings because it provides high nutritional content at a\\
@@
        that neither the nutritional content nor the cost of corn meal is likely to change.)\\
-\\
+
 To refer to the article, thesis or dissertation itself.\\
-\\
+
 Use the present tense in reference to the thesis or dissertation itself and what it contains, shows, etc. For example:\\
-\\
+
        Table 3 shows that the main cause of weight increase was nutritional value of the feed. (Table 3\\
@@
        reads this sentence, so use present tense.)\\
-\\
+
 To discuss your findings and present your conclusions. Also use present tense to discuss your results and their implications.\\
-\\
+
        Weight increased as the nutritional value of feed increased. These results suggest that feeds\\

### 2016-06-27 thesis/intro/intro.tex
removed some temporary tips in intro

--- 2016-06-14 thesis/intro/intro.tex
@@
 Implications of obesity -- disease associated with it and/or cancer.
+
+\subsection{Prevalence of obesity}
+\label{sub:prevalence_of_obesity}
+
+\subsection{Obesity and cancer risk}
+\label{sub:obesity_and_cancer_risk}
 
@@
 
-What is it caused by (10 hallmarks of cancer).
+% TODO: edit the starting sentence so it's not a copy from the beginning
+It is now commonly understood that cancers are caused by dysregulation of various mechanisms or pathways that allow tumour cells to prolilferate, survive and migrate.
+Many of these dysregulated pathways frequently occur in cancer cells, and these characteristic traits, commonly known as the ``Hallmarks of Cancer'', were first described by \citet{Hanahan2000}.
+
+These hallmarks were initially presented in 2000 to provide a common framework for scientists to better understand cancer biology, and also to direct the focus of future research.
+After a decade of advances in cancer research, the ``Hallmarks of Cancer'' were revisited by Hanahan and Weinberg in 2011, and four more hallmarks were added on top of the six presented in the original paper in 2000 \citep{Hanahan2011}.
+
+Together, these ``Hallmarks of Cancer'' (fig 1) have provided a foundation of the general biological knowledge of cancer and its characteristics that allowed the researchers to better understand the underlying cancer biology.
+
+Although much reseach has been conducted on the ``Hallmarks of Cancer'', the molecular processes underlying these hallmarks have still not been fully explained \citep{Hanahan2011}.
+This is due to the fact that cancer cells develop progressively under selective pressures, and are also affected by many environmental factors \citep{Gatza2011}, which makes it difficult to identify causal factors of cancer development.
+Furthermore, these external factors affect how the cancer cells respond to the environment, and may also cause cancer progression and/or drug resistance.
+
+\subsection{Tumour heterogeneity}
+\label{sub:tumour_heterogeneity}
 
@@
 
-\section{Obesity associated genetic signatures}
-\label{sec:obsgene}
+The exact mechanism in which obesity accelerates cancer progression is unknown, but obesity is thought to disrupt the levels of various hormones (such as insulin, leptin, adeponectin, and other adipokines) that eventually activate tumorigenic pathways (reference the overview on obesity and tumour resistance).
 
-Literature review, focussing on \citet{Creighton2012} and \citet{Fuentes-Mattei2014}.
+\section{Genetic signatures}
+\label{sec:genetic_signatures}
+
+Due to advances in sequencing technology over the last few years, many researchers have started to focus on bioinformatics based analyses of cancer cells to better understand the biology of cancer.
+Advantages of this strategy are that very specific gene expression and/or mutational signatures for certain cancer subtypes can be identified from the analyses, which the researchers and clinicians can study in much greater depth and details.
+
+% TODO: have a leading sentence that introduces the gene signatures
+
+\subsection{Obesity associated genetic signatures}
+\label{subsec:obesity_associated_genetic_signatures}
+
+\subsubsection{ \textbf{\citet{Creighton2012} study}}
+\label{ssub:creighton_study}
+
+\subsubsection{ \textbf{\citet{Fuentes-Mattei2014} study}}
+\label{ssub:fuentes_mattei_study}
+
+\subsection{Pathway associated genetic signatures}
+\label{sub:pathway_associated_genetic_signatures}
+
+Talk about \citet{Gatza2010a}
 

### 2016-09-17 thesis/intro/intro.tex
tmp commit

@@
 % TODO: also have a look at the masters talk write up from 2015
-
-% This is taken from the committee report (I think it's a good intro to the intro)
-% TODO: past tense...??
-It is now commonly understood that cancers are caused by dysregulation of various mechanisms or pathways that allow tumour cells to prolilferate, survive and migrate.
-There have been many reports on the association between obesity and worsened cancer prognosis, suggesting that there may be distinct changes that underlie tumorigenesis and/or cancer progression that are specific to patients with higher \gls{bmi}.
-
-% TODO: in the aims section, or here?
-In brief, this research aims to determine whether gene expression signatures exist that are specific to obesity across multiple cancer types, and to investigate whether there aare any common pathways being dysregulated in cancers based on these genetic signatures.
-Better understanding of the pathways being dysregulated in cancer cells in obese patients may lead to improved clinical decisions, and could thus contribute towards personalised treatment in the future.
 
@@
 
-\subsubsection{ \textbf{\citet{Creighton2012} study}}
+\subsubsection{\textbf{\citet{Creighton2012} study}}
 \label{ssub:creighton_study}
 
-\subsubsection{ \textbf{\citet{Fuentes-Mattei2014} study}}
+\subsubsection{\textbf{\citet{Fuentes-Mattei2014} study}}
 \label{ssub:fuentes_mattei_study}

### 2016-09-17 thesis/methods/methods.tex
tmp commit

@@
 
-Samples were classified based on the \gls{who} definition, as shown in Table~\cref{tab:whobmiclass}.
+Samples were classified based on the \gls{who} definition, as shown in \cref{tab:whobmiclass}.
 \begin{table}[hb]
@@
 
-The original raw data from the \citet{Creighton2012} and \citet{Fuentes-Mattei2014} studies were downloaded from the \gls{geo} website.
-Data related to multiple different cancer types were downloaded from free and publicly available sources such as \gls{tcga} and the \gls{icgc}.
+% TODO: add Cris Print's paper
+The raw microarray data from \citet{Creighton2012}, \citet{Fuentes-Mattei2014} and \citet{Gatza2010a}  studies were downloaded from the \gls{geo} website.
+\gls{rnaseq} and clinical data for multiple different cancer types were downloaded from free and publicly available sources such as \gls{icgc} and \gls{tcga}, respectively.
 
-\subsection{Raw data from the \citet{Creighton2012} study}
+\subsection{Raw data from \citet{Creighton2012} study}
 \label{subsec:crrawdat}
 
-The raw microarray gene expression data files of all *** samples from the \citet{Creighton2012} study were downloaded from the \gls{geo} database (\gls{geo} accession ID: GSE24185).
-Clinical data of the samples were obtained from the supplementary table 1 from their paper.
 %TODO: add data/sample composition -- ethinicity, grade, etc -- maybe in table form
-799 obesity associated gene probes identified in the \citet{Creighton2012} study were obtained from the supplementary data file 1 from their paper.
+
+The raw microarray gene expression data files of all 103 samples from the \citet{Creighton2012} study were downloaded from the \gls{geo} database (\gls{geo} accession ID: GSE24185).
+Clinical data of the samples were obtained from the supplementary table 1 from \citet{Creighton2012} paper.
+799 obesity associated gene probes identified in the \citet{Creighton2012} study were obtained from the supplementary data file 1 from \citet{Creighton2012} paper.
 The raw expression data, clinical data, and the list of gene probes were imported into \gls{r}/RStudio.
 
-\subsection{Raw data from the \citet{Fuentes-Mattei2014} study}
+\subsection{Raw data from \citet{Fuentes-Mattei2014} study}
 \label{subsec:fmrawdat}
+
+%TODO: add data/sample composition -- ethinicity, grade, etc -- maybe in table form
+%TODO: find out where  the genes were taken from
 
@@
 130 obesity associated gene probes identified by \citet{Fuentes-Mattei2014} were downloaded from (supplementary data??).
+The raw expression data and the list of gene probes were imported into \gls{r}/RStudio.
+
+\subsection{Raw data from Print et al study}
+\label{subsec:printrawdat}
+
+%TODO: add data/sample composition -- ethinicity, grade, etc -- maybe in table form
+%TODO: find out where  the genes were taken from
+%TODO: Change Fuentes-Mattei citation to Cris Print's citation
+
+The raw microarray gene expression data files of all *** samples from the  \citet{Fuentes-Mattei2014} study were downloaded from the \gls{geo} database (include website name/GEO ID here).
+Clinical data for the samples in this study was not available.
+130 obesity associated gene probes identified by \citet{Fuentes-Mattei2014} were downloaded from (supplementary data??).
+The raw expression data and the list of gene probes were imported into \gls{r}/RStudio.
+
+\subsection{Raw data from \citet{Gatza2010a} study}
+\label{sub:gatzarawdat}
+
+%TODO: add data/sample composition -- ethinicity, grade, etc -- maybe in table form
+%TODO: find out where  the genes were taken from
+%TODO: Change Fuentes-Mattei citation to Cris Print's citation
+
+The raw microarray gene expression data files used in  \citet{Gatza2010a} study were downloaded from the \gls{geo} database (include website name/GEO ID here).
+Clinical data for the samples in this study was not available.
+130 obesity associated gene probes identified by \citet{Gatza2010a} were downloaded from (supplementary data??).
 The raw expression data and the list of gene probes were imported into \gls{r}/RStudio.
@@
 \gls{rnaseq} data for the eight cancer types identified in \cref{subsec:tcga} were downloaded from the \gls{icgc} database.
-The \gls{rnaseq} data for each of the eight cancer types were imported into \gls{r}.
+The \gls{rnaseq} data for each of the eight cancer types were imported into \gls{r}/RStudio.
 
@@
 
-\subsection{Creighton data}
-\label{subsec:crdatproc}
+\subsection{Data normalisation}
+\label{sub:data_normalisation}
 
-% \subsubsection{\textbf{Data processing}}
+\subsubsection{Microarray data}
+\label{ssub:microarray_data}
 
-\subsubsection{\textbf{Raw data normalisation}}
+The raw microarray data were normalised using either the \gls{rma} or \gls{mas} method from the \textit{affy} package in \gls{r}.
+Either \gls{rma} or \gls{mas} normalisation was used to account for the inter-array differences that may have arisen during the data collection process.
 
-The raw microarray data for all 103 samples from \cref{subsec:crrawdat} were normalised using the \gls{rma} method from the \textit{affy} package in \gls{r}.
-\gls{rma} normalisation was used to account for the inter-array differences that may have arisen during the data collection process.
+\subsubsection{RNA-seq data}
+\label{ssub:rna_seq_data}
 
-\subsubsection{\textbf{Raw data standardisation}}
+The raw \gls{rnaseq} data for each cancer types downloaded from \gls{icgc} website had 1 added, and then logged to the power of 10.
 
-The data was standardised so that each gene had a mean ($\mu$) of 0 and a standard deviation ($\sigma$) of 1.
+\subsection{Data standardisation}
+\label{sub:data_standardisation}
+
+The data was standardised so that each gene in the data had a mean ($\mu$) of 0 and a standard deviation ($\sigma$) of 1.
 This ensured that all of the genes were on the same scale as one another, allowing for comparison across all the genes.
 
-\subsubsection{\textbf{Gene probe ID conversion}}
+\subsection{Residual data creation}
+\label{sub:residual_data_creation}
+
+To create the residual data for a given data set, a linear model was fitted to the data with all of the clinical variables that required to be controlled for.
+The \textit{limma} package in \gls{r} was used.
+
+\subsection{Gene probe ID conversion}
+\label{sub:gene_probe_id_conversion}
 
@@
 
-\subsubsection{\textbf{Sample \gls{bmi} calculation}}
+\subsection{Batch correction}
+\label{sub:batch_correction}
 
-\gls{bmi} of the samples were calculated from the clinical data using \cref{eq:bmicalc}.
+When there were more than one microarray data to be used as a single data set, these microarray data were normalised separately then combined together with the \texttt{ComBat} function in the \textit{sva} package in \gls{r}.
+This corrected for the effect of the individual microarray data on the whole data when they were combined together.
 
-\subsection{Fuentes-Mattei data}
-\label{subsec:fmdatproc}
+\subsection{Sample \gls{bmi} calculation}
+\label{sub:sample_bmi_calculation}
 
-% \subsubsection{\textbf{Data processing}}
+Where missing, the \gls{bmi} of the samples were calculated from the clinical data using \cref{eq:bmicalc}.
 
-\subsubsection{\textbf{Raw data normalisation}}
+\subsection{\gls{icgc} data format conversion}
+\label{sub:icgc_data_format_conversion}
 
-The raw microarray data for all *** samples from \cref{subsec:fmrawdat} were normalised using the \gls{rma} method from the \textit{affy} package in \gls{r}.
-\gls{rma} normalisation was used to account for the inter-array differences that may have arisen during the data collection process.
+The raw \gls{rnaseq} data for each cancer types were formatted using the \textit{dplyr} package in \gls{r}, so that the rows corresponded to the genes and the columns to the samples.
 
-\subsubsection{\textbf{Raw data standardisation}}
+\subsection{Sample ID conversion}
+\label{sub:sample_id_conversion}
 
-The data was standardised so that each gene had a mean ($\mu$) of 0 and a standard deviation ($\sigma$) of 1.
-This ensured that all of the genes were on the same scale as one another, allowing for comparison across all the genes.
+\gls{icgc} sample ID in the \gls{rnaseq} data were stripped so that the sample ID matched with the \gls{tcga} ID in the clinical data.
+After the IDs were stripped, samples were checked for any duplicates.
+Where there was a duplicate in the sample ID in the \gls{rnaseq} data, a single sample was chosen to represent that particular ID by using the \texttt{collapseRows} function in the \textit{WGCNA} package in \gls{r}.
 
-\subsubsection{\textbf{Gene probe ID conversion}}
-
-Where appropriate, the gene probe IDs in the raw data were converted into their corresponding gene symbols using the \textit{hgu133a.db} package in \gls{r}.
-Where there was an overlap in the gene symbols, a single data was chosen to represent that particular gene symbol by using the \texttt{collapseRows} function in the \textit{WGCNA} package in \gls{r}.
-
-\subsection{\gls{tcga}}
-\label{subsec:datproctcga}
-
-\subsubsection{\textbf{Raw data normalisation}}
-
-The raw \gls{rnaseq} data for each cancer types from \cref{subsec:tcga} had 1 added, and then logged to the power of 10.
-
-\subsubsection{\textbf{Raw data standardisation}}
-
-The normalised data was standardised so that each gene had a mean ($\mu$) of 0 and a standard deviation ($\sigma$) of 1.
-This ensured that all of the genes were on the same scale as one another, allowing for comparison across all the genes.
-
-\subsection{\gls{icgc}}
-\label{subsec:datprocicgc}
-
-\subsubsection{\textbf{Raw data format conversion}}
-%TODO: add in how I processed the ICGC data format into genes by samples.
-The raw \gls{rnaseq} data for each cancer types were formatted so that the rows corresponded to the genes and the columns to the samples.
-% TODO: find out what functions I used to reformat the icgc data
-This was done using the \textit{dplyr} package in \gls{r}.
-
-\subsubsection{\textbf{Sample ID conversion}}
-%TODO: add details on how I stripped the ICGC ID into TCGA ID
-%TODO: check the actual ID stripping process
-%TODO: check the WGCNA function and check if it uses median/mean/whatever for choosing the sample
-\gls{icgc} sample ID in the \gls{rnaseq} data were stripped so that the ID matched with the \gls{tcga} ID in the clinical data.
-After the IDs were stripped, samples were checked for any overlaps.
-Where there was an overlap in the sample ID in the \gls{rnaseq} data, a single sample was chosen to represent that particular ID by using the \texttt{collapseRows} function in the \textit{WGCNA} package in \gls{r}.
-
-\subsubsection{\textbf{\gls{icgc} sample exclusion criteria}}
-\label{subsec:icgcsampleexclusion}
-
-% \subsubsection{\textbf{Sample BMI availability}}
+\subsection{\gls{icgc} sample exclusion}
+\label{sub:icgc_sample_exclusion}
 
@@
 These samples were removed from the analyses.
-In each of the eight cancer types, the \gls{tcga} sample ID in the clinical data were cross-checked with the \gls{tcga} sample ID in the \gls{rnaseq} data, and vice versa.
+In each of the eight cancer types, the \gls{tcga} sample ID in the clinical data were cross-checked with the sample ID in the \gls{rnaseq} data, and vice versa.
 Any sample that did not have either clinical or \gls{rnaseq} data were removed.
@@
 
+\section{Gene expression analysis}
+\label{sec:gene_expression_analysis}
+
+The samples were grouped based on the sample BMI status (obese or non-obese group), and a linear model was fitted to the data based on the sample grouping.
+Empirical Bayes statisticsc was used to identify the differentially expressed genes between the two groups, and the top scoring genes were listed.
+The error rate of the genes were \gls{fdr} controlled for multiple hypothesis testing.
+
+\section{Pathway enrichment analysis}
+\label{sec:pathway_enrichment_analysis}
+
+For a given list of differentially expressed genes identified in a data set, the \texttt{camera} function from the \textit{limma} package was used to identify the pathways that were enriched in those genes.
+The \gls{go} database was used to identify the enriched pathways.
+
+\section{\gls{svd}}
+\label{sec:singular_value_decomposition}
+
+\gls{svd} was used on a given data set to create the summary metagene scores of a given genetic signature.
+% TODO: add more details on the components (s, v, d) created when svd was applied to the data.
+Any matrix $X$ can be represented in the form:
+\begin{equation}
+	\label{eq:svd}
+	X = UDV'
+\end{equation}
+
+\noindent
+where $U$ is the *****, $D$ is the ********, and $V'$ is the **** (eigenvectors and eigenvalues??).
+
+\section{Transformation matrix}
+\label{sec:transformation_matrix}
+
+Transformation matrix of a genetic signature was created from the \gls{svd} components.
+Rearranging \cref{eq:svd}, you get:
+
+\begin{equation}
+	\label{eq:transmat}
+	V' = U'D^{-1}X
+\end{equation}
+
+\noindent
+where $U'D^{-1}$ represents the transformation matrix.
+The transformation matrix $U'D^{-1}$ was created from the \gls{svd} components, and was used to transform other data set and obtain the metagene in that data set.
+
+\section{Metagene analysis}
+\label{sec:metagene_analysis}
+
+\gls{svd} was applied to the training data set to create the metagene and the transformation matrix in the data set.
+The transformation matrix from the training data was used to transform the other data sets and to create the metagene of the genetic signature in those data sets.
+
+Each of these metagenes were plotted in a heatmap to see the association between the metagene and the overall gene expression of the genes that were used to create the metagene.
+The metagene was also checked for the association between the sample \gls{bmi} status and \gls{bmi} value, which was plotted in a boxplot and scatter plot, respectively.
+
+\section{Gatza pathway metagene direction}
+\label{sec:pathway_metagene_direction}
+
+\section{Obesity metagene prediction}
+\label{sec:obesity_metagene_prediction}
+
 \section{Plot creation}
@@
 
-\subsection{Box plots}
-\label{sub:box_plots}
+% Do I need to explain how to make boxplots and scatter plots?
+% \subsection{Box plots}
+% \label{sub:box_plots}
 
-\subsection{Scatter plots}
-\label{sub:scatter_plots}
+% \subsection{Scatter plots}
+% \label{sub:scatter_plots}
 
-\section{Gene expression analysis}
-\label{sec:gene_expression_analysis}
-
-\section{Metagene analysis}
-\label{sec:metagene_analysis}
-
-\section{Pathway enrichment analysis}
-\label{sec:pathway_enrichment_analysis}
+% TODO: re-write everything so that it's in past tense

### 2016-09-17 thesis/misc/abstract.tex
tmp commit

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Abstract for my thesis
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\vspace*{\fill}

\section*{\centering Abstract}
\addcontentsline{toc}{section}{Abstract}

It is now commonly understood that cancers are caused by dysregulation of various mechanisms or pathways that allow tumour cells to prolilferate, survive and migrate.
There have been many reports on the association between obesity and worsened cancer prognosis, suggesting that there may be distinct changes that underlie tumorigenesis and/or cancer progression that are specific to patients with higher \gls{bmi}.

In brief, this research aims to determine whether gene expression signatures exist that are specific to obesity across multiple cancer types, and to investigate whether there are any common pathways being dysregulated in cancers based on these genetic signatures.
Better understanding of the pathways being dysregulated in cancer cells in obese patients may lead to improved clinical decisions, and could thus contribute towards personalised treatment in the future.

\vfill

### 2016-09-17 thesis/results/results2.tex
tmp commit

\chapter{Results from gatza pathway metagene and direction stuff}

\section{Reproducing results from gatza paper}

### 2016-09-17 thesis/methods/methods.tex
tmp commit

@@
 The samples were grouped based on the sample BMI status (obese or non-obese group), and a linear model was fitted to the data based on the sample grouping.
-Empirical Bayes statisticsc was used to identify the differentially expressed genes between the two groups, and the top scoring genes were listed.
+Empirical Bayes statistics was used to identify the differentially expressed genes between the two groups, and the top scoring genes were listed.
 The error rate of the genes were \gls{fdr} controlled for multiple hypothesis testing.
@@
 
+All 18 pathway signatures from \citet{Gatza2010a} study were used to create both the metagenes and transformation matrices in the Gatza data set.
+Each of the 18 pathway metagenes were compared with the gene expression of the gene that represents the pathway to check whether the metagenes were in the correct direction.
+% TODO: check if it was the AKT1 or AKT2 gene for AKT pathway
+For example, the AKT pathway metagene was compared with the AKT1 gene expression.
+
+The metagenes were flipped depending on the
+
 \section{Obesity metagene prediction}

### 2016-09-19 thesis/methods/methods.tex
Finished very rough draft of the methods section

@@
 
-\gls{bmi} was calculated from the height and weight data of the samples using the following equation:
+\cref{eq:bmicalc}  was used to calculate the \gls{bmi} from the height and weight data of the samples.
 
@@
 			Overweight & 25.0$\sim$29.9\\
-			Obese & \textgreater{} 30.0\\
+			Obese & \geq{} 30.0\\
 		\end{tabular}
@@
 The raw microarray data from \citet{Creighton2012}, \citet{Fuentes-Mattei2014} and \citet{Gatza2010a}  studies were downloaded from the \gls{geo} website.
-\gls{rnaseq} and clinical data for multiple different cancer types were downloaded from free and publicly available sources such as \gls{icgc} and \gls{tcga}, respectively.
+\gls{rnaseq} and clinical data of multiple different cancer types were downloaded from free and publicly available sources such as \gls{icgc} and \gls{tcga}, respectively.
 
@@
 
-The raw microarray gene expression data files of all *** samples from the  \citet{Fuentes-Mattei2014} study were downloaded from the \gls{geo} database (include website name/GEO ID here).
+The raw microarray gene expression data files of all *** samples from the  Print et al. study were downloaded from the \gls{geo} database (include website name/GEO ID here).
 Clinical data for the samples in this study was not available.
-130 obesity associated gene probes identified by \citet{Fuentes-Mattei2014} were downloaded from (supplementary data??).
+*** obesity associated gene probes identified by Print et al. were downloaded from (supplementary data??).
 The raw expression data and the list of gene probes were imported into \gls{r}/RStudio.
@@
 The raw microarray gene expression data files used in  \citet{Gatza2010a} study were downloaded from the \gls{geo} database (include website name/GEO ID here).
+Only the Affymetrix HGU133A microarray samples were included in the project.
 Clinical data for the samples in this study was not available.
-130 obesity associated gene probes identified by \citet{Gatza2010a} were downloaded from (supplementary data??).
+*** obesity associated gene probes identified by \citet{Gatza2010a} were downloaded from (supplementary data??).
 The raw expression data and the list of gene probes were imported into \gls{r}/RStudio.
@@
 
-\gls{rnaseq} data for the eight cancer types identified in \cref{subsec:tcga} were downloaded from the \gls{icgc} database.
+\gls{rnaseq} data for the eight cancer types from \cref{subsec:tcga} were downloaded from the \gls{icgc} database.
 The \gls{rnaseq} data for each of the eight cancer types were imported into \gls{r}/RStudio.
@@
 
-The raw \gls{rnaseq} data for each cancer types downloaded from \gls{icgc} website had 1 added, and then logged to the power of 10.
+The raw \gls{rnaseq} data were normalised in two different ways, depending on the analysis.
+For gene expression analyses, \texttt{voom} function from the \textit{limma} package was used to normalise the data.
+For the purposes of data visualisation or application of metagene transformation matrices, the data had 1 added and then logged to the base of 10.
 
@@
 The data was standardised so that each gene in the data had a mean ($\mu$) of 0 and a standard deviation ($\sigma$) of 1.
-This ensured that all of the genes were on the same scale as one another, allowing for comparison across all the genes.
+This ensured that all of the genes were on the same scale as one another, which allowed direct comparison across all of the genes.
 
@@
 
-To create the residual data for a given data set, a linear model was fitted to the data with all of the clinical variables that required to be controlled for.
-The \textit{limma} package in \gls{r} was used.
+A linear model was constructed from the data with all of the clinical variables controlled for.
+The residual data was taken from this linear model to be used as the residual data for that data set.
 
@@
 When there were more than one microarray data to be used as a single data set, these microarray data were normalised separately then combined together with the \texttt{ComBat} function in the \textit{sva} package in \gls{r}.
-This corrected for the effect of the individual microarray data on the whole data when they were combined together.
+This corrected for the effect of the individual microarray data on the whole data when combined together.
 
@@
 where $U'D^{-1}$ represents the transformation matrix.
-The transformation matrix $U'D^{-1}$ was created from the \gls{svd} components, and was used to transform other data set and obtain the metagene in that data set.
+The transformation matrix $U'D^{-1}$ was created from the \gls{svd} components, and was used to transform other data set and obtain the metagene in the data set.
 
@@
 Each of these metagenes were plotted in a heatmap to see the association between the metagene and the overall gene expression of the genes that were used to create the metagene.
-The metagene was also checked for the association between the sample \gls{bmi} status and \gls{bmi} value, which was plotted in a boxplot and scatter plot, respectively.
+The metagene was also checked for the association between the sample \gls{bmi} status and \gls{bmi} value, which was plotted in a box plot and a scatter plot, respectively.
 
@@
 
-The metagenes were flipped depending on the
+The pathway metagene value was first compared visually with its pathway gene in a heatmap.
+If the direction of the pathway metagene was correct (for example, high pathway metagene value corresponded with high expression of the gene representing that pathway), the metagene values were not changed, but if it was incorrect, the metagene values were flipped for that pathway.
+
+Spearman correlation of the pathway metagene and the expression of the pathway gene was also used to distinguish whether the pathway metagene was in the right direction.
+The direction of the metagene was considered correct when the correlation was positive and incorrect when the correlation was negative.
+
+The clustering of the pathway metagenes were taken into account to decide on the final direction of the metagene.
+The direction of the pathway metagenes were modified until the resulting pathway metagenes were similarly clustered together as in the results from \citet{Gatza2010a} paper.
 
@@
 \label{sec:obesity_metagene_prediction}
+
+% TODO: possibly extra analyses here (gradual pathway inclusion in the model, using all the pathway...?)
+Sample BMI, BMI status and/or pathway metagenes were used to construct a linear model in Cris Print's data.
+The constructed linear model was then used to predict the obesity metagene values in \citet{Creighton2012} data set.
+The predicted obesity metagene values were compared with the true obesity metagene values, and both Spearman and Pearson correlation were calculated.
 
@@
 
-% Do I need to explain how to make boxplots and scatter plots?
-% \subsection{Box plots}
-% \label{sub:box_plots}
-
-% \subsection{Scatter plots}
-% \label{sub:scatter_plots}
+Heatmaps were created using either the \texttt{heatmap.2} function from \textit{gplots} package, or the \texttt{heatmap.2x} function written by Tom Kelly.
+All of the heatmaps that required no or one column bar used the \texttt{heatmap.2} function, and all of the heatmaps that required more than one column bars used the \texttt{heatmap.2x} function.
 

### 2016-09-19 thesis/results/results.tex
Starting the write up of the first part of the results

\chapter{Obesity-associated genetic signatures and cancer}
\label{cha:obesity_genetic_signatures_and_cancer}

\section{Obesity-associated genetic signature from \citet{Creighton2012} study}
\label{sec:creighton_obesity_metagene}

\section{Obesity-associated genetic signature from \citet{Fuentes-Mattei2014} study}
\label{sec:fm_obesity_metagene}

\section{Common genes across multiple cancer types}
\label{sec:common_genes_across_multiple_cancer_types}

\section{Pathways enriched in obesity-associated genes}
\label{sec:pathways_enriched_in_obesity_associated_genes}

### 2016-09-19 thesis/results/results.tex
tmp commit

@@
 \label{cha:obesity_genetic_signatures_and_cancer}
+
+In this chapter, the relationship between the obesity-associated genetic signatures from previous studies and different cancer types were investigated; new obesity-associated genetic signatures were identified in the data set from \citet{Creighton2012} study; and an attempt was made to find any common genes or pathways across different cancer types.
+
+% TODO: figure out what results go into which section, and present it in a logical order
 
@@
 
+\section{New obesity-associated genetic signature in \citet{Creighton2012} data set}
+\label{sec:creighton_obesity_metagene_new}
+
 \section{Common genes across multiple cancer types}

### 2016-09-20 thesis/results/results.tex
tmp commit

--- 2016-09-19 thesis/results/results.tex
@@
 \label{cha:obesity_genetic_signatures_and_cancer}
+
+In this chapter, the relationship between the obesity-associated genetic signatures from previous studies and different cancer types were investigated; new obesity-associated genetic signatures were identified in the data set from \citet{Creighton2012} study; and the presence of common genes or pathways across different cancer types was examined.
 

### 2016-09-20 thesis/methods/methods.tex
tmp commit

@@
 
-%TODO: add data/sample composition -- ethinicity, grade, etc -- maybe in table form
-
-The raw microarray gene expression data files of all 103 samples from the \citet{Creighton2012} study were downloaded from the \gls{geo} database (\gls{geo} accession ID: GSE24185).
+The raw microarray gene expression data files from the \citet{Creighton2012} study were downloaded from the \gls{geo} database (\gls{geo} accession ID: GSE24185).
 Clinical data of the samples were obtained from the supplementary table 1 from \citet{Creighton2012} paper.
 799 obesity associated gene probes identified in the \citet{Creighton2012} study were obtained from the supplementary data file 1 from \citet{Creighton2012} paper.
-The raw expression data, clinical data, and the list of gene probes were imported into \gls{r}/RStudio.
+The raw expression data, clinical data, and the list of gene probes were imported into \gls{r}.
 
@@
 
-%TODO: add data/sample composition -- ethinicity, grade, etc -- maybe in table form
-%TODO: find out where  the genes were taken from
+The raw microarray gene expression data files from the  \citet{Fuentes-Mattei2014} study were downloaded from the \gls{geo} database (\gls{geo} accession ID: GSE20194).
+Clinical data for the samples in this study was also downloaded from the \gls{geo} database (same \gls{geo} accession ID).
+130 obesity associated gene probes identified by \citet{Fuentes-Mattei2014} were taken from the supplementary table 3.
+The raw expression data and the list of gene probes were imported into \gls{r}.
 
-The raw microarray gene expression data files of all *** samples from the  \citet{Fuentes-Mattei2014} study were downloaded from the \gls{geo} database (include website name/GEO ID here).
+\subsection{Raw data from \citet{Gatza2010a} study}
+\label{sub:gatzarawdat}
+
+The raw microarray gene expression data files from the \citet{Gatza2010a} study were downloaded from the \gls{geo} database (\gls{geo} accession ID: GSE1456, GSE1561, GSE2034, GSE3494, GSE4922, and GSE6596).
+Only the Affymetrix HGU133A microarray samples were included in the project.
 Clinical data for the samples in this study was not available.
-130 obesity associated gene probes identified by \citet{Fuentes-Mattei2014} were downloaded from (supplementary data??).
-The raw expression data and the list of gene probes were imported into \gls{r}/RStudio.
+The raw expression data and the list of gene probes were imported into \gls{r}.
 
@@
 
-%TODO: add data/sample composition -- ethinicity, grade, etc -- maybe in table form
-%TODO: find out where  the genes were taken from
-%TODO: Change Fuentes-Mattei citation to Cris Print's citation
+%TODO: Find Cris's paper
 
-The raw microarray gene expression data files of all *** samples from the  Print et al. study were downloaded from the \gls{geo} database (include website name/GEO ID here).
-Clinical data for the samples in this study was not available.
-*** obesity associated gene probes identified by Print et al. were downloaded from (supplementary data??).
-The raw expression data and the list of gene probes were imported into \gls{r}/RStudio.
+The raw microarray gene expression data files from the  Print et al. study were downloaded from the \gls{geo} database (\gls{geo} accession ID: GSE36771).
+Clinical data for the samples in this study was ******.
+The raw expression data and the list of gene probes were imported into \gls{r}.
 
-\subsection{Raw data from \citet{Gatza2010a} study}
-\label{sub:gatzarawdat}
-
-%TODO: add data/sample composition -- ethinicity, grade, etc -- maybe in table form
-%TODO: find out where  the genes were taken from
-%TODO: Change Fuentes-Mattei citation to Cris Print's citation
-
-The raw microarray gene expression data files used in  \citet{Gatza2010a} study were downloaded from the \gls{geo} database (include website name/GEO ID here).
-Only the Affymetrix HGU133A microarray samples were included in the project.
-Clinical data for the samples in this study was not available.
-*** obesity associated gene probes identified by \citet{Gatza2010a} were downloaded from (supplementary data??).
-The raw expression data and the list of gene probes were imported into \gls{r}/RStudio.
+% TODO: Overall summary table of the samples from each data set at the end of this section.
 
@@
 		\begin{tabular}{cc}
-			\textbf{Cancer Type}   & \textbf{Total} \\
+			\textbf{Cancer Type}   & \textbf{Total number of samples} \\
 			\hline

### 2016-10-22 thesis/intro/intro.tex
tmp commit

@@
 
-General intro to obesity.
-Implications of obesity -- disease associated with it and/or cancer.
+% General intro to obesity.
+% Implications of obesity -- disease associated with it and/or cancer.
 
@@
 
-What is cancer.
+% What is cancer.
 
@@
 % TODO: edit the starting sentence so it's not a copy from the beginning
-It is now commonly understood that cancers are caused by dysregulation of various mechanisms or pathways that allow tumour cells to prolilferate, survive and migrate.
-Many of these dysregulated pathways frequently occur in cancer cells, and these characteristic traits, commonly known as the ``Hallmarks of Cancer'', were first described by \citet{Hanahan2000}.
+% It is now commonly understood that cancers are caused by dysregulation of various mechanisms or pathways that allow tumour cells to prolilferate, survive and migrate.
+% Many of these dysregulated pathways frequently occur in cancer cells, and these characteristic traits, commonly known as the ``Hallmarks of Cancer'', were first described by \citet{Hanahan2000}.
 
-These hallmarks were initially presented in 2000 to provide a common framework for scientists to better understand cancer biology, and also to direct the focus of future research.
-After a decade of advances in cancer research, the ``Hallmarks of Cancer'' were revisited by Hanahan and Weinberg in 2011, and four more hallmarks were added on top of the six presented in the original paper in 2000 \citep{Hanahan2011}.
+% These hallmarks were initially presented in 2000 to provide a common framework for scientists to better understand cancer biology, and also to direct the focus of future research.
+% After a decade of advances in cancer research, the ``Hallmarks of Cancer'' were revisited by Hanahan and Weinberg in 2011, and four more hallmarks were added on top of the six presented in the original paper in 2000 \citep{Hanahan2011}.
 
-Together, these ``Hallmarks of Cancer'' (fig 1) have provided a foundation of the general biological knowledge of cancer and its characteristics that allowed the researchers to better understand the underlying cancer biology.
+% Together, these ``Hallmarks of Cancer'' (fig 1) have provided a foundation of the general biological knowledge of cancer and its characteristics that allowed the researchers to better understand the underlying cancer biology.
 
-Although much reseach has been conducted on the ``Hallmarks of Cancer'', the molecular processes underlying these hallmarks have still not been fully explained \citep{Hanahan2011}.
-This is due to the fact that cancer cells develop progressively under selective pressures, and are also affected by many environmental factors \citep{Gatza2011}, which makes it difficult to identify causal factors of cancer development.
-Furthermore, these external factors affect how the cancer cells respond to the environment, and may also cause cancer progression and/or drug resistance.
+% Although much reseach has been conducted on the ``Hallmarks of Cancer'', the molecular processes underlying these hallmarks have still not been fully explained \citep{Hanahan2011}.
+% This is due to the fact that cancer cells develop progressively under selective pressures, and are also affected by many environmental factors \citep{Gatza2011}, which makes it difficult to identify causal factors of cancer development.
+% Furthermore, these external factors affect how the cancer cells respond to the environment, and may also cause cancer progression and/or drug resistance.
 
@@
 
-Why is it important in obesity context.
+% Why is it important in obesity context.
 
-The exact mechanism in which obesity accelerates cancer progression is unknown, but obesity is thought to disrupt the levels of various hormones (such as insulin, leptin, adeponectin, and other adipokines) that eventually activate tumorigenic pathways (reference the overview on obesity and tumour resistance).
+% The exact mechanism in which obesity accelerates cancer progression is unknown, but obesity is thought to disrupt the levels of various hormones (such as insulin, leptin, adeponectin, and other adipokines) that eventually activate tumorigenic pathways (reference the overview on obesity and tumour resistance).
 
@@
 
-Due to advances in sequencing technology over the last few years, many researchers have started to focus on bioinformatics based analyses of cancer cells to better understand the biology of cancer.
-Advantages of this strategy are that very specific gene expression and/or mutational signatures for certain cancer subtypes can be identified from the analyses, which the researchers and clinicians can study in much greater depth and details.
+% Due to advances in sequencing technology over the last few years, many researchers have started to focus on bioinformatics based analyses of cancer cells to better understand the biology of cancer.
+% Advantages of this strategy are that very specific gene expression and/or mutational signatures for certain cancer subtypes can be identified from the analyses, which the researchers and clinicians can study in much greater depth and details.
 
@@
 
-Talk about \citet{Gatza2010a}
+% Talk about \citet{Gatza2010a}
 

### 2016-10-22 thesis/methods/methods.tex
tmp commit

@@
 
-The raw microarray gene expression data files from the  \citet{Fuentes-Mattei2014} study were downloaded from the \gls{geo} database (\gls{geo} accession ID: GSE20194).
+The raw microarray gene expression data files from the  \citet{Fuentes-Mattei2014} study were downloaded from the \gls{geo} database (\gls{geo} accession ID: GSE\-20194).
 Clinical data for the samples in this study was also downloaded from the \gls{geo} database (same \gls{geo} accession ID).
@@
 
-The raw microarray gene expression data files from the \citet{Gatza2010a} study were downloaded from the \gls{geo} database (\gls{geo} accession ID: GSE1456, GSE1561, GSE2034, GSE3494, GSE4922, and GSE6596).
-Only the Affymetrix HGU133A microarray samples were included in the project.
+The raw microarray gene expression data files from the \citet{Gatza2010a} study were downloaded from the \gls{geo} database (\gls{geo} accession ID: GSE1456, GSE\-1561, GSE2034, GSE3494, GSE4922, and GSE6596).
+Only the Affymetrix HGU\-133A microarray samples were included in the project.
 Clinical data for the samples in this study was not available.

### 2016-10-25 thesis/intro/intro.tex
tmp commit
Started my thesis introduction

@@
 % Implications of obesity -- disease associated with it and/or cancer.
+Obesity has been one of the major global problems for more than a decade, associated with many noncommunicable diseases such as diabetes, cardiovascular diseases and certain types of cancers \citep{WHO2014}.
+In fact, the risk of comorbidities increases with the increase in \gls{bmi}, where the risk becomes severe as the \gls{bmi} level approaches the obese category.
+
+The number of overweight and obese population, in both adults and children, have risen in every country of the world, and the trend continues as we speak.
+Estimated to account for 3.4 million deaths per year and 93.6 million \glspl{daly} in 2010, obesity is a serious disease that continues to grow in our society, which the world cannot ignore (citation from 2010 WHO report needed).
+
+\subsection{Definition of obesity}
+\label{sub:definition_of_obesity}
+% TODO: add WHO website to the reference for the definition of obesity (or find proper articles)
+
+Obesity is defined as an abnormal or excessive fat accumulation that may impair health of that individual.
+
+One common and widely used approach to categorise individuals is by measuring the \gls{bmi} of the individual.
+\gls{bmi} is a measurement based on the weight-to-height ratio of an individual that is used by clinicians and epidemiologists to classify adults into underweight, normal weight, overweight or obese categories.
+The unit of \gls{bmi} is defined as the individual's weight in kilograms per square of the height in meters (kg/m$^2$).
+
+\gls{who} (2014) have used \gls{bmi} measurements to define overweight and obesity in adults as an individual with \gls{bmi} $\geq$ 25 kg/m$^2$ and $\geq$ 30 kg/m$^2$, respectively.
+For children under the age of 5, overweight and obesity were defined as weight-for-height greater than 2 standard deviation above \gls{who} Child Growth Standards median and greater than 3 standard deviation above the median, respectively.
+For children aged between 5 to 19 years, overweight and obesity were defined as BMI-for-age greater than 1 standard deviation above \gls{who} Growth Reference Standards median and greater than 2 standard deviation above the median, respectively.
+
+Although the use of \gls{bmi} as an accurate measure of body fat composition and/or representation of the individual's metabolic status remains controversial, many studies have used \gls{bmi} due to the ease of data collection compared to other measurements, such as [method] (citation needed).
+Consequently, most of the publicly available clinical data have height and weight data from which one can calculate the patients' \gls{bmi}, and therefore able to collect the patients' obesity status more conveniently than, for example, waist-to-hip ratio.
 
@@
 
+% \subsection{Preventive measure of obesity}
+% \label{sub:preventive_measure_of_obesity}
+
 \subsection{Obesity and cancer risk}
 \label{sub:obesity_and_cancer_risk}
+
+\subsection{Impact of obesity}
+\label{sub:impact_of_obesity}
 
@@
 
+\subsection{Microarray technology}
+\label{subsec:microarray_technology}
+
+\subsection{Next generation sequencing technology}
+\label{sub:next_generation sequencing_technology}
+
 % Due to advances in sequencing technology over the last few years, many researchers have started to focus on bioinformatics based analyses of cancer cells to better understand the biology of cancer.
@@
 
-\subsection{Obesity associated genetic signatures}
-\label{subsec:obesity_associated_genetic_signatures}
+\section{Obesity associated genetic signatures}
+\label{sec:obesity_associated_genetic_signatures}
 
-\subsubsection{\textbf{\citet{Creighton2012} study}}
-\label{ssub:creighton_study}
+\subsection{\citet{Creighton2012} study}
+\label{sub:creighton_study}
 
-\subsubsection{\textbf{\citet{Fuentes-Mattei2014} study}}
-\label{ssub:fuentes_mattei_study}
+\subsection{\citet{Fuentes-Mattei2014} study}
+\label{sub:fuentes_mattei_study}
 
-\subsection{Pathway associated genetic signatures}
-\label{sub:pathway_associated_genetic_signatures}
+\section{Pathway associated genetic signatures}
+\label{sec:pathway_associated_genetic_signatures}
 
@@
 
+\subsection{\citep{Gatza2011} study}
+\label{sub:gatza_study}
+
 \section{Aim of the project}
 \label{sec:aim}
+
+This research aims to determine whether gene expression signatures exist  that are specific to obesity across multiple cancer types, and to investigate whether there are any common pathways being dysregulated in cancers based on these genetic signatures.
+Better understanding of the pathways being dysregulated in cancer cells in obese patients may lead to improved clinical decisions, and contribute towards personalised treatment in the future.

### 2016-10-25 thesis/misc/glossary.tex
tmp commit
Started my thesis introduction

@@
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%% Glossary terms
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+% TODO: make sure you get a proper definitions for all of these glossary terms
+
+\newglossaryentry{rnaseq}{
+	name=RNA-seq,
+	description={sequence data using transcriptome}
+}
+
+\newglossaryentry{r}{
+	name=R,
+	description={Statistical programming language}
+}
+
+\newglossaryentry{metagene}{
+	name=Metagene,
+	text=metagene,
+	description={Summary score of a sample that represents the overall gene expression of the genes that was used to obtain the scores}
+}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%% Acronym terms
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
 \newacronym{blca}{BLCA}{bladder urothelial carcinoma}
@@
 \newacronym{her2}{HER2}{Human epidermal growth factor receptor 2}
-\newacronym{svd}{SVD}{Singular Value Decomposition}
+\newacronym{svd}{SVD}{Singular value decomposition}
 \newacronym{rma}{RMA}{Robust Multi Array}
+\newacronym{mas}{MAS5}{Microarray Analysis Suite 5.0}
+\newacronym{fdr}{FDR}{False Discovery Rate}
+\newacronym{go}{GO}{Gene Ontology}
+\newacronym[longplural={Disability-Adjusted Life Years}]{daly}{DALY}{Disability-Adjusted Life Year}

### 2016-10-26 thesis/intro/intro.tex
tmp commit

@@
 Obesity has been one of the major global problems for more than a decade, associated with many noncommunicable diseases such as diabetes, cardiovascular diseases and certain types of cancers \citep{WHO2014}.
-In fact, the risk of comorbidities increases with the increase in \gls{bmi}, where the risk becomes severe as the \gls{bmi} level approaches the obese category.
+In fact, the risk of comorbidities increases with the increase in \gls{bmi}, where the risk becomes severe as the \gls{bmi} level approaches the obese category (citation).
 
@@
 
-Obesity is defined as an abnormal or excessive fat accumulation that may impair health of that individual.
+Obesity is defined as an abnormal or excessive fat accumulation that may impair health of that individual (citation).
 
@@
 \gls{who} (2014) have used \gls{bmi} measurements to define overweight and obesity in adults as an individual with \gls{bmi} $\geq$ 25 kg/m$^2$ and $\geq$ 30 kg/m$^2$, respectively.
-For children under the age of 5, overweight and obesity were defined as weight-for-height greater than 2 standard deviation above \gls{who} Child Growth Standards median and greater than 3 standard deviation above the median, respectively.
-For children aged between 5 to 19 years, overweight and obesity were defined as BMI-for-age greater than 1 standard deviation above \gls{who} Growth Reference Standards median and greater than 2 standard deviation above the median, respectively.
+For children under the age of 5, overweight and obesity were defined as weight-for-height greater than 2 standard deviation and 3 standard deviation above \gls{who} Child Growth Standards median, respectively.
+For children aged between 5 to 19 years, overweight and obesity were defined as BMI-for-age greater than 1 standard deviation and 2 standard deviation above \gls{who} Growth Reference Standards median, respectively.
 
-Although the use of \gls{bmi} as an accurate measure of body fat composition and/or representation of the individual's metabolic status remains controversial, many studies have used \gls{bmi} due to the ease of data collection compared to other measurements, such as [method] (citation needed).
-Consequently, most of the publicly available clinical data have height and weight data from which one can calculate the patients' \gls{bmi}, and therefore able to collect the patients' obesity status more conveniently than, for example, waist-to-hip ratio.
+Although the use of \gls{bmi} as an accurate measure of body fat composition and/or representation of the individual's metabolic status remains controversial, many studies have used \gls{bmi} due to the ease of data collection compared to other measurements, such as [example] (citation needed).
+Consequently, most of the publicly available clinical data have height and weight data from which one can calculate the patients' \gls{bmi}, and therefore able to collect the patients' obesity status more conveniently than, for example, [example].
 
@@
 
-% \subsection{Preventive measure of obesity}
-% \label{sub:preventive_measure_of_obesity}
+From the latest global status report on noncommunicable diseases by \citet{WHO2014}:
+\begin{quote}
+	\textit{
+		In 2014, 39\% of adults aged 18 years and older (38\% of men and 40\% of women) were overweight.
+		The worldwide prevalence of obesity nearly doubled between 1980 and 2014.
+		In 2014, 11\% of men and 15\% of women worldwide were obese.
+		Thus, more than half a billion adults worldwide are classed as obese.
+	}
+\end{quote}
 
-\subsection{Obesity and cancer risk}
-\label{sub:obesity_and_cancer_risk}
+\noindent
+This equated to approximately 1 in every 14 people in the world classed as obese in 2014, and therefore, 1 in 14 people around the world were also at severe health risks associated with obesity.
+It was also noted in the report that women were more likely to be obese than men, and that the prevalence of overweight and obesity increased as the income level of the countries increased \citep{WHO2014}.
+
+In addition to this, the worldwide prevalence of childhood overweight and obesity has been increasing steadily since 2000 \citep{WHO2014}.
+Furthermore, the prevalence of overweight in children under 5 years was estimated to reach from 6.3\% in 2013 to 11\% by 2025 if the current trend continued \citep{WHO2014}.
+Although the current obesity status of the adult population sounds daunting, the fact that more of the younger population who are overweight and obese poses a much greater risk for the health of the population.
+
+Until these terrible trends stop, and even after the trend stops, the world will need a plan to live along with this ongoing global epidemic of obesity and its associated diseases.
 
@@
 \label{sub:impact_of_obesity}
+
+% \subsection{Preventive measure of obesity}
+% \label{sub:preventive_measure_of_obesity}
 
@@
 
-\subsection{Relationship between obesity}
-\label{subsec:obsbackground}
+\section{Obesity and cancers}
+\label{sec:obesity_and_cancers}
+
+\subsection{Cancer risks associated with obesity}
+\label{sub:cancer_risks_associated_with_obesity}
+
+\subsection{Mechanism of cancer progression in obese patient}
+\label{sub:mechanism_of_cancer_progression_in_obese_patient}
 

### 2016-10-26 thesis/intro/intro.tex
tmp commit

@@
 
+\citep{Gelber2008, Lee2008, Yusuf2005}
+
 \subsection{Prevalence of obesity}
@@
 
-Until these terrible trends stop, and even after the trend stops, the world will need a plan to live along with this ongoing global epidemic of obesity and its associated diseases.
+Until these terrible trends stop, and even after the trends stop, the world will need a plan to live along with this ongoing global epidemic of obesity and its associated diseases.
+
+\subsection{Causes of Obesity}
+\label{sub:causes_of_obesity}
+
+\citep{Malik2013}
 
@@
 
-% \subsection{Preventive measure of obesity}
-% \label{sub:preventive_measure_of_obesity}
+\citep{Malik2013, Franks2010}
+
+\subsection{Preventive measure of obesity}
+\label{sub:preventive_measure_of_obesity}
+
+\citep{Malik2013}
 

### 2016-10-27 thesis/intro/intro.tex
tmp commit

@@
 Obesity has been one of the major global problems for more than a decade, associated with many noncommunicable diseases such as diabetes, cardiovascular diseases and certain types of cancers \citep{WHO2014}.
-In fact, the risk of comorbidities increases with the increase in \gls{bmi}, where the risk becomes severe as the \gls{bmi} level approaches the obese category (citation).
+In fact, the risk of comorbidities increases with the increase in \gls{bmi}, where the risk becomes severe as the \gls{bmi} level approaches the obese category (citation from 2000 WHO).
 
 The number of overweight and obese population, in both adults and children, have risen in every country of the world, and the trend continues as we speak.
-Estimated to account for 3.4 million deaths per year and 93.6 million \glspl{daly} in 2010, obesity is a serious disease that continues to grow in our society, which the world cannot ignore (citation from 2010 WHO report needed).
+Estimated to account for 3.4 million deaths per year and 93.6 million \glspl{daly} in 2010, obesity is a serious disease that continues to grow in our society (citation from 2010 WHO report needed).
 
@@
 
+Obesity is caused by excessive intake of energy without expending the energy gained.
+As one can suspect, there are many factors involved with the likelihood of an individual to become overweight and obese.
+
+\subsubsection{Individual-level (micro-level) factors}
+\label{ssub:Individual-level (micro-level) factors}
+
+There are many genetic evidence of obesity being caused by individual-level factors that makes some individuals more predisposed (right word??) to obesity than others (citation).
+\textit{Talk about individual-level factors that affect obesity here (need citations on individual factors of obesity; e.g. genetic variation).}
+
+\subsubsection{Population-level (macro-level) factors}
+\label{ssub:Population-level (macro-level) factors}
+
+Even though some individual-level factors are involved with obesity, it is clear that these individual-level factors cannot be accounted for the global epidemic of obesity that the world observes.
+For this reason, obesity is thought to be caused mainly by the environmental factors that are present in the population.
+
+The first key factor that contirbutes to the global rise in obese population is the availability and abundance of certain food in the country.
+
+\textit{(Expand on this -- better trades around the globe (?)) \citep{Kearney2010, Malik2013}}
+
+The second factor to consider is the behavioural changes led by the urbanisation of the population.
+\textit{(Expand on this -- physical activity, sociocultural norm, etc.\citep{Malik2013})}
+
+The third factor is the income and socioeconomic status of the countries.
+\textit{(Expand on this -- rise in income, availability of health care, etc.\citep{Malik2013})}
+
+The last factor, which ties in strongly with the first key factor, is the choice of diet by the population.
+\textit{(Expand on this -- increased variety of food, low quality and high energy food choice, etc.)}
+
 \citep{Malik2013}
@@
 
+As mentioned at the beginning of the current section, obesity has a devastating effect on the health of the population globally.
+
 \citep{Malik2013, Franks2010}
 
-\subsection{Preventive measure of obesity}
-\label{sub:preventive_measure_of_obesity}
+\subsection{Preventive measures of obesity}
+\label{sub:preventive_measures_of_obesity}
 

### 2016-10-28 thesis/intro/intro.tex
tmp commit

@@
 
-\citep{Gelber2008, Lee2008, Yusuf2005}
+possible papers:\citep{Gelber2008, Lee2008, Yusuf2005}
 
@@
 Furthermore, the prevalence of overweight in children under 5 years was estimated to reach from 6.3\% in 2013 to 11\% by 2025 if the current trend continued \citep{WHO2014}.
-Although the current obesity status of the adult population sounds daunting, the fact that more of the younger population who are overweight and obese poses a much greater risk for the health of the population.
+The current obesity status of the adult population is bad already, but the fact that more of the younger population who are overweight and obese poses a much greater risk for the health of the population (discussed in \cref{sub:impact_of_obesity}).
 
@@
 
-Even though some individual-level factors are involved with obesity, it is clear that these individual-level factors cannot be accounted for the global epidemic of obesity that the world observes.
-For this reason, obesity is thought to be caused mainly by the environmental factors that are present in the population.
+Even though some individual-level factors are involved with obesity, it is clear that these individual-level factors cannot be accounted for the global epidemic of obesity that the world currently observes.
+For this reason, obesity is thought to be caused mainly by the environmental factors that are present in the population.\\
 
+\noindent
 The first key factor that contirbutes to the global rise in obese population is the availability and abundance of certain food in the country.
+With the help from government trade policies, food and other goods have become easier to trade between the country, helping the economic growth of many countries around the world (citation -- Kearney?).
+This led to greater abundance of food in general, greater number of choices of food and shifted the overall nutritional status of the country \citep{Malik2013}.
+This shift in the food choice and availability (and the economic benefit associated with the trades) had a positive impact on the country, but the growth in the food market also resulted in a different, much larger problem -- obesity.
 
-\textit{(Expand on this -- better trades around the globe (?)) \citep{Kearney2010, Malik2013}}
+As an example, in the \gls{usa}, the cost of corn and soy is low due to the fact that they are the raw ingredients for most processed food and beverages, such as high fructose corn syrups used to sweeten many soft drinks in the \gls{usa} \citep{Malik2013}.
+Moreover, corn and soy are also the main food for the livestock, which results in lower prices for meat \citep{Malik2013}.
+On the othere hand, the prices of fruits and vegetables remain expensive due to the lack of support by the government to lower the cost associated with the production (more citation...) \citep{Malik2013}.
 
+As a result, more people are inclined to consume food that are cheap, have little nutritional value and high in energy, than the food that are expensive, nutritional and healthy alternatives.
+Thus, the population is more likely to become obese by making worst food choice.\\
+
+\noindent
 The second factor to consider is the behavioural changes led by the urbanisation of the population.
-\textit{(Expand on this -- physical activity, sociocultural norm, etc.\citep{Malik2013})}
 
+\textit{(Expand on this -- physical activity, sociocultural norm, etc.)}
+\citep{Malik2013}\\
+
+\noindent
 The third factor is the income and socioeconomic status of the countries.
-\textit{(Expand on this -- rise in income, availability of health care, etc.\citep{Malik2013})}
+\textit{(Expand on this -- rise in income, availability of health care, etc.)}
+\citep{Malik2013}\\
 
+\noindent
 The last factor, which ties in strongly with the first key factor, is the choice of diet by the population.

### 2016-10-28 thesis/intro/intro.tex
tmp commit

@@
 The second factor to consider is the behavioural changes led by the urbanisation of the population.
+Urbanisation is defined as the gradual increase in the proportion of people living in urban areas.
+As more people take up urban lifestyle, they are exposed to the new environment where there is a better range of food selection, as well as technological and mechanical advancement in transport and other daily chores that may not be available in rural areas where these people come from.
 
-\textit{(Expand on this -- physical activity, sociocultural norm, etc.)}
+Although there are many advantages involved with urbanisation, such as access to developed health care systems, education and advanced technologies, the lifestyle change also impose negative health behaviour that ultimately lead to positive energy balance and therefore obesity \citep{Malik2013}.
+
+One of the obvious factors that is linked to urbanisation is the lack of physical activity.
+Currently, the recommended amount of physical activity to keep an individual fit and healthy is $\geq$ 150 min of moderate-intensity aerobic physical activity per week, or $\geq$ 30 min of physical activity on most days of the week \citep{Pate1995, WHO2010}.
+
+sociocultural norm
+
 \citep{Malik2013}\\
@@
 The third factor is the income and socioeconomic status of the countries.
+As mentioned briefly in the previous point, access to developed health care system will allow weight maintenance of the population, but most likely be of most benefit for the high-income group.
 \textit{(Expand on this -- rise in income, availability of health care, etc.)}

### 2016-10-30 thesis/intro/intro.tex
tmp commit

@@
 \subsubsection{Individual-level (micro-level) factors}
-\label{ssub:Individual-level (micro-level) factors}
+\label{ssub:Individual-level_(micro-level)_factors}
 
@@
 \subsubsection{Population-level (macro-level) factors}
-\label{ssub:Population-level (macro-level) factors}
+\label{ssub:Population-level_(macro-level)_factors}
 
@@
 
-One of the obvious factors that is linked to urbanisation is the lack of physical activity.
+Lack of physical activity is one of the obvious components that contribute to obesity, which is also linked with the urbanisation of the population.
 Currently, the recommended amount of physical activity to keep an individual fit and healthy is $\geq$ 150 min of moderate-intensity aerobic physical activity per week, or $\geq$ 30 min of physical activity on most days of the week \citep{Pate1995, WHO2010}.
+However, in many urban cities, this recommended level of physical exercise is difficult to achieve and maintain, due to the shift towards the sedentary behaviours that are encouraged by the developed transport system, automated household chores, and indoor entertainment \citep{Malik2013}.
 
-sociocultural norm
-
-\citep{Malik2013}\\
+It is unfair for urbanisation to take all of the blame for the lack of exercise in the population, but it is true that urbanisation makes the bar higher for many people to take up the recommended level of exercise and maintain it.\\
 
 \noindent
-The third factor is the income and socioeconomic status of the countries.
-As mentioned briefly in the previous point, access to developed health care system will allow weight maintenance of the population, but most likely be of most benefit for the high-income group.
-\textit{(Expand on this -- rise in income, availability of health care, etc.)}
-\citep{Malik2013}\\
+The third factor is the income and socioeconomic status of the population.
+As the average income increases, more people are able to buy food and are likely to take up the sedentary lifestyle.
+Clearly, these behavioural changes will increase the likelihood of becoming obese.
+
+From this, one would expect the high-income group to have the largest impact from obesity, but on the contrary to this expectation, the biggest impact from obesity is seen in the low- and middle-income groups of the population \citep{Malik2013}.
+As mentioned briefly earlier, access to the developed health care system will allow weight maintenance of the population, but this is likely to be of most benefit for the high-income group in the population.
+The reason for this is, quite simply, the cost associated with the visit to the health care system.
+and therefore high-income group are more likely to make a visit than the low or middle-income group of the population \citep{Malik2013}.
+
+So, for the low- and middle-income groups, the rise in average income in these groups allow the people to have access to food and activities that promotes sedentary bahaviours, but the limited access to the health care system, education and recreational activities that promote physical activites ultimately leads to positive energy balance and obesity \citep{Malik2013}.
+In contrast, the high-income group of the population have better access to the facilities that promote weight maintenance, thus have less of an impact from obesity compared with the low- and middle-income groups of the population.\\
 
 \noindent
-The last factor, which ties in strongly with the first key factor, is the choice of diet by the population.
+The last factor, which ties in strongly with the first key factor, is the choice of diet made by the population.
+The amount of low-quality food have increased significantly in many rapidly growing countries, as these high-energy, low-nutritional food are cheaper and more profitible than the natural produce that have higher nutritional value (citation).
+As a result, more people tend to purchase these low-quality food more often, and consume more energy than they are required.
+
+Many people eat responsibly and avoid these low-quality diet and prevent overconsumption of such food.
+However, the overconsumption of these food is not a big problem as such, but rather the overall quality of the food \citep{Malik2013}.
+These food are usually high in refined sugar and unhealthy fat, highly processed, contain very little nutritional value and low in fibre.
+Thus consuming these food, especially food that are high in saturated fats,
+
+The abundance of food that are high in energy and low in nutritional value have increased significantly.
+In contrast to this, the amount of  natural produces have decreased.
+
 \textit{(Expand on this -- increased variety of food, low quality and high energy food choice, etc.)}
@@
 \citep{Malik2013}
+\\
 

### 2016-10-31 thesis/intro/intro.tex
tmp commit

@@
 The first key factor that contirbutes to the global rise in obese population is the availability and abundance of certain food in the country.
-With the help from government trade policies, food and other goods have become easier to trade between the country, helping the economic growth of many countries around the world (citation -- Kearney?).
+With the help from government trade policies, food and other goods have become easier to trade between the country, helping the economic growth of many countries around the world \citep{Kearney2010}.
 This led to greater abundance of food in general, greater number of choices of food and shifted the overall nutritional status of the country \citep{Malik2013}.
@@
 Moreover, corn and soy are also the main food for the livestock, which results in lower prices for meat \citep{Malik2013}.
-On the othere hand, the prices of fruits and vegetables remain expensive due to the lack of support by the government to lower the cost associated with the production (more citation...) \citep{Malik2013}.
+On the othere hand, the prices of fruits and vegetables remain expensive due to the lack of support by the government to lower the cost associated with the production \citep{Malik2013}.
 
@@
 
-It is unfair for urbanisation to take all of the blame for the lack of exercise in the population, but it is true that urbanisation makes the bar higher for many people to take up the recommended level of exercise and maintain it.\\
+Urbanisation by itself cannot be accounted for the lack of exercise in the population, but it is true that urbanisation makes the bar higher for many people to take up the recommended level of fitness and maintain it.
+\\
 
@@
 The last factor, which ties in strongly with the first key factor, is the choice of diet made by the population.
-The amount of low-quality food have increased significantly in many rapidly growing countries, as these high-energy, low-nutritional food are cheaper and more profitible than the natural produce that have higher nutritional value (citation).
+The amount of low-quality food have increased significantly in many rapidly growing countries, as these high-energy, low-nutritional food are cheaper and more profitible than the natural produce that have higher nutritional value \citep{Kearney2010}.
 As a result, more people tend to purchase these low-quality food more often, and consume more energy than they are required.
 
-Many people eat responsibly and avoid these low-quality diet and prevent overconsumption of such food.
-However, the overconsumption of these food is not a big problem as such, but rather the overall quality of the food \citep{Malik2013}.
-These food are usually high in refined sugar and unhealthy fat, highly processed, contain very little nutritional value and low in fibre.
-Thus consuming these food, especially food that are high in saturated fats,
+However, many people eat responsibly and avoid these low-quality diet and prevent overconsumption of such food, but the overconsumption of these food is not a big problem as such, rather the overall quality of the food \citep{Malik2013}.
+These food are usually high in refined sugar and unhealthy fat, highly processed, contain very little nutritional value and are low in fibre.
+As a result, consumption of these food increases the overall energy intake and achieve no nutritional benefits from the food.
 
-The abundance of food that are high in energy and low in nutritional value have increased significantly.
-In contrast to this, the amount of  natural produces have decreased.
+Whether it is by choice or not, consumption of these low-quality diet leads to obesity and obesity-related diseases. \\
 
-\textit{(Expand on this -- increased variety of food, low quality and high energy food choice, etc.)}
-
-\citep{Malik2013}
-\\
+% All of these population-level, factors together with the individual-level factors, contribute to the obesity status of the population.
 
@@
 
-As mentioned at the beginning of the current section, obesity has a devastating effect on the health of the population globally.
+As mentioned at the beginning of the current section, obesity has a devastating effect on the health of the population in a global scale.
 

### 2016-11-03 thesis/intro/intro.tex
tmp commit

@@
 Obesity has been one of the major global problems for more than a decade, associated with many noncommunicable diseases such as diabetes, cardiovascular diseases and certain types of cancers \citep{WHO2014}.
-In fact, the risk of comorbidities increases with the increase in \gls{bmi}, where the risk becomes severe as the \gls{bmi} level approaches the obese category (citation from 2000 WHO).
+In fact, the risk of comorbidities increases with the increase in \gls{bmi}, where the risk becomes severe as the \gls{bmi} level approaches the obese category \citep{WHO2000}.
 
 The number of overweight and obese population, in both adults and children, have risen in every country of the world, and the trend continues as we speak.
-Estimated to account for 3.4 million deaths per year and 93.6 million \glspl{daly} in 2010, obesity is a serious disease that continues to grow in our society (citation from 2010 WHO report needed).
+Estimated to account for 3.4 million deaths per year and 93.6 million \glspl{daly} in 2010, obesity is a serious disease that continues to grow in our society \citep{Lim2012}.
 
@@
 
-Obesity is defined as an abnormal or excessive fat accumulation that may impair health of that individual (citation).
+Obesity is defined as an abnormal or excessive fat accumulation that may impair health of that individual \citep{Garrow1988}.
 
@@
 
-Obesity is caused by excessive intake of energy without expending the energy gained.
+Obesity is caused by continuous intake of excessive energy without expending the energy gained.
 As one can suspect, there are many factors involved with the likelihood of an individual to become overweight and obese.
@@
 
+There are many environmental factors associated with the global rise in obesity (discussed in \cref{ssub:Population-level_(macro-level)_factors}: \nameref{ssub:Population-level_(macro-level)_factors}).
+Even though these environmental factors are probably the most significant contributor of global obesity epidemic, there are some individual-level, genetic factors that assist in the overall increase in the obese population.
+
 There are many genetic evidence of obesity being caused by individual-level factors that makes some individuals more predisposed (right word??) to obesity than others (citation).
+
+\citep{Neel1962}
+
+LEP, LEPR, POMC, CART, MC4R, etc and its associated genetic/biological mechanisms\\
+
+\citep{Montague1997}
+\citep{Clement1998}
+\citep{Jackson1997}
+\citep{Krude1998}
+\citep{Farooqi2003}
+\citep{Kublaoui2006}
+\citep{Dubern2001}
+\citep{Challis2002}
+
+FTO and other important GWAS studies \\
+
+\citep{Frayling2007}
+\citep{Dina2007}
+\citep{Scuteri2007}
+\citep{Gerken2007}
+
 \textit{Talk about individual-level factors that affect obesity here (need citations on individual factors of obesity; e.g. genetic variation).}
@@
 
+\citep{Calle2003}
+
 \subsection{Mechanism of cancer progression in obese patient}

### 2016-11-03 thesis/intro/intro.tex
tmp commit

@@
 
-\subsection{Causes of Obesity}
-\label{sub:causes_of_obesity}
+\subsection{Physiological mechanism of obesity}
+\label{sub:physiological_mechanism_of_obesity}
 
-Obesity is caused by continuous intake of excessive energy without expending the energy gained.
+Obesity is caused by continuous intake of food, whether they are controlled or not, without expending all of the energy gained.
+The key to the problem is the control of food intake and energy expenditure, and how they are regulated in your body.
+
+\subsubsection{Role of hypothalamus and arcuate nucleus in food intake control}
+\label{ssub:role_of_hypothalamus_and_arcuate_nucleus_in_food_intake_control}
+
+The central role of hypothalamus in the context of obesity is the control of satiety, and therefore food intake.
+The hypothalamus receives both nueronal and hormonal inputs from the body and propagates the signal to the downstream neurons that ultimately affect the food intake of the individual \citep{Bell2005, Spiegelman2001}.
+
+Within the hypothalamus, there is a group of neurons located within the arcuate nucleus that is essential for satiety signal transmission.
+There are two types of neurons in the arcuate nucleus: orexogenic and anorexogenic neurons.
+Orexogenic neurons are responsible for the promotion of food intake and reducing energy expenditure, whereas the anorexogenic neurons have the opposite effect \citep{Barsh2002}.
+
+The orexogenic neurons produce \gls{npy} and \gls{agrp} which induces orexogenic signal to the downstream neurons; anorexogenic neurons produce \gls{pomc} and \gls{cart} which induces anorexogenic signal to the downstream neurons \citep{Barsh2002, Spiegelman2001}.
+When stimulated by, for example, endocrine signals, these two different types of neurons act on the downstream neurons to promote or reduce food intake and energy expenditure.
+
+\subsubsection{Role of hormones on arcuate nucleus neurons}
+\label{ssub:role_of_hormones_on_arcuate_nucleus_neurons}
+
+\subsection{Factors that affect the prevalence of obesity}
+\label{sub:factors_that_affect_the_prevalence_of_obesity}
+
 As one can suspect, there are many factors involved with the likelihood of an individual to become overweight and obese.
@@
 There are many environmental factors associated with the global rise in obesity (discussed in \cref{ssub:Population-level_(macro-level)_factors}: \nameref{ssub:Population-level_(macro-level)_factors}).
-Even though these environmental factors are probably the most significant contributor of global obesity epidemic, there are some individual-level, genetic factors that assist in the overall increase in the obese population.
+Even though these environmental factors are probably the most significant contributor of the current global obesity epidemic, there are some individual-level, genetic factors that assist in the overall increase in the obese population.
 
 There are many genetic evidence of obesity being caused by individual-level factors that makes some individuals more predisposed (right word??) to obesity than others (citation).
-
-\citep{Neel1962}
 
@@
 \citep{Gerken2007}
+
+Talk about all the mutations and contribution of these genes to obesity, and then link it back to ``thrifty'' gene
+\citep{Neel1962}
 

### 2016-11-04 thesis/intro/intro.tex
tmp commit

@@
 
-\subsubsection{Role of hormones on arcuate nucleus neurons}
-\label{ssub:role_of_hormones_on_arcuate_nucleus_neurons}
+\subsubsection{Leptin-melanocortin pathway}
+\label{ssub:leptin_melanocortin_pathway}
+
+Leptin-melanocortin pathway is the pathway in which the level of satiety is controlled and regulated.
+In this pathway, both the orexogenic and anorexogenic neurons have crucial roles in controlling the satiety of the individual, and therefore their eating behaviour.
+
+When a satiety-controlling hormone, such as leptin, reaches the arcuate nucleus region of the hypothalamus, it can stimulate or inhibit orexogenic, anorexogenic, or both neurons.
+Depending on which neurons are stimulated or inhibited, the hormone will ultimately alter the satiety level of the individual.
+
+Firstly, when an orexogenic \gls{npy}/\gls{agrp} neurons are stimulated,
+
+\subsubsection{Role of hormones in leptin-melanocortin pathway}
+\label{ssub:role_of_hormones_in_leptin_melanocortin_pathway}
+
+There are many hormones in your body that regulates various physiological aspects of your body.
+Out of the many, four hormones have an important role in regulating the satiety of an individual: ghrelin, leptin, insulin, and \gls{pyy}.
+
+Ghrelin is a hormone produced mainly in the stomach that has a role in meal initiation and food intake \citep{Kohno2003}.
+
+Leptin, on the otherhand, is a hormone that has an opposite effect as ghrelin, where the hormone signals satiety and reduces food intake.
+
+Insulin, perhaps the most well-known hormone for its role in glucose homeostasis, also has an important role in satiety and food intake regulation.
+
+The last hormone is the \gls{pyy}, a peptide hormone released by the distal gastrointestinal tract that produces an anorexogenic signal.
+\citep{Batterham2002}
 

### 2016-11-08 thesis/intro/intro.tex
tmp commit

@@
 
-When a satiety-controlling hormone, such as leptin, reaches the arcuate nucleus region of the hypothalamus, it can stimulate or inhibit orexogenic, anorexogenic, or both neurons.
+When a satiety-controlling hormone, such as leptin, reaches the arcuate nucleus region of the hypothalamus, it can stimulate or inhibit orexogenic, anorexogenic, or both orexogenic and anorexogenic neurons.
 Depending on which neurons are stimulated or inhibited, the hormone will ultimately alter the satiety level of the individual.
 
-Firstly, when an orexogenic \gls{npy}/\gls{agrp} neurons are stimulated,
+Firstly, when an orexogenic \gls{npy}/\gls{agrp} neurons are stimulated, both \gls{npy} and \gls{agrp} are produced in the cell.
+Both \gls{npy} and \gls{agrp} are neuropeptides that sends a signal to increase food intake and decrease energy expenditure.
+\gls{npy} was first discovered by \citet{Tatemoto1982}, and later associated with altered eating behaviour by \citet{Clark1984}.
+Agouti protein is a protein normally expressed in the skin where it affects the pigmentation, and was first observed over 100 years ago by \citet{Castle1910} in mice that had yellow, or agouti-coloured coat.
+
+It was later found the ubiquitous expression of the protein was caused obesity (citation)
+
+The \gls{npy} and \gls{agrp} neuropeptides are recognised by the downstream neurons via the \gls{nyr} and \gls{mcr}.
 

### 2016-11-09 thesis/intro/intro.tex
tmp commit

--- 2016-11-04 thesis/intro/intro.tex
@@
 
-When a satiety-controlling hormone, such as leptin, reaches the arcuate nucleus region of the hypothalamus, it can stimulate or inhibit orexogenic, anorexogenic, or both neurons.
-Depending on which neurons are stimulated or inhibited, the hormone will ultimately alter the satiety level of the individual.
+When a satiety-controlling hormone such as leptin reaches the arcuate nucleus region of the hypothalamus, it can stimulate or inhibit orexogenic, anorexogenic, or both orexogenic and anorexogenic neurons.
+Depending on which neurons are stimulated or inhibited, the hormone will ultimately alter the satiety level of the individual.\\
 
-Firstly, when an orexogenic \gls{npy}/\gls{agrp} neurons are stimulated,
+\noindent
+Both \gls{pomc} and \gls{cart} are strong anorexogenic neuropeptides that stimulate the downstream effector neurons via the \gls{mc4r} pathway \citep{Barsh2002}.
+
+\gls{pomc} is a precursor protein that is cleaved to produce adrenocorticotropin (ACTH), $\alpha$-, $\beta$- and $\gamma$-\gls{msh}, and $\beta$-endorphin (citation Smith1988, endocr. rev.).
+Perhaps the most important, or most relevant, in the context of obesity and satiety control is the $\alpha$-MSH.
+$\alpha$-MSH is the main ligand that binds to the \gls{mc4r} expressed in the brain, and sends anorexogenic signals to the downstream neurons through this pathway \citep{Krude1998}.
+
+Similarly, \gls{cart} is also a neuropeptide that has a strong anorexogenic effect discovered by (citation Kristensen1998 on CART).
+Unlike $\alpha$-MSH, the receptor for \gls{cart} has not been identified yet, and the exact mechanism in which \gls{cart} acts is unclear (citation... Kristensen1998, Rogge2008).
+
+Since the expression of both neuropeptides are under the control of leptin, and that $\alpha$-MSH is known to stimulate the melanocortin pathway to produce an anorexogenic signal to the downstream neurons, this satiety-controlling signalling pathway is known as the leptin-melanocortin pathway.
+\\
+
+\noindent
+In contrast to the \gls{pomc}/\gls{cart} neurons, orexogenic neuropeptides \gls{npy} and \gls{agrp} are produced in the \gls{npy}/\gls{agrp} neurons when stimulated by hormones that control satiety.
+
+\gls{npy} was first discovered by \citet{Tatemoto1982}, and later associated with altered eating behaviour in rats by \citet{Clark1984}.
+\gls{npy} is recognised by the \gls{nyr}, a G protein-coupled receptor expressed on the downstream effector neurons, which relays the orexogenic signal from the \gls{npy}/\gls{agrp} neurons (citation, Herzog 1992 PNAS) \citep{Barsh2002, Bell2005}
+
+Agouti protein is a protein normally expressed in the skin where it affects the pigmentation, and was first identified over 100 years ago by \citet{Castle1910} in mice that had yellow, or agouti-coloured coat.
+Although the fact that ubiquitous expression of agouti protein caused obesity in mice was known for a long time, the biochemical mechanism that caused such phenotype was not elucidated until 1997 by \citet{Ollmann1997}.
+In brief, \gls{agrp} was shown to be a strong antagonist of the \gls{mc4r}, expressed in the downstream neurons to the \gls{npy}/\gls{agrp} neurons, and thus inhibited the anorexogenic signalling through the melanocortin pathway \citep{Barsh2002,Ollmann1997}.\\
+
+\noindent
+Together, \gls{npy}/\gls{agrp} and \gls{pomc}/\gls{cart} neurons produce contrasting signals that affect the overall satiety of the individual.
+Depending on how strong or weak these signals are, the individual may be more or less susceptible to obesity.
 

### 2016-11-09 thesis/intro/intro.tex
tmp commit

@@
 
-\gls{pomc} is a precursor protein that is cleaved to produce adrenocorticotropin (ACTH), $\alpha$-, $\beta$- and $\gamma$-\gls{msh}, and $\beta$-endorphin (citation Smith1988, endocr. rev.).
+\gls{pomc} is a precursor protein that is cleaved to produce \gls{acth}, $\alpha$-, $\beta$- and $\gamma$-\gls{msh}, and $\beta$-endorphin (citation Smith1988, endocr. rev.).
 Perhaps the most important, or most relevant, in the context of obesity and satiety control is the $\alpha$-MSH.
@@
 There are many hormones in your body that regulates various physiological aspects of your body.
-Out of the many, four hormones have an important role in regulating the satiety of an individual: ghrelin, leptin, insulin, and \gls{pyy}.
+Out of the many, four hormones in particular have an important role in regulating the satiety of an individual: ghrelin, leptin, insulin, and \gls{pyy}.\\
 
-Ghrelin is a hormone produced mainly in the stomach that has a role in meal initiation and food intake \citep{Kohno2003}.
+\noindent
+Ghrelin, originally identified as a peptide that stimulated the release of growth hormone from the pituitary, is a hormone produced mainly in the stomach that has a significant role in meal initiation and food intake, as well as energy homeostasis \citep{Kohno2003} (citation, nakazato2001 nature, kojima1999 nature).
+Ghrelin is known to be a ligand for the \gls{ghsr} which is expressed abundantly on the \gls{npy}/\gls{agrp} neurons, and increases the Ca$^{2+}$ concentration in these neurons \citep{Kohno2003}.
 
+Furthermore, (citation Seaone2003) showed that ghrelin increased the mRNA levels of both \gls{npy} and \gls{agrp} in these neurons and hence their protein expression, indicating the role of ghrelin as the orexogenic hormone that affects the overall outcome of the satiety signal.
+\\
+
+\noindent
 Leptin, on the otherhand, is a hormone that has an opposite effect as ghrelin, where the hormone signals satiety and reduces food intake.
+First identified by \citet{Zhang1994}, the dicovery of the leptin gene, followed by the discovery of leptin receptor gene by \citet{Tartaglia1995}, were probably the major breakthrough in the field of obesity research.
+\\
 
+\noindent
 Insulin, perhaps the most well-known hormone for its role in glucose homeostasis, also has an important role in satiety and food intake regulation.
+\\
 
+\noindent
 The last hormone is the \gls{pyy}, a peptide hormone released by the distal gastrointestinal tract that produces an anorexogenic signal.

### 2016-11-10 thesis/intro/intro.tex
tmp commit

@@
 
-\gls{pomc} is a precursor protein that is cleaved to produce \gls{acth}, $\alpha$-, $\beta$- and $\gamma$-\gls{msh}, and $\beta$-endorphin (citation Smith1988, endocr. rev.).
+\gls{pomc} is a precursor protein that is cleaved to produce \gls{acth}, $\alpha$-, $\beta$- and $\gamma$-\gls{msh}, and $\beta$-endorphin \citep{Smith1988}.
 Perhaps the most important, or most relevant, in the context of obesity and satiety control is the $\alpha$-MSH.
@@
 
-Similarly, \gls{cart} is also a neuropeptide that has a strong anorexogenic effect discovered by (citation Kristensen1998 on CART).
-Unlike $\alpha$-MSH, the receptor for \gls{cart} has not been identified yet, and the exact mechanism in which \gls{cart} acts is unclear (citation... Kristensen1998, Rogge2008).
+Similarly, \gls{cart} is also a neuropeptide that has a strong anorexogenic effect discovered by \citet{Kristensen1998}.
+Unlike $\alpha$-MSH, the receptor for \gls{cart} has not been identified yet, and the exact mechanism in which \gls{cart} acts is unclear \citep{Kristensen1998, Rogge2008}.
 
@@
 \gls{npy} was first discovered by \citet{Tatemoto1982}, and later associated with altered eating behaviour in rats by \citet{Clark1984}.
-\gls{npy} is recognised by the \gls{nyr}, a G protein-coupled receptor expressed on the downstream effector neurons, which relays the orexogenic signal from the \gls{npy}/\gls{agrp} neurons (citation, Herzog 1992 PNAS) \citep{Barsh2002, Bell2005}
+\gls{npy} is recognised by the \gls{nyr}, a G protein-coupled receptor expressed on the downstream effector neurons, which relays the orexogenic signal from the \gls{npy}/\gls{agrp} neurons \citep{Barsh2002, Bell2005, Herzog1992}
 
@@
 \noindent
-Ghrelin, originally identified as a peptide that stimulated the release of growth hormone from the pituitary, is a hormone produced mainly in the stomach that has a significant role in meal initiation and food intake, as well as energy homeostasis \citep{Kohno2003} (citation, nakazato2001 nature, kojima1999 nature).
+Ghrelin, originally identified as a peptide that stimulated the release of growth hormone from the pituitary, is a hormone produced mainly in the stomach that has a significant role in meal initiation and food intake, as well as energy homeostasis \citep{Kohno2003, Kojima1999, Nakazato2001}.
 Ghrelin is known to be a ligand for the \gls{ghsr} which is expressed abundantly on the \gls{npy}/\gls{agrp} neurons, and increases the Ca$^{2+}$ concentration in these neurons \citep{Kohno2003}.
 
-Furthermore, (citation Seaone2003) showed that ghrelin increased the mRNA levels of both \gls{npy} and \gls{agrp} in these neurons and hence their protein expression, indicating the role of ghrelin as the orexogenic hormone that affects the overall outcome of the satiety signal.
+Furthermore, \citet{Seoane2003} showed that ghrelin increased the mRNA levels of both \gls{npy} and \gls{agrp} in these neurons and hence their protein expression, indicating the role of ghrelin as the orexogenic hormone that affects the overall outcome of the satiety signal.
 \\
@@
 \noindent
-Leptin, on the otherhand, is a hormone that has an opposite effect as ghrelin, where the hormone signals satiety and reduces food intake.
+Leptin, on the other hand, is a hormone that has an opposite effect as ghrelin, where the hormone signals satiety and reduces food intake.
 First identified by \citet{Zhang1994}, the dicovery of the leptin gene, followed by the discovery of leptin receptor gene by \citet{Tartaglia1995}, were probably the major breakthrough in the field of obesity research.
+
+Leptin is produced and secreted mainly by the white adipose tissue, and the circulating concentration of leptin in the body is proportional to body fat stores \citep{Barsh2002, Moustafa2013, Zhang1994}.
+Leptin receptor has many splice variants, but the long variant of the leptin receptor is highly expressed only in the hypothalamus \citep{Ghilardi1996, Lee1996}.
+In addition to this, leptin receptor is expressed on both the \gls{npy}/\gls{agrp} and \gls{pomc}/\gls{cart} neurons, and have different effect on the outcome of the satiety signal depending on which neuron leptin stimulates.
+
+In \gls{npy}/\gls{agrp} neuron, it has been shown to reduce the Ca$^{2+}$ concentration, antagonising the effect of ghrelin on this type of neurons \citep{Kohno2003}.
+Furthermore, binding of leptin to the leptin receptor on the \gls{npy}/\gls{agrp} neuron causes a decreased release of \gls{gaba}, which inhibits the \gls{pomc}/\gls{cart} neuron \citep{Cowley2001}.
+
+In \gls{pomc}/\gls{cart} neuron, binding of leptin to its recpetor causes a depolarisation of the neuron \citep{Cowley2001}.
+Together with the indirect effect from the \gls{npy}/\gls{agrp} neuron, leptin inhibits the \gls{pomc}/\gls{cart} neuron to release $\alpha$-\gls{msh}, and therefore sends a satiety signal to its downstream effector neurons.
 \\
@@
 Insulin, perhaps the most well-known hormone for its role in glucose homeostasis, also has an important role in satiety and food intake regulation.
+
+Briefly, when insulin binds to the insulin receptor, \gls{irs} protein family is recruited and phosphorylated by the insulin receptor's intrinsic tyrosine kinase activity \citep{Saltiel2002}.
+The phosphorylated \gls{irs} protein then activates \gls{pi3k}, which in turn activates its downstream targets, causing a rapid signal cascade through the cell \citep{Saltiel2002}.
+
+It was first noticed by \citet{Spanswick1997} that insulin showed a very similar 
 \\

### 2016-11-11 thesis/intro/intro.tex
finished physiological mechanism of obesity

@@
 In \gls{pomc}/\gls{cart} neuron, binding of leptin to its recpetor causes a depolarisation of the neuron \citep{Cowley2001}.
-Together with the indirect effect from the \gls{npy}/\gls{agrp} neuron, leptin inhibits the \gls{pomc}/\gls{cart} neuron to release $\alpha$-\gls{msh}, and therefore sends a satiety signal to its downstream effector neurons.
+Together with the indirect effect from the \gls{npy}/\gls{agrp} neuron, leptin promotes the \gls{pomc}/\gls{cart} neuron to release $\alpha$-\gls{msh}, and therefore sends a anorexogenic signal to its downstream effector neurons.
 \\
@@
 
-It was first noticed by \citet{Spanswick1997} that insulin showed a very similar 
+The evidence of insulin's involvement in the satiety control was not particularly clear; even though there were strong evidence of obesity with neuron-specific deletion of insulin receptor, and that the insulin level, like leptin, was proportional to the fat mass also suggested the role of insulin in satiety and body weight control (more citation from original paper??) \citep{Barsh2002}.
+This was perhaps due to the fact that leptin and insulin used two different signalling pathways in the cell: leptin used the \gls{jak}/\gls{stat} signalling pathway, whereas insulin used the \gls{pi3k} pathway for signalling \citep{Ghilardi1996}.
+This was puzzling to the researchers as there seemed to be no converging point of the two hromones in these pathways.
+
+However, leptin, and subsequently insulin, was shown to activate ATP-dep\-endent K$^+$ channel in glucose-responsive neurons in the hypothalamus \citep{Spanswick1997, Spanswick2000}.
+The activation of K$^+$ channel provided evidence for the two pathways to act similarly on the cell, at least for an acute response, suggesting a convergence in the pathway response by the two hormones, and therefore their role in satiety control.
 \\
@@
 The last hormone is the \gls{pyy}, a peptide hormone released by the distal gastrointestinal tract that produces an anorexogenic signal.
-\citep{Batterham2002}
+
+\gls{pyy} is a gut hormone that is secreted postprandially, and it is released at levels proportional to the calorie content of the meal (original citation, Adrian1985 gastroenterology).
+\gls{y2r} is an inhibitory presynaptic receptor that is highly expressed in the \gls{npy}/\gls{agrp} neuron, and is the target of the \gls{pyy} hormone \citep{Batterham2002}.
+\citet{Batterham2002} showed that \gls{pyy} decreased the expression of \gls{npy} in the \gls{npy}/\gls{agrp} neuron, and activated the \gls{pomc}/\gls{cart} neuron by membrane depolarisation.
+The two mechanisms of \gls{pyy} produces an anorexogenic signal and encourages the individual to stop eating.
 

### 2016-11-12 thesis/intro/intro.tex
tmp commit

@@
 \citet{Batterham2002} showed that \gls{pyy} decreased the expression of \gls{npy} in the \gls{npy}/\gls{agrp} neuron, and activated the \gls{pomc}/\gls{cart} neuron by membrane depolarisation.
-The two mechanisms of \gls{pyy} produces an anorexogenic signal and encourages the individual to stop eating.
+The two mechanisms of \gls{pyy} produces an anorexogenic signal and encourages the individual to stop eating.\\
+
+\noindent
+When any one of these neurons, neuropeptides, hormones or its corresponding receptors become defective, it can lead to some form of eating disorder.
+Therefore the roles of these neurons, neuropeptides, and hormones are crucial for healthy diet and body weight maintenance.
 
@@
 
-There are many genetic evidence of obesity being caused by individual-level factors that makes some individuals more predisposed (right word??) to obesity than others (citation).
+There are many evidence of obesity being caused by individual-level genetic factors that make some individuals more predisposed to obesity than others (citation).
+Generally speaking, there are two classes of genetic-driven obesity: monogenic obesity and polygenic obesity \citep{Moustafa2013}. \\
+
+\noindent
+Monogenic obesity is where a single mutation in one of the genes that is directly involved in the obesity pathway causes severe obesity in the individual.
+Since these mutations are usually directly related to the key neurons/neuropeptides/hor\-mones that were mentioned earlier (\cref{sub:physiological_mechanism_of_obesity}), the resulting phenotype from the mutation causes severe early-onset obesity.
+However, although these mutations cause a devastating effect on the individual, these mutations occur very rarely in the population.
+With that said though, the identification of these mutations have shed light on the mechanisms in which obesity is developed and it is worth mentioning some of these mutations to highlight the importance of the components involved in the obesity pathway.
+
+The first mutation that should be mentioned is in the leptin gene.
+\\
 
@@
 \citep{Challis2002}
+
+\noindent
+Polygenic obesity is where 
+\\
 

### 2016-11-12 thesis/intro/intro.tex
tmp commit

@@
 The first mutation that should be mentioned is in the leptin gene.
+\citet{Montague1997} found a single nucleotide deletion mutation at position 133 of the leptin gene in children with early-onset obesity.
+They further showed that the mutation caused a frameshift mutation in the protein, resulting in an inactive form of leptin that was not able to be secreted \citep{Montague1997}.
+
+In close relation to this mutation is the mutation in the leptin receptor gene.
+\citet{Clement1998} have discovered a single nucleotide substitution mutation that led to the leptin receptor transcript that skipped exon 16 -- leading to a form of the receptor with no transmembrane or intracellular domain.
+Individuals with this rare mutation showed extreme hyperphagia and symptoms of early-onset obesity, even though their circulating leptin levels were present abundantly \citep{Clement1998}.
+
+\citep{Jackson1997}
+\citep{Krude1998}
 \\
 
-LEP, LEPR, POMC, CART, MC4R, etc and its associated genetic/biological mechanisms\\
-
-\citep{Montague1997}
-\citep{Clement1998}
-\citep{Jackson1997}
-\citep{Krude1998}
 \citep{Farooqi2003}
@@
 \citep{Challis2002}
+\\
 

### 2016-11-13 thesis/intro/intro.tex
tmp commit

@@
 \noindent
-When any one of these neurons, neuropeptides, hormones or its corresponding receptors become defective, it can lead to some form of eating disorder.
+When any one of these neurons, neuropeptides, hormones or its corresponding receptors become defective, it can lead to some form of eating disorder (briefly covered in \cref{ssub:Individual-level_(micro-level)_factors}: \nameref{ssub:Individual-level_(micro-level)_factors}).
 Therefore the roles of these neurons, neuropeptides, and hormones are crucial for healthy diet and body weight maintenance.
@@
 In close relation to this mutation is the mutation in the leptin receptor gene.
-\citet{Clement1998} have discovered a single nucleotide substitution mutation that led to the leptin receptor transcript that skipped exon 16 -- leading to a form of the receptor with no transmembrane or intracellular domain.
+\citet{Clement1998} have discovered a single nucleotide substitution mutation that led to the leptin receptor transcript that skipped exon 16, leading to a form of the receptor with no transmembrane or intracellular domain.
 Individuals with this rare mutation showed extreme hyperphagia and symptoms of early-onset obesity, even though their circulating leptin levels were present abundantly \citep{Clement1998}.
 
-\citep{Jackson1997}
-\citep{Krude1998}
+The next couple of mutations are associated with the synthesis and the processing of \gls{pomc}.
+\citet{Krude1998} found single nucleotide mutations in the exonic region of the \gls{pomc} gene that affected the transcription of the $\alpha$-\gls{msh} region (and some other regions) of \gls{pomc}.
+\citet{Jackson1997} identified a mutation in the \gls{pc1}, an enzyme responsible for the processing of many prohormones and neuropeptides, including \gls{pomc}.
+In either case, the mutation resulted in the lack of $\alpha$-\gls{msh} in the neurons for satiety signalling and caused severe early-onset obesity.
+
+As mentioned earlier, even though these mutations are very rare and affects only those individuals that have one of these mutations, it had a profound impact in the field of obesity research and contributed significantly to the knowledge of the mechanism of how obesity develops.
 \\
 
-\citep{Farooqi2003}
-\citep{Kublaoui2006}
-\citep{Dubern2001}
-\citep{Challis2002}
-\\
+% \citep{Farooqi2003}
+% \citep{Kublaoui2006}
+% \citep{Dubern2001}
+% \citep{Challis2002}
+% \\
 
 \noindent
-Polygenic obesity is where 
+Polygenic obesity is where a combination of variants in multiple genes or multiple regions of the genome, together with environmental factors, causes obesity in the individual \citep{Moustafa2013}.
+Unlike in monogenic obesity where a single mutation in a gene causes severe obesity, the variants in some of the genes or genomic regions involved in polygenic obesity have little effect on the individual's obesity status by itself, but in combination with many other variants, together with environmental factors, it has a serious effect on the individual.
+
+Known as the ``thrifty gene'' hypothesis, \citet{Neel1962} suggested that these genes that were important in obesity have been around in the population for many generations, as it would have had a selective advantage in an environment where the population experienced starvation more frequently than the current generation.
+However, as food became more accessible, a subset of the population who were particularly good at storing fat and survive the starvation period ``overreacted'' to this rapid change in the environment \citep{Bell2005, Spiegelman2001}.
+These observations in the rise in obesity in subsets of population cannot be explained simply by external factors alone, and there is a significant genetic component involved \citep{Bell2005}.
 \\
@@
 \citep{Gerken2007}
-
-Talk about all the mutations and contribution of these genes to obesity, and then link it back to ``thrifty'' gene
-\citep{Neel1962}
 

### 2016-11-14 thesis/intro/intro.tex
tmp commit

@@
 
-The number of overweight and obese population, in both adults and children, have risen in every country of the world, and the trend continues as we speak.
+The number of overweight and obese people, in both adults and children, have risen in every country of the world, and the trend continues as we speak.
 Estimated to account for 3.4 million deaths per year and 93.6 million \glspl{daly} in 2010, obesity is a serious disease that continues to grow in our society \citep{Lim2012}.
@@
 
-Although the use of \gls{bmi} as an accurate measure of body fat composition and/or representation of the individual's metabolic status remains controversial, many studies have used \gls{bmi} due to the ease of data collection compared to other measurements, such as [example] (citation needed).
-Consequently, most of the publicly available clinical data have height and weight data from which one can calculate the patients' \gls{bmi}, and therefore able to collect the patients' obesity status more conveniently than, for example, [example].
-
-possible papers:\citep{Gelber2008, Lee2008, Yusuf2005}
+Although the use of \gls{bmi} as an accurate measure of body fat composition and/or representation of the individual's metabolic status remains controversial, many studies have used \gls{bmi} due to the ease of data collection compared to other measurements, such as waist-to-height ratio \citep{Lee2008, Gelber2008}.
+Consequently, most of the publicly available clinical data have height and weight data from which one can calculate the patients' \gls{bmi}, and therefore able to collect the patients' obesity status more conveniently than, for example, waist-to-height ratio.
 
@@
 
-The evidence of insulin's involvement in the satiety control was not particularly clear; even though there were strong evidence of obesity with neuron-specific deletion of insulin receptor, and that the insulin level, like leptin, was proportional to the fat mass also suggested the role of insulin in satiety and body weight control (more citation from original paper??) \citep{Barsh2002}.
+The evidence of insulin's involvement in the satiety control was not particularly clear; even though there were strong evidence of obesity with neuron-specific deletion of insulin receptor, and that the insulin level, like leptin, was proportional to the fat mass also suggested the role of insulin in satiety and body weight control \citep{Barsh2002, Bruning2000, Woods1979}.
 This was perhaps due to the fact that leptin and insulin used two different signalling pathways in the cell: leptin used the \gls{jak}/\gls{stat} signalling pathway, whereas insulin used the \gls{pi3k} pathway for signalling \citep{Ghilardi1996}.
@@
 
-\gls{pyy} is a gut hormone that is secreted postprandially, and it is released at levels proportional to the calorie content of the meal (original citation, Adrian1985 gastroenterology).
+\gls{pyy} is a gut hormone that is secreted postprandially, and it is released at levels proportional to the calorie content of the meal \citep{Adrian1985}.
 \gls{y2r} is an inhibitory presynaptic receptor that is highly expressed in the \gls{npy}/\gls{agrp} neuron, and is the target of the \gls{pyy} hormone \citep{Batterham2002}.
@@
 There are many environmental factors associated with the global rise in obesity (discussed in \cref{ssub:Population-level_(macro-level)_factors}: \nameref{ssub:Population-level_(macro-level)_factors}).
-Even though these environmental factors are probably the most significant contributor of the current global obesity epidemic, there are some individual-level, genetic factors that assist in the overall increase in the obese population.
+Even though these environmental factors are probably the most significant contributor of the current global obesity epidemic, there are some individual-level, genetic factors that assist in the overall increase in the proportion of the population that are obese.
 
-There are many evidence of obesity being caused by individual-level genetic factors that make some individuals more predisposed to obesity than others (citation).
-Generally speaking, there are two classes of genetic-driven obesity: monogenic obesity and polygenic obesity \citep{Moustafa2013}. \\
+There are many evidence of obesity being caused by individual-level genetic factors that make some individuals more predisposed to obesity than others.
+Generally speaking, there are two classes of genetically-driven obesity: monogenic obesity and polygenic obesity \citep{Moustafa2013}. \\
 
@@
 
-Known as the ``thrifty gene'' hypothesis, \citet{Neel1962} suggested that these genes that were important in obesity have been around in the population for many generations, as it would have had a selective advantage in an environment where the population experienced starvation more frequently than the current generation.
+Known as the ``thrifty gene'' hypothesis, \citet{Neel1962} suggested that the genes that were important in obesity have been around in the population for many generations, as it would have had a selective advantage in an environment where the population experienced starvation more frequently than the current generation.
 However, as food became more accessible, a subset of the population who were particularly good at storing fat and survive the starvation period ``overreacted'' to this rapid change in the environment \citep{Bell2005, Spiegelman2001}.
-These observations in the rise in obesity in subsets of population cannot be explained simply by external factors alone, and there is a significant genetic component involved \citep{Bell2005}.
+These observations in the rise in obesity in subsets of population cannot be explained simply by external factors alone, and there is a significant genetic component involved in this phenomenon \citep{Bell2005}.
+
+There have been many studies that have investigated the association of genes and/or genomic regions with obesity.
+Of those, many of the \glspl{gwas} published in 2007 that showed strong evidence of the \gls{fto} gene with \gls{bmi} and obesity were of greatest significance \citep{Dina2007,Frayling2007,Gerken2007,Scuteri2007}.
+Until these studies came out, there has been no common genes or genomic regions that associated with obesity that was reproducible and reliable \citep{Frayling2007}.
+Thus, these studies were the first evidence of common variants that contributed to the polygenic obesity phenotype in the population.
+
+Discovery of the \gls{fto} gene has lead to extensive research on the gene, and the mechanism of action by the \gls{fto} variant has been recently uncovered  from these efforts \citep{Claussnitzer2015a,Smemo2014a}.
+The association between obesity and genetic variations, and their mechanisms that lead to obesity, has not been fully elucidated due to its complexity, but it may not be long before they are fully uncovered.
 \\
-
-FTO and other important GWAS studies \\
-
-\citep{Frayling2007}
-\citep{Dina2007}
-\citep{Scuteri2007}
-\citep{Gerken2007}
-
-\textit{Talk about individual-level factors that affect obesity here (need citations on individual factors of obesity; e.g. genetic variation).}
 
@@
 \noindent
-The first key factor that contirbutes to the global rise in obese population is the availability and abundance of certain food in the country.
+The first key factor that contirbutes to the global rise in the proportion of the population that is obese is the availability and abundance of certain food in the country.
 With the help from government trade policies, food and other goods have become easier to trade between the country, helping the economic growth of many countries around the world \citep{Kearney2010}.
@@
 As mentioned briefly earlier, access to the developed health care system will allow weight maintenance of the population, but this is likely to be of most benefit for the high-income group in the population.
-The reason for this is, quite simply, the cost associated with the visit to the health care system.
-and therefore high-income group are more likely to make a visit than the low or middle-income group of the population \citep{Malik2013}.
+The reason for this is, quite simply, the cost associated with the visit to the health care system, and therefore high-income group is more likely to make a consistent and periodic visit than the low or middle-income group of the population \citep{Malik2013}.
 
@@
 
-% All of these population-level, factors together with the individual-level factors, contribute to the obesity status of the population.
+\noindent
+All of these population-level environmental factors in synergy with the individual-level genetic factors have contributed to the global epidemic of obesity.
 
@@
 As mentioned at the beginning of the current section, obesity has a devastating effect on the health of the population in a global scale.
+
+posing threat to the growing population that are becoming more obese.
 

### 2016-11-14 thesis/misc/glossary.tex
tmp commit
finished physiological mechanism of obesity

@@
 \newacronym{go}{GO}{Gene Ontology}
+\newacronym{usa}{USA}{United States of America}
+\newacronym{npy}{NPY}{neuropeptide Y}
+\newacronym{agrp}{AGRP}{agouti-related protein}
+\newacronym{pomc}{POMC}{pro-opiomelanocortin}
+\newacronym{cart}{CART}{cocaine and amphetamine related transcript}
+\newacronym{mc1r}{MC1R}{melanocortin 1 receptor}
+\newacronym{mc4r}{MC4R}{melanocortin 4 receptor}
+\newacronym{msh}{MSH}{melanocyte-stimulating hormone}
+\newacronym{nyr}{Y1R}{neuropeptide Y1 receptor}
+\newacronym{nyr}{Y2R}{neuropeptide Y2 receptor}
+\newacronym{acth}{ACTH}{adrenocorticotropic hormone}
+\newacronym{ghsr}{GHSR}{growth hormone secretagogue receptor}
+\newacronym{gaba}{GABA}{$\gamma$-aminobutyric acid}
+\newacronym{irs}{IRS}{insulin receptor seubstrate}
+\newacronym{pi3k}{PI3K}{phosphatidylinositol 3-kinase}
+\newacronym{jak}{JAK}{janus kinase}
+\newacronym{pc1}{PC1}{prohormone convertase 1}
+\newacronym{fto}{FTO}{fat mass and obesity associated}
+\newacronym{fto}{FTO}{fat mass and obesity associated}
+\newacronym{snp}{SNP}{single nucleotide polymorphism}
+\newacronym[longplural={genome wide association studies},\glsshortpluralkey={GWAS}]{gwas}{GWAS}{genome wide association study}
+\newacronym{stat}{STAT}{signal transducer and activator of transcription}
+\newacronym{pyy}{PYY$_{3-36}$}{peptide YY$_{3-36}$}
 \newacronym[longplural={Disability-Adjusted Life Years}]{daly}{DALY}{Disability-Adjusted Life Year}

### 2016-11-15 thesis/intro/intro.tex
finished impact of obesity subsection

@@
 
-posing threat to the growing population that are becoming more obese.
+Overweight and obesity have been a risk factor for many diseases, including type 2 diabetes, cardiovascular diseases, hypertension, and many types of cancers \citep{Franks2010, Guh2009, Yach2006}.
+As \citet{Yach2006} mentions in their paper, ``overweight and obesity have become to diabetes what tobacco is to lung cancer'', and it is important to understand that obesity has significant risks and comorbidities associated with it.
 
-\citep{Malik2013, Franks2010}
+Although adulthood obesity is clinically important for risk assessment of diseases, childhood obesity has been focussed more heavily in recent years,  as it has greater public health implications.
+There are many immediate consequences that are related to childhood obesity, but the diseases that develop as these children grow up are more critical than the earlier effects.
+
+In the paper written by \citet{Must1999}, they reviewed many of the clinical studies done in the 1900s and summarised the risks associated with childhood obesity.
+Among the many risks associated with obesity, obese children, compared to lean children, had over 8.5-fold increased risk of developing hypertension as adults \citep{Must1999}.
+On top of this, the relative risks for all-cause mortality and coronary heart disease-specific mortality were 1.5 and 2.5, respectively \citep{Must1999}.
+
+In a more recent paper by \citep{Franks2010}, they have shown that childhood obesity (as well as glucose intolerance and hypertension in childhood) was strongly associated with premature deaths from endogenous causes in American Indian population from Arizona (incidence rate ratio of 1.4 per unit of \gls{bmi} z-score).\\
+
+\noindent
+Health related consequences of obesity are clear from these studies, but it is noteworthy that obesity also has an economical impact on the population as well.
+In the \gls{usa}, within the span of 5 years, the medical cost due to diabetes have risen from \$44 billion to \$92 billion \citep{Yach2006}.
+Although not all diabetes are directly caused by obesity, obesity would have assisted in the rise in the costs associated with diabetes.
+In addition to this, the costs that are associated with diabetes treatment affects the household incomes, increased premature retirement and unemployment which, over time, could have a significant economical impact on the society \citep{Yach2006}.\\
+
+\noindent
+As more of the younger generation is gaining weight and becoming obese, having the risks and disease burden associated with obesity at such early age are not ideal -- both economically and epidemiologically -- and poses a threat to the growing population that are becoming more obese.
 

### 2016-11-16 thesis/intro/intro.tex
finally finished the obesity section of intro

@@
 \noindent
-As more of the younger generation is gaining weight and becoming obese, having the risks and disease burden associated with obesity at such early age are not ideal -- both economically and epidemiologically -- and poses a threat to the growing population that are becoming more obese.
+As more of the younger generation is gaining weight and becoming obese, having the risks and disease burden associated with obesity at such early age are not ideal (both economically and epidemiologically) and poses a threat to the growing population that are becoming more obese.
 
@@
 
-\citep{Malik2013}
+It is apparent from previous sections that obesity has become a serious problem that the world needs to prioritise to solve immediately.
+Even though obesity has been regarded as one of the many global problems to solve, the progress has been slow.
+
+\citet{Sacks2009} proposed an \gls{opa} framework that was based on the framework developed by \citet{WHO2006}.
+In this framework, \citet{Sacks2009} have added extra features so that the policies were targeted to impact three levels of the underlying environment of obesity: the socio-ecological environment (upstream), lifestyle/behavioural changes (midstream), and health services support (downstream).
+
+The upstream approach targets the underlying economic, social and physical norms that the population is currently in.
+By targeting the broad socioeconomic environment that the population lives in, such as taxation, education, employment and housing \citep{Sacks2009}.
+In doing so, it is possible to have a significant change in the population behaviours and eventually improve the health outcomes of the population.
+
+The midstream approach targets the population or subpopulation behaviour change to improve the eating and physical activity behaviours in these groups.
+Unlike in the upstream approach, the midstream approach aims to change the behaviour by implementing social marketing and programmes that encourage such changes \citet{Sacks2009}.
+Again, the result of this approach will be an improvement in the overall positive behaviour of the population towards healthy eating and physical activity.
+
+The downstream approach targets to support the health services and clinical interventions.
+Unlike upstream and midstream approaches, downstream approach generally acts on an individual level, whereby the focus is on the management of the health of an individual and/or the individual's family \citet{Sacks2009}.
+Although not as efficient as upstream or midstream approaches, downstream approach will help improve the habits and behaviours to promote weight management.\\
+
+\noindent
+In the recent years, obesity has been declared as a global priority \citep{Malik2013}.
+Furthermore, comprehensive global monitoring framework, multi-sectoral actions and strengthening of the national policies have been implemented to better control the obesity epidemic \citep{Malik2013}.
+In many countries, the nutritional labelling on products have been adjusted to include the ingredients that have negative impact on obesity (for example, \textit{trans} fat), and even reduce or remove these ingredients from food \citep{Malik2013}.
+
+Food marketing can also be a problem in a highly competitive society that we live in.
+To sell their products that are high in fat, sugar and portion size, food industries market and advertise their products to the population, especially young children, in order to stay ahead in their market place \citep{Chan2010, Malik2013}.
+Even though there is a strong evidence of food marketing influencing the eating behaviours and food preferences in children, very few countries have been able to reduce these marketing and advertising strategies of unhealthy food \citep{Hawkes2007a, Malik2013}.
+
+In contrast to these upstream/midstream approaches, there are some clinical interventions and treatments that help manage weight in an individual.
+One of these therapies is bariatric surgery for obese patients, especially morbidly obese patients \citep{Buchwald2004}.
+Through their meta-analysis, \citet{Buchwald2004} showed that bariatric surgery was able to reduce many of the the comorbidities associated with obesity.
+The fact that these improvements and complete resolution of the comorbidities in high proportion of patients that underwent bariatric surgery shows that, even though it acts on an individual level, downstream approaches such as bariatric surgery may be a highly effective, but not an efficient, method of preventing obesity in the population.\\
+
+\noindent
+It is evident that the most efficient way to reduce the proportion of the population that are obese is to implement national policies that affect the upstream or midstream factors.
+These policies, although they are efficient at targeting the underlying problem, may take a long time to make a significant effect.
+On the otherhand, individual therapeutic plans that affect the downstream factors may be very effective in achieving the desired goal of weight loss, but it is very resource intensive and upscaling this approach would be practically impossible.
+
+The management of obesity at an early stage with midstream and/or downstream approaches while the upstream approach takes effect would be the key to success in controlling for the global obesity epidemic.
 

### 2016-11-16 thesis/intro/intro.tex
Tweaked the BiocRefStyle.bst so it prints web page references correctly

@@
 % What is cancer.
+\begin{quote}
+	\textit{
+	Cancer is a generic term for a large group of diseases that can affect any part of the body \ldots{}
+	One defining feature of cancer is the rapid creation of abnormal cells that grow beyond their usual boundaries, and which can then invade adjoining parts of the body and spread to other organs \ldots{}
+	referred to as metastasizing.
+	Metastases are the major cause of death from cancer.
+	\citep{WHO2016}
+	}
+\end{quote}
 
-\subsection{Hallmarks of Cancer}
+\section{Hallmarks of Cancer}
 \label{subsec:cancerhallmarks}
 
-% TODO: edit the starting sentence so it's not a copy from the beginning
-% It is now commonly understood that cancers are caused by dysregulation of various mechanisms or pathways that allow tumour cells to prolilferate, survive and migrate.
-% Many of these dysregulated pathways frequently occur in cancer cells, and these characteristic traits, commonly known as the ``Hallmarks of Cancer'', were first described by \citet{Hanahan2000}.
+It is now commonly understood that cancers are caused by dysregulation of various mechanisms or pathways that allow tumour cells to prolilferate, survive and migrate.
+Many of these dysregulated pathways frequently occur in cancer cells, and these characteristic traits, commonly known as the ``Hallmarks of Cancer'', were first described by \citet{Hanahan2000}.
 
-% These hallmarks were initially presented in 2000 to provide a common framework for scientists to better understand cancer biology, and also to direct the focus of future research.
-% After a decade of advances in cancer research, the ``Hallmarks of Cancer'' were revisited by Hanahan and Weinberg in 2011, and four more hallmarks were added on top of the six presented in the original paper in 2000 \citep{Hanahan2011}.
+These hallmarks were initially presented in 2000 to provide a common framework for scientists to better understand cancer biology, and also to direct the focus of future research.
+After a decade of advances in cancer research, the ``Hallmarks of Cancer'' were revisited by Hanahan and Weinberg in 2011, and four more hallmarks were added on top of the six presented in the original paper in 2000 \citep{Hanahan2011}.
 
-% Together, these ``Hallmarks of Cancer'' (fig 1) have provided a foundation of the general biological knowledge of cancer and its characteristics that allowed the researchers to better understand the underlying cancer biology.
+Together, these ``Hallmarks of Cancer'' (fig 1) have provided a foundation of the general biological knowledge of cancer and its characteristics that allowed the researchers to better understand the underlying cancer biology.
 
-% Although much reseach has been conducted on the ``Hallmarks of Cancer'', the molecular processes underlying these hallmarks have still not been fully explained \citep{Hanahan2011}.
-% This is due to the fact that cancer cells develop progressively under selective pressures, and are also affected by many environmental factors \citep{Gatza2011}, which makes it difficult to identify causal factors of cancer development.
-% Furthermore, these external factors affect how the cancer cells respond to the environment, and may also cause cancer progression and/or drug resistance.
+Although much reseach has been conducted on the ``Hallmarks of Cancer'', the molecular processes underlying these hallmarks have still not been fully explained \citep{Hanahan2011}.
+This is due to the fact that cancer cells develop progressively under selective pressures, and are also affected by many environmental factors \citep{Gatza2011}, which makes it difficult to identify causal factors of cancer development.
+Furthermore, these external factors affect how the cancer cells respond to the environment, and may also cause cancer progression and/or drug resistance.
 

### 2016-11-16 thesis/misc/abstract.tex
Tweaked the BiocRefStyle.bst so it prints web page references correctly

@@
 
-It is now commonly understood that cancers are caused by dysregulation of various mechanisms or pathways that allow tumour cells to prolilferate, survive and migrate.
-There have been many reports on the association between obesity and worsened cancer prognosis, suggesting that there may be distinct changes that underlie tumorigenesis and/or cancer progression that are specific to patients with higher \gls{bmi}.
+% It is now commonly understood that cancers are caused by dysregulation of various mechanisms or pathways that allow tumour cells to prolilferate, survive and migrate.
+% There have been many reports on the association between obesity and worsened cancer prognosis, suggesting that there may be distinct changes that underlie tumorigenesis and/or cancer progression that are specific to patients with higher \gls{bmi}.
 
-In brief, this research aims to determine whether gene expression signatures exist that are specific to obesity across multiple cancer types, and to investigate whether there are any common pathways being dysregulated in cancers based on these genetic signatures.
-Better understanding of the pathways being dysregulated in cancer cells in obese patients may lead to improved clinical decisions, and could thus contribute towards personalised treatment in the future.
+% In brief, this research aims to determine whether gene expression signatures exist that are specific to obesity across multiple cancer types, and to investigate whether there are any common pathways being dysregulated in cancers based on these genetic signatures.
+% Better understanding of the pathways being dysregulated in cancer cells in obese patients may lead to improved clinical decisions, and could thus contribute towards personalised treatment in the future.
 

### 2016-11-19 thesis/intro/intro.tex
tmp commit

@@
 \label{ch:intro}
-
-% TODO: also have a look at the masters talk write up from 2015
 
@@
 
-% General intro to obesity.
-% Implications of obesity -- disease associated with it and/or cancer.
 Obesity has been one of the major global problems for more than a decade, associated with many noncommunicable diseases such as diabetes, cardiovascular diseases and certain types of cancers \citep{WHO2014}.
@@
 \label{sub:definition_of_obesity}
-% TODO: add WHO website to the reference for the definition of obesity (or find proper articles)
 
@@
 The current obesity status of the adult population is bad already, but the fact that more of the younger population who are overweight and obese poses a much greater risk for the health of the population (discussed in \cref{sub:impact_of_obesity}).
+
+% TODO: Add western-specific and NZ specific statistics on obesity
+% TODO: Add male/female specific stats
+% TODO: Add Maori/PI specific stats
 
@@
 The association between obesity and genetic variations, and their mechanisms that lead to obesity, has not been fully elucidated due to its complexity, but it may not be long before they are fully uncovered.
-\\
 
@@
 	\textit{
-	Cancer is a generic term for a large group of diseases that can affect any part of the body \ldots{}
+	``Cancer is a generic term for a large group of diseases that can affect any part of the body \ldots{}
 	One defining feature of cancer is the rapid creation of abnormal cells that grow beyond their usual boundaries, and which can then invade adjoining parts of the body and spread to other organs \ldots{}
 	referred to as metastasizing.
-	Metastases are the major cause of death from cancer.
+	Metastases are the major cause of death from cancer.''
 	\citep{WHO2016}
@@
 
-\section{Hallmarks of Cancer}
-\label{subsec:cancerhallmarks}
+\subsection{Prevalence and mortality of cancer}
+\label{sub:prevalence_and_mortality_of_cancer}
+
+Cancer is considered as one of the leading causes of deaths around the world, accounting for 8.2 million deaths in 2012 \citep{WHO2014}.
+More than half of cancer deaths are caused by lung, breast, colorectal, stomach and liver cancers \citep{WHO2014}.
+
+The leading cause of deaths in high-income countries is lung cancer in both men and women, then colorectal cancer and breast cancer for men and women, respectively \citep{WHO2014}.
+Although the levels of mortality varies among different countries, cervical, liver and stomach cancers make up a large proportion of cancer deaths among low- and middle-income countries \citep{WHO2014}.\\
+
+\noindent
+\citet{Bray2013} have estimated the global prevalence of cancer in adult population in 2008.
+Since there is no clear definition for cancer prevalence, \citet{Bray2013} used \gls{ldp} as an estimate for cancer prevalence, measuring the incidence of cancer during the period between 2004 and 2008 (5 years) who were still alive at the end of 2008.
+
+From these estimates, Eastern Asia had the greatest 5-year prevalence of cancer with approximately 7 million people with cancer in 2008, followed by North America and Western Europe with approximately 4.7 and 3 million people with cancer \citep{Bray2013}.
+When population of these regions were taken into account and the proportion of cancer prevalence calculated, Western Europe had the highest prevalence proportion of 1.9\%, followed closely by Australia/New Zealand with 1.82\% and North America with 1.7\% \citep{Bray2013}.
+
+Looking at the prevalence of cancer by specific types, breast cancer was the most prevalent type of cancer with 5.2 million cases in 2008 \citep{Bray2013}.
+Colorectal and prostate cancers were second and third most prevalent cancer types, with approximately 3.2 million cases for  both of them \citep{Bray2013}.
+\citet{Bray2013} noted that female-specific cancer types were the most prevalent types of cancer in 75\% of the countries in the world.
+Furthermore, the prevalence of breast, cervix or prostate cancer is the highest in almost 95\% of the countries \citep{Bray2013}.
+
+Although these estimates were a crude measures of cancer prevalence, \citet{Bray2013} provided a useful overview of what the prevalence of cancer is like around the world.\\
+
+\noindent
+The total number of deaths by non-communicable diseases (including cancer) is predicted to increase, reaching 52 million deaths by 2030 \citep{WHO2014}.
+With the prevalence of cancer at such high proportion around the world, and with the cancer mortality increasing in the future, cancer is a serious burden that the world has to face.
+
+\subsection{Mechanisms of cancer development}
+\label{sub:mechanisms_of_cancer_development}
 
 It is now commonly understood that cancers are caused by dysregulation of various mechanisms or pathways that allow tumour cells to prolilferate, survive and migrate.
-Many of these dysregulated pathways frequently occur in cancer cells, and these characteristic traits, commonly known as the ``Hallmarks of Cancer'', were first described by \citet{Hanahan2000}.
+By taking advantage of these pathways, cancer cells are able to grow and proliferate, and eventually metastasise to other parts of the body.
+The environment in which the tumours grow can also be an important factor for their evolution and growth, contributing to the heterogeneity of the cancer cells within the tumour or between tumours at different sites of the body.
 
-These hallmarks were initially presented in 2000 to provide a common framework for scientists to better understand cancer biology, and also to direct the focus of future research.
-After a decade of advances in cancer research, the ``Hallmarks of Cancer'' were revisited by Hanahan and Weinberg in 2011, and four more hallmarks were added on top of the six presented in the original paper in 2000 \citep{Hanahan2011}.
+\subsubsection{Hallmarks of Cancer}
+\label{subsubsec:cancerhallmarks}
 
-Together, these ``Hallmarks of Cancer'' (fig 1) have provided a foundation of the general biological knowledge of cancer and its characteristics that allowed the researchers to better understand the underlying cancer biology.
+% TODO: REWRITE THIS SHIT!!!!!!!
 
-Although much reseach has been conducted on the ``Hallmarks of Cancer'', the molecular processes underlying these hallmarks have still not been fully explained \citep{Hanahan2011}.
-This is due to the fact that cancer cells develop progressively under selective pressures, and are also affected by many environmental factors \citep{Gatza2011}, which makes it difficult to identify causal factors of cancer development.
-Furthermore, these external factors affect how the cancer cells respond to the environment, and may also cause cancer progression and/or drug resistance.
+Early researches in cancer biology have been able to uncover many pathways and regulatory mechanisms in cancer, so much that there were too many possibilities in which cancer could develop from.
+With more than 100 distinct types and subtypes of cancers, each with its own distinct development mechanisms, it became difficult to focus on the overarching mechanisms of cancer development \citep{Hanahan2000}.
+To resolve this, \citet{Hanahan2000} proposed ``six essential alterations'' that cancer cells usually take advantage of during development and progression.
+These six key mechanisms were coined the term ``The Hallmarks of Cancer'', and has been used extensively by many researchers as their guide to target their research in specific areas of cancer biology.
 
-\subsection{Tumour heterogeneity}
-\label{sub:tumour_heterogeneity}
+10 years after the first review where the Hallmarks of Cancer was proposed, Hanahan and Weinberg revisited these hallmarks and added four more hallmarks on top of the six hallmarks presented in 2000 \citep{Hanahan2011}.
+The ten Hallmarks of Cancer are: sustaining proliferative signalling, evading growth suppressors, enabling replicative immortality, activating invasion and metastasis, inducing angiogenesis, resisting cell death, avoiding immune destruction, tumour-promoting inflammation, genome instability and mutation, and deregulating cellular energetics (\cref{fig:hallmarks}).
+Each of these hallmarks will be described very briefly to provide an overview of the significance of these hallmarks in tumour biology.
+
+\begin{figure}[h!]
+	\centering
+	\includegraphics[width=0.8\linewidth]{intro/hallmarks}
+	\caption[Hallmarks of Cancer]{Hallmarks of Cancer. The ten Hallmarks of Cancer as proposed by Hanahan and Weinberg in 2011. (Figure adapted from \citet{Hanahan2011})}
+	\label{fig:hallmarks}
+\end{figure}
+
+\paragraph{Sustaining proliferative signalling}
+
+\noindent
+Cancer is a disease of uncontrolled cell proliferation and division.
+However, without any external mitogenic growth signals, cancer cells cannot proliferate and therefore it is crucial for the cancer cells to maintain a steady proliferative signal.
+
+There are many ways in which cancer cells can maintain this signal.
+Firstly, cancer cells can acquire the ability to produce growth factors and express its corresponding receptor, creating a positive feedback mechanism that continuously stimulates the cells to proliferate \citep{Hanahan2000}.
+Secondly, they can express more receptors for a given growth signals, and thus become ``hyperresponsive'' to the signal \citep{Hanahan2000,Hanahan2011}.
+Thirdly, cancer cells can stimulate the neighbouring normal cells within its microenvironment to produce more growth factors that further assist them to grow \citep{Bhowmick2004, Liotta2001, Wiseman2002}.
+Finally, cancer cells are able to maintain the proliferative signal by mutating the growth factor receptor itself, or the downstream signalling targets of the growth factor receptor pathway \citep{Fuqua1991,SuHuang1997,Satyamoorthy2003}.
+
+By maintaining the the proliferative signals, cancer cells are able to continuously grow in their environment.
+
+\paragraph{Evading growth suppressors}
+
+\noindent
+As well as maintaining proliferative signals, cancer cells must also evade the signals that inhibit cell growth.
+Two tumour suppressor proteins have an essential role in this evasion: the \gls{rb} and p53 proteins.
+
+\Gls{rb} protein is a known tumour suppressor protein that is important in cell cycle regulation \citep{Burkhart2008,Hanahan2011}.
+\Gls{rb} protein is known to arrest the cells in \gls{g1} phase of the cell cycle by regulating E2F transcription factors in the cell, in reponse to molecules such as \gls{tgfb} \citep{Burkhart2008,Hanahan2000}.
+Although \gls{rb} protein is known for its role as a checkpoint in \gls{g1} phase, it also takes part in the regulation of apoptotic cell death, cell differentiation and chromosome stability \citep{Burkhart2008}.
+The contribution of the loss of function of the \gls{rb} protein in tumorigenesis and tumour progression is evident, but at which stage of the development this loss of function mutation is crucial for the cancer cells is yet to be clarified \citep{Burkhart2008}.
+
+p53 protein is another tumour suppressor protein that is closely related to the \gls{rb} protein and its role in cell cycle arrest \citep{Hanahan2011,Levine1997}.
+As \citet{Hanahan2011} mentions in their review, ``whereas \gls{rb} transduces growth-inhibitory signals that originate largely outside of the cell, TP53 receives inputs from stress and abnormality sensors that function within the cell's intracellular operating systems''.
+The level of expression and its stability is crucial for the function of p53 protein, as the abundance of p53 protein in the cell triggers cell cycle arrest or cell \gls{apoptosis}, depending on the extent of the cell signal \citep{Fridman2003,Hanahan2011,Levine1997}.
+When abundant and active, p53 protein transcribes the p21 protein which, through series of events in the p16-cyclin D$_1$-cdk4-\gls{rb} pathway, triggers cell cycle arrest \citep{Levine1997}.
+
+In many cancers, these proteins are usually mutated and selected against to allow continuous growth of the tumour cells.
+
+\paragraph{Resisting cell death}
+
+\noindent
+\Gls{apoptosis}, the natural and controlled death of a cell, is one of many defense mechanisms in the body to prevent the development of cancer.
+
+\paragraph{Enabling replicative immortality}
+
+\noindent
+
+\paragraph{Activating invasion and metastasis}
+
+\noindent
+
+\paragraph{Inducing angiogenesis}
+
+\noindent
+
+\paragraph{Avoiding immune destruction}
+
+\noindent
+
+\paragraph{Tumour-promoting inflammation}
+
+\noindent
+
+\paragraph{Genome instability and mutation}
+
+\noindent
+
+\paragraph{Deregulating cellular energetics}
+
+\noindent
+
+\subsubsection{Metastasis}
+\label{ssub:metastasis}
+
+\subsubsection{Tumour heterogeneity}
+\label{subsub:tumour_heterogeneity}
+
+\subsection{Causes and risk factors of cancer}
+\label{sub:causes_and_risk_factors_of_cancer}
+
+\subsubsection{Genetic mutations}
+\label{ssub:Genetic mutations}
+
+p53 mutations, BRCA?
+
+\subsubsection{Environmental factors}
+\label{ssub:Environmental factors}
+
+\subsection{Treatment of cancer}
+\label{sub:treatment_of_cancer}
 

### 2016-11-19 thesis/misc/glossary.tex
finally finished the obesity section of intro
tmp commit

@@
 	description={Summary score of a sample that represents the overall gene expression of the genes that was used to obtain the scores}
+}
+
+\newglossaryentry{ldp}{
+	name=Limited duration prevalence,
+	text=limited duration prevalence,
+	description={Number of patients diagnosed with a disease within a fixed period of time in the past}
+}
+
+\newglossaryentry{apoptosis}{
+	name=Apoptosis,
+	text=apoptosis,
+	description={The death of cells which occurs as a normal and controlled part of an organism's growth or development}
 }
@@
 \newacronym{snp}{SNP}{single nucleotide polymorphism}
+\newacronym{rb}{RB}{retinoblastoma-associated}
+\newacronym{opa}{OPA}{obesity policy action}
+\newacronym{g1}{G1}{Gap 1}
+\newacronym{tgfb}{TGF-$\beta$}{transforming growth factor-$\beta$}
 \newacronym[longplural={genome wide association studies},\glsshortpluralkey={GWAS}]{gwas}{GWAS}{genome wide association study}

### 2016-11-19 thesis/thesis.tex
tmp commit
tmp commit, added .swp in .gitignore
Tweaked the BiocRefStyle.bst so it prints web page references correctly

@@
-\documentclass[a4paper, 12pt, oneside]{report}
+% \documentclass[a4paper, 12pt, oneside]{report} % for final draft compilation
+\documentclass[a4paper, 12pt, oneside, draft]{report} % for draft compilation
 
@@
 \usepackage[acronym, toc]{glossaries}               % package for making glossaries
+\usepackage{xcolor,colortbl}                        % package for making colours
+\usepackage[noabbrev,capitalise]{cleveref}          % package for referencing multiple referencing in a single \ref command
+\usepackage[font=footnotesize,format=hang]{caption} % package for formatting figure caption
 
-% Set up the hyperlinks - hange the colour of the links to blue:
-\hypersetup{allcolors=blue}
+% Define colour(s) for table rows:
+\definecolor{LightCyan}{rgb}{0.88,1,1}
+
+% Set up the hyperlinks - change the colour of the links to blue:
+% \hypersetup{allcolors=blue}
+\hypersetup{allcolors=black} % uncomment this line to make the hyperlinks black
+
+% Change the default cref name for equations:
+\creflabelformat{equation}{#2\textup{#1}#3}
+
+% Change the default url font:
+\urlstyle{rm}
 
@@
 
-% If you want to compile specific files, uncomment the line below and add the filenames in the {} :
-%\includeonly{}
-
 % Change the title of bibliography to references:
@@
 
-% To generate glossaries:
+% Italicise the subsubsection title:
+\titleformat{\subsubsection}{\normalfont\itshape\bfseries}{\thesubsubsection}{1em}{}
+\titleformat{\paragraph}{\normalfont\itshape}{\theparagraph}{\noindent}{}
+
+% For generating glossaries:
 \include{./misc/glossary}
 \makeglossaries
+
+% Specify graphics path:
+\graphicspath{ {./images/}}
+\DeclareGraphicsExtensions{.pdf,.png,.jpg}
+
+% If you want to compile specific files, uncomment the line below and add the filenames in the {} :
+%\includeonly{}
 
@@
 \pagenumbering{roman} % set page numbering to roman numeral
+
+% Set line spacing to double space:
+\onehalfspacing
 

### 2016-11-20 thesis/intro/intro.tex
tmp commit

@@
 \gls{npy} was first discovered by \citet{Tatemoto1982}, and later associated with altered eating behaviour in rats by \citet{Clark1984}.
-\gls{npy} is recognised by the \gls{nyr}, a G protein-coupled receptor expressed on the downstream effector neurons, which relays the orexogenic signal from the \gls{npy}/\gls{agrp} neurons \citep{Barsh2002, Bell2005, Herzog1992}
+\gls{npy} is recognised by the \gls{nyr1}, a G protein-coupled receptor expressed on the downstream effector neurons, which relays the orexogenic signal from the \gls{npy}/\gls{agrp} neurons \citep{Barsh2002, Bell2005, Herzog1992}
 
@@
 
-% TODO: REWRITE THIS SHIT!!!!!!!
-
 Early researches in cancer biology have been able to uncover many pathways and regulatory mechanisms in cancer, so much that there were too many possibilities in which cancer could develop from.
@@
 The ten Hallmarks of Cancer are: sustaining proliferative signalling, evading growth suppressors, enabling replicative immortality, activating invasion and metastasis, inducing angiogenesis, resisting cell death, avoiding immune destruction, tumour-promoting inflammation, genome instability and mutation, and deregulating cellular energetics (\cref{fig:hallmarks}).
-Each of these hallmarks will be described very briefly to provide an overview of the significance of these hallmarks in tumour biology.
+Each of these hallmarks will be described briefly to provide an overview of the significance of these hallmarks in tumour biology.
 
@@
 \noindent
-\Gls{apoptosis}, the natural and controlled death of a cell, is one of many defense mechanisms in the body to prevent the development of cancer.
+\Gls{apoptosis}, the natural and controlled death of a cell, is one of many defence mechanisms in the body to prevent the development of cancer.
+The triggering of \gls{apoptosis} is controlled and maintained by the balance between the pro- and anti-apoptotic signals that affect the downstream apoptotic proteins \citep{Hanahan2011}.
+
+\Gls{bcl2} family proteins are the major players of apoptosis.
+Embedded in the mitochondrial membrane, the Bax subfamily of \gls{bcl2} proteins are pro-apoptotic, and are under constant repression by the anti-apoptotic proteins of the \gls{bcl2} subfamily proteins \citep{Adams2007,Hanahan2011}.
+The inhibition of the pro-apoptotic proteins are achieved by the interaction between the pro- and anti-apoptotic proteins via the \gls{bh3} domain \citep{Adams2007}.
+In addition to the pro-apoptotic Bax subfamily proteins, there is a family of proteins that contain a single \gls{bh3} domain, and hence named \gls{bh3}-only family proteins \citep{Adams2007}.
+\gls{bh3}-only proteins either disrupt the inhibition of the Bax family proteins by the \gls{bcl2} subfamily proteins, or directly stimulate the Bax subfamily proteins \citep{Adams2007}.
+The activation of the pro-apoptotic proteins disrupt the mitochondrial cell membrane adn releases cytochrome \textit{c}, which in turn triggers the activation of the apoptotic proteases caspase 8 and caspase 9, resulting in apoptosis \citep{Adams2007,Hanahan2011}.
+
+The precise cellular conditions that trigger apoptosis are yet to be uncovered, but there is a clear evidence of p53 tumour suppressor playing a key role in apoptosis initiation.
+When \acrshort{dna} damage is sensed, p53 protein transcribes the pro-apoptotic members of the \gls{bcl2} family proteins, such as Bax, Puma and Noxa \citep{Fridman2003,Hanahan2011}.
+The fact that p53 protein is involved not only in cell cycle regulation, but also in apoptotic pathway shows its importance in the control of cancer development, and perhaps that is the reason why more than 50\% of tumours have a mutation in the \textit{p53} gene \citep{Levine1997}.
 
@@
 \noindent
+When acquired by cancer cells, the previous three hallmarks allow the cells to continuously grow in an uncontrolled manner.
+However, this is not the case as the disruption in cell signalling on its own does not trigger the rapid expansion of the tumour -- only when the cells have the capability of replicating infinitely \citep{Hanahan2000, Hanahan2011}.
+
+There are two barriers in achieving the infinite replicative capacity: senescence and crisis \citep{Hanahan2011}.
+When cells are grown in a culture, cells are able to replicate and grow until a certain point, and the cells reach the stage of senescence where the cells stop growing but are still viable \citep{Hanahan2011}.
+The cells that manage to bypass this senescence phase, by mutating the tumour suppressor proteins such as p53 and \gls{rb} proteins, then undergoes the crisis phase where the majority of the cells in the culture dies \citep{Hanahan2011}.
+Very rarely, some cells are able to survive the crisis phase and acquires the ability to replicate infinitely, and this transition has been termed immortalisation \citep{Hanahan2011, Wright1989}.
+
+Without immortalisation, cancer cells are not able to generate a macroscopic tumours that are capable of killing the host organism \citep{Hanahan2000,Hanahan2011}.
 

### 2016-11-22 thesis/intro/intro.tex
tmp commit

@@
 p53 protein is another tumour suppressor protein that is closely related to the \gls{rb} protein and its role in cell cycle arrest \citep{Hanahan2011,Levine1997}.
-As \citet{Hanahan2011} mentions in their review, ``whereas \gls{rb} transduces growth-inhibitory signals that originate largely outside of the cell, TP53 receives inputs from stress and abnormality sensors that function within the cell's intracellular operating systems''.
+As \citet{Hanahan2011} mentions in their review, ``\gls{rb} transduces growth-inhibitory signals that originate largely outside of the cell, TP53 receives inputs from stress and abnormality sensors that function within the cell's intracellular operating systems''.
 The level of expression and its stability is crucial for the function of p53 protein, as the abundance of p53 protein in the cell triggers cell cycle arrest or cell \gls{apoptosis}, depending on the extent of the cell signal \citep{Fridman2003,Hanahan2011,Levine1997}.
@@
 
+\paragraph{Inducing angiogenesis}
+
+\noindent
+Like normal cells, cancer cells also require access to oxygen and nutrient supply from the blood.
+However, new blood vessel formation, or angiogenesis, only occurs in a handful of occasions, and thus the tumour needs to acquire the ability to stimulate angiogenesis in the microenvironment in which they live in \citep{Hanahan2011}.
+
+It is thought that the angiogenic ability of the tumour is gained at an early stage of its development by disrupting the ``angiogenic switch'' \citep{Hanahan2011}.
+Like many cellular processes, angiogenesis is controlled by the balance between the inducing and inhibiting agents of angiogenesis.
+\Gls{vegfa} is the major growth factor that triggers angiogenesis by signalling through the \acrshort{vegf} receptors (though some receptors have inhibitory effect) \citep{Yancopoulos2000}.
+Opposing the effect of \Gls{vegfa} is the \gls{tsp1}, a protein that is known to inhibit the process of angiogenesis by affecting the bioavailability of \acrshort{vegf}, as well as promoting apoptosis \citep{Kazerounian2008}.
+
+In general, many cancers stimulate the expression of \gls{vegfa} protein and reduces the expression of \gls{tsp1} \citep{Kazerounian2008}.
+
 \paragraph{Activating invasion and metastasis}
@@
 \noindent
+One of the prime reasons that cancers are so difficult to treat and have high mortality is due to their ability to metastasise to other parts of the body.
+Once metastasised to different parts of the body, the clinician not only has to consider the primary tumour site, but also the second, third, or even fourth tumour sites, which may or may not have similar tumour biology to the primary tumour, and thus have to consider multiple treatment plans to cure the patient.
 
-\paragraph{Inducing angiogenesis}
+Metastasis is a multi-step process that requires the selection and survival of cancer cell(s) that is able to metastasise (``seed''), and is also compatible with the specific target tissue where they prosper (``soil'') \citep{Talmadge2010}.
+The downregulation or loss of function of E-cadherin, a protein important in cell-to-cell connection and adherence, is known to trigger tumour invasion and metastasis through the reversible process of \gls{emt} \citep{Hanahan2011,Kalluri2009}.
+Once mobile and invasive, these metastatic cancer cell is able to circulate through the vascular system and eventually invade the target organ, where it must undergo \gls{met} and colonise the site, and eventually metastasise from this site \citep{Hanahan2011,Kalluri2009}.
 
-\noindent
+There are numerous signals, growth factors and pathways that initiate the metastatic phenotype in cancer cell, and it remains an active field of research \citep{Hanahan2011,Kalluri2009}.
 
-\paragraph{Avoiding immune destruction}
+\paragraph{Genome instability and mutation}
 
@@
 
-\paragraph{Genome instability and mutation}
+\paragraph{Deregulating cellular energetics}
 
@@
 
-\paragraph{Deregulating cellular energetics}
+\paragraph{Avoiding immune destruction}
 

### 2016-11-23 thesis/intro/intro.tex
tmp commit

@@
 \noindent
+Many of the hallmarks are acquired by the cancer cells during its growth, through many genomic changes and mutations of the essential genes that allow cancer cells to survive, proliferate, invade and metastasise.
+Thus, it is only logical to induce that many, if not all, cancers are more prone to genetic mutations and restructuring, which allows the cancers to subsequently acquire the already described Hallmarks of Cancer \citep{Hanahan2011}.
+
+Until recently, the evolutionary progress of cancers have been thought to be gradual and stepwise, where each driver mutation is acquired over many years, or even decades \citep{Stephens2011}.
+There are now evidence that show otherwise, whereby a single catastrophic event in the chromosome structure occurs, resulting in major structural rearrangements such as inversions, deletions, duplications and translocations \citep{Leibowitz2015,Stephens2011}.
+This rare event that is estimated to occur in 2$\sim$5\% of all cancers was termed \gls{chromothripsis}, and is thought to be caused by the physical separation of parts of chromosome by a structure called micronuclei \citep{Leibowitz2015,Stephens2011}.
+Owing to its nature of being a one-off chromosomal disruption, chromothripsis (usually) cause disruptions in tumour suppressor genes rather than duplications of oncogenes \citep{Leibowitz2015}.
+
+In addition to this, areas of the chromosome where major structural rearrangements have occured can be prone to a regional hypermutation, termed \gls{kataegis} \citep{Leibowitz2015,Nik-Zainal2012}.
+Although the full mechanism in which \gls{kataegis} occur is not yet defined, \gls{aid}/\acrshort{apobec} family proteins, a family of proteins involved in gene editing, may play a major role in \gls{kataegis} \citep{Leibowitz2015,Nik-Zainal2012}.
+By deaminating cytidine to uracil in the \gls{ssdna} that may occur at the sites of double strand break, \gls{aid}/\acrshort{apobec} family proteins introduces hypermutated regions in the genome \citep{Leibowitz2015,Nik-Zainal2012}.
+
+It is evident that both \gls{chromothripsis} and \gls{kataegis} contribute to the genomic aberrations and mutations that ultimately contribute to many of the tumorigenic phenotypes.
 
@@
 
-\subsubsection{Metastasis}
-\label{ssub:metastasis}
-
+% TODO: do I really need tumour heterogeneity??
 \subsubsection{Tumour heterogeneity}

### 2016-11-23 thesis/misc/glossary.tex
tmp commit

@@
 	text=limited duration prevalence,
-	description={Number of patients diagnosed with a disease within a fixed period of time in the past}
+	description={Number of patients diagnosed with a disease within a fixed period of time in the past \citep{Bray2013}}
 }
@@
 	description={The death of cells which occurs as a normal and controlled part of an organism's growth or development}
+}
+
+\newglossaryentry{chromothripsis}{
+	name=Chromothripsis,
+	text=chromothripsis,
+	description={An all-at-once mutational pattern in which a large number of rearrangements are confined to local regions of one or more chromosomes \citep{Stephens2011}}
+}
+
+\newglossaryentry{kataegis}{
+	name=Kataegis,
+	text=kataegis,
+	description={Showers of single nucleotide changes confined to local regions, often near sites of chromosome rearrangements \citep{Stephens2011}}
 }
@@
 \newacronym{msh}{MSH}{melanocyte-stimulating hormone}
-\newacronym{nyr}{Y1R}{neuropeptide Y1 receptor}
-\newacronym{nyr}{Y2R}{neuropeptide Y2 receptor}
+\newacronym{nyr1}{Y1R}{neuropeptide Y1 receptor}
 \newacronym{acth}{ACTH}{adrenocorticotropic hormone}
@@
 \newacronym{g1}{G1}{Gap 1}
+\newacronym{bcl2}{Bcl-2}{B-cell lymphoma 2}
+\newacronym{bh3}{BH3}{Bcl-2 homology 3}
+\newacronym{dna}{DNA}{deoxyribonucleic acid}
+\newacronym{vegf}{VEGF}{vascular endothelial growth factor}
+\newacronym{vegfa}{VEGF-A}{vascular endothelial growth factor-A}
+\newacronym{tsp1}{TSP-1}{thrombospondin-1}
+\newacronym{emt}{EMT}{endothelial-mesenchymal transition}
+\newacronym{met}{MET}{mesenchymal-endothelial transition}
+\newacronym{aid}{AID}{activation-induced cytidine deaminase}
+\newacronym{apobec}{APOBEC}{apolipoprotein B mRNA editing enzyme, catalytic polypeptide-like}
+\newacronym{ssdna}{ssDNA}{single-stranded DNA}
 \newacronym{tgfb}{TGF-$\beta$}{transforming growth factor-$\beta$}

### 2016-11-25 thesis/intro/intro.tex
tmp commit

@@
 \noindent
+Inflammation and immune response, though it may assist in the control of diseases and injuries, have been implicated in the progression of tumour and their acquisition of many of the hallmarks \citep{Hanahan2011}.
+
+There are two pathways of inflammation that contribute to the cancer survival and progression \citep{Mantovani2008}.
+The intrinsic pathway is caused by oncogenic characteristics of the cancer cells, where these cells produce inflammatory signals themselves and promote advantageous characteristics for the cancer cells, such as increased proliferation and survival \citep{Mantovani2008}.
+On the other hand, the external pathway is caused by infectious or inflammatory conditions in the tumour microenvironment that surround the tumour and assist in the growth of the tumour \citep{Mantovani2008}.
+There are two key transcription factors that connect the two pathways of inflammation that ultimately leads to the prommotion  of many, if not all, of the hallmarks discussed so far: \gls{nfkb} and \gls{stat3} \citep{Mantovani2008}.
+
+The study by \citep{Pikarsky2004} showed that \gls{nfkb} was an essential component for the progression of hepatocellular carcinoma, providing the first evidence as to how inflammation may affect the tumour biology.
+\gls{nfkb} regulates the expression of genes that encode the inflammatory cytokines, angiogenic factors, and many others, including anti-apoptotic genes (through the activation of \gls{stat3}) \citep{Elinav2013,Mantovani2008}.
+\gls{stat3} is a transcription factor that is activated through variety of receptors and pathways, including the \gls{il6} receptor, \glspl{gpcr} and \glspl{tlr} \citep{Yu2007,Yu2014}.
+By increasing the transcription of anti-apoptotic proteins such as \gls{bcl2}, \gls{stat3} promotes tumour survival and proliferation \citep{Yu2007}.
+Furthermore, \gls{stat3} is able to increase the expression of inflammatory cytokines (such as \gls{il6} and \acrshort{il10}) and \acrshort{vegf}, which provides a positive feedback mechanism for further inflammation (and therefore survival and proliferation signals), and advancement of angiogenesis, respectively \citep{Yu2007}.
+
+Though helpful at times, inflammation induced by the immune system must be taken care of and controlled in order to prevent rapid tumour development and progression.
 

### 2016-11-25 thesis/intro/intro.tex
tmp commit

@@
 The downregulation or loss of function of E-cadherin, a protein important in cell-to-cell connection and adherence, is known to trigger tumour invasion and metastasis through the reversible process of \gls{emt} \citep{Hanahan2011,Kalluri2009}.
-Once mobile and invasive, these metastatic cancer cell is able to circulate through the vascular system and eventually invade the target organ, where it must undergo \gls{met} and colonise the site, and eventually metastasise from this site \citep{Hanahan2011,Kalluri2009}.
+Once mobile and invasive, these metastatic cancer cell is able to circulate through the vascular system and eventually invade the target organ, where it must undergo \gls{met} and colonise the site \citep{Hanahan2011,Kalluri2009}.
 
@@
 On the other hand, the external pathway is caused by infectious or inflammatory conditions in the tumour microenvironment that surround the tumour and assist in the growth of the tumour \citep{Mantovani2008}.
-There are two key transcription factors that connect the two pathways of inflammation that ultimately leads to the prommotion  of many, if not all, of the hallmarks discussed so far: \gls{nfkb} and \gls{stat3} \citep{Mantovani2008}.
+There are two key transcription factors that connect these two pathways of inflammation that ultimately leads to the prommotion  of many, if not all, of the hallmarks discussed so far: \gls{nfkb} and \gls{stat3} \citep{Mantovani2008}.
 
@@
 \noindent
+In order to grow and proliferate, cancer cells must adjust their energy metabolism to match their energy requirement \citet{Hanahan2011}.
+There are two ways in which a cell can produce energy: anaerobic respiration, where the cell uses glycolysis to rapidly generate \gls{atp} and lactate; and aerobic respiration where, in the presence of oxygen, the cell utilises the electron transport chain and \gls{atp} synthase in the mitochondria to efficiently produce \gls{atp}.
+However, even with abundant oxygen in the cell, \citet{Wardburg1956} found that cancer cells utilised the glycolytic pathway instead of the more efficient aerobic respiration, which was later termed ``aerobic glycolysis'' \citep{Hanahan2011}.
+
+It seems counterintuitive to limit its full potential of creating \gls{atp} by utilising glycolysis over electron transport chain, as the latter produces almost 18 times more \gls{atp} \citep{Hanahan2011, VanderHeiden2009}.
+However, it is hypothesised  that this apprent inefficiency helps the cell to generate and direct the molecules required for biosynthetic pathways during cell growth and division \citep{Cairns2011,VanderHeiden2009}.
+By deliberately relying purely on glycolysis (known as the ``Wardberg effect'') and slowing the conversion of \gls{pep} to pyruvate, the cell is able to shunt the carbon molecules into reaction pathways that produce constituents used to synthesise macromolecules, such as \acrshort{dna} and lipids \citep{Cairns2011,VanderHeiden2009}.
+
+The fact that the deregulation of energy metabolism is mediated somewhat by the genes involved in many of the other hallmarks (such as p53 and Myc) suggests that the disruption in energy metabolism may be a phenotype caused in conjunction with those core hallmarks \citep{Hanahan2011}.
 

### 2016-11-26 thesis/intro/intro.tex
tmp commit

@@
 \noindent
+Our immune system is constantly monitoring and maintaining normal cell biology within our body, preventing the emergence of cancer \citep{Hanahan2011}.
+That said, cancers are frequently observed in clinics, pointing to the fact that the observed cancer cells must have evaded the diligent monitoring by the immune system \citep{Hanahan2011}.
+
+Immunoediting is a process where the immune system successfully eliminates the highly immunogenic cancer cells during surveillance, but fails to eliminate the weakly immunogenic cells and therefore poses a selective advantage on these cells \citep{Hanahan2011,Teng2008}.
+In fact, there has been reports where cancer cells in immunodeficient mice were not able to initiate tumour growth in the immunocompetent mice, but the cancer cells from the immunocompetent mice were able to do so in immunodeficient mice \citep{Hanahan2011}.
+One possible reason for this was that the cancer cells from the immunodeficient mice were not immunoedited to the extent in which the cells from immunocompetent mice were, and thus unable to evade the immune system as efficiently as those cells in the immunocompetent mice \citep{Hanahan2011}.
+
+Though more experiments must be carried out to confirm its definitive role and mechanisms in which it contributes to cancer biology, immune system evasion is in no doubt a significant component of tumour progression.
+\\
+
+\noindent
 

### 2016-11-26 thesis/intro/intro.tex
finished draft for the hallmarks of cancer section of intro

@@
 \noindent
+By utilising many, if not all, of these major cancer hallmarks, cancer cells are able to initiate growth, proliferate, invade, metastasise and eventually kill the host if untreated.
+There is an evident need for an efficient way to monitor and treat cancer patients, and these Hallmarks of Cancers will prove to be the best target for effective treatments of cancer in the future.
 
 % TODO: do I really need tumour heterogeneity??
-\subsubsection{Tumour heterogeneity}
+\subsubsection{Tumour heterogeneity (?)}
 \label{subsub:tumour_heterogeneity}
+
+\subsubsection{Cancer stem cells (?)}
+\label{ssub:cancer_stem_cells}
 

### 2016-11-28 thesis/intro/intro.tex
tmp commit

@@
 \label{subsub:tumour_heterogeneity}
+TBC
 
@@
 \label{ssub:cancer_stem_cells}
+TBC
 
@@
 
+There are many causes and risk factors associated with the development and progression of cancer.
+Like obesity, cancers are dependent on both the genetic and environmental factors.
+
 \subsubsection{Genetic mutations}
@@
 
-p53 mutations, BRCA?
+It is quite obvious from the previous section (\cref{sub:mechanisms_of_cancer_development}: \nameref{sub:mechanisms_of_cancer_development}) that there are a lot of things that can go wrong in the cell.
+Any one of the ten Hallmarks of Cancers has the potential to initiate cancer growth and progression; every hallmark has their own complex mechanistic pathways; every pathway has many proteins associated with its function or role in the cell; and, though their impacts may differ from protein to protein, mutations in any one of the proteins involved in a pathway has the potential to cause cancer.
+
+% TODO: more citations!!!!
+Broadly speaking, there are three major classes of genes that have significance in the initiation of cancer: oncogenes, tumour suppressor genes and stability genes \citep{Vogelstein2004a}.
+Oncogenes are genes that promote or ``accelerate'' the process of tumorigenesis and/or tumour progression, whereas tumour suppressor genes prevent or ``decelerate'' those effects \citep{Vogelstein2004a}.
+Cancers usually mutate the oncogenes in such a way that it it constitutively activated or expressed at a very high level, and mutate the tumour suppressor genes so that these genes are inactivated and/or have reduced expression \citep{Vogelstein2004a}.
+Another thing to note for these genes is that oncogenes usually only require a ``one-hit'' mutation that affects a single allele, whereas tumour suppressor genes require mutations in both maternal and paternal alleles to have any effect \citep{Stratton2009,Vogelstein2004a}.
+
+Stability genes, also known as caretakers, are genes that are involved in, for example,  \gls{mmr}, \gls{ner} and \gls{ber}; in other words, genes that are involved in \acrshort{dna} repair mechanisms \citep{Vogelstein2004a}.
+These caretaker genes in effect prevents the cell to develop new mutations from \acrshort{dna} replication and mitosis.
+However, only the oncogenes or the tumour suppressor genes are able to directly affect the biology of the cell, and therefore mutations in the stability genes affect the mutation rates of the other classes of genes, ultimately leading to tumorigenesis \citep{Vogelstein2004a}.
+\\
+
+\noindent
+There are two types of mutations that occurs in an individual: germline mutation and somatic mutation.
+Germline mutations are the mutations that are passed on from your parents, and somatic mutations are the mutations acquired by the cell over the course of the individual's lifetime.
+
+Unfortunately, there are germline mutations that make the individual more prediposed to cancer \citep{Vogelstein2004a}.
+These individuals have a mutation in one or more of the classes of cancer-related genes, which makes them more likely to develop tumour than their normal pairs \citep{Vogelstein2004a}.
+As a consequence, these individuals may develop multiple tumours at an early stage of their lives than those who develop tumours from somatic mutations alone, as these individuals already have a mutation that allows all of the cells in their body to have one or more of the Hallmarks of Cancers \citep{Vogelstein2004a}.
+
+In contrast, somatic mutations are acquired during an individual's lifetime and is not passed down to their children.
+Somatic mutations can either be a driver or a passenger mutation, where a driver mutation helps tumorigenesis and its subsequent progression (in other words, mutations in oncogene, tumour suppressor, or stability gene), and passenger mutation is where the mutation has no selective advantage for the cancer and does not contribute to tumour progression \citep{Stratton2009}.
+Since cancers require multiple driver mutations, somatic mutations are an essential part of tumour development.
+Perhaps that is one of the reasons why, in some cancers, germline mutation in the stability genes cause greater susceptibility to cancers than mutations in oncogene or tumour suppressor gene \citep{Vogelstein2004a}.
 
 \subsubsection{Environmental factors}
-\label{ssub:Environmental factors}
+\label{ssub:environmental_factors}
+
+Though mutation rates can differ between individuals,
+
+``The risk of lung cancer diminishes after smoking cessation, but not during the first 5 years, and the relative risk never turns to that of a nonsmoker'' \citep{Hecht1999}. \\
+
+Tobacco carcinogens form DNA adducts (carcinogen metabolite attached covalently to DNA at A or G), which can be activated or detoxfied in the cell \citep{Hecht1999}. \\
+
+If DNA adducts are not detoxfied, it can lead to DNA miscoding and therefore mutation \citep{Hecht1999}. \\
 
@@
 
-\subsection{Microarray technology}
-\label{subsec:microarray_technology}
-
-\subsection{Next generation sequencing technology}
-\label{sub:next_generation sequencing_technology}
+\subsection{Microarray and Next Generation Sequencing technologies}
+\label{sub:microarray_and_next_generation_sequencing_technologies}
 
@@
 
-% TODO: have a leading sentence that introduces the gene signatures
+\subsection{Obesity associated genetic signatures}
+\label{sub:obesity_associated_genetic_signatures}
 
-\section{Obesity associated genetic signatures}
-\label{sec:obesity_associated_genetic_signatures}
+\subsubsection{\citet{Creighton2012} study}
+\label{ssub:creighton_study}
 
-\subsection{\citet{Creighton2012} study}
-\label{sub:creighton_study}
+\subsubsection{\citet{Fuentes-Mattei2014} study}
+\label{ssub:fuentes_mattei_study}
 
-\subsection{\citet{Fuentes-Mattei2014} study}
-\label{sub:fuentes_mattei_study}
-
-\section{Pathway associated genetic signatures}
-\label{sec:pathway_associated_genetic_signatures}
+\subsection{Pathway associated genetic signatures}
+\label{sub:pathway_associated_genetic_signatures}
 
@@
 
-\subsection{\citep{Gatza2011} study}
-\label{sub:gatza_study}
+\subsubsection{\citet{Gatza2011} study}
+\label{ssub:gatza_study}
 

### 2016-11-29 thesis/intro/intro.tex
tmp commit

@@
 
-Though mutation rates can differ between individuals,
+For obesity, since there is a clear evidence of the cause (excessive eating and lack of exercise) and a well defined mechanism of how the disease is developed (excess calories are converted into fat), it is relatively easy to speculate the effect of the environment on the disease (for example, abundance of high calory food).
+On the other hand, cancers are developed through many cellular pathways (Hallmarks of Cancers) which are not fully uncovered yet, with either too many or very few causes associated with those pathways (for example, mutations, inflammation and/or growth factors), and therefore it is difficult to assess how, rather than whether, a given environmental factor causes specific cancer.
 
-``The risk of lung cancer diminishes after smoking cessation, but not during the first 5 years, and the relative risk never turns to that of a nonsmoker'' \citep{Hecht1999}. \\
+With that said, there has been a few environmental factors that have a strong evidence of causing cancer, though the complete mechanism of action for some of these must be elucidated in the coming years.
+% TODO: do I really need infections and inflammations?
+Out of these, infections and inflammation, smoking, \gls{uv} radiation, and diet and obesity will be covered as examples.
+\\
 
-Tobacco carcinogens form DNA adducts (carcinogen metabolite attached covalently to DNA at A or G), which can be activated or detoxfied in the cell \citep{Hecht1999}. \\
+% Infections and inflammmation
 
-If DNA adducts are not detoxfied, it can lead to DNA miscoding and therefore mutation \citep{Hecht1999}. \\
+% Smoking
+\noindent
+Smoking is a known risk factor for many cancers; the obvious one being lung cancer \citep{Gandini2008,Hecht1999}.
+For current smokers, the relative risk is greater than 1.5 in many cancers, including liver, cervix and kidney, and it can rise up to 8.96 in lung cancer \citep{Gandini2008}.
+Even if smokers stop smoking, the relative risk does not decrease until the first 5 years have passed and will never return to the level of non-smokers \citep{Hecht1999}.
+
+In tobacco smoke, there are literally hundreds and thousands of chemical compounds, of which about 20 compounds have been shown to have carcinogenic effects \citep{Hecht1999}.
+Among these, \glspl{pah} and \gls{nnk} have been shown to cause lung cancer in mice when administered systemically \citep{Hecht1999}.
+These carcinogens are processed first by \gls{cyp} enzymes (termed meta\-bolic activation), and the product can either be processed further to be detoxified completely and cause no harm to the cell, or alternatively, the product can covalently bind to the \acrshort{dna}, forming \acrshort{dna} adducts \citep{Hecht1999}.
+Unless these adducts are resolved correctly by the \acrshort{dna} repair mechanisms, it can result in a permanent mutation within the genome and lead to oncogenesis \citep{Hecht1999}.
+
+In addition to this, cigarettes contain free radicals and chemicals that cause tremendous oxidative stress and damage to the cells \citep{Hecht1999}.
+These radicals cause \acrshort{dna} nicks and damages the \acrshort{dna}, again requring the cell's \acrshort{dna} repair mechanisms to remove the damage so it does not become permanent.
+\\
+
+% UV radiation
+\noindent
+Another important environmental factor to consider is \gls{uv} radiation.
+Exposure to \gls{uv} radiation has been significantly associated with the increased risks of skin cancers \citep{Armstrong2001,Gallagher2006}.
+The risks of developing any skin cancers can vary between race (Caucasian people are more likely to get skin cancers), and potentially the age when they were exposed to high amounts of \gls{uv} radiation \citep{Armstrong2001,Gallagher2006}.
+
+Again, the mechanism of how \gls{uv} radiation may cause  skin cancers is related to the \acrshort{dna} repair system.
+\gls{uv} radiation causes the formation of a dimer in the adjacent pyrimidines, resulting in a covalent bond formation between these nucleotides \citep{Friedberg2003,Hoeijmakers2001}.
+These dimers introduce bulky structures in the \acrshort{dna} that prevent the \acrshort{dna} replication and transcription enzymes from functioning properly, and if they are not repaired, it may cause lasting mutations \citep{Friedberg2003,Hoeijmakers2001}.
+These damages caused by \gls{uv} radiation can be relieved by the \gls{ner} system or the photoreactivation process, where a photoreactivating enzyme uses light to monomerize the dimeric pyrimidines \citep{Friedberg2003}.
+\\
+
+\noindent
+As one can imagine, \acrshort{dna} damages caused by smoking and/or \gls{uv} radiation have the potential to damage the genome and initiate cancer if wrong genes are mutated, such as p53 tumour suppressor gene.
+Therefore, limited exposure to both factors is crucial for the prevention of any types of cancers, as well as the more specific types of cancers associated with these factors, such as lung and skin cancers.\\
+
+% Diet and obesity
+\noindent
 

### 2016-11-30 thesis/intro/intro.tex
tmp commit

@@
 With that said, there has been a few environmental factors that have a strong evidence of causing cancer, though the complete mechanism of action for some of these must be elucidated in the coming years.
-% TODO: do I really need infections and inflammations?
-Out of these, infections and inflammation, smoking, \gls{uv} radiation, and diet and obesity will be covered as examples.
+Out of these, smoking, \gls{uv} radiation, and diet and obesity will be covered as examples.
 \\
-
-% Infections and inflammmation
 
@@
 
+% Diet and obesity
+\noindent
+Finally, diet and obesity are known risk factors for many types of cancers \citep{Ames1995a,Calle2004}.
+Though more will be covered in details in \cref{sec:obesity_and_cancers}: \nameref{sec:obesity_and_cancers}, the effect of obesity on cancers will be briefly mentioned here.
+
+Epidemiological evidence of obesity being a risk factor for many cancers has been well established, yet the mechanisms of its contribution to cancer is still  speculative \citep{Calle2003,Kelesidis2006a}.
+Although there are a lot of possible mechcanisms, there is growing evidence of the role of chronic inflammation caused by increased fat mass in the body \citep{Kelesidis2006a,Lumeng2011,Hernandez2013}.
+Free fatty acids from excess adipose tissue and/or diet can be recognised by the \gls{tlr}4, which can then activate \gls{nfkb} \citep{Lumeng2011}.
+Clearly, the result of the activation of \gls{nfkb} can lead to the activation of one of the Hallmarks of Cancers -- tumour-promoting inflammation.
+
+The nutritional content of the food that people consume is also important, as it can both directly and indirectly influence the likelihood of tumorigenesis, as well as its progression \citep{Ames1995a}.
+Many fruits and vegetables contain antioxidants as well as essential viatmins and minerals.
+Antioxidants are able to neutralise the effect of \gls{ros},  and many viatmins have been associated with reduction of many types of cancers \citep{Ames1995a}.
+In fact, ascorbic acid (vitamin C) is able to prevent the lipid peroxidation of lipid plasma caused by tobacco smokes \textit{in vitro}, which could be why smokers, compared to non-smokers, have decreased levels of ascorbic acid \citep{Hecht1999}.
+
+By maintaining healthy diet, it may be possible to reduce the risks of many cancers and obesity, as well as the comorbidities associated with them.
+\\
+
 \noindent
 As one can imagine, \acrshort{dna} damages caused by smoking and/or \gls{uv} radiation have the potential to damage the genome and initiate cancer if wrong genes are mutated, such as p53 tumour suppressor gene.
-Therefore, limited exposure to both factors is crucial for the prevention of any types of cancers, as well as the more specific types of cancers associated with these factors, such as lung and skin cancers.\\
-
-% Diet and obesity
-\noindent
+Healthy diet is also crucial for lowering the risk of cancers through beneficial vitamins and antioxidants, and prevent obesity as well as chronic inflammation that accompanies with it.
+There are many other environmental factors associated with cancers, and limiting the exposure to detrimental environmental factors is crucial for the prevention of any types of cancers, as well as the more specific types of cancers associated with these factors, such as lung and skin cancers.
+\\
 

### 2016-12-01 thesis/intro/intro.tex
finished draft for the cancer part of the intro

@@
 
+Traditionally, the treatment of cancer has been surgery, radiation therapy and chemotherapy.
+Though surgery is an effective method to remove the tumour from the patient, there is always a risk of metastasis and recurrence, which is why combination of surgery with other treatments like chemotherapy is usually considered.
+Earlier chemotherapeutic drugs and radiation therapy had limited effect on tumours due to their resistance against these treatments \citep{Wilhelm2006}.
+However, greater understanding of the pathways and mechanisms that tumour cells use have shed  light on the development of drugs that are more effective against tumours.
+RAF inhibitors are among one of many.
+
+The RAS-RAF-MEK-ERK signalling pathway is a pathway involved in cell survival, growth and proliferation \citep{Samatar2014,Wilhelm2006}.
+The members of this pathway are often mutated in many types of cancers, causing an overactivation of this pathway.
+For example, approximately a third of cancers contain a mutation in the RAS oncogene which allows the constitutive activation of the RAS-ERK pathway \citep{Samatar2014}.
+Development of the drugs that target  the downstream RAF and its mutant forms (such as BRAF$^{V600E}$) have been successful, and many RAF inhibitor drugs such as sorafenib and vemurafenib have been approved for use clinically \citep{Samatar2014,Wilhelm2006}.
+Even though our understandings of signalling pathways and drug development based on these knowledge is improving, there is always the possibility of tumour to acquire resistance to those drugs and there is always a need for new ways to treat cancer patients \citep{Samatar2014}.
+
+In fact, there has been evidence of pathway cross-talks between RAS-ERK pathway and \gls{pi3k}/Akt pathway \citep{Moelling2002,Zimmermann1999a}.
+Furthermore, a study in \textit{Drosophila} showed that the inhibition of just a single pathway does not necessarily stop the effect of the pathway, as the signal can flow through alternate pathways \citep{Dar2012}.
+Pathway cross-talk and increased flux through alternate pathways could be one of the ways in which tumour cells acquire resistance to drugs.
+However, the effectiveness of targeting multiple pathways in humans is yet to be confirmed and further investigations are required to solidify the mode of action and its relevance to tumour biology.
+\\
+
+\noindent
+Though there are difficulties such as resistance to drugs and metastasis, further development of technologies and improved knowledge of tumour biology will only get us closer to the cure of cancer.
+
 \section{Obesity and cancers}

### 2016-12-01 thesis/intro/intro.tex
tmp commit

@@
 
+As mentioned briefly in \cref{sub:causes_and_risk_factors_of_cancer}: \nameref{sub:causes_and_risk_factors_of_cancer}, obesity is considered a major risk factor for many, if not all, types of cancers.
+The fact that the prevalence of obesity is extremely high, it is important to establish the link between the apparent risks associated with cancer.
+
 \subsection{Cancer risks associated with obesity}

### 2016-12-02 thesis/intro/intro.tex
tmp commit

--- 2016-12-01 thesis/intro/intro.tex
@@
 
+As mentioned briefly in \cref{sub:causes_and_risk_factors_of_cancer}: \nameref{sub:causes_and_risk_factors_of_cancer}, obesity is considered a major risk factor for many types of cancers.
+The fact that the prevalence of obesity is extremely high, it is important to establish the link between the apparent risks associated with cancer.
+
 \subsection{Cancer risks associated with obesity}
@@
 
-\citep{Calle2003}
+Although the mechanisms in which obesity causes cancer is still under debate, the association between obesity and cancer has long been established.
+
+Perhaps the most convincing evidence of the association between obesity and cancer was first presented by \citet{Calle2003}.
+In this study, the link between \gls{bmi} and variety of cancers were investigated in 900,000 adults between 1982 and 1998 \citep{Calle2003}.
+The results from this study showed that morbidly obese men and women (\gls{bmi} \textgreater{} 40) had 52\% and 62\% higher chance of dying from all cancers compared to normal weight adults (18.5 \textgreater{} \gls{bmi} \textgreater{} 25), respectively \citep{Calle2003}.
+\gls{bmi} was also associated with higher rates of death from esophageal, colorectal, liver, gallbladder, pancreas, kidney, and some sex-specific cancers \citep{Calle2003}.
+Furthermore, the death rates had a positive linear correlation with increasing \gls{bmi} in all cancers \citep{Calle2003}.
+
+In a more recent meta-analysis study by \citet{Renehan2008} showed that an increase \gls{bmi} by 5 kg/m$^2$ significantly increased the risks of esophageal, thyroid, colon and renal cancers in men; endometrial, gallbladder, esophageal and renal cancers in women.
+The summary of the risks associated with various cancers from this study are shown in \cref{tab:renehan_cancer_risks}.
+This study showed that the association between \gls{bmi} and cancers were sex-specific, applicable to wide range of cancer types, and generally consistent across geographic populations \citep{Renehan2008,Roberts2010}.
+
+Taken together, it is undeniable that obesity is a major risk factor for many types of cancers, and it may even be the cause for some of these cancers.
+
+\begin{table}[h]
+	\centering
+	\begin{threeparttable}
+		\caption{Summary of the risk ratios\tnote{1} that significantly associated with each cancer types from \citet{Renehan2008} study\tnote{2}.}
+		\label{tab:renehan_cancer_risks}
+		\begin{tabular}{lcc}
+													   & Men               & Women\\
+			\hline
+			\rule{0pt}{2.25ex}Colon                 & 1.24 (1.20, 1.28) & 1.09 (1.05, 1.13)\\
+			Rectum                                  & 1.09 (1.06, 1.12) & NA\tnote{3}\\
+			Gallbladder                             & NA                & 1.59 (1.02, 2.47)\\
+			Leukemia                                & 1.08 (1.02, 1.14) & 1.17 (1.04, 1.32)\\
+			Malignant melanoma                      & 1.17 (1.05, 1.30) & NA\\
+			Multiple myeloma                        & 1.11 (1.05, 1.18) & 1.11 (1.07, 1.15)\\
+			Non-Hodgkin lymphoma                    & 1.06 (1.03, 1.09) & 1.07 (1.00, 1.14)\\
+			Esophageal adenocarcinoma               & 1.52 (1.33, 1.74) & 1.51 (1.31, 1.74)\\
+			Pancreatic                              & NA                & 1.12 (1.02, 1.22)\\
+			Renal                                   & 1.24 (1.15, 1.34) & 1.34 (1.25, 1.43)\\
+			Thyroid                                 & 1.33 (1.04, 1.70) & 1.14 (1.06, 1.23)\\
+			Prostate                                & 1.03 (1.00, 1.09) & NA\\
+			Post-menopausal breast                  & NA                & 1.12 (1.08, 1.16)\\
+			Endometrial (\textless{}27 kg/m$^2$)    & NA                & 1.221 (1.084, 1.376)\\
+			Endometrial (\textgreater{}27 kg/m$^2$) & NA                & 1.729 (1.598, 1.872)\\
+		\end{tabular}
+		\begin{tablenotes}
+			\begin{footnotesize}
+				\item [1] All risk ratios are per increase in 5 kg/m$^2$ \gls{bmi}.
+				\item [2] This table is adapted from \citet{Roberts2010} paper, where the original data is taken from \citet{Renehan2008} paper.
+				\item [3] Not applicable.
+			\end{footnotesize}
+		\end{tablenotes}
+	\end{threeparttable}
+\end{table}
 

### 2016-12-06 thesis/intro/intro.tex
tmp commit

@@
 Finally, diet and obesity are known risk factors for many types of cancers \citep{Ames1995a,Calle2004}.
-Though more will be covered in details in \cref{sec:obesity_and_cancers}: \nameref{sec:obesity_and_cancers}, the effect of obesity on cancers will be briefly mentioned here.
+Though more will be covered in details in \cref{sec:obesity_and_cancer}: \nameref{sec:obesity_and_cancer}, the effect of obesity on cancers will be briefly mentioned here.
 
@@
 Earlier chemotherapeutic drugs and radiation therapy had limited effect on tumours due to their resistance against these treatments \citep{Wilhelm2006}.
-However, greater understanding of the pathways and mechanisms that tumour cells use have shed  light on the development of drugs that are more effective against tumours.
-RAF inhibitors are among one of many.
+However, better understanding of the pathways and mechanisms used by the tumour cells have shed  light on the development of drugs that are more effective against tumours.
+RAF inhibitors are among one of the many chemotherapeutic drugs.
 
 The RAS-RAF-MEK-ERK signalling pathway is a pathway involved in cell survival, growth and proliferation \citep{Samatar2014,Wilhelm2006}.
-The members of this pathway are often mutated in many types of cancers, causing an overactivation of this pathway.
-For example, approximately a third of cancers contain a mutation in the RAS oncogene which allows the constitutive activation of the RAS-ERK pathway \citep{Samatar2014}.
-Development of the drugs that target  the downstream RAF and its mutant forms (such as BRAF$^{V600E}$) have been successful, and many RAF inhibitor drugs such as sorafenib and vemurafenib have been approved for use clinically \citep{Samatar2014,Wilhelm2006}.
-Even though our understandings of signalling pathways and drug development based on these knowledge is improving, there is always the possibility of tumour to acquire resistance to those drugs and there is always a need for new ways to treat cancer patients \citep{Samatar2014}.
+The members of this pathway are often mutated in many types of cancers (approximately a third of cancers contain a mutation in the RAS oncogene), causing an overactivation of this pathway \citep{Samatar2014}.
+Development of drugs that target  the downstream RAF and its mutant forms (such as BRAF$^{V600E}$) have been successful, and many RAF inhibitor drugs such as sorafenib and vemurafenib have been approved for clinical use \citep{Samatar2014,Wilhelm2006}.
+Even though the signalling pathways are well understood and the quality of the drugs that are developed based on these knowledge is improving, there is always the possibility of tumour to acquire resistance to those drugs \citep{Samatar2014}.
 
-In fact, there has been evidence of pathway cross-talks between RAS-ERK pathway and \gls{pi3k}/Akt pathway \citep{Moelling2002,Zimmermann1999a}.
-Furthermore, a study in \textit{Drosophila} showed that the inhibition of just a single pathway does not necessarily stop the effect of the pathway, as the signal can flow through alternate pathways \citep{Dar2012}.
+In fact, there has been evidence of pathway cross-talks between RAS-ERK pathway and \gls{pi3k}/Akt pathway, which may help the tumour cells achieve resistance against those drugs \citep{Moelling2002,Zimmermann1999a}.
+A study in a \textit{Drosophila} model showed that the inhibition of just a single pathway does not necessarily stop the effect of the pathway, as the signal can flow through alternate pathways \citep{Dar2012}.
 Pathway cross-talk and increased flux through alternate pathways could be one of the ways in which tumour cells acquire resistance to drugs.
@@
 \noindent
-Though there are difficulties such as resistance to drugs and metastasis, further development of technologies and improved knowledge of tumour biology will only get us closer to the cure of cancer.
+There are many difficulties accompanied by cancer treatment such as resistance to drugs and metastasis, and there is always a need for new ways to treat cancer patients.
+Further understanding of the tumour biology, its mechanism of resistance to drugs, and the effect of targeting multiple pathways (or targeting effective pathways for a patient) will definitely become important in the near future, both for novel drug development and better clinical treatment plan for individual patients.
 
-\section{Obesity and cancers}
-\label{sec:obesity_and_cancers}
+\section{Obesity and cancer}
+\label{sec:obesity_and_cancer}
 
@@
 	\begin{threeparttable}
-		\caption{Summary of the risk ratios\tnote{1} that significantly associated with each cancer types from \citet{Renehan2008} study\tnote{2}.}
+		\caption{Summary of the risk ratios\tnote{1} that significantly associated with each type of cancers from \citet{Renehan2008} study\tnote{2}.}
 		\label{tab:renehan_cancer_risks}
@@
 			Endometrial (\textgreater{}27 kg/m$^2$) & NA                & 1.729 (1.598, 1.872)\\
+			\hline
+			\hline
 		\end{tabular}
@@
 
-\subsection{Mechanism of cancer progression in obese patient}
-\label{sub:mechanism_of_cancer_progression_in_obese_patient}
+\subsection{Mechanisms of cancer progression in obese patient}
+\label{sub:mechanisms_of_cancer_progression_in_obese_patient}
 
-% Why is it important in obesity context.
+The exact mechanisms in which obesity causes cancer have still not been fully elucidated yet.
+There are several hypotheses that links obesity with cancers: disruption of various hormone levels, chronic and sustained inflammation, and activation of the \gls{igf} pathway.
 
-% The exact mechanism in which obesity accelerates cancer progression is unknown, but obesity is thought to disrupt the levels of various hormones (such as insulin, leptin, adeponectin, and other adipokines) that eventually activate tumorigenic pathways (reference the overview on obesity and tumour resistance).
+Previously thought to be a group of cells where the body could store and manage excess fats, it has recently become apparent that adipose tissue is also a metabolically active organ that has a significant role in the endocrine system \citep{Roberts2010}.
+Increased adiposity is known to influence the bioavailability of endogenous sex hormones, including estrogens, androgens and progesterone \citep{Calle2004}.
+\Glspl{adipocyte} synthesise enzymes, such as aromatase, that convert precursor molecules into different forms of androgens and estrogens \citep{Calle2004}.
+Many studies have confirmed the mitogenic effect of sex hormones on certain cell types, mainly breast and endometrial cells \citep{Roberts2010}.
+
+On top of the effects that \glspl{adipocyte} have on circulating concentration of sex hormones, they also produce adipose tissue-specific hormones called adipokines \citep{Roberts2010}.
+Of the many adipokines identified so far, leptin and adipo\-nectin have been studied the most in terms of cancer development \citep{Renehan2006,Roberts2010}.
+Leptin is known to activate not only the satiety signalling pathway, but also the \gls{jak}/\gls{stat}, \gls{pi3k}/Akt signalling pathways \citep{Garofalo2006,Renehan2006}.
+Activation of these pathways lead to the activation of their target genes that are involved in cell proliferation and cell survival, as well as pro-angiogenic factors \citep{Garofalo2006}.
+
+Adiponectin is a hormone that is synthesised and secreted only by the \glspl{adipocyte} \citep{Kelesidis2006a}.
+Unlike leptin, where the concentration of leptin is higher in obese people, adiponectin levels are lower in obese people compared to the normal weight people \citep{Kelesidis2006a}.
 

### 2017-01-02 thesis/intro/intro.tex
Finished Obesity and cancer section draft

@@
 	\begin{threeparttable}
-		\caption{Summary of the risk ratios\tnote{1} that significantly associated with each type of cancers from \citet{Renehan2008} study\tnote{2}.}
+		\caption{Summary of the risk ratios\tnote{1} associated with each type of cancers from the meta-analysis study by \citet{Renehan2008}\tnote{2}.}
 		\label{tab:renehan_cancer_risks}
@@
 			\begin{footnotesize}
-				\item [1] All risk ratios are per increase in 5 kg/m$^2$ \gls{bmi}.
-				\item [2] This table is adapted from \citet{Roberts2010} paper, where the original data is taken from \citet{Renehan2008} paper.
+				\item [1] All risk ratios are per increase in 5 kg/m$^2$ \gls{bmi}; 95\% confidence intervals in brackets.
+				\item [2] Table adapted from \citet{Roberts2010} paper, where the original data in the table was taken from \citet{Renehan2008} study.
 				\item [3] Not applicable.
@@
 The exact mechanisms in which obesity causes cancer have still not been fully elucidated yet.
-There are several hypotheses that links obesity with cancers: disruption of various hormone levels, chronic and sustained inflammation, and activation of the \gls{igf} pathway.
+There are several hypotheses that links obesity with cancers: disruption of various hormone levels, chronic and sustained inflammation, and activation of the insulin/\gls{igf} pathway.
 
@@
 Adiponectin is a hormone that is synthesised and secreted only by the \glspl{adipocyte} \citep{Kelesidis2006a}.
-Unlike leptin, where the concentration of leptin is higher in obese people, adiponectin levels are lower in obese people compared to the normal weight people \citep{Kelesidis2006a}.
+Unlike leptin, where the concentration of leptin is higher in obese people, adiponectin levels are lower in obese people compared to the normal weight people \citep{Kelesidis2006a,Renehan2006}.
+Adiponectin has anti-angiogenic and anti-proliferative effect on the cells, and is secreted by mature adipocytes but not by the premature pre-adipocytes \citep{Gilbert2013}.
+During chronic inflammatory state in obese patients, pro-inflammatory proteins such as \gls{tgfb} and \gls{ifny} can block the maturation of the pre-adipocytes, thus decreasing the level of adiponectin and the accompanying benefecial effects \citep{Gilbert2013}.
+
+Alternatively, or possibly in conjunction with hormone disruption, inflammation caused by obesity may be associated with tumour-promoting environment.
+\Gls{tlr}4 is a type of \gls{prr} and, as mentioned earlier, it is able to recognise circulating free fatty acids and trigger immune response via \gls{nfkb} pathway \citep{Lumeng2011}.
+In addition to the \gls{nfkb} pathway, recent evidence have shown that \gls{tlr}4 could also activate the \gls{jak}/\gls{stat3} pathway \citep{Yu2014}.
+Activation of these pathways will further enhance the development of the tumour-promoting microenvironment caused by these unwanted inflammatory responses of obesity.
+Furthermore, since it is likely that obese people have less adiponectin concentration, obesity-induced inflammation may affect these patients more than those with normal weight.
+
+Finally, insulin/\gls{igf} pathway may have a major role in encouraging tumorigenesis in obese patients.
+\Gls{igf} system involves two ligands (\gls{igf}-I and \gls{igf}-II), its respective receptors (\gls{igf}-IR and \gls{igf}-IIR), and six binding proteins (\gls{igfbp} 1 to 6) \citep{Roberts2010}.
+\gls{igf} signalling results in the activation of \gls{pi3k}/Akt and \gls{erk} pathways, resulting in cell proliferation and reduced apoptosis \citep{Roberts2010}.
+Though insulin is also capable of activating these pathways, it is thought that the mitogenic effects are controlled mainly through the \gls{igf} receptors \citep{Roberts2010}.
+
+Since insulin reduces the level of \gls{igfbp}-3, the main circulating \gls{igfbp}, hyperinsulinemia has been hypothesised to promote tumorigenesis through the mitogenic effect of insulin and free \glspl{igf} \citep{Giovannucci1995,Mckeown1994,Roberts2010}.
+In fact, many studies have consistently associated the increased concentration of free \gls{igf}-I with obesity and increased risk of some cancers; however the results for \gls{igfbp} have been inconsistent \citep{Basen2011}.
+
+It is evident that there are many possible mechanisms in which obesity contributes to the progression of cancer.
+Furthermore, these mechanisms are related to one another and are part of a larger, more complex physiological network \citep{Renehan2006}.
+Due to this complexity, it is difficult to confidently explain how obesity contributes to tumour development.
 

### 2017-01-02 thesis/misc/glossary.tex
tmp commit
Finished Obesity and cancer section draft

@@
 	description={Showers of single nucleotide changes confined to local regions, often near sites of chromosome rearrangements \citep{Stephens2011}}
+}
+
+\newglossaryentry{adipocyte}{
+	name=Adipocyte,
+	text=adipo\-cyte,
+	plural=adipo\-cytes,
+	description={Adipose tissue cell}
 }
@@
 \newacronym{apobec}{APOBEC}{apolipoprotein B mRNA editing enzyme, catalytic polypeptide-like}
+\newacronym{nfkb}{NF-$\kappa$B}{nuclear factor $\kappa$B}
+\newacronym{stat3}{STAT3}{signal transducer and activator of transcription 3}
+\newacronym{il6}{IL-6}{interleukin-6}
+\newacronym{il10}{IL-10}{interleukin-10}
+\newacronym{atp}{ATP}{adenosine triphosphate}
+\newacronym{pep}{PEP}{phosphoenol pyruvate}
+\newacronym{mmr}{MMR}{mismatch repair}
+\newacronym{ner}{NER}{nucleotide-excision repair}
+\newacronym{ber}{BER}{base-excision repair}
+\newacronym{uv}{UV}{ultraviolet}
+\newacronym{pah}{PAH}{polycyclic aromatic hydrocarbon}
+\newacronym{nnk}{NNK}{4-(methylnitrosamino)-1-(3-pyridyl)-1-butanone}
+\newacronym{ros}{ROS}{reactive oxygen species}
+\newacronym{cyp}{CYP}{cytochrome P450}
+\newacronym{igf}{IGF}{insulin-like growth factor}
+\newacronym{igfbp}{IGFBP}{insulin-like growth factor binding protein}
+\newacronym{gpcr}{GPCR}{G-protein coupled receptor}
+\newacronym{tlr}{TLR}{Toll-like receptor}
+\newacronym{prr}{PRR}{pattern recognition receptor}
 \newacronym{ssdna}{ssDNA}{single-stranded DNA}
+\newacronym{erk}{ERK}{extracellular signal regulated kinase}
+\newacronym{ifny}{IFN-$\gamma$}{interferon-$\gamma$}
 \newacronym{tgfb}{TGF-$\beta$}{transforming growth factor-$\beta$}

### 2017-01-06 thesis/intro/intro.tex
Finished draft copy of introduction -- need to cut down on the contents and link everything together

@@
 
+One of the difficulties associated with the treatment of cancers is the identification of the underlying mechanisms or hallmarks that drive the cancer progression.
+As one can imagine, it is difficult for clinicians to treat a patient if they do not know what is causing the disease.
+The reason why it is so difficult to identify the underlying mechanisms is that there are very little reliable biological markers (or ``genetic signatures'') that signify the presence or absence of these pathways and/or hallmarks.
+With that said, significant improvements have been made in recent years owing to the advancement in sequencing technologies and have allowed researchers to study the genetic signatures that represent the hallmarks in a much greater detail.
+
 \subsection{Microarray and Next Generation Sequencing technologies}
@@
 
-% Due to advances in sequencing technology over the last few years, many researchers have started to focus on bioinformatics based analyses of cancer cells to better understand the biology of cancer.
-% Advantages of this strategy are that very specific gene expression and/or mutational signatures for certain cancer subtypes can be identified from the analyses, which the researchers and clinicians can study in much greater depth and details.
+\subsubsection{Microarray technology}
+\label{ssub:microarray_technology}
+
+Microarray technology was first developed by \citet{Schena1995} and has been used extensively to study gene expression patterns and \gls{snp} identification for over two decades.
+In short, microarray technology uses a glass slide with hundreds of thousands of \acrshort{dna} sequences attached to it, where fluorescently labelled \glspl{cdna} taken from the sample \glspl{mrna} are hybridised to these immobilised \acrshort{dna} sequences, and the fluorscent probes are detected \citep{Schena1995,Schulze2001}.
+Since only the exactly matching sequences remain on the microarray slide, researchers are able to identify and quantify the gene expression level of the sample \citep{Schulze2001}.
+
+One disadvantage of microarray technology is that it requires previous knowledge of the genes and \glspl{snp} of the samples being investigated \citep{Hurd2009}.
+This is due to the fact that the detection of sequences relies on the hybridisation of the sample sequences with the microarray sequences and therefore the sequences under investigation must be known before they are invesetigated.
+Another disadvantage of microarrays is the potentially erroneous detection of the sample sequences, both in terms of its sequence and its expression level \citep{Hurd2009}.
+Again, because the microarray slides require pre-defined sequences to be attached to it, there is always a possibility of similar sequences hybridising to the slide.
+Even if the sequences correctly hybridise to the slide, it may be difficult to measure the exact amount of some sequences in the samples.
+Since the amount of sequences in the sample is measured with the relative strength of the fluorescence of the dyes used, it is difficult to detect the exact level of expressions of sequences that are either very rare or highly abundant in the sample \citep{Hurd2009}.
+
+\subsubsection{Next generation sequencing technology}
+\label{ssub:next_generation_sequencing_technology}
+
+\Gls{ngs} technology refers to the high throughput sequencing techniques that allow sequencing of large quantity of genetic material from a sample \citep{Metzker2010}.
+\gls{ngs} technology has been developed and improved in the last decade at a revolutionary pace, making it possible for researchers to study gene expressions and variations in days to weeks, where the same experiment would have taken years to complete before the technology was developed.
+
+Every \gls{ngs} technology uses the same principle process to sequence \acrshort{dna} samples.
+First, \acrshort{dna} samples are fragmented into smaller sizes and universal primers are attached to the ends of the fragments.
+These fragments are then amplified into clusters to make the signals stronger for detection (although this amplification step varies between different techniques) \citep{Metzker2010}.
+After the amplification step, the clusters of \acrshort{dna} fragments are sequenced and ``imaged'' using fluorescently labelled \acrshort{dna} bases, where the incorporation of different bases are detected by its unique fluorescence \citep{Metzker2010}.
+
+\gls{ngs} technology have many advantages over microarray technology.
+Firstly, \gls{ngs} does not require any prior knowledge of the sample like in the microarray technology as all genetic materials are attached to universal primers when they are sequenced -- there is no need to pre-define the sequences to be searched and thus allow the detection of novel genes or transcripts \citep{Hurd2009}.
+This also prevents the introduction of experimental bias, as the sequences are not altered in any way and no specific sequence is searched over the others \citep{Hurd2009}.
+Another advantage is that \gls{ngs} produce count data of the genetic material being sequenced, instead of relying on the intensity of the fluorescence to detect the presence and amount of certan sequence.
+This allows an absolute quantification of the amount of transcripts from the sample, making it possible for a more accurate comparison of the expression levels between samples.
+Lastly, many \gls{ngs} techniques are capable of sequencing multiple samples simultaneously with high throughput, generating tremendous amount of genetic data in a single run.
+
+The use of \gls{ngs} technology have significantly improved the workflow of modern genetic studies and its application in variety of research area have made it possible to dig deeper into the causes of complex diseases, such as cancer.
+With the cost of sequencing declining, personalised genome sequencing and treatment based on the sequence will be a reality in the near future.
 
@@
 
+There is a clear evidence of obesity being a major risk factor for many cancer types, but what genes or pathways directly contribute to tumorigenesis in patients that are obese is yet to be clarified.
+Two studies in particular, one by \citet{Creighton2012} and another by \citet{Fuentes-Mattei2014}, have investigated if there were any difference in genetic composition between obese and normal weight patients, specifically in a group of breast cancer patients.
+In both studies, the authors were able to identify a set of genes that were up or downregulated in breast tumours from obese patients.
+% For simplicity, these genetic signatures will be referred to as ``obesity associated genetic signatures'' hereafter.
+
 \subsubsection{\citet{Creighton2012} study}
 \label{ssub:creighton_study}
+
+In this study, \citet{Creighton2012} used 103 breast tumour samples and analysed these samples using microarrays.
+Of the 103 samples, 35\% were from normal weight, 28\% were from overweight and 37\% were from obese patients.
+They compared the gene expression between the tumours from obese patients with the tumours from non-obese patients and identified 799 probe sets (662 unique genes) with p \textless{} 0.01 and with a \gls{fc} \textgreater{} 1.2.
+Majority of these genes (725 gene probes) were downregulated and 74 gene probes were upregulated in the samples from obese patients compared with the samples from non-obese patients.
 
@@
 
+In the study conducted by \citet{Fuentes-Mattei2014}, they obtained 137 breast tumour samples from \gls{er} positive patients and analysed their transcriptome using microarrays.
+Of the 137 patients, 94 patients were not obese and 43 patients were obese.
+Comparison of the gene expression of the tumour samples from obese patients with the samples from non-obese patients revealed 112 genes that were statistically significantly dysregulated (62 genes upregulated and 50 genes downregulated).
+Furthermore, functional transcriptomic analysis suggested that insulin signalling and inflammation affected human \gls{er}+ breast cancer.
+
+On top of this, \citet{Fuentes-Mattei2014} validated their results using a mouse model, where they compared the expression level in the breast tumours from obese mice with the samples from lean mice.
+Results from this animal model experiment provided strong evidence of obesity significantly accelerating tumour growth through Akt/\gls{mtor} signalling pathway.
+
 \subsection{Pathway associated genetic signatures}
@@
 
-% Talk about \citet{Gatza2010a}
+Like obesity associated genetic signatures, there has also been a focus on the identification of pathways and genes that affect tumour growth.
+About a decade ago, \citet{Bild2006} demonstrated that a collection of genes that are involved in a specific pathway could be used to determine the status of those pathways in specific samples.
+These pathway specific genes were identified experimentally by altering the expression of one specific gene that is central to the pathway in a model cell culture system.
+For example, all of the genes that have altered its expression in the cells that have \textit{Myc} gene induced were noted down as the gene involved in the Myc pathway.
 
-\subsubsection{\citet{Gatza2011} study}
+\subsubsection{\citet{Gatza2010a} study}
 \label{ssub:gatza_study}
+
+In this study, \citet{Gatza2010a} used 18 pathway signatures to classify human breast tumours into more detailed subtypes.
+This was a significant improvement from their previous works on pathway signatures \citep{Bild2006,Bild2009}.
+Using more pathways in their analysis have allowed them to identify certain pathways to coactivate (or cluster) together, suggesting that some pathways are related to one another and may have relevance in identifying the causes.
 
@@
 Better understanding of the pathways being dysregulated in cancer cells in obese patients may lead to improved clinical decisions, and contribute towards personalised treatment in the future.
+
+In the following chapter, the detailed methods of all the analyses carried out for this project is outlined.
+The results from these analyses are split into two parts, each presented in separate chapters: first part of the results clarify the relationship between the obesity associated genetic signatures with various cancer types; and the second part of the results consist of the links between the pathway associated genetic signatures and various cancer types.
+By the end of this thesis, the aim presented here will be answered with clarity.
+However, it is without doubt that there will be many more questions to be answered than what was actually answered -- those questions shall remain unanswered for future investigations.
+
+% TODO: Finished draft copy of introduction -- need to cut down on the contents and link everything together

### 2017-01-10 thesis/methods/methods.tex
tmp commit
Start on data processing section in methods

@@
 \label{ch:methods}
+
+\section{\gls{r} -- statistical programming language}
+\label{sec:r}
+
+All statistical analyses and data manipulation were carried out with \gls{r} (version 3.3.2 -- ``Sincere Pumpkin Patch''), a free open-source programming language and software environment for statistical computing and graphics \citep{R2016}.
 
@@
 Samples were classified based on the \gls{who} definition, as shown in \cref{tab:whobmiclass}.
-\begin{table}[hb]
+\begin{table}[htb]
 	\caption{\gls{who} defined \gls{bmi} classification}
@@
 		\begin{tabular}{lc}
-			\bfseries {Classification} & \bfseries {\gls{bmi} Value}\\
+			Classification & \gls{bmi} Value\\
 			\hline
@@
 			Overweight & 25.0$\sim$29.9\\
-			Obese & \geq{} 30.0\\
+			Obese & $\geq{}$ 30.0\\
+			\hline
+			\hline
 		\end{tabular}
@@
 
-% TODO: add Cris Print's paper
 The raw microarray data from \citet{Creighton2012}, \citet{Fuentes-Mattei2014} and \citet{Gatza2010a}  studies were downloaded from the \gls{geo} website.
 \gls{rnaseq} and clinical data of multiple different cancer types were downloaded from free and publicly available sources such as \gls{icgc} and \gls{tcga}, respectively.
+\\
 
-\subsection{Raw data from \citet{Creighton2012} study}
-\label{subsec:crrawdat}
+\noindent
+The raw Affymetrix HGU\-133A microarray gene expression data files from the \citet{Creighton2012} study were downloaded from the \gls{geo} database (\gls{geo} accession ID: GSE24185).
+Clinical data of the samples (age, ethnicity, tumour grade, menopause status, \gls{bmi}, \gls{er} status, \gls{pr} status, \gls{her2} status, and \gls{ln} status) were obtained from the supplementary table 1 from \citet{Creighton2012} paper.
+799 obesity associated gene probes identified in the \citet{Creighton2012} study were obtained from the supplementary data file 1 from \citet{Creighton2012} paper.
 
-The raw microarray gene expression data files from the \citet{Creighton2012} study were downloaded from the \gls{geo} database (\gls{geo} accession ID: GSE24185).
-Clinical data of the samples were obtained from the supplementary table 1 from \citet{Creighton2012} paper.
-799 obesity associated gene probes identified in the \citet{Creighton2012} study were obtained from the supplementary data file 1 from \citet{Creighton2012} paper.
-The raw expression data, clinical data, and the list of gene probes were imported into \gls{r}.
-
-\subsection{Raw data from \citet{Fuentes-Mattei2014} study}
-\label{subsec:fmrawdat}
-
-The raw microarray gene expression data files from the  \citet{Fuentes-Mattei2014} study were downloaded from the \gls{geo} database (\gls{geo} accession ID: GSE\-20194).
-Clinical data for the samples in this study was also downloaded from the \gls{geo} database (same \gls{geo} accession ID).
-130 obesity associated gene probes identified by \citet{Fuentes-Mattei2014} were taken from the supplementary table 3.
-The raw expression data and the list of gene probes were imported into \gls{r}.
-
-\subsection{Raw data from \citet{Gatza2010a} study}
-\label{sub:gatzarawdat}
+The raw Affymetrix HGU\-133A microarray gene expression data files from the  \citet{Fuentes-Mattei2014} study were downloaded from the \gls{geo} database (\gls{geo} accession ID: GSE\-20194).
+Clinical data for the samples (age, ethnicity, tumour grade, \gls{er}/\gls{pr}/\gls{her2} statuses, and treatments used) in this study was also downloaded from the \gls{geo} database (same \gls{geo} accession ID).
+130 obesity associated gene probes identified by \citet{Fuentes-Mattei2014} were taken from the supplementary table 3 from their paper.
 
 The raw microarray gene expression data files from the \citet{Gatza2010a} study were downloaded from the \gls{geo} database (\gls{geo} accession ID: GSE1456, GSE\-1561, GSE2034, GSE3494, GSE4922, and GSE6596).
-Only the Affymetrix HGU\-133A microarray samples were included in the project.
-Clinical data for the samples in this study was not available.
-The raw expression data and the list of gene probes were imported into \gls{r}.
-
-\subsection{Raw data from Print et al study}
-\label{subsec:printrawdat}
+Only the Affymetrix HGU\-133A microarray samples were included in this project, as other microarray data were analysed using the Affymetrix HGU-133A platform.
+Clinical data for the samples in \citet{Gatza2010a} study was not available, as these samples were a combination of many different datasets.
 
@@
 
-The raw microarray gene expression data files from the  Print et al. study were downloaded from the \gls{geo} database (\gls{geo} accession ID: GSE36771).
-Clinical data for the samples in this study was ******.
-The raw expression data and the list of gene probes were imported into \gls{r}.
-
+The raw microarray gene expression data files from the (Cris' paper) study were downloaded from the \gls{geo} database (\gls{geo} accession ID: GSE36771).
+Clinical data for the samples (age, ethnicity, tumour grade, breast cancer subtype, \gls{er}/\gls{pr} statuses, \gls{ln} status, \gls{bmi} and treatments used) in this study was taken from (Cris' paper).
+\\
 % TODO: Overall summary table of the samples from each data set at the end of this section.
 
-\subsection{\gls{tcga}}
-\label{subsec:tcga}
-
-The clinical data for all available cancer types (33 types in total) were downloaded from \gls{tcga} database and were checked for the height and weight data for each sample.
-Any cancer type with no weight and/or height data of the samples were excluded from the project, as no \gls{bmi} information can be obtained without these data.
-
-This gave a total of eight cancer types: \gls{blca}, \gls{cesc}, \gls{coad}, \gls{kirp}, \gls{lihc}, \gls{read}, \gls{skcm}, \gls{ucec}.
-All eight cancer types were used for further analyses.
-
-\subsection{\gls{icgc}}
-\label{subsec:icgc}
-
-\gls{rnaseq} data for the eight cancer types from \cref{subsec:tcga} were downloaded from the \gls{icgc} database.
-The \gls{rnaseq} data for each of the eight cancer types were imported into \gls{r}/RStudio.
+\noindent
+The clinical data for all available cancer types (33 types in total) were downloaded from \gls{tcga} database (last accessed 1 April 2015) and were checked for both the height and weight data for each sample.
+Any cancer type with no height and/or weight data of the samples were excluded from the project, as no \gls{bmi} information can be obtained without these data.
+Out of these 33 cancer types with clinical data, 14 cancer types had both height and weight data.
+However, only 8 cancer types out of these 14 types had \gls{rnaseq} data available from the \gls{icgc} database (last accessed 7 September 2015), so only those 8 cancer types were downloaded and used in this project.
+The selected cancer types were: \gls{blca}, \gls{cesc}, \gls{coad}, \gls{kirp}, \gls{lihc}, \gls{read}, \gls{skcm}, \gls{ucec}.
 
@@
 \label{sub:data_normalisation}
+
+% TODO: start data normalisation section
 

### 2017-01-11 thesis/methods/methods.tex
Finished data normalisation draft

@@
 Clinical data for the samples in \citet{Gatza2010a} study was not available, as these samples were a combination of many different datasets.
+(Find out where the 18 pathway signatures were downloaded from)
 
@@
 
-% TODO: start data normalisation section
+All of the data were normalised to remove any experimental bias, errors and noise, so only the true biological signals are considered in the analyses.
+Any experimental procedure is prone to errors due to differences in experimental conditions, machinery used to measure signals and technical procedures in different laboratories, just to name a few.
+In order to remove these experimental noises and focus on the true biological signals in the raw data, the data must be normalised one way or another.
 
@@
 
-The raw microarray data were normalised using either the \gls{rma} or \gls{mas} method from the \textit{affy} package in \gls{r}.
-Either \gls{rma} or \gls{mas} normalisation was used to account for the inter-array differences that may have arisen during the data collection process.
+In microarray experiments, there are two types of probes present on the microarray chips: \gls{pm} and \gls{mm} probes \citep{Irizarry2003}.
+As the name suggests, \gls{pm} probes represent the probes that should perfectly match the gene of interest, whereas \gls{mm} probes have there 13th base pair intentionally altered to measure non-specific binding of the gene \citep{Irizarry2003}.
+Various normalisation methods make use of \gls{pm} and \gls{mm}, or ``probe pairs'', to identify true signals from the noise.
+
+\Gls{mas} and \gls{rma} normalisation methods are both part of the \textit{affy} package in \gls{r} \citep{Gautier2004}.
+\Gls{mas} uses a difference-based method where it subtracts a value derived from \gls{mm} from \gls{pm}, but this approach may introduce additional noise and/or errors in some cases \citep{Irizarry2003}.
+\Gls{rma} method is based on the observation that \gls{pm} is a mixture of background and true signal, and uses mathematical models to estimate the expression while correcting for background signals from \gls{pm} only \citep{Irizarry2003}.
+In fact, the \gls{rma} method showed better identification of the true signals than the other methods, including \Gls{mas} method \citep{Irizarry2003}.
+
+The raw microarray data were normalised using both \gls{rma} and \gls{mas} methods, each done separately on a copy of the data.
+The reason why \gls{mas} method was used as well as the \gls{rma} method was becuase the \gls{mas} normalisation method was used in \citet{Gatza2010a} study.
+Though the normalisation methods were not specified in the \citet{Creighton2012} and \citet{Fuentes-Mattei2014} studies, to allow for data comparison between different datasets and accurate result validation of some of the studies, both \gls{rma} and \gls{mas} methods were used to normalise the microarray data.
 
@@
 
+\gls{rnaseq} data are fundamentally different to the microarray data, as \gls{rnaseq} data is a count data of all the sequences, while microarray data is a continuous data of the intensity of the matching probes.
+This means that \gls{rnaseq} data must be processed in a different manner as the microarray data.
+However, by analysing \gls{rnaseq} data as a count data, it limits range of statistical tools and types of analyses that can be done on the data, as many tools were designed for normally distributed data \citep{Law2014}.
+
+To overcome this limitation, \citet{Law2014} developed a method called ``variance modelling at the observational level'', or voom, that allows any \gls{rnaseq} data to be used in any existing statistical analysis pipeline that is precision weight aware, including pipelines used for microarray data analyses.
+In brief, voom first constructs a standard deviation trend from the logged \gls{cpm} value of the genes (experimental design, treatment conditions and other factors are taken into account).
+This trend is then used to interpolate the standard deviation of the observation based on its predicted count size, and the inverse square of the predicted standard deviation is used as the weight  for that observations.
+The weights of the observations and the logged \gls{cpm} can then be used in other statistical pipelines that allows the input of quantitative weights \citep{Law2014}.
+
 The raw \gls{rnaseq} data were normalised in two different ways, depending on the analysis.
-For gene expression analyses, \texttt{voom} function from the \textit{limma} package was used to normalise the data.
-For the purposes of data visualisation or application of metagene transformation matrices, the data had 1 added and then logged to the base of 10.
+For gene expression analyses, voom normalisation method from the \textit{limma} package in \gls{r} was used to normalise the data \citep{Ritchie2015}.
+For the purposes of data visualisation or application of metagene transformation matrices on the \gls{rnaseq} data, the raw data had 1 added (to prevent logging of 0) and then logged to the base of 10 to allow for better visualisation.
 

### 2017-01-13 thesis/methods/methods.tex
Finished batch correction section of methods

@@
 
-\section{\gls{bmi}}
+\section{\Gls{bmi}}
 \label{sec:bmi}
@@
 
-\cref{eq:bmicalc}  was used to calculate the \gls{bmi} from the height and weight data of the samples.
+Where missing, the \gls{bmi} of the samples were calculated from the clinical data using \cref{eq:bmicalc}.
 
@@
 Clinical data for the samples (age, ethnicity, tumour grade, breast cancer subtype, \gls{er}/\gls{pr} statuses, \gls{ln} status, \gls{bmi} and treatments used) in this study was taken from (Cris' paper).
+
+All of the microarray data associate the genes with the corresponding gene probe IDs, and therefore these IDs had to be converted into their corresponding gene symbols.
+The gene probe IDs in the raw data were converted into their corresponding gene symbols using the \textit{hgu133a.db} package in \gls{r} \citep{hgu133}.
+Since multiple gene probes matches back to a single gene of interest in a microarray chip, there were conflicting expression data for some of the genes after the conversion of the gene probe IDs into gene symbols.
+For the gene symbols that had multiple data entries, a single data was chosen to represent the gene symbol by using the \texttt{collapseRows} function in the \textit{WGCNA} package in \gls{r} \citep{Langfelder2008}.
+Likewise, any obesity associated or pathway associated gene probes were converted into gene symbols.
 \\
-% TODO: Overall summary table of the samples from each data set at the end of this section.
+
+% TODO: Overall summary table of the samples from each data set at the end of this section (?).
 
@@
 The selected cancer types were: \gls{blca}, \gls{cesc}, \gls{coad}, \gls{kirp}, \gls{lihc}, \gls{read}, \gls{skcm}, \gls{ucec}.
+
+The raw \gls{rnaseq} data from \gls{icgc} database were formatted in a way that the count data of all the genes were listed for one sample, then the count data for all the genes for the next sample, and so on.
+This data format was highly inconvenient for later analyses so the data were reformatted into gene by sample matrix using the \textit{dplyr} package in \gls{r}.
+
+Another problem with the data was the sample ID in the \gls{rnaseq} data.
+Though similar to the \gls{tcga} sample IDs, the \gls{icgc} IDs in the \gls{rnaseq} data had extra identification code in the sample names.
+To associate each sample in the raw \gls{rnaseq} data from the \gls{icgc} database with the correct samples in the clinical data from the \gls{tcga} database, the \gls{icgc} IDs were stripped so they matched the \gls{tcga} IDs.
+After the IDs were stripped, all of the samples were checked to see if there were any duplicates in either the \gls{icgc} \gls{rnaseq} data or \gls{tcga} clinical data.
+Where there was a duplicate in the sample ID, a single sample was chosen to represent that particular ID by using the \texttt{collapseRows} function in the \textit{WGCNA} package in \gls{r} \citep{Langfelder2008}.
+
+Since some of the samples did not have either height or weight data, or both in the clinical data, these samples were removed from the analyses.
+In each of the eight cancer types, the \gls{tcga} sample ID in the clinical data were cross-checked with the sample ID in the \gls{rnaseq} data, and vice versa.
+Any sample that did not have either clinical information or \gls{rnaseq} data were removed from the analyses.
+This ensured that both clinical and \gls{rnaseq} data were available for all of the samples that were included in the study.
+See \cref{tab:samplesize} for a summary of the number of samples included in the analyses for each cancer type.
+
+%TODO: May need more details on the sample (bmi status, etc.)
+\begin{table}[h]
+	\caption{Summary of the total number of samples included in the analyses for each cancer type}
+	\label{tab:samplesize}
+	\begin{center}
+		\begin{tabular}{lc}
+			Cancer Type   & Number of samples \\
+			\hline
+			\rule{0pt}{2.25ex}BLCA & 261   \\
+			CESC                   & 224   \\
+			COAD                   & 226   \\
+			KIRP                   & 124   \\
+			LIHC                   & 264   \\
+			READ                   & 73    \\
+			SKCM                   & 218   \\
+			UCEC                   & 482   \\
+			\hline
+			\hline
+		\end{tabular}
+	\end{center}
+\end{table}
 
@@
 For gene expression analyses, voom normalisation method from the \textit{limma} package in \gls{r} was used to normalise the data \citep{Ritchie2015}.
-For the purposes of data visualisation or application of metagene transformation matrices on the \gls{rnaseq} data, the raw data had 1 added (to prevent logging of 0) and then logged to the base of 10 to allow for better visualisation.
+For the purposes of data visualisation or application of metagene transformation matrices on the \gls{rnaseq} data, the raw data had 1 added (to prevent logging of 0) and then logged to the base of 10.
 
@@
 
-The data was standardised so that each gene in the data had a mean ($\mu$) of 0 and a standard deviation ($\sigma$) of 1.
-This ensured that all of the genes were on the same scale as one another, which allowed direct comparison across all of the genes.
+For metagene creation and data visualisation that used heatmaps, the data were standardised so that each gene in the data had a \gls{m} of 0 and a \gls{sd} of 1.
+Since the expression levels of the genes could vary significantly (some may have very low expression, whereas another may have very high expression), the direct comparison of the raw expression values between different genes was not feasible.
+Standardisation of each gene allowed the expression levels to be within acceptable range or scale, thus allowed for better visualisation with heatmaps and comparison between different genes was made possible.
 
@@
 
-A linear model was constructed from the data with all of the clinical variables controlled for.
-The residual data was taken from this linear model to be used as the residual data for that data set.
+In order to analyse the data without the effect of certain clinical variables on the data, the residual data of the data was used in some analyses.
+For example, a gene from the raw data might have a strong association with \gls{bmi}, but it is possible that other clinical variables such as \gls{er} status and/or tumour grades are also associated with the gene.
+To focus only on the association with certain variables, the effect of those variables that may affect the analyses must be removed from the data.
 
-\subsection{Gene probe ID conversion}
-\label{sub:gene_probe_id_conversion}
-
-Where appropriate, the gene probe IDs in the raw data were converted into their corresponding gene symbols using the \textit{hgu133a.db} package in \gls{r}.
-Where there was an overlap in the gene symbols, a single data was chosen to represent that particular gene symbol by using the \texttt{collapseRows} function in the \textit{WGCNA} package in \gls{r}.
-Likewise, the obesity associated gene probes were also converted into gene symbols where required.
+To do this, a linear model was created from the data with all of the clinical variables that had to be controlled for; in other words, all of the clincial variables apart from the variables of interest were included in the linear model.
+Once a linear model was fitted to the data, the remaining data, or the ``residual data'', represented the data that had been corrected for all of the unwanted variables.
+This data was used in some analyses that required focus on certain variables without the other variables affecting the result.
 
@@
 
-When there were more than one microarray data to be used as a single data set, these microarray data were normalised separately then combined together with the \texttt{ComBat} function in the \textit{sva} package in \gls{r}.
-This corrected for the effect of the individual microarray data on the whole data when combined together.
+It is most likely that experiments are done at different time period, location and laboratory environment.
+These differences between experiments introduces systematic non-biological differences or ``batch effects'' into the data, making it difficult to directly compare the data from different batches \citep{Johnson2007}.
+The problem with the batch effect is that the normalisation methods do not control and adjust for the effect \citep{Johnson2007}.
+Fortunately, \citet{Johnson2007} developed a method to correct for batch effects in the data, using an empirical Bayesian framework.
 
-\subsection{Sample \gls{bmi} calculation}
-\label{sub:sample_bmi_calculation}
-
-Where missing, the \gls{bmi} of the samples were calculated from the clinical data using \cref{eq:bmicalc}.
-
-\subsection{\gls{icgc} data format conversion}
-\label{sub:icgc_data_format_conversion}
-
-The raw \gls{rnaseq} data for each cancer types were formatted using the \textit{dplyr} package in \gls{r}, so that the rows corresponded to the genes and the columns to the samples.
-
-\subsection{Sample ID conversion}
-\label{sub:sample_id_conversion}
-
-\gls{icgc} sample ID in the \gls{rnaseq} data were stripped so that the sample ID matched with the \gls{tcga} ID in the clinical data.
-After the IDs were stripped, samples were checked for any duplicates.
-Where there was a duplicate in the sample ID in the \gls{rnaseq} data, a single sample was chosen to represent that particular ID by using the \texttt{collapseRows} function in the \textit{WGCNA} package in \gls{r}.
-
-\subsection{\gls{icgc} sample exclusion}
-\label{sub:icgc_sample_exclusion}
-
-Some of the samples did not have either height or weight data, or both.
-These samples were removed from the analyses.
-In each of the eight cancer types, the \gls{tcga} sample ID in the clinical data were cross-checked with the sample ID in the \gls{rnaseq} data, and vice versa.
-Any sample that did not have either clinical or \gls{rnaseq} data were removed.
-This ensured that both clinical and \gls{rnaseq} data were available for all of the samples that were included in the analyses.
-See \cref{tab:samplesize} for a summary of the number of samples included in the analyses for each cancer type.
-
-%TODO: May need more details on the sample (bmi status, etc.)
-\begin{table}[h]
-	\caption{Summary of the total number of samples included in the analyses for each cancer type}
-	\label{tab:samplesize}
-	\begin{center}
-		\begin{tabular}{cc}
-			\textbf{Cancer Type}   & \textbf{Total number of samples} \\
-			\hline
-			\rule{0pt}{2.25ex}BLCA & 261   \\
-			CESC                   & 224   \\
-			COAD                   & 226   \\
-			KIRP                   & 124   \\
-			LIHC                   & 264   \\
-			READ                   & 73    \\
-			SKCM                   & 218   \\
-			UCEC                   & 482   \\
-		\end{tabular}
-	\end{center}
-\end{table}
+Since the data used in \citet{Gatza2010a} study was a combination of multiple microarray data from various studies, the batch effect had to be corrected before any analysis was carried out.
+Each microarray dataset was normalised separately then combined together into a single dataset, and the batch effect was corrected with the \texttt{ComBat} function (an implementation of the batch correcting method by \citet{Johnson2007}) from the \textit{sva} package in \gls{r} \citep{Leek2012}.
 
@@
 
-\section{\gls{svd}}
-\label{sec:singular_value_decomposition}
+\section{Metagene analysis}
+\label{sec:metagene_analysis}
+
+\gls{svd} was applied to the training data set to create the metagene and the transformation matrix in the data set.
+The transformation matrix from the training data was used to transform the other data sets and to create the metagene of the genetic signature in those data sets.
+
+Each of these metagenes were plotted in a heatmap to see the association between the metagene and the overall gene expression of the genes that were used to create the metagene.
+The metagene was also checked for the association between the sample \gls{bmi} status and \gls{bmi} value, which was plotted in a box plot and a scatter plot, respectively.
+
+% \section{\Gls{svd}}
+% \label{sec:singular_value_decomposition}
 
@@
 
-\section{Transformation matrix}
-\label{sec:transformation_matrix}
+% \section{Transformation matrix}
+% \label{sec:transformation_matrix}
 
@@
 The transformation matrix $U'D^{-1}$ was created from the \gls{svd} components, and was used to transform other data set and obtain the metagene in the data set.
-
-\section{Metagene analysis}
-\label{sec:metagene_analysis}
-
-\gls{svd} was applied to the training data set to create the metagene and the transformation matrix in the data set.
-The transformation matrix from the training data was used to transform the other data sets and to create the metagene of the genetic signature in those data sets.
-
-Each of these metagenes were plotted in a heatmap to see the association between the metagene and the overall gene expression of the genes that were used to create the metagene.
-The metagene was also checked for the association between the sample \gls{bmi} status and \gls{bmi} value, which was plotted in a box plot and a scatter plot, respectively.
 

### 2017-01-16 thesis/methods/methods.tex
tmp commit

@@
 
-The samples were grouped based on the sample BMI status (obese or non-obese group), and a linear model was fitted to the data based on the sample grouping.
-Empirical Bayes statistics was used to identify the differentially expressed genes between the two groups, and the top scoring genes were listed.
-The error rate of the genes were \gls{fdr} controlled for multiple hypothesis testing.
+For gene expression analysis, or differential expression analysis, the \textit{limma} package in \gls{r} was used \citep{Ritchie2015}.
+Since there were thousands of genes to be hypothesis tested, adjustment for multiple hypothesis testing had to be considered in order to identify the truly \glspl{deg}.
+
+\subsection{Limma}
+\label{sub:limma}
+
+\textit{limma} is a package that contains variety of tools to analyse microarray data using linear models-based methods, developed by \citet{Ritchie2015}.
+For gene expression analysis, \textit{limma} package fits a linear model in a gene-wise manner and produce test statistics that allow the assessment of whether the gene is differentially expressed or not \citep{Ritchie2015}.
+
+Before the data was analysed using the \textit{limma} analysis pipeline, it was normalised as described in \cref{sub:data_normalisation}: \nameref{sub:data_normalisation}.
+In addition to this, a design matrix that describes the experimental design was created from the clinical data.
+To create the design matrix, the samples were divided into two groups, the obese group and the non-obese group, and the constructed group information was used in the \texttt{model.matrix} function (in \textit{limma} package) to form the design matrix.
+
+\texttt{lmFit} function (\textit{limma}) was used to fit a linear model to the normalised data, using the experiental design information from the design matrix.
+The output of this function was used in the \texttt{eBayes} function (\textit{limma}) to identify the \glspl{deg} from the data.
+In \texttt{eBayes} function, statistical parameters are estimated from the data and these parameters are used in the empirical Bayesian approach to calculate the summary statistics used for the ranking and identification of \glspl{deg} \citep{Smyth2004}.
+
+The summary statistics from the \texttt{eBayes} function can be displayed with the \texttt{topTable} function (\textit{limma}) and includes: the estimate of the fold change of the gene expression in log$_2$, average gene expression level in log$_2$, moderated \textit{t}-statistic, raw p-value of the gene, multiple hypothesis testing adjusted p-value, and \textit{B}-statistic.
+The first value shows the estimate of the log$_2$ fold change of the gene expression compared with the reference group, so this represents the log$_2$ fold change of the gene expression in the obese samples relative to the non-obese group.
+Second value presents the log$_2$ value of the average expression of the gene across all of the arrays/samples.
+Moderated \textit{t}-statistic is the same as normal \textit{t}-statistic, but its standard error has been adjusted by a simple Bayesian method \citep{Smyth2005}.
+Raw p-value and adjusted p-value represent the p-value of the gene before and after it has been corrected for multiple hypothesis testing, respectively.
+Lastly, the value of \textit{B}-statistic represents the log-odds that the gene is differentially expressed, where \textit{B}-statistic of 0 shows that there is a 50\% chance the gene is differentially expressed \citep{Smyth2005}.
+
+From these summary statistics, the most likely \glspl{deg} were chosen from the list of significant genes by setting the threshold of the p-value to be either less than 1\% or 5\%.
+In the case where there were more than 1000 \glspl{deg}, the top 799 genes were picked from the list, as this was the number of genes found in the \citet{Creighton2012} study; otherwise, as many significant genes identified were taken from the list.
+
+\subsection{Multiple hypothesis testing correction}
+\label{sub:multiple_hypothesis_testing_correction}
+
+In an experiment where there are multiple hypotheses being tested, the rate or proportion of \gls{type1} appearing from the experiment must be controlled.
+The probability of \gls{type1} occuring for a single hypopthesis is usually controlled at some significance threshold $\alpha$, which is usually set at 0.05 \citep{Shaffer1995}.
+In a typical microarray experiment, there are over 20,000 gene probes to be tested for differential expression, and with a significance threshold of $\alpha = 0.05$, this would yield approximately 1,000 \glspl{type1}.
+In other words, 1,000 gene probes would be identified as differentially expressed, when in fact they are not.
+
+There are two broad classes of methods to correct for this problem: \gls{fwer} control and \gls{fdr} control.
+With \gls{fwer} control, the method primarily aims to set the $\alpha$-value for each hypothesis testing ($\alpha_i$) such that the sum of all the $\alpha_i$ is equal to $\alpha$ \citep{Hochberg1987,Shaffer1995}.
+Usually, $\alpha_i$ is set to $\frac{\alpha}{n}$, where $n$ is the number of hypothesis tests carried out in the experiment \citep{Shaffer1995}.
+This highly conservative approach significantly improves the certainty of the result from the experiment, but at the same time it significantly increases the likelihood of missing the truly \glspl{deg}, or \glspl{type2}.
+
+In contrast to the conservative \gls{fwer} control methods, the \gls{fdr} method developed by \citet{Benjamini1995a} control the \glspl{type1} while maintaining statistical \gls{power}.
+The \gls{fdr} method controls the ``expected proportion of errors among the rejected hypotheses'' by adjusting the $\alpha$ to an alternative value, $q^*$, which is calculated from the p-values of the hypothesis tests \citep{Benjamini1995a}.
+With \gls{fdr}, the p-values are ranked and ordered from the lowest to the highest p-value,
 

### 2017-01-16 thesis/misc/glossary.tex
Finished draft copy of introduction -- need to cut down on the contents and link everything together
Finished data normalisation draft
Finished batch correction section of methods
tmp commit

@@
 
+\newglossaryentry{type1}{
+	name=Type I error,
+	text=type I error,
+	plural=type I errors,
+	description={Incorrect rejection of a true null hypothesis; False positive}
+}
+
+\newglossaryentry{type2}{
+	name=Type II error,
+	text=type II error,
+	plural=type II errors,
+	description={Incorrect retaining of a false null hypothesis; False negative}
+}
+
+\newglossaryentry{power}{
+	name=Power,
+	text=power,
+	description={Probability that it will reject a false null hypothesis; It is inversely related to the probability of making a type II error}
+}
+
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 \newacronym{icgc}{ICGC}{International Cancer Genome Consortium}
-\newacronym{bmi}{BMI}{Body Mass Index}
+\newacronym{bmi}{BMI}{body mass index}
 \newacronym{who}{WHO}{World Health Organization}
-\newacronym{ln}{LN}{Lymph node}
-\newacronym{pr}{PR}{Progesterone receptor}
-\newacronym{er}{ER}{Estrogen receptor}
-\newacronym{her2}{HER2}{Human epidermal growth factor receptor 2}
-\newacronym{svd}{SVD}{Singular value decomposition}
-\newacronym{rma}{RMA}{Robust Multi Array}
-\newacronym{mas}{MAS5}{Microarray Analysis Suite 5.0}
-\newacronym{fdr}{FDR}{False Discovery Rate}
+\newacronym{ln}{LN}{lymph node}
+\newacronym{pr}{PR}{progesterone receptor}
+\newacronym{er}{ER}{estrogen receptor}
+\newacronym{her2}{HER2}{human epidermal growth factor receptor 2}
+\newacronym{svd}{SVD}{singular value decomposition}
+\newacronym{rma}{RMA}{robust multi array}
+\newacronym{mas}{MAS5}{microarray analysis suite 5.0}
+\newacronym{fdr}{FDR}{false discovery rate}
+\newacronym{fwer}{FWER}{family-wise error rate}
 \newacronym{go}{GO}{Gene Ontology}
@@
 \newacronym{mmr}{MMR}{mismatch repair}
+\newacronym{mtor}{mTOR}{mammalian target of rapamycin}
 \newacronym{ner}{NER}{nucleotide-excision repair}
@@
 \newacronym{uv}{UV}{ultraviolet}
+\newacronym{deg}{DEG}{differentially expressed gene}
 \newacronym{pah}{PAH}{polycyclic aromatic hydrocarbon}
@@
 \newacronym{gpcr}{GPCR}{G-protein coupled receptor}
+\newacronym{pm}{PM}{perfect match}
+\newacronym{mm}{MM}{mismatch}
 \newacronym{tlr}{TLR}{Toll-like receptor}
+\newacronym{cpm}{cpm}{counts per million}
+\newacronym{m}{$\mu$}{mean}
+\newacronym{sd}{$\sigma$}{standard deviation}
 \newacronym{prr}{PRR}{pattern recognition receptor}
@@
 \newacronym{erk}{ERK}{extracellular signal regulated kinase}
+\newacronym{cdna}{cDNA}{complimentary DNA}
+\newacronym{mrna}{mRNA}{messenger RNA}
+\newacronym{ngs}{NGS}{next generation sequencing}
+\newacronym{fc}{FC}{fold change}
 \newacronym{ifny}{IFN-$\gamma$}{interferon-$\gamma$}

### 2017-01-16 thesis/methods/methods.tex
Finished expression analysis section

@@
 In contrast to the conservative \gls{fwer} control methods, the \gls{fdr} method developed by \citet{Benjamini1995a} control the \glspl{type1} while maintaining statistical \gls{power}.
-The \gls{fdr} method controls the ``expected proportion of errors among the rejected hypotheses'' by adjusting the $\alpha$ to an alternative value, $q^*$, which is calculated from the p-values of the hypothesis tests \citep{Benjamini1995a}.
-With \gls{fdr}, the p-values are ranked and ordered from the lowest to the highest p-value,
+The \gls{fdr} method controls the ``expected proportion of errors among the rejected hypotheses'' by adjusting the $\alpha$-level (denoted as $q^*$ in \gls{fdr}) depending on the rank of the p-value \citep{Benjamini1995a}.
+With \gls{fdr}, the p-values are ordered and ranked from the lowest to the highest p-value, and for each hypothesis the p-value is compared with the adjusted threshold value:
+\begin{equation}
+	\label{eq:fdr}
+	P_{(i)} \leq \frac{i}{m}q^*
+\end{equation}
+where $P_{(i)}$ is the p-value of the ordered and ranked $i$th hypothesis, $m$ is the total number of hypotheses, and $q^*$ is the threshold value \citep{Benjamini1995a}.
+From \cref{eq:fdr}, the adjusted p-value from \gls{fdr} control is the product of the p-value with the number of hypotheses, divided by its rank.
+
+Since \gls{fdr} method controls \gls{type1} and have greater statistical \gls{power} than the \gls{fwer} methods, \gls{fdr} method was used to identify the \glspl{deg} for gene expression analysis.
 

### 2017-01-18 thesis/methods/methods.tex
tmp commit

@@
 \textit{limma} is a package that contains variety of tools to analyse microarray data using linear models-based methods, developed by \citet{Ritchie2015}.
-For gene expression analysis, \textit{limma} package fits a linear model in a gene-wise manner and produce test statistics that allow the assessment of whether the gene is differentially expressed or not \citep{Ritchie2015}.
+For gene expression analysis, \textit{limma} package fits a linear model in a genewise manner and produce test statistics that allow the assessment of whether the gene is differentially expressed or not \citep{Ritchie2015}.
 
@@
 \label{sec:pathway_enrichment_analysis}
+
+Identification of \glspl{deg} provides great deal of information about the effect of treatment and provides a list of genes that may have a role in that particular experimental setting.
+With that said, it is often difficult for the investigators to provide a plausible biological explanation with just a long list of \glspl{deg}, as the list lacks the link between the genes with the biological cause \citep{Khatri2012}.
+Therefore, given a list of \glspl{deg}, it is crucial for any researcher to undertake pathway enrichment analysis in order to provide useful insights into the underlying biological mechanisms the \glspl{deg} may be involved in.
+
+\subsection{Competitive and self-contained tests}
+\label{sub:competitive_and_self_contained_tests}
+
+In general, there are two major classes of methods for pathway enrichment analysis: self-contained or competitive tests \citep{Goeman2007}.
+The main difference between the two classes is the definition of the null hypothesis.
+Letting $G$ be the gene set of interest and $G^c$ its compliment, then in self-contained tests, the null hypothesis is formulated as follows:
+
+\begin{quote}
+	\textit{$H_0^{\textrm{self}}$: No genes in G are differentially expressed. }
+\end{quote}
+
+\noindent
+whereas in competitive tests, the null hypothesis is defined as:
+
+\begin{quote}
+	\textit{$H_0^{\textrm{comp}}$: The genes in G are at most as often differentially expressed as the genes in $G^c$.
+	}
+\end{quote}
+
+These hypotheses show that self-contained tests care only about the genes defined in the gene set, but competitive tests examine the genes in the defined set as well as the genes not present in the set \citep{Wu2012}.
+Due to the fact that self-contained tests do not take into account for the genes not defined in the gene set, self-contained tests are more restrictive compared to competitive tests.
+The restrictive property of self-contained tests give it greater power, as they are able to reject the null hypothesis at a higher accuracy for more gene sets than the competitive tests \citep{Goeman2007}.
+As \citet{Goeman2007} stated in their paper, ``the competitive types of test can be said to voluntarily relinquish some power in order to make a stronger statement''.
+
+Taken together, self-contained tests are more appropriate in assessing whether a biological process is significantly involved in an expreiment, and in contrast, competitive tests are better for selecting the most relevant biological processes from those that are not \citep{Wu2012}.
+
+\subsection{\Gls{ora} and \gls{fcs}}
+\label{sub:ora_and_fcs}
+
+Though the approaches differ between the two, both \acrfull{ora} and \acrfull{fcs} are used to identify whether certain pathways are enriched, given a list of \glspl{deg}.
+The common approach taken by \gls{ora} is to count the \glspl{deg} that are part of a biological pathway and perform a statistical test to decide whether the pathway is over- or under-represented in the list of genes \citep{Khatri2012}.
+Statistical test used in \gls{ora} includes $\chi^2$-test, hypergeometric test, and binomial test \citep{Khatri2012}.
+Depending on the methods, \gls{ora} uses either self-contained or competitive tests.
+
+There are several limitations to the \gls{ora} approach.
+Firstly, the statistical tests are independent of the measured changes and therefore ignores the values associated with the genes, such as intensities and significance of the change \citep{Khatri2012}.
+Secondly, only the most significant genes are selected for the input list of genes, which means that the almost significant genes are discarded from the analysis and results in a loss of information \citep{Khatri2012}.
+Lastly, by treating each gene individually, the analysis ignores the biological interaction and complexity of the genes with other genes, as well as between pathways \citep{Khatri2012}.
+
+In general, \gls{fcs} methods measure gene-level statistics from the given data, then aggregate these statistics into pathway-level statistics, which is then used to calculate the statistical significance of the pathway \citep{Khatri2012}.
+Either self-contained or competitive tests can be used to calculate the statistical significance of the pathways.
+
+\Gls{fcs} approach addresses some of the limitations presented by the \gls{ora} methods.
+By calculating the statistical significance per pathway, \gls{fcs} approach takes into account of all of the genes involved in the pathways, not just the genes that are differentially expressed.
+Furthermore, \gls{fcs} approach also detect small but consistant and coordinated changes in the expression of the genes, unlike in \gls{ora} where molecular measurements are completely ignored \citep{Khatri2012}.
+
+Though \gls{fcs} approach improves on \gls{ora} approach, there are still some limitations.
+Since the analyses that use \gls{fcs} compare the pathways independently from one another, it ignores the interactive nature of biological pathways.
+Another limitation is that \gls{fcs} methods use the molecular measurements to rank the genes, but they do not consider these changes into further analysis \citep{Khatri2012}.
+For example, if one gene was expressed with a 2-fold change, whereas another gene was expressed with a 20-fold change, the \gls{fcs} method will rank these genes accordingly and disregard the fact that the second gene should be weighted more than the first gene \citep{Khatri2012}.
+
+Although both approaches have their own limitations, it is clear that \gls{fcs} approach have more advantages over \gls{ora} methods.
+
+\subsection{\Gls{camera}}
+\label{sub:camera}
+
+\Acrfull{camera} is a competitive \gls{fcs}-based method developed by \citet{Wu2012}, implemented as the \texttt{camera} function in the \textit{limma} package \citep{Ritchie2015}.
+Briefly, \gls{camera} fits a linear model in a genewise manner and calculates the genewise test statistics using the log\gls{fc} between the conditions.
+These genewise test statistics are used in the \gls{wmw} rank sum test to question whether a pathway is significantly enriched in the data \citep{Wu2012}.
+
+One problem with the other \gls{fcs}-based methods is that these methods do not consider the inter-gene correlation of the gene set being tested \citep{Wu2012}.
+Since other methods assume that all the genes are equivalent under the null hypothesis, inter-gene correlation of the genes in a gene set will violate this assumption \citep{Wu2012}.
+As a consequence, the \gls{type1} rate is increased in these methods.
+\Gls{camera} accounts for this correlation by estimating the \gls{vif} from the genewise correlation and the number of genes in the gene set \citep{Wu2012}.
+
+\gls{camera} was used in this project to identify the pathways that are enriched in samples from obese patients compared with the samples from non-obese patients, as it provides a competitive \gls{fcs}-based method.
+
+\subsection{\gls{go} database}
+\label{sub:go_database}
 

### 2017-01-18 thesis/methods/methods.tex
Finished pathway enrichment analysis section

@@
 
-For a given list of differentially expressed genes identified in a data set, the \texttt{camera} function from the \textit{limma} package was used to identify the pathways that were enriched in those genes.
-The \gls{go} database was used to identify the enriched pathways.
+\Gls{ontology} is defined as a set of concepts and categories in a subject area or domain that shows their properties and the relations between them.
+\Acrfull{go}, as the name suggests, is an \gls{ontology} of the genes and biological pathways, curated and maintained by the \gls{go} Consortium \citep{GO2000,GO2004}.
+The goal of the \gls{go} Consortium is to ``produce a structured, precisely defined, common, controlled vocabulary for describing the roles of genes and gene products in any organism'' \citep{GO2000}.
+The \gls{go} database includes not only the human genetic information, but collates variety of information from eukaryotic cells, including \textit{Arabidopsis thaliana} and \textit{Drosophila melanogaster} \citep{GO2000,GO2004}.
+
+\gls{go} provides three distinct ontologies to describe the role of a gene or its gene product in the cell: biological process, molecular function, and cellular component \citep{GO2000}.
+Biological process categorises the genes or its gene products into their overarching biological purpose or goal in the cell; for example, ``cell death'' and ``cell growth and maintenance''.
+Molecular function describes the biochemical activity a gene or protein has, ignoring when or where the activity takes place.
+The definition of ``biochemical activity'' is very broad as it includes specific binding to ligands and structure.
+Examples of terms used in molecular function are ``kinase activity'', ``transporter'' and ``ligands''.
+Lastly, cellular components refers to the location in which the gene product is active, such as ``Golgi apparatus'' and ``nuclear membrane''.
+
+Due to the sheer amount of terms and definitions available from the data, \gls{go} was chosen as the database for pathway reference.
+The \textit{GO.db} package was used to load the \gls{go} database into \gls{r} \citep{Carlson2016}.
+Only the human related \gls{go} terms were used in pathway enrichment analysis, and these terms were manipulated in a form which the \texttt{camera} function was able to recognise.
 

### 2017-01-18 thesis/misc/glossary.tex
tmp commit
Finished pathway enrichment analysis section

@@
 
+\newglossaryentry{ontology}{
+	name=Ontology,
+	text=ontology,
+	description={A set of concepts and categories in a subject area or domain that shows their properties and the relations between them}
+}
+
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 \newacronym{go}{GO}{Gene Ontology}
+\newacronym{kegg}{KEGG}{Kyoto Encyclopedia of Genes and Genomes}
 \newacronym{usa}{USA}{United States of America}
@@
 \newacronym{ssdna}{ssDNA}{single-stranded DNA}
+\newacronym{camera}{CAMERA}{correlation adjusted mean rank gene set test}
+\newacronym{ora}{ORA}{over representation analysis}
+\newacronym{fcs}{FCS}{functional class scoring}
+\newacronym{wmw}{WMW}{Wilcoxon-Mann-Whitney}
+\newacronym{vif}{VIF}{variance inflation factor}
 \newacronym{erk}{ERK}{extracellular signal regulated kinase}

### 2017-01-19 thesis/methods/methods.tex
Finished metagene analysis section

@@
 Letting $G$ be the gene set of interest and $G^c$ its compliment, then in self-contained tests, the null hypothesis is formulated as follows:
-
 \begin{quote}
@@
 \end{quote}
-
-\noindent
 whereas in competitive tests, the null hypothesis is defined as:
-
 \begin{quote}
@@
 \end{quote}
-
 These hypotheses show that self-contained tests care only about the genes defined in the gene set, but competitive tests examine the genes in the defined set as well as the genes not present in the set \citep{Wu2012}.
@@
 
-\gls{svd} was applied to the training data set to create the metagene and the transformation matrix in the data set.
-The transformation matrix from the training data was used to transform the other data sets and to create the metagene of the genetic signature in those data sets.
-
-Each of these metagenes were plotted in a heatmap to see the association between the metagene and the overall gene expression of the genes that were used to create the metagene.
-The metagene was also checked for the association between the sample \gls{bmi} status and \gls{bmi} value, which was plotted in a box plot and a scatter plot, respectively.
-
-% \section{\Gls{svd}}
-% \label{sec:singular_value_decomposition}
-
-\gls{svd} was used on a given data set to create the summary metagene scores of a given genetic signature.
-% TODO: add more details on the components (s, v, d) created when svd was applied to the data.
-Any matrix $X$ can be represented in the form:
+In order to assess whether a certain set of genes are up or down regulated by a sample, some sort of score based on these genes must be calculated.
+\Gls{svd} is a mathematical method that splits a matrix into several matrices that represent the original matrix \citep{Golub1970}.
+With \gls{svd}, a matrix $X$ of size $n$ genes by $m$ arrays (or samples) can be represented as:
 \begin{equation}
 	\label{eq:svd}
-	X = UDV'
+	X = UDV^T
 \end{equation}
+where the columns of $U$ and $V$ contain the left and right singular vectors of $X$ respectively, and $D$ contains the singular values of $X$.
 
-\noindent
-where $U$ is the *****, $D$ is the ********, and $V'$ is the **** (eigenvectors and eigenvalues??).
+The term that is of the greatest importance is $V$, as this matrix contains the principal components of the original data matrix $X$.
+Previous studies have used the principal components to summarise the expression of the set of genes for each sample in the experiment \citep{Alter2000,West2001}.
+This allows direct comparison of the expressions of multiple different genes in different arrays or experiments \citep{Alter2000}.
+Furthermore, this summary gene, or metagene, can be used to sort the samples and provide a meaningful grouping of the data which may help understand the underlying biology \citep{Alter2000}.
 
-% \section{Transformation matrix}
-% \label{sec:transformation_matrix}
+With respect to this project, \gls{svd} was used to assess whether the metagene created from various obesity or pathway associated genes were significantly associated with certain clinical variables, such as \gls{bmi}.
+To do this, the normalised original expression data was reduced in size so that it only included the genes of ineterest (either obesity or pathway associated) and the samples, then \texttt{svd} function from the \textit{base} package in \gls{r} was used to apply \gls{svd} to the matrix.
+The first principal component was taken from the $V$ matrix and was used as the metagene scores for the samples in the data.
 
-Transformation matrix of a genetic signature was created from the \gls{svd} components.
-Rearranging \cref{eq:svd}, you get:
-
+Another important property of \gls{svd} is that the signal used to create the metagene in the first data set can be directly applied to other data set to obtain the metagene scores in that data set.
+Rearrangement of \cref{eq:svd} for $V^T$ gives the following equation:
 \begin{equation}
 	\label{eq:transmat}
-	V' = U'D^{-1}X
+	V^T = U^{T}D^{-1}X
 \end{equation}
-
-\noindent
-where $U'D^{-1}$ represents the transformation matrix.
-The transformation matrix $U'D^{-1}$ was created from the \gls{svd} components, and was used to transform other data set and obtain the metagene in the data set.
+where $U^{T}D^{-1}$ will be referred to as the ``transformation matrix'' hereafter.
+Clearly, by substituting the data matrix $X$ with another data matrix, the transformation matrix allows for the generation of the metagene for that data set.
+The reason why the transformation matrix is used instead of independently applying \gls{svd} to the other data set is because the signature, and therefore the metagene, from a single data set is dependent on the signal within its data set.
+Therefore, the metagenes created independently from different data sets using \gls{svd} does not have the same weighting as the metagene from the original data set, and consequently does not provide a fair comparison of the metagenes across different data sets.
+For this reason, the transformation matrix for a given set of genes was created in the data set in which the genes were first identified in, and then transformed into other data sets to obatain the metagenes.
 

### 2017-01-20 thesis/methods/methods.tex
Finished obesity metagene direction subsection

@@
 
-In order to assess whether a certain set of genes are up or down regulated by a sample, some sort of score based on these genes must be calculated.
+\subsection{\Gls{svd}}
+\label{sub:svd}
+
+In order to ertain set of genes are up or down regulated by a sample, some sort of score based on these genes must be calculated.
 \Gls{svd} is a mathematical method that splits a matrix into several matrices that represent the original matrix \citep{Golub1970}.
@@
 
-\section{Gatza pathway metagene direction}
-\label{sec:pathway_metagene_direction}
+\subsection{Metagene direction}
+\label{sub:metagene_direction}
 
-All 18 pathway signatures from \citet{Gatza2010a} study were used to create both the metagenes and transformation matrices in the Gatza data set.
-Each of the 18 pathway metagenes were compared with the gene expression of the gene that represents the pathway to check whether the metagenes were in the correct direction.
-% TODO: check if it was the AKT1 or AKT2 gene for AKT pathway
-For example, the AKT pathway metagene was compared with the AKT1 gene expression.
+One thing to consider when analysing data with metagenes is the direction of the metagene.
+When \gls{svd} creates the metagene from the given data, it does not consider the direction of the signal, but only the magnitude of the signal in the data.
+Therefore, the metagene created by \gls{svd} can be either positively or negatively correlated to the phenotype that the set of genes reflect.
+For example, a metagene created from a set of genes that are associated with \gls{er} status may have higher metagene scores in the \gls{er} negative samples and low metagene scores in the \gls{er} positive samples.
+In this case, the generated metagene must be flipped in order to reflect the phenotype with the metagene scores (low metagene score with low phenotype and high metagene score with high phenotype).
 
-The pathway metagene value was first compared visually with its pathway gene in a heatmap.
-If the direction of the pathway metagene was correct (for example, high pathway metagene value corresponded with high expression of the gene representing that pathway), the metagene values were not changed, but if it was incorrect, the metagene values were flipped for that pathway.
+In this project, two types of metagenes were created: obesity associated and pathway associated metagenes.
+For obesity associated metagenes, the direction of the metagenes were checked with the obesity status and \gls{bmi} value of the samples from the data in which the metagene was derived from.
+Checking the direction of the pathway metagenes from \citet{Gatza2010a} study is not as simple as the obesity associated metagenes, as the experimental design that identify which samples were treated to generate the pathway signature was not available.
+Due to this, it was not possible to validate whether the generated pathway metagene was in the right direction.
 
-Spearman correlation of the pathway metagene and the expression of the pathway gene was also used to distinguish whether the pathway metagene was in the right direction.
-The direction of the metagene was considered correct when the correlation was positive and incorrect when the correlation was negative.
-
-The clustering of the pathway metagenes were taken into account to decide on the final direction of the metagene.
-The direction of the pathway metagenes were modified until the resulting pathway metagenes were similarly clustered together as in the results from \citet{Gatza2010a} paper.
+As an alternative, the pathway metagenes were visually compared with the gene that represent the pathway using a heatmat, as the expression of this gene would most likely be affected the most by the pathway signature.
+For example, the Akt pathway metagene was compared with the AKT1 gene expression, and if the metagene was not in the same direction as the AKT1 gene expression, the metagene was flipped.
+\cref{tab:metagene_direction} summarises the genes used to check the direction of the pathway metagenes.
+\begin{table}[htb]
+	\caption{18 pathways from \citet{Gatza2010a} and its respective genes used to check the direction of the pathway metagene}
+	\label{tab:metagene_direction}
+	\begin{center}
+		\begin{tabular}{lc}
+			Pathway & Representative gene\\
+			\hline
+			\rule{0pt}{2.25ex}Akt & AKT1\\
+			\gls{bcat}            & CTNNB1\\
+			E2F1                  & E2F1\\
+			\gls{egfr}            & EGFR\\
+			\gls{er}              & ESR1\\
+			\gls{her2}            & ERBB2\\
+			\gls{ifna}            & IFNA1\\
+			\gls{ifny}            & IFNG\\
+			Myc                   & MYC\\
+			p53                   & TP53\\
+			p63                   & TP63\\
+			\gls{pi3k}            & PIK3CA\\
+			\gls{pr}              & PGR\\
+			Ras                   & HRAS\\
+			\gls{stat3}           & STAT3\\
+			Src                   & SRC\\
+			\gls{tgfb}            & TGFB1\\
+			\gls{tnfa}            & TNF\\
+			\hline
+			\hline
+		\end{tabular}
+	\end{center}
+\end{table}
+In addition to this, the correlation of the pathway metagene with its repective gene was checked to further aid the decision on whether the metagene is going in the correct direction.
+This was due to the fact that some of the metagenes were visually difficult to determine whether it was in the right direction.
+Therefore, the direction as well as the magnitude of the correlation value was also taken into account in making the decision whether to flip the metagene.
+Lastly, further on top of this, the clustering of the pathway metagenes with the other pathway metagenes were also considered.
+It was known from \citet{Gatza2010a} study that some pathways clustered and grouped together with specific pathways, so the metagenes were flipped so that the grouping of the pathways matched the results from their study.
 

### 2017-01-21 thesis/methods/methods.tex
Finished obesity metagene prediction section

@@
 
-\section{Obesity metagene prediction}
-\label{sec:obesity_metagene_prediction}
+\section{Obesity metagene prediction with pathway metagenes}
+\label{sec:obesity_metagene_prediction_with_pathway_metagenes}
+
+To determine whether the obesity metagene was associated with any of the pathway metagenes, a linear model was constructed with some clinical variables and pathway metagenes, and then this model was used to predict the obesity metagene.
+This was based on the idea that, if a variable or a pathway metagene was a significant part of the model for the obesity metagene, then the variable or a pathway metagene may be associated with the obesity metagene.
+Furthermore, the model that was created based on the significant variables can be used to predict the obesity metagene, which can be used to compare with the original obesity metagene; if the predicted metagene was similar to the original metagene, then it provides an evidence that the pathway used to construct the model was associated with the obesity metagene.
+
+Four linear models were created in (Cris' data -- get reference) to predict the obesity associated metagene from the \citet{Creighton2012} study.
+The first model used the sample \gls{bmi} value or \gls{bmi} status by itself, the second model with sample \gls{bmi} value, \gls{bmi} status and selected pathway associated metagenes, the third model with selected pathway associated metagenes, and the last model was constructed with just the \gls{pr} pathway.
+The selected pathways used in the linear model construction were \gls{bcat}, \gls{er}, \gls{ifna}, \gls{ifny}, Myc and \gls{pr}.
+
+In deciding on the pathways to be used for the construction of the linear model, the correlation of different versions of pathway metagenes, \gls{svd}-derived and transformation matrix-derived, in different data sets were used.
+When a transformation matrix is applied to the data set in which it was derived from, the resulting metagene is identical to the one created from the direct application of \gls{svd}.
+Therefore, theoretically, if the metagene derived from a transformation matrix is similar to the metagene created from \gls{svd}, then it shows that the genetic signature is consistent and reliable across different data sets.
+The above mentioned pathways had high correlation value between the \gls{svd}-derived and transformation matrix-derived pathway metagenes, and so it was used in the construction of the linear model.
 
 % TODO: possibly extra analyses here (gradual pathway inclusion in the model, using all the pathway...?)
-Sample BMI, BMI status and/or pathway metagenes were used to construct a linear model in Cris Print's data.
-The constructed linear model was then used to predict the obesity metagene values in \citet{Creighton2012} data set.
-The predicted obesity metagene values were compared with the true obesity metagene values, and both Spearman and Pearson correlation were calculated.
+
+The predicted obesity metagenes that were created in (Cris' data -- get reference) was compared with the original obesity metagene from the \citet{Creighton2012} data set.
+Both Pearson and Spearman correlation was calculated to assess the relationship between the predicted and original metagenes.
+Scatter plots were also created to visually compare the two metagene scores.
 
@@
 All of the heatmaps that required no or one column bar used the \texttt{heatmap.2} function, and all of the heatmaps that required more than one column bars used the \texttt{heatmap.2x} function.
-
-% TODO: re-write everything so that it's in past tense

### 2017-01-22 thesis/intro/intro.tex
Finished obesity metagene direction subsection
Started editing introduction -- tmp commit

@@
 Obesity is defined as an abnormal or excessive fat accumulation that may impair health of that individual \citep{Garrow1988}.
-
 One common and widely used approach to categorise individuals is by measuring the \gls{bmi} of the individual.
@@
 Furthermore, the prevalence of overweight in children under 5 years was estimated to reach from 6.3\% in 2013 to 11\% by 2025 if the current trend continued \citep{WHO2014}.
-The current obesity status of the adult population is bad already, but the fact that more of the younger population who are overweight and obese poses a much greater risk for the health of the population (discussed in \cref{sub:impact_of_obesity}).
+The current obesity status of the adult population is bad already, but the fact that more of the younger population who are overweight and obese poses a much greater risk for the health of the population.
 
@@
 
-Until these terrible trends stop, and even after the trends stop, the world will need a plan to live along with this ongoing global epidemic of obesity and its associated diseases.
+% Until these terrible trends stop, and even after the trends stop, the world will need a plan to live along with this ongoing global epidemic of obesity and its associated diseases.
 
@@
 
-Obesity is caused by continuous intake of food, whether they are controlled or not, without expending all of the energy gained.
+Obesity is caused by continuous intake of food without expending all of the energy gained.
 The key to the problem is the control of food intake and energy expenditure, and how they are regulated in your body.
-
-\subsubsection{Role of hypothalamus and arcuate nucleus in food intake control}
-\label{ssub:role_of_hypothalamus_and_arcuate_nucleus_in_food_intake_control}
-
-The central role of hypothalamus in the context of obesity is the control of satiety, and therefore food intake.
+The role of hypothalamus is central to the control of satiety and many mutations that cause obesity affects the satiety signals to and/or from hypothalamus.
 The hypothalamus receives both nueronal and hormonal inputs from the body and propagates the signal to the downstream neurons that ultimately affect the food intake of the individual \citep{Bell2005, Spiegelman2001}.
 
-Within the hypothalamus, there is a group of neurons located within the arcuate nucleus that is essential for satiety signal transmission.
-There are two types of neurons in the arcuate nucleus: orexogenic and anorexogenic neurons.
+There are two types of neurons in the arcuate nucleus (located in the hypothalamus): orexogenic and anorexogenic neurons.
 Orexogenic neurons are responsible for the promotion of food intake and reducing energy expenditure, whereas the anorexogenic neurons have the opposite effect \citep{Barsh2002}.
-
-The orexogenic neurons produce \gls{npy} and \gls{agrp} which induces orexogenic signal to the downstream neurons; anorexogenic neurons produce \gls{pomc} and \gls{cart} which induces anorexogenic signal to the downstream neurons \citep{Barsh2002, Spiegelman2001}.
 When stimulated by, for example, endocrine signals, these two different types of neurons act on the downstream neurons to promote or reduce food intake and energy expenditure.
 
-\subsubsection{Leptin-melanocortin pathway}
-\label{ssub:leptin_melanocortin_pathway}
+Leptin-melanocortin pathway is the main pathway in which the level of satiety is controlled and regulated \citep{Spiegelman2001}.
+In this pathway, both the orexogenic and anorexogenic neurons (located in the arcuate nucleus of the hypothalamus) have crucial roles in controlling the satiety of the individual, and therefore their eating behaviour \citep{Barsh2002,Bell2005}.
+When a satiety-controlling hormone such as leptin reaches the arcuate nucleus region of the hypothalamus, it can stimulate or inhibit orexogenic, anorexogenic, or both orexogenic and anorexogenic neurons.
+Depending on which neurons are stimulated or inhibited, the hormone will ultimately alter the satiety level of the individual.
 
-Leptin-melanocortin pathway is the pathway in which the level of satiety is controlled and regulated.
-In this pathway, both the orexogenic and anorexogenic neurons have crucial roles in controlling the satiety of the individual, and therefore their eating behaviour.
+Though there are many hormones that regulate satiety, only leptin and insulin will be covered here.
+Leptin gene was first discovered by \citet{Zhang1994} and the discovery of the leptin receptor gene was followed shortly after by \citet{Tartaglia1995}.
+Together, the identification of the leptin and leptin receptor genes was a major break through in the field of obesity research.
 
-When a satiety-controlling hormone such as leptin reaches the arcuate nucleus region of the hypothalamus, it can stimulate or inhibit orexogenic, anorexogenic, or both orexogenic and anorexogenic neurons.
-Depending on which neurons are stimulated or inhibited, the hormone will ultimately alter the satiety level of the individual.\\
+Leptin is a hormone that signals satiety and reduces food intake, and is secreted mainly by the white adipose tissue \citep{Moustafa2013,Zhang1994}.
+The circulating concentration of leptin in the body is proportional to the body fat stores \citep{Barsh2002, Moustafa2013}.
+Leptin signals through the \gls{jak}/\gls{stat} pathway via the long form of the leptin receptor that is present in high expression in the hypothalamus \citep{Ghilardi1996,Lee1996}.
+Depending on the type of neurons the leptin receptor is expressed, leptin stimulate or inhibit these neurons and ultimately produces an anorexogenic signal to the downstream effector neurons \citep{Bell2005}.
 
-\noindent
-Both \gls{pomc} and \gls{cart} are strong anorexogenic neuropeptides that stimulate the downstream effector neurons via the \gls{mc4r} pathway \citep{Barsh2002}.
-
-\gls{pomc} is a precursor protein that is cleaved to produce \gls{acth}, $\alpha$-, $\beta$- and $\gamma$-\gls{msh}, and $\beta$-endorphin \citep{Smith1988}.
-Perhaps the most important, or most relevant, in the context of obesity and satiety control is the $\alpha$-MSH.
-$\alpha$-MSH is the main ligand that binds to the \gls{mc4r} expressed in the brain, and sends anorexogenic signals to the downstream neurons through this pathway \citep{Krude1998}.
-
-Similarly, \gls{cart} is also a neuropeptide that has a strong anorexogenic effect discovered by \citet{Kristensen1998}.
-Unlike $\alpha$-MSH, the receptor for \gls{cart} has not been identified yet, and the exact mechanism in which \gls{cart} acts is unclear \citep{Kristensen1998, Rogge2008}.
-
-Since the expression of both neuropeptides are under the control of leptin, and that $\alpha$-MSH is known to stimulate the melanocortin pathway to produce an anorexogenic signal to the downstream neurons, this satiety-controlling signalling pathway is known as the leptin-melanocortin pathway.
-\\
-
-\noindent
-In contrast to the \gls{pomc}/\gls{cart} neurons, orexogenic neuropeptides \gls{npy} and \gls{agrp} are produced in the \gls{npy}/\gls{agrp} neurons when stimulated by hormones that control satiety.
-
-\gls{npy} was first discovered by \citet{Tatemoto1982}, and later associated with altered eating behaviour in rats by \citet{Clark1984}.
-\gls{npy} is recognised by the \gls{nyr1}, a G protein-coupled receptor expressed on the downstream effector neurons, which relays the orexogenic signal from the \gls{npy}/\gls{agrp} neurons \citep{Barsh2002, Bell2005, Herzog1992}
-
-Agouti protein is a protein normally expressed in the skin where it affects the pigmentation, and was first identified over 100 years ago by \citet{Castle1910} in mice that had yellow, or agouti-coloured coat.
-Although the fact that ubiquitous expression of agouti protein caused obesity in mice was known for a long time, the biochemical mechanism that caused such phenotype was not elucidated until 1997 by \citet{Ollmann1997}.
-In brief, \gls{agrp} was shown to be a strong antagonist of the \gls{mc4r}, expressed in the downstream neurons to the \gls{npy}/\gls{agrp} neurons, and thus inhibited the anorexogenic signalling through the melanocortin pathway \citep{Barsh2002,Ollmann1997}.\\
-
-\noindent
-Together, \gls{npy}/\gls{agrp} and \gls{pomc}/\gls{cart} neurons produce contrasting signals that affect the overall satiety of the individual.
-Depending on how strong or weak these signals are, the individual may be more or less susceptible to obesity.
-
-\subsubsection{Role of hormones in leptin-melanocortin pathway}
-\label{ssub:role_of_hormones_in_leptin_melanocortin_pathway}
-
-There are many hormones in your body that regulates various physiological aspects of your body.
-Out of the many, four hormones in particular have an important role in regulating the satiety of an individual: ghrelin, leptin, insulin, and \gls{pyy}.\\
-
-\noindent
-Ghrelin, originally identified as a peptide that stimulated the release of growth hormone from the pituitary, is a hormone produced mainly in the stomach that has a significant role in meal initiation and food intake, as well as energy homeostasis \citep{Kohno2003, Kojima1999, Nakazato2001}.
-Ghrelin is known to be a ligand for the \gls{ghsr} which is expressed abundantly on the \gls{npy}/\gls{agrp} neurons, and increases the Ca$^{2+}$ concentration in these neurons \citep{Kohno2003}.
-
-Furthermore, \citet{Seoane2003} showed that ghrelin increased the mRNA levels of both \gls{npy} and \gls{agrp} in these neurons and hence their protein expression, indicating the role of ghrelin as the orexogenic hormone that affects the overall outcome of the satiety signal.
-\\
-
-\noindent
-Leptin, on the other hand, is a hormone that has an opposite effect as ghrelin, where the hormone signals satiety and reduces food intake.
-First identified by \citet{Zhang1994}, the dicovery of the leptin gene, followed by the discovery of leptin receptor gene by \citet{Tartaglia1995}, were probably the major breakthrough in the field of obesity research.
-
-Leptin is produced and secreted mainly by the white adipose tissue, and the circulating concentration of leptin in the body is proportional to body fat stores \citep{Barsh2002, Moustafa2013, Zhang1994}.
-Leptin receptor has many splice variants, but the long variant of the leptin receptor is highly expressed only in the hypothalamus \citep{Ghilardi1996, Lee1996}.
-In addition to this, leptin receptor is expressed on both the \gls{npy}/\gls{agrp} and \gls{pomc}/\gls{cart} neurons, and have different effect on the outcome of the satiety signal depending on which neuron leptin stimulates.
-
-In \gls{npy}/\gls{agrp} neuron, it has been shown to reduce the Ca$^{2+}$ concentration, antagonising the effect of ghrelin on this type of neurons \citep{Kohno2003}.
-Furthermore, binding of leptin to the leptin receptor on the \gls{npy}/\gls{agrp} neuron causes a decreased release of \gls{gaba}, which inhibits the \gls{pomc}/\gls{cart} neuron \citep{Cowley2001}.
-
-In \gls{pomc}/\gls{cart} neuron, binding of leptin to its recpetor causes a depolarisation of the neuron \citep{Cowley2001}.
-Together with the indirect effect from the \gls{npy}/\gls{agrp} neuron, leptin promotes the \gls{pomc}/\gls{cart} neuron to release $\alpha$-\gls{msh}, and therefore sends a anorexogenic signal to its downstream effector neurons.
-\\
-
-\noindent
 Insulin, perhaps the most well-known hormone for its role in glucose homeostasis, also has an important role in satiety and food intake regulation.
-
 Briefly, when insulin binds to the insulin receptor, \gls{irs} protein family is recruited and phosphorylated by the insulin receptor's intrinsic tyrosine kinase activity \citep{Saltiel2002}.
@@
 
-The evidence of insulin's involvement in the satiety control was not particularly clear; even though there were strong evidence of obesity with neuron-specific deletion of insulin receptor, and that the insulin level, like leptin, was proportional to the fat mass also suggested the role of insulin in satiety and body weight control \citep{Barsh2002, Bruning2000, Woods1979}.
-This was perhaps due to the fact that leptin and insulin used two different signalling pathways in the cell: leptin used the \gls{jak}/\gls{stat} signalling pathway, whereas insulin used the \gls{pi3k} pathway for signalling \citep{Ghilardi1996}.
+The evidence of insulin's involvement in the satiety control was not particularly clear even though there were strong evidence of obesity with neuron-specific deletion of insulin receptor, and that the insulin level, like leptin, was proportional to the fat mass also suggested the role of insulin in satiety and body weight control \citep{Barsh2002, Bruning2000, Woods1979}.
+This was perhaps due to the fact that leptin and insulin used two different signalling pathways in the cell; leptin used the \gls{jak}/\gls{stat} signalling pathway, whereas insulin used the \gls{pi3k} pathway for signalling \citep{Ghilardi1996}.
 This was puzzling to the researchers as there seemed to be no converging point of the two hromones in these pathways.
@@
 The activation of K$^+$ channel provided evidence for the two pathways to act similarly on the cell, at least for an acute response, suggesting a convergence in the pathway response by the two hormones, and therefore their role in satiety control.
-\\
 
-\noindent
-The last hormone is the \gls{pyy}, a peptide hormone released by the distal gastrointestinal tract that produces an anorexogenic signal.
-
-\gls{pyy} is a gut hormone that is secreted postprandially, and it is released at levels proportional to the calorie content of the meal \citep{Adrian1985}.
-\gls{y2r} is an inhibitory presynaptic receptor that is highly expressed in the \gls{npy}/\gls{agrp} neuron, and is the target of the \gls{pyy} hormone \citep{Batterham2002}.
-\citet{Batterham2002} showed that \gls{pyy} decreased the expression of \gls{npy} in the \gls{npy}/\gls{agrp} neuron, and activated the \gls{pomc}/\gls{cart} neuron by membrane depolarisation.
-The two mechanisms of \gls{pyy} produces an anorexogenic signal and encourages the individual to stop eating.\\
-
-\noindent
-When any one of these neurons, neuropeptides, hormones or its corresponding receptors become defective, it can lead to some form of eating disorder (briefly covered in \cref{ssub:Individual-level_(micro-level)_factors}: \nameref{ssub:Individual-level_(micro-level)_factors}).
-Therefore the roles of these neurons, neuropeptides, and hormones are crucial for healthy diet and body weight maintenance.
+Satiety control is crucial in order to manage  and maintain the body weight in a healthy state.
+As described above, there are many components that regulate satiety and disruption of any part of this satiety controlling pathway can encourage the development of obesity.
 
@@
 
-There are many environmental factors associated with the global rise in obesity (discussed in \cref{ssub:Population-level_(macro-level)_factors}: \nameref{ssub:Population-level_(macro-level)_factors}).
-Even though these environmental factors are probably the most significant contributor of the current global obesity epidemic, there are some individual-level, genetic factors that assist in the overall increase in the proportion of the population that are obese.
+There are many environmental factors associated with the global rise in obesity.
+Even though these environmental factors are probably the most significant contributor to the current global obesity epidemic, there are some individual-level, genetic factors that contribute to the overall increase in the proportion of the population that are obese.
 
-There are many evidence of obesity being caused by individual-level genetic factors that make some individuals more predisposed to obesity than others.
+There is much evidence of obesity being caused by individual-level genetic factors that make some individuals more predisposed to obesity than others.
 Generally speaking, there are two classes of genetically-driven obesity: monogenic obesity and polygenic obesity \citep{Moustafa2013}. \\
@@
 \noindent
-Monogenic obesity is where a single mutation in one of the genes that is directly involved in the obesity pathway causes severe obesity in the individual.
-Since these mutations are usually directly related to the key neurons/neuropeptides/hor\-mones that were mentioned earlier (\cref{sub:physiological_mechanism_of_obesity}), the resulting phenotype from the mutation causes severe early-onset obesity.
-However, although these mutations cause a devastating effect on the individual, these mutations occur very rarely in the population.
-With that said though, the identification of these mutations have shed light on the mechanisms in which obesity is developed and it is worth mentioning some of these mutations to highlight the importance of the components involved in the obesity pathway.
+Monogenic obesity is where a single mutation in one of the genes that is directly involved in the satiety controlling pathway causes severe obesity in that individual.
+Since these mutations are usually directly related to the key constituents in the pathway mentioned earlier (\cref{sub:physiological_mechanism_of_obesity}), the resulting phenotype from the mutation causes severe early-onset obesity.
+However, although these mutations cause a devastating effect on the individual carrying the mutation, these mutations occur very rarely in the population.
+% With that said though, the identification of these mutations have shed light on the mechanisms in which obesity is developed and it is worth mentioning some of these mutations to highlight the importance of the components involved in the obesity pathway.
 
-The first mutation that should be mentioned is in the leptin gene.
+An example of a mutation that causes severe early-onset obesity is a mutation in the leptin gene.
 \citet{Montague1997} found a single nucleotide deletion mutation at position 133 of the leptin gene in children with early-onset obesity.
 They further showed that the mutation caused a frameshift mutation in the protein, resulting in an inactive form of leptin that was not able to be secreted \citep{Montague1997}.
-
-In close relation to this mutation is the mutation in the leptin receptor gene.
-\citet{Clement1998} have discovered a single nucleotide substitution mutation that led to the leptin receptor transcript that skipped exon 16, leading to a form of the receptor with no transmembrane or intracellular domain.
-Individuals with this rare mutation showed extreme hyperphagia and symptoms of early-onset obesity, even though their circulating leptin levels were present abundantly \citep{Clement1998}.
-
-The next couple of mutations are associated with the synthesis and the processing of \gls{pomc}.
-\citet{Krude1998} found single nucleotide mutations in the exonic region of the \gls{pomc} gene that affected the transcription of the $\alpha$-\gls{msh} region (and some other regions) of \gls{pomc}.
-\citet{Jackson1997} identified a mutation in the \gls{pc1}, an enzyme responsible for the processing of many prohormones and neuropeptides, including \gls{pomc}.
-In either case, the mutation resulted in the lack of $\alpha$-\gls{msh} in the neurons for satiety signalling and caused severe early-onset obesity.
-
-As mentioned earlier, even though these mutations are very rare and affects only those individuals that have one of these mutations, it had a profound impact in the field of obesity research and contributed significantly to the knowledge of the mechanism of how obesity develops.
+As a result, children with this mutation was constantly hungry due to the lack of anorexogenic signals, and became obese at a very early stage of their lives.
 \\
 
-% \citep{Farooqi2003}
-% \citep{Kublaoui2006}
-% \citep{Dubern2001}
-% \citep{Challis2002}
-% \\
+% In close relation to this mutation is the mutation in the leptin receptor gene.
+% \citet{Clement1998} have discovered a single nucleotide substitution mutation that led to the leptin receptor transcript that skipped exon 16, leading to a form of the receptor with no transmembrane or intracellular domain.
+% Individuals with this rare mutation showed extreme hyperphagia and symptoms of early-onset obesity, even though their circulating leptin levels were present abundantly \citep{Clement1998}.
+
+% The next couple of mutations are associated with the synthesis and the processing of \gls{pomc}.
+% \citet{Krude1998} found single nucleotide mutations in the exonic region of the \gls{pomc} gene that affected the transcription of the $\alpha$-\gls{msh} region (and some other regions) of \gls{pomc}.
+% \citet{Jackson1997} identified a mutation in the \gls{pc1}, an enzyme responsible for the processing of many prohormones and neuropeptides, including \gls{pomc}.
+% In either case, the mutation resulted in the lack of $\alpha$-\gls{msh} in the neurons for satiety signalling and caused severe early-onset obesity.
+
+% As mentioned earlier, even though these mutations are very rare and affects only those individuals that have one of these mutations, it had a profound impact in the field of obesity research and contributed significantly to the knowledge of the mechanism of how obesity develops.
 
@@
 Polygenic obesity is where a combination of variants in multiple genes or multiple regions of the genome, together with environmental factors, causes obesity in the individual \citep{Moustafa2013}.
-Unlike in monogenic obesity where a single mutation in a gene causes severe obesity, the variants in some of the genes or genomic regions involved in polygenic obesity have little effect on the individual's obesity status by itself, but in combination with many other variants, together with environmental factors, it has a serious effect on the individual.
+Unlike in monogenic obesity where a single mutation in a gene causes severe obesity, the variants in some of the genes or genomic regions involved in polygenic obesity have little effect on the individual's obesity status by itself, but in combination with many other variants and together with environmental factors, it has a serious obesogenic effect on the individual.
 
-Known as the ``thrifty gene'' hypothesis, \citet{Neel1962} suggested that the genes that were important in obesity have been around in the population for many generations, as it would have had a selective advantage in an environment where the population experienced starvation more frequently than the current generation.
-However, as food became more accessible, a subset of the population who were particularly good at storing fat and survive the starvation period ``overreacted'' to this rapid change in the environment \citep{Bell2005, Spiegelman2001}.
-These observations in the rise in obesity in subsets of population cannot be explained simply by external factors alone, and there is a significant genetic component involved in this phenomenon \citep{Bell2005}.
+% Known as the ``thrifty gene'' hypothesis, \citet{Neel1962} suggested that the genes that were important in obesity have been around in the population for many generations, as it would have had a selective advantage in an environment where the population experienced starvation more frequently than the current generation.
+% However, as food became more accessible, a subset of the population who were particularly good at storing fat and survive the starvation period ``overreacted'' to this rapid change in the environment \citep{Bell2005, Spiegelman2001}.
+% These observations in the rise in obesity in subsets of population cannot be explained simply by external factors alone, and there is a significant genetic component involved in this phenomenon \citep{Bell2005}.
 
@@
 
-Discovery of the \gls{fto} gene has lead to extensive research on the gene, and the mechanism of action by the \gls{fto} variant has been recently uncovered  from these efforts \citep{Claussnitzer2015a,Smemo2014a}.
-The association between obesity and genetic variations, and their mechanisms that lead to obesity, has not been fully elucidated due to its complexity, but it may not be long before they are fully uncovered.
+% Discovery of the \gls{fto} gene has lead to extensive research on the gene, and the mechanism of action by the \gls{fto} variant has been recently uncovered  from these efforts \citep{Claussnitzer2015a,Smemo2014a}.
+% The association between obesity and genetic variations, and their mechanisms that lead to obesity, has not been fully elucidated due to its complexity, but it may not be long before they are fully uncovered.
 
@@
 Even though some individual-level factors are involved with obesity, it is clear that these individual-level factors cannot be accounted for the global epidemic of obesity that the world currently observes.
-For this reason, obesity is thought to be caused mainly by the environmental factors that are present in the population.\\
+For this reason, obesity is thought to be caused mainly by the environmental factors that act at a population-level.\\
 
@@
 This led to greater abundance of food in general, greater number of choices of food and shifted the overall nutritional status of the country \citep{Malik2013}.
-This shift in the food choice and availability (and the economic benefit associated with the trades) had a positive impact on the country, but the growth in the food market also resulted in a different, much larger problem -- obesity.
+This shift in the food choice and availability (and the economic benefit associated with the trades) had a positive impact on the country, but the growth in the food market also resulted in population-wide rise in obesity.
+% This shift in the food choice and availability (and the economic benefit associated with the trades) had a positive impact on the country, but the growth in the food market also resulted in a different, much larger problem -- obesity.
 
@@
 Moreover, corn and soy are also the main food for the livestock, which results in lower prices for meat \citep{Malik2013}.
-On the othere hand, the prices of fruits and vegetables remain expensive due to the lack of support by the government to lower the cost associated with the production \citep{Malik2013}.
+On the other hand, the prices of fruits and vegetables remain expensive due to the lack of support by the government to lower the cost associated with the production \citep{Malik2013}.
 
 As a result, more people are inclined to consume food that are cheap, have little nutritional value and high in energy, than the food that are expensive, nutritional and healthy alternatives.
-Thus, the population is more likely to become obese by making worst food choice.\\
+Thus, the population is more likely to become obese by making poor food choice.\\
 
@@
 The second factor to consider is the behavioural changes led by the urbanisation of the population.
-Urbanisation is defined as the gradual increase in the proportion of people living in urban areas.
-As more people take up urban lifestyle, they are exposed to the new environment where there is a better range of food selection, as well as technological and mechanical advancement in transport and other daily chores that may not be available in rural areas where these people come from.
+% Urbanisation is defined as the gradual increase in the proportion of people living in urban areas.
+As more people take up urban lifestyle, they are exposed to the new environment where there is a better range of food selection, as well as technological and mechanical advancement in transport and other daily chores that may not be available in rural areas.
 
 Although there are many advantages involved with urbanisation, such as access to developed health care systems, education and advanced technologies, the lifestyle change also impose negative health behaviour that ultimately lead to positive energy balance and therefore obesity \citep{Malik2013}.
-
-Lack of physical activity is one of the obvious components that contribute to obesity, which is also linked with the urbanisation of the population.
-Currently, the recommended amount of physical activity to keep an individual fit and healthy is $\geq$ 150 min of moderate-intensity aerobic physical activity per week, or $\geq$ 30 min of physical activity on most days of the week \citep{Pate1995, WHO2010}.
-However, in many urban cities, this recommended level of physical exercise is difficult to achieve and maintain, due to the shift towards the sedentary behaviours that are encouraged by the developed transport system, automated household chores, and indoor entertainment \citep{Malik2013}.
-
-Urbanisation by itself cannot be accounted for the lack of exercise in the population, but it is true that urbanisation makes the bar higher for many people to take up the recommended level of fitness and maintain it.
+Lack of physical activity is one of the obvious components that is linked with the urbanisation of the population.
+% Currently, the recommended amount of physical activity to keep an individual fit and healthy is $\geq$ 150 min of moderate-intensity aerobic physical activity per week, or $\geq$ 30 min of physical activity on most days of the week \citep{Pate1995, WHO2010}.
+In many urban cities, the recommended level of physical exercise is difficult to achieve and maintain due to the shift towards the sedentary behaviours that are encouraged by the developed transport system, automated household chores, and indoor entertainment \citep{Malik2013}.
+\\
+% Urbanisation by itself cannot be accounted for the lack of exercise in the population, but it is true that urbanisation makes the bar higher for many people to take up the recommended level of fitness and maintain it.
 \\
@@
 \noindent
-The third factor is the income and socioeconomic status of the population.
+The last factor to be mentioned is the income and socioeconomic status of the population.
 As the average income increases, more people are able to buy food and are likely to take up the sedentary lifestyle.
-Clearly, these behavioural changes will increase the likelihood of becoming obese.
 
 From this, one would expect the high-income group to have the largest impact from obesity, but on the contrary to this expectation, the biggest impact from obesity is seen in the low- and middle-income groups of the population \citep{Malik2013}.
-As mentioned briefly earlier, access to the developed health care system will allow weight maintenance of the population, but this is likely to be of most benefit for the high-income group in the population.
-The reason for this is, quite simply, the cost associated with the visit to the health care system, and therefore high-income group is more likely to make a consistent and periodic visit than the low or middle-income group of the population \citep{Malik2013}.
+Access to the developed health care system will allow weight maintenance of the population, but this is likely to be of most benefit for the high-income group in the population.
+The reason for this, quite simply, is the cost associated with the visit to the health care system, as high-income group is more likely to make a consistent and periodic visit than the low or middle-income group of the population \citep{Malik2013}.
 
-So, for the low- and middle-income groups, the rise in average income in these groups allow the people to have access to food and activities that promotes sedentary bahaviours, but the limited access to the health care system, education and recreational activities that promote physical activites ultimately leads to positive energy balance and obesity \citep{Malik2013}.
-In contrast, the high-income group of the population have better access to the facilities that promote weight maintenance, thus have less of an impact from obesity compared with the low- and middle-income groups of the population.\\
+So, for the low- and middle-income groups, the rise in average income in these groups allow the people to have access to food and activities that promotes sedentary bahaviours, but the limited access to the health care system, education and recreational activities that promote physical activites ultimately lead to positive energy balance \citep{Malik2013}.
+In contrast, the high-income group of the population have better access to the facilities that promote weight maintenance, thus have less of an impact compared with the low- and middle-income groups of the population.
+\\
 
-\noindent
-The last factor, which ties in strongly with the first key factor, is the choice of diet made by the population.
-The amount of low-quality food have increased significantly in many rapidly growing countries, as these high-energy, low-nutritional food are cheaper and more profitible than the natural produce that have higher nutritional value \citep{Kearney2010}.
-As a result, more people tend to purchase these low-quality food more often, and consume more energy than they are required.
+% \noindent
+% The last factor, which ties in strongly with the first key factor, is the choice of diet made by the population.
+% The amount of low-quality food have increased significantly in many rapidly growing countries, as these high-energy, low-nutritional food are cheaper and more profitible than the natural produce that have higher nutritional value \citep{Kearney2010}.
+% As a result, more people tend to purchase these low-quality food more often, and consume more energy than they are required.
 
-However, many people eat responsibly and avoid these low-quality diet and prevent overconsumption of such food, but the overconsumption of these food is not a big problem as such, rather the overall quality of the food \citep{Malik2013}.
-These food are usually high in refined sugar and unhealthy fat, highly processed, contain very little nutritional value and are low in fibre.
-As a result, consumption of these food increases the overall energy intake and achieve no nutritional benefits from the food.
+% However, many people eat responsibly and avoid these low-quality diet and prevent overconsumption of such food, but the overconsumption of these food is not a big problem as such, rather the overall quality of the food \citep{Malik2013}.
+% These food are usually high in refined sugar and unhealthy fat, highly processed, contain very little nutritional value and are low in fibre.
+% As a result, consumption of these food increases the overall energy intake and achieve no nutritional benefits from the food.
 
-Whether it is by choice or not, consumption of these low-quality diet leads to obesity and obesity-related diseases. \\
+% Whether it is by choice or not, consumption of these low-quality diet leads to obesity and obesity-related diseases. \\
 
@@
 
-\subsection{Impact of obesity}
-\label{sub:impact_of_obesity}
+% \subsection{Impact of obesity}
+% \label{sub:impact_of_obesity}
 
-As mentioned at the beginning of the current section, obesity has a devastating effect on the health of the population in a global scale.
+% As mentioned at the beginning of the current section, obesity has a devastating effect on the health of the population in a global scale.
+% Overweight and obesity have been a risk factor for many diseases, including type 2 diabetes, cardiovascular diseases, hypertension, and many types of cancers \citep{Franks2010, Guh2009, Yach2006}.
+% As \citet{Yach2006} mentions in their paper, ``overweight and obesity have become to diabetes what tobacco is to lung cancer'', and it is important to understand that obesity has significant risks and comorbidities associated with it.
 
-Overweight and obesity have been a risk factor for many diseases, including type 2 diabetes, cardiovascular diseases, hypertension, and many types of cancers \citep{Franks2010, Guh2009, Yach2006}.
-As \citet{Yach2006} mentions in their paper, ``overweight and obesity have become to diabetes what tobacco is to lung cancer'', and it is important to understand that obesity has significant risks and comorbidities associated with it.
+% Although adulthood obesity is clinically important for risk assessment of diseases, childhood obesity has been focussed more heavily in recent years,  as it has greater public health implications.
+% There are many immediate consequences that are related to childhood obesity, but the diseases that develop as these children grow up are more critical than the earlier effects.
+% A paper by \citet{Must1999} reviewed many of the clinical studies done in the last several decades and summarised the risks associated with childhood obesity.
+% Among the many risks associated with obesity, obese children, compared to lean children, had over 8.5-fold increased risk of developing hypertension as adults \citep{Must1999}.
+% On top of this, the relative risks for all-cause mortality and coronary heart disease-specific mortality were 1.5 and 2.5, respectively \citep{Must1999}.
 
-Although adulthood obesity is clinically important for risk assessment of diseases, childhood obesity has been focussed more heavily in recent years,  as it has greater public health implications.
-There are many immediate consequences that are related to childhood obesity, but the diseases that develop as these children grow up are more critical than the earlier effects.
+% In a more recent study by \citep{Franks2010}, they have shown that childhood obesity (as well as glucose intolerance and hypertension in childhood) was strongly associated with premature deaths from endogenous causes in American Indian population from Arizona (incidence rate ratio of 1.4 per unit of \gls{bmi} z-score).
 
-In the paper written by \citet{Must1999}, they reviewed many of the clinical studies done in the 1900s and summarised the risks associated with childhood obesity.
-Among the many risks associated with obesity, obese children, compared to lean children, had over 8.5-fold increased risk of developing hypertension as adults \citep{Must1999}.
-On top of this, the relative risks for all-cause mortality and coronary heart disease-specific mortality were 1.5 and 2.5, respectively \citep{Must1999}.
+% Health related consequences of obesity are clear from many epidemiological studies, but it is noteworthy that obesity also has an economical impact on the population as well.
+% In the \gls{usa}, within the span of 5 years, the medical cost due to diabetes have risen from \$44 billion to \$92 billion \citep{Yach2006}.
+% Although not all diabetes are directly caused by obesity, obesity would have assisted in the rise in the costs associated with diabetes.
+% In addition to this,
+% The costs that are associated with the treatments of comorbidities of obesity affects the household incomes, increased premature retirement and unemployment, which could have a significant economical impact on the society over time \citep{Yach2006}.
 
-In a more recent paper by \citep{Franks2010}, they have shown that childhood obesity (as well as glucose intolerance and hypertension in childhood) was strongly associated with premature deaths from endogenous causes in American Indian population from Arizona (incidence rate ratio of 1.4 per unit of \gls{bmi} z-score).\\
+% As more of the younger generation is gaining weight and becoming obese, having the risks and disease burden associated with obesity at such early age are not ideal (both economically and epidemiologically) and poses a threat to the growing population that are becoming more obese.
 
-\noindent
-Health related consequences of obesity are clear from these studies, but it is noteworthy that obesity also has an economical impact on the population as well.
-In the \gls{usa}, within the span of 5 years, the medical cost due to diabetes have risen from \$44 billion to \$92 billion \citep{Yach2006}.
-Although not all diabetes are directly caused by obesity, obesity would have assisted in the rise in the costs associated with diabetes.
-In addition to this, the costs that are associated with diabetes treatment affects the household incomes, increased premature retirement and unemployment which, over time, could have a significant economical impact on the society \citep{Yach2006}.\\
+% \subsection{Preventive measures of obesity}
+% \label{sub:preventive_measures_of_obesity}
 
-\noindent
-As more of the younger generation is gaining weight and becoming obese, having the risks and disease burden associated with obesity at such early age are not ideal (both economically and epidemiologically) and poses a threat to the growing population that are becoming more obese.
+% It is apparent from previous sections that obesity has become a serious problem that the world needs to prioritise to solve immediately.
+% Even though obesity has been regarded as one of the many global problems to solve, the progress has been slow.
 
-\subsection{Preventive measures of obesity}
-\label{sub:preventive_measures_of_obesity}
+% \citet{Sacks2009} proposed an \gls{opa} framework that was based on the framework developed by \citet{WHO2006}.
+% In this framework, \citet{Sacks2009} have added extra features so that the policies were targeted to impact three levels of the underlying environment of obesity: the socio-ecological environment (upstream), lifestyle/behavioural changes (midstream), and health services support (downstream).
 
-It is apparent from previous sections that obesity has become a serious problem that the world needs to prioritise to solve immediately.
-Even though obesity has been regarded as one of the many global problems to solve, the progress has been slow.
+% The upstream approach targets the underlying economic, social and physical norms that the population is currently in.
+% By targeting the broad socioeconomic environment that the population lives in, such as taxation, education, employment and housing \citep{Sacks2009}.
+% In doing so, it is possible to have a significant change in the population behaviours and eventually improve the health outcomes of the population.
 
-\citet{Sacks2009} proposed an \gls{opa} framework that was based on the framework developed by \citet{WHO2006}.
-In this framework, \citet{Sacks2009} have added extra features so that the policies were targeted to impact three levels of the underlying environment of obesity: the socio-ecological environment (upstream), lifestyle/behavioural changes (midstream), and health services support (downstream).
+% The midstream approach targets the population or subpopulation behaviour change to improve the eating and physical activity behaviours in these groups.
+% Unlike in the upstream approach, the midstream approach aims to change the behaviour by implementing social marketing and programmes that encourage such changes \citet{Sacks2009}.
+% Again, the result of this approach will be an improvement in the overall positive behaviour of the population towards healthy eating and physical activity.
 
-The upstream approach targets the underlying economic, social and physical norms that the population is currently in.
-By targeting the broad socioeconomic environment that the population lives in, such as taxation, education, employment and housing \citep{Sacks2009}.
-In doing so, it is possible to have a significant change in the population behaviours and eventually improve the health outcomes of the population.
+% The downstream approach targets to support the health services and clinical interventions.
+% Unlike upstream and midstream approaches, downstream approach generally acts on an individual level, whereby the focus is on the management of the health of an individual and/or the individual's family \citet{Sacks2009}.
+% Although not as efficient as upstream or midstream approaches, downstream approach will help improve the habits and behaviours to promote weight management.\\
 
-The midstream approach targets the population or subpopulation behaviour change to improve the eating and physical activity behaviours in these groups.
-Unlike in the upstream approach, the midstream approach aims to change the behaviour by implementing social marketing and programmes that encourage such changes \citet{Sacks2009}.
-Again, the result of this approach will be an improvement in the overall positive behaviour of the population towards healthy eating and physical activity.
+% \noindent
+% In the recent years, obesity has been declared as a global priority \citep{Malik2013}.
+% Furthermore, comprehensive global monitoring framework, multi-sectoral actions and strengthening of the national policies have been implemented to better control the obesity epidemic \citep{Malik2013}.
+% In many countries, the nutritional labelling on products have been adjusted to include the ingredients that have negative impact on obesity (for example, \textit{trans} fat), and even reduce or remove these ingredients from food \citep{Malik2013}.
 
-The downstream approach targets to support the health services and clinical interventions.
-Unlike upstream and midstream approaches, downstream approach generally acts on an individual level, whereby the focus is on the management of the health of an individual and/or the individual's family \citet{Sacks2009}.
-Although not as efficient as upstream or midstream approaches, downstream approach will help improve the habits and behaviours to promote weight management.\\
+% Food marketing can also be a problem in a highly competitive society that we live in.
+% To sell their products that are high in fat, sugar and portion size, food industries market and advertise their products to the population, especially young children, in order to stay ahead in their market place \citep{Chan2010, Malik2013}.
+% Even though there is a strong evidence of food marketing influencing the eating behaviours and food preferences in children, very few countries have been able to reduce these marketing and advertising strategies of unhealthy food \citep{Hawkes2007a, Malik2013}.
 
-\noindent
-In the recent years, obesity has been declared as a global priority \citep{Malik2013}.
-Furthermore, comprehensive global monitoring framework, multi-sectoral actions and strengthening of the national policies have been implemented to better control the obesity epidemic \citep{Malik2013}.
-In many countries, the nutritional labelling on products have been adjusted to include the ingredients that have negative impact on obesity (for example, \textit{trans} fat), and even reduce or remove these ingredients from food \citep{Malik2013}.
+% In contrast to these upstream/midstream approaches, there are some clinical interventions and treatments that help manage weight in an individual.
+% One of these therapies is bariatric surgery for obese patients, especially morbidly obese patients \citep{Buchwald2004}.
+% Through their meta-analysis, \citet{Buchwald2004} showed that bariatric surgery was able to reduce many of the the comorbidities associated with obesity.
+% The fact that these improvements and complete resolution of the comorbidities in high proportion of patients that underwent bariatric surgery shows that, even though it acts on an individual level, downstream approaches such as bariatric surgery may be a highly effective, but not an efficient, method of preventing obesity in the population.\\
 
-Food marketing can also be a problem in a highly competitive society that we live in.
-To sell their products that are high in fat, sugar and portion size, food industries market and advertise their products to the population, especially young children, in order to stay ahead in their market place \citep{Chan2010, Malik2013}.
-Even though there is a strong evidence of food marketing influencing the eating behaviours and food preferences in children, very few countries have been able to reduce these marketing and advertising strategies of unhealthy food \citep{Hawkes2007a, Malik2013}.
+% \noindent
+% It is evident that the most efficient way to reduce the proportion of the population that are obese is to implement national policies that affect the upstream or midstream factors.
+% These policies, although they are efficient at targeting the underlying problem, may take a long time to make a significant effect.
+% On the otherhand, individual therapeutic plans that affect the downstream factors may be very effective in achieving the desired goal of weight loss, but it is very resource intensive and upscaling this approach would be practically impossible.
 
-In contrast to these upstream/midstream approaches, there are some clinical interventions and treatments that help manage weight in an individual.
-One of these therapies is bariatric surgery for obese patients, especially morbidly obese patients \citep{Buchwald2004}.
-Through their meta-analysis, \citet{Buchwald2004} showed that bariatric surgery was able to reduce many of the the comorbidities associated with obesity.
-The fact that these improvements and complete resolution of the comorbidities in high proportion of patients that underwent bariatric surgery shows that, even though it acts on an individual level, downstream approaches such as bariatric surgery may be a highly effective, but not an efficient, method of preventing obesity in the population.\\
-
-\noindent
-It is evident that the most efficient way to reduce the proportion of the population that are obese is to implement national policies that affect the upstream or midstream factors.
-These policies, although they are efficient at targeting the underlying problem, may take a long time to make a significant effect.
-On the otherhand, individual therapeutic plans that affect the downstream factors may be very effective in achieving the desired goal of weight loss, but it is very resource intensive and upscaling this approach would be practically impossible.
-
-The management of obesity at an early stage with midstream and/or downstream approaches while the upstream approach takes effect would be the key to success in controlling for the global obesity epidemic.
+% The management of obesity at an early stage with midstream and/or downstream approaches while the upstream approach takes effect would be the key to success in controlling for the global obesity epidemic.
 
@@
 
+% TODO: Add western-specific and NZ specific statistics on cancer
+% TODO: Add male/female specific stats
+% TODO: Add Maori/PI specific stats
+
 Cancer is considered as one of the leading causes of deaths around the world, accounting for 8.2 million deaths in 2012 \citep{WHO2014}.
 More than half of cancer deaths are caused by lung, breast, colorectal, stomach and liver cancers \citep{WHO2014}.
-
 The leading cause of deaths in high-income countries is lung cancer in both men and women, then colorectal cancer and breast cancer for men and women, respectively \citep{WHO2014}.
-Although the levels of mortality varies among different countries, cervical, liver and stomach cancers make up a large proportion of cancer deaths among low- and middle-income countries \citep{WHO2014}.\\
+Although the levels of mortality vary among different countries, cervical, liver and stomach cancers make up a large proportion of cancer deaths among low- and middle-income countries \citep{WHO2014}.
+\\
 
 \noindent
-\citet{Bray2013} have estimated the global prevalence of cancer in adult population in 2008.
-Since there is no clear definition for cancer prevalence, \citet{Bray2013} used \gls{ldp} as an estimate for cancer prevalence, measuring the incidence of cancer during the period between 2004 and 2008 (5 years) who were still alive at the end of 2008.
-
-From these estimates, Eastern Asia had the greatest 5-year prevalence of cancer with approximately 7 million people with cancer in 2008, followed by North America and Western Europe with approximately 4.7 and 3 million people with cancer \citep{Bray2013}.
-When population of these regions were taken into account and the proportion of cancer prevalence calculated, Western Europe had the highest prevalence proportion of 1.9\%, followed closely by Australia/New Zealand with 1.82\% and North America with 1.7\% \citep{Bray2013}.
+\citet{Bray2013} estimated the global prevalence of cancer in adult population in 2008.
+Since there is no clear definition for cancer prevalence, \citet{Bray2013} used \gls{ldp} as an estimate for cancer prevalence, and measured the incidence of cancer during the period between 2004 and 2008 (5 years).
+% measuring the incidence of cancer during the period between 2004 and 2008 (5 years) who were still alive at the end of 2008.
+From these estimates, Eastern Asia had the greatest 5-year prevalence of cancer, with approximately 7 million people with cancer in 2008, followed by North America and Western Europe with approximately 4.7 and 3 million people with cancer \citep{Bray2013}.
+When the population of these regions was taken into account and the proportion of cancer prevalence calculated, Western Europe had the highest prevalence proportion of 1.9\%, followed closely by Australia/New Zealand with 1.82\% and North America with 1.7\% \citep{Bray2013}.
 
@@
 Furthermore, the prevalence of breast, cervix or prostate cancer is the highest in almost 95\% of the countries \citep{Bray2013}.
-
-Although these estimates were a crude measures of cancer prevalence, \citet{Bray2013} provided a useful overview of what the prevalence of cancer is like around the world.\\
+Although these estimates provide a relatively crude measure of cancer prevalence, \citet{Bray2013} provided a useful overview of what the prevalence of cancer is like around the world.\\
 
@@
 The total number of deaths by non-communicable diseases (including cancer) is predicted to increase, reaching 52 million deaths by 2030 \citep{WHO2014}.
-With the prevalence of cancer at such high proportion around the world, and with the cancer mortality increasing in the future, cancer is a serious burden that the world has to face.
+With the prevalence of cancer at such high levels around the world, and with cancer mortality increasing in the future, cancer is a serious burden that the world has to face.
 
@@
 
-It is now commonly understood that cancers are caused by dysregulation of various mechanisms or pathways that allow tumour cells to prolilferate, survive and migrate.
-By taking advantage of these pathways, cancer cells are able to grow and proliferate, and eventually metastasise to other parts of the body.
-The environment in which the tumours grow can also be an important factor for their evolution and growth, contributing to the heterogeneity of the cancer cells within the tumour or between tumours at different sites of the body.
+It is now commonly understood that cancers are caused by dysregulation of various molecular mechanisms or pathways that allow tumour cells to prolilferate, survive and migrate.
+By taking advantage of these pathways, cancer cells are able to grow, and eventually metastasise to other parts of the body.
+The environment in which the tumours arise can also be an important factor for their evolution and growth, contributing to the heterogeneity of the cancer cells within the tumour or between tumours at different sites of the body.
 
@@
 
-Early researches in cancer biology have been able to uncover many pathways and regulatory mechanisms in cancer, so much that there were too many possibilities in which cancer could develop from.
+Extensive research in cancer biology have been able to uncover many pathways and regulatory mechanisms involved in cancer.
+% , so much that there were too many possibilities in which cancer could develop from.
 With more than 100 distinct types and subtypes of cancers, each with its own distinct development mechanisms, it became difficult to focus on the overarching mechanisms of cancer development \citep{Hanahan2000}.
 To resolve this, \citet{Hanahan2000} proposed ``six essential alterations'' that cancer cells usually take advantage of during development and progression.
-These six key mechanisms were coined the term ``The Hallmarks of Cancer'', and has been used extensively by many researchers as their guide to target their research in specific areas of cancer biology.
+These six key mechanisms were coined as being ``The Hallmarks of Cancer'', and have been used extensively by many researchers as their guide to target their research in specific areas of cancer biology.
 
-10 years after the first review where the Hallmarks of Cancer was proposed, Hanahan and Weinberg revisited these hallmarks and added four more hallmarks on top of the six hallmarks presented in 2000 \citep{Hanahan2011}.
+Ten years after the first review where the concept of Hallmarks of Cancer was proposed, Hanahan and Weinberg revisited these hallmarks and added four more on top of the six hallmarks presented in 2000 \citep{Hanahan2011}.
 The ten Hallmarks of Cancer are: sustaining proliferative signalling, evading growth suppressors, enabling replicative immortality, activating invasion and metastasis, inducing angiogenesis, resisting cell death, avoiding immune destruction, tumour-promoting inflammation, genome instability and mutation, and deregulating cellular energetics (\cref{fig:hallmarks}).
-Each of these hallmarks will be described briefly to provide an overview of the significance of these hallmarks in tumour biology.
+Each of these hallmarks will be described briefly to provide an overview of the significance of these mechanisms in tumour biology.
 
@@
 Cancer is a disease of uncontrolled cell proliferation and division.
-However, without any external mitogenic growth signals, cancer cells cannot proliferate and therefore it is crucial for the cancer cells to maintain a steady proliferative signal.
+Normal cells cannot proliferate without any external mitogenic growth signals, but cancer cells are able to deregulate these signals to maintain a steady proliferative signal \citep{Hanahan2011}.
 
 There are many ways in which cancer cells can maintain this signal.
-Firstly, cancer cells can acquire the ability to produce growth factors and express its corresponding receptor, creating a positive feedback mechanism that continuously stimulates the cells to proliferate \citep{Hanahan2000}.
-Secondly, they can express more receptors for a given growth signals, and thus become ``hyperresponsive'' to the signal \citep{Hanahan2000,Hanahan2011}.
-Thirdly, cancer cells can stimulate the neighbouring normal cells within its microenvironment to produce more growth factors that further assist them to grow \citep{Bhowmick2004, Liotta2001, Wiseman2002}.
-Finally, cancer cells are able to maintain the proliferative signal by mutating the growth factor receptor itself, or the downstream signalling targets of the growth factor receptor pathway \citep{Fuqua1991,SuHuang1997,Satyamoorthy2003}.
-
+Firstly, cancer cells can acquire the ability to produce growth factors and express the corresponding receptor, creating a positive feedback mechanism that continuously stimulates the cells to proliferate \citep{Hanahan2000}.
+Secondly, they can express more receptors for a given growth signal, and thus become ``hyperresponsive'' to the signal \citep{Hanahan2000,Hanahan2011}.
+Thirdly, cancer cells can stimulate the neighbouring normal cells within the tumour microenvironment to produce more growth factors that further assist them to grow \citep{Bhowmick2004, Liotta2001, Wiseman2002}.
+Finally, cancer cells are able to maintain proliferative signalling by mutating the growth factor receptor itself, or the downstream signalling targets of the growth factor receptor pathway \citep{Fuqua1991,SuHuang1997,Satyamoorthy2003}.
 By maintaining the the proliferative signals, cancer cells are able to continuously grow in their environment.
@@
 \noindent
-As well as maintaining proliferative signals, cancer cells must also evade the signals that inhibit cell growth.
-Two tumour suppressor proteins have an essential role in this evasion: the \gls{rb} and p53 proteins.
-
-\Gls{rb} protein is a known tumour suppressor protein that is important in cell cycle regulation \citep{Burkhart2008,Hanahan2011}.
-\Gls{rb} protein is known to arrest the cells in \gls{g1} phase of the cell cycle by regulating E2F transcription factors in the cell, in reponse to molecules such as \gls{tgfb} \citep{Burkhart2008,Hanahan2000}.
-Although \gls{rb} protein is known for its role as a checkpoint in \gls{g1} phase, it also takes part in the regulation of apoptotic cell death, cell differentiation and chromosome stability \citep{Burkhart2008}.
-The contribution of the loss of function of the \gls{rb} protein in tumorigenesis and tumour progression is evident, but at which stage of the development this loss of function mutation is crucial for the cancer cells is yet to be clarified \citep{Burkhart2008}.
-
-p53 protein is another tumour suppressor protein that is closely related to the \gls{rb} protein and its role in cell cycle arrest \citep{Hanahan2011,Levine1997}.
-As \citet{Hanahan2011} mentions in their review, ``\gls{rb} transduces growth-inhibitory signals that originate largely outside of the cell, TP53 receives inputs from stress and abnormality sensors that function within the cell's intracellular operating systems''.
+As well as maintaining proliferative signals, cancer cells must also evade signals that inhibit cell growth.
+p53 protein is a tumour suppressor protein that has a  role in cell cycle arrest \citep{Hanahan2011,Levine1997}.
+As \citet{Hanahan2011} mention in their review, ``TP53 receives inputs from stress and abnormality sensors that function within the cell's intracellular operating systems''.
 The level of expression and its stability is crucial for the function of p53 protein, as the abundance of p53 protein in the cell triggers cell cycle arrest or cell \gls{apoptosis}, depending on the extent of the cell signal \citep{Fridman2003,Hanahan2011,Levine1997}.
 When abundant and active, p53 protein transcribes the p21 protein which, through series of events in the p16-cyclin D$_1$-cdk4-\gls{rb} pathway, triggers cell cycle arrest \citep{Levine1997}.
+In many cancers, p53 protein is usually mutated and selected against to allow continuous growth of the tumour cells.
 
-In many cancers, these proteins are usually mutated and selected against to allow continuous growth of the tumour cells.
+% Two tumour suppressor proteins have an essential role in this evasion: the \gls{rb} and p53 proteins.
+
+% \Gls{rb} protein is a known tumour suppressor protein that is important in cell cycle regulation \citep{Burkhart2008,Hanahan2011}.
+% \Gls{rb} protein is known to arrest the cells in \gls{g1} phase of the cell cycle by regulating E2F transcription factors in the cell, in reponse to molecules such as \gls{tgfb} \citep{Burkhart2008,Hanahan2000}.
+% Although \gls{rb} protein is known for its role as a checkpoint in \gls{g1} phase, it also takes part in the regulation of apoptotic cell death, cell differentiation and chromosome stability \citep{Burkhart2008}.
+% The contribution of the loss of function of the \gls{rb} protein in tumorigenesis and tumour progression is evident, but at which stage of the development this loss of function mutation is crucial for the cancer cells is yet to be clarified \citep{Burkhart2008}.
+
+% p53 protein is another tumour suppressor protein that is closely related to the \gls{rb} protein and its role in cell cycle arrest \citep{Hanahan2011,Levine1997}.
+% As \citet{Hanahan2011} mentions in their review, ``\gls{rb} transduces growth-inhibitory signals that originate largely outside of the cell, TP53 receives inputs from stress and abnormality sensors that function within the cell's intracellular operating systems''.
+% The level of expression and its stability is crucial for the function of p53 protein, as the abundance of p53 protein in the cell triggers cell cycle arrest or cell \gls{apoptosis}, depending on the extent of the cell signal \citep{Fridman2003,Hanahan2011,Levine1997}.
+% When abundant and active, p53 protein transcribes the p21 protein which, through series of events in the p16-cyclin D$_1$-cdk4-\gls{rb} pathway, triggers cell cycle arrest \citep{Levine1997}.
 
@@
 
-\Gls{bcl2} family proteins are the major players of apoptosis.
-Embedded in the mitochondrial membrane, the Bax subfamily of \gls{bcl2} proteins are pro-apoptotic, and are under constant repression by the anti-apoptotic proteins of the \gls{bcl2} subfamily proteins \citep{Adams2007,Hanahan2011}.
-The inhibition of the pro-apoptotic proteins are achieved by the interaction between the pro- and anti-apoptotic proteins via the \gls{bh3} domain \citep{Adams2007}.
-In addition to the pro-apoptotic Bax subfamily proteins, there is a family of proteins that contain a single \gls{bh3} domain, and hence named \gls{bh3}-only family proteins \citep{Adams2007}.
-\gls{bh3}-only proteins either disrupt the inhibition of the Bax family proteins by the \gls{bcl2} subfamily proteins, or directly stimulate the Bax subfamily proteins \citep{Adams2007}.
-The activation of the pro-apoptotic proteins disrupt the mitochondrial cell membrane adn releases cytochrome \textit{c}, which in turn triggers the activation of the apoptotic proteases caspase 8 and caspase 9, resulting in apoptosis \citep{Adams2007,Hanahan2011}.
+% \Gls{bcl2} family proteins are the major players of apoptosis.
+% Embedded in the mitochondrial membrane, the Bax subfamily of \gls{bcl2} proteins are pro-apoptotic, and are under constant repression by the anti-apoptotic proteins of the \gls{bcl2} subfamily proteins \citep{Adams2007,Hanahan2011}.
+% The inhibition of the pro-apoptotic proteins are achieved by the interaction between the pro- and anti-apoptotic proteins via the \gls{bh3} domain \citep{Adams2007}.
+% In addition to the pro-apoptotic Bax subfamily proteins, there is a family of proteins that contain a single \gls{bh3} domain, and hence named \gls{bh3}-only family proteins \citep{Adams2007}.
+% \gls{bh3}-only proteins either disrupt the inhibition of the Bax family proteins by the \gls{bcl2} subfamily proteins, or directly stimulate the Bax subfamily proteins \citep{Adams2007}.
+% The activation of the pro-apoptotic proteins disrupt the mitochondrial cell membrane adn releases cytochrome \textit{c}, which in turn triggers the activation of the apoptotic proteases caspase 8 and caspase 9, resulting in apoptosis \citep{Adams2007,Hanahan2011}.
 
@@
 When \acrshort{dna} damage is sensed, p53 protein transcribes the pro-apoptotic members of the \gls{bcl2} family proteins, such as Bax, Puma and Noxa \citep{Fridman2003,Hanahan2011}.
-The fact that p53 protein is involved not only in cell cycle regulation, but also in apoptotic pathway shows its importance in the control of cancer development, and perhaps that is the reason why more than 50\% of tumours have a mutation in the \textit{p53} gene \citep{Levine1997}.
+The fact that p53 protein is involved not only in cell cycle regulation, but also in apoptotic pathway shows its importance in the control of cancer development, and perhaps that is the reason why more than 50\% of tumours have a mutation in the \textit{TP53} gene \citep{Levine1997}.
 
@@
 When cells are grown in a culture, cells are able to replicate and grow until a certain point, and the cells reach the stage of senescence where the cells stop growing but are still viable \citep{Hanahan2011}.
-The cells that manage to bypass this senescence phase, by mutating the tumour suppressor proteins such as p53 and \gls{rb} proteins, then undergoes the crisis phase where the majority of the cells in the culture dies \citep{Hanahan2011}.
-Very rarely, some cells are able to survive the crisis phase and acquires the ability to replicate infinitely, and this transition has been termed immortalisation \citep{Hanahan2011, Wright1989}.
-
+The cells that manage to bypass this senescence phase by mutating the tumour suppressor proteins such as p53 protein, undergoes the crisis phase where the majority of the cells in the culture dies \citep{Hanahan2011}.
+Very rarely, some cells are able to survive the crisis phase and acquire the ability to replicate infinitely, and this transition has been termed ``immortalisation'' \citep{Hanahan2011, Wright1989}.
 Without immortalisation, cancer cells are not able to generate a macroscopic tumours that are capable of killing the host organism \citep{Hanahan2000,Hanahan2011}.
@@
 Like normal cells, cancer cells also require access to oxygen and nutrient supply from the blood.
-However, new blood vessel formation, or angiogenesis, only occurs in a handful of occasions, and thus the tumour needs to acquire the ability to stimulate angiogenesis in the microenvironment in which they live in \citep{Hanahan2011}.
+However, new blood vessel formation, or angiogenesis, only occur in a handful of occasions, and thus the tumour needs to acquire the ability to stimulate angiogenesis in the microenvironment in which they live in \citep{Hanahan2011}.
 
-It is thought that the angiogenic ability of the tumour is gained at an early stage of its development by disrupting the ``angiogenic switch'' \citep{Hanahan2011}.
-Like many cellular processes, angiogenesis is controlled by the balance between the inducing and inhibiting agents of angiogenesis.
+It is thought that the angiogenic ability of the tumour is gained at an early stage of tumour development by disrupting the ``angiogenic switch'' \citep{Hanahan2011}.
+% Like many cellular processes, angiogenesis is controlled by the balance between the inducing and inhibiting agents of angiogenesis.
 \Gls{vegfa} is the major growth factor that triggers angiogenesis by signalling through the \acrshort{vegf} receptors (though some receptors have inhibitory effect) \citep{Yancopoulos2000}.
 Opposing the effect of \Gls{vegfa} is the \gls{tsp1}, a protein that is known to inhibit the process of angiogenesis by affecting the bioavailability of \acrshort{vegf}, as well as promoting apoptosis \citep{Kazerounian2008}.
-
 In general, many cancers stimulate the expression of \gls{vegfa} protein and reduces the expression of \gls{tsp1} \citep{Kazerounian2008}.
@@
 Metastasis is a multi-step process that requires the selection and survival of cancer cell(s) that is able to metastasise (``seed''), and is also compatible with the specific target tissue where they prosper (``soil'') \citep{Talmadge2010}.
-The downregulation or loss of function of E-cadherin, a protein important in cell-to-cell connection and adherence, is known to trigger tumour invasion and metastasis through the reversible process of \gls{emt} \citep{Hanahan2011,Kalluri2009}.
-Once mobile and invasive, these metastatic cancer cell is able to circulate through the vascular system and eventually invade the target organ, where it must undergo \gls{met} and colonise the site \citep{Hanahan2011,Kalluri2009}.
-
+% The downregulation or loss of function of E-cadherin, a protein important in cell-to-cell connection and adherence, is known to trigger tumour invasion and metastasis through the reversible process of \gls{emt} \citep{Hanahan2011,Kalluri2009}.
+Once mobile and invasive, metastatic cancer cell is able to circulate in the vascular system and eventually invade the target organ, where it must undergo \gls{met} and colonise the site \citep{Hanahan2011,Kalluri2009}.
 There are numerous signals, growth factors and pathways that initiate the metastatic phenotype in cancer cell, and it remains an active field of research \citep{Hanahan2011,Kalluri2009}.
@@
 \noindent
-Many of the hallmarks are acquired by the cancer cells during its growth, through many genomic changes and mutations of the essential genes that allow cancer cells to survive, proliferate, invade and metastasise.
+Many of the hallmarks are acquired by the cancer cells during tumout growth, through many genomic changes and mutations of the essential genes that allow cancer cells to survive, proliferate, invade and metastasise.
 Thus, it is only logical to induce that many, if not all, cancers are more prone to genetic mutations and restructuring, which allows the cancers to subsequently acquire the already described Hallmarks of Cancer \citep{Hanahan2011}.
@@
 There are now evidence that show otherwise, whereby a single catastrophic event in the chromosome structure occurs, resulting in major structural rearrangements such as inversions, deletions, duplications and translocations \citep{Leibowitz2015,Stephens2011}.
-This rare event that is estimated to occur in 2$\sim$5\% of all cancers was termed \gls{chromothripsis}, and is thought to be caused by the physical separation of parts of chromosome by a structure called micronuclei \citep{Leibowitz2015,Stephens2011}.
+% This rare event that is estimated to occur in 2$\sim$5\% of all cancers was termed \gls{chromothripsis}, and is thought to be caused by the physical separation of parts of chromosome by a structure called micronuclei \citep{Leibowitz2015,Stephens2011}.
 Owing to its nature of being a one-off chromosomal disruption, chromothripsis (usually) cause disruptions in tumour suppressor genes rather than duplications of oncogenes \citep{Leibowitz2015}.
@@
 Although the full mechanism in which \gls{kataegis} occur is not yet defined, \gls{aid}/\acrshort{apobec} family proteins, a family of proteins involved in gene editing, may play a major role in \gls{kataegis} \citep{Leibowitz2015,Nik-Zainal2012}.
-By deaminating cytidine to uracil in the \gls{ssdna} that may occur at the sites of double strand break, \gls{aid}/\acrshort{apobec} family proteins introduces hypermutated regions in the genome \citep{Leibowitz2015,Nik-Zainal2012}.
-
-It is evident that both \gls{chromothripsis} and \gls{kataegis} contribute to the genomic aberrations and mutations that ultimately contribute to many of the tumorigenic phenotypes.
+% By deaminating cytidine to uracil in the \gls{ssdna} that may occur at the sites of double strand break, \gls{aid}/\acrshort{apobec} family proteins introduces hypermutated regions in the genome \citep{Leibowitz2015,Nik-Zainal2012}.
+It is evident that both \gls{chromothripsis} and \gls{kataegis} contribute to the genomic aberrations and mutations that ultimately encourage the acquisition of many of the tumorigenic phenotypes.
 
@@
 \noindent
-Inflammation and immune response, though it may assist in the control of diseases and injuries, have been implicated in the progression of tumour and their acquisition of many of the hallmarks \citep{Hanahan2011}.
-
-There are two pathways of inflammation that contribute to the cancer survival and progression \citep{Mantovani2008}.
-The intrinsic pathway is caused by oncogenic characteristics of the cancer cells, where these cells produce inflammatory signals themselves and promote advantageous characteristics for the cancer cells, such as increased proliferation and survival \citep{Mantovani2008}.
-On the other hand, the external pathway is caused by infectious or inflammatory conditions in the tumour microenvironment that surround the tumour and assist in the growth of the tumour \citep{Mantovani2008}.
-There are two key transcription factors that connect these two pathways of inflammation that ultimately leads to the prommotion  of many, if not all, of the hallmarks discussed so far: \gls{nfkb} and \gls{stat3} \citep{Mantovani2008}.
+Inflammation and immune response, though it may assist in the control of diseases and injuries, have been implicated in the progression of tumour and their acquisition of many of the cancer hallmarks \citep{Hanahan2011}.
+% There are two pathways of inflammation that contribute to the cancer survival and progression \citep{Mantovani2008}.
+% The intrinsic pathway is caused by oncogenic characteristics of the cancer cells, where these cells produce inflammatory signals themselves and promote advantageous characteristics for the cancer cells, such as increased proliferation and survival \citep{Mantovani2008}.
+% On the other hand, the external pathway is caused by infectious or inflammatory conditions in the tumour microenvironment that surround the tumour and assist in the growth of the tumour \citep{Mantovani2008}.
+% There are two key transcription factors that connect these two pathways of inflammation that ultimately leads to the prommotion  of many, if not all, of the hallmarks discussed so far: \gls{nfkb} and \gls{stat3} \citep{Mantovani2008}.
+There are two key transcription factors of inflammation that ultimately leads to the prommotion  of many, if not all, of the hallmarks discussed so far: \gls{nfkb} and \gls{stat3} \citep{Mantovani2008}.
 
@@
 \gls{nfkb} regulates the expression of genes that encode the inflammatory cytokines, angiogenic factors, and many others, including anti-apoptotic genes (through the activation of \gls{stat3}) \citep{Elinav2013,Mantovani2008}.
-\gls{stat3} is a transcription factor that is activated through variety of receptors and pathways, including the \gls{il6} receptor, \glspl{gpcr} and \glspl{tlr} \citep{Yu2007,Yu2014}.
+% \gls{stat3} is a transcription factor that is activated through variety of receptors and pathways, including the \gls{il6} receptor, \glspl{gpcr} and \glspl{tlr} \citep{Yu2007,Yu2014}.
+\gls{stat3} is a transcription factor that is activated through variety of receptors and pathways \citep{Yu2007,Yu2014}.
 By increasing the transcription of anti-apoptotic proteins such as \gls{bcl2}, \gls{stat3} promotes tumour survival and proliferation \citep{Yu2007}.
-Furthermore, \gls{stat3} is able to increase the expression of inflammatory cytokines (such as \gls{il6} and \acrshort{il10}) and \acrshort{vegf}, which provides a positive feedback mechanism for further inflammation (and therefore survival and proliferation signals), and advancement of angiogenesis, respectively \citep{Yu2007}.
+Furthermore, \gls{stat3} is able to increase the expression of inflammatory cytokines (such as \gls{il6} and \acrshort{il10}) and \acrshort{vegf}, which provides a positive feedback mechanism for further inflammation (and therefore survival and proliferation signals) and advancement of angiogenesis, respectively \citep{Yu2007}.
 
-Though helpful at times, inflammation induced by the immune system must be taken care of and controlled in order to prevent rapid tumour development and progression.
+% Though helpful at times, inflammation induced by the immune system must be taken care of and controlled in order to prevent rapid tumour development and progression.
 
@@
 In order to grow and proliferate, cancer cells must adjust their energy metabolism to match their energy requirement \citet{Hanahan2011}.
-There are two ways in which a cell can produce energy: anaerobic respiration, where the cell uses glycolysis to rapidly generate \gls{atp} and lactate; and aerobic respiration where, in the presence of oxygen, the cell utilises the electron transport chain and \gls{atp} synthase in the mitochondria to efficiently produce \gls{atp}.
-However, even with abundant oxygen in the cell, \citet{Wardburg1956} found that cancer cells utilised the glycolytic pathway instead of the more efficient aerobic respiration, which was later termed ``aerobic glycolysis'' \citep{Hanahan2011}.
-
+% There are two ways in which a cell can produce energy: anaerobic respiration, where the cell uses glycolysis to rapidly generate \gls{atp} and lactate; and aerobic respiration where, in the presence of oxygen, the cell utilises the electron transport chain and \gls{atp} synthase in the mitochondria to efficiently produce \gls{atp}.
+% However, even with abundant oxygen in the cell, \citet{Wardburg1956} found that cancer cells utilised the glycolytic pathway instead of the more efficient aerobic respiration, which was later termed ``aerobic glycolysis'' \citep{Hanahan2011}.
+Even with abundant oxygen in the cell, \citet{Wardburg1956} found that cancer cells utilised the glycolytic pathway instead of the more efficient aerobic respiration, which was later termed ``aerobic glycolysis'' \citep{Hanahan2011}.
 It seems counterintuitive to limit its full potential of creating \gls{atp} by utilising glycolysis over electron transport chain, as the latter produces almost 18 times more \gls{atp} \citep{Hanahan2011, VanderHeiden2009}.
@@
 Our immune system is constantly monitoring and maintaining normal cell biology within our body, preventing the emergence of cancer \citep{Hanahan2011}.
-That said, cancers are frequently observed in clinics, pointing to the fact that the observed cancer cells must have evaded the diligent monitoring by the immune system \citep{Hanahan2011}.
+With that said, cancers are frequently observed in clinics, pointing to the fact that the observed cancer cells must have evaded the diligent monitoring by the immune system \citep{Hanahan2011}.
 
@@
 One possible reason for this was that the cancer cells from the immunodeficient mice were not immunoedited to the extent in which the cells from immunocompetent mice were, and thus unable to evade the immune system as efficiently as those cells in the immunocompetent mice \citep{Hanahan2011}.
-
 Though more experiments must be carried out to confirm its definitive role and mechanisms in which it contributes to cancer biology, immune system evasion is in no doubt a significant component of tumour progression.
@@
 There is an evident need for an efficient way to monitor and treat cancer patients, and these Hallmarks of Cancers will prove to be the best target for effective treatments of cancer in the future.
-
-% TODO: do I really need tumour heterogeneity??
-\subsubsection{Tumour heterogeneity (?)}
-\label{subsub:tumour_heterogeneity}
-TBC
-
-\subsubsection{Cancer stem cells (?)}
-\label{ssub:cancer_stem_cells}
-TBC
 
@@
 
-It is quite obvious from the previous section (\cref{sub:mechanisms_of_cancer_development}: \nameref{sub:mechanisms_of_cancer_development}) that there are a lot of things that can go wrong in the cell.
+% It is quite obvious from the previous section (\cref{sub:mechanisms_of_cancer_development}) that there are a lot of things that can go wrong in the cell.
 Any one of the ten Hallmarks of Cancers has the potential to initiate cancer growth and progression; every hallmark has their own complex mechanistic pathways; every pathway has many proteins associated with its function or role in the cell; and, though their impacts may differ from protein to protein, mutations in any one of the proteins involved in a pathway has the potential to cause cancer.
 
-% TODO: more citations!!!!
-Broadly speaking, there are three major classes of genes that have significance in the initiation of cancer: oncogenes, tumour suppressor genes and stability genes \citep{Vogelstein2004a}.
 Oncogenes are genes that promote or ``accelerate'' the process of tumorigenesis and/or tumour progression, whereas tumour suppressor genes prevent or ``decelerate'' those effects \citep{Vogelstein2004a}.
-Cancers usually mutate the oncogenes in such a way that it it constitutively activated or expressed at a very high level, and mutate the tumour suppressor genes so that these genes are inactivated and/or have reduced expression \citep{Vogelstein2004a}.
+Cancers usually mutate oncogenes in such a way that they become constitutively activated, and mutate the tumour suppressor genes so that these genes are inactivated and/or have reduced expression \citep{Vogelstein2004a}.
 Another thing to note for these genes is that oncogenes usually only require a ``one-hit'' mutation that affects a single allele, whereas tumour suppressor genes require mutations in both maternal and paternal alleles to have any effect \citep{Stratton2009,Vogelstein2004a}.
@@
 Finally, diet and obesity are known risk factors for many types of cancers \citep{Ames1995a,Calle2004}.
-Though more will be covered in details in \cref{sec:obesity_and_cancer}: \nameref{sec:obesity_and_cancer}, the effect of obesity on cancers will be briefly mentioned here.
+Though more will be covered in details in \cref{sec:obesity_and_cancer}, the effect of obesity on cancers will be briefly mentioned here.
 
@@
 
-As mentioned briefly in \cref{sub:causes_and_risk_factors_of_cancer}: \nameref{sub:causes_and_risk_factors_of_cancer}, obesity is considered a major risk factor for many types of cancers.
+As mentioned briefly in \cref{sub:causes_and_risk_factors_of_cancer}, obesity is considered a major risk factor for many types of cancers.
 The fact that the prevalence of obesity is extremely high, it is important to establish the link between the apparent risks associated with cancer.
@@
 In this study, \citet{Gatza2010a} used 18 pathway signatures to classify human breast tumours into more detailed subtypes.
+The 18 pathways used in this study were: Akt, \gls{bcat}, E2F1, \gls{egfr}, \gls{er}, \gls{her2}, \gls{ifna}, \gls{ifny}, Myc, p53, p63, \gls{pi3k}, \gls{pr}, Ras, \gls{stat3}, Src, \gls{tgfb}, and \gls{tnfa} pathways.
 This was a significant improvement from their previous works on pathway signatures \citep{Bild2006,Bild2009}.
 Using more pathways in their analysis have allowed them to identify certain pathways to coactivate (or cluster) together, suggesting that some pathways are related to one another and may have relevance in identifying the causes.
+% TODO: list 18 pathways.
 

### 2017-01-23 thesis/intro/intro.tex
Finished editing introduction - tmp commit

@@
 \noindent
-Many of the hallmarks are acquired by the cancer cells during tumout growth, through many genomic changes and mutations of the essential genes that allow cancer cells to survive, proliferate, invade and metastasise.
+Many of the hallmarks are acquired by the cancer cells during tumour growth, through many genomic changes and mutations of the essential genes that allow cancer cells to survive, proliferate, invade and metastasise.
 Thus, it is only logical to induce that many, if not all, cancers are more prone to genetic mutations and restructuring, which allows the cancers to subsequently acquire the already described Hallmarks of Cancer \citep{Hanahan2011}.
@@
 Oncogenes are genes that promote or ``accelerate'' the process of tumorigenesis and/or tumour progression, whereas tumour suppressor genes prevent or ``decelerate'' those effects \citep{Vogelstein2004a}.
-Cancers usually mutate oncogenes in such a way that they become constitutively activated, and mutate the tumour suppressor genes so that these genes are inactivated and/or have reduced expression \citep{Vogelstein2004a}.
-Another thing to note for these genes is that oncogenes usually only require a ``one-hit'' mutation that affects a single allele, whereas tumour suppressor genes require mutations in both maternal and paternal alleles to have any effect \citep{Stratton2009,Vogelstein2004a}.
+Cancers usually mutate oncogenes in such a way that they become constitutively activated and mutate the tumour suppressor genes so that these genes are inactivated and/or have reduced expression \citep{Vogelstein2004a}.
+Another thing to note for these genes is that oncogenes usually only require a ``one-hit'' mutation that affects a single allele, whereas tumour suppressor genes generally require mutations in both maternal and paternal alleles to have any effect \citep{Stratton2009,Vogelstein2004a}.
 
 Stability genes, also known as caretakers, are genes that are involved in, for example,  \gls{mmr}, \gls{ner} and \gls{ber}; in other words, genes that are involved in \acrshort{dna} repair mechanisms \citep{Vogelstein2004a}.
-These caretaker genes in effect prevents the cell to develop new mutations from \acrshort{dna} replication and mitosis.
-However, only the oncogenes or the tumour suppressor genes are able to directly affect the biology of the cell, and therefore mutations in the stability genes affect the mutation rates of the other classes of genes, ultimately leading to tumorigenesis \citep{Vogelstein2004a}.
-\\
+These caretaker genes help prevent the cell from developing new mutations during \acrshort{dna} replication and mitosis.
+However, only the oncogenes or the tumour suppressor genes are able to directly affect the biology of the cell, and therefore mutations in stability genes tend to affect the mutation rates of the other classes of genes, ultimately leading to tumorigenesis \citep{Vogelstein2004a}.
 
-\noindent
 There are two types of mutations that occurs in an individual: germline mutation and somatic mutation.
-Germline mutations are the mutations that are passed on from your parents, and somatic mutations are the mutations acquired by the cell over the course of the individual's lifetime.
+Germline mutations are the mutations that are passed on from an individual's parents, and somatic mutations are the mutations acquired by the cell over the course of the individual's lifetime.
 
 Unfortunately, there are germline mutations that make the individual more prediposed to cancer \citep{Vogelstein2004a}.
-These individuals have a mutation in one or more of the classes of cancer-related genes, which makes them more likely to develop tumour than their normal pairs \citep{Vogelstein2004a}.
-As a consequence, these individuals may develop multiple tumours at an early stage of their lives than those who develop tumours from somatic mutations alone, as these individuals already have a mutation that allows all of the cells in their body to have one or more of the Hallmarks of Cancers \citep{Vogelstein2004a}.
+These individuals have a mutation in one or more of the classes of cancer-related genes, which makes them more likely to develop a tumour than individuals without the mutation \citep{Vogelstein2004a}.
+In contrast, somatic mutations are acquired during an individual's lifetime and is not passed down to their children.
+Somatic mutations can either be a driver or a passenger mutation, where a driver mutation helps tumorigenesis and its subsequent progression (in other words, mutations in oncogene, tumour suppressor, or stability gene), and passenger mutation have neither a selective advantage nor contribute to tumour progression \citep{Stratton2009}.
 
-In contrast, somatic mutations are acquired during an individual's lifetime and is not passed down to their children.
-Somatic mutations can either be a driver or a passenger mutation, where a driver mutation helps tumorigenesis and its subsequent progression (in other words, mutations in oncogene, tumour suppressor, or stability gene), and passenger mutation is where the mutation has no selective advantage for the cancer and does not contribute to tumour progression \citep{Stratton2009}.
 Since cancers require multiple driver mutations, somatic mutations are an essential part of tumour development.
@@
 
-For obesity, since there is a clear evidence of the cause (excessive eating and lack of exercise) and a well defined mechanism of how the disease is developed (excess calories are converted into fat), it is relatively easy to speculate the effect of the environment on the disease (for example, abundance of high calory food).
-On the other hand, cancers are developed through many cellular pathways (Hallmarks of Cancers) which are not fully uncovered yet, with either too many or very few causes associated with those pathways (for example, mutations, inflammation and/or growth factors), and therefore it is difficult to assess how, rather than whether, a given environmental factor causes specific cancer.
+% For obesity, since there is a clear evidence of the cause (excessive eating and lack of exercise) and a well defined mechanism of how the disease is developed (excess calories are converted into fat), it is relatively easy to speculate the effect of the environment on the disease (for example, abundance of high calory food).
+Cancers are developed through many cellular pathways (Hallmarks of Cancers) which are not fully uncovered yet, with either too many or very few causes associated with those pathways.
+Therefore it is difficult to assess how, rather than whether, a given environmental factor causes specific cancer.
+With that said, several environmental factors have a strong evidence of causing cancer.
+% , though the complete mechanism of action for some of these must be elucidated in the coming years.
+% Out of these diet and obesity will be covered as an example.
 
-With that said, there has been a few environmental factors that have a strong evidence of causing cancer, though the complete mechanism of action for some of these must be elucidated in the coming years.
-Out of these, smoking, \gls{uv} radiation, and diet and obesity will be covered as examples.
-\\
-
-% Smoking
-\noindent
 Smoking is a known risk factor for many cancers; the obvious one being lung cancer \citep{Gandini2008,Hecht1999}.
 For current smokers, the relative risk is greater than 1.5 in many cancers, including liver, cervix and kidney, and it can rise up to 8.96 in lung cancer \citep{Gandini2008}.
-Even if smokers stop smoking, the relative risk does not decrease until the first 5 years have passed and will never return to the level of non-smokers \citep{Hecht1999}.
+% Even if smokers stop smoking, the relative risk does not decrease until the first 5 years have passed and will never return to the level of non-smokers \citep{Hecht1999}.
+% In tobacco smoke, there are literally hundreds and thousands of chemical compounds, of which about 20 compounds have been shown to have carcinogenic effects \citep{Hecht1999}.
+% Among these, \glspl{pah} and \gls{nnk} have been shown to cause lung cancer in mice when administered systemically \citep{Hecht1999}.
+Tobacco smoke contain many carcinogens which are processed first by \gls{cyp} enzymes (termed meta\-bolic activation), and the product can either be processed further to be detoxified completely and cause no harm to the cell, or alternatively, the product can covalently bind to the \acrshort{dna}, forming \acrshort{dna} adducts \citep{Hecht1999}.
+Unless these adducts are resolved correctly by the \acrshort{dna} repair mechanisms, it can result in a permanent mutation within the genome and lead to oncogenesis \citep{Hecht1999}.
+% In addition to this, cigarettes contain free radicals and chemicals that cause tremendous oxidative stress and damage to the cells \citep{Hecht1999}.
+% These radicals cause \acrshort{dna} nicks and damages the \acrshort{dna}, again requring the cell's \acrshort{dna} repair mechanisms to remove the damage so it does not become permanent.
 
-In tobacco smoke, there are literally hundreds and thousands of chemical compounds, of which about 20 compounds have been shown to have carcinogenic effects \citep{Hecht1999}.
-Among these, \glspl{pah} and \gls{nnk} have been shown to cause lung cancer in mice when administered systemically \citep{Hecht1999}.
-These carcinogens are processed first by \gls{cyp} enzymes (termed meta\-bolic activation), and the product can either be processed further to be detoxified completely and cause no harm to the cell, or alternatively, the product can covalently bind to the \acrshort{dna}, forming \acrshort{dna} adducts \citep{Hecht1999}.
-Unless these adducts are resolved correctly by the \acrshort{dna} repair mechanisms, it can result in a permanent mutation within the genome and lead to oncogenesis \citep{Hecht1999}.
+% Another important environmental factor to consider is \gls{uv} radiation.
+Exposure to \gls{uv} radiation has been significantly associated with the increased risks of skin cancers \citep{Armstrong2001,Gallagher2006}.
+% The risks of developing any skin cancers can vary between race (Caucasian people are more likely to get skin cancers), and potentially the age when they were exposed to high amounts of \gls{uv} radiation \citep{Armstrong2001,Gallagher2006}.
+Similar to smoking, the mechanism of how \gls{uv} radiation may cause  skin cancers is related to the \acrshort{dna} repair system.
+\gls{uv} radiation causes the formation of a dimer in the adjacent pyrimidines, resulting in a covalent bond formation between these nucleotides \citep{Friedberg2003,Hoeijmakers2001}.
+These dimers introduce bulky structures in the \acrshort{dna} that prevent the \acrshort{dna} replication and transcription enzymes from functioning properly, and it may cause lasting mutations if not repaired \citep{Friedberg2003,Hoeijmakers2001}.
+These damages caused by \gls{uv} radiation can be relieved by the \gls{ner} system or the photoreactivation process, where a photoreactivating enzyme uses light to monomerize the dimeric pyrimidines \citep{Friedberg2003}.
 
-In addition to this, cigarettes contain free radicals and chemicals that cause tremendous oxidative stress and damage to the cells \citep{Hecht1999}.
-These radicals cause \acrshort{dna} nicks and damages the \acrshort{dna}, again requring the cell's \acrshort{dna} repair mechanisms to remove the damage so it does not become permanent.
-\\
-
-% UV radiation
-\noindent
-Another important environmental factor to consider is \gls{uv} radiation.
-Exposure to \gls{uv} radiation has been significantly associated with the increased risks of skin cancers \citep{Armstrong2001,Gallagher2006}.
-The risks of developing any skin cancers can vary between race (Caucasian people are more likely to get skin cancers), and potentially the age when they were exposed to high amounts of \gls{uv} radiation \citep{Armstrong2001,Gallagher2006}.
-
-Again, the mechanism of how \gls{uv} radiation may cause  skin cancers is related to the \acrshort{dna} repair system.
-\gls{uv} radiation causes the formation of a dimer in the adjacent pyrimidines, resulting in a covalent bond formation between these nucleotides \citep{Friedberg2003,Hoeijmakers2001}.
-These dimers introduce bulky structures in the \acrshort{dna} that prevent the \acrshort{dna} replication and transcription enzymes from functioning properly, and if they are not repaired, it may cause lasting mutations \citep{Friedberg2003,Hoeijmakers2001}.
-These damages caused by \gls{uv} radiation can be relieved by the \gls{ner} system or the photoreactivation process, where a photoreactivating enzyme uses light to monomerize the dimeric pyrimidines \citep{Friedberg2003}.
-\\
-
-% Diet and obesity
-\noindent
-Finally, diet and obesity are known risk factors for many types of cancers \citep{Ames1995a,Calle2004}.
-Though more will be covered in details in \cref{sec:obesity_and_cancer}, the effect of obesity on cancers will be briefly mentioned here.
-
+Diet and obesity are known risk factors for many types of cancers as well \citep{Ames1995a,Calle2004}.
+% Though more will be covered in details in \cref{sec:obesity_and_cancer}, the effect of obesity on cancers will be briefly mentioned here.
 Epidemiological evidence of obesity being a risk factor for many cancers has been well established, yet the mechanisms of its contribution to cancer is still  speculative \citep{Calle2003,Kelesidis2006a}.
 Although there are a lot of possible mechcanisms, there is growing evidence of the role of chronic inflammation caused by increased fat mass in the body \citep{Kelesidis2006a,Lumeng2011,Hernandez2013}.
-Free fatty acids from excess adipose tissue and/or diet can be recognised by the \gls{tlr}4, which can then activate \gls{nfkb} \citep{Lumeng2011}.
-Clearly, the result of the activation of \gls{nfkb} can lead to the activation of one of the Hallmarks of Cancers -- tumour-promoting inflammation.
+Free fatty acids from excess adipose tissue and/or diet can be recognised by \gls{tlr} (specifically \gls{tlr}4), which can then activate \gls{nfkb}, leading to the activation of one of the Hallmarks of Cancer -- tumour-promoting inflammation \citep{Lumeng2011}.
+% The nutritional content of the food that people consume is also important, as it can both directly and indirectly influence the likelihood of tumorigenesis, as well as its progression \citep{Ames1995a}.
+% Many fruits and vegetables contain antioxidants as well as essential viatmins and minerals.
+% Antioxidants are able to neutralise the effect of \gls{ros},  and many viatmins have been associated with reduction of many types of cancers \citep{Ames1995a}.
+% In fact, ascorbic acid (vitamin C) is able to prevent the lipid peroxidation of lipid plasma caused by tobacco smokes \textit{in vitro}, which could be why smokers, compared to non-smokers, have decreased levels of ascorbic acid \citep{Hecht1999}.
+% By maintaining healthy diet, it may be possible to reduce the risks of many cancers and obesity, as well as the comorbidities associated with them.
 
-The nutritional content of the food that people consume is also important, as it can both directly and indirectly influence the likelihood of tumorigenesis, as well as its progression \citep{Ames1995a}.
-Many fruits and vegetables contain antioxidants as well as essential viatmins and minerals.
-Antioxidants are able to neutralise the effect of \gls{ros},  and many viatmins have been associated with reduction of many types of cancers \citep{Ames1995a}.
-In fact, ascorbic acid (vitamin C) is able to prevent the lipid peroxidation of lipid plasma caused by tobacco smokes \textit{in vitro}, which could be why smokers, compared to non-smokers, have decreased levels of ascorbic acid \citep{Hecht1999}.
-
-By maintaining healthy diet, it may be possible to reduce the risks of many cancers and obesity, as well as the comorbidities associated with them.
-\\
-
-\noindent
-As one can imagine, \acrshort{dna} damages caused by smoking and/or \gls{uv} radiation have the potential to damage the genome and initiate cancer if wrong genes are mutated, such as p53 tumour suppressor gene.
-Healthy diet is also crucial for lowering the risk of cancers through beneficial vitamins and antioxidants, and prevent obesity as well as chronic inflammation that accompanies with it.
-There are many other environmental factors associated with cancers, and limiting the exposure to detrimental environmental factors is crucial for the prevention of any types of cancers, as well as the more specific types of cancers associated with these factors, such as lung and skin cancers.
-\\
+% As one can imagine, \acrshort{dna} damages caused by smoking and/or \gls{uv} radiation have the potential to damage the genome and initiate cancer if wrong genes are mutated, such as p53 tumour suppressor gene.
+% Healthy diet is also crucial for lowering the risk of cancers through beneficial vitamins and antioxidants, and prevent obesity as well as chronic inflammation that accompanies with it.
+There are many other environmental factors associated with cancers, and limiting the exposure to detrimental environmental factors is crucial for the prevention of any type of cancers, as well as the more specific type of cancers associated with these factors, such as lung and skin cancers.
 
@@
 
-The RAS-RAF-MEK-ERK signalling pathway is a pathway involved in cell survival, growth and proliferation \citep{Samatar2014,Wilhelm2006}.
-The members of this pathway are often mutated in many types of cancers (approximately a third of cancers contain a mutation in the RAS oncogene), causing an overactivation of this pathway \citep{Samatar2014}.
+The RAS-RAF-MEK-ERK signalling pathway is  involved in cell survival, growth and proliferation \citep{Samatar2014,Wilhelm2006}.
+The members of this pathway are often mutated in many types of cancer (approximately a third of cancers contain a mutation in the RAS oncogene), causing an overactivation of this pathway \citep{Samatar2014}.
 Development of drugs that target  the downstream RAF and its mutant forms (such as BRAF$^{V600E}$) have been successful, and many RAF inhibitor drugs such as sorafenib and vemurafenib have been approved for clinical use \citep{Samatar2014,Wilhelm2006}.
@@
 However, the effectiveness of targeting multiple pathways in humans is yet to be confirmed and further investigations are required to solidify the mode of action and its relevance to tumour biology.
-\\
 
-\noindent
-There are many difficulties accompanied by cancer treatment such as resistance to drugs and metastasis, and there is always a need for new ways to treat cancer patients.
 Further understanding of the tumour biology, its mechanism of resistance to drugs, and the effect of targeting multiple pathways (or targeting effective pathways for a patient) will definitely become important in the near future, both for novel drug development and better clinical treatment plan for individual patients.
@@
 
-As mentioned briefly in \cref{sub:causes_and_risk_factors_of_cancer}, obesity is considered a major risk factor for many types of cancers.
+As mentioned briefly in \cref{sub:causes_and_risk_factors_of_cancer}, obesity is considered a major risk factor for many types of cancer.
 The fact that the prevalence of obesity is extremely high, it is important to establish the link between the apparent risks associated with cancer.
@@
 Although the mechanisms in which obesity causes cancer is still under debate, the association between obesity and cancer has long been established.
-
 Perhaps the most convincing evidence of the association between obesity and cancer was first presented by \citet{Calle2003}.
@@
 
-In a more recent meta-analysis study by \citet{Renehan2008} showed that an increase \gls{bmi} by 5 kg/m$^2$ significantly increased the risks of esophageal, thyroid, colon and renal cancers in men; endometrial, gallbladder, esophageal and renal cancers in women.
+In a more recent meta-analysis study, \citet{Renehan2008} showed that an increase in \gls{bmi} by 5 kg/m$^2$ significantly increased the risks of esophageal, thyroid, colon and renal cancers in men, and endometrial, gallbladder, esophageal and renal cancers in women.
 The summary of the risks associated with various cancers from this study are shown in \cref{tab:renehan_cancer_risks}.
-This study showed that the association between \gls{bmi} and cancers were sex-specific, applicable to wide range of cancer types, and generally consistent across geographic populations \citep{Renehan2008,Roberts2010}.
+This meant that the association between \gls{bmi} and cancers were sex-specific, applicable to wide range of cancer types, and generally consistent across different geographic populations \citep{Renehan2008,Roberts2010}.
+It is undeniable that obesity is a major risk factor for many types of cancer and it may even be the cause for some of these cancers.
 
-Taken together, it is undeniable that obesity is a major risk factor for many types of cancers, and it may even be the cause for some of these cancers.
+% Taken together, it is undeniable that obesity is a major risk factor for many types of cancers, and it may even be the cause for some of these cancers.
 
-\begin{table}[h]
+\begin{table}[htb]
 	\centering
@@
 The exact mechanisms in which obesity causes cancer have still not been fully elucidated yet.
-There are several hypotheses that links obesity with cancers: disruption of various hormone levels, chronic and sustained inflammation, and activation of the insulin/\gls{igf} pathway.
+There are several hypotheses that link obesity with cancer: disruption of various hormone levels, chronic and sustained inflammation, and activation of the insulin/\gls{igf} pathway.
 
-Previously thought to be a group of cells where the body could store and manage excess fats, it has recently become apparent that adipose tissue is also a metabolically active organ that has a significant role in the endocrine system \citep{Roberts2010}.
+% Previously thought to be a group of cells where the body could store and manage excess fats, it has recently become apparent that adipose tissue is also a metabolically active organ that has a significant role in the endocrine system \citep{Roberts2010}.
+Adipose tissue has long been thought to be a group of cells to store excess fats, but it has recently become apparent that adipose tissue is also a metabolically active endocrine organ \citep{Roberts2010}.
 Increased adiposity is known to influence the bioavailability of endogenous sex hormones, including estrogens, androgens and progesterone \citep{Calle2004}.
@@
 Of the many adipokines identified so far, leptin and adipo\-nectin have been studied the most in terms of cancer development \citep{Renehan2006,Roberts2010}.
-Leptin is known to activate not only the satiety signalling pathway, but also the \gls{jak}/\gls{stat}, \gls{pi3k}/Akt signalling pathways \citep{Garofalo2006,Renehan2006}.
-Activation of these pathways lead to the activation of their target genes that are involved in cell proliferation and cell survival, as well as pro-angiogenic factors \citep{Garofalo2006}.
+Leptin is known to activate not only the satiety signalling pathway, but also the \gls{jak}/\gls{stat} and \gls{pi3k}/Akt signalling pathways \citep{Garofalo2006,Renehan2006}.
+Activation of these pathways lead to the activation of target genes that are involved in cell proliferation and cell survival, as well as pro-angiogenic factors \citep{Garofalo2006}.
 
@@
 Adiponectin has anti-angiogenic and anti-proliferative effect on the cells, and is secreted by mature adipocytes but not by the premature pre-adipocytes \citep{Gilbert2013}.
-During chronic inflammatory state in obese patients, pro-inflammatory proteins such as \gls{tgfb} and \gls{ifny} can block the maturation of the pre-adipocytes, thus decreasing the level of adiponectin and the accompanying benefecial effects \citep{Gilbert2013}.
+During chronic inflammatory state in obese patients, pro-inflammatory proteins such as \gls{tgfb} and \gls{ifny} can block the maturation of the pre-adipocytes, thus decreasing the level of adiponectin and the accompanying beneficial effects \citep{Gilbert2013}.
 
 Alternatively, or possibly in conjunction with hormone disruption, inflammation caused by obesity may be associated with tumour-promoting environment.
-\Gls{tlr}4 is a type of \gls{prr} and, as mentioned earlier, it is able to recognise circulating free fatty acids and trigger immune response via \gls{nfkb} pathway \citep{Lumeng2011}.
-In addition to the \gls{nfkb} pathway, recent evidence have shown that \gls{tlr}4 could also activate the \gls{jak}/\gls{stat3} pathway \citep{Yu2014}.
+\Gls{tlr}4 is a type of \gls{prr} and, as mentioned earlier, is able to recognise circulating free fatty acids and trigger immune response via \gls{nfkb} pathway \citep{Lumeng2011}.
+In addition to \gls{nfkb} pathway, recent evidence have shown that \gls{tlr}4 could also activate the \gls{jak}/\gls{stat3} pathway \citep{Yu2014}.
 Activation of these pathways will further enhance the development of the tumour-promoting microenvironment caused by these unwanted inflammatory responses of obesity.
@@
 Finally, insulin/\gls{igf} pathway may have a major role in encouraging tumorigenesis in obese patients.
-\Gls{igf} system involves two ligands (\gls{igf}-I and \gls{igf}-II), its respective receptors (\gls{igf}-IR and \gls{igf}-IIR), and six binding proteins (\gls{igfbp} 1 to 6) \citep{Roberts2010}.
+% \Gls{igf} system involves two ligands (\gls{igf}-I and \gls{igf}-II), its respective receptors (\gls{igf}-IR and \gls{igf}-IIR), and six binding proteins (\gls{igfbp} 1 to 6) \citep{Roberts2010}.
 \gls{igf} signalling results in the activation of \gls{pi3k}/Akt and \gls{erk} pathways, resulting in cell proliferation and reduced apoptosis \citep{Roberts2010}.
 Though insulin is also capable of activating these pathways, it is thought that the mitogenic effects are controlled mainly through the \gls{igf} receptors \citep{Roberts2010}.
-
 Since insulin reduces the level of \gls{igfbp}-3, the main circulating \gls{igfbp}, hyperinsulinemia has been hypothesised to promote tumorigenesis through the mitogenic effect of insulin and free \glspl{igf} \citep{Giovannucci1995,Mckeown1994,Roberts2010}.
@@
 As one can imagine, it is difficult for clinicians to treat a patient if they do not know what is causing the disease.
-The reason why it is so difficult to identify the underlying mechanisms is that there are very little reliable biological markers (or ``genetic signatures'') that signify the presence or absence of these pathways and/or hallmarks.
+The reason why it is so difficult to identify the underlying biological mechanisms is that there are very little reliable biological markers (or ``genetic signatures'') that signify the presence or absence of these pathways.
 With that said, significant improvements have been made in recent years owing to the advancement in sequencing technologies and have allowed researchers to study the genetic signatures that represent the hallmarks in a much greater detail.
@@
 Microarray technology was first developed by \citet{Schena1995} and has been used extensively to study gene expression patterns and \gls{snp} identification for over two decades.
-In short, microarray technology uses a glass slide with hundreds of thousands of \acrshort{dna} sequences attached to it, where fluorescently labelled \glspl{cdna} taken from the sample \glspl{mrna} are hybridised to these immobilised \acrshort{dna} sequences, and the fluorscent probes are detected \citep{Schena1995,Schulze2001}.
-Since only the exactly matching sequences remain on the microarray slide, researchers are able to identify and quantify the gene expression level of the sample \citep{Schulze2001}.
+In short, microarray technology uses a glass slide with hundreds of thousands of short \acrshort{dna} sequences attached to it, where fluorescently labelled \glspl{cdna} taken from the sample \glspl{mrna} are hybridised to these immobilised \acrshort{dna} sequences, and the fluorscent probes are detected to measure the gene expression \citep{Schena1995,Schulze2001}.
+% Since only the exactly matching sequences remain on the microarray slide, researchers are able to identify and quantify the gene expression level of the sample \citep{Schulze2001}.
 
 One disadvantage of microarray technology is that it requires previous knowledge of the genes and \glspl{snp} of the samples being investigated \citep{Hurd2009}.
-This is due to the fact that the detection of sequences relies on the hybridisation of the sample sequences with the microarray sequences and therefore the sequences under investigation must be known before they are invesetigated.
+This is due to the fact that the detection of sequences rely on the hybridisation of the sample sequences with the microarray sequences and therefore the sequences under investigation must be known to attach the correct sequences to the microarray.
 Another disadvantage of microarrays is the potentially erroneous detection of the sample sequences, both in terms of its sequence and its expression level \citep{Hurd2009}.
-Again, because the microarray slides require pre-defined sequences to be attached to it, there is always a possibility of similar sequences hybridising to the slide.
-Even if the sequences correctly hybridise to the slide, it may be difficult to measure the exact amount of some sequences in the samples.
-Since the amount of sequences in the sample is measured with the relative strength of the fluorescence of the dyes used, it is difficult to detect the exact level of expressions of sequences that are either very rare or highly abundant in the sample \citep{Hurd2009}.
+Again, because the microarray slides require pre-defined sequences to be attached to it, there is always a possibility of hybridisation of similar but different sequence to the slide.
+Even if the sequences correctly hybridise to the slide, it may be difficult to measure the exact amount of some sequences in the samples, since the amount of sequences in the sample is measured with the relative strength of the fluorescence of the dyes used \citep{Hurd2009}.
+Therefore, it is difficult to detect the exact level of expressions of sequences that are either very rare or highly abundant in the sample \citep{Hurd2009}.
 
@@
 
-\Gls{ngs} technology refers to the high throughput sequencing techniques that allow sequencing of large quantity of genetic material from a sample \citep{Metzker2010}.
+\Gls{ngs} technology refers to the high throughput sequencing techniques that allow sequencing of large quantity of genetic material \citep{Metzker2010}.
 \gls{ngs} technology has been developed and improved in the last decade at a revolutionary pace, making it possible for researchers to study gene expressions and variations in days to weeks, where the same experiment would have taken years to complete before the technology was developed.
@@
 First, \acrshort{dna} samples are fragmented into smaller sizes and universal primers are attached to the ends of the fragments.
-These fragments are then amplified into clusters to make the signals stronger for detection (although this amplification step varies between different techniques) \citep{Metzker2010}.
+% These fragments are then amplified into clusters to make the signals stronger for detection (although this amplification step vary between different techniques) \citep{Metzker2010}.
+These fragments are then amplified into clusters to make the signals stronger for detection \citep{Metzker2010}.
 After the amplification step, the clusters of \acrshort{dna} fragments are sequenced and ``imaged'' using fluorescently labelled \acrshort{dna} bases, where the incorporation of different bases are detected by its unique fluorescence \citep{Metzker2010}.
@@
 \gls{ngs} technology have many advantages over microarray technology.
-Firstly, \gls{ngs} does not require any prior knowledge of the sample like in the microarray technology as all genetic materials are attached to universal primers when they are sequenced -- there is no need to pre-define the sequences to be searched and thus allow the detection of novel genes or transcripts \citep{Hurd2009}.
-This also prevents the introduction of experimental bias, as the sequences are not altered in any way and no specific sequence is searched over the others \citep{Hurd2009}.
-Another advantage is that \gls{ngs} produce count data of the genetic material being sequenced, instead of relying on the intensity of the fluorescence to detect the presence and amount of certan sequence.
+Firstly, \gls{ngs} does not require any prior knowledge of the sample like in the microarray technology, as all genetic materials are attached to universal primers when they are sequenced; there is no need to pre-define the sequences to be searched and thus allow the detection of novel genes or transcripts \citep{Hurd2009}.
+This also prevents the introduction of experimental bias, as the sequences are not altered in any way and no specific sequence is searched in a greater detail over another \citep{Hurd2009}.
+Another advantage is that \gls{ngs} produce count data of the genetic material being sequenced instead of relying on the intensity of the fluorescence to detect the presence and amount of certan sequence.
 This allows an absolute quantification of the amount of transcripts from the sample, making it possible for a more accurate comparison of the expression levels between samples.
@@
 
-In this study, \citet{Creighton2012} used 103 breast tumour samples and analysed these samples using microarrays.
+\citet{Creighton2012} used 103 breast tumour samples and analysed these samples using microarrays.
 Of the 103 samples, 35\% were from normal weight, 28\% were from overweight and 37\% were from obese patients.
@@
 
-In the study conducted by \citet{Fuentes-Mattei2014}, they obtained 137 breast tumour samples from \gls{er} positive patients and analysed their transcriptome using microarrays.
-Of the 137 patients, 94 patients were not obese and 43 patients were obese.
+\citet{Fuentes-Mattei2014} studied 137 breast tumour samples from \gls{er} positive patients and analysed their transcriptome using microarrays.
+Of the 137 patients, 94 patients were not obese and 43 patients were.
 Comparison of the gene expression of the tumour samples from obese patients with the samples from non-obese patients revealed 112 genes that were statistically significantly dysregulated (62 genes upregulated and 50 genes downregulated).
@@
 On top of this, \citet{Fuentes-Mattei2014} validated their results using a mouse model, where they compared the expression level in the breast tumours from obese mice with the samples from lean mice.
-Results from this animal model experiment provided strong evidence of obesity significantly accelerating tumour growth through Akt/\gls{mtor} signalling pathway.
+Results from this animal model provided strong evidence that obesity significantly accelerates tumour growth through Akt/\gls{mtor} signalling pathway.
 
@@
 Like obesity associated genetic signatures, there has also been a focus on the identification of pathways and genes that affect tumour growth.
-About a decade ago, \citet{Bild2006} demonstrated that a collection of genes that are involved in a specific pathway could be used to determine the status of those pathways in specific samples.
+About a decade ago, \citet{Bild2006} demonstrated that a collection of genes that are involved in a specific pathway could be used to determine the status of those pathways in a sample.
 These pathway specific genes were identified experimentally by altering the expression of one specific gene that is central to the pathway in a model cell culture system.
-For example, all of the genes that have altered its expression in the cells that have \textit{Myc} gene induced were noted down as the gene involved in the Myc pathway.
+For example, all of the genes that have altered its expression in the cells that have \textit{Myc} gene induced were noted down as the genes involved in the Myc pathway.
 
@@
 
-In this study, \citet{Gatza2010a} used 18 pathway signatures to classify human breast tumours into more detailed subtypes.
+\citet{Gatza2010a} used 18 pathway signatures to classify human breast tumours into more detailed subtypes.
 The 18 pathways used in this study were: Akt, \gls{bcat}, E2F1, \gls{egfr}, \gls{er}, \gls{her2}, \gls{ifna}, \gls{ifny}, Myc, p53, p63, \gls{pi3k}, \gls{pr}, Ras, \gls{stat3}, Src, \gls{tgfb}, and \gls{tnfa} pathways.
 This was a significant improvement from their previous works on pathway signatures \citep{Bild2006,Bild2009}.
-Using more pathways in their analysis have allowed them to identify certain pathways to coactivate (or cluster) together, suggesting that some pathways are related to one another and may have relevance in identifying the causes.
-% TODO: list 18 pathways.
+Using more pathways in their analysis have allowed them to identify certain pathways to be coactivated (or cluster) together, suggesting that some pathways are related to one another and may have relevance in identifying the causes.
 
@@
 In the following chapter, the detailed methods of all the analyses carried out for this project is outlined.
-The results from these analyses are split into two parts, each presented in separate chapters: first part of the results clarify the relationship between the obesity associated genetic signatures with various cancer types; and the second part of the results consist of the links between the pathway associated genetic signatures and various cancer types.
-By the end of this thesis, the aim presented here will be answered with clarity.
-However, it is without doubt that there will be many more questions to be answered than what was actually answered -- those questions shall remain unanswered for future investigations.
+The results from these analyses are split into two parts, each presented in separate chapters: first part of the results clarify the relationship between the obesity associated genetic signatures with various cancer types; and the second part of the results consist of the links between the pathway associated genetic signatures and various cancer types, as well as the previously identified obesity associated signatures.
+\\
+
+\noindent
+(By the end of this thesis, the aim presented here will be answered with clarity.
+However, it is without doubt that there will be many more questions to be answered than what was actually answered -- those questions shall remain unanswered for future investigations.)
+include this bit or not \ldots
 

### 2017-01-23 thesis/intro/intro.tex
Minor edit in intro and methods

@@
 
+One thing to note here is that most of these pathway associated genetic signatures are involved in one or more of the Hallmarks of Cancer.
+This means that these pathway associated signatures can be used to provide insights into the underlying biological mechanisms of the samples used in this project.
+
 \section{Aim of the project}
@@
 
+\section{Thesis outline}
+\label{sec:thesis_outline}
+
 In the following chapter, the detailed methods of all the analyses carried out for this project is outlined.
 The results from these analyses are split into two parts, each presented in separate chapters: first part of the results clarify the relationship between the obesity associated genetic signatures with various cancer types; and the second part of the results consist of the links between the pathway associated genetic signatures and various cancer types, as well as the previously identified obesity associated signatures.
-\\
-
-\noindent
-(By the end of this thesis, the aim presented here will be answered with clarity.
-However, it is without doubt that there will be many more questions to be answered than what was actually answered -- those questions shall remain unanswered for future investigations.)
-include this bit or not \ldots
+By the end of this thesis, the aim presented here will be answered with clarity.
 

### 2017-01-23 thesis/methods/methods.tex
Started editing introduction -- tmp commit
Minor edit in intro and methods

@@
 
+\subsection{Box and scatter plots}
+\label{sub:box_and_scatter_plots}
+
+Box plots and scatter plots were used to visualise the association of a metagene with some of the clinical variables of the data.
+Box plots and scatter plots were plotted using the \texttt{boxplot} and \texttt{plot} functions from the \textit{graphics} package in \gls{r}, respectively.
+The p-value and \gls{anova} p-value in the box plots were calculated using the \texttt{t.test} and \texttt{aov} functions from the \textit{stats} package in \gls{r}, respectively.
+The line of best fit and the $R^2$-value of the line were taken from the summary statistics produced by the \texttt{lm} function from the \textit{stats} package in \gls{r}.
+
 \subsection{Heatmaps}
@@
 
-Heatmaps were created using either the \texttt{heatmap.2} function from \textit{gplots} package, or the \texttt{heatmap.2x} function written by Tom Kelly.
-All of the heatmaps that required no or one column bar used the \texttt{heatmap.2} function, and all of the heatmaps that required more than one column bars used the \texttt{heatmap.2x} function.
+To visualise the association of the sample metagene scores with the sample gene expression, heatmaps were used.
+Heatmap is an effective method to visualise such data, as the ordering of the genes and/or samples may be controlled based on certain values, for example metagene scores.
+Furthermore, heatmaps are able to cluster the samples and/or genes based on the ``distance'' between the data, or in other words, how closely related the data are.
+Heatmaps were created using the \texttt{heatmap.2} function from the \textit{gplots} package in \gls{r}.

### 2017-01-23 thesis/results/results.tex
tmp commit

--- 2016-09-19 thesis/results/results.tex
@@
-\chapter{Obesity-associated genetic signatures and cancer}
+\chapter{Obesity associated genetic signatures and cancer}
 \label{cha:obesity_genetic_signatures_and_cancer}
 
-\section{Obesity-associated genetic signature from \citet{Creighton2012} study}
+In this chapter, the relationship between the obesity associated genetic signatures from previous studies and different cancer types were investigated; new obesity associated genetic signatures were identified in the data set from \citet{Creighton2012} study; and the presence of common genes or pathways across different cancer types was examined.
+
+\section{Obesity associated genetic signature from \citet{Creighton2012} study}
 \label{sec:creighton_obesity_metagene}
 
-\section{Obesity-associated genetic signature from \citet{Fuentes-Mattei2014} study}
+% \begin{figure}[h!]
+% 	\centering
+% 	\includegraphics[width=0.8\linewidth]{intro/hallmarks}
+% 	\caption[Hallmarks of Cancer]{Hallmarks of Cancer. The ten Hallmarks of Cancer as proposed by Hanahan and Weinberg in 2011. (Figure adapted from \citet{Hanahan2011})}
+% 	\label{fig:hallmarks}
+% \end{figure}
+
+\section{Obesity associated genetic signature from \citet{Fuentes-Mattei2014} study}
 \label{sec:fm_obesity_metagene}
+
+\section{New obesity associated genetic signature in \citet{Creighton2012} data set}
+\label{sec:creighton_obesity_metagene_new}
 
@@
 
-\section{Pathways enriched in obesity-associated genes}
+\section{Pathways enriched in obesity associated genes}
 \label{sec:pathways_enriched_in_obesity_associated_genes}

### 2017-01-23 thesis/results/results2.tex
Starting the write up of the first part of the results
tmp commit

@@
-\chapter{Results from gatza pathway metagene and direction stuff}
+\chapter{Obesity associated genetic signatures and pathway signatures}
+\label{cha:obesity_associated_genetic_signature_and_pathway_signatures}
 
-\section{Reproducing results from gatza paper}
+In this chapter, the relationship between the obesity-associated genetic signatures from previous studies and different cancer types were investigated; new obesity-associated genetic signatures were identified in the data set from \citet{Creighton2012} study; and the presence of common genes or pathways across different cancer types was examined.

### 2017-01-28 thesis/discussion/discussion.tex
Finished Creighton obesity metagene section draft

\chapter{Discussion (draft)}

\section{something}

% List of things to discuss

% RESULTS 1

Lack of association of obesity metagenes with different cancer data sets: due to metagene derivation from a specific data

\begin{figure}[ht]
\caption{Dummy}
\label{fig:dum1}
\end{figure}

\begin{table}[ht]
\caption{Dummy}
\begin{tabular}{|c|c|}
\end{tabular}
\label{tab:dum1}
\end{table}

### 2017-01-28 thesis/methods/methods.tex
Finished Creighton obesity metagene section draft

@@
 
-In order to ertain set of genes are up or down regulated by a sample, some sort of score based on these genes must be calculated.
+In order to determine whether a set of genes are up or down regulated by a sample, some sort of score based on these genes must be calculated.
 \Gls{svd} is a mathematical method that splits a matrix into several matrices that represent the original matrix \citep{Golub1970}.
@@
 With respect to this project, \gls{svd} was used to assess whether the metagene created from various obesity or pathway associated genes were significantly associated with certain clinical variables, such as \gls{bmi}.
-To do this, the normalised original expression data was reduced in size so that it only included the genes of ineterest (either obesity or pathway associated) and the samples, then \texttt{svd} function from the \textit{base} package in \gls{r} was used to apply \gls{svd} to the matrix.
+To do this, the normalised original expression data was reduced in size so that it only included the genes of interest (either obesity or pathway associated) and the samples, then \texttt{svd} function from the \textit{base} package in \gls{r} was used to apply \gls{svd} to the matrix.
 The first principal component was taken from the $V$ matrix and was used as the metagene scores for the samples in the data.

### 2017-01-28 thesis/results/results.tex
Finished Creighton obesity metagene section draft

@@
 
-In this chapter, the relationship between the obesity associated genetic signatures from previous studies and different cancer types were investigated; new obesity associated genetic signatures were identified in the data set from \citet{Creighton2012} study; and the presence of common genes or pathways across different cancer types was examined.
+The obesity associated genetic signatures are central to this project in order to clarify the relationship between \gls{bmi} and cancer.
+Firstly in this chapter, the previously identified obesity associated signatures from the studies conducted by \citet{Creighton2012} and \citet{Fuentes-Mattei2014} were examined in turn to judge the agreement of these signatures with the sample \gls{bmi}, as presented in thier results.
+Secondly, novel obesity associated genetic signatures were identified in the \citet{Creighton2012} data set and was compared with the signatures.
+Lastly, the presence of common genes or pathways that were associated with obesity in multiple types of cancer was explored using the \gls{icgc} cancer data sets.
 
@@
 
-% \begin{figure}[h!]
-% 	\centering
-% 	\includegraphics[width=0.8\linewidth]{intro/hallmarks}
-% 	\caption[Hallmarks of Cancer]{Hallmarks of Cancer. The ten Hallmarks of Cancer as proposed by Hanahan and Weinberg in 2011. (Figure adapted from \citet{Hanahan2011})}
-% 	\label{fig:hallmarks}
-% \end{figure}
+Obesity metagene was created using the obesity associated genetic signatures from \citet{Creighton2012} study.
+Since there was no description about the normalisation method used by Creighton's group, the \gls{rma} method was used to  normalise the Creighton data set.
+The obesity metagene was plotted in a heatmap with the sample gene expression (\cref{fig:crmetaheat1}) to check whether the metagene scores were in accordance with the overall gene expression of the samples.
+
+\begin{figure}[htb]
+	\centering
+	\includegraphics[width=0.8\linewidth]{results1/crmeta_original}
+	\caption[Heatmap of Creighton's metagene in Creighton's data]{Heatmap showing Creighton's obesity associated metagene with the gene expressions of obesity associated genes in Creighton's data.
+Level of expression is represented in the top right histogram, where low and high gene expression were colour-coded with blue and red, respectively.
+Each row of the heatmap represents a gene from the obesity associated genetic signature, and each column of the heatmap represents a sample from Creighton's data.
+The obesity associated metagene scores of the samples are shown in a separate row at top of the heatmap, and the tree diagram of the heirarchical clustering of the genes is shown to the left of the heatmap.
+For clarity, the metagene scores were flipped in order to match the results from \citet{Creighton2012} paper.}
+	\label{fig:crmetaheat1}
+\end{figure}
+
+As shown in \cref{fig:crmetaheat1}, high obesity associated metagene score of the sample reflected low expression in majority of the genes in the signature, and in contrast, low obesity associated metagene score of the sample reflected high expression in majority of the genes in the signature.
+This was consistent with the reported property of the obesity associated signature by \citet{Creighton2012}.
+To provide further evidence that the obesity associated metagenes were in fact associated with the \gls{bmi} status and \gls{bmi} value of the samples, a box plot and a scatter plot were created, respectively (\cref{fig:crmetaboxplot1}).
+
+\begin{figure}[htp!]
+	\centering
+	\includegraphics[width=0.45\linewidth]{results1/crmeta_original}
+	\hfill
+	\includegraphics[width=0.45\linewidth]{results1/crmeta_original}
+	\caption[Creighton metagene and sample \gls{bmi}/\gls{bmi} status]{(insert figure of box plot and scatter plot)}
+	\label{fig:crmetaboxplot1}
+\end{figure}
+
+\cref{fig:crmetaboxplot1} clearly showed that the obesity associated metagene from \citet{Creighton2012} study significantly associated with the sample \gls{bmi} status, as well as the \gls{bmi} value of the samples (include p-value and r-squared).
+It should be noted here that the obesity associated metagene significantly associated with the samples that were obese, but not the samples that were overweight.
+This was due to the fact that the obesity associated genetic signatures were originally identified from the comparison of the samples that were obese with the samples that were not obese, and therefore the metagene scores were significant with obese group, but not the overweight group.
+\\
+
+\noindent
+Now that the association of the obesity metagene from \citet{Creighton2012} study was established in Creighton's data set, the obesity associated signature was transferred to the \gls{icgc} cancer data.
+The direction of the obesity associated metagene was checked in the Creighton's data, so that high metagene scores reflected high sample \gls{bmi} and low metagene scores reflected low sample \gls{bmi} (\cref{fig:crmetaboxplot1}).
+The transformation matrix was created in Creighton's data, as described in \cref{sub:svd}.
+
+All of the \gls{icgc} data were normalised as described in \cref{ssub:rna_seq_data}.
+Before the transformation matrix was applied to the log$_{10}$-normalised cancer data, the suitability of standardised data or untouched (non-standardised) data was determined (appendix).
+From these results, the standardised data was the most suitable to use for matrix transformation.
+The transformation matrix was applied to each cancer data set in turn to obtain the obesity metagenes for each of the data set.
+Each obesity metagenes were plotted on a heatmap with the corresponding data set in which the metagene was taken from \cref{fig:crmetaicgc1}.
+These heatmaps confirmed that the obesity associated metagene derived from Creighton's data was able to capture the overall gene expression pattern, where the metagene scores reflected the expression levels of majority of the genes in the signature.
+
+\begin{figure}[htp!]
+	\centering
+	\includegraphics[width=0.45\linewidth]{results1/crmeta_original}
+	\hfill
+	\includegraphics[width=0.45\linewidth]{results1/crmeta_original}
+	\caption[Creighton metagene and \acrshort{icgc} cancer data]{(insert ICGC heatmaps)}
+	\label{fig:crmetaicgc1}
+\end{figure}
+
+\begin{figure}[htp!]
+	\centering
+	\includegraphics[width=0.45\linewidth]{results1/crmeta_original}
+	\hfill
+	\includegraphics[width=0.45\linewidth]{results1/crmeta_original}
+	\caption[Creighton metagene and sample \gls{bmi}/\gls{bmi} status in \acrshort{icgc} data]{(insert figure of box plot and scatter plot for ICGC data)}
+	\label{fig:crmetaboxicgc1}
+\end{figure}
+
+As before, association of the obesity metagene with the sample \gls{bmi} and \gls{bmi} status was checked in their respective cancer data set (\cref{fig:crmetaboxicgc1}).
+In contrast to the strong association of the obesity metagene with sample \gls{bmi} and \gls{bmi} status in Creighton's data, the obesity metagene taken from the \gls{icgc} cancer data did not significantly associate with the sample \gls{bmi} or \gls{bmi} status in any of the \gls{icgc} cancer data.
+
+There could be several reasons for the apparent lack of association of the metagene with the sample \gls{bmi} or \gls{bmi} status.
+First, transformation matrix was derived from Creighton's microarray data, but the \gls{icgc} cancer data were \gls{rnaseq} data.
+Though the log$_{10}$ normalisation and standardisation of the data was the most appropriate adjustment to be made to the \gls{rnaseq} data, this adjustment may not be equivalent to the \gls{rma} normalisation method that was used on the microarray data.
+Secondly, none of the \gls{icgc} cancer was from breast as in Creighton's data.
+Since the obesity associated signature was identified in breast cancer data, the signature may be specific to breast cancer and may not be suitable in other cancer types.
+\\
+
+\noindent
+To check whether the metagene was specific to breast cancer microarray data, the same transformation matrix was applied to Print's breast cancer microarray data.
+Print's data was normalised with \gls{rma} method as described in \cref{ssub:microarray_data}, and the transformation matrix was applied to the normalised data to obtain the metagene.
+The metagene was again compared with the gene expression of the samples with a heatmap (\cref{fig:crmetaprint1}), and the association of the metagene with the sample \gls{bmi} and \gls{bmi} status was examined with box and scatter plots (\cref{fig:crmetaboxprint1}).
+
+\begin{figure}[htp!]
+	\centering
+	\includegraphics[width=0.45\linewidth]{results1/crmeta_original}
+	\hfill
+	\includegraphics[width=0.45\linewidth]{results1/crmeta_original}
+	\caption[Creighton metagene and Print's cancer data]{(insert Print heatmaps)}
+	\label{fig:crmetaprint1}
+\end{figure}
+
+\begin{figure}[htp!]
+	\centering
+	\includegraphics[width=0.45\linewidth]{results1/crmeta_original}
+	\hfill
+	\includegraphics[width=0.45\linewidth]{results1/crmeta_original}
+	\caption[Creighton metagene and sample \gls{bmi}/\gls{bmi} status in Print's data]{(insert figure of box plot and scatter plot for Print data)}
+	\label{fig:crmetaboxprint1}
+\end{figure}
+
+The obesity metagene managed to reflect the overall gene expression of the samples in Print's data.
+However, like in \gls{icgc} cancer data, the obesity metagene scores did not significantly associate with the sample \gls{bmi} or \gls{bmi} status.
+These results confirmed that the lack of association of the obesity metagene from Creighton's data was not due to the technology in which the data was gathered (mcroarray or \gls{rnaseq}), nor the type of cancer in which the transformation matrix was applied to.
+
+All together, all of these results suggest that the obesity metagene identified in the study conducted by \citet{Creighton2012} associated with sample \gls{bmi} and \gls{bmi} status in Creighton's data, but did not transfer well in other cancer data sets.
+The lack of association with \gls{bmi} was not due to the type of technology platform in which the data was gathered, as neither the \gls{icgc} \gls{rnaseq} data nor Print's microarray data did not show significant association with the obesity metagene.
+Furthermore, Creighton's metagene was not dependent on the cancer type since the obesity associated genetic signature did not show significant association in neither the \gls{icgc} cancer data nor in Print's breast cancer data.
+
+One possible reason why Creighton's obesity metagene did not show significant association with other data sets could be because the genetic signature was in fact not an obesity specific signature, but a signature that detected another clinical variable.
+Another reason for this apparent lack of association could be that the genetic signature was too specific to the data and was not a broad obesity associated genetic signature, but an obesity associated signature specifically for Creighton's data set.
 
@@
 
+It was surprising that Creighton's metagene did not show significant association with \gls{bmi} in other data sets.
+
 \section{New obesity associated genetic signature in \citet{Creighton2012} data set}
 \label{sec:creighton_obesity_metagene_new}
+
+Both the metagenes generated from Creighton's data and Fuentes-Mattei's data were not associated with sample \gls{bmi}.
 
@@
 
+It was clear that the metagenes created from obesity signatures were not consistent across different data sets.
+To identify genes that were associated with sample \gls{bmi} in multiple cancer types, \glspl{deg} were searched (?).
+
 \section{Pathways enriched in obesity associated genes}

### 2017-01-28 thesis/thesis.tex
tmp commit
Finished Obesity and cancer section draft
Minor edit to the table of contents and glossary style
Finished Creighton obesity metagene section draft

@@
-% \documentclass[a4paper, 12pt, oneside]{report} % for final draft compilation
-\documentclass[a4paper, 12pt, oneside, draft]{report} % for draft compilation
+\documentclass[a4paper, 12pt, oneside]{report} % for final draft compilation
+% \documentclass[a4paper, 12pt, oneside, draft]{report} % for draft compilation
 
@@
 \usepackage{chngcntr}                                  % package for making continuous figure numbering, instead of section-based numbering
+\usepackage[flushleft]{threeparttable}                 % package to add notes at the bottom of the table
 \usepackage{setspace}                                  % package for setting line space (used in title page)
@@
 \usepackage[md]{titlesec}                              % package for tweaking the chapter/section/subsection titles format
-\usepackage[acronym, toc]{glossaries}               % package for making glossaries
+\usepackage[acronym]{glossaries} % package for making glossaries
 \usepackage{xcolor,colortbl}                           % package for making colours
@@
 \usepackage[font=footnotesize,format=hang]{caption}    % package for formatting figure caption
+\usepackage{subcaption}    % package for formatting figure caption
 
@@
 
+% Change the title of Table of Contents:
+\renewcommand{\contentsname}{Table of Contents}
+
+% Change the title of Table of Contents:
+\renewcommand{\thesubfigure}{\Alph{subfigure}}
+
 % Italicise the subsubsection title:
@@
 % Specify graphics path:
-\graphicspath{ {./images/}}
+\graphicspath{{./images/}}
 \DeclareGraphicsExtensions{.pdf,.png,.jpg}
@@
 % If you want to compile specific files, uncomment the line below and add the filenames in the {} :
-%\includeonly{}
+\includeonly{./results/results}
 
@@
 
+\addcontentsline{toc}{section}{\contentsname}
+
 \tableofcontents
-
-\addcontentsline{toc}{section}{\contentsname}
 
@@
 
-\printglossaries
+\addcontentsline{toc}{section}{Glossary}
+
+\printglossary[style=altlist]
 
@@
 
-% \listoftables
+\addcontentsline{toc}{section}{Acronyms}
 
-% \clearpage
+\printglossary[type=\acronymtype,style=super]
+
+\clearpage
 
@@
 
+\addcontentsline{toc}{chapter}{\bibname}
+
 % Print out the bibliography:
-\bibliography{MSc}
+\bibliography{./MSc,./MSc2}
 

### 2017-01-28 thesis/results/results.tex
tmp commit

@@
 
-It was surprising that Creighton's metagene did not show significant association with \gls{bmi} in other data sets.
+Since Creighton's obesity associated genetic signature was not associated with sample \gls{bmi} or \gls{bmi} status in other data sets, Fuentes-Mattei's (FM) obesity associated genetic signature was examined to see whether FM's obesity metagene associated with sample \gls{bmi} or \gls{bmi} status.
+Since FM's data set did not have sample \gls{bmi} information, FM's obesity metagene was not able to be compared with the sample \gls{bmi} or \gls{bmi} status in the original FM's data.
+However, the transformation matrix was still applied to other cancer data to see whether FM's obesity metagene associated with sample \gls{bmi}/\gls{bmi} status
+
+Firstly, FM's data was normalised with the \gls{rma} method and \gls{svd} was applied to the normalised data to get the transformation matrix.
+The transformation matrix was used to transform the \gls{rma}-normalised Creighton's data to extract FM's obesity metagene scores.
+FM's obesity metagene scores were compared with gene expressions and sample \gls{bmi}/\gls{bmi} status in Creighton's data, as shown in \cref{fig:fmmetacr1,fig:fmmetaboxcr1}.
+Clearly, like with Creighton's obesity metagene, FM's obesity metagene was reflective of the overall gene expression of the samples, but did not associate with the sample \gls{bmi} or \gls{bmi} status.
+
+\begin{figure}[htp!]
+	\centering
+	\includegraphics[width=0.45\linewidth]{results1/crmeta_original}
+	\hfill
+	\includegraphics[width=0.45\linewidth]{results1/crmeta_original}
+	\caption[FM metagene and Creighton's cancer data]{(insert Creighton heatmaps)}
+	\label{fig:fmmetacr1}
+\end{figure}
+
+\begin{figure}[htp!]
+	\centering
+	\includegraphics[width=0.45\linewidth]{results1/crmeta_original}
+	\hfill
+	\includegraphics[width=0.45\linewidth]{results1/crmeta_original}
+	\caption[FM metagene and sample \gls{bmi}/\gls{bmi} status in Creighton's data]{(insert figure of box plot and scatter plot for Creighton data)}
+	\label{fig:fmmetaboxcr1}
+\end{figure}
+
+Next, the transformation matrix was applied to the \gls{icgc} cancer data and resulting metagenes were compared with the gene expression and sample \gls{bmi}/\gls{bmi} status.
+As evident in \cref{fig:fmmetaicgc1}, FM's obesity metagene scores appeared to reflect the gene expression of FM's obesity associated genetic signature.
+However, as with all the results in this section so far, FM's obesity metagene did not significantly associate with any of the \gls{icgc} cancer data (\cref{fig:fmmetaboxicgc1}).
+
+\begin{figure}[htp!]
+	\centering
+	\includegraphics[width=0.45\linewidth]{results1/crmeta_original}
+	\hfill
+	\includegraphics[width=0.45\linewidth]{results1/crmeta_original}
+	\caption[FM metagene and \gls{icgc} cancer data]{(insert ICGC heatmaps)}
+	\label{fig:fmmetaicgc1}
+\end{figure}
+
+\begin{figure}[htp!]
+	\centering
+	\includegraphics[width=0.45\linewidth]{results1/crmeta_original}
+	\hfill
+	\includegraphics[width=0.45\linewidth]{results1/crmeta_original}
+	\caption[FM metagene and sample \gls{bmi}/\gls{bmi} status in \gls{icgc} data]{(insert figure of box plot and scatter plot for Creighton data)}
+	\label{fig:fmmetaboxicgc1}
+\end{figure}
+
+Lastly, the transformation matrix was applied to Print's data.
+
+\\
+
+\begin{figure}[htp!]
+	\centering
+	\includegraphics[width=0.45\linewidth]{results1/crmeta_original}
+	\hfill
+	\includegraphics[width=0.45\linewidth]{results1/crmeta_original}
+	\caption[FM metagene and Print's cancer data]{(insert Print heatmaps)}
+	\label{fig:fmmetacris1}
+\end{figure}
+
+\begin{figure}[htp!]
+	\centering
+	\includegraphics[width=0.45\linewidth]{results1/crmeta_original}
+	\hfill
+	\includegraphics[width=0.45\linewidth]{results1/crmeta_original}
+	\caption[FM metagene and sample \gls{bmi}/\gls{bmi} status in Print's data]{(insert figure of box plot and scatter plot for Print's data)}
+	\label{fig:fmmetaboxcris1}
+\end{figure}
+
+\noindent
 

### 2017-01-28 thesis/results/results.tex
Finished FM metagene section (but very rough)

@@
 
-Lastly, the transformation matrix was applied to Print's data.
-
+Lastly, the transformation matrix was applied to Print's data (\cref{fig:fmmetaboxcris1,fig:fmmetacris1}).
+Again, FM's obesity metagene scores reflected the gene expression of the samples, did not significantly associate with sample \gls{bmi}/\gls{bmi} status.
 \\
@@
 \noindent
+These results showed that FM's obesity associated metagene was not transferable to other cancer data sets, similar to Creighton's obesity associated metagene.
+This meant that neither Creighton's nor FM's obesity associate metagenes may be too specific to the orignial data set in which the signature was identified in.
+Furthermore, it may be possible that these obesity associated metagenes may not be related to obesity, but another clinical variable which may be closely related to \gls{bmi}.
 

### 2017-01-28 thesis/results/results.tex
tmp commit

@@
 
-Both the metagenes generated from Creighton's data and Fuentes-Mattei's data were not associated with sample \gls{bmi}.
+Both metagenes generated from Creighton's data and FM's data were able to capture the overall pattern of gene expression of the samples, but did not associate with sample \gls{bmi}/\gls{bmi} status.
+One possible reason for this result could be that the obesity associated genetic signatures from Creighton's and FM's study may not have been truly associated with sample \gls{bmi} status, but another clinical variable.
+To remove this possibility, obesity associated genetic signature was identified in Creighton's data after controlling for all the clinical variables in the data set.
+FM's data set was not used to get obesity associated genetic signature, as no \gls{bmi} information was available for the samples in FM's data.
 

### 2017-01-30 thesis/results/results.tex
tmp commit

@@
 Obesity metagene was created using the obesity associated genetic signatures from \citet{Creighton2012} study.
-Since there was no description about the normalisation method used by Creighton's group, the \gls{rma} method was used to  normalise the Creighton data set.
+% As the genetic signatures were described as gene probe IDs, the gene probes had to be converted into gene symbols (see \cref{sec:data}).
+% This resulted in 644 unique gene symbols, compared to 799 gene probes IDs in the original signature.
+There was no description about the normalisation method used by Creighton's group when they first analysed their data, so the more reliable \gls{rma} method was used to  normalise the Creighton data set.
 The obesity metagene was plotted in a heatmap with the sample gene expression (\cref{fig:crmetaheat1}) to check whether the metagene scores were in accordance with the overall gene expression of the samples.
@@
 
-\section{New obesity associated genetic signature in \citet{Creighton2012} data set}
+\section{Novel obesity associated genetic signatures from \citet{Creighton2012} data set}
 \label{sec:creighton_obesity_metagene_new}
+
+\subsection{Identification of obesity associated genetic signatures}
+\label{sub:identification_of_obesity_associated_genetic_signatures}
 
@@
 One possible reason for this result could be that the obesity associated genetic signatures from Creighton's and FM's study may not have been truly associated with sample \gls{bmi} status, but another clinical variable.
-To remove this possibility, obesity associated genetic signature was identified in Creighton's data after controlling for all the clinical variables in the data set.
-FM's data set was not used to get obesity associated genetic signature, as no \gls{bmi} information was available for the samples in FM's data.
+To remove this possibility, obesity associated genetic signatures were identified in Creighton's data after controlling for all the clinical variables in the data set.
+FM's data set was not used to get obesity associated genetic signatures, as no \gls{bmi} information was available for the samples in FM's data.
+
+Firstly, to validate whether the identified obesity associated genetic signature was similar to the signature that Creighton's groups had originally found, the identified \glspl{deg} were compared with the set of genes identified by Creighton's group.
+\glspl{deg}  were identified between the samples from obese patient and non-obese patient in the \gls{rma}-normalised Creighton's data, as described in \cref{sec:gene_expression_analysis}.
+Without adjusting the p-value with FDR, 5278 gene probes and 1781 gene probes were significant at p \textless{} 0.05 and p \textless{} 0.01, respectively.
+After afjustment, there were only 9 gene probes significant at p \textless{} 0.05 and no gene was significant at p \textless{} 0.01.
+Furthermore, there were only 61 gene probes that were significantly differentially expressed at p \textless{} 0.05 with a log$_2$ \gls{fc} greater than 1.2.
+From these observations, the log$_2$ fold change of the gene probes were ignored and the threshold p-value was set to 0.01 (without adjustment) for the identification of gene probes.
+This was to include as many gene probes as possible without losing power (?).
+Additionally, when there were more than 799 gene probes identified, only the most significant 799 gene probes were taken as the obesity associated genetic signature, as this many genes were originally identified by Creighton's group.
+These criteria were applied for the identification of other genetic signatures as well.
+
+The above analysis was repeated with the residual data (\gls{rma}-normalised Creighton's data that had been controlled for other clinical variables; see \cref{sub:residual_data_creation}).
+The clinical variables controlled were age, ethnicity, menopause status, tumour grade, hormone (\gls{er}, \gls{pr}, and \gls{her2}) statuses  and \gls{ln} status.
+In the residual data, 1104 gene probes were significant with unadjusted p-value (p \textless{} 0.01).
+Again, the most significant 799 genes were taken as the obesity associated genetic signature from this data.
+
+In addition to the above two obesity associated genetic signatures, two more sets of genetic signatures were identified by taking the common genes between the above two signatures with Creighton's original obesity associated genetic signatures.
+There were 239 common gene probes between Creighton's original genetic signature and the gene probes identified from the unadjusted Creighton's data, and 168 common gene probes between the original signature and gene probes from residual data in Creighton's data.
+The genetic signatures identified were summarised as a Venn diagram, as shown in \cref{fig:venn1}.
+\\
+
+\begin{figure}[htp!]
+	\centering
+	\includegraphics[width=0.45\linewidth]{results1/crmeta_original}
+	\caption[Summary of the \glspl{deg} identified from Creighton's data]{(add venn diagram)}
+	\label{fig:venn1}
+\end{figure}
+
+\noindent
+There was a possible bias between different ethnic groups, where African American patients were more likely to be obese compared to the Caucasian patients (see \cref{ssub:creighton_study}).
+Though ethnicity was controlled in the residual data, the effect of ethnicity on the data was completely removed to prevent any possibility of ethnicity influencing the analysis.
+Therefore, the effect of ethnicity was completely removed by considering only the Caucasian patients in Creighton's data, which left a total of 77 Caucasian patients in the data set.
+With this Caucasian-only data set, obesity associated genetic signatures were identified as described above; first with unadjusted data, then with the data with clinical variables adjusted (except ethnicity, as it has already been controlled for), and lastly the common genes with Creighton's original signature were identified.
+
+2129 gene probes and 1558 gene probes were identified in the unadjusted and clinical variable-adjusted Caucasian patient data, respectively.
+As before, the most significant 799 gene probes were selected from these gene probes.
+There were 148 and 92 common gene probes with Creighton's original genetic signatures and unadjusted or clinical variable-adjusted Caucasian patient data set, respectively.
+Again, Venn diagram was used to summarise the genes identified from the Caucasian patient data set (\cref{fig:venn2}).
+\\
+
+\begin{figure}[htp!]
+	\centering
+	\includegraphics[width=0.45\linewidth]{results1/crmeta_original}
+	\caption[Summary of the \glspl{deg} identified from Creighton's data (Caucasian patient data)]{(add venn diagram)}
+	\label{fig:venn2}
+\end{figure}
+
+\noindent
+For simplicity, Creighton's original obesity associated signature will be referred to as ``original'' signature, new signatures from unadjusted and clinical variable-adjusted Creighton's data as ``Cr'' and ``Res'' signature respectively, and the signatures from unadjusted and clinical variable-adjusted Caucasian patient data set as ``Ca'' and ``CaRes'' signature respectively.
+All of the overlap signatures will be referred to as their respective signature, but with ``Ol'' added at the end, for example ``CaResOl'' for the overlap signature identified in clinical variable-adjusted Caucasian patient data set.
+All of these obesity associated genetic signatures (eight in total) were checked to see whether these signatures had significant association with sample \gls{bmi}/\gls{bmi} status.
+
+\subsection{Novel obesity associated signatures and sample \gls{bmi}/\gls{bmi} status}
+\label{sub:_novel_obesity_associated_signatures_and_sample_bmi}
+
+As with the original signature, all eight signatures were validated in Creighton's data set first, and then compared in other cancer data sets by using the transformation matrix generated in Creighton's data.
+% Before the metagenes were created, all of the signatures were converted into gene symbols, as described in \cref{sec:data} (\cref{tab:signature_gene_num}).
+Creighton's data was normalised with \gls{rma} method and \gls{svd} was applied to generate the metagene for the corresponding genetic signature.
+
+% \begin{table}[htbp]
+% 	\centering
+% 	\caption{Summary of the number of gene probes and genes in each obesity associated genetic signature}
+% 	\label{tab:signature_gene_num}
+% 	\begin{tabular}{lcc}
+% 		Obesity associated genetic signature & No. gene probes   & No. genes\\
+% 		\hline
+% 		\rule{0pt}{2.25ex}Cr & 799 & 678\\
+% 		Res                  & 799 & 655\\
+% 		Ca                   & 799 & 657\\
+% 		CaRes                & 799 & 651\\
+% 		CrOl                 & 239 & 199\\
+% 		ResOl                & 168 & 147\\
+% 		CaOl                 & 148 & 128\\
+% 		CaResOl              &  92 & 86\\
+% 		\hline
+% 		\hline
+% 	\end{tabular}
+% \end{table}
 

### 2017-01-31 thesis/methods/methods.tex
tmp commit
tmp commit; added appendix file

@@
 Each microarray dataset was normalised separately then combined together into a single dataset, and the batch effect was corrected with the \texttt{ComBat} function (an implementation of the batch correcting method by \citet{Johnson2007}) from the \textit{sva} package in \gls{r} \citep{Leek2012}.
+
+\subsection{Sample randomisation in simulation analysis}
+\label{sub:sample_randomisation_in_simulation_analysis}
+
+In some cases, an analysis was repeated a number of times to provide statistical evidence that the observed results did not occur by chance.
+All of the simulated results must be derived from a randomised data, and by comparing the true results with the results that occured by chance, simulation analysis provides statistical evidence for the true results.
+When a simulation was required, none of the values in the original data were modified but the values in the original data were assigned randomly to different samples; in other words, the samples were shuffled.
+This provided a random set of data in which a simulation could be conducted with.
+The samples were shuffled with the \texttt{sample} function from the \textit{base} package in \gls{r}, and were reshuffled for each simulation.
 
@@
 Heatmaps were created using the \texttt{heatmap.2} function from the \textit{gplots} package in \gls{r}.
+
+\subsection{Venn diagrams}
+\label{sub:venn_diagrams}

### 2017-01-31 thesis/misc/appendix.tex
tmp commit; added appendix file

\begin{appendices}
	\crefalias{chapter}{appcha}
	\crefalias{section}{appsec}
	\renewcommand{\thesection}{\Alph{section}\arabic{section}}
	\chapter{Additional results from \cref{cha:obesity_genetic_signatures_and_cancer}}
	\label{app:a}

	\section{Number of gene probes/genes in each signature}
	\label{app:number_of_gene_probes_genes_in_each_signature}

\end{appendices}

### 2017-01-31 thesis/results/results.tex
tmp commit; added appendix file

@@
 \noindent
+In addition to these eight obesity associated genetic signatures, a genetic signature was identified based on the correlation of gene probe expression with sample \gls{bmi} as continuous values, rather than discrete values as in \gls{bmi} status.
+However, no significant genes were identified and thus was not used for further analysis (results detailed in appendix).
+
 For simplicity, Creighton's original obesity associated signature will be referred to as ``original'' signature, new signatures from unadjusted and clinical variable-adjusted Creighton's data as ``Cr'' and ``Res'' signature respectively, and the signatures from unadjusted and clinical variable-adjusted Caucasian patient data set as ``Ca'' and ``CaRes'' signature respectively.
 All of the overlap signatures will be referred to as their respective signature, but with ``Ol'' added at the end, for example ``CaResOl'' for the overlap signature identified in clinical variable-adjusted Caucasian patient data set.
-All of these obesity associated genetic signatures (eight in total) were checked to see whether these signatures had significant association with sample \gls{bmi}/\gls{bmi} status.
+All of these obesity associated genetic signatures (eight in total) were checked to see whether these signatures had significant association with sample \gls{bmi}/\gls{bmi} status in various data set.
 
@@
 % Before the metagenes were created, all of the signatures were converted into gene symbols, as described in \cref{sec:data} (\cref{tab:signature_gene_num}).
-Creighton's data was normalised with \gls{rma} method and \gls{svd} was applied to generate the metagene for the corresponding genetic signature.
+Creighton's original data (unadjusted, all samples included) was normalised with \gls{rma} method and \gls{svd} was applied to generate the metagene for the corresponding genetic signature.
+The direction of the metagenes were first examined to make sure that all of the metagenes were in line with one another (see \cref{sub:metagene_direction}; results shown in appendix).
+The comparison of the metagenes with the sample gene expression in Creighton's data were displayed as heatmaps (\cref{fig:degmetacr}).
+It was clear from the heatmaps that all of the metagenes reflected the overall expression of the corresponding genetic signatures.
+The association of the metagenes with sample \gls{bmi}/\gls{bmi} status was significant for all eight of the metagenes identified (\cref{fig:degmetacrbmi}).
+These results confirmed that all of the obesity associated genetic signatures identified in \cref{sub:identification_of_obesity_associated_genetic_signatures} significantly associated with sample \gls{bmi}/\gls{bmi} status in the data in which the metagenes were derived from.
+
+\begin{figure}[htp!]
+	\centering
+	\includegraphics[width=0.45\linewidth]{results1/crmeta_original}
+	\caption[Novel obesity associated metagenes in Creighton's data]{(add all eight metagene heatmaps)}
+	\label{fig:degmetacr}
+\end{figure}
+
+\begin{figure}[htp!]
+	\centering
+	\includegraphics[width=0.45\linewidth]{results1/crmeta_original}
+	\caption[Novel obesity associated metagenes in Creighton's data]{(add all eight metagene heatmaps)}
+	\label{fig:degmetacrbmi}
+\end{figure}
+
+To confirm whether these metagenes showed significant association with sample \gls{bmi}/\gls{bmi} status in other cancer data sets, transformation matrix was applied to the \gls{icgc} cancer data and Print's breast cancer data.
+Firstly in \gls{icgc} cancer data sets, all eight metagene scores were reflective of the sample gene expression of the genetic signatures (\cref{fig:degmetaicgc}), but none of the metagenes associated significantly with sample \gls{bmi}/\gls{bmi} status (\cref{fig:degmetaicgcbmi}).
+Likewise in Print's data, the higher the metagenes scores were the more expressed the genes were (and vice versa; \cref{fig:degmetaprint}), but lacked the association with sample \gls{bmi}/\gls{bmi} status (\cref{fig:degmetaprintbmi}).
+
+These results showed that, even though all of the metagenes significantly associated with sample \gls{bmi}/\gls{bmi} status in the data the genetic signatures were derived from, the metagenes did not transfer well in other cancer data sets.
+Furthermore, these results showed that the lack of association with sample \gls{bmi}/\gls{bmi} status were not due to other clinical variables in Creighton's data set.
+This raised a question of whether there are any obesity associated genetic signature that are common in multiple cancer types.
+
+\begin{figure}[htp!]
+	\centering
+	\includegraphics[width=0.45\linewidth]{results1/crmeta_original}
+	\caption[Novel obesity associated metagenes in \gls{icgc} data]{(add all eight metagene heatmaps)}
+	\label{fig:degmetaicgc}
+\end{figure}
+
+\begin{figure}[htp!]
+	\centering
+	\includegraphics[width=0.45\linewidth]{results1/crmeta_original}
+	\caption[Novel obesity associated metagenes in \gls{icgc} data]{(add all eight metagene heatmaps)}
+	\label{fig:degmetaicgcbmi}
+\end{figure}
+
+\begin{figure}[htp!]
+	\centering
+	\includegraphics[width=0.45\linewidth]{results1/crmeta_original}
+	\caption[Novel obesity associated metagenes in Print's data]{(add all eight metagene heatmaps)}
+	\label{fig:degmetaprint}
+\end{figure}
+
+\begin{figure}[htp!]
+	\centering
+	\includegraphics[width=0.45\linewidth]{results1/crmeta_original}
+	\caption[Novel obesity associated metagenes in Print's data]{(add all eight metagene heatmaps)}
+	\label{fig:degmetaprintbmi}
+\end{figure}
 
@@
 
-It was clear that the metagenes created from obesity signatures were not consistent across different data sets.
-To identify genes that were associated with sample \gls{bmi} in multiple cancer types, \glspl{deg} were searched (?).
+It was clear that the obesity associated genetic signatures created from Creighton's data were not consistent across different data sets.
+To determine whether there was any obesity associated genetic signature that was expressed in multiple different cancer types, gene expression analysis was carried out on each of the eight \gls{icgc} cancer types and common genes were searched from the \glspl{deg} that were identified.
+All of the \gls{icgc} cancer data sets were normalised with voom method (\cref{ssub:rna_seq_data}), which were then put through the gene expression analysis pipeline to identify \glspl{deg} between the obese group of samples and non-obese group of samples (\cref{sec:gene_expression_analysis}).
+\cref{tab:icgcdegnum} summarised the number of genes found from each cancer type with p \textless{} 0.05.
+
+\begin{table}[htbp]
+	\centering
+	\caption{Summary of the number of \glspl{deg} identified in each \gls{icgc} cancer data set}
+	\label{tab:icgcdegnum}
+	\begin{tabular}{lc}
+		Cancer type & No. of \glspl{deg}\\
+		\hline
+		\rule{0pt}{2.25ex}\gls{blca} & 679 \\
+		\gls{cesc} & 1229\\
+		\gls{coad} & 974\\
+		\gls{kirp} & 687\\
+		\gls{lihc} & 3340\\
+		\gls{read} & 796\\
+		\gls{skcm} & 1137\\
+		\gls{ucec} & 2934\\
+		\hline
+		\hline
+	\end{tabular}
+\end{table}
+
+There were 9695 unique \glspl{deg} across the eight cancer types, and these genes were checked for any commonalities across the different cancer types.
+There were 7330 genes that were differentially expressed by one cancer type, 2024 genes expressed by any two cancer types, 320 genes expressed by any three cancer types, and 21 genes expressed by any four cancer types.
+There were no genes differentially expressed by six or more cancer types (see \cref{tab:icgcdegtab} for a summary).
+To confirm that these results were statistically significant, the gene expression analysis was repeated 1000 times for each cancer type after the samples were randomised in each analysis (see \cref{sub:sample_randomisation_in_simulation_analysis}).
+The results from the simulation were summarised together with the earlier results in \cref{tab:icgcdegtab}.
+
+(insert description of the simulation results).
+The results from the simulation showed that the \glspl{deg} identified from the gene expression analysis of \gls{icgc} cancer data were significant.
+However, the results also confirmed that there were no obesity associated \glspl{deg} that were common across different cancer types.
+
+\begin{table}[htbp]
+	\centering
+	\begin{threeparttable}
+		\caption{Summary of the number of \glspl{deg} identified by the gene expression analysis and simulation analysis in \gls{icgc} cancer data}
+		\label{tab:icgcdegtab}
+		\begin{tabular}{>{\quad}lcccccccc}
+			& \multicolumn{8}{c}{No. of cancer types expressing the \glspl{deg}}\\
+			& 1 & 2 & 3 & 4 & 5 & 6 & 7 & 8\\
+			\hline
+			\rule{0pt}{2.25ex}\hspace{-1em}Results from gene expression analysis & 7330 & 2024 & 320 & 21 & 0 & 0 & 0 & 0  \\
+			\hspace{-1em}Results from simulation:                                &      &      &     &    &   &   &   &  \\
+			Mean no. of \glspl{deg} identified                                   & 0    & 0    & 0   & 0  & 0 & 0 & 0 & 0  \\
+			$95^{th}$ percentile                                                 & 0    & 0    & 0   & 0  & 0 & 0 & 0 & 0  \\
+			\hline
+			\hline
+		\end{tabular}
+		% \begin{tablenotes}
+		% 	\begin{footnotesize}
+		% 	\item [1] The numbers represent the number of cancer types a gene was expressed in.
+		% 	\item [2] The simulation was repeated 1000 times.
+		% 	\end{footnotesize}
+		% \end{tablenotes}
+	\end{threeparttable}
+\end{table}
 

### 2017-02-01 thesis/methods/methods.tex
Added figure images into results; tmp commit

@@
 
-\subsection{\gls{go} database}
-\label{sub:go_database}
+\subsection{Pathway databases}
+\label{sub:pathway_databases}
+
+\subsubsection{\gls{go} database}
+\label{ssub:go_database}
 
@@
 
-Due to the sheer amount of terms and definitions available from the data, \gls{go} was chosen as the database for pathway reference.
-The \textit{GO.db} package was used to load the \gls{go} database into \gls{r} \citep{Carlson2016}.
-Only the human related \gls{go} terms were used in pathway enrichment analysis, and these terms were manipulated in a form which the \texttt{camera} function was able to recognise.
+% Due to the sheer amount of terms and definitions available from the data, \gls{go} was chosen as the database for pathway reference.
+% The \textit{GO.db} package was used to load the \gls{go} database into \gls{r} \citep{Carlson2016}.
+% Only the human related \gls{go} terms were used in pathway enrichment analysis, and these terms were manipulated in a form which the \texttt{camera} function was able to recognise.
+
+\subsubsection{\gls{kegg} database}
+\label{ssub:kegg_database}
+
+\subsubsection{Reactome database}
+\label{ssub:Reactome_database}
 

### 2017-02-01 thesis/misc/appendix.tex
Added figure images into results; tmp commit

@@
 	\crefalias{section}{appsec}
-	\renewcommand{\thesection}{\Alph{section}\arabic{section}}
+	\renewcommand{\thesection}{\Alph{chapter}\arabic{section}}
 	\chapter{Additional results from \cref{cha:obesity_genetic_signatures_and_cancer}}
 	\label{app:a}
+
+	\section{Metagenes created from raw data vs. standardised data (Creighton)}
+	\label{sec:metagenes_created_from_raw_data_vs_standardised_data}
+
+	\begin{figure}[htpb]
+		\centering
+		\includegraphics[width=0.45\linewidth,page=1]{appendix/check_raw_vs_std_mg}
+		\hfill
+		\includegraphics[width=0.45\linewidth,page=2]{appendix/check_raw_vs_std_mg}
+		\caption{appendix/Check_raw_vs_std}
+		\label{fig:appendix/check_raw_vs_std}
+	\end{figure}
+
+	\section{Metagenes created from raw data vs. standardised data (ICGC)}
+	\label{sec:metagenes_created_from_raw_data_vs_standardised_data_icgc}
+
+	\begin{figure}[htpb]
+		\centering
+		\includegraphics[width=0.45\linewidth,page=1]{appendix/crtcga_raw_vs_std}
+		\hfill
+		\includegraphics[width=0.45\linewidth,page=2]{appendix/crtcga_raw_vs_std}
+		\caption{appendix/Check_raw_vs_std}
+		\label{fig:appendix/check_raw_vs_std}
+	\end{figure}
+
+	\section{Rest of the ICGC cancer heatmap resutls}
+	\label{sec:rest_of_the_icgc_cancer_heatmap_resutls}
+
+	\begin{figure}[htpb]
+		\centering
+		\includegraphics[width=0.45\linewidth,page=1]{appendix/crtcga_std}
+		\hfill
+		\includegraphics[width=0.45\linewidth,page=2]{appendix/crtcga_std}
+		\caption{appendix/Check_raw_vs_std}
+		\label{fig:appendix/check_raw_vs_std}
+	\end{figure}
+
+	\section{FM metagene, raw vs std}
+	\label{sec:fm_metagene_raw_vs_std}
+
+	\begin{figure}[htpb]
+		\centering
+		\includegraphics[page=1,width=0.8\linewidth]{appendix/fm_meta_check}
+		\caption{appendix/fm_meta_check}
+		\label{fig:appendix/fm_meta_check}
+	\end{figure}
+
+	\section{Creighton DEG metagene direction}
+	\label{sec:creighton_deg_metagene_direction}
+
+	\begin{figure}[htpb]
+		\centering
+		\includegraphics[page=1,width=0.8\linewidth]{cr_meta_direction}
+		\caption{cr_meta_direction}
+		\label{fig:cr_meta_direction}
+	\end{figure}
 

### 2017-02-01 thesis/results/results.tex
Added figure images into results; tmp commit

@@
 There was no description about the normalisation method used by Creighton's group when they first analysed their data, so the more reliable \gls{rma} method was used to  normalise the Creighton data set.
-The obesity metagene was plotted in a heatmap with the sample gene expression (\cref{fig:crmetaheat1}) to check whether the metagene scores were in accordance with the overall gene expression of the samples.
+The obesity metagene was plotted in a heatmap with the sample gene expression (\cref{fig:crmetaheat}) to check whether the metagene scores were in accordance with the overall gene expression of the samples.
 
@@
 	\centering
-	\includegraphics[width=0.8\linewidth]{results1/crmeta_original}
-	\caption[Heatmap of Creighton's metagene in Creighton's data]{Heatmap showing Creighton's obesity associated metagene with the gene expressions of obesity associated genes in Creighton's data.
+	\includegraphics[page=3,width=0.8\linewidth]{results1/creighton_mg_heatmap1}
+	\caption[Creighton's metagene and sample gene expression in Creighton's data]{Heatmap showing Creighton's obesity associated metagene with the gene expressions of obesity associated genes in Creighton's data.
 	Level of expression is represented in the top right histogram, where low and high gene expression were colour-coded with blue and red, respectively.
@@
 	For clarity, the metagene scores were flipped in order to match the results from \citet{Creighton2012} paper.}
-	\label{fig:crmetaheat1}
+	\label{fig:crmetaheat}
 \end{figure}
 
-As shown in \cref{fig:crmetaheat1}, high obesity associated metagene score of the sample reflected low expression in majority of the genes in the signature, and in contrast, low obesity associated metagene score of the sample reflected high expression in majority of the genes in the signature.
+As shown in \cref{fig:crmetaheat}, high obesity associated metagene score of the sample reflected low expression in majority of the genes in the signature, and in contrast, low obesity associated metagene score of the sample reflected high expression in majority of the genes in the signature.
 This was consistent with the reported property of the obesity associated signature by \citet{Creighton2012}.
-To provide further evidence that the obesity associated metagenes were in fact associated with the \gls{bmi} status and \gls{bmi} value of the samples, a box plot and a scatter plot were created, respectively (\cref{fig:crmetaboxplot1}).
+To provide further evidence that the obesity associated metagenes were in fact associated with the \gls{bmi} status and \gls{bmi} value of the samples, a box plot and a scatter plot were created, respectively (\cref{fig:crmetaboxplot}).
 
@@
 	\centering
-	\includegraphics[width=0.45\linewidth]{results1/crmeta_original}
+	\includegraphics[page=4,width=0.45\linewidth]{results1/creighton_mg_heatmap1}
 	\hfill
-	\includegraphics[width=0.45\linewidth]{results1/crmeta_original}
-	\caption[Creighton metagene and sample \gls{bmi}/\gls{bmi} status]{(insert figure of box plot and scatter plot)}
-	\label{fig:crmetaboxplot1}
+	\includegraphics[page=5,width=0.45\linewidth]{results1/creighton_mg_heatmap1}
+	\caption[Creighton's metagene and sample \gls{bmi}/\gls{bmi} status in Creighton's data]{Box plot and scatter plot showing the association of Creighton's obesity associated metagene with sample \gls{bmi} status and \gls{bmi} from Creighton's data, respectively.
+	In the box plot, the p-values above the groups represent the statistical significance of the association of the metagene with the overweight or obese group compared with the normal weight group.
+	The \gls{anova} p-value shows the statistical significance of the association of the metagene with the sample \gls{bmi} groups.
+	In the scatter plot, $R^2$- and p-values describe the adjusted coefficient of determination of the regression line and the statistical significance of the linear model used to draw the regression line, respectively.}
+	\label{fig:crmetaboxplot}
 \end{figure}
 
-\cref{fig:crmetaboxplot1} clearly showed that the obesity associated metagene from \citet{Creighton2012} study significantly associated with the sample \gls{bmi} status, as well as the \gls{bmi} value of the samples (include p-value and r-squared).
+\cref{fig:crmetaboxplot} clearly showed that the obesity associated metagene from \citet{Creighton2012} study significantly associated with the sample \gls{bmi} status, as well as the \gls{bmi} value of the samples (include p-value and r-squared).
 It should be noted here that the obesity associated metagene significantly associated with the samples that were obese, but not the samples that were overweight.
@@
 Now that the association of the obesity metagene from \citet{Creighton2012} study was established in Creighton's data set, the obesity associated signature was transferred to the \gls{icgc} cancer data.
-The direction of the obesity associated metagene was checked in the Creighton's data, so that high metagene scores reflected high sample \gls{bmi} and low metagene scores reflected low sample \gls{bmi} (\cref{fig:crmetaboxplot1}).
+The direction of the obesity associated metagene was checked in the Creighton's data, so that high metagene scores reflected high sample \gls{bmi} and low metagene scores reflected low sample \gls{bmi} (\cref{fig:crmetaboxplot}).
 The transformation matrix was created in Creighton's data, as described in \cref{sub:svd}.
@@
 The transformation matrix was applied to each cancer data set in turn to obtain the obesity metagenes for each of the data set.
-Each obesity metagenes were plotted on a heatmap with the corresponding data set in which the metagene was taken from \cref{fig:crmetaicgc1}.
+Each obesity metagenes were plotted on a heatmap with the corresponding data set in which the metagene was taken from \cref{fig:crmetaicgc}.
 These heatmaps confirmed that the obesity associated metagene derived from Creighton's data was able to capture the overall gene expression pattern, where the metagene scores reflected the expression levels of majority of the genes in the signature.
+
+As before, association of the obesity metagene with the sample \gls{bmi} and \gls{bmi} status was checked in their respective cancer data set (\cref{fig:crmetaicgc}).
+In contrast to the strong association of the obesity metagene with sample \gls{bmi} and \gls{bmi} status in Creighton's data, the obesity metagene taken from the \gls{icgc} cancer data did not significantly associate with the sample \gls{bmi} or \gls{bmi} status in any of the \gls{icgc} cancer data.
 
@@
 	\centering
-	\includegraphics[width=0.45\linewidth]{results1/crmeta_original}
+	\includegraphics[page=3,width=0.8\linewidth]{results1/crtcga_std}\\
+	\vspace{1em}
+	\includegraphics[page=4,width=0.45\linewidth]{results1/crtcga_std}
 	\hfill
-	\includegraphics[width=0.45\linewidth]{results1/crmeta_original}
-	\caption[Creighton metagene and \acrshort{icgc} cancer data]{(insert ICGC heatmaps)}
-	\label{fig:crmetaicgc1}
+	\includegraphics[page=5,width=0.45\linewidth]{results1/crtcga_std}
+	\caption[Creighton's metagene in \acrshort{icgc} \acrshort{blca} data]{Heatmap, box plot and scatter plot showing the association of Creighton's obesity associated metagene with sample gene expression, \gls{bmi} and \gls{bmi} status from \gls{icgc} \gls{blca} data, respectively.
+	The results for other cancer types are in \cref{app:a}.
+	Scales, p-values and $R^2$-value are as described in previous figures.}
+	\label{fig:crmetaicgc}
 \end{figure}
-
-\begin{figure}[htp!]
-	\centering
-	\includegraphics[width=0.45\linewidth]{results1/crmeta_original}
-	\hfill
-	\includegraphics[width=0.45\linewidth]{results1/crmeta_original}
-	\caption[Creighton metagene and sample \gls{bmi}/\gls{bmi} status in \acrshort{icgc} data]{(insert figure of box plot and scatter plot for ICGC data)}
-	\label{fig:crmetaboxicgc1}
-\end{figure}
-
-As before, association of the obesity metagene with the sample \gls{bmi} and \gls{bmi} status was checked in their respective cancer data set (\cref{fig:crmetaboxicgc1}).
-In contrast to the strong association of the obesity metagene with sample \gls{bmi} and \gls{bmi} status in Creighton's data, the obesity metagene taken from the \gls{icgc} cancer data did not significantly associate with the sample \gls{bmi} or \gls{bmi} status in any of the \gls{icgc} cancer data.
 
@@
 Print's data was normalised with \gls{rma} method as described in \cref{ssub:microarray_data}, and the transformation matrix was applied to the normalised data to obtain the metagene.
-The metagene was again compared with the gene expression of the samples with a heatmap (\cref{fig:crmetaprint1}), and the association of the metagene with the sample \gls{bmi} and \gls{bmi} status was examined with box and scatter plots (\cref{fig:crmetaboxprint1}).
+The metagene was again compared with the gene expression of the samples with a heatmap (\cref{fig:crmetaprint}), and the association of the metagene with the sample \gls{bmi} and \gls{bmi} status was examined with box and scatter plots (\cref{fig:crmetaprint}).
+
+The obesity metagene managed to reflect the overall gene expression of the samples in Print's data.
+However, as in \gls{icgc} cancer data, the obesity metagene scores did not significantly associate with the sample \gls{bmi} or \gls{bmi} status.
+These results confirmed that the lack of association of the obesity metagene from Creighton's data was not due to the technology in which the data was gathered (mcroarray or \gls{rnaseq}), nor the type of cancer in which the transformation matrix was applied to.
 
@@
 	\centering
-	\includegraphics[width=0.45\linewidth]{results1/crmeta_original}
+	\includegraphics[width=0.8\linewidth,page=3]{results1/cris_cr_trans_meta}\\
+	\vspace{1em}
+	\includegraphics[width=0.45\linewidth,page=4]{results1/cris_cr_trans_meta}
 	\hfill
-	\includegraphics[width=0.45\linewidth]{results1/crmeta_original}
-	\caption[Creighton metagene and Print's cancer data]{(insert Print heatmaps)}
-	\label{fig:crmetaprint1}
+	\includegraphics[width=0.45\linewidth,page=5]{results1/cris_cr_trans_meta}
+	\caption[Creighton's metagene in Print's breast cancer data]{Heatmap, box plot and scatter plot showing the association of Creighton's obesity associated metagene with sample gene expression, \gls{bmi} and \gls{bmi} status from Print's breast cancer data, respectively.
+	Scales, p-values and $R^2$-value are as described in previous figures.}
+	\label{fig:crmetaprint}
 \end{figure}
-
-\begin{figure}[htp!]
-	\centering
-	\includegraphics[width=0.45\linewidth]{results1/crmeta_original}
-	\hfill
-	\includegraphics[width=0.45\linewidth]{results1/crmeta_original}
-	\caption[Creighton metagene and sample \gls{bmi}/\gls{bmi} status in Print's data]{(insert figure of box plot and scatter plot for Print data)}
-	\label{fig:crmetaboxprint1}
-\end{figure}
-
-The obesity metagene managed to reflect the overall gene expression of the samples in Print's data.
-However, like in \gls{icgc} cancer data, the obesity metagene scores did not significantly associate with the sample \gls{bmi} or \gls{bmi} status.
-These results confirmed that the lack of association of the obesity metagene from Creighton's data was not due to the technology in which the data was gathered (mcroarray or \gls{rnaseq}), nor the type of cancer in which the transformation matrix was applied to.
 
@@
 The transformation matrix was used to transform the \gls{rma}-normalised Creighton's data to extract FM's obesity metagene scores.
-FM's obesity metagene scores were compared with gene expressions and sample \gls{bmi}/\gls{bmi} status in Creighton's data, as shown in \cref{fig:fmmetacr1,fig:fmmetaboxcr1}.
+FM's obesity metagene scores were compared with gene expressions and sample \gls{bmi}/\gls{bmi} status in Creighton's data, as shown in \cref{fig:fmmetacr}.
 Clearly, like with Creighton's obesity metagene, FM's obesity metagene was reflective of the overall gene expression of the samples, but did not associate with the sample \gls{bmi} or \gls{bmi} status.
@@
 	\centering
-	\includegraphics[width=0.45\linewidth]{results1/crmeta_original}
+	\includegraphics[page=3,width=0.8\linewidth]{results1/cr_fm_meta}\\
+	\vspace{1em}
+	\includegraphics[page=4,width=0.45\linewidth]{results1/cr_fm_meta}
 	\hfill
-	\includegraphics[width=0.45\linewidth]{results1/crmeta_original}
-	\caption[FM metagene and Creighton's cancer data]{(insert Creighton heatmaps)}
-	\label{fig:fmmetacr1}
+	\includegraphics[page=5,width=0.45\linewidth]{results1/cr_fm_meta}
+	\caption[FM's metagene in Creighton's  data]{Heatmap, box plot and scatter plot showing the association of FM's obesity associated metagene with sample gene expression, \gls{bmi} and \gls{bmi} status from Creighton's  data, respectively.
+	Scales, p-values and $R^2$-value are as described in previous figures.}
+	\label{fig:fmmetacr}
 \end{figure}
+
+Next, the transformation matrix was applied to the \gls{icgc} cancer data and resulting metagenes were compared with the gene expression and sample \gls{bmi}/\gls{bmi} status.
+As evident in \cref{fig:fmmetaicgc}, FM's obesity metagene scores appeared to reflect the gene expression of FM's obesity associated genetic signature.
+However, as with all the results in this section so far, FM's obesity metagene did not significantly associate with any of the \gls{icgc} cancer data (\cref{fig:fmmetaicgc}).
 
@@
 	\centering
-	\includegraphics[width=0.45\linewidth]{results1/crmeta_original}
+	\includegraphics[page=3,width=0.8\linewidth]{results1/fm_meta_ICGC}\\
+	\vspace{1em}
+	\includegraphics[page=4,width=0.45\linewidth]{results1/fm_meta_ICGC}
 	\hfill
-	\includegraphics[width=0.45\linewidth]{results1/crmeta_original}
-	\caption[FM metagene and sample \gls{bmi}/\gls{bmi} status in Creighton's data]{(insert figure of box plot and scatter plot for Creighton data)}
-	\label{fig:fmmetaboxcr1}
+	\includegraphics[page=5,width=0.45\linewidth]{results1/fm_meta_ICGC}
+	\caption[FM's metagene in \acrshort{icgc} \acrshort{blca} data]{Heatmap, box plot and scatter plot showing the association of FM's obesity associated metagene with sample gene expression, \gls{bmi} and \gls{bmi} status from \acrshort{icgc} \acrshort{blca} data, respectively.
+	The results for other cancer types are in \cref{app:a}.
+	Scales, p-values and $R^2$-value are as described in previous figures.}
+	\label{fig:fmmetaicgc}
 \end{figure}
 
-Next, the transformation matrix was applied to the \gls{icgc} cancer data and resulting metagenes were compared with the gene expression and sample \gls{bmi}/\gls{bmi} status.
-As evident in \cref{fig:fmmetaicgc1}, FM's obesity metagene scores appeared to reflect the gene expression of FM's obesity associated genetic signature.
-However, as with all the results in this section so far, FM's obesity metagene did not significantly associate with any of the \gls{icgc} cancer data (\cref{fig:fmmetaboxicgc1}).
-
-\begin{figure}[htp!]
-	\centering
-	\includegraphics[width=0.45\linewidth]{results1/crmeta_original}
-	\hfill
-	\includegraphics[width=0.45\linewidth]{results1/crmeta_original}
-	\caption[FM metagene and \gls{icgc} cancer data]{(insert ICGC heatmaps)}
-	\label{fig:fmmetaicgc1}
-\end{figure}
-
-\begin{figure}[htp!]
-	\centering
-	\includegraphics[width=0.45\linewidth]{results1/crmeta_original}
-	\hfill
-	\includegraphics[width=0.45\linewidth]{results1/crmeta_original}
-	\caption[FM metagene and sample \gls{bmi}/\gls{bmi} status in \gls{icgc} data]{(insert figure of box plot and scatter plot for Creighton data)}
-	\label{fig:fmmetaboxicgc1}
-\end{figure}
-
-Lastly, the transformation matrix was applied to Print's data (\cref{fig:fmmetaboxcris1,fig:fmmetacris1}).
+Lastly, the transformation matrix was applied to Print's data (\cref{fig:fmmetacris}).
 Again, FM's obesity metagene scores reflected the gene expression of the samples, did not significantly associate with sample \gls{bmi}/\gls{bmi} status.
@@
 	\centering
-	\includegraphics[width=0.45\linewidth]{results1/crmeta_original}
+	\includegraphics[page=3,width=0.8\linewidth]{results1/cris_fm_meta}\\
+	\vspace{1em}
+	\includegraphics[page=4,width=0.45\linewidth]{results1/cris_fm_meta}
 	\hfill
-	\includegraphics[width=0.45\linewidth]{results1/crmeta_original}
-	\caption[FM metagene and Print's cancer data]{(insert Print heatmaps)}
-	\label{fig:fmmetacris1}
-\end{figure}
-
-\begin{figure}[htp!]
-	\centering
-	\includegraphics[width=0.45\linewidth]{results1/crmeta_original}
-	\hfill
-	\includegraphics[width=0.45\linewidth]{results1/crmeta_original}
-	\caption[FM metagene and sample \gls{bmi}/\gls{bmi} status in Print's data]{(insert figure of box plot and scatter plot for Print's data)}
-	\label{fig:fmmetaboxcris1}
+	\includegraphics[page=5,width=0.45\linewidth]{results1/cris_fm_meta}
+	\caption[FM's metagene in Print's breast cancer data]{Heatmap, box plot and scatter plot showing the association of FM's obesity associated metagene with sample gene expression, \gls{bmi} and \gls{bmi} status from Print's breast cancer data, respectively.
+	Scales, p-values and $R^2$-value are as described in previous figures.}
+	\label{fig:fmmetacris}
 \end{figure}
@@
 	\centering
-	\includegraphics[width=0.45\linewidth]{results1/crmeta_original}
+	\includegraphics[page=1,width=0.8\linewidth]{results1/deg_venn}
 	\caption[Summary of the \glspl{deg} identified from Creighton's data]{(add venn diagram)}
@@
 	\centering
-	\includegraphics[width=0.45\linewidth]{results1/crmeta_original}
+	\includegraphics[page=2,width=0.8\linewidth]{results1/deg_venn}
 	\caption[Summary of the \glspl{deg} identified from Creighton's data (Caucasian patient data)]{(add venn diagram)}
@@
 It was clear from the heatmaps that all of the metagenes reflected the overall expression of the corresponding genetic signatures.
-The association of the metagenes with sample \gls{bmi}/\gls{bmi} status was significant for all eight of the metagenes identified (\cref{fig:degmetacrbmi}).
+The association of the metagenes with sample \gls{bmi}/\gls{bmi} status was significant for all eight of the metagenes identified (\cref{fig:degmetacr}).
 These results confirmed that all of the obesity associated genetic signatures identified in \cref{sub:identification_of_obesity_associated_genetic_signatures} significantly associated with sample \gls{bmi}/\gls{bmi} status in the data in which the metagenes were derived from.
@@
 	\centering
-	\includegraphics[width=0.45\linewidth]{results1/crmeta_original}
-	\caption[Novel obesity associated metagenes in Creighton's data]{(add all eight metagene heatmaps)}
+	\includegraphics[page=3,width=0.8\linewidth]{results1/cr_deg_meta_vs_clin}\\
+	\vspace{1em}
+	\includegraphics[page=4,width=0.45\linewidth]{results1/cr_deg_meta_vs_clin}
+	\hfill
+	\includegraphics[page=5,width=0.45\linewidth]{results1/cr_deg_meta_vs_clin}
+	\caption[Cr obesity associated metagene in Creighton's data]{Heatmap, box plot and scatter plot showing the association of Cr obesity associated metagene with sample gene expression, \gls{bmi} and \gls{bmi} status from Creighton's data, respectively.
+	The results for other metagenes are in \cref{app:a}.
+	Scales, p-values and $R^2$-value are as described in previous figures.}
 	\label{fig:degmetacr}
@@
 
-\begin{figure}[htp!]
-	\centering
-	\includegraphics[width=0.45\linewidth]{results1/crmeta_original}
-	\caption[Novel obesity associated metagenes in Creighton's data]{(add all eight metagene heatmaps)}
-	\label{fig:degmetacrbmi}
-\end{figure}
-
 To confirm whether these metagenes showed significant association with sample \gls{bmi}/\gls{bmi} status in other cancer data sets, transformation matrix was applied to the \gls{icgc} cancer data and Print's breast cancer data.
-Firstly in \gls{icgc} cancer data sets, all eight metagene scores were reflective of the sample gene expression of the genetic signatures (\cref{fig:degmetaicgc}), but none of the metagenes associated significantly with sample \gls{bmi}/\gls{bmi} status (\cref{fig:degmetaicgcbmi}).
-Likewise in Print's data, the higher the metagenes scores were the more expressed the genes were (and vice versa; \cref{fig:degmetaprint}), but lacked the association with sample \gls{bmi}/\gls{bmi} status (\cref{fig:degmetaprintbmi}).
+Firstly in \gls{icgc} cancer data sets, all eight metagene scores were reflective of the sample gene expression of the genetic signatures (\cref{fig:degmetaicgc}), but none of the metagenes associated significantly with sample \gls{bmi}/\gls{bmi} status (\cref{fig:degmetaicgc}).
+Likewise in Print's data, the higher the metagenes scores were the more expressed the genes were (and vice versa; \cref{fig:degmetaprint}), but lacked the association with sample \gls{bmi}/\gls{bmi} status (\cref{fig:degmetaprint}).
 
@@
 	\centering
-	\includegraphics[width=0.45\linewidth]{results1/crmeta_original}
-	\caption[Novel obesity associated metagenes in \gls{icgc} data]{(add all eight metagene heatmaps)}
+	\includegraphics[page=3,width=0.8\linewidth]{results1/rawobsgene_ICGC}\\
+	\vspace{1em}
+	\includegraphics[page=4,width=0.45\linewidth]{results1/rawobsgene_ICGC}
+	\hfill
+	\includegraphics[page=5,width=0.45\linewidth]{results1/rawobsgene_ICGC}
+	\caption[Cr obesity associated metagene in \acrshort{icgc} \acrshort{blca} data]{Heatmap, box plot and scatter plot showing the association of Cr obesity associated metagene with sample gene expression, \gls{bmi} and \gls{bmi} status from \acrshort{icgc} \acrshort{blca} data, respectively.
+	The results for other metagenes and other cancer types are in \cref{app:a}.
+	Scales, p-values and $R^2$-value are as described in previous figures.}
 	\label{fig:degmetaicgc}
@@
 	\centering
-	\includegraphics[width=0.45\linewidth]{results1/crmeta_original}
-	\caption[Novel obesity associated metagenes in \gls{icgc} data]{(add all eight metagene heatmaps)}
-	\label{fig:degmetaicgcbmi}
-\end{figure}
-
-\begin{figure}[htp!]
-	\centering
-	\includegraphics[width=0.45\linewidth]{results1/crmeta_original}
-	\caption[Novel obesity associated metagenes in Print's data]{(add all eight metagene heatmaps)}
+	\includegraphics[page=3,width=0.8\linewidth]{results1/cris_crdeg_trans_meta}\\
+	\vspace{1em}
+	\includegraphics[page=4,width=0.45\linewidth]{results1/cris_cr_trans_meta}
+	\hfill
+	\includegraphics[page=5,width=0.45\linewidth]{results1/cris_cr_trans_meta}
+	\caption[Cr obesity associated metagene in Print's data]{Heatmap, box plot and scatter plot showing the association of Cr obesity associated metagene with sample gene expression, \gls{bmi} and \gls{bmi} status from Print's data, respectively.
+	The results for other metagenes are in \cref{app:a}.
+	Scales, p-values and $R^2$-value are as described in previous figures.}
 	\label{fig:degmetaprint}
-\end{figure}
-
-\begin{figure}[htp!]
-	\centering
-	\includegraphics[width=0.45\linewidth]{results1/crmeta_original}
-	\caption[Novel obesity associated metagenes in Print's data]{(add all eight metagene heatmaps)}
-	\label{fig:degmetaprintbmi}
 \end{figure}
@@
 \label{sec:pathways_enriched_in_obesity_associated_genes}
+
+To check whether there were any significant pathways enriched in any of the \gls{icgc} cancer data sets that were associated with obesity, pathway enrichment analysis was carried out on each cancer type separately and then all cancer types combined (see \cref{sec:pathway_enrichment_analysis}).

### 2017-02-01 thesis/results/results2.tex
Added figure images into results; tmp commit

--- 2016-09-17 thesis/results/results2.tex
@@
-\chapter{Results from gatza pathway metagene and direction stuff}
+\chapter{Obesity associated genetic signatures and pathway signatures}
+\label{cha:obesity_associated_genetic_signature_and_pathway_signatures}
 
-\section{Reproducing results from gatza paper}
+In this chapter, the underlying biological mechanism of the obesity associated signatures from Creighton's data were investigated.
+In doing so, Gatza's pathway genetic signatures were utilised to determine which biological pathways the obesity associated genetic signatures were most similar to.
+First, the direction of Gatza's pathway associated genetic signatures were resolved; then the pathway associated metagenes were compared with the obesity associated metagenes; and lastly linear models were constructed based on the pathway metagenes and sample \gls{bmi}/\gls{bmi} status to predict the obesity associated metagenes.
+
+\section{Pathway associated genetic signatures from \citet{Gatza2010a} study}
+\label{sec:pathway_associated_genetic_signatures_from_gatza2010a_study}
+
+First, the direction of Gatza's pathway associated genetic signatures were resolved
+
+\section{Pathway associated metagenes and obesity associated metagenes}
+\label{sec:pathway_associated_metagenes_and_obesity_associated_metagenes}
+
+then the pathway associated metagenes were compared with the obesity associated metagenes
+
+\section{Prediction of obesity associated metagene with pathway associate metagene}
+\label{sec:prediction_of_obesity_associated_metagene_with_pathway_associate_metagene}
+
+and lastly linear models were constructed based on the pathway metagenes and sample \gls{bmi}/\gls{bmi} status to predict the obesity associated metagenes.

### 2017-02-02 thesis/misc/appendix.tex
Mostly finished results1 - tmp commit

@@
 
+	\begin{table}[htbp]
+		\centering
+		\caption{Summary of the number of gene probes and genes in each obesity associated genetic signature}
+		\label{tab:signature_gene_num}
+		\begin{tabular}{lcc}
+			Obesity associated genetic signature & No. gene probes   & No. genes\\
+			\hline
+			\rule{0pt}{2.25ex}Original & 799 & 644\\
+			Cr                         & 799 & 678\\
+			Res                        & 799 & 655\\
+			Ca                         & 799 & 657\\
+			CaRes                      & 799 & 651\\
+			CrOl                       & 239 & 199\\
+			ResOl                      & 168 & 147\\
+			CaOl                       & 148 & 128\\
+			CaResOl                    & 92  & 86\\
+			\hline
+			\hline
+		\end{tabular}
+	\end{table}
+
 \end{appendices}

### 2017-02-02 thesis/results/results.tex
Mostly finished results1 - tmp commit

@@
 The obesity associated genetic signatures are central to this project in order to clarify the relationship between \gls{bmi} and cancer.
-Firstly in this chapter, the previously identified obesity associated signatures from the studies conducted by \citet{Creighton2012} and \citet{Fuentes-Mattei2014} were examined in turn to judge the agreement of these signatures with the sample \gls{bmi}, as presented in thier results.
-Secondly, novel obesity associated genetic signatures were identified in the \citet{Creighton2012} data set and was compared with the signatures.
+Firstly in this chapter, the previously identified obesity associated signatures from the studies conducted by \citet{Creighton2012} and \citet{Fuentes-Mattei2014} were examined in turn to judge the agreement of these signatures with the sample \gls{bmi}/\gls{bmi} status, as presented in thier results.
+Secondly, novel obesity associated genetic signatures were identified in the \citet{Creighton2012} data set and was compared with Creighton's and Fuentes-Mattei's  signatures.
 Lastly, the presence of common genes or pathways that were associated with obesity in multiple types of cancer was explored using the \gls{icgc} cancer data sets.
@@
 There was no description about the normalisation method used by Creighton's group when they first analysed their data, so the more reliable \gls{rma} method was used to  normalise the Creighton data set.
-The obesity metagene was plotted in a heatmap with the sample gene expression (\cref{fig:crmetaheat}) to check whether the metagene scores were in accordance with the overall gene expression of the samples.
+The obesity metagene was plotted in a heatmap with the sample gene expression  to check whether the metagene scores were in accordance with the overall gene expression of the samples (\cref{fig:crmetaheat}).
 
@@
 As shown in \cref{fig:crmetaheat}, high obesity associated metagene score of the sample reflected low expression in majority of the genes in the signature, and in contrast, low obesity associated metagene score of the sample reflected high expression in majority of the genes in the signature.
-This was consistent with the reported property of the obesity associated signature by \citet{Creighton2012}.
+This was consistent with the reported property of the obesity associated signature by \citet{Creighton2012} (see \cref{ssub:creighton_study}).
 To provide further evidence that the obesity associated metagenes were in fact associated with the \gls{bmi} status and \gls{bmi} value of the samples, a box plot and a scatter plot were created, respectively (\cref{fig:crmetaboxplot}).
@@
 
-\cref{fig:crmetaboxplot} clearly showed that the obesity associated metagene from \citet{Creighton2012} study significantly associated with the sample \gls{bmi} status, as well as the \gls{bmi} value of the samples (include p-value and r-squared).
-It should be noted here that the obesity associated metagene significantly associated with the samples that were obese, but not the samples that were overweight.
-This was due to the fact that the obesity associated genetic signatures were originally identified from the comparison of the samples that were obese with the samples that were not obese, and therefore the metagene scores were significant with obese group, but not the overweight group.
+\cref{fig:crmetaboxplot} clearly showed that the obesity associated metagene from \citet{Creighton2012} study significantly associated with the sample \gls{bmi} status, as well as sample \gls{bmi} value (include p-value and r-squared).
+It should be noted here that the obesity associated metagene significantly associated with the samples that were obese, but not with the samples that were overweight.
+This was due to the fact that the obesity associated genetic signatures were originally identified from the comparison of the samples that were obese with the samples that were not obese, and therefore the metagene scores were significant with the obese group, but not with the overweight group.
+In addition to this, although the regression line in the scatter plot showed statistically significant association with sample \gls{bmi}, the sample \gls{bmi} values seemed to be randomly dispersed across the metagene scores.
+This suggested that perhaps Creighton's obesity metagene was not associated with sample \gls{bmi} as strongly as expected.
 \\
@@
 \noindent
-Now that the association of the obesity metagene from \citet{Creighton2012} study was established in Creighton's data set, the obesity associated signature was transferred to the \gls{icgc} cancer data.
-The direction of the obesity associated metagene was checked in the Creighton's data, so that high metagene scores reflected high sample \gls{bmi} and low metagene scores reflected low sample \gls{bmi} (\cref{fig:crmetaboxplot}).
-The transformation matrix was created in Creighton's data, as described in \cref{sub:svd}.
+Now that the association of the obesity metagene from \citet{Creighton2012} study was established in Creighton's data set, the obesity associated signature was generated in the \gls{icgc} cancer data.
+The direction of the obesity associated metagene was checked in the Creighton's data first, so that high metagene scores reflected high sample \gls{bmi} and low metagene scores reflected low sample \gls{bmi} (\cref{fig:crmetaboxplot}).
+The transformation matrix was then created in Creighton's data, as described in \cref{sub:svd}.
 
 All of the \gls{icgc} data were normalised as described in \cref{ssub:rna_seq_data}.
-Before the transformation matrix was applied to the log$_{10}$-normalised cancer data, the suitability of standardised data or untouched (non-standardised) data was determined (appendix).
-From these results, the standardised data was the most suitable to use for matrix transformation.
-The transformation matrix was applied to each cancer data set in turn to obtain the obesity metagenes for each of the data set.
-Each obesity metagenes were plotted on a heatmap with the corresponding data set in which the metagene was taken from \cref{fig:crmetaicgc}.
+Before the transformation matrix was applied to the log$_{10}$-normalised cancer data, the suitability of standardised data or untouched (non-standardised) data was determined (see \cref{sec:metagenes_created_from_raw_data_vs_standardised_data_icgc}).
+From these results, the standardised data was found to be the most suitable to use for matrix transformation.
+The transformation matrix was applied to each cancer data set in turn to obtain Creighton's obesity metagenes for each of the data set.
+Each obesity metagenes were plotted in a heatmap with the corresponding data set in which the metagene was taken from (\cref{fig:crmetaicgc}; \cref{sec:rest_of_the_icgc_cancer_heatmap_resutls}).
 These heatmaps confirmed that the obesity associated metagene derived from Creighton's data was able to capture the overall gene expression pattern, where the metagene scores reflected the expression levels of majority of the genes in the signature.
-
-As before, association of the obesity metagene with the sample \gls{bmi} and \gls{bmi} status was checked in their respective cancer data set (\cref{fig:crmetaicgc}).
-In contrast to the strong association of the obesity metagene with sample \gls{bmi} and \gls{bmi} status in Creighton's data, the obesity metagene taken from the \gls{icgc} cancer data did not significantly associate with the sample \gls{bmi} or \gls{bmi} status in any of the \gls{icgc} cancer data.
+As before, association of the obesity metagene with the sample \gls{bmi} and \gls{bmi} status was checked in their respective cancer data set (\cref{fig:crmetaicgc}; \cref{sec:rest_of_the_icgc_cancer_heatmap_resutls}).
+Out of all the cancer types, only \gls{blca} data set showed significant association of Creighton's obesity metagene with the overweight group (but not with the obese group).
+However, neither the \gls{anova} p-value nor the regression line in the scatter plot were statistically significant, suggesting that this apparent association with the overweight group was not due to the Creighton's obesity metagene (?).
 
@@
 
-There could be several reasons for the apparent lack of association of the metagene with the sample \gls{bmi} or \gls{bmi} status.
+There could be several reasons for the apparent lack of association of the metagene with sample \gls{bmi}/\gls{bmi} status.
 First, transformation matrix was derived from Creighton's microarray data, but the \gls{icgc} cancer data were \gls{rnaseq} data.
-Though the log$_{10}$ normalisation and standardisation of the data was the most appropriate adjustment to be made to the \gls{rnaseq} data, this adjustment may not be equivalent to the \gls{rma} normalisation method that was used on the microarray data.
-Secondly, none of the \gls{icgc} cancer was from breast as in Creighton's data.
+Though the log$_{10}$ normalisation and standardisation of the data was the most appropriate adjustment to be made to the \gls{rnaseq} data, this adjustment was not equivalent to the \gls{rma} normalisation method that was used on the microarray data.
+Secondly, none of the \gls{icgc} cancer originated from breast as in Creighton's data.
 Since the obesity associated signature was identified in breast cancer data, the signature may be specific to breast cancer and may not be suitable in other cancer types.
@@
 \noindent
-To check whether the metagene was specific to breast cancer microarray data, the same transformation matrix was applied to Print's breast cancer microarray data.
-Print's data was normalised with \gls{rma} method as described in \cref{ssub:microarray_data}, and the transformation matrix was applied to the normalised data to obtain the metagene.
-The metagene was again compared with the gene expression of the samples with a heatmap (\cref{fig:crmetaprint}), and the association of the metagene with the sample \gls{bmi} and \gls{bmi} status was examined with box and scatter plots (\cref{fig:crmetaprint}).
+To check whether the metagene was specific to breast cancer microarray data, the same transformation matrix was applied to breast cancer microarray data from \citet{Print2016} study (referred to as Print's data).
+Print's data was normalised with \gls{rma} method and the transformation matrix was applied to the normalised data to obtain the metagene.
+The metagene was again compared with the gene expression of the samples with a heatmap and the association of the metagene with the sample \gls{bmi}/\gls{bmi} status was examined with box and scatter plots (\cref{fig:crmetaprint}).
 
 The obesity metagene managed to reflect the overall gene expression of the samples in Print's data.
-However, as in \gls{icgc} cancer data, the obesity metagene scores did not significantly associate with the sample \gls{bmi} or \gls{bmi} status.
-These results confirmed that the lack of association of the obesity metagene from Creighton's data was not due to the technology in which the data was gathered (mcroarray or \gls{rnaseq}), nor the type of cancer in which the transformation matrix was applied to.
+However, as with \gls{icgc} cancer data, the obesity metagene scores did not significantly associate with the sample \gls{bmi} or \gls{bmi} status.
+These results confirmed that the lack of association of the obesity metagene from Creighton's data was not due to the technology in which the data was gathered (mcroarray or \gls{rnaseq}), nor the cancer type in which the transformation matrix was applied to.
 
@@
 
-All together, all of these results suggest that the obesity metagene identified in the study conducted by \citet{Creighton2012} associated with sample \gls{bmi} and \gls{bmi} status in Creighton's data, but did not transfer well in other cancer data sets.
-The lack of association with \gls{bmi} was not due to the type of technology platform in which the data was gathered, as neither the \gls{icgc} \gls{rnaseq} data nor Print's microarray data did not show significant association with the obesity metagene.
+All together, all of these results suggest that the obesity metagene identified in the study conducted by \citet{Creighton2012} associated with sample \gls{bmi} and \gls{bmi} status only in Creighton's data.
+The lack of association with \gls{bmi} was not due to the type of technology platform in which the data was gathered, as neither the \gls{icgc} \gls{rnaseq} data nor Print's microarray data showed no significant association with the obesity metagene.
 Furthermore, Creighton's metagene was not dependent on the cancer type since the obesity associated genetic signature did not show significant association in neither the \gls{icgc} cancer data nor in Print's breast cancer data.
 
-One possible reason why Creighton's obesity metagene did not show significant association with other data sets could be because the genetic signature was in fact not an obesity specific signature, but a signature that detected another clinical variable.
+One possible reason why Creighton's obesity metagene did not show significant association with other data sets could be because the genetic signature was in fact not an obesity specific signature, but a signature that was detected due to another clinical variable.
 Another reason for this apparent lack of association could be that the genetic signature was too specific to the data and was not a broad obesity associated genetic signature, but an obesity associated signature specifically for Creighton's data set.
@@
 Firstly, FM's data was normalised with the \gls{rma} method and \gls{svd} was applied to the normalised data to get the transformation matrix.
-The transformation matrix was used to transform the \gls{rma}-normalised Creighton's data to extract FM's obesity metagene scores.
+The transformation matrix was used to transform the \gls{rma}-normalised Creighton's data to extract FM's obesity metagene scores in Creighton's data.
 FM's obesity metagene scores were compared with gene expressions and sample \gls{bmi}/\gls{bmi} status in Creighton's data, as shown in \cref{fig:fmmetacr}.
@@
 Next, the transformation matrix was applied to the \gls{icgc} cancer data and resulting metagenes were compared with the gene expression and sample \gls{bmi}/\gls{bmi} status.
-As evident in \cref{fig:fmmetaicgc}, FM's obesity metagene scores appeared to reflect the gene expression of FM's obesity associated genetic signature.
-However, as with all the results in this section so far, FM's obesity metagene did not significantly associate with any of the \gls{icgc} cancer data (\cref{fig:fmmetaicgc}).
+As evident in \cref{fig:fmmetaicgc}, FM's obesity metagene scores appeared to reflect the overall gene expression of FM's obesity associated genetic signature.
+As with all the results in this chapter so far, FM's obesity metagene did not significantly associate with any of the \gls{icgc} cancer data, except in \gls{blca} data set (\cref{fig:fmmetaicgc}).
+FM's obesity metagene significantly associated with the overweight group (but not with the obese group), and also had a significant \gls{anova} p-value.
+On the contrary to the association of the metagene with the sample \gls{bmi} status, FM's obesity metagene was not associated with sample \gls{bmi}.
+These results suggested that the samples that were overweight in the \gls{blca} data set had similar biological properties as the samples that were obese in FM's data set.
+However, due to the fact that FM's metagene lacked association with sample \gls{bmi} in \gls{blca} data set and that the metagene did not show any significant association in any other cancer type, it was difficult to determine whether the observed association of FM's obesity metagene with the overweight group of samples was truly reflective of the effect resulted from FM's metagene.
+Lastly, the transformation matrix was applied to Print's data (\cref{fig:fmmetacris}).
+Again, FM's obesity metagene scores reflected the gene expression of the samples, did not significantly associate with sample \gls{bmi}/\gls{bmi} status.
+\\
 
@@
 \end{figure}
-
-Lastly, the transformation matrix was applied to Print's data (\cref{fig:fmmetacris}).
-Again, FM's obesity metagene scores reflected the gene expression of the samples, did not significantly associate with sample \gls{bmi}/\gls{bmi} status.
-\\
 
@@
 These results showed that FM's obesity associated metagene was not transferable to other cancer data sets, similar to Creighton's obesity associated metagene.
-This meant that neither Creighton's nor FM's obesity associate metagenes may be too specific to the orignial data set in which the signature was identified in.
-Furthermore, it may be possible that these obesity associated metagenes may not be related to obesity, but another clinical variable which may be closely related to \gls{bmi}.
+This meant that both Creighton's and FM's obesity associate metagenes may have been too specific to the orignial data set in which the signatures were identified in.
+Furthermore, there was a possibility that these obesity associated metagenes were not related to obesity, but associated with a different clinical variable that may be closely related to \gls{bmi}.
 
@@
 
-Both metagenes generated from Creighton's data and FM's data were able to capture the overall pattern of gene expression of the samples, but did not associate with sample \gls{bmi}/\gls{bmi} status.
-One possible reason for this result could be that the obesity associated genetic signatures from Creighton's and FM's study may not have been truly associated with sample \gls{bmi} status, but another clinical variable.
-To remove this possibility, obesity associated genetic signatures were identified in Creighton's data after controlling for all the clinical variables in the data set.
+Both the obesity associated metagenes generated from Creighton's data and FM's data were able to capture the overall pattern of gene expression of the genetic signature in the samples, but did not associate with sample \gls{bmi}/\gls{bmi} status.
+One possible reason for this result could be that the obesity associated genetic signatures from Creighton's and FM's study may not have been truly associated with sample \gls{bmi} status, but with another clinical variable.
+To reject this possibility, obesity associated genetic signatures were identified in Creighton's data after controlling for all the clinical variables in the data set.
 FM's data set was not used to get obesity associated genetic signatures, as no \gls{bmi} information was available for the samples in FM's data.
@@
 	\centering
-	\includegraphics[page=1,width=0.8\linewidth]{results1/deg_venn}
+	\includegraphics[width=0.5\linewidth]{results1/deg_venn1}
 	\caption[Summary of the \glspl{deg} identified from Creighton's data]{(add venn diagram)}
@@
 Therefore, the effect of ethnicity was completely removed by considering only the Caucasian patients in Creighton's data, which left a total of 77 Caucasian patients in the data set.
-With this Caucasian-only data set, obesity associated genetic signatures were identified as described above; first with unadjusted data, then with the data with clinical variables adjusted (except ethnicity, as it has already been controlled for), and lastly the common genes with Creighton's original signature were identified.
+With this Caucasian-only data set, obesity associated genetic signatures were identified as described above; first with unadjusted data, then with the data with clinical variables adjusted (except ethnicity, as it has already been controlled for by considering Caucasian samples only), and lastly the common genes with Creighton's original signature were identified.
 
-2129 gene probes and 1558 gene probes were identified in the unadjusted and clinical variable-adjusted Caucasian patient data, respectively.
+2129 and 1558 gene probes were identified in the unadjusted and clinical variable-adjusted Caucasian patient data, respectively.
 As before, the most significant 799 gene probes were selected from these gene probes.
@@
 	\centering
-	\includegraphics[page=2,width=0.8\linewidth]{results1/deg_venn}
+	\includegraphics[width=0.5\linewidth]{results1/deg_venn2}
 	\caption[Summary of the \glspl{deg} identified from Creighton's data (Caucasian patient data)]{(add venn diagram)}
@@
 
-As with the original signature, all eight signatures were validated in Creighton's data set first, and then compared in other cancer data sets by using the transformation matrix generated in Creighton's data.
+As with the original signature, all eight obesity associated genetic signatures were validated in Creighton's data set first, and then compared in other cancer data sets by using the transformation matrix generated in Creighton's data.
 % Before the metagenes were created, all of the signatures were converted into gene symbols, as described in \cref{sec:data} (\cref{tab:signature_gene_num}).
@@
 The association of the metagenes with sample \gls{bmi}/\gls{bmi} status was significant for all eight of the metagenes identified (\cref{fig:degmetacr}).
-These results confirmed that all of the obesity associated genetic signatures identified in \cref{sub:identification_of_obesity_associated_genetic_signatures} significantly associated with sample \gls{bmi}/\gls{bmi} status in the data in which the metagenes were derived from.
+These results confirmed that all of the obesity associated genetic signatures identified in \cref{sub:identification_of_obesity_associated_genetic_signatures} significantly associated with sample \gls{bmi}/\gls{bmi} status in Creighton's data, where the metagenes were derived from.
 
@@
 	\label{fig:degmetacr}
+\end{figure}
+
+Additionally, the correlation of all of the obesity associated metagenes were examined to see whether these metagenes were similar to one another.
+As clearly shown in \cref{fig:cr_meta_cor}, there were two distinct groups within the eight metagenes; the first group contained the metagenes that were not overlapped with the original metagene (Cr, Res, Ca and CaRes), while the other group had the overlapped metagenes (CrOl, ResOl, CaOl and CaResOl).
+With that said, all eight metagenes showed high correlation with one another (lowest correlation approximately at 0.85), which suggested that all of these metagenes were detecting similar underlying biological mechanism from the data.
+
+\begin{figure}[htpb]
+	\centering
+	\includegraphics[width=0.8\linewidth]{results1/cr_meta_cor}
+	\caption[Pearson correlation of all eight obesity associated metagenes identified in Creighton's data]{Heatmap showing the Pearson correlation of all eight obesity associated metagenes from Creighton's data with one another.
+	\gls{svd} was applied to \gls{rma}-normalised Creighton's data to generate each of the eight metagenes.
+	High and low correlation were represented as red and blue, respectively, where the colours were matched with the values on the scale shown in top right.}
+	\label{fig:cr_meta_cor}
 \end{figure}
@@
 
-% \begin{table}[htbp]
-% 	\centering
-% 	\caption{Summary of the number of gene probes and genes in each obesity associated genetic signature}
-% 	\label{tab:signature_gene_num}
-% 	\begin{tabular}{lcc}
-% 		Obesity associated genetic signature & No. gene probes   & No. genes\\
-% 		\hline
-% 		\rule{0pt}{2.25ex}Cr & 799 & 678\\
-% 		Res                  & 799 & 655\\
-% 		Ca                   & 799 & 657\\
-% 		CaRes                & 799 & 651\\
-% 		CrOl                 & 239 & 199\\
-% 		ResOl                & 168 & 147\\
-% 		CaOl                 & 148 & 128\\
-% 		CaResOl              &  92 & 86\\
-% 		\hline
-% 		\hline
-% 	\end{tabular}
-% \end{table}
-
 \section{Common genes across multiple cancer types}
@@
 	\begin{tabular}{lc}
-		Cancer type & No. of \glspl{deg}\\
+		Cancer type & No. of \glspl{deg} identified\\
 		\hline
@@
 
-(insert description of the simulation results).
-The results from the simulation showed that the \glspl{deg} identified from the gene expression analysis of \gls{icgc} cancer data were significant.
-However, the results also confirmed that there were no obesity associated \glspl{deg} that were common across different cancer types.
+The results from the simulation showed that, on average, 5732 genes were found to be differentially expressed by a cancer type, 1057 by two cancer types, 111 by three cancer types, and 7 by four cancer types.
+When the results from \gls{icgc} gene expression analysis were compared with the simulation results, the number of \glspl{deg} found were statistically significant, as the numbers of genes identified exceeded the 95$^{th}$ percentile values for up to 5 cancer types.
+This result confirmed that the \glspl{deg} from the \gls{icgc} cancer data were not identified by chance.
+Unfortunately, there were no \glspl{deg} expressed in all eight cancer types, which confirmed that there was no common genes that were differentially expressed between the samples that were obese and normal weight.
 
@@
 		\begin{tabular}{>{\quad}lcccccccc}
-			& \multicolumn{8}{c}{No. of cancer types expressing the \glspl{deg}}\\
+			& \multicolumn{8}{c}{\small No. of cancer types expressing the \glspl{deg}\tnote{1}}\\
 			& 1 & 2 & 3 & 4 & 5 & 6 & 7 & 8\\
 			\hline
-			\rule{0pt}{2.25ex}\hspace{-1em}Results from gene expression analysis & 7330 & 2024 & 320 & 21 & 0 & 0 & 0 & 0  \\
-			\hspace{-1em}Results from simulation:                                &      &      &     &    &   &   &   &  \\
-			Mean no. of \glspl{deg} identified                                   & 0    & 0    & 0   & 0  & 0 & 0 & 0 & 0  \\
-			$95^{th}$ percentile                                                 & 0    & 0    & 0   & 0  & 0 & 0 & 0 & 0  \\
+			\rule{0pt}{2.25ex}\hspace{-1em}{\small Results from gene expression analysis} & 7330 & 2024 & 320 & 21 & 0 & 0 & 0 & 0 \\
+			\hspace{-1em}{\small Results from simulation:\tnote{2}}                       &      &      &     &    &   &   &   &   \\
+			{\small Mean no. of \glspl{deg} identified}                                   & 5732 & 1057 & 111 & 7  & 0 & 0 & 0 & 0 \\
+			{\small $95^{th}$ percentile}                                                 & 6965 & 1722 & 227 & 20 & 2 & 0 & 0 & 0 \\
 			\hline
@@
 		\end{tabular}
-		% \begin{tablenotes}
-		% 	\begin{footnotesize}
-		% 	\item [1] The numbers represent the number of cancer types a gene was expressed in.
-		% 	\item [2] The simulation was repeated 1000 times.
-		% 	\end{footnotesize}
-		% \end{tablenotes}
+		\begin{tablenotes}
+			\begin{footnotesize}
+			\item [1] The numbers represent the number of cancer types a gene was expressed in.
+			\item [2] The simulation was repeated 1000 times, each with randomised samples.
+			\end{footnotesize}
+		\end{tablenotes}
 	\end{threeparttable}
@@
 To check whether there were any significant pathways enriched in any of the \gls{icgc} cancer data sets that were associated with obesity, pathway enrichment analysis was carried out on each cancer type separately and then all cancer types combined (see \cref{sec:pathway_enrichment_analysis}).
+
+(unfinished)

### 2017-02-02 thesis/results/results.tex
tmp commit

@@
 	\includegraphics[width=0.5\linewidth]{results1/deg_venn1}
-	\caption[Summary of the \glspl{deg} identified from Creighton's data]{(add venn diagram)}
+	\caption[Veen diagram of the \glspl{deg} identified from Creighton's data (all samples)]{Venn diagram showing the common genes between the signatures obtained from unadjusted and clinical variable-adjusted Creighton's data and Creighton's original obesity associated genetic signature.}
 	\label{fig:venn1}
@@
 	\includegraphics[width=0.5\linewidth]{results1/deg_venn2}
-	\caption[Summary of the \glspl{deg} identified from Creighton's data (Caucasian patient data)]{(add venn diagram)}
+	\caption[Summary of the \glspl{deg} identified from Creighton's data (Caucasian patient data)]{Venn diagram showing the common genes between the signatures obtained from unadjusted and clinical variable-adjusted Creighton's data (Caucasian patients only) and Creighton's original obesity associated genetic signature.}
 	\label{fig:venn2}
@@
 To confirm whether these metagenes showed significant association with sample \gls{bmi}/\gls{bmi} status in other cancer data sets, transformation matrix was applied to the \gls{icgc} cancer data and Print's breast cancer data.
-Firstly in \gls{icgc} cancer data sets, all eight metagene scores were reflective of the sample gene expression of the genetic signatures (\cref{fig:degmetaicgc}), but none of the metagenes associated significantly with sample \gls{bmi}/\gls{bmi} status (\cref{fig:degmetaicgc}).
-Likewise in Print's data, the higher the metagenes scores were the more expressed the genes were (and vice versa; \cref{fig:degmetaprint}), but lacked the association with sample \gls{bmi}/\gls{bmi} status (\cref{fig:degmetaprint}).
+Firstly in \gls{icgc} cancer data sets, all eight metagene scores were reflective of the sample gene expression of the genetic signatures (\cref{fig:degmetaicgc}).
+In all cancer types but \gls{blca}, none of the metagenes significantly associated with the sample \gls{bmi}/\gls{bmi} status (results shown in \cref{app:a}).
+As shown in \cref{fig:degmetaicgc}, only the \gls{blca} data set showed significant association of the Cr metagene with sample \gls{bmi}/\gls{bmi} status.
+With the metagenes other than Cr metagene, Res, CrOl and Ca metagenes showed statistical significance in overweight group, \gls{anova} and the regression line p-values; CaRes metagene was significant in overweight group and \gls{anova} p-value; CrOl and ResOl metagenes were significant in overweight group; and ResOl and CaOl metagenes were not significant in \gls{blca} data set.
 
+This was unexpected since all of the genetic signatures were results of gene expression analysis between the non-obese group and the obese group of samples in Creighton's data, and yet the metagenes showed significant association with the overweight group in \gls{blca} data set.
+These results suggested that the samples that were overweight in \gls{blca} data were similar in phenotype with the samples that were obese in Creighton's data.
+Again, due to the fact that all of the metagenes lacked association with sample \gls{bmi}/\gls{bmi} status in all other \gls{icgc} data set, it was difficult to conclude whether the observed association of many of the obesity metagenes with the overweight group of samples was truly reflective of the effect resulted from these metagenes.
+
+\begin{table}[htpb]
+	\centering
+	\caption{P-values of all the metagenes in \gls{blca} cancer data}
+	\label{tab:degmetablca}
+	\begin{threeparttable}
+		\begin{tabular}{lccccc}
+			& \multicolumn{3}{c}{ P-values} & \multicolumn{2}{c}{ Regression line statistics}\\
+			\cmidrule(r){2-4} \cmidrule(r){5-6}
+			 Metagenes &  Overweight &  Obese &  \gls{anova} &  R$^2$ &  P \\
+			\hline
+			\rule{0pt}{2.25ex}Cr & {\bfseries 0.0134}\tnote{1} & 0.1365 & {\bfseries 0.0387} & 0.0171 & {\bfseries 0.0195} \\
+			Res                  & {\bfseries 0.0055}          & 0.1974 & {\bfseries 0.0196} & 0.0117 & {\bfseries 0.0446} \\
+			CrOl                 & {\bfseries 0.0383}          & 0.4583 & 0.1096             & 0.0047 & 0.1374             \\
+			ResOl                & 0.0909                      & 0.7092 & 0.2125             & 0.0018 & 0.2254             \\
+			Ca                   & {\bfseries 0.0077}          & 0.1973 & {\bfseries 0.0231} & 0.0116 & {\bfseries 0.0456} \\
+			CaRes                & {\bfseries 0.0104}          & 0.2712 & {\bfseries 0.0322} & 0.0101 & 0.0572             \\
+			CaOl                 & 0.0575                      & 0.6185 & 0.1487             & 0.0022 & 0.2126             \\
+			CaResOl              & {\bfseries 0.0463}          & 0.5820 & 0.1263             & 0.0036 & 0.1660             \\
+			\hline
+			\hline
+		\end{tabular}
+		\begin{tablenotes}
+			\item [1] All values in bold are statistically significant (p \textless{} 0.05).
+		\end{tablenotes}
+	\end{threeparttable}
+\end{table}
+
+In Print's data, the higher the metagenes scores were the more expressed the genes were (and vice versa), but lacked the association with sample \gls{bmi}/\gls{bmi} status (\cref{fig:degmetaprint}).
 These results showed that, even though all of the metagenes significantly associated with sample \gls{bmi}/\gls{bmi} status in the data the genetic signatures were derived from, the metagenes did not transfer well in other cancer data sets.
 Furthermore, these results showed that the lack of association with sample \gls{bmi}/\gls{bmi} status were not due to other clinical variables in Creighton's data set.
-This raised a question of whether there are any obesity associated genetic signature that are common in multiple cancer types.
+This raised a question of whether there was any obesity associated genetic signature that was common in multiple cancer types.
 
@@
 
-\section{Pathways enriched in obesity associated genes}
+\section{Pathways enriched in obesity associated genes (haven't finished)}
 \label{sec:pathways_enriched_in_obesity_associated_genes}
@@
 To check whether there were any significant pathways enriched in any of the \gls{icgc} cancer data sets that were associated with obesity, pathway enrichment analysis was carried out on each cancer type separately and then all cancer types combined (see \cref{sec:pathway_enrichment_analysis}).
-
-(unfinished)

### 2017-02-07 thesis/results/results2.tex
finished first section of results2 - tmp commit

@@
 
-First, the direction of Gatza's pathway associated genetic signatures were resolved
+Pathway associated genetic signatures from the \citet{Gatza2010a} study were examined for their consistency with the reported results.
+In the \citet{Gatza2010a} study, their data comprised of samples from different data sets from other studies, \gls{mas}-normalised and the metagene scores were ranked with probit.
+However, the analyses so far have used \gls{rma}-normalisation method and ranked based on the number of samples present in the data.
+To decide which normalisation or ranking methods were suitable for the analysis, the different methods were compared in Gatza's data (see \cref{app:b}).
+Results shown in \cref{app:b} clarified that there was no significant difference in the ranking methods used.
+The same went for the normalisation methods, as long as the normalisation methods used to get the transformation matrix and the data in which the transformation matrix was applied to were the same.
+For example, transformation matrix was taken from \gls{rma}-normalised data and then applied to \gls{rma}-normalised data set.
+From these results, Gatza's data set was batch corrected first (\cref{sub:batch_correction}), then normalised with \gls{rma} method and the scores were ranked based on the number of samples in the data set.
+
+Before Gatza's pathway metagenes were compared with the obesity associated metagenes, the direction of Gatza's pathway metagenes were checked to make sure the metagenes were in the correct direction (\cref{sub:metagene_direction}).
+Each pathway metagenes was generated from Gatza's data set and the direction of the metagene was checked with the expression of the representative  pathway gene (\cref{tab:metagene_direction}).
+The correlation of all the pathway metagenes with one another were plotted as a heatmap in \cref{fig:gatza_meta_dir}.
+% The pathway metagenes showed a few distinct groups of pathways: a group that showed high correlation with E2F1, \gls{pi3k}, Myc, \gls{bcat} and Ras pathways; a group with \gls{ifna}, \gls{ifny} and \gls{tnfa} pathways;
+The most prominent group had five pathways (E2F1, \gls{pi3k}, Myc, \gls{bcat} and Ras) that clustered at the top right hand corner of the heatmap.
+% TODO: add additional info about why these pathways clustered together?
+Other groups included \gls{ifna}/\gls{ifny}/\gls{tnfa} pathways, \gls{er}/\gls{pr}/p53 pathway and p63/\gls{her2} pathways.
+In addition to these highly correlated groups, \gls{stat3}/\gls{tgfb}/Src/\gls{egfr}/Akt pathways showed little correlation with one another.
+Comparing these groups with the results presented by \citet{Gatza2010a} (see \cref{app:b}), the identified groups approximately resembled the groups identified in their study, which confirmed that the directions of Gatza's pathway metagenes were similar to those used in the \citet{Gatza2010a} study.
+
+\begin{figure}[htpb]
+	\centering
+	\includegraphics[page=15,width=0.7\linewidth]{results2/gatza_meta_trans}
+	\caption{Gatza metagene direction}
+	\label{fig:gatza_meta_dir}
+\end{figure}
+
+To see whether the directionality of Gatza's pathway metagenes were being transferred across to other data sets properly, pathway metagenes were generated with the transformation matrices and the groupings of the metagenes were examined.
+Creighton's, FM's and Cris' data were \gls{rma}-normalised and transformation matrices of the pathway genetic signatures were applied to the data.
+The metagenes were plotted in a heatmap (shown in \cref{app:b}).
+These heatmaps showed similar groupings  as seen in \cref{fig:gatza_meta_dir}, which proved that the directions of the pathway metagenes were independent of the data sets.
+
+% In order to investigate what the underlying biological mechanism of the obesity associated genetic signatures was, Gatza's pathway associated genetic signatures were used to create linear models to predict the obesity associated metagene scores.
+% Before linear models were created, Gatza's pathway metagenes were checked for their directionality to ensure that the pathway metagenes were acting as reported by \citet{Gatza2010a} (see \cref{sub:metagene_direction}).
 
@@
 \label{sec:pathway_associated_metagenes_and_obesity_associated_metagenes}
+
+Results from previous section confirmed that the pathway metagenes from the \citet{Gatza2010a} study were in the correct directions.
 

### 2017-02-07 thesis/results/results2.tex
tmp commit

@@
 The same went for the normalisation methods, as long as the normalisation methods used to get the transformation matrix and the data in which the transformation matrix was applied to were the same.
-For example, transformation matrix was taken from \gls{rma}-normalised data and then applied to \gls{rma}-normalised data set.
+For example, transformation matrix taken from \gls{rma}-normalised data applied to \gls{rma}-normalised data set.
+With that said, \gls{rma}-normalised data sets had a more consistent correlation of the \gls{svd}-generated metagenes with the transformation matrix-generated metagenes across different data sets than the \gls{mas}-normalised data sets (see \cref{app:b}; more details in \cref{sec:pathway_associated_metagenes_and_obesity_associated_metagenes}).
 From these results, Gatza's data set was batch corrected first (\cref{sub:batch_correction}), then normalised with \gls{rma} method and the scores were ranked based on the number of samples in the data set.
 
-Before Gatza's pathway metagenes were compared with the obesity associated metagenes, the direction of Gatza's pathway metagenes were checked to make sure the metagenes were in the correct direction (\cref{sub:metagene_direction}).
-Each pathway metagenes was generated from Gatza's data set and the direction of the metagene was checked with the expression of the representative  pathway gene (\cref{tab:metagene_direction}).
+Before Gatza's pathway metagenes were compared with the obesity associated metagenes, the direction of Gatza's pathway metagenes had to be checked to make sure the metagenes were in the correct direction (\cref{sub:metagene_direction}).
+% Each pathway metagenes was generated from Gatza's data set and the direction of the metagene was checked with the expression of the representative pathway gene (\cref{tab:metagene_direction}), and the groupings of the pathways were considered as well.
 The correlation of all the pathway metagenes with one another were plotted as a heatmap in \cref{fig:gatza_meta_dir}.
-% The pathway metagenes showed a few distinct groups of pathways: a group that showed high correlation with E2F1, \gls{pi3k}, Myc, \gls{bcat} and Ras pathways; a group with \gls{ifna}, \gls{ifny} and \gls{tnfa} pathways;
 The most prominent group had five pathways (E2F1, \gls{pi3k}, Myc, \gls{bcat} and Ras) that clustered at the top right hand corner of the heatmap.
@@
 
-To see whether the directionality of Gatza's pathway metagenes were being transferred across to other data sets properly, pathway metagenes were generated with the transformation matrices and the groupings of the metagenes were examined.
-Creighton's, FM's and Cris' data were \gls{rma}-normalised and transformation matrices of the pathway genetic signatures were applied to the data.
-The metagenes were plotted in a heatmap (shown in \cref{app:b}).
-These heatmaps showed similar groupings  as seen in \cref{fig:gatza_meta_dir}, which proved that the directions of the pathway metagenes were independent of the data sets.
+To see whether the directionality of Gatza's pathway metagenes were being transferred across to other data sets properly, pathway metagenes were generated in other data sets with the transformation matrices and the groupings of the metagenes were examined.
+Creighton's, FM's and Cris' data were \gls{rma}-normalised and transformation matrices of the pathway genetic signatures (derived from Gatza's data set) were applied to the data sets.
+The metagenes were plotted in a heatmap (shown in \cref{app:b}), which showed similar groupings  as seen in \cref{fig:gatza_meta_dir}.
+This result confirmed that the pathway metagenes were acting similarly in all the data sets in which the transformation matrices have been applied to.
 
@@
 
-Results from previous section confirmed that the pathway metagenes from the \citet{Gatza2010a} study were in the correct directions.
+Results from previous section confirmed that the pathway metagenes from the \citet{Gatza2010a} study were in the correct directions, and that the pathway metagenes were behaving as expected in different data sets.
+Now that the directions of the pathway metagenes were established, these metagenes were ready for comparison with the obesity associted metagenes.
+However, all of the pathway metagenes were derived from Gatza's data, whereas the majority of obesity associted metagenes were derived from Creighton's data, which presents a problem of deciding on which data set the transformation matrices should be generated in.
 
-then the pathway associated metagenes were compared with the obesity associated metagenes
+In theory, the metagenes generated from the application of \gls{svd} and the metagenes generated from transformation matrix would be exactly the same if the transformation matrix was derived from the same data set.
+For example, an obesity associted metagene generated from Creighton's data with \gls{svd} would have the same values as the metagenes generated in Creighton's data with the transformation matrix (that was derived in Creighton's data).
+Furthermore, if the \gls{svd}-derived metagenes and transformation matrix-derived metagenes were the same (or at least similar) in a different data set, this suggests that there is no difference in the data sets, at least in terms of the expression of the genetic signatures being investigated.
+In other words, the transformation matrix can be made in either data sets, in the original data set or in a different data set, if there is no difference in the \gls{svd}-derived or transformation matrix-derived metagenes.
 

### 2017-02-12 thesis/results/results2.tex
tmp commit

@@
 Results shown in \cref{app:b} clarified that there was no significant difference in the ranking methods used.
-The same went for the normalisation methods, as long as the normalisation methods used to get the transformation matrix and the data in which the transformation matrix was applied to were the same.
-For example, transformation matrix taken from \gls{rma}-normalised data applied to \gls{rma}-normalised data set.
-With that said, \gls{rma}-normalised data sets had a more consistent correlation of the \gls{svd}-generated metagenes with the transformation matrix-generated metagenes across different data sets than the \gls{mas}-normalised data sets (see \cref{app:b}; more details in \cref{sec:pathway_associated_metagenes_and_obesity_associated_metagenes}).
-From these results, Gatza's data set was batch corrected first (\cref{sub:batch_correction}), then normalised with \gls{rma} method and the scores were ranked based on the number of samples in the data set.
+Correlation and scatter plots of the different combinations of transformation matrices (derived from either \gls{rma}- or \gls{mas}-normalised data) with \gls{rma}- or \gls{mas}-normalised data.
+It seemed like the normalisation methods of the data in which the transformation matrices were applied to had the most significant effect on the resulting metagenes, rather than the transformation matrices themselves (\cref{app:b}).
+This meant that all of the data sets had to be normalised with a single normalisation method, so \gls{rma} normalisation method was used as this method was more reliable than the \gls{mas} method.
+Therefore, Gatza's data set was batch corrected first (\cref{sub:batch_correction}), then normalised with \gls{rma} method and the scores were ranked based on the number of samples in the data set.
 
+One thing to be noted from these scatter plots was that there were some pathway signatures that were more variable than the others.
+As an example, the \gls{tgfb} metagenes were significantly more dispersed compared to the \gls{pr} metagenes, even though these metagenes were both generated similarly in the same data set (add figure).
+These differences were seen in other data sets as well (\cref{app:b}).
+This result provided evidence that some of the pathway genetic signatures from the \citet{Gatza2010a} study were more consistent across different data sets than the others.
+\\
+
+\noindent
 Before Gatza's pathway metagenes were compared with the obesity associated metagenes, the direction of Gatza's pathway metagenes had to be checked to make sure the metagenes were in the correct direction (\cref{sub:metagene_direction}).
@@
 Now that the directions of the pathway metagenes were established, these metagenes were ready for comparison with the obesity associted metagenes.
-However, all of the pathway metagenes were derived from Gatza's data, whereas the majority of obesity associted metagenes were derived from Creighton's data, which presents a problem of deciding on which data set the transformation matrices should be generated in.
+However, all of the pathway metagenes were derived from Gatza's data, whereas the majority of the obesity associted metagenes were derived from Creighton's data, which presents a problem of deciding on which data set the transformation matrices should be generated in.
 
 In theory, the metagenes generated from the application of \gls{svd} and the metagenes generated from transformation matrix would be exactly the same if the transformation matrix was derived from the same data set.
-For example, an obesity associted metagene generated from Creighton's data with \gls{svd} would have the same values as the metagenes generated in Creighton's data with the transformation matrix (that was derived in Creighton's data).
-Furthermore, if the \gls{svd}-derived metagenes and transformation matrix-derived metagenes were the same (or at least similar) in a different data set, this suggests that there is no difference in the data sets, at least in terms of the expression of the genetic signatures being investigated.
-In other words, the transformation matrix can be made in either data sets, in the original data set or in a different data set, if there is no difference in the \gls{svd}-derived or transformation matrix-derived metagenes.
+For example, an obesity associted metagene generated from Creighton's data with \gls{svd} would have the same values as the metagenes generated in Creighton's data with the transformation matrix (where the transformation matrix was derived from Creighton's data).
+Furthermore, if the \gls{svd}-derived metagenes and transformation matrix-derived metagenes were the same (or at least similar) in a different data set, this suggests that there was no difference in the data sets, at least in terms of the expression of the genetic signatures being investigated.
+In other words, the transformation matrix can be made in either the original data set or in a different data set, if there is no difference in the \gls{svd}-derived or transformation matrix-derived metagenes.
+
+% TODO: start from here
+To decide which data set the transformation matrices for obesity and pathway associated genetic signatures should be generated in, the \gls{svd}- and transformation matrix(TM)-generated metagene scores were compared in all of the data sets.
+As in \cref{sec:pathway_associated_genetic_signatures_from_gatza2010a_study}, all data sets were normalised with the \gls{rma} method and metagenes were ranked based on the number of samples in each data set.
+Transformation matrices for the obesity associted genetic signatures were made in Creighton's data set and pathway associted genetic signatures were made in Gatza's data set.
+Metagenes for all of the obesity and pathway associated genetic signatures were generated in all of the data sets with \gls{svd} and transformation matrices.
+The Spearman correlation of the \gls{svd}-generated and TM-generated metagene scores were calculated for each genetic signatures.
+
+\begin{table}[htpb]
+	\centering
+	\caption{Spearman correlation of the svd vs tm metagenes}
+	\label{tab:svd_vs_tm}
+	\begin{tabular}{c}
+		gtrmatransres (RMA)
+		\hline
+		FM & Creighton   &     Cris
+		akt   -0.5563 -0.5190 -0.5900
+		bcat   0.9905  0.9897  0.9977
+		e2f1   0.9193  0.9646  0.8438
+		egfr  -0.4040 -0.0430 -0.3358
+		er     0.9966  0.9978  0.9942
+		her2  -0.9794 -0.9553 -0.5817
+		ifna   0.9951 -0.9830 -0.9991
+		ifng   0.9718 -0.9086  0.9950
+		myc    0.9689  0.9878 -0.9852
+		p53    0.7923  0.3808  0.9981
+		p63    0.8368  0.8319  0.2951
+		pi3k   0.9365 -0.9543  0.5989
+		pr     0.9845 -0.9511 -0.9887
+		ras   -0.7229 -0.9078 -0.9125
+		src    0.9548 -0.9575 -0.7173
+		stat3  0.6167 -0.1902  0.9159
+		tgfb   0.9890 -0.9918 -0.2543
+		tnfa   0.4315 -0.6046  0.9365
+
+			   obsrmatransres (RMA)
+		   \hline
+		   Gatza        FM       Cris
+		rawobs     0.9985 0.9872 -0.9161
+		crol       0.9982 0.9926  0.9715
+		resobs    -0.9987 0.9898 -0.9652
+		rescrol    0.9981 0.9927  0.9571
+		caobs      0.9985 0.9893 -0.9468
+		cacrol     0.9983 0.9937  0.9677
+		caresobs  -0.9988 0.9939  0.9865
+		carescrol  0.9984 0.9952  0.9642
+		crobs     -0.9928 0.9862 -0.9344
+
+	\end{tabular}
+\end{table}
 

### 2017-02-13 thesis/results/results2.tex
tmp commit

@@
 Results from previous section confirmed that the pathway metagenes from the \citet{Gatza2010a} study were in the correct directions, and that the pathway metagenes were behaving as expected in different data sets.
-Now that the directions of the pathway metagenes were established, these metagenes were ready for comparison with the obesity associted metagenes.
-However, all of the pathway metagenes were derived from Gatza's data, whereas the majority of the obesity associted metagenes were derived from Creighton's data, which presents a problem of deciding on which data set the transformation matrices should be generated in.
+Now that the directions of the pathway metagenes were established, these metagenes were ready for comparison with the obesity associated metagenes.
+However, all of the pathway metagenes were derived from Gatza's data, whereas the majority of the obesity associated metagenes were derived from Creighton's data, which presents a problem of deciding on which data set the transformation matrices should be generated in.
 
 In theory, the metagenes generated from the application of \gls{svd} and the metagenes generated from transformation matrix would be exactly the same if the transformation matrix was derived from the same data set.
-For example, an obesity associted metagene generated from Creighton's data with \gls{svd} would have the same values as the metagenes generated in Creighton's data with the transformation matrix (where the transformation matrix was derived from Creighton's data).
+For example, an obesity associated metagene generated from Creighton's data with \gls{svd} would have the same values as the metagenes generated in Creighton's data with the transformation matrix (where the transformation matrix was derived from Creighton's data).
 Furthermore, if the \gls{svd}-derived metagenes and transformation matrix-derived metagenes were the same (or at least similar) in a different data set, this suggests that there was no difference in the data sets, at least in terms of the expression of the genetic signatures being investigated.
@@
 As in \cref{sec:pathway_associated_genetic_signatures_from_gatza2010a_study}, all data sets were normalised with the \gls{rma} method and metagenes were ranked based on the number of samples in each data set.
-Transformation matrices for the obesity associted genetic signatures were made in Creighton's data set and pathway associted genetic signatures were made in Gatza's data set.
+Transformation matrices for the obesity associated genetic signatures were made in Creighton's data set and pathway associated genetic signatures were made in Gatza's data set.
 Metagenes for all of the obesity and pathway associated genetic signatures were generated in all of the data sets with \gls{svd} and transformation matrices.
-The Spearman correlation of the \gls{svd}-generated and TM-generated metagene scores were calculated for each genetic signatures.
+The Spearman correlation of the \gls{svd}-generated and TM-generated metagene scores were calculated for each genetic signatures (\cref{tab:svd_vs_tm_path,tab:svd_vs_tm_obs}).
 
@@
 	\centering
-	\caption{Spearman correlation of the svd vs tm metagenes}
-	\label{tab:svd_vs_tm}
-	\begin{tabular}{c}
-		gtrmatransres (RMA)
+	\caption{Spearman correlation of the svd vs tm metagenes (pathway)}
+	\label{tab:svd_vs_tm_path}
+	\begin{tabular}{lcccc}
+		& Gatza & Creighton & Cris & Fuentes-Mattei\\
 		\hline
-		FM & Creighton   &     Cris
-		akt   -0.5563 -0.5190 -0.5900
-		bcat   0.9905  0.9897  0.9977
-		e2f1   0.9193  0.9646  0.8438
-		egfr  -0.4040 -0.0430 -0.3358
-		er     0.9966  0.9978  0.9942
-		her2  -0.9794 -0.9553 -0.5817
-		ifna   0.9951 -0.9830 -0.9991
-		ifng   0.9718 -0.9086  0.9950
-		myc    0.9689  0.9878 -0.9852
-		p53    0.7923  0.3808  0.9981
-		p63    0.8368  0.8319  0.2951
-		pi3k   0.9365 -0.9543  0.5989
-		pr     0.9845 -0.9511 -0.9887
-		ras   -0.7229 -0.9078 -0.9125
-		src    0.9548 -0.9575 -0.7173
-		stat3  0.6167 -0.1902  0.9159
-		tgfb   0.9890 -0.9918 -0.2543
-		tnfa   0.4315 -0.6046  0.9365
-
-			   obsrmatransres (RMA)
-		   \hline
-		   Gatza        FM       Cris
-		rawobs     0.9985 0.9872 -0.9161
-		crol       0.9982 0.9926  0.9715
-		resobs    -0.9987 0.9898 -0.9652
-		rescrol    0.9981 0.9927  0.9571
-		caobs      0.9985 0.9893 -0.9468
-		cacrol     0.9983 0.9937  0.9677
-		caresobs  -0.9988 0.9939  0.9865
-		carescrol  0.9984 0.9952  0.9642
-		crobs     -0.9928 0.9862 -0.9344
+		\rule{0pt}{2.25ex}Akt & 1.000 & 0.5190 & 0.5900 & 0.5563 \\
+		\gls{bcat}            & 1.000 & 0.9897 & 0.9977 & 0.9905 \\
+		E2F1                  & 1.000 & 0.9646 & 0.8438 & 0.9193 \\
+		\gls{egfr}            & 1.000 & 0.0430 & 0.3358 & 0.4040 \\
+		\gls{er}              & 1.000 & 0.9978 & 0.9942 & 0.9966 \\
+		\gls{her2}            & 1.000 & 0.9553 & 0.5817 & 0.9794 \\
+		\gls{ifna}            & 1.000 & 0.9830 & 0.9991 & 0.9951 \\
+		\gls{ifny}            & 1.000 & 0.9086 & 0.9950 & 0.9718 \\
+		Myc                   & 1.000 & 0.9878 & 0.9852 & 0.9689 \\
+		p53                   & 1.000 & 0.3808 & 0.9981 & 0.7923 \\
+		p63                   & 1.000 & 0.8319 & 0.2951 & 0.8368 \\
+		\gls{pi3k}            & 1.000 & 0.9543 & 0.5989 & 0.9365 \\
+		\gls{pr}              & 1.000 & 0.9511 & 0.9887 & 0.9845 \\
+		Ras                   & 1.000 & 0.9078 & 0.9125 & 0.7229 \\
+		Src                   & 1.000 & 0.9575 & 0.7173 & 0.9548 \\
+		\gls{stat3}           & 1.000 & 0.1902 & 0.9159 & 0.6167 \\
+		\gls{tgfb}            & 1.000 & 0.9918 & 0.2543 & 0.9890 \\
+		\gls{tnfa}            & 1.000 & 0.6046 & 0.9365 & 0.4315 \\
 
@@
 \end{table}
+
+\begin{table}[htpb]
+	\centering
+	\caption{Spearman correlation of the svd vs tm metagenes (obesity)}
+	\label{tab:svd_vs_tm_obs}
+	\begin{tabular}{lcccc}
+		& Creighton & Gatza  & Fuentes-Mattei & Cris\\
+		\hline
+		\rule{0pt}{2.25ex} Cr & 1.000     & 0.9985 & 0.9872 & 0.9161 \\
+		Res                   & 1.000     & 0.9987 & 0.9898 & 0.9652 \\
+		CrOl                  & 1.000     & 0.9982 & 0.9926 & 0.9715 \\
+		ResOl                 & 1.000     & 0.9981 & 0.9927 & 0.9571 \\
+		Ca                    & 1.000     & 0.9985 & 0.9893 & 0.9468 \\
+		CaRes                 & 1.000     & 0.9988 & 0.9939 & 0.9865 \\
+		CaOl                  & 1.000     & 0.9983 & 0.9937 & 0.9677 \\
+		CaResOl               & 1.000     & 0.9984 & 0.9952 & 0.9642 \\
+		Original              & 1.000     & 0.9928 & 0.9862 & 0.9344 \\
+	\end{tabular}
+\end{table}
+
+In \cref{tab:svd_vs_tm_path,tab:svd_vs_tm_obs}, the obesity and pathway associated genetic signatures had a correlatin of 1 in Creighton's and Gatza's data sets respectively, as the transformation matrices were generated in those data sets.
+The correlations of the obesity and pathway metagenes were variable across different data sets, which was as expected since all the data sets were different from one another.
+However, what was unexpected was the fact that there were some genetic signatures that were highly correlated in all of the data sets, whereas the metagene correlation of other signatures were highly variable across different data sets.
+For example, the \gls{bcat} pathway metagene was highly correlated in all of the data sets (\textgreater{} 0.98), whereas the \gls{stat3} pathway metagene was variable across different data sets, ranging from 0.1902 to 0.9159 (\cref{tab:svd_vs_tm_path,fig:allmetacor_bar}).
+The fact that some pathway associated metagenes were consistent across different data sets suggested that some of these genetic signatures were reliable and did not depend on the data set the transformation matrices were derived from.
+On the other hand, the genetic signatures that were not consistent across the data sets were likely to be dependent on the data set in which they were derived from, and therefore the transformation matrices for these signatures must be derived from Gatza's data set.
+
+% TODO: start from here
+In contrast to the pathway associated genetic signatures, all of the obesity associated genetic signatures showed high correlation of the metagenes across the different data sets.
+
+\begin{figure}[htpb]
+	\centering
+	\includegraphics[page=2,width=0.45\linewidth]{results2/allmetacor_bar}
+	\includegraphics[page=16,width=0.45\linewidth]{results2/allmetacor_bar}
+	\caption{Visual comparison of the correlation of the metagenes across different data sets}
+	\label{fig:allmetacor_bar}
+\end{figure}
 

### 2017-02-14 thesis/methods/methods.tex
tmp commit
finished pathway and obesity metagene section - tmp commit

@@
 
+% TODO: add example plots for metagene direction
+
 One thing to consider when analysing data with metagenes is the direction of the metagene.
@@
 For obesity associated metagenes, the direction of the metagenes were checked with the obesity status and \gls{bmi} value of the samples from the data in which the metagene was derived from.
+When the data set did not have \gls{bmi}/\gls{bmi} status information for the samples, a set of common genes were taken from all of the obesity associated genetic signatures and their expressions compared with the metagene scores.
+Since all of the genetic signatures are associated with obesity, the genes that are common to all of the obesity associated genetic signatures will be related to the metagenes produced by these signatures, and therefore the expressions of these genes should be in line with the metagene scores.
+
 Checking the direction of the pathway metagenes from \citet{Gatza2010a} study is not as simple as the obesity associated metagenes, as the experimental design that identify which samples were treated to generate the pathway signature was not available.
 Due to this, it was not possible to validate whether the generated pathway metagene was in the right direction.
-
 As an alternative, the pathway metagenes were visually compared with the gene that represent the pathway using a heatmat, as the expression of this gene would most likely be affected the most by the pathway signature.
@@
 \cref{tab:metagene_direction} summarises the genes used to check the direction of the pathway metagenes.
+
 \begin{table}[htb]
@@
 			E2F1                  & E2F1\\
-			\gls{egfr}            & EGFR\\
-			\gls{er}              & ESR1\\
-			\gls{her2}            & ERBB2\\
-			\gls{ifna}            & IFNA1\\
-			\gls{ifny}            & IFNG\\
+			\Gls{egfr}            & EGFR\\
+			\Gls{er}              & ESR1\\
+			\Gls{her2}            & ERBB2\\
+			\Gls{ifna}            & IFNA1\\
+			\Gls{ifny}            & IFNG\\
 			Myc                   & MYC\\
@@
 			p63                   & TP63\\
-			\gls{pi3k}            & PIK3CA\\
-			\gls{pr}              & PGR\\
+			\Gls{pi3k}            & PIK3CA\\
+			\Gls{pr}              & PGR\\
 			Ras                   & HRAS\\
-			\gls{stat3}           & STAT3\\
+			\Gls{stat3}           & STAT3\\
 			Src                   & SRC\\
-			\gls{tgfb}            & TGFB1\\
-			\gls{tnfa}            & TNF\\
+			\Gls{tgfb}            & TGFB1\\
+			\Gls{tnfa}            & TNF\\
 			\hline
@@
 \end{table}
+
 In addition to this, the correlation of the pathway metagene with its repective gene was checked to further aid the decision on whether the metagene is going in the correct direction.
@@
 Therefore, the direction as well as the magnitude of the correlation value was also taken into account in making the decision whether to flip the metagene.
-Lastly, further on top of this, the clustering of the pathway metagenes with the other pathway metagenes were also considered.
+Lastly, the clustering of the pathway metagenes with the other pathway metagenes were also considered.
 It was known from \citet{Gatza2010a} study that some pathways clustered and grouped together with specific pathways, so the metagenes were flipped so that the grouping of the pathways matched the results from their study.
@@
 \label{sub:venn_diagrams}
+
+% TODO: talk about spearman correlation?

### 2017-02-14 thesis/results/results2.tex
finished pathway and obesity metagene section - tmp commit

@@
 
-% TODO: start from here
-In contrast to the pathway associated genetic signatures, all of the obesity associated genetic signatures showed high correlation of the metagenes across the different data sets.
+In contrast to the pathway associated genetic signatures, all of the obesity associated genetic signatures showed high correlation of the \gls{svd}- and transformation matrix-generated metagenes across the different data sets.
+This suggested that obesity associated genetic signatures were consistent across all data sets and the transformation matrices for these signatures could be made in any of the data sets.
+Taken together, transformation matrices for all the genetic  signatures were decided to be made in the Gatza's data set, since there were some pathway associated signatures that were specific to Gatza's data, but none of the obesity associated genetic signatures were specific to a data set.
+\\
 
@@
 
+\noindent
+To visually determine which pathway associated genetic signatures were most similar to the obesity associated genetic signatures, heatmaps were created with the metagenes for all of the genetic signatures.
+The directions of the pathway associated metagenes have already been determined in \cref{sec:pathway_associated_genetic_signatures_from_gatza2010a_study}, and the directions of the obesity associated metagenes were checked in Gatza's data set as described in \cref{sub:metagene_direction}.
+All of the metagenes were created in \gls{rma}-normalised Gatza's data set with \gls{svd}, and the metagene scores were plotted in a heatmap and clustered into groups (\cref{fig:gatza_allmeta}).
+
+\begin{figure}[htpb]
+	\centering
+	\includegraphics[page=15,width=0.8\linewidth]{results2/all_meta_trans}
+	\caption{All metagenes in Gatza's data}
+	\label{fig:gatza_allmeta}
+\end{figure}
+
+From \cref{fig:gatza_allmeta}, all of the obesity associated metagenes from Creighton's data set clustered together into a group, but none of the pathway associated metagenes associated strongly with the obesity metagenes.
+Though \gls{her2} and p63 pathway metagenes were grouped next to the cluster of obesity metagenes, the similarity was at most about 0.6.
+Fuentes-Mattei's obesity associated metagene did not cluster together with the obesity metagenes from Creighton's data set, nor with any of the other pathway metagenes.
+The fact that Fuentes-Mattei's metagene did not cluster with Creighton's metagenes showed that the nature of Fuentes-Mattei's metagene was different from those metagenes found in Creighton's data set.
+Furthermore, their metagene did not cluster with Gatza's Akt pathway in the heatmap, even though \citet{Fuentes-Mattei2014} provided clear evidence that their metagene specifically activated the Akt/\gls{mtor} signalling pathway in a mouse model.
+This may have been due to the lack of consistency of the Akt pathway signature in different data sets (\cref{tab:svd_vs_tm_path}).
+Another likely reason could be that Fuentes-Mattei's genetic signature was not comprised only of the genes from the Akt pathway, but genes from other pathways as well, and so the metagene did not cluster together with the Akt pathway signature in the heatmap.
+
+The results in this section suggested that none of the obesity associated genetic signatures were similar or related to any of the pathway associated genetic signatures from the \citet{Gatza2010a} study.
+
+(add heatmaps of other data as well? or maybe in appendix)
+
+% Transformation matrices for all of the obesity and pathway associated genetic signatures were generated in Gatza's data.
+
 \section{Prediction of obesity associated metagene with pathway associate metagene}

### 2017-02-15 thesis/results/results2.tex
tmp commit

@@
 Furthermore, if the \gls{svd}-derived metagenes and transformation matrix-derived metagenes were the same (or at least similar) in a different data set, this suggests that there was no difference in the data sets, at least in terms of the expression of the genetic signatures being investigated.
-In other words, the transformation matrix can be made in either the original data set or in a different data set, if there is no difference in the \gls{svd}-derived or transformation matrix-derived metagenes.
+In other words, the transformation matrix can be made in either the original data set or in a different data set if there was no difference in the \gls{svd}-derived or transformation matrix-derived metagenes.
 
-% TODO: start from here
 To decide which data set the transformation matrices for obesity and pathway associated genetic signatures should be generated in, the \gls{svd}- and transformation matrix(TM)-generated metagene scores were compared in all of the data sets.
@@
 
-% Transformation matrices for all of the obesity and pathway associated genetic signatures were generated in Gatza's data.
-
 \section{Prediction of obesity associated metagene with pathway associate metagene}
@@
 
-and lastly linear models were constructed based on the pathway metagenes and sample \gls{bmi}/\gls{bmi} status to predict the obesity associated metagenes.
+% and lastly linear models were constructed based on the pathway metagenes and sample \gls{bmi}/\gls{bmi} status to predict the obesity associated metagenes.
+
+To confirm that the obesity associated metagenes were not related to any of the biological pathway signatures from the \citet{Gatza2010a} study, linear models were created to predict the obesity metagene scores with the pathway metagene scores.
+If any of the pathway metagenes were significant in the linear model, this provides a strong evidence that the significant pathways in the model were related to the obesity metagene.
+% TODO: start from here

### 2017-02-16 thesis/results/results2.tex
tmp commit

@@
 
-(add heatmaps of other data as well? or maybe in appendix)
+(add heatmaps of other data (Creighton/FM/Cris) as well? or maybe in appendix)
 
@@
 To confirm that the obesity associated metagenes were not related to any of the biological pathway signatures from the \citet{Gatza2010a} study, linear models were created to predict the obesity metagene scores with the pathway metagene scores.
-If any of the pathway metagenes were significant in the linear model, this provides a strong evidence that the significant pathways in the model were related to the obesity metagene.
-% TODO: start from here
+If any of the pathway metagenes were significant in the linear model, this provided evidence that the significant pathways in the linear model were related to the obesity metagene.
+Since most of the obesity metagenes were found in Creighton's data set and only Creighton's and Cris' data sets had \gls{bmi}/\gls{bmi} status information for the samples, Cris' data set was used as the training data set to create linear models for obesity metagene score prediction.
+
+Metagene scores used for the construction of linear models were generated from the application of the transformation matrices derived from Gatza's data (from \cref{sec:pathway_associated_metagenes_and_obesity_associated_metagenes}) to \gls{rma}-normalised Cris' data.
+For each of the obesity metagenes identified, seven linear models were created: \gls{bmi}-only, \gls{bmi} status-only, \gls{bmi} and \gls{bmi} status, pathway metagenes-only, and all of the combinations of pathway metagenes with \gls{bmi} and/or \gls{bmi} status.

### 2017-02-16 thesis/results/results2.tex
tmp commit

@@
 For each of the obesity metagenes identified, seven linear models were created: \gls{bmi}-only, \gls{bmi} status-only, \gls{bmi} and \gls{bmi} status, pathway metagenes-only, and all of the combinations of pathway metagenes with \gls{bmi} and/or \gls{bmi} status.
+
+\begin{table}[htpb]
+	\centering
+	\caption{caption}
+	\label{tab:lm_sig_var}
+	\begin{tabular}{llcc}
+		Linear Model & Variables & Estimate & P-value\\
+		\hline
+		\gls{\gls{bmi}}-only                                 & \gls{bmi}  & -0.001481 & 0.717      \\
+		\gls{\gls{bmi}} status-only                          & Obese      & -0.09809  & 0.166      \\
+                                                             & Overweight & -0.01263  & 0.871      \\
+		\gls{\gls{bmi}} and \gls{\gls{bmi}} status           & \gls{bmi}  & 0.01087   & 0.1232     \\
+                                                             & Obese      & -0.24812  & 0.0398 *   \\
+                                                             & Overweight & -0.06873  & 0.4199     \\
+		Pathways only                                        & \gls{bcat} & 0.091552  & 0.6765     \\
+                                                             & \gls{er}   & 0.052266  & 0.8342     \\
+                                                             & \gls{ifna} & -0.499303 & 0.3070     \\
+                                                             & \gls{ifny} & 0.422406  & 0.3983     \\
+                                                             & myc        & -0.003752 & 0.9867     \\
+                                                             & \gls{pr}   & 0.583569  & 0.0164   * \\
+		\gls{\gls{bmi}} and Pathways                         & \gls{bmi}  & -0.002368 & 0.5271     \\
+                                                             & \gls{bcat} & 0.112118  & 0.6144     \\
+                                                             & \gls{er}   & 0.054001  & 0.8294     \\
+                                                             & \gls{ifna} & -0.538290 & 0.2763     \\
+                                                             & \gls{ifny} & 0.466144  & 0.3576     \\
+                                                             & myc        & 0.016832  & 0.9411     \\
+                                                             & \gls{pr}   & 0.592696  & 0.0154   * \\
+		\gls{\gls{bmi}} and Pathways                         & Obese      & -0.10346  & 0.1125     \\
+                                                             & Overweight & -0.01200  & 0.8641     \\
+                                                             & \gls{bcat} & 0.14950   & 0.5008     \\
+                                                             & \gls{er}   & 0.03263   & 0.8955     \\
+                                                             & \gls{ifna} & -0.67067  & 0.1767     \\
+                                                             & \gls{ifny} & 0.58879   & 0.2453     \\
+                                                             & myc        & 0.04388   & 0.8457     \\
+                                                             & \gls{pr}   & 0.57993   & 0.0164   * \\
+		\gls{\gls{bmi}}, \gls{\gls{bmi}} status and pathways & \gls{bmi}  & 0.009057  & 0.1614     \\
+                                                             & Obese      & -0.229223 & 0.0397   * \\
+                                                             & Overweight & -0.057586 & 0.4546     \\
+                                                             & \gls{bcat} & 0.145953  & 0.5087     \\
+                                                             & \gls{er}   & 0.009057  & 0.9708     \\
+                                                             & \gls{ifna} & -0.693675 & 0.1605     \\
+                                                             & \gls{ifny} & 0.588958  & 0.2426     \\
+                                                             & myc        & 0.026253  & 0.9070     \\
+                                                             & \gls{pr}   & 0.544864  & 0.0240   * \\
+
+	\end{tabular}
+\end{table}
+
+% [1] "rawobsgene"
+
+% Call:
+% lm(formula = as.formula(formula_txt[j]), data = data)
+
+% Coefficients:
+%              Estimate Pr(>|t|)
+% [1] "crolgene"
+
+% Call:
+% lm(formula = as.formula(formula_txt[j]), data = data)
+
+% Coefficients:
+%              Estimate Std. Error t value Pr(>|t|)
+% (Intercept)         0.553214  0.123891 4.465  2.16e-05 ***
+% bmi                 -0.001629 0.004071 -0.400 0.69
+% (Intercept)         0.55700   0.05464  10.194 <2e-16   ***
+% bmiStatusobese      -0.10551  0.07021  -1.503 0.136
+% bmiStatusoverweight -0.02165  0.07727  -0.280 0.780
+% (Intercept)         0.307513  0.163530 1.880  0.0631   .
+% bmi                 0.011279  0.006976 1.617  0.1092
+% bmiStatusobese      -0.261145 0.118795 -2.198 0.0304   *
+% bmiStatusoverweight -0.079845 0.084661 -0.943 0.3480
+% (Intercept)         0.007141  0.382725 0.019  0.9852
+% bcat_probes         0.156062  0.223779 0.697  0.4873
+% er_probes           0.123403  0.254800 0.484  0.6293
+% ifna_probes         -0.430229 0.497301 -0.865 0.3892
+% ifng_probes         0.378437  0.509352 0.743  0.4594
+% myc_probes          0.160785  0.229142 0.702  0.4846
+% pr_probes           0.597403  0.244252 2.446  0.0164   *
+% (Intercept)         0.055088  0.389472 0.141  0.8878
+% bmi                 -0.002745 0.003814 -0.720 0.4736
+% bcat_probes         0.179900  0.226799 0.793  0.4297
+% er_probes           0.125413  0.255485 0.491  0.6247
+% ifna_probes         -0.475417 0.502546 -0.946 0.3466
+% ifng_probes         0.429131  0.515526 0.832  0.4074
+% myc_probes          0.184642  0.232124 0.795  0.4284
+% pr_probes           0.607982  0.245336 2.478  0.0151   *
+% (Intercept)         0.01666   0.37989  0.044  0.9651
+% bmiStatusobese      -0.11197  0.06602  -1.696 0.0933   .
+% bmiStatusoverweight -0.02348  0.07151  -0.328 0.7434
+% bcat_probes         0.22035   0.22616  0.974  0.3325
+% er_probes           0.10450   0.25330  0.413  0.6809
+% ifna_probes         -0.60343  0.50372  -1.198 0.2341
+% ifng_probes         0.54671   0.51484  1.062  0.2911
+% myc_probes          0.21341   0.22998  0.928  0.3559
+% pr_probes           0.59491   0.24257  2.453  0.0161   *
+% (Intercept)         -0.128149 0.393204 -0.326 0.7453
+% bmi                 0.008834  0.006565 1.346  0.1819
+% bmiStatusobese      -0.234634 0.112386 -2.088 0.0397   *
+% bmiStatusoverweight -0.067942 0.078484 -0.866 0.3890
+% bcat_probes         0.216892  0.225167 0.963  0.3380
+% er_probes           0.081506  0.252745 0.322  0.7478
+% ifna_probes         -0.625864 0.501740 -1.247 0.2155
+% ifng_probes         0.546873  0.512538 1.067  0.2889
+% myc_probes          0.196215  0.229305 0.856  0.3945
+% pr_probes           0.560706  0.242816 2.309  0.0232   *
+
+% [1] "resobsgene"
+
+% Call:
+% lm(formula = as.formula(formula_txt[j]), data = data)
+
+% Coefficients:
+%              Estimate Std. Error t value Pr(>|t|)
+% (Intercept)         0.547104  0.123915 4.415  2.62e-05 ***
+% bmi                 -0.001422 0.004072 -0.349 0.728
+% (Intercept)         0.547619  0.054641 10.022 <2e-16   ***
+% bmiStatusobese      -0.096832 0.070212 -1.379 0.171
+% bmiStatusoverweight -0.001804 0.077273 -0.023 0.981
+% (Intercept)         0.298589  0.163546 1.826  0.0710   .
+% bmi                 0.011259  0.006976 1.614  0.1099
+% bmiStatusobese      -0.252186 0.118807 -2.123 0.0364   *
+% bmiStatusoverweight -0.059898 0.084669 -0.707 0.4810
+% (Intercept)         0.18694   0.36963  0.506  0.6142
+% bcat_probes         0.09717   0.21612  0.450  0.6541
+% er_probes           0.07879   0.24608  0.320  0.7496
+% ifna_probes         -0.17576  0.48029  -0.366 0.7152
+% ifng_probes         0.07930   0.49193  0.161  0.8723
+% myc_probes          -0.04517  0.22130  -0.204 0.8387
+% pr_probes           0.59554   0.23590  2.525  0.0133   *
+% (Intercept)         0.221757  0.376612 0.589  0.5574
+% bmi                 -0.001994 0.003688 -0.540 0.5902
+% bcat_probes         0.114479  0.219311 0.522  0.6029
+% er_probes           0.080247  0.247050 0.325  0.7461
+% ifna_probes         -0.208580 0.485954 -0.429 0.6688
+% ifng_probes         0.116119  0.498505 0.233  0.8163
+% myc_probes          -0.027845 0.224460 -0.124 0.9015
+% pr_probes           0.603222  0.237235 2.543  0.0127   *
+% (Intercept)         0.195965  0.367384 0.533  0.5951
+% bmiStatusobese      -0.096834 0.063844 -1.517 0.1328
+% bmiStatusoverweight -0.004412 0.069155 -0.064 0.9493
+% bcat_probes         0.150379  0.218719 0.688  0.4935
+% er_probes           0.058880  0.244961 0.240  0.8106
+% ifna_probes         -0.344126 0.487134 -0.706 0.4817
+% ifng_probes         0.242697  0.497893 0.487  0.6271
+% myc_probes          -0.001285 0.222407 -0.006 0.9954
+% pr_probes           0.591191  0.234582 2.520  0.0135   *
+% (Intercept)         0.041459  0.379420 0.109  0.9132
+% bmi                 0.009426  0.006335 1.488  0.1403
+% bmiStatusobese      -0.227717 0.108446 -2.100 0.0386   *
+% bmiStatusoverweight -0.051853 0.075733 -0.685 0.4953
+% bcat_probes         0.146687  0.217273 0.675  0.5013
+% er_probes           0.034350  0.243884 0.141  0.8883
+% ifna_probes         -0.368065 0.484150 -0.760 0.4491
+% ifng_probes         0.242870  0.494570 0.491  0.6246
+% myc_probes          -0.019634 0.221267 -0.089 0.9295
+% pr_probes           0.554701  0.234304 2.367  0.0201   *
+
+% [1] "rescrolgene"
+
+% Call:
+% lm(formula = as.formula(formula_txt[j]), data = data)
+
+% Coefficients:
+%               Estimate Std. Error t value Pr(>|t|)
+% (Intercept)         0.5343899  0.1239554 4.311  3.9e-05 ***
+% bmi                 -0.0009922 0.0040730 -0.244 0.808
+% (Intercept)         0.54870    0.05482   10.009 <2e-16  ***
+% bmiStatusobese      -0.09063   0.07044   -1.287 0.201
+% bmiStatusoverweight -0.01515   0.07753   -0.195 0.845
+% (Intercept)         0.304200   0.164179  1.853  0.0670  .
+% bmi                 0.011054   0.007003  1.578  0.1178
+% bmiStatusobese      -0.243160  0.119267  -2.039 0.0442  *
+% bmiStatusoverweight -0.072189  0.084997  -0.849 0.3978
+% (Intercept)         0.17069    0.37929   0.450  0.6537
+% bcat_probes         0.09532    0.22177   0.430  0.6683
+% er_probes           0.03171    0.25251   0.126  0.9004
+% ifna_probes         -0.57965   0.49283   -1.176 0.2426
+% ifng_probes         0.50919    0.50478   1.009  0.3157
+% myc_probes          0.06345    0.22708   0.279  0.7805
+% pr_probes           0.54200    0.24206   2.239  0.0276  *
+% (Intercept)         0.204758   0.386507  0.530  0.5976
+% bmi                 -0.001950  0.003785  -0.515 0.6076
+% bcat_probes         0.112257   0.225072  0.499  0.6192
+% er_probes           0.033134   0.253540  0.131  0.8963
+% ifna_probes         -0.611750  0.498720  -1.227 0.2231
+% ifng_probes         0.545210   0.511601  1.066  0.2894
+% myc_probes          0.080405   0.230357  0.349  0.7279
+% pr_probes           0.549516   0.243468  2.257  0.0264  *
+% (Intercept)         0.17933    0.37777   0.475  0.636
+% bmiStatusobese      -0.09681   0.06565   -1.475 0.144
+% bmiStatusoverweight -0.01219   0.07111   -0.171 0.864
+% bcat_probes         0.14969    0.22490   0.666  0.507
+% er_probes           0.01354    0.25189   0.054  0.957
+% ifna_probes         -0.73888   0.50091   -1.475 0.144
+% ifng_probes         0.66381    0.51197   1.297  0.198
+% myc_probes          0.10813    0.22869   0.473  0.638
+% pr_probes           0.53872    0.24121   2.233  0.028   *
+% (Intercept)         0.027729   0.390580  0.071  0.9436
+% bmi                 0.009248   0.006522  1.418  0.1596
+% bmiStatusobese      -0.225233  0.111636  -2.018 0.0466  *
+% bmiStatusoverweight -0.058734  0.077961  -0.753 0.4532
+% bcat_probes         0.146066   0.223664  0.653  0.5154
+% er_probes           -0.010526  0.251058  -0.042 0.9667
+% ifna_probes         -0.762373  0.498391  -1.530 0.1296
+% ifng_probes         0.663981   0.509117  1.304  0.1955
+% myc_probes          0.090122   0.227775  0.396  0.6933
+% pr_probes           0.502919   0.241196  2.085  0.0399  *
+
+% [1] "caobsgene"
+
+% Call:
+% lm(formula = as.formula(formula_txt[j]), data = data)
+
+% Coefficients:
+%              Estimate Std. Error t value Pr(>|t|)
+% (Intercept)         0.538968  0.123943 4.349  3.39e-05 ***
+% bmi                 -0.001147 0.004073 -0.282 0.779
+% (Intercept)         0.53896   0.05458  9.875  2.76e-16 ***
+% bmiStatusobese      -0.08982  0.07013  -1.281 0.203
+% bmiStatusoverweight 0.01804   0.07719  0.234  0.816
+% (Intercept)         0.27986   0.16317  1.715  0.0896   .
+% bmi                 0.01171   0.00696  1.683  0.0957   .
+% bmiStatusobese      -0.25146  0.11853  -2.121 0.0365   *
+% bmiStatusoverweight -0.04241  0.08447  -0.502 0.6168
+% (Intercept)         0.12454   0.38803  0.321  0.7490
+% bcat_probes         0.14068   0.22688  0.620  0.5367
+% er_probes           0.11597   0.25833  0.449  0.6545
+% ifna_probes         -0.33319  0.50420  -0.661 0.5104
+% ifng_probes         0.30871   0.51641  0.598  0.5515
+% myc_probes          -0.04569  0.23232  -0.197 0.8445
+% pr_probes           0.56692   0.24764  2.289  0.0243   *
+% (Intercept)         0.162063  0.395326 0.410  0.6828
+% bmi                 -0.002148 0.003872 -0.555 0.5804
+% bcat_probes         0.159335  0.230208 0.692  0.4906
+% er_probes           0.117547  0.259326 0.453  0.6514
+% ifna_probes         -0.368552 0.510101 -0.723 0.4718
+% ifng_probes         0.348376  0.523276 0.666  0.5072
+% myc_probes          -0.027017 0.235613 -0.115 0.9090
+% pr_probes           0.575201  0.249024 2.310  0.0232   *
+% (Intercept)         0.134514  0.384960 0.349  0.7276
+% bmiStatusobese      -0.099695 0.066899 -1.490 0.1397
+% bmiStatusoverweight 0.008987  0.072463 0.124  0.9016
+% bcat_probes         0.193430  0.229183 0.844  0.4009
+% er_probes           0.092451  0.256681 0.360  0.7196
+% ifna_probes         -0.522336 0.510440 -1.023 0.3089
+% ifng_probes         0.492130  0.521713 0.943  0.3481
+% myc_probes          -0.001892 0.233048 -0.008 0.9935
+% pr_probes           0.560583  0.245805 2.281  0.0249   *
+% (Intercept)         -0.028482 0.397505 -0.072 0.9430
+% bmi                 0.009944  0.006637 1.498  0.1376
+% bmiStatusobese      -0.237770 0.113616 -2.093 0.0392   *
+% bmiStatusoverweight -0.041061 0.079343 -0.518 0.6061
+% bcat_probes         0.189535  0.227630 0.833  0.4073
+% er_probes           0.066573  0.255509 0.261  0.7950
+% ifna_probes         -0.547591 0.507227 -1.080 0.2832
+% ifng_probes         0.492312  0.518143 0.950  0.3446
+% myc_probes          -0.021250 0.231813 -0.092 0.9272
+% pr_probes           0.522087  0.245472 2.127  0.0362   *
+
+% [1] "cacrolgene"
+
+% Call:
+% lm(formula = as.formula(formula_txt[j]), data = data)
+
+% Coefficients:
+%               Estimate Std. Error t value Pr(>|t|)
+% (Intercept)         5.062e-01  1.240e-01 4.083  9.16e-05 ***
+% bmi                 -3.953e-05 4.074e-03 -0.010 0.992
+% (Intercept)         0.52453    0.05485   9.564  1.29e-15 ***
+% bmiStatusobese      -0.06411   0.07047   -0.910 0.365
+% bmiStatusoverweight 0.02958    0.07756   0.381  0.704
+% (Intercept)         0.273274   0.164135  1.665  0.0992   .
+% bmi                 0.011360   0.007001  1.622  0.1080
+% bmiStatusobese      -0.220856  0.119235  -1.852 0.0671   .
+% bmiStatusoverweight -0.029032  0.084974  -0.342 0.7334
+% (Intercept)         0.02932    0.40048   0.073  0.9418
+% bcat_probes         0.23660    0.23416   1.010  0.3150
+% er_probes           0.09521    0.26662   0.357  0.7218
+% ifna_probes         -0.49129   0.52037   -0.944 0.3476
+% ifng_probes         0.44466    0.53298   0.834  0.4063
+% myc_probes          0.16111    0.23977   0.672  0.5033
+% pr_probes           0.49566    0.25558   1.939  0.0555   .
+% (Intercept)         0.051698   0.408467  0.127  0.8996
+% bmi                 -0.001281  0.004000  -0.320 0.7495
+% bcat_probes         0.247720   0.237861  1.041  0.3004
+% er_probes           0.096149   0.267946  0.359  0.7205
+% ifna_probes         -0.512383  0.527057  -0.972 0.3335
+% ifng_probes         0.468320   0.540670  0.866  0.3887
+% myc_probes          0.172240   0.243445  0.708  0.4811
+% pr_probes           0.500601   0.257301  1.946  0.0548   .
+% (Intercept)         0.03806    0.39931   0.095  0.9243
+% bmiStatusobese      -0.07973   0.06939   -1.149 0.2536
+% bmiStatusoverweight 0.02249    0.07516   0.299  0.7655
+% bcat_probes         0.27648    0.23772   1.163  0.2479
+% er_probes           0.07298    0.26625   0.274  0.7846
+% ifna_probes         -0.66043   0.52946   -1.247 0.2155
+% ifng_probes         0.60854    0.54116   1.125  0.2638
+% myc_probes          0.19456    0.24173   0.805  0.4230
+% pr_probes           0.48849    0.25497   1.916  0.0586   .
+% (Intercept)         -0.123986  0.412743  -0.300 0.7646
+% bmi                 0.009886   0.006892  1.434  0.1549
+% bmiStatusobese      -0.216997  0.117971  -1.839 0.0692   .
+% bmiStatusoverweight -0.027267  0.082384  -0.331 0.7414
+% bcat_probes         0.272606   0.236356  1.153  0.2518
+% er_probes           0.047249   0.265304  0.178  0.8591
+% ifna_probes         -0.685542  0.526672  -1.302 0.1964
+% ifng_probes         0.608724   0.538006  1.131  0.2609
+% myc_probes          0.175314   0.240700  0.728  0.4683
+% pr_probes           0.450215   0.254882  1.766  0.0808   .
+
+% [1] "caresobsgene"
+
+% Call:
+% lm(formula = as.formula(formula_txt[j]), data = data)
+
+% Coefficients:
+%             Estimate Std. Error t value Pr(>|t|)
+% (Intercept)         0.462424  0.123913 3.732  0.00032  ***
+% bmi                 0.001442  0.004072 0.354  0.72408
+% (Intercept)         0.46970   0.05452  8.616  1.39e-13 ***
+% bmiStatusobese      0.09338   0.07005  1.333  0.186
+% bmiStatusoverweight -0.01840  0.07710  -0.239 0.812
+% (Intercept)         0.722057  0.163100 4.427  2.55e-05 ***
+% bmi                 -0.011409 0.006957 -1.640 0.1043
+% bmiStatusobese      0.250806  0.118483 2.117  0.0369   *
+% bmiStatusoverweight 0.040473  0.084438 0.479  0.6328
+% (Intercept)         1.1017    0.3815   2.888  0.00483  **
+% bcat_probes         -0.2217   0.2230   -0.994 0.32284
+% er_probes           -0.2308   0.2540   -0.909 0.36579
+% ifna_probes         0.1853    0.4957   0.374  0.70938
+% ifng_probes         -0.1528   0.5077   -0.301 0.76408
+% myc_probes          -0.0737   0.2284   -0.323 0.74765
+% pr_probes           -0.6876   0.2434   -2.825 0.00581  **
+% (Intercept)         1.056654  0.388317 2.721  0.00780  **
+% bmi                 0.002580  0.003803 0.678  0.49929
+% bcat_probes         -0.244104 0.226127 -1.079 0.28322
+% er_probes           -0.232713 0.254728 -0.914 0.36335
+% ifna_probes         0.227770  0.501056 0.455  0.65049
+% ifng_probes         -0.200460 0.513998 -0.390 0.69745
+% myc_probes          -0.096122 0.231435 -0.415 0.67888
+% pr_probes           -0.697585 0.244608 -2.852 0.00538  **
+% (Intercept)         1.09174   0.37821  2.887  0.00488  **
+% bmiStatusobese      0.10153   0.06573  1.545  0.12592
+% bmiStatusoverweight -0.00544  0.07119  -0.076 0.93926
+% bcat_probes         -0.27598  0.22517  -1.226 0.22351
+% er_probes           -0.20770  0.25218  -0.824 0.41234
+% ifna_probes         0.37360   0.50149  0.745  0.45823
+% ifng_probes         -0.33545  0.51257  -0.654 0.51449
+% myc_probes          -0.11868  0.22896  -0.518 0.60548
+% pr_probes           -0.68170  0.24150  -2.823 0.00586  **
+% (Intercept)         1.235747  0.391476 3.157  0.00218  **
+% bmi                 -0.008785 0.006537 -1.344 0.18237
+% bmiStatusobese      0.223518  0.111892 1.998  0.04881  *
+% bmiStatusoverweight 0.038776  0.078139 0.496  0.62094
+% bcat_probes         -0.272540 0.224177 -1.216 0.22730
+% er_probes           -0.184837 0.251634 -0.735 0.46455
+% ifna_probes         0.395907  0.499535 0.793  0.43015
+% ifng_probes         -0.335611 0.510285 -0.658 0.51243
+% myc_probes          -0.101582 0.228298 -0.445 0.65743
+% pr_probes           -0.647688 0.241749 -2.679 0.00879  **
+
+% [1] "carescrolgene"
+
+% Call:
+% lm(formula = as.formula(formula_txt[j]), data = data)
+
+% Coefficients:
+%              Estimate Std. Error t value Pr(>|t|)
+% (Intercept)         0.4928357 0.1239867 3.975  0.000136 ***
+% bmi                 0.0004131 0.0040740 0.101  0.919450
+% (Intercept)         0.52814   0.05483   9.633  9.13e-16 ***
+% bmiStatusobese      -0.06913  0.07045   -0.981 0.329
+% bmiStatusoverweight 0.02453   0.07753   0.316  0.752
+% (Intercept)         0.232845  0.163204  1.427  0.1569
+% bmi                 0.013351  0.006962  1.918  0.0582   .
+% bmiStatusobese      -0.253344 0.118558  -2.137 0.0352   *
+% bmiStatusoverweight -0.044356 0.084492  -0.525 0.6008
+% (Intercept)         -0.1205   0.3891    -0.310 0.7576
+% bcat_probes         0.2439    0.2275    1.072  0.2864
+% er_probes           0.1383    0.2590    0.534  0.5948
+% ifna_probes         -0.4364   0.5056    -0.863 0.3903
+% ifng_probes         0.4301    0.5178    0.831  0.4084
+% myc_probes          0.2837    0.2330    1.218  0.2264
+% pr_probes           0.5789    0.2483    2.331  0.0219   *
+% (Intercept)         -0.102321 0.396926  -0.258 0.7972
+% bmi                 -0.001038 0.003887  -0.267 0.7900
+% bcat_probes         0.252943  0.231140  1.094  0.2767
+% er_probes           0.139029  0.260375  0.534  0.5947
+% ifna_probes         -0.453502 0.512164  -0.885 0.3782
+% ifng_probes         0.449266  0.525393  0.855  0.3947
+% myc_probes          0.292758  0.236566  1.238  0.2191
+% pr_probes           0.582903  0.250031  2.331  0.0219   *
+% (Intercept)         -0.11188  0.38790   -0.288 0.7737
+% bmiStatusobese      -0.08019  0.06741   -1.190 0.2373
+% bmiStatusoverweight 0.01833   0.07302   0.251  0.8024
+% bcat_probes         0.28469   0.23093   1.233  0.2209
+% er_probes           0.11686   0.25864   0.452  0.6525
+% ifna_probes         -0.60153  0.51434   -1.170 0.2453
+% ifng_probes         0.59011   0.52570   1.123  0.2646
+% myc_probes          0.31782   0.23483   1.353  0.1793
+% pr_probes           0.57227   0.24768   2.311  0.0231   *
+% (Intercept)         -0.284628 0.400004  -0.712 0.4786
+% bmi                 0.010538  0.006679  1.578  0.1181
+% bmiStatusobese      -0.226528 0.114330  -1.981 0.0506   .
+% bmiStatusoverweight -0.034711 0.079842  -0.435 0.6648
+% bcat_probes         0.280559  0.229061  1.225  0.2239
+% er_probes           0.089437  0.257116  0.348  0.7288
+% ifna_probes         -0.628292 0.510417  -1.231 0.2216
+% ifng_probes         0.590302  0.521401  1.132  0.2606
+% myc_probes          0.297308  0.233271  1.275  0.2058
+% pr_probes           0.531475  0.247015  2.152  0.0341   *
+
+% [1] "crobsgenes"
+
+% Call:
+% lm(formula = as.formula(formula_txt[j]), data = data)
+
+% Coefficients:
+%             Estimate Std. Error t value Pr(>|t|)
+% (Intercept)         0.456126  0.123888 3.682  0.000381 ***
+% bmi                 0.001654  0.004071 0.406  0.685322
+% (Intercept)         0.459596  0.054606 8.417  3.7e-13  ***
+% bmiStatusobese      0.100893  0.070167 1.438  0.154
+% bmiStatusoverweight 0.005772  0.077224 0.075  0.941
+% (Intercept)         0.704210  0.163518 4.307  4.04e-05 ***
+% bmi                 -0.011059 0.006975 -1.586 0.1162
+% bmiStatusobese      0.253491  0.118786 2.134  0.0354   *
+% bmiStatusoverweight 0.062836  0.084654 0.742  0.4598
+% (Intercept)         1.0667    0.3806   2.803  0.00618  **
+% bcat_probes         -0.2284   0.2225   -1.026 0.30740
+% er_probes           -0.1481   0.2534   -0.584 0.56041
+% ifna_probes         0.4504    0.4945   0.911  0.36477
+% ifng_probes         -0.4013   0.5065   -0.792 0.43018
+% myc_probes          -0.1402   0.2279   -0.615 0.53986
+% pr_probes           -0.6446   0.2429   -2.654 0.00938  **
+% (Intercept)         1.014279  0.387058 2.620  0.01029  *
+% bmi                 0.003004  0.003791 0.792  0.43014
+% bcat_probes         -0.254482 0.225394 -1.129 0.26184
+% er_probes           -0.150261 0.253902 -0.592 0.55545
+% ifna_probes         0.499868  0.499432 1.001  0.31954
+% ifng_probes         -0.456812 0.512332 -0.892 0.37494
+% myc_probes          -0.166314 0.230685 -0.721 0.47278
+% pr_probes           -0.656157 0.243815 -2.691 0.00847  **
+% (Intercept)         1.05662   0.37693  2.803  0.0062   **
+% bmiStatusobese      0.11346   0.06550  1.732  0.0867   .
+% bmiStatusoverweight 0.01412   0.07095  0.199  0.8427
+% bcat_probes         -0.29209  0.22440  -1.302 0.1964
+% er_probes           -0.12674  0.25133  -0.504 0.6153
+% ifna_probes         0.63723   0.49979  1.275  0.2056
+% ifng_probes         -0.58273  0.51083  -1.141 0.2570
+% myc_probes          -0.19254  0.22819  -0.844 0.4010
+% pr_probes           -0.64072  0.24068  -2.662 0.0092   **
+% (Intercept)         1.198471  0.390241 3.071  0.00283  **
+% bmi                 -0.008654 0.006516 -1.328 0.18754
+% bmiStatusobese      0.233624  0.111539 2.095  0.03905  *
+% bmiStatusoverweight 0.057676  0.077893 0.740  0.46097
+% bcat_probes         -0.288700 0.223470 -1.292 0.19974
+% er_probes           -0.104217 0.250840 -0.415 0.67880
+% ifna_probes         0.659208  0.497958 1.324  0.18895
+% ifng_probes         -0.582884 0.508675 -1.146 0.25491
+% myc_probes          -0.175698 0.227577 -0.772 0.44214
+% pr_probes           -0.607216 0.240986 -2.520 0.01353  *
+
+% [1] "fmobsgenes"
+
+% Call:
+% lm(formula = as.formula(formula_txt[j]), data = data)
+
+% Coefficients:
+%             Estimate Std. Error t value Pr(>|t|)
+% (Intercept)         0.436628  0.123787 3.527  0.000643 ***
+% bmi                 0.002314  0.004067 0.569  0.570763
+% (Intercept)         0.58369   0.05435  10.740 <2e-16   ***
+% bmiStatusobese      -0.08452  0.06983  -1.210 0.2291
+% bmiStatusoverweight -0.14827  0.07686  -1.929 0.0567   .
+% (Intercept)         0.251170  0.160863 1.561  0.12176
+% bmi                 0.015034  0.006862 2.191  0.03091  *
+% bmiStatusobese      -0.291956 0.116858 -2.498 0.01419  *
+% bmiStatusoverweight -0.225840 0.083280 -2.712 0.00794  **
+% (Intercept)         0.33317   0.36900  0.903  0.3689
+% bcat_probes         -0.13489  0.21576  -0.625 0.5334
+% er_probes           0.11973   0.24567  0.487  0.6271
+% ifna_probes         -0.56244  0.47947  -1.173 0.2438
+% ifng_probes         0.58838   0.49109  1.198  0.2340
+% myc_probes          0.41768   0.22093  1.891  0.0618   .
+% pr_probes           -0.08815  0.23550  -0.374 0.7090
+% (Intercept)         0.291390  0.375705 0.776  0.4400
+% bmi                 0.002392  0.003679 0.650  0.5172
+% bcat_probes         -0.155669 0.218783 -0.712 0.4786
+% er_probes           0.117981  0.246455 0.479  0.6333
+% ifna_probes         -0.523056 0.484783 -1.079 0.2835
+% ifng_probes         0.544202  0.497304 1.094  0.2767
+% myc_probes          0.396892  0.223919 1.772  0.0797   .
+% pr_probes           -0.097371 0.236664 -0.411 0.6817
+% (Intercept)         0.33459   0.36629  0.913  0.3634
+% bmiStatusobese      -0.08029  0.06366  -1.261 0.2104
+% bmiStatusoverweight -0.12492  0.06895  -1.812 0.0733   .
+% bcat_probes         -0.07254  0.21807  -0.333 0.7402
+% er_probes           0.13036   0.24424  0.534  0.5948
+% ifna_probes         -0.56035  0.48569  -1.154 0.2517
+% ifng_probes         0.58760   0.49642  1.184  0.2397
+% myc_probes          0.46651   0.22175  2.104  0.0382   *
+% pr_probes           -0.07506  0.23389  -0.321 0.7490
+% (Intercept)         0.082260  0.370293 0.222  0.82471
+% bmi                 0.015393  0.006183 2.490  0.01464  *
+% bmiStatusobese      -0.294047 0.105838 -2.778 0.00666  **
+% bmiStatusoverweight -0.202402 0.073911 -2.738 0.00746  **
+% bcat_probes         -0.078569 0.212047 -0.371 0.71187
+% er_probes           0.090296  0.238018 0.379  0.70532
+% ifna_probes         -0.599445 0.472504 -1.269 0.20787
+% ifng_probes         0.587878  0.482673 1.218  0.22646
+% myc_probes          0.436540  0.215944 2.022  0.04623  *
+% pr_probes           -0.134650 0.228668 -0.589 0.55746

### 2017-02-21 thesis/results/results2.tex
tmp commit

--- 2017-02-16 thesis/results/results2.tex
@@
 For each of the obesity metagenes identified, seven linear models were created: \gls{bmi}-only, \gls{bmi} status-only, \gls{bmi} and \gls{bmi} status, pathway metagenes-only, and all of the combinations of pathway metagenes with \gls{bmi} and/or \gls{bmi} status.
+The only variables that showed significance in any of the models that predicted the obesity metagenes taken from the Creighton \textit{et al.} data set were \gls{pr} metagene and whether samples were obese or not (\cref{tab:lm_sig_var}).
+In the model that predicted the FM metagene scores, the obesity status of the samples and Myc metagene score were significant, but the \gls{pr} metagene score was not (\cref{app:b}).
+
+\begin{table}[htpb]
+	\centering
+	\caption{Description of the linear models used to predict the CR obesity metagene}
+	\label{tab:lm_sig_var}
+	\begin{threeparttable}
+		\begin{tabular}{llrr}
+			Linear Model & Variables & Estimate & P-value\\
+			\hline
+			\hline
+			\rule{0pt}{2.25ex}\gls{bmi}-only                           & \gls{bmi}  & -0.0016 & 0.717\\
+			\hline
+			\rule{0pt}{2.25ex}\gls{bmi} status-only                    & Overweight & -0.0126 & 0.871\\
+                                                                       & Obese      & -0.0981 & 0.166\\
+			\hline
+			\rule{0pt}{2.25ex}\gls{bmi} and \gls{bmi} status           & \gls{bmi}  & 0.0109  & 0.123\\
+                                                                       & Overweight & -0.0687 & 0.420\\
+                                                                       & Obese      & -0.2481 & \textbf{0.040}\tnote{1}\\
+			\hline
+			\rule{0pt}{2.25ex}Pathways only                            & \gls{bcat} & 0.0916  & 0.677\\
+                                                                       & \gls{er}   & 0.0523  & 0.834\\
+                                                                       & \gls{ifna} & -0.4993 & 0.307\\
+                                                                       & \gls{ifny} & 0.4224  & 0.398\\
+                                                                       & Myc        & -0.0038 & 0.987\\
+                                                                       & \gls{pr}   & 0.5836  & \textbf{0.016}\\
+			\hline
+			\rule{0pt}{2.25ex}\gls{bmi} and Pathways                   & \gls{bmi}  & -0.0024 & 0.527\\
+                                                                       & \gls{bcat} & 0.1121  & 0.614\\
+                                                                       & \gls{er}   & 0.0540  & 0.829\\
+                                                                       & \gls{ifna} & -0.5383 & 0.276\\
+                                                                       & \gls{ifny} & 0.4661  & 0.358\\
+                                                                       & Myc        & 0.0168  & 0.941\\
+                                                                       & \gls{pr}   & 0.5927  & \textbf{0.015}\\
+			\hline
+			\rule{0pt}{2.25ex}\gls{bmi} status and Pathways            & Overweight & -0.0120 & 0.864\\
+                                                                       & Obese      & -0.1035 & 0.113\\
+                                                                       & \gls{bcat} & 0.1495  & 0.501\\
+                                                                       & \gls{er}   & 0.0326  & 0.896\\
+                                                                       & \gls{ifna} & -0.6707 & 0.177\\
+                                                                       & \gls{ifny} & 0.5888  & 0.245\\
+                                                                       & Myc        & 0.0439  & 0.846\\
+                                                                       & \gls{pr}   & 0.5799  & \textbf{0.016}\\
+			\hline
+			\rule{0pt}{2.25ex}\gls{bmi}, \gls{bmi} status and pathways & \gls{bmi}  & 0.0091  & 0.161\\
+                                                                       & Overweight & -0.0576 & 0.455\\
+                                                                       & Obese      & -0.2292 & \textbf{0.040}\\
+                                                                       & \gls{bcat} & 0.1460  & 0.509\\
+                                                                       & \gls{er}   & 0.0091  & 0.971\\
+                                                                       & \gls{ifna} & -0.6937 & 0.161\\
+                                                                       & \gls{ifny} & 0.5890  & 0.243\\
+                                                                       & Myc        & 0.0263  & 0.907\\
+                                                                       & \gls{pr}   & 0.5449  & \textbf{0.024}\\
+			\hline
+			\hline
+		\end{tabular}
+		\begin{tablenotes}
+			\item [1] All values in bold are statistically significant (p \textless{} 0.05).
+		\end{tablenotes}
+	\end{threeparttable}
+\end{table}
+
+All of the models constructed from sample \gls{bmi}, \gls{bmi} status and/or Gatza's pathway metagene scores confirmed that none of the pathways were significant, except for the \gls{pr} and Myc pathway metagene scores in the models for Creighton metagenes and FM metagene, respectively.
+It was not surprising that the \gls{pr} pathway metagene showed significance in the models generated in Cris' data set, as \gls{pr} status was very strongly correlated with Cris' data set.
+The significance of the Myc metagene in the model for FM metagene was difficult to interpret, as there was no reported association of FM metagene with Myc pathway by \citet{Fuentes-Mattei2014}.
+However, ``sustained cell proliferation'' was one of the cancer hallmarks that \citet{Fuentes-Mattei2014} had shown to be related to their obesity genetic signature, so the Myc pathway metagene may have picked up the proliferation signal in the FM metagene.
+
+Since the \gls{pr} pathway metagene showed significant association in many of the linear models, linear models were made with all of the combinations of \gls{bmi}, \gls{bmi} status and \gls{pr} pathway metagene for all of the obesity associated metagenes (\cref{tab:lm_pr_only} and \cref{app:b}).
+
+\begin{table}[htpb]
+	\centering
+	\caption{Description of the linear models used to predict the CR obesity metagene, using only the sample \gls{bmi}, \gls{bmi} status and \gls{pr} pathway metagene scores}
+	\label{tab:lm_pr_only}
+	\begin{threeparttable}
+			\begin{tabular}{llrS[detect-weight]}
+				Linear Model & Variables & Estimate &P-value\\
+				\hline
+				\hline
+				\rule{0pt}{2.25ex}\gls{pr}-only & \gls{pr}   & 0.477  & {\num{6.06d-7}}\tnote{1}\\
+			\end{tabular}
+			\begin{tablenotes}
+				\item [1] All values in bold are statistically significant (p \textless{} 0.05).
+			\end{tablenotes}
+	\end{threeparttable}
+\end{table}
+% \begin{table}[htpb]
+% 	\centering
+% 	\caption{Description of the linear models used to predict the CR obesity metagene, using only the sample \gls{bmi}, \gls{bmi} status and \gls{pr} pathway metagene scores}
+% 	\label{tab:lm_pr_only}
+% 	\begin{threeparttable}
+% 			\begin{tabular}{llrS}
+% 				Linear Model & Variables & Estimate &P-value\\
+% 				\hline
+% 				\hline
+% 				\rule{0pt}{2.25ex}\gls{pr}-only                            & \gls{pr}   & 0.477  & \num{6.06d-7}\tnote{1}\\
+% 				\hline
+% 				\rule{0pt}{2.25ex}\gls{bmi} and \gls{pr}                   & \gls{bmi}  & -0.002 & 0.586\\
+%                                                                            & \gls{pr}   & 0.478  & 6.35 \times10^{-7}\\
+% 				\hline
+% 				\rule{0pt}{2.25ex}\gls{bmi} status and \gls{pr}            & Obese      & -0.085 & 0.176\\
+%                                                                            & Overweight & -0.011 & 0.875\\
+%                                                                            & \gls{pr}   & 0.471  & 8.25 \times10^{-7}\\
+% 				\hline
+% 				\rule{0pt}{2.25ex}\gls{bmi}, \gls{bmi} status and \gls{pr} & \gls{bmi}  & 0.007  & 0.237\\
+%                                                                            & Obese      & -0.188 & 0.081\\
+%                                                                            & Overweight & -0.049 & 0.516\\
+%                                                                            & \gls{pr}   & 0.459  & 1.54 \times10^{-6}\\
+% 				\hline
+% 				\hline
+% 			\end{tabular}
+% 			\begin{tablenotes}
+% 				\item [1] All values in bold are statistically significant (p \textless{} 0.05).
+% 			\end{tablenotes}
+% 	\end{threeparttable}
+% \end{table}

### 2017-02-21 thesis/thesis.tex
tmp commit
tmp commit; added appendix file
Added figure images into results; tmp commit
finished pathway and obesity metagene section - tmp commit

@@
 \usepackage[T1]{fontenc}                            % package for specifying output font encoding
+\usepackage[acronym]{glossaries}                    % package for making glossaries
+\usepackage[flushleft]{threeparttable}              % package to add notes at the bottom of the table
+\usepackage[font=footnotesize,format=hang]{caption} % package for formatting figure caption
+\usepackage[hidelinks, colorlinks]{hyperref}        % package for hyperlinking references - hidelinks option removes the ugly boxes around the huperlinks
+\usepackage[md]{titlesec}                           % package for tweaking the chapter/section/subsection titles format
+\usepackage[noabbrev,capitalise]{cleveref}          % package for referencing multiple referencing in a single \ref command
+\usepackage[toc]{appendix}                          % package for making appendices
 \usepackage[utf8]{inputenc}                         % package for specifying input font encoding
-\usepackage{natbib}                                    % package for reference management
+\usepackage{chngcntr}                               % package for making continuous figure numbering, instead of section-based numbering
 \usepackage{graphicx}                               % package for graphics handling (i.e. figure images)
 \usepackage{mathptmx}                               % package for Times New Roman font
-\usepackage{chngcntr}                                  % package for making continuous figure numbering, instead of section-based numbering
-\usepackage[flushleft]{threeparttable}                 % package to add notes at the bottom of the table
+\usepackage{natbib}                                 % package for reference management
 \usepackage{setspace}                               % package for setting line space (used in title page)
-\usepackage[hidelinks, colorlinks]{hyperref}           % package for hyperlinking references - hidelinks option removes the ugly boxes around the huperlinks
-\usepackage[md]{titlesec}                              % package for tweaking the chapter/section/subsection titles format
-\usepackage[acronym]{glossaries} % package for making glossaries
 \usepackage{xcolor,colortbl}                        % package for making colours
-\usepackage[noabbrev,capitalise]{cleveref}             % package for referencing multiple referencing in a single \ref command
-\usepackage[font=footnotesize,format=hang]{caption}    % package for formatting figure caption
-\usepackage{subcaption}    % package for formatting figure caption
+\usepackage{booktabs}                               % package for better looking tables
+\usepackage{longtable}                              % package for tables that span multiple pages
+\usepackage{siunitx}                                % package for scientific notations and units
+\usepackage{bm}                                % package for scientific notations and units
 
@@
 \creflabelformat{equation}{#2\textup{#1}#3}
+
+% Change the default cref name for appendix:
+\crefname{appcha}{Appendix}{Appendices}
+\crefname{appsec}{Appendix}{Appendices}
+
+% Make siunitx detect bold/italics/etc:
+\sisetup{detect-weight=true,detect-family=true,detect-inline-weight=text,table-parse-only}
 
@@
 % If you want to compile specific files, uncomment the line below and add the filenames in the {} :
-\includeonly{./results/results}
+\includeonly{{./methods/methods},{./results/results},{./results/results2},{./misc/appendix}}
+% \includeonly{{./results/results},{./results/results2},{./misc/appendix}}
 
@@
 
+\include{./misc/appendix}
+
 % Use the reference style specified in BiocRefStyle.bst:

### 2017-02-21 thesis/results/results2.tex
tmp commit

@@
 
+\bfseries \num{6.06d-7}
+
 \begin{table}[htpb]
@@
 	\begin{threeparttable}
-			\begin{tabular}{llrS[detect-weight]}
-				Linear Model & Variables & Estimate &P-value\\
+			\begin{tabular}{llrS}
+				Linear Model & Variables & Estimate & {P-value}\\
 				\hline
 				\hline
-				\rule{0pt}{2.25ex}\gls{pr}-only & \gls{pr}   & 0.477  & {\num{6.06d-7}}\tnote{1}\\
+				\rule{0pt}{2.25ex}\gls{pr}-only                            & \gls{pr}   & 0.477  & {\bfseries \num{6.06d-7}}\tnote{*}\\
+				\hline
+				\rule{0pt}{2.25ex}\gls{bmi} and \gls{pr}                   & \gls{bmi}  & -0.002 & 0.586\\
+                                                                           & \gls{pr}   & 0.478  & {\bfseries \num{6.35d-7}}\\
+				\hline
+				\rule{0pt}{2.25ex}\gls{bmi} status and \gls{pr}            & Obese      & -0.085 & 0.176\\
+                                                                           & Overweight & -0.011 & 0.875\\
+                                                                           & \gls{pr}   & 0.471  & {\bfseries \num{8.25d-7}}\\
+				\hline
+				\rule{0pt}{2.25ex}\gls{bmi}, \gls{bmi} status and \gls{pr} & \gls{bmi}  & 0.007  & 0.237\\
+                                                                           & Obese      & -0.188 & {\bfseries 0.081}\\
+                                                                           & Overweight & -0.049 & 0.516\\
+                                                                           & \gls{pr}   & 0.459  & {\bfseries \num{1.54d-6}}\\
+				\hline
+				\hline
 			\end{tabular}
 			\begin{tablenotes}
-				\item [1] All values in bold are statistically significant (p \textless{} 0.05).
+				\item [*] All values in bold are statistically significant (p \textless{} 0.05).
 			\end{tablenotes}

### 2017-02-22 thesis/results/results2.tex
tmp commit

--- 2017-02-21 thesis/results/results2.tex
@@
 		\hline
+		\hline
 		\rule{0pt}{2.25ex}Akt & 1.000 & 0.5190 & 0.5900 & 0.5563 \\
@@
 		\gls{tnfa}            & 1.000 & 0.6046 & 0.9365 & 0.4315 \\
-
+		\hline
+		\hline
 	\end{tabular}
@@
 		\hline
+		\hline
 		\rule{0pt}{2.25ex} Cr & 1.000     & 0.9985 & 0.9872 & 0.9161 \\
@@
 		Original              & 1.000     & 0.9928 & 0.9862 & 0.9344 \\
+		\hline
+		\hline
 	\end{tabular}
@@
 			\hline
-			\rule{0pt}{2.25ex}\gls{bmi}-only                           & \gls{bmi}  & -0.0016 & 0.717\\
+			\rule{0pt}{2.25ex}\gls{bmi} only                           & \gls{bmi}  & -0.0016 & 0.717\\
 			\hline
-			\rule{0pt}{2.25ex}\gls{bmi} status-only                    & Overweight & -0.0126 & 0.871\\
+			\rule{0pt}{2.25ex}\gls{bmi} status only                    & Overweight & -0.0126 & 0.871\\
                                                                        & Obese      & -0.0981 & 0.166\\
@@
 
-Since the \gls{pr} pathway metagene showed significant association in many of the linear models, linear models were made with all of the combinations of \gls{bmi}, \gls{bmi} status and \gls{pr} pathway metagene for all of the obesity associated metagenes (\cref{tab:lm_pr_only} and \cref{app:b}).
+Since the \gls{pr} pathway metagene showed significant association in many of the linear models, linear models were made with each of the combinations of \gls{bmi}, \gls{bmi} status and \gls{pr} pathway metagene for all of the obesity associated metagenes (\cref{tab:lm_pr_only} and \cref{app:b}).
+As expected, all of the linear models showed significant contribution of the \gls{pr} metagene scores to the model for all obesity metagenes (except FM metagene; \cref{app:b}).
 
@@
 	\begin{threeparttable}
-			\begin{tabular}{llrS[detect-weight]}
-				Linear Model & Variables & Estimate &P-value\\
+			\begin{tabular}{llr{\bfseries}S}
+				Linear Model & Variables & Estimate & {P-value}\\
 				\hline
 				\hline
-				\rule{0pt}{2.25ex}\gls{pr}-only & \gls{pr}   & 0.477  & {\num{6.06d-7}}\tnote{1}\\
+				\rule{0pt}{2.25ex}\gls{pr} only                            & \gls{pr}   & 0.477  & {\bfseries \num{6.06d-7}}\tnote{1}\\
+				\hline
+				\rule{0pt}{2.25ex}\gls{bmi} and \gls{pr}                   & \gls{bmi}  & -0.002 & 0.586\\
+                                                                           & \gls{pr}   & 0.478  & {\bfseries \num{6.35d-7}}\\
+				\hline
+				\rule{0pt}{2.25ex}\gls{bmi} status and \gls{pr}            & Obese      & -0.085 & 0.176\\
+                                                                           & Overweight & -0.011 & 0.875\\
+                                                                           & \gls{pr}   & 0.471  & {\bfseries \num{8.25d-7}}\\
+				\hline
+				\rule{0pt}{2.25ex}\gls{bmi}, \gls{bmi} status and \gls{pr} & \gls{bmi}  & 0.007  & 0.237\\
+                                                                           & Obese      & -0.188 & 0.081\\
+                                                                           & Overweight & -0.049 & 0.516\\
+                                                                           & \gls{pr}   & 0.459  & {\bfseries \num{1.54d-6}}\\
+				\hline
+				\hline
 			\end{tabular}
@@
 \end{table}
-% \begin{table}[htpb]
-% 	\centering
-% 	\caption{Description of the linear models used to predict the CR obesity metagene, using only the sample \gls{bmi}, \gls{bmi} status and \gls{pr} pathway metagene scores}
-% 	\label{tab:lm_pr_only}
-% 	\begin{threeparttable}
-% 			\begin{tabular}{llrS}
-% 				Linear Model & Variables & Estimate &P-value\\
-% 				\hline
-% 				\hline
-% 				\rule{0pt}{2.25ex}\gls{pr}-only                            & \gls{pr}   & 0.477  & \num{6.06d-7}\tnote{1}\\
-% 				\hline
-% 				\rule{0pt}{2.25ex}\gls{bmi} and \gls{pr}                   & \gls{bmi}  & -0.002 & 0.586\\
-%                                                                            & \gls{pr}   & 0.478  & 6.35 \times10^{-7}\\
-% 				\hline
-% 				\rule{0pt}{2.25ex}\gls{bmi} status and \gls{pr}            & Obese      & -0.085 & 0.176\\
-%                                                                            & Overweight & -0.011 & 0.875\\
-%                                                                            & \gls{pr}   & 0.471  & 8.25 \times10^{-7}\\
-% 				\hline
-% 				\rule{0pt}{2.25ex}\gls{bmi}, \gls{bmi} status and \gls{pr} & \gls{bmi}  & 0.007  & 0.237\\
-%                                                                            & Obese      & -0.188 & 0.081\\
-%                                                                            & Overweight & -0.049 & 0.516\\
-%                                                                            & \gls{pr}   & 0.459  & 1.54 \times10^{-6}\\
-% 				\hline
-% 				\hline
-% 			\end{tabular}
-% 			\begin{tablenotes}
-% 				\item [1] All values in bold are statistically significant (p \textless{} 0.05).
-% 			\end{tablenotes}
-% 	\end{threeparttable}
-% \end{table}
+
+To determine whether any of the linear models were able to predict the corresponding obesity metagene scores, the predicted metagene scores were generated from the models and then compared with the original scores.
+The predictions were first made in Cris' data, where the metagene scores generated from the application of transformation matrices (from Gatza's data set) were used to make the predictions.
+The predicted metagenes were then compared with the true metagenes that were originally generated from the transformation matrices (\cref{fig:predict_cris_cris}).
+
+The predicted metagenes from the \gls{bmi} only, \gls{bmi} status only and \gls{bmi} and \gls{bmi} status linear models were not significantly related to the true metagene in many of the obesity metagenes; though \gls{bmi} and \gls{bmi} status model showed significance in some metagenes (\cref{app:b}).
+On the other hand, all of the predicted metagenes generated from the models that contained any of the pathway metagenes showed significant association with the true metagene scores in all of the obesity metagenes (\cref{fig:predict_cris_cris,app:b}).
+However, the $R^2$ values for all of the predicted metagenes against the true metagenes were very low ($R^2$ \textless{} 0.4).
+Furthermore, all of the plots showed clearly that the data points in the scatter plots were highly variable, suggesting that even though the association between the predicted and the true metagene scores were statistically significant, the linear models generated may not be as good as it looks in \cref{fig:predict_cris_cris}.
+
+\begin{figure}[htpb]
+	\centering
+	\includegraphics[page=1,width=0.32\linewidth]{results2/prediction_cris_with_cris(cris_model)}
+	\includegraphics[page=2,width=0.32\linewidth]{results2/prediction_cris_with_cris(cris_model)}
+	\includegraphics[page=3,width=0.32\linewidth]{results2/prediction_cris_with_cris(cris_model)}
+	\includegraphics[page=4,width=0.32\linewidth]{results2/prediction_cris_with_cris(cris_model)}
+	\includegraphics[page=5,width=0.32\linewidth]{results2/prediction_cris_with_cris(cris_model)}
+	\includegraphics[page=6,width=0.32\linewidth]{results2/prediction_cris_with_cris(cris_model)}
+	\includegraphics[page=7,width=0.32\linewidth]{results2/prediction_cris_with_cris(cris_model)}
+	\includegraphics[page=74,width=0.32\linewidth]{results2/prediction_cris_with_cris(cris_model)}
+	\includegraphics[page=71,width=0.32\linewidth]{results2/prediction_cris_with_cris(cris_model)}
+	\includegraphics[page=72,width=0.32\linewidth]{results2/prediction_cris_with_cris(cris_model)}
+	\includegraphics[page=73,width=0.32\linewidth]{results2/prediction_cris_with_cris(cris_model)}
+	\caption{Comparison of predicted Cr metagenes with the true Cr metagene in Cris' data}
+	\label{fig:predict_cris_cris}
+\end{figure}
+
+To confirm the results shown in Cris' data set (training set), the linear models were applied to Creighton's data (test set).
+As shown in \cref{fig:predict_cr_cris}, all of the predicted metagenes generated from the linear models were statistically significantly associated with the true metagenes, although there were some exceptions in the models for other obesity metagenes (see \cref{app:b}).
+Similar to the results from Cris' data set, all of the $R^2$ values were low ($R^2$ \textless{} 0.45) and the data points were widely spread out in the plots.
+Again, these results suggested that the predicted metagene scores from the linear models may be statistically significant, but may not be clinically significant, as the association of the predicted values were only weakly associated with the true values.
+\\
+
+\begin{figure}[htpb]
+	\centering
+	\includegraphics[page=1,width=0.32\linewidth]{results2/prediction_cr_with_cris}
+	\includegraphics[page=2,width=0.32\linewidth]{results2/prediction_cr_with_cris}
+	\includegraphics[page=3,width=0.32\linewidth]{results2/prediction_cr_with_cris}
+	\includegraphics[page=4,width=0.32\linewidth]{results2/prediction_cr_with_cris}
+	\includegraphics[page=5,width=0.32\linewidth]{results2/prediction_cr_with_cris}
+	\includegraphics[page=6,width=0.32\linewidth]{results2/prediction_cr_with_cris}
+	\includegraphics[page=7,width=0.32\linewidth]{results2/prediction_cr_with_cris}
+	\includegraphics[page=74,width=0.32\linewidth]{results2/prediction_cr_with_cris}
+	\includegraphics[page=71,width=0.32\linewidth]{results2/prediction_cr_with_cris}
+	\includegraphics[page=72,width=0.32\linewidth]{results2/prediction_cr_with_cris}
+	\includegraphics[page=73,width=0.32\linewidth]{results2/prediction_cr_with_cris}
+	\caption{Comparison of predicted Cr metagenes with the true Cr metagene in Creighton's data}
+	\label{fig:predict_cr_cris}
+\end{figure}
+
+\noindent

### 2017-02-22 thesis/results/results2.tex
tmp commit

@@
 \noindent
+Lastly, since only the ``consistent'' pathway metagenes were picked from all of the available pathway metagene scores, linear models were constructed with all of the pathway metagenes in a step wise fashion.
+Step wise construction of a linear model is where the variables for a model are selected computationally depending on the contribution of the variable to the resuting linear model, and thus contains only the variables that ``matter'' to the linear model.
+Sample \gls{bmi}, \gls{bmi} status and all of the pathway metagene scores were used to generate the step wise linear model for each of the obesity metagenes.

### 2017-02-23 thesis/results/results2.tex
Finished draft of results2 - tmp commit

@@
 However, the $R^2$ values for all of the predicted metagenes against the true metagenes were very low ($R^2$ \textless{} 0.4).
-Furthermore, all of the plots showed clearly that the data points in the scatter plots were highly variable, suggesting that even though the association between the predicted and the true metagene scores were statistically significant, the linear models generated may not be as good as it looks in \cref{fig:predict_cris_cris}.
+Furthermore, all of the plots in \cref{fig:predict_cris_cris, app:b} showed clearly that the data points in the scatter plots were highly variable, suggesting that even though the association between the predicted and the true metagene scores were statistically significant, the variables in the linear models may not be strongly associated to the obesity metagene.
 
@@
 Similar to the results from Cris' data set, all of the $R^2$ values were low ($R^2$ \textless{} 0.45) and the data points were widely spread out in the plots.
-Again, these results suggested that the predicted metagene scores from the linear models may be statistically significant, but may not be clinically significant, as the association of the predicted values were only weakly associated with the true values.
+Again, these results suggested that the predicted metagene scores from the linear models may be statistically significant, but may not be relevant to the obesity metagenes which the models predict, as the association of the predicted values were only weakly associated with the true values.
 \\
@@
 Sample \gls{bmi}, \gls{bmi} status and all of the pathway metagene scores were used to generate the step wise linear model for each of the obesity metagenes.
+The linear models that resulted from the step wise selection of the variables were summarised in \cref{app:b}.
+
+These models were used to predict the obesity metagenes in Cris' data to test the performance of the models.
+All of the models were able to make statistically significant prediction of the obesity metagenes in Cris' data set (\cref{fig:stepwise_cris}).
+Unlike the previous models, the step wise models showed very high $R^2$ values, where most predictions had $R^2$ \textgreater{} 0.85.
+The statistical significance and the strong association of the predicted metagenes were observed in Creighton's data set as well (\cref{fig:stepwise_cr}).
+Though the $R^2$ values were slightly lower than those from Cris' data set, the $R^2$ values were more promising than the non-step wise linear models.
+
+\begin{figure}[htpb]
+	\centering
+	\includegraphics[page=1,width=0.32\linewidth]{results2/stepwise_cris_cris}
+	\includegraphics[page=2,width=0.32\linewidth]{results2/stepwise_cris_cris}
+	\includegraphics[page=3,width=0.32\linewidth]{results2/stepwise_cris_cris}
+	\includegraphics[page=4,width=0.32\linewidth]{results2/stepwise_cris_cris}
+	\includegraphics[page=5,width=0.32\linewidth]{results2/stepwise_cris_cris}
+	\includegraphics[page=6,width=0.32\linewidth]{results2/stepwise_cris_cris}
+	\includegraphics[page=7,width=0.32\linewidth]{results2/stepwise_cris_cris}
+	\includegraphics[page=8,width=0.32\linewidth]{results2/stepwise_cris_cris}
+	\includegraphics[page=9,width=0.32\linewidth]{results2/stepwise_cris_cris}
+	\includegraphics[page=10,width=0.32\linewidth]{results2/stepwise_cris_cris}
+	\caption{Comparison of predicted obesity metagenes with the true obesity metagene in Cris' data}
+	\label{fig:stepwise_cris}
+\end{figure}
+
+\begin{figure}[htpb]
+	\centering
+	\includegraphics[page=1,width=0.32\linewidth]{results2/stepwise_cr_cris}
+	\includegraphics[page=2,width=0.32\linewidth]{results2/stepwise_cr_cris}
+	\includegraphics[page=3,width=0.32\linewidth]{results2/stepwise_cr_cris}
+	\includegraphics[page=4,width=0.32\linewidth]{results2/stepwise_cr_cris}
+	\includegraphics[page=5,width=0.32\linewidth]{results2/stepwise_cr_cris}
+	\includegraphics[page=6,width=0.32\linewidth]{results2/stepwise_cr_cris}
+	\includegraphics[page=7,width=0.32\linewidth]{results2/stepwise_cr_cris}
+	\includegraphics[page=8,width=0.32\linewidth]{results2/stepwise_cr_cris}
+	\includegraphics[page=9,width=0.32\linewidth]{results2/stepwise_cr_cris}
+	\includegraphics[page=10,width=0.32\linewidth]{results2/stepwise_cr_cris}
+	\caption{Comparison of predicted obesity metagenes with the true obesity metagene in Creighton's data}
+	\label{fig:stepwise_cr}
+\end{figure}
+
+Taken together, these results suggested that the variables included in these step wise models were associated with the obesity metagenes identified in Creighton's and FM's data sets.
+However, there were no single common pathway that was present in all of the models; although Akt, \gls{egfr} and \gls{pr} pathway metagenes were frequently included in many of the models.
+The role of the pathways included in the models and the possible biological links of these pathways to the obesity metagenes remain unclear, and should be the focus for future investigations.

### 2017-02-23 thesis/results/results.tex
tmp commit
Finished the last bit of results1 - tmp commit

@@
 			\hline
+			\hline
 			\rule{0pt}{2.25ex}Cr & {\bfseries 0.0134}\tnote{1} & 0.1365 & {\bfseries 0.0387} & 0.0171 & {\bfseries 0.0195} \\
@@
 		\hline
+		\hline
 		\rule{0pt}{2.25ex}\gls{blca} & 679 \\
@@
 			\hline
+			\hline
 			\rule{0pt}{2.25ex}\hspace{-1em}{\small Results from gene expression analysis} & 7330 & 2024 & 320 & 21 & 0 & 0 & 0 & 0 \\
@@
 
-\section{Pathways enriched in obesity associated genes (haven't finished)}
-\label{sec:pathways_enriched_in_obesity_associated_genes}
+\section{Pathways enriched in \gls{icgc} data sets}
+\label{sec:pathways_enriched_in_icgc_data_sets}
 
-To check whether there were any significant pathways enriched in any of the \gls{icgc} cancer data sets that were associated with obesity, pathway enrichment analysis was carried out on each cancer type separately and then all cancer types combined (see \cref{sec:pathway_enrichment_analysis}).
+To check whether there were any significant pathways enriched in any of the \gls{icgc} cancer data sets that were associated with obesity, pathway enrichment analysis was carried out on each cancer type separately and then all cancer types combined (\cref{sec:pathway_enrichment_analysis}).
+Each cancer type was normalised with voom (\cref{ssub:rna_seq_data}) and pathway enrichment analysis was carried out as described in \cref{sec:pathway_enrichment_analysis}.
+Unfortunately, there were no pathways significantly enriched in any of the \gls{icgc} cancer types when analysed independent of one another.
+
+Since there were no pathways enriched in any of the cancer types when analysed individually, all of the \gls{icgc} cancer data sets were combined to see if there were any pathways enriched in the combined data set.
+The combined data set was generated in two ways: first combined data set was created by voom normalising each cancer data set separately, combine all of the normalised cancer data sets into one, then applied batch correction to the data set (\cref{sub:batch_correction}); and the second combined data set was created by combining the cancer data sets into one first, then the combined data was voom normalised, and finally batch correction was applied to obtain the final data set.
+Each of the combined data sets were analysed for enriched pathways, but there were no pathways significantly enriched in either data sets.
+These results suggested that there were no biological pathways associated with obesity in the \gls{icgc} cancer data sets.

### 2017-02-23 thesis/results/results2.tex
Finished the last bit of results1 - tmp commit

@@
 
-One thing to be noted from these scatter plots was that there were some pathway signatures that were more variable than the others.
-As an example, the \gls{tgfb} metagenes were significantly more dispersed compared to the \gls{pr} metagenes, even though these metagenes were both generated similarly in the same data set (add figure).
+One thing noted from these scatter plots was that there were some pathway signatures that were more variable than the others.
+As an example, the \gls{tgfb} metagenes were significantly more dispersed compared to the \gls{pr} metagenes, even though these metagenes were both generated similarly in the same data set (\cref{fig:gt_rma_vs_mas}).
 These differences were seen in other data sets as well (\cref{app:b}).
@@
 \\
+
+\begin{figure}[htpb]
+	\centering
+	\includegraphics[page=75,width=0.45\linewidth]{results2/allmeta_rma_vs_mas}
+	\includegraphics[page=97,width=0.45\linewidth]{results2/allmeta_rma_vs_mas}
+	\caption{Example plot}
+	\label{fig:gt_rma_vs_mas}
+\end{figure}
 
@@
 However, the $R^2$ values for all of the predicted metagenes against the true metagenes were very low ($R^2$ \textless{} 0.4).
-Furthermore, all of the plots in \cref{fig:predict_cris_cris, app:b} showed clearly that the data points in the scatter plots were highly variable, suggesting that even though the association between the predicted and the true metagene scores were statistically significant, the variables in the linear models may not be strongly associated to the obesity metagene.
+Furthermore, all of the plots in \cref{fig:predict_cris_cris,app:b} showed clearly that the data points in the scatter plots were highly variable, suggesting that even though the association between the predicted and the true metagene scores were statistically significant, the variables in the linear models may not be strongly associated to the obesity metagene.
 
@@
 Similar to the results from Cris' data set, all of the $R^2$ values were low ($R^2$ \textless{} 0.45) and the data points were widely spread out in the plots.
-Again, these results suggested that the predicted metagene scores from the linear models may be statistically significant, but may not be relevant to the obesity metagenes which the models predict, as the association of the predicted values were only weakly associated with the true values.
+Again, these results suggested that the predicted metagene scores from the linear models may be statistically significant, but may not be relevant to the obesity metagenes which the models predict, as the predicted values were only weakly associated with the true values.
 \\

### 2017-02-24 thesis/results/results2.tex
edited all of the tables and figure captions to something a bit more useful - tmp commit

@@
 Results shown in \cref{app:b} clarified that there was no significant difference in the ranking methods used.
-Correlation and scatter plots of the different combinations of transformation matrices (derived from either \gls{rma}- or \gls{mas}-normalised data) with \gls{rma}- or \gls{mas}-normalised data.
-It seemed like the normalisation methods of the data in which the transformation matrices were applied to had the most significant effect on the resulting metagenes, rather than the transformation matrices themselves (\cref{app:b}).
-This meant that all of the data sets had to be normalised with a single normalisation method, so \gls{rma} normalisation method was used as this method was more reliable than the \gls{mas} method.
+Correlation and scatter plots of the different combinations of transformation matrices (derived from either \gls{rma}- or \gls{mas}-normalised data) with \gls{rma}- or \gls{mas}-normalised data were calculated and plotted.
+From these results, it seemed like the normalisation methods of the data in which the transformation matrices were applied to had the most significant effect on the resulting metagenes, rather than the transformation matrices themselves (\cref{app:b}).
+This meant that all of the data sets had to be normalised with a single normalisation method, so the \gls{rma} normalisation method was used as this method was more reliable than the \gls{mas} method.
 Therefore, Gatza's data set was batch corrected first (\cref{sub:batch_correction}), then normalised with \gls{rma} method and the scores were ranked based on the number of samples in the data set.
@@
 	\centering
-	\includegraphics[page=75,width=0.45\linewidth]{results2/allmeta_rma_vs_mas}
-	\includegraphics[page=97,width=0.45\linewidth]{results2/allmeta_rma_vs_mas}
-	\caption{Example plot}
+	\includegraphics[page=76,width=0.45\linewidth]{results2/allmeta_rma_vs_mas}
+	\includegraphics[page=100,width=0.45\linewidth]{results2/allmeta_rma_vs_mas}
+	\caption{Comparison of the metagenes generated in the \gls{mas} normalised Gatza \textit{et al.} data set from the application of the transformation matrices derived from either the \gls{rma} or \gls{mas} normalised data}
 	\label{fig:gt_rma_vs_mas}
@@
 	\includegraphics[page=15,width=0.7\linewidth]{results2/gatza_meta_trans}
-	\caption{Gatza metagene direction}
+	\caption{Heatmap of all the pathway metagenes from the \citet{Gatza2010a} study in Gatza \textit{et al.} data set}
 	\label{fig:gatza_meta_dir}
@@
 	\centering
-	\caption{Spearman correlation of the svd vs tm metagenes (pathway)}
+	\begin{threeparttable}
+	\caption{Summary of the Spearman correlations of the pathway metagenes generated from \gls{svd} and the application of the transformation matrices\tnote{1} in Gatza \textit{et al.}, Creighton \textit{et al.}, Print \textit{et al.} and Fuentes-Mattei \textit{et al.} data sets}
 	\label{tab:svd_vs_tm_path}
@@
 		\end{tabular}
+		\begin{tablenotes}
+			\begin{footnotesize}
+				\item [1] Transformation matrices were derived from the Gatza \textit{et al.} data set.
+			\end{footnotesize}
+		\end{tablenotes}
+	\end{threeparttable}
 \end{table}
@@
 	\centering
-	\caption{Spearman correlation of the svd vs tm metagenes (obesity)}
+	\begin{threeparttable}
+		\caption{Summary of the Spearman correlations of the obesity metagenes generated from \gls{svd} and the application of the transformation matrices\tnote{1} in Gatza \textit{et al.}, Creighton \textit{et al.}, Print \textit{et al.} and Fuentes-Mattei \textit{et al.} data sets}
 		\label{tab:svd_vs_tm_obs}
@@
 		\end{tabular}
+			\begin{tablenotes}
+				\begin{footnotesize}
+				\item [1] Transformation matrices were derived from the Creighton \textit{et al.} data set.
+				\end{footnotesize}
+			\end{tablenotes}
+	\end{threeparttable}
 \end{table}
@@
 	\includegraphics[page=16,width=0.45\linewidth]{results2/allmetacor_bar}
-	\caption{Visual comparison of the correlation of the metagenes across different data sets}
+	\caption{Visual comparison of the Spearman correlations of \gls{svd}-derived and transformation matrix-derived \gls{bcat} and \gls{stat3} metagenes across different data sets}
 	\label{fig:allmetacor_bar}
@@
 	\includegraphics[page=15,width=0.8\linewidth]{results2/all_meta_trans}
-	\caption{All metagenes in Gatza's data}
+	\caption{Heatmap of all the pathway and obesity metagenes in Gatza \textit{et al.} data set}
 	\label{fig:gatza_allmeta}
@@
 	\centering
-	\caption{Description of the linear models used to predict the CR obesity metagene}
+	\caption{Description of the linear models used to predict the CR obesity metagene in Print \textit{et al.} data set}
 	\label{tab:lm_sig_var}
@@
 		\begin{tablenotes}
+			\begin{footnotesize}
 				\item [1] All values in bold are statistically significant (p \textless{} 0.05).
+			\end{footnotesize}
 		\end{tablenotes}
@@
 	\centering
-	\caption{Description of the linear models used to predict the CR obesity metagene, using only the sample \gls{bmi}, \gls{bmi} status and \gls{pr} pathway metagene scores}
+	\caption{Description of the linear models used to predict the CR obesity metagene in Print \textit{et al.} data set, using only the sample \gls{bmi}, \gls{bmi} status and \gls{pr} pathway metagene scores}
 	\label{tab:lm_pr_only}
@@
 			\begin{tablenotes}
+				\begin{footnotesize}
 					\item [1] All values in bold are statistically significant (p \textless{} 0.05).
+				\end{footnotesize}
 			\end{tablenotes}
@@
 	\includegraphics[page=73,width=0.32\linewidth]{results2/prediction_cris_with_cris(cris_model)}
-	\caption{Comparison of predicted Cr metagenes with the true Cr metagene in Cris' data}
+	\caption{Comparison of the predicted Cr metagenes with the true Cr metagene in the Print \textit{et al.} data}
 	\label{fig:predict_cris_cris}
@@
 	\includegraphics[page=73,width=0.32\linewidth]{results2/prediction_cr_with_cris}
-	\caption{Comparison of predicted Cr metagenes with the true Cr metagene in Creighton's data}
+	\caption{Comparison of the predicted Cr metagenes with the true Cr metagene in the Creighton \textit{et al.} data}
 	\label{fig:predict_cr_cris}
@@
 	\includegraphics[page=10,width=0.32\linewidth]{results2/stepwise_cris_cris}
-	\caption{Comparison of predicted obesity metagenes with the true obesity metagene in Cris' data}
+	\caption{Comparison of the predicted obesity metagenes with the true obesity metagene in Print \textit{et al.} data}
 	\label{fig:stepwise_cris}
@@
 	\includegraphics[page=10,width=0.32\linewidth]{results2/stepwise_cr_cris}
-	\caption{Comparison of predicted obesity metagenes with the true obesity metagene in Creighton's data}
+	\caption{Comparison of the predicted obesity metagenes with the true obesity metagene in Creighton \textit{et al.} data}
 	\label{fig:stepwise_cr}

### 2017-02-26 thesis/thesis.tex
tmp commit
Finished the last bit of results1 - tmp commit

@@
 \usepackage{siunitx}                                % package for scientific notations and units
-\usepackage{bm}                                % package for scientific notations and units
+\usepackage{etoolbox}                               % package to 'robustify' \bfseries
 
@@
 % Make siunitx detect bold/italics/etc:
-\sisetup{detect-weight=true,detect-family=true,detect-inline-weight=text,table-parse-only}
+\robustify\bfseries
+\sisetup{detect-all}
+
+% Set the width of captions for threeparttable to \linewidth:
+\captionsetup[threeparttable]{width=\linewidth}
 
@@
 % If you want to compile specific files, uncomment the line below and add the filenames in the {} :
-\includeonly{{./methods/methods},{./results/results},{./results/results2},{./misc/appendix}}
+\includeonly{{./methods/methods},{./results/results},{./results/results2},{./discussion/discussion},{./misc/appendix}}
 % \includeonly{{./results/results},{./results/results2},{./misc/appendix}}
+% \includeonly{{./results/results},{./results/results2},{./discussion/discussion},{./misc/appendix}}
 

### 2017-02-27 thesis/discussion/discussion.tex
tmp commit
Finished FM metagene section (but very rough)
finished pathway and obesity metagene section - tmp commit
Finished draft of results2 - tmp commit

@@
-\chapter{Discussion (draft)}
+\chapter{Discussion}
+\label{cha:discussion}
 
-\section{something}
+To conclude this thesis, all of the results were re-examined and summarised to highlight the key findings from this project.
+%TODO: add a little more content to the next sentence
+There were many limitations to this project and these limitations were adressed in \cref{sec:limitations}.
+Lastly, a conclusion was drawn from all of the evidences from this project and suggestions for future aims and experiments were made.
 
-% List of things to discuss
+\section{Summary of the results}
+\label{sec:summary_of_the_results}
+
+% TODO: -- Recap work
+% TODO: -- Put into context of very current literature
+% TODO:   	==> How does this project fit into the literature
+
+There were two main aims for this project: firstly to determine whether there were any obesity specific genetic signatures that could be tranferred across multiple cancer types; and secondly to investigate whether there were any biological pathways dysregulated in the samples that were obese compared to the smaples that were not obese.
+These aims were addressed in \cref{cha:obesity_genetic_signatures_and_cancer,cha:obesity_associated_genetic_signature_and_pathway_signatures}, resepectively.
+\\
+
+\noindent
+\cref{cha:obesity_genetic_signatures_and_cancer} focussed mainly on establishing the link between the obesity associated genetic signatures from the \citet{Creighton2012} and the \citet{Fuentes-Mattei2014} studies with the sample \gls{bmi} and \gls{bmi} status, as reported by their studies.
+The obesity metagene generated from the obesity signature from the \citet{Creighton2012} study was able to ``capture'' the overall genetic expression patterns of the samples and significantly correlated with the sample \gls{bmi} and \gls{bmi} status in the CR data set.
+This result was in accordance with the description provided by Creighton \textit{et al.}.
+
+To see whether this obesity signature was able to show similar association with the sample \gls{bmi} and \gls{bmi} status in other cancer data sets, metagenes were created in each of the cancer data sets with the transformation matrix generated in the Creighton \textit{et al.} data set.
+The obesity metagenes were reflective of the gene expression patterns of the genes in the Creighton \textit{et al.} genetic signature where low and high metagene scores  corresponded to low and high gene expressions, respectively.
+Although obesity metagene scores reflected the gene expression patterns, the metagene was not significantly associated with the sample \gls{bmi}/\gls{bmi} status in any of the other cancer data sets, except for  the \gls{blca} data set which showed significant association only with the overweight group (discussed later in this section).
+
+Initially, these results were thought to be due to the obesity associated genetic signature being specific only to the Creighton \textit{et al.} data set.
+Therefore, a different obesity associated genetic signature from the \citet{Fuentes-Mattei2014} study was used to see whether this signature was able to show significant association with the sample \gls{bmi} and \gls{bmi} status in other cancer data sets.
+The results were similar to the obesity metagene from the \citet{Creighton2012} study; obesity metagene scores reflected the gene expression patterns but the scores were not significantly associated with the sample \gls{bmi}/\gls{bmi} status in any of the cancer data sets (except the \gls{blca} data set).
+Together with the results from CR obesity metagene, these results suggested that both CR and FM genetic signatures were only significant in the data set in which the signature was derived in (though this was not confirmed in the Fuentes-Mattei \textit{et al.} data set, as no sample \gls{bmi} information was available).
+
+The fact that both obesity associated genetic signatures identified by Creighton \textit{et al.} and Fuentes-Mattei \textit{et al.} showed no sign of significant association with sample \gls{bmi}/\gls{bmi} status in majority of the cancer data sets raised a question of whether these signatures were actually related to obesity, or a different clinical variable within the data set.
+This question led to the identification of the various versions of obesity associated genetic signatures in \cref{sub:identification_of_obesity_associated_genetic_signatures} that were controlled for clinical variables in the Creighton \textit{et al.} data set.
+All of the metagenes created from these clinical variable-controlled genetic signatures were consistent with the gene expression patterns but none of the metagenes showed significant association with the sample \gls{bmi}/\gls{bmi} status in the data set other than the Creighton \textit{et al.} data set, with the exception of \gls{blca} data set.
+
+In the \gls{blca} data set, many of the metagenes were associated with the overweight group with significant P-value and/or \gls{anova} P-value, but never with the obese group.
+This was unexpected, as all of the metagenes were generated based on the list of \glspl{deg} between the obese and the non-obese groups in the Creighton \textit{et al.} data set, and not the overweight group.
+In addition to the association of the metagenes with the overweight group, some metagenes showed significant association with sample \gls{bmi} in the \gls{blca} data set.
+It was difficult to conclude with confidence that those metagenes that showed significant association with the overweight group in \gls{blca} data set was truly due to the effect of the metagenes, as these metagenes were generated from the obese group and not the overweight group in the original data sets,  and there were no significant association of these metagenes in any other \gls{icgc} cancer data sets either.
+With that said, there is a possibility that the genotypes of the samples that are overweight in \gls{blca} data set are similar to the genotypes of the samples that are obese in the Creighton \textit{et al.} data set.
+
+\section{Limitations}
+\label{sec:limitations}
+
+% -- Limitations of this project
 
@@
 
-Lack of association of obesity metagenes with different cancer data sets: due to metagene derivation from a specific data
+% Creighton metagene section
+% Lack of association of obesity metagenes with different cancer data sets
+% -- due to metagene derivation from a specific data and therefore cannot be applied to other data set
+% -- don't forget about the possibility of having samples that may not be driven by obesity at all in the data set (and therefore the metagene may be affected by these samples).
 
-\begin{figure}[ht]
-\caption{Dummy}
-\label{fig:dum1}
-\end{figure}
+% FM metagene section
+% -- don't forget about the possibility of having samples that may not be driven by obesity at all in the data set (and therefore the metagene may be affected by these samples).
 
-\begin{table}[ht]
-\caption{Dummy}
-\begin{tabular}{|c|c|}
-\end{tabular}
-\label{tab:dum1}
-\end{table}
+% gene expression section
+% -- mention that Creighton's group probably didn't control for FDR
+
+% RESULTS 2
+
+% Pathway associated genetic signature directions
+
+% Pathway and obesity associated genetic signatures
+% -- some metagenes had consistently high correlation across different data sets.
+%		-- showed that these signatures were more reliable than the others
+%		-- the qualities of these signatures were good as they showed similar metagene scores across different data sets
+%			-- quality of the signatures may be due to the difference in experimental conditions, or the samples used by Gatza et al...?
+
+% -- FM metagene not clustering with AKT
+%		-- due to shitty akt pathway signature, or difference in mouse model.
+
+% Linear model prediction stuff
+% -- PR may have been significant only because it was from Cris' data?
+
+\section{Conclusion}
+\label{sec:conclusion}
+
+\section{Future directions}
+\label{sec:future_directions}
+
+% -- future directions

### 2017-02-27 thesis/discussion/discussion.tex
tmp commit

@@
 \cref{cha:obesity_genetic_signatures_and_cancer} focussed mainly on establishing the link between the obesity associated genetic signatures from the \citet{Creighton2012} and the \citet{Fuentes-Mattei2014} studies with the sample \gls{bmi} and \gls{bmi} status, as reported by their studies.
-The obesity metagene generated from the obesity signature from the \citet{Creighton2012} study was able to ``capture'' the overall genetic expression patterns of the samples and significantly correlated with the sample \gls{bmi} and \gls{bmi} status in the CR data set.
-This result was in accordance with the description provided by Creighton \textit{et al.}.
+The obesity associated genetic signature from the \citet{Creighton2012} study was examined first.
+The obesity metagene generated from the obesity signature from the \citet{Creighton2012} study was able to ``capture'' the overall genetic expression patterns of the samples, where low and high metagene scores  corresponded to low and high gene expressions, respectively.
+This result was in accordance with the characteristic of the obesity metagene provided by Creighton \textit{et al.}.
+Furthermore, the obesity metagene significantly correlated with the sample \gls{bmi} and \gls{bmi} status in the CR data set.
 
 To see whether this obesity signature was able to show similar association with the sample \gls{bmi} and \gls{bmi} status in other cancer data sets, metagenes were created in each of the cancer data sets with the transformation matrix generated in the Creighton \textit{et al.} data set.
-The obesity metagenes were reflective of the gene expression patterns of the genes in the Creighton \textit{et al.} genetic signature where low and high metagene scores  corresponded to low and high gene expressions, respectively.
+Like in Creighton \textit{et al.} data set, the obesity metagenes were reflective of the gene expression patterns of the genes in the genetic signature.
 Although obesity metagene scores reflected the gene expression patterns, the metagene was not significantly associated with the sample \gls{bmi}/\gls{bmi} status in any of the other cancer data sets, except for  the \gls{blca} data set which showed significant association only with the overweight group (discussed later in this section).
@@
 The results were similar to the obesity metagene from the \citet{Creighton2012} study; obesity metagene scores reflected the gene expression patterns but the scores were not significantly associated with the sample \gls{bmi}/\gls{bmi} status in any of the cancer data sets (except the \gls{blca} data set).
-Together with the results from CR obesity metagene, these results suggested that both CR and FM genetic signatures were only significant in the data set in which the signature was derived in (though this was not confirmed in the Fuentes-Mattei \textit{et al.} data set, as no sample \gls{bmi} information was available).
+Together with the results from CR obesity metagene, these results suggested that both CR and FM genetic signatures were only significant in the original data set in which the signature was derived in (though this was not confirmed in the Fuentes-Mattei \textit{et al.} data set, as no sample \gls{bmi} information was available).
 
 The fact that both obesity associated genetic signatures identified by Creighton \textit{et al.} and Fuentes-Mattei \textit{et al.} showed no sign of significant association with sample \gls{bmi}/\gls{bmi} status in majority of the cancer data sets raised a question of whether these signatures were actually related to obesity, or a different clinical variable within the data set.
-This question led to the identification of the various versions of obesity associated genetic signatures in \cref{sub:identification_of_obesity_associated_genetic_signatures} that were controlled for clinical variables in the Creighton \textit{et al.} data set.
-All of the metagenes created from these clinical variable-controlled genetic signatures were consistent with the gene expression patterns but none of the metagenes showed significant association with the sample \gls{bmi}/\gls{bmi} status in the data set other than the Creighton \textit{et al.} data set, with the exception of \gls{blca} data set.
+This question led to the identification of the various versions of obesity associated genetic signatures in \cref{sub:identification_of_obesity_associated_genetic_signatures} that were controlled for clinical variables (sex, age, ethnicity, menopause status, tumour grade, \gls{er}/\gls{her2}/\gls{pr} statuses and \gls{ln} status) in the Creighton \textit{et al.} data set.
+All of the metagenes created from these clinical variable-controlled genetic signatures were consistent with the gene expression patterns but none of the metagenes showed significant association with the sample \gls{bmi}/\gls{bmi} status in the data set other than the Creighton \textit{et al.} data set, with the exception of the overweight group in the \gls{blca} data set.
 
@@
 In addition to the association of the metagenes with the overweight group, some metagenes showed significant association with sample \gls{bmi} in the \gls{blca} data set.
-It was difficult to conclude with confidence that those metagenes that showed significant association with the overweight group in \gls{blca} data set was truly due to the effect of the metagenes, as these metagenes were generated from the obese group and not the overweight group in the original data sets,  and there were no significant association of these metagenes in any other \gls{icgc} cancer data sets either.
-With that said, there is a possibility that the genotypes of the samples that are overweight in \gls{blca} data set are similar to the genotypes of the samples that are obese in the Creighton \textit{et al.} data set.
+It was difficult to conclude with confidence that those metagenes that showed significant association with the overweight group in \gls{blca} data set was truly due to the effect of the metagenes, as these metagenes were generated from the obese group and not the overweight group in the original data sets.
+Furthermore, there were no significant association of these metagenes in any other \gls{icgc} cancer data sets and provided no further evidence that supported the results presented in the \gls{blca} data set.
+
+% TODO: search for any evidence of BLCA with breast cancers - add to this paragraph
+With that said, there is a possibility that the genotypes of the samples that are overweight in \gls{blca} data set are similar to those samples that are obese in the Creighton \textit{et al.} data set.
+\textit{(TODO: search for any evidence of relationship/connection between BLCA and breast cancers - add to this paragraph)}
+
+The results so far have provided strong evidence that the obesity associated genetic signatures from the Creighton \textit{et al.} and Fuentes-Mattei \textit{et al.} data sets were not associated with the sample \gls{bmi}/\gls{bmi} status.
+In the last attempt to identify any genes that were common across multiple cancer types that were also associated with obesity, gene expression analyses were carried out on the \gls{icgc} cancer data sets (\cref{sec:common_genes_across_multiple_cancer_types}).
+There were many \glspl{deg} identified by each of the \gls{icgc} cancer types, though there were no genes in common for all eight cancer types.
+The results from the simulation have confirmed this observation where there were no \glspl{deg} associated with obesity that were common across all eight cancer types.
+
+The simulation results also showed that the number of \glspl{deg} identified in these cancer types were greater than one would expect by chance.
+However, as observed by the results from the simulation, there were many genes identified by chance alone, which suggested that many of the genes found were false positives.
 

### 2017-02-28 thesis/discussion/discussion.tex
tmp commit

@@
 
-There were two main aims for this project: firstly to determine whether there were any obesity specific genetic signatures that could be tranferred across multiple cancer types; and secondly to investigate whether there were any biological pathways dysregulated in the samples that were obese compared to the smaples that were not obese.
+There were two main aims to this project: firstly to determine whether there were any obesity specific genetic signatures that could be tranferred across multiple cancer types; and secondly to investigate whether there were any biological pathways dysregulated in the samples that were obese compared to the smaples that were not obese.
 These aims were addressed in \cref{cha:obesity_genetic_signatures_and_cancer,cha:obesity_associated_genetic_signature_and_pathway_signatures}, resepectively.
@@
 The simulation results also showed that the number of \glspl{deg} identified in these cancer types were greater than one would expect by chance.
-However, as observed by the results from the simulation, there were many genes identified by chance alone, which suggested that many of the genes found were false positives.
+However, as observed by the results from the simulation, there were many genes identified by chance alone, which suggested that majority of the \glspl{deg} found in the eight cancer types were false positives.
+This apparently high level of \Glspl{type1} was not only observed in the \gls{icgc} data sets, but also in the Creighton \textit{et al.} data set when gene expression analyses were carried out to generate various obesity associated genetic signatures in \cref{sub:identification_of_obesity_associated_genetic_signatures}.
+
+The fact that there were so many false positives in the \glspl{deg} from the \gls{icgc} data sets indicated that the use of the sample \gls{bmi}/\gls{bmi} status as the variable for gene expression analysis was prone to \Glspl{type1}.
+This meant that there were very little evidence of true obesity associated genetic signatures in the data sets that have been explored in this project, or perhaps the sample \gls{bmi}/\gls{bmi} status was not enough to pin point the underlying biological relationship between obesity and cancer.
+Even if \gls{bmi} and/or \gls{bmi} status was enough to identify the genes that were truly associated with obesity, it would not be a trivial task to identify the truly obesity associated genes from those that were not.
+This raises an important question of whether the sample \gls{bmi}/\gls{bmi} status was enough to uncover the true association between obesity and cancer (adressed in \cref{sec:limitations,sec:future_directions}).
+
+Finally, pathway enrichment analysis was carried out in the \gls{icgc} data sets to investigate whether there were any pathways that significantly associated with obesity (\cref{sec:pathways_enriched_in_icgc_data_sets}).
+There were no pathways enriched in any of the \gls{icgc} data sets, nor the combined \gls{icgc} data set.
+Even though there was no significantly enriched  pathway in the \gls{icgc} data sets, it was evident from the obesity metagenes from Creighton \textit{et al.} and Fuentes-Mattei \textit{et al.} data sets that these metagenes were associated with the gene expressions.
+This suggested that the metagenes were able to identify some sort of genetic pattern in the data, even though these metagenes were not significantly associated with obesity.
+\\
+
+\noindent
+In \cref{cha:obesity_associated_genetic_signature_and_pathway_signatures}, analyses were carried out to determine what the genetic pattern the obesity metagenes from \cref{cha:obesity_genetic_signatures_and_cancer} were associating with.
 

### 2017-02-28 thesis/discussion/discussion.tex
tmp commit

@@
 In \cref{cha:obesity_associated_genetic_signature_and_pathway_signatures}, analyses were carried out to determine what the genetic pattern the obesity metagenes from \cref{cha:obesity_genetic_signatures_and_cancer} were associating with.
+The pathway associated genetic signatures from the \citet{Gatza2010a} study were used to establish the biological connection with the obesity associated genetic signatures.
+In order to make the obesity and pathway associated genetic signatures comparable with one another, the orientations of the metagenes were examined so that all of the metagenes were ``correct'' (\cref{sec:pathway_associated_genetic_signatures_from_gatza2010a_study}).
+
+Once the directions of the metagenes were determined, the obesity metagenes were clustered together in the heatmap to visualise the similarity of the obesity metagenes with the pathway metagenes.
+The heatmap revealed that all of the obesity metagenes that have originated from the Creighton \textit{et al.} study clustered together in a group with no other pathway metagenes correlating with the metagenes, which meant that the obesity metagenes were not similar to any of the pathway metagenes from the \citet{Gatza2010a} study.
+The clustering of the obesity metagene was not surprising, as the results from \cref{sub:_novel_obesity_associated_signatures_and_sample_bmi} already showed that the metagenes were highly correlated with one another.
+This result showed that all of the obesity metagenes created from the Creighton \textit{et al.} data associated with an unknown genetic signature that was not related to any of the pathway associated genetic signatures from the \citet{Gatza2010a} study.
+
+The obesity metagene from the \citet{Fuentes-Mattei2014} study was a cluster on its own, where the metagene did not cluster with any of the Creighton \textit{et al.} obesity metagenes nor with the Gatza \textit{et al.} pathway metagene.
+Though Fuentes-Mattei \textit{et al.} suggested a biological link with the Akt/\gls{mtor} pathway in their study, the result from the heatmap showed otherwise.
 

### 2017-03-01 thesis/discussion/discussion.tex
finished rough draft of the result summary section in discussion - tmp commit

@@
 
-There were two main aims to this project: firstly to determine whether there were any obesity specific genetic signatures that could be tranferred across multiple cancer types; and secondly to investigate whether there were any biological pathways dysregulated in the samples that were obese compared to the smaples that were not obese.
+There were two main aims to this project: firstly to determine whether there were any obesity specific genetic signatures that could be tranferred across multiple cancer types; and secondly to investigate whether there were any biological pathways dysregulated in the samples that were obese compared to the samples that were not obese.
 These aims were addressed in \cref{cha:obesity_genetic_signatures_and_cancer,cha:obesity_associated_genetic_signature_and_pathway_signatures}, resepectively.
@@
 
-The obesity metagene from the \citet{Fuentes-Mattei2014} study was a cluster on its own, where the metagene did not cluster with any of the Creighton \textit{et al.} obesity metagenes nor with the Gatza \textit{et al.} pathway metagene.
+The obesity metagene from the \citet{Fuentes-Mattei2014} study was a cluster of its own, where the metagene did not group with any of the Creighton \textit{et al.} obesity metagenes nor with any of the Gatza \textit{et al.} pathway metagene.
 Though Fuentes-Mattei \textit{et al.} suggested a biological link with the Akt/\gls{mtor} pathway in their study, the result from the heatmap showed otherwise.
+This may have been due to the inconsistency of the pathway metagene scores across different data as mentioned in \cref{sec:pathway_associated_metagenes_and_obesity_associated_metagenes}.
+In fact, the Akt pathway metagene scores were variable across different cancer data sets (\cref{app:b}), which suggested the Akt pathway genetic signature was of poor quality.
+
+Though how unreliable the Akt pathway metagene scores may be, there is still a possibility that the FM obesity associated genetic signature may be unrelated to the Akt pathway, as Fuentes-Mattei \textit{et al.} validated their human breast tumour-derived genetic signature in a mouse model \citep{Fuentes-Mattei2014}.
+Mouse models are indeed helpful in studying biological processes \textit{in vivo}, especially when the cause and the outcome of the disease is well established.
+However, cancer is a complex genetic disorder which may be difficult to model accurately in mice.
+Furthermore, obesity is a disease that have a variety of biological effects (for example the release of adipokines and the disruption of the hormone levels) that could be acting in a different way in mice compared to humans, which further complicates the causal relationship between obesity and cancer.
+Nevertheless, the evidence shown by \citet{Fuentes-Mattei2014} was convincing and the role of Akt/\gls{mtor} pathway in obesity and cancer should be explored in greater details.
+(include this paragraph or not... tbh, I think this is a weak argument)
+
+To further investigate whether there were any evidence of a pathway signature being associated with the obesity associated genetic signatures, linear models were created in Cris' data set with the pathway metagene scores.
+First, linear models were created with the sample \gls{bmi}, \gls{bmi} status and a selection of the most ``consistent'' pathway metagene scores (based on the consistency across different data sets).
+From these linear models, \gls{pr} pathway metagene scores showed significant contribution to the model in predicting the obesity associated metagene scores in all of the linear models (apart from the FM obesity metagene).
+Therefore, linear models were also created with \gls{pr} pathway metagene in combination with the sample \gls{bmi}/\gls{bmi} status.
+
+The predictions of the obesity metagene scores were made using the linear models created in Cris' data set.
+All of the predicted obesity metagene scores significantly correlated with the true obesity metagene scores in both Cris' and Creighton's data sets.
+However, the $R^2$-values for all of these predictions were low (highest $R^2$ $\approx$ 0.42) with highly variable data points which suggested that, although the predictions reflected the true values significantly, the variables included in the linear models may not be as relevant to the obesity metagenes as it may seem from these plots.
+
+Since linear models were created only with the ``consistent'' pathway metagene scores, a step wise method was used to generate linear models that included all of the variables that allowed the model to predict the true obesity metagene the best.
+As shown by the results in \cref{sec:prediction_of_obesity_associated_metagene_with_pathway_associate_metagene}, all of the models predicted the true obesity metagene scores with high $R^2$-value in Creighton's data set (highest $R^2$ $\approx$ 0.88) and were statistically significant.
+This meant that the variables that were significant in the linear models played a role in the obesity associated genetic signatures.
+However, out of all of the step wise linear models created, there was no single variable that was common to all of the linear models; although Akt, \gls{egfr} and \gls{pr} were included in many of the models, which suggested that these pathway signatures may have an important role in the obesity associated genetic signatures.
+
+With that said, even if these variables were related to the obesity associated genetic signatures, the fact that these results depend on the pathway associated genetic signatures from the \citet{Gatza2010a} study makes the results questionable.
+This was due to the lack of consistency of the metagenes across different cancer data sets, where more than half of the pathway metagenes showed a correlation of less than 0.8 in at least one data set (\cref{sec:pathway_associated_metagenes_and_obesity_associated_metagenes,app:b}), which meant that at least half of the signatures were specific to the Gatza \textit{et al.} data set.
+Thus, without further validation of the quality of the pathway associated genetic signatures and verification of the roles of these signatures in the biological relationship between obesity and cancer, these results should be interpreted with great caution.
 

### 2017-03-01 thesis/results/results2.tex
tmp commit
finished rough draft of the result summary section in discussion - tmp commit

@@
 
-\section{Prediction of obesity associated metagene with pathway associate metagene}
+\section{Prediction of obesity associated metagenes with pathway associate metagenes}
 \label{sec:prediction_of_obesity_associated_metagene_with_pathway_associate_metagene}
@@
 As shown in \cref{fig:predict_cr_cris}, all of the predicted metagenes generated from the linear models were statistically significantly associated with the true metagenes, although there were some exceptions in the models for other obesity metagenes (see \cref{app:b}).
-Similar to the results from Cris' data set, all of the $R^2$ values were low ($R^2$ \textless{} 0.45) and the data points were widely spread out in the plots.
+Similar to the results from Cris' data set, all of the $R^2$ values were low ($R^2$ \textless{} 0.43) and the data points were widely spread out in the plots.
 Again, these results suggested that the predicted metagene scores from the linear models may be statistically significant, but may not be relevant to the obesity metagenes which the models predict, as the predicted values were only weakly associated with the true values.

### 2017-03-01 thesis/discussion/discussion.tex
minor rearrangement of the discussion section - tmp commit

@@
 
-There were two main aims to this project: firstly to determine whether there were any obesity specific genetic signatures that could be tranferred across multiple cancer types; and secondly to investigate whether there were any biological pathways dysregulated in the samples that were obese compared to the samples that were not obese.
-These aims were addressed in \cref{cha:obesity_genetic_signatures_and_cancer,cha:obesity_associated_genetic_signature_and_pathway_signatures}, resepectively.
-\\
+\subsection{Obesity associated genetic signatures}
+\label{sub:obesity_associated_genetic_signatures}
 
-\noindent
 \cref{cha:obesity_genetic_signatures_and_cancer} focussed mainly on establishing the link between the obesity associated genetic signatures from the \citet{Creighton2012} and the \citet{Fuentes-Mattei2014} studies with the sample \gls{bmi} and \gls{bmi} status, as reported by their studies.
@@
 Like in Creighton \textit{et al.} data set, the obesity metagenes were reflective of the gene expression patterns of the genes in the genetic signature.
-Although obesity metagene scores reflected the gene expression patterns, the metagene was not significantly associated with the sample \gls{bmi}/\gls{bmi} status in any of the other cancer data sets, except for  the \gls{blca} data set which showed significant association only with the overweight group (discussed later in this section).
+Although obesity metagene scores reflected the gene expression patterns, the metagene was not significantly associated with the sample \gls{bmi}/\gls{bmi} status in any of the other cancer data sets, except for  the \gls{blca} data set which showed significant association only with the overweight group (discussed later in \cref{sub:blca_and_obesity_metagenes}).
 
@@
 
-In the \gls{blca} data set, many of the metagenes were associated with the overweight group with significant P-value and/or \gls{anova} P-value, but never with the obese group.
+\subsection{\gls{blca} data set and obesity metagenes}
+\label{sub:blca_and_obesity_metagenes}
+
+In the \gls{blca} data set, many of the obesity metagenes were associated with the overweight group with significant P-value and/or \gls{anova} P-value, but never with the obese group.
 This was unexpected, as all of the metagenes were generated based on the list of \glspl{deg} between the obese and the non-obese groups in the Creighton \textit{et al.} data set, and not the overweight group.
@@
 \textit{(TODO: search for any evidence of relationship/connection between BLCA and breast cancers - add to this paragraph)}
+
+\subsection{Common obesity associated genes and pathways across multiple cancer types}
+\label{sub:common_obesity_associated_genes_and_pathways_across_multiple_cancer_types}
 
@@
 
-The fact that there were so many false positives in the \glspl{deg} from the \gls{icgc} data sets indicated that the use of the sample \gls{bmi}/\gls{bmi} status as the variable for gene expression analysis was prone to \Glspl{type1}.
-This meant that there were very little evidence of true obesity associated genetic signatures in the data sets that have been explored in this project, or perhaps the sample \gls{bmi}/\gls{bmi} status was not enough to pin point the underlying biological relationship between obesity and cancer.
-Even if \gls{bmi} and/or \gls{bmi} status was enough to identify the genes that were truly associated with obesity, it would not be a trivial task to identify the truly obesity associated genes from those that were not.
-This raises an important question of whether the sample \gls{bmi}/\gls{bmi} status was enough to uncover the true association between obesity and cancer (adressed in \cref{sec:limitations,sec:future_directions}).
-
 Finally, pathway enrichment analysis was carried out in the \gls{icgc} data sets to investigate whether there were any pathways that significantly associated with obesity (\cref{sec:pathways_enriched_in_icgc_data_sets}).
@@
 This suggested that the metagenes were able to identify some sort of genetic pattern in the data, even though these metagenes were not significantly associated with obesity.
-\\
 
-\noindent
+\subsection{False positives in gene expression analyses}
+\label{sub:false_positives_in_gene_expression_analyses}
+
+The fact that there were so many false positives in the \glspl{deg} from the \gls{icgc} data sets indicated that the use of the sample \gls{bmi}/\gls{bmi} status as the ``treatment'' conditions for gene expression analysis was prone to \Glspl{type1}.
+This meant that there were very little evidence of true obesity associated genetic signatures in any of the data sets that have been explored in this project, or perhaps the sample \gls{bmi}/\gls{bmi} status was not enough to pinpoint the underlying biological relationship between obesity and cancer.
+Even if the sample \gls{bmi} and/or \gls{bmi} status was enough to identify the genes that were truly associated with obesity, it would not be a trivial task to identify the truly obesity associated genes from those that were not.
+This raises an important question of whether the sample \gls{bmi}/\gls{bmi} status was appropriate to uncover the true association between obesity and cancer (adressed in \cref{sec:limitations,sec:future_directions}).
+
+\textit{(need to add more content here...?)}
+
+\subsection{Genetic signature captured by the obesity metagenes}
+\label{sub:genetic_signature_captured_by_the_obesity_metagenes}
+
 In \cref{cha:obesity_associated_genetic_signature_and_pathway_signatures}, analyses were carried out to determine what the genetic pattern the obesity metagenes from \cref{cha:obesity_genetic_signatures_and_cancer} were associating with.
@@
 
+\subsection{Linear models to predict obesity metagenes}
+\label{sub:linear_models_to_predict_obesity_metagenes}
+
 To further investigate whether there were any evidence of a pathway signature being associated with the obesity associated genetic signatures, linear models were created in Cris' data set with the pathway metagene scores.
 First, linear models were created with the sample \gls{bmi}, \gls{bmi} status and a selection of the most ``consistent'' pathway metagene scores (based on the consistency across different data sets).
-From these linear models, \gls{pr} pathway metagene scores showed significant contribution to the model in predicting the obesity associated metagene scores in all of the linear models (apart from the FM obesity metagene).
-Therefore, linear models were also created with \gls{pr} pathway metagene in combination with the sample \gls{bmi}/\gls{bmi} status.
+From these linear models, \gls{pr} pathway metagene scores showed significant contribution to the model in predicting the obesity associated metagene scores in all of the linear models, with the exception of the FM obesity metagene (where Myc was the significant variable; \cref{app:b}).
+
+This was not surprising since all but FM obesity metagene were derived from the Creighton \textit{et al.} data set, and therefore the FM obesity metagene was likely to be different to those from the Creighton \textit{et al.} data set.
+Adding to this, difference in these metagenes were also shown in the heatmap of the obesity and pathway metagenes mentioned earlier, where the FM metagene clustered with neither the CR obesity metagenes nor with any other pathway metagenes.
+This implied that the FM metagene differed fundamentally from the CR obesity metagenes.
+Nevertheless, to clarify the significance of the \gls{pr} pathway with the obesity metagenes, linear models were also created with \gls{pr} pathway metagene in combination with the sample \gls{bmi}/\gls{bmi} status.
 
@@
 
+\subsection{Significance of the pathway associated signatures}
+\label{sub:significance_of_pr_pathway}
+
+% -- PR may have been significant only because it was from Cris' data?
+
 With that said, even if these variables were related to the obesity associated genetic signatures, the fact that these results depend on the pathway associated genetic signatures from the \citet{Gatza2010a} study makes the results questionable.
@@
 \label{sec:limitations}
+
+\subsection{Definition of obesity}
+\label{sub:definition_of_obesity}
+
+\subsection{Quality of the data}
+\label{sub:quality_of_the_data}
+
+% talk about number of samples...?
 
@@
 
-% Linear model prediction stuff
-% -- PR may have been significant only because it was from Cris' data?
+\subsection{Quality of the genetic signatures}
+\label{sub:quality_of_the_genetic_signatures}
 
@@
 
+There were two main aims to this project: firstly to determine whether there were any obesity specific genetic signatures that could be tranferred across multiple cancer types; and secondly to investigate whether there were any biological pathways dysregulated in the samples that were obese compared to the samples that were not obese.
+These aims were addressed in \cref{cha:obesity_genetic_signatures_and_cancer,cha:obesity_associated_genetic_signature_and_pathway_signatures}, resepectively.
+\\
+
+\noindent
+
 \section{Future directions}
 \label{sec:future_directions}
-
-% -- future directions

### 2017-03-01 thesis/discussion/discussion.tex
tmp commit

@@
 
-% TODO: -- Recap work
 % TODO: -- Put into context of very current literature
@@
 However, as observed by the results from the simulation, there were many genes identified by chance alone, which suggested that majority of the \glspl{deg} found in the eight cancer types were false positives.
-This apparently high level of \Glspl{type1} was not only observed in the \gls{icgc} data sets, but also in the Creighton \textit{et al.} data set when gene expression analyses were carried out to generate various obesity associated genetic signatures in \cref{sub:identification_of_obesity_associated_genetic_signatures}.
+This apparently high level of \Glspl{type1} was not only observed in the \gls{icgc} data sets, but also in the Creighton \textit{et al.} data set when gene expression analyses were carried out to generate various obesity associated genetic signatures in \cref{sub:identification_of_obesity_associated_genetic_signatures} (discussed in more details in \cref{sub:false_positives_in_gene_expression_analyses}).
 
@@
 There were no pathways enriched in any of the \gls{icgc} data sets, nor the combined \gls{icgc} data set.
-Even though there was no significantly enriched  pathway in the \gls{icgc} data sets, it was evident from the obesity metagenes from Creighton \textit{et al.} and Fuentes-Mattei \textit{et al.} data sets that these metagenes were associated with the gene expressions.
+Even though there was no significantly enriched  pathway in the \gls{icgc} data sets, it was evident from the obesity metagenes from Creighton \textit{et al.} and Fuentes-Mattei \textit{et al.} data sets that the obesity metagenes were associated with the sample gene expressions.
 This suggested that the metagenes were able to identify some sort of genetic pattern in the data, even though these metagenes were not significantly associated with obesity.
@@
 The pathway associated genetic signatures from the \citet{Gatza2010a} study were used to establish the biological connection with the obesity associated genetic signatures.
-In order to make the obesity and pathway associated genetic signatures comparable with one another, the orientations of the metagenes were examined so that all of the metagenes were ``correct'' (\cref{sec:pathway_associated_genetic_signatures_from_gatza2010a_study}).
+In order to make the obesity and pathway associated genetic signatures comparable with one another, the orientations of the metagenes were examined so that all of the metagenes were in the ``correct'' orientation (see \cref{sec:pathway_associated_genetic_signatures_from_gatza2010a_study}).
 
@@
 
-The obesity metagene from the \citet{Fuentes-Mattei2014} study was a cluster of its own, where the metagene did not group with any of the Creighton \textit{et al.} obesity metagenes nor with any of the Gatza \textit{et al.} pathway metagene.
+The obesity metagene from the \citet{Fuentes-Mattei2014} study was a cluster of its own, where the metagene did not group with any of the obesity metagenes derived from the Creighton \textit{et al.} data set nor with any of the Gatza \textit{et al.} pathway metagenes.
 Though Fuentes-Mattei \textit{et al.} suggested a biological link with the Akt/\gls{mtor} pathway in their study, the result from the heatmap showed otherwise.
-This may have been due to the inconsistency of the pathway metagene scores across different data as mentioned in \cref{sec:pathway_associated_metagenes_and_obesity_associated_metagenes}.
+This may have been due to the inconsistency of the pathway metagene scores across different data sets as mentioned in \cref{sec:pathway_associated_metagenes_and_obesity_associated_metagenes}.
 In fact, the Akt pathway metagene scores were variable across different cancer data sets (\cref{app:b}), which suggested the Akt pathway genetic signature was of poor quality.
@@
 Nevertheless, the evidence shown by \citet{Fuentes-Mattei2014} was convincing and the role of Akt/\gls{mtor} pathway in obesity and cancer should be explored in greater details.
-(include this paragraph or not... tbh, I think this is a weak argument)
+\textit{(include this paragraph or not... tbh, I think this is a weak argument)}
 
@@
 
-This was not surprising since all but FM obesity metagene were derived from the Creighton \textit{et al.} data set, and therefore the FM obesity metagene was likely to be different to those from the Creighton \textit{et al.} data set.
-Adding to this, difference in these metagenes were also shown in the heatmap of the obesity and pathway metagenes mentioned earlier, where the FM metagene clustered with neither the CR obesity metagenes nor with any other pathway metagenes.
+This was not surprising since all but the FM obesity metagene were derived from the Creighton \textit{et al.} data set, and therefore the FM obesity metagene was likely to be different to those from the Creighton \textit{et al.} data set.
+Adding to this, difference between the CR and FM metagenes were also shown in the heatmap of the obesity and pathway metagenes mentioned earlier, where the FM metagene clustered with neither the CR obesity metagenes nor with any other pathway metagenes.
 This implied that the FM metagene differed fundamentally from the CR obesity metagenes.
@@
 The predictions of the obesity metagene scores were made using the linear models created in Cris' data set.
-All of the predicted obesity metagene scores significantly correlated with the true obesity metagene scores in both Cris' and Creighton's data sets.
-However, the $R^2$-values for all of these predictions were low (highest $R^2$ $\approx$ 0.42) with highly variable data points which suggested that, although the predictions reflected the true values significantly, the variables included in the linear models may not be as relevant to the obesity metagenes as it may seem from these plots.
+All of the predicted obesity metagene scores significantly correlated with the true obesity metagene scores in both Cris' (training) and Creighton's (testing) data sets.
+However, the $R^2$-values for all of these predictions were low (highest $R^2$ $\approx$ 0.42; \cref{app:b}) with highly variable data points which suggested that, although the predictions reflected the true values significantly, the variables included in the linear models may not be as relevant to the obesity metagenes as it may seem from these plots.
 
@@
 This meant that the variables that were significant in the linear models played a role in the obesity associated genetic signatures.
-However, out of all of the step wise linear models created, there was no single variable that was common to all of the linear models; although Akt, \gls{egfr} and \gls{pr} were included in many of the models, which suggested that these pathway signatures may have an important role in the obesity associated genetic signatures.
+However, there was no single variable that was common to all of the linear models, but Akt, \gls{egfr} and \gls{pr} were included in many of the models, which suggested that these pathway signatures may have an important role in the obesity associated genetic signatures.
 
-\subsection{Significance of the pathway associated signatures}
+\subsection{Significance of the pathway associated signatures with obesity associated signatures}
 \label{sub:significance_of_pr_pathway}
@@
 
+The linear models that were created from the ``consistent'' pathway signatures in Cris' data set showed significant association of the \gls{pr} pathway with the obesity metagenes.
+But was this due to the true biological relationship between the \gls{pr} pathway and the obesity metagenes, or only because the models were created in Cris' data set?
+In fact, the proportion of the samples that were \gls{pr}$^+$ was greater in Cris' data set (62 \gls{pr}$^+$ and 37 \gls{pr}$^-$) compared with the Creighton's data set (48 \gls{pr}$^+$ and 50 \gls{pr}$^-$), which implied that the Cris' data set was biased towards \gls{pr} status and hence \gls{pr} pathway signature.
+
+However, with this logic, \gls{er} pathway signature would have also been biased, as the  sample  \gls{er} status was similar to the \gls{pr} status in either data sets: 72 and 58 \gls{er}$^+$ and 27 and 42 \gls{er}$^-$ samples in Cris' and Creighton's data sets, respectively.
+The fact that the \gls{er} pathway signature was not significant in the models suggested that the obesity metagenes were related to the \gls{pr} pathway irrespective of the sample \gls{pr} status.
+With that said, the \gls{pr} status of a sample does not necessarily reflect the level of the \gls{pr} pathway activity, the samples in Cris' data may have stronger \gls{pr} characteristics than the Creighton's data and therefore cannot deny the possibility of the bias.
+
+In the step wise linear models, many variables were identified to be significantly associated with the obesity metagenes.
 With that said, even if these variables were related to the obesity associated genetic signatures, the fact that these results depend on the pathway associated genetic signatures from the \citet{Gatza2010a} study makes the results questionable.

### 2017-03-01 thesis/intro/intro.tex
tmp commit

@@
 % Though more will be covered in details in \cref{sec:obesity_and_cancer}, the effect of obesity on cancers will be briefly mentioned here.
-Epidemiological evidence of obesity being a risk factor for many cancers has been well established, yet the mechanisms of its contribution to cancer is still  speculative \citep{Calle2003,Kelesidis2006a}.
-Although there are a lot of possible mechcanisms, there is growing evidence of the role of chronic inflammation caused by increased fat mass in the body \citep{Kelesidis2006a,Lumeng2011,Hernandez2013}.
+Epidemiological evidence of obesity being a risk factor for many cancers has been well established, yet the mechanisms of its contribution to cancer is still  speculative \citep{Calle2003,Kelesidis2006}.
+Although there are a lot of possible mechcanisms, there is growing evidence of the role of chronic inflammation caused by increased fat mass in the body \citep{Kelesidis2006,Lumeng2011,Hernandez2013}.
 Free fatty acids from excess adipose tissue and/or diet can be recognised by \gls{tlr} (specifically \gls{tlr}4), which can then activate \gls{nfkb}, leading to the activation of one of the Hallmarks of Cancer -- tumour-promoting inflammation \citep{Lumeng2011}.
@@
 
-Adiponectin is a hormone that is synthesised and secreted only by the \glspl{adipocyte} \citep{Kelesidis2006a}.
-Unlike leptin, where the concentration of leptin is higher in obese people, adiponectin levels are lower in obese people compared to the normal weight people \citep{Kelesidis2006a,Renehan2006}.
+Adiponectin is a hormone that is synthesised and secreted only by the \glspl{adipocyte} \citep{Kelesidis2006}.
+Unlike leptin, where the concentration of leptin is higher in obese people, adiponectin levels are lower in obese people compared to the normal weight people \citep{Kelesidis2006,Renehan2006}.
 Adiponectin has anti-angiogenic and anti-proliferative effect on the cells, and is secreted by mature adipocytes but not by the premature pre-adipocytes \citep{Gilbert2013}.
@@
 Of the 103 samples, 35\% were from normal weight, 28\% were from overweight and 37\% were from obese patients.
+It was also noted that African American patients were more likely to be obese than the Caucasian patients.
 They compared the gene expression between the tumours from obese patients with the tumours from non-obese patients and identified 799 probe sets (662 unique genes) with p \textless{} 0.01 and with a \gls{fc} \textgreater{} 1.2.
@@
 Comparison of the gene expression of the tumour samples from obese patients with the samples from non-obese patients revealed 112 genes that were statistically significantly dysregulated (62 genes upregulated and 50 genes downregulated).
-Furthermore, functional transcriptomic analysis suggested that insulin signalling and inflammation affected human \gls{er}+ breast cancer.
+Furthermore, functional transcriptomic analysis suggested that insulin signalling and inflammation affected human \gls{er}$^+$ breast cancer.
 

### 2017-03-02 thesis/discussion/discussion.tex
Started on limitations section - tmp commit

@@
 % TODO:   	==> How does this project fit into the literature
+
+% There were two main aims to this project: firstly to determine whether there were any obesity specific genetic signatures that could be tranferred across multiple cancer types; and secondly to investigate whether there were any biological pathways dysregulated in the samples that were obese compared to the samples that were not obese.
+% These aims were addressed in \cref{cha:obesity_genetic_signatures_and_cancer,cha:obesity_associated_genetic_signature_and_pathway_signatures}, resepectively, and will be answered in \cref{sec:conclusion}.
 
@@
 
-\subsection{\gls{blca} data set and obesity metagenes}
+\subsection{\gls{blca} data set and the obesity metagenes}
 \label{sub:blca_and_obesity_metagenes}
@@
 
-\subsection{False positives in gene expression analyses}
+\subsection{False positives in the gene expression analyses}
 \label{sub:false_positives_in_gene_expression_analyses}
@@
 
-\subsection{Linear models to predict obesity metagenes}
+\subsection{Linear models to predict the obesity metagenes}
 \label{sub:linear_models_to_predict_obesity_metagenes}
@@
 
-\subsection{Significance of the pathway associated signatures with obesity associated signatures}
+\subsection{Significance of the pathway associated signatures with the obesity associated signatures}
 \label{sub:significance_of_pr_pathway}
-
-% -- PR may have been significant only because it was from Cris' data?
 
@@
 But was this due to the true biological relationship between the \gls{pr} pathway and the obesity metagenes, or only because the models were created in Cris' data set?
-In fact, the proportion of the samples that were \gls{pr}$^+$ was greater in Cris' data set (62 \gls{pr}$^+$ and 37 \gls{pr}$^-$) compared with the Creighton's data set (48 \gls{pr}$^+$ and 50 \gls{pr}$^-$), which implied that the Cris' data set was biased towards \gls{pr} status and hence \gls{pr} pathway signature.
+In fact, the proportion of the samples that were \gls{pr}$^+$ was greater in Cris' data set (62 \gls{pr}$^+$ and 37 \gls{pr}$^-$ samples) compared with the Creighton's data set (48 \gls{pr}$^+$ and 50 \gls{pr}$^-$ samples), which implied that the Cris' data set was biased towards \gls{pr} status and hence \gls{pr} pathway signature.
 
@@
 With that said, even if these variables were related to the obesity associated genetic signatures, the fact that these results depend on the pathway associated genetic signatures from the \citet{Gatza2010a} study makes the results questionable.
-This was due to the lack of consistency of the metagenes across different cancer data sets, where more than half of the pathway metagenes showed a correlation of less than 0.8 in at least one data set (\cref{sec:pathway_associated_metagenes_and_obesity_associated_metagenes,app:b}), which meant that at least half of the signatures were specific to the Gatza \textit{et al.} data set.
+This was due to the lack of consistency of the metagenes across different cancer data sets, where more than half of the pathway metagenes showed a correlation of less than 0.8 in at least one data set (\cref{sec:pathway_associated_metagenes_and_obesity_associated_metagenes,app:b}).
+This meant that at least half of the signatures were specific to the Gatza \textit{et al.} data set and could be highly variable in other data sets.
 Thus, without further validation of the quality of the pathway associated genetic signatures and verification of the roles of these signatures in the biological relationship between obesity and cancer, these results should be interpreted with great caution.
@@
 
+% There were many limitations to this project, though some have already been mentioned in \cref{sec:summary_of_the_results}.
+
 \subsection{Definition of obesity}
-\label{sub:definition_of_obesity}
+\label{sub:discussion_definition_of_obesity}
+
+First and possibly the most important limitation to this project was the use of \gls{bmi} as a measurement for obesity and its subsequent classification of the samples based on \gls{bmi}.
+It was important to define the obese phenotype accurately, as all of the obesity associated genetic signatures, gene expression analyses and pathway enrichment analysis were dependent on the patient \gls{bmi} status.
+\Gls{bmi} is perhaps the most popular measurement to estimate the central adiposity owing to the relative ease of measurement in clinical settings, and hence used in various clinical trials and large epidemiological studies as an indication of the patients' metabolic state.
+However, there have been studies that showed other measurements, such as waist-to-height ratio and waist-to-hip ratio, to be a better representation of the central adiposity of the patient \citep{Dalton2003,Lee2008}.
+
+Perhaps the reason why there were so many false positives when gene expression analyses were carried out in \cref{sub:false_positives_in_gene_expression_analyses} was because \gls{bmi} was used to classify the patients into their corresponding \gls{bmi} statuses, instead of another measurement.
+It was likely that the sample \gls{bmi} (and therefore \gls{bmi} status) was not good enough a phenotypic characteristic to describe the underlying biology driven by obesity.
+Since the gene expression analyses were based on such ``vague'' parameter, it may have picked up so many genes that were apparently related to obesity when in fact the genes were not related to obesity at all.
+
+Ideally, multiple measurements that are indicative of obesity should be made in order to determine the obesity status of the patients to make a more accurate classification.
+% Taking these extra measurements could be the difference between 
+Additional measurements such as blood lipid profiles and insulin levels may also help understand the underlying biological state in which the patients are in, and would greatly help focus on the genes that are truly affecting the patients.
+However, these measurements will be difficult to carry out in a systematic fashion in a clinical setting, and to collate existing patient data with these measurements to make a data set that is large enough for the analyses done in this project may be impossible.
+Nevertheless, even the measurement of waist circumference will significantly help describe the obesity phenotype more accurately in future investigations.
 
@@
 \label{sub:quality_of_the_data}
+
+Second limitation was the quality of the data sets in which this project have used.
+All of the data sets used in this project were from an online and publicly available source, and the types of data sets used were either microarray or \gls{rnaseq} data.
+
+First problem -- different studies generated different data sets in a different way/method
+
+second problem -- micorarry vs. rna seq data
+
+third problem -- pure quality of the data
 
@@
 
+The final limitation to consider was the quality of the genetic signatures used in this project.
+This limitation is partly related to the previous section, as all of the genetic signatures were derived from the data sets used by other groups.
+
 \section{Conclusion}
@@
 These aims were addressed in \cref{cha:obesity_genetic_signatures_and_cancer,cha:obesity_associated_genetic_signature_and_pathway_signatures}, resepectively.
-\\
 
-\noindent
+As shown clearly by the results from \cref{cha:obesity_genetic_signatures_and_cancer}, there seemed to be no  genetic signature that were able to differntiate between the samples that were obese from those that were not, across different cancer types.
+
+Furthermore, there seemed to be no genes that were differentially expressed between the obese and the non-obese patients that were also common between different cancer types.
 

### 2017-03-03 thesis/discussion/discussion.tex
tmp commit

@@
 
-\subsection{\gls{blca} data set and the obesity metagenes}
+\subsection{\Gls{blca} data set and the obesity metagenes}
 \label{sub:blca_and_obesity_metagenes}
@@
 Ideally, multiple measurements that are indicative of obesity should be made in order to determine the obesity status of the patients to make a more accurate classification.
-% Taking these extra measurements could be the difference between 
 Additional measurements such as blood lipid profiles and insulin levels may also help understand the underlying biological state in which the patients are in, and would greatly help focus on the genes that are truly affecting the patients.
 However, these measurements will be difficult to carry out in a systematic fashion in a clinical setting, and to collate existing patient data with these measurements to make a data set that is large enough for the analyses done in this project may be impossible.
-Nevertheless, even the measurement of waist circumference will significantly help describe the obesity phenotype more accurately in future investigations.
+Nevertheless, even the measurement of waist circumference in addition to \gls{bmi} may significantly help describe the obesity phenotype more accurately in future investigations.
 
@@
 All of the data sets used in this project were from an online and publicly available source, and the types of data sets used were either microarray or \gls{rnaseq} data.
+First potential problem with using an open source data from other research groups is that the data is generated by multiple different groups with different technologies and experimental protocols.
+Since different laboratories have their own way to carry out certain experiments, the sequence data from each of the studies used in this project may have been produced and processed in a completely different manner, not to mention about the use of different patient cohorts in each experiment.
 
-First problem -- different studies generated different data sets in a different way/method
+For example, in the \citet{Creighton2012} study, the patient breast tumour biopsies were ``trimmed such that all the samples had $\geq$ 70\% tumour cells'' in the samples.
+On the other hand, \citet{Fuentes-Mattei2014} only mentioned that they used breast tumour biopsies from the patient with no comment about the quality or quantity of the biopsies.
+However, they took one third of the tumours from the mouse models to analyse the transcriptome of the mouse tumours, so perhaps Fuentes-Mattei \textit{et al.} used similar approach with the human breast tumour samples.
+Without a doubt, the difference in how the samples are handled will affect the quality of the data, making it challenging to compare different data sets and it may even be possible for the samples within the data set to be incomparable.
 
-second problem -- micorarry vs. rna seq data
+% TODO: move some of this stuff into the false positive section
+In fact, a couple of observations were made in this project that suggested the Creighton \textit{et al.} data set was of poor quality.
+First in \cref{sub:identification_of_obesity_associated_genetic_signatures}, there were many \glspl{deg} identified from the gene expression analyses when obesity associated genetic signatures were sought for in the Creighton \textit{et al.} data set.
+This suggested that there was a significant number of genes that were not related to obesity (false positives) in the genetic signatures.
+Furthermore, the fact that about a third of the genes in any one of the obesity signatures were ``unique'' supported the apparent abundance of false positives in the signatures (\cref{fig:venn1,fig:venn2}).
+From these results, it was likely that the obesity signatures that resulted from the Creighton \textit{et al.} data set were full of false positives due to the poor quality of the data.
 
-third problem -- pure quality of the data
+Although the underlying quality of the data provided by these studies could not be refined in any way, best efforts were made to make sure the data sets were free of other factors that may affect the results.
+% Although the underlying quality of the data provided by these studies could not be refined in any way, best efforts were made to make sure the results from these data sets were reliable.
+This leads to the second problem: the different sequencing technologies used in the studies.
+All of Creighton \textit{et al.}, Fuentes-Mattei \textit{et al.}, Print \textit{et al.} and Gatza \textit{et al.} data sets were analysed with the microarray technology, whereas all of the \gls{icgc} cancer data sets were sequenced with \gls{ngs} technology.
+As mentioned in \cref{sub:data_normalisation}, the two technologies use fundamentally different principals to output sequence information of the sample; microarray uses light intensity and \gls{ngs} uses count data.
+
+In order to make these two different types of data comparable to one another, normalisation methods from the \textit{limma} package were used (see \cref{sub:data_normalisation}).
+This enabled the standard \textit{limma} analysis pipelines, normally used for the analysis of microarray data, to be used by the \gls{icgc} cancer \gls{rnaseq} data sets.
+In this way, the results from the two types of data sets were fairly reliable and consistent as the same analysis methods were applied to the data sets.
+In addition to the difference in the sequencing technologies, the Gatza \textit{et al.} data set comprised of multiple different data sets, each from different sources.
+Since individual data set had their own experimental differences (``batch effect''), it was important to correct for these differences when multiple data sets were combined into one (see \cref{sub:batch_correction}).
+In this project, batch corrections were made in the Gatza \textit{et al.} data set and the combined \gls{icgc} cancer data sets (used in \cref{sec:pathways_enriched_in_icgc_data_sets}) to eliminate the batch effects.
+
+It was evident from some of the results that the Creighton \textit{et al.} data set was of poor quality, and the validity of the other data sets (and therefore the results generated from these data sets) were also debatable, despite all of the attempts made to control for other factors that may have affected the results.
 
@@
 The final limitation to consider was the quality of the genetic signatures used in this project.
-This limitation is partly related to the previous section, as all of the genetic signatures were derived from the data sets used by other groups.
+This limitation is partly related to \cref{sub:quality_of_the_data}, as  the identification of the  genetic signatures depended  on the quality of the data as well as the definition  of the signature in the first instance (for example, obesity or Akt overexpression).
+
+Firstly, the obesity signatures identified by both Creighton \textit{et al.} and Fuentes-Mattei \textit{et al.} relied on the definition of obesity based on the sample \gls{bmi} values.
+As discussed in \cref{sub:discussion_definition_of_obesity}, the use of \gls{bmi} may not have been the best measurement for central adiposity.
+In fact, both Creighton \textit{et al.} and Fuentes-Mattei \textit{et al.} signatures associated with different pathway associated signatures in \cref{sec:pathway_associated_metagenes_and_obesity_associated_metagenes,sec:prediction_of_obesity_associated_metagene_with_pathway_associate_metagene}, which suggested that obesity as defined by \gls{bmi} was not enough to reveal the underlying biological characteristic of obesity, and hence showed association with different pathway signatures.
+Furthermore, as mentioned in \cref{sub:quality_of_the_data}, it was likely that the Creighton \textit{et al.} data set, and perhaps other data sets as well, was of poor quality and may have affected the ability of the signatures as a marker for obesity in cancer.
+
+Secondly for the pathway associated genetic signatures,
 

### 2017-03-08 thesis/discussion/discussion.tex
Future directions left to complete - tmp commit

@@
 The fact that the \gls{er} pathway signature was not significant in the models suggested that the obesity metagenes were related to the \gls{pr} pathway irrespective of the sample \gls{pr} status.
-With that said, the \gls{pr} status of a sample does not necessarily reflect the level of the \gls{pr} pathway activity, the samples in Cris' data may have stronger \gls{pr} characteristics than the Creighton's data and therefore cannot deny the possibility of the bias.
+With that said, the \gls{pr} status of a sample does not necessarily reflect the level of the \gls{pr} pathway activity; the samples in Cris' data may have stronger \gls{pr} characteristics than the Creighton's data and therefore cannot deny the possibility of the bias.
 
@@
 In fact, both Creighton \textit{et al.} and Fuentes-Mattei \textit{et al.} signatures associated with different pathway associated signatures in \cref{sec:pathway_associated_metagenes_and_obesity_associated_metagenes,sec:prediction_of_obesity_associated_metagene_with_pathway_associate_metagene}, which suggested that obesity as defined by \gls{bmi} was not enough to reveal the underlying biological characteristic of obesity, and hence showed association with different pathway signatures.
-Furthermore, as mentioned in \cref{sub:quality_of_the_data}, it was likely that the Creighton \textit{et al.} data set, and perhaps other data sets as well, was of poor quality and may have affected the ability of the signatures as a marker for obesity in cancer.
+Furthermore, as mentioned in \cref{sub:quality_of_the_data}, it was likely that the Creighton \textit{et al.} data set (and perhaps other data sets as well) was of poor quality and may have affected the ability of the generated signatures as a marker for obesity in cancer.
 
-Secondly for the pathway associated genetic signatures,
+Secondly, the method in which the pathway associated genetic signatures were defined in the Gatza \textit{et al.} study may have been questionable.
+In their studies, Gatza \textit{et al.} used model cell lines to identify the pathway associated genetic signatures by altering the expression of the representative gene for the pathway, and decided all of the genes affected by the change in representative gene expression were associated with that pathway \citep{Gatza2010a} (add Gatza's old paper).
+However, some of these pathway associated genetic signatures were derived in different cell lines to one another, which meant that it was possible for some of the pathway signatures to be defined incorrectly.
+In fact, as seen in \cref{sec:pathway_associated_genetic_signatures_from_gatza2010a_study,sec:pathway_associated_metagenes_and_obesity_associated_metagenes}, some pathway signatures produced more consistent metagenes from the data sets than the other signatures.
+These results were merely an indication that the pathway signatures were generated incorrectly and it was not possible to determine whether these signatures were truly associated to the corresponding pathways.
+
+Taken together, the quality of both the obesity and the pathway associated genetic signatures may not have been good and, in combination with poor quality data, may have produced unexpected results.
 
@@
 As shown clearly by the results from \cref{cha:obesity_genetic_signatures_and_cancer}, there seemed to be no  genetic signature that were able to differntiate between the samples that were obese from those that were not, across different cancer types.
+Furthermore, there were no genes differentially expressed between obese and non-obese patients that were common across multiple cancer types.
+These results have suggested that there were no obesity specific genetic signatures that were common across different cancer types.
 
-Furthermore, there seemed to be no genes that were differentially expressed between the obese and the non-obese patients that were also common between different cancer types.
+The results from \cref{cha:obesity_associated_genetic_signature_and_pathway_signatures} showed that the obesity associated genetic signatures generated from the Creighton \textit{et al.} data set were different from those generated from the Fuentes-Mattei \textit{et al.} data set.
+None of the pathway associated genetic signatures was similar to any of the obesity associated genetic signatures in the heatmap.
+The results from the linear model prediction suggested a possible association of the obesity associated genetic signatures with the Akt, \gls{egfr} and \gls{pr} pathways.
+
+To summarise, there were no common genetic signatures that associated with the obesity status of the patients.
+However, the Akt, \gls{egfr} and \gls{pr} pathways may have a role in promoting the tumour progression in obese patients.
+Further investigations with good quality data sets and larger patient cohorts are required to clarify the biological relationship between obesity and cancer.
 

### 2017-03-08 thesis/results/results2.tex
Future directions left to complete - tmp commit

@@
 One thing noted from these scatter plots was that there were some pathway signatures that were more variable than the others.
-As an example, the \gls{tgfb} metagenes were significantly more dispersed compared to the \gls{pr} metagenes, even though these metagenes were both generated similarly in the same data set (\cref{fig:gt_rma_vs_mas}).
+As an example, the \gls{tgfb} metagenes were significantly more dispersed compared to the \gls{pr} metagenes, even though both of these metagenes were generated in the same data set through similar processes (\cref{fig:gt_rma_vs_mas}).
 These differences were seen in other data sets as well (\cref{app:b}).

### 2017-03-09 thesis/discussion/discussion.tex
finished discussion...? Start editing other parts - tmp commit

@@
 To conclude this thesis, all of the results were re-examined and summarised to highlight the key findings from this project.
-%TODO: add a little more content to the next sentence
-There were many limitations to this project and these limitations were adressed in \cref{sec:limitations}.
+There were many limitations to this project and these limitations were adressed either with the results summary in \cref{sec:summary_of_the_results}, or seaprately in \cref{sec:limitations}.
 Lastly, a conclusion was drawn from all of the evidences from this project and suggestions for future aims and experiments were made.
@@
 
-The fact that both obesity associated genetic signatures identified by Creighton \textit{et al.} and Fuentes-Mattei \textit{et al.} showed no sign of significant association with sample \gls{bmi}/\gls{bmi} status in majority of the cancer data sets raised a question of whether these signatures were actually related to obesity, or a different clinical variable within the data set.
+The fact that both CR and FM obesity associated genetic signatures showed no sign of significant association with sample \gls{bmi}/\gls{bmi} status in majority of the cancer data sets raised a question of whether these signatures were actually related to obesity, or a different clinical variable within the data set.
 This question led to the identification of the various versions of obesity associated genetic signatures in \cref{sub:identification_of_obesity_associated_genetic_signatures} that were controlled for clinical variables (sex, age, ethnicity, menopause status, tumour grade, \gls{er}/\gls{her2}/\gls{pr} statuses and \gls{ln} status) in the Creighton \textit{et al.} data set.
 All of the metagenes created from these clinical variable-controlled genetic signatures were consistent with the gene expression patterns but none of the metagenes showed significant association with the sample \gls{bmi}/\gls{bmi} status in the data set other than the Creighton \textit{et al.} data set, with the exception of the overweight group in the \gls{blca} data set.
+
+It was possible that all of these results were due to the sample quality (and hence due to the quality of the data).
+Perhaps many of the samples in these data sets had breast cancers due to a cause unrelated to their obesity status; in other words, some of the breast tumour samples were not driven by obesity.
+This may affect the detection of the truly obesity associated genes, as some of the tumour samples from the patients will have nothing to do with obesity and cancer, and hence misclassified into wrong groups.
 
@@
 The results so far have provided strong evidence that the obesity associated genetic signatures from the Creighton \textit{et al.} and Fuentes-Mattei \textit{et al.} data sets were not associated with the sample \gls{bmi}/\gls{bmi} status.
-In the last attempt to identify any genes that were common across multiple cancer types that were also associated with obesity, gene expression analyses were carried out on the \gls{icgc} cancer data sets (\cref{sec:common_genes_across_multiple_cancer_types}).
+In the last attempt to identify any obesity associated genes that were common across multiple cancer types, gene expression analyses were carried out on the \gls{icgc} cancer data sets (\cref{sec:common_genes_across_multiple_cancer_types}).
 There were many \glspl{deg} identified by each of the \gls{icgc} cancer types, though there were no genes in common for all eight cancer types.
@@
 This meant that there were very little evidence of true obesity associated genetic signatures in any of the data sets that have been explored in this project, or perhaps the sample \gls{bmi}/\gls{bmi} status was not enough to pinpoint the underlying biological relationship between obesity and cancer.
+
+Evidence of large number of false positives was also seen in \cref{sub:identification_of_obesity_associated_genetic_signatures}, where there were many \glspl{deg} identified from the gene expression analyses when obesity associated genetic signatures were sought for in the Creighton \textit{et al.} data set.
+This suggested that there was a significant number of genes that were not related to obesity (false positives) in the genetic signatures.
+Furthermore, the fact that about a third of the genes in any one of the obesity signatures were ``unique'' to each of the individual obesity signatures supported the apparent abundance of false positives in the signatures (\cref{fig:venn1,fig:venn2}).
+From these results, it was likely that the obesity signatures that resulted from the Creighton \textit{et al.} data set had many genes that were false positives, perhaps due to the poor quality of the data set (\cref{sub:quality_of_the_data}).
+
 Even if the sample \gls{bmi} and/or \gls{bmi} status was enough to identify the genes that were truly associated with obesity, it would not be a trivial task to identify the truly obesity associated genes from those that were not.
-This raises an important question of whether the sample \gls{bmi}/\gls{bmi} status was appropriate to uncover the true association between obesity and cancer (adressed in \cref{sec:limitations,sec:future_directions}).
-
-\textit{(need to add more content here...?)}
+This raises an important question of whether the sample \gls{bmi}/\gls{bmi} status was an appropriate clinical variable to uncover the true association between obesity and cancer (more in \cref{sec:limitations,sec:future_directions}).
 
@@
 Once the directions of the metagenes were determined, the obesity metagenes were clustered together in the heatmap to visualise the similarity of the obesity metagenes with the pathway metagenes.
-The heatmap revealed that all of the obesity metagenes that have originated from the Creighton \textit{et al.} study clustered together in a group with no other pathway metagenes correlating with the metagenes, which meant that the obesity metagenes were not similar to any of the pathway metagenes from the \citet{Gatza2010a} study.
+The heatmap revealed that all of the obesity metagenes that have originated from the Creighton \textit{et al.} study clustered together in a group with no other pathway metagenes correlating with these metagenes, which meant that the obesity metagenes were not similar to any of the pathway metagenes from the \citet{Gatza2010a} study.
 The clustering of the obesity metagene was not surprising, as the results from \cref{sub:_novel_obesity_associated_signatures_and_sample_bmi} already showed that the metagenes were highly correlated with one another.
@@
 
-The obesity metagene from the \citet{Fuentes-Mattei2014} study was a cluster of its own, where the metagene did not group with any of the obesity metagenes derived from the Creighton \textit{et al.} data set nor with any of the Gatza \textit{et al.} pathway metagenes.
+The obesity metagene from the \citet{Fuentes-Mattei2014} study was a cluster of its own, where the metagene did not group with any of the obesity metagenes derived from the Creighton \textit{et al.} data set, nor with any of the Gatza \textit{et al.} pathway metagenes.
 Though Fuentes-Mattei \textit{et al.} suggested a biological link with the Akt/\gls{mtor} pathway in their study, the result from the heatmap showed otherwise.
@@
 All of the predicted obesity metagene scores significantly correlated with the true obesity metagene scores in both Cris' (training) and Creighton's (testing) data sets.
-However, the $R^2$-values for all of these predictions were low (highest $R^2$ $\approx$ 0.42; \cref{app:b}) with highly variable data points which suggested that, although the predictions reflected the true values significantly, the variables included in the linear models may not be as relevant to the obesity metagenes as it may seem from these plots.
+However, the $R^2$-values for all of these predictions were low (highest $R^2$ $\approx$ 0.42; \cref{app:b}) with highly variable data points in the scatter plots which suggested that, although the predictions reflected the true values significantly, the variables included in the linear models may not be relevant to the obesity metagenes as it may seem from these plots.
 
@@
 This meant that the variables that were significant in the linear models played a role in the obesity associated genetic signatures.
-However, there was no single variable that was common to all of the linear models, but Akt, \gls{egfr} and \gls{pr} were included in many of the models, which suggested that these pathway signatures may have an important role in the obesity associated genetic signatures.
+There was no single pathway signature that was common to all of the linear models, but Akt, \gls{egfr} and \gls{pr} pathway metagenes were included in many of the models, which suggested that these pathway signatures may have an important role in the obesity associated genetic signatures.
 
@@
 In the step wise linear models, many variables were identified to be significantly associated with the obesity metagenes.
-With that said, even if these variables were related to the obesity associated genetic signatures, the fact that these results depend on the pathway associated genetic signatures from the \citet{Gatza2010a} study makes the results questionable.
+With that said, even if these variables were related to the obesity associated genetic signatures, the fact that these results depended on the pathway associated genetic signatures from the \citet{Gatza2010a} study makes the results questionable.
 This was due to the lack of consistency of the metagenes across different cancer data sets, where more than half of the pathway metagenes showed a correlation of less than 0.8 in at least one data set (\cref{sec:pathway_associated_metagenes_and_obesity_associated_metagenes,app:b}).
@@
 
-First and possibly the most important limitation to this project was the use of \gls{bmi} as a measurement for obesity and its subsequent classification of the samples based on \gls{bmi}.
+First and possibly the most important limitation to this project was the use of \gls{bmi} as a measurement for central adiposity and its subsequent classification of the samples based on \gls{bmi}.
 It was important to define the obese phenotype accurately, as all of the obesity associated genetic signatures, gene expression analyses and pathway enrichment analysis were dependent on the patient \gls{bmi} status.
@@
 
-Second limitation was the quality of the data sets in which this project have used.
+Second limitation was the quality of the data sets this project have used.
 All of the data sets used in this project were from an online and publicly available source, and the types of data sets used were either microarray or \gls{rnaseq} data.
@@
 However, they took one third of the tumours from the mouse models to analyse the transcriptome of the mouse tumours, so perhaps Fuentes-Mattei \textit{et al.} used similar approach with the human breast tumour samples.
-Without a doubt, the difference in how the samples are handled will affect the quality of the data, making it challenging to compare different data sets and it may even be possible for the samples within the data set to be incomparable.
-
-% TODO: move some of this stuff into the false positive section
-In fact, a couple of observations were made in this project that suggested the Creighton \textit{et al.} data set was of poor quality.
-First in \cref{sub:identification_of_obesity_associated_genetic_signatures}, there were many \glspl{deg} identified from the gene expression analyses when obesity associated genetic signatures were sought for in the Creighton \textit{et al.} data set.
-This suggested that there was a significant number of genes that were not related to obesity (false positives) in the genetic signatures.
-Furthermore, the fact that about a third of the genes in any one of the obesity signatures were ``unique'' supported the apparent abundance of false positives in the signatures (\cref{fig:venn1,fig:venn2}).
-From these results, it was likely that the obesity signatures that resulted from the Creighton \textit{et al.} data set were full of false positives due to the poor quality of the data.
+Without a doubt, the difference in how the samples were handled affected the quality of the data, which made it challenging to compare different data sets and it may even be possible for the samples within the data set to be incomparable.
+In fact, a couple of observations were made in this project that suggested the Creighton \textit{et al.} data set was of poor quality (see \cref{sub:false_positives_in_gene_expression_analyses}).
 
@@
 
-% talk about number of samples...?
-
-% -- Limitations of this project
-
-% RESULTS 1
-
-% Creighton metagene section
-% Lack of association of obesity metagenes with different cancer data sets
-% -- due to metagene derivation from a specific data and therefore cannot be applied to other data set
-% -- don't forget about the possibility of having samples that may not be driven by obesity at all in the data set (and therefore the metagene may be affected by these samples).
-
-% FM metagene section
-% -- don't forget about the possibility of having samples that may not be driven by obesity at all in the data set (and therefore the metagene may be affected by these samples).
-
-% gene expression section
-% -- mention that Creighton's group probably didn't control for FDR
-
-% RESULTS 2
-
-% Pathway associated genetic signature directions
-
-% Pathway and obesity associated genetic signatures
-% -- some metagenes had consistently high correlation across different data sets.
-%		-- showed that these signatures were more reliable than the others
-%		-- the qualities of these signatures were good as they showed similar metagene scores across different data sets
-%			-- quality of the signatures may be due to the difference in experimental conditions, or the samples used by Gatza et al...?
-
-% -- FM metagene not clustering with AKT
-%		-- due to shitty akt pathway signature, or difference in mouse model.
-
 \subsection{Quality of the genetic signatures}
@@
 
-Secondly, the method in which the pathway associated genetic signatures were defined in the Gatza \textit{et al.} study may have been questionable.
-In their studies, Gatza \textit{et al.} used model cell lines to identify the pathway associated genetic signatures by altering the expression of the representative gene for the pathway, and decided all of the genes affected by the change in representative gene expression were associated with that pathway \citep{Gatza2010a} (add Gatza's old paper).
+Secondly, the method in which the pathway associated genetic signatures was defined in the Gatza \textit{et al.} study may have been questionable.
+In their studies, Gatza \textit{et al.} used model cell lines to identify the pathway associated genetic signatures by altering the expression of the representative gene for the pathway and decided all of the genes affected by the change in representative gene expression were associated with that pathway \citep{Gatza2010a} (add Gatza's old paper).
 However, some of these pathway associated genetic signatures were derived in different cell lines to one another, which meant that it was possible for some of the pathway signatures to be defined incorrectly.
 In fact, as seen in \cref{sec:pathway_associated_genetic_signatures_from_gatza2010a_study,sec:pathway_associated_metagenes_and_obesity_associated_metagenes}, some pathway signatures produced more consistent metagenes from the data sets than the other signatures.
-These results were merely an indication that the pathway signatures were generated incorrectly and it was not possible to determine whether these signatures were truly associated to the corresponding pathways.
+These results were merely an indication that the pathway signatures may have been generated incorrectly and it was not possible to determine whether these signatures were truly associated to the corresponding pathways as the original data was not available to reference the signatures back for validation.
 
@@
 These results have suggested that there were no obesity specific genetic signatures that were common across different cancer types.
-
 The results from \cref{cha:obesity_associated_genetic_signature_and_pathway_signatures} showed that the obesity associated genetic signatures generated from the Creighton \textit{et al.} data set were different from those generated from the Fuentes-Mattei \textit{et al.} data set.
@@
 
-To summarise, there were no common genetic signatures that associated with the obesity status of the patients.
-However, the Akt, \gls{egfr} and \gls{pr} pathways may have a role in promoting the tumour progression in obese patients.
+To summarise, there were no common genetic signatures that significantly associated with the obesity status of the patients.
+However, the Akt, \gls{egfr} and \gls{pr} pathways may have a role in promoting the tumour progression in patients that are obese.
 Further investigations with good quality data sets and larger patient cohorts are required to clarify the biological relationship between obesity and cancer.
@@
 \label{sec:future_directions}
+
+This project have managed to show that there were no obesity specific genes  that were common across multiple cancer types.
+Although this study failed to identify obesity specific genes that affected tumour biology, it is likely that there is some complex mechanism underlying the relationship between obesity and cancer; otherwise many of the clinical and epidemiological associations between obesity and cancer would not make any sense.
+Since Akt, \gls{egfr} and \gls{pr} pathways were signicant in the linear models that predicted the obesity metagenes, more thorough investigation of these pathways may be a good start to clarify the relationship.
+The apparent significance of the obesity associated genetic signatures in \gls{blca} data set was interesting as well, and deeper analysis of this data set may help further understand the biology of \gls{blca} and its association with obesity.
+
+In this project, the lack of association of the obesity associated genetic signatures with any of the cancer types could have been due to the quality of both the raw data sets and the genetic signatures used in this project.
+Therefore, careful selection of the raw data sets with sufficient, and if possible, additional information (such as waist circumference) about the samples and well-validated genetic signatures will be essential for future investigations.

### 2017-03-10 thesis/discussion/discussion.tex
Finished discussion, now edit everything - tmp commit

@@
 % TODO:   	==> How does this project fit into the literature
-
-% There were two main aims to this project: firstly to determine whether there were any obesity specific genetic signatures that could be tranferred across multiple cancer types; and secondly to investigate whether there were any biological pathways dysregulated in the samples that were obese compared to the samples that were not obese.
-% These aims were addressed in \cref{cha:obesity_genetic_signatures_and_cancer,cha:obesity_associated_genetic_signature_and_pathway_signatures}, resepectively, and will be answered in \cref{sec:conclusion}.
 
@@
 
-% TODO: search for any evidence of BLCA with breast cancers - add to this paragraph
 With that said, there is a possibility that the genotypes of the samples that are overweight in \gls{blca} data set are similar to those samples that are obese in the Creighton \textit{et al.} data set.
-\textit{(TODO: search for any evidence of relationship/connection between BLCA and breast cancers - add to this paragraph)}
+In fact, \citet{Damrauer2014} have identified ``basal-like'' and ``luminal''  subtypes in bladder cancer that resembled the corresponding subtypes in breast cancer, and suggested that bladder  and breast cancers may have common physiological properties.
+Since all of the obesity associated genetic signatures were derived from breast cancer data sets, it made sense that these signatures were significant in the \gls{blca} data set, as bladder cancer and breast cancer could have been biologically similar to one another.
+However, the degree of similarity between these two cancer types are not yet known and further analyses are required for a definitive explanation of the results.
 
@@
 
-Though how unreliable the Akt pathway metagene scores may be, there is still a possibility that the FM obesity associated genetic signature may be unrelated to the Akt pathway, as Fuentes-Mattei \textit{et al.} validated their human breast tumour-derived genetic signature in a mouse model \citep{Fuentes-Mattei2014}.
+(Though how unreliable the Akt pathway metagene scores may be, there is still a possibility that the FM obesity associated genetic signature may be unrelated to the Akt pathway, as Fuentes-Mattei \textit{et al.} validated their human breast tumour-derived genetic signature in a mouse model \citep{Fuentes-Mattei2014}.
 	Mouse models are indeed helpful in studying biological processes \textit{in vivo}, especially when the cause and the outcome of the disease is well established.
@@
 	Furthermore, obesity is a disease that have a variety of biological effects (for example the release of adipokines and the disruption of the hormone levels) that could be acting in a different way in mice compared to humans, which further complicates the causal relationship between obesity and cancer.
-Nevertheless, the evidence shown by \citet{Fuentes-Mattei2014} was convincing and the role of Akt/\gls{mtor} pathway in obesity and cancer should be explored in greater details.
+	Nevertheless, the evidence shown by \citet{Fuentes-Mattei2014} was convincing and the role of Akt/\gls{mtor} pathway in obesity and cancer should be explored in greater details.)
 \textit{(include this paragraph or not... tbh, I think this is a weak argument)}
@@
 Since Akt, \gls{egfr} and \gls{pr} pathways were signicant in the linear models that predicted the obesity metagenes, more thorough investigation of these pathways may be a good start to clarify the relationship.
-The apparent significance of the obesity associated genetic signatures in \gls{blca} data set was interesting as well, and deeper analysis of this data set may help further understand the biology of \gls{blca} and its association with obesity.
+Further analyses of \gls{blca} may help clarify the similarity between bladder and breast cancer types and provide an insight into the association between the obesity associated genetic signatures and \gls{blca}.
 

### 2017-03-10 thesis/misc/appendix.tex
tmp commit
finished first section of results2 - tmp commit
Changed all the data set names and removed "'s" - tmp commit

@@
 	\renewcommand{\thesection}{\Alph{chapter}\arabic{section}}
-	\chapter{Additional results from \cref{cha:obesity_genetic_signatures_and_cancer}}
+
+	\chapter{Additional results for \cref{cha:obesity_genetic_signatures_and_cancer}}
 	\label{app:a}
 
-	\section{Metagenes created from raw data vs. standardised data (Creighton)}
+	\section{Metagenes created from raw data and standardised data in Creighton \textit{et al.} data set}
 	\label{sec:metagenes_created_from_raw_data_vs_standardised_data}
@@
 		\includegraphics[width=0.45\linewidth,page=2]{appendix/check_raw_vs_std_mg}
-		\caption{appendix/Check_raw_vs_std}
+		\caption{Scatter plots showing the raw and ranked obesity metagene scores generated from raw or standardised Creighton \textit{et al.} data set. }
 		\label{fig:appendix/check_raw_vs_std}
@@
 
-	\section{Metagenes created from raw data vs. standardised data (ICGC)}
+	\section{Metagenes created from raw data and standardised data (ICGC)}
 	\label{sec:metagenes_created_from_raw_data_vs_standardised_data_icgc}
 
-	\begin{figure}[htpb]
-		\centering
-		\includegraphics[width=0.45\linewidth,page=1]{appendix/crtcga_raw_vs_std}
-		\hfill
-		\includegraphics[width=0.45\linewidth,page=2]{appendix/crtcga_raw_vs_std}
-		\caption{appendix/Check_raw_vs_std}
-		\label{fig:appendix/check_raw_vs_std}
-	\end{figure}
+	% \begin{figure}[h]
+	% 	\centering
+	% 	\includegraphics[width=0.45\linewidth,page=1]{appendix/crtcga_raw_vs_std}
+	% 	\hfill
+	% 	\includegraphics[width=0.45\linewidth,page=2]{appendix/crtcga_raw_vs_std}
+	% 	\caption{appendix/Check_raw_vs_std}
+	% 	\label{fig:appendix/check_raw_vs_std}
+	% \end{figure}
 
@@
 
+	\chapter{Additional results from \cref{cha:obesity_associated_genetic_signature_and_pathway_signatures}}
+	\label{app:b}
+
+	\begin{figure}[htpb]
+		\centering
+		\includegraphics[width=0.7\linewidth]{results2/gatzares1}
+		\caption{results from gatza's study}
+		\label{fig:gatza_paper_res}
+	\end{figure}
+
+% [1] "crolgene"
+
+% Call:
+% lm(formula = as.formula(formula_txt[j]), data = data)
+
+% Coefficients:
+%              Estimate Std. Error t value Pr(>|t|)
+% (Intercept)         0.553214  0.123891 4.465  2.16e-05 ***
+% bmi                 -0.001629 0.004071 -0.400 0.69
+% (Intercept)         0.55700   0.05464  10.194 <2e-16   ***
+% bmiStatusobese      -0.10551  0.07021  -1.503 0.136
+% bmiStatusoverweight -0.02165  0.07727  -0.280 0.780
+% (Intercept)         0.307513  0.163530 1.880  0.0631   .
+% bmi                 0.011279  0.006976 1.617  0.1092
+% bmiStatusobese      -0.261145 0.118795 -2.198 0.0304   *
+% bmiStatusoverweight -0.079845 0.084661 -0.943 0.3480
+% (Intercept)         0.007141  0.382725 0.019  0.9852
+% bcat_probes         0.156062  0.223779 0.697  0.4873
+% er_probes           0.123403  0.254800 0.484  0.6293
+% ifna_probes         -0.430229 0.497301 -0.865 0.3892
+% ifng_probes         0.378437  0.509352 0.743  0.4594
+% myc_probes          0.160785  0.229142 0.702  0.4846
+% pr_probes           0.597403  0.244252 2.446  0.0164   *
+% (Intercept)         0.055088  0.389472 0.141  0.8878
+% bmi                 -0.002745 0.003814 -0.720 0.4736
+% bcat_probes         0.179900  0.226799 0.793  0.4297
+% er_probes           0.125413  0.255485 0.491  0.6247
+% ifna_probes         -0.475417 0.502546 -0.946 0.3466
+% ifng_probes         0.429131  0.515526 0.832  0.4074
+% myc_probes          0.184642  0.232124 0.795  0.4284
+% pr_probes           0.607982  0.245336 2.478  0.0151   *
+% (Intercept)         0.01666   0.37989  0.044  0.9651
+% bmiStatusobese      -0.11197  0.06602  -1.696 0.0933   .
+% bmiStatusoverweight -0.02348  0.07151  -0.328 0.7434
+% bcat_probes         0.22035   0.22616  0.974  0.3325
+% er_probes           0.10450   0.25330  0.413  0.6809
+% ifna_probes         -0.60343  0.50372  -1.198 0.2341
+% ifng_probes         0.54671   0.51484  1.062  0.2911
+% myc_probes          0.21341   0.22998  0.928  0.3559
+% pr_probes           0.59491   0.24257  2.453  0.0161   *
+% (Intercept)         -0.128149 0.393204 -0.326 0.7453
+% bmi                 0.008834  0.006565 1.346  0.1819
+% bmiStatusobese      -0.234634 0.112386 -2.088 0.0397   *
+% bmiStatusoverweight -0.067942 0.078484 -0.866 0.3890
+% bcat_probes         0.216892  0.225167 0.963  0.3380
+% er_probes           0.081506  0.252745 0.322  0.7478
+% ifna_probes         -0.625864 0.501740 -1.247 0.2155
+% ifng_probes         0.546873  0.512538 1.067  0.2889
+% myc_probes          0.196215  0.229305 0.856  0.3945
+% pr_probes           0.560706  0.242816 2.309  0.0232   *
+
+% [1] "resobsgene"
+
+% Call:
+% lm(formula = as.formula(formula_txt[j]), data = data)
+
+% Coefficients:
+%              Estimate Std. Error t value Pr(>|t|)
+% (Intercept)         0.547104  0.123915 4.415  2.62e-05 ***
+% bmi                 -0.001422 0.004072 -0.349 0.728
+% (Intercept)         0.547619  0.054641 10.022 <2e-16   ***
+% bmiStatusobese      -0.096832 0.070212 -1.379 0.171
+% bmiStatusoverweight -0.001804 0.077273 -0.023 0.981
+% (Intercept)         0.298589  0.163546 1.826  0.0710   .
+% bmi                 0.011259  0.006976 1.614  0.1099
+% bmiStatusobese      -0.252186 0.118807 -2.123 0.0364   *
+% bmiStatusoverweight -0.059898 0.084669 -0.707 0.4810
+% (Intercept)         0.18694   0.36963  0.506  0.6142
+% bcat_probes         0.09717   0.21612  0.450  0.6541
+% er_probes           0.07879   0.24608  0.320  0.7496
+% ifna_probes         -0.17576  0.48029  -0.366 0.7152
+% ifng_probes         0.07930   0.49193  0.161  0.8723
+% myc_probes          -0.04517  0.22130  -0.204 0.8387
+% pr_probes           0.59554   0.23590  2.525  0.0133   *
+% (Intercept)         0.221757  0.376612 0.589  0.5574
+% bmi                 -0.001994 0.003688 -0.540 0.5902
+% bcat_probes         0.114479  0.219311 0.522  0.6029
+% er_probes           0.080247  0.247050 0.325  0.7461
+% ifna_probes         -0.208580 0.485954 -0.429 0.6688
+% ifng_probes         0.116119  0.498505 0.233  0.8163
+% myc_probes          -0.027845 0.224460 -0.124 0.9015
+% pr_probes           0.603222  0.237235 2.543  0.0127   *
+% (Intercept)         0.195965  0.367384 0.533  0.5951
+% bmiStatusobese      -0.096834 0.063844 -1.517 0.1328
+% bmiStatusoverweight -0.004412 0.069155 -0.064 0.9493
+% bcat_probes         0.150379  0.218719 0.688  0.4935
+% er_probes           0.058880  0.244961 0.240  0.8106
+% ifna_probes         -0.344126 0.487134 -0.706 0.4817
+% ifng_probes         0.242697  0.497893 0.487  0.6271
+% myc_probes          -0.001285 0.222407 -0.006 0.9954
+% pr_probes           0.591191  0.234582 2.520  0.0135   *
+% (Intercept)         0.041459  0.379420 0.109  0.9132
+% bmi                 0.009426  0.006335 1.488  0.1403
+% bmiStatusobese      -0.227717 0.108446 -2.100 0.0386   *
+% bmiStatusoverweight -0.051853 0.075733 -0.685 0.4953
+% bcat_probes         0.146687  0.217273 0.675  0.5013
+% er_probes           0.034350  0.243884 0.141  0.8883
+% ifna_probes         -0.368065 0.484150 -0.760 0.4491
+% ifng_probes         0.242870  0.494570 0.491  0.6246
+% myc_probes          -0.019634 0.221267 -0.089 0.9295
+% pr_probes           0.554701  0.234304 2.367  0.0201   *
+
+% [1] "rescrolgene"
+
+% Call:
+% lm(formula = as.formula(formula_txt[j]), data = data)
+
+% Coefficients:
+%               Estimate Std. Error t value Pr(>|t|)
+% (Intercept)         0.5343899  0.1239554 4.311  3.9e-05 ***
+% bmi                 -0.0009922 0.0040730 -0.244 0.808
+% (Intercept)         0.54870    0.05482   10.009 <2e-16  ***
+% bmiStatusobese      -0.09063   0.07044   -1.287 0.201
+% bmiStatusoverweight -0.01515   0.07753   -0.195 0.845
+% (Intercept)         0.304200   0.164179  1.853  0.0670  .
+% bmi                 0.011054   0.007003  1.578  0.1178
+% bmiStatusobese      -0.243160  0.119267  -2.039 0.0442  *
+% bmiStatusoverweight -0.072189  0.084997  -0.849 0.3978
+% (Intercept)         0.17069    0.37929   0.450  0.6537
+% bcat_probes         0.09532    0.22177   0.430  0.6683
+% er_probes           0.03171    0.25251   0.126  0.9004
+% ifna_probes         -0.57965   0.49283   -1.176 0.2426
+% ifng_probes         0.50919    0.50478   1.009  0.3157
+% myc_probes          0.06345    0.22708   0.279  0.7805
+% pr_probes           0.54200    0.24206   2.239  0.0276  *
+% (Intercept)         0.204758   0.386507  0.530  0.5976
+% bmi                 -0.001950  0.003785  -0.515 0.6076
+% bcat_probes         0.112257   0.225072  0.499  0.6192
+% er_probes           0.033134   0.253540  0.131  0.8963
+% ifna_probes         -0.611750  0.498720  -1.227 0.2231
+% ifng_probes         0.545210   0.511601  1.066  0.2894
+% myc_probes          0.080405   0.230357  0.349  0.7279
+% pr_probes           0.549516   0.243468  2.257  0.0264  *
+% (Intercept)         0.17933    0.37777   0.475  0.636
+% bmiStatusobese      -0.09681   0.06565   -1.475 0.144
+% bmiStatusoverweight -0.01219   0.07111   -0.171 0.864
+% bcat_probes         0.14969    0.22490   0.666  0.507
+% er_probes           0.01354    0.25189   0.054  0.957
+% ifna_probes         -0.73888   0.50091   -1.475 0.144
+% ifng_probes         0.66381    0.51197   1.297  0.198
+% myc_probes          0.10813    0.22869   0.473  0.638
+% pr_probes           0.53872    0.24121   2.233  0.028   *
+% (Intercept)         0.027729   0.390580  0.071  0.9436
+% bmi                 0.009248   0.006522  1.418  0.1596
+% bmiStatusobese      -0.225233  0.111636  -2.018 0.0466  *
+% bmiStatusoverweight -0.058734  0.077961  -0.753 0.4532
+% bcat_probes         0.146066   0.223664  0.653  0.5154
+% er_probes           -0.010526  0.251058  -0.042 0.9667
+% ifna_probes         -0.762373  0.498391  -1.530 0.1296
+% ifng_probes         0.663981   0.509117  1.304  0.1955
+% myc_probes          0.090122   0.227775  0.396  0.6933
+% pr_probes           0.502919   0.241196  2.085  0.0399  *
+
+% [1] "caobsgene"
+
+% Call:
+% lm(formula = as.formula(formula_txt[j]), data = data)
+
+% Coefficients:
+%              Estimate Std. Error t value Pr(>|t|)
+% (Intercept)         0.538968  0.123943 4.349  3.39e-05 ***
+% bmi                 -0.001147 0.004073 -0.282 0.779
+% (Intercept)         0.53896   0.05458  9.875  2.76e-16 ***
+% bmiStatusobese      -0.08982  0.07013  -1.281 0.203
+% bmiStatusoverweight 0.01804   0.07719  0.234  0.816
+% (Intercept)         0.27986   0.16317  1.715  0.0896   .
+% bmi                 0.01171   0.00696  1.683  0.0957   .
+% bmiStatusobese      -0.25146  0.11853  -2.121 0.0365   *
+% bmiStatusoverweight -0.04241  0.08447  -0.502 0.6168
+% (Intercept)         0.12454   0.38803  0.321  0.7490
+% bcat_probes         0.14068   0.22688  0.620  0.5367
+% er_probes           0.11597   0.25833  0.449  0.6545
+% ifna_probes         -0.33319  0.50420  -0.661 0.5104
+% ifng_probes         0.30871   0.51641  0.598  0.5515
+% myc_probes          -0.04569  0.23232  -0.197 0.8445
+% pr_probes           0.56692   0.24764  2.289  0.0243   *
+% (Intercept)         0.162063  0.395326 0.410  0.6828
+% bmi                 -0.002148 0.003872 -0.555 0.5804
+% bcat_probes         0.159335  0.230208 0.692  0.4906
+% er_probes           0.117547  0.259326 0.453  0.6514
+% ifna_probes         -0.368552 0.510101 -0.723 0.4718
+% ifng_probes         0.348376  0.523276 0.666  0.5072
+% myc_probes          -0.027017 0.235613 -0.115 0.9090
+% pr_probes           0.575201  0.249024 2.310  0.0232   *
+% (Intercept)         0.134514  0.384960 0.349  0.7276
+% bmiStatusobese      -0.099695 0.066899 -1.490 0.1397
+% bmiStatusoverweight 0.008987  0.072463 0.124  0.9016
+% bcat_probes         0.193430  0.229183 0.844  0.4009
+% er_probes           0.092451  0.256681 0.360  0.7196
+% ifna_probes         -0.522336 0.510440 -1.023 0.3089
+% ifng_probes         0.492130  0.521713 0.943  0.3481
+% myc_probes          -0.001892 0.233048 -0.008 0.9935
+% pr_probes           0.560583  0.245805 2.281  0.0249   *
+% (Intercept)         -0.028482 0.397505 -0.072 0.9430
+% bmi                 0.009944  0.006637 1.498  0.1376
+% bmiStatusobese      -0.237770 0.113616 -2.093 0.0392   *
+% bmiStatusoverweight -0.041061 0.079343 -0.518 0.6061
+% bcat_probes         0.189535  0.227630 0.833  0.4073
+% er_probes           0.066573  0.255509 0.261  0.7950
+% ifna_probes         -0.547591 0.507227 -1.080 0.2832
+% ifng_probes         0.492312  0.518143 0.950  0.3446
+% myc_probes          -0.021250 0.231813 -0.092 0.9272
+% pr_probes           0.522087  0.245472 2.127  0.0362   *
+
+% [1] "cacrolgene"
+
+% Call:
+% lm(formula = as.formula(formula_txt[j]), data = data)
+
+% Coefficients:
+%               Estimate Std. Error t value Pr(>|t|)
+% (Intercept)         5.062e-01  1.240e-01 4.083  9.16e-05 ***
+% bmi                 -3.953e-05 4.074e-03 -0.010 0.992
+% (Intercept)         0.52453    0.05485   9.564  1.29e-15 ***
+% bmiStatusobese      -0.06411   0.07047   -0.910 0.365
+% bmiStatusoverweight 0.02958    0.07756   0.381  0.704
+% (Intercept)         0.273274   0.164135  1.665  0.0992   .
+% bmi                 0.011360   0.007001  1.622  0.1080
+% bmiStatusobese      -0.220856  0.119235  -1.852 0.0671   .
+% bmiStatusoverweight -0.029032  0.084974  -0.342 0.7334
+% (Intercept)         0.02932    0.40048   0.073  0.9418
+% bcat_probes         0.23660    0.23416   1.010  0.3150
+% er_probes           0.09521    0.26662   0.357  0.7218
+% ifna_probes         -0.49129   0.52037   -0.944 0.3476
+% ifng_probes         0.44466    0.53298   0.834  0.4063
+% myc_probes          0.16111    0.23977   0.672  0.5033
+% pr_probes           0.49566    0.25558   1.939  0.0555   .
+% (Intercept)         0.051698   0.408467  0.127  0.8996
+% bmi                 -0.001281  0.004000  -0.320 0.7495
+% bcat_probes         0.247720   0.237861  1.041  0.3004
+% er_probes           0.096149   0.267946  0.359  0.7205
+% ifna_probes         -0.512383  0.527057  -0.972 0.3335
+% ifng_probes         0.468320   0.540670  0.866  0.3887
+% myc_probes          0.172240   0.243445  0.708  0.4811
+% pr_probes           0.500601   0.257301  1.946  0.0548   .
+% (Intercept)         0.03806    0.39931   0.095  0.9243
+% bmiStatusobese      -0.07973   0.06939   -1.149 0.2536
+% bmiStatusoverweight 0.02249    0.07516   0.299  0.7655
+% bcat_probes         0.27648    0.23772   1.163  0.2479
+% er_probes           0.07298    0.26625   0.274  0.7846
+% ifna_probes         -0.66043   0.52946   -1.247 0.2155
+% ifng_probes         0.60854    0.54116   1.125  0.2638
+% myc_probes          0.19456    0.24173   0.805  0.4230
+% pr_probes           0.48849    0.25497   1.916  0.0586   .
+% (Intercept)         -0.123986  0.412743  -0.300 0.7646
+% bmi                 0.009886   0.006892  1.434  0.1549
+% bmiStatusobese      -0.216997  0.117971  -1.839 0.0692   .
+% bmiStatusoverweight -0.027267  0.082384  -0.331 0.7414
+% bcat_probes         0.272606   0.236356  1.153  0.2518
+% er_probes           0.047249   0.265304  0.178  0.8591
+% ifna_probes         -0.685542  0.526672  -1.302 0.1964
+% ifng_probes         0.608724   0.538006  1.131  0.2609
+% myc_probes          0.175314   0.240700  0.728  0.4683
+% pr_probes           0.450215   0.254882  1.766  0.0808   .
+
+% [1] "caresobsgene"
+
+% Call:
+% lm(formula = as.formula(formula_txt[j]), data = data)
+
+% Coefficients:
+%             Estimate Std. Error t value Pr(>|t|)
+% (Intercept)         0.462424  0.123913 3.732  0.00032  ***
+% bmi                 0.001442  0.004072 0.354  0.72408
+% (Intercept)         0.46970   0.05452  8.616  1.39e-13 ***
+% bmiStatusobese      0.09338   0.07005  1.333  0.186
+% bmiStatusoverweight -0.01840  0.07710  -0.239 0.812
+% (Intercept)         0.722057  0.163100 4.427  2.55e-05 ***
+% bmi                 -0.011409 0.006957 -1.640 0.1043
+% bmiStatusobese      0.250806  0.118483 2.117  0.0369   *
+% bmiStatusoverweight 0.040473  0.084438 0.479  0.6328
+% (Intercept)         1.1017    0.3815   2.888  0.00483  **
+% bcat_probes         -0.2217   0.2230   -0.994 0.32284
+% er_probes           -0.2308   0.2540   -0.909 0.36579
+% ifna_probes         0.1853    0.4957   0.374  0.70938
+% ifng_probes         -0.1528   0.5077   -0.301 0.76408
+% myc_probes          -0.0737   0.2284   -0.323 0.74765
+% pr_probes           -0.6876   0.2434   -2.825 0.00581  **
+% (Intercept)         1.056654  0.388317 2.721  0.00780  **
+% bmi                 0.002580  0.003803 0.678  0.49929
+% bcat_probes         -0.244104 0.226127 -1.079 0.28322
+% er_probes           -0.232713 0.254728 -0.914 0.36335
+% ifna_probes         0.227770  0.501056 0.455  0.65049
+% ifng_probes         -0.200460 0.513998 -0.390 0.69745
+% myc_probes          -0.096122 0.231435 -0.415 0.67888
+% pr_probes           -0.697585 0.244608 -2.852 0.00538  **
+% (Intercept)         1.09174   0.37821  2.887  0.00488  **
+% bmiStatusobese      0.10153   0.06573  1.545  0.12592
+% bmiStatusoverweight -0.00544  0.07119  -0.076 0.93926
+% bcat_probes         -0.27598  0.22517  -1.226 0.22351
+% er_probes           -0.20770  0.25218  -0.824 0.41234
+% ifna_probes         0.37360   0.50149  0.745  0.45823
+% ifng_probes         -0.33545  0.51257  -0.654 0.51449
+% myc_probes          -0.11868  0.22896  -0.518 0.60548
+% pr_probes           -0.68170  0.24150  -2.823 0.00586  **
+% (Intercept)         1.235747  0.391476 3.157  0.00218  **
+% bmi                 -0.008785 0.006537 -1.344 0.18237
+% bmiStatusobese      0.223518  0.111892 1.998  0.04881  *
+% bmiStatusoverweight 0.038776  0.078139 0.496  0.62094
+% bcat_probes         -0.272540 0.224177 -1.216 0.22730
+% er_probes           -0.184837 0.251634 -0.735 0.46455
+% ifna_probes         0.395907  0.499535 0.793  0.43015
+% ifng_probes         -0.335611 0.510285 -0.658 0.51243
+% myc_probes          -0.101582 0.228298 -0.445 0.65743
+% pr_probes           -0.647688 0.241749 -2.679 0.00879  **
+
+% [1] "carescrolgene"
+
+% Call:
+% lm(formula = as.formula(formula_txt[j]), data = data)
+
+% Coefficients:
+%              Estimate Std. Error t value Pr(>|t|)
+% (Intercept)         0.4928357 0.1239867 3.975  0.000136 ***
+% bmi                 0.0004131 0.0040740 0.101  0.919450
+% (Intercept)         0.52814   0.05483   9.633  9.13e-16 ***
+% bmiStatusobese      -0.06913  0.07045   -0.981 0.329
+% bmiStatusoverweight 0.02453   0.07753   0.316  0.752
+% (Intercept)         0.232845  0.163204  1.427  0.1569
+% bmi                 0.013351  0.006962  1.918  0.0582   .
+% bmiStatusobese      -0.253344 0.118558  -2.137 0.0352   *
+% bmiStatusoverweight -0.044356 0.084492  -0.525 0.6008
+% (Intercept)         -0.1205   0.3891    -0.310 0.7576
+% bcat_probes         0.2439    0.2275    1.072  0.2864
+% er_probes           0.1383    0.2590    0.534  0.5948
+% ifna_probes         -0.4364   0.5056    -0.863 0.3903
+% ifng_probes         0.4301    0.5178    0.831  0.4084
+% myc_probes          0.2837    0.2330    1.218  0.2264
+% pr_probes           0.5789    0.2483    2.331  0.0219   *
+% (Intercept)         -0.102321 0.396926  -0.258 0.7972
+% bmi                 -0.001038 0.003887  -0.267 0.7900
+% bcat_probes         0.252943  0.231140  1.094  0.2767
+% er_probes           0.139029  0.260375  0.534  0.5947
+% ifna_probes         -0.453502 0.512164  -0.885 0.3782
+% ifng_probes         0.449266  0.525393  0.855  0.3947
+% myc_probes          0.292758  0.236566  1.238  0.2191
+% pr_probes           0.582903  0.250031  2.331  0.0219   *
+% (Intercept)         -0.11188  0.38790   -0.288 0.7737
+% bmiStatusobese      -0.08019  0.06741   -1.190 0.2373
+% bmiStatusoverweight 0.01833   0.07302   0.251  0.8024
+% bcat_probes         0.28469   0.23093   1.233  0.2209
+% er_probes           0.11686   0.25864   0.452  0.6525
+% ifna_probes         -0.60153  0.51434   -1.170 0.2453
+% ifng_probes         0.59011   0.52570   1.123  0.2646
+% myc_probes          0.31782   0.23483   1.353  0.1793
+% pr_probes           0.57227   0.24768   2.311  0.0231   *
+% (Intercept)         -0.284628 0.400004  -0.712 0.4786
+% bmi                 0.010538  0.006679  1.578  0.1181
+% bmiStatusobese      -0.226528 0.114330  -1.981 0.0506   .
+% bmiStatusoverweight -0.034711 0.079842  -0.435 0.6648
+% bcat_probes         0.280559  0.229061  1.225  0.2239
+% er_probes           0.089437  0.257116  0.348  0.7288
+% ifna_probes         -0.628292 0.510417  -1.231 0.2216
+% ifng_probes         0.590302  0.521401  1.132  0.2606
+% myc_probes          0.297308  0.233271  1.275  0.2058
+% pr_probes           0.531475  0.247015  2.152  0.0341   *
+
+% [1] "crobsgenes"
+
+% Call:
+% lm(formula = as.formula(formula_txt[j]), data = data)
+
+% Coefficients:
+%             Estimate Std. Error t value Pr(>|t|)
+% (Intercept)         0.456126  0.123888 3.682  0.000381 ***
+% bmi                 0.001654  0.004071 0.406  0.685322
+% (Intercept)         0.459596  0.054606 8.417  3.7e-13  ***
+% bmiStatusobese      0.100893  0.070167 1.438  0.154
+% bmiStatusoverweight 0.005772  0.077224 0.075  0.941
+% (Intercept)         0.704210  0.163518 4.307  4.04e-05 ***
+% bmi                 -0.011059 0.006975 -1.586 0.1162
+% bmiStatusobese      0.253491  0.118786 2.134  0.0354   *
+% bmiStatusoverweight 0.062836  0.084654 0.742  0.4598
+% (Intercept)         1.0667    0.3806   2.803  0.00618  **
+% bcat_probes         -0.2284   0.2225   -1.026 0.30740
+% er_probes           -0.1481   0.2534   -0.584 0.56041
+% ifna_probes         0.4504    0.4945   0.911  0.36477
+% ifng_probes         -0.4013   0.5065   -0.792 0.43018
+% myc_probes          -0.1402   0.2279   -0.615 0.53986
+% pr_probes           -0.6446   0.2429   -2.654 0.00938  **
+% (Intercept)         1.014279  0.387058 2.620  0.01029  *
+% bmi                 0.003004  0.003791 0.792  0.43014
+% bcat_probes         -0.254482 0.225394 -1.129 0.26184
+% er_probes           -0.150261 0.253902 -0.592 0.55545
+% ifna_probes         0.499868  0.499432 1.001  0.31954
+% ifng_probes         -0.456812 0.512332 -0.892 0.37494
+% myc_probes          -0.166314 0.230685 -0.721 0.47278
+% pr_probes           -0.656157 0.243815 -2.691 0.00847  **
+% (Intercept)         1.05662   0.37693  2.803  0.0062   **
+% bmiStatusobese      0.11346   0.06550  1.732  0.0867   .
+% bmiStatusoverweight 0.01412   0.07095  0.199  0.8427
+% bcat_probes         -0.29209  0.22440  -1.302 0.1964
+% er_probes           -0.12674  0.25133  -0.504 0.6153
+% ifna_probes         0.63723   0.49979  1.275  0.2056
+% ifng_probes         -0.58273  0.51083  -1.141 0.2570
+% myc_probes          -0.19254  0.22819  -0.844 0.4010
+% pr_probes           -0.64072  0.24068  -2.662 0.0092   **
+% (Intercept)         1.198471  0.390241 3.071  0.00283  **
+% bmi                 -0.008654 0.006516 -1.328 0.18754
+% bmiStatusobese      0.233624  0.111539 2.095  0.03905  *
+% bmiStatusoverweight 0.057676  0.077893 0.740  0.46097
+% bcat_probes         -0.288700 0.223470 -1.292 0.19974
+% er_probes           -0.104217 0.250840 -0.415 0.67880
+% ifna_probes         0.659208  0.497958 1.324  0.18895
+% ifng_probes         -0.582884 0.508675 -1.146 0.25491
+% myc_probes          -0.175698 0.227577 -0.772 0.44214
+% pr_probes           -0.607216 0.240986 -2.520 0.01353  *
+
+% [1] "fmobsgenes"
+
+% Call:
+% lm(formula = as.formula(formula_txt[j]), data = data)
+
+% Coefficients:
+%             Estimate Std. Error t value Pr(>|t|)
+% (Intercept)         0.436628  0.123787 3.527  0.000643 ***
+% bmi                 0.002314  0.004067 0.569  0.570763
+% (Intercept)         0.58369   0.05435  10.740 <2e-16   ***
+% bmiStatusobese      -0.08452  0.06983  -1.210 0.2291
+% bmiStatusoverweight -0.14827  0.07686  -1.929 0.0567   .
+% (Intercept)         0.251170  0.160863 1.561  0.12176
+% bmi                 0.015034  0.006862 2.191  0.03091  *
+% bmiStatusobese      -0.291956 0.116858 -2.498 0.01419  *
+% bmiStatusoverweight -0.225840 0.083280 -2.712 0.00794  **
+% (Intercept)         0.33317   0.36900  0.903  0.3689
+% bcat_probes         -0.13489  0.21576  -0.625 0.5334
+% er_probes           0.11973   0.24567  0.487  0.6271
+% ifna_probes         -0.56244  0.47947  -1.173 0.2438
+% ifng_probes         0.58838   0.49109  1.198  0.2340
+% myc_probes          0.41768   0.22093  1.891  0.0618   .
+% pr_probes           -0.08815  0.23550  -0.374 0.7090
+% (Intercept)         0.291390  0.375705 0.776  0.4400
+% bmi                 0.002392  0.003679 0.650  0.5172
+% bcat_probes         -0.155669 0.218783 -0.712 0.4786
+% er_probes           0.117981  0.246455 0.479  0.6333
+% ifna_probes         -0.523056 0.484783 -1.079 0.2835
+% ifng_probes         0.544202  0.497304 1.094  0.2767
+% myc_probes          0.396892  0.223919 1.772  0.0797   .
+% pr_probes           -0.097371 0.236664 -0.411 0.6817
+% (Intercept)         0.33459   0.36629  0.913  0.3634
+% bmiStatusobese      -0.08029  0.06366  -1.261 0.2104
+% bmiStatusoverweight -0.12492  0.06895  -1.812 0.0733   .
+% bcat_probes         -0.07254  0.21807  -0.333 0.7402
+% er_probes           0.13036   0.24424  0.534  0.5948
+% ifna_probes         -0.56035  0.48569  -1.154 0.2517
+% ifng_probes         0.58760   0.49642  1.184  0.2397
+% myc_probes          0.46651   0.22175  2.104  0.0382   *
+% pr_probes           -0.07506  0.23389  -0.321 0.7490
+% (Intercept)         0.082260  0.370293 0.222  0.82471
+% bmi                 0.015393  0.006183 2.490  0.01464  *
+% bmiStatusobese      -0.294047 0.105838 -2.778 0.00666  **
+% bmiStatusoverweight -0.202402 0.073911 -2.738 0.00746  **
+% bcat_probes         -0.078569 0.212047 -0.371 0.71187
+% er_probes           0.090296  0.238018 0.379  0.70532
+% ifna_probes         -0.599445 0.472504 -1.269 0.20787
+% ifng_probes         0.587878  0.482673 1.218  0.22646
+% myc_probes          0.436540  0.215944 2.022  0.04623  *
+% pr_probes           -0.134650 0.228668 -0.589 0.55746
+
+\begin{longtable}{llr{\bfseries}S}
+	\centering
+	\caption{Table showing the variables in the step wise linear models to predict the obesity metagenes}
+	\label{tab:stepwise_model}\\
+		Obesity metagene & Variables used in the model & Estimate & {P-value}\\
+		\hline
+	\endfirsthead
+	\multicolumn{4}{c}{\tablename\ \thetable{}\ (continued)}\\
+		\hline
+		\hline
+	\endhead
+		\hline
+		\hline
+		Cr         & Akt         & -0.3415   & \bfseries \num{4.43d-8}\\
+				   & \gls{egfr}  & 0.33852   & \bfseries \num{1.94d-05}\\
+				   & \gls{er}    & -0.20237  & \bfseries 0.00419\\
+				   & \gls{ifny}  & 0.12233   & 0.08700\\
+				   & \gls{pr}    & -0.33920  & \bfseries \num{2.09d-5}\\
+				   & Ras         & -0.22611  & \bfseries 0.00193\\
+				   & Src         & -0.11452  & 0.08177\\
+				   & \gls{tgfb}  & 0.55638   & \bfseries \textless{} \num{2d-16}\\
+				   & \gls{tnfa}  & -0.25501  & \bfseries 0.00616\\
+		\hline
+		CrOl       & Overweight  & 0.040918  & 0.074865\\
+				   & Obese       & -0.006381 & 0.756121\\
+				   & Akt         & -0.356713 & \bfseries \num{1.89d-07}\\
+				   & E2F1        & 0.160229  & \bfseries 0.005411\\
+				   & \gls{egfr}  & 0.148618  & 0.062683\\
+				   & p53         & 0.336552  & \bfseries \num{4.35d-05}\\
+				   & \gls{pr}    & -0.225423 & \bfseries 0.000639\\
+				   & Src         & -0.428758 & \bfseries \num{5.85d-10}\\
+				   & \gls{stat3} & -0.124118 & \bfseries 0.009658\\
+		\hline
+		Res        & Akt         & 0.24613   & \bfseries 0.000289\\
+				   & \gls{egfr}  & -0.46827  & \bfseries \num{2.16d-05}\\
+				   & \gls{er}    & 0.18400   & 0.061446\\
+				   & \gls{her2}  & 0.13835   & 0.118228\\
+				   & \gls{ifny}  & -0.21019  & \bfseries 0.044084\\
+				   & \gls{pr}    & 0.41195   & \bfseries 0.000300\\
+				   & Ras         & 0.26506   & \bfseries 0.010465\\
+				   & \gls{tgfb}  & -0.58075  & \bfseries \num{7.64d-11}\\
+				   & \gls{tnfa}  & 0.42200   & \bfseries 0.001834\\
+		\hline
+		ResOl      & Akt         & -0.33596  & \bfseries 0.000294\\
+				   & \gls{bcat}  & -0.19866  & 0.138953\\
+				   & E2F1        & 0.36059   & \bfseries 0.009325\\
+				   & \gls{egfr}  & 0.23421   & 0.056865\\
+				   & \gls{her2}  & -0.13652  & 0.121689\\
+				   & p53         & 0.38746   & \bfseries 0.000519\\
+				   & \gls{pi3k}  & -0.21881  & 0.148676\\
+				   & \gls{pr}    & -0.32228  & \bfseries 0.001656\\
+				   & Ras         & -0.28984  & \bfseries 0.033975\\
+				   & Src         & -0.40779  & \bfseries \num{5.26d-06}\\
+				   & \gls{stat3} & -0.11452  & 0.125609\\
+				   & \gls{tnfa}  & -0.11442  & 0.117023\\
+		\hline
+		Ca         & Akt         & -0.26243  & \bfseries \num{7.04d-06}\\
+				   & \gls{egfr}  & 0.36305   & \bfseries 0.000139\\
+				   & \gls{er}    & -0.18614  & \bfseries 0.030771\\
+				   & \gls{ifny}  & 0.22118   & \bfseries 0.013034\\
+				   & \gls{pr}    & -0.42911  & \bfseries \num{1.07d-05}\\
+				   & Ras         & -0.28124  & \bfseries 0.000898\\
+				   & \gls{tgfb}  & 0.76139   & \bfseries \textless{} \num{2d-16}\\
+				   & \gls{tnfa}  & -0.33200  & \bfseries 0.003919\\
+		\hline
+		CaOl       & Overweight  & 0.07613   & \bfseries 0.006875\\
+				   & Obese       & 0.01408   & 0.564622\\
+				   & Akt         & -0.44727  & \bfseries \num{2.43d-07}\\
+				   & \gls{bcat}  & -0.13463  & 0.140261\\
+				   & E2F1        & 0.33864   & \bfseries 0.001249\\
+				   & p53         & 0.28777   & \bfseries 0.000996\\
+				   & \gls{pr}    & -0.28255  & \bfseries \num{2.18d-05}\\
+				   & Src         & -0.44149  & \bfseries \num{2.02d-08}\\
+				   & \gls{tnfa}  & -0.14215  & \bfseries 0.013897\\
+		\hline
+		CaRes      & \gls{bcat}  & 0.26899   & 0.05238  \\
+				   & E2F1        & -0.22115  & 0.08848  \\
+				   & \gls{egfr}  & 0.44400   & \bfseries \num{6.67d-05 }\\
+				   & \gls{her2}  & -0.20304  & \bfseries 0.04615  \\
+				   & \gls{ifny}  & 0.32874   & \bfseries 0.00574  \\
+				   & \gls{pr}    & -0.28764  & \bfseries 0.00104  \\
+				   & Src         & -0.17989  & 0.05404  \\
+				   & \gls{tgfb}  & 0.52898   & \bfseries \num{2.15d-06 }\\
+				   & \gls{tnfa}  & -0.38677  & \bfseries 0.00564  \\
+		\hline
+		CaResOl    & \gls{bmi}   & 0.004314  & 0.108546\\
+				   & Overweight  & 0.033086  & 0.306390\\
+				   & Obese       & -0.062257 & 0.175776\\
+				   & Akt         & -0.309706 & \bfseries 0.000339\\
+				   & \gls{bcat}  & -0.198165 & 0.140502\\
+				   & E2F1        & 0.270471  & \bfseries 0.011450\\
+				   & \gls{egfr}  & 0.171219  & 0.113691\\
+				   & Myc         & 0.152007  & 0.056987\\
+				   & p53         & 0.294895  & \bfseries 0.003734\\
+				   & \gls{pr}    & -0.230896 & \bfseries 0.000823\\
+				   & Ras         & -0.270824 & \bfseries 0.044788\\
+				   & Src         & -0.512567 & \bfseries \num{2.72d-08}\\
+				   & \gls{tnfa}  & -0.103302 & 0.083449\\
+		\hline
+		Original   & \gls{bmi}   & 0.002696  & 0.14407\\
+				   & Overweight  & 0.010522  & 0.62856\\
+				   & Obese       & -0.061020 & 0.05542\\
+				   & Akt         & -0.238412 & \bfseries \num{4.93d-05}\\
+				   & E2F1        & 0.185085  & \bfseries 0.02754\\
+				   & \gls{egfr}  & 0.147219  & 0.07423\\
+				   & \gls{her2}  & -0.086804 & 0.14070\\
+				   & p53         & 0.416414  & \bfseries \num{5.83d-08}\\
+				   & \gls{pi3k}  & -0.216179 & \bfseries 0.03528\\
+				   & \gls{pr}    & -0.199097 & \bfseries 0.00253\\
+				   & Ras         & -0.200544 & \bfseries 0.00898\\
+				   & Src         & -0.543834 & \bfseries \num{2.10d-15}\\
+				   & \gls{stat3} & -0.093363 & \bfseries 0.02749\\
+		\hline
+		FM         & \gls{bmi}   & 0.011093  & \bfseries 0.036229\\
+				   & Overweight  & -0.153753 & \bfseries 0.014277\\
+				   & Obese       & -0.212483 & \bfseries 0.017378\\
+				   & p53         & 0.460800  & \bfseries 0.002512\\
+				   & \gls{pi3k}  & -0.426246 & \bfseries 0.000118\\
+				   & \gls{pr}    & -0.373119 & \bfseries 0.020737\\
+				   & \gls{stat3} & 0.248745  & \bfseries 0.019073\\
+				   & \gls{tgfb}  & 0.152925  & 0.128933\\
+		\hline
+		\hline
+	\end{longtable}
+
 \end{appendices}

### 2017-03-10 thesis/results/results.tex
Changed all the data set names and removed "'s" - tmp commit

@@
 Firstly in this chapter, the previously identified obesity associated signatures from the studies conducted by \citet{Creighton2012} and \citet{Fuentes-Mattei2014} were examined in turn to judge the agreement of these signatures with the sample \gls{bmi}/\gls{bmi} status, as presented in thier results.
-Secondly, novel obesity associated genetic signatures were identified in the \citet{Creighton2012} data set and was compared with Creighton's and Fuentes-Mattei's  signatures.
+Secondly, novel obesity associated genetic signatures were identified in the Creighton \textit{et al.} (CR) data set and was compared with the signatures from the Creighton \textit{et al.} and Fuentes-Mattei \textit{et al.} studies.
 Lastly, the presence of common genes or pathways that were associated with obesity in multiple types of cancer was explored using the \gls{icgc} cancer data sets.
@@
 
-Obesity metagene was created using the obesity associated genetic signatures from \citet{Creighton2012} study.
-% As the genetic signatures were described as gene probe IDs, the gene probes had to be converted into gene symbols (see \cref{sec:data}).
-% This resulted in 644 unique gene symbols, compared to 799 gene probes IDs in the original signature.
-There was no description about the normalisation method used by Creighton's group when they first analysed their data, so the more reliable \gls{rma} method was used to  normalise the Creighton data set.
+Obesity metagene was created using the obesity associated genetic signatures from the \citet{Creighton2012} study.
+There was no description about the normalisation method used by Creighton \textit{et al.} when they first analysed their data, so the more reliable \gls{rma} method was used to  normalise the CR data set.
 The obesity metagene was plotted in a heatmap with the sample gene expression  to check whether the metagene scores were in accordance with the overall gene expression of the samples (\cref{fig:crmetaheat}).
@@
 	\includegraphics[page=3,width=0.8\linewidth]{results1/creighton_mg_heatmap1}
-	\caption[Creighton's metagene and sample gene expression in Creighton's data]{Heatmap showing Creighton's obesity associated metagene with the gene expressions of obesity associated genes in Creighton's data.
+	\caption[Obesity metagene from \citet{Creighton2012} study and sample gene expression in CR data]{Heatmap showing the obesity metagene from the \citet{Creighton2012} study with the gene expressions of obesity associated genes in CR data.
 	Level of expression is represented in the top right histogram, where low and high gene expression were colour-coded with blue and red, respectively.
-	Each row of the heatmap represents a gene from the obesity associated genetic signature, and each column of the heatmap represents a sample from Creighton's data.
+	Each row of the heatmap represents a gene from the obesity associated genetic signature, and each column of the heatmap represents a sample from CR data.
 	The obesity associated metagene scores of the samples are shown in a separate row at top of the heatmap, and the tree diagram of the heirarchical clustering of the genes is shown to the left of the heatmap.
-	For clarity, the metagene scores were flipped in order to match the results from \citet{Creighton2012} paper.}
+	For clarity, the metagene scores were flipped in order to match the results from the \citet{Creighton2012} study.}
 	\label{fig:crmetaheat}
@@
 	\includegraphics[page=5,width=0.45\linewidth]{results1/creighton_mg_heatmap1}
-	\caption[Creighton's metagene and sample \gls{bmi}/\gls{bmi} status in Creighton's data]{Box plot and scatter plot showing the association of Creighton's obesity associated metagene with sample \gls{bmi} status and \gls{bmi} from Creighton's data, respectively.
+	\caption[Obesity metagene from the \citet{Creighton2012} study and sample \gls{bmi}/\gls{bmi} status in CR data]{Box plot and scatter plot showing the association of the obesity metagene from the \citet{Creighton2012} study with sample \gls{bmi} status and \gls{bmi} from CR data, respectively.
 	In the box plot, the p-values above the groups represent the statistical significance of the association of the metagene with the overweight or obese group compared with the normal weight group.
@@
 
-\cref{fig:crmetaboxplot} clearly showed that the obesity associated metagene from \citet{Creighton2012} study significantly associated with the sample \gls{bmi} status, as well as sample \gls{bmi} value (include p-value and r-squared).
+\cref{fig:crmetaboxplot} clearly showed that the obesity metagene from the \citet{Creighton2012} study significantly associated with the sample \gls{bmi} status, as well as sample \gls{bmi} value (include p-value and r-squared).
 It should be noted here that the obesity associated metagene significantly associated with the samples that were obese, but not with the samples that were overweight.
@@
 In addition to this, although the regression line in the scatter plot showed statistically significant association with sample \gls{bmi}, the sample \gls{bmi} values seemed to be randomly dispersed across the metagene scores.
-This suggested that perhaps Creighton's obesity metagene was not associated with sample \gls{bmi} as strongly as expected.
+This suggested that perhaps the obesity metagene from the \citet{Creighton2012} study was not associated with sample \gls{bmi} as strongly as expected.
 \\
@@
 \noindent
-Now that the association of the obesity metagene from \citet{Creighton2012} study was established in Creighton's data set, the obesity associated signature was generated in the \gls{icgc} cancer data.
-The direction of the obesity associated metagene was checked in the Creighton's data first, so that high metagene scores reflected high sample \gls{bmi} and low metagene scores reflected low sample \gls{bmi} (\cref{fig:crmetaboxplot}).
-The transformation matrix was then created in Creighton's data, as described in \cref{sub:svd}.
+Now that the association of the obesity metagene from the \citet{Creighton2012} study was established in CR data set, the obesity associated signature was generated in the \gls{icgc} cancer data.
+The direction of the obesity associated metagene was checked in the CR data first, so that high metagene scores reflected high sample \gls{bmi} and low metagene scores reflected low sample \gls{bmi} (\cref{fig:crmetaboxplot}).
+The transformation matrix was then created in CR data, as described in \cref{sub:svd}.
 
@@
 From these results, the standardised data was found to be the most suitable to use for matrix transformation.
-The transformation matrix was applied to each cancer data set in turn to obtain Creighton's obesity metagenes for each of the data set.
+The transformation matrix was applied to each cancer data set in turn to obtain the Creighton \textit{et al.} obesity metagenes from each of the data set.
 Each obesity metagenes were plotted in a heatmap with the corresponding data set in which the metagene was taken from (\cref{fig:crmetaicgc}; \cref{sec:rest_of_the_icgc_cancer_heatmap_resutls}).
-These heatmaps confirmed that the obesity associated metagene derived from Creighton's data was able to capture the overall gene expression pattern, where the metagene scores reflected the expression levels of majority of the genes in the signature.
+These heatmaps confirmed that the obesity associated metagene derived from CR data was able to capture the overall gene expression pattern, where the metagene scores reflected the expression levels of majority of the genes in the signature.
 As before, association of the obesity metagene with the sample \gls{bmi} and \gls{bmi} status was checked in their respective cancer data set (\cref{fig:crmetaicgc}; \cref{sec:rest_of_the_icgc_cancer_heatmap_resutls}).
-Out of all the cancer types, only \gls{blca} data set showed significant association of Creighton's obesity metagene with the overweight group (but not with the obese group).
-However, neither the \gls{anova} p-value nor the regression line in the scatter plot were statistically significant, suggesting that this apparent association with the overweight group was not due to the Creighton's obesity metagene (?).
+Out of all the cancer types, only \gls{blca} data set showed significant association of obesity metagene with the overweight group (but not with the obese group).
+However, neither the \gls{anova} p-value nor the regression line in the scatter plot were statistically significant, suggesting that this apparent association with the overweight group was not due to the obesity metagene (?).
 
@@
 	\includegraphics[page=5,width=0.45\linewidth]{results1/crtcga_std}
-	\caption[Creighton's metagene in \acrshort{icgc} \acrshort{blca} data]{Heatmap, box plot and scatter plot showing the association of Creighton's obesity associated metagene with sample gene expression, \gls{bmi} and \gls{bmi} status from \gls{icgc} \gls{blca} data, respectively.
+	\caption[Obesity metagene from the \citet{Creighton2012} study in \acrshort{icgc} \acrshort{blca} data]{Heatmap, box plot and scatter plot showing the association of obesity metagene from the \citet{Creighton2012} study with sample gene expression, \gls{bmi} and \gls{bmi} status from \gls{icgc} \gls{blca} data, respectively.
 	The results for other cancer types are in \cref{app:a}.
@@
 There could be several reasons for the apparent lack of association of the metagene with sample \gls{bmi}/\gls{bmi} status.
-First, transformation matrix was derived from Creighton's microarray data, but the \gls{icgc} cancer data were \gls{rnaseq} data.
+First, transformation matrix was derived from CR microarray data, but the \gls{icgc} cancer data were \gls{rnaseq} data.
 Though the log$_{10}$ normalisation and standardisation of the data was the most appropriate adjustment to be made to the \gls{rnaseq} data, this adjustment was not equivalent to the \gls{rma} normalisation method that was used on the microarray data.
-Secondly, none of the \gls{icgc} cancer originated from breast as in Creighton's data.
+Secondly, none of the \gls{icgc} cancer originated from breast as in CR data.
 Since the obesity associated signature was identified in breast cancer data, the signature may be specific to breast cancer and may not be suitable in other cancer types.
@@
 \noindent
-To check whether the metagene was specific to breast cancer microarray data, the same transformation matrix was applied to breast cancer microarray data from \citet{Print2016} study (referred to as Print's data).
-Print's data was normalised with \gls{rma} method and the transformation matrix was applied to the normalised data to obtain the metagene.
+To check whether the metagene was specific to breast cancer microarray data, the same transformation matrix was applied to the \gls{nzbc} microarray data from \citet{Print2016} study (referred to as \gls{nzbc} data hereafter).
+\gls{nzbc} data was normalised with \gls{rma} method and the transformation matrix was applied to the normalised data to obtain the metagene.
 The metagene was again compared with the gene expression of the samples with a heatmap and the association of the metagene with the sample \gls{bmi}/\gls{bmi} status was examined with box and scatter plots (\cref{fig:crmetaprint}).
 
-The obesity metagene managed to reflect the overall gene expression of the samples in Print's data.
+The obesity metagene managed to reflect the overall gene expression of the samples in \gls{nzbc} data.
 However, as with \gls{icgc} cancer data, the obesity metagene scores did not significantly associate with the sample \gls{bmi} or \gls{bmi} status.
-These results confirmed that the lack of association of the obesity metagene from Creighton's data was not due to the technology in which the data was gathered (mcroarray or \gls{rnaseq}), nor the cancer type in which the transformation matrix was applied to.
+These results confirmed that the lack of association of the obesity metagene from CR data was not due to the technology in which the data was gathered (mcroarray or \gls{rnaseq}), nor the cancer type in which the transformation matrix was applied to.
 
@@
 	\includegraphics[width=0.45\linewidth,page=5]{results1/cris_cr_trans_meta}
-	\caption[Creighton's metagene in Print's breast cancer data]{Heatmap, box plot and scatter plot showing the association of Creighton's obesity associated metagene with sample gene expression, \gls{bmi} and \gls{bmi} status from Print's breast cancer data, respectively.
+	\caption[Obesity metagene from \citet{Creighton2012} study in \gls{nzbc} data]{Heatmap, box plot and scatter plot showing the association of obesity metagene from the \citet{Creighton2012} study with sample gene expression, \gls{bmi} and \gls{bmi} status from \gls{nzbc} data, respectively.
 	Scales, p-values and $R^2$-value are as described in previous figures.}
@@
 
-All together, all of these results suggest that the obesity metagene identified in the study conducted by \citet{Creighton2012} associated with sample \gls{bmi} and \gls{bmi} status only in Creighton's data.
-The lack of association with \gls{bmi} was not due to the type of technology platform in which the data was gathered, as neither the \gls{icgc} \gls{rnaseq} data nor Print's microarray data showed no significant association with the obesity metagene.
-Furthermore, Creighton's metagene was not dependent on the cancer type since the obesity associated genetic signature did not show significant association in neither the \gls{icgc} cancer data nor in Print's breast cancer data.
+All together, all of these results suggest that the obesity metagene identified in the study conducted by \citet{Creighton2012} associated with sample \gls{bmi} and \gls{bmi} status only in CR data.
+The lack of association with \gls{bmi} was not due to the type of technology platform in which the data was gathered, as neither the \gls{icgc} \gls{rnaseq} data nor \gls{nzbc} microarray data showed no significant association with the obesity metagene.
+Furthermore, the obesity metagene was not dependent on the cancer type since the obesity associated genetic signature did not show significant association in neither the \gls{icgc} cancer data nor in \gls{nzbc} data.
 
-One possible reason why Creighton's obesity metagene did not show significant association with other data sets could be because the genetic signature was in fact not an obesity specific signature, but a signature that was detected due to another clinical variable.
-Another reason for this apparent lack of association could be that the genetic signature was too specific to the data and was not a broad obesity associated genetic signature, but an obesity associated signature specifically for Creighton's data set.
+One possible reason why the obesity metagene from the \citet{Creighton2012} study did not show significant association with other data sets could be because the genetic signature was in fact not an obesity specific signature, but a signature that was detected due to another clinical variable.
+Another reason for this apparent lack of association could be that the genetic signature was too specific to the data and was not a broad obesity associated genetic signature, but an obesity associated signature specifically for CR data set.
 
@@
 
-Since Creighton's obesity associated genetic signature was not associated with sample \gls{bmi} or \gls{bmi} status in other data sets, Fuentes-Mattei's (FM) obesity associated genetic signature was examined to see whether FM's obesity metagene associated with sample \gls{bmi} or \gls{bmi} status.
-Since FM's data set did not have sample \gls{bmi} information, FM's obesity metagene was not able to be compared with the sample \gls{bmi} or \gls{bmi} status in the original FM's data.
-However, the transformation matrix was still applied to other cancer data to see whether FM's obesity metagene associated with sample \gls{bmi}/\gls{bmi} status
+Since the obesity associated genetic signature from the \citet{Creighton2012} study was not associated with sample \gls{bmi} or \gls{bmi} status in other data sets, the obesity associated genetic signature from the \citet{Fuentes-Matter2014} study (FM) was examined to see whether FM obesity metagene associated with sample \gls{bmi} or \gls{bmi} status.
+Since FM data set did not have sample \gls{bmi} information, FM obesity metagene was not able to be compared with the sample \gls{bmi} or \gls{bmi} status in the original FM data.
+However, the transformation matrix was still applied to other cancer data to see whether FM obesity metagene associated with sample \gls{bmi}/\gls{bmi} status
 
-Firstly, FM's data was normalised with the \gls{rma} method and \gls{svd} was applied to the normalised data to get the transformation matrix.
-The transformation matrix was used to transform the \gls{rma}-normalised Creighton's data to extract FM's obesity metagene scores in Creighton's data.
-FM's obesity metagene scores were compared with gene expressions and sample \gls{bmi}/\gls{bmi} status in Creighton's data, as shown in \cref{fig:fmmetacr}.
-Clearly, like with Creighton's obesity metagene, FM's obesity metagene was reflective of the overall gene expression of the samples, but did not associate with the sample \gls{bmi} or \gls{bmi} status.
+Firstly, FM data was normalised with the \gls{rma} method and \gls{svd} was applied to the normalised data to get the transformation matrix.
+The transformation matrix was used to transform the \gls{rma}-normalised CR data to extract FM obesity metagene scores in CR data.
+FM obesity metagene scores were compared with gene expressions and sample \gls{bmi}/\gls{bmi} status in CR data, as shown in \cref{fig:fmmetacr}.
+Clearly, like with the obesity metagene identified by Creighton \textit{et al.}, FM obesity metagene was reflective of the overall gene expression of the samples, but did not associate with the sample \gls{bmi} or \gls{bmi} status.
 
@@
 	\includegraphics[page=5,width=0.45\linewidth]{results1/cr_fm_meta}
-	\caption[FM's metagene in Creighton's  data]{Heatmap, box plot and scatter plot showing the association of FM's obesity associated metagene with sample gene expression, \gls{bmi} and \gls{bmi} status from Creighton's  data, respectively.
+	\caption[FM metagene in CR data]{Heatmap, box plot and scatter plot showing the association of FM obesity associated metagene with sample gene expression, \gls{bmi} and \gls{bmi} status from CR  data, respectively.
 	Scales, p-values and $R^2$-value are as described in previous figures.}
@@
 Next, the transformation matrix was applied to the \gls{icgc} cancer data and resulting metagenes were compared with the gene expression and sample \gls{bmi}/\gls{bmi} status.
-As evident in \cref{fig:fmmetaicgc}, FM's obesity metagene scores appeared to reflect the overall gene expression of FM's obesity associated genetic signature.
-As with all the results in this chapter so far, FM's obesity metagene did not significantly associate with any of the \gls{icgc} cancer data, except in \gls{blca} data set (\cref{fig:fmmetaicgc}).
-FM's obesity metagene significantly associated with the overweight group (but not with the obese group), and also had a significant \gls{anova} p-value.
-On the contrary to the association of the metagene with the sample \gls{bmi} status, FM's obesity metagene was not associated with sample \gls{bmi}.
-These results suggested that the samples that were overweight in the \gls{blca} data set had similar biological properties as the samples that were obese in FM's data set.
-However, due to the fact that FM's metagene lacked association with sample \gls{bmi} in \gls{blca} data set and that the metagene did not show any significant association in any other cancer type, it was difficult to determine whether the observed association of FM's obesity metagene with the overweight group of samples was truly reflective of the effect resulted from FM's metagene.
-Lastly, the transformation matrix was applied to Print's data (\cref{fig:fmmetacris}).
-Again, FM's obesity metagene scores reflected the gene expression of the samples, did not significantly associate with sample \gls{bmi}/\gls{bmi} status.
+As evident in \cref{fig:fmmetaicgc}, FM obesity metagene scores appeared to reflect the overall gene expression of FM obesity associated genetic signature.
+As with all the results in this chapter so far, FM obesity metagene did not significantly associate with any of the \gls{icgc} cancer data, except in \gls{blca} data set (\cref{fig:fmmetaicgc}).
+FM obesity metagene significantly associated with the overweight group (but not with the obese group), and also had a significant \gls{anova} p-value.
+On the contrary to the association of the metagene with the sample \gls{bmi} status, FM obesity metagene was not associated with sample \gls{bmi}.
+These results suggested that the samples that were overweight in the \gls{blca} data set had similar biological properties as the samples that were obese in FM data set.
+However, due to the fact that FM metagene lacked association with sample \gls{bmi} in \gls{blca} data set and that the metagene did not show any significant association in any other cancer type, it was difficult to determine whether the observed association of FM obesity metagene with the overweight group of samples was truly reflective of the effect resulted from FM metagene.
+Lastly, the transformation matrix was applied to \gls{nzbc} data (\cref{fig:fmmetacris}).
+Again, FM obesity metagene scores reflected the gene expression of the samples, did not significantly associate with sample \gls{bmi}/\gls{bmi} status.
 \\
@@
 	\includegraphics[page=5,width=0.45\linewidth]{results1/fm_meta_ICGC}
-	\caption[FM's metagene in \acrshort{icgc} \acrshort{blca} data]{Heatmap, box plot and scatter plot showing the association of FM's obesity associated metagene with sample gene expression, \gls{bmi} and \gls{bmi} status from \acrshort{icgc} \acrshort{blca} data, respectively.
+	\caption[FM metagene in \acrshort{icgc} \acrshort{blca} data]{Heatmap, box plot and scatter plot showing the association of FM obesity associated metagene with sample gene expression, \gls{bmi} and \gls{bmi} status from \acrshort{icgc} \acrshort{blca} data, respectively.
 	The results for other cancer types are in \cref{app:a}.
@@
 	\includegraphics[page=5,width=0.45\linewidth]{results1/cris_fm_meta}
-	\caption[FM's metagene in Print's breast cancer data]{Heatmap, box plot and scatter plot showing the association of FM's obesity associated metagene with sample gene expression, \gls{bmi} and \gls{bmi} status from Print's breast cancer data, respectively.
+	\caption[FM metagene in \gls{nzbc} data]{Heatmap, box plot and scatter plot showing the association of FM obesity associated metagene with sample gene expression, \gls{bmi} and \gls{bmi} status from \gls{nzbc} data, respectively.
 	Scales, p-values and $R^2$-value are as described in previous figures.}
@@
 \noindent
-These results showed that FM's obesity associated metagene was not transferable to other cancer data sets, similar to Creighton's obesity associated metagene.
-This meant that both Creighton's and FM's obesity associate metagenes may have been too specific to the orignial data set in which the signatures were identified in.
+These results showed that FM obesity associated metagene was not transferable to other cancer data sets, similar to the obesity metagene identified by Creighton \textit{et al.}.
+This meant that both the obesity metagenes identified by Creighton \textit{et al.} and Fuentes-Mattei \textit{et al.} may have been too specific to the orignial data set in which the signatures were identified in.
 Furthermore, there was a possibility that these obesity associated metagenes were not related to obesity, but associated with a different clinical variable that may be closely related to \gls{bmi}.
@@
 
-Both the obesity associated metagenes generated from Creighton's data and FM's data were able to capture the overall pattern of gene expression of the genetic signature in the samples, but did not associate with sample \gls{bmi}/\gls{bmi} status.
-One possible reason for this result could be that the obesity associated genetic signatures from Creighton's and FM's study may not have been truly associated with sample \gls{bmi} status, but with another clinical variable.
-To reject this possibility, obesity associated genetic signatures were identified in Creighton's data after controlling for all the clinical variables in the data set.
-FM's data set was not used to get obesity associated genetic signatures, as no \gls{bmi} information was available for the samples in FM's data.
+Both the obesity associated metagenes generated from CR data and FM data were able to capture the overall pattern of gene expression of the genetic signature in the samples, but did not associate with sample \gls{bmi}/\gls{bmi} status.
+One possible reason for this result could be that the obesity associated genetic signatures from the \citet{Creighton2012} and \citet{Fuente-Matter2014} studies may not have been truly associated with sample \gls{bmi}  status, but with another clinical variable.
+To reject this possibility, obesity associated genetic signatures were identified in CR data after controlling for all the clinical variables in the data set.
+FM data set was not used to get obesity associated genetic signatures, as no \gls{bmi} information was available for the samples in FM data.
 
-Firstly, to validate whether the identified obesity associated genetic signature was similar to the signature that Creighton's groups had originally found, the identified \glspl{deg} were compared with the set of genes identified by Creighton's group.
-\glspl{deg}  were identified between the samples from obese patient and non-obese patient in the \gls{rma}-normalised Creighton's data, as described in \cref{sec:gene_expression_analysis}.
+Firstly, to validate whether the identified obesity associated genetic signature was similar to the signature that Creighton \textit{et al.} had originally found, the identified \glspl{deg} were compared with the original genetic signature identified by Creighton \textit{et al.}.
+\glspl{deg}  were identified between the samples from obese patient and non-obese patient in the \gls{rma}-normalised CR data, as described in \cref{sec:gene_expression_analysis}.
 Without adjusting the p-value with FDR, 5278 gene probes and 1781 gene probes were significant at p \textless{} 0.05 and p \textless{} 0.01, respectively.
@@
 This was to include as many gene probes as possible without losing power (?).
-Additionally, when there were more than 799 gene probes identified, only the most significant 799 gene probes were taken as the obesity associated genetic signature, as this many genes were originally identified by Creighton's group.
+Additionally, when there were more than 799 gene probes identified, only the most significant 799 gene probes were taken as the obesity associated genetic signature, as this many genes were originally identified by Creighton \textit{et al.}.
 These criteria were applied for the identification of other genetic signatures as well.
 
-The above analysis was repeated with the residual data (\gls{rma}-normalised Creighton's data that had been controlled for other clinical variables; see \cref{sub:residual_data_creation}).
+The above analysis was repeated with the residual data (\gls{rma}-normalised CR data that had been controlled for other clinical variables; see \cref{sub:residual_data_creation}).
 The clinical variables controlled were age, ethnicity, menopause status, tumour grade, hormone (\gls{er}, \gls{pr}, and \gls{her2}) statuses  and \gls{ln} status.
@@
 
-In addition to the above two obesity associated genetic signatures, two more sets of genetic signatures were identified by taking the common genes between the above two signatures with Creighton's original obesity associated genetic signatures.
-There were 239 common gene probes between Creighton's original genetic signature and the gene probes identified from the unadjusted Creighton's data, and 168 common gene probes between the original signature and gene probes from residual data in Creighton's data.
+In addition to the above two obesity associated genetic signatures, two more sets of genetic signatures were identified by taking the common genes between the above two signatures with the original obesity associated genetic signatures.
+There were 239 common gene probes between the original genetic signature and the gene probes identified from the unadjusted CR data, and 168 common gene probes between the original signature and gene probes from residual data in CR data.
 The genetic signatures identified were summarised as a Venn diagram, as shown in \cref{fig:venn1}.
@@
 	\includegraphics[width=0.5\linewidth]{results1/deg_venn1}
-	\caption[Veen diagram of the \glspl{deg} identified from Creighton's data (all samples)]{Venn diagram showing the common genes between the signatures obtained from unadjusted and clinical variable-adjusted Creighton's data and Creighton's original obesity associated genetic signature.}
+	\caption[Veen diagram of the \glspl{deg} identified from CR data (all samples)]{Venn diagram showing the common genes between the signatures obtained from unadjusted and clinical variable-adjusted CR data and the original obesity associated genetic signature.}
 	\label{fig:venn1}
@@
 Though ethnicity was controlled in the residual data, the effect of ethnicity on the data was completely removed to prevent any possibility of ethnicity influencing the analysis.
-Therefore, the effect of ethnicity was completely removed by considering only the Caucasian patients in Creighton's data, which left a total of 77 Caucasian patients in the data set.
-With this Caucasian-only data set, obesity associated genetic signatures were identified as described above; first with unadjusted data, then with the data with clinical variables adjusted (except ethnicity, as it has already been controlled for by considering Caucasian samples only), and lastly the common genes with Creighton's original signature were identified.
+Therefore, the effect of ethnicity was completely removed by considering only the Caucasian patients in CR data, which left a total of 77 Caucasian patients in the data set.
+With this Caucasian-only data set, obesity associated genetic signatures were identified as described above; first with unadjusted data, then with the data with clinical variables adjusted (except ethnicity, as it has already been controlled for by considering Caucasian samples only), and lastly the common genes with the original signature were identified.
 
@@
 As before, the most significant 799 gene probes were selected from these gene probes.
-There were 148 and 92 common gene probes with Creighton's original genetic signatures and unadjusted or clinical variable-adjusted Caucasian patient data set, respectively.
+There were 148 and 92 common gene probes with the original genetic signatures and unadjusted or clinical variable-adjusted Caucasian patient data set, respectively.
 Again, Venn diagram was used to summarise the genes identified from the Caucasian patient data set (\cref{fig:venn2}).
@@
 	\includegraphics[width=0.5\linewidth]{results1/deg_venn2}
-	\caption[Summary of the \glspl{deg} identified from Creighton's data (Caucasian patient data)]{Venn diagram showing the common genes between the signatures obtained from unadjusted and clinical variable-adjusted Creighton's data (Caucasian patients only) and Creighton's original obesity associated genetic signature.}
+	\caption[Summary of the \glspl{deg} identified from CR data (Caucasian patient data)]{Venn diagram showing the common genes between the signatures obtained from unadjusted and clinical variable-adjusted CR data (Caucasian patients only) and the original obesity associated genetic signature.}
 	\label{fig:venn2}
@@
 
-For simplicity, Creighton's original obesity associated signature will be referred to as ``original'' signature, new signatures from unadjusted and clinical variable-adjusted Creighton's data as ``Cr'' and ``Res'' signature respectively, and the signatures from unadjusted and clinical variable-adjusted Caucasian patient data set as ``Ca'' and ``CaRes'' signature respectively.
-All of the overlap signatures will be referred to as their respective signature, but with ``Ol'' added at the end, for example ``CaResOl'' for the overlap signature identified in clinical variable-adjusted Caucasian patient data set.
 All of these obesity associated genetic signatures (eight in total) were checked to see whether these signatures had significant association with sample \gls{bmi}/\gls{bmi} status in various data set.
+For simplicity, the abbreviations in \cref{tab:mg_abbrev} will be used to refer to the appropriate genetic signatures.
+
+\begin{table}[htpb]
+	\centering
+	\caption{Summary of the abbreviations used to refer to the different obesity associated genetic signatures}
+	\label{tab:mg_abbrev}
+	\begin{tabular}{lp{0.51\textwidth}c}
+		\hline
+		\hline
+		Abbreviation & Definition & No. of gene probes\\
+		\hline
+		\rule{0pt}{2.25ex}Or      & Original obesity associated genetic signature identified by \citet{Creighton2012}                       & 799\\
+		\rule{0pt}{2.25ex}Cr      & Obesity associated genetic signature identified from unadjusted CR data                                & 799 \\
+		\rule{0pt}{2.25ex}CrOl    & Genes common between Or and Cr genetic signatures                                                       & 239\\
+		\rule{0pt}{2.25ex}Res     & Obesity associated genetic signature identified from clinical variable-adjusted CR data                & 799\\
+		\rule{0pt}{2.25ex}ResOl   & Genes common between Or and Res genetic signatures                                                      & 168\\
+		\rule{0pt}{2.25ex}Ca      & Obesity associated genetic signature identified from unadjusted Caucasian-only CR data                 & 799\\
+		\rule{0pt}{2.25ex}CaOl    & Genes common between Or and Ca genetic signatures                                                       & 148\\
+		\rule{0pt}{2.25ex}CaRes   & Obesity associated genetic signature identified from clinical variable-adjusted Caucasian-only CR data & 799\\
+		\rule{0pt}{2.25ex}CaResOl & Genes common between Or and CaRes genetic signatures                                                    & 92\\
+		\hline
+		\hline
+	\end{tabular}
+\end{table}
 
@@
 
-As with the original signature, all eight obesity associated genetic signatures were validated in Creighton's data set first, and then compared in other cancer data sets by using the transformation matrix generated in Creighton's data.
+As with the Or signature, all eight obesity associated genetic signatures were validated in CR data set first, and then compared in other cancer data sets by using the transformation matrix generated in CR data.
 % Before the metagenes were created, all of the signatures were converted into gene symbols, as described in \cref{sec:data} (\cref{tab:signature_gene_num}).
-Creighton's original data (unadjusted, all samples included) was normalised with \gls{rma} method and \gls{svd} was applied to generate the metagene for the corresponding genetic signature.
+CR data (unadjusted, all samples included) was normalised with \gls{rma} method and \gls{svd} was applied to generate the metagene for the corresponding genetic signature.
 The direction of the metagenes were first examined to make sure that all of the metagenes were in line with one another (see \cref{sub:metagene_direction}; results shown in appendix).
-The comparison of the metagenes with the sample gene expression in Creighton's data were displayed as heatmaps (\cref{fig:degmetacr}).
+The comparison of the metagenes with the sample gene expression in CR data were displayed as heatmaps (\cref{fig:degmetacr}).
 It was clear from the heatmaps that all of the metagenes reflected the overall expression of the corresponding genetic signatures.
 The association of the metagenes with sample \gls{bmi}/\gls{bmi} status was significant for all eight of the metagenes identified (\cref{fig:degmetacr}).
-These results confirmed that all of the obesity associated genetic signatures identified in \cref{sub:identification_of_obesity_associated_genetic_signatures} significantly associated with sample \gls{bmi}/\gls{bmi} status in Creighton's data, where the metagenes were derived from.
+These results confirmed that all of the obesity associated genetic signatures identified in \cref{sub:identification_of_obesity_associated_genetic_signatures} significantly associated with sample \gls{bmi}/\gls{bmi} status in CR data, where the metagenes were derived from.
 
@@
 	\includegraphics[page=5,width=0.45\linewidth]{results1/cr_deg_meta_vs_clin}
-	\caption[Cr obesity associated metagene in Creighton's data]{Heatmap, box plot and scatter plot showing the association of Cr obesity associated metagene with sample gene expression, \gls{bmi} and \gls{bmi} status from Creighton's data, respectively.
+	\caption[Cr obesity associated metagene in CR data]{Heatmap, box plot and scatter plot showing the association of Cr obesity associated metagene with sample gene expression, \gls{bmi} and \gls{bmi} status from CR data, respectively.
 	The results for other metagenes are in \cref{app:a}.
@@
 Additionally, the correlation of all of the obesity associated metagenes were examined to see whether these metagenes were similar to one another.
-As clearly shown in \cref{fig:cr_meta_cor}, there were two distinct groups within the eight metagenes; the first group contained the metagenes that were not overlapped with the original metagene (Cr, Res, Ca and CaRes), while the other group had the overlapped metagenes (CrOl, ResOl, CaOl and CaResOl).
+As clearly shown in \cref{fig:cr_meta_cor}, there were two distinct groups within the eight metagenes; the first group contained the metagenes that were not overlapped with the Or metagene (Cr, Res, Ca and CaRes), while the other group had the overlapped metagenes (CrOl, ResOl, CaOl and CaResOl).
 With that said, all eight metagenes showed high correlation with one another (lowest correlation approximately at 0.85), which suggested that all of these metagenes were detecting similar underlying biological mechanism from the data.
@@
 	\includegraphics[width=0.8\linewidth]{results1/cr_meta_cor}
-	\caption[Pearson correlation of all eight obesity associated metagenes identified in Creighton's data]{Heatmap showing the Pearson correlation of all eight obesity associated metagenes from Creighton's data with one another.
-	\gls{svd} was applied to \gls{rma}-normalised Creighton's data to generate each of the eight metagenes.
+	\caption[Pearson correlation of all eight obesity associated metagenes identified in CR data]{Heatmap showing the Pearson correlation of all eight obesity associated metagenes from CR data with one another.
+	\gls{svd} was applied to \gls{rma}-normalised CR data to generate each of the eight metagenes.
 	High and low correlation were represented as red and blue, respectively, where the colours were matched with the values on the scale shown in top right.}
@@
 
-To confirm whether these metagenes showed significant association with sample \gls{bmi}/\gls{bmi} status in other cancer data sets, transformation matrix was applied to the \gls{icgc} cancer data and Print's breast cancer data.
+To confirm whether these metagenes showed significant association with sample \gls{bmi}/\gls{bmi} status in other cancer data sets, transformation matrix was applied to the \gls{icgc} cancer data and \gls{nzbc} data.
 Firstly in \gls{icgc} cancer data sets, all eight metagene scores were reflective of the sample gene expression of the genetic signatures (\cref{fig:degmetaicgc}).
@@
 
-This was unexpected since all of the genetic signatures were results of gene expression analysis between the non-obese group and the obese group of samples in Creighton's data, and yet the metagenes showed significant association with the overweight group in \gls{blca} data set.
-These results suggested that the samples that were overweight in \gls{blca} data were similar in phenotype with the samples that were obese in Creighton's data.
+This was unexpected since all of the genetic signatures were results of gene expression analysis between the non-obese group and the obese group of samples in CR data, and yet the metagenes showed significant association with the overweight group in \gls{blca} data set.
+These results suggested that the samples that were overweight in \gls{blca} data were similar in phenotype with the samples that were obese in CR data.
 Again, due to the fact that all of the metagenes lacked association with sample \gls{bmi}/\gls{bmi} status in all other \gls{icgc} data set, it was difficult to conclude whether the observed association of many of the obesity metagenes with the overweight group of samples was truly reflective of the effect resulted from these metagenes.
@@
 
-In Print's data, the higher the metagenes scores were the more expressed the genes were (and vice versa), but lacked the association with sample \gls{bmi}/\gls{bmi} status (\cref{fig:degmetaprint}).
+In \gls{nzbc} data, the higher the metagenes scores were the more expressed the genes were (and vice versa), but lacked the association with sample \gls{bmi}/\gls{bmi} status (\cref{fig:degmetaprint}).
 These results showed that, even though all of the metagenes significantly associated with sample \gls{bmi}/\gls{bmi} status in the data the genetic signatures were derived from, the metagenes did not transfer well in other cancer data sets.
-Furthermore, these results showed that the lack of association with sample \gls{bmi}/\gls{bmi} status were not due to other clinical variables in Creighton's data set.
+Furthermore, these results showed that the lack of association with sample \gls{bmi}/\gls{bmi} status were not due to other clinical variables in CR data set.
 This raised a question of whether there was any obesity associated genetic signature that was common in multiple cancer types.
@@
 	\includegraphics[page=5,width=0.45\linewidth]{results1/cris_cr_trans_meta}
-	\caption[Cr obesity associated metagene in Print's data]{Heatmap, box plot and scatter plot showing the association of Cr obesity associated metagene with sample gene expression, \gls{bmi} and \gls{bmi} status from Print's data, respectively.
+	\caption[Cr obesity associated metagene in \gls{nzbc} data]{Heatmap, box plot and scatter plot showing the association of Cr obesity associated metagene with sample gene expression, \gls{bmi} and \gls{bmi} status from \gls{nzbc} data, respectively.
 	The results for other metagenes are in \cref{app:a}.
@@
 
-It was clear that the obesity associated genetic signatures created from Creighton's data were not consistent across different data sets.
+It was clear that the obesity associated genetic signatures created from CR data were not consistent across different data sets.
 To determine whether there was any obesity associated genetic signature that was expressed in multiple different cancer types, gene expression analysis was carried out on each of the eight \gls{icgc} cancer types and common genes were searched from the \glspl{deg} that were identified.

### 2017-03-10 thesis/results/results2.tex
Changed all the data set names and removed "'s" - tmp commit

@@
 
-In this chapter, the underlying biological mechanism of the obesity associated signatures from Creighton's data were investigated.
-In doing so, Gatza's pathway genetic signatures were utilised to determine which biological pathways the obesity associated genetic signatures were most similar to.
-First, the direction of Gatza's pathway associated genetic signatures were resolved; then the pathway associated metagenes were compared with the obesity associated metagenes; and lastly linear models were constructed based on the pathway metagenes and sample \gls{bmi}/\gls{bmi} status to predict the obesity associated metagenes.
+In this chapter, the underlying biological mechanism of the obesity associated signatures from CR data were investigated.
+In doing so, the pathway genetic signatures from the \citet{Gatza2010a} study (GT) were utilised to determine which biological pathways the obesity associated genetic signatures were most similar to.
+First, the direction of GT pathway associated genetic signatures were resolved; then the pathway associated metagenes were compared with the obesity associated metagenes; and lastly linear models were constructed based on the pathway metagenes and sample \gls{bmi}/\gls{bmi} status to predict the obesity associated metagenes.
 
@@
 However, the analyses so far have used \gls{rma}-normalisation method and ranked based on the number of samples present in the data.
-To decide which normalisation or ranking methods were suitable for the analysis, the different methods were compared in Gatza's data (see \cref{app:b}).
+To decide which normalisation or ranking methods were suitable for the analysis, the different methods were compared in GT data (see \cref{app:b}).
 Results shown in \cref{app:b} clarified that there was no significant difference in the ranking methods used.
@@
 This meant that all of the data sets had to be normalised with a single normalisation method, so the \gls{rma} normalisation method was used as this method was more reliable than the \gls{mas} method.
-Therefore, Gatza's data set was batch corrected first (\cref{sub:batch_correction}), then normalised with \gls{rma} method and the scores were ranked based on the number of samples in the data set.
+Therefore, GT data set was batch corrected first (\cref{sub:batch_correction}), then normalised with \gls{rma} method and the scores were ranked based on the number of samples in the data set.
 
@@
 	\includegraphics[page=100,width=0.45\linewidth]{results2/allmeta_rma_vs_mas}
-	\caption{Comparison of the metagenes generated in the \gls{mas} normalised Gatza \textit{et al.} data set from the application of the transformation matrices derived from either the \gls{rma} or \gls{mas} normalised data}
+	\caption{Comparison of the metagenes generated in the \gls{mas} normalised GT data set from the application of the transformation matrices derived from either the \gls{rma} or \gls{mas} normalised data}
 	\label{fig:gt_rma_vs_mas}
@@
 \noindent
-Before Gatza's pathway metagenes were compared with the obesity associated metagenes, the direction of Gatza's pathway metagenes had to be checked to make sure the metagenes were in the correct direction (\cref{sub:metagene_direction}).
-% Each pathway metagenes was generated from Gatza's data set and the direction of the metagene was checked with the expression of the representative pathway gene (\cref{tab:metagene_direction}), and the groupings of the pathways were considered as well.
+Before GT pathway metagenes were compared with the obesity associated metagenes, the direction of GT pathway metagenes had to be checked to make sure the metagenes were in the correct direction (\cref{sub:metagene_direction}).
+% Each pathway metagenes was generated from GT data set and the direction of the metagene was checked with the expression of the representative pathway gene (\cref{tab:metagene_direction}), and the groupings of the pathways were considered as well.
 The correlation of all the pathway metagenes with one another were plotted as a heatmap in \cref{fig:gatza_meta_dir}.
@@
 In addition to these highly correlated groups, \gls{stat3}/\gls{tgfb}/Src/\gls{egfr}/Akt pathways showed little correlation with one another.
-Comparing these groups with the results presented by \citet{Gatza2010a} (see \cref{app:b}), the identified groups approximately resembled the groups identified in their study, which confirmed that the directions of Gatza's pathway metagenes were similar to those used in the \citet{Gatza2010a} study.
+Comparing these groups with the results presented by \citet{Gatza2010a} (see \cref{app:b}), the identified groups approximately resembled the groups identified in their study, which confirmed that the directions of GT pathway metagenes were similar to those used by Gatza \textit{et al.}.
 
@@
 	\includegraphics[page=15,width=0.7\linewidth]{results2/gatza_meta_trans}
-	\caption{Heatmap of all the pathway metagenes from the \citet{Gatza2010a} study in Gatza \textit{et al.} data set}
+	\caption{Heatmap of all the pathway metagenes from the \citet{Gatza2010a} study in GT data set}
 	\label{fig:gatza_meta_dir}
@@
 
-To see whether the directionality of Gatza's pathway metagenes were being transferred across to other data sets properly, pathway metagenes were generated in other data sets with the transformation matrices and the groupings of the metagenes were examined.
-Creighton's, FM's and Cris' data were \gls{rma}-normalised and transformation matrices of the pathway genetic signatures (derived from Gatza's data set) were applied to the data sets.
+To see whether the directionality of GT pathway metagenes were being transferred across to other data sets properly, pathway metagenes were generated in other data sets with the transformation matrices and the groupings of the metagenes were examined.
+CR, FM and \gls{nzbc} data were \gls{rma}-normalised and transformation matrices of the pathway genetic signatures (derived from GT data set) were applied to the data sets.
 The metagenes were plotted in a heatmap (shown in \cref{app:b}), which showed similar groupings  as seen in \cref{fig:gatza_meta_dir}.
@@
 
-% In order to investigate what the underlying biological mechanism of the obesity associated genetic signatures was, Gatza's pathway associated genetic signatures were used to create linear models to predict the obesity associated metagene scores.
-% Before linear models were created, Gatza's pathway metagenes were checked for their directionality to ensure that the pathway metagenes were acting as reported by \citet{Gatza2010a} (see \cref{sub:metagene_direction}).
+% In order to investigate what the underlying biological mechanism of the obesity associated genetic signatures was, GT pathway associated genetic signatures were used to create linear models to predict the obesity associated metagene scores.
+% Before linear models were created, GT pathway metagenes were checked for their directionality to ensure that the pathway metagenes were acting as reported by \citet{Gatza2010a} (see \cref{sub:metagene_direction}).
 
@@
 Now that the directions of the pathway metagenes were established, these metagenes were ready for comparison with the obesity associated metagenes.
-However, all of the pathway metagenes were derived from Gatza's data, whereas the majority of the obesity associated metagenes were derived from Creighton's data, which presents a problem of deciding on which data set the transformation matrices should be generated in.
+However, all of the pathway metagenes were derived from GT data, whereas the majority of the obesity associated metagenes were derived from CR data, which presents a problem of deciding on which data set the transformation matrices should be generated in.
 
 In theory, the metagenes generated from the application of \gls{svd} and the metagenes generated from transformation matrix would be exactly the same if the transformation matrix was derived from the same data set.
-For example, an obesity associated metagene generated from Creighton's data with \gls{svd} would have the same values as the metagenes generated in Creighton's data with the transformation matrix (where the transformation matrix was derived from Creighton's data).
+For example, an obesity associated metagene generated from CR data with \gls{svd} would have the same values as the metagenes generated in CR data with the transformation matrix (where the transformation matrix was derived from CR data).
 Furthermore, if the \gls{svd}-derived metagenes and transformation matrix-derived metagenes were the same (or at least similar) in a different data set, this suggests that there was no difference in the data sets, at least in terms of the expression of the genetic signatures being investigated.
@@
 As in \cref{sec:pathway_associated_genetic_signatures_from_gatza2010a_study}, all data sets were normalised with the \gls{rma} method and metagenes were ranked based on the number of samples in each data set.
-Transformation matrices for the obesity associated genetic signatures were made in Creighton's data set and pathway associated genetic signatures were made in Gatza's data set.
+Transformation matrices for the obesity associated genetic signatures were made in CR data set and pathway associated genetic signatures were made in GT data set.
 Metagenes for all of the obesity and pathway associated genetic signatures were generated in all of the data sets with \gls{svd} and transformation matrices.
@@
 	\begin{threeparttable}
-	\caption{Summary of the Spearman correlations of the pathway metagenes generated from \gls{svd} and the application of the transformation matrices\tnote{1} in Gatza \textit{et al.}, Creighton \textit{et al.}, Print \textit{et al.} and Fuentes-Mattei \textit{et al.} data sets}
+	\caption{Summary of the Spearman correlations of the pathway metagenes generated from \gls{svd} and the application of the transformation matrices\tnote{1} in GT, CR, \gls{nzbc} and FM data sets}
 	\label{tab:svd_vs_tm_path}
 		\begin{tabular}{lcccc}
-			& Gatza & Creighton & Cris & Fuentes-Mattei\\
+			& GT & CR & \gls{nzbc} & FM\\
 			\hline
@@
 			\begin{footnotesize}
-				\item [1] Transformation matrices were derived from the Gatza \textit{et al.} data set.
+				\item [1] Transformation matrices were derived from the GT data set.
 			\end{footnotesize}
@@
 	\begin{threeparttable}
-		\caption{Summary of the Spearman correlations of the obesity metagenes generated from \gls{svd} and the application of the transformation matrices\tnote{1} in Gatza \textit{et al.}, Creighton \textit{et al.}, Print \textit{et al.} and Fuentes-Mattei \textit{et al.} data sets}
+		\caption{Summary of the Spearman correlations of the obesity metagenes generated from \gls{svd} and the application of the transformation matrices\tnote{1} in GT, CR, \gls{nzbc} and FM data sets}
 		\label{tab:svd_vs_tm_obs}
 		\begin{tabular}{lcccc}
-			& Creighton & Gatza  & Fuentes-Mattei & Cris\\
+			& CR & GT & FM & \gls{nzbc}\\
 			\hline
@@
 				\begin{footnotesize}
-				\item [1] Transformation matrices were derived from the Creighton \textit{et al.} data set.
+				\item [1] Transformation matrices were derived from the CR data set.
 				\end{footnotesize}
@@
 
-In \cref{tab:svd_vs_tm_path,tab:svd_vs_tm_obs}, the obesity and pathway associated genetic signatures had a correlatin of 1 in Creighton's and Gatza's data sets respectively, as the transformation matrices were generated in those data sets.
+In \cref{tab:svd_vs_tm_path,tab:svd_vs_tm_obs}, the obesity and pathway associated genetic signatures had a correlatin of 1 in CRs and GT data sets respectively, as the transformation matrices were generated in those data sets.
 The correlations of the obesity and pathway metagenes were variable across different data sets, which was as expected since all the data sets were different from one another.
@@
 The fact that some pathway associated metagenes were consistent across different data sets suggested that some of these genetic signatures were reliable and did not depend on the data set the transformation matrices were derived from.
-On the other hand, the genetic signatures that were not consistent across the data sets were likely to be dependent on the data set in which they were derived from, and therefore the transformation matrices for these signatures must be derived from Gatza's data set.
+On the other hand, the genetic signatures that were not consistent across the data sets were likely to be dependent on the data set in which they were derived from, and therefore the transformation matrices for these signatures must be derived from GT data set.
 
@@
 This suggested that obesity associated genetic signatures were consistent across all data sets and the transformation matrices for these signatures could be made in any of the data sets.
-Taken together, transformation matrices for all the genetic  signatures were decided to be made in the Gatza's data set, since there were some pathway associated signatures that were specific to Gatza's data, but none of the obesity associated genetic signatures were specific to a data set.
+Taken together, transformation matrices for all the genetic  signatures were decided to be made in the GT data set, since there were some pathway associated signatures that were specific to GT data, but none of the obesity associated genetic signatures were specific to a data set.
 \\
@@
 To visually determine which pathway associated genetic signatures were most similar to the obesity associated genetic signatures, heatmaps were created with the metagenes for all of the genetic signatures.
-The directions of the pathway associated metagenes have already been determined in \cref{sec:pathway_associated_genetic_signatures_from_gatza2010a_study}, and the directions of the obesity associated metagenes were checked in Gatza's data set as described in \cref{sub:metagene_direction}.
-All of the metagenes were created in \gls{rma}-normalised Gatza's data set with \gls{svd}, and the metagene scores were plotted in a heatmap and clustered into groups (\cref{fig:gatza_allmeta}).
+The directions of the pathway associated metagenes have already been determined in \cref{sec:pathway_associated_genetic_signatures_from_gatza2010a_study}, and the directions of the obesity associated metagenes were checked in GT data set as described in \cref{sub:metagene_direction}.
+All of the metagenes were created in \gls{rma}-normalised GT data set with \gls{svd}, and the metagene scores were plotted in a heatmap and clustered into groups (\cref{fig:gatza_allmeta}).
 
@@
 	\includegraphics[page=15,width=0.8\linewidth]{results2/all_meta_trans}
-	\caption{Heatmap of all the pathway and obesity metagenes in Gatza \textit{et al.} data set}
+	\caption{Heatmap of all the pathway and obesity metagenes in GT data set}
 	\label{fig:gatza_allmeta}
@@
 
-From \cref{fig:gatza_allmeta}, all of the obesity associated metagenes from Creighton's data set clustered together into a group, but none of the pathway associated metagenes associated strongly with the obesity metagenes.
+From \cref{fig:gatza_allmeta}, all of the obesity associated metagenes from CR data set clustered together into a group, but none of the pathway associated metagenes associated strongly with the obesity metagenes.
 Though \gls{her2} and p63 pathway metagenes were grouped next to the cluster of obesity metagenes, the similarity was at most about 0.6.
-Fuentes-Mattei's obesity associated metagene did not cluster together with the obesity metagenes from Creighton's data set, nor with any of the other pathway metagenes.
-The fact that Fuentes-Mattei's metagene did not cluster with Creighton's metagenes showed that the nature of Fuentes-Mattei's metagene was different from those metagenes found in Creighton's data set.
-Furthermore, their metagene did not cluster with Gatza's Akt pathway in the heatmap, even though \citet{Fuentes-Mattei2014} provided clear evidence that their metagene specifically activated the Akt/\gls{mtor} signalling pathway in a mouse model.
+FM obesity associated metagene did not cluster together with the obesity metagenes from CR data set, nor with any of the other pathway metagenes.
+The fact that FM metagene did not cluster with any of the CR metagenes showed that the nature of FM metagene was different from those metagenes found in CR data set.
+Furthermore, their metagene did not cluster with GT Akt pathway in the heatmap, even though \citet{Fuentes-Mattei2014} provided clear evidence that their metagene specifically activated the Akt/\gls{mtor} signalling pathway in a mouse model.
 This may have been due to the lack of consistency of the Akt pathway signature in different data sets (\cref{tab:svd_vs_tm_path}).
-Another likely reason could be that Fuentes-Mattei's genetic signature was not comprised only of the genes from the Akt pathway, but genes from other pathways as well, and so the metagene did not cluster together with the Akt pathway signature in the heatmap.
+Another likely reason could be that FM genetic signature was not comprised only of the genes from the Akt pathway, but genes from other pathways as well, and so the metagene did not cluster together with the Akt pathway signature in the heatmap.
 
@@
 If any of the pathway metagenes were significant in the linear model, this provided evidence that the significant pathways in the linear model were related to the obesity metagene.
-Since most of the obesity metagenes were found in Creighton's data set and only Creighton's and Cris' data sets had \gls{bmi}/\gls{bmi} status information for the samples, Cris' data set was used as the training data set to create linear models for obesity metagene score prediction.
+Since most of the obesity metagenes were found in CR data set and only CR and \gls{nzbc} data sets had \gls{bmi}/\gls{bmi} status information for the samples, \gls{nzbc} data set was used as the training data set to create linear models for obesity metagene score prediction.
 
-Metagene scores used for the construction of linear models were generated from the application of the transformation matrices derived from Gatza's data (from \cref{sec:pathway_associated_metagenes_and_obesity_associated_metagenes}) to \gls{rma}-normalised Cris' data.
+Metagene scores used for the construction of linear models were generated from the application of the transformation matrices derived from GT data (from \cref{sec:pathway_associated_metagenes_and_obesity_associated_metagenes}) to \gls{rma}-normalised \gls{nzbc} data.
 For each of the obesity metagenes identified, seven linear models were created: \gls{bmi}-only, \gls{bmi} status-only, \gls{bmi} and \gls{bmi} status, pathway metagenes-only, and all of the combinations of pathway metagenes with \gls{bmi} and/or \gls{bmi} status.
-The only variables that showed significance in any of the models that predicted the obesity metagenes taken from the Creighton \textit{et al.} data set were \gls{pr} metagene and whether samples were obese or not (\cref{tab:lm_sig_var}).
+The only variables that showed significance in any of the models that predicted the obesity metagenes taken from the CR data set were \gls{pr} metagene and whether samples were obese or not (\cref{tab:lm_sig_var}).
 In the model that predicted the FM metagene scores, the obesity status of the samples and Myc metagene score were significant, but the \gls{pr} metagene score was not (\cref{app:b}).
@@
 	\centering
-	\caption{Description of the linear models used to predict the CR obesity metagene in Print \textit{et al.} data set}
+	\caption{Description of the linear models used to predict the CR obesity metagene in \gls{nzbc} data set}
 	\label{tab:lm_sig_var}
@@
 
-All of the models constructed from sample \gls{bmi}, \gls{bmi} status and/or Gatza's pathway metagene scores confirmed that none of the pathways were significant, except for the \gls{pr} and Myc pathway metagene scores in the models for Creighton metagenes and FM metagene, respectively.
-It was not surprising that the \gls{pr} pathway metagene showed significance in the models generated in Cris' data set, as \gls{pr} status was very strongly correlated with Cris' data set.
+All of the models constructed from sample \gls{bmi}, \gls{bmi} status and/or GT pathway metagene scores confirmed that none of the pathways were significant, except for the \gls{pr} and Myc pathway metagene scores in the models for CR metagenes and FM metagene, respectively.
+It was not surprising that the \gls{pr} pathway metagene showed significance in the models generated in \gls{nzbc} data set, as \gls{pr} status was very strongly correlated with \gls{nzbc} data set.
 The significance of the Myc metagene in the model for FM metagene was difficult to interpret, as there was no reported association of FM metagene with Myc pathway by \citet{Fuentes-Mattei2014}.
@@
 	\centering
-	\caption{Description of the linear models used to predict the CR obesity metagene in Print \textit{et al.} data set, using only the sample \gls{bmi}, \gls{bmi} status and \gls{pr} pathway metagene scores}
+	\caption{Description of the linear models used to predict the CR obesity metagene in \gls{nzbc} data set, using only the sample \gls{bmi}, \gls{bmi} status and \gls{pr} pathway metagene scores}
 	\label{tab:lm_pr_only}
@@
 To determine whether any of the linear models were able to predict the corresponding obesity metagene scores, the predicted metagene scores were generated from the models and then compared with the original scores.
-The predictions were first made in Cris' data, where the metagene scores generated from the application of transformation matrices (from Gatza's data set) were used to make the predictions.
+The predictions were first made in \gls{nzbc} data, where the metagene scores generated from the application of transformation matrices (from GT data set) were used to make the predictions.
 The predicted metagenes were then compared with the true metagenes that were originally generated from the transformation matrices (\cref{fig:predict_cris_cris}).
@@
 	\includegraphics[page=73,width=0.32\linewidth]{results2/prediction_cris_with_cris(cris_model)}
-	\caption{Comparison of the predicted Cr metagenes with the true Cr metagene in the Print \textit{et al.} data}
+	\caption{Comparison of the predicted Cr metagenes with the true Cr metagene in the \gls{nzbc} data}
 	\label{fig:predict_cris_cris}
@@
 
-To confirm the results shown in Cris' data set (training set), the linear models were applied to Creighton's data (test set).
+To confirm the results shown in \gls{nzbc} data set (training set), the linear models were applied to CR data (test set).
 As shown in \cref{fig:predict_cr_cris}, all of the predicted metagenes generated from the linear models were statistically significantly associated with the true metagenes, although there were some exceptions in the models for other obesity metagenes (see \cref{app:b}).
-Similar to the results from Cris' data set, all of the $R^2$ values were low ($R^2$ \textless{} 0.43) and the data points were widely spread out in the plots.
+Similar to the results from \gls{nzbc} data set, all of the $R^2$ values were low ($R^2$ \textless{} 0.43) and the data points were widely spread out in the plots.
 Again, these results suggested that the predicted metagene scores from the linear models may be statistically significant, but may not be relevant to the obesity metagenes which the models predict, as the predicted values were only weakly associated with the true values.
@@
 	\includegraphics[page=73,width=0.32\linewidth]{results2/prediction_cr_with_cris}
-	\caption{Comparison of the predicted Cr metagenes with the true Cr metagene in the Creighton \textit{et al.} data}
+	\caption{Comparison of the predicted Cr metagenes with the true Cr metagene in the CR data}
 	\label{fig:predict_cr_cris}
@@
 
-These models were used to predict the obesity metagenes in Cris' data to test the performance of the models.
-All of the models were able to make statistically significant prediction of the obesity metagenes in Cris' data set (\cref{fig:stepwise_cris}).
+These models were used to predict the obesity metagenes in \gls{nzbc} data to test the performance of the models.
+All of the models were able to make statistically significant prediction of the obesity metagenes in \gls{nzbc} data set (\cref{fig:stepwise_cris}).
 Unlike the previous models, the step wise models showed very high $R^2$ values, where most predictions had $R^2$ \textgreater{} 0.85.
-The statistical significance and the strong association of the predicted metagenes were observed in Creighton's data set as well (\cref{fig:stepwise_cr}).
-Though the $R^2$ values were slightly lower than those from Cris' data set, the $R^2$ values were more promising than the non-step wise linear models.
+The statistical significance and the strong association of the predicted metagenes were observed in CR data set as well (\cref{fig:stepwise_cr}).
+Though the $R^2$ values were slightly lower than those from \gls{nzbc} data set, the $R^2$ values were more promising than the non-step wise linear models.
 
@@
 	\includegraphics[page=10,width=0.32\linewidth]{results2/stepwise_cris_cris}
-	\caption{Comparison of the predicted obesity metagenes with the true obesity metagene in Print \textit{et al.} data}
+	\caption{Comparison of the predicted obesity metagenes with the true obesity metagene in \gls{nzbc} data}
 	\label{fig:stepwise_cris}
@@
 	\includegraphics[page=10,width=0.32\linewidth]{results2/stepwise_cr_cris}
-	\caption{Comparison of the predicted obesity metagenes with the true obesity metagene in Creighton \textit{et al.} data}
+	\caption{Comparison of the predicted obesity metagenes with the true obesity metagene in CR data}
 	\label{fig:stepwise_cr}
@@
 
-Taken together, these results suggested that the variables included in these step wise models were associated with the obesity metagenes identified in Creighton's and FM's data sets.
+Taken together, these results suggested that the variables included in these step wise models were associated with the obesity metagenes identified in CR and FM data sets.
 However, there were no single common pathway that was present in all of the models; although Akt, \gls{egfr} and \gls{pr} pathway metagenes were frequently included in many of the models.

### 2017-03-11 thesis/misc/appendix.tex
Finished draft appendix A - tmp commit

@@
 
-	\section{Metagenes created from raw data and standardised data in Creighton \textit{et al.} data set}
+	\section{Comparison of the Creighton \textit{et al.} obesity metagene in standardised or non-standardised CR data}
 	\label{sec:metagenes_created_from_raw_data_vs_standardised_data}
@@
 		\includegraphics[width=0.45\linewidth,page=2]{appendix/check_raw_vs_std_mg}
-		\caption{Scatter plots showing the raw and ranked obesity metagene scores generated from raw or standardised Creighton \textit{et al.} data set. }
+		\caption{Scatter plots showing the raw and ranked Creighton \textit{et al.} obesity metagene values generated from standardised or non-standardised CR data set. }
 		\label{fig:appendix/check_raw_vs_std}
@@
 
-	\section{Metagenes created from raw data and standardised data (ICGC)}
-	\label{sec:metagenes_created_from_raw_data_vs_standardised_data_icgc}
+	\section{Comparison of the Creighton \textit{et al.} obesity metagene in standardised or non-standardised \gls{icgc} data}
+	\label{sec:comp_cr_raw_std_icgc}
 
-	% \begin{figure}[h]
-	% 	\centering
-	% 	\includegraphics[width=0.45\linewidth,page=1]{appendix/crtcga_raw_vs_std}
-	% 	\hfill
-	% 	\includegraphics[width=0.45\linewidth,page=2]{appendix/crtcga_raw_vs_std}
-	% 	\caption{appendix/Check_raw_vs_std}
-	% 	\label{fig:appendix/check_raw_vs_std}
-	% \end{figure}
+	\begin{figure}[h]
+		\centering
+		\includegraphics[width=0.45\linewidth,page=1]{appendix/crtcga_raw_vs_std}
+		\hfill
+		\includegraphics[width=0.45\linewidth,page=2]{appendix/crtcga_raw_vs_std}\\
+		\includegraphics[width=0.45\linewidth,page=5]{appendix/crtcga_raw_vs_std}
+		\hfill
+		\includegraphics[width=0.45\linewidth,page=6]{appendix/crtcga_raw_vs_std}\\
+		\caption{Scatter plots comparing the Creighton \textit{et al.} obesity metagene generated from transformation matrix (from standardised or non-standardised CR data) in non-standardised (left) and standardised (right) \gls{icgc} data.}
+		\label{fig:appendix/check_raw_vs_std}
+	\end{figure}
 
-	\section{Rest of the ICGC cancer heatmap resutls}
-	\label{sec:rest_of_the_icgc_cancer_heatmap_resutls}
+	\begin{figure}[htpb]
+		\ContinuedFloat
+		\captionsetup{list=off,format=cont}
+		\centering
+		\includegraphics[width=0.45\linewidth,page=9]{appendix/crtcga_raw_vs_std}
+		\hfill
+		\includegraphics[width=0.45\linewidth,page=10]{appendix/crtcga_raw_vs_std}\\
+		\includegraphics[width=0.45\linewidth,page=13]{appendix/crtcga_raw_vs_std}
+		\hfill
+		\includegraphics[width=0.45\linewidth,page=14]{appendix/crtcga_raw_vs_std}\\
+		\includegraphics[width=0.45\linewidth,page=17]{appendix/crtcga_raw_vs_std}
+		\hfill
+		\includegraphics[width=0.45\linewidth,page=18]{appendix/crtcga_raw_vs_std}\\
+		\caption{Figure continued}
+	\end{figure}
+
+	\begin{figure}[htpb]
+		\ContinuedFloat
+		\captionsetup{list=off,format=cont}
+		\centering
+		\includegraphics[width=0.45\linewidth,page=21]{appendix/crtcga_raw_vs_std}
+		\hfill
+		\includegraphics[width=0.45\linewidth,page=22]{appendix/crtcga_raw_vs_std}\\
+		\includegraphics[width=0.45\linewidth,page=25]{appendix/crtcga_raw_vs_std}
+		\hfill
+		\includegraphics[width=0.45\linewidth,page=26]{appendix/crtcga_raw_vs_std}\\
+		\includegraphics[width=0.45\linewidth,page=29]{appendix/crtcga_raw_vs_std}
+		\hfill
+		\includegraphics[width=0.45\linewidth,page=30]{appendix/crtcga_raw_vs_std}\\
+		\caption{Figure continued}
+	\end{figure}
+
+	\noindent
+	As shown in \cref{fig:appendix/check_raw_vs_std}, the Creighton \textit{et al.} obesity metagene values generated from the transformation matrices (from standardised or non-standardised CR data) were more consistent (or less variable) in the standardised \gls{icgc} data, compared with the non-standardised data.
+	Results from \cref{fig:appendix/check_raw_vs_std_tm} suggested that there was no obvious difference in the quality of the metagenes generated from standardised or non-standardised TM.
+
+	\begin{figure}[h]
+		\centering
+		\includegraphics[width=0.45\linewidth,page=3]{appendix/crtcga_raw_vs_std}
+		\hfill
+		\includegraphics[width=0.45\linewidth,page=4]{appendix/crtcga_raw_vs_std}\\
+		\includegraphics[width=0.45\linewidth,page=7]{appendix/crtcga_raw_vs_std}
+		\hfill
+		\includegraphics[width=0.45\linewidth,page=8]{appendix/crtcga_raw_vs_std}\\
+		\caption{Scatter plots comparing the Creighton \textit{et al.} obesity metagene generated from transformation matrix from non-standardisedright (left) or standardised (right) CR data in non-standardised  and standardised  \gls{icgc} data.}
+		\label{fig:appendix/check_raw_vs_std_tm}
+	\end{figure}
+
+	\begin{figure}[htpb]
+		\ContinuedFloat
+		\captionsetup{list=off,format=cont}
+		\centering
+		\includegraphics[width=0.45\linewidth,page=11]{appendix/crtcga_raw_vs_std}
+		\hfill
+		\includegraphics[width=0.45\linewidth,page=12]{appendix/crtcga_raw_vs_std}\\
+		\includegraphics[width=0.45\linewidth,page=15]{appendix/crtcga_raw_vs_std}
+		\hfill
+		\includegraphics[width=0.45\linewidth,page=16]{appendix/crtcga_raw_vs_std}\\
+		\includegraphics[width=0.45\linewidth,page=19]{appendix/crtcga_raw_vs_std}
+		\hfill
+		\includegraphics[width=0.45\linewidth,page=20]{appendix/crtcga_raw_vs_std}\\
+		\caption{Figure continued}
+	\end{figure}
+
+	\begin{figure}[htpb]
+		\ContinuedFloat
+		\captionsetup{list=off,format=cont}
+		\centering
+		\includegraphics[width=0.45\linewidth,page=23]{appendix/crtcga_raw_vs_std}
+		\hfill
+		\includegraphics[width=0.45\linewidth,page=24]{appendix/crtcga_raw_vs_std}\\
+		\includegraphics[width=0.45\linewidth,page=27]{appendix/crtcga_raw_vs_std}
+		\hfill
+		\includegraphics[width=0.45\linewidth,page=28]{appendix/crtcga_raw_vs_std}\\
+		\includegraphics[width=0.45\linewidth,page=31]{appendix/crtcga_raw_vs_std}
+		\hfill
+		\includegraphics[width=0.45\linewidth,page=32]{appendix/crtcga_raw_vs_std}\\
+		\caption{Figure continued}
+	\end{figure}
+
+	\section{Remainder of the results of Creighton \textit{et al.} obesity metagene in \gls{icgc} cancer data sets}
+	\label{sec:rest_of_the_cr_icgc_cancer_heatmap_results}
+
+	\begin{figure}[htp!]
+		\centering
+		\includegraphics[width=0.45\linewidth,page=8]{appendix/crtcga_std}\\
+		\vspace{1em}
+		\includegraphics[width=0.45\linewidth,page=13]{appendix/crtcga_std}
+		\hfill
+		\includegraphics[width=0.45\linewidth,page=18]{appendix/crtcga_std}\\
+		\caption{Heatmaps showing the association of obesity metagene from the \citet{Creighton2012} study with sample gene expression from the other \gls{icgc} data sets.
+	Level of expression is represented in the top right histogram, where low and high gene expression were colour-coded with blue and red, respectively.
+	Each row of the heatmap represents a gene from the obesity associated genetic signature, and each column of the heatmap represents a sample from the \gls{icgc} data.
+	The obesity associated metagene scores of the samples are shown in a separate row at top of the heatmap, and the tree diagram of the heirarchical clustering of the genes is shown to the left of the heatmap.
+}
+		\label{fig:appendix/cr_icgc_heatmap}
+	\end{figure}
+
+	\begin{figure}[htpb]
+		\ContinuedFloat
+		\captionsetup{list=off,format=cont}
+		\centering
+		\includegraphics[width=0.45\linewidth,page=23]{appendix/crtcga_std}
+		\hfill
+		\includegraphics[width=0.45\linewidth,page=28]{appendix/crtcga_std}\\
+		\vspace{1em}
+		\includegraphics[width=0.45\linewidth,page=33]{appendix/crtcga_std}
+		\hfill
+		\includegraphics[width=0.45\linewidth,page=38]{appendix/crtcga_std}\\
+		\vspace{1em}
+		\caption{figure}
+	\end{figure}
+
+	\noindent
+	As shown in \cref{fig:appendix/cr_icgc_heatmap}, the obesity metagene from the Creighton \textit{et al.} study was reflective of the sample gene expression levels of the obesity associated genetic signature in all of the other \gls{icgc} cancer data sets.
+	However, none of the \gls{icgc} cancer data sets (other than the \gls{blca} data set; \cref{fig:crmetaicgc} in \cref{sec:creighton_obesity_metagene}) showed any significant association of the sample \gls{bmi} or \gls{bmi} status with the Creighton \textit{et al.} obesity metagene (\cref{fig:appendix/cr_icgc_box_scatter}).
 
@@
 		\centering
-		\includegraphics[width=0.45\linewidth,page=1]{appendix/crtcga_std}
+		\includegraphics[page=9,width=0.45\linewidth]{appendix/crtcga_std}
 		\hfill
-		\includegraphics[width=0.45\linewidth,page=2]{appendix/crtcga_std}
-		\caption{appendix/Check_raw_vs_std}
-		\label{fig:appendix/check_raw_vs_std}
+		\includegraphics[page=10,width=0.45\linewidth]{appendix/crtcga_std}\\
+		\includegraphics[page=14,width=0.45\linewidth]{appendix/crtcga_std}
+		\hfill
+		\includegraphics[page=15,width=0.45\linewidth]{appendix/crtcga_std}\\
+		\caption{Box plots and scatter plots showing the association of obesity metagene from the \citet{Creighton2012} study with the sample \gls{bmi} status  and \gls{bmi} from the other \gls{icgc} data sets, respectively.
+	In the box plot, the p-values above the groups represent the statistical significance of the association of the metagene with the overweight or obese group compared with the normal weight group.
+	The \gls{anova} p-value shows the statistical significance of the association of the metagene with the sample \gls{bmi} groups.
+	In the scatter plot, $R^2$- and p-values describe the adjusted coefficient of determination of the regression line and the statistical significance of the linear model used to draw the regression line, respectively.}
+		\label{fig:appendix/cr_icgc_box_scatter}
 	\end{figure}
 
-	\section{FM metagene, raw vs std}
+	\begin{figure}[htpb]
+		\ContinuedFloat
+		\captionsetup{list=off,format=cont}
+		\centering
+		\includegraphics[page=19,width=0.45\linewidth]{appendix/crtcga_std}
+		\hfill
+		\includegraphics[page=20,width=0.45\linewidth]{appendix/crtcga_std}\\
+		\includegraphics[page=24,width=0.45\linewidth]{appendix/crtcga_std}
+		\hfill
+		\includegraphics[page=25,width=0.45\linewidth]{appendix/crtcga_std}\\
+		\includegraphics[page=29,width=0.45\linewidth]{appendix/crtcga_std}
+		\hfill
+		\includegraphics[page=30,width=0.45\linewidth]{appendix/crtcga_std}\\
+		\caption{Name}
+	\end{figure}
+
+	\begin{figure}[htpb]
+		\ContinuedFloat
+		\captionsetup{list=off,format=cont}
+		\centering
+		\includegraphics[page=34,width=0.45\linewidth]{appendix/crtcga_std}
+		\hfill
+		\includegraphics[page=35,width=0.45\linewidth]{appendix/crtcga_std}\\
+		\includegraphics[page=39,width=0.45\linewidth]{appendix/crtcga_std}
+		\hfill
+		\includegraphics[page=40,width=0.45\linewidth]{appendix/crtcga_std}\\
+		\caption{Name}
+	\end{figure}
+
+	\section{Comparison of the FM obesity metagene in standardised or non-standardised FM data}
 	\label{sec:fm_metagene_raw_vs_std}
+
+	\noindent
+	\cref{fig:appendix/fm_meta_check} suggested that the FM obesity associated genetic signature was noisy, unlike the Creighton \textit{et al.} signature (\cref{fig:appendix/check_raw_vs_std}).
+	The heatmaps in \cref{fig:appendix/fm_meta_heat} showed further evidence of the FM genetic signature being noisy and of poor quality, as the FM metagene did not reflect the sample gene expression patterns as well as the obesity associated genetic signature from the Creighton \textit{et al.} study.
 
@@
 		\centering
-		\includegraphics[page=1,width=0.8\linewidth]{appendix/fm_meta_check}
-		\caption{appendix/fm_meta_check}
+		\includegraphics[page=1,width=0.45\linewidth]{appendix/fm_meta_check}
+		\hfill
+		\includegraphics[page=2,width=0.45\linewidth]{appendix/fm_meta_check}\\
+		\caption{Scatter plots showing the raw and ranked FM obesity metagene values generated from standardised or non-standardised FM data set. }
 		\label{fig:appendix/fm_meta_check}
+	\end{figure}
+
+	\begin{figure}[htpb]
+		\centering
+		\includegraphics[page=5,width=0.45\linewidth]{appendix/fm_meta_check}
+		\hfill
+		\includegraphics[page=8,width=0.45\linewidth]{appendix/fm_meta_check}\\
+		\caption{Heatmaps showing the association of obesity metagene from the Fuentes-Mattei \textit{et al.} study with sample gene expression from the FM data set.
+		Scales are as described in previous figures.}
+		\label{fig:appendix/fm_meta_heat}
+	\end{figure}
+
+	\section{Remainder of the results of FM obesity metagene in \gls{icgc} cancer data sets}
+	\label{sec:rest_of_the_fm_icgc_cancer_heatmap_results}
+
+	\noindent
+	The FM obesity metagene from was reflective of the sample gene expression levels of the obesity associated genetic signature in all of the other \gls{icgc} cancer data sets (\cref{fig:appendix/fm_icgc_heatmap}).
+	However, none of the \gls{icgc} cancer data sets (other than the \gls{blca} data set; \cref{fig:fmmetaicgc} in \cref{sec:fm_obesity_metagene}) showed any significant association of the sample \gls{bmi} or \gls{bmi} status with the FM obesity metagene (\cref{fig:appendix/fm_icgc_box_scatter}).
+
+	\begin{figure}[htp!]
+		\centering
+		\includegraphics[width=0.45\linewidth,page=8]{appendix/fm_meta_ICGC}
+		\vspace{1em}
+		\includegraphics[width=0.45\linewidth,page=13]{appendix/fm_meta_ICGC}\\
+		\hfill
+		\caption{Heatmaps showing the association of obesity metagene from the \citet{Creighton2012} study with sample gene expression from the other \gls{icgc} data sets.
+	Scales are as described in previous figures.}
+		\label{fig:appendix/fm_icgc_heatmap}
+	\end{figure}
+
+	\begin{figure}[htpb]
+		\ContinuedFloat
+		\captionsetup{list=off,format=cont}
+		\centering
+		\includegraphics[width=0.45\linewidth,page=18]{appendix/fm_meta_ICGC}\\
+		\vspace{1em}
+		\includegraphics[width=0.45\linewidth,page=23]{appendix/fm_meta_ICGC}
+		\hfill
+		\includegraphics[width=0.45\linewidth,page=28]{appendix/fm_meta_ICGC}\\
+		\vspace{1em}
+		\includegraphics[width=0.45\linewidth,page=33]{appendix/fm_meta_ICGC}
+		\hfill
+		\includegraphics[width=0.45\linewidth,page=38]{appendix/fm_meta_ICGC}\\
+		\vspace{1em}
+		\caption{figure}
+	\end{figure}
+
+	\begin{figure}[htpb]
+		\centering
+		\includegraphics[page=9,width=0.45\linewidth]{appendix/fm_meta_icgc}
+		\hfill
+		\includegraphics[page=10,width=0.45\linewidth]{appendix/fm_meta_icgc}\\
+		\includegraphics[page=14,width=0.45\linewidth]{appendix/fm_meta_icgc}
+		\hfill
+		\includegraphics[page=15,width=0.45\linewidth]{appendix/fm_meta_icgc}\\
+		\caption{Box plots and scatter plots showing the association of obesity metagene from the \citet{Creighton2012} study with the sample \gls{bmi} status  and \gls{bmi} from the other \gls{icgc} data sets, respectively.
+	P-values and $R^2$-value are as described in previous figures.}
+		\label{fig:appendix/fm_icgc_box_scatter}
+	\end{figure}
+
+	\begin{figure}[htpb]
+		\ContinuedFloat
+		\captionsetup{list=off,format=cont}
+		\centering
+		\includegraphics[page=19,width=0.45\linewidth]{appendix/fm_meta_icgc}
+		\hfill
+		\includegraphics[page=20,width=0.45\linewidth]{appendix/fm_meta_icgc}\\
+		\includegraphics[page=24,width=0.45\linewidth]{appendix/fm_meta_icgc}
+		\hfill
+		\includegraphics[page=25,width=0.45\linewidth]{appendix/fm_meta_icgc}\\
+		\includegraphics[page=29,width=0.45\linewidth]{appendix/fm_meta_icgc}
+		\hfill
+		\includegraphics[page=30,width=0.45\linewidth]{appendix/fm_meta_icgc}\\
+		\caption{Name}
+	\end{figure}
+
+	\begin{figure}[htpb]
+		\ContinuedFloat
+		\captionsetup{list=off,format=cont}
+		\centering
+		\includegraphics[page=34,width=0.45\linewidth]{appendix/fm_meta_icgc}
+		\hfill
+		\includegraphics[page=35,width=0.45\linewidth]{appendix/fm_meta_icgc}\\
+		\includegraphics[page=39,width=0.45\linewidth]{appendix/fm_meta_icgc}
+		\hfill
+		\includegraphics[page=40,width=0.45\linewidth]{appendix/fm_meta_icgc}\\
+		\caption{Name}
 	\end{figure}
@@
 
+	To determine the correct direction of each of the novel obesity metagenes obtained from the CR data, heatmaps were created in the CR data.
+	Genes that were common to all eight obesity associated genetic signatures were used in the heatmaps to make sure the genes and their expressions were the same in every heatmap.
+	The heatmaps from \cref{fig:cr_meta_direction} clearly showed that the Res, ResOl and CaRes obesity metagenes were in the opposite direction compared to the other obesity metagenes, and therefore the three obesity metagenes were flipped in the subsequent analyses.
+
 	\begin{figure}[htpb]
 		\centering
-		\includegraphics[page=1,width=0.8\linewidth]{cr_meta_direction}
-		\caption{cr_meta_direction}
+		\includegraphics[page=3,width=0.45\linewidth]{appendix/cr_meta_direction_raw}
+		\hfill
+		\includegraphics[page=6,width=0.45\linewidth]{appendix/cr_meta_direction_raw}\\
+		\caption{Heatmaps showing the association of all the novel obesity metagenes from the CR data set with the sample gene expression in CR the data set.
+		Genes that were common to all of the novel obesity associated genetic signatures were used in the heatmap.
+		Scales are as described in previous figures.}
 		\label{fig:cr_meta_direction}
@@
 
-	\section{Number of gene probes/genes in each signature}
+	\begin{figure}[htpb]
+		\ContinuedFloat
+		\captionsetup{list=off,format=cont}
+		\centering
+		\includegraphics[page=9,width=0.45\linewidth]{appendix/cr_meta_direction_raw}
+		\hfill
+		\includegraphics[page=12,width=0.45\linewidth]{appendix/cr_meta_direction_raw}\\
+		\vspace{1em}
+		\includegraphics[page=15,width=0.45\linewidth]{appendix/cr_meta_direction_raw}
+		\hfill
+		\includegraphics[page=18,width=0.45\linewidth]{appendix/cr_meta_direction_raw}\\
+		\vspace{1em}
+		\includegraphics[page=21,width=0.45\linewidth]{appendix/cr_meta_direction_raw}
+		\hfill
+		\includegraphics[page=24,width=0.45\linewidth]{appendix/cr_meta_direction_raw}\\
+		\caption{Name}
+	\end{figure}
+
+	\section{Number of gene probes/genes in each of the obesity associated genetic signature}
 	\label{app:number_of_gene_probes_genes_in_each_signature}
+
+	Since all of the genes in the obesity associated genetic signatures from the CR data were in terms of the microarray gene probes rather than gene symbols, the gene probes were converted into gene symbols.
+	Furthermore, since some gene symbols were not present in the \gls{icgc} data, these genes were omitted from the analyses.
+	Therefore, the final number of genes used in the analyses in this project were slightly less than the number of genes originally identified.
+	The final number of genes for each of the obesity associated genetic signatures were summarised in \cref{tab:signature_gene_num}.
 
@@
 		\centering
-		\caption{Summary of the number of gene probes and genes in each obesity associated genetic signature}
+		\caption{Summary of the number of gene probes and gene symbols for each of the obesity associated genetic signatures identified from the CR data}
 		\label{tab:signature_gene_num}
 		\begin{tabular}{lcc}
-			Obesity associated genetic signature & No. gene probes   & No. genes\\
+			Obesity associated genetic signature & No. gene probes   & No. gene symbols\\
 			\hline
-			\rule{0pt}{2.25ex}Original & 799 & 644\\
-			Cr                         & 799 & 678\\
-			Res                        & 799 & 655\\
-			Ca                         & 799 & 657\\
-			CaRes                      & 799 & 651\\
-			CrOl                       & 239 & 199\\
-			ResOl                      & 168 & 147\\
-			CaOl                       & 148 & 128\\
-			CaResOl                    & 92  & 86\\
+			\rule{0pt}{2.25ex}Or & 799 & 644 \\
+			Cr                   & 799 & 678 \\
+			Res                  & 799 & 655 \\
+			Ca                   & 799 & 657 \\
+			CaRes                & 799 & 651 \\
+			CrOl                 & 239 & 199 \\
+			ResOl                & 168 & 147 \\
+			CaOl                 & 148 & 128 \\
+			CaResOl              & 92  & 86  \\
 			\hline

### 2017-03-11 thesis/results/results.tex
Finished draft appendix A - tmp commit

@@
 
-Since the obesity associated genetic signature from the \citet{Creighton2012} study was not associated with sample \gls{bmi} or \gls{bmi} status in other data sets, the obesity associated genetic signature from the \citet{Fuentes-Matter2014} study (FM) was examined to see whether FM obesity metagene associated with sample \gls{bmi} or \gls{bmi} status.
+Since the obesity associated genetic signature from the \citet{Creighton2012} study was not associated with sample \gls{bmi} or \gls{bmi} status in other data sets, the obesity associated genetic signature from the \citet{Fuentes-Mattei2014} study (FM) was examined to see whether FM obesity metagene associated with sample \gls{bmi} or \gls{bmi} status.
 Since FM data set did not have sample \gls{bmi} information, FM obesity metagene was not able to be compared with the sample \gls{bmi} or \gls{bmi} status in the original FM data.
@@
 
-Firstly, FM data was normalised with the \gls{rma} method and \gls{svd} was applied to the normalised data to get the transformation matrix.
+Firstly, FM data was normalised with the \gls{rma} method and \gls{svd} was applied to the normalised FM data to get the transformation matrix.
 The transformation matrix was used to transform the \gls{rma}-normalised CR data to extract FM obesity metagene scores in CR data.

### 2017-03-11 thesis/results/results2.tex
Finished draft appendix A - tmp commit

@@
 The fact that FM metagene did not cluster with any of the CR metagenes showed that the nature of FM metagene was different from those metagenes found in CR data set.
-Furthermore, their metagene did not cluster with GT Akt pathway in the heatmap, even though \citet{Fuentes-Mattei2014} provided clear evidence that their metagene specifically activated the Akt/\gls{mtor} signalling pathway in a mouse model.
+Furthermore, their metagene did not cluster with the Akt pathway in the heatmap, even though \citet{Fuentes-Mattei2014} provided clear evidence that their metagene specifically activated the Akt/\gls{mtor} signalling pathway in a mouse model.
 This may have been due to the lack of consistency of the Akt pathway signature in different data sets (\cref{tab:svd_vs_tm_path}).

### 2017-03-11 thesis/misc/appendix.tex
Really finished the draft for appendix A - tmp commit

@@
 	\end{table}
+
+	\section{Remainder of the results of the other CR obesity metagenes in CR data set}
+	\label{sec:rest_of_the_cr_ob_meta_heatmap_results}
+
+	As already mentioned in \cref{sub:_novel_obesity_associated_signatures_and_sample_bmi}, all of the obesity associated geneti signatures derived from the CR data captured the overall gene expressions in the CR data, and significantly correlated with the sample \gls{bmi}/\gls{bmi} status (\cref{fig:appendix/cr_ob_meta_heatmap,fig:appendix/cr_ob_meta_box_scatter}).
+
+	\begin{figure}[htp!]
+		\centering
+		\includegraphics[width=0.45\linewidth,page=8]{appendix/cr_deg_meta_vs_clin}\\
+		\vspace{1em}
+		\includegraphics[width=0.45\linewidth,page=13]{appendix/cr_deg_meta_vs_clin}
+		\hfill
+		\includegraphics[width=0.45\linewidth,page=18]{appendix/cr_deg_meta_vs_clin}\\
+		\vspace{1em}
+		\includegraphics[width=0.45\linewidth,page=23]{appendix/cr_deg_meta_vs_clin}
+		\hfill
+		\includegraphics[width=0.45\linewidth,page=28]{appendix/cr_deg_meta_vs_clin}\\
+		\caption{Heatmaps showing the association of the other CR obesity metagenes with the sample gene expression from the CR data set.
+	Scales are as described in previous figures.}
+		\label{fig:appendix/cr_ob_meta_heatmap}
+	\end{figure}
+
+	\begin{figure}[htpb]
+		\ContinuedFloat
+		\captionsetup{list=off,format=cont}
+		\centering
+		\includegraphics[width=0.45\linewidth,page=33]{appendix/cr_deg_meta_vs_clin}
+		\hfill
+		\includegraphics[width=0.45\linewidth,page=38]{appendix/cr_deg_meta_vs_clin}\\
+		\vspace{1em}
+		\caption{figure}
+	\end{figure}
+
+	\begin{figure}[htpb]
+		\centering
+		\includegraphics[page=9,width=0.45\linewidth]{appendix/cr_deg_meta_vs_clin}
+		\hfill
+		\includegraphics[page=10,width=0.45\linewidth]{appendix/cr_deg_meta_vs_clin}\\
+		\caption{Box plots and scatter plots showing the association of the other CR obesity metagenes with the sample \gls{bmi} status  and \gls{bmi} from the CR data set, respectively.
+	P-values and $R^2$-value are as described in previous figures.}
+		\label{fig:appendix/cr_ob_meta_box_scatter}
+	\end{figure}
+
+	\begin{figure}[htpb]
+		\ContinuedFloat
+		\captionsetup{list=off,format=cont}
+		\centering
+		\includegraphics[page=14,width=0.45\linewidth]{appendix/cr_deg_meta_vs_clin}
+		\hfill
+		\includegraphics[page=15,width=0.45\linewidth]{appendix/cr_deg_meta_vs_clin}\\
+		\includegraphics[page=19,width=0.45\linewidth]{appendix/cr_deg_meta_vs_clin}
+		\hfill
+		\includegraphics[page=20,width=0.45\linewidth]{appendix/cr_deg_meta_vs_clin}\\
+		\includegraphics[page=24,width=0.45\linewidth]{appendix/cr_deg_meta_vs_clin}
+		\hfill
+		\includegraphics[page=25,width=0.45\linewidth]{appendix/cr_deg_meta_vs_clin}\\
+		\caption{Name}
+	\end{figure}
+
+	\begin{figure}[htpb]
+		\ContinuedFloat
+		\captionsetup{list=off,format=cont}
+		\centering
+		\includegraphics[page=29,width=0.45\linewidth]{appendix/cr_deg_meta_vs_clin}
+		\hfill
+		\includegraphics[page=30,width=0.45\linewidth]{appendix/cr_deg_meta_vs_clin}\\
+		\includegraphics[page=34,width=0.45\linewidth]{appendix/cr_deg_meta_vs_clin}
+		\hfill
+		\includegraphics[page=35,width=0.45\linewidth]{appendix/cr_deg_meta_vs_clin}\\
+		\includegraphics[page=39,width=0.45\linewidth]{appendix/cr_deg_meta_vs_clin}
+		\hfill
+		\includegraphics[page=40,width=0.45\linewidth]{appendix/cr_deg_meta_vs_clin}\\
+		\caption{Name}
+	\end{figure}
+
+	\section{Remainder of the results of the other CR obesity metagenes in \gls{icgc} data sets}
+	\label{sec:rest_of_the_cr_ob_meta_heatmap_results_icgc}
+
+	In the \gls{icgc} data sets, \gls{blca} data set was the only \gls{icgc} data set to show any significant association with the sample \gls{bmi} status.
+	All of the other \gls{icgc} cancer data sets showed no significant association with the sample \gls{bmi}/\gls{bmi} status, but the metagenes were reflective of the sample gene expressions in all of the cancer types.
+	For clarity and convenience, results will be split based on the \gls{icgc} cancer type.
+
+	\subsubsection{\gls{blca}}
+	\label{ssub:blca}
+
+	\begin{figure}[htp!]
+		\centering
+		\includegraphics[page=3,width=0.45\linewidth]{appendix/crolgene_ICGC}\\
+		\vspace{1em}
+		\includegraphics[page=3,width=0.45\linewidth]{appendix/resobsgene_ICGC}
+		\hfill
+		\includegraphics[page=3,width=0.45\linewidth]{appendix/rescrolgene_ICGC}\\
+		\caption{Heatmaps showing the association of the other CR obesity metagenes with the sample gene expression from the \gls{icgc} \gls{blca} data set.
+		Scales, p-values and $R^2$-value are as described in previous figures.}
+		\label{fig:degmetaicgc_blca}
+	\end{figure}
+
+	\begin{figure}[htpb]
+		\ContinuedFloat
+		\captionsetup{list=off,format=cont}
+		\centering
+		\includegraphics[page=3,height=0.43\linewidth,width=0.45\linewidth]{appendix/caobsgene_ICGC}
+		\hfill
+		\includegraphics[page=3,height=0.43\linewidth,width=0.45\linewidth]{appendix/cacrolgene_ICGC}\\
+		\includegraphics[page=3,height=0.43\linewidth,width=0.45\linewidth]{appendix/caresobsgene_ICGC}
+		\hfill
+		\includegraphics[page=3,height=0.43\linewidth,width=0.45\linewidth]{appendix/carescrolgene_ICGC}\\
+		\caption{Name}
+	\end{figure}
+
+	\begin{figure}[htpb]
+		\centering
+		\includegraphics[page=4,width=0.45\linewidth]{appendix/crolgene_ICGC}
+		\hfill
+		\includegraphics[page=5,width=0.45\linewidth]{appendix/crolgene_ICGC}\\
+		\caption{Box plots and scatter plots showing the association of the other CR obesity metagenes with the sample \gls{bmi} status  and \gls{bmi} from the \gls{icgc} \gls{blca} data set, respectively.
+	P-values and $R^2$-value are as described in previous figures.}
+		\label{fig:appendix/cr_ob_meta_box_scatter_blca}
+	\end{figure}
+
+	\begin{figure}[htpb]
+		\ContinuedFloat
+		\captionsetup{list=off,format=cont}
+		\centering
+		\includegraphics[page=4,width=0.45\linewidth]{appendix/resobsgene_ICGC}
+		\hfill
+		\includegraphics[page=5,width=0.45\linewidth]{appendix/resobsgene_ICGC}\\
+		\includegraphics[page=4,width=0.45\linewidth]{appendix/rescrolgene_ICGC}
+		\hfill
+		\includegraphics[page=5,width=0.45\linewidth]{appendix/rescrolgene_ICGC}\\
+		\includegraphics[page=4,width=0.45\linewidth]{appendix/caobsgene_ICGC}
+		\hfill
+		\includegraphics[page=5,width=0.45\linewidth]{appendix/caobsgene_ICGC}\\
+		\caption{Name}
+	\end{figure}
+
+	\begin{figure}[htpb]
+		\ContinuedFloat
+		\captionsetup{list=off,format=cont}
+		\centering
+		\includegraphics[page=4,width=0.45\linewidth]{appendix/cacrolgene_ICGC}
+		\hfill
+		\includegraphics[page=5,width=0.45\linewidth]{appendix/cacrolgene_ICGC}\\
+		\includegraphics[page=4,width=0.45\linewidth]{appendix/caresobsgene_ICGC}
+		\hfill
+		\includegraphics[page=5,width=0.45\linewidth]{appendix/caresobsgene_ICGC}\\
+		\includegraphics[page=4,width=0.45\linewidth]{appendix/carescrolgene_ICGC}
+		\hfill
+		\includegraphics[page=5,width=0.45\linewidth]{appendix/carescrolgene_ICGC}\\
+		\caption{Name}
+	\end{figure}
+
+	\newpage
+
+	\subsubsection{\gls{cesc}}
+	\label{ssub:cesc}
+
+	\begin{figure}[htp!]
+		\centering
+		\includegraphics[page=8,height=0.43\linewidth,width=0.45\linewidth]{appendix/rawobsgene_ICGC}
+		\hfill
+		\includegraphics[page=8,height=0.43\linewidth,width=0.45\linewidth]{appendix/crolgene_ICGC}\\
+		\includegraphics[page=8,height=0.43\linewidth,width=0.45\linewidth]{appendix/resobsgene_ICGC}
+		\hfill
+		\includegraphics[page=8,height=0.43\linewidth,width=0.45\linewidth]{appendix/rescrolgene_ICGC}\\
+		\includegraphics[page=8,height=0.43\linewidth,width=0.45\linewidth]{appendix/caobsgene_ICGC}
+		\hfill
+		\includegraphics[page=8,height=0.43\linewidth,width=0.45\linewidth]{appendix/cacrolgene_ICGC}\\
+		\caption{Heatmaps showing the association of the other CR obesity metagenes with the sample gene expression from the \gls{icgc} \gls{cesc} data set.
+		Scales, p-values and $R^2$-value are as described in previous figures.}
+		\label{fig:degmetaicgc_cesc}
+	\end{figure}
+
+	\begin{figure}[htpb]
+		\ContinuedFloat
+		\captionsetup{list=off,format=cont}
+		\centering
+		\includegraphics[page=8,height=0.42\linewidth,width=0.45\linewidth]{appendix/caresobsgene_ICGC}
+		\hfill
+		\includegraphics[page=8,height=0.42\linewidth,width=0.45\linewidth]{appendix/carescrolgene_ICGC}\\
+		\caption{Name}
+	\end{figure}
+
+	\begin{figure}[!htpb]
+		\centering
+		\includegraphics[page=9,width=0.40\linewidth]{appendix/rawobsgene_ICGC}
+		\hfill
+		\includegraphics[page=10,width=0.40\linewidth]{appendix/rawobsgene_ICGC}\\
+		\includegraphics[page=9,width=0.40\linewidth]{appendix/crolgene_ICGC}
+		\hfill
+		\includegraphics[page=10,width=0.40\linewidth]{appendix/crolgene_ICGC}\\
+		\caption{Box plots and scatter plots showing the association of the other CR obesity metagenes with the sample \gls{bmi} status  and \gls{bmi} from the \gls{icgc} \gls{cesc} data set, respectively.
+	P-values and $R^2$-value are as described in previous figures.}
+		\label{fig:appendix/cr_ob_meta_box_scatter_cesc}
+	\end{figure}
+
+	\begin{figure}[htpb]
+		\ContinuedFloat
+		\captionsetup{list=off,format=cont}
+		\centering
+		\includegraphics[page=9,width=0.45\linewidth]{appendix/resobsgene_ICGC}
+		\hfill
+		\includegraphics[page=10,width=0.45\linewidth]{appendix/resobsgene_ICGC}\\
+		\includegraphics[page=9,width=0.45\linewidth]{appendix/rescrolgene_ICGC}
+		\hfill
+		\includegraphics[page=10,width=0.45\linewidth]{appendix/rescrolgene_ICGC}\\
+		\includegraphics[page=9,width=0.45\linewidth]{appendix/caobsgene_ICGC}
+		\hfill
+		\includegraphics[page=10,width=0.45\linewidth]{appendix/caobsgene_ICGC}\\
+		\caption{Name}
+	\end{figure}
+
+	\begin{figure}[htpb]
+		\ContinuedFloat
+		\captionsetup{list=off,format=cont}
+		\centering
+		\includegraphics[page=9,width=0.45\linewidth]{appendix/cacrolgene_ICGC}
+		\hfill
+		\includegraphics[page=10,width=0.45\linewidth]{appendix/cacrolgene_ICGC}\\
+		\includegraphics[page=9,width=0.45\linewidth]{appendix/caresobsgene_ICGC}
+		\hfill
+		\includegraphics[page=10,width=0.45\linewidth]{appendix/caresobsgene_ICGC}\\
+		\includegraphics[page=9,width=0.45\linewidth]{appendix/carescrolgene_ICGC}
+		\hfill
+		\includegraphics[page=10,width=0.45\linewidth]{appendix/carescrolgene_ICGC}\\
+		\caption{Name}
+	\end{figure}
+
+	\newpage
+
+	\subsubsection{\gls{coad}}
+	\label{ssub:coad}
+
+	\begin{figure}[htp!]
+		\centering
+		\includegraphics[page=13,height=0.43\linewidth,width=0.45\linewidth]{appendix/rawobsgene_ICGC}
+		\hfill
+		\includegraphics[page=13,height=0.43\linewidth,width=0.45\linewidth]{appendix/crolgene_ICGC}\\
+		\includegraphics[page=13,height=0.43\linewidth,width=0.45\linewidth]{appendix/resobsgene_ICGC}
+		\hfill
+		\includegraphics[page=13,height=0.43\linewidth,width=0.45\linewidth]{appendix/rescrolgene_ICGC}\\
+		\includegraphics[page=13,height=0.43\linewidth,width=0.45\linewidth]{appendix/caobsgene_ICGC}
+		\hfill
+		\includegraphics[page=13,height=0.43\linewidth,width=0.45\linewidth]{appendix/cacrolgene_ICGC}\\
+		\caption{Heatmaps showing the association of the other CR obesity metagenes with the sample gene expression from the \gls{icgc} \gls{coad} data set.
+		Scales, p-values and $R^2$-value are as described in previous figures.}
+		\label{fig:degmetaicgc_coad}
+	\end{figure}
+
+	\begin{figure}[htpb]
+		\ContinuedFloat
+		\captionsetup{list=off,format=cont}
+		\centering
+		\includegraphics[page=13,height=0.42\linewidth,width=0.45\linewidth]{appendix/caresobsgene_ICGC}
+		\hfill
+		\includegraphics[page=13,height=0.42\linewidth,width=0.45\linewidth]{appendix/carescrolgene_ICGC}\\
+		\caption{Name}
+	\end{figure}
+
+	\begin{figure}[!htpb]
+		\centering
+		\includegraphics[page=14,width=0.40\linewidth]{appendix/rawobsgene_ICGC}
+		\hfill
+		\includegraphics[page=15,width=0.40\linewidth]{appendix/rawobsgene_ICGC}\\
+		\includegraphics[page=14,width=0.40\linewidth]{appendix/crolgene_ICGC}
+		\hfill
+		\includegraphics[page=15,width=0.40\linewidth]{appendix/crolgene_ICGC}\\
+		\caption{Box plots and scatter plots showing the association of the other CR obesity metagenes with the sample \gls{bmi} status  and \gls{bmi} from the \gls{icgc} \gls{coad} data set, respectively.
+	P-values and $R^2$-value are as described in previous figures.}
+		\label{fig:appendix/cr_ob_meta_box_scatter_coad}
+	\end{figure}
+
+	\begin{figure}[htpb]
+		\ContinuedFloat
+		\captionsetup{list=off,format=cont}
+		\centering
+		\includegraphics[page=14,width=0.45\linewidth]{appendix/resobsgene_ICGC}
+		\hfill
+		\includegraphics[page=15,width=0.45\linewidth]{appendix/resobsgene_ICGC}\\
+		\includegraphics[page=14,width=0.45\linewidth]{appendix/rescrolgene_ICGC}
+		\hfill
+		\includegraphics[page=15,width=0.45\linewidth]{appendix/rescrolgene_ICGC}\\
+		\includegraphics[page=14,width=0.45\linewidth]{appendix/caobsgene_ICGC}
+		\hfill
+		\includegraphics[page=15,width=0.45\linewidth]{appendix/caobsgene_ICGC}\\
+		\caption{Name}
+	\end{figure}
+
+	\begin{figure}[htpb]
+		\ContinuedFloat
+		\captionsetup{list=off,format=cont}
+		\centering
+		\includegraphics[page=14,width=0.45\linewidth]{appendix/cacrolgene_ICGC}
+		\hfill
+		\includegraphics[page=15,width=0.45\linewidth]{appendix/cacrolgene_ICGC}\\
+		\includegraphics[page=14,width=0.45\linewidth]{appendix/caresobsgene_ICGC}
+		\hfill
+		\includegraphics[page=15,width=0.45\linewidth]{appendix/caresobsgene_ICGC}\\
+		\includegraphics[page=14,width=0.45\linewidth]{appendix/carescrolgene_ICGC}
+		\hfill
+		\includegraphics[page=15,width=0.45\linewidth]{appendix/carescrolgene_ICGC}\\
+		\caption{Name}
+	\end{figure}
+
+	\newpage
+
+	\subsubsection{\gls{kirp}}
+	\label{ssub:kirp}
+
+	\begin{figure}[htp!]
+		\centering
+		\includegraphics[page=18,height=0.43\linewidth,width=0.45\linewidth]{appendix/rawobsgene_ICGC}
+		\hfill
+		\includegraphics[page=18,height=0.43\linewidth,width=0.45\linewidth]{appendix/crolgene_ICGC}\\
+		\includegraphics[page=18,height=0.43\linewidth,width=0.45\linewidth]{appendix/resobsgene_ICGC}
+		\hfill
+		\includegraphics[page=18,height=0.43\linewidth,width=0.45\linewidth]{appendix/rescrolgene_ICGC}\\
+		\includegraphics[page=18,height=0.43\linewidth,width=0.45\linewidth]{appendix/caobsgene_ICGC}
+		\hfill
+		\includegraphics[page=18,height=0.43\linewidth,width=0.45\linewidth]{appendix/cacrolgene_ICGC}\\
+		\caption{Heatmaps showing the association of the other CR obesity metagenes with the sample gene expression from the \gls{icgc} \gls{kirp} data set.
+		Scales, p-values and $R^2$-value are as described in previous figures.}
+		\label{fig:degmetaicgc_kirp}
+	\end{figure}
+
+	\begin{figure}[htpb]
+		\ContinuedFloat
+		\captionsetup{list=off,format=cont}
+		\centering
+		\includegraphics[page=18,height=0.42\linewidth,width=0.45\linewidth]{appendix/caresobsgene_ICGC}
+		\hfill
+		\includegraphics[page=18,height=0.42\linewidth,width=0.45\linewidth]{appendix/carescrolgene_ICGC}\\
+		\caption{Name}
+	\end{figure}
+
+	\begin{figure}[!htpb]
+		\centering
+		\includegraphics[page=19,width=0.40\linewidth]{appendix/rawobsgene_ICGC}
+		\hfill
+		\includegraphics[page=20,width=0.40\linewidth]{appendix/rawobsgene_ICGC}\\
+		\includegraphics[page=19,width=0.40\linewidth]{appendix/crolgene_ICGC}
+		\hfill
+		\includegraphics[page=20,width=0.40\linewidth]{appendix/crolgene_ICGC}\\
+		\caption{Box plots and scatter plots showing the association of the other CR obesity metagenes with the sample \gls{bmi} status  and \gls{bmi} from the \gls{icgc} \gls{kirp} data set, respectively.
+	P-values and $R^2$-value are as described in previous figures.}
+		\label{fig:appendix/cr_ob_meta_box_scatter_kirp}
+	\end{figure}
+
+	\begin{figure}[htpb]
+		\ContinuedFloat
+		\captionsetup{list=off,format=cont}
+		\centering
+		\includegraphics[page=19,width=0.45\linewidth]{appendix/resobsgene_ICGC}
+		\hfill
+		\includegraphics[page=20,width=0.45\linewidth]{appendix/resobsgene_ICGC}\\
+		\includegraphics[page=19,width=0.45\linewidth]{appendix/rescrolgene_ICGC}
+		\hfill
+		\includegraphics[page=20,width=0.45\linewidth]{appendix/rescrolgene_ICGC}\\
+		\includegraphics[page=19,width=0.45\linewidth]{appendix/caobsgene_ICGC}
+		\hfill
+		\includegraphics[page=20,width=0.45\linewidth]{appendix/caobsgene_ICGC}\\
+		\caption{Name}
+	\end{figure}
+
+	\begin{figure}[htpb]
+		\ContinuedFloat
+		\captionsetup{list=off,format=cont}
+		\centering
+		\includegraphics[page=19,width=0.45\linewidth]{appendix/cacrolgene_ICGC}
+		\hfill
+		\includegraphics[page=20,width=0.45\linewidth]{appendix/cacrolgene_ICGC}\\
+		\includegraphics[page=19,width=0.45\linewidth]{appendix/caresobsgene_ICGC}
+		\hfill
+		\includegraphics[page=20,width=0.45\linewidth]{appendix/caresobsgene_ICGC}\\
+		\includegraphics[page=19,width=0.45\linewidth]{appendix/carescrolgene_ICGC}
+		\hfill
+		\includegraphics[page=20,width=0.45\linewidth]{appendix/carescrolgene_ICGC}\\
+		\caption{Name}
+	\end{figure}
+
+	\newpage
+
+	\subsubsection{\gls{lihc}}
+	\label{ssub:lihc}
+
+	\begin{figure}[htp!]
+		\centering
+		\includegraphics[page=23,height=0.43\linewidth,width=0.45\linewidth]{appendix/rawobsgene_ICGC}
+		\hfill
+		\includegraphics[page=23,height=0.43\linewidth,width=0.45\linewidth]{appendix/crolgene_ICGC}\\
+		\includegraphics[page=23,height=0.43\linewidth,width=0.45\linewidth]{appendix/resobsgene_ICGC}
+		\hfill
+		\includegraphics[page=23,height=0.43\linewidth,width=0.45\linewidth]{appendix/rescrolgene_ICGC}\\
+		\includegraphics[page=23,height=0.43\linewidth,width=0.45\linewidth]{appendix/caobsgene_ICGC}
+		\hfill
+		\includegraphics[page=23,height=0.43\linewidth,width=0.45\linewidth]{appendix/cacrolgene_ICGC}\\
+		\caption{Heatmaps showing the association of the other CR obesity metagenes with the sample gene expression from the \gls{icgc} \gls{lihc} data set.
+		Scales, p-values and $R^2$-value are as described in previous figures.}
+		\label{fig:degmetaicgc_lihc}
+	\end{figure}
+
+	\begin{figure}[htpb]
+		\ContinuedFloat
+		\captionsetup{list=off,format=cont}
+		\centering
+		\includegraphics[page=23,height=0.42\linewidth,width=0.45\linewidth]{appendix/caresobsgene_ICGC}
+		\hfill
+		\includegraphics[page=23,height=0.42\linewidth,width=0.45\linewidth]{appendix/carescrolgene_ICGC}\\
+		\caption{Name}
+	\end{figure}
+
+	\begin{figure}[!htpb]
+		\centering
+		\includegraphics[page=24,width=0.40\linewidth]{appendix/rawobsgene_ICGC}
+		\hfill
+		\includegraphics[page=25,width=0.40\linewidth]{appendix/rawobsgene_ICGC}\\
+		\includegraphics[page=24,width=0.40\linewidth]{appendix/crolgene_ICGC}
+		\hfill
+		\includegraphics[page=25,width=0.40\linewidth]{appendix/crolgene_ICGC}\\
+		\caption{Box plots and scatter plots showing the association of the other CR obesity metagenes with the sample \gls{bmi} status  and \gls{bmi} from the \gls{icgc} \gls{lihc} data set, respectively.
+	P-values and $R^2$-value are as described in previous figures.}
+		\label{fig:appendix/cr_ob_meta_box_scatter_lihc}
+	\end{figure}
+
+	\begin{figure}[htpb]
+		\ContinuedFloat
+		\captionsetup{list=off,format=cont}
+		\centering
+		\includegraphics[page=24,width=0.45\linewidth]{appendix/resobsgene_ICGC}
+		\hfill
+		\includegraphics[page=25,width=0.45\linewidth]{appendix/resobsgene_ICGC}\\
+		\includegraphics[page=24,width=0.45\linewidth]{appendix/rescrolgene_ICGC}
+		\hfill
+		\includegraphics[page=25,width=0.45\linewidth]{appendix/rescrolgene_ICGC}\\
+		\includegraphics[page=24,width=0.45\linewidth]{appendix/caobsgene_ICGC}
+		\hfill
+		\includegraphics[page=25,width=0.45\linewidth]{appendix/caobsgene_ICGC}\\
+		\caption{Name}
+	\end{figure}
+
+	\begin{figure}[htpb]
+		\ContinuedFloat
+		\captionsetup{list=off,format=cont}
+		\centering
+		\includegraphics[page=24,width=0.45\linewidth]{appendix/cacrolgene_ICGC}
+		\hfill
+		\includegraphics[page=25,width=0.45\linewidth]{appendix/cacrolgene_ICGC}\\
+		\includegraphics[page=24,width=0.45\linewidth]{appendix/caresobsgene_ICGC}
+		\hfill
+		\includegraphics[page=25,width=0.45\linewidth]{appendix/caresobsgene_ICGC}\\
+		\includegraphics[page=24,width=0.45\linewidth]{appendix/carescrolgene_ICGC}
+		\hfill
+		\includegraphics[page=25,width=0.45\linewidth]{appendix/carescrolgene_ICGC}\\
+		\caption{Name}
+	\end{figure}
+
+	\newpage
+
+	\subsubsection{\gls{read}}
+	\label{ssub:read}
+
+	\begin{figure}[htp!]
+		\centering
+		\includegraphics[page=28,height=0.43\linewidth,width=0.45\linewidth]{appendix/rawobsgene_ICGC}
+		\hfill
+		\includegraphics[page=28,height=0.43\linewidth,width=0.45\linewidth]{appendix/crolgene_ICGC}\\
+		\includegraphics[page=28,height=0.43\linewidth,width=0.45\linewidth]{appendix/resobsgene_ICGC}
+		\hfill
+		\includegraphics[page=28,height=0.43\linewidth,width=0.45\linewidth]{appendix/rescrolgene_ICGC}\\
+		\includegraphics[page=28,height=0.43\linewidth,width=0.45\linewidth]{appendix/caobsgene_ICGC}
+		\hfill
+		\includegraphics[page=28,height=0.43\linewidth,width=0.45\linewidth]{appendix/cacrolgene_ICGC}\\
+		\caption{Heatmaps showing the association of the other CR obesity metagenes with the sample gene expression from the \gls{icgc} \gls{read} data set.
+		Scales, p-values and $R^2$-value are as described in previous figures.}
+		\label{fig:degmetaicgc_read}
+	\end{figure}
+
+	\begin{figure}[htpb]
+		\ContinuedFloat
+		\captionsetup{list=off,format=cont}
+		\centering
+		\includegraphics[page=28,height=0.42\linewidth,width=0.45\linewidth]{appendix/caresobsgene_ICGC}
+		\hfill
+		\includegraphics[page=28,height=0.42\linewidth,width=0.45\linewidth]{appendix/carescrolgene_ICGC}\\
+		\caption{Name}
+	\end{figure}
+
+	\begin{figure}[!htpb]
+		\centering
+		\includegraphics[page=29,width=0.40\linewidth]{appendix/rawobsgene_ICGC}
+		\hfill
+		\includegraphics[page=30,width=0.40\linewidth]{appendix/rawobsgene_ICGC}\\
+		\includegraphics[page=29,width=0.40\linewidth]{appendix/crolgene_ICGC}
+		\hfill
+		\includegraphics[page=30,width=0.40\linewidth]{appendix/crolgene_ICGC}\\
+		\caption{Box plots and scatter plots showing the association of the other CR obesity metagenes with the sample \gls{bmi} status  and \gls{bmi} from the \gls{icgc} \gls{read} data set, respectively.
+	P-values and $R^2$-value are as described in previous figures.}
+		\label{fig:appendix/cr_ob_meta_box_scatter_read}
+	\end{figure}
+
+	\begin{figure}[htpb]
+		\ContinuedFloat
+		\captionsetup{list=off,format=cont}
+		\centering
+		\includegraphics[page=29,width=0.45\linewidth]{appendix/resobsgene_ICGC}
+		\hfill
+		\includegraphics[page=30,width=0.45\linewidth]{appendix/resobsgene_ICGC}\\
+		\includegraphics[page=29,width=0.45\linewidth]{appendix/rescrolgene_ICGC}
+		\hfill
+		\includegraphics[page=30,width=0.45\linewidth]{appendix/rescrolgene_ICGC}\\
+		\includegraphics[page=29,width=0.45\linewidth]{appendix/caobsgene_ICGC}
+		\hfill
+		\includegraphics[page=30,width=0.45\linewidth]{appendix/caobsgene_ICGC}\\
+		\caption{Name}
+	\end{figure}
+
+	\begin{figure}[htpb]
+		\ContinuedFloat
+		\captionsetup{list=off,format=cont}
+		\centering
+		\includegraphics[page=29,width=0.45\linewidth]{appendix/cacrolgene_ICGC}
+		\hfill
+		\includegraphics[page=30,width=0.45\linewidth]{appendix/cacrolgene_ICGC}\\
+		\includegraphics[page=29,width=0.45\linewidth]{appendix/caresobsgene_ICGC}
+		\hfill
+		\includegraphics[page=30,width=0.45\linewidth]{appendix/caresobsgene_ICGC}\\
+		\includegraphics[page=29,width=0.45\linewidth]{appendix/carescrolgene_ICGC}
+		\hfill
+		\includegraphics[page=30,width=0.45\linewidth]{appendix/carescrolgene_ICGC}\\
+		\caption{Name}
+	\end{figure}
+
+	\newpage
+
+	\subsubsection{\gls{skcm}}
+	\label{ssub:skcm}
+
+	\begin{figure}[htp!]
+		\centering
+		\includegraphics[page=33,height=0.43\linewidth,width=0.45\linewidth]{appendix/rawobsgene_ICGC}
+		\hfill
+		\includegraphics[page=33,height=0.43\linewidth,width=0.45\linewidth]{appendix/crolgene_ICGC}\\
+		\includegraphics[page=33,height=0.43\linewidth,width=0.45\linewidth]{appendix/resobsgene_ICGC}
+		\hfill
+		\includegraphics[page=33,height=0.43\linewidth,width=0.45\linewidth]{appendix/rescrolgene_ICGC}\\
+		\includegraphics[page=33,height=0.43\linewidth,width=0.45\linewidth]{appendix/caobsgene_ICGC}
+		\hfill
+		\includegraphics[page=33,height=0.43\linewidth,width=0.45\linewidth]{appendix/cacrolgene_ICGC}\\
+		\caption{Heatmaps showing the association of the other CR obesity metagenes with the sample gene expression from the \gls{icgc} \gls{skcm} data set.
+		Scales, p-values and $R^2$-value are as described in previous figures.}
+		\label{fig:degmetaicgc_skcm}
+	\end{figure}
+
+	\begin{figure}[htpb]
+		\ContinuedFloat
+		\captionsetup{list=off,format=cont}
+		\centering
+		\includegraphics[page=33,height=0.42\linewidth,width=0.45\linewidth]{appendix/caresobsgene_ICGC}
+		\hfill
+		\includegraphics[page=33,height=0.42\linewidth,width=0.45\linewidth]{appendix/carescrolgene_ICGC}\\
+		\caption{Name}
+	\end{figure}
+
+	\begin{figure}[!htpb]
+		\centering
+		\includegraphics[page=34,width=0.40\linewidth]{appendix/rawobsgene_ICGC}
+		\hfill
+		\includegraphics[page=35,width=0.40\linewidth]{appendix/rawobsgene_ICGC}\\
+		\includegraphics[page=34,width=0.40\linewidth]{appendix/crolgene_ICGC}
+		\hfill
+		\includegraphics[page=35,width=0.40\linewidth]{appendix/crolgene_ICGC}\\
+		\caption{Box plots and scatter plots showing the association of the other CR obesity metagenes with the sample \gls{bmi} status  and \gls{bmi} from the \gls{icgc} \gls{skcm} data set, respectively.
+	P-values and $R^2$-value are as described in previous figures.}
+		\label{fig:appendix/cr_ob_meta_box_scatter_skcm}
+	\end{figure}
+
+	\begin{figure}[htpb]
+		\ContinuedFloat
+		\captionsetup{list=off,format=cont}
+		\centering
+		\includegraphics[page=34,width=0.45\linewidth]{appendix/resobsgene_ICGC}
+		\hfill
+		\includegraphics[page=35,width=0.45\linewidth]{appendix/resobsgene_ICGC}\\
+		\includegraphics[page=34,width=0.45\linewidth]{appendix/rescrolgene_ICGC}
+		\hfill
+		\includegraphics[page=35,width=0.45\linewidth]{appendix/rescrolgene_ICGC}\\
+		\includegraphics[page=34,width=0.45\linewidth]{appendix/caobsgene_ICGC}
+		\hfill
+		\includegraphics[page=35,width=0.45\linewidth]{appendix/caobsgene_ICGC}\\
+		\caption{Name}
+	\end{figure}
+
+	\begin{figure}[htpb]
+		\ContinuedFloat
+		\captionsetup{list=off,format=cont}
+		\centering
+		\includegraphics[page=34,width=0.45\linewidth]{appendix/cacrolgene_ICGC}
+		\hfill
+		\includegraphics[page=35,width=0.45\linewidth]{appendix/cacrolgene_ICGC}\\
+		\includegraphics[page=34,width=0.45\linewidth]{appendix/caresobsgene_ICGC}
+		\hfill
+		\includegraphics[page=35,width=0.45\linewidth]{appendix/caresobsgene_ICGC}\\
+		\includegraphics[page=34,width=0.45\linewidth]{appendix/carescrolgene_ICGC}
+		\hfill
+		\includegraphics[page=35,width=0.45\linewidth]{appendix/carescrolgene_ICGC}\\
+		\caption{Name}
+	\end{figure}
+
+	\newpage
+
+	\subsubsection{\gls{ucec}}
+	\label{ssub:ucec}
+
+	\begin{figure}[htp!]
+		\centering
+		\includegraphics[page=38,height=0.43\linewidth,width=0.45\linewidth]{appendix/rawobsgene_ICGC}
+		\hfill
+		\includegraphics[page=38,height=0.43\linewidth,width=0.45\linewidth]{appendix/crolgene_ICGC}\\
+		\includegraphics[page=38,height=0.43\linewidth,width=0.45\linewidth]{appendix/resobsgene_ICGC}
+		\hfill
+		\includegraphics[page=38,height=0.43\linewidth,width=0.45\linewidth]{appendix/rescrolgene_ICGC}\\
+		\includegraphics[page=38,height=0.43\linewidth,width=0.45\linewidth]{appendix/caobsgene_ICGC}
+		\hfill
+		\includegraphics[page=38,height=0.43\linewidth,width=0.45\linewidth]{appendix/cacrolgene_ICGC}\\
+		\caption{Heatmaps showing the association of the other CR obesity metagenes with the sample gene expression from the \gls{icgc} \gls{ucec} data set.
+		Scales, p-values and $R^2$-value are as described in previous figures.}
+		\label{fig:degmetaicgc_ucec}
+	\end{figure}
+
+	\begin{figure}[htpb]
+		\ContinuedFloat
+		\captionsetup{list=off,format=cont}
+		\centering
+		\includegraphics[page=38,height=0.42\linewidth,width=0.45\linewidth]{appendix/caresobsgene_ICGC}
+		\hfill
+		\includegraphics[page=38,height=0.42\linewidth,width=0.45\linewidth]{appendix/carescrolgene_ICGC}\\
+		\caption{Name}
+	\end{figure}
+
+	\begin{figure}[!htpb]
+		\centering
+		\includegraphics[page=39,width=0.40\linewidth]{appendix/rawobsgene_ICGC}
+		\hfill
+		\includegraphics[page=40,width=0.40\linewidth]{appendix/rawobsgene_ICGC}\\
+		\includegraphics[page=39,width=0.40\linewidth]{appendix/crolgene_ICGC}
+		\hfill
+		\includegraphics[page=40,width=0.40\linewidth]{appendix/crolgene_ICGC}\\
+		\caption{Box plots and scatter plots showing the association of the other CR obesity metagenes with the sample \gls{bmi} status  and \gls{bmi} from the \gls{icgc} \gls{ucec} data set, respectively.
+	P-values and $R^2$-value are as described in previous figures.}
+		\label{fig:appendix/cr_ob_meta_box_scatter_ucec}
+	\end{figure}
+
+	\begin{figure}[htpb]
+		\ContinuedFloat
+		\captionsetup{list=off,format=cont}
+		\centering
+		\includegraphics[page=39,width=0.45\linewidth]{appendix/resobsgene_ICGC}
+		\hfill
+		\includegraphics[page=40,width=0.45\linewidth]{appendix/resobsgene_ICGC}\\
+		\includegraphics[page=39,width=0.45\linewidth]{appendix/rescrolgene_ICGC}
+		\hfill
+		\includegraphics[page=40,width=0.45\linewidth]{appendix/rescrolgene_ICGC}\\
+		\includegraphics[page=39,width=0.45\linewidth]{appendix/caobsgene_ICGC}
+		\hfill
+		\includegraphics[page=40,width=0.45\linewidth]{appendix/caobsgene_ICGC}\\
+		\caption{Name}
+	\end{figure}
+
+	\begin{figure}[htpb]
+		\ContinuedFloat
+		\captionsetup{list=off,format=cont}
+		\centering
+		\includegraphics[page=39,width=0.45\linewidth]{appendix/cacrolgene_ICGC}
+		\hfill
+		\includegraphics[page=40,width=0.45\linewidth]{appendix/cacrolgene_ICGC}\\
+		\includegraphics[page=39,width=0.45\linewidth]{appendix/caresobsgene_ICGC}
+		\hfill
+		\includegraphics[page=40,width=0.45\linewidth]{appendix/caresobsgene_ICGC}\\
+		\includegraphics[page=39,width=0.45\linewidth]{appendix/carescrolgene_ICGC}
+		\hfill
+		\includegraphics[page=40,width=0.45\linewidth]{appendix/carescrolgene_ICGC}\\
+		\caption{Name}
+	\end{figure}
+
+	\newpage
+
+	\section{Remainder of the results of the other CR obesity metagenes in \gls{nzbc} data set}
+	\label{sec:rest_of_the_cr_ob_meta_heatmap_results_cris}
+
+	All of the other obesity metagenes from the CR data set were able to capture the overall gene expressions of the samples in the \gls{nzbc} data set (\cref{fig:appendix/cr_ob_meta_heatmap_cris}).
+	However, none of the CR obesity metagenes were able to significantly associate with the sample \gls{bmi}/\gls{bmi} status in the \gls{nzbc} data set (\cref{fig:appendix/cr_ob_meta_box_scatter_cris}).
+
+	\begin{figure}[htp!]
+		\centering
+		\includegraphics[width=0.45\linewidth,page=8]{appendix/cris_crdeg_trans_meta}\\
+		\vspace{1em}
+		\includegraphics[width=0.45\linewidth,page=13]{appendix/cris_crdeg_trans_meta}
+		\hfill
+		\includegraphics[width=0.45\linewidth,page=18]{appendix/cris_crdeg_trans_meta}\\
+		\caption{Heatmaps showing the association of the other CR obesity metagenes with the sample gene expression from the \gls{nzbc} data set.
+	Scales are as described in previous figures.}
+		\label{fig:appendix/cr_ob_meta_heatmap_cris}
+	\end{figure}
+
+	\begin{figure}[htpb]
+		\ContinuedFloat
+		\captionsetup{list=off,format=cont}
+		\centering
+		\includegraphics[width=0.45\linewidth,page=23]{appendix/cris_crdeg_trans_meta}
+		\hfill
+		\includegraphics[width=0.45\linewidth,page=28]{appendix/cris_crdeg_trans_meta}\\
+		\includegraphics[width=0.45\linewidth,page=33]{appendix/cris_crdeg_trans_meta}
+		\hfill
+		\includegraphics[width=0.45\linewidth,page=38]{appendix/cris_crdeg_trans_meta}\\
+		\caption{figure}
+	\end{figure}
+
+	\begin{figure}[htpb]
+		\centering
+		\includegraphics[page=9,width=0.42\linewidth]{appendix/cris_crdeg_trans_meta}
+		\hfill
+		\includegraphics[page=10,width=0.42\linewidth]{appendix/cris_crdeg_trans_meta}\\
+		\caption{Box plots and scatter plots showing the association of the other CR obesity metagenes with the sample \gls{bmi} status  and \gls{bmi} from the \gls{nzbc} data set, respectively.
+	P-values and $R^2$-value are as described in previous figures.}
+		\label{fig:appendix/cr_ob_meta_box_scatter_cris}
+	\end{figure}
+
+	\begin{figure}[htpb]
+		\ContinuedFloat
+		\captionsetup{list=off,format=cont}
+		\centering
+		\includegraphics[page=14,width=0.45\linewidth]{appendix/cris_crdeg_trans_meta}
+		\hfill
+		\includegraphics[page=15,width=0.45\linewidth]{appendix/cris_crdeg_trans_meta}\\
+		\includegraphics[page=19,width=0.45\linewidth]{appendix/cris_crdeg_trans_meta}
+		\hfill
+		\includegraphics[page=20,width=0.45\linewidth]{appendix/cris_crdeg_trans_meta}\\
+		\includegraphics[page=24,width=0.45\linewidth]{appendix/cris_crdeg_trans_meta}
+		\hfill
+		\includegraphics[page=25,width=0.45\linewidth]{appendix/cris_crdeg_trans_meta}\\
+		\caption{Name}
+	\end{figure}
+
+	\begin{figure}[htpb]
+		\ContinuedFloat
+		\captionsetup{list=off,format=cont}
+		\centering
+		\includegraphics[page=29,width=0.45\linewidth]{appendix/cris_crdeg_trans_meta}
+		\hfill
+		\includegraphics[page=30,width=0.45\linewidth]{appendix/cris_crdeg_trans_meta}\\
+		\includegraphics[page=34,width=0.45\linewidth]{appendix/cris_crdeg_trans_meta}
+		\hfill
+		\includegraphics[page=35,width=0.45\linewidth]{appendix/cris_crdeg_trans_meta}\\
+		\includegraphics[page=39,width=0.45\linewidth]{appendix/cris_crdeg_trans_meta}
+		\hfill
+		\includegraphics[page=40,width=0.45\linewidth]{appendix/cris_crdeg_trans_meta}\\
+		\caption{Name}
+	\end{figure}
 

### 2017-03-11 thesis/results/results.tex
Really finished the draft for appendix A - tmp commit

@@
 	\vspace{1em}
-	\includegraphics[page=4,width=0.45\linewidth]{results1/cris_cr_trans_meta}
+	\includegraphics[page=4,width=0.45\linewidth]{results1/cris_crdeg_trans_meta}
 	\hfill
-	\includegraphics[page=5,width=0.45\linewidth]{results1/cris_cr_trans_meta}
+	\includegraphics[page=5,width=0.45\linewidth]{results1/cris_crdeg_trans_meta}
 	\caption[Cr obesity associated metagene in \gls{nzbc} data]{Heatmap, box plot and scatter plot showing the association of Cr obesity associated metagene with sample gene expression, \gls{bmi} and \gls{bmi} status from \gls{nzbc} data, respectively.

### 2017-03-11 thesis/misc/appendix2.tex
Divided the appendix file into two - tmp commit

	\chapter{Additional results from \cref{cha:obesity_associated_genetic_signature_and_pathway_signatures}}
	\label{app:b}

	\begin{figure}[htpb]
		\centering
		\includegraphics[width=0.7\linewidth]{results2/gatzares1}
		\caption{results from gatza's study}
		\label{fig:gatza_paper_res}
	\end{figure}

	% [1] "crolgene"

	% Call:
	% lm(formula = as.formula(formula_txt[j]), data = data)

	% Coefficients:
	%              Estimate Std. Error t value Pr(>|t|)
	% (Intercept)         0.553214  0.123891 4.465  2.16e-05 ***
	% bmi                 -0.001629 0.004071 -0.400 0.69
	% (Intercept)         0.55700   0.05464  10.194 <2e-16   ***
	% bmiStatusobese      -0.10551  0.07021  -1.503 0.136
	% bmiStatusoverweight -0.02165  0.07727  -0.280 0.780
	% (Intercept)         0.307513  0.163530 1.880  0.0631   .
	% bmi                 0.011279  0.006976 1.617  0.1092
	% bmiStatusobese      -0.261145 0.118795 -2.198 0.0304   *
	% bmiStatusoverweight -0.079845 0.084661 -0.943 0.3480
	% (Intercept)         0.007141  0.382725 0.019  0.9852
	% bcat_probes         0.156062  0.223779 0.697  0.4873
	% er_probes           0.123403  0.254800 0.484  0.6293
	% ifna_probes         -0.430229 0.497301 -0.865 0.3892
	% ifng_probes         0.378437  0.509352 0.743  0.4594
	% myc_probes          0.160785  0.229142 0.702  0.4846
	% pr_probes           0.597403  0.244252 2.446  0.0164   *
	% (Intercept)         0.055088  0.389472 0.141  0.8878
	% bmi                 -0.002745 0.003814 -0.720 0.4736
	% bcat_probes         0.179900  0.226799 0.793  0.4297
	% er_probes           0.125413  0.255485 0.491  0.6247
	% ifna_probes         -0.475417 0.502546 -0.946 0.3466
	% ifng_probes         0.429131  0.515526 0.832  0.4074
	% myc_probes          0.184642  0.232124 0.795  0.4284
	% pr_probes           0.607982  0.245336 2.478  0.0151   *
	% (Intercept)         0.01666   0.37989  0.044  0.9651
	% bmiStatusobese      -0.11197  0.06602  -1.696 0.0933   .
	% bmiStatusoverweight -0.02348  0.07151  -0.328 0.7434
	% bcat_probes         0.22035   0.22616  0.974  0.3325
	% er_probes           0.10450   0.25330  0.413  0.6809
	% ifna_probes         -0.60343  0.50372  -1.198 0.2341
	% ifng_probes         0.54671   0.51484  1.062  0.2911
	% myc_probes          0.21341   0.22998  0.928  0.3559
	% pr_probes           0.59491   0.24257  2.453  0.0161   *
	% (Intercept)         -0.128149 0.393204 -0.326 0.7453
	% bmi                 0.008834  0.006565 1.346  0.1819
	% bmiStatusobese      -0.234634 0.112386 -2.088 0.0397   *
	% bmiStatusoverweight -0.067942 0.078484 -0.866 0.3890
	% bcat_probes         0.216892  0.225167 0.963  0.3380
	% er_probes           0.081506  0.252745 0.322  0.7478
	% ifna_probes         -0.625864 0.501740 -1.247 0.2155
	% ifng_probes         0.546873  0.512538 1.067  0.2889
	% myc_probes          0.196215  0.229305 0.856  0.3945
	% pr_probes           0.560706  0.242816 2.309  0.0232   *

	% [1] "resobsgene"

	% Call:
	% lm(formula = as.formula(formula_txt[j]), data = data)

	% Coefficients:
	%              Estimate Std. Error t value Pr(>|t|)
	% (Intercept)         0.547104  0.123915 4.415  2.62e-05 ***
	% bmi                 -0.001422 0.004072 -0.349 0.728
	% (Intercept)         0.547619  0.054641 10.022 <2e-16   ***
	% bmiStatusobese      -0.096832 0.070212 -1.379 0.171
	% bmiStatusoverweight -0.001804 0.077273 -0.023 0.981
	% (Intercept)         0.298589  0.163546 1.826  0.0710   .
	% bmi                 0.011259  0.006976 1.614  0.1099
	% bmiStatusobese      -0.252186 0.118807 -2.123 0.0364   *
	% bmiStatusoverweight -0.059898 0.084669 -0.707 0.4810
	% (Intercept)         0.18694   0.36963  0.506  0.6142
	% bcat_probes         0.09717   0.21612  0.450  0.6541
	% er_probes           0.07879   0.24608  0.320  0.7496
	% ifna_probes         -0.17576  0.48029  -0.366 0.7152
	% ifng_probes         0.07930   0.49193  0.161  0.8723
	% myc_probes          -0.04517  0.22130  -0.204 0.8387
	% pr_probes           0.59554   0.23590  2.525  0.0133   *
	% (Intercept)         0.221757  0.376612 0.589  0.5574
	% bmi                 -0.001994 0.003688 -0.540 0.5902
	% bcat_probes         0.114479  0.219311 0.522  0.6029
	% er_probes           0.080247  0.247050 0.325  0.7461
	% ifna_probes         -0.208580 0.485954 -0.429 0.6688
	% ifng_probes         0.116119  0.498505 0.233  0.8163
	% myc_probes          -0.027845 0.224460 -0.124 0.9015
	% pr_probes           0.603222  0.237235 2.543  0.0127   *
	% (Intercept)         0.195965  0.367384 0.533  0.5951
	% bmiStatusobese      -0.096834 0.063844 -1.517 0.1328
	% bmiStatusoverweight -0.004412 0.069155 -0.064 0.9493
	% bcat_probes         0.150379  0.218719 0.688  0.4935
	% er_probes           0.058880  0.244961 0.240  0.8106
	% ifna_probes         -0.344126 0.487134 -0.706 0.4817
	% ifng_probes         0.242697  0.497893 0.487  0.6271
	% myc_probes          -0.001285 0.222407 -0.006 0.9954
	% pr_probes           0.591191  0.234582 2.520  0.0135   *
	% (Intercept)         0.041459  0.379420 0.109  0.9132
	% bmi                 0.009426  0.006335 1.488  0.1403
	% bmiStatusobese      -0.227717 0.108446 -2.100 0.0386   *
	% bmiStatusoverweight -0.051853 0.075733 -0.685 0.4953
	% bcat_probes         0.146687  0.217273 0.675  0.5013
	% er_probes           0.034350  0.243884 0.141  0.8883
	% ifna_probes         -0.368065 0.484150 -0.760 0.4491
	% ifng_probes         0.242870  0.494570 0.491  0.6246
	% myc_probes          -0.019634 0.221267 -0.089 0.9295
	% pr_probes           0.554701  0.234304 2.367  0.0201   *

	% [1] "rescrolgene"

	% Call:
	% lm(formula = as.formula(formula_txt[j]), data = data)

	% Coefficients:
	%               Estimate Std. Error t value Pr(>|t|)
	% (Intercept)         0.5343899  0.1239554 4.311  3.9e-05 ***
	% bmi                 -0.0009922 0.0040730 -0.244 0.808
	% (Intercept)         0.54870    0.05482   10.009 <2e-16  ***
	% bmiStatusobese      -0.09063   0.07044   -1.287 0.201
	% bmiStatusoverweight -0.01515   0.07753   -0.195 0.845
	% (Intercept)         0.304200   0.164179  1.853  0.0670  .
	% bmi                 0.011054   0.007003  1.578  0.1178
	% bmiStatusobese      -0.243160  0.119267  -2.039 0.0442  *
	% bmiStatusoverweight -0.072189  0.084997  -0.849 0.3978
	% (Intercept)         0.17069    0.37929   0.450  0.6537
	% bcat_probes         0.09532    0.22177   0.430  0.6683
	% er_probes           0.03171    0.25251   0.126  0.9004
	% ifna_probes         -0.57965   0.49283   -1.176 0.2426
	% ifng_probes         0.50919    0.50478   1.009  0.3157
	% myc_probes          0.06345    0.22708   0.279  0.7805
	% pr_probes           0.54200    0.24206   2.239  0.0276  *
	% (Intercept)         0.204758   0.386507  0.530  0.5976
	% bmi                 -0.001950  0.003785  -0.515 0.6076
	% bcat_probes         0.112257   0.225072  0.499  0.6192
	% er_probes           0.033134   0.253540  0.131  0.8963
	% ifna_probes         -0.611750  0.498720  -1.227 0.2231
	% ifng_probes         0.545210   0.511601  1.066  0.2894
	% myc_probes          0.080405   0.230357  0.349  0.7279
	% pr_probes           0.549516   0.243468  2.257  0.0264  *
	% (Intercept)         0.17933    0.37777   0.475  0.636
	% bmiStatusobese      -0.09681   0.06565   -1.475 0.144
	% bmiStatusoverweight -0.01219   0.07111   -0.171 0.864
	% bcat_probes         0.14969    0.22490   0.666  0.507
	% er_probes           0.01354    0.25189   0.054  0.957
	% ifna_probes         -0.73888   0.50091   -1.475 0.144
	% ifng_probes         0.66381    0.51197   1.297  0.198
	% myc_probes          0.10813    0.22869   0.473  0.638
	% pr_probes           0.53872    0.24121   2.233  0.028   *
	% (Intercept)         0.027729   0.390580  0.071  0.9436
	% bmi                 0.009248   0.006522  1.418  0.1596
	% bmiStatusobese      -0.225233  0.111636  -2.018 0.0466  *
	% bmiStatusoverweight -0.058734  0.077961  -0.753 0.4532
	% bcat_probes         0.146066   0.223664  0.653  0.5154
	% er_probes           -0.010526  0.251058  -0.042 0.9667
	% ifna_probes         -0.762373  0.498391  -1.530 0.1296
	% ifng_probes         0.663981   0.509117  1.304  0.1955
	% myc_probes          0.090122   0.227775  0.396  0.6933
	% pr_probes           0.502919   0.241196  2.085  0.0399  *

	% [1] "caobsgene"

	% Call:
	% lm(formula = as.formula(formula_txt[j]), data = data)

	% Coefficients:
	%              Estimate Std. Error t value Pr(>|t|)
	% (Intercept)         0.538968  0.123943 4.349  3.39e-05 ***
	% bmi                 -0.001147 0.004073 -0.282 0.779
	% (Intercept)         0.53896   0.05458  9.875  2.76e-16 ***
	% bmiStatusobese      -0.08982  0.07013  -1.281 0.203
	% bmiStatusoverweight 0.01804   0.07719  0.234  0.816
	% (Intercept)         0.27986   0.16317  1.715  0.0896   .
	% bmi                 0.01171   0.00696  1.683  0.0957   .
	% bmiStatusobese      -0.25146  0.11853  -2.121 0.0365   *
	% bmiStatusoverweight -0.04241  0.08447  -0.502 0.6168
	% (Intercept)         0.12454   0.38803  0.321  0.7490
	% bcat_probes         0.14068   0.22688  0.620  0.5367
	% er_probes           0.11597   0.25833  0.449  0.6545
	% ifna_probes         -0.33319  0.50420  -0.661 0.5104
	% ifng_probes         0.30871   0.51641  0.598  0.5515
	% myc_probes          -0.04569  0.23232  -0.197 0.8445
	% pr_probes           0.56692   0.24764  2.289  0.0243   *
	% (Intercept)         0.162063  0.395326 0.410  0.6828
	% bmi                 -0.002148 0.003872 -0.555 0.5804
	% bcat_probes         0.159335  0.230208 0.692  0.4906
	% er_probes           0.117547  0.259326 0.453  0.6514
	% ifna_probes         -0.368552 0.510101 -0.723 0.4718
	% ifng_probes         0.348376  0.523276 0.666  0.5072
	% myc_probes          -0.027017 0.235613 -0.115 0.9090
	% pr_probes           0.575201  0.249024 2.310  0.0232   *
	% (Intercept)         0.134514  0.384960 0.349  0.7276
	% bmiStatusobese      -0.099695 0.066899 -1.490 0.1397
	% bmiStatusoverweight 0.008987  0.072463 0.124  0.9016
	% bcat_probes         0.193430  0.229183 0.844  0.4009
	% er_probes           0.092451  0.256681 0.360  0.7196
	% ifna_probes         -0.522336 0.510440 -1.023 0.3089
	% ifng_probes         0.492130  0.521713 0.943  0.3481
	% myc_probes          -0.001892 0.233048 -0.008 0.9935
	% pr_probes           0.560583  0.245805 2.281  0.0249   *
	% (Intercept)         -0.028482 0.397505 -0.072 0.9430
	% bmi                 0.009944  0.006637 1.498  0.1376
	% bmiStatusobese      -0.237770 0.113616 -2.093 0.0392   *
	% bmiStatusoverweight -0.041061 0.079343 -0.518 0.6061
	% bcat_probes         0.189535  0.227630 0.833  0.4073
	% er_probes           0.066573  0.255509 0.261  0.7950
	% ifna_probes         -0.547591 0.507227 -1.080 0.2832
	% ifng_probes         0.492312  0.518143 0.950  0.3446
	% myc_probes          -0.021250 0.231813 -0.092 0.9272
	% pr_probes           0.522087  0.245472 2.127  0.0362   *

	% [1] "cacrolgene"

	% Call:
	% lm(formula = as.formula(formula_txt[j]), data = data)

	% Coefficients:
	%               Estimate Std. Error t value Pr(>|t|)
	% (Intercept)         5.062e-01  1.240e-01 4.083  9.16e-05 ***
	% bmi                 -3.953e-05 4.074e-03 -0.010 0.992
	% (Intercept)         0.52453    0.05485   9.564  1.29e-15 ***
	% bmiStatusobese      -0.06411   0.07047   -0.910 0.365
	% bmiStatusoverweight 0.02958    0.07756   0.381  0.704
	% (Intercept)         0.273274   0.164135  1.665  0.0992   .
	% bmi                 0.011360   0.007001  1.622  0.1080
	% bmiStatusobese      -0.220856  0.119235  -1.852 0.0671   .
	% bmiStatusoverweight -0.029032  0.084974  -0.342 0.7334
	% (Intercept)         0.02932    0.40048   0.073  0.9418
	% bcat_probes         0.23660    0.23416   1.010  0.3150
	% er_probes           0.09521    0.26662   0.357  0.7218
	% ifna_probes         -0.49129   0.52037   -0.944 0.3476
	% ifng_probes         0.44466    0.53298   0.834  0.4063
	% myc_probes          0.16111    0.23977   0.672  0.5033
	% pr_probes           0.49566    0.25558   1.939  0.0555   .
	% (Intercept)         0.051698   0.408467  0.127  0.8996
	% bmi                 -0.001281  0.004000  -0.320 0.7495
	% bcat_probes         0.247720   0.237861  1.041  0.3004
	% er_probes           0.096149   0.267946  0.359  0.7205
	% ifna_probes         -0.512383  0.527057  -0.972 0.3335
	% ifng_probes         0.468320   0.540670  0.866  0.3887
	% myc_probes          0.172240   0.243445  0.708  0.4811
	% pr_probes           0.500601   0.257301  1.946  0.0548   .
	% (Intercept)         0.03806    0.39931   0.095  0.9243
	% bmiStatusobese      -0.07973   0.06939   -1.149 0.2536
	% bmiStatusoverweight 0.02249    0.07516   0.299  0.7655
	% bcat_probes         0.27648    0.23772   1.163  0.2479
	% er_probes           0.07298    0.26625   0.274  0.7846
	% ifna_probes         -0.66043   0.52946   -1.247 0.2155
	% ifng_probes         0.60854    0.54116   1.125  0.2638
	% myc_probes          0.19456    0.24173   0.805  0.4230
	% pr_probes           0.48849    0.25497   1.916  0.0586   .
	% (Intercept)         -0.123986  0.412743  -0.300 0.7646
	% bmi                 0.009886   0.006892  1.434  0.1549
	% bmiStatusobese      -0.216997  0.117971  -1.839 0.0692   .
	% bmiStatusoverweight -0.027267  0.082384  -0.331 0.7414
	% bcat_probes         0.272606   0.236356  1.153  0.2518
	% er_probes           0.047249   0.265304  0.178  0.8591
	% ifna_probes         -0.685542  0.526672  -1.302 0.1964
	% ifng_probes         0.608724   0.538006  1.131  0.2609
	% myc_probes          0.175314   0.240700  0.728  0.4683
	% pr_probes           0.450215   0.254882  1.766  0.0808   .

	% [1] "caresobsgene"

	% Call:
	% lm(formula = as.formula(formula_txt[j]), data = data)

	% Coefficients:
	%             Estimate Std. Error t value Pr(>|t|)
	% (Intercept)         0.462424  0.123913 3.732  0.00032  ***
	% bmi                 0.001442  0.004072 0.354  0.72408
	% (Intercept)         0.46970   0.05452  8.616  1.39e-13 ***
	% bmiStatusobese      0.09338   0.07005  1.333  0.186
	% bmiStatusoverweight -0.01840  0.07710  -0.239 0.812
	% (Intercept)         0.722057  0.163100 4.427  2.55e-05 ***
	% bmi                 -0.011409 0.006957 -1.640 0.1043
	% bmiStatusobese      0.250806  0.118483 2.117  0.0369   *
	% bmiStatusoverweight 0.040473  0.084438 0.479  0.6328
	% (Intercept)         1.1017    0.3815   2.888  0.00483  **
	% bcat_probes         -0.2217   0.2230   -0.994 0.32284
	% er_probes           -0.2308   0.2540   -0.909 0.36579
	% ifna_probes         0.1853    0.4957   0.374  0.70938
	% ifng_probes         -0.1528   0.5077   -0.301 0.76408
	% myc_probes          -0.0737   0.2284   -0.323 0.74765
	% pr_probes           -0.6876   0.2434   -2.825 0.00581  **
	% (Intercept)         1.056654  0.388317 2.721  0.00780  **
	% bmi                 0.002580  0.003803 0.678  0.49929
	% bcat_probes         -0.244104 0.226127 -1.079 0.28322
	% er_probes           -0.232713 0.254728 -0.914 0.36335
	% ifna_probes         0.227770  0.501056 0.455  0.65049
	% ifng_probes         -0.200460 0.513998 -0.390 0.69745
	% myc_probes          -0.096122 0.231435 -0.415 0.67888
	% pr_probes           -0.697585 0.244608 -2.852 0.00538  **
	% (Intercept)         1.09174   0.37821  2.887  0.00488  **
	% bmiStatusobese      0.10153   0.06573  1.545  0.12592
	% bmiStatusoverweight -0.00544  0.07119  -0.076 0.93926
	% bcat_probes         -0.27598  0.22517  -1.226 0.22351
	% er_probes           -0.20770  0.25218  -0.824 0.41234
	% ifna_probes         0.37360   0.50149  0.745  0.45823
	% ifng_probes         -0.33545  0.51257  -0.654 0.51449
	% myc_probes          -0.11868  0.22896  -0.518 0.60548
	% pr_probes           -0.68170  0.24150  -2.823 0.00586  **
	% (Intercept)         1.235747  0.391476 3.157  0.00218  **
	% bmi                 -0.008785 0.006537 -1.344 0.18237
	% bmiStatusobese      0.223518  0.111892 1.998  0.04881  *
	% bmiStatusoverweight 0.038776  0.078139 0.496  0.62094
	% bcat_probes         -0.272540 0.224177 -1.216 0.22730
	% er_probes           -0.184837 0.251634 -0.735 0.46455
	% ifna_probes         0.395907  0.499535 0.793  0.43015
	% ifng_probes         -0.335611 0.510285 -0.658 0.51243
	% myc_probes          -0.101582 0.228298 -0.445 0.65743
	% pr_probes           -0.647688 0.241749 -2.679 0.00879  **

	% [1] "carescrolgene"

	% Call:
	% lm(formula = as.formula(formula_txt[j]), data = data)

	% Coefficients:
	%              Estimate Std. Error t value Pr(>|t|)
	% (Intercept)         0.4928357 0.1239867 3.975  0.000136 ***
	% bmi                 0.0004131 0.0040740 0.101  0.919450
	% (Intercept)         0.52814   0.05483   9.633  9.13e-16 ***
	% bmiStatusobese      -0.06913  0.07045   -0.981 0.329
	% bmiStatusoverweight 0.02453   0.07753   0.316  0.752
	% (Intercept)         0.232845  0.163204  1.427  0.1569
	% bmi                 0.013351  0.006962  1.918  0.0582   .
	% bmiStatusobese      -0.253344 0.118558  -2.137 0.0352   *
	% bmiStatusoverweight -0.044356 0.084492  -0.525 0.6008
	% (Intercept)         -0.1205   0.3891    -0.310 0.7576
	% bcat_probes         0.2439    0.2275    1.072  0.2864
	% er_probes           0.1383    0.2590    0.534  0.5948
	% ifna_probes         -0.4364   0.5056    -0.863 0.3903
	% ifng_probes         0.4301    0.5178    0.831  0.4084
	% myc_probes          0.2837    0.2330    1.218  0.2264
	% pr_probes           0.5789    0.2483    2.331  0.0219   *
	% (Intercept)         -0.102321 0.396926  -0.258 0.7972
	% bmi                 -0.001038 0.003887  -0.267 0.7900
	% bcat_probes         0.252943  0.231140  1.094  0.2767
	% er_probes           0.139029  0.260375  0.534  0.5947
	% ifna_probes         -0.453502 0.512164  -0.885 0.3782
	% ifng_probes         0.449266  0.525393  0.855  0.3947
	% myc_probes          0.292758  0.236566  1.238  0.2191
	% pr_probes           0.582903  0.250031  2.331  0.0219   *
	% (Intercept)         -0.11188  0.38790   -0.288 0.7737
	% bmiStatusobese      -0.08019  0.06741   -1.190 0.2373
	% bmiStatusoverweight 0.01833   0.07302   0.251  0.8024
	% bcat_probes         0.28469   0.23093   1.233  0.2209
	% er_probes           0.11686   0.25864   0.452  0.6525
	% ifna_probes         -0.60153  0.51434   -1.170 0.2453
	% ifng_probes         0.59011   0.52570   1.123  0.2646
	% myc_probes          0.31782   0.23483   1.353  0.1793
	% pr_probes           0.57227   0.24768   2.311  0.0231   *
	% (Intercept)         -0.284628 0.400004  -0.712 0.4786
	% bmi                 0.010538  0.006679  1.578  0.1181
	% bmiStatusobese      -0.226528 0.114330  -1.981 0.0506   .
	% bmiStatusoverweight -0.034711 0.079842  -0.435 0.6648
	% bcat_probes         0.280559  0.229061  1.225  0.2239
	% er_probes           0.089437  0.257116  0.348  0.7288
	% ifna_probes         -0.628292 0.510417  -1.231 0.2216
	% ifng_probes         0.590302  0.521401  1.132  0.2606
	% myc_probes          0.297308  0.233271  1.275  0.2058
	% pr_probes           0.531475  0.247015  2.152  0.0341   *

	% [1] "crobsgenes"

	% Call:
	% lm(formula = as.formula(formula_txt[j]), data = data)

	% Coefficients:
	%             Estimate Std. Error t value Pr(>|t|)
	% (Intercept)         0.456126  0.123888 3.682  0.000381 ***
	% bmi                 0.001654  0.004071 0.406  0.685322
	% (Intercept)         0.459596  0.054606 8.417  3.7e-13  ***
	% bmiStatusobese      0.100893  0.070167 1.438  0.154
	% bmiStatusoverweight 0.005772  0.077224 0.075  0.941
	% (Intercept)         0.704210  0.163518 4.307  4.04e-05 ***
	% bmi                 -0.011059 0.006975 -1.586 0.1162
	% bmiStatusobese      0.253491  0.118786 2.134  0.0354   *
	% bmiStatusoverweight 0.062836  0.084654 0.742  0.4598
	% (Intercept)         1.0667    0.3806   2.803  0.00618  **
	% bcat_probes         -0.2284   0.2225   -1.026 0.30740
	% er_probes           -0.1481   0.2534   -0.584 0.56041
	% ifna_probes         0.4504    0.4945   0.911  0.36477
	% ifng_probes         -0.4013   0.5065   -0.792 0.43018
	% myc_probes          -0.1402   0.2279   -0.615 0.53986
	% pr_probes           -0.6446   0.2429   -2.654 0.00938  **
	% (Intercept)         1.014279  0.387058 2.620  0.01029  *
	% bmi                 0.003004  0.003791 0.792  0.43014
	% bcat_probes         -0.254482 0.225394 -1.129 0.26184
	% er_probes           -0.150261 0.253902 -0.592 0.55545
	% ifna_probes         0.499868  0.499432 1.001  0.31954
	% ifng_probes         -0.456812 0.512332 -0.892 0.37494
	% myc_probes          -0.166314 0.230685 -0.721 0.47278
	% pr_probes           -0.656157 0.243815 -2.691 0.00847  **
	% (Intercept)         1.05662   0.37693  2.803  0.0062   **
	% bmiStatusobese      0.11346   0.06550  1.732  0.0867   .
	% bmiStatusoverweight 0.01412   0.07095  0.199  0.8427
	% bcat_probes         -0.29209  0.22440  -1.302 0.1964
	% er_probes           -0.12674  0.25133  -0.504 0.6153
	% ifna_probes         0.63723   0.49979  1.275  0.2056
	% ifng_probes         -0.58273  0.51083  -1.141 0.2570
	% myc_probes          -0.19254  0.22819  -0.844 0.4010
	% pr_probes           -0.64072  0.24068  -2.662 0.0092   **
	% (Intercept)         1.198471  0.390241 3.071  0.00283  **
	% bmi                 -0.008654 0.006516 -1.328 0.18754
	% bmiStatusobese      0.233624  0.111539 2.095  0.03905  *
	% bmiStatusoverweight 0.057676  0.077893 0.740  0.46097
	% bcat_probes         -0.288700 0.223470 -1.292 0.19974
	% er_probes           -0.104217 0.250840 -0.415 0.67880
	% ifna_probes         0.659208  0.497958 1.324  0.18895
	% ifng_probes         -0.582884 0.508675 -1.146 0.25491
	% myc_probes          -0.175698 0.227577 -0.772 0.44214
	% pr_probes           -0.607216 0.240986 -2.520 0.01353  *

	% [1] "fmobsgenes"

	% Call:
	% lm(formula = as.formula(formula_txt[j]), data = data)

	% Coefficients:
	%             Estimate Std. Error t value Pr(>|t|)
	% (Intercept)         0.436628  0.123787 3.527  0.000643 ***
	% bmi                 0.002314  0.004067 0.569  0.570763
	% (Intercept)         0.58369   0.05435  10.740 <2e-16   ***
	% bmiStatusobese      -0.08452  0.06983  -1.210 0.2291
	% bmiStatusoverweight -0.14827  0.07686  -1.929 0.0567   .
	% (Intercept)         0.251170  0.160863 1.561  0.12176
	% bmi                 0.015034  0.006862 2.191  0.03091  *
	% bmiStatusobese      -0.291956 0.116858 -2.498 0.01419  *
	% bmiStatusoverweight -0.225840 0.083280 -2.712 0.00794  **
	% (Intercept)         0.33317   0.36900  0.903  0.3689
	% bcat_probes         -0.13489  0.21576  -0.625 0.5334
	% er_probes           0.11973   0.24567  0.487  0.6271
	% ifna_probes         -0.56244  0.47947  -1.173 0.2438
	% ifng_probes         0.58838   0.49109  1.198  0.2340
	% myc_probes          0.41768   0.22093  1.891  0.0618   .
	% pr_probes           -0.08815  0.23550  -0.374 0.7090
	% (Intercept)         0.291390  0.375705 0.776  0.4400
	% bmi                 0.002392  0.003679 0.650  0.5172
	% bcat_probes         -0.155669 0.218783 -0.712 0.4786
	% er_probes           0.117981  0.246455 0.479  0.6333
	% ifna_probes         -0.523056 0.484783 -1.079 0.2835
	% ifng_probes         0.544202  0.497304 1.094  0.2767
	% myc_probes          0.396892  0.223919 1.772  0.0797   .
	% pr_probes           -0.097371 0.236664 -0.411 0.6817
	% (Intercept)         0.33459   0.36629  0.913  0.3634
	% bmiStatusobese      -0.08029  0.06366  -1.261 0.2104
	% bmiStatusoverweight -0.12492  0.06895  -1.812 0.0733   .
	% bcat_probes         -0.07254  0.21807  -0.333 0.7402
	% er_probes           0.13036   0.24424  0.534  0.5948
	% ifna_probes         -0.56035  0.48569  -1.154 0.2517
	% ifng_probes         0.58760   0.49642  1.184  0.2397
	% myc_probes          0.46651   0.22175  2.104  0.0382   *
	% pr_probes           -0.07506  0.23389  -0.321 0.7490
	% (Intercept)         0.082260  0.370293 0.222  0.82471
	% bmi                 0.015393  0.006183 2.490  0.01464  *
	% bmiStatusobese      -0.294047 0.105838 -2.778 0.00666  **
	% bmiStatusoverweight -0.202402 0.073911 -2.738 0.00746  **
	% bcat_probes         -0.078569 0.212047 -0.371 0.71187
	% er_probes           0.090296  0.238018 0.379  0.70532
	% ifna_probes         -0.599445 0.472504 -1.269 0.20787
	% ifng_probes         0.587878  0.482673 1.218  0.22646
	% myc_probes          0.436540  0.215944 2.022  0.04623  *
	% pr_probes           -0.134650 0.228668 -0.589 0.55746

\begin{longtable}{llr{\bfseries}S}
	\centering
	\caption{Table showing the variables in the step wise linear models to predict the obesity metagenes}
	\label{tab:stepwise_model}\\
		Obesity metagene & Variables used in the model & Estimate & {P-value}\\
		\hline
	\endfirsthead
	\multicolumn{4}{c}{\tablename\ \thetable{}\ (continued)}\\
		\hline
		\hline
	\endhead
		\hline
		\hline
		Cr         & Akt         & -0.3415   & \bfseries \num{4.43d-8}\\
				   & \gls{egfr}  & 0.33852   & \bfseries \num{1.94d-05}\\
				   & \gls{er}    & -0.20237  & \bfseries 0.00419\\
				   & \gls{ifny}  & 0.12233   & 0.08700\\
				   & \gls{pr}    & -0.33920  & \bfseries \num{2.09d-5}\\
				   & Ras         & -0.22611  & \bfseries 0.00193\\
				   & Src         & -0.11452  & 0.08177\\
				   & \gls{tgfb}  & 0.55638   & \bfseries \textless{} \num{2d-16}\\
				   & \gls{tnfa}  & -0.25501  & \bfseries 0.00616\\
		\hline
		CrOl       & Overweight  & 0.040918  & 0.074865\\
				   & Obese       & -0.006381 & 0.756121\\
				   & Akt         & -0.356713 & \bfseries \num{1.89d-07}\\
				   & E2F1        & 0.160229  & \bfseries 0.005411\\
				   & \gls{egfr}  & 0.148618  & 0.062683\\
				   & p53         & 0.336552  & \bfseries \num{4.35d-05}\\
				   & \gls{pr}    & -0.225423 & \bfseries 0.000639\\
				   & Src         & -0.428758 & \bfseries \num{5.85d-10}\\
				   & \gls{stat3} & -0.124118 & \bfseries 0.009658\\
		\hline
		Res        & Akt         & 0.24613   & \bfseries 0.000289\\
				   & \gls{egfr}  & -0.46827  & \bfseries \num{2.16d-05}\\
				   & \gls{er}    & 0.18400   & 0.061446\\
				   & \gls{her2}  & 0.13835   & 0.118228\\
				   & \gls{ifny}  & -0.21019  & \bfseries 0.044084\\
				   & \gls{pr}    & 0.41195   & \bfseries 0.000300\\
				   & Ras         & 0.26506   & \bfseries 0.010465\\
				   & \gls{tgfb}  & -0.58075  & \bfseries \num{7.64d-11}\\
				   & \gls{tnfa}  & 0.42200   & \bfseries 0.001834\\
		\hline
		ResOl      & Akt         & -0.33596  & \bfseries 0.000294\\
				   & \gls{bcat}  & -0.19866  & 0.138953\\
				   & E2F1        & 0.36059   & \bfseries 0.009325\\
				   & \gls{egfr}  & 0.23421   & 0.056865\\
				   & \gls{her2}  & -0.13652  & 0.121689\\
				   & p53         & 0.38746   & \bfseries 0.000519\\
				   & \gls{pi3k}  & -0.21881  & 0.148676\\
				   & \gls{pr}    & -0.32228  & \bfseries 0.001656\\
				   & Ras         & -0.28984  & \bfseries 0.033975\\
				   & Src         & -0.40779  & \bfseries \num{5.26d-06}\\
				   & \gls{stat3} & -0.11452  & 0.125609\\
				   & \gls{tnfa}  & -0.11442  & 0.117023\\
		\hline
		Ca         & Akt         & -0.26243  & \bfseries \num{7.04d-06}\\
				   & \gls{egfr}  & 0.36305   & \bfseries 0.000139\\
				   & \gls{er}    & -0.18614  & \bfseries 0.030771\\
				   & \gls{ifny}  & 0.22118   & \bfseries 0.013034\\
				   & \gls{pr}    & -0.42911  & \bfseries \num{1.07d-05}\\
				   & Ras         & -0.28124  & \bfseries 0.000898\\
				   & \gls{tgfb}  & 0.76139   & \bfseries \textless{} \num{2d-16}\\
				   & \gls{tnfa}  & -0.33200  & \bfseries 0.003919\\
		\hline
		CaOl       & Overweight  & 0.07613   & \bfseries 0.006875\\
				   & Obese       & 0.01408   & 0.564622\\
				   & Akt         & -0.44727  & \bfseries \num{2.43d-07}\\
				   & \gls{bcat}  & -0.13463  & 0.140261\\
				   & E2F1        & 0.33864   & \bfseries 0.001249\\
				   & p53         & 0.28777   & \bfseries 0.000996\\
				   & \gls{pr}    & -0.28255  & \bfseries \num{2.18d-05}\\
				   & Src         & -0.44149  & \bfseries \num{2.02d-08}\\
				   & \gls{tnfa}  & -0.14215  & \bfseries 0.013897\\
		\hline
		CaRes      & \gls{bcat}  & 0.26899   & 0.05238  \\
				   & E2F1        & -0.22115  & 0.08848  \\
				   & \gls{egfr}  & 0.44400   & \bfseries \num{6.67d-05 }\\
				   & \gls{her2}  & -0.20304  & \bfseries 0.04615  \\
				   & \gls{ifny}  & 0.32874   & \bfseries 0.00574  \\
				   & \gls{pr}    & -0.28764  & \bfseries 0.00104  \\
				   & Src         & -0.17989  & 0.05404  \\
				   & \gls{tgfb}  & 0.52898   & \bfseries \num{2.15d-06 }\\
				   & \gls{tnfa}  & -0.38677  & \bfseries 0.00564  \\
		\hline
		CaResOl    & \gls{bmi}   & 0.004314  & 0.108546\\
				   & Overweight  & 0.033086  & 0.306390\\
				   & Obese       & -0.062257 & 0.175776\\
				   & Akt         & -0.309706 & \bfseries 0.000339\\
				   & \gls{bcat}  & -0.198165 & 0.140502\\
				   & E2F1        & 0.270471  & \bfseries 0.011450\\
				   & \gls{egfr}  & 0.171219  & 0.113691\\
				   & Myc         & 0.152007  & 0.056987\\
				   & p53         & 0.294895  & \bfseries 0.003734\\
				   & \gls{pr}    & -0.230896 & \bfseries 0.000823\\
				   & Ras         & -0.270824 & \bfseries 0.044788\\
				   & Src         & -0.512567 & \bfseries \num{2.72d-08}\\
				   & \gls{tnfa}  & -0.103302 & 0.083449\\
		\hline
		Original   & \gls{bmi}   & 0.002696  & 0.14407\\
				   & Overweight  & 0.010522  & 0.62856\\
				   & Obese       & -0.061020 & 0.05542\\
				   & Akt         & -0.238412 & \bfseries \num{4.93d-05}\\
				   & E2F1        & 0.185085  & \bfseries 0.02754\\
				   & \gls{egfr}  & 0.147219  & 0.07423\\
				   & \gls{her2}  & -0.086804 & 0.14070\\
				   & p53         & 0.416414  & \bfseries \num{5.83d-08}\\
				   & \gls{pi3k}  & -0.216179 & \bfseries 0.03528\\
				   & \gls{pr}    & -0.199097 & \bfseries 0.00253\\
				   & Ras         & -0.200544 & \bfseries 0.00898\\
				   & Src         & -0.543834 & \bfseries \num{2.10d-15}\\
				   & \gls{stat3} & -0.093363 & \bfseries 0.02749\\
		\hline
		FM         & \gls{bmi}   & 0.011093  & \bfseries 0.036229\\
				   & Overweight  & -0.153753 & \bfseries 0.014277\\
				   & Obese       & -0.212483 & \bfseries 0.017378\\
				   & p53         & 0.460800  & \bfseries 0.002512\\
				   & \gls{pi3k}  & -0.426246 & \bfseries 0.000118\\
				   & \gls{pr}    & -0.373119 & \bfseries 0.020737\\
				   & \gls{stat3} & 0.248745  & \bfseries 0.019073\\
				   & \gls{tgfb}  & 0.152925  & 0.128933\\
		\hline
		\hline
	\end{longtable}

\end{appendices}

### 2017-03-13 thesis/misc/appendix.tex
Divided the appendix file into two - tmp commit
Need to add linear model prediction scatter plots - tmp commit

--- 2017-03-10 thesis/misc/appendix.tex
@@
 
-	\section{Metagenes created from raw data and standardised data (ICGC)}
-	\label{sec:metagenes_created_from_raw_data_vs_standardised_data_icgc}
+	\section{Comparison of the Creighton \textit{et al.} obesity metagene in standardised or non-standardised \gls{icgc} data}
+	\label{sec:comp_cr_raw_std_icgc}
 
-	% \begin{figure}[h]
-	% 	\centering
-	% 	\includegraphics[width=0.45\linewidth,page=1]{appendix/crtcga_raw_vs_std}
-	% 	\hfill
-	% 	\includegraphics[width=0.45\linewidth,page=2]{appendix/crtcga_raw_vs_std}
-	% 	\caption{appendix/Check_raw_vs_std}
-	% 	\label{fig:appendix/check_raw_vs_std}
-	% \end{figure}
+	As shown in \cref{fig:appendix/check_raw_vs_std}, the Creighton \textit{et al.} obesity metagene values generated from the transformation matrices (from standardised or non-standardised CR data) were more consistent (or less variable) in the standardised \gls{icgc} \gls{blca} data, compared with the non-standardised data.
+	Results from \cref{fig:appendix/check_raw_vs_std_tm} suggested that there was no obvious difference in the quality of the metagenes generated from standardised or non-standardised TM.
+	Only the results from \gls{blca} data set are shown, as the results from the other \gls{icgc} cancer types showed very similar results.
 
-	\section{Rest of the ICGC cancer heatmap resutls}
-	\label{sec:rest_of_the_icgc_cancer_heatmap_resutls}
+	\begin{figure}[h]
+		\centering
+		\includegraphics[width=0.45\linewidth,page=1]{appendix/crtcga_raw_vs_std}
+		\hfill
+		\includegraphics[width=0.45\linewidth,page=2]{appendix/crtcga_raw_vs_std}\\
+		\caption{Scatter plots comparing the Creighton \textit{et al.} obesity metagene generated from transformation matrix (from standardised or non-standardised CR data) in non-standardised (left) and standardised (right) \gls{icgc} \gls{blca} data.}
+		\label{fig:appendix/check_raw_vs_std}
+	\end{figure}
+
+	\begin{figure}[h]
+		\centering
+		\includegraphics[width=0.45\linewidth,page=3]{appendix/crtcga_raw_vs_std}
+		\hfill
+		\includegraphics[width=0.45\linewidth,page=4]{appendix/crtcga_raw_vs_std}\\
+		\caption{Scatter plots comparing the Creighton \textit{et al.} obesity metagene generated from transformation matrix from non-standardisedright (left) or standardised (right) CR data in non-standardised  and standardised  \gls{icgc} \gls{blca} data.}
+		\label{fig:appendix/check_raw_vs_std_tm}
+	\end{figure}
+
+	\section{Remainder of the results of Creighton \textit{et al.} obesity metagenes in the \gls{icgc} cancer data sets}
+	\label{sec:rest_of_the_cr_icgc_cancer_heatmap_results}
+
+	As shown in \cref{fig:appendix/cr_icgc_heatmap}, the original obesity metagene from the Creighton \textit{et al.} study was reflective of the sample gene expression levels of the obesity associated genetic signature in all of the other \gls{icgc} cancer data sets.
+	However, none of the \gls{icgc} cancer data sets (other than the \gls{blca} data set; \cref{fig:crmetaicgc} in \cref{sec:creighton_obesity_metagene}) showed any significant association of the sample \gls{bmi} or \gls{bmi} status with the Or obesity metagene (\cref{fig:appendix/cr_icgc_box_scatter}).
+	The heatmaps, box plots and scatter plots for the other CR obesity metagenes in other \gls{icgc} cancer types are not shown, as many of the plots were very similar to the plots from the \gls{blca} data set.
+	The statistics from the results of the box and scatter plots in the other \gls{icgc} cancer types were summarised in \cref{tab:degmetacesc,tab:degmetacoad,tab:degmetakirp,tab:degmetalihc,tab:degmetaread,tab:degmetaskcm,tab:degmetaucec}.
+
+	\begin{figure}[htp!]
+		\centering
+		\includegraphics[width=0.32\linewidth,page=8]{appendix/crtcga_std}\\
+		\vspace{1em}
+		\includegraphics[width=0.32\linewidth,page=13]{appendix/crtcga_std}
+		\includegraphics[width=0.32\linewidth,page=18]{appendix/crtcga_std}
+		\includegraphics[width=0.32\linewidth,page=23]{appendix/crtcga_std}\\
+		\vspace{1em}
+		\includegraphics[width=0.32\linewidth,page=28]{appendix/crtcga_std}
+		\includegraphics[width=0.32\linewidth,page=33]{appendix/crtcga_std}
+		\includegraphics[width=0.32\linewidth,page=38]{appendix/crtcga_std}\\
+		\caption{Heatmaps showing the association of obesity metagene from the \citet{Creighton2012} study with sample gene expression from the other \gls{icgc} data sets.
+	Level of expression is represented in the top right histogram, where low and high gene expression were colour-coded with blue and red, respectively.
+	Each row of the heatmap represents a gene from the obesity associated genetic signature, and each column of the heatmap represents a sample from the \gls{icgc} data.
+	The obesity associated metagene scores of the samples are shown in a separate row at top of the heatmap, and the tree diagram of the heirarchical clustering of the genes is shown to the left of the heatmap.
+}
+		\label{fig:appendix/cr_icgc_heatmap}
+	\end{figure}
 
@@
 		\centering
-		\includegraphics[width=0.45\linewidth,page=1]{appendix/crtcga_std}
+		\includegraphics[page=9,width=0.45\linewidth]{appendix/crtcga_std}
 		\hfill
-		\includegraphics[width=0.45\linewidth,page=2]{appendix/crtcga_std}
-		\caption{appendix/Check_raw_vs_std}
-		\label{fig:appendix/check_raw_vs_std}
+		\includegraphics[page=10,width=0.45\linewidth]{appendix/crtcga_std}\\
+		\includegraphics[page=14,width=0.45\linewidth]{appendix/crtcga_std}
+		\hfill
+		\includegraphics[page=15,width=0.45\linewidth]{appendix/crtcga_std}\\
+		\caption{Box plots and scatter plots showing the association of obesity metagene from the \citet{Creighton2012} study with the sample \gls{bmi} status  and \gls{bmi} from the other \gls{icgc} data sets, respectively.
+	In the box plot, the p-values above the groups represent the statistical significance of the association of the metagene with the overweight or obese group compared with the normal weight group.
+	The \gls{anova} p-value shows the statistical significance of the association of the metagene with the sample \gls{bmi} groups.
+	In the scatter plot, $R^2$- and p-values describe the adjusted coefficient of determination of the regression line and the statistical significance of the linear model used to draw the regression line, respectively.}
+		\label{fig:appendix/cr_icgc_box_scatter}
 	\end{figure}
 
-	\section{FM metagene, raw vs std}
-	\label{sec:fm_metagene_raw_vs_std}
+	\begin{figure}[htpb]
+		\ContinuedFloat
+		\captionsetup{list=off,format=cont}
+		\centering
+		\includegraphics[page=19,width=0.45\linewidth]{appendix/crtcga_std}
+		\hfill
+		\includegraphics[page=20,width=0.45\linewidth]{appendix/crtcga_std}\\
+		\includegraphics[page=24,width=0.45\linewidth]{appendix/crtcga_std}
+		\hfill
+		\includegraphics[page=25,width=0.45\linewidth]{appendix/crtcga_std}\\
+		\includegraphics[page=29,width=0.45\linewidth]{appendix/crtcga_std}
+		\hfill
+		\includegraphics[page=30,width=0.45\linewidth]{appendix/crtcga_std}\\
+		\caption{Name}
+	\end{figure}
+
+	\begin{figure}[htpb]
+		\ContinuedFloat
+		\captionsetup{list=off,format=cont}
+		\centering
+		\includegraphics[page=34,width=0.45\linewidth]{appendix/crtcga_std}
+		\hfill
+		\includegraphics[page=35,width=0.45\linewidth]{appendix/crtcga_std}\\
+		\includegraphics[page=39,width=0.45\linewidth]{appendix/crtcga_std}
+		\hfill
+		\includegraphics[page=40,width=0.45\linewidth]{appendix/crtcga_std}\\
+		\caption{Name}
+	\end{figure}
+
+\begin{table}[htpb]
+	\centering
+	\caption{Statistics of all the obesity metagenes in \gls{cesc} cancer data}
+	\label{tab:degmetacesc}
+	\begin{threeparttable}
+		\begin{tabular}{lccccc}
+			& \multicolumn{3}{c}{ P-values} & \multicolumn{2}{c}{ Regression line statistics}\\
+			\cmidrule(r){2-4} \cmidrule(r){5-6}
+			Metagenes &  Overweight &  Obese &  \gls{anova} &  R$^2$ &  P \\
+			\hline
+			\hline
+			\rule{0pt}{2.25ex}Cr      & 0.9332                      & 0.5302  & 0.7622             & -0.003950  & 0.7265              \\
+            Res     & 0.9205                      & 0.3369  & 0.6043             & 0.004249   & 0.8122              \\
+            CrOl    & 0.8940                      & 0.1430  & 0.2959             & -0.0004927 & 0.3465              \\
+            ResOl   & 0.9930                      & 0.1705  & 0.3078             & 0.003263   & 0.1898              \\
+            Ca      & 0.8861                      & 0.3616  & 0.6426             & -0.003307  & 0.6073              \\
+            CaRes   & 0.8323                      & 0.3330  & 0.6167             & -0.003832  & 0.7001              \\
+            CaOl    & 0.9124                      & 0.1602  & 0.2600             & -0.001612  & 0.4242              \\
+            CaResOl & 0.9646                      & 0.1663  & 0.2863             & 0.003097   & 0.1946              \\
+			\hline
+			\hline
+		\end{tabular}
+	\end{threeparttable}
+\end{table}
+
+\begin{table}[htpb]
+	\centering
+	\caption{Statistics of all the obesity metagenes in \gls{coad} cancer data}
+	\label{tab:degmetacoad}
+	\begin{threeparttable}
+		\begin{tabular}{lccccc}
+			& \multicolumn{3}{c}{ P-values} & \multicolumn{2}{c}{ Regression line statistics}\\
+			\cmidrule(r){2-4} \cmidrule(r){5-6}
+			Metagenes &  Overweight &  Obese &  \gls{anova} &  R$^2$ &  P \\
+			\hline
+			\hline
+			\rule{0pt}{2.25ex}Cr      & 0.2843                      & 0.8176  & 0.5460             & -0.004373  & 0.8865              \\
+            Res     & 0.4091                      & 0.8473  & 0.7064             & -0.004243  & 0.8245              \\
+            CrOl    & 0.4063                      & 0.9295  & 0.6150             & -0.00397   & 0.7400              \\
+            ResOl   & 0.3748                      & 0.8484  & 0.5295             & -0.004020  & 0.7532              \\
+            Ca      & 0.6507                      & 0.7692  & 0.7579             & -0.002493  & 0.5075              \\
+            CaRes   & 0.4565                      & 0.9808  & 0.7161             & -0.004278  & 0.8385              \\
+            CaOl    & 0.2931                      & 0.9217  & 0.4621             & -0.00425   & 0.8273              \\
+            CaResOl & 0.4203                      & 0.9087  & 0.7017             & -0.004461  & 0.9800              \\
+			\hline
+			\hline
+		\end{tabular}
+	\end{threeparttable}
+\end{table}
+
+\begin{table}[htpb]
+	\centering
+	\caption{Statistics of all the obesity metagenes in \gls{kirp} cancer data}
+	\label{tab:degmetakirp}
+	\begin{threeparttable}
+		\begin{tabular}{lccccc}
+			& \multicolumn{3}{c}{ P-values} & \multicolumn{2}{c}{ Regression line statistics}\\
+			\cmidrule(r){2-4} \cmidrule(r){5-6}
+			Metagenes &  Overweight &  Obese &  \gls{anova} &  R$^2$ &  P \\
+			\hline
+			\hline
+			\rule{0pt}{2.25ex}Cr      & 0.1853                      & 0.2590  & 0.3387             & -0.005657  & 0.5798              \\
+            Res     & 0.1063                      & 0.1825  & 0.2128             & -0.005037  & 0.5368              \\
+            CrOl    & 0.2576                      & 0.1757  & 0.3162             & -0.005763  & 0.5879              \\
+            ResOl   & 0.3120                      & 0.1738  & 0.3198             & -0.004513  & 0.5048              \\
+            Ca      & 0.2431                      & 0.2142  & 0.3591             & -0.005582  & 0.5743              \\
+            CaRes   & 0.2327                      & 0.2344  & 0.3738             & -0.006250  & 0.6279              \\
+            CaOl    & 0.3426                      & 0.2775  & 0.4693             & -0.007333  & 0.77469             \\
+            CaResOl & 0.2643                      & 0.1315  & 0.2410             & -0.001408  & 0.3649              \\
+			\hline
+			\hline
+		\end{tabular}
+	\end{threeparttable}
+\end{table}
+
+\begin{table}[htpb]
+	\centering
+	\caption{Statistics of all the obesity metagenes in \gls{lihc} cancer data}
+	\label{tab:degmetalihc}
+	\begin{threeparttable}
+		\begin{tabular}{lccccc}
+			& \multicolumn{3}{c}{ P-values} & \multicolumn{2}{c}{ Regression line statistics}\\
+			\cmidrule(r){2-4} \cmidrule(r){5-6}
+			Metagenes &  Overweight &  Obese &  \gls{anova} &  R$^2$ &  P \\
+			\hline
+			\hline
+			\rule{0pt}{2.25ex}Cr      & 0.4033                      & 0.2900  & 0.5221             & 0.006348   & 0.1028              \\
+            Res     & 0.5008                      & 0.8743  & 0.8004             & -0.0003709 & 0.3430              \\
+            CrOl    & 0.7599                      & 0.8562  & 0.9189             & -0.001554  & 0.4424              \\
+            ResOl   & 0.9238                      & 0.9773  & 0.9938             & -0.001283  & 0.4162              \\
+            Ca      & 0.4898                      & 0.8221  & 0.7917             & 0.0005899  & 0.2834              \\
+            CaRes   & 0.5976                      & 0.9929  & 0.8638             & -0.001090  & 0.3990              \\
+            CaOl    & 0.8717                      & 0.8834  & 0.9679             & -0.002010  & 0.4925              \\
+            CaResOl & 0.5369                      & 0.7185  & 0.7073             & -0.0004212 & 0.3465              \\
+			\hline
+			\hline
+		\end{tabular}
+	\end{threeparttable}
+\end{table}
+
+\begin{table}[htpb]
+	\centering
+	\caption{Statistics of all the obesity metagenes in \gls{read} cancer data}
+	\label{tab:degmetaread}
+	\begin{threeparttable}
+		\begin{tabular}{lccccc}
+			& \multicolumn{3}{c}{ P-values} & \multicolumn{2}{c}{ Regression line statistics}\\
+			\cmidrule(r){2-4} \cmidrule(r){5-6}
+			Metagenes &  Overweight &  Obese &  \gls{anova} &  R$^2$ &  P \\
+			\hline
+			\hline
+			\rule{0pt}{2.25ex}Cr      & 0.9129                      & 0.4008  & 0.5942             & -0.0132    & 0.8038              \\
+            Res     & 0.7737                      & 0.2244  & 0.4308             & -0.008917  & 0.5484              \\
+            CrOl    & 0.8439                      & 0.1283  & 0.2643             & 0.001024   & 0.3036              \\
+            ResOl   & 0.6953                      & 0.06841 & 0.1766             & 0.01099    & 0.1840              \\
+            Ca      & 0.6298                      & 0.1326  & 0.2960             & -0.001242  & 0.3432              \\
+            CaRes   & 0.6302                      & 0.1053  & 0.2453             & 0.001989   & 0.2885              \\
+            CaOl    & 0.6871                      & 0.1411  & 0.3358             & -0.004847  & 0.4219              \\
+            CaResOl & 0.8821                      & 0.08474 & 0.1817             & 0.001095   & 0.3025              \\
+			\hline
+			\hline
+		\end{tabular}
+	\end{threeparttable}
+\end{table}
+
+\begin{table}[htpb]
+	\centering
+	\caption{Statistics of all the obesity metagenes in \gls{skcm} cancer data}
+	\label{tab:degmetaskcm}
+	\begin{threeparttable}
+		\begin{tabular}{lccccc}
+			& \multicolumn{3}{c}{ P-values} & \multicolumn{2}{c}{ Regression line statistics}\\
+			\cmidrule(r){2-4} \cmidrule(r){5-6}
+			Metagenes &  Overweight &  Obese &  \gls{anova} &  R$^2$ &  P \\
+			\hline
+			\hline
+			\rule{0pt}{2.25ex}Cr      & 0.7115                      & 0.3329  & 0.6264             & 0.002887   & 0.2033              \\
+            Res     & 0.7604                      & 0.5408  & 0.6396             & -0.001725  & 0.4296              \\
+            CrOl    & 0.4719                      & 0.1224  & 0.2913             & 0.01261    & 0.05347             \\
+            ResOl   & 0.8640                      & 0.1668  & 0.3051             & 0.01541    & {\bfseries 0.0372}\tnote{1}  \\
+            Ca      & 0.8792                      & 0.2084  & 0.3739             & 0.006674   & 0.1184              \\
+            CaRes   & 0.8098                      & 0.1950  & 0.3677             & 0.005746   & 0.1347              \\
+            CaOl    & 0.9478                      & 0.2726  & 0.4471             & 0.01009    & 0.07449             \\
+            CaResOl & 0.7309                      & 0.2561  & 0.2830             & 0.01606    & {\bfseries 0.03421} \\
+			\hline
+			\hline
+		\end{tabular}
+		\begin{tablenotes}
+			\item [1] All values in bold are statistically significant (p \textless{} 0.05).
+		\end{tablenotes}
+	\end{threeparttable}
+\end{table}
+
+\begin{table}[htpb]
+	\centering
+	\caption{Statistics of all the obesity metagenes in \gls{ucec} cancer data}
+	\label{tab:degmetaucec}
+	\begin{threeparttable}
+		\begin{tabular}{lccccc}
+			& \multicolumn{3}{c}{ P-values} & \multicolumn{2}{c}{ Regression line statistics}\\
+			\cmidrule(r){2-4} \cmidrule(r){5-6}
+			Metagenes &  Overweight &  Obese &  \gls{anova} &  R$^2$ &  P \\
+			\hline
+			\hline
+			\rule{0pt}{2.25ex}Cr      & 0.8357                      & 0.08117 & 0.1039             & 0.006952   & {\bfseries 0.03716}\tnote{1}\\
+            Res     & 0.8664                      & 0.2023  & 0.3131             & 0.001998   & 0.1619              \\
+            CrOl    & 0.9251                      & 0.3800  & 0.5800             & -0.000311  & 0.3569              \\
+            ResOl   & 0.5826                      & 0.09608 & 0.2156             & 0.003487   & 0.1021              \\
+            Ca      & 0.7491                      & 0.1103  & 0.1905             & 0.003811   & 0.09260             \\
+            CaRes   & 0.8086                      & 0.2002  & 0.3419             & 0.0005824  & 0.2584              \\
+            CaOl    & 0.9068                      & 0.2672  & 0.4153             & 0.00305    & 0.1166              \\
+            CaResOl & 0.8360                      & 0.2440  & 0.4034             & 0.00022898 & 0.2863              \\
+			\hline
+			\hline
+		\end{tabular}
+		\begin{tablenotes}
+			\item [1] All values in bold are statistically significant (p \textless{} 0.05).
+		\end{tablenotes}
+	\end{threeparttable}
+\end{table}
+
+	\section{Remainder of the results of the other CR obesity metagenes in \gls{nzbc} data set}
+	\label{sec:rest_of_the_cr_ob_meta_heatmap_results_cris}
+
+	All of the other obesity metagenes from the CR data set were able to capture the overall gene expressions of the samples in the \gls{nzbc} data set (\cref{fig:appendix/cr_ob_meta_heatmap_cris}).
+	However, none of the CR obesity metagenes were able to significantly associate with the sample \gls{bmi}/\gls{bmi} status in the \gls{nzbc} data set (\cref{fig:appendix/cr_ob_meta_box_scatter_cris}).
+
+	\begin{figure}[htp!]
+		\centering
+		\includegraphics[width=0.32\linewidth,page=8]{appendix/cris_crdeg_trans_meta}\\
+		\includegraphics[width=0.32\linewidth,page=13]{appendix/cris_crdeg_trans_meta}
+		\includegraphics[width=0.32\linewidth,page=18]{appendix/cris_crdeg_trans_meta}
+		\includegraphics[width=0.32\linewidth,page=23]{appendix/cris_crdeg_trans_meta}\\
+		\includegraphics[width=0.32\linewidth,page=28]{appendix/cris_crdeg_trans_meta}
+		\includegraphics[width=0.32\linewidth,page=33]{appendix/cris_crdeg_trans_meta}
+		\includegraphics[width=0.32\linewidth,page=38]{appendix/cris_crdeg_trans_meta}\\
+		\caption{Heatmaps showing the association of the other CR obesity metagenes with the sample gene expression from the \gls{nzbc} data set.
+	Scales are as described in previous figures.}
+		\label{fig:appendix/cr_ob_meta_heatmap_cris}
+	\end{figure}
 
@@
 		\centering
-		\includegraphics[page=1,width=0.8\linewidth]{appendix/fm_meta_check}
-		\caption{appendix/fm_meta_check}
-		\label{fig:appendix/fm_meta_check}
+		\includegraphics[page=9,width=0.42\linewidth]{appendix/cris_crdeg_trans_meta}
+		\hfill
+		\includegraphics[page=10,width=0.42\linewidth]{appendix/cris_crdeg_trans_meta}\\
+		\caption{Box plots and scatter plots showing the association of the other CR obesity metagenes with the sample \gls{bmi} status  and \gls{bmi} from the \gls{nzbc} data set, respectively.
+	P-values and $R^2$-value are as described in previous figures.}
+		\label{fig:appendix/cr_ob_meta_box_scatter_cris}
 	\end{figure}
 
-	\section{Creighton DEG metagene direction}
-	\label{sec:creighton_deg_metagene_direction}
+	\begin{figure}[htpb]
+		\ContinuedFloat
+		\captionsetup{list=off,format=cont}
+		\centering
+		\includegraphics[page=14,width=0.45\linewidth]{appendix/cris_crdeg_trans_meta}
+		\hfill
+		\includegraphics[page=15,width=0.45\linewidth]{appendix/cris_crdeg_trans_meta}\\
+		\includegraphics[page=19,width=0.45\linewidth]{appendix/cris_crdeg_trans_meta}
+		\hfill
+		\includegraphics[page=20,width=0.45\linewidth]{appendix/cris_crdeg_trans_meta}\\
+		\includegraphics[page=24,width=0.45\linewidth]{appendix/cris_crdeg_trans_meta}
+		\hfill
+		\includegraphics[page=25,width=0.45\linewidth]{appendix/cris_crdeg_trans_meta}\\
+		\caption{Name}
+	\end{figure}
 
 	\begin{figure}[htpb]
+		\ContinuedFloat
+		\captionsetup{list=off,format=cont}
 		\centering
-		\includegraphics[page=1,width=0.8\linewidth]{cr_meta_direction}
-		\caption{cr_meta_direction}
-		\label{fig:cr_meta_direction}
+		\includegraphics[page=29,width=0.45\linewidth]{appendix/cris_crdeg_trans_meta}
+		\hfill
+		\includegraphics[page=30,width=0.45\linewidth]{appendix/cris_crdeg_trans_meta}\\
+		\includegraphics[page=34,width=0.45\linewidth]{appendix/cris_crdeg_trans_meta}
+		\hfill
+		\includegraphics[page=35,width=0.45\linewidth]{appendix/cris_crdeg_trans_meta}\\
+		\includegraphics[page=39,width=0.45\linewidth]{appendix/cris_crdeg_trans_meta}
+		\hfill
+		\includegraphics[page=40,width=0.45\linewidth]{appendix/cris_crdeg_trans_meta}\\
+		\caption{Name}
 	\end{figure}
-
-	\section{Number of gene probes/genes in each signature}
-	\label{app:number_of_gene_probes_genes_in_each_signature}
-
-	\begin{table}[htbp]
-		\centering
-		\caption{Summary of the number of gene probes and genes in each obesity associated genetic signature}
-		\label{tab:signature_gene_num}
-		\begin{tabular}{lcc}
-			Obesity associated genetic signature & No. gene probes   & No. genes\\
-			\hline
-			\rule{0pt}{2.25ex}Original & 799 & 644\\
-			Cr                         & 799 & 678\\
-			Res                        & 799 & 655\\
-			Ca                         & 799 & 657\\
-			CaRes                      & 799 & 651\\
-			CrOl                       & 239 & 199\\
-			ResOl                      & 168 & 147\\
-			CaOl                       & 148 & 128\\
-			CaResOl                    & 92  & 86\\
-			\hline
-			\hline
-		\end{tabular}
-	\end{table}
-
-	\chapter{Additional results from \cref{cha:obesity_associated_genetic_signature_and_pathway_signatures}}
-	\label{app:b}
-
-	\begin{figure}[htpb]
-		\centering
-		\includegraphics[width=0.7\linewidth]{results2/gatzares1}
-		\caption{results from gatza's study}
-		\label{fig:gatza_paper_res}
-	\end{figure}
-
-% [1] "crolgene"
-
-% Call:
-% lm(formula = as.formula(formula_txt[j]), data = data)
-
-% Coefficients:
-%              Estimate Std. Error t value Pr(>|t|)
-% (Intercept)         0.553214  0.123891 4.465  2.16e-05 ***
-% bmi                 -0.001629 0.004071 -0.400 0.69
-% (Intercept)         0.55700   0.05464  10.194 <2e-16   ***
-% bmiStatusobese      -0.10551  0.07021  -1.503 0.136
-% bmiStatusoverweight -0.02165  0.07727  -0.280 0.780
-% (Intercept)         0.307513  0.163530 1.880  0.0631   .
-% bmi                 0.011279  0.006976 1.617  0.1092
-% bmiStatusobese      -0.261145 0.118795 -2.198 0.0304   *
-% bmiStatusoverweight -0.079845 0.084661 -0.943 0.3480
-% (Intercept)         0.007141  0.382725 0.019  0.9852
-% bcat_probes         0.156062  0.223779 0.697  0.4873
-% er_probes           0.123403  0.254800 0.484  0.6293
-% ifna_probes         -0.430229 0.497301 -0.865 0.3892
-% ifng_probes         0.378437  0.509352 0.743  0.4594
-% myc_probes          0.160785  0.229142 0.702  0.4846
-% pr_probes           0.597403  0.244252 2.446  0.0164   *
-% (Intercept)         0.055088  0.389472 0.141  0.8878
-% bmi                 -0.002745 0.003814 -0.720 0.4736
-% bcat_probes         0.179900  0.226799 0.793  0.4297
-% er_probes           0.125413  0.255485 0.491  0.6247
-% ifna_probes         -0.475417 0.502546 -0.946 0.3466
-% ifng_probes         0.429131  0.515526 0.832  0.4074
-% myc_probes          0.184642  0.232124 0.795  0.4284
-% pr_probes           0.607982  0.245336 2.478  0.0151   *
-% (Intercept)         0.01666   0.37989  0.044  0.9651
-% bmiStatusobese      -0.11197  0.06602  -1.696 0.0933   .
-% bmiStatusoverweight -0.02348  0.07151  -0.328 0.7434
-% bcat_probes         0.22035   0.22616  0.974  0.3325
-% er_probes           0.10450   0.25330  0.413  0.6809
-% ifna_probes         -0.60343  0.50372  -1.198 0.2341
-% ifng_probes         0.54671   0.51484  1.062  0.2911
-% myc_probes          0.21341   0.22998  0.928  0.3559
-% pr_probes           0.59491   0.24257  2.453  0.0161   *
-% (Intercept)         -0.128149 0.393204 -0.326 0.7453
-% bmi                 0.008834  0.006565 1.346  0.1819
-% bmiStatusobese      -0.234634 0.112386 -2.088 0.0397   *
-% bmiStatusoverweight -0.067942 0.078484 -0.866 0.3890
-% bcat_probes         0.216892  0.225167 0.963  0.3380
-% er_probes           0.081506  0.252745 0.322  0.7478
-% ifna_probes         -0.625864 0.501740 -1.247 0.2155
-% ifng_probes         0.546873  0.512538 1.067  0.2889
-% myc_probes          0.196215  0.229305 0.856  0.3945
-% pr_probes           0.560706  0.242816 2.309  0.0232   *
-
-% [1] "resobsgene"
-
-% Call:
-% lm(formula = as.formula(formula_txt[j]), data = data)
-
-% Coefficients:
-%              Estimate Std. Error t value Pr(>|t|)
-% (Intercept)         0.547104  0.123915 4.415  2.62e-05 ***
-% bmi                 -0.001422 0.004072 -0.349 0.728
-% (Intercept)         0.547619  0.054641 10.022 <2e-16   ***
-% bmiStatusobese      -0.096832 0.070212 -1.379 0.171
-% bmiStatusoverweight -0.001804 0.077273 -0.023 0.981
-% (Intercept)         0.298589  0.163546 1.826  0.0710   .
-% bmi                 0.011259  0.006976 1.614  0.1099
-% bmiStatusobese      -0.252186 0.118807 -2.123 0.0364   *
-% bmiStatusoverweight -0.059898 0.084669 -0.707 0.4810
-% (Intercept)         0.18694   0.36963  0.506  0.6142
-% bcat_probes         0.09717   0.21612  0.450  0.6541
-% er_probes           0.07879   0.24608  0.320  0.7496
-% ifna_probes         -0.17576  0.48029  -0.366 0.7152
-% ifng_probes         0.07930   0.49193  0.161  0.8723
-% myc_probes          -0.04517  0.22130  -0.204 0.8387
-% pr_probes           0.59554   0.23590  2.525  0.0133   *
-% (Intercept)         0.221757  0.376612 0.589  0.5574
-% bmi                 -0.001994 0.003688 -0.540 0.5902
-% bcat_probes         0.114479  0.219311 0.522  0.6029
-% er_probes           0.080247  0.247050 0.325  0.7461
-% ifna_probes         -0.208580 0.485954 -0.429 0.6688
-% ifng_probes         0.116119  0.498505 0.233  0.8163
-% myc_probes          -0.027845 0.224460 -0.124 0.9015
-% pr_probes           0.603222  0.237235 2.543  0.0127   *
-% (Intercept)         0.195965  0.367384 0.533  0.5951
-% bmiStatusobese      -0.096834 0.063844 -1.517 0.1328
-% bmiStatusoverweight -0.004412 0.069155 -0.064 0.9493
-% bcat_probes         0.150379  0.218719 0.688  0.4935
-% er_probes           0.058880  0.244961 0.240  0.8106
-% ifna_probes         -0.344126 0.487134 -0.706 0.4817
-% ifng_probes         0.242697  0.497893 0.487  0.6271
-% myc_probes          -0.001285 0.222407 -0.006 0.9954
-% pr_probes           0.591191  0.234582 2.520  0.0135   *
-% (Intercept)         0.041459  0.379420 0.109  0.9132
-% bmi                 0.009426  0.006335 1.488  0.1403
-% bmiStatusobese      -0.227717 0.108446 -2.100 0.0386   *
-% bmiStatusoverweight -0.051853 0.075733 -0.685 0.4953
-% bcat_probes         0.146687  0.217273 0.675  0.5013
-% er_probes           0.034350  0.243884 0.141  0.8883
-% ifna_probes         -0.368065 0.484150 -0.760 0.4491
-% ifng_probes         0.242870  0.494570 0.491  0.6246
-% myc_probes          -0.019634 0.221267 -0.089 0.9295
-% pr_probes           0.554701  0.234304 2.367  0.0201   *
-
-% [1] "rescrolgene"
-
-% Call:
-% lm(formula = as.formula(formula_txt[j]), data = data)
-
-% Coefficients:
-%               Estimate Std. Error t value Pr(>|t|)
-% (Intercept)         0.5343899  0.1239554 4.311  3.9e-05 ***
-% bmi                 -0.0009922 0.0040730 -0.244 0.808
-% (Intercept)         0.54870    0.05482   10.009 <2e-16  ***
-% bmiStatusobese      -0.09063   0.07044   -1.287 0.201
-% bmiStatusoverweight -0.01515   0.07753   -0.195 0.845
-% (Intercept)         0.304200   0.164179  1.853  0.0670  .
-% bmi                 0.011054   0.007003  1.578  0.1178
-% bmiStatusobese      -0.243160  0.119267  -2.039 0.0442  *
-% bmiStatusoverweight -0.072189  0.084997  -0.849 0.3978
-% (Intercept)         0.17069    0.37929   0.450  0.6537
-% bcat_probes         0.09532    0.22177   0.430  0.6683
-% er_probes           0.03171    0.25251   0.126  0.9004
-% ifna_probes         -0.57965   0.49283   -1.176 0.2426
-% ifng_probes         0.50919    0.50478   1.009  0.3157
-% myc_probes          0.06345    0.22708   0.279  0.7805
-% pr_probes           0.54200    0.24206   2.239  0.0276  *
-% (Intercept)         0.204758   0.386507  0.530  0.5976
-% bmi                 -0.001950  0.003785  -0.515 0.6076
-% bcat_probes         0.112257   0.225072  0.499  0.6192
-% er_probes           0.033134   0.253540  0.131  0.8963
-% ifna_probes         -0.611750  0.498720  -1.227 0.2231
-% ifng_probes         0.545210   0.511601  1.066  0.2894
-% myc_probes          0.080405   0.230357  0.349  0.7279
-% pr_probes           0.549516   0.243468  2.257  0.0264  *
-% (Intercept)         0.17933    0.37777   0.475  0.636
-% bmiStatusobese      -0.09681   0.06565   -1.475 0.144
-% bmiStatusoverweight -0.01219   0.07111   -0.171 0.864
-% bcat_probes         0.14969    0.22490   0.666  0.507
-% er_probes           0.01354    0.25189   0.054  0.957
-% ifna_probes         -0.73888   0.50091   -1.475 0.144
-% ifng_probes         0.66381    0.51197   1.297  0.198
-% myc_probes          0.10813    0.22869   0.473  0.638
-% pr_probes           0.53872    0.24121   2.233  0.028   *
-% (Intercept)         0.027729   0.390580  0.071  0.9436
-% bmi                 0.009248   0.006522  1.418  0.1596
-% bmiStatusobese      -0.225233  0.111636  -2.018 0.0466  *
-% bmiStatusoverweight -0.058734  0.077961  -0.753 0.4532
-% bcat_probes         0.146066   0.223664  0.653  0.5154
-% er_probes           -0.010526  0.251058  -0.042 0.9667
-% ifna_probes         -0.762373  0.498391  -1.530 0.1296
-% ifng_probes         0.663981   0.509117  1.304  0.1955
-% myc_probes          0.090122   0.227775  0.396  0.6933
-% pr_probes           0.502919   0.241196  2.085  0.0399  *
-
-% [1] "caobsgene"
-
-% Call:
-% lm(formula = as.formula(formula_txt[j]), data = data)
-
-% Coefficients:
-%              Estimate Std. Error t value Pr(>|t|)
-% (Intercept)         0.538968  0.123943 4.349  3.39e-05 ***
-% bmi                 -0.001147 0.004073 -0.282 0.779
-% (Intercept)         0.53896   0.05458  9.875  2.76e-16 ***
-% bmiStatusobese      -0.08982  0.07013  -1.281 0.203
-% bmiStatusoverweight 0.01804   0.07719  0.234  0.816
-% (Intercept)         0.27986   0.16317  1.715  0.0896   .
-% bmi                 0.01171   0.00696  1.683  0.0957   .
-% bmiStatusobese      -0.25146  0.11853  -2.121 0.0365   *
-% bmiStatusoverweight -0.04241  0.08447  -0.502 0.6168
-% (Intercept)         0.12454   0.38803  0.321  0.7490
-% bcat_probes         0.14068   0.22688  0.620  0.5367
-% er_probes           0.11597   0.25833  0.449  0.6545
-% ifna_probes         -0.33319  0.50420  -0.661 0.5104
-% ifng_probes         0.30871   0.51641  0.598  0.5515
-% myc_probes          -0.04569  0.23232  -0.197 0.8445
-% pr_probes           0.56692   0.24764  2.289  0.0243   *
-% (Intercept)         0.162063  0.395326 0.410  0.6828
-% bmi                 -0.002148 0.003872 -0.555 0.5804
-% bcat_probes         0.159335  0.230208 0.692  0.4906
-% er_probes           0.117547  0.259326 0.453  0.6514
-% ifna_probes         -0.368552 0.510101 -0.723 0.4718
-% ifng_probes         0.348376  0.523276 0.666  0.5072
-% myc_probes          -0.027017 0.235613 -0.115 0.9090
-% pr_probes           0.575201  0.249024 2.310  0.0232   *
-% (Intercept)         0.134514  0.384960 0.349  0.7276
-% bmiStatusobese      -0.099695 0.066899 -1.490 0.1397
-% bmiStatusoverweight 0.008987  0.072463 0.124  0.9016
-% bcat_probes         0.193430  0.229183 0.844  0.4009
-% er_probes           0.092451  0.256681 0.360  0.7196
-% ifna_probes         -0.522336 0.510440 -1.023 0.3089
-% ifng_probes         0.492130  0.521713 0.943  0.3481
-% myc_probes          -0.001892 0.233048 -0.008 0.9935
-% pr_probes           0.560583  0.245805 2.281  0.0249   *
-% (Intercept)         -0.028482 0.397505 -0.072 0.9430
-% bmi                 0.009944  0.006637 1.498  0.1376
-% bmiStatusobese      -0.237770 0.113616 -2.093 0.0392   *
-% bmiStatusoverweight -0.041061 0.079343 -0.518 0.6061
-% bcat_probes         0.189535  0.227630 0.833  0.4073
-% er_probes           0.066573  0.255509 0.261  0.7950
-% ifna_probes         -0.547591 0.507227 -1.080 0.2832
-% ifng_probes         0.492312  0.518143 0.950  0.3446
-% myc_probes          -0.021250 0.231813 -0.092 0.9272
-% pr_probes           0.522087  0.245472 2.127  0.0362   *
-
-% [1] "cacrolgene"
-
-% Call:
-% lm(formula = as.formula(formula_txt[j]), data = data)
-
-% Coefficients:
-%               Estimate Std. Error t value Pr(>|t|)
-% (Intercept)         5.062e-01  1.240e-01 4.083  9.16e-05 ***
-% bmi                 -3.953e-05 4.074e-03 -0.010 0.992
-% (Intercept)         0.52453    0.05485   9.564  1.29e-15 ***
-% bmiStatusobese      -0.06411   0.07047   -0.910 0.365
-% bmiStatusoverweight 0.02958    0.07756   0.381  0.704
-% (Intercept)         0.273274   0.164135  1.665  0.0992   .
-% bmi                 0.011360   0.007001  1.622  0.1080
-% bmiStatusobese      -0.220856  0.119235  -1.852 0.0671   .
-% bmiStatusoverweight -0.029032  0.084974  -0.342 0.7334
-% (Intercept)         0.02932    0.40048   0.073  0.9418
-% bcat_probes         0.23660    0.23416   1.010  0.3150
-% er_probes           0.09521    0.26662   0.357  0.7218
-% ifna_probes         -0.49129   0.52037   -0.944 0.3476
-% ifng_probes         0.44466    0.53298   0.834  0.4063
-% myc_probes          0.16111    0.23977   0.672  0.5033
-% pr_probes           0.49566    0.25558   1.939  0.0555   .
-% (Intercept)         0.051698   0.408467  0.127  0.8996
-% bmi                 -0.001281  0.004000  -0.320 0.7495
-% bcat_probes         0.247720   0.237861  1.041  0.3004
-% er_probes           0.096149   0.267946  0.359  0.7205
-% ifna_probes         -0.512383  0.527057  -0.972 0.3335
-% ifng_probes         0.468320   0.540670  0.866  0.3887
-% myc_probes          0.172240   0.243445  0.708  0.4811
-% pr_probes           0.500601   0.257301  1.946  0.0548   .
-% (Intercept)         0.03806    0.39931   0.095  0.9243
-% bmiStatusobese      -0.07973   0.06939   -1.149 0.2536
-% bmiStatusoverweight 0.02249    0.07516   0.299  0.7655
-% bcat_probes         0.27648    0.23772   1.163  0.2479
-% er_probes           0.07298    0.26625   0.274  0.7846
-% ifna_probes         -0.66043   0.52946   -1.247 0.2155
-% ifng_probes         0.60854    0.54116   1.125  0.2638
-% myc_probes          0.19456    0.24173   0.805  0.4230
-% pr_probes           0.48849    0.25497   1.916  0.0586   .
-% (Intercept)         -0.123986  0.412743  -0.300 0.7646
-% bmi                 0.009886   0.006892  1.434  0.1549
-% bmiStatusobese      -0.216997  0.117971  -1.839 0.0692   .
-% bmiStatusoverweight -0.027267  0.082384  -0.331 0.7414
-% bcat_probes         0.272606   0.236356  1.153  0.2518
-% er_probes           0.047249   0.265304  0.178  0.8591
-% ifna_probes         -0.685542  0.526672  -1.302 0.1964
-% ifng_probes         0.608724   0.538006  1.131  0.2609
-% myc_probes          0.175314   0.240700  0.728  0.4683
-% pr_probes           0.450215   0.254882  1.766  0.0808   .
-
-% [1] "caresobsgene"
-
-% Call:
-% lm(formula = as.formula(formula_txt[j]), data = data)
-
-% Coefficients:
-%             Estimate Std. Error t value Pr(>|t|)
-% (Intercept)         0.462424  0.123913 3.732  0.00032  ***
-% bmi                 0.001442  0.004072 0.354  0.72408
-% (Intercept)         0.46970   0.05452  8.616  1.39e-13 ***
-% bmiStatusobese      0.09338   0.07005  1.333  0.186
-% bmiStatusoverweight -0.01840  0.07710  -0.239 0.812
-% (Intercept)         0.722057  0.163100 4.427  2.55e-05 ***
-% bmi                 -0.011409 0.006957 -1.640 0.1043
-% bmiStatusobese      0.250806  0.118483 2.117  0.0369   *
-% bmiStatusoverweight 0.040473  0.084438 0.479  0.6328
-% (Intercept)         1.1017    0.3815   2.888  0.00483  **
-% bcat_probes         -0.2217   0.2230   -0.994 0.32284
-% er_probes           -0.2308   0.2540   -0.909 0.36579
-% ifna_probes         0.1853    0.4957   0.374  0.70938
-% ifng_probes         -0.1528   0.5077   -0.301 0.76408
-% myc_probes          -0.0737   0.2284   -0.323 0.74765
-% pr_probes           -0.6876   0.2434   -2.825 0.00581  **
-% (Intercept)         1.056654  0.388317 2.721  0.00780  **
-% bmi                 0.002580  0.003803 0.678  0.49929
-% bcat_probes         -0.244104 0.226127 -1.079 0.28322
-% er_probes           -0.232713 0.254728 -0.914 0.36335
-% ifna_probes         0.227770  0.501056 0.455  0.65049
-% ifng_probes         -0.200460 0.513998 -0.390 0.69745
-% myc_probes          -0.096122 0.231435 -0.415 0.67888
-% pr_probes           -0.697585 0.244608 -2.852 0.00538  **
-% (Intercept)         1.09174   0.37821  2.887  0.00488  **
-% bmiStatusobese      0.10153   0.06573  1.545  0.12592
-% bmiStatusoverweight -0.00544  0.07119  -0.076 0.93926
-% bcat_probes         -0.27598  0.22517  -1.226 0.22351
-% er_probes           -0.20770  0.25218  -0.824 0.41234
-% ifna_probes         0.37360   0.50149  0.745  0.45823
-% ifng_probes         -0.33545  0.51257  -0.654 0.51449
-% myc_probes          -0.11868  0.22896  -0.518 0.60548
-% pr_probes           -0.68170  0.24150  -2.823 0.00586  **
-% (Intercept)         1.235747  0.391476 3.157  0.00218  **
-% bmi                 -0.008785 0.006537 -1.344 0.18237
-% bmiStatusobese      0.223518  0.111892 1.998  0.04881  *
-% bmiStatusoverweight 0.038776  0.078139 0.496  0.62094
-% bcat_probes         -0.272540 0.224177 -1.216 0.22730
-% er_probes           -0.184837 0.251634 -0.735 0.46455
-% ifna_probes         0.395907  0.499535 0.793  0.43015
-% ifng_probes         -0.335611 0.510285 -0.658 0.51243
-% myc_probes          -0.101582 0.228298 -0.445 0.65743
-% pr_probes           -0.647688 0.241749 -2.679 0.00879  **
-
-% [1] "carescrolgene"
-
-% Call:
-% lm(formula = as.formula(formula_txt[j]), data = data)
-
-% Coefficients:
-%              Estimate Std. Error t value Pr(>|t|)
-% (Intercept)         0.4928357 0.1239867 3.975  0.000136 ***
-% bmi                 0.0004131 0.0040740 0.101  0.919450
-% (Intercept)         0.52814   0.05483   9.633  9.13e-16 ***
-% bmiStatusobese      -0.06913  0.07045   -0.981 0.329
-% bmiStatusoverweight 0.02453   0.07753   0.316  0.752
-% (Intercept)         0.232845  0.163204  1.427  0.1569
-% bmi                 0.013351  0.006962  1.918  0.0582   .
-% bmiStatusobese      -0.253344 0.118558  -2.137 0.0352   *
-% bmiStatusoverweight -0.044356 0.084492  -0.525 0.6008
-% (Intercept)         -0.1205   0.3891    -0.310 0.7576
-% bcat_probes         0.2439    0.2275    1.072  0.2864
-% er_probes           0.1383    0.2590    0.534  0.5948
-% ifna_probes         -0.4364   0.5056    -0.863 0.3903
-% ifng_probes         0.4301    0.5178    0.831  0.4084
-% myc_probes          0.2837    0.2330    1.218  0.2264
-% pr_probes           0.5789    0.2483    2.331  0.0219   *
-% (Intercept)         -0.102321 0.396926  -0.258 0.7972
-% bmi                 -0.001038 0.003887  -0.267 0.7900
-% bcat_probes         0.252943  0.231140  1.094  0.2767
-% er_probes           0.139029  0.260375  0.534  0.5947
-% ifna_probes         -0.453502 0.512164  -0.885 0.3782
-% ifng_probes         0.449266  0.525393  0.855  0.3947
-% myc_probes          0.292758  0.236566  1.238  0.2191
-% pr_probes           0.582903  0.250031  2.331  0.0219   *
-% (Intercept)         -0.11188  0.38790   -0.288 0.7737
-% bmiStatusobese      -0.08019  0.06741   -1.190 0.2373
-% bmiStatusoverweight 0.01833   0.07302   0.251  0.8024
-% bcat_probes         0.28469   0.23093   1.233  0.2209
-% er_probes           0.11686   0.25864   0.452  0.6525
-% ifna_probes         -0.60153  0.51434   -1.170 0.2453
-% ifng_probes         0.59011   0.52570   1.123  0.2646
-% myc_probes          0.31782   0.23483   1.353  0.1793
-% pr_probes           0.57227   0.24768   2.311  0.0231   *
-% (Intercept)         -0.284628 0.400004  -0.712 0.4786
-% bmi                 0.010538  0.006679  1.578  0.1181
-% bmiStatusobese      -0.226528 0.114330  -1.981 0.0506   .
-% bmiStatusoverweight -0.034711 0.079842  -0.435 0.6648
-% bcat_probes         0.280559  0.229061  1.225  0.2239
-% er_probes           0.089437  0.257116  0.348  0.7288
-% ifna_probes         -0.628292 0.510417  -1.231 0.2216
-% ifng_probes         0.590302  0.521401  1.132  0.2606
-% myc_probes          0.297308  0.233271  1.275  0.2058
-% pr_probes           0.531475  0.247015  2.152  0.0341   *
-
-% [1] "crobsgenes"
-
-% Call:
-% lm(formula = as.formula(formula_txt[j]), data = data)
-
-% Coefficients:
-%             Estimate Std. Error t value Pr(>|t|)
-% (Intercept)         0.456126  0.123888 3.682  0.000381 ***
-% bmi                 0.001654  0.004071 0.406  0.685322
-% (Intercept)         0.459596  0.054606 8.417  3.7e-13  ***
-% bmiStatusobese      0.100893  0.070167 1.438  0.154
-% bmiStatusoverweight 0.005772  0.077224 0.075  0.941
-% (Intercept)         0.704210  0.163518 4.307  4.04e-05 ***
-% bmi                 -0.011059 0.006975 -1.586 0.1162
-% bmiStatusobese      0.253491  0.118786 2.134  0.0354   *
-% bmiStatusoverweight 0.062836  0.084654 0.742  0.4598
-% (Intercept)         1.0667    0.3806   2.803  0.00618  **
-% bcat_probes         -0.2284   0.2225   -1.026 0.30740
-% er_probes           -0.1481   0.2534   -0.584 0.56041
-% ifna_probes         0.4504    0.4945   0.911  0.36477
-% ifng_probes         -0.4013   0.5065   -0.792 0.43018
-% myc_probes          -0.1402   0.2279   -0.615 0.53986
-% pr_probes           -0.6446   0.2429   -2.654 0.00938  **
-% (Intercept)         1.014279  0.387058 2.620  0.01029  *
-% bmi                 0.003004  0.003791 0.792  0.43014
-% bcat_probes         -0.254482 0.225394 -1.129 0.26184
-% er_probes           -0.150261 0.253902 -0.592 0.55545
-% ifna_probes         0.499868  0.499432 1.001  0.31954
-% ifng_probes         -0.456812 0.512332 -0.892 0.37494
-% myc_probes          -0.166314 0.230685 -0.721 0.47278
-% pr_probes           -0.656157 0.243815 -2.691 0.00847  **
-% (Intercept)         1.05662   0.37693  2.803  0.0062   **
-% bmiStatusobese      0.11346   0.06550  1.732  0.0867   .
-% bmiStatusoverweight 0.01412   0.07095  0.199  0.8427
-% bcat_probes         -0.29209  0.22440  -1.302 0.1964
-% er_probes           -0.12674  0.25133  -0.504 0.6153
-% ifna_probes         0.63723   0.49979  1.275  0.2056
-% ifng_probes         -0.58273  0.51083  -1.141 0.2570
-% myc_probes          -0.19254  0.22819  -0.844 0.4010
-% pr_probes           -0.64072  0.24068  -2.662 0.0092   **
-% (Intercept)         1.198471  0.390241 3.071  0.00283  **
-% bmi                 -0.008654 0.006516 -1.328 0.18754
-% bmiStatusobese      0.233624  0.111539 2.095  0.03905  *
-% bmiStatusoverweight 0.057676  0.077893 0.740  0.46097
-% bcat_probes         -0.288700 0.223470 -1.292 0.19974
-% er_probes           -0.104217 0.250840 -0.415 0.67880
-% ifna_probes         0.659208  0.497958 1.324  0.18895
-% ifng_probes         -0.582884 0.508675 -1.146 0.25491
-% myc_probes          -0.175698 0.227577 -0.772 0.44214
-% pr_probes           -0.607216 0.240986 -2.520 0.01353  *
-
-% [1] "fmobsgenes"
-
-% Call:
-% lm(formula = as.formula(formula_txt[j]), data = data)
-
-% Coefficients:
-%             Estimate Std. Error t value Pr(>|t|)
-% (Intercept)         0.436628  0.123787 3.527  0.000643 ***
-% bmi                 0.002314  0.004067 0.569  0.570763
-% (Intercept)         0.58369   0.05435  10.740 <2e-16   ***
-% bmiStatusobese      -0.08452  0.06983  -1.210 0.2291
-% bmiStatusoverweight -0.14827  0.07686  -1.929 0.0567   .
-% (Intercept)         0.251170  0.160863 1.561  0.12176
-% bmi                 0.015034  0.006862 2.191  0.03091  *
-% bmiStatusobese      -0.291956 0.116858 -2.498 0.01419  *
-% bmiStatusoverweight -0.225840 0.083280 -2.712 0.00794  **
-% (Intercept)         0.33317   0.36900  0.903  0.3689
-% bcat_probes         -0.13489  0.21576  -0.625 0.5334
-% er_probes           0.11973   0.24567  0.487  0.6271
-% ifna_probes         -0.56244  0.47947  -1.173 0.2438
-% ifng_probes         0.58838   0.49109  1.198  0.2340
-% myc_probes          0.41768   0.22093  1.891  0.0618   .
-% pr_probes           -0.08815  0.23550  -0.374 0.7090
-% (Intercept)         0.291390  0.375705 0.776  0.4400
-% bmi                 0.002392  0.003679 0.650  0.5172
-% bcat_probes         -0.155669 0.218783 -0.712 0.4786
-% er_probes           0.117981  0.246455 0.479  0.6333
-% ifna_probes         -0.523056 0.484783 -1.079 0.2835
-% ifng_probes         0.544202  0.497304 1.094  0.2767
-% myc_probes          0.396892  0.223919 1.772  0.0797   .
-% pr_probes           -0.097371 0.236664 -0.411 0.6817
-% (Intercept)         0.33459   0.36629  0.913  0.3634
-% bmiStatusobese      -0.08029  0.06366  -1.261 0.2104
-% bmiStatusoverweight -0.12492  0.06895  -1.812 0.0733   .
-% bcat_probes         -0.07254  0.21807  -0.333 0.7402
-% er_probes           0.13036   0.24424  0.534  0.5948
-% ifna_probes         -0.56035  0.48569  -1.154 0.2517
-% ifng_probes         0.58760   0.49642  1.184  0.2397
-% myc_probes          0.46651   0.22175  2.104  0.0382   *
-% pr_probes           -0.07506  0.23389  -0.321 0.7490
-% (Intercept)         0.082260  0.370293 0.222  0.82471
-% bmi                 0.015393  0.006183 2.490  0.01464  *
-% bmiStatusobese      -0.294047 0.105838 -2.778 0.00666  **
-% bmiStatusoverweight -0.202402 0.073911 -2.738 0.00746  **
-% bcat_probes         -0.078569 0.212047 -0.371 0.71187
-% er_probes           0.090296  0.238018 0.379  0.70532
-% ifna_probes         -0.599445 0.472504 -1.269 0.20787
-% ifng_probes         0.587878  0.482673 1.218  0.22646
-% myc_probes          0.436540  0.215944 2.022  0.04623  *
-% pr_probes           -0.134650 0.228668 -0.589 0.55746
-
-\begin{longtable}{llr{\bfseries}S}
-	\centering
-	\caption{Table showing the variables in the step wise linear models to predict the obesity metagenes}
-	\label{tab:stepwise_model}\\
-		Obesity metagene & Variables used in the model & Estimate & {P-value}\\
-		\hline
-	\endfirsthead
-	\multicolumn{4}{c}{\tablename\ \thetable{}\ (continued)}\\
-		\hline
-		\hline
-	\endhead
-		\hline
-		\hline
-		Cr         & Akt         & -0.3415   & \bfseries \num{4.43d-8}\\
-				   & \gls{egfr}  & 0.33852   & \bfseries \num{1.94d-05}\\
-				   & \gls{er}    & -0.20237  & \bfseries 0.00419\\
-				   & \gls{ifny}  & 0.12233   & 0.08700\\
-				   & \gls{pr}    & -0.33920  & \bfseries \num{2.09d-5}\\
-				   & Ras         & -0.22611  & \bfseries 0.00193\\
-				   & Src         & -0.11452  & 0.08177\\
-				   & \gls{tgfb}  & 0.55638   & \bfseries \textless{} \num{2d-16}\\
-				   & \gls{tnfa}  & -0.25501  & \bfseries 0.00616\\
-		\hline
-		CrOl       & Overweight  & 0.040918  & 0.074865\\
-				   & Obese       & -0.006381 & 0.756121\\
-				   & Akt         & -0.356713 & \bfseries \num{1.89d-07}\\
-				   & E2F1        & 0.160229  & \bfseries 0.005411\\
-				   & \gls{egfr}  & 0.148618  & 0.062683\\
-				   & p53         & 0.336552  & \bfseries \num{4.35d-05}\\
-				   & \gls{pr}    & -0.225423 & \bfseries 0.000639\\
-				   & Src         & -0.428758 & \bfseries \num{5.85d-10}\\
-				   & \gls{stat3} & -0.124118 & \bfseries 0.009658\\
-		\hline
-		Res        & Akt         & 0.24613   & \bfseries 0.000289\\
-				   & \gls{egfr}  & -0.46827  & \bfseries \num{2.16d-05}\\
-				   & \gls{er}    & 0.18400   & 0.061446\\
-				   & \gls{her2}  & 0.13835   & 0.118228\\
-				   & \gls{ifny}  & -0.21019  & \bfseries 0.044084\\
-				   & \gls{pr}    & 0.41195   & \bfseries 0.000300\\
-				   & Ras         & 0.26506   & \bfseries 0.010465\\
-				   & \gls{tgfb}  & -0.58075  & \bfseries \num{7.64d-11}\\
-				   & \gls{tnfa}  & 0.42200   & \bfseries 0.001834\\
-		\hline
-		ResOl      & Akt         & -0.33596  & \bfseries 0.000294\\
-				   & \gls{bcat}  & -0.19866  & 0.138953\\
-				   & E2F1        & 0.36059   & \bfseries 0.009325\\
-				   & \gls{egfr}  & 0.23421   & 0.056865\\
-				   & \gls{her2}  & -0.13652  & 0.121689\\
-				   & p53         & 0.38746   & \bfseries 0.000519\\
-				   & \gls{pi3k}  & -0.21881  & 0.148676\\
-				   & \gls{pr}    & -0.32228  & \bfseries 0.001656\\
-				   & Ras         & -0.28984  & \bfseries 0.033975\\
-				   & Src         & -0.40779  & \bfseries \num{5.26d-06}\\
-				   & \gls{stat3} & -0.11452  & 0.125609\\
-				   & \gls{tnfa}  & -0.11442  & 0.117023\\
-		\hline
-		Ca         & Akt         & -0.26243  & \bfseries \num{7.04d-06}\\
-				   & \gls{egfr}  & 0.36305   & \bfseries 0.000139\\
-				   & \gls{er}    & -0.18614  & \bfseries 0.030771\\
-				   & \gls{ifny}  & 0.22118   & \bfseries 0.013034\\
-				   & \gls{pr}    & -0.42911  & \bfseries \num{1.07d-05}\\
-				   & Ras         & -0.28124  & \bfseries 0.000898\\
-				   & \gls{tgfb}  & 0.76139   & \bfseries \textless{} \num{2d-16}\\
-				   & \gls{tnfa}  & -0.33200  & \bfseries 0.003919\\
-		\hline
-		CaOl       & Overweight  & 0.07613   & \bfseries 0.006875\\
-				   & Obese       & 0.01408   & 0.564622\\
-				   & Akt         & -0.44727  & \bfseries \num{2.43d-07}\\
-				   & \gls{bcat}  & -0.13463  & 0.140261\\
-				   & E2F1        & 0.33864   & \bfseries 0.001249\\
-				   & p53         & 0.28777   & \bfseries 0.000996\\
-				   & \gls{pr}    & -0.28255  & \bfseries \num{2.18d-05}\\
-				   & Src         & -0.44149  & \bfseries \num{2.02d-08}\\
-				   & \gls{tnfa}  & -0.14215  & \bfseries 0.013897\\
-		\hline
-		CaRes      & \gls{bcat}  & 0.26899   & 0.05238  \\
-				   & E2F1        & -0.22115  & 0.08848  \\
-				   & \gls{egfr}  & 0.44400   & \bfseries \num{6.67d-05 }\\
-				   & \gls{her2}  & -0.20304  & \bfseries 0.04615  \\
-				   & \gls{ifny}  & 0.32874   & \bfseries 0.00574  \\
-				   & \gls{pr}    & -0.28764  & \bfseries 0.00104  \\
-				   & Src         & -0.17989  & 0.05404  \\
-				   & \gls{tgfb}  & 0.52898   & \bfseries \num{2.15d-06 }\\
-				   & \gls{tnfa}  & -0.38677  & \bfseries 0.00564  \\
-		\hline
-		CaResOl    & \gls{bmi}   & 0.004314  & 0.108546\\
-				   & Overweight  & 0.033086  & 0.306390\\
-				   & Obese       & -0.062257 & 0.175776\\
-				   & Akt         & -0.309706 & \bfseries 0.000339\\
-				   & \gls{bcat}  & -0.198165 & 0.140502\\
-				   & E2F1        & 0.270471  & \bfseries 0.011450\\
-				   & \gls{egfr}  & 0.171219  & 0.113691\\
-				   & Myc         & 0.152007  & 0.056987\\
-				   & p53         & 0.294895  & \bfseries 0.003734\\
-				   & \gls{pr}    & -0.230896 & \bfseries 0.000823\\
-				   & Ras         & -0.270824 & \bfseries 0.044788\\
-				   & Src         & -0.512567 & \bfseries \num{2.72d-08}\\
-				   & \gls{tnfa}  & -0.103302 & 0.083449\\
-		\hline
-		Original   & \gls{bmi}   & 0.002696  & 0.14407\\
-				   & Overweight  & 0.010522  & 0.62856\\
-				   & Obese       & -0.061020 & 0.05542\\
-				   & Akt         & -0.238412 & \bfseries \num{4.93d-05}\\
-				   & E2F1        & 0.185085  & \bfseries 0.02754\\
-				   & \gls{egfr}  & 0.147219  & 0.07423\\
-				   & \gls{her2}  & -0.086804 & 0.14070\\
-				   & p53         & 0.416414  & \bfseries \num{5.83d-08}\\
-				   & \gls{pi3k}  & -0.216179 & \bfseries 0.03528\\
-				   & \gls{pr}    & -0.199097 & \bfseries 0.00253\\
-				   & Ras         & -0.200544 & \bfseries 0.00898\\
-				   & Src         & -0.543834 & \bfseries \num{2.10d-15}\\
-				   & \gls{stat3} & -0.093363 & \bfseries 0.02749\\
-		\hline
-		FM         & \gls{bmi}   & 0.011093  & \bfseries 0.036229\\
-				   & Overweight  & -0.153753 & \bfseries 0.014277\\
-				   & Obese       & -0.212483 & \bfseries 0.017378\\
-				   & p53         & 0.460800  & \bfseries 0.002512\\
-				   & \gls{pi3k}  & -0.426246 & \bfseries 0.000118\\
-				   & \gls{pr}    & -0.373119 & \bfseries 0.020737\\
-				   & \gls{stat3} & 0.248745  & \bfseries 0.019073\\
-				   & \gls{tgfb}  & 0.152925  & 0.128933\\
-		\hline
-		\hline
-	\end{longtable}
-
-\end{appendices}

### 2017-03-13 thesis/misc/appendix2.tex
Need to add linear model prediction scatter plots - tmp commit

@@
 	\label{app:b}
+
+	\section{Ranking method for the pathway associated genetic signatures}
+	\label{sec:ranking_method_for_the_pathway_associated_genetic_signatures}
+
+	To determine whether it was the best to rank the metagene scores based on the number of samples or rank the scores with the probit method, the two ranking methods were compared in scatter plots with the GT pathway metagenes in \gls{rma} or \gls{mas} normalised GT data.
+	Since the results from the \gls{rma} and \gls{mas} normalised GT data were so similar to one another, only the results from the \gls{rma} normalised data are presented (\cref{fig:appendix/rank_scatter}).
+	\cref{fig:appendix/rank_scatter} showed that the metagene scores for all of the GT pathway associated genetic signatures were approximately the same in either of the ranking methods used.
+
+	\begin{figure}[htp!]
+		\centering
+		\includegraphics[width=0.32\linewidth,page=1]{appendix/gatza_rma_meta_rank_vs_probit}
+		\includegraphics[width=0.32\linewidth,page=2]{appendix/gatza_rma_meta_rank_vs_probit}
+		\includegraphics[width=0.32\linewidth,page=3]{appendix/gatza_rma_meta_rank_vs_probit}\\
+		\includegraphics[width=0.32\linewidth,page=4]{appendix/gatza_rma_meta_rank_vs_probit}
+		\includegraphics[width=0.32\linewidth,page=5]{appendix/gatza_rma_meta_rank_vs_probit}
+		\includegraphics[width=0.32\linewidth,page=6]{appendix/gatza_rma_meta_rank_vs_probit}\\
+		\includegraphics[width=0.32\linewidth,page=7]{appendix/gatza_rma_meta_rank_vs_probit}
+		\includegraphics[width=0.32\linewidth,page=8]{appendix/gatza_rma_meta_rank_vs_probit}
+		\includegraphics[width=0.32\linewidth,page=9]{appendix/gatza_rma_meta_rank_vs_probit}\\
+		\includegraphics[width=0.32\linewidth,page=10]{appendix/gatza_rma_meta_rank_vs_probit}
+		\includegraphics[width=0.32\linewidth,page=11]{appendix/gatza_rma_meta_rank_vs_probit}
+		\includegraphics[width=0.32\linewidth,page=12]{appendix/gatza_rma_meta_rank_vs_probit}\\
+		\caption[]{Scatter plots comparing the GT pathway metagenes ranked with probit and ranked based on number of samples in \gls{rma} normalised GT data. }
+		\label{fig:appendix/rank_scatter}
+	\end{figure}
+
+	\begin{figure}[htpb]
+		\ContinuedFloat
+		\captionsetup{list=off,format=cont}
+		\centering
+		\includegraphics[width=0.32\linewidth,page=13]{appendix/gatza_rma_meta_rank_vs_probit}
+		\includegraphics[width=0.32\linewidth,page=14]{appendix/gatza_rma_meta_rank_vs_probit}
+		\includegraphics[width=0.32\linewidth,page=15]{appendix/gatza_rma_meta_rank_vs_probit}\\
+		\includegraphics[width=0.32\linewidth,page=16]{appendix/gatza_rma_meta_rank_vs_probit}
+		\includegraphics[width=0.32\linewidth,page=17]{appendix/gatza_rma_meta_rank_vs_probit}
+		\includegraphics[width=0.32\linewidth,page=18]{appendix/gatza_rma_meta_rank_vs_probit}\\
+		\vspace{1em}
+		\caption[]{figure}
+	\end{figure}
+
+	\section{Normalisation method for the pathway associated genetic signature transformation matrices}
+	\label{sec:normalisation_method_for_pathway_associated_genetic_signature_transformation_matrices}
+
+	To investigate whether the normalisation methods were important in the creation of the pathway associated genetic signature transformation matrices, and whether the normalisation methods of the data affected the resulting metagenes, transformation matrices were generated in \gls{rma} or \gls{mas} normalised GT data set and then applied to the \gls{rma} or \gls{mas} normalised GT data set.
+	The results in \cref{fig:appendix/gt_meta_rma_mas} showed less variability of the metagene scores when the metagenes were derived from the data with the same normalisation method (for example, metagenes derived from \gls{rma} normalised data), regardless of the transformation matrix used to produce the metagenes.
+	This was most clearly demonstrated by the \gls{bcat}, \gls{er}, \gls{ifna}, \gls{ifny}, p53 and \gls{pr} pathway associated genetic signatures.
+
+	\begin{figure}[htp!]
+		\centering
+		\includegraphics[width=0.32\linewidth,page=1]{appendix/allmeta_rma_vs_mas}
+		\includegraphics[width=0.32\linewidth,page=2]{appendix/allmeta_rma_vs_mas}
+		\includegraphics[width=0.32\linewidth,page=3]{appendix/allmeta_rma_vs_mas}\\
+		\includegraphics[width=0.32\linewidth,page=4]{appendix/allmeta_rma_vs_mas}
+		\includegraphics[width=0.32\linewidth,page=5]{appendix/allmeta_rma_vs_mas}
+		\includegraphics[width=0.32\linewidth,page=6]{appendix/allmeta_rma_vs_mas}\\
+		\includegraphics[width=0.32\linewidth,page=7]{appendix/allmeta_rma_vs_mas}
+		\includegraphics[width=0.32\linewidth,page=8]{appendix/allmeta_rma_vs_mas}
+		\includegraphics[width=0.32\linewidth,page=9]{appendix/allmeta_rma_vs_mas}\\
+		\includegraphics[width=0.32\linewidth,page=10]{appendix/allmeta_rma_vs_mas}
+		\includegraphics[width=0.32\linewidth,page=11]{appendix/allmeta_rma_vs_mas}
+		\includegraphics[width=0.32\linewidth,page=12]{appendix/allmeta_rma_vs_mas}\\
+		\caption[]{Scatter plots comparing the GT pathway metagenes derived from the transformation matrices (TM) in \gls{rma} or \gls{mas} normalised GT data set.
+		All of the transformation matrices were generated in either the \gls{rma} or \gls{mas} normalised GT data set.
+		}
+		\label{fig:appendix/gt_meta_rma_mas}
+	\end{figure}
+
+	\begin{figure}[htpb]
+		\ContinuedFloat
+		\captionsetup{list=off,format=cont}
+		\centering
+		\includegraphics[width=0.32\linewidth,page=13]{appendix/allmeta_rma_vs_mas}
+		\includegraphics[width=0.32\linewidth,page=14]{appendix/allmeta_rma_vs_mas}
+		\includegraphics[width=0.32\linewidth,page=15]{appendix/allmeta_rma_vs_mas}\\
+		\includegraphics[width=0.32\linewidth,page=16]{appendix/allmeta_rma_vs_mas}
+		\includegraphics[width=0.32\linewidth,page=17]{appendix/allmeta_rma_vs_mas}
+		\includegraphics[width=0.32\linewidth,page=18]{appendix/allmeta_rma_vs_mas}\\
+		\includegraphics[width=0.32\linewidth,page=19]{appendix/allmeta_rma_vs_mas}
+		\includegraphics[width=0.32\linewidth,page=20]{appendix/allmeta_rma_vs_mas}
+		\includegraphics[width=0.32\linewidth,page=21]{appendix/allmeta_rma_vs_mas}\\
+		\includegraphics[width=0.32\linewidth,page=22]{appendix/allmeta_rma_vs_mas}
+		\includegraphics[width=0.32\linewidth,page=23]{appendix/allmeta_rma_vs_mas}
+		\includegraphics[width=0.32\linewidth,page=24]{appendix/allmeta_rma_vs_mas}\\
+		\caption[]{figure}
+	\end{figure}
+
+	\begin{figure}[htpb]
+		\ContinuedFloat
+		\captionsetup{list=off,format=cont}
+		\centering
+		\includegraphics[width=0.32\linewidth,page=25]{appendix/allmeta_rma_vs_mas}
+		\includegraphics[width=0.32\linewidth,page=26]{appendix/allmeta_rma_vs_mas}
+		\includegraphics[width=0.32\linewidth,page=27]{appendix/allmeta_rma_vs_mas}\\
+		\includegraphics[width=0.32\linewidth,page=28]{appendix/allmeta_rma_vs_mas}
+		\includegraphics[width=0.32\linewidth,page=29]{appendix/allmeta_rma_vs_mas}
+		\includegraphics[width=0.32\linewidth,page=30]{appendix/allmeta_rma_vs_mas}\\
+		\includegraphics[width=0.32\linewidth,page=31]{appendix/allmeta_rma_vs_mas}
+		\includegraphics[width=0.32\linewidth,page=32]{appendix/allmeta_rma_vs_mas}
+		\includegraphics[width=0.32\linewidth,page=33]{appendix/allmeta_rma_vs_mas}\\
+		\includegraphics[width=0.32\linewidth,page=34]{appendix/allmeta_rma_vs_mas}
+		\includegraphics[width=0.32\linewidth,page=35]{appendix/allmeta_rma_vs_mas}
+		\includegraphics[width=0.32\linewidth,page=36]{appendix/allmeta_rma_vs_mas}\\
+		\caption[]{figure}
+	\end{figure}
+
+	\begin{figure}[htpb]
+		\ContinuedFloat
+		\captionsetup{list=off,format=cont}
+		\centering
+		\includegraphics[width=0.32\linewidth,page=37]{appendix/allmeta_rma_vs_mas}
+		\includegraphics[width=0.32\linewidth,page=38]{appendix/allmeta_rma_vs_mas}
+		\includegraphics[width=0.32\linewidth,page=39]{appendix/allmeta_rma_vs_mas}\\
+		\includegraphics[width=0.32\linewidth,page=40]{appendix/allmeta_rma_vs_mas}
+		\includegraphics[width=0.32\linewidth,page=41]{appendix/allmeta_rma_vs_mas}
+		\includegraphics[width=0.32\linewidth,page=42]{appendix/allmeta_rma_vs_mas}\\
+		\includegraphics[width=0.32\linewidth,page=43]{appendix/allmeta_rma_vs_mas}
+		\includegraphics[width=0.32\linewidth,page=44]{appendix/allmeta_rma_vs_mas}
+		\includegraphics[width=0.32\linewidth,page=45]{appendix/allmeta_rma_vs_mas}\\
+		\includegraphics[width=0.32\linewidth,page=46]{appendix/allmeta_rma_vs_mas}
+		\includegraphics[width=0.32\linewidth,page=47]{appendix/allmeta_rma_vs_mas}
+		\includegraphics[width=0.32\linewidth,page=48]{appendix/allmeta_rma_vs_mas}\\
+		\caption[]{figure}
+	\end{figure}
+
+	\begin{figure}[htpb]
+		\ContinuedFloat
+		\captionsetup{list=off,format=cont}
+		\centering
+		\includegraphics[width=0.32\linewidth,page=49]{appendix/allmeta_rma_vs_mas}
+		\includegraphics[width=0.32\linewidth,page=50]{appendix/allmeta_rma_vs_mas}
+		\includegraphics[width=0.32\linewidth,page=51]{appendix/allmeta_rma_vs_mas}\\
+		\includegraphics[width=0.32\linewidth,page=52]{appendix/allmeta_rma_vs_mas}
+		\includegraphics[width=0.32\linewidth,page=53]{appendix/allmeta_rma_vs_mas}
+		\includegraphics[width=0.32\linewidth,page=54]{appendix/allmeta_rma_vs_mas}\\
+		\includegraphics[width=0.32\linewidth,page=55]{appendix/allmeta_rma_vs_mas}
+		\includegraphics[width=0.32\linewidth,page=56]{appendix/allmeta_rma_vs_mas}
+		\includegraphics[width=0.32\linewidth,page=57]{appendix/allmeta_rma_vs_mas}\\
+		\includegraphics[width=0.32\linewidth,page=58]{appendix/allmeta_rma_vs_mas}
+		\includegraphics[width=0.32\linewidth,page=59]{appendix/allmeta_rma_vs_mas}
+		\includegraphics[width=0.32\linewidth,page=60]{appendix/allmeta_rma_vs_mas}\\
+		\caption[]{figure}
+	\end{figure}
+
+	\begin{figure}[htpb]
+		\ContinuedFloat
+		\captionsetup{list=off,format=cont}
+		\centering
+		\includegraphics[width=0.32\linewidth,page=61]{appendix/allmeta_rma_vs_mas}
+		\includegraphics[width=0.32\linewidth,page=62]{appendix/allmeta_rma_vs_mas}
+		\includegraphics[width=0.32\linewidth,page=63]{appendix/allmeta_rma_vs_mas}\\
+		\includegraphics[width=0.32\linewidth,page=64]{appendix/allmeta_rma_vs_mas}
+		\includegraphics[width=0.32\linewidth,page=65]{appendix/allmeta_rma_vs_mas}
+		\includegraphics[width=0.32\linewidth,page=66]{appendix/allmeta_rma_vs_mas}\\
+		\includegraphics[width=0.32\linewidth,page=67]{appendix/allmeta_rma_vs_mas}
+		\includegraphics[width=0.32\linewidth,page=68]{appendix/allmeta_rma_vs_mas}
+		\includegraphics[width=0.32\linewidth,page=69]{appendix/allmeta_rma_vs_mas}\\
+		\includegraphics[width=0.32\linewidth,page=70]{appendix/allmeta_rma_vs_mas}
+		\includegraphics[width=0.32\linewidth,page=71]{appendix/allmeta_rma_vs_mas}
+		\includegraphics[width=0.32\linewidth,page=72]{appendix/allmeta_rma_vs_mas}\\
+		\caption[]{figure}
+	\end{figure}
+
+	\begin{figure}[htpb]
+		\ContinuedFloat
+		\captionsetup{list=off,format=cont}
+		\centering
+		\includegraphics[width=0.32\linewidth,page=73]{appendix/allmeta_rma_vs_mas}
+		\includegraphics[width=0.32\linewidth,page=74]{appendix/allmeta_rma_vs_mas}
+		\includegraphics[width=0.32\linewidth,page=75]{appendix/allmeta_rma_vs_mas}\\
+		\includegraphics[width=0.32\linewidth,page=76]{appendix/allmeta_rma_vs_mas}
+		\includegraphics[width=0.32\linewidth,page=77]{appendix/allmeta_rma_vs_mas}
+		\includegraphics[width=0.32\linewidth,page=78]{appendix/allmeta_rma_vs_mas}\\
+		\includegraphics[width=0.32\linewidth,page=79]{appendix/allmeta_rma_vs_mas}
+		\includegraphics[width=0.32\linewidth,page=80]{appendix/allmeta_rma_vs_mas}
+		\includegraphics[width=0.32\linewidth,page=81]{appendix/allmeta_rma_vs_mas}\\
+		\includegraphics[width=0.32\linewidth,page=82]{appendix/allmeta_rma_vs_mas}
+		\includegraphics[width=0.32\linewidth,page=83]{appendix/allmeta_rma_vs_mas}
+		\includegraphics[width=0.32\linewidth,page=84]{appendix/allmeta_rma_vs_mas}\\
+		\caption[]{figure}
+	\end{figure}
+
+	\begin{figure}[htpb]
+		\ContinuedFloat
+		\captionsetup{list=off,format=cont}
+		\centering
+		\includegraphics[width=0.32\linewidth,page=85]{appendix/allmeta_rma_vs_mas}
+		\includegraphics[width=0.32\linewidth,page=86]{appendix/allmeta_rma_vs_mas}
+		\includegraphics[width=0.32\linewidth,page=87]{appendix/allmeta_rma_vs_mas}\\
+		\includegraphics[width=0.32\linewidth,page=88]{appendix/allmeta_rma_vs_mas}
+		\includegraphics[width=0.32\linewidth,page=89]{appendix/allmeta_rma_vs_mas}
+		\includegraphics[width=0.32\linewidth,page=90]{appendix/allmeta_rma_vs_mas}\\
+		\includegraphics[width=0.32\linewidth,page=91]{appendix/allmeta_rma_vs_mas}
+		\includegraphics[width=0.32\linewidth,page=92]{appendix/allmeta_rma_vs_mas}
+		\includegraphics[width=0.32\linewidth,page=93]{appendix/allmeta_rma_vs_mas}\\
+		\includegraphics[width=0.32\linewidth,page=94]{appendix/allmeta_rma_vs_mas}
+		\includegraphics[width=0.32\linewidth,page=95]{appendix/allmeta_rma_vs_mas}
+		\includegraphics[width=0.32\linewidth,page=96]{appendix/allmeta_rma_vs_mas}\\
+		\caption[]{figure}
+	\end{figure}
+
+	\begin{figure}[htpb]
+		\ContinuedFloat
+		\captionsetup{list=off,format=cont}
+		\centering
+		\includegraphics[width=0.32\linewidth,page=97]{appendix/allmeta_rma_vs_mas}
+		\includegraphics[width=0.32\linewidth,page=98]{appendix/allmeta_rma_vs_mas}
+		\includegraphics[width=0.32\linewidth,page=99]{appendix/allmeta_rma_vs_mas}\\
+		\includegraphics[width=0.32\linewidth,page=100]{appendix/allmeta_rma_vs_mas}
+		\includegraphics[width=0.32\linewidth,page=101]{appendix/allmeta_rma_vs_mas}
+		\includegraphics[width=0.32\linewidth,page=102]{appendix/allmeta_rma_vs_mas}\\
+		\includegraphics[width=0.32\linewidth,page=103]{appendix/allmeta_rma_vs_mas}
+		\includegraphics[width=0.32\linewidth,page=104]{appendix/allmeta_rma_vs_mas}
+		\includegraphics[width=0.32\linewidth,page=105]{appendix/allmeta_rma_vs_mas}\\
+		\includegraphics[width=0.32\linewidth,page=106]{appendix/allmeta_rma_vs_mas}
+		\includegraphics[width=0.32\linewidth,page=107]{appendix/allmeta_rma_vs_mas}
+		\includegraphics[width=0.32\linewidth,page=108]{appendix/allmeta_rma_vs_mas}\\
+		\caption[]{figure}
+	\end{figure}
+
+	\newpage
+
+	\section{Additional results used to determine the directionality of the pathway metagenes}
+	\label{sec:additional_results_to_determine_the_directionality_of_the_pathway_metagenes}
+
+	The directionality of the pathway metagenes were determined by comparing the metagene values with the gene expression of the representative gene for the pathway associated genetic signature, as well as the gene expression pattern of the pathway signature (\cref{fig:appendix/gt_meta_rma_mas}).
+	Together with the original results presented by \citet{Gatza2010a} (\cref{sec:result_from_gatza2010a_study}), the final directions of the pathway metagenes were decided.
+
+	\begin{figure}[htp!]
+		\centering
+		\includegraphics[width=0.32\linewidth,page=1]{appendix/gtrmastdrank}
+		\includegraphics[width=0.32\linewidth,page=2]{appendix/gtrmastdrank}
+		\includegraphics[width=0.32\linewidth,page=3]{appendix/gtrmastdrank}\\
+		\includegraphics[width=0.32\linewidth,page=4]{appendix/gtrmastdrank}
+		\includegraphics[width=0.32\linewidth,page=5]{appendix/gtrmastdrank}
+		\includegraphics[width=0.32\linewidth,page=6]{appendix/gtrmastdrank}\\
+		\caption[]{Heatmaps showing the gene expression of the pathway associated genetic signatures, pathway metagenes, and the expression of the representative pathway gene in \gls{rma} normalised GT data set.
+		For each pathway associated genetic signature, the gene expression pattern of the signature was plotted in the heatmap.
+		The expression of the representative gene for the pathway signature (see \cref{tab:metagene_direction} in \cref{sub:metagene_direction}) was plotted seaparately as a bar above the pathway metagene scores.
+		}
+		\label{fig:appendix/gt_meta_rma_mas}
+	\end{figure}
+
+	\begin{figure}[htpb]
+		\ContinuedFloat
+		\captionsetup{list=off,format=cont}
+		\centering
+		\includegraphics[width=0.32\linewidth,page=7]{appendix/gtrmastdrank}
+		\includegraphics[width=0.32\linewidth,page=8]{appendix/gtrmastdrank}
+		\includegraphics[width=0.32\linewidth,page=9]{appendix/gtrmastdrank}\\
+		\includegraphics[width=0.32\linewidth,page=10]{appendix/gtrmastdrank}
+		\includegraphics[width=0.32\linewidth,page=11]{appendix/gtrmastdrank}
+		\includegraphics[width=0.32\linewidth,page=12]{appendix/gtrmastdrank}\\
+		\includegraphics[width=0.32\linewidth,page=13]{appendix/gtrmastdrank}
+		\includegraphics[width=0.32\linewidth,page=14]{appendix/gtrmastdrank}
+		\includegraphics[width=0.32\linewidth,page=15]{appendix/gtrmastdrank}\\
+		\includegraphics[width=0.32\linewidth,page=16]{appendix/gtrmastdrank}
+		\includegraphics[width=0.32\linewidth,page=17]{appendix/gtrmastdrank}
+		\includegraphics[width=0.32\linewidth,page=18]{appendix/gtrmastdrank}\\
+		\caption[]{figure}
+	\end{figure}
+
+	\newpage
+
+	\section{Original result from \citet{Gatza2010a} study}
+	\label{sec:result_from_gatza2010a_study}
+
+	The original result from the \citet{Gatza2010a} study were used to determine the directions of the GT pathway associated genetic signatures in \cref{sec:pathway_associated_genetic_signatures_from_gatza2010a_study} (\cref{fig:gatza_paper_res}).
 
@@
 		\includegraphics[width=0.7\linewidth]{results2/gatzares1}
-		\caption{results from gatza's study}
+		\caption[]{Heatmap showing the correlation of all of the GT pathway associated genetic signatures with one another.
+		Figure adapted from the \citet{Gatza2010a} study.}
 		\label{fig:gatza_paper_res}
@@
 
-	% [1] "crolgene"
+	\section{Correlation of the GT pathway metagenes in other cancer data sets}
+	\label{sec:correlation_of_the_gt_pathway_metagenes_in_other_cancer_data_sets}
 
-	% Call:
-	% lm(formula = as.formula(formula_txt[j]), data = data)
+	The correlation of the pathway associated genetic signatures were plotted as heatmaps for the CR, FM and \gls{nzbc} data sets to confirm that the pathway transformation matrices from the GT data set were generating metagenes that were similar to those in the original \gls{rma} normalised GT data.
+	Most importantly, the pathway metagenes generated from the transformation matrices had to cluster in a similar pattern as the pathway metagenes in the GT data set.
+	This was to confirm that the directionality of the metagenes were conserved even after the transformation matrices were applied to the data set
 
-	% Coefficients:
-	%              Estimate Std. Error t value Pr(>|t|)
-	% (Intercept)         0.553214  0.123891 4.465  2.16e-05 ***
-	% bmi                 -0.001629 0.004071 -0.400 0.69
-	% (Intercept)         0.55700   0.05464  10.194 <2e-16   ***
-	% bmiStatusobese      -0.10551  0.07021  -1.503 0.136
-	% bmiStatusoverweight -0.02165  0.07727  -0.280 0.780
-	% (Intercept)         0.307513  0.163530 1.880  0.0631   .
-	% bmi                 0.011279  0.006976 1.617  0.1092
-	% bmiStatusobese      -0.261145 0.118795 -2.198 0.0304   *
-	% bmiStatusoverweight -0.079845 0.084661 -0.943 0.3480
-	% (Intercept)         0.007141  0.382725 0.019  0.9852
-	% bcat_probes         0.156062  0.223779 0.697  0.4873
-	% er_probes           0.123403  0.254800 0.484  0.6293
-	% ifna_probes         -0.430229 0.497301 -0.865 0.3892
-	% ifng_probes         0.378437  0.509352 0.743  0.4594
-	% myc_probes          0.160785  0.229142 0.702  0.4846
-	% pr_probes           0.597403  0.244252 2.446  0.0164   *
-	% (Intercept)         0.055088  0.389472 0.141  0.8878
-	% bmi                 -0.002745 0.003814 -0.720 0.4736
-	% bcat_probes         0.179900  0.226799 0.793  0.4297
-	% er_probes           0.125413  0.255485 0.491  0.6247
-	% ifna_probes         -0.475417 0.502546 -0.946 0.3466
-	% ifng_probes         0.429131  0.515526 0.832  0.4074
-	% myc_probes          0.184642  0.232124 0.795  0.4284
-	% pr_probes           0.607982  0.245336 2.478  0.0151   *
-	% (Intercept)         0.01666   0.37989  0.044  0.9651
-	% bmiStatusobese      -0.11197  0.06602  -1.696 0.0933   .
-	% bmiStatusoverweight -0.02348  0.07151  -0.328 0.7434
-	% bcat_probes         0.22035   0.22616  0.974  0.3325
-	% er_probes           0.10450   0.25330  0.413  0.6809
-	% ifna_probes         -0.60343  0.50372  -1.198 0.2341
-	% ifng_probes         0.54671   0.51484  1.062  0.2911
-	% myc_probes          0.21341   0.22998  0.928  0.3559
-	% pr_probes           0.59491   0.24257  2.453  0.0161   *
-	% (Intercept)         -0.128149 0.393204 -0.326 0.7453
-	% bmi                 0.008834  0.006565 1.346  0.1819
-	% bmiStatusobese      -0.234634 0.112386 -2.088 0.0397   *
-	% bmiStatusoverweight -0.067942 0.078484 -0.866 0.3890
-	% bcat_probes         0.216892  0.225167 0.963  0.3380
-	% er_probes           0.081506  0.252745 0.322  0.7478
-	% ifna_probes         -0.625864 0.501740 -1.247 0.2155
-	% ifng_probes         0.546873  0.512538 1.067  0.2889
-	% myc_probes          0.196215  0.229305 0.856  0.3945
-	% pr_probes           0.560706  0.242816 2.309  0.0232   *
+	\begin{figure}[htp!]
+		\centering
+		\includegraphics[width=0.65\linewidth,page=3]{appendix/gatza_meta_trans}
+		\includegraphics[width=0.65\linewidth,page=7]{appendix/gatza_meta_trans}
+		\caption[]{Heatmaps showing the correlation of all of the pathway metagenes with one another in \gls{rma} normalised CR, \gls{nzbc} and FM data sets.
+		}
+		\label{fig:appendix/gt_meta_rma_mas}
+	\end{figure}
 
-	% [1] "resobsgene"
+	\begin{figure}[htpb]
+		\ContinuedFloat
+		\captionsetup{list=off,format=cont}
+		\centering
+		\includegraphics[width=0.7\linewidth,page=11]{appendix/gatza_meta_trans}
+		\caption[]{figure}
+	\end{figure}
 
-	% Call:
-	% lm(formula = as.formula(formula_txt[j]), data = data)
+	\section{Directionality of the obesity metagenes in the GT data}
+	\label{sec:directionality_of_the_obesity_metagenes_in_the_gt_data}
 
-	% Coefficients:
-	%              Estimate Std. Error t value Pr(>|t|)
-	% (Intercept)         0.547104  0.123915 4.415  2.62e-05 ***
-	% bmi                 -0.001422 0.004072 -0.349 0.728
-	% (Intercept)         0.547619  0.054641 10.022 <2e-16   ***
-	% bmiStatusobese      -0.096832 0.070212 -1.379 0.171
-	% bmiStatusoverweight -0.001804 0.077273 -0.023 0.981
-	% (Intercept)         0.298589  0.163546 1.826  0.0710   .
-	% bmi                 0.011259  0.006976 1.614  0.1099
-	% bmiStatusobese      -0.252186 0.118807 -2.123 0.0364   *
-	% bmiStatusoverweight -0.059898 0.084669 -0.707 0.4810
-	% (Intercept)         0.18694   0.36963  0.506  0.6142
-	% bcat_probes         0.09717   0.21612  0.450  0.6541
-	% er_probes           0.07879   0.24608  0.320  0.7496
-	% ifna_probes         -0.17576  0.48029  -0.366 0.7152
-	% ifng_probes         0.07930   0.49193  0.161  0.8723
-	% myc_probes          -0.04517  0.22130  -0.204 0.8387
-	% pr_probes           0.59554   0.23590  2.525  0.0133   *
-	% (Intercept)         0.221757  0.376612 0.589  0.5574
-	% bmi                 -0.001994 0.003688 -0.540 0.5902
-	% bcat_probes         0.114479  0.219311 0.522  0.6029
-	% er_probes           0.080247  0.247050 0.325  0.7461
-	% ifna_probes         -0.208580 0.485954 -0.429 0.6688
-	% ifng_probes         0.116119  0.498505 0.233  0.8163
-	% myc_probes          -0.027845 0.224460 -0.124 0.9015
-	% pr_probes           0.603222  0.237235 2.543  0.0127   *
-	% (Intercept)         0.195965  0.367384 0.533  0.5951
-	% bmiStatusobese      -0.096834 0.063844 -1.517 0.1328
-	% bmiStatusoverweight -0.004412 0.069155 -0.064 0.9493
-	% bcat_probes         0.150379  0.218719 0.688  0.4935
-	% er_probes           0.058880  0.244961 0.240  0.8106
-	% ifna_probes         -0.344126 0.487134 -0.706 0.4817
-	% ifng_probes         0.242697  0.497893 0.487  0.6271
-	% myc_probes          -0.001285 0.222407 -0.006 0.9954
-	% pr_probes           0.591191  0.234582 2.520  0.0135   *
-	% (Intercept)         0.041459  0.379420 0.109  0.9132
-	% bmi                 0.009426  0.006335 1.488  0.1403
-	% bmiStatusobese      -0.227717 0.108446 -2.100 0.0386   *
-	% bmiStatusoverweight -0.051853 0.075733 -0.685 0.4953
-	% bcat_probes         0.146687  0.217273 0.675  0.5013
-	% er_probes           0.034350  0.243884 0.141  0.8883
-	% ifna_probes         -0.368065 0.484150 -0.760 0.4491
-	% ifng_probes         0.242870  0.494570 0.491  0.6246
-	% myc_probes          -0.019634 0.221267 -0.089 0.9295
-	% pr_probes           0.554701  0.234304 2.367  0.0201   *
+	The directionality of the CR obesity associated genetic signatures were determined in the \gls{rma} normalised GT data set to make sure these signatures were in the ``correct'' direction, and therefore able to be compared with the pathway associated genetic signatures appropriately.
+	As clearly shown in \cref{fig:appendix/ob_meta_dir_gt}, the Res metagene was flipped as the Res metagene was in the oppsite direction compared with the other obesity metagenes.
 
-	% [1] "rescrolgene"
+	\begin{figure}[htp!]
+		\centering
+		\includegraphics[width=0.32\linewidth,page=1]{appendix/gtobrma}
+		\includegraphics[width=0.32\linewidth,page=2]{appendix/gtobrma}
+		\includegraphics[width=0.32\linewidth,page=3]{appendix/gtobrma}\\
+		\includegraphics[width=0.32\linewidth,page=4]{appendix/gtobrma}
+		\includegraphics[width=0.32\linewidth,page=5]{appendix/gtobrma}
+		\includegraphics[width=0.32\linewidth,page=6]{appendix/gtobrma}\\
+		\includegraphics[width=0.32\linewidth,page=7]{appendix/gtobrma}
+		\includegraphics[width=0.32\linewidth,page=8]{appendix/gtobrma}
+		\includegraphics[width=0.32\linewidth,page=9]{appendix/gtobrma}\\
+		\caption[]{Heatmaps showing the obesity metagenes from the CR data set with the gene expressions of the obesity associated genetic signatures in the \gls{rma} normalised GT data set.
+		}
+		\label{fig:appendix/ob_meta_dir_gt}
+	\end{figure}
 
-	% Call:
-	% lm(formula = as.formula(formula_txt[j]), data = data)
+	\newpage
 
-	% Coefficients:
-	%               Estimate Std. Error t value Pr(>|t|)
-	% (Intercept)         0.5343899  0.1239554 4.311  3.9e-05 ***
-	% bmi                 -0.0009922 0.0040730 -0.244 0.808
-	% (Intercept)         0.54870    0.05482   10.009 <2e-16  ***
-	% bmiStatusobese      -0.09063   0.07044   -1.287 0.201
-	% bmiStatusoverweight -0.01515   0.07753   -0.195 0.845
-	% (Intercept)         0.304200   0.164179  1.853  0.0670  .
-	% bmi                 0.011054   0.007003  1.578  0.1178
-	% bmiStatusobese      -0.243160  0.119267  -2.039 0.0442  *
-	% bmiStatusoverweight -0.072189  0.084997  -0.849 0.3978
-	% (Intercept)         0.17069    0.37929   0.450  0.6537
-	% bcat_probes         0.09532    0.22177   0.430  0.6683
-	% er_probes           0.03171    0.25251   0.126  0.9004
-	% ifna_probes         -0.57965   0.49283   -1.176 0.2426
-	% ifng_probes         0.50919    0.50478   1.009  0.3157
-	% myc_probes          0.06345    0.22708   0.279  0.7805
-	% pr_probes           0.54200    0.24206   2.239  0.0276  *
-	% (Intercept)         0.204758   0.386507  0.530  0.5976
-	% bmi                 -0.001950  0.003785  -0.515 0.6076
-	% bcat_probes         0.112257   0.225072  0.499  0.6192
-	% er_probes           0.033134   0.253540  0.131  0.8963
-	% ifna_probes         -0.611750  0.498720  -1.227 0.2231
-	% ifng_probes         0.545210   0.511601  1.066  0.2894
-	% myc_probes          0.080405   0.230357  0.349  0.7279
-	% pr_probes           0.549516   0.243468  2.257  0.0264  *
-	% (Intercept)         0.17933    0.37777   0.475  0.636
-	% bmiStatusobese      -0.09681   0.06565   -1.475 0.144
-	% bmiStatusoverweight -0.01219   0.07111   -0.171 0.864
-	% bcat_probes         0.14969    0.22490   0.666  0.507
-	% er_probes           0.01354    0.25189   0.054  0.957
-	% ifna_probes         -0.73888   0.50091   -1.475 0.144
-	% ifng_probes         0.66381    0.51197   1.297  0.198
-	% myc_probes          0.10813    0.22869   0.473  0.638
-	% pr_probes           0.53872    0.24121   2.233  0.028   *
-	% (Intercept)         0.027729   0.390580  0.071  0.9436
-	% bmi                 0.009248   0.006522  1.418  0.1596
-	% bmiStatusobese      -0.225233  0.111636  -2.018 0.0466  *
-	% bmiStatusoverweight -0.058734  0.077961  -0.753 0.4532
-	% bcat_probes         0.146066   0.223664  0.653  0.5154
-	% er_probes           -0.010526  0.251058  -0.042 0.9667
-	% ifna_probes         -0.762373  0.498391  -1.530 0.1296
-	% ifng_probes         0.663981   0.509117  1.304  0.1955
-	% myc_probes          0.090122   0.227775  0.396  0.6933
-	% pr_probes           0.502919   0.241196  2.085  0.0399  *
+	\section{Comparison of the correlation of \gls{svd} and TM generated pathway metagenes}
+	\label{sec:comparison_of_the_correlation_of_svd_and_tm_generated_pathway_metagenes}
 
-	% [1] "caobsgene"
+	This section provides visual comparison of the Spearman correlation of the \gls{svd}- and TM-generated pathway and obesity metagenes in the CR, FM, GT and \gls{nzbc} data sets (\cref{fig:allmetacor_bar_appendix}).
+	The Akt, \gls{egfr}, \gls{her2}, p53, p63, \gls{pi3k}, Ras, Src, \gls{stat3}, \gls{tgfb} and \gls{tnfa} pathway signatures showed a correlation of 0.8 or less in at least one of the four data sets.
 
-	% Call:
-	% lm(formula = as.formula(formula_txt[j]), data = data)
+	\begin{figure}[htpb]
+		\centering
+		\includegraphics[width=0.24\linewidth,page=1]{results2/allmetacor_bar}
+		\includegraphics[width=0.24\linewidth,page=2]{results2/allmetacor_bar}
+		\includegraphics[width=0.24\linewidth,page=3]{results2/allmetacor_bar}
+		\includegraphics[width=0.24\linewidth,page=4]{results2/allmetacor_bar}\\
+		\caption[]{Visual comparison of the Spearman correlations of the \gls{svd}-derived and transformation matrix-derived pathway and obesity metagenes across different data sets}
+		\label{fig:allmetacor_bar_appendix}
+	\end{figure}
 
-	% Coefficients:
-	%              Estimate Std. Error t value Pr(>|t|)
-	% (Intercept)         0.538968  0.123943 4.349  3.39e-05 ***
-	% bmi                 -0.001147 0.004073 -0.282 0.779
-	% (Intercept)         0.53896   0.05458  9.875  2.76e-16 ***
-	% bmiStatusobese      -0.08982  0.07013  -1.281 0.203
-	% bmiStatusoverweight 0.01804   0.07719  0.234  0.816
-	% (Intercept)         0.27986   0.16317  1.715  0.0896   .
-	% bmi                 0.01171   0.00696  1.683  0.0957   .
-	% bmiStatusobese      -0.25146  0.11853  -2.121 0.0365   *
-	% bmiStatusoverweight -0.04241  0.08447  -0.502 0.6168
-	% (Intercept)         0.12454   0.38803  0.321  0.7490
-	% bcat_probes         0.14068   0.22688  0.620  0.5367
-	% er_probes           0.11597   0.25833  0.449  0.6545
-	% ifna_probes         -0.33319  0.50420  -0.661 0.5104
-	% ifng_probes         0.30871   0.51641  0.598  0.5515
-	% myc_probes          -0.04569  0.23232  -0.197 0.8445
-	% pr_probes           0.56692   0.24764  2.289  0.0243   *
-	% (Intercept)         0.162063  0.395326 0.410  0.6828
-	% bmi                 -0.002148 0.003872 -0.555 0.5804
-	% bcat_probes         0.159335  0.230208 0.692  0.4906
-	% er_probes           0.117547  0.259326 0.453  0.6514
-	% ifna_probes         -0.368552 0.510101 -0.723 0.4718
-	% ifng_probes         0.348376  0.523276 0.666  0.5072
-	% myc_probes          -0.027017 0.235613 -0.115 0.9090
-	% pr_probes           0.575201  0.249024 2.310  0.0232   *
-	% (Intercept)         0.134514  0.384960 0.349  0.7276
-	% bmiStatusobese      -0.099695 0.066899 -1.490 0.1397
-	% bmiStatusoverweight 0.008987  0.072463 0.124  0.9016
-	% bcat_probes         0.193430  0.229183 0.844  0.4009
-	% er_probes           0.092451  0.256681 0.360  0.7196
-	% ifna_probes         -0.522336 0.510440 -1.023 0.3089
-	% ifng_probes         0.492130  0.521713 0.943  0.3481
-	% myc_probes          -0.001892 0.233048 -0.008 0.9935
-	% pr_probes           0.560583  0.245805 2.281  0.0249   *
-	% (Intercept)         -0.028482 0.397505 -0.072 0.9430
-	% bmi                 0.009944  0.006637 1.498  0.1376
-	% bmiStatusobese      -0.237770 0.113616 -2.093 0.0392   *
-	% bmiStatusoverweight -0.041061 0.079343 -0.518 0.6061
-	% bcat_probes         0.189535  0.227630 0.833  0.4073
-	% er_probes           0.066573  0.255509 0.261  0.7950
-	% ifna_probes         -0.547591 0.507227 -1.080 0.2832
-	% ifng_probes         0.492312  0.518143 0.950  0.3446
-	% myc_probes          -0.021250 0.231813 -0.092 0.9272
-	% pr_probes           0.522087  0.245472 2.127  0.0362   *
+	\begin{figure}[htpb]
+		\ContinuedFloat
+		\captionsetup{list=off,format=cont}
+		\centering
+		\includegraphics[width=0.24\linewidth,page=5]{results2/allmetacor_bar}
+		\includegraphics[width=0.24\linewidth,page=6]{results2/allmetacor_bar}
+		\includegraphics[width=0.24\linewidth,page=7]{results2/allmetacor_bar}
+		\includegraphics[width=0.24\linewidth,page=8]{results2/allmetacor_bar}\\
+		\includegraphics[width=0.24\linewidth,page=9]{results2/allmetacor_bar}
+		\includegraphics[width=0.24\linewidth,page=10]{results2/allmetacor_bar}
+		\includegraphics[width=0.24\linewidth,page=11]{results2/allmetacor_bar}
+		\includegraphics[width=0.24\linewidth,page=12]{results2/allmetacor_bar}\\
+		\includegraphics[width=0.24\linewidth,page=13]{results2/allmetacor_bar}
+		\includegraphics[width=0.24\linewidth,page=14]{results2/allmetacor_bar}
+		\includegraphics[width=0.24\linewidth,page=15]{results2/allmetacor_bar}
+		\includegraphics[width=0.24\linewidth,page=16]{results2/allmetacor_bar}\\
+		\includegraphics[width=0.24\linewidth,page=17]{results2/allmetacor_bar}
+		\includegraphics[width=0.24\linewidth,page=18]{results2/allmetacor_bar}
+		\includegraphics[width=0.24\linewidth,page=19]{results2/allmetacor_bar}
+		\includegraphics[width=0.24\linewidth,page=20]{results2/allmetacor_bar}\\
+		\includegraphics[width=0.24\linewidth,page=21]{results2/allmetacor_bar}
+		\includegraphics[width=0.24\linewidth,page=22]{results2/allmetacor_bar}
+		\includegraphics[width=0.24\linewidth,page=23]{results2/allmetacor_bar}
+		\includegraphics[width=0.24\linewidth,page=24]{results2/allmetacor_bar}\\
+		\includegraphics[width=0.24\linewidth,page=25]{results2/allmetacor_bar}
+		\includegraphics[width=0.24\linewidth,page=26]{results2/allmetacor_bar}
+		\includegraphics[width=0.24\linewidth,page=27]{results2/allmetacor_bar}
+		\includegraphics[width=0.24\linewidth,page=28]{results2/allmetacor_bar}\\
+		\caption{figure}
+	\end{figure}
+	\newpage
 
-	% [1] "cacrolgene"
+	\section{Correlation of the pathway and obesity metagenes in other cancer data sets}
+	\label{sec:correlation_of_the_pathway_and_obesity_metagenes_in_other_cancer_data_sets}
 
-	% Call:
-	% lm(formula = as.formula(formula_txt[j]), data = data)
+	Correlation of the pathway and obesity associated genetic signatures were plotted in a heatmap for the GT data in \cref{sec:pathway_associated_metagenes_and_obesity_associated_metagenes}.
+	In this section, the correlation of these genetic signatures were plotted as heatmaps for the CR, FM and \gls{nzbc} data sets.
+	Though each of the heatmaps show slightly different clustering patterns, the results were similar to that in the GT data set; the CR obesity metagenes clustered together in a single group, and FM metagene did not cluster with any other genetic signatures.
 
-	% Coefficients:
-	%               Estimate Std. Error t value Pr(>|t|)
-	% (Intercept)         5.062e-01  1.240e-01 4.083  9.16e-05 ***
-	% bmi                 -3.953e-05 4.074e-03 -0.010 0.992
-	% (Intercept)         0.52453    0.05485   9.564  1.29e-15 ***
-	% bmiStatusobese      -0.06411   0.07047   -0.910 0.365
-	% bmiStatusoverweight 0.02958    0.07756   0.381  0.704
-	% (Intercept)         0.273274   0.164135  1.665  0.0992   .
-	% bmi                 0.011360   0.007001  1.622  0.1080
-	% bmiStatusobese      -0.220856  0.119235  -1.852 0.0671   .
-	% bmiStatusoverweight -0.029032  0.084974  -0.342 0.7334
-	% (Intercept)         0.02932    0.40048   0.073  0.9418
-	% bcat_probes         0.23660    0.23416   1.010  0.3150
-	% er_probes           0.09521    0.26662   0.357  0.7218
-	% ifna_probes         -0.49129   0.52037   -0.944 0.3476
-	% ifng_probes         0.44466    0.53298   0.834  0.4063
-	% myc_probes          0.16111    0.23977   0.672  0.5033
-	% pr_probes           0.49566    0.25558   1.939  0.0555   .
-	% (Intercept)         0.051698   0.408467  0.127  0.8996
-	% bmi                 -0.001281  0.004000  -0.320 0.7495
-	% bcat_probes         0.247720   0.237861  1.041  0.3004
-	% er_probes           0.096149   0.267946  0.359  0.7205
-	% ifna_probes         -0.512383  0.527057  -0.972 0.3335
-	% ifng_probes         0.468320   0.540670  0.866  0.3887
-	% myc_probes          0.172240   0.243445  0.708  0.4811
-	% pr_probes           0.500601   0.257301  1.946  0.0548   .
-	% (Intercept)         0.03806    0.39931   0.095  0.9243
-	% bmiStatusobese      -0.07973   0.06939   -1.149 0.2536
-	% bmiStatusoverweight 0.02249    0.07516   0.299  0.7655
-	% bcat_probes         0.27648    0.23772   1.163  0.2479
-	% er_probes           0.07298    0.26625   0.274  0.7846
-	% ifna_probes         -0.66043   0.52946   -1.247 0.2155
-	% ifng_probes         0.60854    0.54116   1.125  0.2638
-	% myc_probes          0.19456    0.24173   0.805  0.4230
-	% pr_probes           0.48849    0.25497   1.916  0.0586   .
-	% (Intercept)         -0.123986  0.412743  -0.300 0.7646
-	% bmi                 0.009886   0.006892  1.434  0.1549
-	% bmiStatusobese      -0.216997  0.117971  -1.839 0.0692   .
-	% bmiStatusoverweight -0.027267  0.082384  -0.331 0.7414
-	% bcat_probes         0.272606   0.236356  1.153  0.2518
-	% er_probes           0.047249   0.265304  0.178  0.8591
-	% ifna_probes         -0.685542  0.526672  -1.302 0.1964
-	% ifng_probes         0.608724   0.538006  1.131  0.2609
-	% myc_probes          0.175314   0.240700  0.728  0.4683
-	% pr_probes           0.450215   0.254882  1.766  0.0808   .
+	\begin{figure}[htp!]
+		\centering
+		\includegraphics[width=0.7\linewidth,page=3]{appendix/all_meta_trans}
+		\caption[]{Heatmaps showing the correlation of all of the pathway and obesity metagenes with one another in \gls{rma} normalised CR, \gls{nzbc} and FM data sets.
+		}
+		\label{fig:appendix/gt_meta_rma_mas}
+	\end{figure}
 
-	% [1] "caresobsgene"
+	\begin{figure}[htpb]
+		\ContinuedFloat
+		\captionsetup{list=off,format=cont}
+		\centering
+		\includegraphics[width=0.7\linewidth,page=7]{appendix/all_meta_trans}
+		\includegraphics[width=0.7\linewidth,page=11]{appendix/all_meta_trans}
+		\caption[]{figure}
+	\end{figure}
 
-	% Call:
-	% lm(formula = as.formula(formula_txt[j]), data = data)
+	\newpage
 
-	% Coefficients:
-	%             Estimate Std. Error t value Pr(>|t|)
-	% (Intercept)         0.462424  0.123913 3.732  0.00032  ***
-	% bmi                 0.001442  0.004072 0.354  0.72408
-	% (Intercept)         0.46970   0.05452  8.616  1.39e-13 ***
-	% bmiStatusobese      0.09338   0.07005  1.333  0.186
-	% bmiStatusoverweight -0.01840  0.07710  -0.239 0.812
-	% (Intercept)         0.722057  0.163100 4.427  2.55e-05 ***
-	% bmi                 -0.011409 0.006957 -1.640 0.1043
-	% bmiStatusobese      0.250806  0.118483 2.117  0.0369   *
-	% bmiStatusoverweight 0.040473  0.084438 0.479  0.6328
-	% (Intercept)         1.1017    0.3815   2.888  0.00483  **
-	% bcat_probes         -0.2217   0.2230   -0.994 0.32284
-	% er_probes           -0.2308   0.2540   -0.909 0.36579
-	% ifna_probes         0.1853    0.4957   0.374  0.70938
-	% ifng_probes         -0.1528   0.5077   -0.301 0.76408
-	% myc_probes          -0.0737   0.2284   -0.323 0.74765
-	% pr_probes           -0.6876   0.2434   -2.825 0.00581  **
-	% (Intercept)         1.056654  0.388317 2.721  0.00780  **
-	% bmi                 0.002580  0.003803 0.678  0.49929
-	% bcat_probes         -0.244104 0.226127 -1.079 0.28322
-	% er_probes           -0.232713 0.254728 -0.914 0.36335
-	% ifna_probes         0.227770  0.501056 0.455  0.65049
-	% ifng_probes         -0.200460 0.513998 -0.390 0.69745
-	% myc_probes          -0.096122 0.231435 -0.415 0.67888
-	% pr_probes           -0.697585 0.244608 -2.852 0.00538  **
-	% (Intercept)         1.09174   0.37821  2.887  0.00488  **
-	% bmiStatusobese      0.10153   0.06573  1.545  0.12592
-	% bmiStatusoverweight -0.00544  0.07119  -0.076 0.93926
-	% bcat_probes         -0.27598  0.22517  -1.226 0.22351
-	% er_probes           -0.20770  0.25218  -0.824 0.41234
-	% ifna_probes         0.37360   0.50149  0.745  0.45823
-	% ifng_probes         -0.33545  0.51257  -0.654 0.51449
-	% myc_probes          -0.11868  0.22896  -0.518 0.60548
-	% pr_probes           -0.68170  0.24150  -2.823 0.00586  **
-	% (Intercept)         1.235747  0.391476 3.157  0.00218  **
-	% bmi                 -0.008785 0.006537 -1.344 0.18237
-	% bmiStatusobese      0.223518  0.111892 1.998  0.04881  *
-	% bmiStatusoverweight 0.038776  0.078139 0.496  0.62094
-	% bcat_probes         -0.272540 0.224177 -1.216 0.22730
-	% er_probes           -0.184837 0.251634 -0.735 0.46455
-	% ifna_probes         0.395907  0.499535 0.793  0.43015
-	% ifng_probes         -0.335611 0.510285 -0.658 0.51243
-	% myc_probes          -0.101582 0.228298 -0.445 0.65743
-	% pr_probes           -0.647688 0.241749 -2.679 0.00879  **
+	\section{Summary of the remainder of the linear models in \gls{nzbc} data}
+	\label{sec:summary_of_the_linear_models_in_nzbc_data}
 
-	% [1] "carescrolgene"
+	All of the linear models for the other CR and FM obesity metagenes were summarised in this section.
+	In all of the CR data set-derived obesity metagenes, the \gls{pr} metagene showed up as a significant variable in the linear models (\cref{tab:lm_sig_var_crol,tab:lm_sig_var_res,tab:lm_sig_var_resol,tab:lm_sig_var_ca,tab:lm_sig_var_caol,tab:lm_sig_var_cares,tab:lm_sig_var_caresol,tab:lm_sig_var_or}).
+	For the linear model that predicted the FM obesity metagene scores, the Myc pathway metagene was significant (\cref{tab:lm_sig_var_fm}).
 
-	% Call:
-	% lm(formula = as.formula(formula_txt[j]), data = data)
-
-	% Coefficients:
-	%              Estimate Std. Error t value Pr(>|t|)
-	% (Intercept)         0.4928357 0.1239867 3.975  0.000136 ***
-	% bmi                 0.0004131 0.0040740 0.101  0.919450
-	% (Intercept)         0.52814   0.05483   9.633  9.13e-16 ***
-	% bmiStatusobese      -0.06913  0.07045   -0.981 0.329
-	% bmiStatusoverweight 0.02453   0.07753   0.316  0.752
-	% (Intercept)         0.232845  0.163204  1.427  0.1569
-	% bmi                 0.013351  0.006962  1.918  0.0582   .
-	% bmiStatusobese      -0.253344 0.118558  -2.137 0.0352   *
-	% bmiStatusoverweight -0.044356 0.084492  -0.525 0.6008
-	% (Intercept)         -0.1205   0.3891    -0.310 0.7576
-	% bcat_probes         0.2439    0.2275    1.072  0.2864
-	% er_probes           0.1383    0.2590    0.534  0.5948
-	% ifna_probes         -0.4364   0.5056    -0.863 0.3903
-	% ifng_probes         0.4301    0.5178    0.831  0.4084
-	% myc_probes          0.2837    0.2330    1.218  0.2264
-	% pr_probes           0.5789    0.2483    2.331  0.0219   *
-	% (Intercept)         -0.102321 0.396926  -0.258 0.7972
-	% bmi                 -0.001038 0.003887  -0.267 0.7900
-	% bcat_probes         0.252943  0.231140  1.094  0.2767
-	% er_probes           0.139029  0.260375  0.534  0.5947
-	% ifna_probes         -0.453502 0.512164  -0.885 0.3782
-	% ifng_probes         0.449266  0.525393  0.855  0.3947
-	% myc_probes          0.292758  0.236566  1.238  0.2191
-	% pr_probes           0.582903  0.250031  2.331  0.0219   *
-	% (Intercept)         -0.11188  0.38790   -0.288 0.7737
-	% bmiStatusobese      -0.08019  0.06741   -1.190 0.2373
-	% bmiStatusoverweight 0.01833   0.07302   0.251  0.8024
-	% bcat_probes         0.28469   0.23093   1.233  0.2209
-	% er_probes           0.11686   0.25864   0.452  0.6525
-	% ifna_probes         -0.60153  0.51434   -1.170 0.2453
-	% ifng_probes         0.59011   0.52570   1.123  0.2646
-	% myc_probes          0.31782   0.23483   1.353  0.1793
-	% pr_probes           0.57227   0.24768   2.311  0.0231   *
-	% (Intercept)         -0.284628 0.400004  -0.712 0.4786
-	% bmi                 0.010538  0.006679  1.578  0.1181
-	% bmiStatusobese      -0.226528 0.114330  -1.981 0.0506   .
-	% bmiStatusoverweight -0.034711 0.079842  -0.435 0.6648
-	% bcat_probes         0.280559  0.229061  1.225  0.2239
-	% er_probes           0.089437  0.257116  0.348  0.7288
-	% ifna_probes         -0.628292 0.510417  -1.231 0.2216
-	% ifng_probes         0.590302  0.521401  1.132  0.2606
-	% myc_probes          0.297308  0.233271  1.275  0.2058
-	% pr_probes           0.531475  0.247015  2.152  0.0341   *
-
-	% [1] "crobsgenes"
-
-	% Call:
-	% lm(formula = as.formula(formula_txt[j]), data = data)
-
-	% Coefficients:
-	%             Estimate Std. Error t value Pr(>|t|)
-	% (Intercept)         0.456126  0.123888 3.682  0.000381 ***
-	% bmi                 0.001654  0.004071 0.406  0.685322
-	% (Intercept)         0.459596  0.054606 8.417  3.7e-13  ***
-	% bmiStatusobese      0.100893  0.070167 1.438  0.154
-	% bmiStatusoverweight 0.005772  0.077224 0.075  0.941
-	% (Intercept)         0.704210  0.163518 4.307  4.04e-05 ***
-	% bmi                 -0.011059 0.006975 -1.586 0.1162
-	% bmiStatusobese      0.253491  0.118786 2.134  0.0354   *
-	% bmiStatusoverweight 0.062836  0.084654 0.742  0.4598
-	% (Intercept)         1.0667    0.3806   2.803  0.00618  **
-	% bcat_probes         -0.2284   0.2225   -1.026 0.30740
-	% er_probes           -0.1481   0.2534   -0.584 0.56041
-	% ifna_probes         0.4504    0.4945   0.911  0.36477
-	% ifng_probes         -0.4013   0.5065   -0.792 0.43018
-	% myc_probes          -0.1402   0.2279   -0.615 0.53986
-	% pr_probes           -0.6446   0.2429   -2.654 0.00938  **
-	% (Intercept)         1.014279  0.387058 2.620  0.01029  *
-	% bmi                 0.003004  0.003791 0.792  0.43014
-	% bcat_probes         -0.254482 0.225394 -1.129 0.26184
-	% er_probes           -0.150261 0.253902 -0.592 0.55545
-	% ifna_probes         0.499868  0.499432 1.001  0.31954
-	% ifng_probes         -0.456812 0.512332 -0.892 0.37494
-	% myc_probes          -0.166314 0.230685 -0.721 0.47278
-	% pr_probes           -0.656157 0.243815 -2.691 0.00847  **
-	% (Intercept)         1.05662   0.37693  2.803  0.0062   **
-	% bmiStatusobese      0.11346   0.06550  1.732  0.0867   .
-	% bmiStatusoverweight 0.01412   0.07095  0.199  0.8427
-	% bcat_probes         -0.29209  0.22440  -1.302 0.1964
-	% er_probes           -0.12674  0.25133  -0.504 0.6153
-	% ifna_probes         0.63723   0.49979  1.275  0.2056
-	% ifng_probes         -0.58273  0.51083  -1.141 0.2570
-	% myc_probes          -0.19254  0.22819  -0.844 0.4010
-	% pr_probes           -0.64072  0.24068  -2.662 0.0092   **
-	% (Intercept)         1.198471  0.390241 3.071  0.00283  **
-	% bmi                 -0.008654 0.006516 -1.328 0.18754
-	% bmiStatusobese      0.233624  0.111539 2.095  0.03905  *
-	% bmiStatusoverweight 0.057676  0.077893 0.740  0.46097
-	% bcat_probes         -0.288700 0.223470 -1.292 0.19974
-	% er_probes           -0.104217 0.250840 -0.415 0.67880
-	% ifna_probes         0.659208  0.497958 1.324  0.18895
-	% ifng_probes         -0.582884 0.508675 -1.146 0.25491
-	% myc_probes          -0.175698 0.227577 -0.772 0.44214
-	% pr_probes           -0.607216 0.240986 -2.520 0.01353  *
-
-	% [1] "fmobsgenes"
-
-	% Call:
-	% lm(formula = as.formula(formula_txt[j]), data = data)
-
-	% Coefficients:
-	%             Estimate Std. Error t value Pr(>|t|)
-	% (Intercept)         0.436628  0.123787 3.527  0.000643 ***
-	% bmi                 0.002314  0.004067 0.569  0.570763
-	% (Intercept)         0.58369   0.05435  10.740 <2e-16   ***
-	% bmiStatusobese      -0.08452  0.06983  -1.210 0.2291
-	% bmiStatusoverweight -0.14827  0.07686  -1.929 0.0567   .
-	% (Intercept)         0.251170  0.160863 1.561  0.12176
-	% bmi                 0.015034  0.006862 2.191  0.03091  *
-	% bmiStatusobese      -0.291956 0.116858 -2.498 0.01419  *
-	% bmiStatusoverweight -0.225840 0.083280 -2.712 0.00794  **
-	% (Intercept)         0.33317   0.36900  0.903  0.3689
-	% bcat_probes         -0.13489  0.21576  -0.625 0.5334
-	% er_probes           0.11973   0.24567  0.487  0.6271
-	% ifna_probes         -0.56244  0.47947  -1.173 0.2438
-	% ifng_probes         0.58838   0.49109  1.198  0.2340
-	% myc_probes          0.41768   0.22093  1.891  0.0618   .
-	% pr_probes           -0.08815  0.23550  -0.374 0.7090
-	% (Intercept)         0.291390  0.375705 0.776  0.4400
-	% bmi                 0.002392  0.003679 0.650  0.5172
-	% bcat_probes         -0.155669 0.218783 -0.712 0.4786
-	% er_probes           0.117981  0.246455 0.479  0.6333
-	% ifna_probes         -0.523056 0.484783 -1.079 0.2835
-	% ifng_probes         0.544202  0.497304 1.094  0.2767
-	% myc_probes          0.396892  0.223919 1.772  0.0797   .
-	% pr_probes           -0.097371 0.236664 -0.411 0.6817
-	% (Intercept)         0.33459   0.36629  0.913  0.3634
-	% bmiStatusobese      -0.08029  0.06366  -1.261 0.2104
-	% bmiStatusoverweight -0.12492  0.06895  -1.812 0.0733   .
-	% bcat_probes         -0.07254  0.21807  -0.333 0.7402
-	% er_probes           0.13036   0.24424  0.534  0.5948
-	% ifna_probes         -0.56035  0.48569  -1.154 0.2517
-	% ifng_probes         0.58760   0.49642  1.184  0.2397
-	% myc_probes          0.46651   0.22175  2.104  0.0382   *
-	% pr_probes           -0.07506  0.23389  -0.321 0.7490
-	% (Intercept)         0.082260  0.370293 0.222  0.82471
-	% bmi                 0.015393  0.006183 2.490  0.01464  *
-	% bmiStatusobese      -0.294047 0.105838 -2.778 0.00666  **
-	% bmiStatusoverweight -0.202402 0.073911 -2.738 0.00746  **
-	% bcat_probes         -0.078569 0.212047 -0.371 0.71187
-	% er_probes           0.090296  0.238018 0.379  0.70532
-	% ifna_probes         -0.599445 0.472504 -1.269 0.20787
-	% ifng_probes         0.587878  0.482673 1.218  0.22646
-	% myc_probes          0.436540  0.215944 2.022  0.04623  *
-	% pr_probes           -0.134650 0.228668 -0.589 0.55746
-
-\begin{longtable}{llr{\bfseries}S}
+	\begin{table}[htpb]
 		\centering
-	\caption{Table showing the variables in the step wise linear models to predict the obesity metagenes}
-	\label{tab:stepwise_model}\\
-		Obesity metagene & Variables used in the model & Estimate & {P-value}\\
-		\hline
-	\endfirsthead
-	\multicolumn{4}{c}{\tablename\ \thetable{}\ (continued)}\\
+		\caption[]{Description of the linear models used to predict the CrOl obesity metagene in \gls{nzbc} data set}
+		\label{tab:lm_sig_var_crol}
+		\begin{threeparttable}
+			\begin{tabular}{llr{\bfseries}S}
+				Linear Model & Variables & Estimate & P-value\\
 				\hline
 				\hline
-	\endhead
+				\rule{0pt}{2.25ex}\gls{bmi} only                           & \gls{bmi}  & -0.001629 & 0.6900 \\
+				\hline
+				\rule{0pt}{2.25ex}\gls{bmi} status only                    & Overweight & -0.02165  & 0.7800 \\
+                                                                           & Obese      & -0.1055   & 0.1360 \\
+				\hline
+				\rule{0pt}{2.25ex}\gls{bmi} and \gls{bmi} status           & \gls{bmi}  & 0.01128   & 0.1092 \\
+                                                                           & Overweight & -0.07985  & 0.3480 \\
+                                                                           & Obese      & -0.2611   & \bfseries 0.0304  \\
+				\hline
+				\rule{0pt}{2.25ex}Pathways only                            & \gls{bcat} & 0.1561    & 0.4873 \\
+                                                                           & \gls{er}   & 0.1234    & 0.6293 \\
+                                                                           & \gls{ifna} & -0.4302   & 0.3892 \\
+                                                                           & \gls{ifny} & 0.3784    & 0.4594 \\
+                                                                           & Myc        & 0.1608    & 0.4846 \\
+                                                                           & \gls{pr}   & 0.5974    & \bfseries 0.0164  \\
+				\hline
+				\rule{0pt}{2.25ex}\gls{bmi} and Pathways                   & \gls{bmi}  & -0.002745 & 0.4736 \\
+                                                                           & \gls{bcat} & 0.1799    & 0.4297 \\
+                                                                           & \gls{er}   & 0.1254    & 0.6247 \\
+                                                                           & \gls{ifna} & -0.4754   & 0.3466 \\
+                                                                           & \gls{ifny} & 0.4291    & 0.4074 \\
+                                                                           & Myc        & 0.1846    & 0.4284 \\
+                                                                           & \gls{pr}   & 0.6080    & \bfseries 0.0151  \\
+				\hline
+				\rule{0pt}{2.25ex}\gls{bmi} status and Pathways            & Overweight & -0.02348  & 0.7434 \\
+                                                                           & Obese      & -0.1120   & \bfseries 0.0933  \\
+                                                                           & \gls{bcat} & 0.2204    & 0.3325 \\
+                                                                           & \gls{er}   & 0.1045    & 0.6809 \\
+                                                                           & \gls{ifna} & -0.6034   & 0.2341 \\
+                                                                           & \gls{ifny} & 0.5467    & 0.2911 \\
+                                                                           & Myc        & 0.2134    & 0.3559 \\
+                                                                           & \gls{pr}   & 0.5949    & \bfseries 0.0161  \\
+				\hline
+				\rule{0pt}{2.25ex}\gls{bmi}, \gls{bmi} status and pathways & \gls{bmi}  & 0.008834  & 0.1819 \\
+                                                                           & Overweight & -0.06794  & 0.3890 \\
+                                                                           & Obese      & -0.2346   & \bfseries 0.0397  \\
+                                                                           & \gls{bcat} & 0.2169    & 0.3380 \\
+                                                                           & \gls{er}   & 0.08151   & 0.7478 \\
+                                                                           & \gls{ifna} & -0.6259   & 0.2155 \\
+                                                                           & \gls{ifny} & 0.5469    & 0.2889 \\
+                                                                           & Myc        & 0.1962    & 0.3945 \\
+                                                                           & \gls{pr}   & 0.5607    & \bfseries 0.0232  \\
 				\hline
 				\hline
-		Cr         & Akt         & -0.3415   & \bfseries \num{4.43d-8}\\
-				   & \gls{egfr}  & 0.33852   & \bfseries \num{1.94d-05}\\
-				   & \gls{er}    & -0.20237  & \bfseries 0.00419\\
-				   & \gls{ifny}  & 0.12233   & 0.08700\\
-				   & \gls{pr}    & -0.33920  & \bfseries \num{2.09d-5}\\
-				   & Ras         & -0.22611  & \bfseries 0.00193\\
-				   & Src         & -0.11452  & 0.08177\\
-				   & \gls{tgfb}  & 0.55638   & \bfseries \textless{} \num{2d-16}\\
-				   & \gls{tnfa}  & -0.25501  & \bfseries 0.00616\\
-		\hline
-		CrOl       & Overweight  & 0.040918  & 0.074865\\
-				   & Obese       & -0.006381 & 0.756121\\
-				   & Akt         & -0.356713 & \bfseries \num{1.89d-07}\\
-				   & E2F1        & 0.160229  & \bfseries 0.005411\\
-				   & \gls{egfr}  & 0.148618  & 0.062683\\
-				   & p53         & 0.336552  & \bfseries \num{4.35d-05}\\
-				   & \gls{pr}    & -0.225423 & \bfseries 0.000639\\
-				   & Src         & -0.428758 & \bfseries \num{5.85d-10}\\
-				   & \gls{stat3} & -0.124118 & \bfseries 0.009658\\
-		\hline
-		Res        & Akt         & 0.24613   & \bfseries 0.000289\\
-				   & \gls{egfr}  & -0.46827  & \bfseries \num{2.16d-05}\\
-				   & \gls{er}    & 0.18400   & 0.061446\\
-				   & \gls{her2}  & 0.13835   & 0.118228\\
-				   & \gls{ifny}  & -0.21019  & \bfseries 0.044084\\
-				   & \gls{pr}    & 0.41195   & \bfseries 0.000300\\
-				   & Ras         & 0.26506   & \bfseries 0.010465\\
-				   & \gls{tgfb}  & -0.58075  & \bfseries \num{7.64d-11}\\
-				   & \gls{tnfa}  & 0.42200   & \bfseries 0.001834\\
-		\hline
-		ResOl      & Akt         & -0.33596  & \bfseries 0.000294\\
-				   & \gls{bcat}  & -0.19866  & 0.138953\\
-				   & E2F1        & 0.36059   & \bfseries 0.009325\\
-				   & \gls{egfr}  & 0.23421   & 0.056865\\
-				   & \gls{her2}  & -0.13652  & 0.121689\\
-				   & p53         & 0.38746   & \bfseries 0.000519\\
-				   & \gls{pi3k}  & -0.21881  & 0.148676\\
-				   & \gls{pr}    & -0.32228  & \bfseries 0.001656\\
-				   & Ras         & -0.28984  & \bfseries 0.033975\\
-				   & Src         & -0.40779  & \bfseries \num{5.26d-06}\\
-				   & \gls{stat3} & -0.11452  & 0.125609\\
-				   & \gls{tnfa}  & -0.11442  & 0.117023\\
-		\hline
-		Ca         & Akt         & -0.26243  & \bfseries \num{7.04d-06}\\
-				   & \gls{egfr}  & 0.36305   & \bfseries 0.000139\\
-				   & \gls{er}    & -0.18614  & \bfseries 0.030771\\
-				   & \gls{ifny}  & 0.22118   & \bfseries 0.013034\\
-				   & \gls{pr}    & -0.42911  & \bfseries \num{1.07d-05}\\
-				   & Ras         & -0.28124  & \bfseries 0.000898\\
-				   & \gls{tgfb}  & 0.76139   & \bfseries \textless{} \num{2d-16}\\
-				   & \gls{tnfa}  & -0.33200  & \bfseries 0.003919\\
-		\hline
-		CaOl       & Overweight  & 0.07613   & \bfseries 0.006875\\
-				   & Obese       & 0.01408   & 0.564622\\
-				   & Akt         & -0.44727  & \bfseries \num{2.43d-07}\\
-				   & \gls{bcat}  & -0.13463  & 0.140261\\
-				   & E2F1        & 0.33864   & \bfseries 0.001249\\
-				   & p53         & 0.28777   & \bfseries 0.000996\\
-				   & \gls{pr}    & -0.28255  & \bfseries \num{2.18d-05}\\
-				   & Src         & -0.44149  & \bfseries \num{2.02d-08}\\
-				   & \gls{tnfa}  & -0.14215  & \bfseries 0.013897\\
-		\hline
-		CaRes      & \gls{bcat}  & 0.26899   & 0.05238  \\
-				   & E2F1        & -0.22115  & 0.08848  \\
-				   & \gls{egfr}  & 0.44400   & \bfseries \num{6.67d-05 }\\
-				   & \gls{her2}  & -0.20304  & \bfseries 0.04615  \\
-				   & \gls{ifny}  & 0.32874   & \bfseries 0.00574  \\
-				   & \gls{pr}    & -0.28764  & \bfseries 0.00104  \\
-				   & Src         & -0.17989  & 0.05404  \\
-				   & \gls{tgfb}  & 0.52898   & \bfseries \num{2.15d-06 }\\
-				   & \gls{tnfa}  & -0.38677  & \bfseries 0.00564  \\
-		\hline
-		CaResOl    & \gls{bmi}   & 0.004314  & 0.108546\\
-				   & Overweight  & 0.033086  & 0.306390\\
-				   & Obese       & -0.062257 & 0.175776\\
-				   & Akt         & -0.309706 & \bfseries 0.000339\\
-				   & \gls{bcat}  & -0.198165 & 0.140502\\
-				   & E2F1        & 0.270471  & \bfseries 0.011450\\
-				   & \gls{egfr}  & 0.171219  & 0.113691\\
-				   & Myc         & 0.152007  & 0.056987\\
-				   & p53         & 0.294895  & \bfseries 0.003734\\
-				   & \gls{pr}    & -0.230896 & \bfseries 0.000823\\
-				   & Ras         & -0.270824 & \bfseries 0.044788\\
-				   & Src         & -0.512567 & \bfseries \num{2.72d-08}\\
-				   & \gls{tnfa}  & -0.103302 & 0.083449\\
-		\hline
-		Original   & \gls{bmi}   & 0.002696  & 0.14407\\
-				   & Overweight  & 0.010522  & 0.62856\\
-				   & Obese       & -0.061020 & 0.05542\\
-				   & Akt         & -0.238412 & \bfseries \num{4.93d-05}\\
-				   & E2F1        & 0.185085  & \bfseries 0.02754\\
-				   & \gls{egfr}  & 0.147219  & 0.07423\\
-				   & \gls{her2}  & -0.086804 & 0.14070\\
-				   & p53         & 0.416414  & \bfseries \num{5.83d-08}\\
-				   & \gls{pi3k}  & -0.216179 & \bfseries 0.03528\\
-				   & \gls{pr}    & -0.199097 & \bfseries 0.00253\\
-				   & Ras         & -0.200544 & \bfseries 0.00898\\
-				   & Src         & -0.543834 & \bfseries \num{2.10d-15}\\
-				   & \gls{stat3} & -0.093363 & \bfseries 0.02749\\
-		\hline
-		FM         & \gls{bmi}   & 0.011093  & \bfseries 0.036229\\
-				   & Overweight  & -0.153753 & \bfseries 0.014277\\
-				   & Obese       & -0.212483 & \bfseries 0.017378\\
-				   & p53         & 0.460800  & \bfseries 0.002512\\
-				   & \gls{pi3k}  & -0.426246 & \bfseries 0.000118\\
-				   & \gls{pr}    & -0.373119 & \bfseries 0.020737\\
-				   & \gls{stat3} & 0.248745  & \bfseries 0.019073\\
-				   & \gls{tgfb}  & 0.152925  & 0.128933\\
+			\end{tabular}
+			\begin{tablenotes}
+				\begin{footnotesize}
+				\item [1] All values in bold are statistically significant (p \textless{} 0.05).
+				\end{footnotesize}
+			\end{tablenotes}
+		\end{threeparttable}
+	\end{table}
+
+	\begin{table}[htpb]
+		\centering
+		\caption[]{Description of the linear models used to predict the Res obesity metagene in \gls{nzbc} data set}
+		\label{tab:lm_sig_var_res}
+		\begin{threeparttable}
+			\begin{tabular}{llr{\bfseries}S}
+				Linear Model & Variables & Estimate & P-value\\
 				\hline
 				\hline
-	\end{longtable}
+				\rule{0pt}{2.25ex}\gls{bmi} only                           & \gls{bmi}  & -0.001422 & 0.7280 \\
+				\hline
+				\rule{0pt}{2.25ex}\gls{bmi} status only                    & Overweight & -0.001804 & 0.9810 \\
+                                                                           & Obese      & -0.09683  & 0.1710 \\
+				\hline
+				\rule{0pt}{2.25ex}\gls{bmi} and \gls{bmi} status           & \gls{bmi}  & 0.01126   & 0.1099 \\
+                                                                           & Overweight & -0.05990  & 0.4810 \\
+                                                                           & Obese      & -0.2522   & \bfseries 0.0364  \\
+				\hline
+				\rule{0pt}{2.25ex}Pathways only                            & \gls{bcat} & 0.09717   & 0.6541 \\
+                                                                           & \gls{er}   & 0.07879   & 0.7496 \\
+                                                                           & \gls{ifna} & -0.1758   & 0.7152 \\
+                                                                           & \gls{ifny} & 0.07930   & 0.8723 \\
+                                                                           & Myc        & -0.04517  & 0.8387 \\
+                                                                           & \gls{pr}   & 0.5955    & \bfseries 0.0133  \\
+				\hline
+				\rule{0pt}{2.25ex}\gls{bmi} and Pathways                   & \gls{bmi}  & -0.001994 & 0.5902 \\
+                                                                           & \gls{bcat} & 0.1145    & 0.6029 \\
+                                                                           & \gls{er}   & 0.08025   & 0.7461 \\
+                                                                           & \gls{ifna} & -0.2086   & 0.6688 \\
+                                                                           & \gls{ifny} & 0.1161    & 0.8163 \\
+                                                                           & Myc        & -0.02785  & 0.9015 \\
+                                                                           & \gls{pr}   & 0.6032    & \bfseries 0.0127  \\
+				\hline
+				\rule{0pt}{2.25ex}\gls{bmi} status and Pathways            & Overweight & -0.004412 & 0.9493 \\
+                                                                           & Obese      & -0.09683  & 0.1328 \\
+                                                                           & \gls{bcat} & 0.1504    & 0.4935 \\
+                                                                           & \gls{er}   & 0.05888   & 0.8106 \\
+                                                                           & \gls{ifna} & -0.3441   & 0.4817 \\
+                                                                           & \gls{ifny} & 0.2427    & 0.6271 \\
+                                                                           & Myc        & -0.001285 & 0.9954 \\
+                                                                           & \gls{pr}   & 0.5912    & \bfseries 0.0135  \\
+				\hline
+				\rule{0pt}{2.25ex}\gls{bmi}, \gls{bmi} status and pathways & \gls{bmi}  & 0.009426  & 0.1403 \\
+                                                                           & Overweight & -0.05185  & 0.4953 \\
+                                                                           & Obese      & -0.2277   & \bfseries 0.0386  \\
+                                                                           & \gls{bcat} & 0.1467    & 0.5013 \\
+                                                                           & \gls{er}   & 0.03435   & 0.8883 \\
+                                                                           & \gls{ifna} & -0.3681   & 0.4491 \\
+                                                                           & \gls{ifny} & 0.2429    & 0.6246 \\
+                                                                           & Myc        & -0.01963  & 0.9295 \\
+                                                                           & \gls{pr}   & 0.5547    & \bfseries 0.0201  \\
+				\hline
+				\hline
+			\end{tabular}
+			\begin{tablenotes}
+				\begin{footnotesize}
+				\item [1] All values in bold are statistically significant (p \textless{} 0.05).
+				\end{footnotesize}
+			\end{tablenotes}
+		\end{threeparttable}
+	\end{table}
+
+	\begin{table}[htpb]
+		\centering
+		\caption[]{Description of the linear models used to predict the ResOl obesity metagene in \gls{nzbc} data set}
+		\label{tab:lm_sig_var_resol}
+		\begin{threeparttable}
+			\begin{tabular}{llr{\bfseries}S}
+				Linear Model & Variables & Estimate & P-value\\
+				\hline
+				\hline
+				\rule{0pt}{2.25ex}\gls{bmi} only                           & \gls{bmi}  & -0.0009922 & 0.8080 \\
+				\hline
+				\rule{0pt}{2.25ex}\gls{bmi} status only                    & Overweight & -0.01515   & 0.8450 \\
+                                                                           & Obese      & -0.09063   & 0.2010 \\
+				\hline
+				\rule{0pt}{2.25ex}\gls{bmi} and \gls{bmi} status           & \gls{bmi}  & 0.01105    & 0.1178 \\
+                                                                           & Overweight & -0.07219   & 0.3978 \\
+                                                                           & Obese      & -0.2432    & \bfseries 0.0442  \\
+				\hline
+				\rule{0pt}{2.25ex}Pathways only                            & \gls{bcat} & 0.09532    & 0.6683 \\
+                                                                           & \gls{er}   & 0.03171    & 0.9004 \\
+                                                                           & \gls{ifna} & -0.5797    & 0.2426 \\
+                                                                           & \gls{ifny} & 0.5092     & 0.3157 \\
+                                                                           & Myc        & 0.06345    & 0.7805 \\
+                                                                           & \gls{pr}   & 0.5420     & \bfseries 0.0276  \\
+				\hline
+				\rule{0pt}{2.25ex}\gls{bmi} and Pathways                   & \gls{bmi}  & -0.001950  & 0.6076 \\
+                                                                           & \gls{bcat} & 0.1123     & 0.6192 \\
+                                                                           & \gls{er}   & 0.03313    & 0.8963 \\
+                                                                           & \gls{ifna} & -0.6118    & 0.2231 \\
+                                                                           & \gls{ifny} & 0.5452     & 0.2894 \\
+                                                                           & Myc        & 0.08041    & 0.7279 \\
+                                                                           & \gls{pr}   & 0.5495     & \bfseries 0.0264  \\
+				\hline
+				\rule{0pt}{2.25ex}\gls{bmi} status and Pathways            & Overweight & -0.01219   & 0.8640 \\
+                                                                           & Obese      & -0.09681   & 0.1440 \\
+                                                                           & \gls{bcat} & 0.1497     & 0.5070 \\
+                                                                           & \gls{er}   & 0.01354    & 0.9570 \\
+                                                                           & \gls{ifna} & -0.7389    & 0.1440 \\
+                                                                           & \gls{ifny} & 0.6638     & 0.1980 \\
+                                                                           & Myc        & 0.1081     & 0.6380 \\
+                                                                           & \gls{pr}   & 0.5387     & \bfseries 0.0280  \\
+				\hline
+				\rule{0pt}{2.25ex}\gls{bmi}, \gls{bmi} status and pathways & \gls{bmi}  & 0.009248   & 0.1596 \\
+                                                                           & Overweight & -0.05873   & 0.4532 \\
+                                                                           & Obese      & -0.2252    & \bfseries 0.0466  \\
+                                                                           & \gls{bcat} & 0.1461     & 0.5154 \\
+                                                                           & \gls{er}   & -0.01053   & 0.9667 \\
+                                                                           & \gls{ifna} & -0.7624    & 0.1296 \\
+                                                                           & \gls{ifny} & 0.6640     & 0.1955 \\
+                                                                           & Myc        & 0.09012    & 0.6933 \\
+                                                                           & \gls{pr}   & 0.5029     & \bfseries 0.0399  \\
+				\hline
+				\hline
+			\end{tabular}
+			\begin{tablenotes}
+				\begin{footnotesize}
+				\item [1] All values in bold are statistically significant (p \textless{} 0.05).
+				\end{footnotesize}
+			\end{tablenotes}
+		\end{threeparttable}
+	\end{table}
+
+	\begin{table}[htpb]
+		\centering
+		\caption[]{Description of the linear models used to predict the Ca obesity metagene in \gls{nzbc} data set}
+		\label{tab:lm_sig_var_ca}
+		\begin{threeparttable}
+			\begin{tabular}{llr{\bfseries}S}
+				Linear Model & Variables & Estimate & P-value\\
+				\hline
+				\hline
+				\rule{0pt}{2.25ex}\gls{bmi} only                           & \gls{bmi}  & -0.001147 & 0.7790 \\
+				\hline
+				\rule{0pt}{2.25ex}\gls{bmi} status only                    & Overweight & 0.01804   & 0.8160 \\
+                                                                           & Obese      & -0.08982  & 0.2030 \\
+				\hline
+				\rule{0pt}{2.25ex}\gls{bmi} and \gls{bmi} status           & \gls{bmi}  & 0.01171   & 0.0957 \\
+                                                                           & Overweight & -0.04241  & 0.6168 \\
+                                                                           & Obese      & -0.2515   & \bfseries 0.0365  \\
+				\hline
+				\rule{0pt}{2.25ex}Pathways only                            & \gls{bcat} & 0.1407    & 0.5367 \\
+                                                                           & \gls{er}   & 0.1160    & 0.6545 \\
+                                                                           & \gls{ifna} & -0.3332   & 0.5104 \\
+                                                                           & \gls{ifny} & 0.3087    & 0.5515 \\
+                                                                           & Myc        & -0.04569  & 0.8445 \\
+                                                                           & \gls{pr}   & 0.5669    & \bfseries 0.0243  \\
+				\hline
+				\rule{0pt}{2.25ex}\gls{bmi} and Pathways                   & \gls{bmi}  & -0.002148 & 0.5804 \\
+                                                                           & \gls{bcat} & 0.1593    & 0.4906 \\
+                                                                           & \gls{er}   & 0.1175    & 0.6514 \\
+                                                                           & \gls{ifna} & -0.3686   & 0.4718 \\
+                                                                           & \gls{ifny} & 0.3484    & 0.5072 \\
+                                                                           & Myc        & -0.02702  & 0.9090 \\
+                                                                           & \gls{pr}   & 0.5752    & \bfseries 0.0232  \\
+				\hline
+				\rule{0pt}{2.25ex}\gls{bmi} status and Pathways            & Overweight & 0.008987  & 0.9016 \\
+                                                                           & Obese      & -0.09970  & 0.1397 \\
+                                                                           & \gls{bcat} & 0.1934    & 0.4009 \\
+                                                                           & \gls{er}   & 0.09245   & 0.7196 \\
+                                                                           & \gls{ifna} & -0.5223   & 0.3089 \\
+                                                                           & \gls{ifny} & 0.4921    & 0.3481 \\
+                                                                           & Myc        & -0.001892 & 0.9935 \\
+                                                                           & \gls{pr}   & 0.5606    & \bfseries 0.0249  \\
+				\hline
+				\rule{0pt}{2.25ex}\gls{bmi}, \gls{bmi} status and pathways & \gls{bmi}  & 0.009944  & 0.1376 \\
+                                                                           & Overweight & -0.04106  & 0.6061 \\
+                                                                           & Obese      & -0.2378   & \bfseries 0.0392  \\
+                                                                           & \gls{bcat} & 0.1895    & 0.4073 \\
+                                                                           & \gls{er}   & 0.06657   & 0.7950 \\
+                                                                           & \gls{ifna} & -0.5476   & 0.2832 \\
+                                                                           & \gls{ifny} & 0.4923    & 0.3446 \\
+                                                                           & Myc        & -0.02125  & 0.9272 \\
+                                                                           & \gls{pr}   & 0.5221    & \bfseries 0.0362  \\
+				\hline
+				\hline
+			\end{tabular}
+			\begin{tablenotes}
+				\begin{footnotesize}
+				\item [1] All values in bold are statistically significant (p \textless{} 0.05).
+				\end{footnotesize}
+			\end{tablenotes}
+		\end{threeparttable}
+	\end{table}
+
+	\begin{table}[htpb]
+		\centering
+		\caption[]{Description of the linear models used to predict the CaOl obesity metagene in \gls{nzbc} data set}
+		\label{tab:lm_sig_var_caol}
+		\begin{threeparttable}
+			\begin{tabular}{llr{\bfseries}S}
+				Linear Model & Variables & Estimate & P-value\\
+				\hline
+				\hline
+				\rule{0pt}{2.25ex}\gls{bmi} only                           & \gls{bmi}  & \num{-3.953d-05} & 0.9920 \\
+				\hline
+				\rule{0pt}{2.25ex}\gls{bmi} status only                    & Overweight & 0.02958    & 0.7040 \\
+                                                                           & Obese      & -0.06411   & 0.3650 \\
+				\hline
+				\rule{0pt}{2.25ex}\gls{bmi} and \gls{bmi} status           & \gls{bmi}  & 0.01136    & 0.1080 \\
+                                                                           & Overweight & -0.02903   & 0.7334 \\
+                                                                           & Obese      & -0.2209    & 0.0671 \\
+				\hline
+				\rule{0pt}{2.25ex}Pathways only                            & \gls{bcat} & 0.2366     & 0.3150 \\
+                                                                           & \gls{er}   & 0.09521    & 0.7218 \\
+                                                                           & \gls{ifna} & -0.4913    & 0.3476 \\
+                                                                           & \gls{ifny} & 0.4447     & 0.4063 \\
+                                                                           & Myc        & 0.1611     & 0.5033 \\
+                                                                           & \gls{pr}   & 0.4957     & 0.0555 \\
+				\hline
+				\rule{0pt}{2.25ex}\gls{bmi} and Pathways                   & \gls{bmi}  & -0.001281  & 0.7495 \\
+                                                                           & \gls{bcat} & 0.2477     & 0.3004 \\
+                                                                           & \gls{er}   & 0.09615    & 0.7205 \\
+                                                                           & \gls{ifna} & -0.5124    & 0.3335 \\
+                                                                           & \gls{ifny} & 0.4683     & 0.3887 \\
+                                                                           & Myc        & 0.1722     & 0.4811 \\
+                                                                           & \gls{pr}   & 0.5006     & 0.0548 \\
+				\hline
+				\rule{0pt}{2.25ex}\gls{bmi} status and Pathways            & Overweight & 0.02249    & 0.7655 \\
+                                                                           & Obese      & -0.07973   & 0.2536 \\
+                                                                           & \gls{bcat} & 0.2765     & 0.2479 \\
+                                                                           & \gls{er}   & 0.07298    & 0.7846 \\
+                                                                           & \gls{ifna} & -0.6604    & 0.2155 \\
+                                                                           & \gls{ifny} & 0.6085     & 0.2638 \\
+                                                                           & Myc        & 0.1946     & 0.4230 \\
+                                                                           & \gls{pr}   & 0.4885     & 0.0586 \\
+				\hline
+				\rule{0pt}{2.25ex}\gls{bmi}, \gls{bmi} status and pathways & \gls{bmi}  & 0.009886   & 0.1549 \\
+                                                                           & Overweight & -0.02727   & 0.7414 \\
+                                                                           & Obese      & -0.2170    & 0.0692 \\
+                                                                           & \gls{bcat} & 0.2726     & 0.2518 \\
+                                                                           & \gls{er}   & 0.04725    & 0.8591 \\
+                                                                           & \gls{ifna} & -0.6855    & 0.1964 \\
+                                                                           & \gls{ifny} & 0.6087     & 0.2609 \\
+                                                                           & Myc        & 0.1753     & 0.4683 \\
+                                                                           & \gls{pr}   & 0.4502     & 0.0808 \\
+				\hline
+				\hline
+			\end{tabular}
+			\begin{tablenotes}
+				\begin{footnotesize}
+				\item [1] All values in bold are statistically significant (p \textless{} 0.05).
+				\end{footnotesize}
+			\end{tablenotes}
+		\end{threeparttable}
+	\end{table}
+
+	\begin{table}[htpb]
+		\centering
+		\caption[]{Description of the linear models used to predict the CaRes obesity metagene in \gls{nzbc} data set}
+		\label{tab:lm_sig_var_cares}
+		\begin{threeparttable}
+			\begin{tabular}{llr{\bfseries}S}
+				Linear Model & Variables & Estimate & P-value\\
+				\hline
+				\hline
+				\rule{0pt}{2.25ex}\gls{bmi} only                           & \gls{bmi}  & 0.001442  & 0.72408 \\
+				\hline
+				\rule{0pt}{2.25ex}\gls{bmi} status only                    & Overweight & -0.01840  & 0.8120  \\
+                                                                           & Obese      & 0.09338   & 0.1860  \\
+				\hline
+				\rule{0pt}{2.25ex}\gls{bmi} and \gls{bmi} status           & \gls{bmi}  & -0.01141  & 0.1043  \\
+                                                                           & Overweight & 0.04047   & 0.6328  \\
+                                                                           & Obese      & 0.2508    & \bfseries 0.0369   \\
+				\hline
+				\rule{0pt}{2.25ex}Pathways only                            & \gls{bcat} & -0.2217   & 0.32284 \\
+                                                                           & \gls{er}   & -0.2308   & 0.36579 \\
+                                                                           & \gls{ifna} & 0.1853    & 0.70938 \\
+                                                                           & \gls{ifny} & -0.1528   & 0.76408 \\
+                                                                           & Myc        & -0.0737   & 0.74765 \\
+                                                                           & \gls{pr}   & -0.6876   & \bfseries 0.00581  \\
+				\hline
+				\rule{0pt}{2.25ex}\gls{bmi} and Pathways                   & \gls{bmi}  & 0.002580  & 0.49929 \\
+                                                                           & \gls{bcat} & -0.2441   & 0.28322 \\
+                                                                           & \gls{er}   & -0.2327   & 0.36335 \\
+                                                                           & \gls{ifna} & 0.2278    & 0.65049 \\
+                                                                           & \gls{ifny} & -0.2005   & 0.69745 \\
+                                                                           & Myc        & -0.09612  & 0.67888 \\
+                                                                           & \gls{pr}   & -0.6976   & \bfseries 0.00538  \\
+				\hline
+				\rule{0pt}{2.25ex}\gls{bmi} status and Pathways            & Overweight & -0.005440 & 0.93926 \\
+                                                                           & Obese      & 0.1015    & 0.12592 \\
+                                                                           & \gls{bcat} & -0.2760   & 0.22351 \\
+                                                                           & \gls{er}   & -0.2077   & 0.41234 \\
+                                                                           & \gls{ifna} & 0.3736    & 0.45823 \\
+                                                                           & \gls{ifny} & -0.3355   & 0.51449 \\
+                                                                           & Myc        & -0.1187   & 0.60548 \\
+                                                                           & \gls{pr}   & -0.6817   & \bfseries 0.00586  \\
+				\hline
+				\rule{0pt}{2.25ex}\gls{bmi}, \gls{bmi} status and pathways & \gls{bmi}  & -0.008785 & 0.18237 \\
+                                                                           & Overweight & 0.0388    & 0.62094 \\
+                                                                           & Obese      & 0.2235    & \bfseries 0.04881  \\
+                                                                           & \gls{bcat} & -0.2725   & 0.22730 \\
+                                                                           & \gls{er}   & -0.1848   & 0.46455 \\
+                                                                           & \gls{ifna} & 0.3959    & 0.43015 \\
+                                                                           & \gls{ifny} & -0.3356   & 0.51243 \\
+                                                                           & Myc        & -0.1016   & 0.65743 \\
+                                                                           & \gls{pr}   & -0.6477   & \bfseries 0.00879  \\
+				\hline
+				\hline
+			\end{tabular}
+			\begin{tablenotes}
+				\begin{footnotesize}
+				\item [1] All values in bold are statistically significant (p \textless{} 0.05).
+				\end{footnotesize}
+			\end{tablenotes}
+		\end{threeparttable}
+	\end{table}
+
+	\begin{table}[htpb]
+		\centering
+		\caption[]{Description of the linear models used to predict the CaResOl obesity metagene in \gls{nzbc} data set}
+		\label{tab:lm_sig_var_caresol}
+		\begin{threeparttable}
+			\begin{tabular}{llr{\bfseries}S}
+				Linear Model & Variables & Estimate & P-value\\
+				\hline
+				\hline
+				\rule{0pt}{2.25ex}\gls{bmi} only                           & \gls{bmi}  & 0.0004131 & 0.9194 \\
+				\hline
+				\rule{0pt}{2.25ex}\gls{bmi} status only                    & Overweight & 0.02453   & 0.7520 \\
+                                                                           & Obese      & -0.06913  & 0.3290 \\
+				\hline
+				\rule{0pt}{2.25ex}\gls{bmi} and \gls{bmi} status           & \gls{bmi}  & 0.01335   & 0.0582 \\
+                                                                           & Overweight & -0.04436  & 0.6008 \\
+                                                                           & Obese      & -0.2533   & \bfseries 0.0352  \\
+				\hline
+				\rule{0pt}{2.25ex}Pathways only                            & \gls{bcat} & 0.2439    & 0.2864 \\
+                                                                           & \gls{er}   & 0.1383    & 0.5948 \\
+                                                                           & \gls{ifna} & -0.4364   & 0.3903 \\
+                                                                           & \gls{ifny} & 0.4301    & 0.4084 \\
+                                                                           & Myc        & 0.2837    & 0.2264 \\
+                                                                           & \gls{pr}   & 0.5789    & \bfseries 0.0219  \\
+				\hline
+				\rule{0pt}{2.25ex}\gls{bmi} and Pathways                   & \gls{bmi}  & -0.001038 & 0.7900 \\
+                                                                           & \gls{bcat} & 0.2529    & 0.2767 \\
+                                                                           & \gls{er}   & 0.1390    & 0.5947 \\
+                                                                           & \gls{ifna} & -0.4535   & 0.3782 \\
+                                                                           & \gls{ifny} & 0.4493    & 0.3947 \\
+                                                                           & Myc        & 0.2928    & 0.2191 \\
+                                                                           & \gls{pr}   & 0.5829    & \bfseries 0.0219  \\
+				\hline
+				\rule{0pt}{2.25ex}\gls{bmi} status and Pathways            & Overweight & 0.01833   & 0.8024 \\
+                                                                           & Obese      & -0.08019  & 0.2373 \\
+                                                                           & \gls{bcat} & 0.2847    & 0.2209 \\
+                                                                           & \gls{er}   & 0.1169    & 0.6525 \\
+                                                                           & \gls{ifna} & -0.6015   & 0.2453 \\
+                                                                           & \gls{ifny} & 0.5901    & 0.2646 \\
+                                                                           & Myc        & 0.3178    & 0.1793 \\
+                                                                           & \gls{pr}   & 0.5723    & \bfseries 0.0231  \\
+				\hline
+				\rule{0pt}{2.25ex}\gls{bmi}, \gls{bmi} status and pathways & \gls{bmi}  & 0.01054   & 0.1181 \\
+                                                                           & Overweight & -0.03471  & 0.6648 \\
+                                                                           & Obese      & -0.2265   & 0.0506 \\
+                                                                           & \gls{bcat} & 0.2806    & 0.2239 \\
+                                                                           & \gls{er}   & 0.08944   & 0.7288 \\
+                                                                           & \gls{ifna} & -0.6283   & 0.2216 \\
+                                                                           & \gls{ifny} & 0.5903    & 0.2606 \\
+                                                                           & Myc        & 0.2973    & 0.2058 \\
+                                                                           & \gls{pr}   & 0.5315    & \bfseries 0.0341  \\
+				\hline
+				\hline
+			\end{tabular}
+			\begin{tablenotes}
+				\begin{footnotesize}
+				\item [1] All values in bold are statistically significant (p \textless{} 0.05).
+				\end{footnotesize}
+			\end{tablenotes}
+		\end{threeparttable}
+	\end{table}
+
+	\begin{table}[htpb]
+		\centering
+		\caption[]{Description of the linear models used to predict the Or obesity metagene in \gls{nzbc} data set}
+		\label{tab:lm_sig_var_or}
+		\begin{threeparttable}
+			\begin{tabular}{llr{\bfseries}S}
+				Linear Model & Variables & Estimate & P-value\\
+				\hline
+				\hline
+				\rule{0pt}{2.25ex}\gls{bmi} only                           & \gls{bmi}  & 0.001654  & 0.68532 \\
+				\hline
+				\rule{0pt}{2.25ex}\gls{bmi} status only                    & Overweight & 0.005772  & 0.9410  \\
+                                                                           & Obese      & 0.1009    & 0.1540  \\
+				\hline
+				\rule{0pt}{2.25ex}\gls{bmi} and \gls{bmi} status           & \gls{bmi}  & -0.01106  & 0.1162  \\
+                                                                           & Overweight & 0.06284   & 0.4598  \\
+                                                                           & Obese      & 0.2535    & \bfseries 0.0354   \\
+				\hline
+				\rule{0pt}{2.25ex}Pathways only                            & \gls{bcat} & -0.2284   & 0.3074 \\
+                                                                           & \gls{er}   & -0.1481   & 0.5604 \\
+                                                                           & \gls{ifna} & 0.4504    & 0.3647 \\
+                                                                           & \gls{ifny} & -0.4013   & 0.4302 \\
+                                                                           & Myc        & -0.1402   & 0.5399 \\
+                                                                           & \gls{pr}   & -0.6446   & \bfseries 0.009380  \\
+				\hline
+				\rule{0pt}{2.25ex}\gls{bmi} and Pathways                   & \gls{bmi}  & 0.003004  & 0.4301 \\
+                                                                           & \gls{bcat} & -0.2545   & 0.2618 \\
+                                                                           & \gls{er}   & -0.1502   & 0.5555 \\
+                                                                           & \gls{ifna} & 0.4999    & 0.3195 \\
+                                                                           & \gls{ifny} & -0.4568   & 0.3749 \\
+                                                                           & Myc        & -0.1663   & 0.4728 \\
+                                                                           & \gls{pr}   & -0.6562   & \bfseries 0.0085  \\
+				\hline
+				\rule{0pt}{2.25ex}\gls{bmi} status and Pathways            & Overweight & 0.01412   & 0.8427  \\
+                                                                           & Obese      & 0.1135    & 0.0867  \\
+                                                                           & \gls{bcat} & -0.2921   & 0.1964  \\
+                                                                           & \gls{er}   & -0.1267   & 0.6153  \\
+                                                                           & \gls{ifna} & 0.6372    & 0.2056  \\
+                                                                           & \gls{ifny} & -0.5827   & 0.2570  \\
+                                                                           & Myc        & -0.1925   & 0.4010  \\
+                                                                           & \gls{pr}   & -0.6407   & \bfseries 0.0092   \\
+				\hline
+				\rule{0pt}{2.25ex}\gls{bmi}, \gls{bmi} status and pathways & \gls{bmi}  & -0.008654 & 0.1875 \\
+                                                                           & Overweight & 0.05768   & 0.4610 \\
+                                                                           & Obese      & 0.2336    & \bfseries 0.03905  \\
+                                                                           & \gls{bcat} & -0.2887   & 0.1997 \\
+                                                                           & \gls{er}   & -0.1042   & 0.6788 \\
+                                                                           & \gls{ifna} & 0.6592    & 0.1890 \\
+                                                                           & \gls{ifny} & -0.5829   & 0.2549 \\
+                                                                           & Myc        & -0.1757   & 0.4421 \\
+                                                                           & \gls{pr}   & -0.6072   & \bfseries 0.01353  \\
+				\hline
+				\hline
+			\end{tabular}
+			\begin{tablenotes}
+				\begin{footnotesize}
+				\item [1] All values in bold are statistically significant (p \textless{} 0.05).
+				\end{footnotesize}
+			\end{tablenotes}
+		\end{threeparttable}
+	\end{table}
+
+	\begin{table}[htpb]
+		\centering
+		\caption[]{Description of the linear models used to predict the FM obesity metagene in \gls{nzbc} data set}
+		\label{tab:lm_sig_var_fm}
+		\begin{threeparttable}
+			\begin{tabular}{llr{\bfseries}S}
+				Linear Model & Variables & Estimate & P-value\\
+				\hline
+				\hline
+				\rule{0pt}{2.25ex}\gls{bmi} only                           & \gls{bmi}  & 0.002314 & 0.5708 \\
+				\hline
+				\rule{0pt}{2.25ex}\gls{bmi} status only                    & Overweight & -0.1483  & 0.05670 \\
+                                                                           & Obese      & -0.08452 & 0.22910 \\
+				\hline
+				\rule{0pt}{2.25ex}\gls{bmi} and \gls{bmi} status           & \gls{bmi}  & 0.01503  & \bfseries 0.03091  \\
+                                                                           & Overweight & -0.2258  & \bfseries 0.007940\\ 
+                                                                           & Obese      & -0.2920  & \bfseries 0.01419  \\
+				\hline
+				\rule{0pt}{2.25ex}Pathways only                            & \gls{bcat} & -0.1349  & 0.5334  \\
+                                                                           & \gls{er}   & 0.1197   & 0.6271  \\
+                                                                           & \gls{ifna} & -0.5624  & 0.2438  \\
+                                                                           & \gls{ifny} & 0.5884   & 0.2340  \\
+                                                                           & Myc        & 0.4177   & 0.06180 \\
+                                                                           & \gls{pr}   & -0.08815 & 0.7090  \\
+				\hline
+				\rule{0pt}{2.25ex}\gls{bmi} and Pathways                   & \gls{bmi}  & 0.002392 & 0.5172  \\
+                                                                           & \gls{bcat} & -0.1557  & 0.4786  \\
+                                                                           & \gls{er}   & 0.1180   & 0.6333  \\
+                                                                           & \gls{ifna} & -0.5231  & 0.2835  \\
+                                                                           & \gls{ifny} & 0.5442   & 0.2767  \\
+                                                                           & Myc        & 0.3969   & 0.07970 \\
+                                                                           & \gls{pr}   & -0.09737 & 0.6817  \\
+				\hline
+				\rule{0pt}{2.25ex}\gls{bmi} status and Pathways            & Overweight & -0.1249  & 0.07330 \\
+                                                                           & Obese      & -0.08029 & 0.2104  \\
+                                                                           & \gls{bcat} & -0.07254 & 0.7402  \\
+                                                                           & \gls{er}   & 0.1304   & 0.5948  \\
+                                                                           & \gls{ifna} & -0.5604  & 0.2517  \\
+                                                                           & \gls{ifny} & 0.5876   & 0.2397  \\
+                                                                           & Myc        & 0.4665   & \bfseries 0.03820  \\
+                                                                           & \gls{pr}   & -0.07506 & 0.7490  \\
+				\hline
+				\rule{0pt}{2.25ex}\gls{bmi}, \gls{bmi} status and pathways & \gls{bmi}  & 0.01539  & \bfseries 0.01464  \\
+                                                                           & Overweight & -0.2024  & \bfseries 0.007460  \\
+                                                                           & Obese      & -0.2940  & \bfseries 0.006660  \\
+                                                                           & \gls{bcat} & -0.0786  & 0.7119 \\
+                                                                           & \gls{er}   & 0.09030  & 0.7053 \\
+                                                                           & \gls{ifna} & -0.5994  & 0.2079 \\
+                                                                           & \gls{ifny} & 0.5878   & 0.2265 \\
+                                                                           & Myc        & 0.4365   & \bfseries{0.04623} \\
+                                                                           & \gls{pr}   & -0.1347  & 0.5575 \\
+				\hline
+				\hline
+			\end{tabular}
+			\begin{tablenotes}
+				\begin{footnotesize}
+				\item [1] All values in bold are statistically significant (p \textless{} 0.05).
+				\end{footnotesize}
+			\end{tablenotes}
+		\end{threeparttable}
+	\end{table}
+
+	\newpage
+
+	\section{Summary of the remainder of the \gls{pr}-only linear models in \gls{nzbc} data}
+	\label{sec:summary_of_the_pr_linear_models_in_nzbc_data}
+
+	The \gls{pr}-only linear models for the other obesity associated genetic signatures that were generated in the \gls{nzbc} data were summarised in \cref{tab:lm_pr_only_crol,tab:lm_pr_only_res,tab:lm_pr_only_resol,tab:lm_pr_only_ca,tab:lm_pr_only_caol,tab:lm_pr_only_cares,tab:lm_pr_only_caresol,tab:lm_pr_only_or,tab:lm_pr_only_fm}.
+	The \gls{pr} pathway metagene was significant in all of the linear models that predicted the obesity metagenes derived from the CR data set \cref{tab:lm_pr_only_crol,tab:lm_pr_only_res,tab:lm_pr_only_resol,tab:lm_pr_only_ca,tab:lm_pr_only_caol,tab:lm_pr_only_cares,tab:lm_pr_only_caresol,tab:lm_pr_only_or}.
+	On the other hand, the sample \gls{bmi}/\gls{bmi} status were significant in the linear model that predicted the FM obesity metagene \cref{tab:lm_pr_only_fm}.
+
+	\begin{table}[htpb]
+		\centering
+		\caption[]{Description of the linear models used to predict the CrOl obesity metagene in \gls{nzbc} data set, using only the sample \gls{bmi}, \gls{bmi} status and \gls{pr} pathway metagene scores}
+		\label{tab:lm_pr_only_crol}
+		\begin{threeparttable}
+			\begin{tabular}{llr{\bfseries}S}
+				Linear Model & Variables & Estimate & {P-value}\\
+					\hline
+					\hline
+					\rule{0pt}{2.25ex}\gls{pr} only                            & \gls{pr}   & 0.4484    & \bfseries \num{3.26d-06} \tnote{1} \\
+					\hline
+					\rule{0pt}{2.25ex}\gls{bmi} and \gls{pr}                   & \gls{bmi}  & -0.002084 & 0.5699  \\
+                                                                               & \gls{pr}   & 0.4497    & \bfseries \num{3.36d-06} \\
+					\hline
+					\rule{0pt}{2.25ex}\gls{bmi} status and \gls{pr}            & Obese      & -0.09307  & 0.1440   \\
+                                                                               & Overweight & -0.01989  & 0.7750   \\
+                                                                               & \gls{pr}   & 0.4418    & \bfseries \num{4.32d-06} \\
+					\hline
+					\rule{0pt}{2.25ex}\gls{bmi}, \gls{bmi} status and \gls{pr} & \gls{bmi}  & 0.008072  & 0.2060   \\
+                                                                               & Obese      & -0.2048   & 0.06100   \\
+                                                                               & Overweight & -0.06159  & 0.4230   \\
+                                                                               & \gls{pr}   & 0.4294    & \bfseries \num{7.93d-06} \\
+					\hline
+					\hline
+			\end{tabular}
+				\begin{tablenotes}
+					\begin{footnotesize}
+					\item [1] All values in bold are statistically significant (p \textless{} 0.05).
+					\end{footnotesize}
+				\end{tablenotes}
+		\end{threeparttable}
+	\end{table}
+
+	\begin{table}[htpb]
+		\centering
+		\caption[]{Description of the linear models used to predict the Res obesity metagene in \gls{nzbc} data set, using only the sample \gls{bmi}, \gls{bmi} status and \gls{pr} pathway metagene scores}
+		\label{tab:lm_pr_only_res}
+		\begin{threeparttable}
+			\begin{tabular}{llr \bfseries S}
+				Linear Model & Variables & Estimate & {P-value}\\
+					\hline
+					\hline
+					\rule{0pt}{2.25ex}\gls{pr} only                            & \gls{pr}   & 0.4926    & \bfseries \num{2.22d-07} \tnote{1}\\
+					\hline
+					\rule{0pt}{2.25ex}\gls{bmi} and \gls{pr}                   & \gls{bmi}  & -0.001923 & 0.5904   \\
+                                                                               & \gls{pr}   & 0.4938    & \bfseries \num{2.35d-07} \\
+					\hline
+					\rule{0pt}{2.25ex}\gls{bmi} status and \gls{pr}            & Obese      & -0.08315  & 0.1790   \\
+                                                                               & Overweight & 0.0001257 & 0.9990   \\
+                                                                               & \gls{pr}   & 0.4862    & \bfseries \num{3.00d-07} \\
+					\hline
+					\rule{0pt}{2.25ex}\gls{bmi}, \gls{bmi} status and \gls{pr} & \gls{bmi}  & 0.007716  & 0.2140   \\
+                                                                               & Obese      & -0.1899   & 0.0740   \\
+                                                                               & Overweight & -0.03974  & 0.5950   \\
+                                                                               & \gls{pr}   & 0.4744    & \bfseries \num{5.78d-07} \\
+					\hline
+					\hline
+			\end{tabular}
+				\begin{tablenotes}
+					\begin{footnotesize}
+					\item [1] All values in bold are statistically significant (p \textless{} 0.05).
+					\end{footnotesize}
+				\end{tablenotes}
+		\end{threeparttable}
+	\end{table}
+
+	\begin{table}[htpb]
+		\centering
+		\caption[]{Description of the linear models used to predict the ResOl obesity metagene in \gls{nzbc} data set, using only the sample \gls{bmi}, \gls{bmi} status and \gls{pr} pathway metagene scores}
+		\label{tab:lm_pr_only_resol}
+		\begin{threeparttable}
+			\begin{tabular}{llr{\bfseries}S}
+				Linear Model & Variables & Estimate & {P-value}\\
+					\hline
+					\hline
+					\rule{0pt}{2.25ex}\gls{pr} only                            & \gls{pr}   & 0.4581    & \bfseries \num{1.86d-06} \tnote{1}\\
+					\hline
+					\rule{0pt}{2.25ex}\gls{bmi} and \gls{pr}                   & \gls{bmi}  & -0.001457 & 0.6897   \\
+                                                                               & \gls{pr}   & 0.4590    & \bfseries \num{1.99d-06} \\
+					\hline
+					\rule{0pt}{2.25ex}\gls{bmi} status and \gls{pr}            & Obese      & -0.07789  & 0.2200   \\
+                                                                               & Overweight & -0.01336  & 0.8480   \\
+                                                                               & \gls{pr}   & 0.4526    & \bfseries \num{2.56d-06} \\
+					\hline
+					\rule{0pt}{2.25ex}\gls{bmi}, \gls{bmi} status and \gls{pr} & \gls{bmi}  & 0.007763  & 0.2230   \\
+                                                                               & Obese      & -0.1853   & 0.0890   \\
+                                                                               & Overweight & -0.05346  & 0.4860   \\
+                                                                               & \gls{pr}   & 0.4407    & \bfseries \num{4.71d-06} \\
+					\hline
+					\hline
+			\end{tabular}
+				\begin{tablenotes}
+					\begin{footnotesize}
+					\item [1] All values in bold are statistically significant (p \textless{} 0.05).
+					\end{footnotesize}
+				\end{tablenotes}
+		\end{threeparttable}
+	\end{table}
+
+	\begin{table}[htpb]
+		\centering
+		\caption[]{Description of the linear models used to predict the Ca obesity metagene in \gls{nzbc} data set, using only the sample \gls{bmi}, \gls{bmi} status and \gls{pr} pathway metagene scores}
+		\label{tab:lm_pr_only_ca}
+		\begin{threeparttable}
+			\begin{tabular}{llr{\bfseries}S}
+				Linear Model & Variables & Estimate & {P-value}\\
+					\hline
+					\hline
+					\rule{0pt}{2.25ex}\gls{pr} only                            & \gls{pr}   & 0.3850    & \bfseries \num{8.34d-05} \tnote{1}\\
+					\hline
+					\rule{0pt}{2.25ex}\gls{bmi} and \gls{pr}                   & \gls{bmi}  & -0.001538 & 0.6848    \\
+                                                                               & \gls{pr}   & 0.3859    & \bfseries \num{8.67d-05}  \\
+					\hline
+					\rule{0pt}{2.25ex}\gls{bmi} status and \gls{pr}            & Obese      & -0.07917  & 0.2274    \\
+                                                                               & Overweight & 0.01954   & 0.7857    \\
+                                                                               & \gls{pr}   & 0.3782    & \bfseries 0.0001060  \\
+					\hline
+					\rule{0pt}{2.25ex}\gls{bmi}, \gls{bmi} status and \gls{pr} & \gls{bmi}  & 0.008992  & 0.1718    \\
+                                                                               & Obese      & -0.2036   & 0.07034   \\
+                                                                               & Overweight & -0.02692  & 0.7338    \\
+                                                                               & \gls{pr}   & 0.3644    & \bfseries 0.0001830  \\
+					\hline
+					\hline
+			\end{tabular}
+				\begin{tablenotes}
+					\begin{footnotesize}
+					\item [1] All values in bold are statistically significant (p \textless{} 0.05).
+					\end{footnotesize}
+				\end{tablenotes}
+		\end{threeparttable}
+	\end{table}
+
+	\begin{table}[htpb]
+		\centering
+		\caption[]{Description of the linear models used to predict the CaOl obesity metagene in \gls{nzbc} data set, using only the sample \gls{bmi}, \gls{bmi} status and \gls{pr} pathway metagene scores}
+		\label{tab:lm_pr_only_caol}
+		\begin{threeparttable}
+			\begin{tabular}{llr{\bfseries}S}
+				Linear Model & Variables & Estimate & {P-value}\\
+					\hline
+					\hline
+					\rule{0pt}{2.25ex}\gls{pr} only                            & \gls{pr}   & 0.3425     & \bfseries 0.0005200  \tnote{1}\\
+					\hline
+					\rule{0pt}{2.25ex}\gls{bmi} and \gls{pr}                   & \gls{bmi}  & -0.0003869 & 0.9201    \\
+                                                                               & \gls{pr}   & 0.3428     & \bfseries 0.0005530  \\
+					\hline
+					\rule{0pt}{2.25ex}\gls{bmi} status and \gls{pr}            & Obese      & -0.05462   & 0.4147    \\
+                                                                               & Overweight & 0.03092    & 0.6742    \\
+                                                                               & \gls{pr}   & 0.3373     & \bfseries 0.0006470  \\
+					\hline
+					\rule{0pt}{2.25ex}\gls{bmi}, \gls{bmi} status and \gls{pr} & \gls{bmi}  & 0.008943   & 0.1842    \\
+                                                                               & Obese      & -0.1784    & 0.1206    \\
+                                                                               & Overweight & -0.01528   & 0.8504    \\
+                                                                               & \gls{pr}   & 0.3236     & \bfseries 0.001060   \\
+					\hline
+					\hline
+			\end{tabular}
+				\begin{tablenotes}
+					\begin{footnotesize}
+					\item [1] All values in bold are statistically significant (p \textless{} 0.05).
+					\end{footnotesize}
+				\end{tablenotes}
+		\end{threeparttable}
+	\end{table}
+
+	\begin{table}[htpb]
+		\centering
+		\caption[]{Description of the linear models used to predict the CaRes obesity metagene in \gls{nzbc} data set, using only the sample \gls{bmi}, \gls{bmi} status and \gls{pr} pathway metagene scores}
+		\label{tab:lm_pr_only_cares}
+		\begin{threeparttable}
+			\begin{tabular}{llr{\bfseries}S}
+				Linear Model & Variables & Estimate & {P-value}\\
+					\hline
+					\hline
+					\rule{0pt}{2.25ex}\gls{pr} only                            & \gls{pr}   & -0.4255   & \bfseries \num{1.13d-05} \tnote{1}\\
+					\hline
+					\rule{0pt}{2.25ex}\gls{bmi} and \gls{pr}                   & \gls{bmi}  & 0.001874  & 0.6140   \\
+                                                                               & \gls{pr}   & -0.4267   & \bfseries \num{1.17d-05} \\
+					\hline
+					\rule{0pt}{2.25ex}\gls{bmi} status and \gls{pr}            & Obese      & 0.08159   & 0.2040   \\
+                                                                               & Overweight & -0.02006  & 0.7760   \\
+                                                                               & \gls{pr}   & -0.4186   & \bfseries \num{1.44d-05} \\
+					\hline
+					\rule{0pt}{2.25ex}\gls{bmi}, \gls{bmi} status and \gls{pr} & \gls{bmi}  & -0.008379 & 0.1936   \\
+                                                                               & Obese      & 0.1976    & 0.07310  \\
+                                                                               & Overweight & 0.02323   & 0.7645   \\
+                                                                               & \gls{pr}   & -0.4057   & \bfseries \num{2.59d-05} \\
+					\hline
+					\hline
+			\end{tabular}
+				\begin{tablenotes}
+					\begin{footnotesize}
+					\item [1] All values in bold are statistically significant (p \textless{} 0.05).
+					\end{footnotesize}
+				\end{tablenotes}
+		\end{threeparttable}
+	\end{table}
+
+	\begin{table}[htpb]
+		\centering
+		\caption[]{Description of the linear models used to predict the CaResOl obesity metagene in \gls{nzbc} data set, using only the sample \gls{bmi}, \gls{bmi} status and \gls{pr} pathway metagene scores}
+		\label{tab:lm_pr_only_caresol}
+		\begin{threeparttable}
+			\begin{tabular}{llr{\bfseries}S}
+				Linear Model & Variables & Estimate & {P-value}\\
+					\hline
+					\hline
+					\rule{0pt}{2.25ex}\gls{pr} only                            & \gls{pr}   & 0.4094     & \bfseries \num{2.58d-05} \tnote{1}\\
+					\hline
+					\rule{0pt}{2.25ex}\gls{bmi} and \gls{pr}                   & \gls{bmi}  & -1.809d-06 & 0.9996   \\
+                                                                               & \gls{pr}   & 4.094d-01  & \bfseries \num{2.85d-05} \\
+					\hline
+					\rule{0pt}{2.25ex}\gls{bmi} status and \gls{pr}            & Obese      & -0.05775   & 0.3740   \\
+                                                                               & Overweight & 0.02613    & 0.7140   \\
+                                                                               & \gls{pr}   & 0.4041     & \bfseries \num{3.36d-05} \\
+					\hline
+					\rule{0pt}{2.25ex}\gls{bmi}, \gls{bmi} status and \gls{pr} & \gls{bmi}  & 0.01045    & 0.1091   \\
+                                                                               & Obese      & -0.2024    & 0.06910  \\
+                                                                               & Overweight & -0.02786   & 0.7220   \\
+                                                                               & \gls{pr}   & 0.3881     & \bfseries \num{6.21d-05} \\
+					\hline
+					\hline
+			\end{tabular}
+				\begin{tablenotes}
+					\begin{footnotesize}
+					\item [1] All values in bold are statistically significant (p \textless{} 0.05).
+					\end{footnotesize}
+				\end{tablenotes}
+		\end{threeparttable}
+	\end{table}
+
+	\begin{table}[htpb]
+		\centering
+		\caption[]{Description of the linear models used to predict the Or obesity metagene in \gls{nzbc} data set, using only the sample \gls{bmi}, \gls{bmi} status and \gls{pr} pathway metagene scores}
+		\label{tab:lm_pr_only_or}
+		\begin{threeparttable}
+			\begin{tabular}{llr{\bfseries}S}
+				Linear Model & Variables & Estimate & {P-value}\\
+					\hline
+					\hline
+					\rule{0pt}{2.25ex}\gls{pr} only                            & \gls{pr}   & -0.4463   & \bfseries \num{3.66d-06} \tnote{1}\\
+					\hline
+					\rule{0pt}{2.25ex}\gls{bmi} and \gls{pr}                   & \gls{bmi}  & 0.002108  & 0.5660   \\
+                                                                               & \gls{pr}   & -0.4476   & \bfseries \num{3.77d-06} \\
+					\hline
+					\rule{0pt}{2.25ex}\gls{bmi} status and \gls{pr}            & Obese      & 0.088520  & 0.1650   \\
+                                                                               & Overweight & 0.004028  & 0.9540   \\
+                                                                               & \gls{pr}   & -0.4396   & \bfseries \num{4.82d-06} \\
+					\hline
+					\rule{0pt}{2.25ex}\gls{bmi}, \gls{bmi} status and \gls{pr} & \gls{bmi}  & -0.007866 & 0.2180   \\
+                                                                               & Obese      & 0.1974    & 0.07100  \\
+                                                                               & Overweight & 0.04466   & 0.5620   \\
+                                                                               & \gls{pr}   & -0.4275   & \bfseries \num{8.77d-06} \\
+					\hline
+					\hline
+			\end{tabular}
+				\begin{tablenotes}
+					\begin{footnotesize}
+					\item [1] All values in bold are statistically significant (p \textless{} 0.05).
+					\end{footnotesize}
+				\end{tablenotes}
+		\end{threeparttable}
+	\end{table}
+
+	\begin{table}[htpb]
+		\centering
+		\caption[]{Description of the linear models used to predict the FM obesity metagene in \gls{nzbc} data set, using only the sample \gls{bmi}, \gls{bmi} status and \gls{pr} pathway metagene scores}
+		\label{tab:lm_pr_only_fm}
+		\begin{threeparttable}
+			\begin{tabular}{llr{\bfseries}S}
+				Linear Model & Variables & Estimate & {P-value}\\
+					\hline
+					\hline
+					\rule{0pt}{2.25ex}\gls{pr} only                            & \gls{pr}   & -0.1285  & 0.2050   \\
+					\hline
+					\rule{0pt}{2.25ex}\gls{bmi} and \gls{pr}                   & \gls{bmi}  & 0.002446 & 0.5479   \\
+                                                                               & \gls{pr}   & -0.1301  & 0.2012   \\
+					\hline
+					\rule{0pt}{2.25ex}\gls{bmi} status and \gls{pr}            & Obese      & -0.08819 & 0.2080   \\
+                                                                               & Overweight & -0.1488  & 0.05500  \\
+                                                                               & \gls{pr}   & -0.1306  & 0.1940   \\
+					\hline
+					\rule{0pt}{2.25ex}\gls{bmi}, \gls{bmi} status and \gls{pr} & \gls{bmi}  & 0.01619  & \bfseries 0.02008 \tnote{1}\\
+                                                                               & Obese      & -0.3123  & \bfseries 0.008750  \\
+                                                                               & Overweight & -0.2324  & \bfseries 0.006030  \\
+                                                                               & \gls{pr}   & -0.1554  & 0.1165   \\
+					\hline
+					\hline
+			\end{tabular}
+				\begin{tablenotes}
+					\begin{footnotesize}
+					\item [1] All values in bold are statistically significant (p \textless{} 0.05).
+					\end{footnotesize}
+				\end{tablenotes}
+		\end{threeparttable}
+	\end{table}
+
+	\newpage
 

### 2017-03-13 thesis/results/results2.tex
Need to add linear model prediction scatter plots - tmp commit

@@
 	\centering
-	\caption{Description of the linear models used to predict the CR obesity metagene in \gls{nzbc} data set}
+	\caption{Description of the linear models used to predict the Cr obesity metagene in \gls{nzbc} data set}
 	\label{tab:lm_sig_var}
@@
 	\centering
-	\caption{Description of the linear models used to predict the CR obesity metagene in \gls{nzbc} data set, using only the sample \gls{bmi}, \gls{bmi} status and \gls{pr} pathway metagene scores}
+	\caption{Description of the linear models used to predict the Cr obesity metagene in \gls{nzbc} data set, using only the sample \gls{bmi}, \gls{bmi} status and \gls{pr} pathway metagene scores}
 	\label{tab:lm_pr_only}
@@
 Sample \gls{bmi}, \gls{bmi} status and all of the pathway metagene scores were used to generate the step wise linear model for each of the obesity metagenes.
-The linear models that resulted from the step wise selection of the variables were summarised in \cref{app:b}.
+The linear models that resulted from the step wise selection of the variables were summarised in \cref{tab:stepwise_model}.
+
+\begin{longtable}{llr{\bfseries}S}
+	\centering
+	\caption{Description of the step wise linear models used to predict all of the obesity metagenes in \gls{nzbc} data set}
+	\label{tab:stepwise_model}\\
+		Obesity metagene & Variables used in the model & Estimate & {P-value}\\
+		\hline
+	\endfirsthead
+	\multicolumn{4}{c}{\tablename\ \thetable{}\ (continued)}\\
+		\hline
+		\hline
+	\endhead
+		\hline
+		\hline
+		Cr         & Akt         & -0.3415   & \bfseries \num{4.43d-8}\\
+				   & \gls{egfr}  & 0.33852   & \bfseries \num{1.94d-05}\\
+				   & \gls{er}    & -0.20237  & \bfseries 0.00419\\
+				   & \gls{ifny}  & 0.12233   & 0.08700\\
+				   & \gls{pr}    & -0.33920  & \bfseries \num{2.09d-5}\\
+				   & Ras         & -0.22611  & \bfseries 0.00193\\
+				   & Src         & -0.11452  & 0.08177\\
+				   & \gls{tgfb}  & 0.55638   & \bfseries \textless{} \num{2d-16}\\
+				   & \gls{tnfa}  & -0.25501  & \bfseries 0.00616\\
+		\hline
+		CrOl       & Overweight  & 0.040918  & 0.074865\\
+				   & Obese       & -0.006381 & 0.756121\\
+				   & Akt         & -0.356713 & \bfseries \num{1.89d-07}\\
+				   & E2F1        & 0.160229  & \bfseries 0.005411\\
+				   & \gls{egfr}  & 0.148618  & 0.062683\\
+				   & p53         & 0.336552  & \bfseries \num{4.35d-05}\\
+				   & \gls{pr}    & -0.225423 & \bfseries 0.000639\\
+				   & Src         & -0.428758 & \bfseries \num{5.85d-10}\\
+				   & \gls{stat3} & -0.124118 & \bfseries 0.009658\\
+		\hline
+		Res        & Akt         & 0.24613   & \bfseries 0.000289\\
+				   & \gls{egfr}  & -0.46827  & \bfseries \num{2.16d-05}\\
+				   & \gls{er}    & 0.18400   & 0.061446\\
+				   & \gls{her2}  & 0.13835   & 0.118228\\
+				   & \gls{ifny}  & -0.21019  & \bfseries 0.044084\\
+				   & \gls{pr}    & 0.41195   & \bfseries 0.000300\\
+				   & Ras         & 0.26506   & \bfseries 0.010465\\
+				   & \gls{tgfb}  & -0.58075  & \bfseries \num{7.64d-11}\\
+				   & \gls{tnfa}  & 0.42200   & \bfseries 0.001834\\
+		\hline
+		ResOl      & Akt         & -0.33596  & \bfseries 0.000294\\
+				   & \gls{bcat}  & -0.19866  & 0.138953\\
+				   & E2F1        & 0.36059   & \bfseries 0.009325\\
+				   & \gls{egfr}  & 0.23421   & 0.056865\\
+				   & \gls{her2}  & -0.13652  & 0.121689\\
+				   & p53         & 0.38746   & \bfseries 0.000519\\
+				   & \gls{pi3k}  & -0.21881  & 0.148676\\
+				   & \gls{pr}    & -0.32228  & \bfseries 0.001656\\
+				   & Ras         & -0.28984  & \bfseries 0.033975\\
+				   & Src         & -0.40779  & \bfseries \num{5.26d-06}\\
+				   & \gls{stat3} & -0.11452  & 0.125609\\
+				   & \gls{tnfa}  & -0.11442  & 0.117023\\
+		\hline
+		Ca         & Akt         & -0.26243  & \bfseries \num{7.04d-06}\\
+				   & \gls{egfr}  & 0.36305   & \bfseries 0.000139\\
+				   & \gls{er}    & -0.18614  & \bfseries 0.030771\\
+				   & \gls{ifny}  & 0.22118   & \bfseries 0.013034\\
+				   & \gls{pr}    & -0.42911  & \bfseries \num{1.07d-05}\\
+				   & Ras         & -0.28124  & \bfseries 0.000898\\
+				   & \gls{tgfb}  & 0.76139   & \bfseries \textless{} \num{2d-16}\\
+				   & \gls{tnfa}  & -0.33200  & \bfseries 0.003919\\
+		\hline
+		CaOl       & Overweight  & 0.07613   & \bfseries 0.006875\\
+				   & Obese       & 0.01408   & 0.564622\\
+				   & Akt         & -0.44727  & \bfseries \num{2.43d-07}\\
+				   & \gls{bcat}  & -0.13463  & 0.140261\\
+				   & E2F1        & 0.33864   & \bfseries 0.001249\\
+				   & p53         & 0.28777   & \bfseries 0.000996\\
+				   & \gls{pr}    & -0.28255  & \bfseries \num{2.18d-05}\\
+				   & Src         & -0.44149  & \bfseries \num{2.02d-08}\\
+				   & \gls{tnfa}  & -0.14215  & \bfseries 0.013897\\
+		\hline
+		CaRes      & \gls{bcat}  & 0.26899   & 0.05238  \\
+				   & E2F1        & -0.22115  & 0.08848  \\
+				   & \gls{egfr}  & 0.44400   & \bfseries \num{6.67d-05 }\\
+				   & \gls{her2}  & -0.20304  & \bfseries 0.04615  \\
+				   & \gls{ifny}  & 0.32874   & \bfseries 0.00574  \\
+				   & \gls{pr}    & -0.28764  & \bfseries 0.00104  \\
+				   & Src         & -0.17989  & 0.05404  \\
+				   & \gls{tgfb}  & 0.52898   & \bfseries \num{2.15d-06 }\\
+				   & \gls{tnfa}  & -0.38677  & \bfseries 0.00564  \\
+		\hline
+		CaResOl    & \gls{bmi}   & 0.004314  & 0.108546\\
+				   & Overweight  & 0.033086  & 0.306390\\
+				   & Obese       & -0.062257 & 0.175776\\
+				   & Akt         & -0.309706 & \bfseries 0.000339\\
+				   & \gls{bcat}  & -0.198165 & 0.140502\\
+				   & E2F1        & 0.270471  & \bfseries 0.011450\\
+				   & \gls{egfr}  & 0.171219  & 0.113691\\
+				   & Myc         & 0.152007  & 0.056987\\
+				   & p53         & 0.294895  & \bfseries 0.003734\\
+				   & \gls{pr}    & -0.230896 & \bfseries 0.000823\\
+				   & Ras         & -0.270824 & \bfseries 0.044788\\
+				   & Src         & -0.512567 & \bfseries \num{2.72d-08}\\
+				   & \gls{tnfa}  & -0.103302 & 0.083449\\
+		\hline
+		Or         & \gls{bmi}   & 0.002696  & 0.14407\\
+				   & Overweight  & 0.010522  & 0.62856\\
+				   & Obese       & -0.061020 & 0.05542\\
+				   & Akt         & -0.238412 & \bfseries \num{4.93d-05}\\
+				   & E2F1        & 0.185085  & \bfseries 0.02754\\
+				   & \gls{egfr}  & 0.147219  & 0.07423\\
+				   & \gls{her2}  & -0.086804 & 0.14070\\
+				   & p53         & 0.416414  & \bfseries \num{5.83d-08}\\
+				   & \gls{pi3k}  & -0.216179 & \bfseries 0.03528\\
+				   & \gls{pr}    & -0.199097 & \bfseries 0.00253\\
+				   & Ras         & -0.200544 & \bfseries 0.00898\\
+				   & Src         & -0.543834 & \bfseries \num{2.10d-15}\\
+				   & \gls{stat3} & -0.093363 & \bfseries 0.02749\\
+		\hline
+		FM         & \gls{bmi}   & 0.011093  & \bfseries 0.036229\\
+				   & Overweight  & -0.153753 & \bfseries 0.014277\\
+				   & Obese       & -0.212483 & \bfseries 0.017378\\
+				   & p53         & 0.460800  & \bfseries 0.002512\\
+				   & \gls{pi3k}  & -0.426246 & \bfseries 0.000118\\
+				   & \gls{pr}    & -0.373119 & \bfseries 0.020737\\
+				   & \gls{stat3} & 0.248745  & \bfseries 0.019073\\
+				   & \gls{tgfb}  & 0.152925  & 0.128933\\
+		\hline
+		\hline
+	\end{longtable}
 

### 2017-03-13 thesis/misc/appendix2.tex
Added all of the extra results in appendix2; clean up the titles - tmp commit

@@
 
+	\begin{TableNotes}
+		\begin{footnotesize}
+		\item [1] All values in bold are statistically significant (p \textless{} 0.05).
+		\end{footnotesize}
+	\end{TableNotes}
+
+	\begin{longtable}{llc{\bfseries}S}
+		\centering
+		\caption{tmp}
+		\label{tab:prediction_summary_cris}\\
+		Obesity metagene & Model                         & R$^2$      & P-value   \\
+		\endfirsthead
+		\multicolumn{4}{c}{\tablename\ \thetable{}\ (continued)}\\
+		\hline
+		\hline
+		\endhead
+		\hline
+		\hline
+		\rule{0pt}{2.25ex}Cr      & \gls{bmi} only                           & 0.001852   & 0.2797              \\
+                                  & \gls{bmi} and \gls{bmi} status           & 0.02834    & 0.05237             \\
+                                  & Pathways only                            & 0.2416     & 1.431e-07 \tnote{1} \\
+                                  & \gls{bmi} and pathways                   & 0.2466     & 1.031e-07           \\
+                                  & \gls{bmi} status and pathways            & 0.2719     & 1.883e-08           \\
+                                  & \gls{bmi}, \gls{bmi} status and pathways & 0.2814     & 9.811e-09           \\
+                                  & \gls{pr} only                            & 0.2041     & 1.604e-06           \\
+                                  & \gls{bmi} and \gls{pr}                   & 0.2047     & 1.54e-06            \\
+                                  & \gls{bmi} status and \gls{pr}            & 0.2249     & 4.237e-07           \\
+                                  & \gls{bmi}, \gls{bmi} status and \gls{pr} & 0.2372     & 1.908e-07           \\
+		\hline
+		\rule{0pt}{2.25ex}CrOl    & \gls{bmi} only                           & 0.001295   & 0.2910              \\
+                                  & \gls{bmi} and \gls{bmi} status           & 0.03509    & 0.03517             \\
+                                  & Pathways only                            & 0.2041     & 1.604e-06           \\
+                                  & \gls{bmi} and pathways                   & 0.2173     & 6.941e-07           \\
+                                  & \gls{bmi} status and pathways            & 0.2411     & 1.48e-07            \\
+                                  & \gls{bmi}, \gls{bmi} status and pathways & 0.2597     & 4.308e-08           \\
+                                  & \gls{pr} only                            & 0.2024     & 1.786e-06           \\
+                                  & \gls{bmi} and \gls{pr}                   & 0.2056     & 1.458e-06           \\
+                                  & \gls{bmi} status and \gls{pr}            & 0.2212     & 5.398e-07           \\
+                                  & \gls{bmi}, \gls{bmi} status and \gls{pr} & 0.2336     & 2.417e-07           \\
+		\hline
+		\rule{0pt}{2.25ex}Res     & \gls{bmi} only                           & 0.0008245  & 0.3011              \\
+                                  & \gls{bmi} and \gls{bmi} status           & 0.02452    & 0.06579             \\
+                                  & Pathways only                            & 0.2785     & 1.196e-08           \\
+                                  & \gls{bmi} and pathways                   & 0.2780     & 1.233e-08           \\
+                                  & \gls{bmi} status and pathways            & 0.2914     & 4.887e-09           \\
+                                  & \gls{bmi}, \gls{bmi} status and pathways & 0.3195     & 6.576e-10           \\
+                                  & \gls{pr} only                            & 0.2331     & 2.492e-07           \\
+                                  & \gls{bmi} and \gls{pr}                   & 0.2349     & 2.215e-07           \\
+                                  & \gls{bmi} status and \gls{pr}            & 0.2538     & 6.38e-08            \\
+                                  & \gls{bmi}, \gls{bmi} status and \gls{pr} & 0.2631     & 3.421e-08           \\
+		\hline
+		\rule{0pt}{2.25ex}ResOl   & \gls{bmi} only                           & -0.002442  & 0.3851              \\
+                                  & \gls{bmi} and \gls{bmi} status           & 0.03878    & 0.02835             \\
+                                  & Pathways only                            & 0.2384     & 1.762e-07           \\
+                                  & \gls{bmi} and pathways                   & 0.2371     & 1.927e-07           \\
+                                  & \gls{bmi} status and pathways            & 0.2670     & 2.616e-08           \\
+                                  & \gls{bmi}, \gls{bmi} status and pathways & 0.2878     & 6.263e-09           \\
+                                  & \gls{pr} only                            & 0.2251     & 4.198e-07           \\
+                                  & \gls{bmi} and \gls{pr}                   & 0.2257     & 4.044e-07           \\
+                                  & \gls{bmi} status and \gls{pr}            & 0.2392     & 1.675e-07           \\
+                                  & \gls{bmi}, \gls{bmi} status and \gls{pr} & 0.2568     & 5.207e-08           \\
+		\hline
+		\rule{0pt}{2.25ex}Ca      & \gls{bmi} only                           & -0.0004067 & 0.3296              \\
+                                  & \gls{bmi} and \gls{bmi} status           & 0.03576    & 0.03383             \\
+                                  & Pathways only                            & 0.2043     & 1.585e-06           \\
+                                  & \gls{bmi} and pathways                   & 0.2077     & 1.28e-06            \\
+                                  & \gls{bmi} status and pathways            & 0.2342     & 2.321e-07           \\
+                                  & \gls{bmi}, \gls{bmi} status and pathways & 0.2487     & 8.927e-08           \\
+                                  & \gls{pr} only                            & 0.1362     & 0.0001009           \\
+                                  & \gls{bmi} and \gls{pr}                   & 0.1364     & 9.944e-05           \\
+                                  & \gls{bmi} status and \gls{pr}            & 0.1643     & 1.876e-05           \\
+                                  & \gls{bmi}, \gls{bmi} status and \gls{pr} & 0.1744     & 1.012e-05           \\
+		\hline
+		\rule{0pt}{2.25ex}CaOl    & \gls{bmi} only                           & -0.006593  & 0.5509              \\
+                                  & \gls{bmi} and \gls{bmi} status           & 0.0301     & 0.04718             \\
+                                  & Pathways only                            & 0.1323     & 0.0001273           \\
+                                  & \gls{bmi} and pathways                   & 0.1347     & 0.0001106           \\
+                                  & \gls{bmi} status and pathways            & 0.1487     & 4.802e-05           \\
+                                  & \gls{bmi}, \gls{bmi} status and pathways & 0.1657     & 1.725e-05           \\
+                                  & \gls{pr} only                            & 0.1103     & 0.0004599           \\
+                                  & \gls{bmi} and \gls{pr}                   & 0.1106     & 0.0004539           \\
+                                  & \gls{bmi} status and \gls{pr}            & 0.1235     & 0.0002134           \\
+                                  & \gls{bmi}, \gls{bmi} status and \gls{pr} & 0.1383     & 8.92e-05            \\
+		\hline
+		\rule{0pt}{2.25ex}CaRes   & \gls{bmi} only                           & 0.001142   & 0.2943              \\
+                                  & \gls{bmi} and \gls{bmi} status           & 0.03048    & 0.04612             \\
+                                  & Pathways only                            & 0.2181     & 6.586e-07           \\
+                                  & \gls{bmi} and pathways                   & 0.2373     & 1.902e-07           \\
+                                  & \gls{bmi} status and pathways            & 0.2431     & 1.298e-07           \\
+                                  & \gls{bmi}, \gls{bmi} status and pathways & 0.2501     & 8.152e-08           \\
+                                  & \gls{pr} only                            & 0.1566     & 2.979e-05           \\
+                                  & \gls{bmi} and \gls{pr}                   & 0.1586     & 2.654e-05           \\
+                                  & \gls{bmi} status and \gls{pr}            & 0.1844     & 5.485e-06           \\
+                                  & \gls{bmi}, \gls{bmi} status and \gls{pr} & 0.1957     & 2.707e-06           \\
+		\hline
+		\rule{0pt}{2.25ex}CaResOl & \gls{bmi} only                           & -0.006037  & 0.5225              \\
+                                  & \gls{bmi} and \gls{bmi} status           & 0.0413     & 0.02449             \\
+                                  & Pathways only                            & 0.1719     & 1.183e-05           \\
+                                  & \gls{bmi} and pathways                   & 0.1721     & 1.167e-05           \\
+                                  & \gls{bmi} status and pathways            & 0.1961     & 2.642e-06           \\
+                                  & \gls{bmi}, \gls{bmi} status and pathways & 0.2095     & 1.141e-06           \\
+                                  & \gls{pr} only                            & 0.1571     & 2.894e-05           \\
+                                  & \gls{bmi} and \gls{pr}                   & 0.1574     & 2.843e-05           \\
+                                  & \gls{bmi} status and \gls{pr}            & 0.17       & 1.327e-05           \\
+                                  & \gls{bmi}, \gls{bmi} status and \gls{pr} & 0.1939     & 3.044e-06           \\
+		\hline
+		\rule{0pt}{2.25ex}Or      & \gls{bmi} only                           & 0.0008035  & 0.3015              \\
+                                  & \gls{bmi} and \gls{bmi} status           & 0.03699    & 0.03147             \\
+                                  & Pathways only                            & 0.2554     & 5.738e-08           \\
+                                  & \gls{bmi} and pathways                   & 0.2568     & 5.234e-08           \\
+                                  & \gls{bmi} status and pathways            & 0.28       & 1.075e-08           \\
+                                  & \gls{bmi}, \gls{bmi} status and pathways & 0.2974     & 3.194e-09           \\
+                                  & \gls{pr} only                            & 0.2388     & 1.717e-07           \\
+                                  & \gls{bmi} and \gls{pr}                   & 0.2382     & 1.79e-07            \\
+                                  & \gls{bmi} status and \gls{pr}            & 0.2531     & 6.663e-08           \\
+                                  & \gls{bmi}, \gls{bmi} status and \gls{pr} & 0.2684     & 2.382e-08           \\
+		\hline
+		\rule{0pt}{2.25ex}FM      & \gls{bmi} only                           & -0.008238  & 0.6563              \\
+                                  & \gls{bmi} and \gls{bmi} status           & 0.07438    & 0.003652            \\
+                                  & Pathways only                            & 0.268      & 2.449e-08           \\
+                                  & \gls{bmi} and pathways                   & 0.2645     & 3.106e-08           \\
+                                  & \gls{bmi} status and pathways            & 0.3067     & 1.654e-09           \\
+                                  & \gls{bmi}, \gls{bmi} status and pathways & 0.3653     & 2.112e-11           \\
+                                  & \gls{pr} only                            & -0.004177  & 0.4434              \\
+                                  & \gls{bmi} and \gls{pr}                   & 0.007149   & 0.1946              \\
+                                  & \gls{bmi} status and \gls{pr}            & 0.0394     & 0.02734             \\
+                                  & \gls{bmi}, \gls{bmi} status and \gls{pr} & 0.1166     & 0.0003183           \\
+		\hline
+		\hline
+		\insertTableNotes
+	\end{longtable}
+
+	\begin{table}[htpb]
+		\centering
+		\caption{tmp table 2}
+		\label{tab:bmi_status_cris}
+		\begin{tabular}{lccc}
+			& \multicolumn{3}{c}{ P-values} \\
+			\cmidrule(r){2-4}
+			Obesity metagene & Overweight & Obese & \gls{anova} \\
+			\hline
+			\hline
+			\rule{0pt}{2.25ex}Cr & 0.2002 & 0.1501  & 0.2591 \\
+			CrOl                 & 0.3126 & 0.1532  & 0.3134 \\
+			Res                  & 0.9708 & 0.1764  & 0.2576 \\
+			ResOl                & 0.4001 & 0.1919  & 0.3950 \\
+			Ca                   & 0.1847 & 0.1099  & 0.2000 \\
+			CaOl                 & 0.3466 & 0.2325  & 0.4273 \\
+			CaRes                & 0.1713 & 0.1154  & 0.2017 \\
+			CaResOl              & 0.2976 & 0.1742  & 0.3439 \\
+			Or                   & 0.2669 & 0.1621  & 0.3074 \\
+			FM                   & 0.4191 & 0.05207 & 0.1206 \\
+			\hline
+			\hline
+		\end{tabular}
+	\end{table}
+
+	\begin{longtable}{llc{\bfseries}S}
+		\centering
+		\caption{tmp3}
+		\label{tab:prediction_summary_cr}\\
+		Obesity metagene & Model                         & R$^2$      & P-value   \\
+		\endfirsthead
+		\multicolumn{4}{c}{\tablename\ \thetable{}\ (continued)}\\
+		\hline
+		\hline
+		\endhead
+		\hline
+		\hline
+		\rule{0pt}{2.25ex}Cr      & \gls{bmi} only                           & 0.1239     & 0.000157  \\
+                                  & \gls{bmi} and \gls{bmi} status           & 0.0471     & 0.01568   \\
+                                  & Pathways only                            & 0.1756     & 6.307e-06 \\
+                                  & \gls{bmi} and pathways                   & 0.1407     & 5.596e-05 \\
+                                  & \gls{bmi} status and pathways            & 0.05509    & 0.009721  \\
+                                  & \gls{bmi}, \gls{bmi} status and pathways & 0.04186    & 0.02147   \\
+                                  & \gls{pr} only                            & 0.1327     & 9.191e-05 \\
+                                  & \gls{bmi} and \gls{pr}                   & 0.1116     & 0.0003317 \\
+                                  & \gls{bmi} status and \gls{pr}            & 0.03887    & 0.02571   \\
+                                  & \gls{bmi}, \gls{bmi} status and \gls{pr} & 0.04176    & 0.02161   \\
+		\hline
+		\rule{0pt}{2.25ex}CrOl    & \gls{bmi} only                           & 0.18       & 4.738e-06 \\
+                                  & \gls{bmi} and \gls{bmi} status           & 0.06523    & 0.005312  \\
+                                  & Pathways only                            & 0.1296     & 0.000111  \\
+                                  & \gls{bmi} and pathways                   & 0.09006    & 0.00121   \\
+                                  & \gls{bmi} status and pathways            & 0.02455    & 0.06179   \\
+                                  & \gls{bmi}, \gls{bmi} status and pathways & 0.01941    & 0.08532   \\
+                                  & \gls{pr} only                            & 0.1817     & 4.249e-06 \\
+                                  & \gls{bmi} and \gls{pr}                   & 0.1558     & 2.188e-05 \\
+                                  & \gls{bmi} status and \gls{pr}            & 0.07293    & 0.003359  \\
+                                  & \gls{bmi}, \gls{bmi} status and \gls{pr} & 0.07116    & 0.003732  \\
+		\hline
+		\rule{0pt}{2.25ex}Res     & \gls{bmi} only                           & 0.09072    & 0.001163  \\
+                                  & \gls{bmi} and \gls{bmi} status           & 0.04082    & 0.02287   \\
+                                  & Pathways only                            & 0.2263     & 2.275e-07 \\
+                                  & \gls{bmi} and pathways                   & 0.2012     & 1.206e-06 \\
+                                  & \gls{bmi} status and pathways            & 0.1184     & 0.0002191 \\
+                                  & \gls{bmi}, \gls{bmi} status and pathways & 0.1005     & 0.0006487 \\
+                                  & \gls{pr} only                            & 0.1419     & 5.195e-05 \\
+                                  & \gls{bmi} and \gls{pr}                   & 0.1224     & 0.0001718 \\
+                                  & \gls{bmi} status and \gls{pr}            & 0.05342    & 0.01074   \\
+                                  & \gls{bmi}, \gls{bmi} status and \gls{pr} & 0.05419    & 0.01026   \\
+		\hline
+		\rule{0pt}{2.25ex}ResOl   & \gls{bmi} only                           & 0.1535     & 2.536e-05 \\
+                                  & \gls{bmi} and \gls{bmi} status           & 0.0466     & 0.01615   \\
+                                  & Pathways only                            & 0.1581     & 1.899e-05 \\
+                                  & \gls{bmi} and pathways                   & 0.1329     & 9.083e-05 \\
+                                  & \gls{bmi} status and pathways            & 0.05085    & 0.01252   \\
+                                  & \gls{bmi}, \gls{bmi} status and pathways & 0.04523    & 0.01754   \\
+                                  & \gls{pr} only                            & 0.2067     & 8.421e-07 \\
+                                  & \gls{bmi} and \gls{pr}                   & 0.1936     & 1.979e-06 \\
+                                  & \gls{bmi} status and \gls{pr}            & 0.1149     & 0.0002723 \\
+                                  & \gls{bmi}, \gls{bmi} status and \gls{pr} & 0.1097     & 0.0003713 \\
+		\hline
+		\rule{0pt}{2.25ex}Ca      & \gls{bmi} only                           & 0.07979    & 0.002233  \\
+                                  & \gls{bmi} and \gls{bmi} status           & 0.05026    & 0.01297   \\
+                                  & Pathways only                            & 0.2379     & 1.042e-07 \\
+                                  & \gls{bmi} and pathways                   & 0.2081     & 7.653e-07 \\
+                                  & \gls{bmi} status and pathways            & 0.09383    & 0.0009655 \\
+                                  & \gls{bmi}, \gls{bmi} status and pathways & 0.07756    & 0.00255   \\
+                                  & \gls{pr} only                            & 0.09938    & 0.000692  \\
+                                  & \gls{bmi} and \gls{pr}                   & 0.08114    & 0.00206   \\
+                                  & \gls{bmi} status and \gls{pr}            & 0.01059    & 0.1512    \\
+                                  & \gls{bmi}, \gls{bmi} status and \gls{pr} & 0.008212   & 0.1774    \\
+		\hline
+		\rule{0pt}{2.25ex}CaOl    & \gls{bmi} only                           & 0.1242     & 0.0001542 \\
+                                  & \gls{bmi} and \gls{bmi} status           & 0.05609    & 0.009156  \\
+                                  & Pathways only                            & 0.08453    & 0.001683  \\
+                                  & \gls{bmi} and pathways                   & 0.07397    & 0.003157  \\
+                                  & \gls{bmi} status and pathways            & 0.003647   & 0.244     \\
+                                  & \gls{bmi}, \gls{bmi} status and pathways & 0.0003557  & 0.3111    \\
+                                  & \gls{pr} only                            & 0.1253     & 0.0001444 \\
+                                  & \gls{bmi} and \gls{pr}                   & 0.1243     & 0.0001534 \\
+                                  & \gls{bmi} status and \gls{pr}            & 0.02838    & 0.04876   \\
+                                  & \gls{bmi}, \gls{bmi} status and \gls{pr} & 0.02625    & 0.0556    \\
+		\hline
+		\rule{0pt}{2.25ex}CaRes   & \gls{bmi} only                           & 0.05877    & 0.007805  \\
+                                  & \gls{bmi} and \gls{bmi} status           & 0.0415     & 0.02194   \\
+                                  & Pathways only                            & 0.2438     & 6.952e-08 \\
+                                  & \gls{bmi} and pathways                   & 0.2037     & 1.023e-06 \\
+                                  & \gls{bmi} status and pathways            & 0.1132     & 0.0003018 \\
+                                  & \gls{bmi}, \gls{bmi} status and pathways & 0.0995     & 0.0006872 \\
+                                  & \gls{pr} only                            & 0.1096     & 0.0003748 \\
+                                  & \gls{bmi} and \gls{pr}                   & 0.08751    & 0.001409  \\
+                                  & \gls{bmi} status and \gls{pr}            & 0.02671    & 0.05406   \\
+                                  & \gls{bmi}, \gls{bmi} status and \gls{pr} & 0.02392    & 0.06429   \\
+		\hline
+		\rule{0pt}{2.25ex}CaResOl & \gls{bmi} only                           & 0.1167     & 0.0002439 \\
+                                  & \gls{bmi} and \gls{bmi} status           & 0.04905    & 0.01395   \\
+                                  & Pathways only                            & 0.07821    & 0.002453  \\
+                                  & \gls{bmi} and pathways                   & 0.06736    & 0.004681  \\
+                                  & \gls{bmi} status and pathways            & -0.0005575 & 0.3338    \\
+                                  & \gls{bmi}, \gls{bmi} status and pathways & -0.003248  & 0.4151    \\
+                                  & \gls{pr} only                            & 0.1632     & 1.377e-05 \\
+                                  & \gls{bmi} and \gls{pr}                   & 0.1628     & 1.412e-05 \\
+                                  & \gls{bmi} status and \gls{pr}            & 0.05905    & 0.007678  \\
+                                  & \gls{bmi}, \gls{bmi} status and \gls{pr} & 0.05391    & 0.01043   \\
+		\hline
+		\rule{0pt}{2.25ex}Or      & \gls{bmi} only                           & 0.1246     & 0.0001508 \\
+                                  & \gls{bmi} and \gls{bmi} status           & 0.03912    & 0.02533   \\
+                                  & Pathways only                            & 0.1649     & 1.237e-05 \\
+                                  & \gls{bmi} and pathways                   & 0.1311     & 0.0001014 \\
+                                  & \gls{bmi} status and pathways            & 0.05614    & 0.009131  \\
+                                  & \gls{bmi}, \gls{bmi} status and pathways & 0.04897    & 0.01402   \\
+                                  & \gls{pr} only                            & 0.1805     & 4.604e-06 \\
+                                  & \gls{bmi} and \gls{pr}                   & 0.1626     & 1.431e-05 \\
+                                  & \gls{bmi} status and \gls{pr}            & 0.08503    & 0.001634  \\
+                                  & \gls{bmi}, \gls{bmi} status and \gls{pr} & 0.08819    & 0.001353  \\
+		\hline
+		\rule{0pt}{2.25ex}FM      & \gls{bmi} only                           & -0.004507  & 0.4632    \\
+                                  & \gls{bmi} and \gls{bmi} status           & 0.01947    & 0.08502   \\
+                                  & Pathways only                            & 0.4202     & 8.085e-14 \\
+                                  & \gls{bmi} and pathways                   & 0.4213     & 7.317e-14 \\
+                                  & \gls{bmi} status and pathways            & 0.3386     & 6.885e-11 \\
+                                  & \gls{bmi}, \gls{bmi} status and pathways & 0.2734     & 8.759e-09 \\
+                                  & \gls{pr} only                            & 0.09798    & 0.0007529 \\
+                                  & \gls{bmi} and \gls{pr}                   & 0.09642    & 0.0008267 \\
+                                  & \gls{bmi} status and \gls{pr}            & -0.00988   & 0.9634    \\
+                                  & \gls{bmi}, \gls{bmi} status and \gls{pr} & -0.008862  & 0.7478    \\
+		\hline
+		\hline
+		\insertTableNotes
+	\end{longtable}
+
+	\begin{table}[htpb]
+		\centering
+		\caption{tmp table 4}
+		\label{tab:bmi_status_cr}
+		\begin{threeparttable}
+			\begin{tabular}{lccc}
+				& \multicolumn{3}{c}{ P-values} \\
+				\cmidrule(r){2-4}
+				Obesity metagene & Overweight & Obese & \gls{anova} \\
+				\hline
+				\hline
+				\rule{0pt}{2.25ex}Cr & 3.005e-06 & 0.0002419 & 6.931e-06  \\
+				CrOl                 & 2.924e-07 & 8.516e-06 & 1.467e-07  \\
+				Res                  & 0.2783    & 2.146e-05 & 5.715e-05  \\
+				ResOl                & 2.023e-06 & 3.31e-05  & 1.052e-06  \\
+				Ca                   & 0.00283   & 8.769e-05 & 0.0002354  \\
+				CaOl                 & 0.0001869 & 3.328e-06 & 4.371e-06  \\
+				CaRes                & 0.008582  & 0.0004552 & 0.001334   \\
+				CaResOl              & 0.0004192 & 5.338e-06 & 8.806e-06  \\
+				Or                   & 5.806e-06 & 0.0002258 & 8.092e-06  \\
+				FM                   & 0.8573    & 0.185     & 0.3568     \\
+				\hline
+				\hline
+			\end{tabular}
+			\begin{tablenotes}
+				\begin{footnotesize}
+				\item [1] All values in bold are statistically significant (p \textless{} 0.05).
+				\end{footnotesize}
+			\end{tablenotes}
+		\end{threeparttable}
+	\end{table}
+
 \end{appendices}

### 2017-03-14 thesis/misc/appendix.tex
Draft thesis completed; now edit the whole thing - tmp commit

@@
 
-	\chapter{Additional results for \cref{cha:obesity_genetic_signatures_and_cancer}}
+	\chapter{Additional results from \cref{cha:obesity_genetic_signatures_and_cancer}}
 	\label{app:a}
@@
 		\includegraphics[width=0.45\linewidth,page=2]{appendix/check_raw_vs_std_mg}
-		\caption{Scatter plots showing the raw and ranked Creighton \textit{et al.} obesity metagene values generated from standardised or non-standardised CR data set. }
+		\caption[Comparison of the raw and ranked Creighton \textit{et al.} obesity metagene scores from the standardised or non-standardised CR data]{Scatter plots showing the raw and ranked Creighton \textit{et al.} obesity metagene scores from the standardised or non-standardised CR data.
+		The \gls{rma}-normalised CR data was standardised or used as is to generate the obesity metagene scores.
+		The metagene scores were ranked based on the number of samples in the CR data set.
+		}
 		\label{fig:appendix/check_raw_vs_std}
@@
 		\includegraphics[width=0.45\linewidth,page=2]{appendix/crtcga_raw_vs_std}\\
-		\caption{Scatter plots comparing the Creighton \textit{et al.} obesity metagene generated from transformation matrix (from standardised or non-standardised CR data) in non-standardised (left) and standardised (right) \gls{icgc} \gls{blca} data.}
+		\caption[Comparison of the Creighton \textit{et al.} obesity metagene generated from the standardised or non-standardised TM]{Scatter plots comparing the Creighton \textit{et al.} obesity metagene generated from the transformation matrix (from standardised or non-standardised CR data) in the non-standardised (left) and standardised (right) \gls{icgc} \gls{blca} data.}
 		\label{fig:appendix/check_raw_vs_std}
@@
 		\includegraphics[width=0.45\linewidth,page=4]{appendix/crtcga_raw_vs_std}\\
-		\caption{Scatter plots comparing the Creighton \textit{et al.} obesity metagene generated from transformation matrix from non-standardisedright (left) or standardised (right) CR data in non-standardised  and standardised  \gls{icgc} \gls{blca} data.}
+		\caption[Comparison of the Creighton \textit{et al.} obesity metagene generated from the standardised or non-standardised  \gls{icgc} \gls{blca} data]{Scatter plots comparing the Creighton \textit{et al.} obesity metagene from the non-standardised  and standardised  \gls{icgc} \gls{blca} data, generated by the application of the transformation matrix from the non-standardised (left) or standardised (right) CR data.}
 		\label{fig:appendix/check_raw_vs_std_tm}
@@
 
-	\section{Remainder of the results of Creighton \textit{et al.} obesity metagenes in the \gls{icgc} cancer data sets}
+	\section{Remainder of the results of the other Creighton \textit{et al.} obesity metagenes in the \gls{icgc} cancer data sets}
 	\label{sec:rest_of_the_cr_icgc_cancer_heatmap_results}
@@
 		\includegraphics[width=0.32\linewidth,page=38]{appendix/crtcga_std}\\
-		\caption{Heatmaps showing the association of obesity metagene from the \citet{Creighton2012} study with sample gene expression from the other \gls{icgc} data sets.
+		\caption[Obesity metagene from the \citet{Creighton2012} study and the sample gene expressions in the other \gls{icgc} cancer types]{Heatmaps showing the association of obesity metagene from the \citet{Creighton2012} study with sample gene expression from the other \gls{icgc} data sets.
 	Level of expression is represented in the top right histogram, where low and high gene expression were colour-coded with blue and red, respectively.
@@
 		\includegraphics[page=15,width=0.45\linewidth]{appendix/crtcga_std}\\
-		\caption{Box plots and scatter plots showing the association of obesity metagene from the \citet{Creighton2012} study with the sample \gls{bmi} status  and \gls{bmi} from the other \gls{icgc} data sets, respectively.
+		\caption[Obesity metagene from the \citet{Creighton2012} study and the sample \gls{bmi}/\gls{bmi} status in the other \gls{icgc} cancer types]{Box plots and scatter plots showing the association of obesity metagene from the \citet{Creighton2012} study with the sample \gls{bmi} status  and \gls{bmi} from the other \gls{icgc} data sets, respectively.
 	In the box plot, the p-values above the groups represent the statistical significance of the association of the metagene with the overweight or obese group compared with the normal weight group.
@@
 	\centering
-	\caption{Statistics of all the obesity metagenes in \gls{cesc} cancer data}
+	\caption[Statistics of all the obesity metagenes in the \gls{icgc} \acrshort{cesc} cancer data]{Statistics of all the obesity metagenes in the \gls{icgc} \gls{cesc} cancer data}
 	\label{tab:degmetacesc}
@@
 	\centering
-	\caption{Statistics of all the obesity metagenes in \gls{coad} cancer data}
+	\caption[Statistics of all the obesity metagenes in the \gls{icgc} \acrshort{coad} cancer data]{Statistics of all the obesity metagenes in the \gls{icgc} \gls{coad} cancer data}
 	\label{tab:degmetacoad}
@@
 	\centering
-	\caption{Statistics of all the obesity metagenes in \gls{kirp} cancer data}
+	\caption[Statistics of all the obesity metagenes in the \gls{icgc} \acrshort{kirp} cancer data]{Statistics of all the obesity metagenes in the \gls{icgc} \gls{kirp} cancer data}
 	\label{tab:degmetakirp}
@@
 	\centering
-	\caption{Statistics of all the obesity metagenes in \gls{lihc} cancer data}
+	\caption[Statistics of all the obesity metagenes in the \gls{icgc} \acrshort{lihc} cancer data]{Statistics of all the obesity metagenes in the \gls{icgc} \gls{lihc} cancer data}
 	\label{tab:degmetalihc}
@@
 	\centering
-	\caption{Statistics of all the obesity metagenes in \gls{read} cancer data}
+	\caption[Statistics of all the obesity metagenes in the \gls{icgc} \acrshort{read} cancer data]{Statistics of all the obesity metagenes in the \gls{icgc} \gls{read} cancer data}
 	\label{tab:degmetaread}
@@
 	\centering
-	\caption{Statistics of all the obesity metagenes in \gls{skcm} cancer data}
+	\caption[Statistics of all the obesity metagenes in the \gls{icgc} \acrshort{skcm} cancer data]{Statistics of all the obesity metagenes in the \gls{icgc} \gls{skcm} cancer data}
 	\label{tab:degmetaskcm}
@@
 	\centering
-	\caption{Statistics of all the obesity metagenes in \gls{ucec} cancer data}
+	\caption[Statistics of all the obesity metagenes in the \gls{icgc} \acrshort{ucec} cancer data]{Statistics of all the obesity metagenes in the \gls{icgc} \gls{ucec} cancer data}
 	\label{tab:degmetaucec}
@@
 
+	\section{Remainder of the results of FM obesity metagenes in the \gls{icgc} cancer data sets}
+	\label{sec:rest_of_the_fm_icgc_cancer_heatmap_results}
+
+%TODO: add FM metagene plots for ICGC
+
 	\section{Remainder of the results of the other CR obesity metagenes in \gls{nzbc} data set}
@@
 		\includegraphics[width=0.32\linewidth,page=38]{appendix/cris_crdeg_trans_meta}\\
-		\caption{Heatmaps showing the association of the other CR obesity metagenes with the sample gene expression from the \gls{nzbc} data set.
+		\caption[Association of the other CR obesity metagenes with the sample gene expressions in the \gls{nzbc} data]{Heatmaps showing the association of the other obesity metagenes identified from the CR data with the sample gene expressions from the \gls{nzbc} data set.
 	Scales are as described in previous figures.}
@@
 		\includegraphics[page=10,width=0.42\linewidth]{appendix/cris_crdeg_trans_meta}\\
-		\caption{Box plots and scatter plots showing the association of the other CR obesity metagenes with the sample \gls{bmi} status  and \gls{bmi} from the \gls{nzbc} data set, respectively.
+		\caption[Association of the other CR obesity metagenes with the sample \gls{bmi}/\gls{bmi} status in the \gls{nzbc} data]{Box plots and scatter plots showing the association of the other CR obesity metagenes with the sample \gls{bmi} status  and \gls{bmi} from the \gls{nzbc} data set, respectively.
 	P-values and $R^2$-value are as described in previous figures.}

### 2017-03-14 thesis/misc/appendix2.tex
Draft thesis completed; now edit the whole thing - tmp commit

@@
 		\includegraphics[width=0.32\linewidth,page=12]{appendix/gatza_rma_meta_rank_vs_probit}\\
-		\caption[]{Scatter plots comparing the GT pathway metagenes ranked with probit and ranked based on number of samples in \gls{rma} normalised GT data. }
+		\caption[Comparison of the ranking methods for the pathway metagenes in the \acrshort{rma}-normalised GT data]{Scatter plots comparing the GT pathway metagenes ranked with probit and ranked based on number of samples in \gls{rma} normalised GT data.
+}
 		\label{fig:appendix/rank_scatter}
@@
 		\includegraphics[width=0.32\linewidth,page=12]{appendix/allmeta_rma_vs_mas}\\
-		\caption[]{Scatter plots comparing the GT pathway metagenes derived from the transformation matrices (TM) in \gls{rma} or \gls{mas} normalised GT data set.
+		\caption[Comparison of the normalisation methods used for the pathway metagene TM and the GT data]{Scatter plots comparing the GT pathway metagenes derived from the transformation matrices in \gls{rma}- or \gls{mas}-normalised GT data set.
 		All of the transformation matrices were generated in either the \gls{rma} or \gls{mas} normalised GT data set.
@@
 		\includegraphics[width=0.32\linewidth,page=6]{appendix/gtrmastdrank}\\
-		\caption[]{Heatmaps showing the gene expression of the pathway associated genetic signatures, pathway metagenes, and the expression of the representative pathway gene in \gls{rma} normalised GT data set.
+		\caption[Directionality of the pathway metagenes in the GT data]{Heatmaps showing the gene expression of the pathway associated genetic signatures, pathway metagenes, and the expression of the representative pathway gene in \gls{rma}-normalised GT data set.
 		For each pathway associated genetic signature, the gene expression pattern of the signature was plotted in the heatmap.
@@
 		\includegraphics[width=0.7\linewidth]{results2/gatzares1}
-		\caption[]{Heatmap showing the correlation of all of the GT pathway associated genetic signatures with one another.
+		\caption[Original result presented in the \citet{Gatza2010a} study]{Heatmap showing the correlation of all of the GT pathway associated genetic signatures with one another.
 		Figure adapted from the \citet{Gatza2010a} study.}
@@
 		\includegraphics[width=0.65\linewidth,page=7]{appendix/gatza_meta_trans}
-		\caption[]{Heatmaps showing the correlation of all of the pathway metagenes with one another in \gls{rma} normalised CR, \gls{nzbc} and FM data sets.
-		}
+		\caption[Heatmaps of the Pearson correlation of all the pathway metagenes in the \gls{rma}-normalised CR, \gls{nzbc} and FM data]{Heatmaps showing the correlation of all of the pathway metagenes with one another in \gls{rma} normalised CR, \gls{nzbc} and FM data sets.
+		High and low correlation were represented as red and blue, respectively, where the colours were matched with the values on the scale shown in the top left histogram.}
 		\label{fig:appendix/gt_meta_rma_mas}
@@
 		\includegraphics[width=0.32\linewidth,page=9]{appendix/gtobrma}\\
-		\caption[]{Heatmaps showing the obesity metagenes from the CR data set with the gene expressions of the obesity associated genetic signatures in the \gls{rma} normalised GT data set.
-		}
+		\caption[Heatmaps of all the obesity and the pathway metagenes in the CR, \gls{nzbc} and FM data]{Heatmaps showing the obesity metagenes from the CR data set with the gene expressions of the obesity associated genetic signatures in the \gls{rma} normalised GT data set.
+		High and low correlation were represented as red and blue, respectively, where the colours were matched with the values on the scale shown in the top left histogram.}
 		\label{fig:appendix/ob_meta_dir_gt}
@@
 		\includegraphics[width=0.24\linewidth,page=4]{results2/allmetacor_bar}\\
-		\caption[]{Visual comparison of the Spearman correlations of the \gls{svd}-derived and transformation matrix-derived pathway and obesity metagenes across different data sets}
+		\caption[Bar plots showing the Spearman correlation of the \gls{svd}- and TM-derived obesity and pathway metagenes across different data sets]{Visual comparison of the Spearman correlations of the \gls{svd}-derived and transformation matrix-derived pathway and obesity metagenes across different data sets}
 		\label{fig:allmetacor_bar_appendix}
@@
 		\includegraphics[width=0.7\linewidth,page=3]{appendix/all_meta_trans}
-		\caption[]{Heatmaps showing the correlation of all of the pathway and obesity metagenes with one another in \gls{rma} normalised CR, \gls{nzbc} and FM data sets.
+		\caption[Heatmaps of the Pearson correlation of all the obesity and the pathway metagenes with one another in the \gls{rma}-normalised CR, \gls{nzbc} and FM data]{Heatmaps showing the correlation of all of the pathway and obesity metagenes with one another in the \gls{rma}-normalised CR, \gls{nzbc} and FM data sets.
 		}
@@
 		\centering
-		\caption[]{Description of the linear models used to predict the CrOl obesity metagene in \gls{nzbc} data set}
+		\caption{Description of the linear models constructed from the \gls{nzbc} data to predict the CrOl obesity metagene}
 		\label{tab:lm_sig_var_crol}
@@
 		\centering
-		\caption[]{Description of the linear models used to predict the Res obesity metagene in \gls{nzbc} data set}
+		\caption{Description of the linear models constructed from the \gls{nzbc} data to predict the Res obesity metagene}
 		\label{tab:lm_sig_var_res}
@@
 		\centering
-		\caption[]{Description of the linear models used to predict the ResOl obesity metagene in \gls{nzbc} data set}
+		\caption{Description of the linear models constructed from the \gls{nzbc} data to predict the ResOl obesity metagene}
 		\label{tab:lm_sig_var_resol}
@@
 		\centering
-		\caption[]{Description of the linear models used to predict the Ca obesity metagene in \gls{nzbc} data set}
+		\caption{Description of the linear models constructed from the \gls{nzbc} data to predict the Ca obesity metagene}
 		\label{tab:lm_sig_var_ca}
@@
 		\centering
-		\caption[]{Description of the linear models used to predict the CaOl obesity metagene in \gls{nzbc} data set}
+		\caption{Description of the linear models constructed from the \gls{nzbc} data to predict the CaOl obesity metagene}
 		\label{tab:lm_sig_var_caol}
@@
 		\centering
-		\caption[]{Description of the linear models used to predict the CaRes obesity metagene in \gls{nzbc} data set}
+		\caption{Description of the linear models constructed from the \gls{nzbc} data to predict the CaRes obesity metagene}
 		\label{tab:lm_sig_var_cares}
@@
 		\centering
-		\caption[]{Description of the linear models used to predict the CaResOl obesity metagene in \gls{nzbc} data set}
+		\caption{Description of the linear models constructed from the \gls{nzbc} data to predict the CaResOl obesity metagene}
 		\label{tab:lm_sig_var_caresol}
@@
 		\centering
-		\caption[]{Description of the linear models used to predict the Or obesity metagene in \gls{nzbc} data set}
+		\caption{Description of the linear models constructed from the \gls{nzbc} data to predict the Or obesity metagene}
 		\label{tab:lm_sig_var_or}
@@
 		\centering
-		\caption[]{Description of the linear models used to predict the FM obesity metagene in \gls{nzbc} data set}
+		\caption{Description of the linear models constructed from the \gls{nzbc} data to predict the FM obesity metagene}
 		\label{tab:lm_sig_var_fm}
@@
 		\centering
-		\caption[]{Description of the linear models used to predict the CrOl obesity metagene in \gls{nzbc} data set, using only the sample \gls{bmi}, \gls{bmi} status and \gls{pr} pathway metagene scores}
+		\caption[Description of the linear models constructed from the \gls{nzbc} data to predict the CrOl obesity, using only the sample \gls{bmi}, \gls{bmi} status and the \acrshort{pr} pathway metagene score]{Description of the linear models constructed from the \gls{nzbc} data to predict the CrOl obesity, using only the sample \gls{bmi}, \gls{bmi} status and the \gls{pr} pathway metagene score}
 		\label{tab:lm_pr_only_crol}
@@
 		\centering
-		\caption[]{Description of the linear models used to predict the Res obesity metagene in \gls{nzbc} data set, using only the sample \gls{bmi}, \gls{bmi} status and \gls{pr} pathway metagene scores}
+		\caption[Description of the linear models constructed from the \gls{nzbc} data to predict the Res obesity, using only the sample \gls{bmi}, \gls{bmi} status and the \acrshort{pr} pathway metagene score]{Description of the linear models constructed from the \gls{nzbc} data to predict the Res obesity, using only the sample \gls{bmi}, \gls{bmi} status and the \gls{pr} pathway metagene score}
 		\label{tab:lm_pr_only_res}
@@
 		\centering
-		\caption[]{Description of the linear models used to predict the ResOl obesity metagene in \gls{nzbc} data set, using only the sample \gls{bmi}, \gls{bmi} status and \gls{pr} pathway metagene scores}
+		\caption[Description of the linear models constructed from the \gls{nzbc} data to predict the ResOl obesity, using only the sample \gls{bmi}, \gls{bmi} status and the \acrshort{pr} pathway metagene score]{Description of the linear models constructed from the \gls{nzbc} data to predict the ResOl obesity, using only the sample \gls{bmi}, \gls{bmi} status and the \gls{pr} pathway metagene score}
 		\label{tab:lm_pr_only_resol}
@@
 		\centering
-		\caption[]{Description of the linear models used to predict the Ca obesity metagene in \gls{nzbc} data set, using only the sample \gls{bmi}, \gls{bmi} status and \gls{pr} pathway metagene scores}
+		\caption[Description of the linear models constructed from the \gls{nzbc} data to predict the Ca obesity, using only the sample \gls{bmi}, \gls{bmi} status and the \acrshort{pr} pathway metagene score]{Description of the linear models constructed from the \gls{nzbc} data to predict the Ca obesity, using only the sample \gls{bmi}, \gls{bmi} status and the \gls{pr} pathway metagene score}
 		\label{tab:lm_pr_only_ca}
@@
 		\centering
-		\caption[]{Description of the linear models used to predict the CaOl obesity metagene in \gls{nzbc} data set, using only the sample \gls{bmi}, \gls{bmi} status and \gls{pr} pathway metagene scores}
+		\caption[Description of the linear models constructed from the \gls{nzbc} data to predict the CaOl obesity, using only the sample \gls{bmi}, \gls{bmi} status and the \acrshort{pr} pathway metagene score]{Description of the linear models constructed from the \gls{nzbc} data to predict the CaOl obesity, using only the sample \gls{bmi}, \gls{bmi} status and the \gls{pr} pathway metagene score}
 		\label{tab:lm_pr_only_caol}
@@
 		\centering
-		\caption[]{Description of the linear models used to predict the CaRes obesity metagene in \gls{nzbc} data set, using only the sample \gls{bmi}, \gls{bmi} status and \gls{pr} pathway metagene scores}
+		\caption[Description of the linear models constructed from the \gls{nzbc} data to predict the CaRes obesity, using only the sample \gls{bmi}, \gls{bmi} status and the \acrshort{pr} pathway metagene score]{Description of the linear models constructed from the \gls{nzbc} data to predict the CaRes obesity, using only the sample \gls{bmi}, \gls{bmi} status and the \gls{pr} pathway metagene score}
 		\label{tab:lm_pr_only_cares}
@@
 		\centering
-		\caption[]{Description of the linear models used to predict the CaResOl obesity metagene in \gls{nzbc} data set, using only the sample \gls{bmi}, \gls{bmi} status and \gls{pr} pathway metagene scores}
+		\caption[Description of the linear models constructed from the \gls{nzbc} data to predict the CaResOl obesity, using only the sample \gls{bmi}, \gls{bmi} status and the \acrshort{pr} pathway metagene score]{Description of the linear models constructed from the \gls{nzbc} data to predict the CaResOl obesity, using only the sample \gls{bmi}, \gls{bmi} status and the \gls{pr} pathway metagene score}
 		\label{tab:lm_pr_only_caresol}
@@
 		\centering
-		\caption[]{Description of the linear models used to predict the Or obesity metagene in \gls{nzbc} data set, using only the sample \gls{bmi}, \gls{bmi} status and \gls{pr} pathway metagene scores}
+		\caption[Description of the linear models constructed from the \gls{nzbc} data to predict the Or obesity, using only the sample \gls{bmi}, \gls{bmi} status and the \acrshort{pr} pathway metagene score]{Description of the linear models constructed from the \gls{nzbc} data to predict the Or obesity, using only the sample \gls{bmi}, \gls{bmi} status and the \gls{pr} pathway metagene score}
 		\label{tab:lm_pr_only_or}
@@
 		\centering
-		\caption[]{Description of the linear models used to predict the FM obesity metagene in \gls{nzbc} data set, using only the sample \gls{bmi}, \gls{bmi} status and \gls{pr} pathway metagene scores}
+		\caption[Description of the linear models constructed from the \gls{nzbc} data to predict the FM obesity, using only the sample \gls{bmi}, \gls{bmi} status and the \acrshort{pr} pathway metagene score]{Description of the linear models constructed from the \gls{nzbc} data to predict the FM obesity, using only the sample \gls{bmi}, \gls{bmi} status and the \gls{pr} pathway metagene score}
 		\label{tab:lm_pr_only_fm}
@@
 
+	\section{Summary statistics of the predicted obesity metagenes with sample \gls{bmi}/\gls{bmi} status in \gls{nzbc} and CR data}
+	\label{sec:summary_statistics_of_the_predicted_obesity_metagenes_with_sample_bmi_bmi_status_in_nzbc_and_cr_data}
+
 	\begin{TableNotes}
@@
 		\centering
-		\caption{tmp}
+		\caption[Summary of the statistics from the comparison of all the predicted obesity metagene scores with the corresponding obesity metagenes from the \gls{nzbc} data]{Table summarising the linear model statistics from the comparison of the various linear model-predicted obesity metagenes scores with the corresponding obesity metagene scores from the \gls{nzbc} data set for all of the obesity metagenes}
 		\label{tab:prediction_summary_cris}\\
@@
 		\centering
-		\caption{tmp table 2}
+		\caption[Summary of the statistics from the comparison of all the predicted obesity metagene scores from the \gls{bmi} status-only model with the corresponding obesity metagenes from the \gls{nzbc} data]{Table summarising the linear model statistics from the comparison of the predicted obesity metagenes scores from the \gls{bmi} status-only model with the corresponding obesity metagene scores from the \gls{nzbc} data set for all of the obesity metagenes}
 		\label{tab:bmi_status_cris}
@@
 		\centering
-		\caption{tmp3}
+		\caption[Summary of the statistics from the comparison of all the predicted obesity metagene scores with the corresponding obesity metagenes from the CR data]{Table summarising the linear model statistics from the comparison of the various linear model-predicted obesity metagenes scores with the corresponding obesity metagene scores from the CR data set for all of the obesity metagenes}
 		\label{tab:prediction_summary_cr}\\
@@
 		\centering
-		\caption{tmp table 4}
+		\caption[Summary of the statistics from the comparison of all the predicted obesity metagene scores from the \gls{bmi} status-only model with the corresponding obesity metagenes from the CR data]{Table summarising the linear model statistics from the comparison of the predicted obesity metagenes scores from the \gls{bmi} status-only model with the corresponding obesity metagene scores from the CR data set for all of the obesity metagenes}
 		\label{tab:bmi_status_cr}

### 2017-03-14 thesis/results/results.tex
Added all of the extra results in appendix2; clean up the titles - tmp commit
Draft thesis completed; now edit the whole thing - tmp commit

@@
 	\includegraphics[page=3,width=0.8\linewidth]{results1/creighton_mg_heatmap1}
-	\caption[Obesity metagene from \citet{Creighton2012} study and sample gene expression in CR data]{Heatmap showing the obesity metagene from the \citet{Creighton2012} study with the gene expressions of obesity associated genes in CR data.
-	Level of expression is represented in the top right histogram, where low and high gene expression were colour-coded with blue and red, respectively.
-	Each row of the heatmap represents a gene from the obesity associated genetic signature, and each column of the heatmap represents a sample from CR data.
-	The obesity associated metagene scores of the samples are shown in a separate row at top of the heatmap, and the tree diagram of the heirarchical clustering of the genes is shown to the left of the heatmap.
+	\caption[Obesity metagene from \citet{Creighton2012} study and sample gene expression in CR data] {Heatmap showing the obesity metagene from the \citet{Creighton2012} study with the sample gene expressions of the obesity associated genes in the CR data.
+	Level of expression is represented in the top right histogram, where low and high gene expression are colour-coded with blue and red, respectively.
+	Each row of the heatmap represents a gene from the obesity associated genetic signature, and each column of the heatmap represents a sample from the CR data.
+	The obesity associated metagene scores of the samples are shown as a separate row at top of the heatmap, and the tree diagram of the heirarchical clustering of the genes is shown on the left of the heatmap.
 	For clarity, the metagene scores were flipped in order to match the results from the \citet{Creighton2012} study.}
@@
 	\includegraphics[page=5,width=0.45\linewidth]{results1/creighton_mg_heatmap1}
-	\caption[Obesity metagene from the \citet{Creighton2012} study and sample \gls{bmi}/\gls{bmi} status in CR data]{Box plot and scatter plot showing the association of the obesity metagene from the \citet{Creighton2012} study with sample \gls{bmi} status and \gls{bmi} from CR data, respectively.
-	In the box plot, the p-values above the groups represent the statistical significance of the association of the metagene with the overweight or obese group compared with the normal weight group.
-	The \gls{anova} p-value shows the statistical significance of the association of the metagene with the sample \gls{bmi} groups.
+	\caption[Obesity metagene from the \citet{Creighton2012} study and sample \gls{bmi}/\gls{bmi} status in CR data]{Box plot and scatter plot showing the association of the obesity metagene from the \citet{Creighton2012} study with the sample \gls{bmi} status and \gls{bmi}, respectively, from the CR data.
+	In the box plot, the p-values above the groups represent the statistical significance of the association of the obesity metagene with the overweight or obese group compared with the normal weight group.
+	The \gls{anova} p-value shows the statistical significance of the association of the obesity metagene with the sample \gls{bmi} groups.
 	In the scatter plot, $R^2$- and p-values describe the adjusted coefficient of determination of the regression line and the statistical significance of the linear model used to draw the regression line, respectively.}
@@
 	\includegraphics[page=5,width=0.45\linewidth]{results1/crtcga_std}
-	\caption[Obesity metagene from the \citet{Creighton2012} study in \acrshort{icgc} \acrshort{blca} data]{Heatmap, box plot and scatter plot showing the association of obesity metagene from the \citet{Creighton2012} study with sample gene expression, \gls{bmi} and \gls{bmi} status from \gls{icgc} \gls{blca} data, respectively.
-	The results for other cancer types are in \cref{app:a}.
+	\caption[Obesity metagene from the \citet{Creighton2012} study in the \acrshort{icgc} \acrshort{blca} data]{Heatmap, box plot and scatter plot showing the association of the obesity metagene from the \citet{Creighton2012} study with the sample gene expression, \gls{bmi} and \gls{bmi} status, respectively, from the \gls{icgc} \gls{blca} data.
+	The results for other \gls{icgc} cancer types are shown in \cref{sec:rest_of_the_cr_icgc_cancer_heatmap_results}.
 	Scales, p-values and $R^2$-value are as described in previous figures.}
@@
 	\includegraphics[width=0.45\linewidth,page=5]{results1/cris_cr_trans_meta}
-	\caption[Obesity metagene from \citet{Creighton2012} study in \gls{nzbc} data]{Heatmap, box plot and scatter plot showing the association of obesity metagene from the \citet{Creighton2012} study with sample gene expression, \gls{bmi} and \gls{bmi} status from \gls{nzbc} data, respectively.
+	\caption[Obesity metagene from \citet{Creighton2012} study in the \gls{nzbc} data]{Heatmap, box plot and scatter plot showing the association of the obesity metagene from the \citet{Creighton2012} study with the sample gene expression, \gls{bmi} and \gls{bmi} status from the \gls{nzbc} data, respectively.
 	Scales, p-values and $R^2$-value are as described in previous figures.}
@@
 	\includegraphics[page=5,width=0.45\linewidth]{results1/cr_fm_meta}
-	\caption[FM metagene in CR data]{Heatmap, box plot and scatter plot showing the association of FM obesity associated metagene with sample gene expression, \gls{bmi} and \gls{bmi} status from CR  data, respectively.
+	\caption[FM obesity metagene in the CR data]{Heatmap, box plot and scatter plot showing the association of the FM obesity associated metagene with the sample gene expression, \gls{bmi} and \gls{bmi} status from the CR  data, respectively.
 	Scales, p-values and $R^2$-value are as described in previous figures.}
@@
 	\includegraphics[page=5,width=0.45\linewidth]{results1/fm_meta_ICGC}
-	\caption[FM metagene in \acrshort{icgc} \acrshort{blca} data]{Heatmap, box plot and scatter plot showing the association of FM obesity associated metagene with sample gene expression, \gls{bmi} and \gls{bmi} status from \acrshort{icgc} \acrshort{blca} data, respectively.
-	The results for other cancer types are in \cref{app:a}.
+	\caption[FM obesity metagene in the \acrshort{icgc} \acrshort{blca} data]{Heatmap, box plot and scatter plot showing the association of the FM obesity associated metagene with the sample gene expression, \gls{bmi} and \gls{bmi} status from the \acrshort{icgc} \acrshort{blca} data, respectively.
+	The results for other \gls{icgc} cancer types are shown in \cref{sec:rest_of_the_fm_icgc_cancer_heatmap_results}.
 	Scales, p-values and $R^2$-value are as described in previous figures.}
@@
 	\includegraphics[page=5,width=0.45\linewidth]{results1/cris_fm_meta}
-	\caption[FM metagene in \gls{nzbc} data]{Heatmap, box plot and scatter plot showing the association of FM obesity associated metagene with sample gene expression, \gls{bmi} and \gls{bmi} status from \gls{nzbc} data, respectively.
+	\caption[FM metagene in the \gls{nzbc} data]{Heatmap, box plot and scatter plot showing the association of the FM obesity associated metagene with the sample gene expression, \gls{bmi} and \gls{bmi} status from the \gls{nzbc} data, respectively.
 	Scales, p-values and $R^2$-value are as described in previous figures.}
@@
 	\includegraphics[width=0.5\linewidth]{results1/deg_venn1}
-	\caption[Veen diagram of the \glspl{deg} identified from CR data (all samples)]{Venn diagram showing the common genes between the signatures obtained from unadjusted and clinical variable-adjusted CR data and the original obesity associated genetic signature.}
+	\caption[Venn diagram of the \glspl{deg} identified from the CR data (all patients included)]{Venn diagram showing the common gene probes between the signatures obtained from the unadjusted and the clinical variable-adjusted CR data and the original obesity associated genetic signature.
+}
 	\label{fig:venn1}
@@
 	\includegraphics[width=0.5\linewidth]{results1/deg_venn2}
-	\caption[Summary of the \glspl{deg} identified from CR data (Caucasian patient data)]{Venn diagram showing the common genes between the signatures obtained from unadjusted and clinical variable-adjusted CR data (Caucasian patients only) and the original obesity associated genetic signature.}
+	\caption[Venn diagram of the \glspl{deg} identified from the CR data (Caucasian patients only)]{Venn diagram showing the common gene probes between the signatures obtained from the unadjusted and the clinical variable-adjusted CR data (Caucasian patients only) and the original obesity associated genetic signature.}
 	\label{fig:venn2}
@@
 	\centering
-	\caption{Summary of the abbreviations used to refer to the different obesity associated genetic signatures}
+	\caption{Summary of the abbreviations used to refer to the different obesity associated genetic signatures identified in the CR data}
 	\label{tab:mg_abbrev}
@@
 	\includegraphics[page=5,width=0.45\linewidth]{results1/cr_deg_meta_vs_clin}
-	\caption[Cr obesity associated metagene in CR data]{Heatmap, box plot and scatter plot showing the association of Cr obesity associated metagene with sample gene expression, \gls{bmi} and \gls{bmi} status from CR data, respectively.
-	The results for other metagenes are in \cref{app:a}.
+	\caption[Cr obesity metagene in the CR data]{Heatmap, box plot and scatter plot showing the association of the Cr obesity associated metagene with the sample gene expression, \gls{bmi} and \gls{bmi} status from the CR data, respectively.
+	The results for the other obesity metagenes are shown in \cref{sec:rest_of_the_cr_ob_meta_heatmap_results_cris}.
 	Scales, p-values and $R^2$-value are as described in previous figures.}
@@
 	\includegraphics[width=0.8\linewidth]{results1/cr_meta_cor}
-	\caption[Pearson correlation of all eight obesity associated metagenes identified in CR data]{Heatmap showing the Pearson correlation of all eight obesity associated metagenes from CR data with one another.
-	\gls{svd} was applied to \gls{rma}-normalised CR data to generate each of the eight metagenes.
-	High and low correlation were represented as red and blue, respectively, where the colours were matched with the values on the scale shown in top right.}
+	\caption[Pearson correlation of all eight obesity metagenes identified in the CR data]{Heatmap showing the Pearson correlation of all eight obesity associated metagenes from the CR data with one another.
+	\gls{svd} was applied to the \gls{rma}-normalised CR data to generate each of the eight metagenes.
+	High and low correlation were represented as red and blue, respectively, where the colours were matched with the values on the scale shown in the top right histogram.}
 	\label{fig:cr_meta_cor}
@@
 	\centering
-	\caption{P-values of all the metagenes in \gls{blca} cancer data}
+	\caption{Statistics of all the obesity metagenes with the sample \gls{bmi}/\gls{bmi} status in the \gls{icgc} \gls{blca} cancer data}
 	\label{tab:degmetablca}
@@
 	\includegraphics[page=5,width=0.45\linewidth]{results1/rawobsgene_ICGC}
-	\caption[Cr obesity associated metagene in \acrshort{icgc} \acrshort{blca} data]{Heatmap, box plot and scatter plot showing the association of Cr obesity associated metagene with sample gene expression, \gls{bmi} and \gls{bmi} status from \acrshort{icgc} \acrshort{blca} data, respectively.
-	The results for other metagenes and other cancer types are in \cref{app:a}.
+	\caption[Cr obesity metagene in the \acrshort{icgc} \acrshort{blca} data]{Heatmap, box plot and scatter plot showing the association of the Cr obesity metagene with the sample gene expression, \gls{bmi} and \gls{bmi} status from the \acrshort{icgc} \acrshort{blca} data, respectively.
+	The results for other other metagenes and other \gls{icgc} cancer types are shown in \cref{sec:rest_of_the_cr_icgc_cancer_heatmap_results}.
 	Scales, p-values and $R^2$-value are as described in previous figures.}
@@
 	\includegraphics[page=5,width=0.45\linewidth]{results1/cris_crdeg_trans_meta}
-	\caption[Cr obesity associated metagene in \gls{nzbc} data]{Heatmap, box plot and scatter plot showing the association of Cr obesity associated metagene with sample gene expression, \gls{bmi} and \gls{bmi} status from \gls{nzbc} data, respectively.
-	The results for other metagenes are in \cref{app:a}.
+	\caption[Cr obesity metagene in the \gls{nzbc} data]{Heatmap, box plot and scatter plot showing the association of the Cr obesity associated metagene with the sample gene expression, \gls{bmi} and \gls{bmi} status from the \gls{nzbc} data, respectively.
+	The results for the other obesity metagenes are shown in \cref{sec:rest_of_the_cr_ob_meta_heatmap_results_cris}.
 	Scales, p-values and $R^2$-value are as described in previous figures.}
@@
 	\centering
-	\caption{Summary of the number of \glspl{deg} identified in each \gls{icgc} cancer data set}
+	\caption{Summary of the number of \glspl{deg} identified in each of the \gls{icgc} cancer data set}
 	\label{tab:icgcdegnum}
@@
 	\begin{threeparttable}
-		\caption{Summary of the number of \glspl{deg} identified by the gene expression analysis and simulation analysis in \gls{icgc} cancer data}
+		\caption{Summary of the number of \glspl{deg} identified by the gene expression analysis and simulation analysis in the \gls{icgc} cancer data}
 		\label{tab:icgcdegtab}

### 2017-03-14 thesis/results/results2.tex
Draft thesis completed; now edit the whole thing - tmp commit

@@
 	\includegraphics[page=100,width=0.45\linewidth]{results2/allmeta_rma_vs_mas}
-	\caption{Comparison of the metagenes generated in the \gls{mas} normalised GT data set from the application of the transformation matrices derived from either the \gls{rma} or \gls{mas} normalised data}
+	\caption[Comparison of the pathway metagenes generated in the \acrshort{mas} normalised GT data set from the application of the transformation matrices derived from either the \acrshort{rma} or \acrshort{mas} normalised GT data]{Comparison of the pathway metagenes generated in the \acrshort{mas} normalised GT data set from the application of the transformation matrices derived from either the \acrshort{rma} or \acrshort{mas} normalised GT data.
+	Only the \gls{bcat} (left) and the \gls{tgfb}  (right) pathway metagenes are shown.
+	The scatter plots for the other pathway metagenes are shown in \cref{sec:normalisation_method_for_pathway_associated_genetic_signature_transformation_matrices}.
+	}
 	\label{fig:gt_rma_vs_mas}
@@
 	\includegraphics[page=15,width=0.7\linewidth]{results2/gatza_meta_trans}
-	\caption{Heatmap of all the pathway metagenes from the \citet{Gatza2010a} study in GT data set}
+	\caption[Heatmap of the Pearson correlation of all the pathway metagenes with one another in the \acrshort{rma}-normalised GT data]{Heatmap showing the Pearson correlation of all the pathway metagenes from the \citet{Gatza2010a} study with one another  in the GT data set.
+		High and low correlation were represented as red and blue, respectively, where the colours were matched with the values on the scale shown in the top left histogram.}
 	\label{fig:gatza_meta_dir}
@@
 	\begin{threeparttable}
-	\caption{Summary of the Spearman correlations of the pathway metagenes generated from \gls{svd} and the application of the transformation matrices\tnote{1} in GT, CR, \gls{nzbc} and FM data sets}
+	\caption[Summary of the Spearman correlations of the \gls{svd}- and TM-derived pathway metagenes in the GT, CR, \gls{nzbc} and FM data sets][Summary of the Spearman correlations of the \gls{svd}- and transformation matrix\tnote{1}-derived pathway metagenes in the GT, CR, \gls{nzbc} and FM data sets]
 	\label{tab:svd_vs_tm_path}
@@
 	\begin{threeparttable}
-		\caption{Summary of the Spearman correlations of the obesity metagenes generated from \gls{svd} and the application of the transformation matrices\tnote{1} in GT, CR, \gls{nzbc} and FM data sets}
+	\caption[Summary of the Spearman correlations of the \gls{svd}- and TM-derived obesity metagenes in the GT, CR, \gls{nzbc} and FM data sets][Summary of the Spearman correlations of the \gls{svd}- and transformation matrix\tnote{1}-derived obesity metagenes in the GT, CR, \gls{nzbc} and FM data sets]
 		\label{tab:svd_vs_tm_obs}
@@
 	\includegraphics[page=16,width=0.45\linewidth]{results2/allmetacor_bar}
-	\caption{Visual comparison of the Spearman correlations of \gls{svd}-derived and transformation matrix-derived \gls{bcat} and \gls{stat3} metagenes across different data sets}
+	\caption[Bar plots showing the Spearman correlation of the \gls{svd}- and TM-derived pathway metagenes across different data sets]{Bar plots showing the Spearman correlations of the \gls{svd}-derived and the transformation matrix-derived \gls{bcat} and \gls{stat3} metagenes across different data sets.
+	Only the \gls{bcat} (left) and the \gls{tgfb}  (right) pathway metagenes are shown.
+	The scatter plots for the other pathway metagenes are shown in \cref{sec:comparison_of_the_correlation_of_svd_and_tm_generated_pathway_metagenes}.
+	}
 	\label{fig:allmetacor_bar}
@@
 	\includegraphics[page=15,width=0.8\linewidth]{results2/all_meta_trans}
-	\caption{Heatmap of all the pathway and obesity metagenes in GT data set}
+	\caption[Heatmap of the Pearson correlation of all the pathway and obesity metagenes with one another in the \acrshort{rma}-normalised GT data]{Heatmap showing the Pearson correlation of all the pathway and obesity metagenes  with one another  in the \gls{rma}-normalised GT data set.
+		High and low correlation were represented as red and blue, respectively, where the colours were matched with the values on the scale shown in the top left histogram.}
 	\label{fig:gatza_allmeta}
@@
 	\centering
-	\caption{Description of the linear models used to predict the Cr obesity metagene in \gls{nzbc} data set}
+	\caption{Description of the linear models constructed from the \gls{nzbc} data to predict the Cr obesity metagene}
 	\label{tab:lm_sig_var}
@@
 	\centering
-	\caption{Description of the linear models used to predict the Cr obesity metagene in \gls{nzbc} data set, using only the sample \gls{bmi}, \gls{bmi} status and \gls{pr} pathway metagene scores}
+	\caption[Description of the linear models constructed from the \gls{nzbc} data to predict the Cr obesity, using only the sample \gls{bmi}, \gls{bmi} status and the \acrshort{pr} pathway metagene score]{Description of the linear models constructed from the \gls{nzbc} data to predict the Cr obesity, using only the sample \gls{bmi}, \gls{bmi} status and the \gls{pr} pathway metagene score}
 	\label{tab:lm_pr_only}
@@
 	\includegraphics[page=73,width=0.32\linewidth]{results2/prediction_cris_with_cris(cris_model)}
-	\caption{Comparison of the predicted Cr metagenes with the true Cr metagene in the \gls{nzbc} data}
+	\caption[Comparison of the predicted Cr metagene scores with the Cr metagene score from the \gls{nzbc} data]{ Box plot and scatter plots comparing the Cr metagene score predicted by the different linear models constructed from the \gls{nzbc} data, with the Cr metagene score from the \gls{nzbc} data.
+	Only the results from the models used to predict the Cr metagene scores are shown.
+	The summary of the statistics for the other obesity metagenes in the \gls{nzbc} data are shown in \cref{sec:summary_statistics_of_the_predicted_obesity_metagenes_with_sample_bmi_bmi_status_in_nzbc_and_cr_data}.
+	P-values and $R^2$-value are as described in previous figures.
+	}
 	\label{fig:predict_cris_cris}
@@
 	\includegraphics[page=73,width=0.32\linewidth]{results2/prediction_cr_with_cris}
-	\caption{Comparison of the predicted Cr metagenes with the true Cr metagene in the CR data}
+	\caption[Comparison of the predicted Cr metagene scores with the Cr metagene score from the CR data]{ Box plot and scatter plots comparing the Cr metagene score predicted by the different linear models constructed from the \gls{nzbc} data, with the Cr metagene score from the CR data.
+	Only the results from the models used to predict the Cr metagene scores are shown.
+	The summary of the statistics for the other obesity metagenes in the CR data are shown in \cref{sec:summary_statistics_of_the_predicted_obesity_metagenes_with_sample_bmi_bmi_status_in_nzbc_and_cr_data}.
+	P-values and $R^2$-value are as described in previous figures.
+	}
 	\label{fig:predict_cr_cris}
@@
 	\centering
-	\caption{Description of the step wise linear models used to predict all of the obesity metagenes in \gls{nzbc} data set}
+	\caption[Description of the step wise linear models constructed from the \gls{nzbc} data to predict all of the obesity metagenes]{Description of the step wise linear models constructed from the \gls{nzbc} data to predict all of the obesity metagenes}
 	\label{tab:stepwise_model}\\
@@
 	\includegraphics[page=10,width=0.32\linewidth]{results2/stepwise_cris_cris}
-	\caption{Comparison of the predicted obesity metagenes with the true obesity metagene in \gls{nzbc} data}
+	\caption[Comparison of all the obesity metagene scores predicted from the step wise linear models with the obesity metagene scores from the \gls{nzbc} data]{Scatter plots comparing all of the obesity metagene scores predicted by the step wise linear models constructed from the \gls{nzbc} data, with the corresponding obesity metagene scores from the \gls{nzbc} data.
+	P-values and $R^2$-value are as described in previous figures.}
 	\label{fig:stepwise_cris}
@@
 	\includegraphics[page=10,width=0.32\linewidth]{results2/stepwise_cr_cris}
-	\caption{Comparison of the predicted obesity metagenes with the true obesity metagene in CR data}
+	\caption[Comparison of all the obesity metagene scores predicted from the step wise linear models with the obesity metagene scores from the CR data]{Scatter plots comparing all of the obesity metagene scores predicted by the step wise linear models constructed from the \gls{nzbc} data, with the corresponding obesity metagene scores from the CR data.
+	P-values and $R^2$-value are as described in previous figures.}
 	\label{fig:stepwise_cr}

### 2017-03-14 thesis/intro/intro.tex
Draft thesis completed; now edit the whole thing - tmp commit
Edited Intro

@@
-\chapter{Introduction (draft)}
+\chapter{Introduction}
 \label{ch:intro}
@@
 In fact, the risk of comorbidities increases with the increase in \gls{bmi}, where the risk becomes severe as the \gls{bmi} level approaches the obese category \citep{WHO2000}.
-
-The number of overweight and obese people, in both adults and children, have risen in every country of the world, and the trend continues as we speak.
+The number of overweight and obese people, in both adults and children, have risen in every country of the world and the trend will likely to continue in the future.
 Estimated to account for 3.4 million deaths per year and 93.6 million \glspl{daly} in 2010, obesity is a serious disease that continues to grow in our society \citep{Lim2012}.
@@
 One common and widely used approach to categorise individuals is by measuring the \gls{bmi} of the individual.
-\gls{bmi} is a measurement based on the weight-to-height ratio of an individual that is used by clinicians and epidemiologists to classify adults into underweight, normal weight, overweight or obese categories.
-The unit of \gls{bmi} is defined as the individual's weight in kilograms per square of the height in meters (kg/m$^2$).
+\gls{bmi} is a measurement based on the weight-to-height ratio of an individual that is often used by clinicians and epidemiologists to classify adults into underweight, normal weight, overweight or obese categories.
+The unit of \gls{bmi} is defined as the weight in kilograms per square of the height in meters (kg/m$^2$).
 
-\gls{who} (2014) have used \gls{bmi} measurements to define overweight and obesity in adults as an individual with \gls{bmi} $\geq$ 25 kg/m$^2$ and $\geq$ 30 kg/m$^2$, respectively.
+\gls{who} (2014) have used \gls{bmi} measurements to define overweight and obesity in adults as individuals with \gls{bmi} $\geq$ 25 kg/m$^2$ and $\geq$ 30 kg/m$^2$, respectively.
 For children under the age of 5, overweight and obesity were defined as weight-for-height greater than 2 standard deviation and 3 standard deviation above \gls{who} Child Growth Standards median, respectively.
@@
 
-Although the use of \gls{bmi} as an accurate measure of body fat composition and/or representation of the individual's metabolic status remains controversial, many studies have used \gls{bmi} due to the ease of data collection compared to other measurements, such as waist-to-height ratio \citep{Lee2008, Gelber2008}.
-Consequently, most of the publicly available clinical data have height and weight data from which one can calculate the patients' \gls{bmi}, and therefore able to collect the patients' obesity status more conveniently than, for example, waist-to-height ratio.
+However, the use of \gls{bmi} as an accurate measure of body fat composition and/or representation of the individual's metabolic status remains controversial.
+In fact, many studies have suggested the use of other measurements such as waist-to-height and waist-to-hip ratios to classify the patients into different categories, as these measurements were better predictors of the obesity related diseases than \gls{bmi} \citep{Dalton2003,Gelber2008,Lee2008}.
+With that said, \gls{bmi} is still widely used due to the ease of data collection compared to other measurements and, as a result, most of the publicly available clinical data have  height and weight data of the patients from which one can calculate the patients' \gls{bmi}.
 
@@
 This equated to approximately 1 in every 14 people in the world classed as obese in 2014, and therefore, 1 in 14 people around the world were also at severe health risks associated with obesity.
+The highest prevalence of overweight and obesity was in the American regions, where 61\% and 27\% of the population were overweight and obese, respectively \citep{WHO2014}.
 It was also noted in the report that women were more likely to be obese than men, and that the prevalence of overweight and obesity increased as the income level of the countries increased \citep{WHO2014}.
@@
 
-% TODO: Add western-specific and NZ specific statistics on obesity
+In \gls{nz}, obesity is a serious problem as approximately one third (31.6\%) of the adults in \gls{nz} were classified as obese in 2015/2016 \citep{Health2016}.
+The Pacific and M\=aori populations were affected the most, where 67\% and 47\% of the adults, respectively, were obese \citep{Health2016}.
+\gls{nz} not only have a high prevalence of obesity compared to many of the other countries of the world, but also have a unique health problem that affects the Pacific and M\=aori populations more than the rest of the \gls{nz} population.
+
+% TODO: Add western-specific statistics on obesity (wang2007)
 % TODO: Add male/female specific stats
-% TODO: Add Maori/PI specific stats
-
-% Until these terrible trends stop, and even after the trends stop, the world will need a plan to live along with this ongoing global epidemic of obesity and its associated diseases.
 
@@
 
-Obesity is caused by continuous intake of food without expending all of the energy gained.
-The key to the problem is the control of food intake and energy expenditure, and how they are regulated in your body.
-The role of hypothalamus is central to the control of satiety and many mutations that cause obesity affects the satiety signals to and/or from hypothalamus.
+Obesity is caused by continuous intake of food without expending all of the energy gained from the food.
+The key to the problem is the control of food intake and energy expenditure, and how they are regulated in the body.
+The role of hypothalamus is central to the control of satiety and many of the mutations that cause obesity affects the satiety signals to and/or from hypothalamus.
 The hypothalamus receives both nueronal and hormonal inputs from the body and propagates the signal to the downstream neurons that ultimately affect the food intake of the individual \citep{Bell2005, Spiegelman2001}.
@@
 Insulin, perhaps the most well-known hormone for its role in glucose homeostasis, also has an important role in satiety and food intake regulation.
-Briefly, when insulin binds to the insulin receptor, \gls{irs} protein family is recruited and phosphorylated by the insulin receptor's intrinsic tyrosine kinase activity \citep{Saltiel2002}.
+Briefly, when insulin binds to the insulin receptor, \gls{irs} protein family is recruited and phosphorylated by the  intrinsic tyrosine kinase activity of the insulin receptor \citep{Saltiel2002}.
 The phosphorylated \gls{irs} protein then activates \gls{pi3k}, which in turn activates its downstream targets, causing a rapid signal cascade through the cell \citep{Saltiel2002}.
 
-The evidence of insulin's involvement in the satiety control was not particularly clear even though there were strong evidence of obesity with neuron-specific deletion of insulin receptor, and that the insulin level, like leptin, was proportional to the fat mass also suggested the role of insulin in satiety and body weight control \citep{Barsh2002, Bruning2000, Woods1979}.
+The evidence of the involvement of insulin in the satiety control was not particularly clear even though there were strong evidence of obesity with neuron-specific deletion of insulin receptor, and the fact that the insulin level, like leptin, was proportional to the fat mass also suggested the role of insulin in satiety and body weight control \citep{Barsh2002, Bruning2000, Woods1979}.
 This was perhaps due to the fact that leptin and insulin used two different signalling pathways in the cell; leptin used the \gls{jak}/\gls{stat} signalling pathway, whereas insulin used the \gls{pi3k} pathway for signalling \citep{Ghilardi1996}.
-This was puzzling to the researchers as there seemed to be no converging point of the two hromones in these pathways.
-
 However, leptin, and subsequently insulin, was shown to activate ATP-dep\-endent K$^+$ channel in glucose-responsive neurons in the hypothalamus \citep{Spanswick1997, Spanswick2000}.
@@
 
-There are many environmental factors associated with the global rise in obesity.
-Even though these environmental factors are probably the most significant contributor to the current global obesity epidemic, there are some individual-level, genetic factors that contribute to the overall increase in the proportion of the population that are obese.
+Even though environmental factors are probably the most significant contributor to the current global obesity epidemic, there are some individual-level, genetic factors that contribute to the overall increase in the proportion of the population that are obese.
+There is much evidence of obesity being caused by individual-level genetic factors that make some individuals more predisposed to obesity than others.
+Generally speaking, there are two classes of genetically-driven obesity: monogenic obesity and polygenic obesity \citep{Moustafa2013}.
 
-There is much evidence of obesity being caused by individual-level genetic factors that make some individuals more predisposed to obesity than others.
-Generally speaking, there are two classes of genetically-driven obesity: monogenic obesity and polygenic obesity \citep{Moustafa2013}. \\
-
-\noindent
 Monogenic obesity is where a single mutation in one of the genes that is directly involved in the satiety controlling pathway causes severe obesity in that individual.
@@
 However, although these mutations cause a devastating effect on the individual carrying the mutation, these mutations occur very rarely in the population.
-% With that said though, the identification of these mutations have shed light on the mechanisms in which obesity is developed and it is worth mentioning some of these mutations to highlight the importance of the components involved in the obesity pathway.
 
@@
 As a result, children with this mutation was constantly hungry due to the lack of anorexogenic signals, and became obese at a very early stage of their lives.
-\\
+Though these mutations are very rare and affects only those individuals that have one of these mutations, it had a profound impact in the field of obesity research and contributed significantly to the knowledge of the mechanism of the development of obesity.
 
-% In close relation to this mutation is the mutation in the leptin receptor gene.
-% \citet{Clement1998} have discovered a single nucleotide substitution mutation that led to the leptin receptor transcript that skipped exon 16, leading to a form of the receptor with no transmembrane or intracellular domain.
-% Individuals with this rare mutation showed extreme hyperphagia and symptoms of early-onset obesity, even though their circulating leptin levels were present abundantly \citep{Clement1998}.
-
-% The next couple of mutations are associated with the synthesis and the processing of \gls{pomc}.
-% \citet{Krude1998} found single nucleotide mutations in the exonic region of the \gls{pomc} gene that affected the transcription of the $\alpha$-\gls{msh} region (and some other regions) of \gls{pomc}.
-% \citet{Jackson1997} identified a mutation in the \gls{pc1}, an enzyme responsible for the processing of many prohormones and neuropeptides, including \gls{pomc}.
-% In either case, the mutation resulted in the lack of $\alpha$-\gls{msh} in the neurons for satiety signalling and caused severe early-onset obesity.
-
-% As mentioned earlier, even though these mutations are very rare and affects only those individuals that have one of these mutations, it had a profound impact in the field of obesity research and contributed significantly to the knowledge of the mechanism of how obesity develops.
-
-\noindent
-Polygenic obesity is where a combination of variants in multiple genes or multiple regions of the genome, together with environmental factors, causes obesity in the individual \citep{Moustafa2013}.
-Unlike in monogenic obesity where a single mutation in a gene causes severe obesity, the variants in some of the genes or genomic regions involved in polygenic obesity have little effect on the individual's obesity status by itself, but in combination with many other variants and together with environmental factors, it has a serious obesogenic effect on the individual.
-
-% Known as the ``thrifty gene'' hypothesis, \citet{Neel1962} suggested that the genes that were important in obesity have been around in the population for many generations, as it would have had a selective advantage in an environment where the population experienced starvation more frequently than the current generation.
-% However, as food became more accessible, a subset of the population who were particularly good at storing fat and survive the starvation period ``overreacted'' to this rapid change in the environment \citep{Bell2005, Spiegelman2001}.
-% These observations in the rise in obesity in subsets of population cannot be explained simply by external factors alone, and there is a significant genetic component involved in this phenomenon \citep{Bell2005}.
+Polygenic obesity is where a combination of variants in multiple genes or multiple regions of the genome, together with the environmental factors, causes obesity in the individual \citep{Moustafa2013}.
+Unlike in monogenic obesity where a single mutation in a gene causes severe obesity, the variants in some of the genes or genomic regions involved in polygenic obesity have little effect on the obesity status of the individual by itself, but in combination with many other variants and together with environmental factors, it has a serious obesogenic effect on the individual.
 
 There have been many studies that have investigated the association of genes and/or genomic regions with obesity.
-Of those, many of the \glspl{gwas} published in 2007 that showed strong evidence of the \gls{fto} gene with \gls{bmi} and obesity were of greatest significance \citep{Dina2007,Frayling2007,Gerken2007,Scuteri2007}.
+Of those, many of the \glspl{gwas} published in 2007 showed strong evidence of the \gls{fto} gene with \gls{bmi} and obesity  \citep{Dina2007,Frayling2007,Gerken2007,Scuteri2007}.
 Until these studies came out, there has been no common genes or genomic regions that associated with obesity that was reproducible and reliable \citep{Frayling2007}.
 Thus, these studies were the first evidence of common variants that contributed to the polygenic obesity phenotype in the population.
-
-% Discovery of the \gls{fto} gene has lead to extensive research on the gene, and the mechanism of action by the \gls{fto} variant has been recently uncovered  from these efforts \citep{Claussnitzer2015a,Smemo2014a}.
-% The association between obesity and genetic variations, and their mechanisms that lead to obesity, has not been fully elucidated due to its complexity, but it may not be long before they are fully uncovered.
 
@@
 
-Even though some individual-level factors are involved with obesity, it is clear that these individual-level factors cannot be accounted for the global epidemic of obesity that the world currently observes.
-For this reason, obesity is thought to be caused mainly by the environmental factors that act at a population-level.\\
+It is clear that the individual-level factors cannot be accounted for the global epidemic of obesity that the world currently observes.
+For this reason, obesity is thought to be caused mainly by the environmental factors that act on a population-level.
+The first key factor that contributes to the global rise in the proportion of the population that is obese is the availability and abundance of certain food in the country.
+With the help from the government trade policies, food and other goods have become easier to trade between countries, helping the economic growth of many countries around the world \citep{Kearney2010}.
+This led to greater abundance of food in general, greater number of choices of food and shifted the overall nutritional status of the country \citep{Malik2013}.
+This shift in the food choice and availability (and the economic benefit associated with the trades) had a positive impact on the country, but the growth in the food market also resulted in the population-wide rise in obesity.
 
-\noindent
-The first key factor that contirbutes to the global rise in the proportion of the population that is obese is the availability and abundance of certain food in the country.
-With the help from government trade policies, food and other goods have become easier to trade between the country, helping the economic growth of many countries around the world \citep{Kearney2010}.
-This led to greater abundance of food in general, greater number of choices of food and shifted the overall nutritional status of the country \citep{Malik2013}.
-This shift in the food choice and availability (and the economic benefit associated with the trades) had a positive impact on the country, but the growth in the food market also resulted in population-wide rise in obesity.
-% This shift in the food choice and availability (and the economic benefit associated with the trades) had a positive impact on the country, but the growth in the food market also resulted in a different, much larger problem -- obesity.
-
-As an example, in the \gls{usa}, the cost of corn and soy is low due to the fact that they are the raw ingredients for most processed food and beverages, such as high fructose corn syrups used to sweeten many soft drinks in the \gls{usa} \citep{Malik2013}.
+As an example, in the \gls{usa}, the costs of corn and soy are low due to the fact that they are the raw ingredients for most processed food and beverages, such as high fructose corn syrups used to sweeten many soft drinks in the \gls{usa} \citep{Malik2013}.
 Moreover, corn and soy are also the main food for the livestock, which results in lower prices for meat \citep{Malik2013}.
@@
 As a result, more people are inclined to consume food that are cheap, have little nutritional value and high in energy, than the food that are expensive, nutritional and healthy alternatives.
-Thus, the population is more likely to become obese by making poor food choice.\\
+Thus, the population is more likely to become obese by making poor food choice.
 
-\noindent
 The second factor to consider is the behavioural changes led by the urbanisation of the population.
-% Urbanisation is defined as the gradual increase in the proportion of people living in urban areas.
 As more people take up urban lifestyle, they are exposed to the new environment where there is a better range of food selection, as well as technological and mechanical advancement in transport and other daily chores that may not be available in rural areas.
@@
 Lack of physical activity is one of the obvious components that is linked with the urbanisation of the population.
-% Currently, the recommended amount of physical activity to keep an individual fit and healthy is $\geq$ 150 min of moderate-intensity aerobic physical activity per week, or $\geq$ 30 min of physical activity on most days of the week \citep{Pate1995, WHO2010}.
 In many urban cities, the recommended level of physical exercise is difficult to achieve and maintain due to the shift towards the sedentary behaviours that are encouraged by the developed transport system, automated household chores, and indoor entertainment \citep{Malik2013}.
-\\
-% Urbanisation by itself cannot be accounted for the lack of exercise in the population, but it is true that urbanisation makes the bar higher for many people to take up the recommended level of fitness and maintain it.
-\\
 
-\noindent
 The last factor to be mentioned is the income and socioeconomic status of the population.
 As the average income increases, more people are able to buy food and are likely to take up the sedentary lifestyle.
-
 From this, one would expect the high-income group to have the largest impact from obesity, but on the contrary to this expectation, the biggest impact from obesity is seen in the low- and middle-income groups of the population \citep{Malik2013}.
@@
 In contrast, the high-income group of the population have better access to the facilities that promote weight maintenance, thus have less of an impact compared with the low- and middle-income groups of the population.
-\\
-
-% \noindent
-% The last factor, which ties in strongly with the first key factor, is the choice of diet made by the population.
-% The amount of low-quality food have increased significantly in many rapidly growing countries, as these high-energy, low-nutritional food are cheaper and more profitible than the natural produce that have higher nutritional value \citep{Kearney2010}.
-% As a result, more people tend to purchase these low-quality food more often, and consume more energy than they are required.
-
-% However, many people eat responsibly and avoid these low-quality diet and prevent overconsumption of such food, but the overconsumption of these food is not a big problem as such, rather the overall quality of the food \citep{Malik2013}.
-% These food are usually high in refined sugar and unhealthy fat, highly processed, contain very little nutritional value and are low in fibre.
-% As a result, consumption of these food increases the overall energy intake and achieve no nutritional benefits from the food.
-
-% Whether it is by choice or not, consumption of these low-quality diet leads to obesity and obesity-related diseases. \\
-
-\noindent
 All of these population-level environmental factors in synergy with the individual-level genetic factors have contributed to the global epidemic of obesity.
-
-% \subsection{Impact of obesity}
-% \label{sub:impact_of_obesity}
-
-% As mentioned at the beginning of the current section, obesity has a devastating effect on the health of the population in a global scale.
-% Overweight and obesity have been a risk factor for many diseases, including type 2 diabetes, cardiovascular diseases, hypertension, and many types of cancers \citep{Franks2010, Guh2009, Yach2006}.
-% As \citet{Yach2006} mentions in their paper, ``overweight and obesity have become to diabetes what tobacco is to lung cancer'', and it is important to understand that obesity has significant risks and comorbidities associated with it.
-
-% Although adulthood obesity is clinically important for risk assessment of diseases, childhood obesity has been focussed more heavily in recent years,  as it has greater public health implications.
-% There are many immediate consequences that are related to childhood obesity, but the diseases that develop as these children grow up are more critical than the earlier effects.
-% A paper by \citet{Must1999} reviewed many of the clinical studies done in the last several decades and summarised the risks associated with childhood obesity.
-% Among the many risks associated with obesity, obese children, compared to lean children, had over 8.5-fold increased risk of developing hypertension as adults \citep{Must1999}.
-% On top of this, the relative risks for all-cause mortality and coronary heart disease-specific mortality were 1.5 and 2.5, respectively \citep{Must1999}.
-
-% In a more recent study by \citep{Franks2010}, they have shown that childhood obesity (as well as glucose intolerance and hypertension in childhood) was strongly associated with premature deaths from endogenous causes in American Indian population from Arizona (incidence rate ratio of 1.4 per unit of \gls{bmi} z-score).
-
-% Health related consequences of obesity are clear from many epidemiological studies, but it is noteworthy that obesity also has an economical impact on the population as well.
-% In the \gls{usa}, within the span of 5 years, the medical cost due to diabetes have risen from \$44 billion to \$92 billion \citep{Yach2006}.
-% Although not all diabetes are directly caused by obesity, obesity would have assisted in the rise in the costs associated with diabetes.
-% In addition to this,
-% The costs that are associated with the treatments of comorbidities of obesity affects the household incomes, increased premature retirement and unemployment, which could have a significant economical impact on the society over time \citep{Yach2006}.
-
-% As more of the younger generation is gaining weight and becoming obese, having the risks and disease burden associated with obesity at such early age are not ideal (both economically and epidemiologically) and poses a threat to the growing population that are becoming more obese.
-
-% \subsection{Preventive measures of obesity}
-% \label{sub:preventive_measures_of_obesity}
-
-% It is apparent from previous sections that obesity has become a serious problem that the world needs to prioritise to solve immediately.
-% Even though obesity has been regarded as one of the many global problems to solve, the progress has been slow.
-
-% \citet{Sacks2009} proposed an \gls{opa} framework that was based on the framework developed by \citet{WHO2006}.
-% In this framework, \citet{Sacks2009} have added extra features so that the policies were targeted to impact three levels of the underlying environment of obesity: the socio-ecological environment (upstream), lifestyle/behavioural changes (midstream), and health services support (downstream).
-
-% The upstream approach targets the underlying economic, social and physical norms that the population is currently in.
-% By targeting the broad socioeconomic environment that the population lives in, such as taxation, education, employment and housing \citep{Sacks2009}.
-% In doing so, it is possible to have a significant change in the population behaviours and eventually improve the health outcomes of the population.
-
-% The midstream approach targets the population or subpopulation behaviour change to improve the eating and physical activity behaviours in these groups.
-% Unlike in the upstream approach, the midstream approach aims to change the behaviour by implementing social marketing and programmes that encourage such changes \citet{Sacks2009}.
-% Again, the result of this approach will be an improvement in the overall positive behaviour of the population towards healthy eating and physical activity.
-
-% The downstream approach targets to support the health services and clinical interventions.
-% Unlike upstream and midstream approaches, downstream approach generally acts on an individual level, whereby the focus is on the management of the health of an individual and/or the individual's family \citet{Sacks2009}.
-% Although not as efficient as upstream or midstream approaches, downstream approach will help improve the habits and behaviours to promote weight management.\\
-
-% \noindent
-% In the recent years, obesity has been declared as a global priority \citep{Malik2013}.
-% Furthermore, comprehensive global monitoring framework, multi-sectoral actions and strengthening of the national policies have been implemented to better control the obesity epidemic \citep{Malik2013}.
-% In many countries, the nutritional labelling on products have been adjusted to include the ingredients that have negative impact on obesity (for example, \textit{trans} fat), and even reduce or remove these ingredients from food \citep{Malik2013}.
-
-% Food marketing can also be a problem in a highly competitive society that we live in.
-% To sell their products that are high in fat, sugar and portion size, food industries market and advertise their products to the population, especially young children, in order to stay ahead in their market place \citep{Chan2010, Malik2013}.
-% Even though there is a strong evidence of food marketing influencing the eating behaviours and food preferences in children, very few countries have been able to reduce these marketing and advertising strategies of unhealthy food \citep{Hawkes2007a, Malik2013}.
-
-% In contrast to these upstream/midstream approaches, there are some clinical interventions and treatments that help manage weight in an individual.
-% One of these therapies is bariatric surgery for obese patients, especially morbidly obese patients \citep{Buchwald2004}.
-% Through their meta-analysis, \citet{Buchwald2004} showed that bariatric surgery was able to reduce many of the the comorbidities associated with obesity.
-% The fact that these improvements and complete resolution of the comorbidities in high proportion of patients that underwent bariatric surgery shows that, even though it acts on an individual level, downstream approaches such as bariatric surgery may be a highly effective, but not an efficient, method of preventing obesity in the population.\\
-
-% \noindent
-% It is evident that the most efficient way to reduce the proportion of the population that are obese is to implement national policies that affect the upstream or midstream factors.
-% These policies, although they are efficient at targeting the underlying problem, may take a long time to make a significant effect.
-% On the otherhand, individual therapeutic plans that affect the downstream factors may be very effective in achieving the desired goal of weight loss, but it is very resource intensive and upscaling this approach would be practically impossible.
-
-% The management of obesity at an early stage with midstream and/or downstream approaches while the upstream approach takes effect would be the key to success in controlling for the global obesity epidemic.
 
@@
 
-% What is cancer.
 \begin{quote}
@@
 
-% TODO: Add western-specific and NZ specific statistics on cancer
-% TODO: Add male/female specific stats
-% TODO: Add Maori/PI specific stats
-
 Cancer is considered as one of the leading causes of deaths around the world, accounting for 8.2 million deaths in 2012 \citep{WHO2014}.
@@
 Although the levels of mortality vary among different countries, cervical, liver and stomach cancers make up a large proportion of cancer deaths among low- and middle-income countries \citep{WHO2014}.
-\\
 
-\noindent
 \citet{Bray2013} estimated the global prevalence of cancer in adult population in 2008.
 Since there is no clear definition for cancer prevalence, \citet{Bray2013} used \gls{ldp} as an estimate for cancer prevalence, and measured the incidence of cancer during the period between 2004 and 2008 (5 years).
-% measuring the incidence of cancer during the period between 2004 and 2008 (5 years) who were still alive at the end of 2008.
 From these estimates, Eastern Asia had the greatest 5-year prevalence of cancer, with approximately 7 million people with cancer in 2008, followed by North America and Western Europe with approximately 4.7 and 3 million people with cancer \citep{Bray2013}.
@@
 Furthermore, the prevalence of breast, cervix or prostate cancer is the highest in almost 95\% of the countries \citep{Bray2013}.
-Although these estimates provide a relatively crude measure of cancer prevalence, \citet{Bray2013} provided a useful overview of what the prevalence of cancer is like around the world.\\
+Although these estimates provide a relatively crude measure of cancer prevalence, \citet{Bray2013} provided a useful overview of the prevalence of cancer  around the world.
 
-\noindent
+In \gls{nz}, the most common cancer types diagnosed for males are prostate, colorectal, melanoma, lung and non-Hodgkin lymphoma, which all together accounted for 65\% of newly registered cancers in males in 2013 \citep{Health2016b}.
+For females, breast (which accounted for 28.3\% of all new female cancer cases), colorectal, melanoma, lung and uterine cancers were the most common cancer types diagnosed in 2013 \citep{Health2016b}.
+The cancer registration rates were higher in females aged 25-54 years than males of the same age group, but the registration rates were significantly higher in males for people who were 60 years or older \citep{Health2016b}.
+Like with obesity, there is a significant difference in the cancer registration rates between the M\=aori and the non-M\=aori populations.
+The rates were 1.2 and 1.4 times greater in the M\=aori population compared to the non-M\=aori population for males and females, respectively \citep{Health2016b}.
+Again, this difference between the M\=aori and non-M\=aori population is unique to \gls{nz}, and improved understanding of the underlying problem is essential for better treatment plans and outcomes.
+
 The total number of deaths by non-communicable diseases (including cancer) is predicted to increase, reaching 52 million deaths by 2030 \citep{WHO2014}.
@@
 It is now commonly understood that cancers are caused by dysregulation of various molecular mechanisms or pathways that allow tumour cells to prolilferate, survive and migrate.
-By taking advantage of these pathways, cancer cells are able to grow, and eventually metastasise to other parts of the body.
-The environment in which the tumours arise can also be an important factor for their evolution and growth, contributing to the heterogeneity of the cancer cells within the tumour or between tumours at different sites of the body.
+By taking advantage of these pathways, cancer cells are able to grow and eventually metastasise to other parts of the body.
+The environment in which the tumours arise can also be an important factor for their evolution and growth.
 
@@
 
-Extensive research in cancer biology have been able to uncover many pathways and regulatory mechanisms involved in cancer.
-% , so much that there were too many possibilities in which cancer could develop from.
+Extensive research in cancer biology has uncovered many pathways and regulatory mechanisms involved in cancer.
 With more than 100 distinct types and subtypes of cancers, each with its own distinct development mechanisms, it became difficult to focus on the overarching mechanisms of cancer development \citep{Hanahan2000}.
@@
 Finally, cancer cells are able to maintain proliferative signalling by mutating the growth factor receptor itself, or the downstream signalling targets of the growth factor receptor pathway \citep{Fuqua1991,SuHuang1997,Satyamoorthy2003}.
-By maintaining the the proliferative signals, cancer cells are able to continuously grow in their environment.
+By maintaining the proliferative signals, cancer cells are able to continuously grow in their environment.
 
@@
 As well as maintaining proliferative signals, cancer cells must also evade signals that inhibit cell growth.
-p53 protein is a tumour suppressor protein that has a  role in cell cycle arrest \citep{Hanahan2011,Levine1997}.
+The p53 protein is a tumour suppressor protein that has a  role in cell cycle arrest \citep{Hanahan2011,Levine1997}.
 As \citet{Hanahan2011} mention in their review, ``TP53 receives inputs from stress and abnormality sensors that function within the cell's intracellular operating systems''.
@@
 When abundant and active, p53 protein transcribes the p21 protein which, through series of events in the p16-cyclin D$_1$-cdk4-\gls{rb} pathway, triggers cell cycle arrest \citep{Levine1997}.
-In many cancers, p53 protein is usually mutated and selected against to allow continuous growth of the tumour cells.
-
-% Two tumour suppressor proteins have an essential role in this evasion: the \gls{rb} and p53 proteins.
-
-% \Gls{rb} protein is a known tumour suppressor protein that is important in cell cycle regulation \citep{Burkhart2008,Hanahan2011}.
-% \Gls{rb} protein is known to arrest the cells in \gls{g1} phase of the cell cycle by regulating E2F transcription factors in the cell, in reponse to molecules such as \gls{tgfb} \citep{Burkhart2008,Hanahan2000}.
-% Although \gls{rb} protein is known for its role as a checkpoint in \gls{g1} phase, it also takes part in the regulation of apoptotic cell death, cell differentiation and chromosome stability \citep{Burkhart2008}.
-% The contribution of the loss of function of the \gls{rb} protein in tumorigenesis and tumour progression is evident, but at which stage of the development this loss of function mutation is crucial for the cancer cells is yet to be clarified \citep{Burkhart2008}.
-
-% p53 protein is another tumour suppressor protein that is closely related to the \gls{rb} protein and its role in cell cycle arrest \citep{Hanahan2011,Levine1997}.
-% As \citet{Hanahan2011} mentions in their review, ``\gls{rb} transduces growth-inhibitory signals that originate largely outside of the cell, TP53 receives inputs from stress and abnormality sensors that function within the cell's intracellular operating systems''.
-% The level of expression and its stability is crucial for the function of p53 protein, as the abundance of p53 protein in the cell triggers cell cycle arrest or cell \gls{apoptosis}, depending on the extent of the cell signal \citep{Fridman2003,Hanahan2011,Levine1997}.
-% When abundant and active, p53 protein transcribes the p21 protein which, through series of events in the p16-cyclin D$_1$-cdk4-\gls{rb} pathway, triggers cell cycle arrest \citep{Levine1997}.
+In many cancers, p53 protein is mutated and selected against to allow continuous growth of the tumour cells.
 
@@
 
-% \Gls{bcl2} family proteins are the major players of apoptosis.
-% Embedded in the mitochondrial membrane, the Bax subfamily of \gls{bcl2} proteins are pro-apoptotic, and are under constant repression by the anti-apoptotic proteins of the \gls{bcl2} subfamily proteins \citep{Adams2007,Hanahan2011}.
-% The inhibition of the pro-apoptotic proteins are achieved by the interaction between the pro- and anti-apoptotic proteins via the \gls{bh3} domain \citep{Adams2007}.
-% In addition to the pro-apoptotic Bax subfamily proteins, there is a family of proteins that contain a single \gls{bh3} domain, and hence named \gls{bh3}-only family proteins \citep{Adams2007}.
-% \gls{bh3}-only proteins either disrupt the inhibition of the Bax family proteins by the \gls{bcl2} subfamily proteins, or directly stimulate the Bax subfamily proteins \citep{Adams2007}.
-% The activation of the pro-apoptotic proteins disrupt the mitochondrial cell membrane adn releases cytochrome \textit{c}, which in turn triggers the activation of the apoptotic proteases caspase 8 and caspase 9, resulting in apoptosis \citep{Adams2007,Hanahan2011}.
-
 The precise cellular conditions that trigger apoptosis are yet to be uncovered, but there is a clear evidence of p53 tumour suppressor playing a key role in apoptosis initiation.
-When \acrshort{dna} damage is sensed, p53 protein transcribes the pro-apoptotic members of the \gls{bcl2} family proteins, such as Bax, Puma and Noxa \citep{Fridman2003,Hanahan2011}.
-The fact that p53 protein is involved not only in cell cycle regulation, but also in apoptotic pathway shows its importance in the control of cancer development, and perhaps that is the reason why more than 50\% of tumours have a mutation in the \textit{TP53} gene \citep{Levine1997}.
+In short, when \acrshort{dna} damage is sensed, p53 protein transcribes the pro-apoptotic members of the \gls{bcl2} family proteins, such as Bax, Puma and Noxa \citep{Fridman2003,Hanahan2011}.
+The fact that p53 protein is involved not only in the regulation of cell cycle, but also in the apoptotic pathway shows its importance in the control of cancer development, and perhaps that is the reason why more than 50\% of tumours have a mutation in the \textit{TP53} gene \citep{Levine1997}.
 
@@
 When acquired by cancer cells, the previous three hallmarks allow the cells to continuously grow in an uncontrolled manner.
-However, this is not the case as the disruption in cell signalling on its own does not trigger the rapid expansion of the tumour -- only when the cells have the capability of replicating infinitely \citep{Hanahan2000, Hanahan2011}.
+However, this is not the case as the disruption in cell signalling on its own does not trigger the rapid expansion of the tumour; this occurs only when the cells have the capability of replicating infinitely \citep{Hanahan2000, Hanahan2011}.
 
@@
 When cells are grown in a culture, cells are able to replicate and grow until a certain point, and the cells reach the stage of senescence where the cells stop growing but are still viable \citep{Hanahan2011}.
-The cells that manage to bypass this senescence phase by mutating the tumour suppressor proteins such as p53 protein, undergoes the crisis phase where the majority of the cells in the culture dies \citep{Hanahan2011}.
+The cells that manage to bypass this senescence phase by mutating the tumour suppressor proteins such as the p53 protein, undergoes the crisis phase where the majority of the cells in the culture dies \citep{Hanahan2011}.
 Very rarely, some cells are able to survive the crisis phase and acquire the ability to replicate infinitely, and this transition has been termed ``immortalisation'' \citep{Hanahan2011, Wright1989}.
@@
 It is thought that the angiogenic ability of the tumour is gained at an early stage of tumour development by disrupting the ``angiogenic switch'' \citep{Hanahan2011}.
-% Like many cellular processes, angiogenesis is controlled by the balance between the inducing and inhibiting agents of angiogenesis.
 \Gls{vegfa} is the major growth factor that triggers angiogenesis by signalling through the \acrshort{vegf} receptors (though some receptors have inhibitory effect) \citep{Yancopoulos2000}.
@@
 \noindent
-One of the prime reasons that cancers are so difficult to treat and have high mortality is due to their ability to metastasise to other parts of the body.
+One of the prime reasons that cancers are so difficult to treat and have high mortality rate is due to their ability to metastasise to other parts of the body.
 Once metastasised to different parts of the body, the clinician not only has to consider the primary tumour site, but also the second, third, or even fourth tumour sites, which may or may not have similar tumour biology to the primary tumour, and thus have to consider multiple treatment plans to cure the patient.
@@
 Metastasis is a multi-step process that requires the selection and survival of cancer cell(s) that is able to metastasise (``seed''), and is also compatible with the specific target tissue where they prosper (``soil'') \citep{Talmadge2010}.
-% The downregulation or loss of function of E-cadherin, a protein important in cell-to-cell connection and adherence, is known to trigger tumour invasion and metastasis through the reversible process of \gls{emt} \citep{Hanahan2011,Kalluri2009}.
 Once mobile and invasive, metastatic cancer cell is able to circulate in the vascular system and eventually invade the target organ, where it must undergo \gls{met} and colonise the site \citep{Hanahan2011,Kalluri2009}.
@@
 Until recently, the evolutionary progress of cancers have been thought to be gradual and stepwise, where each driver mutation is acquired over many years, or even decades \citep{Stephens2011}.
-There are now evidence that show otherwise, whereby a single catastrophic event in the chromosome structure occurs, resulting in major structural rearrangements such as inversions, deletions, duplications and translocations \citep{Leibowitz2015,Stephens2011}.
-% This rare event that is estimated to occur in 2$\sim$5\% of all cancers was termed \gls{chromothripsis}, and is thought to be caused by the physical separation of parts of chromosome by a structure called micronuclei \citep{Leibowitz2015,Stephens2011}.
+There are now evidence that show otherwise, whereby a single catastrophic event (termed \gls{chromothripsis}) occurs in the chromosome structure, resulting in major structural rearrangements such as inversions, deletions, duplications and translocations \citep{Leibowitz2015,Stephens2011}.
 Owing to its nature of being a one-off chromosomal disruption, chromothripsis (usually) cause disruptions in tumour suppressor genes rather than duplications of oncogenes \citep{Leibowitz2015}.
@@
 Although the full mechanism in which \gls{kataegis} occur is not yet defined, \gls{aid}/\acrshort{apobec} family proteins, a family of proteins involved in gene editing, may play a major role in \gls{kataegis} \citep{Leibowitz2015,Nik-Zainal2012}.
-% By deaminating cytidine to uracil in the \gls{ssdna} that may occur at the sites of double strand break, \gls{aid}/\acrshort{apobec} family proteins introduces hypermutated regions in the genome \citep{Leibowitz2015,Nik-Zainal2012}.
 It is evident that both \gls{chromothripsis} and \gls{kataegis} contribute to the genomic aberrations and mutations that ultimately encourage the acquisition of many of the tumorigenic phenotypes.
@@
 Inflammation and immune response, though it may assist in the control of diseases and injuries, have been implicated in the progression of tumour and their acquisition of many of the cancer hallmarks \citep{Hanahan2011}.
-% There are two pathways of inflammation that contribute to the cancer survival and progression \citep{Mantovani2008}.
-% The intrinsic pathway is caused by oncogenic characteristics of the cancer cells, where these cells produce inflammatory signals themselves and promote advantageous characteristics for the cancer cells, such as increased proliferation and survival \citep{Mantovani2008}.
-% On the other hand, the external pathway is caused by infectious or inflammatory conditions in the tumour microenvironment that surround the tumour and assist in the growth of the tumour \citep{Mantovani2008}.
-% There are two key transcription factors that connect these two pathways of inflammation that ultimately leads to the prommotion  of many, if not all, of the hallmarks discussed so far: \gls{nfkb} and \gls{stat3} \citep{Mantovani2008}.
 There are two key transcription factors of inflammation that ultimately leads to the prommotion  of many, if not all, of the hallmarks discussed so far: \gls{nfkb} and \gls{stat3} \citep{Mantovani2008}.
@@
 \gls{nfkb} regulates the expression of genes that encode the inflammatory cytokines, angiogenic factors, and many others, including anti-apoptotic genes (through the activation of \gls{stat3}) \citep{Elinav2013,Mantovani2008}.
-% \gls{stat3} is a transcription factor that is activated through variety of receptors and pathways, including the \gls{il6} receptor, \glspl{gpcr} and \glspl{tlr} \citep{Yu2007,Yu2014}.
-\gls{stat3} is a transcription factor that is activated through variety of receptors and pathways \citep{Yu2007,Yu2014}.
+\gls{stat3} is a transcription factor that is activated through a variety of receptors and pathways \citep{Yu2007,Yu2014}.
 By increasing the transcription of anti-apoptotic proteins such as \gls{bcl2}, \gls{stat3} promotes tumour survival and proliferation \citep{Yu2007}.
 Furthermore, \gls{stat3} is able to increase the expression of inflammatory cytokines (such as \gls{il6} and \acrshort{il10}) and \acrshort{vegf}, which provides a positive feedback mechanism for further inflammation (and therefore survival and proliferation signals) and advancement of angiogenesis, respectively \citep{Yu2007}.
-
-% Though helpful at times, inflammation induced by the immune system must be taken care of and controlled in order to prevent rapid tumour development and progression.
 
@@
 In order to grow and proliferate, cancer cells must adjust their energy metabolism to match their energy requirement \citet{Hanahan2011}.
-% There are two ways in which a cell can produce energy: anaerobic respiration, where the cell uses glycolysis to rapidly generate \gls{atp} and lactate; and aerobic respiration where, in the presence of oxygen, the cell utilises the electron transport chain and \gls{atp} synthase in the mitochondria to efficiently produce \gls{atp}.
-% However, even with abundant oxygen in the cell, \citet{Wardburg1956} found that cancer cells utilised the glycolytic pathway instead of the more efficient aerobic respiration, which was later termed ``aerobic glycolysis'' \citep{Hanahan2011}.
-Even with abundant oxygen in the cell, \citet{Wardburg1956} found that cancer cells utilised the glycolytic pathway instead of the more efficient aerobic respiration, which was later termed ``aerobic glycolysis'' \citep{Hanahan2011}.
+\citet{Wardburg1956} found that cancer cells, even with an abundance of oxygen in the cell, utilised the glycolytic pathway instead of the more efficient aerobic respiration, and was later termed ``aerobic glycolysis'' \citep{Hanahan2011}.
 It seems counterintuitive to limit its full potential of creating \gls{atp} by utilising glycolysis over electron transport chain, as the latter produces almost 18 times more \gls{atp} \citep{Hanahan2011, VanderHeiden2009}.
 However, it is hypothesised  that this apprent inefficiency helps the cell to generate and direct the molecules required for biosynthetic pathways during cell growth and division \citep{Cairns2011,VanderHeiden2009}.
-By deliberately relying purely on glycolysis (known as the ``Wardberg effect'') and slowing the conversion of \gls{pep} to pyruvate, the cell is able to shunt the carbon molecules into reaction pathways that produce constituents used to synthesise macromolecules, such as \acrshort{dna} and lipids \citep{Cairns2011,VanderHeiden2009}.
+By deliberately relying purely on glycolysis (known as the ``Wardberg effect'') and slowing the conversion of \gls{pep} to pyruvate, the cell is able to direct the carbon molecules into reaction pathways that produce the constituents used to synthesise macromolecules, such as \acrshort{dna} and lipids \citep{Cairns2011,VanderHeiden2009}.
 
@@
 Immunoediting is a process where the immune system successfully eliminates the highly immunogenic cancer cells during surveillance, but fails to eliminate the weakly immunogenic cells and therefore poses a selective advantage on these cells \citep{Hanahan2011,Teng2008}.
-In fact, there has been reports where cancer cells in immunodeficient mice were not able to initiate tumour growth in the immunocompetent mice, but the cancer cells from the immunocompetent mice were able to do so in immunodeficient mice \citep{Hanahan2011}.
+In fact, there have been reports where the cancer cells in the immunodeficient mice were not able to initiate tumour growth in the immunocompetent mice, but the cancer cells from the immunocompetent mice were able to do so in the immunodeficient mice \citep{Hanahan2011}.
 One possible reason for this was that the cancer cells from the immunodeficient mice were not immunoedited to the extent in which the cells from immunocompetent mice were, and thus unable to evade the immune system as efficiently as those cells in the immunocompetent mice \citep{Hanahan2011}.
-Though more experiments must be carried out to confirm its definitive role and mechanisms in which it contributes to cancer biology, immune system evasion is in no doubt a significant component of tumour progression.
+Though more experiments must be carried out to confirm its definitive role and mechanisms in which immunoediting contributes to cancer biology, immune system evasion is no doubt a significant factor that contribute to tumour progression.
 \\
@@
 There are many causes and risk factors associated with the development and progression of cancer.
-Like obesity, cancers are dependent on both the genetic and environmental factors.
+Like obesity, cancers are dependent on both genetic and environmental factors.
 
@@
 
-% It is quite obvious from the previous section (\cref{sub:mechanisms_of_cancer_development}) that there are a lot of things that can go wrong in the cell.
-Any one of the ten Hallmarks of Cancers has the potential to initiate cancer growth and progression; every hallmark has their own complex mechanistic pathways; every pathway has many proteins associated with its function or role in the cell; and, though their impacts may differ from protein to protein, mutations in any one of the proteins involved in a pathway has the potential to cause cancer.
+Any one of the ten Hallmarks of Cancers has the potential to initiate cancer growth and progression and every hallmark has their own complex mechanistic pathways.
+Furthermore, every pathway has many proteins associated with its function or role in the cell and, though their impacts may differ from protein to protein, mutations in any one of the proteins involved in a pathway have the potential to cause cancer.
 
 Oncogenes are genes that promote or ``accelerate'' the process of tumorigenesis and/or tumour progression, whereas tumour suppressor genes prevent or ``decelerate'' those effects \citep{Vogelstein2004a}.
-Cancers usually mutate oncogenes in such a way that they become constitutively activated and mutate the tumour suppressor genes so that these genes are inactivated and/or have reduced expression \citep{Vogelstein2004a}.
-Another thing to note for these genes is that oncogenes usually only require a ``one-hit'' mutation that affects a single allele, whereas tumour suppressor genes generally require mutations in both maternal and paternal alleles to have any effect \citep{Stratton2009,Vogelstein2004a}.
+Cancers usually mutate oncogenes in such a way that they become constitutively activated, and mutate the tumour suppressor genes so that these genes are inactivated and/or have reduced expression \citep{Vogelstein2004a}.
+Another thing to note for these genes is that the oncogenes usually only require a ``one-hit'' mutation that affects a single allele, whereas the tumour suppressor genes generally require mutations in both maternal and paternal alleles to have any effect \citep{Stratton2009,Vogelstein2004a}.
 
@@
 There are two types of mutations that occurs in an individual: germline mutation and somatic mutation.
-Germline mutations are the mutations that are passed on from an individual's parents, and somatic mutations are the mutations acquired by the cell over the course of the individual's lifetime.
+Germline mutations are the mutations that are passed on from the parents of an individual, and somatic mutations are the mutations acquired by the cell over the course of the lifetime of an individual.
 
@@
 These individuals have a mutation in one or more of the classes of cancer-related genes, which makes them more likely to develop a tumour than individuals without the mutation \citep{Vogelstein2004a}.
-In contrast, somatic mutations are acquired during an individual's lifetime and is not passed down to their children.
+In contrast, somatic mutations are acquired during the lifetime of an individual, and is not passed down to their children.
 Somatic mutations can either be a driver or a passenger mutation, where a driver mutation helps tumorigenesis and its subsequent progression (in other words, mutations in oncogene, tumour suppressor, or stability gene), and passenger mutation have neither a selective advantage nor contribute to tumour progression \citep{Stratton2009}.
@@
 Since cancers require multiple driver mutations, somatic mutations are an essential part of tumour development.
-Perhaps that is one of the reasons why, in some cancers, germline mutation in the stability genes cause greater susceptibility to cancers than mutations in oncogene or tumour suppressor gene \citep{Vogelstein2004a}.
+Perhaps that is one of the reasons why, in some cancers, germline mutation in the stability genes (such as \textit{BRCA1} and \textit{BRCA2} genes) cause greater susceptibility to cancers than mutations in the oncogene or the tumour suppressor gene \citep{Vogelstein2004a}.
 
@@
 
-% For obesity, since there is a clear evidence of the cause (excessive eating and lack of exercise) and a well defined mechanism of how the disease is developed (excess calories are converted into fat), it is relatively easy to speculate the effect of the environment on the disease (for example, abundance of high calory food).
-Cancers are developed through many cellular pathways (Hallmarks of Cancers) which are not fully uncovered yet, with either too many or very few causes associated with those pathways.
+Cancers are developed through many cellular pathways (Hallmarks of Cancers) that are not fully uncovered yet, with either too many or very few causes associated with those pathways.
 Therefore it is difficult to assess how, rather than whether, a given environmental factor causes specific cancer.
 With that said, several environmental factors have a strong evidence of causing cancer.
-% , though the complete mechanism of action for some of these must be elucidated in the coming years.
-% Out of these diet and obesity will be covered as an example.
 
@@
 For current smokers, the relative risk is greater than 1.5 in many cancers, including liver, cervix and kidney, and it can rise up to 8.96 in lung cancer \citep{Gandini2008}.
-% Even if smokers stop smoking, the relative risk does not decrease until the first 5 years have passed and will never return to the level of non-smokers \citep{Hecht1999}.
-% In tobacco smoke, there are literally hundreds and thousands of chemical compounds, of which about 20 compounds have been shown to have carcinogenic effects \citep{Hecht1999}.
-% Among these, \glspl{pah} and \gls{nnk} have been shown to cause lung cancer in mice when administered systemically \citep{Hecht1999}.
 Tobacco smoke contain many carcinogens which are processed first by \gls{cyp} enzymes (termed meta\-bolic activation), and the product can either be processed further to be detoxified completely and cause no harm to the cell, or alternatively, the product can covalently bind to the \acrshort{dna}, forming \acrshort{dna} adducts \citep{Hecht1999}.
 Unless these adducts are resolved correctly by the \acrshort{dna} repair mechanisms, it can result in a permanent mutation within the genome and lead to oncogenesis \citep{Hecht1999}.
-% In addition to this, cigarettes contain free radicals and chemicals that cause tremendous oxidative stress and damage to the cells \citep{Hecht1999}.
-% These radicals cause \acrshort{dna} nicks and damages the \acrshort{dna}, again requring the cell's \acrshort{dna} repair mechanisms to remove the damage so it does not become permanent.
 
-% Another important environmental factor to consider is \gls{uv} radiation.
 Exposure to \gls{uv} radiation has been significantly associated with the increased risks of skin cancers \citep{Armstrong2001,Gallagher2006}.
-% The risks of developing any skin cancers can vary between race (Caucasian people are more likely to get skin cancers), and potentially the age when they were exposed to high amounts of \gls{uv} radiation \citep{Armstrong2001,Gallagher2006}.
 Similar to smoking, the mechanism of how \gls{uv} radiation may cause  skin cancers is related to the \acrshort{dna} repair system.
@@
 
-Diet and obesity are known risk factors for many types of cancers as well \citep{Ames1995a,Calle2004}.
-% Though more will be covered in details in \cref{sec:obesity_and_cancer}, the effect of obesity on cancers will be briefly mentioned here.
+Diet and obesity are known risk factors for many types of cancers as well \citep{Ames1995,Calle2004}.
 Epidemiological evidence of obesity being a risk factor for many cancers has been well established, yet the mechanisms of its contribution to cancer is still  speculative \citep{Calle2003,Kelesidis2006}.
 Although there are a lot of possible mechcanisms, there is growing evidence of the role of chronic inflammation caused by increased fat mass in the body \citep{Kelesidis2006,Lumeng2011,Hernandez2013}.
-Free fatty acids from excess adipose tissue and/or diet can be recognised by \gls{tlr} (specifically \gls{tlr}4), which can then activate \gls{nfkb}, leading to the activation of one of the Hallmarks of Cancer -- tumour-promoting inflammation \citep{Lumeng2011}.
-% The nutritional content of the food that people consume is also important, as it can both directly and indirectly influence the likelihood of tumorigenesis, as well as its progression \citep{Ames1995a}.
-% Many fruits and vegetables contain antioxidants as well as essential viatmins and minerals.
-% Antioxidants are able to neutralise the effect of \gls{ros},  and many viatmins have been associated with reduction of many types of cancers \citep{Ames1995a}.
-% In fact, ascorbic acid (vitamin C) is able to prevent the lipid peroxidation of lipid plasma caused by tobacco smokes \textit{in vitro}, which could be why smokers, compared to non-smokers, have decreased levels of ascorbic acid \citep{Hecht1999}.
-% By maintaining healthy diet, it may be possible to reduce the risks of many cancers and obesity, as well as the comorbidities associated with them.
-
-% As one can imagine, \acrshort{dna} damages caused by smoking and/or \gls{uv} radiation have the potential to damage the genome and initiate cancer if wrong genes are mutated, such as p53 tumour suppressor gene.
-% Healthy diet is also crucial for lowering the risk of cancers through beneficial vitamins and antioxidants, and prevent obesity as well as chronic inflammation that accompanies with it.
-There are many other environmental factors associated with cancers, and limiting the exposure to detrimental environmental factors is crucial for the prevention of any type of cancers, as well as the more specific type of cancers associated with these factors, such as lung and skin cancers.
+Free fatty acids from excess adipose tissue and/or diet can be recognised by \gls{tlr} (specifically \gls{tlr}4), which can then activate the \gls{nfkb} pathway, leading to the activation of one of the Hallmarks of Cancer (tumour-promoting inflammation) \citep{Lumeng2011}.
+There are many other environmental factors associated with cancers and limiting the exposure to detrimental environmental factors is crucial for the prevention of many cancers.
 
@@
 Traditionally, the treatment of cancer has been surgery, radiation therapy and chemotherapy.
-Though surgery is an effective method to remove the tumour from the patient, there is always a risk of metastasis and recurrence, which is why combination of surgery with other treatments like chemotherapy is usually considered.
+Though surgery is an effective method to remove the tumour from the patient, there is always a risk of metastasis and recurrence, which is why a combination of surgery with other treatments like chemotherapy is usually considered.
 Earlier chemotherapeutic drugs and radiation therapy had limited effect on tumours due to their resistance against these treatments \citep{Wilhelm2006}.
@@
 Development of drugs that target  the downstream RAF and its mutant forms (such as BRAF$^{V600E}$) have been successful, and many RAF inhibitor drugs such as sorafenib and vemurafenib have been approved for clinical use \citep{Samatar2014,Wilhelm2006}.
-Even though the signalling pathways are well understood and the quality of the drugs that are developed based on these knowledge is improving, there is always the possibility of tumour to acquire resistance to those drugs \citep{Samatar2014}.
+Even though the signalling pathways are well understood and the quality of the drugs that are developed based on these knowledge is improving, there is always the possibility of the tumour to acquire resistance to those drugs \citep{Samatar2014}.
 
-In fact, there has been evidence of pathway cross-talks between RAS-ERK pathway and \gls{pi3k}/Akt pathway, which may help the tumour cells achieve resistance against those drugs \citep{Moelling2002,Zimmermann1999a}.
-A study in a \textit{Drosophila} model showed that the inhibition of just a single pathway does not necessarily stop the effect of the pathway, as the signal can flow through alternate pathways \citep{Dar2012}.
-Pathway cross-talk and increased flux through alternate pathways could be one of the ways in which tumour cells acquire resistance to drugs.
-However, the effectiveness of targeting multiple pathways in humans is yet to be confirmed and further investigations are required to solidify the mode of action and its relevance to tumour biology.
+In fact, there have been evidence of pathway cross-talks between the RAS-ERK pathway and the \gls{pi3k}/Akt pathway, which may help the tumour cells achieve resistance to the  drugs that block the RAS-ERK pathway \citep{Moelling2002,Zimmermann1999a}.
+A study in a \textit{Drosophila} model showed that the inhibition of just a single pathway does not necessarily stop the effect of the pathway, as the signal can flow through other pathways \citep{Dar2012}.
+Pathway cross-talk and increased flux through alternative pathways could be one of the ways in which tumour cells acquire resistance to drugs.
+The effectiveness of targeting multiple pathways in humans is yet to be confirmed and further investigations are required to solidify the mode of action and its relevance to tumour biology.
 
-Further understanding of the tumour biology, its mechanism of resistance to drugs, and the effect of targeting multiple pathways (or targeting effective pathways for a patient) will definitely become important in the near future, both for novel drug development and better clinical treatment plan for individual patients.
+Further understanding of the tumour biology, its mechanism of resistance to drugs, and the effect of targeting multiple pathways (or targeting effective pathways for a patient) will definitely become important in the near future, both for novel drug development and for better clinical treatment plans for the patients.
 
@@
 As mentioned briefly in \cref{sub:causes_and_risk_factors_of_cancer}, obesity is considered a major risk factor for many types of cancer.
-The fact that the prevalence of obesity is extremely high, it is important to establish the link between the apparent risks associated with cancer.
+The prevalence of obesity is extremely high around the world and  it is important to establish the link between the apparent risks associated with cancer.
 
@@
 
-Although the mechanisms in which obesity causes cancer is still under debate, the association between obesity and cancer has long been established.
+Although the mechanisms in which obesity causes cancer are still under debate, the association between obesity and cancer has long been established.
 Perhaps the most convincing evidence of the association between obesity and cancer was first presented by \citet{Calle2003}.
@@
 
-% Taken together, it is undeniable that obesity is a major risk factor for many types of cancers, and it may even be the cause for some of these cancers.
-
 \begin{table}[htb]
@@
 	\begin{threeparttable}
-		\caption{Summary of the risk ratios\tnote{1} associated with each type of cancers from the meta-analysis study by \citet{Renehan2008}\tnote{2}.}
+		\caption[Summary of the risk ratios associated with each cancer type from the meta-analysis study by \citet{Renehan2008}]{Summary of the risk ratios\tnote{1} associated with each cancer type from the meta-analysis study by \citet{Renehan2008}\tnote{2}.}
 		\label{tab:renehan_cancer_risks}
@@
 				\item [1] All risk ratios are per increase in 5 kg/m$^2$ \gls{bmi}; 95\% confidence intervals in brackets.
-				\item [2] Table adapted from \citet{Roberts2010} paper, where the original data in the table was taken from \citet{Renehan2008} study.
+				\item [2] Table adapted from \citet{Roberts2010}, where the original data in the table was from \citet{Renehan2008}.
 				\item [3] Not applicable.
@@
 The exact mechanisms in which obesity causes cancer have still not been fully elucidated yet.
-There are several hypotheses that link obesity with cancer: disruption of various hormone levels, chronic and sustained inflammation, and activation of the insulin/\gls{igf} pathway.
+There are several hypotheses that link obesity with cancer: disruption of various hormone levels, chronic and sustained inflammation, and the activation of the insulin/\gls{igf} pathway.
 
-% Previously thought to be a group of cells where the body could store and manage excess fats, it has recently become apparent that adipose tissue is also a metabolically active organ that has a significant role in the endocrine system \citep{Roberts2010}.
 Adipose tissue has long been thought to be a group of cells to store excess fats, but it has recently become apparent that adipose tissue is also a metabolically active endocrine organ \citep{Roberts2010}.
 Increased adiposity is known to influence the bioavailability of endogenous sex hormones, including estrogens, androgens and progesterone \citep{Calle2004}.
-\Glspl{adipocyte} synthesise enzymes, such as aromatase, that convert precursor molecules into different forms of androgens and estrogens \citep{Calle2004}.
-Many studies have confirmed the mitogenic effect of sex hormones on certain cell types, mainly breast and endometrial cells \citep{Roberts2010}.
+\Glspl{adipocyte} synthesises enzymes, such as aromatase, that convert precursor molecules into different forms of androgens and estrogens \citep{Calle2004}.
+Many studies have confirmed the mitogenic effect of sex hormones on certain cell types; mainly breast and endometrial cell types \citep{Roberts2010}.
 
-On top of the effects that \glspl{adipocyte} have on circulating concentration of sex hormones, they also produce adipose tissue-specific hormones called adipokines \citep{Roberts2010}.
+On top of the effects that \glspl{adipocyte} have on circulating concentration of sex hormones, \glspl{adipocyte} also produce adipose tissue-specific hormones called adipokines \citep{Roberts2010}.
 Of the many adipokines identified so far, leptin and adipo\-nectin have been studied the most in terms of cancer development \citep{Renehan2006,Roberts2010}.
 Leptin is known to activate not only the satiety signalling pathway, but also the \gls{jak}/\gls{stat} and \gls{pi3k}/Akt signalling pathways \citep{Garofalo2006,Renehan2006}.
-Activation of these pathways lead to the activation of target genes that are involved in cell proliferation and cell survival, as well as pro-angiogenic factors \citep{Garofalo2006}.
+Activation of these pathways lead to the activation of the target genes that are involved in cell proliferation and cell survival, as well as pro-angiogenic factors \citep{Garofalo2006}.
 
 Adiponectin is a hormone that is synthesised and secreted only by the \glspl{adipocyte} \citep{Kelesidis2006}.
-Unlike leptin, where the concentration of leptin is higher in obese people, adiponectin levels are lower in obese people compared to the normal weight people \citep{Kelesidis2006,Renehan2006}.
-Adiponectin has anti-angiogenic and anti-proliferative effect on the cells, and is secreted by mature adipocytes but not by the premature pre-adipocytes \citep{Gilbert2013}.
+Unlike leptin, where the concentration of leptin is higher in those who are obese, adiponectin levels are lower in obese people compared to the normal weight people \citep{Kelesidis2006,Renehan2006}.
+Adiponectin have anti-angiogenic and anti-proliferative effects on the cells, and is secreted by mature adipocytes but not by the premature pre-adipocytes \citep{Gilbert2013}.
 During chronic inflammatory state in obese patients, pro-inflammatory proteins such as \gls{tgfb} and \gls{ifny} can block the maturation of the pre-adipocytes, thus decreasing the level of adiponectin and the accompanying beneficial effects \citep{Gilbert2013}.
 
-Alternatively, or possibly in conjunction with hormone disruption, inflammation caused by obesity may be associated with tumour-promoting environment.
+Alternatively, or possibly in conjunction with the disruption of hormone levels, inflammation caused by obesity may be associated with tumour-promoting environment.
 \Gls{tlr}4 is a type of \gls{prr} and, as mentioned earlier, is able to recognise circulating free fatty acids and trigger immune response via \gls{nfkb} pathway \citep{Lumeng2011}.
 In addition to \gls{nfkb} pathway, recent evidence have shown that \gls{tlr}4 could also activate the \gls{jak}/\gls{stat3} pathway \citep{Yu2014}.
-Activation of these pathways will further enhance the development of the tumour-promoting microenvironment caused by these unwanted inflammatory responses of obesity.
-Furthermore, since it is likely that obese people have less adiponectin concentration, obesity-induced inflammation may affect these patients more than those with normal weight.
+Activation of these pathways will further enhance the development of the tumour-promoting microenvironment caused by these unwanted inflammatory responses brought upon by obesity.
+Furthermore, since it is likely that the people who are obese have less adiponectin concentration, obesity-induced inflammation may affect the patients that are obese more than those with normal weight.
 
 Finally, insulin/\gls{igf} pathway may have a major role in encouraging tumorigenesis in obese patients.
-% \Gls{igf} system involves two ligands (\gls{igf}-I and \gls{igf}-II), its respective receptors (\gls{igf}-IR and \gls{igf}-IIR), and six binding proteins (\gls{igfbp} 1 to 6) \citep{Roberts2010}.
 \gls{igf} signalling results in the activation of \gls{pi3k}/Akt and \gls{erk} pathways, resulting in cell proliferation and reduced apoptosis \citep{Roberts2010}.
 Though insulin is also capable of activating these pathways, it is thought that the mitogenic effects are controlled mainly through the \gls{igf} receptors \citep{Roberts2010}.
-Since insulin reduces the level of \gls{igfbp}-3, the main circulating \gls{igfbp}, hyperinsulinemia has been hypothesised to promote tumorigenesis through the mitogenic effect of insulin and free \glspl{igf} \citep{Giovannucci1995,Mckeown1994,Roberts2010}.
+Since insulin reduces the level of \gls{igfbp}-3, the main circulating \gls{igfbp}, hyperinsulinemia has been hypothesised to promote tumorigenesis through the mitogenic effect of insulin and by increasing the level of free \glspl{igf} \citep{Giovannucci1995,Mckeown1994,Roberts2010}.
 In fact, many studies have consistently associated the increased concentration of free \gls{igf}-I with obesity and increased risk of some cancers; however the results for \gls{igfbp} have been inconsistent \citep{Basen2011}.
@@
 Furthermore, these mechanisms are related to one another and are part of a larger, more complex physiological network \citep{Renehan2006}.
-Due to this complexity, it is difficult to confidently explain how obesity contributes to tumour development.
+Due to this complexity, it is difficult to confidently explain how obesity contributes to the tumour development.
 
@@
 One of the difficulties associated with the treatment of cancers is the identification of the underlying mechanisms or hallmarks that drive the cancer progression.
-As one can imagine, it is difficult for clinicians to treat a patient if they do not know what is causing the disease.
-The reason why it is so difficult to identify the underlying biological mechanisms is that there are very little reliable biological markers (or ``genetic signatures'') that signify the presence or absence of these pathways.
-With that said, significant improvements have been made in recent years owing to the advancement in sequencing technologies and have allowed researchers to study the genetic signatures that represent the hallmarks in a much greater detail.
+As one can imagine, it is difficult for the clinicians to treat a patient if they do not know what is causing the disease.
+The reason why it is so difficult to identify the underlying biological mechanisms is that there are very little reliable biological markers (or ``genetic signatures'') that signify the presence or absence of the causal pathways.
+With that said, significant improvements have been made in recent years owing to the advancement in the sequencing technologies and have allowed the researchers to study the genetic signatures that represent the hallmarks in a much greater detail.
 
@@
 
-Microarray technology was first developed by \citet{Schena1995} and has been used extensively to study gene expression patterns and \gls{snp} identification for over two decades.
-In short, microarray technology uses a glass slide with hundreds of thousands of short \acrshort{dna} sequences attached to it, where fluorescently labelled \glspl{cdna} taken from the sample \glspl{mrna} are hybridised to these immobilised \acrshort{dna} sequences, and the fluorscent probes are detected to measure the gene expression \citep{Schena1995,Schulze2001}.
-% Since only the exactly matching sequences remain on the microarray slide, researchers are able to identify and quantify the gene expression level of the sample \citep{Schulze2001}.
+Microarray technology was first developed by \citet{Schena1995} and has been used extensively to study the gene expression patterns and \gls{snp} identification for over two decades.
+In short, microarray technology uses a glass slide with hundreds of thousands of short \acrshort{dna} sequences attached to it, where fluorescently labelled \glspl{cdna} taken from the sample \glspl{mrna} are hybridised to these immobilised \acrshort{dna} sequences \citep{Schena1995,Schulze2001}.
+The fluorscent probes are then detected to measure the gene expression \citep{Schena1995,Schulze2001}.
 
-One disadvantage of microarray technology is that it requires previous knowledge of the genes and \glspl{snp} of the samples being investigated \citep{Hurd2009}.
-This is due to the fact that the detection of sequences rely on the hybridisation of the sample sequences with the microarray sequences and therefore the sequences under investigation must be known to attach the correct sequences to the microarray.
+One disadvantage of the microarray technology is that it requires previous knowledge of the genes and \glspl{snp} of the samples being investigated \citep{Hurd2009}.
+This is due to the fact that the detection of sequences rely on the hybridisation of the sample sequences with the microarray sequences and therefore the sequences under investigation must be known prior to the investigation to attach the correct sequences to the microarray.
 Another disadvantage of microarrays is the potentially erroneous detection of the sample sequences, both in terms of its sequence and its expression level \citep{Hurd2009}.
 Again, because the microarray slides require pre-defined sequences to be attached to it, there is always a possibility of hybridisation of similar but different sequence to the slide.
-Even if the sequences correctly hybridise to the slide, it may be difficult to measure the exact amount of some sequences in the samples, since the amount of sequences in the sample is measured with the relative strength of the fluorescence of the dyes used \citep{Hurd2009}.
-Therefore, it is difficult to detect the exact level of expressions of sequences that are either very rare or highly abundant in the sample \citep{Hurd2009}.
+Even if the sequences correctly hybridises to the slide, it may be difficult to measure the exact amount of some sequences in the samples, since the amount of sequences in the sample is measured with the relative strength of the fluorescence of the dyes \citep{Hurd2009}.
+Therefore, it is difficult to detect the exact level of expressions of the sequences that are either very rare or highly abundant in the sample \citep{Hurd2009}.
 
@@
 First, \acrshort{dna} samples are fragmented into smaller sizes and universal primers are attached to the ends of the fragments.
-% These fragments are then amplified into clusters to make the signals stronger for detection (although this amplification step vary between different techniques) \citep{Metzker2010}.
 These fragments are then amplified into clusters to make the signals stronger for detection \citep{Metzker2010}.
@@
 
-There is a clear evidence of obesity being a major risk factor for many cancer types, but what genes or pathways directly contribute to tumorigenesis in patients that are obese is yet to be clarified.
-Two studies in particular, one by \citet{Creighton2012} and another by \citet{Fuentes-Mattei2014}, have investigated if there were any difference in genetic composition between obese and normal weight patients, specifically in a group of breast cancer patients.
-In both studies, the authors were able to identify a set of genes that were up or downregulated in breast tumours from obese patients.
-% For simplicity, these genetic signatures will be referred to as ``obesity associated genetic signatures'' hereafter.
+There is a clear evidence of obesity being a major risk factor for many cancer types (\cref{sec:obesity_and_cancer}), but the genes or pathways that directly contribute to tumorigenesis in patients that are obese is yet to be clarified.
+Two studies in particular, one by \citet{Creighton2012} and another by \citet{Fuentes-Mattei2014}, have investigated whether there were any difference in genetic composition between the obese and the non-obese patients, specifically in a group of breast cancer patients.
+In both studies, the authors were able to identify a set of genes that were up or downregulated in breast tumours from the patients that were obese.
 
@@
 It was also noted that African American patients were more likely to be obese than the Caucasian patients.
-They compared the gene expression between the tumours from obese patients with the tumours from non-obese patients and identified 799 probe sets (662 unique genes) with p \textless{} 0.01 and with a \gls{fc} \textgreater{} 1.2.
-Majority of these genes (725 gene probes) were downregulated and 74 gene probes were upregulated in the samples from obese patients compared with the samples from non-obese patients.
+They compared the gene expression between the tumours from obese patients with the tumours from non-obese patients and identified 799 gene probe sets (662 unique genes) with p \textless{} 0.01 and with a \gls{fc} \textgreater{} 1.2.
+Majority of these genes (725 gene probes) were downregulated and 74 gene probes were upregulated in the breast tumour samples from the obese patients compared with the breast tumour samples from the non-obese patients.
 
@@
 \citet{Fuentes-Mattei2014} studied 137 breast tumour samples from \gls{er} positive patients and analysed their transcriptome using microarrays.
-Of the 137 patients, 94 patients were not obese and 43 patients were.
-Comparison of the gene expression of the tumour samples from obese patients with the samples from non-obese patients revealed 112 genes that were statistically significantly dysregulated (62 genes upregulated and 50 genes downregulated).
-Furthermore, functional transcriptomic analysis suggested that insulin signalling and inflammation affected human \gls{er}$^+$ breast cancer.
+Of the 137 patients, 43 patients were obese and 94 patients were not.
+Comparison of the gene expressions of the tumour samples from the obese patients with the tumour samples from the non-obese patients revealed 112 genes that were statistically significantly dysregulated (62 genes upregulated and 50 genes downregulated).
+Furthermore, functional transcriptomic analysis suggested that the insulin signalling and inflammation affected human \gls{er}$^+$ breast cancer.
 
-On top of this, \citet{Fuentes-Mattei2014} validated their results using a mouse model, where they compared the expression level in the breast tumours from obese mice with the samples from lean mice.
-Results from this animal model provided strong evidence that obesity significantly accelerates tumour growth through Akt/\gls{mtor} signalling pathway.
+On top of this, \citet{Fuentes-Mattei2014} validated their results using a mouse model, where they compared the expression level in the breast tumours from the obese mice with those from the lean mice.
+Results from this animal model provided strong evidence that obesity significantly accelerated tumour growth through Akt/\gls{mtor} signalling pathway.
 
@@
 Like obesity associated genetic signatures, there has also been a focus on the identification of pathways and genes that affect tumour growth.
-About a decade ago, \citet{Bild2006} demonstrated that a collection of genes that are involved in a specific pathway could be used to determine the status of those pathways in a sample.
+About a decade ago, \citet{Bild2006} demonstrated that a collection of genes that are involved in a specific pathway could be used to determine the status of those pathways in tumour samples.
 These pathway specific genes were identified experimentally by altering the expression of one specific gene that is central to the pathway in a model cell culture system.
-For example, all of the genes that have altered its expression in the cells that have \textit{Myc} gene induced were noted down as the genes involved in the Myc pathway.
+For example, all of the genes that have altered its expression in the cells that had  the \textit{Myc} gene induced were noted down as the genes involved in the Myc pathway.
 
@@
 This was a significant improvement from their previous works on pathway signatures \citep{Bild2006,Bild2009}.
-Using more pathways in their analysis have allowed them to identify certain pathways to be coactivated (or cluster) together, suggesting that some pathways are related to one another and may have relevance in identifying the causes.
+Using more pathways in their analysis have allowed them to identify certain pathways to be coactivated (or cluster) together, suggesting that some of the pathways are related to one another and may have relevance in identifying the biological causes of cancer.
 
-One thing to note here is that most of these pathway associated genetic signatures are involved in one or more of the Hallmarks of Cancer.
-This means that these pathway associated signatures can be used to provide insights into the underlying biological mechanisms of the samples used in this project.
+One thing to note here is that most of these pathway associated genetic signatures are involved in one or more of the Hallmarks of Cancer (\cref{subsubsec:cancerhallmarks}).
+This means that these pathway associated signatures can be used to provide insights into the underlying biological mechanisms of the samples used in this project and help understand the underlying biology of cancers in obese patients.
 
-\section{Aim of the project}
+\section{Aims of the project}
 \label{sec:aim}
@@
 In the following chapter, the detailed methods of all the analyses carried out for this project is outlined.
-The results from these analyses are split into two parts, each presented in separate chapters: first part of the results clarify the relationship between the obesity associated genetic signatures with various cancer types; and the second part of the results consist of the links between the pathway associated genetic signatures and various cancer types, as well as the previously identified obesity associated signatures.
-By the end of this thesis, the aim presented here will be answered with clarity.
-
-% TODO: Finished draft copy of introduction -- need to cut down on the contents and link everything together
+The results from these analyses are split into two parts, each presented in a separate chapter.
+First part of the results clarify the relationship between the obesity associated genetic signatures with various cancer types.
+Second part of the results consist of the links between the pathway associated genetic signatures and various cancer types, as well as the previously identified obesity associated signatures.
+Lastly, the results will be summarised and the conclusion will be drawn from these results.

### 2017-03-16 thesis/methods/methods.tex
Draft thesis completed; now edit the whole thing - tmp commit
tmp commit

@@
-\chapter{Methods (draft)}
+\chapter{Methods}
 \label{ch:methods}
@@
 
+\section{Cancer data}
+\label{sec:data}
+
+The raw microarray data from the \citet{Creighton2012}, \citet{Fuentes-Mattei2014} and \citet{Gatza2010a}  studies were downloaded from the \gls{geo} website (\citealp{Edgar2002}; \citetalias{GEO}).
+In addition to these breast cancer data sets, the raw microarray data set from a cohort of \gls{nz} breast cancer patients used in the \citet{Print2016} study was also downloaded from the \gls{geo} website.
+\gls{rnaseq} and clinical data of multiple different cancer types were downloaded from the \gls{icgc} (\citetalias{ICGC}; \citealp{Zhang2011}) and \gls{tcga} \citepalias{TCGA}, respectively.
+
+\subsection{Breast cancer data}
+\label{sub:breast_cancer_data}
+
+\subsubsection{\citet{Creighton2012} data}
+\label{ssub:creighton2012_data}
+
+The raw Affymetrix HGU\-133A microarray gene expression data files from the \citet{Creighton2012} study were downloaded from the \gls{geo} database (\gls{geo} accession ID: GSE24185).
+Clinical data of the samples (age, ethnicity, tumour grade, menopause status, \gls{bmi}, \gls{er} status, \gls{pr} status, \gls{her2} status, and \gls{ln} status) were obtained from the Supplementary Table 1 from the \citet{Creighton2012} study.
+799 obesity associated gene probes identified in the \citet{Creighton2012} study were obtained from the Supplementary Data File 1 from the \citet{Creighton2012} study.
+
+\subsubsection{\citet{Fuentes-Mattei2014} data}
+\label{ssub:fuentes-mattei2014_data}
+
+The raw Affymetrix HGU\-133A microarray gene expression data files from the  \citet{Fuentes-Mattei2014} study were downloaded from the \gls{geo} database (\gls{geo} accession ID: GSE\-20194).
+Clinical data for the samples (age, ethnicity, tumour grade, \gls{er}/\gls{pr}/\gls{her2} statuses, and treatments used) in this study was also downloaded from the \gls{geo} database (same \gls{geo} accession ID).
+130 obesity associated gene probes identified by \citet{Fuentes-Mattei2014} were taken from the Supplementary Table 3 from their paper.
+
+\subsubsection{\gls{nzbc} data}
+\label{ssub:nzbc_data}
+
+The raw Affymetrix HGU133A microarray gene expression data files from the \citet{Print2016} study were downloaded from the \gls{geo} database (\gls{geo} accession ID: GSE3\-6771).
+Clinical data for the samples (age, ethnicity, tumour grade, breast cancer subtype, \gls{er}/\gls{pr} statuses, \gls{ln} status, \gls{bmi} and treatments used) in this study was obtained from Professor Cris Print (personal communication with Assoc.
+Prof.
+Mik Black).
+
+\subsubsection{\citet{Gatza2010a} data}
+\label{ssub:gatza2010a_data}
+
+The raw microarray gene expression data files from the \citet{Gatza2010a} study were downloaded from the \gls{geo} database (\gls{geo} accession ID: GSE1456, GSE\-1561, GSE2034, GSE3494, GSE4922 and GSE6596).
+Only the Affymetrix HGU\-133A microarray samples were included in this project, as the other microarray data from the \citet{Creighton2012}, \citet{Fuentes-Mattei2014} and \citet{Print2016} studies were analysed using the Affymetrix HGU-133A platform.
+Clinical data for the samples in \citet{Gatza2010a} study was not available, as these samples were a combination of many different datasets.
+The Akt, \gls{bcat}, E2F1, \gls{egfr}, \gls{er}, \gls{her2}, \gls{ifna}, \gls{ifny}, Myc, p53, p63, \gls{pi3k}, \gls{pr}, Ras, \gls{stat3}, Src, \gls{tgfb}, and \gls{tnfa} pathway associated genetic signatures were used by Gatza \textit{et al.}, and the gene probes of all the pathway associated genetic signatures were obtained from the Supplemental Table 10 from their paper.
+
+\subsubsection{Gene probe ID conversion}
+\label{ssub:gene_probe_id_conversion}
+
+All of the microarray data used gene probe IDs to refer to the genes, and therefore these probe IDs had to be converted into their corresponding gene symbols.
+The gene probe IDs in the raw data were converted into their corresponding gene symbols using the \textit{hgu133a.db} package in \gls{r} \citep{hgu133}.
+Since multiple gene probe sets match back to a single gene of interest in a microarray chip, there were conflicting expression data for some of the genes after the conversion of the gene probe sets into gene symbols.
+For the gene symbols that had multiple probe set entries, a single probe set was chosen to represent the gene symbol by using the \texttt{collapseRows} function in the \textit{WGCNA} package in \gls{r}, using the default parameters (which chose the probe set with a maximum mean value) \citep{Langfelder2008}.
+Likewise, any obesity associated or pathway associated gene probes were converted into gene symbols.
+
+\begin{table}[htpb]
+	\centering
+	\caption{Number of samples in each of the breast cancer microarray data}
+	\label{tab:num_sample_microarray}
+	\begin{tabular}{lc}
+		Data set & Number of samples\\
+		\hline
+		\hline
+		\rule{0pt}{2.25ex}Creighton \textit{et al.} & 103\\
+		Fuentes-Mattei \textit{et al.}              & 278\\
+		\gls{nzbc}                                  & 99 \\
+		Gatza \textit{et al.}                       & 1060\\
+		\hline
+		\hline
+	\end{tabular}
+\end{table}
+
+\begin{ThreePartTable}
+		\centering
+	\begin{TableNotes}
+		\begin{footnotesize}
+			\item [1] Not available.
+		\end{footnotesize}
+	\end{TableNotes}
+
+	\begin{longtable}{lcc}
+		\caption{Summary of the clinical variables in the Creighton \textit{et al.} and \gls{nzbc} microarray data}
+		\label{tab:clin_summary}\\
+		Clinical variable & Creighton \textit{et al.} & \gls{nzbc} \\
+		\endfirsthead
+		\multicolumn{3}{c}{\tablename\ \thetable{}\ (continued)}\\
+		\endhead
+		\hline
+		\hline
+		\rule{0pt}{2.25ex}Age \\
+		\hspace{1em} Min.                    & 30 & 31 \\
+		\hspace{1em} Max.                    & 72 & 92 \\
+		\hspace{1em} Mean                    & 49 & 59 \\
+		\hspace{1em} Median                  & 48 & 60 \\
+		\hline
+		\rule{0pt}{2.25ex}Ethnicity         \\
+		\hspace{1em} Caucasian               & 77 & 71 \\
+		\hspace{1em} African-American        & 16 & 0  \\
+		\hspace{1em} \gls{nz} M\=aori        & 0  & 10 \\
+		\hspace{1em} Pacific Islands         & 0  & 14 \\
+		\hspace{1em} Asian                   & 10 & 4  \\
+		\hline
+		\rule{0pt}{2.25ex}\gls{bmi} status  \\
+		\hspace{1em} Normal weight           & 36 & 28 \\
+		\hspace{1em} Overweight              & 29 & 28 \\
+		\hspace{1em} Obese                   & 38 & 43 \\
+		\hline
+		\rule{0pt}{2.25ex}Menopause status  \\
+		\hspace{1em} Premenopause            & 49 & NA\tnote{1}\\
+		\hspace{1em} Perimenopause           & 9  & NA \\
+		\hspace{1em} Postmenopause           & 45 & NA \\
+		\hline
+		\rule{0pt}{2.25ex}Tumour grade      \\
+		\hspace{1em} 1                       & 14 & 11 \\
+		\hspace{1em} 2                       & 35 & 39 \\
+		\hspace{1em} 3                       & 54 & 49 \\
+		\hline
+		\rule{0pt}{2.25ex}\gls{er} status   \\
+		\hspace{1em} \gls{er}$^+$            & 58 & 72 \\
+		\hspace{1em} \gls{er}$^-$            & 42 & 27 \\
+		\hspace{1em} Unknown                 & 3  & 0  \\
+		\hline
+		\rule{0pt}{2.25ex}\gls{pr} status   \\
+		\hspace{1em} \gls{pr}$^+$            & 48 & 62 \\
+		\hspace{1em} \gls{pr}$^-$            & 50 & 37 \\
+		\hspace{1em} Unknown                 & 5  & 0  \\
+		\hline
+		\rule{0pt}{2.25ex}\gls{her2} status \\
+		\hspace{1em} \gls{her2}$^+$          & 12 & NA \\
+		\hspace{1em} \gls{her2}$^-$          & 54 & NA \\
+		\hspace{1em} Unknown                 & 37 & NA \\
+		\hline
+		\rule{0pt}{2.25ex}\gls{ln} status   \\
+		\hspace{1em} \gls{ln}$^+$            & 14 & 56\\
+		\hspace{1em} \gls{ln}$^-$            & 89 & 40\\
+		\hspace{1em} Unknown                 & 0  & 3 \\
+		\hline
+		\hline
+		\hspace{12em}\insertTableNotes
+	\end{longtable}
+\end{ThreePartTable}
+
+\subsection{\gls{icgc} cancer data}
+\label{sub:icgc_cancer_data}
+
+\subsubsection{\gls{rnaseq} and clinical data}
+\label{ssub:rnaseq_and_clinical_data}
+
+The clinical data for all available cancer types (33 types in total) were downloaded from \gls{tcga} database (last accessed 1 April 2015) and were checked for both the height and weight data for each sample.
+Any cancer type with no height and/or weight data for the samples was excluded from the project, as no \gls{bmi} information can be obtained without these data.
+
+Out of these 33 cancer types with clinical data, 14 cancer types had both height and weight data.
+However, only 8 cancer types out of these 14 types had \gls{rnaseq} data available from the \gls{icgc} database (last accessed 7 September 2015), so only those 8 cancer types were downloaded and used in this project.
+The selected cancer types were: \gls{blca}, \gls{cesc}, \gls{coad}, \gls{kirp}, \gls{lihc}, \gls{read}, \gls{skcm}, \gls{ucec}.
+
+\subsubsection{Data formatting and processing}
+\label{ssub:data_formatting_and_processing}
+
+The raw \gls{rnaseq} data from the \gls{icgc} database were formatted in a way that the count data of all the genes were listed for one sample, then the count data for all the genes for the next sample, and so on.
+This data format was highly inconvenient for later analyses so the data were reformatted into a ``gene by sample'' matrix format using the \textit{dplyr} package in \gls{r}.
+
+Another problem with the data was the sample ID in the \gls{rnaseq} data.
+Though similar to the \gls{tcga} sample IDs, the \gls{icgc} IDs in the \gls{rnaseq} data had an extra identification code in the sample names.
+To associate each sample in the raw \gls{rnaseq} data from the \gls{icgc} database with the correct samples in the clinical data from the \gls{tcga} database, the \gls{icgc} IDs were re-coded so they matched the \gls{tcga} IDs.
+After the IDs were re-coded, all of the samples were checked to see if there were any duplicates in either the \gls{icgc} \gls{rnaseq} data or \gls{tcga} clinical data.
+Where there was a duplicate in the sample ID, a single sample with the highest mean expression value was chosen to represent that particular ID by using the \texttt{collapseRows} function in the \textit{WGCNA} package in \gls{r} \citep{Langfelder2008}.
+
+Since some of the samples did not have either height or weight data, these samples were removed from the analyses.
+In each of the eight cancer types, the \gls{tcga} sample IDs in the clinical data were cross-checked with the sample IDs in the \gls{rnaseq} data, and vice versa.
+Any samples that did not have either the clinical information or the \gls{rnaseq} data were removed from the analyses.
+This ensured that both clinical and \gls{rnaseq} data were available for all of the samples that were included in this project.
+See \cref{tab:samplesize} for a summary of the number of samples included in the analyses for each cancer type.
+
+\begin{table}[htpb]
+	\caption{Summary of the total number of samples included in the analyses for each cancer type}
+	\label{tab:samplesize}
+	\begin{center}
+		\begin{tabular}{lcc}
+			Cancer Type   & Initial number of samples & Number of samples after exclusion\\
+			\hline
+			\hline
+			\rule{0pt}{2.25ex}BLCA & 295 & 261   \\
+			CESC                   & 259 & 224   \\
+			COAD                   & 428 & 226   \\
+			KIRP                   & 222 & 124   \\
+			LIHC                   & 294 & 264   \\
+			READ                   & 154 & 73    \\
+			SKCM                   & 430 & 218   \\
+			UCEC                   & 508 & 482   \\
+			\hline
+			\hline
+		\end{tabular}
+	\end{center}
+\end{table}
+
+\begin{table}[htpb]
+	\caption{Summary of the sample \gls{bmi} status in the \gls{icgc} cancer data}
+	\label{tab:icgc_bmi}
+	\begin{center}
+		\begin{tabular}{lcccc}
+			& \multicolumn{4}{c}{Number of samples}\\
+			\cmidrule(r){2-5}
+			Cancer type & Normal weight & Overweight & Obese & Total\\
+			\hline
+			\hline
+			\rule{0pt}{2.25ex}BLCA & 103 & 97  & 61  & 261 \\
+			CESC                   & 83  & 64  & 77  & 224 \\
+			COAD                   & 75  & 81  & 70  & 226 \\
+			KIRP                   & 26  & 53  & 45  & 124 \\
+			LIHC                   & 139 & 76  & 49  & 264 \\
+			READ                   & 22  & 36  & 15  & 73  \\
+			SKCM                   & 71  & 74  & 73  & 218 \\
+			UCEC                   & 88  & 108 & 286 & 482 \\
+			\hline
+			\hline
+		\end{tabular}
+	\end{center}
+\end{table}
+
 \section{\Gls{bmi}}
@@
 
-Where missing, the \gls{bmi} of the samples were calculated from the clinical data using \cref{eq:bmicalc}.
+Where missing, \cref{eq:bmicalc} was used to calculate the \gls{bmi} of the samples in all of the cancer data with height and weight data available (\cref{sec:data}).
 
@@
 
-Samples were classified based on the \gls{who} definition, as shown in \cref{tab:whobmiclass}.
+Samples were classified as normal weight (\gls{bmi} \textless{} 25), overweight (25 $\leq$ \gls{bmi} \textless{} 30) or obese (\gls{bmi} $\geq$ 30) based on the \gls{who} definition (\cref{tab:whobmiclass}).
+
 \begin{table}[htb]
@@
 
-\section{Publicly available cancer data}
-\label{sec:data}
-
-The raw microarray data from \citet{Creighton2012}, \citet{Fuentes-Mattei2014} and \citet{Gatza2010a}  studies were downloaded from the \gls{geo} website.
-\gls{rnaseq} and clinical data of multiple different cancer types were downloaded from free and publicly available sources such as \gls{icgc} and \gls{tcga}, respectively.
-\\
-
-\noindent
-The raw Affymetrix HGU\-133A microarray gene expression data files from the \citet{Creighton2012} study were downloaded from the \gls{geo} database (\gls{geo} accession ID: GSE24185).
-Clinical data of the samples (age, ethnicity, tumour grade, menopause status, \gls{bmi}, \gls{er} status, \gls{pr} status, \gls{her2} status, and \gls{ln} status) were obtained from the supplementary table 1 from \citet{Creighton2012} paper.
-799 obesity associated gene probes identified in the \citet{Creighton2012} study were obtained from the supplementary data file 1 from \citet{Creighton2012} paper.
-
-The raw Affymetrix HGU\-133A microarray gene expression data files from the  \citet{Fuentes-Mattei2014} study were downloaded from the \gls{geo} database (\gls{geo} accession ID: GSE\-20194).
-Clinical data for the samples (age, ethnicity, tumour grade, \gls{er}/\gls{pr}/\gls{her2} statuses, and treatments used) in this study was also downloaded from the \gls{geo} database (same \gls{geo} accession ID).
-130 obesity associated gene probes identified by \citet{Fuentes-Mattei2014} were taken from the supplementary table 3 from their paper.
-
-The raw microarray gene expression data files from the \citet{Gatza2010a} study were downloaded from the \gls{geo} database (\gls{geo} accession ID: GSE1456, GSE\-1561, GSE2034, GSE3494, GSE4922, and GSE6596).
-Only the Affymetrix HGU\-133A microarray samples were included in this project, as other microarray data were analysed using the Affymetrix HGU-133A platform.
-Clinical data for the samples in \citet{Gatza2010a} study was not available, as these samples were a combination of many different datasets.
-(Find out where the 18 pathway signatures were downloaded from)
-
-%TODO: Find Cris's paper
-
-The raw microarray gene expression data files from the (Cris' paper) study were downloaded from the \gls{geo} database (\gls{geo} accession ID: GSE36771).
-Clinical data for the samples (age, ethnicity, tumour grade, breast cancer subtype, \gls{er}/\gls{pr} statuses, \gls{ln} status, \gls{bmi} and treatments used) in this study was taken from (Cris' paper).
-
-All of the microarray data associate the genes with the corresponding gene probe IDs, and therefore these IDs had to be converted into their corresponding gene symbols.
-The gene probe IDs in the raw data were converted into their corresponding gene symbols using the \textit{hgu133a.db} package in \gls{r} \citep{hgu133}.
-Since multiple gene probes matches back to a single gene of interest in a microarray chip, there were conflicting expression data for some of the genes after the conversion of the gene probe IDs into gene symbols.
-For the gene symbols that had multiple data entries, a single data was chosen to represent the gene symbol by using the \texttt{collapseRows} function in the \textit{WGCNA} package in \gls{r} \citep{Langfelder2008}.
-Likewise, any obesity associated or pathway associated gene probes were converted into gene symbols.
-\\
-
-% TODO: Overall summary table of the samples from each data set at the end of this section (?).
-
-\noindent
-The clinical data for all available cancer types (33 types in total) were downloaded from \gls{tcga} database (last accessed 1 April 2015) and were checked for both the height and weight data for each sample.
-Any cancer type with no height and/or weight data of the samples were excluded from the project, as no \gls{bmi} information can be obtained without these data.
-Out of these 33 cancer types with clinical data, 14 cancer types had both height and weight data.
-However, only 8 cancer types out of these 14 types had \gls{rnaseq} data available from the \gls{icgc} database (last accessed 7 September 2015), so only those 8 cancer types were downloaded and used in this project.
-The selected cancer types were: \gls{blca}, \gls{cesc}, \gls{coad}, \gls{kirp}, \gls{lihc}, \gls{read}, \gls{skcm}, \gls{ucec}.
-
-The raw \gls{rnaseq} data from \gls{icgc} database were formatted in a way that the count data of all the genes were listed for one sample, then the count data for all the genes for the next sample, and so on.
-This data format was highly inconvenient for later analyses so the data were reformatted into gene by sample matrix using the \textit{dplyr} package in \gls{r}.
-
-Another problem with the data was the sample ID in the \gls{rnaseq} data.
-Though similar to the \gls{tcga} sample IDs, the \gls{icgc} IDs in the \gls{rnaseq} data had extra identification code in the sample names.
-To associate each sample in the raw \gls{rnaseq} data from the \gls{icgc} database with the correct samples in the clinical data from the \gls{tcga} database, the \gls{icgc} IDs were stripped so they matched the \gls{tcga} IDs.
-After the IDs were stripped, all of the samples were checked to see if there were any duplicates in either the \gls{icgc} \gls{rnaseq} data or \gls{tcga} clinical data.
-Where there was a duplicate in the sample ID, a single sample was chosen to represent that particular ID by using the \texttt{collapseRows} function in the \textit{WGCNA} package in \gls{r} \citep{Langfelder2008}.
-
-Since some of the samples did not have either height or weight data, or both in the clinical data, these samples were removed from the analyses.
-In each of the eight cancer types, the \gls{tcga} sample ID in the clinical data were cross-checked with the sample ID in the \gls{rnaseq} data, and vice versa.
-Any sample that did not have either clinical information or \gls{rnaseq} data were removed from the analyses.
-This ensured that both clinical and \gls{rnaseq} data were available for all of the samples that were included in the study.
-See \cref{tab:samplesize} for a summary of the number of samples included in the analyses for each cancer type.
-
-%TODO: May need more details on the sample (bmi status, etc.)
-\begin{table}[h]
-	\caption{Summary of the total number of samples included in the analyses for each cancer type}
-	\label{tab:samplesize}
-	\begin{center}
-		\begin{tabular}{lc}
-			Cancer Type   & Number of samples \\
-			\hline
-			\rule{0pt}{2.25ex}BLCA & 261   \\
-			CESC                   & 224   \\
-			COAD                   & 226   \\
-			KIRP                   & 124   \\
-			LIHC                   & 264   \\
-			READ                   & 73    \\
-			SKCM                   & 218   \\
-			UCEC                   & 482   \\
-			\hline
-			\hline
-		\end{tabular}
-	\end{center}
-\end{table}
-
 \section{Data processing}
@@
 Any experimental procedure is prone to errors due to differences in experimental conditions, machinery used to measure signals and technical procedures in different laboratories, just to name a few.
-In order to remove these experimental noises and focus on the true biological signals in the raw data, the data must be normalised one way or another.
+In order to remove these experimental noises and focus on the true biological signals in the raw data.
 
@@
 
-In microarray experiments, there are two types of probes present on the microarray chips: \gls{pm} and \gls{mm} probes \citep{Irizarry2003}.
+In Affymetrix HGU133A microarray experiments, there are two types of probes present on the microarray chips: \gls{pm} and \gls{mm} probes \citep{Irizarry2003}.
 As the name suggests, \gls{pm} probes represent the probes that should perfectly match the gene of interest, whereas \gls{mm} probes have there 13th base pair intentionally altered to measure non-specific binding of the gene \citep{Irizarry2003}.
@@
 \Gls{mas} uses a difference-based method where it subtracts a value derived from \gls{mm} from \gls{pm}, but this approach may introduce additional noise and/or errors in some cases \citep{Irizarry2003}.
-\Gls{rma} method is based on the observation that \gls{pm} is a mixture of background and true signal, and uses mathematical models to estimate the expression while correcting for background signals from \gls{pm} only \citep{Irizarry2003}.
+The \gls{rma} method is based on the observation that \gls{pm} is a mixture of background and true signal, and uses mathematical models to estimate the expression while correcting for background signals from \gls{pm} only \citep{Irizarry2003}.
 In fact, the \gls{rma} method showed better identification of the true signals than the other methods, including \Gls{mas} method \citep{Irizarry2003}.
@@
 The raw microarray data were normalised using both \gls{rma} and \gls{mas} methods, each done separately on a copy of the data.
-The reason why \gls{mas} method was used as well as the \gls{rma} method was becuase the \gls{mas} normalisation method was used in \citet{Gatza2010a} study.
+The reason why the \gls{mas} method was used as well as the \gls{rma} method was becuase the \gls{mas} normalisation method was used in the \citet{Gatza2010a} study.
 Though the normalisation methods were not specified in the \citet{Creighton2012} and \citet{Fuentes-Mattei2014} studies, to allow for data comparison between different datasets and accurate result validation of some of the studies, both \gls{rma} and \gls{mas} methods were used to normalise the microarray data.
@@
 
-\gls{rnaseq} data are fundamentally different to the microarray data, as \gls{rnaseq} data is a count data of all the sequences, while microarray data is a continuous data of the intensity of the matching probes.
-This means that \gls{rnaseq} data must be processed in a different manner as the microarray data.
-However, by analysing \gls{rnaseq} data as a count data, it limits range of statistical tools and types of analyses that can be done on the data, as many tools were designed for normally distributed data \citep{Law2014}.
+\gls{rnaseq} data are fundamentally different to microarray data, as \gls{rnaseq} data is a count data of transcript fragments, while microarray data is a continuous data based on the intensity of the probe pairs.
+This means that \gls{rnaseq} data must be processed in a different manner to microarray data.
+However, by analysing \gls{rnaseq} data as a count data, it limits the range of statistical tools and types of analyses that can be done on the data, as many tools are designed for normally distributed data, like the light intensity measures from the microarray data \citep{Law2014}.
 
-To overcome this limitation, \citet{Law2014} developed a method called ``variance modelling at the observational level'', or voom, that allows any \gls{rnaseq} data to be used in any existing statistical analysis pipeline that is precision weight aware, including pipelines used for microarray data analyses.
+To overcome this limitation, \citet{Law2014} developed a method called ``variance modelling at the observational level'', or ``voom'', that allows any \gls{rnaseq} data to be used in any existing statistical analysis pipeline that is precision weight aware, including pipelines used for microarray data analyses.
+% TODO: check the meaning of precision weight in the above sentence
 In brief, voom first constructs a standard deviation trend from the logged \gls{cpm} value of the genes (experimental design, treatment conditions and other factors are taken into account).
-This trend is then used to interpolate the standard deviation of the observation based on its predicted count size, and the inverse square of the predicted standard deviation is used as the weight  for that observations.
-The weights of the observations and the logged \gls{cpm} can then be used in other statistical pipelines that allows the input of quantitative weights \citep{Law2014}.
+This trend is then used to interpolate the standard deviation of the observation based on its predicted count size, and the inverse square of the predicted standard deviation is used as the weight  for that observation.
+The weights of the observations and the logged \gls{cpm} can then be used in other statistical pipelines that allow the input of quantitative weights \citep{Law2014}.
 
-The raw \gls{rnaseq} data were normalised in two different ways, depending on the analysis.
-For gene expression analyses, voom normalisation method from the \textit{limma} package in \gls{r} was used to normalise the data \citep{Ritchie2015}.
-For the purposes of data visualisation or application of metagene transformation matrices on the \gls{rnaseq} data, the raw data had 1 added (to prevent logging of 0) and then logged to the base of 10.
+The raw \gls{rnaseq} data were normalised in two different ways depending on the analysis.
+For gene expression analyses (\cref{sec:gene_expression_analysis}) and pathway enrichment analyses (\cref{sec:pathway_enrichment_analysis}), voom normalisation from the \textit{limma} package in \gls{r} was used to normalise the data \citep{Ritchie2015}.
+For the purposes of data visualisation (\cref{sec:plot_creation}) or application of the metagene transformation matrices (\cref{sub:svd}) on the \gls{rnaseq} data, the raw data had 1 added (to prevent logging of 0) and then logged to the base of 10.
 
@@
 For metagene creation and data visualisation that used heatmaps, the data were standardised so that each gene in the data had a \gls{m} of 0 and a \gls{sd} of 1.
-Since the expression levels of the genes could vary significantly (some may have very low expression, whereas another may have very high expression), the direct comparison of the raw expression values between different genes was not feasible.
-Standardisation of each gene allowed the expression levels to be within acceptable range or scale, thus allowed for better visualisation with heatmaps and comparison between different genes was made possible.
+Since the expression levels of the genes could vary substantially (some may have very low expression, whereas another may have very high expression), the direct comparison of the raw expression values between different genes was not feasible.
+Standardisation of each gene allowed the expression levels to be on the same scale, thus allowing better visualisation with heatmaps and comparison between different genes was made possible.
 
@@
 
-In order to analyse the data without the effect of certain clinical variables on the data, the residual data of the data was used in some analyses.
+In some analyses, the Creighton \textit{et al.} data was adjusted for clinical variables to remove any confounding effects from the variables.
 For example, a gene from the raw data might have a strong association with \gls{bmi}, but it is possible that other clinical variables such as \gls{er} status and/or tumour grades are also associated with the gene.
-To focus only on the association with certain variables, the effect of those variables that may affect the analyses must be removed from the data.
+To prevent such confounding effect, all of the clinical variables except those that are of interest were adjusted with the use of linear models.
 
-To do this, a linear model was created from the data with all of the clinical variables that had to be controlled for; in other words, all of the clincial variables apart from the variables of interest were included in the linear model.
-Once a linear model was fitted to the data, the remaining data, or the ``residual data'', represented the data that had been corrected for all of the unwanted variables.
-This data was used in some analyses that required focus on certain variables without the other variables affecting the result.
+A linear model was created from the Creighton \textit{et al.} data with age, ethnicity, menopause status, tumour grade, hormone (\gls{er}, \gls{pr}, and \gls{her2}) statuses  and \gls{ln} status included in the model.
+In addition to this model, another linear model was constructed with the Creighton \textit{et al.} data with only the Caucasian patients included in the data (same clinical variables, except, ethnicity were used in the model); this was to completely remove the effect of ethnicity.
+Once a linear model was fitted to the Creighton \textit{et al.} data, the remaining data, or the ``residual data'', represented the data that had been corrected for all of the unwanted variables.
+This residual data was used in the analyses that required focus on certain variables without the other variables affecting the result.
 
@@
 Since the data used in \citet{Gatza2010a} study was a combination of multiple microarray data from various studies, the batch effect had to be corrected before any analysis was carried out.
-Each microarray dataset was normalised separately then combined together into a single dataset, and the batch effect was corrected with the \texttt{ComBat} function (an implementation of the batch correcting method by \citet{Johnson2007}) from the \textit{sva} package in \gls{r} \citep{Leek2012}.
+Each microarray dataset was normalised with either \gls{rma} or \Gls{mas} separately then combined together into a single dataset, and the batch effect was corrected with the \texttt{ComBat} function (an implementation of the batch correcting method by \citet{Johnson2007}) from the \textit{sva} package in \gls{r} \citep{Leek2012}.
+
+\gls{icgc} cancer data were also corrected for batch effect for the pathway enrichment analysis (\cref{sec:pathway_enrichment_analysis}).
+Two combined data sets were created with the \gls{icgc} data.
+The first combined data set was created by voom normalising each cancer data set separately, combine all of the normalised cancer data sets into one, then applied batch correction to the data set (\cref{sub:batch_correction}).
+The second combined data set was created by first combining the cancer data sets into one, then the combined data was voom normalised, and then the batch effect was corrected to obtain the final data set.
 
@@
 In some cases, an analysis was repeated a number of times to provide statistical evidence that the observed results did not occur by chance.
-All of the simulated results must be derived from a randomised data, and by comparing the true results with the results that occured by chance, simulation analysis provides statistical evidence for the true results.
+All of the simulated results must be derived from a randomised data, and by comparing the observed results with the results that occured by chance, simulation analysis provides statistical evidence for the observed results.
 When a simulation was required, none of the values in the original data were modified but the values in the original data were assigned randomly to different samples; in other words, the samples were shuffled.
@@
 
-\textit{limma} is a package that contains variety of tools to analyse microarray data using linear models-based methods, developed by \citet{Ritchie2015}.
-For gene expression analysis, \textit{limma} package fits a linear model in a genewise manner and produce test statistics that allow the assessment of whether the gene is differentially expressed or not \citep{Ritchie2015}.
+\textit{limma} is a package that contains variety of tools to analyse gene expression data using linear models-based methods, developed by \citet{Ritchie2015}.
+For gene expression analysis, \textit{limma} package fits a linear model in a genewise manner and produce test statistics that allow the assessment of whether the gene is differentially expressed or not between certain groups of samples \citep{Ritchie2015}.
+In this project, the gene expression analysis was carried out to identify \glspl{deg} in the samples that were obese compared to the samples that were non-obese group.
 
 Before the data was analysed using the \textit{limma} analysis pipeline, it was normalised as described in \cref{sub:data_normalisation}: \nameref{sub:data_normalisation}.
-In addition to this, a design matrix that describes the experimental design was created from the clinical data.
+In addition to this, a design matrix that describes the experimental design (obese versus non-obese) was created from the clinical data.
 To create the design matrix, the samples were divided into two groups, the obese group and the non-obese group, and the constructed group information was used in the \texttt{model.matrix} function (in \textit{limma} package) to form the design matrix.
 
-\texttt{lmFit} function (\textit{limma}) was used to fit a linear model to the normalised data, using the experiental design information from the design matrix.
+The \texttt{lmFit} function (\textit{limma}) was used to fit a linear model to the normalised data, using the experiental design information from the design matrix.
 The output of this function was used in the \texttt{eBayes} function (\textit{limma}) to identify the \glspl{deg} from the data.
-In \texttt{eBayes} function, statistical parameters are estimated from the data and these parameters are used in the empirical Bayesian approach to calculate the summary statistics used for the ranking and identification of \glspl{deg} \citep{Smyth2004}.
+In the \texttt{eBayes} function, statistical parameters are estimated from the data and these parameters are used in the empirical Bayes approach to calculate the summary statistics used for the ranking and identification of \glspl{deg} \citep{Smyth2004}.
 
-The summary statistics from the \texttt{eBayes} function can be displayed with the \texttt{topTable} function (\textit{limma}) and includes: the estimate of the fold change of the gene expression in log$_2$, average gene expression level in log$_2$, moderated \textit{t}-statistic, raw p-value of the gene, multiple hypothesis testing adjusted p-value, and \textit{B}-statistic.
+The summary statistics from the \texttt{eBayes} function can be displayed with the \texttt{topTable} function (\textit{limma}) and includes the estimate of the fold change of the gene expression in log$_2$, average gene expression level in log$_2$, moderated \textit{t}-statistic, raw p-value of the gene, multiple hypothesis testing adjusted p-value, and \textit{B}-statistic.
 The first value shows the estimate of the log$_2$ fold change of the gene expression compared with the reference group, so this represents the log$_2$ fold change of the gene expression in the obese samples relative to the non-obese group.
-Second value presents the log$_2$ value of the average expression of the gene across all of the arrays/samples.
-Moderated \textit{t}-statistic is the same as normal \textit{t}-statistic, but its standard error has been adjusted by a simple Bayesian method \citep{Smyth2005}.
+The second value presents the log$_2$ value of the average expression of the gene across all of the arrays/samples.
+%TODO: more details on simple Bayesian method
+The moderated \textit{t}-statistic is the same as normal \textit{t}-statistic, but its standard error has been adjusted by a simple Bayesian method \citep{Smyth2005}.
 Raw p-value and adjusted p-value represent the p-value of the gene before and after it has been corrected for multiple hypothesis testing, respectively.
@@
 From these summary statistics, the most likely \glspl{deg} were chosen from the list of significant genes by setting the threshold of the p-value to be either less than 1\% or 5\%.
-In the case where there were more than 1000 \glspl{deg}, the top 799 genes were picked from the list, as this was the number of genes found in the \citet{Creighton2012} study; otherwise, as many significant genes identified were taken from the list.
+In the case where there were more than 1000 \glspl{deg}, the top 799 probe sets were picked from the list, as this was the number of genes found in the \citet{Creighton2012} study; otherwise, as many significant genes identified were taken from the list.
 
@@
 
-Since \gls{fdr} method controls \gls{type1} and have greater statistical \gls{power} than the \gls{fwer} methods, \gls{fdr} method was used to identify the \glspl{deg} for gene expression analysis.
+Since \gls{fdr} method controls \gls{type1} in a less conservative manner than the \gls{fwer} methods, the \gls{fdr} method was used to identify the \glspl{deg} for gene expression analyses.
 
@@
 
-Identification of \glspl{deg} provides great deal of information about the effect of treatment and provides a list of genes that may have a role in that particular experimental setting.
-With that said, it is often difficult for the investigators to provide a plausible biological explanation with just a long list of \glspl{deg}, as the list lacks the link between the genes with the biological cause \citep{Khatri2012}.
-Therefore, given a list of \glspl{deg}, it is crucial for any researcher to undertake pathway enrichment analysis in order to provide useful insights into the underlying biological mechanisms the \glspl{deg} may be involved in.
-
-\subsection{Competitive and self-contained tests}
-\label{sub:competitive_and_self_contained_tests}
-
-In general, there are two major classes of methods for pathway enrichment analysis: self-contained or competitive tests \citep{Goeman2007}.
-The main difference between the two classes is the definition of the null hypothesis.
-Letting $G$ be the gene set of interest and $G^c$ its compliment, then in self-contained tests, the null hypothesis is formulated as follows:
-\begin{quote}
-	\textit{$H_0^{\textrm{self}}$: No genes in G are differentially expressed. }
-\end{quote}
-whereas in competitive tests, the null hypothesis is defined as:
-\begin{quote}
-	\textit{$H_0^{\textrm{comp}}$: The genes in G are at most as often differentially expressed as the genes in $G^c$.
-	}
-\end{quote}
-These hypotheses show that self-contained tests care only about the genes defined in the gene set, but competitive tests examine the genes in the defined set as well as the genes not present in the set \citep{Wu2012}.
-Due to the fact that self-contained tests do not take into account for the genes not defined in the gene set, self-contained tests are more restrictive compared to competitive tests.
-The restrictive property of self-contained tests give it greater power, as they are able to reject the null hypothesis at a higher accuracy for more gene sets than the competitive tests \citep{Goeman2007}.
-As \citet{Goeman2007} stated in their paper, ``the competitive types of test can be said to voluntarily relinquish some power in order to make a stronger statement''.
-
-Taken together, self-contained tests are more appropriate in assessing whether a biological process is significantly involved in an expreiment, and in contrast, competitive tests are better for selecting the most relevant biological processes from those that are not \citep{Wu2012}.
+Identification of \glspl{deg} provides a list of genes that may have a role in a particular experimental setting.
+With that said, it is often difficult for the investigators to provide a plausible biological explanation with just a long list of \glspl{deg}, as the list lacks the link between the genes and the biological cause \citep{Khatri2012}.
+Therefore, given a list of \glspl{deg}, it is important for any researcher to undertake pathway enrichment analysis in order to obtain useful insights into the underlying biological mechanisms the \glspl{deg} may be involved in.
 
@@
 Though the approaches differ between the two, both \acrfull{ora} and \acrfull{fcs} are used to identify whether certain pathways are enriched, given a list of \glspl{deg}.
+
+\subsubsection{\Gls{ora}}
+\label{ssub:ora}
+
 The common approach taken by \gls{ora} is to count the \glspl{deg} that are part of a biological pathway and perform a statistical test to decide whether the pathway is over- or under-represented in the list of genes \citep{Khatri2012}.
 Statistical test used in \gls{ora} includes $\chi^2$-test, hypergeometric test, and binomial test \citep{Khatri2012}.
-Depending on the methods, \gls{ora} uses either self-contained or competitive tests.
 
@@
 Secondly, only the most significant genes are selected for the input list of genes, which means that the almost significant genes are discarded from the analysis and results in a loss of information \citep{Khatri2012}.
-Lastly, by treating each gene individually, the analysis ignores the biological interaction and complexity of the genes with other genes, as well as between pathways \citep{Khatri2012}.
+Lastly, by treating each gene individually, the analysis ignores the biological interaction and complexity of the genes with the other genes, as well as between pathways \citep{Khatri2012}.
 
-In general, \gls{fcs} methods measure gene-level statistics from the given data, then aggregate these statistics into pathway-level statistics, which is then used to calculate the statistical significance of the pathway \citep{Khatri2012}.
-Either self-contained or competitive tests can be used to calculate the statistical significance of the pathways.
+\subsubsection{\gls{fcs}}
+\label{ssub:fcs}
 
-\Gls{fcs} approach addresses some of the limitations presented by the \gls{ora} methods.
-By calculating the statistical significance per pathway, \gls{fcs} approach takes into account of all of the genes involved in the pathways, not just the genes that are differentially expressed.
-Furthermore, \gls{fcs} approach also detect small but consistant and coordinated changes in the expression of the genes, unlike in \gls{ora} where molecular measurements are completely ignored \citep{Khatri2012}.
+In general, the \gls{fcs} methods measure gene-level statistics from the given data, then aggregate these statistics into pathway-level statistics, which is then used to calculate the statistical significance of the pathway \citep{Khatri2012}.
+There are two major classes of methods when calculating the statistical significance: self-contained or competitive tests \citep{Goeman2007}.
+Self-contained tests are more appropriate in assessing whether a biological process is significantly involved in an expreiment, and in contrast, competitive tests are better for selecting the most relevant biological processes from those that are not \citep{Wu2012}.
 
-Though \gls{fcs} approach improves on \gls{ora} approach, there are still some limitations.
-Since the analyses that use \gls{fcs} compare the pathways independently from one another, it ignores the interactive nature of biological pathways.
-Another limitation is that \gls{fcs} methods use the molecular measurements to rank the genes, but they do not consider these changes into further analysis \citep{Khatri2012}.
-For example, if one gene was expressed with a 2-fold change, whereas another gene was expressed with a 20-fold change, the \gls{fcs} method will rank these genes accordingly and disregard the fact that the second gene should be weighted more than the first gene \citep{Khatri2012}.
+The main difference between the two classes is the definition of the null hypothesis.
+Letting $G$ be the gene set of interest and $G^c$ its compliment, then in the self-contained tests, the null hypothesis is formulated as follows:
+\begin{quote}
+	\textit{$H_0^{\textrm{self}}$: No genes in G are differentially expressed. }
+\end{quote}
+whereas in the competitive tests, the null hypothesis is defined as:
+\begin{quote}
+	\textit{$H_0^{\textrm{comp}}$: The genes in G are at most as often differentially expressed as the genes in $G^c$.
+	}
+\end{quote}
+These hypotheses show that the self-contained tests care only about the genes defined in the gene set, but the competitive tests examine the genes in the defined set as well as the genes not present in the set \citep{Wu2012}.
+Due to the fact that self-contained tests do not take into account for the genes not defined in the gene set, self-contained tests are more restrictive compared to competitive tests.
+The restrictive property of the self-contained tests give it greater power, as they are able to reject the null hypothesis at a higher accuracy for more gene sets than the competitive tests \citep{Goeman2007}.
+As \citet{Goeman2007} stated in their paper, ``the competitive types of test can be said to voluntarily relinquish some power in order to make a stronger statement''.
+
+%TODO: add GSEA here (mootha?)
+
+The \Gls{fcs} approach addresses some of the limitations presented by the \gls{ora} methods.
+By calculating the statistical significance per pathway, the \gls{fcs} approach takes into account of all of the genes involved in the pathways, and not just the genes that are differentially expressed.
+Furthermore, the \gls{fcs} approach also detect small but consistant and coordinated changes in the expression of the genes, unlike in the \gls{ora} where molecular measurements are completely ignored \citep{Khatri2012}.
+
+Though the \gls{fcs} approach improves on the \gls{ora} approach, there are still some limitations.
+Since the analyses that use \gls{fcs} compare the pathways independently from one another, it ignores the interactive nature of the biological pathways.
+Another limitation is that the \gls{fcs} methods use the molecular measurements to rank the genes, but they do not consider these changes into further analysis \citep{Khatri2012}.
+For example, if one gene was expressed with a 2-fold change, whereas another gene was expressed with a 20-fold change, the \gls{fcs} method will rank these genes accordingly (first, second, and so on) and disregard the fact that the second gene should be weighted ten times as more than the first gene \citep{Khatri2012}.
 
@@
 \Acrfull{camera} is a competitive \gls{fcs}-based method developed by \citet{Wu2012}, implemented as the \texttt{camera} function in the \textit{limma} package \citep{Ritchie2015}.
-Briefly, \gls{camera} fits a linear model in a genewise manner and calculates the genewise test statistics using the log\gls{fc} between the conditions.
+Briefly, \gls{camera} fits a linear model in a genewise manner and calculates the genewise test statistics using the log\gls{fc} between the conditions (for example, obese and non-obese samples).
 These genewise test statistics are used in the \gls{wmw} rank sum test to question whether a pathway is significantly enriched in the data \citep{Wu2012}.
@@
 As a consequence, the \gls{type1} rate is increased in these methods.
-\Gls{camera} accounts for this correlation by estimating the \gls{vif} from the genewise correlation and the number of genes in the gene set \citep{Wu2012}.
+However, \Gls{camera} accounts for this correlation by estimating the \gls{vif} from the genewise correlation and the number of genes in the gene set \citep{Wu2012}.
 
-\gls{camera} was used in this project to identify the pathways that are enriched in samples from obese patients compared with the samples from non-obese patients, as it provides a competitive \gls{fcs}-based method.
+\gls{camera} was used in this project to identify the pathways that were enriched in the samples from obese patients compared with the samples from non-obese patients, as \gls{camera} provided a competitive \gls{fcs}-based method.
 
@@
 
-% Due to the sheer amount of terms and definitions available from the data, \gls{go} was chosen as the database for pathway reference.
-% The \textit{GO.db} package was used to load the \gls{go} database into \gls{r} \citep{Carlson2016}.
-% Only the human related \gls{go} terms were used in pathway enrichment analysis, and these terms were manipulated in a form which the \texttt{camera} function was able to recognise.
+The \textit{GO.db} package was used to load the \gls{go} database into \gls{r} \citep{Carlson2016}.
+Only the human related \gls{go} terms were used in pathway enrichment analysis, and these terms were organised into a format so that the pathways were able to be queried by using the gene symbol.
 
@@
 
+\gls{kegg} is a genomic database constructed by \citet{Kanehisa2000} that store information about genes, pathways and ligands for a variety of species.
+The \gls{kegg} database also curates other information such as the approved drugs in US and Japan and genes related to diseases, making it a great tool for exploring the biological significance of genes and pathways \citep{Kanehisa2008}.
+
+The \textit{KEGG.db} package was used to load the \gls{kegg} database into \gls{r} \citep{kegg}.
+Only the human related \gls{kegg} terms were used in pathway enrichment analysis, and these terms were organised into a format so that the pathways were able to be queried by using the gene symbol.
+
 \subsubsection{Reactome database}
 \label{ssub:Reactome_database}
+
+The Reactome database is a ``curated, peer-reviewed resource of humana biological processes'' \citep{Joshi2005}.
+Reactome database is constructed base on the reactions that have direct evidence from the literature, which are connected into a higher order pathway structures \citep{Joshi2005}.
+
+The \textit{reactome.db} package was used to load the Reactome database into \gls{r} \citep{reactome}.
+Only the human related Reactome terms were used in pathway enrichment analysis, and these terms were organised into a format so that the pathways were able to be queried by using the gene symbol.
 
@@
 
-In order to determine whether a set of genes are up or down regulated by a sample, some sort of score based on these genes must be calculated.
-\Gls{svd} is a mathematical method that splits a matrix into several matrices that represent the original matrix \citep{Golub1970}.
+In order to determine whether a set of genes is up or down regulated by a sample, some sort of score based on these genes must be calculated.
+\Gls{svd} is a mathematical method that splits a data matrix into several smaller matrices \citep{Golub1970}.
+These matrices, when multiplied together, are able to recreate the original data matrix.
 With \gls{svd}, a matrix $X$ of size $n$ genes by $m$ arrays (or samples) can be represented as:
@@
 This allows direct comparison of the expressions of multiple different genes in different arrays or experiments \citep{Alter2000}.
-Furthermore, this summary gene, or metagene, can be used to sort the samples and provide a meaningful grouping of the data which may help understand the underlying biology \citep{Alter2000}.
+Furthermore, this summary gene, or ``metagene'', can be used to sort the samples and provide a meaningful grouping of the data which may help understand the underlying biology \citep{Alter2000}.
 
-With respect to this project, \gls{svd} was used to assess whether the metagene created from various obesity or pathway associated genes were significantly associated with certain clinical variables, such as \gls{bmi}.
+With respect to this project, \gls{svd} was used to assess whether the metagenes created from various obesity or pathway associated genes were significantly associated with certain clinical variables, such as \gls{bmi}.
 To do this, the normalised original expression data was reduced in size so that it only included the genes of interest (either obesity or pathway associated) and the samples, then \texttt{svd} function from the \textit{base} package in \gls{r} was used to apply \gls{svd} to the matrix.
-The first principal component was taken from the $V$ matrix and was used as the metagene scores for the samples in the data.
+The first principal component was taken from the $V$ matrix and was used as the metagene scores for each samples in the data set.
 
@@
 The reason why the transformation matrix is used instead of independently applying \gls{svd} to the other data set is because the signature, and therefore the metagene, from a single data set is dependent on the signal within its data set.
-Therefore, the metagenes created independently from different data sets using \gls{svd} does not have the same weighting as the metagene from the original data set, and consequently does not provide a fair comparison of the metagenes across different data sets.
+Therefore, the metagenes created independently from different data sets using \gls{svd} do not have the same weighting as the metagene from the original data set, and consequently does not provide a fair comparison of the metagenes across different data sets.
 For this reason, the transformation matrix for a given set of genes was created in the data set in which the genes were first identified in, and then transformed into other data sets to obatain the metagenes.
+
+\subsection{Ranking of the metagene scores}
+\label{sub:ranking_of_the_metagene_scores}
+
+Two different approaches were taken in order to rank the metagene scores of the samples.
+The first approach was to rank the metagene scores with the \texttt{rank} function from the \textit{base} package in \gls{r}, then divided the ranked scores with the total number of samples in the data set to obtain a ranked metagene scores between 0 and 1.
+In most analyses, this approach was used to assess the association of the metagene with the sample gene expressions and clinical variables.
+
+An alternative ranking method was used in the Gatza \textit{et al.} data to compare the different ranking methods.
+In this method, the sample metagene scores were ranked based on the probit function.
+To do this, the sample metagene scores were scaled and centered using the \texttt{scale} function from the \textit{base} pakage in \gls{r}, then the \texttt{pnorm} function from the \textit{stats} package was used rank the metagene scores.
+Unless stated otherwise, all of the metagenes in this project were ranked based on the number of samples and not with the probit method.
 
@@
 
-% TODO: add example plots for metagene direction
+One thing to consider when analysing data with metagenes is the direction of the metagene.
+When \gls{svd} creates the metagene from a given data set, it does not consider the direction of the signal, but only the magnitude of the signal in the data.
+Therefore, the metagene created by \gls{svd} can be either positively or negatively correlated to the phenotype that the set of genes reflect.
+For example, a metagene created from a set of genes that are associated with the \gls{er} pathway may have higher metagene scores in the samples with low expression of the \gls{er} gene, and low metagene scores in the samples with high expression of the \gls{er} gene (\cref{fig:methods/meta_dir}, left heatmap).
+In this case, the direction of the generated metagene must be corrected in order to reflect the phenotype with the metagene scores (low metagene score with low gene expression and high metagene score with high gene expression; \cref{fig:methods/meta_dir}, right heatmap).
 
-One thing to consider when analysing data with metagenes is the direction of the metagene.
-When \gls{svd} creates the metagene from the given data, it does not consider the direction of the signal, but only the magnitude of the signal in the data.
-Therefore, the metagene created by \gls{svd} can be either positively or negatively correlated to the phenotype that the set of genes reflect.
-For example, a metagene created from a set of genes that are associated with \gls{er} status may have higher metagene scores in the \gls{er} negative samples and low metagene scores in the \gls{er} positive samples.
-In this case, the generated metagene must be flipped in order to reflect the phenotype with the metagene scores (low metagene score with low phenotype and high metagene score with high phenotype).
+\begin{figure}[htpb]
+	\centering
+	\includegraphics[width=0.45\linewidth,page=5]{methods/gtrmastdrank_incorrect}
+	\hfill
+	\includegraphics[width=0.45\linewidth,page=5]{methods/gtrmastdrank_corrected}
+	\caption[Example heatmaps showing the direction of the uncorrected and corrected \acrshort{er} pathway metagene with the \acrshort{er} gene expression in the \acrshort{rma}-normalised Gatza \textit{et al.} data.]{Heatmaps showing the direction of the unflipped (left) and flipped (right) \gls{er} pathway metagene with the \gls{er} gene expression in the \gls{rma}-normalised Gatza \textit{et al.} data.
+	Level of expression is represented in the top right histogram, where low and high gene expression are colour-coded with blue and red, respectively.
+	Each row of the heatmap represents a gene from the \gls{er} genetic signature, and each column of the heatmap represents a sample from the Gatza \textit{et al.} data.
+	The \gls{er} gene expressions of the samples are shown as a separate row at top of the heatmap with the sample \gls{er} metagene scores below the \gls{er} gene expression, and the tree diagram of the heirarchical clustering of the genes is shown on the left of the heatmap.
+	}
+	\label{fig:methods/meta_dir}
+\end{figure}
 
@@
 Checking the direction of the pathway metagenes from \citet{Gatza2010a} study is not as simple as the obesity associated metagenes, as the experimental design that identify which samples were treated to generate the pathway signature was not available.
-Due to this, it was not possible to validate whether the generated pathway metagene was in the right direction.
-As an alternative, the pathway metagenes were visually compared with the gene that represent the pathway using a heatmat, as the expression of this gene would most likely be affected the most by the pathway signature.
-For example, the Akt pathway metagene was compared with the AKT1 gene expression, and if the metagene was not in the same direction as the AKT1 gene expression, the metagene was flipped.
+Due to this, it was not possible to validate whether the generated pathway metagene was in the ``right'' direction.
+As an alternative, the pathway metagenes were visually compared with the gene that represent the pathway using a heatmap, as the expression of this gene would likely be affected the most by the pathway signature.
+For example, the \gls{er} pathway metagene was compared with \textit{ESR1} gene expression, and if the metagene was not in the same direction as \textit{ESR1} gene expression, the direction of the metagene was corrected (\cref{fig:methods/meta_dir}).
 \cref{tab:metagene_direction} summarises the genes used to check the direction of the pathway metagenes.
@@
 	\begin{center}
-		\begin{tabular}{lc}
-			Pathway & Representative gene\\
+		\begin{tabular}{lcl}
+			Pathway & Representative gene & Pathway description\\
 			\hline
-			\rule{0pt}{2.25ex}Akt & AKT1\\
-			\gls{bcat}            & CTNNB1\\
-			E2F1                  & E2F1\\
-			\Gls{egfr}            & EGFR\\
-			\Gls{er}              & ESR1\\
-			\Gls{her2}            & ERBB2\\
-			\Gls{ifna}            & IFNA1\\
-			\Gls{ifny}            & IFNG\\
-			Myc                   & MYC\\
-			p53                   & TP53\\
-			p63                   & TP63\\
-			\Gls{pi3k}            & PIK3CA\\
-			\Gls{pr}              & PGR\\
-			Ras                   & HRAS\\
-			\Gls{stat3}           & STAT3\\
-			Src                   & SRC\\
-			\Gls{tgfb}            & TGFB1\\
-			\Gls{tnfa}            & TNF\\
+			\rule{0pt}{2.25ex}Akt & AKT1   & \\
+			\gls{bcat}            & CTNNB1 & \\
+			E2F1                  & E2F1   & \\
+			\Gls{egfr}            & EGFR   & \\
+			\Gls{er}              & ESR1   & \\
+			\Gls{her2}            & ERBB2  & \\
+			\Gls{ifna}            & IFNA1  & \\
+			\Gls{ifny}            & IFNG   & \\
+			Myc                   & MYC    & \\
+			p53                   & TP53   & \\
+			p63                   & TP63   & \\
+			\Gls{pi3k}            & PIK3CA & \\
+			\Gls{pr}              & PGR    & \\
+			Ras                   & HRAS   & \\
+			\Gls{stat3}           & STAT3  & \\
+			Src                   & SRC    & \\
+			\Gls{tgfb}            & TGFB1  & \\
+			\Gls{tnfa}            & TNF    & \\
 			\hline
@@
 Lastly, the clustering of the pathway metagenes with the other pathway metagenes were also considered.
-It was known from \citet{Gatza2010a} study that some pathways clustered and grouped together with specific pathways, so the metagenes were flipped so that the grouping of the pathways matched the results from their study.
+It was known from \citet{Gatza2010a} study that some pathways clustered and grouped together with specific pathways, so the direction of some of the metagenes were altered so that the grouping of the pathways matched the results of their study.
 
@@
 
-To determine whether the obesity metagene was associated with any of the pathway metagenes, a linear model was constructed with some clinical variables and pathway metagenes, and then this model was used to predict the obesity metagene.
-This was based on the idea that, if a variable or a pathway metagene was a significant part of the model for the obesity metagene, then the variable or a pathway metagene may be associated with the obesity metagene.
-Furthermore, the model that was created based on the significant variables can be used to predict the obesity metagene, which can be used to compare with the original obesity metagene; if the predicted metagene was similar to the original metagene, then it provides an evidence that the pathway used to construct the model was associated with the obesity metagene.
+To determine whether the obesity metagene was associated with any of the pathway metagenes, a linear model was constructed with some clinical variables and pathway metagenes, which was then used to predict the obesity metagene.
+This was based on the idea that, if a clinical variable or a pathway metagene was a significant part of the model for the obesity metagene, then that clinical variable or pathway metagene may be associated with the obesity metagene.
+Furthermore, the model that was created based on the significant variables can be used to predict the obesity metagene, which can be used to compare the values with the original obesity metagene.
+If the predicted metagene was similar to the original metagene, then it provides an evidence that the pathway used to construct the model was associated with the obesity metagene.
 
-Four linear models were created in (Cris' data -- get reference) to predict the obesity associated metagene from the \citet{Creighton2012} study.
-The first model used the sample \gls{bmi} value or \gls{bmi} status by itself, the second model with sample \gls{bmi} value, \gls{bmi} status and selected pathway associated metagenes, the third model with selected pathway associated metagenes, and the last model was constructed with just the \gls{pr} pathway.
+%TODO: re-write the linear model stuff:
+Four linear models were created in the \gls{nzbc} to predict the obesity associated metagene from the \citet{Creighton2012} study.
+The first model used the sample \gls{bmi} value or \gls{bmi} status by itself, the second model with the sample \gls{bmi} value, \gls{bmi} status and selected pathway associated metagenes, the third model with selected pathway associated metagenes, and the last model was constructed with just the \gls{pr} pathway.
 The selected pathways used in the linear model construction were \gls{bcat}, \gls{er}, \gls{ifna}, \gls{ifny}, Myc and \gls{pr}.
@@
 
-The predicted obesity metagenes that were created in (Cris' data -- get reference) was compared with the original obesity metagene from the \citet{Creighton2012} data set.
+The predicted obesity metagenes that were created in the \gls{nzbc} data set was compared with the original obesity metagene from the \citet{Creighton2012} data set.
 Both Pearson and Spearman correlation was calculated to assess the relationship between the predicted and original metagenes.
@@
 Furthermore, heatmaps are able to cluster the samples and/or genes based on the ``distance'' between the data, or in other words, how closely related the data are.
-Heatmaps were created using the \texttt{heatmap.2} function from the \textit{gplots} package in \gls{r}.
+Heatmaps were created using the \texttt{heatmap.2} function from the \textit{gplots} package in \gls{r} \citep{gplots}.
 
@@
 
+\subsection{Additional colours}
+\label{sub:additional_colours}
+
 % TODO: talk about spearman correlation?

### 2017-03-16 thesis/misc/glossary.tex
Finished obesity metagene direction subsection
Started editing introduction -- tmp commit
Minor edit in intro and methods
Changed all the data set names and removed "'s" - tmp commit
Edited Intro
tmp commit

@@
 	name=Type I error,
-	text=type I error,
-	plural=type I errors,
+	text=Type I error,
+	plural=Type I errors,
 	description={Incorrect rejection of a true null hypothesis; False positive}
@@
 	name=Type II error,
-	text=type II error,
-	plural=type II errors,
+	text=Type II error,
+	plural=Type II errors,
 	description={Incorrect retaining of a false null hypothesis; False negative}
@@
 	text=power,
-	description={Probability that it will reject a false null hypothesis; It is inversely related to the probability of making a type II error}
+	description={Probability that it will reject a false null hypothesis; It is inversely related to the probability of making a Type II error}
 }
@@
 \newacronym{gaba}{GABA}{$\gamma$-aminobutyric acid}
-\newacronym{irs}{IRS}{insulin receptor seubstrate}
+\newacronym{irs}{IRS}{insulin receptor substrate}
+\newacronym{anova}{ANOVA}{analysis of variance}
 \newacronym{pi3k}{PI3K}{phosphatidylinositol 3-kinase}
@@
 \newacronym{fc}{FC}{fold change}
-\newacronym{ifny}{IFN-$\gamma$}{interferon-$\gamma$}
-\newacronym{tgfb}{TGF-$\beta$}{transforming growth factor-$\beta$}
+\newacronym{ifny}{IFN$\gamma$}{interferon-$\gamma$}
+\newacronym{ifna}{IFN$\alpha$}{interferon-$\alpha$}
+\newacronym{tgfb}{TGF$\beta$}{transforming growth factor-$\beta$}
+\newacronym{tnfa}{TNF$\alpha$}{tumour necrosis factor-$\alpha$}
+\newacronym{bcat}{BCAT}{$\beta$-catenin}
+\newacronym{egfr}{EGFR}{epidermal growth factor receptor}
 \newacronym[longplural={genome wide association studies},\glsshortpluralkey={GWAS}]{gwas}{GWAS}{genome wide association study}
@@
 \newacronym{pyy}{PYY$_{3-36}$}{peptide YY$_{3-36}$}
+\newacronym{nzbc}{NZBC}{New Zealand Breast Cancer}
+\newacronym{nz}{NZ}{New Zealand}
 \newacronym[longplural={Disability-Adjusted Life Years}]{daly}{DALY}{Disability-Adjusted Life Year}

### 2017-03-16 thesis/methods/methods.tex
tmp commit

@@
 There are two major classes of methods when calculating the statistical significance: self-contained or competitive tests \citep{Goeman2007}.
-Self-contained tests are more appropriate in assessing whether a biological process is significantly involved in an expreiment, and in contrast, competitive tests are better for selecting the most relevant biological processes from those that are not \citep{Wu2012}.
+Self-contained tests are more appropriate in assessing whether a biological process is significantly involved in an expreiment, whereas the competitive tests are better for selecting the most relevant biological processes from those that are not \citep{Wu2012}.
 
@@
 The restrictive property of the self-contained tests give it greater power, as they are able to reject the null hypothesis at a higher accuracy for more gene sets than the competitive tests \citep{Goeman2007}.
-As \citet{Goeman2007} stated in their paper, ``the competitive types of test can be said to voluntarily relinquish some power in order to make a stronger statement''.
-
-%TODO: add GSEA here (mootha?)
+However, as \citet{Goeman2007} stated in their paper, ``the competitive types of test can be said to voluntarily relinquish some power in order to make a stronger statement''.
+In fact, the \gls{gsea} method (developed by \citet{Subramanian2005}), one of the first and perhaps the most popular \gls{fcs} methods, uses the competitive test approach.
 
@@
 For example, if one gene was expressed with a 2-fold change, whereas another gene was expressed with a 20-fold change, the \gls{fcs} method will rank these genes accordingly (first, second, and so on) and disregard the fact that the second gene should be weighted ten times as more than the first gene \citep{Khatri2012}.
-
 Although both approaches have their own limitations, it is clear that \gls{fcs} approach have more advantages over \gls{ora} methods.
@@
 	\includegraphics[width=0.45\linewidth,page=5]{methods/gtrmastdrank_corrected}
-	\caption[Example heatmaps showing the direction of the uncorrected and corrected \acrshort{er} pathway metagene with the \acrshort{er} gene expression in the \acrshort{rma}-normalised Gatza \textit{et al.} data.]{Heatmaps showing the direction of the unflipped (left) and flipped (right) \gls{er} pathway metagene with the \gls{er} gene expression in the \gls{rma}-normalised Gatza \textit{et al.} data.
-	Level of expression is represented in the top right histogram, where low and high gene expression are colour-coded with blue and red, respectively.
+	\caption[Example heatmaps showing the direction of the uncorrected and corrected \acrshort{er} pathway metagene with the gene expression of the \acrshort{er} pathway genetic signature in the \acrshort{rma}-normalised Gatza \textit{et al.} data.]{Heatmaps showing the direction of the unflipped (left) and flipped (right) \gls{er} pathway metagene with the gene expression of the \gls{er} pathway genetic signature in the \gls{rma}-normalised Gatza \textit{et al.} data.
+	Level of gene expression is represented in the top right histogram, where low and high gene expression are colour-coded with blue and red, respectively.
 	Each row of the heatmap represents a gene from the \gls{er} genetic signature, and each column of the heatmap represents a sample from the Gatza \textit{et al.} data.
-	The \gls{er} gene expressions of the samples are shown as a separate row at top of the heatmap with the sample \gls{er} metagene scores below the \gls{er} gene expression, and the tree diagram of the heirarchical clustering of the genes is shown on the left of the heatmap.
+	The \textit{ESR1} gene expressions of the samples are shown as a separate row at top of the heatmap with the sample \gls{er} metagene scores below the \textit{ESR1} gene expression, and the tree diagram of the heirarchical clustering of the genes is shown on the left of the heatmap.
 	}
@@
 As an alternative, the pathway metagenes were visually compared with the gene that represent the pathway using a heatmap, as the expression of this gene would likely be affected the most by the pathway signature.
-For example, the \gls{er} pathway metagene was compared with \textit{ESR1} gene expression, and if the metagene was not in the same direction as \textit{ESR1} gene expression, the direction of the metagene was corrected (\cref{fig:methods/meta_dir}).
-\cref{tab:metagene_direction} summarises the genes used to check the direction of the pathway metagenes.
+For example, the \gls{er} pathway metagene was compared with the \textit{ESR1} gene expression, and if the metagene was not in the same direction as the \textit{ESR1} gene expression, the direction of the metagene was corrected (\cref{fig:methods/meta_dir}).
+\cref{tab:metagene_direction} summarises the genes used to check the direction of each of the pathway metagenes.
 
-\begin{table}[htb]
+In addition to this, the correlation of the pathway metagene with its repective gene was checked to further aid the decision on whether the metagene is going in the correct direction.
+This was due to the fact that some of the metagenes were visually difficult to determine whether it was in the right direction or not.
+Therefore, the direction as well as the magnitude of the correlation value was also taken into account in making the decision of whether to change the direction of the metagene.
+Lastly, the clustering of the pathway metagenes with the other pathway metagenes were also considered.
+It was known from the \citet{Gatza2010a} study that some pathways clustered and grouped together with certain pathways, so the direction of some of the metagenes were altered in order to keep the grouping of the pathways matched to the results in their paper.
+
+\begin{longtable}{llp{8cm}}
 	\caption{18 pathways from \citet{Gatza2010a} and its respective genes used to check the direction of the pathway metagene}
-	\label{tab:metagene_direction}
-	\begin{center}
-		\begin{tabular}{lcl}
-			Pathway & Representative gene & Pathway description\\
-			\hline
-			\rule{0pt}{2.25ex}Akt & AKT1   & \\
-			\gls{bcat}            & CTNNB1 & \\
-			E2F1                  & E2F1   & \\
-			\Gls{egfr}            & EGFR   & \\
-			\Gls{er}              & ESR1   & \\
-			\Gls{her2}            & ERBB2  & \\
-			\Gls{ifna}            & IFNA1  & \\
-			\Gls{ifny}            & IFNG   & \\
-			Myc                   & MYC    & \\
-			p53                   & TP53   & \\
-			p63                   & TP63   & \\
-			\Gls{pi3k}            & PIK3CA & \\
-			\Gls{pr}              & PGR    & \\
-			Ras                   & HRAS   & \\
-			\Gls{stat3}           & STAT3  & \\
-			Src                   & SRC    & \\
-			\Gls{tgfb}            & TGFB1  & \\
-			\Gls{tnfa}            & TNF    & \\
+	\label{tab:metagene_direction}\\
+	\centering
+	Pathway & Representative gene & Brief pathway description\\
 	\hline
 	\hline
-		\end{tabular}
-	\end{center}
-\end{table}
-
-In addition to this, the correlation of the pathway metagene with its repective gene was checked to further aid the decision on whether the metagene is going in the correct direction.
-This was due to the fact that some of the metagenes were visually difficult to determine whether it was in the right direction.
-Therefore, the direction as well as the magnitude of the correlation value was also taken into account in making the decision whether to flip the metagene.
-Lastly, the clustering of the pathway metagenes with the other pathway metagenes were also considered.
-It was known from \citet{Gatza2010a} study that some pathways clustered and grouped together with specific pathways, so the direction of some of the metagenes were altered so that the grouping of the pathways matched the results of their study.
+	\endfirsthead
+	\multicolumn{3}{c}{\tablename\ \thetable{}\ (continued)}\\
+	\hline
+	\hline
+	\endhead
+	\hline
+	\hline
+	\endlastfoot
+	\rule{0pt}{2.25ex}Akt         & AKT1   & Also known as protein kinase B; involved in cell survival\\
+	\rule{0pt}{2.25ex}\gls{bcat}  & CTNNB1 & $\beta$-catenin; involved in the Wnt signalling pathway and cell-cell adhesion\\
+	\rule{0pt}{2.25ex}E2F1        & E2F1   & Member of the E2F transcription factor family in higher eukaryotes; involved in cell cycle regulation\\
+	\rule{0pt}{2.25ex}\Gls{egfr}  & EGFR   & Also known as ERBB1 or HER1; involved in signalling the pathways for cell proliferation \\
+	\rule{0pt}{2.25ex}\Gls{er}    & ESR1   & Involved in cell cycle regulation, cell proliferation and cell survival\\
+	\rule{0pt}{2.25ex}\Gls{her2}  & ERBB2  & Also known as ERBB2; involved in cell proliferaion and apoptosis\\
+	\rule{0pt}{2.25ex}\Gls{ifna}  & IFNA1  & Cytokine involved in innate immune response against viral infection\\
+	\rule{0pt}{2.25ex}\Gls{ifny}  & IFNG   & Cytokine critical for innate and adaptive immune response\\
+	\rule{0pt}{2.25ex}Myc         & MYC    & Transcription factor involved in cell proliferaion, cell growth and apoptosis\\
+	\rule{0pt}{2.25ex}p53         & TP53   & Tumour suppressor protein involved in \acrshort{dna} repair, cell cycle regulation and apoptosis\\
+	\rule{0pt}{2.25ex}p63         & TP63   & Involved in development\\
+	\rule{0pt}{2.25ex}\Gls{pi3k}  & PIK3CA & Involved in cell growth, cell proliferaion, cell differentiation and cell survival\\
+	\rule{0pt}{2.25ex}\Gls{pr}    & PGR    & Signal transducer that activates other signalling pathways; involved in cell proliferation\\
+	\rule{0pt}{2.25ex}Ras         & HRAS   & Involved in cell proliferaion, cell differentiation, cell adhesion, apoptosis and cell migration\\
+	\rule{0pt}{2.25ex}\Gls{stat3} & STAT3  & Involved in cell growth and apoptosis\\
+	\rule{0pt}{2.25ex}Src         & SRC    & Involved in cell survival, angiogenesis, cell proliferation and cell invasion\\
+	\rule{0pt}{2.25ex}\Gls{tgfb}  & TGFB1  & Involved in cell proliferation, cell differentiation and apoptosis\\
+	\rule{0pt}{2.25ex}\Gls{tnfa}  & TNF    & Involved in immune response, inflammation and apoptosis\\
+\end{longtable}
 

### 2017-03-16 thesis/methods/methods.tex
Finished editing methods

@@
 %TODO: re-write the linear model stuff:
-Four linear models were created in the \gls{nzbc} to predict the obesity associated metagene from the \citet{Creighton2012} study.
-The first model used the sample \gls{bmi} value or \gls{bmi} status by itself, the second model with the sample \gls{bmi} value, \gls{bmi} status and selected pathway associated metagenes, the third model with selected pathway associated metagenes, and the last model was constructed with just the \gls{pr} pathway.
+
+Eleven linear models were created in the \gls{nzbc} to predict the obesity associated metagene from the \citet{Creighton2012} study.
+First seven models created were combinations of the sample \gls{bmi}, \gls{bmi} status and a selection of pathway associated metagene scores.
 The selected pathways used in the linear model construction were \gls{bcat}, \gls{er}, \gls{ifna}, \gls{ifny}, Myc and \gls{pr}.
+In deciding on the pathways to be used for the construction of the linear model, the correlation of the \gls{svd}- and transformation matrix-derived pathway metagene scores in different data sets were examined.
 
-In deciding on the pathways to be used for the construction of the linear model, the correlation of different versions of pathway metagenes, \gls{svd}-derived and transformation matrix-derived, in different data sets were used.
 When a transformation matrix is applied to the data set in which it was derived from, the resulting metagene is identical to the one created from the direct application of \gls{svd}.
@@
 
-% TODO: possibly extra analyses here (gradual pathway inclusion in the model, using all the pathway...?)
+The last four linear models were constructed with the different combinations of the sample \gls{bmi}, \gls{bmi} status and the \gls{pr} pathway metagene.
+The reason the \gls{pr} pathway metagene was used separately was because in the initial analysis, the \gls{pr} pathway metagene was the only pathway associated genetic signature that came up as significant in the other linear models (\cref{sec:prediction_of_obesity_associated_metagene_with_pathway_associate_metagene}).
+All of the created linear models are summarised in \cref{tab:lm_model_method}.
 
-The predicted obesity metagenes that were created in the \gls{nzbc} data set was compared with the original obesity metagene from the \citet{Creighton2012} data set.
-Both Pearson and Spearman correlation was calculated to assess the relationship between the predicted and original metagenes.
-Scatter plots were also created to visually compare the two metagene scores.
+\begin{table}[htpb]
+	\centering
+	\begin{threeparttable}
+		\caption{Summary of all the linear models constructed in the \gls{nzbc} data}
+		\label{tab:lm_model_method}
+		\begin{tabular}{l}
+			Linear models\\
+			\hline
+			\hline
+			\gls{bmi} only\\
+			\gls{bmi} status only\\
+			Selected pathways only\tnote{1}\\
+			\gls{bmi} and \gls{bmi} status\\
+			\gls{bmi} and selected pathways\tnote{1}\\
+			\gls{bmi} status and selected pathways\tnote{1}\\
+			\gls{bmi}, \gls{bmi} status and selected pathways\tnote{1}\\
+			\gls{pr} pathway only\\
+			\gls{pr} pathway and \gls{bmi}\\
+			\gls{pr} pathway and \gls{bmi} status\\
+			\gls{pr} pathway, \gls{bmi} and \gls{bmi} status\\
+			\hline
+			\hline
+		\end{tabular}
+		\begin{tablenotes}
+			\begin{footnotesize}
+			\item [1] \gls{bcat}, \gls{er}, \gls{ifna}, \Gls{ifny}, Myc and \gls{pr} pathway metagene scores were used.
+			\end{footnotesize}
+		\end{tablenotes}
+	\end{threeparttable}
+\end{table}
+
+In addition to these eleven linear models, a linear model was created with the sample \gls{bmi}, \gls{bmi} status and all of the pathway associated metagenes using a stepwise approach.
+Step wise approach allows the construction of a linear model that best fit the data, given a set of variables.
+There are three approaches in creating a linear model in a step wise method: top-down, bottom-up, and both.
+
+In a top-down approach, all of the variables are first included in the linear model.
+At each step, a variable is removed from the model and the \gls{aic} of the model is compared with the original model to determine whether the variable should remain in the model or not.
+This will result in a model with the variables that are most significant to the model.
+In a bottom-up approach, the model begins with no variable and adds one variable at a time to the model.
+If the variable increased the ability of the model to predict the response variable, then the variable is included in the model.
+Again, this will result in a model with only the variables that are significant in predicting the response variable.
+Alternatively, both methods in combination can be used, where the model begins with all of the variables and at each step, a variable can be added to or removed from the linear model.
+In this project, the combination of top-down and bottom-up approaches was used to create the step wise linear model.
+
+The obesity metagenes that were predicted from the linear models in the \gls{nzbc} data set was compared with the original obesity metagene scores.
+Where appropriate, p-values, \gls{anova} p-values and  $R^2$-values were calculated to assess the relationship between the predicted and the original metagene scores.
+Scatter plots and box plots were also created to visually compare the two metagene scores.
 
@@
 
-\subsection{Box and scatter plots}
+\subsection{Bar, box and scatter plots}
 \label{sub:box_and_scatter_plots}
 
+Bar plots were plotted to visualise the correlation between the \gls{svd}- and transformation matrix-derived obesity and pathway metagene scores in different data sets.
+The \texttt{barplot} function from the \textit{graphics} package in \gls{r} was used to plot the bar plots.
 Box plots and scatter plots were used to visualise the association of a metagene with some of the clinical variables of the data.
-Box plots and scatter plots were plotted using the \texttt{boxplot} and \texttt{plot} functions from the \textit{graphics} package in \gls{r}, respectively.
+Box plots and scatter plots were plotted using the \texttt{boxplot} and \texttt{plot} functions, respectively, from the \textit{graphics} package in \gls{r}.
 The p-value and \gls{anova} p-value in the box plots were calculated using the \texttt{t.test} and \texttt{aov} functions from the \textit{stats} package in \gls{r}, respectively.
@@
 Heatmaps were created using the \texttt{heatmap.2} function from the \textit{gplots} package in \gls{r} \citep{gplots}.
+For the heatmaps that required two bars above the heatmap (like those that investigate the directionality of the pathway metagenes), the heatmaps were created with the \texttt{heatmap.2x} function developed by Tom Kelly (\url{https://github.com/TomKellyGenetics/heatmap.2x}).
 
@@
 
+To summarise the number of \glspl{deg} from the gene expression analyses of the Creighton \textit{et al.} data that overlapped with the original Creighton \textit{et al.} obesity associated genetic signature, Venn diagrams were plotted.
+Venn diagrams were plotted using the \texttt{venn} function from the \textit{gplots} package in \gls{r} \citep{gplots}.
+The \texttt{overLapper} function (from \url{http://faculty.ucr.edu/~tgirke/Documents/R_BioCond/My_R_Scripts/overLapper.R}) was also used to help plot the Venn diagram \citep{Girke2016}.
+
 \subsection{Additional colours}
@@
 
-% TODO: talk about spearman correlation?
+Some colour palettes such as the colours used by MATLAB are not installed in \gls{r} by default.
+For a colour palette used in MATLAB, the \texttt{matlab.like} function from the \textit{colorRamps} package in \gls{r} was used \citep{Keitt2012}.
+For the pastel colour palette used in the bar plots, the \texttt{brewer.pal} function from the \textit{RColorBrewer} package in \gls{r} was used \citep{Neuwirth2014}.
+
+%TODO: add Spearman and Pearson correlation?

### 2017-03-16 thesis/results/results2.tex
Finished editing methods

@@
 As in \cref{sec:pathway_associated_genetic_signatures_from_gatza2010a_study}, all data sets were normalised with the \gls{rma} method and metagenes were ranked based on the number of samples in each data set.
-Transformation matrices for the obesity associated genetic signatures were made in CR data set and pathway associated genetic signatures were made in GT data set.
+Transformation matrices for the obesity associated genetic signatures were made in CR data set and pathway associated genetic signatures were made in the GT data set.
 Metagenes for all of the obesity and pathway associated genetic signatures were generated in all of the data sets with \gls{svd} and transformation matrices.
@@
 	\begin{threeparttable}
-	\caption[Summary of the Spearman correlations of the \gls{svd}- and TM-derived pathway metagenes in the GT, CR, \gls{nzbc} and FM data sets][Summary of the Spearman correlations of the \gls{svd}- and transformation matrix\tnote{1}-derived pathway metagenes in the GT, CR, \gls{nzbc} and FM data sets]
+	\caption[Summary of the Spearman correlations of the \gls{svd}- and TM-derived pathway metagenes in the GT, CR, \gls{nzbc} and FM data sets]{Summary of the Spearman correlations of the \gls{svd}- and transformation matrix\tnote{1}-derived pathway metagenes in the GT, CR, \gls{nzbc} and FM data sets}
 	\label{tab:svd_vs_tm_path}
@@
 			\begin{footnotesize}
-				\item [1] Transformation matrices were derived from the GT data set.
+				\item [1] Transformation matrices were derived from the \gls{rma}-normalised GT data set.
 			\end{footnotesize}
@@
 	\begin{threeparttable}
-	\caption[Summary of the Spearman correlations of the \gls{svd}- and TM-derived obesity metagenes in the GT, CR, \gls{nzbc} and FM data sets][Summary of the Spearman correlations of the \gls{svd}- and transformation matrix\tnote{1}-derived obesity metagenes in the GT, CR, \gls{nzbc} and FM data sets]
+	\caption[Summary of the Spearman correlations of the \gls{svd}- and TM-derived obesity metagenes in the GT, CR, \gls{nzbc} and FM data sets]{Summary of the Spearman correlations of the \gls{svd}- and transformation matrix\tnote{1}-derived obesity metagenes in the GT, CR, \gls{nzbc} and FM data sets}
 		\label{tab:svd_vs_tm_obs}
@@
 				\begin{footnotesize}
-				\item [1] Transformation matrices were derived from the CR data set.
+				\item [1] Transformation matrices were derived from the \gls{rma}-normalised CR data set.
 				\end{footnotesize}

### 2017-03-20 thesis/methods/methods.tex
Finished editing results1

@@
 
+To provide an insight into the relevance of the obesity associated genetic signatures in \gls{nz} patient cohort, the \gls{nzbc} data set used by \citet{Print2016} were used.
 The raw Affymetrix HGU133A microarray gene expression data files from the \citet{Print2016} study were downloaded from the \gls{geo} database (\gls{geo} accession ID: GSE3\-6771).
-Clinical data for the samples (age, ethnicity, tumour grade, breast cancer subtype, \gls{er}/\gls{pr} statuses, \gls{ln} status, \gls{bmi} and treatments used) in this study was obtained from Professor Cris Print (personal communication with Assoc.
+Clinical data for the samples (age, ethnicity, tumour grade, breast cancer subtype, \gls{er}/\gls{pr} statuses, \gls{ln} status, \gls{bmi} and treatments used) in this study was obtained from Prof.
+Cris Print (personal communication with Assoc.
 Prof.
@@
 
+	\newpage
 	\begin{longtable}{lcc}
@@
 		\hline
-		\hspace{12em}\insertTableNotes
+		\insertTableNotes
 	\end{longtable}
@@
 
-To overcome this limitation, \citet{Law2014} developed a method called ``variance modelling at the observational level'', or ``voom'', that allows any \gls{rnaseq} data to be used in any existing statistical analysis pipeline that is precision weight aware, including pipelines used for microarray data analyses.
-% TODO: check the meaning of precision weight in the above sentence
+To overcome this limitation, \citet{Law2014} developed a method called ``variance modelling at the observational level'', or ``voom'', that allows any \gls{rnaseq} data to be used in ``any statistical pipeline for microarray data that is precision weight aware''.
 In brief, voom first constructs a standard deviation trend from the logged \gls{cpm} value of the genes (experimental design, treatment conditions and other factors are taken into account).
@@
 The second value presents the log$_2$ value of the average expression of the gene across all of the arrays/samples.
-%TODO: more details on simple Bayesian method
-The moderated \textit{t}-statistic is the same as normal \textit{t}-statistic, but its standard error has been adjusted by a simple Bayesian method \citep{Smyth2005}.
+The moderated \textit{t}-statistic is the same as normal \textit{t}-statistic, but its standard error has been adjusted (or ``shrunk'') toward a common value by a simple Bayesian method \citep{Smyth2005}.
 Raw p-value and adjusted p-value represent the p-value of the gene before and after it has been corrected for multiple hypothesis testing, respectively.
@@
 	\rule{0pt}{2.25ex}Akt         & AKT1   & Also known as protein kinase B; involved in cell survival\\
+	\hline
 	\rule{0pt}{2.25ex}\gls{bcat}  & CTNNB1 & $\beta$-catenin; involved in the Wnt signalling pathway and cell-cell adhesion\\
+	\hline
 	\rule{0pt}{2.25ex}E2F1        & E2F1   & Member of the E2F transcription factor family in higher eukaryotes; involved in cell cycle regulation\\
+	\hline
 	\rule{0pt}{2.25ex}\Gls{egfr}  & EGFR   & Also known as ERBB1 or HER1; involved in signalling the pathways for cell proliferation \\
+	\hline
 	\rule{0pt}{2.25ex}\Gls{er}    & ESR1   & Involved in cell cycle regulation, cell proliferation and cell survival\\
+	\hline
 	\rule{0pt}{2.25ex}\Gls{her2}  & ERBB2  & Also known as ERBB2; involved in cell proliferaion and apoptosis\\
+	\hline
 	\rule{0pt}{2.25ex}\Gls{ifna}  & IFNA1  & Cytokine involved in innate immune response against viral infection\\
+	\hline
 	\rule{0pt}{2.25ex}\Gls{ifny}  & IFNG   & Cytokine critical for innate and adaptive immune response\\
+	\hline
 	\rule{0pt}{2.25ex}Myc         & MYC    & Transcription factor involved in cell proliferaion, cell growth and apoptosis\\
+	\hline
 	\rule{0pt}{2.25ex}p53         & TP53   & Tumour suppressor protein involved in \acrshort{dna} repair, cell cycle regulation and apoptosis\\
+	\hline
 	\rule{0pt}{2.25ex}p63         & TP63   & Involved in development\\
+	\hline
 	\rule{0pt}{2.25ex}\Gls{pi3k}  & PIK3CA & Involved in cell growth, cell proliferaion, cell differentiation and cell survival\\
+	\hline
 	\rule{0pt}{2.25ex}\Gls{pr}    & PGR    & Signal transducer that activates other signalling pathways; involved in cell proliferation\\
+	\hline
 	\rule{0pt}{2.25ex}Ras         & HRAS   & Involved in cell proliferaion, cell differentiation, cell adhesion, apoptosis and cell migration\\
+	\hline
 	\rule{0pt}{2.25ex}\Gls{stat3} & STAT3  & Involved in cell growth and apoptosis\\
+	\hline
 	\rule{0pt}{2.25ex}Src         & SRC    & Involved in cell survival, angiogenesis, cell proliferation and cell invasion\\
+	\hline
 	\rule{0pt}{2.25ex}\Gls{tgfb}  & TGFB1  & Involved in cell proliferation, cell differentiation and apoptosis\\
+	\hline
 	\rule{0pt}{2.25ex}\Gls{tnfa}  & TNF    & Involved in immune response, inflammation and apoptosis\\
@@
 If the predicted metagene was similar to the original metagene, then it provides an evidence that the pathway used to construct the model was associated with the obesity metagene.
-
-%TODO: re-write the linear model stuff:
 

### 2017-03-20 thesis/misc/appendix.tex
Finished editing results1

@@
 		\label{fig:appendix/check_raw_vs_std_tm}
+	\end{figure}
+
+	\section{Remainder of the results of FM obesity metagenes in the \gls{icgc} cancer data sets}
+	\label{sec:rest_of_the_fm_icgc_cancer_heatmap_results}
+
+	\cref{fig:appendix/fm_ob_meta_heatmap_icgc} shows the association of the FM obesity metagene with the gene expression pattern of the FM obesity associated genetic signature in the other \gls{icgc} cancer data sets.
+	\cref{fig:appendix/fm_ob_meta_box_scatter_icgc} shows the association of the FM obesity metagene scores with the patient \gls{bmi} and \gls{bmi} status.
+	Clearly, the FM metagene was able to capture the overall gene expression patterns, but was not significantly associated with the patient \gls{bmi} or \gls{bmi} status.
+
+	\begin{figure}[htpb]
+		\centering
+		\includegraphics[width=0.32\linewidth,page=8]{appendix/fm_meta_ICGC}\\
+		\includegraphics[width=0.32\linewidth,page=13]{appendix/fm_meta_ICGC}
+		\includegraphics[width=0.32\linewidth,page=18]{appendix/fm_meta_ICGC}
+		\includegraphics[width=0.32\linewidth,page=23]{appendix/fm_meta_ICGC}\\
+		\includegraphics[width=0.32\linewidth,page=28]{appendix/fm_meta_ICGC}
+		\includegraphics[width=0.32\linewidth,page=33]{appendix/fm_meta_ICGC}
+		\includegraphics[width=0.32\linewidth,page=38]{appendix/fm_meta_ICGC}\\
+		\caption[Association of the FM obesity metagene with the sample gene expressions in the other \gls{icgc} data]{Heatmaps showing the association of the FM obesity metagenes  with the sample gene expressions in the other \gls{icgc} data sets.
+		Scales are as described in previous figures.}
+		\label{fig:appendix/fm_ob_meta_heatmap_icgc}
+	\end{figure}
+
+	\begin{figure}[htpb]
+		\centering
+		\includegraphics[page=9,width=0.42\linewidth]{appendix/fm_meta_ICGC}
+		\hfill
+		\includegraphics[page=10,width=0.42\linewidth]{appendix/fm_meta_ICGC}\\
+		\caption[Association of the FM obesity metagene with the patient \gls{bmi}/\gls{bmi} status in the other \gls{icgc} data]{Box plots and scatter plots showing the association of the FM obesity metagene with the patient \gls{bmi} status  and \gls{bmi}, respectively, in the \gls{icgc} data sets.
+		P-values and $R^2$-value are as described in previous figures.}
+		\label{fig:appendix/fm_ob_meta_box_scatter_icgc}
+	\end{figure}
+
+	\begin{figure}[htpb]
+		\ContinuedFloat
+		\captionsetup{list=off,format=cont}
+		\centering
+		\includegraphics[page=14,width=0.45\linewidth]{appendix/fm_meta_ICGC}
+		\hfill
+		\includegraphics[page=15,width=0.45\linewidth]{appendix/fm_meta_ICGC}\\
+		\includegraphics[page=19,width=0.45\linewidth]{appendix/fm_meta_ICGC}
+		\hfill
+		\includegraphics[page=20,width=0.45\linewidth]{appendix/fm_meta_ICGC}\\
+		\includegraphics[page=24,width=0.45\linewidth]{appendix/fm_meta_ICGC}
+		\hfill
+		\includegraphics[page=25,width=0.45\linewidth]{appendix/fm_meta_ICGC}\\
+		\caption{Name}
+	\end{figure}
+
+	\begin{figure}[htpb]
+		\ContinuedFloat
+		\captionsetup{list=off,format=cont}
+		\centering
+		\includegraphics[page=29,width=0.45\linewidth]{appendix/fm_meta_ICGC}
+		\hfill
+		\includegraphics[page=30,width=0.45\linewidth]{appendix/fm_meta_ICGC}\\
+		\includegraphics[page=34,width=0.45\linewidth]{appendix/fm_meta_ICGC}
+		\hfill
+		\includegraphics[page=35,width=0.45\linewidth]{appendix/fm_meta_ICGC}\\
+		\includegraphics[page=39,width=0.45\linewidth]{appendix/fm_meta_ICGC}
+		\hfill
+		\includegraphics[page=40,width=0.45\linewidth]{appendix/fm_meta_ICGC}\\
+		\caption{Name}
+	\end{figure}
+
+	\section{Direction of all the obesity metagenes from the CR data}
+	\label{sec:direction_of_all_the_obesity_metagenes_from_the_cr_data}
+
+	The direction of the obesity metagenes identified from the CR data were examined and corrected for (\cref{fig:crob_meta_direction}).
+	Gene probes that were common to all obesity associated genetic signatures were used to examine whether the metagene was in the correct direction, relative to the other obesity metagenes
+
+	\begin{figure}[htpb]
+		\centering
+		\includegraphics[width=0.32\linewidth,page=3]{appendix/cr_meta_direction_raw}
+		\includegraphics[width=0.32\linewidth,page=6]{appendix/cr_meta_direction_raw}\\
+		\includegraphics[width=0.32\linewidth,page=9]{appendix/cr_meta_direction_raw}
+		\includegraphics[width=0.32\linewidth,page=12]{appendix/cr_meta_direction_raw}
+		\includegraphics[width=0.32\linewidth,page=15]{appendix/cr_meta_direction_raw}\\
+		\includegraphics[width=0.32\linewidth,page=18]{appendix/cr_meta_direction_raw}
+		\includegraphics[width=0.32\linewidth,page=21]{appendix/cr_meta_direction_raw}
+		\includegraphics[width=0.32\linewidth,page=24]{appendix/cr_meta_direction_raw}\\
+		\caption[Directionality of the obesity metagenes in the CR data]{Heatmaps showing the gene expression of the obesity associated genetic signatures with the obesity metagene in \gls{rma}-normalised CR data set.
+		The gene probes that were common to all of the obesity associated genetic signatures were used to determine the direction of the obesity metagenes.
+		The scales and axes are the same as the previous heatmaps.
+		}
+		\label{fig:crob_meta_direction}
+	\end{figure}
+
+	\section{Remainder of the results of the other CR obesity metagenes in \gls{nzbc} data set}
+	\label{sec:rest_of_the_cr_ob_meta_heatmap_results_cris}
+
+	All of the other obesity metagenes from the CR data set were able to capture the overall gene expressions of the samples in the \gls{nzbc} data set (\cref{fig:appendix/cr_ob_meta_heatmap_cris}).
+	However, none of the CR obesity metagenes were able to significantly associate with the sample \gls{bmi}/\gls{bmi} status in the \gls{nzbc} data set (\cref{fig:appendix/cr_ob_meta_box_scatter_cris}).
+
+	\begin{figure}[htp!]
+		\centering
+		\includegraphics[width=0.32\linewidth,page=8]{appendix/cris_crdeg_trans_meta}\\
+		\includegraphics[width=0.32\linewidth,page=13]{appendix/cris_crdeg_trans_meta}
+		\includegraphics[width=0.32\linewidth,page=18]{appendix/cris_crdeg_trans_meta}
+		\includegraphics[width=0.32\linewidth,page=23]{appendix/cris_crdeg_trans_meta}\\
+		\includegraphics[width=0.32\linewidth,page=28]{appendix/cris_crdeg_trans_meta}
+		\includegraphics[width=0.32\linewidth,page=33]{appendix/cris_crdeg_trans_meta}
+		\includegraphics[width=0.32\linewidth,page=38]{appendix/cris_crdeg_trans_meta}\\
+		\caption[Association of the other CR obesity metagenes with the sample gene expressions in the \gls{nzbc} data]{Heatmaps showing the association of the other obesity metagenes identified from the CR data with the sample gene expressions from the \gls{nzbc} data set.
+		Scales are as described in previous figures.}
+		\label{fig:appendix/cr_ob_meta_heatmap_cris}
+	\end{figure}
+
+	\begin{figure}[htpb]
+		\centering
+		\includegraphics[page=9,width=0.42\linewidth]{appendix/cris_crdeg_trans_meta}
+		\hfill
+		\includegraphics[page=10,width=0.42\linewidth]{appendix/cris_crdeg_trans_meta}\\
+		\caption[Association of the other CR obesity metagenes with the sample \gls{bmi}/\gls{bmi} status in the \gls{nzbc} data]{Box plots and scatter plots showing the association of the other CR obesity metagenes with the sample \gls{bmi} status  and \gls{bmi} from the \gls{nzbc} data set, respectively.
+		P-values and $R^2$-value are as described in previous figures.}
+		\label{fig:appendix/cr_ob_meta_box_scatter_cris}
+	\end{figure}
+
+	\begin{figure}[htpb]
+		\ContinuedFloat
+		\captionsetup{list=off,format=cont}
+		\centering
+		\includegraphics[page=14,width=0.45\linewidth]{appendix/cris_crdeg_trans_meta}
+		\hfill
+		\includegraphics[page=15,width=0.45\linewidth]{appendix/cris_crdeg_trans_meta}\\
+		\includegraphics[page=19,width=0.45\linewidth]{appendix/cris_crdeg_trans_meta}
+		\hfill
+		\includegraphics[page=20,width=0.45\linewidth]{appendix/cris_crdeg_trans_meta}\\
+		\includegraphics[page=24,width=0.45\linewidth]{appendix/cris_crdeg_trans_meta}
+		\hfill
+		\includegraphics[page=25,width=0.45\linewidth]{appendix/cris_crdeg_trans_meta}\\
+		\caption{Name}
+	\end{figure}
+
+	\begin{figure}[htpb]
+		\ContinuedFloat
+		\captionsetup{list=off,format=cont}
+		\centering
+		\includegraphics[page=29,width=0.45\linewidth]{appendix/cris_crdeg_trans_meta}
+		\hfill
+		\includegraphics[page=30,width=0.45\linewidth]{appendix/cris_crdeg_trans_meta}\\
+		\includegraphics[page=34,width=0.45\linewidth]{appendix/cris_crdeg_trans_meta}
+		\hfill
+		\includegraphics[page=35,width=0.45\linewidth]{appendix/cris_crdeg_trans_meta}\\
+		\includegraphics[page=39,width=0.45\linewidth]{appendix/cris_crdeg_trans_meta}
+		\hfill
+		\includegraphics[page=40,width=0.45\linewidth]{appendix/cris_crdeg_trans_meta}\\
+		\caption{Name}
 	\end{figure}
@@
 
-	\section{Remainder of the results of FM obesity metagenes in the \gls{icgc} cancer data sets}
-	\label{sec:rest_of_the_fm_icgc_cancer_heatmap_results}
+	\section{Obesity associated genetic signature using sample \gls{bmi} values}
+	\label{sec:obesity_associated_genetic_signature_using_sample_bmi_values}
 
-%TODO: add FM metagene plots for ICGC
+	Since all of these obesity associated genetic signatures were derived based on the discrete values of sample \gls{bmi} staus, a genetic signature was identified using the sample \gls{bmi} value (continuous variable).
+	Gene probe expressions were correlated with the sample \gls{bmi} values and was corrected for multiple hypothesis testing with \gls{fdr}.
+	There were no significant gene that correlated with the sample \gls{bmi} (p \textless{} 0.05).
 
-	\section{Remainder of the results of the other CR obesity metagenes in \gls{nzbc} data set}
-	\label{sec:rest_of_the_cr_ob_meta_heatmap_results_cris}
-
-	All of the other obesity metagenes from the CR data set were able to capture the overall gene expressions of the samples in the \gls{nzbc} data set (\cref{fig:appendix/cr_ob_meta_heatmap_cris}).
-	However, none of the CR obesity metagenes were able to significantly associate with the sample \gls{bmi}/\gls{bmi} status in the \gls{nzbc} data set (\cref{fig:appendix/cr_ob_meta_box_scatter_cris}).
+	With that said, the top 799 gene probes that correlated the most with the sample \gls{bmi} were examined to see if there was any association.
+	The results were similar to the other obesity associated genetic signatures: the continuous \gls{bmi} signature significantly associated with sample \gls{bmi} and \gls{bmi} status in the CR data, but did not show any significance in any of the other cancer data.
 
@@
 		\centering
-		\includegraphics[width=0.32\linewidth,page=8]{appendix/cris_crdeg_trans_meta}\\
-		\includegraphics[width=0.32\linewidth,page=13]{appendix/cris_crdeg_trans_meta}
-		\includegraphics[width=0.32\linewidth,page=18]{appendix/cris_crdeg_trans_meta}
-		\includegraphics[width=0.32\linewidth,page=23]{appendix/cris_crdeg_trans_meta}\\
-		\includegraphics[width=0.32\linewidth,page=28]{appendix/cris_crdeg_trans_meta}
-		\includegraphics[width=0.32\linewidth,page=33]{appendix/cris_crdeg_trans_meta}
-		\includegraphics[width=0.32\linewidth,page=38]{appendix/cris_crdeg_trans_meta}\\
-		\caption[Association of the other CR obesity metagenes with the sample gene expressions in the \gls{nzbc} data]{Heatmaps showing the association of the other obesity metagenes identified from the CR data with the sample gene expressions from the \gls{nzbc} data set.
-	Scales are as described in previous figures.}
-		\label{fig:appendix/cr_ob_meta_heatmap_cris}
+		\includegraphics[page=3,width=0.8\linewidth]{appendix/contbmi_cr}\\
+		\includegraphics[page=4,width=0.45\linewidth]{appendix/contbmi_cr}
+		\hfill
+		\includegraphics[page=5,width=0.45\linewidth]{appendix/contbmi_cr}
+		\caption[Continuous \gls{bmi} metagene in various cancer data]{Heatmap, box plot and scatter plot showing the association of the continuous \gls{bmi} metagene with the sample gene expression, \gls{bmi} and \gls{bmi} status, respectively, in various cancer data.
+		Scales, p-values and $R^2$-value are as described in previous figures.}
+		\label{fig:cont_bmi}
 	\end{figure}
 
-	\begin{figure}[htpb]
-		\centering
-		\includegraphics[page=9,width=0.42\linewidth]{appendix/cris_crdeg_trans_meta}
-		\hfill
-		\includegraphics[page=10,width=0.42\linewidth]{appendix/cris_crdeg_trans_meta}\\
-		\caption[Association of the other CR obesity metagenes with the sample \gls{bmi}/\gls{bmi} status in the \gls{nzbc} data]{Box plots and scatter plots showing the association of the other CR obesity metagenes with the sample \gls{bmi} status  and \gls{bmi} from the \gls{nzbc} data set, respectively.
-	P-values and $R^2$-value are as described in previous figures.}
-		\label{fig:appendix/cr_ob_meta_box_scatter_cris}
-	\end{figure}
-
-	\begin{figure}[htpb]
+	\begin{figure}[htp!]
 		\ContinuedFloat
@@
 		\centering
-		\includegraphics[page=14,width=0.45\linewidth]{appendix/cris_crdeg_trans_meta}
+		\includegraphics[page=3,width=0.8\linewidth]{appendix/contbmi_ICGC}\\
+		\includegraphics[page=4,width=0.45\linewidth]{appendix/contbmi_ICGC}
 		\hfill
-		\includegraphics[page=15,width=0.45\linewidth]{appendix/cris_crdeg_trans_meta}\\
-		\includegraphics[page=19,width=0.45\linewidth]{appendix/cris_crdeg_trans_meta}
-		\hfill
-		\includegraphics[page=20,width=0.45\linewidth]{appendix/cris_crdeg_trans_meta}\\
-		\includegraphics[page=24,width=0.45\linewidth]{appendix/cris_crdeg_trans_meta}
-		\hfill
-		\includegraphics[page=25,width=0.45\linewidth]{appendix/cris_crdeg_trans_meta}\\
-		\caption{Name}
+		\includegraphics[page=5,width=0.45\linewidth]{appendix/contbmi_ICGC}
+		\caption{hello}
 	\end{figure}
 
-	\begin{figure}[htpb]
+	\begin{figure}[htp!]
 		\ContinuedFloat
@@
 		\centering
-		\includegraphics[page=29,width=0.45\linewidth]{appendix/cris_crdeg_trans_meta}
+		\includegraphics[page=3,width=0.8\linewidth]{appendix/contbmi_cris}\\
+		\includegraphics[page=4,width=0.45\linewidth]{appendix/contbmi_cris}
 		\hfill
-		\includegraphics[page=30,width=0.45\linewidth]{appendix/cris_crdeg_trans_meta}\\
-		\includegraphics[page=34,width=0.45\linewidth]{appendix/cris_crdeg_trans_meta}
-		\hfill
-		\includegraphics[page=35,width=0.45\linewidth]{appendix/cris_crdeg_trans_meta}\\
-		\includegraphics[page=39,width=0.45\linewidth]{appendix/cris_crdeg_trans_meta}
-		\hfill
-		\includegraphics[page=40,width=0.45\linewidth]{appendix/cris_crdeg_trans_meta}\\
-		\caption{Name}
+		\includegraphics[page=5,width=0.45\linewidth]{appendix/contbmi_cris}
+		\caption{hello}
 	\end{figure}
+
+\section{Pathways significant in each of the cancer types}
+\label{sec:pathways_significant_in_each_of_the_cancer_types}
+
+\begin{table}[htpb]
+	\centering
+	\caption{Significantly enriched \gls{kegg} pathways in \gls{icgc} cancer data}
+	\label{tab:sig_path_kegg}
+	\begin{threeparttable}
+		\begin{tabular}{ll}
+			Cancer type                      & Pathway         \\
+			\hline
+			\hline
+				\rule{0pt}{2.25ex}\gls{blca} & NA\tnote{1}             \\
+				\gls{cesc}                   & ABC transporter \\
+				\gls{coad}                   & NA              \\
+				\gls{kirp}                   & NA              \\
+				\gls{lihc}                   & NA              \\
+				\gls{read}                   & NA              \\
+				\gls{skcm}                   & NA              \\
+				\gls{ucec}                   & NA              \\
+			\hline
+			\hline
+		\end{tabular}
+		\begin{tablenotes}
+			\item [1] Not available (no pathway enriched)
+		\end{tablenotes}
+	\end{threeparttable}
+\end{table}
+
+\begin{table}[htpb]
+	\centering
+	\caption{Significantly enriched Reactome pathways in \gls{icgc} cancer data}
+	\label{tab:sig_path_reactome}
+	\begin{threeparttable}
+		\begin{tabular}{ll}
+			Cancer type & Pathway \\
+			\hline
+			\hline
+				\rule{0pt}{2.25ex}\gls{blca} & Phosphate bond hydrolysis by NUDT proteins \\
+				\gls{cesc}                   & Mitochondrial ABC transporters             \\
+				\gls{coad}   & NA\tnote{1}               \\
+				\gls{kirp}   & NA             \\
+				\gls{lihc}   & NA             \\
+				\gls{read}   & NA             \\
+				\gls{skcm}   & NA             \\
+				\gls{ucec}   & NA             \\
+			\hline
+			\hline
+		\end{tabular}
+		\begin{tablenotes}
+			\item [1] Not available (no pathway enriched)
+		\end{tablenotes}
+	\end{threeparttable}
+\end{table}
+
+\newpage
+
+	\begin{TableNotes}
+		\begin{footnotesize}
+			\item [1] Not available (no pathway enriched)
+		\end{footnotesize}
+	\end{TableNotes}
+
+\begin{longtable}{lp{0.8\textwidth}}
+	\centering
+	\label{tab:sig_path_go}\\
+	Cancer type & Pathway\\
+	\endfirsthead
+	\multicolumn{2}{c}{\tablename\ \thetable{}\ (continued)} \\
+	\hline
+	\hline
+	\endhead
+	\hline
+	\hline
+	\rule{0pt}{2.25ex}\gls{blca} & Nucleobase-containing small molecule catabolic process                                          \\
+                                 & SUMO-specific protease activity                                                                 \\
+                                 & Diphosphoinositol-polyphosphate diphosphatase activity                                          \\
+                                 & Inositol diphosphate tetrakisphosphate diphosphatase activity                                   \\
+                                 & Inositol-1-diphosphate-2,3,4,5,6-pentakisphosphate diphosphatase activity                       \\
+                                 & Inositol-3-diphosphate-1,2,4,5,6-pentakisphosphate diphosphatase activity                       \\
+                                 & Inositol-5-diphosphate-1,2,3,4,6-pentakisphosphate diphosphatase activity                       \\
+                                 & Inositol-1,5-bisdiphosphate-2,3,4,6-tetrakisphosphate 1-diphosphatase activity                  \\
+                                 & Inositol-1,5-bisdiphosphate-2,3,4,6-tetrakisphosphate 5-diphosphatase activity                  \\
+                                 & Inositol-3,5-bisdiphosphate-2,3,4,6-tetrakisphosphate 5-diphosphatase activity                  \\
+                                 & Phosphatidylinositol 3-kinase complex, class IB                                                 \\
+                                 & Neurotrophin binding                                                                            \\
+                                 & 1-phosphatidylinositol-5-phosphate 4-kinase activity                                            \\
+                                 & Coagulation                                                                                     \\
+                                 & Cholesterol O-acyltransferase activity                                                          \\
+                                 & Pyrimidine deoxyribonucleotide catabolic process                                                \\
+                                 & Nucleotidase activity                                                                           \\
+                                 & G-protein coupled neurotensin receptor activity                                                 \\
+                                 & 8-oxo-7,8-dihydroguanosine triphosphate pyrophosphatase activity                                \\
+                                 & 8-oxo-7,8-dihydrodeoxyguanosine triphosphate pyrophosphatase activity                           \\
+                                 & IDP catabolic process                                                                           \\
+                                 & ADP-sugar diphosphatase activity                                                                \\
+	\hline
+	\gls{cesc}                   & ATP-binding cassette (ABC) transporter complex                                                  \\
+                                 & Protein delipidation                                                                            \\
+                                 & Atg8 ligase activity                                                                            \\
+                                 & Cyclin-dependent protein kinase 5 holoenzyme complex                                            \\
+                                 & Molybdopterin synthase complex                                                                  \\
+                                 & Negative regulation of nodal signaling pathway                                                  \\
+                                 & Aspartate-tRNA ligase activity                                                                  \\
+                                 & Peroxisomal long-chain fatty acid import                                                        \\
+                                 & Sterol-transporting ATPase activity                                                             \\
+                                 & Negative regulation of intestinal phytosterol absorption                                        \\
+                                 & Negative regulation of intestinal cholesterol absorption                                        \\
+                                 & Apolipoprotein A-I receptor activity                                                            \\
+                                 & Sulfonylurea receptor activity                                                                  \\
+                                 & Atg12 transferase activity                                                                      \\
+                                 & Superior olivary nucleus maturation                                                             \\
+                                 & Cyclin-dependent protein kinase 5 activator activity                                            \\
+                                 & ATPase activity, coupled to transmembrane movement of substances                                \\
+	\hline
+	\gls{coad}                   & NA\tnote{1}                                                                                     \\
+	\hline
+	\gls{kirp}                   & Trace-amine receptor activity                                                                   \\
+                                 & Regulation of viral release from host cell                                                      \\
+                                 & Nuclear import                                                                                  \\
+	\hline
+	\gls{lihc}                   & NA                                                                                              \\
+	\hline
+	\gls{read}                   & Nuclear polyadenylation-dependent rRNA catabolic process                                        \\
+                                 & Nuclear polyadenylation-dependent tRNA catabolic process                                        \\
+                                 & Sodium channel complex                                                                          \\
+                                 & U1 snRNA 3'-end processing                                                                      \\
+                                 & U5 snRNA 3'-end processing                                                                      \\
+                                 & Glucuronosyl-N-acetylglucosaminyl-proteoglycan 4-alpha-N-acetylglucosaminyltransferase activity \\
+                                 & Pyrroline-5-carboxylate reductase activity                                                      \\
+                                 & Hydroxymethylglutaryl-CoA reductase (NADPH) kinase activity                                     \\
+                                 & Acetyl-CoA carboxylase kinase activity                                                          \\
+                                 & rRNA modification                                                                               \\
+                                 & Rab GDP-dissociation inhibitor activity                                                         \\
+                                 & Nuclear retention of pre-mRNA with aberrant 3'-ends at the site of transcription                \\
+                                 & Cellular polysaccharide biosynthetic process                                                    \\
+                                 & Heparan sulfate N-acetylglucosaminyltransferase activity                                        \\
+                                 & N-acetylglucosaminyl-proteoglycan 4-beta-glucuronosyltransferase activity                       \\
+                                 & Glucuronyl-galactosyl-proteoglycan 4-alpha-N-acetylglucosaminyltransferase activity             \\
+                                 & Regulation of RIG-I signaling pathway                                                           \\
+                                 & Regulation of cysteine-type endopeptidase activity                                              \\
+                                 & Adenylate cyclase-inhibiting serotonin receptor signaling pathway                               \\
+                                 & Peptide-O-fucosyltransferase activity                                                           \\
+                                 & Interferon-gamma receptor activity                                                              \\
+	\hline
+	\gls{skcm}                   & Tissue morphogenesis                                                                            \\
+                                 & Beta-1,3-galactosyl-O-glycosyl-glycoprotein beta-1,6-N-acetylglucosaminyltransferase activity   \\
+                                 & N-acetyllactosaminide beta-1,6-N-acetylglucosaminyltransferase activity                         \\
+                                 & Posterior mesonephric tubule development                                                        \\
+                                 & Regulation of stem cell differentiation                                                         \\
+                                 & Negative regulation of nodal signaling pathway                                                  \\
+                                 & Aspartate-tRNA ligase activity                                                                  \\
+                                 & Glutamate-cysteine ligase complex                                                               \\
+                                 & Glutamate-cysteine ligase activity                                                              \\
+                                 & Rab GDP-dissociation inhibitor activity                                                         \\
+	\hline
+	\gls{ucec}                   & SUMO-specific protease activity                                                                 \\
+                                 & Regulation of dendritic cell cytokine production                                                \\
+                                 & Detection of triacyl bacterial lipopeptide                                                      \\
+                                 & Cellular response to triacyl bacterial lipopeptide                                              \\
+                                 & Toll-like receptor 1-Toll-like receptor 2 protein complex                                       \\
+                                 & Detection of diacyl bacterial lipopeptide                                                       \\
+                                 & Cellular response to diacyl bacterial lipopeptide                                               \\
+                                 & Toll-like receptor 2-Toll-like receptor 6 protein complex                                       \\
+                                 & Diacyl lipopeptide binding                                                                      \\
+                                 & Coagulation                                                                                     \\
+                                 & Polo kinase kinase activity                                                                     \\
+                                 & Regulation of protein serine/threonine kinase activity                                          \\
+                                 & DNA-dependent protein kinase complex                                                            \\
+                                 & Methylosome                                                                                     \\
+	\hline
+	\hline
+	\insertTableNotes
+\end{longtable}

### 2017-03-20 thesis/results/results.tex
Finished editing results1

@@
 
-The obesity associated genetic signatures are central to this project in order to clarify the relationship between \gls{bmi} and cancer.
-Firstly in this chapter, the previously identified obesity associated signatures from the studies conducted by \citet{Creighton2012} and \citet{Fuentes-Mattei2014} were examined in turn to judge the agreement of these signatures with the sample \gls{bmi}/\gls{bmi} status, as presented in thier results.
-Secondly, novel obesity associated genetic signatures were identified in the Creighton \textit{et al.} (CR) data set and was compared with the signatures from the Creighton \textit{et al.} and Fuentes-Mattei \textit{et al.} studies.
-Lastly, the presence of common genes or pathways that were associated with obesity in multiple types of cancer was explored using the \gls{icgc} cancer data sets.
+The obesity associated genetic signatures are central to this project, as a means to clarify the relationship between \gls{bmi} and tumour biology.
+In this chapter, the previously identified obesity associated genetic signatures from the studies conducted by \citet{Creighton2012} and \citet{Fuentes-Mattei2014} are examined in turn to judge the agreement of these signatures with the sample \gls{bmi}/\gls{bmi} status, as presented in thier results.
+After this, novel obesity associated genetic signatures are identified in the Creighton \textit{et al.} (CR) data set and was compared with the signatures from the Creighton \textit{et al.} and the Fuentes-Mattei \textit{et al.} studies.
+Lastly, the presence of common genes or pathways associated with obesity in multiple types of cancer is explored using the data sets from \gls{icgc} (\citetalias{ICGC}; \citealp{Zhang2011}).
 
@@
 
-Obesity metagene was created using the obesity associated genetic signatures from the \citet{Creighton2012} study.
-There was no description about the normalisation method used by Creighton \textit{et al.} when they first analysed their data, so the more reliable \gls{rma} method was used to  normalise the CR data set.
-The obesity metagene was plotted in a heatmap with the sample gene expression  to check whether the metagene scores were in accordance with the overall gene expression of the samples (\cref{fig:crmetaheat}).
+An obesity metagene was created using the obesity associated genetic signature from the \citet{Creighton2012} study.
+There was no description about the normalisation method used by Creighton \textit{et al.} when they first analysed their data, so the popular \gls{rma} method \citep{Irizarry2003} was used to  normalise the CR data set.
+The obesity metagene was created in the standardised \gls{rma}-normalised CR data set, and  was plotted above a heatmap with the sample gene expression  to check whether the metagene scores were in accordance with the overall gene expression of the samples (\cref{fig:crmetaheat}).
 
@@
 	The obesity associated metagene scores of the samples are shown as a separate row at top of the heatmap, and the tree diagram of the heirarchical clustering of the genes is shown on the left of the heatmap.
-	For clarity, the metagene scores were flipped in order to match the results from the \citet{Creighton2012} study.}
+	For clarity, the sign of the metagene scores were reversed in order to match the results from the \citet{Creighton2012} study.}
 	\label{fig:crmetaheat}
@@
 
-As shown in \cref{fig:crmetaheat}, high obesity associated metagene score of the sample reflected low expression in majority of the genes in the signature, and in contrast, low obesity associated metagene score of the sample reflected high expression in majority of the genes in the signature.
+As shown in \cref{fig:crmetaheat}, a high obesity associated metagene score  reflects low expression in majority of the genes in the signature, and in contrast, a low obesity associated metagene score reflects high expression in the majority of the genes in the signature.
 This was consistent with the reported property of the obesity associated signature by \citet{Creighton2012} (see \cref{ssub:creighton_study}).
-To provide further evidence that the obesity associated metagenes were in fact associated with the \gls{bmi} status and \gls{bmi} value of the samples, a box plot and a scatter plot were created, respectively (\cref{fig:crmetaboxplot}).
+To provide further evidence that the obesity associated metagenes were in fact associated with both the \gls{bmi} status and the \gls{bmi} value of the samples, a box plot and a scatter plot were created, respectively (\cref{fig:crmetaboxplot}).
 
@@
 This was due to the fact that the obesity associated genetic signatures were originally identified from the comparison of the samples that were obese with the samples that were not obese, and therefore the metagene scores were significant with the obese group, but not with the overweight group.
-In addition to this, although the regression line in the scatter plot showed statistically significant association with sample \gls{bmi}, the sample \gls{bmi} values seemed to be randomly dispersed across the metagene scores.
-This suggested that perhaps the obesity metagene from the \citet{Creighton2012} study was not associated with sample \gls{bmi} as strongly as expected.
+In addition to this, although the regression line in the scatter plot showed statistically significant association with sample \gls{bmi}, the sample \gls{bmi} values seem to be randomly dispersed across the metagene scores.
+This suggests that perhaps the obesity metagene from the \citet{Creighton2012} study was not associated with the sample \gls{bmi} as strongly as reported.
 \\
@@
 All of the \gls{icgc} data were normalised as described in \cref{ssub:rna_seq_data}.
-Before the transformation matrix was applied to the log$_{10}$-normalised cancer data, the suitability of standardised data or untouched (non-standardised) data was determined (see \cref{sec:metagenes_created_from_raw_data_vs_standardised_data_icgc}).
+Before the transformation matrix was applied to the log$_{10}$-normalised cancer data, the obesity metagene was created from the standardised data or untouched (non-standardised) \gls{icgc} data and compared with one another to determine which was better for the application of the transformation matrix (see \cref{sec:comp_cr_raw_std_icgc}).
 From these results, the standardised data was found to be the most suitable to use for matrix transformation.
-The transformation matrix was applied to each cancer data set in turn to obtain the Creighton \textit{et al.} obesity metagenes from each of the data set.
-Each obesity metagenes were plotted in a heatmap with the corresponding data set in which the metagene was taken from (\cref{fig:crmetaicgc}; \cref{sec:rest_of_the_icgc_cancer_heatmap_resutls}).
-These heatmaps confirmed that the obesity associated metagene derived from CR data was able to capture the overall gene expression pattern, where the metagene scores reflected the expression levels of majority of the genes in the signature.
-As before, association of the obesity metagene with the sample \gls{bmi} and \gls{bmi} status was checked in their respective cancer data set (\cref{fig:crmetaicgc}; \cref{sec:rest_of_the_icgc_cancer_heatmap_resutls}).
-Out of all the cancer types, only \gls{blca} data set showed significant association of obesity metagene with the overweight group (but not with the obese group).
-However, neither the \gls{anova} p-value nor the regression line in the scatter plot were statistically significant, suggesting that this apparent association with the overweight group was not due to the obesity metagene (?).
+The transformation matrix was applied to each cancer data set in turn to generate an obesity metagene from each of the data sets.
+Each obesity metagene was plotted in a heatmap with the corresponding data set from which the metagene was taken (\cref{fig:crmetaicgc}; \cref{sec:rest_of_the_cr_icgc_cancer_heatmap_results}).
+These heatmaps confirmed that the obesity associated metagene derived from the CR data was able to capture the overall gene expression pattern, where the metagene scores reflected the expression levels of the majority of the genes in the signature.
+As before, association of the obesity metagene with the sample \gls{bmi} and \gls{bmi} status was checked in their respective cancer data set (\cref{fig:crmetaicgc}; \cref{sec:rest_of_the_cr_icgc_cancer_heatmap_results}).
+Out of all the cancer types, only the \gls{blca} data set showed a significant association with the obesity metagene,  and only for the overweight group (not the obese group).
+% However, neither the \gls{anova} p-value nor the regression line in the scatter plot were statistically significant, suggesting that this apparent association with the overweight group was not due to the obesity metagene (?).
 
@@
 	\includegraphics[page=3,width=0.8\linewidth]{results1/crtcga_std}\\
-	\vspace{1em}
 	\includegraphics[page=4,width=0.45\linewidth]{results1/crtcga_std}
@@
 There could be several reasons for the apparent lack of association of the metagene with sample \gls{bmi}/\gls{bmi} status.
-First, transformation matrix was derived from CR microarray data, but the \gls{icgc} cancer data were \gls{rnaseq} data.
-Though the log$_{10}$ normalisation and standardisation of the data was the most appropriate adjustment to be made to the \gls{rnaseq} data, this adjustment was not equivalent to the \gls{rma} normalisation method that was used on the microarray data.
-Secondly, none of the \gls{icgc} cancer originated from breast as in CR data.
-Since the obesity associated signature was identified in breast cancer data, the signature may be specific to breast cancer and may not be suitable in other cancer types.
+First, the transformation matrix was derived from the CR microarray data, but the \gls{icgc} cancer data were generated via \gls{rnaseq}.
+Though the log$_{10}$-normalisation and standardisation of the data was the most appropriate adjustments to be made to the \gls{rnaseq} data, these adjustments were not equivalent to the \gls{rma} normalisation method that was used on the microarray data.
+Secondly, none of the \gls{icgc} cancer originated from breast tumours as in the CR data.
+Since the obesity associated signature was identified in the breast cancer data, the signature may be specific to breast cancer and may not be generalisable to other cancer types.
 \\
@@
 \noindent
-To check whether the metagene was specific to breast cancer microarray data, the same transformation matrix was applied to the \gls{nzbc} microarray data from \citet{Print2016} study (referred to as \gls{nzbc} data hereafter).
-\gls{nzbc} data was normalised with \gls{rma} method and the transformation matrix was applied to the normalised data to obtain the metagene.
+To check whether the metagene was specific to breast cancer microarray data, the same transformation matrix was applied to the \gls{nzbc} microarray data from the \citet{Print2016} study (referred to as \gls{nzbc} data hereafter).
+\gls{nzbc} data was normalised with the \gls{rma} method and the transformation matrix was applied to the normalised data to obtain the metagene.
 The metagene was again compared with the gene expression of the samples with a heatmap and the association of the metagene with the sample \gls{bmi}/\gls{bmi} status was examined with box and scatter plots (\cref{fig:crmetaprint}).
 
-The obesity metagene managed to reflect the overall gene expression of the samples in \gls{nzbc} data.
-However, as with \gls{icgc} cancer data, the obesity metagene scores did not significantly associate with the sample \gls{bmi} or \gls{bmi} status.
-These results confirmed that the lack of association of the obesity metagene from CR data was not due to the technology in which the data was gathered (mcroarray or \gls{rnaseq}), nor the cancer type in which the transformation matrix was applied to.
+The Creighton \textit{et al.} obesity metagene managed to reflect the overall gene expression of the samples in \gls{nzbc}  data.
+However, as with the \gls{icgc} cancer data, the obesity metagene scores did not significantly associate with the sample \gls{bmi} or \gls{bmi} status (\cref{fig:crmetaprint}).
+These results confirmed that the lack of association of the obesity metagene from the CR data was not due to the technology in which the data was gathered (mcroarray or \gls{rnaseq}), nor the cancer type in which the transformation matrix was applied to.
 
@@
 	\includegraphics[width=0.8\linewidth,page=3]{results1/cris_cr_trans_meta}\\
-	\vspace{1em}
 	\includegraphics[width=0.45\linewidth,page=4]{results1/cris_cr_trans_meta}
@@
 
-All together, all of these results suggest that the obesity metagene identified in the study conducted by \citet{Creighton2012} associated with sample \gls{bmi} and \gls{bmi} status only in CR data.
-The lack of association with \gls{bmi} was not due to the type of technology platform in which the data was gathered, as neither the \gls{icgc} \gls{rnaseq} data nor \gls{nzbc} microarray data showed no significant association with the obesity metagene.
-Furthermore, the obesity metagene was not dependent on the cancer type since the obesity associated genetic signature did not show significant association in neither the \gls{icgc} cancer data nor in \gls{nzbc} data.
+Taken together, these results suggest that the obesity metagene identified in the study conducted by \citet{Creighton2012} associated with sample \gls{bmi} and \gls{bmi} status only in that particular data set.
+The lack of association with \gls{bmi} was not due to the type of technology platform in which the data was gathered, as neither the \gls{icgc} \gls{rnaseq} data nor the \gls{nzbc} microarray data showed a significant association with the obesity metagene.
+Furthermore, the obesity metagene was not dependent on the cancer type since the obesity associated genetic signature did not show significant association in either the \gls{icgc} cancer data or in \gls{nzbc} data.
 
-One possible reason why the obesity metagene from the \citet{Creighton2012} study did not show significant association with other data sets could be because the genetic signature was in fact not an obesity specific signature, but a signature that was detected due to another clinical variable.
-Another reason for this apparent lack of association could be that the genetic signature was too specific to the data and was not a broad obesity associated genetic signature, but an obesity associated signature specifically for CR data set.
+One possible reason why the obesity metagene from the \citet{Creighton2012} study did not show significant association with other data sets could be because the genetic signature was in fact not an obesity specific signature, but a signature that was detected due to another clinical variable (investigated in \cref{sec:creighton_obesity_metagene_new}).
+Another reason for this apparent lack of association could be that the genetic signature was too specific to the data and was not a broad obesity associated genetic signature, but an obesity associated signature specific to the samples that were profiled in the Creighton \textit{et al.} publication.
 
@@
 
-Since the obesity associated genetic signature from the \citet{Creighton2012} study was not associated with sample \gls{bmi} or \gls{bmi} status in other data sets, the obesity associated genetic signature from the \citet{Fuentes-Mattei2014} study (FM) was examined to see whether FM obesity metagene associated with sample \gls{bmi} or \gls{bmi} status.
-Since FM data set did not have sample \gls{bmi} information, FM obesity metagene was not able to be compared with the sample \gls{bmi} or \gls{bmi} status in the original FM data.
-However, the transformation matrix was still applied to other cancer data to see whether FM obesity metagene associated with sample \gls{bmi}/\gls{bmi} status
+Since the obesity associated genetic signature from the \citet{Creighton2012} study was not associated with sample \gls{bmi} or \gls{bmi} status in other data sets, the obesity associated genetic signature from the \citet{Fuentes-Mattei2014} study (FM) was examined to see whether the FM obesity metagene associated with the sample \gls{bmi} or \gls{bmi} status.
+Since the FM data set did not have sample \gls{bmi} information, the FM obesity metagene was not able to be compared with the sample \gls{bmi} or \gls{bmi} status in the original data.
+However, the transformation matrix was still applied to other tumour types to see whether the FM obesity metagene associated with sample \gls{bmi}/\gls{bmi} status in those data sets.
 
-Firstly, FM data was normalised with the \gls{rma} method and \gls{svd} was applied to the normalised FM data to get the transformation matrix.
-The transformation matrix was used to transform the \gls{rma}-normalised CR data to extract FM obesity metagene scores in CR data.
-FM obesity metagene scores were compared with gene expressions and sample \gls{bmi}/\gls{bmi} status in CR data, as shown in \cref{fig:fmmetacr}.
-Clearly, like with the obesity metagene identified by Creighton \textit{et al.}, FM obesity metagene was reflective of the overall gene expression of the samples, but did not associate with the sample \gls{bmi} or \gls{bmi} status.
+Firstly, FM data was normalised with the \gls{rma} method and \gls{svd} was performed on the normalised FM data to obtain the transformation matrix.
+The transformation matrix was used to transform the \gls{rma}-normalised CR data to extract the FM obesity metagene scores in the CR data.
+The FM obesity metagene scores were compared with gene expression and sample \gls{bmi}/\gls{bmi} status in the CR data, as shown in \cref{fig:fmmetacr}.
+Clearly, as with the CR obesity metagene, the FM obesity metagene was reflective of the overall gene expression of the samples, but did not associate with the sample \gls{bmi} or \gls{bmi} status, suggesting that this signature also does not generalise to other data sets.
+The transformation matrix was applied to \gls{nzbc} data  as well (\cref{fig:fmmetacris}).
+Again, FM obesity metagene scores reflected the gene expression of the samples, did not significantly associate with sample \gls{bmi}/\gls{bmi} status.
+
+Next, the transformation matrix was applied to the \gls{icgc} cancer data and the resulting metagenes were compared with the gene expression and sample \gls{bmi}/\gls{bmi} status.
+As evident in \cref{fig:fmmetaicgc}, the FM obesity metagene scores appeared to reflect the overall gene expression of the FM obesity associated genetic signature.
+As with all the results in this chapter so far, the FM obesity metagene did not significantly associate with any of the \gls{icgc} cancer data, except in the \gls{blca} data set (\cref{fig:fmmetaicgc}).
+The FM obesity metagene significantly associated with the overweight group (but not with the obese group), and also had a significant \gls{anova} p-value.
+On the contrary to the association of the metagene with the sample \gls{bmi} status, the FM obesity metagene was not associated with sample \gls{bmi}.
+These results suggested that the samples that were overweight in the \gls{blca} data set had similar biological properties as the samples that were obese in the FM data set.
+However, due to the fact that the FM metagene lacked association with the sample \gls{bmi} in the \gls{blca} data set and that the metagene did not show any significant association in any other cancer type, it was difficult to determine whether the observed association of the FM obesity metagene with the overweight group of samples was truly reflective of the effect resulted from the FM metagene.
+\\
 
@@
 	\includegraphics[page=3,width=0.8\linewidth]{results1/cr_fm_meta}\\
-	\vspace{1em}
 	\includegraphics[page=4,width=0.45\linewidth]{results1/cr_fm_meta}
@@
 
-Next, the transformation matrix was applied to the \gls{icgc} cancer data and resulting metagenes were compared with the gene expression and sample \gls{bmi}/\gls{bmi} status.
-As evident in \cref{fig:fmmetaicgc}, FM obesity metagene scores appeared to reflect the overall gene expression of FM obesity associated genetic signature.
-As with all the results in this chapter so far, FM obesity metagene did not significantly associate with any of the \gls{icgc} cancer data, except in \gls{blca} data set (\cref{fig:fmmetaicgc}).
-FM obesity metagene significantly associated with the overweight group (but not with the obese group), and also had a significant \gls{anova} p-value.
-On the contrary to the association of the metagene with the sample \gls{bmi} status, FM obesity metagene was not associated with sample \gls{bmi}.
-These results suggested that the samples that were overweight in the \gls{blca} data set had similar biological properties as the samples that were obese in FM data set.
-However, due to the fact that FM metagene lacked association with sample \gls{bmi} in \gls{blca} data set and that the metagene did not show any significant association in any other cancer type, it was difficult to determine whether the observed association of FM obesity metagene with the overweight group of samples was truly reflective of the effect resulted from FM metagene.
-Lastly, the transformation matrix was applied to \gls{nzbc} data (\cref{fig:fmmetacris}).
-Again, FM obesity metagene scores reflected the gene expression of the samples, did not significantly associate with sample \gls{bmi}/\gls{bmi} status.
-\\
+\begin{figure}[htp!]
+	\centering
+	\includegraphics[page=3,width=0.8\linewidth]{results1/cris_fm_meta}\\
+	\includegraphics[page=4,width=0.45\linewidth]{results1/cris_fm_meta}
+	\hfill
+	\includegraphics[page=5,width=0.45\linewidth]{results1/cris_fm_meta}
+	\caption[FM metagene in the \gls{nzbc} data]{Heatmap, box plot and scatter plot showing the association of the FM obesity associated metagene with the sample gene expression, \gls{bmi} and \gls{bmi} status from the \gls{nzbc} data, respectively.
+	Scales, p-values and $R^2$-value are as described in previous figures.}
+	\label{fig:fmmetacris}
+\end{figure}
 
@@
 	\includegraphics[page=3,width=0.8\linewidth]{results1/fm_meta_ICGC}\\
-	\vspace{1em}
 	\includegraphics[page=4,width=0.45\linewidth]{results1/fm_meta_ICGC}
@@
 
-\begin{figure}[htp!]
-	\centering
-	\includegraphics[page=3,width=0.8\linewidth]{results1/cris_fm_meta}\\
-	\vspace{1em}
-	\includegraphics[page=4,width=0.45\linewidth]{results1/cris_fm_meta}
-	\hfill
-	\includegraphics[page=5,width=0.45\linewidth]{results1/cris_fm_meta}
-	\caption[FM metagene in the \gls{nzbc} data]{Heatmap, box plot and scatter plot showing the association of the FM obesity associated metagene with the sample gene expression, \gls{bmi} and \gls{bmi} status from the \gls{nzbc} data, respectively.
-	Scales, p-values and $R^2$-value are as described in previous figures.}
-	\label{fig:fmmetacris}
-\end{figure}
-
 \noindent
-These results showed that FM obesity associated metagene was not transferable to other cancer data sets, similar to the obesity metagene identified by Creighton \textit{et al.}.
-This meant that both the obesity metagenes identified by Creighton \textit{et al.} and Fuentes-Mattei \textit{et al.} may have been too specific to the orignial data set in which the signatures were identified in.
-Furthermore, there was a possibility that these obesity associated metagenes were not related to obesity, but associated with a different clinical variable that may be closely related to \gls{bmi}.
+These results showed that FM obesity associated metagene was not generalisable to other cancer data sets, similar to the obesity metagene identified by Creighton \textit{et al.}.
+This meant that both the obesity metagenes identified by Creighton \textit{et al.} and Fuentes-Mattei \textit{et al.} may have been too specific to the orignial data set in which the signatures were identified.
+Furthermore, there was a possibility that these obesity associated metagenes were not related to obesity, but associated with a different clinical variable that may be closely related to \gls{bmi} (\cref{sec:creighton_obesity_metagene_new}).
 
@@
 
-Both the obesity associated metagenes generated from CR data and FM data were able to capture the overall pattern of gene expression of the genetic signature in the samples, but did not associate with sample \gls{bmi}/\gls{bmi} status.
-One possible reason for this result could be that the obesity associated genetic signatures from the \citet{Creighton2012} and \citet{Fuente-Matter2014} studies may not have been truly associated with sample \gls{bmi}  status, but with another clinical variable.
-To reject this possibility, obesity associated genetic signatures were identified in CR data after controlling for all the clinical variables in the data set.
+Both the obesity associated metagenes generated from CR data and FM data were able to capture the overall pattern of gene expression of the genetic signature in the samples, but did not associate with sample \gls{bmi} or \gls{bmi} status.
+One possible reason for this result could be that the obesity associated genetic signatures from the \citet{Creighton2012} and \citet{Fuente-Mattei2014} studies may not have been truly associated with sample \gls{bmi}  status, but with another clinical variable.
+To invevstigate this possibility, obesity associated genetic signatures were identified in CR data after controlling for all the clinical variables in the data set.
 FM data set was not used to get obesity associated genetic signatures, as no \gls{bmi} information was available for the samples in FM data.
 
-Firstly, to validate whether the identified obesity associated genetic signature was similar to the signature that Creighton \textit{et al.} had originally found, the identified \glspl{deg} were compared with the original genetic signature identified by Creighton \textit{et al.}.
-\glspl{deg}  were identified between the samples from obese patient and non-obese patient in the \gls{rma}-normalised CR data, as described in \cref{sec:gene_expression_analysis}.
-Without adjusting the p-value with FDR, 5278 gene probes and 1781 gene probes were significant at p \textless{} 0.05 and p \textless{} 0.01, respectively.
-After afjustment, there were only 9 gene probes significant at p \textless{} 0.05 and no gene was significant at p \textless{} 0.01.
-Furthermore, there were only 61 gene probes that were significantly differentially expressed at p \textless{} 0.05 with a log$_2$ \gls{fc} greater than 1.2.
-From these observations, the log$_2$ fold change of the gene probes were ignored and the threshold p-value was set to 0.01 (without adjustment) for the identification of gene probes.
-This was to include as many gene probes as possible without losing power (?).
+Firstly, to determine whether the identified obesity associated genetic signature was similar to the signature that Creighton \textit{et al.} had originally found, the identified \glspl{deg} were compared with the original genetic signature identified by Creighton \textit{et al.}.
+\citet{Creighton2012} originally found their obesity associated genetic signature by doing a gene expression analysis between the breast tumour samples from the obese and non-obese patients, and from this list, they selected the 799 significant \glspl{deg} with log$_2$-fold change greater than 1.2.
+
+Therefore, \glspl{deg}  were identified between the samples from obese and non-obese patients in the \gls{rma}-normalised CR data, as described in \cref{sec:gene_expression_analysis}.
+Without adjusting the p-value to account for multiple hupothesis testing, 5278 gene probes and 1781 gene probes were significant at p \textless{} 0.05 and p \textless{} 0.01, respectively (\cref{tab:ob_deg_summary}).
+After adjustment, there were only 9 gene probes significant at p \textless{} 0.05 and no gene was significant at p \textless{} 0.01, which suggests that the the majority of the 799 probe sets reported by Creighton \textit{et al.} may actually be false positives.
+Furthermore, there were only 61 gene probes that were significantly differentially expressed at p \textless{} 0.05 with a log$_2$-\gls{fc} greater than 1.2.
+From these observations, the log$_2$-fold change of the gene probes were ignored and the threshold p-value was set to 0.01 (without adjustment) for the identification of probe sets.
 Additionally, when there were more than 799 gene probes identified, only the most significant 799 gene probes were taken as the obesity associated genetic signature, as this many genes were originally identified by Creighton \textit{et al.}.
+This was to include as many probe set as possible so the novel genetic signatures identified were able to be compared with the original Creighton \textit{et al.} signature.
 These criteria were applied for the identification of other genetic signatures as well.
@@
 The clinical variables controlled were age, ethnicity, menopause status, tumour grade, hormone (\gls{er}, \gls{pr}, and \gls{her2}) statuses  and \gls{ln} status.
-In the residual data, 1104 gene probes were significant with unadjusted p-value (p \textless{} 0.01).
+In the residual data, 1104 gene probes were significant with unadjusted p-value (p \textless{} 0.01; additional results shown in \cref{tab:ob_deg_summary}).
 Again, the most significant 799 genes were taken as the obesity associated genetic signature from this data.
+
+\begin{table}[htpb]
+	\centering
+	\caption[Summary of the number of \acrshort{deg}s identified using different unadjusted and \acrshort{fdr}-adjusted p-value threshold in different versions of the \gls{rma}-normalised CR data set]{Summary of the number of \glspl{deg} identified using different unadjusted and \gls{fdr}-adjusted p-value threshold in different versions of the \gls{rma}-normalised CR data set}
+	\label{tab:ob_deg_summary}
+	\begin{tabular}{lcccc}
+		& \multicolumn{4}{c}{Number of \glspl{deg} identified}\\
+		\cmidrule(r){2-5}
+		& \multicolumn{2}{c}{Unadjusted p-value} & \multicolumn{2}{c}{\gls{fdr}-adjusted p-value}\\
+		\cmidrule(r){2-3} \cmidrule(r){4-5}
+		& 0.01 & 0.05 & 0.01 & 0.05\\
+		\hline
+		\hline
+		\rule{0pt}{2.25ex}CR data       & 1781 & 5278 & 0 & 9 \\
+		Residual CR data                & 1104 & 4371 & 0 & 0 \\
+		Caucasian-only CR data          & 2129 & 6029 & 0 & 0 \\
+		Residual Caucasion-only CR data & 1558 & 5427 & 0 & 0 \\
+		\hline
+		\hline
+	\end{tabular}
+\end{table}
 
@@
 
-\begin{figure}[htp!]
+\begin{figure}[htpb]
 	\centering
@@
 \noindent
-There was a possible bias between different ethnic groups, where African American patients were more likely to be obese compared to the Caucasian patients (see \cref{ssub:creighton_study}).
+There was a possible bias between different ethnic groups, where African American patients were more likely to be obese compared to the Caucasian patients (13 out of 16 were obese in African American group, whereas 22 out of 77 were obese in Caucasian group; see \cref{ssub:creighton_study}).
 Though ethnicity was controlled in the residual data, the effect of ethnicity on the data was completely removed to prevent any possibility of ethnicity influencing the analysis.
-Therefore, the effect of ethnicity was completely removed by considering only the Caucasian patients in CR data, which left a total of 77 Caucasian patients in the data set.
-With this Caucasian-only data set, obesity associated genetic signatures were identified as described above; first with unadjusted data, then with the data with clinical variables adjusted (except ethnicity, as it has already been controlled for by considering Caucasian samples only), and lastly the common genes with the original signature were identified.
+Therefore, the effect of ethnicity was ignored by considering only the Caucasian patients in CR data, which left a total of 77 Caucasian patients in the data set.
+With this Caucasian-only data set, obesity associated genetic signatures were identified as described above; first with the unadjusted data, then with the data with clinical variables adjusted (except ethnicity, as it has already been controlled for by considering Caucasian samples only), and lastly the common genes with the original signature.
 
@@
 Again, Venn diagram was used to summarise the genes identified from the Caucasian patient data set (\cref{fig:venn2}).
-\\
 
-\begin{figure}[htp!]
+Since all of these obesity associated genetic signatures were derived based on the discrete values of sample \gls{bmi} staus, a genetic signature was identified by using the correlation of the gene probe expression with the sample \gls{bmi} value (continuous variable).
+However, no significant genes were identified and thus was not used for further analysis (\cref{sec:obesity_associated_genetic_signature_using_sample_bmi_values}).
+
+\begin{figure}[htpb]
 	\centering
@@
 \end{figure}
-
-\noindent
-In addition to these eight obesity associated genetic signatures, a genetic signature was identified based on the correlation of gene probe expression with sample \gls{bmi} as continuous values, rather than discrete values as in \gls{bmi} status.
-However, no significant genes were identified and thus was not used for further analysis (results detailed in appendix).
 
@@
 
-As with the Or signature, all eight obesity associated genetic signatures were validated in CR data set first, and then compared in other cancer data sets by using the transformation matrix generated in CR data.
-% Before the metagenes were created, all of the signatures were converted into gene symbols, as described in \cref{sec:data} (\cref{tab:signature_gene_num}).
-CR data (unadjusted, all samples included) was normalised with \gls{rma} method and \gls{svd} was applied to generate the metagene for the corresponding genetic signature.
-The direction of the metagenes were first examined to make sure that all of the metagenes were in line with one another (see \cref{sub:metagene_direction}; results shown in appendix).
+As with the Or signature, all eight obesity associated genetic signatures were validated in CR data set first, and then compared in other cancer data sets by using the transformation matrix that was generated in \gls{rma}-normalised CR data.
+First, the CR data (unadjusted, all samples included) was normalised with the \gls{rma} method and \gls{svd} was applied to generate the all of the obesity metagenes.
+The direction of the metagenes were examined to make sure that all of the metagenes were in line with one another (see \cref{sub:metagene_direction}; \cref{sec:direction_of_all_the_obesity_metagenes_from_the_cr_data}).
 The comparison of the metagenes with the sample gene expression in CR data were displayed as heatmaps (\cref{fig:degmetacr}).
-It was clear from the heatmaps that all of the metagenes reflected the overall expression of the corresponding genetic signatures.
+It was clear from the heatmaps that all of the metagenes reflected the overall expression of the corresponding obesity associated genetic signatures.
 The association of the metagenes with sample \gls{bmi}/\gls{bmi} status was significant for all eight of the metagenes identified (\cref{fig:degmetacr}).
 These results confirmed that all of the obesity associated genetic signatures identified in \cref{sub:identification_of_obesity_associated_genetic_signatures} significantly associated with sample \gls{bmi}/\gls{bmi} status in CR data, where the metagenes were derived from.
+However in \gls{nzbc} data, the higher the metagenes scores of the samples were, the more expressed the genes were (and vice versa), but lacked the association with the sample \gls{bmi}/\gls{bmi} status (\cref{fig:degmetaprint}; \cref{sec:rest_of_the_cr_ob_meta_heatmap_results_cris}).
 
@@
 
-Additionally, the correlation of all of the obesity associated metagenes were examined to see whether these metagenes were similar to one another.
+Additionally, the correlation of all of the obesity metagenes were examined to see whether these metagenes were similar to one another.
 As clearly shown in \cref{fig:cr_meta_cor}, there were two distinct groups within the eight metagenes; the first group contained the metagenes that were not overlapped with the Or metagene (Cr, Res, Ca and CaRes), while the other group had the overlapped metagenes (CrOl, ResOl, CaOl and CaResOl).
-With that said, all eight metagenes showed high correlation with one another (lowest correlation approximately at 0.85), which suggested that all of these metagenes were detecting similar underlying biological mechanism from the data.
+With that said, all eight metagenes showed high correlation with one another (lowest correlation approximately at 0.85), which suggested that all of these metagenes were detecting similar underlying biological mechanism from the data, as would be expected.
+
+\begin{figure}[htp!]
+	\centering
+	\includegraphics[page=3,width=0.8\linewidth]{results1/cris_crdeg_trans_meta}\\
+	\vspace{1em}
+	\includegraphics[page=4,width=0.45\linewidth]{results1/cris_crdeg_trans_meta}
+	\hfill
+	\includegraphics[page=5,width=0.45\linewidth]{results1/cris_crdeg_trans_meta}
+	\caption[Cr obesity metagene in the \gls{nzbc} data]{Heatmap, box plot and scatter plot showing the association of the Cr obesity associated metagene with the sample gene expression, \gls{bmi} and \gls{bmi} status from the \gls{nzbc} data, respectively.
+	The results for the other obesity metagenes are shown in \cref{sec:rest_of_the_cr_ob_meta_heatmap_results_cris}.
+	Scales, p-values and $R^2$-value are as described in previous figures.}
+	\label{fig:degmetaprint}
+\end{figure}
 
@@
 	\gls{svd} was applied to the \gls{rma}-normalised CR data to generate each of the eight metagenes.
-	High and low correlation were represented as red and blue, respectively, where the colours were matched with the values on the scale shown in the top right histogram.}
+	High and low correlation were represented as red and blue, respectively, where the colours were matched with the values on the scale shown in the top right histogram.
+	The lowest correlation was 0.8469, between the CrOl and the CaRes metagenes.
+	}
 	\label{fig:cr_meta_cor}
@@
 
-To confirm whether these metagenes showed significant association with sample \gls{bmi}/\gls{bmi} status in other cancer data sets, transformation matrix was applied to the \gls{icgc} cancer data and \gls{nzbc} data.
-Firstly in \gls{icgc} cancer data sets, all eight metagene scores were reflective of the sample gene expression of the genetic signatures (\cref{fig:degmetaicgc}).
-In all cancer types but \gls{blca}, none of the metagenes significantly associated with the sample \gls{bmi}/\gls{bmi} status (results shown in \cref{app:a}).
-As shown in \cref{fig:degmetaicgc}, only the \gls{blca} data set showed significant association of the Cr metagene with sample \gls{bmi}/\gls{bmi} status.
-With the metagenes other than Cr metagene, Res, CrOl and Ca metagenes showed statistical significance in overweight group, \gls{anova} and the regression line p-values; CaRes metagene was significant in overweight group and \gls{anova} p-value; CrOl and ResOl metagenes were significant in overweight group; and ResOl and CaOl metagenes were not significant in \gls{blca} data set.
+To confirm whether these metagenes showed significant association with sample \gls{bmi}/\gls{bmi} status in other cancer types, the transformation matrix was applied to the \gls{icgc} cancer data.
+Firstly in the \gls{icgc} cancer data sets, all eight metagene scores were reflective of the sample gene expression of the corresponding genetic signatures (\cref{fig:degmetaicgc}; \cref{sec:rest_of_the_cr_icgc_cancer_heatmap_results}).
+In all cancer types but \gls{blca}, none of the metagenes significantly associated with the sample \gls{bmi}/\gls{bmi} status (\cref{sec:rest_of_the_cr_icgc_cancer_heatmap_results}).
+As shown in \cref{fig:degmetaicgc}, only the \gls{blca} data set showed significant association of the Cr metagene with the sample \gls{bmi}/\gls{bmi} status.
+The Res and Ca metagenes also showed statistical significance in the overweight group, \gls{anova} and the regression line p-values; CaRes and Res metagenes were significant in overweight group and \gls{anova} p-value; CrOl and CaResOl metagenes were significant in overweight group; and ResOl and CaOl metagenes were not significantly associated with the sample \gls{bmi} or \gls{bmi} status in \gls{blca} data set.
 
-This was unexpected since all of the genetic signatures were results of gene expression analysis between the non-obese group and the obese group of samples in CR data, and yet the metagenes showed significant association with the overweight group in \gls{blca} data set.
-These results suggested that the samples that were overweight in \gls{blca} data were similar in phenotype with the samples that were obese in CR data.
-Again, due to the fact that all of the metagenes lacked association with sample \gls{bmi}/\gls{bmi} status in all other \gls{icgc} data set, it was difficult to conclude whether the observed association of many of the obesity metagenes with the overweight group of samples was truly reflective of the effect resulted from these metagenes.
+This was unexpected since all of the genetic signatures resulted from the  gene expression analysis between the non-obese group and the obese group of the samples in CR data, and yet the metagenes showed significant association with the overweight group rather than the obese group in the \gls{blca} data set.
+These results suggested that the samples that were overweight in the \gls{blca} data were similar in phenotype with the samples that were obese in CR data.
+Again, due to the fact that all of the metagenes lacked association with the sample \gls{bmi}or \gls{bmi} status in all other \gls{icgc} cancer types, it was difficult to conclude whether the observed association of many of the obesity metagenes with the overweight group of the samples was truly reflective of the effect resulted from these metagenes.
+
+Taken together, these results showed that, even though all of the obesity metagenes significantly associated with the sample \gls{bmi}/\gls{bmi} status in the original data where the genetic signatures were derived from, the metagenes were not generalisable in other cancer data sets.
+Furthermore, these results showed that the lack of association with sample \gls{bmi} and \gls{bmi} status were not due to other clinical variables in the CR data set.
+This raised a question of whether there was any obesity associated genetic signature that was common in multiple types of cancer.
 
@@
 
-In \gls{nzbc} data, the higher the metagenes scores were the more expressed the genes were (and vice versa), but lacked the association with sample \gls{bmi}/\gls{bmi} status (\cref{fig:degmetaprint}).
-These results showed that, even though all of the metagenes significantly associated with sample \gls{bmi}/\gls{bmi} status in the data the genetic signatures were derived from, the metagenes did not transfer well in other cancer data sets.
-Furthermore, these results showed that the lack of association with sample \gls{bmi}/\gls{bmi} status were not due to other clinical variables in CR data set.
-This raised a question of whether there was any obesity associated genetic signature that was common in multiple cancer types.
-
 \begin{figure}[htp!]
@@
 
-\begin{figure}[htp!]
-	\centering
-	\includegraphics[page=3,width=0.8\linewidth]{results1/cris_crdeg_trans_meta}\\
-	\vspace{1em}
-	\includegraphics[page=4,width=0.45\linewidth]{results1/cris_crdeg_trans_meta}
-	\hfill
-	\includegraphics[page=5,width=0.45\linewidth]{results1/cris_crdeg_trans_meta}
-	\caption[Cr obesity metagene in the \gls{nzbc} data]{Heatmap, box plot and scatter plot showing the association of the Cr obesity associated metagene with the sample gene expression, \gls{bmi} and \gls{bmi} status from the \gls{nzbc} data, respectively.
-	The results for the other obesity metagenes are shown in \cref{sec:rest_of_the_cr_ob_meta_heatmap_results_cris}.
-	Scales, p-values and $R^2$-value are as described in previous figures.}
-	\label{fig:degmetaprint}
-\end{figure}
-
 \section{Common genes across multiple cancer types}
@@
 
-It was clear that the obesity associated genetic signatures created from CR data were not consistent across different data sets.
-To determine whether there was any obesity associated genetic signature that was expressed in multiple different cancer types, gene expression analysis was carried out on each of the eight \gls{icgc} cancer types and common genes were searched from the \glspl{deg} that were identified.
-All of the \gls{icgc} cancer data sets were normalised with voom method (\cref{ssub:rna_seq_data}), which were then put through the gene expression analysis pipeline to identify \glspl{deg} between the obese group of samples and non-obese group of samples (\cref{sec:gene_expression_analysis}).
+It was clear that the obesity associated genetic signatures created from the CR data set were not consistent across different data sets.
+To determine whether there was any obesity associated genetic signature that was expressed in multiple different cancer types, gene expression analysis was carried out on each of the eight \gls{icgc} cancer types, and common genes were searched from the \glspl{deg} that were identified.
+All of the \gls{icgc} cancer data sets were normalised with voom (\cref{ssub:rna_seq_data}), which were then put through the gene expression analysis pipeline to identify the \glspl{deg} between the obese group of samples and non-obese group of samples (\cref{sec:gene_expression_analysis}).
 \cref{tab:icgcdegnum} summarised the number of genes found from each cancer type with p \textless{} 0.05.
@@
 There were 9695 unique \glspl{deg} across the eight cancer types, and these genes were checked for any commonalities across the different cancer types.
-There were 7330 genes that were differentially expressed by one cancer type, 2024 genes expressed by any two cancer types, 320 genes expressed by any three cancer types, and 21 genes expressed by any four cancer types.
-There were no genes differentially expressed by six or more cancer types (see \cref{tab:icgcdegtab} for a summary).
+There were 7330 genes that were differentially expressed in one cancer type, 2024 genes differentially expressed in any two cancer types, 320 genes expressed in any three cancer types, and 21 genes expressed in any four cancer types.
+There were no genes differentially expressed by five or more cancer types (see \cref{tab:icgcdegtab} for a summary).
 To confirm that these results were statistically significant, the gene expression analysis was repeated 1000 times for each cancer type after the samples were randomised in each analysis (see \cref{sub:sample_randomisation_in_simulation_analysis}).
@@
 
-The results from the simulation showed that, on average, 5732 genes were found to be differentially expressed by a cancer type, 1057 by two cancer types, 111 by three cancer types, and 7 by four cancer types.
-When the results from \gls{icgc} gene expression analysis were compared with the simulation results, the number of \glspl{deg} found were statistically significant, as the numbers of genes identified exceeded the 95$^{th}$ percentile values for up to 5 cancer types.
+The results from the simulation showed that, on average, 5732 genes were found to be differentially expressed in any single cancer type, 1057 in any two cancer types, 111 in any three cancer types, and 7 in any four cancer types.
+Unfortunately, there were no \glspl{deg} expressed in all eight cancer types, which confirmed that there was no common genes that were differentially expressed between the samples that were obese and normal weight.
+When the results from \gls{icgc} gene expression analysis were compared with the simulation results, the number of \glspl{deg} found were statistically significant, as the numbers of genes identified exceeded the 95$^{th}$ percentile values for up to 4 cancer types.
 This result confirmed that the \glspl{deg} from the \gls{icgc} cancer data were not identified by chance.
-Unfortunately, there were no \glspl{deg} expressed in all eight cancer types, which confirmed that there was no common genes that were differentially expressed between the samples that were obese and normal weight.
+However, this result also showed that many of the genes identified in the gene expression analysis of the cancer types were observed by chance and that the majority of these genes were likely to be false positives.
 
@@
 			\rule{0pt}{2.25ex}\hspace{-1em}{\small Results from gene expression analysis} & 7330 & 2024 & 320 & 21 & 0 & 0 & 0 & 0 \\
-			\hspace{-1em}{\small Results from simulation:\tnote{2}}                       &      &      &     &    &   &   &   &   \\
+			\hspace{-1em}{\small Results from the simulation:\tnote{2}}                       &      &      &     &    &   &   &   &   \\
 			{\small Mean no. of \glspl{deg} identified}                                   & 5732 & 1057 & 111 & 7  & 0 & 0 & 0 & 0 \\
@@
 Each cancer type was normalised with voom (\cref{ssub:rna_seq_data}) and pathway enrichment analysis was carried out as described in \cref{sec:pathway_enrichment_analysis}.
-Unfortunately, there were no pathways significantly enriched in any of the \gls{icgc} cancer types when analysed independent of one another.
+
+When the pathway enrichment analysis was carried out with the \gls{kegg} pathway data base, only the ``ABC transporter'' was significantly enriched (\gls{fdr}-adjusted p-value \textless{} 0.05) in the \gls{cesc} data, and no other pathways were enriched in any of the other \gls{icgc} cancer types (\cref{sec:pathways_significant_in_each_of_the_cancer_types}).
+With the Reactome data base, ``Phosphatase bond hydrolysis by NUDT proteins'' (in \gls{blca}) and ``Mitochondrial ABC transporters'' (in \gls{cesc}) were significantly enriched (\cref{sec:pathways_significant_in_each_of_the_cancer_types}).
+With the \gls{go} data base, there were 22 pathways significantly enriched in \gls{blca}, 17 pathways in \gls{cesc}, 3 pathways in \gls{kirp}, 21 pathways in \gls{read}, 10 pathways in \gls{skcm}, 14 pathways in \gls{ucec}, and no pathways were significant in the \gls{coad} and \gls{lihc} cancer types (\cref{sec:pathways_significant_in_each_of_the_cancer_types}).
+
+Though there were many pathways identified as significantly enriched with the \gls{go} data base in each of the cancer types, there were no pathways that were common across the different cancer types.
+Furthermore, these pathways were not similar in terms of the biological activities involved; for example, pathways enriched in the \gls{blca} data suggested possible activation of the \gls{pi3k} pathway, but the pathways enriched in the \gls{read} data were mainly involved in \acrshort{rna} processing (\cref{sec:pathways_significant_in_each_of_the_cancer_types}).
+One interesting result from this enrichment analysis was that the ``ABC transport'' pathway was identified as significantly enriched in the \gls{cesc} data set by all three pathway databases, suggesting that the ``ABC transport'' pathway may be a core component in the \gls{cesc} tumour biology.
+All of these results indicated that the pathways that drive tumour progression were unique to each cancer type, and no one pathway was associated with obesity and tumour biology across multiple cancer types.
 

### 2017-03-20 thesis/thesis.tex
finished rough draft of the result summary section in discussion - tmp commit
Started on limitations section - tmp commit
tmp commit
Finished discussion, now edit everything - tmp commit
Changed all the data set names and removed "'s" - tmp commit
Finished draft appendix A - tmp commit
Divided the appendix file into two - tmp commit
Need to add linear model prediction scatter plots - tmp commit
Added all of the extra results in appendix2; clean up the titles - tmp commit
Draft thesis completed; now edit the whole thing - tmp commit
Edited Intro
Finished editing results1

@@
 \usepackage[acronym]{glossaries}                    % package for making glossaries
-\usepackage[flushleft]{threeparttable}              % package to add notes at the bottom of the table
+\usepackage[flushleft]{threeparttablex}              % package to add notes at the bottom of the table
 \usepackage[font=footnotesize,format=hang]{caption} % package for formatting figure caption
@@
 
+% Set up continuous figures:
+\DeclareCaptionFormat{cont}{#1 (cont.)\par}
+
 % Change the default url font:
@@
 % If you want to compile specific files, uncomment the line below and add the filenames in the {} :
-\includeonly{{./methods/methods},{./results/results},{./results/results2},{./discussion/discussion},{./misc/appendix}}
-% \includeonly{{./results/results},{./results/results2},{./misc/appendix}}
-% \includeonly{{./results/results},{./results/results2},{./discussion/discussion},{./misc/appendix}}
+% \includeonly{{./intro/intro}}
+\includeonly{{./methods/methods},{./results/results2},{./misc/appendix2}}
+% \includeonly{{./results/results}}
+% \includeonly{{./results/results2}}
+% \includeonly{{./discussion/discussion}}
+% \includeonly{{./misc/appendix}}
+% \includeonly{{./misc/appendix2}}
+
+\defcitealias{TCGA}{NCI and NHGRI, 2016}
+\defcitealias{ICGC}{ICGC, 2016}
+\defcitealias{GEO}{NCBI, 2016}
 
@@
 
+\include{./misc/appendix2}
+
 % Use the reference style specified in BiocRefStyle.bst:
@@
 % Print out the bibliography:
-\bibliography{./MSc,./MSc2}
+\bibliography{./MSc,./msc2,./msc3}
 

### 2017-03-22 thesis/discussion/discussion.tex
Almost finished editing discussion - tmp commit

@@
 
-To conclude this thesis, all of the results were re-examined and summarised to highlight the key findings from this project.
-There were many limitations to this project and these limitations were adressed either with the results summary in \cref{sec:summary_of_the_results}, or seaprately in \cref{sec:limitations}.
-Lastly, a conclusion was drawn from all of the evidences from this project and suggestions for future aims and experiments were made.
+To conclude this thesis, all of the results were re-examined and are summarised here to highlight the key findings from this project.
+There were many limitations to this project and these limitations are adressed either in the summary in \cref{sec:summary_of_the_results}, or seaprately in \cref{sec:limitations}.
+Lastly, conclusions are drawn from all of the evidenc from this project and suggestions for future aims and experiments are made.
 
@@
 \label{sec:summary_of_the_results}
-
-% TODO: -- Put into context of very current literature
-% TODO:   	==> How does this project fit into the literature
 
@@
 The obesity metagene generated from the obesity signature from the \citet{Creighton2012} study was able to ``capture'' the overall genetic expression patterns of the samples, where low and high metagene scores  corresponded to low and high gene expressions, respectively.
-This result was in accordance with the characteristic of the obesity metagene provided by Creighton \textit{et al.}.
+This result was in accordance with the characteristics of the obesity metagene provided by \citet{Creighton2012}.
 Furthermore, the obesity metagene significantly correlated with the sample \gls{bmi} and \gls{bmi} status in the CR data set.
 
-To see whether this obesity signature was able to show similar association with the sample \gls{bmi} and \gls{bmi} status in other cancer data sets, metagenes were created in each of the cancer data sets with the transformation matrix generated in the Creighton \textit{et al.} data set.
-Like in Creighton \textit{et al.} data set, the obesity metagenes were reflective of the gene expression patterns of the genes in the genetic signature.
-Although obesity metagene scores reflected the gene expression patterns, the metagene was not significantly associated with the sample \gls{bmi}/\gls{bmi} status in any of the other cancer data sets, except for  the \gls{blca} data set which showed significant association only with the overweight group (discussed later in \cref{sub:blca_and_obesity_metagenes}).
+To see whether this obesity signature was able to show similar association with the sample \gls{bmi} and \gls{bmi} status in other cancer data sets, metagenes were created in each of the cancer data sets with the transformation matrix generated in the CR  data set.
+Like in the CR  data set, the obesity metagenes were reflective of the gene expression patterns of the genes in the genetic signature.
+However, the obesity metagene was not significantly associated with the patient \gls{bmi} or \gls{bmi} status in any of the other cancer data sets, except for  the \gls{blca} data set which showed significant association only with the overweight group (discussed later in \cref{sub:blca_and_obesity_metagenes}).
 
-Initially, these results were thought to be due to the obesity associated genetic signature being specific only to the Creighton \textit{et al.} data set.
-Therefore, a different obesity associated genetic signature from the \citet{Fuentes-Mattei2014} study was used to see whether this signature was able to show significant association with the sample \gls{bmi} and \gls{bmi} status in other cancer data sets.
-The results were similar to the obesity metagene from the \citet{Creighton2012} study; obesity metagene scores reflected the gene expression patterns but the scores were not significantly associated with the sample \gls{bmi}/\gls{bmi} status in any of the cancer data sets (except the \gls{blca} data set).
-Together with the results from CR obesity metagene, these results suggested that both CR and FM genetic signatures were only significant in the original data set in which the signature was derived in (though this was not confirmed in the Fuentes-Mattei \textit{et al.} data set, as no sample \gls{bmi} information was available).
+Initially, these results were thought to be due to the obesity associated genetic signature being specific only to the CR data set.
+Therefore, a different obesity associated genetic signature from the \citet{Fuentes-Mattei2014} study was used to see whether this signature was able to show significant association with the patient \gls{bmi} in other cancer data sets.
+The results were similar to the obesity metagene from the \citet{Creighton2012} study; FM obesity metagene scores reflected the gene expressions of the signature but the scores were not significantly associated with the patient \gls{bmi} or \gls{bmi} status in any of the cancer data sets (except the \gls{blca} data set).
+Together with the results from the CR obesity metagene, these results suggested that both CR and FM genetic signatures were only significant in the original data set in which the signature was derived in (though this was not confirmed in the FM data set, as no patient \gls{bmi} information was available).
 
-The fact that both CR and FM obesity associated genetic signatures showed no sign of significant association with sample \gls{bmi}/\gls{bmi} status in majority of the cancer data sets raised a question of whether these signatures were actually related to obesity, or a different clinical variable within the data set.
-This question led to the identification of the various versions of obesity associated genetic signatures in \cref{sub:identification_of_obesity_associated_genetic_signatures} that were controlled for clinical variables (sex, age, ethnicity, menopause status, tumour grade, \gls{er}/\gls{her2}/\gls{pr} statuses and \gls{ln} status) in the Creighton \textit{et al.} data set.
-All of the metagenes created from these clinical variable-controlled genetic signatures were consistent with the gene expression patterns but none of the metagenes showed significant association with the sample \gls{bmi}/\gls{bmi} status in the data set other than the Creighton \textit{et al.} data set, with the exception of the overweight group in the \gls{blca} data set.
+The fact that both CR and FM obesity associated genetic signatures showed no sign of significant association with the patient \gls{bmi} in majority of the cancer data sets raised a question of whether these signatures were actually related to obesity, or a different clinical variable within the data set.
+This question led to the identification of the various versions of obesity associated genetic signatures in \cref{sub:identification_of_obesity_associated_genetic_signatures} that were controlled by clinical variables (sex, age, ethnicity, menopause status, tumour grade, \gls{er}/\gls{her2}/\gls{pr} statuses and \gls{ln} status) in the CR data set.
+All of the metagenes created from these clinical variable-controlled genetic signatures were consistent with the gene expression patterns, but none of the metagenes showed significant association with the sample \gls{bmi}/\gls{bmi} status in the data set other than the CR data set, with the exception of the overweight group in the \gls{blca} data set.
 
-It was possible that all of these results were due to the sample quality (and hence due to the quality of the data).
-Perhaps many of the samples in these data sets had breast cancers due to a cause unrelated to their obesity status; in other words, some of the breast tumour samples were not driven by obesity.
-This may affect the detection of the truly obesity associated genes, as some of the tumour samples from the patients will have nothing to do with obesity and cancer, and hence misclassified into wrong groups.
+It was possible that all of these results were due to the sample quality (and hence due to the quality of the data), or perhaps many of the samples in these data sets had breast cancers due to a cause unrelated to their obesity status
+As many epidemiological studies have pointed out, obesity is a strong risk factor for cancer (covered in \cref{sec:obesity_and_cancer}), but this does not necessarily mean that all cancer in obese patients developed as a result of obesity.
+In other words, the cancer identified in obese patients may have been caused by another, completely different biological mechanism.
+Thus the tumour samples taken from obese and overweight patients will be a mixture of samples that are driven by obesity-related mechanisms and mechanisms that are not related to obesity.
+These different subgroups of tumours (obesity-driven or not) cannot be easily identified and isolated, which may affect the detection of the truly obesity associated genes, as some of the tumour samples from the patients will have nothing to do with obesity and cancer, and hence misclassified into wrong groups.
 
@@
 In the \gls{blca} data set, many of the obesity metagenes were associated with the overweight group with significant P-value and/or \gls{anova} P-value, but never with the obese group.
-This was unexpected, as all of the metagenes were generated based on the list of \glspl{deg} between the obese and the non-obese groups in the Creighton \textit{et al.} data set, and not the overweight group.
+This was unexpected, as all of the metagenes were generated based on the list of \glspl{deg} between the obese and the non-obese groups in the CR data set, and not the overweight group.
 In addition to the association of the metagenes with the overweight group, some metagenes showed significant association with sample \gls{bmi} in the \gls{blca} data set.
-It was difficult to conclude with confidence that those metagenes that showed significant association with the overweight group in \gls{blca} data set was truly due to the effect of the metagenes, as these metagenes were generated from the obese group and not the overweight group in the original data sets.
+It was difficult to conclude with confidence that those metagenes that showed significant association with the overweight group in the \gls{blca} data set was truly due to the effect of the metagenes, as these metagenes were generated from the obese group and not the overweight group in the original data sets.
 Furthermore, there were no significant association of these metagenes in any other \gls{icgc} cancer data sets and provided no further evidence that supported the results presented in the \gls{blca} data set.
 
-With that said, there is a possibility that the genotypes of the samples that are overweight in \gls{blca} data set are similar to those samples that are obese in the Creighton \textit{et al.} data set.
+With that said, there is a possibility that the genotypes of the samples that are overweight in the \gls{blca} data set are similar to those samples that are obese in the CR data set.
 In fact, \citet{Damrauer2014} have identified ``basal-like'' and ``luminal''  subtypes in bladder cancer that resembled the corresponding subtypes in breast cancer, and suggested that bladder  and breast cancers may have common physiological properties.
@@
 
-The results so far have provided strong evidence that the obesity associated genetic signatures from the Creighton \textit{et al.} and Fuentes-Mattei \textit{et al.} data sets were not associated with the sample \gls{bmi}/\gls{bmi} status.
-In the last attempt to identify any obesity associated genes that were common across multiple cancer types, gene expression analyses were carried out on the \gls{icgc} cancer data sets (\cref{sec:common_genes_across_multiple_cancer_types}).
-There were many \glspl{deg} identified by each of the \gls{icgc} cancer types, though there were no genes in common for all eight cancer types.
-The results from the simulation have confirmed this observation where there were no \glspl{deg} associated with obesity that were common across all eight cancer types.
+The results so far have provided strong evidence that the obesity associated genetic signatures from the CR and the FM data sets were not associated with the patient \gls{bmi}.
+In the last attempt to identify any obesity associated genes that were common across multiple cancer types, gene expression analyses were carried out on the \gls{icgc} cancer data sets using samples for which \gls{bmi} data were available (\cref{sec:common_genes_across_multiple_cancer_types}).
+There were many \glspl{deg} identified by each of the \gls{icgc} cancer types, but there were no genes in common for all eight cancer types, and this observation was supported by the results from the simulation.
 
@@
 However, as observed by the results from the simulation, there were many genes identified by chance alone, which suggested that majority of the \glspl{deg} found in the eight cancer types were false positives.
-This apparently high level of \Glspl{type1} was not only observed in the \gls{icgc} data sets, but also in the Creighton \textit{et al.} data set when gene expression analyses were carried out to generate various obesity associated genetic signatures in \cref{sub:identification_of_obesity_associated_genetic_signatures} (discussed in more details in \cref{sub:false_positives_in_gene_expression_analyses}).
+This apparently high level of \Glspl{type1} was not only observed in the \gls{icgc} data sets, but also in the CR data set when gene expression analyses were carried out to generate various obesity associated genetic signatures in \cref{sub:identification_of_obesity_associated_genetic_signatures} (discussed in more detail in \cref{sub:false_positives_in_gene_expression_analyses}).
 
@@
 There were no pathways enriched in any of the \gls{icgc} data sets, nor the combined \gls{icgc} data set.
-Even though there was no significantly enriched  pathway in the \gls{icgc} data sets, it was evident from the obesity metagenes from Creighton \textit{et al.} and Fuentes-Mattei \textit{et al.} data sets that the obesity metagenes were associated with the sample gene expressions.
+Although there was no significantly enriched  pathway in the \gls{icgc} data sets, it was evident from the obesity metagenes from the CR and FM data sets that the obesity metagenes were associated with the gene expressions of the obesity associated genetic signatures.
 This suggested that the metagenes were able to identify some sort of genetic pattern in the data, even though these metagenes were not significantly associated with obesity.
@@
 
-The fact that there were so many false positives in the \glspl{deg} from the \gls{icgc} data sets indicated that the use of the sample \gls{bmi}/\gls{bmi} status as the ``treatment'' conditions for gene expression analysis was prone to \Glspl{type1}.
-This meant that there were very little evidence of true obesity associated genetic signatures in any of the data sets that have been explored in this project, or perhaps the sample \gls{bmi}/\gls{bmi} status was not enough to pinpoint the underlying biological relationship between obesity and cancer.
+The fact that there were so many false positives in the \glspl{deg} from the \gls{icgc} data sets indicated that the use of the patient \gls{bmi}/\gls{bmi} status as the ``treatment'' conditions (in other words independent variable) for gene expression analysis was prone to \Glspl{type1}.
+This meant that there were very little evidence of true obesity associated genetic signatures in any of the data sets that have been explored in this project, or perhaps the patient \gls{bmi}/\gls{bmi} status was not enough to pinpoint the underlying biological relationship between obesity and cancer (see \cref{sub:discussion_definition_of_obesity} as well).
 
-Evidence of large number of false positives was also seen in \cref{sub:identification_of_obesity_associated_genetic_signatures}, where there were many \glspl{deg} identified from the gene expression analyses when obesity associated genetic signatures were sought for in the Creighton \textit{et al.} data set.
+Evidence of large number of false positives was also seen in \cref{sub:identification_of_obesity_associated_genetic_signatures}, where there were many \glspl{deg} identified from the gene expression analyses when obesity associated genetic signatures were sought for in the CR data set.
 This suggested that there was a significant number of genes that were not related to obesity (false positives) in the genetic signatures.
 Furthermore, the fact that about a third of the genes in any one of the obesity signatures were ``unique'' to each of the individual obesity signatures supported the apparent abundance of false positives in the signatures (\cref{fig:venn1,fig:venn2}).
-From these results, it was likely that the obesity signatures that resulted from the Creighton \textit{et al.} data set had many genes that were false positives, perhaps due to the poor quality of the data set (\cref{sub:quality_of_the_data}).
+From these results, it was likely that the obesity signatures that resulted from the CR data set had many genes that were false positives, perhaps due to the poor quality of the data set (\cref{sub:quality_of_the_data}).
 
-Even if the sample \gls{bmi} and/or \gls{bmi} status was enough to identify the genes that were truly associated with obesity, it would not be a trivial task to identify the truly obesity associated genes from those that were not.
-This raises an important question of whether the sample \gls{bmi}/\gls{bmi} status was an appropriate clinical variable to uncover the true association between obesity and cancer (more in \cref{sec:limitations,sec:future_directions}).
+Even if the patient \gls{bmi} and/or \gls{bmi} status was enough to identify the genes that were truly associated with obesity, it would not be a trivial task to identify the truly obesity associated genes from those that were not.
+This raises an important question of whether the patient \gls{bmi} was an appropriate clinical variable to uncover the true association between obesity and cancer (more in \cref{sec:limitations,sec:future_directions}).
 
@@
 Once the directions of the metagenes were determined, the obesity metagenes were clustered together in the heatmap to visualise the similarity of the obesity metagenes with the pathway metagenes.
-The heatmap revealed that all of the obesity metagenes that have originated from the Creighton \textit{et al.} study clustered together in a group with no other pathway metagenes correlating with these metagenes, which meant that the obesity metagenes were not similar to any of the pathway metagenes from the \citet{Gatza2010a} study.
+The heatmap revealed that all of the obesity metagenes that have originated from the CR data clustered together in a group with no other pathway metagenes correlating with these metagenes, which meant that the obesity metagenes were not similar to any of the pathway metagenes from the \citet{Gatza2010a} study.
 The clustering of the obesity metagene was not surprising, as the results from \cref{sub:_novel_obesity_associated_signatures_and_sample_bmi} already showed that the metagenes were highly correlated with one another.
-This result showed that all of the obesity metagenes created from the Creighton \textit{et al.} data associated with an unknown genetic signature that was not related to any of the pathway associated genetic signatures from the \citet{Gatza2010a} study.
+This result showed that all of the obesity metagenes created from the CR data associated with an unknown genetic signature that was not related to any of the pathway associated genetic signatures from the \citet{Gatza2010a} study.
 
-The obesity metagene from the \citet{Fuentes-Mattei2014} study was a cluster of its own, where the metagene did not group with any of the obesity metagenes derived from the Creighton \textit{et al.} data set, nor with any of the Gatza \textit{et al.} pathway metagenes.
-Though Fuentes-Mattei \textit{et al.} suggested a biological link with the Akt/\gls{mtor} pathway in their study, the result from the heatmap showed otherwise.
+The obesity metagene from the \citet{Fuentes-Mattei2014} study was a cluster of its own, where the metagene did not group with any of the obesity metagenes derived from the CR data set, nor with any of the GT pathway metagenes.
+Though Fuentes-Mattei \textit{et al.} suggested a biological link with the Akt/\gls{mtor} pathway in their study, the result from the heatmap did not support that finding.
 This may have been due to the inconsistency of the pathway metagene scores across different data sets as mentioned in \cref{sec:pathway_associated_metagenes_and_obesity_associated_metagenes}.
@@
 
-To further investigate whether there were any evidence of a pathway signature being associated with the obesity associated genetic signatures, linear models were created in Cris' data set with the pathway metagene scores.
-First, linear models were created with the sample \gls{bmi}, \gls{bmi} status and a selection of the most ``consistent'' pathway metagene scores (based on the consistency across different data sets).
-From these linear models, \gls{pr} pathway metagene scores showed significant contribution to the model in predicting the obesity associated metagene scores in all of the linear models, with the exception of the FM obesity metagene (where Myc was the significant variable; \cref{app:b}).
+To further investigate whether there was any evidence of a pathway signature being related to the obesity associated genetic signatures, linear models were created in the \gls{nzbc} data set with the pathway metagene scores.
+First, linear models were created with the patient \gls{bmi}, \gls{bmi} status and a selection of the most ``consistent'' pathway metagene scores (based on the consistency across different data sets).
+From these linear models, \gls{pr} pathway metagene scores showed significant contribution to the model in predicting the obesity associated metagene scores in all of the linear models, with the exception of the FM obesity metagene (where Myc was the significant variable; \cref{sec:summary_of_the_linear_models_in_nzbc_data}).
 
-This was not surprising since all but the FM obesity metagene were derived from the Creighton \textit{et al.} data set, and therefore the FM obesity metagene was likely to be different to those from the Creighton \textit{et al.} data set.
-Adding to this, difference between the CR and FM metagenes were also shown in the heatmap of the obesity and pathway metagenes mentioned earlier, where the FM metagene clustered with neither the CR obesity metagenes nor with any other pathway metagenes.
+This was not surprising since all but the FM obesity metagene were derived from the CR data set, and therefore the FM obesity metagene was likely to be different to those from the CR data set.
+Adding to this, differences between the CR and FM metagenes were also shown in the heatmap of the obesity and pathway metagenes mentioned earlier, where the FM metagene clustered with neither the CR obesity metagenes nor with any other pathway metagenes.
 This implied that the FM metagene differed fundamentally from the CR obesity metagenes.
-Nevertheless, to clarify the significance of the \gls{pr} pathway with the obesity metagenes, linear models were also created with \gls{pr} pathway metagene in combination with the sample \gls{bmi}/\gls{bmi} status.
+Nevertheless, to clarify the significance of the \gls{pr} pathway with the obesity metagenes, linear models were also created with \gls{pr} pathway metagene in combination with the patient \gls{bmi} and/or \gls{bmi} status.
 
-The predictions of the obesity metagene scores were made using the linear models created in Cris' data set.
-All of the predicted obesity metagene scores significantly correlated with the true obesity metagene scores in both Cris' (training) and Creighton's (testing) data sets.
-However, the $R^2$-values for all of these predictions were low (highest $R^2$ $\approx$ 0.42; \cref{app:b}) with highly variable data points in the scatter plots which suggested that, although the predictions reflected the true values significantly, the variables included in the linear models may not be relevant to the obesity metagenes as it may seem from these plots.
+The predictions of the obesity metagene scores were made using the linear models created in the \gls{nzbc} data set.
+All of the predicted obesity metagene scores significantly correlated with the true obesity metagene scores in both the \gls{nzbc} (training) and CR (testing) data sets.
+However, the $R^2$-values for all of these predictions were low (highest $R^2$ $\approx$ 0.42; \cref{sec:summary_statistics_of_the_predicted_obesity_metagenes_with_sample_bmi_bmi_status_in_nzbc_and_cr_data}) with highly variable data points in the scatter plots.
+It was unclear from these predictions and plots whether the models explained the true underlying biological mechanisms of the obesity associated genetic signatures.
+Nevertheless, the models did explain some variations in the obesity metagenes, but there may be some other unknown biological processes that provide more information about the signatures.
 
-Since linear models were created only with the ``consistent'' pathway metagene scores, a step wise method was used to generate linear models that included all of the variables that allowed the model to predict the true obesity metagene the best.
+Since linear models were created only with the ``consistent'' pathway metagene scores, a step wise method was used to generate linear models that included the patient \gls{bmi}, \gls{bmi} status and all of the pathway metagenes that allowed the model to predict the true obesity metagene the best.
 As shown by the results in \cref{sec:prediction_of_obesity_associated_metagene_with_pathway_associate_metagene}, all of the models predicted the true obesity metagene scores with high $R^2$-value in Creighton's data set (highest $R^2$ $\approx$ 0.88) and were statistically significant.
 This meant that the variables that were significant in the linear models played a role in the obesity associated genetic signatures.
-There was no single pathway signature that was common to all of the linear models, but Akt, \gls{egfr} and \gls{pr} pathway metagenes were included in many of the models, which suggested that these pathway signatures may have an important role in the obesity associated genetic signatures.
+There was no single pathway signature that was common to all of the linear models, but Akt, \gls{egfr} and \gls{pr} pathway metagenes were included in many of the models, which suggested that these pathway signatures may have an important role in the obesity associated genetic signatures (\cref{sub:stepwise_linear_model_prediction_in_nzbc_and_cr_data_sets}).
 
@@
 
-The linear models that were created from the ``consistent'' pathway signatures in Cris' data set showed significant association of the \gls{pr} pathway with the obesity metagenes.
-But was this due to the true biological relationship between the \gls{pr} pathway and the obesity metagenes, or only because the models were created in Cris' data set?
-In fact, the proportion of the samples that were \gls{pr}$^+$ was greater in Cris' data set (62 \gls{pr}$^+$ and 37 \gls{pr}$^-$ samples) compared with the Creighton's data set (48 \gls{pr}$^+$ and 50 \gls{pr}$^-$ samples), which implied that the Cris' data set was biased towards \gls{pr} status and hence \gls{pr} pathway signature.
+The linear models that were created from the ``consistent'' pathway signatures in the \gls{nzbc} data set showed significant association of the \gls{pr} pathway with the obesity metagenes.
+But was this due to the true biological relationship between the \gls{pr} pathway and the obesity metagenes, or only because the models were created in the \gls{nzbc} data set?
+In fact, the proportion of the samples that were \gls{pr}$^+$ was greater in the \gls{nzbc} data set (62 \gls{pr}$^+$ and 37 \gls{pr}$^-$ samples) compared with the CR data set (48 \gls{pr}$^+$ and 50 \gls{pr}$^-$ samples), which implied that the patients in the \gls{nzbc} data set was more likely to show association with the \gls{pr} pathway metagene.
 
-However, with this logic, \gls{er} pathway signature would have also been biased, as the  sample  \gls{er} status was similar to the \gls{pr} status in either data sets: 72 and 58 \gls{er}$^+$ and 27 and 42 \gls{er}$^-$ samples in Cris' and Creighton's data sets, respectively.
+However, with this logic, the \gls{er} pathway signature would as likely to be associated, as the  sample  \gls{er} status was similar to the \gls{pr} status in either data sets: 72 \gls{er}$^+$ and 27 \gls{er}$^-$ samples in the \gls{nzbc} data, and 58 \gls{er}$^+$ and  42 \gls{er}$^-$ samples in the  CR data set.
 The fact that the \gls{er} pathway signature was not significant in the models suggested that the obesity metagenes were related to the \gls{pr} pathway irrespective of the sample \gls{pr} status.
-With that said, the \gls{pr} status of a sample does not necessarily reflect the level of the \gls{pr} pathway activity; the samples in Cris' data may have stronger \gls{pr} characteristics than the Creighton's data and therefore cannot deny the possibility of the bias.
+Furthermore, both the \gls{nzbc} and the CR data sets showed strong association between the \gls{pr} and \gls{er} pathway metagenes with the patient \gls{pr} and \gls{er} status, respectively (\cref{})(add reference to pr and er graph).
+Since both \gls{er} and \gls{pr} signatures were significantly associated with the patient \gls{er} and \gls{pr} satuses, both signatures would have been as equally likely to show up significant in the models, supporting the idea that the \gls{er} and \gls{pr} statuses were independent of the apparent significance of these signatures in the models.
 
@@
 With that said, even if these variables were related to the obesity associated genetic signatures, the fact that these results depended on the pathway associated genetic signatures from the \citet{Gatza2010a} study makes the results questionable.
-This was due to the lack of consistency of the metagenes across different cancer data sets, where more than half of the pathway metagenes showed a correlation of less than 0.8 in at least one data set (\cref{sec:pathway_associated_metagenes_and_obesity_associated_metagenes,app:b}).
-This meant that at least half of the signatures were specific to the Gatza \textit{et al.} data set and could be highly variable in other data sets.
-Thus, without further validation of the quality of the pathway associated genetic signatures and verification of the roles of these signatures in the biological relationship between obesity and cancer, these results should be interpreted with great caution.
+This was due to the lack of consistency of the metagenes across different cancer data sets, where more than half of the pathway metagenes showed a correlation of less than 0.8 in at least one data set (\cref{sec:pathway_associated_metagenes_and_obesity_associated_metagenes}).
+This meant that at least half of the signatures were specific to the GT data set and could be highly variable in the other data sets, reflecting the potential differences in the cohorts (at the molecular level).
+Thus, without further validation of the quality of the pathway associated genetic signatures and verification of the roles of these signatures in the biological relationship between obesity and cancer, these results should be interpreted with care.
 
@@
 
-% There were many limitations to this project, though some have already been mentioned in \cref{sec:summary_of_the_results}.
-
 \subsection{Definition of obesity}
@@
 
-First and possibly the most important limitation to this project was the use of \gls{bmi} as a measurement for central adiposity and its subsequent classification of the samples based on \gls{bmi}.
+First and possibly the most important limitation to this project was the use of \gls{bmi} as a measurement for obesity and its subsequent classification of the patients based on \gls{bmi}.
+Obesity is a term given to those individuals with excessive amount of fat and can be defined in various ways.
 It was important to define the obese phenotype accurately, as all of the obesity associated genetic signatures, gene expression analyses and pathway enrichment analysis were dependent on the patient \gls{bmi} status.
-\Gls{bmi} is perhaps the most popular measurement to estimate the central adiposity owing to the relative ease of measurement in clinical settings, and hence used in various clinical trials and large epidemiological studies as an indication of the patients' metabolic state.
-However, there have been studies that showed other measurements, such as waist-to-height ratio and waist-to-hip ratio, to be a better representation of the central adiposity of the patient \citep{Dalton2003,Lee2008}.
+\Gls{bmi} is perhaps the most popular measurement to estimate obesity owing to the relative ease of measurement in clinical settings, and hence used in various clinical trials and large epidemiological studies as an indication of the patients' metabolic state.
+However, there have been studies that showed other measurements, such as waist-to-height ratio and waist-to-hip ratio, to be a better representation of the patients' obesity status \citep{Dalton2003,Lee2008}.
 
 Perhaps the reason why there were so many false positives when gene expression analyses were carried out in \cref{sub:false_positives_in_gene_expression_analyses} was because \gls{bmi} was used to classify the patients into their corresponding \gls{bmi} statuses, instead of another measurement.
-It was likely that the sample \gls{bmi} (and therefore \gls{bmi} status) was not good enough a phenotypic characteristic to describe the underlying biology driven by obesity.
-Since the gene expression analyses were based on such ``vague'' parameter, it may have picked up so many genes that were apparently related to obesity when in fact the genes were not related to obesity at all.
+It was likely that the patient \gls{bmi} (and therefore \gls{bmi} status) was not good enough a phenotypic characteristic to describe the underlying biology driven by obesity.
+Since the gene expression analyses were based on such ``vague'' parameter, it may have picked up many genes that were apparently related to obesity when in fact the genes were not related to obesity at all.
 
@@
 Additional measurements such as blood lipid profiles and insulin levels may also help understand the underlying biological state in which the patients are in, and would greatly help focus on the genes that are truly affecting the patients.
-However, these measurements will be difficult to carry out in a systematic fashion in a clinical setting, and to collate existing patient data with these measurements to make a data set that is large enough for the analyses done in this project may be impossible.
+However, these measurements are difficult to carry out in a systematic fashion in a clinical setting, and to collate existing patient data with these measurements to make a data set that is large enough for the analyses done in this project may be impossible.
 Nevertheless, even the measurement of waist circumference in addition to \gls{bmi} may significantly help describe the obesity phenotype more accurately in future investigations.
@@
 
-Second limitation was the quality of the data sets this project have used.
+Another limitation of this work was the quality of the data sets.
 All of the data sets used in this project were from an online and publicly available source, and the types of data sets used were either microarray or \gls{rnaseq} data.
-First potential problem with using an open source data from other research groups is that the data is generated by multiple different groups with different technologies and experimental protocols.
-Since different laboratories have their own way to carry out certain experiments, the sequence data from each of the studies used in this project may have been produced and processed in a completely different manner, not to mention about the use of different patient cohorts in each experiment.
+The first potential problem with using open source data from other research groups is that the data are generated by multiple different groups with different technologies and experimental protocols.
+Since different laboratories have their own way to carry out certain experiments, the sequence data from each of the studies used in this project may have been produced and processed in a completely different manner, not to mention the use of different patient cohorts in each experiment.
 
 For example, in the \citet{Creighton2012} study, the patient breast tumour biopsies were ``trimmed such that all the samples had $\geq$ 70\% tumour cells'' in the samples.
-On the other hand, \citet{Fuentes-Mattei2014} only mentioned that they used breast tumour biopsies from the patient with no comment about the quality or quantity of the biopsies.
-However, they took one third of the tumours from the mouse models to analyse the transcriptome of the mouse tumours, so perhaps Fuentes-Mattei \textit{et al.} used similar approach with the human breast tumour samples.
-Without a doubt, the difference in how the samples were handled affected the quality of the data, which made it challenging to compare different data sets and it may even be possible for the samples within the data set to be incomparable.
-In fact, a couple of observations were made in this project that suggested the Creighton \textit{et al.} data set was of poor quality (see \cref{sub:false_positives_in_gene_expression_analyses}).
+\citet{Fuentes-Mattei2014} only mentioned that they used breast tumour biopsies from patients with no comment about the quality or quantity of the biopsies.
+Without a doubt, the difference in how the samples were handled can affect the quality of the data, and make it challenging to compare different data sets \citep{Irizarry2005}.
+In fact, observations were made in this project that suggested the data set from the \citet{Creighton2012} study was of poor quality (see \cref{sub:false_positives_in_gene_expression_analyses}).
 
-Although the underlying quality of the data provided by these studies could not be refined in any way, best efforts were made to make sure the data sets were free of other factors that may affect the results.
-% Although the underlying quality of the data provided by these studies could not be refined in any way, best efforts were made to make sure the results from these data sets were reliable.
-This leads to the second problem: the different sequencing technologies used in the studies.
-All of Creighton \textit{et al.}, Fuentes-Mattei \textit{et al.}, Print \textit{et al.} and Gatza \textit{et al.} data sets were analysed with the microarray technology, whereas all of the \gls{icgc} cancer data sets were sequenced with \gls{ngs} technology.
+Although the underlying quality of the data provided by these studies could not be improved in any way, best efforts were made to  ensure the data sets were free of other factors that may affect the results.
+This leads to the second problem: the different technologies used in the studies.
+All of Creighton \textit{et al.}, Fuentes-Mattei \textit{et al.}, Print \textit{et al.} and Gatza \textit{et al.} data sets were analysed using microarray technology, whereas all of the \gls{icgc} cancer data sets were generated with \gls{ngs} technology.
 As mentioned in \cref{sub:data_normalisation}, the two technologies use fundamentally different principals to output sequence information of the sample; microarray uses light intensity and \gls{ngs} uses count data.
@@
 This enabled the standard \textit{limma} analysis pipelines, normally used for the analysis of microarray data, to be used by the \gls{icgc} cancer \gls{rnaseq} data sets.
-In this way, the results from the two types of data sets were fairly reliable and consistent as the same analysis methods were applied to the data sets.
-In addition to the difference in the sequencing technologies, the Gatza \textit{et al.} data set comprised of multiple different data sets, each from different sources.
-Since individual data set had their own experimental differences (``batch effect''), it was important to correct for these differences when multiple data sets were combined into one (see \cref{sub:batch_correction}).
+In this way, the results from the two types of data sets were fairly reliable and consistent, as the same analysis methods were applied to the data sets.
+In addition to the difference in the sequencing technologies, the Gatza \textit{et al.} data set comprised multiple different data sets, each from different sources.
+Since individual data sets had their own experimental differences (for example, ``batch effects''), it was important to correct for these differences when multiple data sets were combined into one (see \cref{sub:batch_correction}).
 In this project, batch corrections were made in the Gatza \textit{et al.} data set and the combined \gls{icgc} cancer data sets (used in \cref{sec:pathways_enriched_in_icgc_data_sets}) to eliminate the batch effects.
-
-It was evident from some of the results that the Creighton \textit{et al.} data set was of poor quality, and the validity of the other data sets (and therefore the results generated from these data sets) were also debatable, despite all of the attempts made to control for other factors that may have affected the results.
 
@@
 Firstly, the obesity signatures identified by both Creighton \textit{et al.} and Fuentes-Mattei \textit{et al.} relied on the definition of obesity based on the sample \gls{bmi} values.
-As discussed in \cref{sub:discussion_definition_of_obesity}, the use of \gls{bmi} may not have been the best measurement for central adiposity.
-In fact, both Creighton \textit{et al.} and Fuentes-Mattei \textit{et al.} signatures associated with different pathway associated signatures in \cref{sec:pathway_associated_metagenes_and_obesity_associated_metagenes,sec:prediction_of_obesity_associated_metagene_with_pathway_associate_metagene}, which suggested that obesity as defined by \gls{bmi} was not enough to reveal the underlying biological characteristic of obesity, and hence showed association with different pathway signatures.
-Furthermore, as mentioned in \cref{sub:quality_of_the_data}, it was likely that the Creighton \textit{et al.} data set (and perhaps other data sets as well) was of poor quality and may have affected the ability of the generated signatures as a marker for obesity in cancer.
+As discussed in \cref{sub:discussion_definition_of_obesity}, the use of \gls{bmi} may not have been the best measurement for obesity.
+In fact, both CR and FM obesity signatures associated with different pathway signatures in \cref{sec:pathway_associated_metagenes_and_obesity_associated_metagenes,sec:prediction_of_obesity_associated_metagene_with_pathway_associate_metagene}, which suggested that obesity as defined by \gls{bmi} was not enough to reveal the underlying biological characteristics of obesity, and hence showed association with different pathway signatures.
+Furthermore, as mentioned in \cref{sub:quality_of_the_data}, it was likely that the CR data set (and perhaps other data sets as well) was of poor quality and may have affected the ability of the generated signatures as a marker for obesity in cancer.
 
-Secondly, the method in which the pathway associated genetic signatures was defined in the Gatza \textit{et al.} study may have been questionable.
-In their studies, Gatza \textit{et al.} used model cell lines to identify the pathway associated genetic signatures by altering the expression of the representative gene for the pathway and decided all of the genes affected by the change in representative gene expression were associated with that pathway \citep{Gatza2010a} (add Gatza's old paper).
-However, some of these pathway associated genetic signatures were derived in different cell lines to one another, which meant that it was possible for some of the pathway signatures to be defined incorrectly.
+Secondly, the method in which the pathway associated genetic signatures was defined in the \citet{Gatza2010a} study may have been questionable.
+In their studies, Gatza \textit{et al.} used model cell lines to identify the pathway associated genetic signatures by altering the expression of the representative gene for the pathway and assigned all of the genes affected by the change in representative gene expression to be associated with that pathway \citep{Bild2006,Gatza2010a}.
+However, some of these pathway associated genetic signatures were derived in different cell lines to one another, which meant that some pathway signatures may not reflect the true pathway activity \textit{in vivo}.
 In fact, as seen in \cref{sec:pathway_associated_genetic_signatures_from_gatza2010a_study,sec:pathway_associated_metagenes_and_obesity_associated_metagenes}, some pathway signatures produced more consistent metagenes from the data sets than the other signatures.
-These results were merely an indication that the pathway signatures may have been generated incorrectly and it was not possible to determine whether these signatures were truly associated to the corresponding pathways as the original data was not available to reference the signatures back for validation.
-
-Taken together, the quality of both the obesity and the pathway associated genetic signatures may not have been good and, in combination with poor quality data, may have produced unexpected results.
+These results were merely an indication that the pathway signatures may have reflected the pathway activities poorly, and it was not possible to determine whether these signatures were truly related to the corresponding pathways, as the original data was not available to reference the signatures back for validation.
+With that said, few of the pathway signatures correlated well with some clinical variables (such as \gls{er} and \gls{pr}), so not all signatures may have been of poor quality (\cref{}) (add reference to pr and er graph).
 
@@
 
-There were two main aims to this project: firstly to determine whether there were any obesity specific genetic signatures that could be tranferred across multiple cancer types; and secondly to investigate whether there were any biological pathways dysregulated in the samples that were obese compared to the samples that were not obese.
+There were two main aims to this project: firstly to determine whether there were any obesity specific genetic signatures that could be tranferred across multiple cancer types; and secondly to investigate whether there were any biological pathways dysregulated in the tumour samples from the patients that were obese compared to the samples from non-obese patients.
 These aims were addressed in \cref{cha:obesity_genetic_signatures_and_cancer,cha:obesity_associated_genetic_signature_and_pathway_signatures}, resepectively.
 
-As shown clearly by the results from \cref{cha:obesity_genetic_signatures_and_cancer}, there seemed to be no  genetic signature that were able to differntiate between the samples that were obese from those that were not, across different cancer types.
+As shown clearly by the results from \cref{cha:obesity_genetic_signatures_and_cancer}, there seemed to be no  genetic signatures that were able to differntiate between the tumour samples from the patients that were obese from those that were not, across different cancer types.
 Furthermore, there were no genes differentially expressed between obese and non-obese patients that were common across multiple cancer types.
@@
 
-This project have managed to show that there were no obesity specific genes  that were common across multiple cancer types.
-Although this study failed to identify obesity specific genes that affected tumour biology, it is likely that there is some complex mechanism underlying the relationship between obesity and cancer; otherwise many of the clinical and epidemiological associations between obesity and cancer would not make any sense.
+This project has managed to show that there were no obesity specific genes  that were common across multiple cancer types.
+Although this study failed to identify obesity specific genes that affected tumour biology, it is likely that there is some complex mechanism underlying the relationship between obesity and cancer; otherwise many of the clinical and epidemiological associations between obesity and cancer risk would not make any sense.
 Since Akt, \gls{egfr} and \gls{pr} pathways were signicant in the linear models that predicted the obesity metagenes, more thorough investigation of these pathways may be a good start to clarify the relationship.
-Further analyses of \gls{blca} may help clarify the similarity between bladder and breast cancer types and provide an insight into the association between the obesity associated genetic signatures and \gls{blca}.
+Further analyses of \gls{blca} may help clarify the similarity between bladder and breast cancer types and provide insight into the association between the obesity associated genetic signatures and \gls{blca}.
 

### 2017-03-22 thesis/intro/intro.tex
Finished editing results1
Almost finished editing discussion - tmp commit

@@
 				\item [2] Table adapted from \citet{Roberts2010}, where the original data in the table was from \citet{Renehan2008}.
-				\item [3] Not applicable.
+				\item [3] Not available.
 			\end{footnotesize}
@@
 These pathway specific genes were identified experimentally by altering the expression of one specific gene that is central to the pathway in a model cell culture system.
-For example, all of the genes that have altered its expression in the cells that had  the \textit{Myc} gene induced were noted down as the genes involved in the Myc pathway.
+For example, all of the genes that have altered its expression in the cells that had  the \textit{MYC} gene induced were noted down as the genes involved in the Myc pathway.
 

### 2017-03-22 thesis/misc/appendix2.tex
Almost finished editing discussion - tmp commit

@@
+\begin{appendices}
+	\crefalias{chapter}{appcha}
+	\crefalias{section}{appsec}
+	\renewcommand{\thesection}{\Alph{chapter}\arabic{section}}
+
 	\chapter{Additional results from \cref{cha:obesity_associated_genetic_signature_and_pathway_signatures}}
@@
 	To investigate whether the normalisation methods were important in the creation of the pathway associated genetic signature transformation matrices, and whether the normalisation methods of the data affected the resulting metagenes, transformation matrices were generated in \gls{rma} or \gls{mas} normalised GT data set and then applied to the \gls{rma} or \gls{mas} normalised GT data set.
-	The results in \cref{fig:appendix/gt_meta_rma_mas} showed less variability of the metagene scores when the metagenes were derived from the data with the same normalisation method (for example, metagenes derived from \gls{rma} normalised data), regardless of the transformation matrix used to produce the metagenes.
+	The results in \cref{fig:appendix/gt_meta_rma_mas} showed less variability of the metagene scores when the metagenes were derived from the data with the same normalisation method, regardless of the transformation matrix used to produce the metagenes.
 	This was most clearly demonstrated by the \gls{bcat}, \gls{er}, \gls{ifna}, \gls{ifny}, p53 and \gls{pr} pathway associated genetic signatures.
+
+	As mentioned in \cref{sec:pathway_associated_genetic_signatures_from_gatza2010a_study} some of the pathway metagenes showed greater variability than the other pathway metagene scores.
+	This was also seen in other data sets as well (only the \gls{pr} and \gls{tgfb} metagenes shown; \cref{fig:appendix/gt_meta_rma_mas_other_data}).
 
@@
 
+	\begin{figure}[htp!]
+		\centering
+		\includegraphics[width=0.32\linewidth,page=244]{appendix/allmeta_rma_vs_mas}
+		\includegraphics[width=0.32\linewidth,page=412]{appendix/allmeta_rma_vs_mas}
+		\includegraphics[width=0.32\linewidth,page=580]{appendix/allmeta_rma_vs_mas}\\
+		\includegraphics[width=0.32\linewidth,page=268]{appendix/allmeta_rma_vs_mas}
+		\includegraphics[width=0.32\linewidth,page=436]{appendix/allmeta_rma_vs_mas}
+		\includegraphics[width=0.32\linewidth,page=604]{appendix/allmeta_rma_vs_mas}
+		\caption[Comparison of the pathway metagenes generated in the \acrshort{mas} normalised CR, FM and \gls{nzbc} data sets from the application of the transformation matrices derived from either the \acrshort{rma} or \acrshort{mas} normalised GT data]{Comparison of the pathway metagenes generated in the \acrshort{mas} normalised CR, FM and \gls{nzbc} data sets from the application of the transformation matrices derived from either the \acrshort{rma} or \acrshort{mas} normalised GT data.
+	Only the \gls{pr} (top) and the \gls{tgfb}  (bottom) pathway metagenes are shown, due to the large number of scatter plots generated.
+	}
+		\label{fig:appendix/gt_meta_rma_mas_other_data}
+	\end{figure}
+
 	\newpage
@@
 
-	The directionality of the pathway metagenes were determined by comparing the metagene values with the gene expression of the representative gene for the pathway associated genetic signature, as well as the gene expression pattern of the pathway signature (\cref{fig:appendix/gt_meta_rma_mas}).
+	The directionality of the pathway metagenes were determined by comparing the metagene values with the gene expression of the representative gene for the pathway associated genetic signature, as well as the gene expression pattern of the pathway signature (\cref{fig:appendix/gt_pathmeta_rank}).
 	Together with the original results presented by \citet{Gatza2010a} (\cref{sec:result_from_gatza2010a_study}), the final directions of the pathway metagenes were decided.
@@
 		}
-		\label{fig:appendix/gt_meta_rma_mas}
+		\label{fig:appendix/gt_pathmeta_rank}
 	\end{figure}
@@
 		High and low correlation were represented as red and blue, respectively, where the colours were matched with the values on the scale shown in the top left histogram.}
-		\label{fig:appendix/gt_meta_rma_mas}
+		\label{fig:appendix/gt_pathmeta_other_data}
 	\end{figure}
@@
 		}
-		\label{fig:appendix/gt_meta_rma_mas}
+		\label{fig:appendix/all_meta_other_data}
 	\end{figure}
@@
 	\label{sec:summary_statistics_of_the_predicted_obesity_metagenes_with_sample_bmi_bmi_status_in_nzbc_and_cr_data}
+
+	Instead of presenting 11 scatter plots for the 10 obesity metagenes used in this project, all of the R$^2$- and p-values were summarised in \cref{tab:prediction_summary_cris}.
 

### 2017-03-22 thesis/results/results2.tex
Almost finished editing discussion - tmp commit

@@
 In doing so, the pathway genetic signatures from the \citet{Gatza2010a} study (GT) were utilised to determine which biological pathways the obesity associated genetic signatures were most similar to.
-First, the direction of GT pathway associated genetic signatures were resolved; then the pathway associated metagenes were compared with the obesity associated metagenes; and lastly linear models were constructed based on the pathway metagenes and sample \gls{bmi}/\gls{bmi} status to predict the obesity associated metagenes.
+First, the direction of GT pathway associated genetic signatures were resolved, then the pathway associated metagenes were compared with the obesity associated metagenes, and lastly linear models were constructed based on the pathway metagenes and patient \gls{bmi} to predict the obesity associated metagenes.
 
@@
 
-Pathway associated genetic signatures from the \citet{Gatza2010a} study were examined for their consistency with the reported results.
-In the \citet{Gatza2010a} study, their data comprised of samples from different data sets from other studies, \gls{mas}-normalised and the metagene scores were ranked with probit.
-However, the analyses so far have used \gls{rma}-normalisation method and ranked based on the number of samples present in the data.
-To decide which normalisation or ranking methods were suitable for the analysis, the different methods were compared in GT data (see \cref{app:b}).
-Results shown in \cref{app:b} clarified that there was no significant difference in the ranking methods used.
-Correlation and scatter plots of the different combinations of transformation matrices (derived from either \gls{rma}- or \gls{mas}-normalised data) with \gls{rma}- or \gls{mas}-normalised data were calculated and plotted.
-From these results, it seemed like the normalisation methods of the data in which the transformation matrices were applied to had the most significant effect on the resulting metagenes, rather than the transformation matrices themselves (\cref{app:b}).
+Pathway associated genetic signatures from the \citet{Gatza2010a} study were first examined in the GT data to make sure the results from their paper was reproducible.
+
+\subsection{Ranking and normalisation methods}
+\label{sub:ranking_and_normalisation_methods}
+
+In the \citet{Gatza2010a} study, their data comprised of samples from different data sets from other studies, \gls{mas}-normalised and the metagene scores were transformed to a scale between 0 to 1 using a probit function.
+However, the analyses so far have used \gls{rma} normalisation method and ranked based on the number of samples present in the data (fractional ranking; \cref{sub:ranking_of_the_metagene_scores}).
+To decide which normalisation methods or ranking approaches were suitable for the analysis, the different methods were compared in the GT data (\cref{sec:ranking_method_for_the_pathway_associated_genetic_signatures,sec:normalisation_method_for_pathway_associated_genetic_signature_transformation_matrices}).
+Results shown in \cref{sec:ranking_method_for_the_pathway_associated_genetic_signatures} clarified that there was no significant difference in the ranking approaches used.
+Correlation and scatter plots of the different combinations of transformation matrices (derived from either the \gls{rma}- or \gls{mas}-normalised GT data) with the  GT data normalised with the different methods were calculated and plotted.
+From these results, it was clear the normalisation methods used had the most significant effect on the resulting metagenes, rather than the transformation matrices themselves (\cref{sec:normalisation_method_for_pathway_associated_genetic_signature_transformation_matrices}).
 This meant that all of the data sets had to be normalised with a single normalisation method, so the \gls{rma} normalisation method was used as this method was more reliable than the \gls{mas} method.
-Therefore, GT data set was batch corrected first (\cref{sub:batch_correction}), then normalised with \gls{rma} method and the scores were ranked based on the number of samples in the data set.
+%TODO: add reference for the above sentence
+Therefore, the GT data set was batch corrected first (\cref{sub:batch_correction}), then normalised with the \gls{rma} method and the metagene scores were ranked with fractional ranking (\cref{sub:ranking_of_the_metagene_scores}).
 
-One thing noted from these scatter plots was that there were some pathway signatures that were more variable than the others.
+One thing noted from these scatter plots was that there were some pathway signatures that were more variable than the others (\cref{sec:normalisation_method_for_pathway_associated_genetic_signature_transformation_matrices}).
 As an example, the \gls{tgfb} metagenes were significantly more dispersed compared to the \gls{pr} metagenes, even though both of these metagenes were generated in the same data set through similar processes (\cref{fig:gt_rma_vs_mas}).
-These differences were seen in other data sets as well (\cref{app:b}).
-This result provided evidence that some of the pathway genetic signatures from the \citet{Gatza2010a} study were more consistent across different data sets than the others.
-\\
+These differences were seen in other data sets as well (\cref{fig:appendix/gt_meta_rma_mas_other_data}).
+This result provided evidence that some of the pathway genetic signatures from the \citet{Gatza2010a} study were less variable across different data sets than the others.
 
@@
 	\caption[Comparison of the pathway metagenes generated in the \acrshort{mas} normalised GT data set from the application of the transformation matrices derived from either the \acrshort{rma} or \acrshort{mas} normalised GT data]{Comparison of the pathway metagenes generated in the \acrshort{mas} normalised GT data set from the application of the transformation matrices derived from either the \acrshort{rma} or \acrshort{mas} normalised GT data.
-	Only the \gls{bcat} (left) and the \gls{tgfb}  (right) pathway metagenes are shown.
+	Only the \gls{pr} (left) and the \gls{tgfb}  (right) pathway metagenes are shown.
 	The scatter plots for the other pathway metagenes are shown in \cref{sec:normalisation_method_for_pathway_associated_genetic_signature_transformation_matrices}.
@@
 
-\noindent
-Before GT pathway metagenes were compared with the obesity associated metagenes, the direction of GT pathway metagenes had to be checked to make sure the metagenes were in the correct direction (\cref{sub:metagene_direction}).
-% Each pathway metagenes was generated from GT data set and the direction of the metagene was checked with the expression of the representative pathway gene (\cref{tab:metagene_direction}), and the groupings of the pathways were considered as well.
+\subsection{Pathway metagene directionality}
+\label{sub:pathway_metagene_directionality}
+
+Before GT pathway metagenes were compared with the obesity associated metagenes, the direction of the GT pathway metagenes had to be checked to make sure the metagenes were in the ``correct'' direction (\cref{sub:metagene_direction}).
 The correlation of all the pathway metagenes with one another were plotted as a heatmap in \cref{fig:gatza_meta_dir}.
 The most prominent group had five pathways (E2F1, \gls{pi3k}, Myc, \gls{bcat} and Ras) that clustered at the top right hand corner of the heatmap.
-% TODO: add additional info about why these pathways clustered together?
-Other groups included \gls{ifna}/\gls{ifny}/\gls{tnfa} pathways, \gls{er}/\gls{pr}/p53 pathway and p63/\gls{her2} pathways.
+Other groups included \gls{ifna}/\gls{ifny}/\gls{tnfa} pathways, \gls{er}/\gls{pr}/p53 pathways and p63/\gls{her2} pathways.
 In addition to these highly correlated groups, \gls{stat3}/\gls{tgfb}/Src/\gls{egfr}/Akt pathways showed little correlation with one another.
-Comparing these groups with the results presented by \citet{Gatza2010a} (see \cref{app:b}), the identified groups approximately resembled the groups identified in their study, which confirmed that the directions of GT pathway metagenes were similar to those used by Gatza \textit{et al.}.
+Comparing these groups with the results presented by \citet{Gatza2010a} (\cref{sec:result_from_gatza2010a_study}), the identified clusters approximately resembled the pathway groups identified in by Gatza \textit{et al.}, which confirmed that the directions of GT pathway metagenes were consistent with those used by Gatza \textit{et al.}.
 
@@
 	\includegraphics[page=15,width=0.7\linewidth]{results2/gatza_meta_trans}
-	\caption[Heatmap of the Pearson correlation of all the pathway metagenes with one another in the \acrshort{rma}-normalised GT data]{Heatmap showing the Pearson correlation of all the pathway metagenes from the \citet{Gatza2010a} study with one another  in the GT data set.
-		High and low correlation were represented as red and blue, respectively, where the colours were matched with the values on the scale shown in the top left histogram.}
+	\caption[Heatmap of the Pearson correlation of all the pathway metagenes with one another in the \acrshort{rma} normalised GT data]{Heatmap showing the Pearson correlation of all the pathway metagenes from the \citet{Gatza2010a} study with one another  in the \gls{rma} normalised GT data set.
+		High and low correlation were represented as red and blue, respectively, where the colours were matched with the values on the scale shown in the top left histogram.
+		Refer to \cref{tab:metagene_direction} for the detailed descriptions of the abbreviations used.
+		}
 	\label{fig:gatza_meta_dir}
@@
 
-To see whether the directionality of GT pathway metagenes were being transferred across to other data sets properly, pathway metagenes were generated in other data sets with the transformation matrices and the groupings of the metagenes were examined.
+To see whether the directionality of the GT pathway metagenes were being transferred across to other data sets properly, pathway metagenes were generated in other data sets with the transformation matrices and the groupings of the metagenes were examined.
 CR, FM and \gls{nzbc} data were \gls{rma}-normalised and transformation matrices of the pathway genetic signatures (derived from GT data set) were applied to the data sets.
@@
 
-% In order to investigate what the underlying biological mechanism of the obesity associated genetic signatures was, GT pathway associated genetic signatures were used to create linear models to predict the obesity associated metagene scores.
-% Before linear models were created, GT pathway metagenes were checked for their directionality to ensure that the pathway metagenes were acting as reported by \citet{Gatza2010a} (see \cref{sub:metagene_direction}).
-
 \section{Pathway associated metagenes and obesity associated metagenes}
@@
 
-Results from previous section confirmed that the pathway metagenes from the \citet{Gatza2010a} study were in the correct directions, and that the pathway metagenes were behaving as expected in different data sets.
+Results from the previous section confirmed that the pathway metagenes from the \citet{Gatza2010a} study were in the correct directions, and that the pathway metagenes were behaving as expected in the different data sets.
 Now that the directions of the pathway metagenes were established, these metagenes were ready for comparison with the obesity associated metagenes.
-However, all of the pathway metagenes were derived from GT data, whereas the majority of the obesity associated metagenes were derived from CR data, which presents a problem of deciding on which data set the transformation matrices should be generated in.
+However, all of the pathway metagenes were derived from the GT data, whereas the majority of the obesity associated metagenes were derived from the CR data, which presents a problem in terms of deciding which data set the transformation matrices should be generated in.
 
-In theory, the metagenes generated from the application of \gls{svd} and the metagenes generated from transformation matrix would be exactly the same if the transformation matrix was derived from the same data set.
-For example, an obesity associated metagene generated from CR data with \gls{svd} would have the same values as the metagenes generated in CR data with the transformation matrix (where the transformation matrix was derived from CR data).
-Furthermore, if the \gls{svd}-derived metagenes and transformation matrix-derived metagenes were the same (or at least similar) in a different data set, this suggests that there was no difference in the data sets, at least in terms of the expression of the genetic signatures being investigated.
+\subsection{Obesity and pathway metagene transformation matrices}
+\label{sub:obesity_and_pathway_metagene_transformation_matrices}
+
+Metagenes generated from the application of \gls{svd} and metagenes generated from the transformation matrix are exactly the same if the transformation matrix is derived from the same data set.
+For example, an obesity associated metagene generated from the CR data with \gls{svd} would have the same values as the metagenes generated in the CR data with the transformation matrix (where the transformation matrix was derived from CR data).
+Furthermore, if the \gls{svd}-derived metagenes and transformation matrix-derived metagenes were the same (or at least similar) in a different data set, this suggests that the metagene-related biology contained in the two data sets is highly similar.
 In other words, the transformation matrix can be made in either the original data set or in a different data set if there was no difference in the \gls{svd}-derived or transformation matrix-derived metagenes.
@@
 To decide which data set the transformation matrices for obesity and pathway associated genetic signatures should be generated in, the \gls{svd}- and transformation matrix(TM)-generated metagene scores were compared in all of the data sets.
-As in \cref{sec:pathway_associated_genetic_signatures_from_gatza2010a_study}, all data sets were normalised with the \gls{rma} method and metagenes were ranked based on the number of samples in each data set.
-Transformation matrices for the obesity associated genetic signatures were made in CR data set and pathway associated genetic signatures were made in the GT data set.
+As in \cref{sec:pathway_associated_genetic_signatures_from_gatza2010a_study}, all data sets were normalised with the \gls{rma} method and metagenes were ranked with fractional ranking.
+Transformation matrices for the obesity associated genetic signatures were made in the CR data set and the pathway associated genetic signatures were made in the GT data set.
 Metagenes for all of the obesity and pathway associated genetic signatures were generated in all of the data sets with \gls{svd} and transformation matrices.
@@
 			\begin{footnotesize}
-				\item [1] Transformation matrices were derived from the \gls{rma}-normalised GT data set.
+				\item [1] Transformation matrices were derived from the \gls{rma} normalised GT data set.
 			\end{footnotesize}
@@
 				\begin{footnotesize}
-				\item [1] Transformation matrices were derived from the \gls{rma}-normalised CR data set.
+				\item [1] Transformation matrices were derived from the \gls{rma} normalised CR data set.
 				\end{footnotesize}
@@
 
-In \cref{tab:svd_vs_tm_path,tab:svd_vs_tm_obs}, the obesity and pathway associated genetic signatures had a correlatin of 1 in CRs and GT data sets respectively, as the transformation matrices were generated in those data sets.
+In \cref{tab:svd_vs_tm_path,tab:svd_vs_tm_obs}, the obesity and pathway associated genetic signatures had a correlation of 1 in CR and GT data sets respectively, as the transformation matrices were generated in those data sets.
 The correlations of the obesity and pathway metagenes were variable across different data sets, which was as expected since all the data sets were different from one another.
 However, what was unexpected was the fact that there were some genetic signatures that were highly correlated in all of the data sets, whereas the metagene correlation of other signatures were highly variable across different data sets.
-For example, the \gls{bcat} pathway metagene was highly correlated in all of the data sets (\textgreater{} 0.98), whereas the \gls{stat3} pathway metagene was variable across different data sets, ranging from 0.1902 to 0.9159 (\cref{tab:svd_vs_tm_path,fig:allmetacor_bar}).
+For example, the \gls{bcat} pathway metagene was highly correlated in all of the data sets (\textgreater{} 0.98), whereas the \gls{stat3} pathway metagene was variable across different data sets, ranging from 0.1902 to 0.9159 (\cref{tab:svd_vs_tm_path}).
+% For example, the \gls{bcat} pathway metagene was highly correlated in all of the data sets (\textgreater{} 0.98), whereas the \gls{stat3} pathway metagene was variable across different data sets, ranging from 0.1902 to 0.9159 (\cref{tab:svd_vs_tm_path,fig:allmetacor_bar}).
 The fact that some pathway associated metagenes were consistent across different data sets suggested that some of these genetic signatures were reliable and did not depend on the data set the transformation matrices were derived from.
-On the other hand, the genetic signatures that were not consistent across the data sets were likely to be dependent on the data set in which they were derived from, and therefore the transformation matrices for these signatures must be derived from GT data set.
+On the other hand, the genetic signatures that were not consistent across the data sets were likely to be dependent on the data set in which they were derived from, and therefore the transformation matrices for these signatures must be derived from the GT data set.
+This also showed that the tumour biology may be different for different data sets, likely due to the difference in the patient cohorts that were selected for each of the studies.
 
-In contrast to the pathway associated genetic signatures, all of the obesity associated genetic signatures showed high correlation of the \gls{svd}- and transformation matrix-generated metagenes across the different data sets.
-This suggested that obesity associated genetic signatures were consistent across all data sets and the transformation matrices for these signatures could be made in any of the data sets.
-Taken together, transformation matrices for all the genetic  signatures were decided to be made in the GT data set, since there were some pathway associated signatures that were specific to GT data, but none of the obesity associated genetic signatures were specific to a data set.
-\\
+In contrast to the pathway associated genetic signatures, all of the obesity associated genetic signatures showed high correlation for the \gls{svd}- and transformation matrix-generated metagenes across the different data sets.
+This suggested that the obesity associated genetic signatures were relatively consistent across all data sets, and the transformation matrices for these signatures could be made in any of the data sets.
+For simplicity, it was decided that the transformation matrices for all the genetic  signatures would be made in the GT data set, since there were some pathway associated signatures that were specific to the GT data, but none of the obesity associated genetic signatures were specific to a data set.
 
-\begin{figure}[htpb]
-	\centering
-	\includegraphics[page=2,width=0.45\linewidth]{results2/allmetacor_bar}
-	\includegraphics[page=16,width=0.45\linewidth]{results2/allmetacor_bar}
-	\caption[Bar plots showing the Spearman correlation of the \gls{svd}- and TM-derived pathway metagenes across different data sets]{Bar plots showing the Spearman correlations of the \gls{svd}-derived and the transformation matrix-derived \gls{bcat} and \gls{stat3} metagenes across different data sets.
-	Only the \gls{bcat} (left) and the \gls{tgfb}  (right) pathway metagenes are shown.
-	The scatter plots for the other pathway metagenes are shown in \cref{sec:comparison_of_the_correlation_of_svd_and_tm_generated_pathway_metagenes}.
-	}
-	\label{fig:allmetacor_bar}
-\end{figure}
+% \begin{figure}[htpb]
+% 	\centering
+% 	\includegraphics[page=2,width=0.45\linewidth]{results2/allmetacor_bar}
+% 	\includegraphics[page=16,width=0.45\linewidth]{results2/allmetacor_bar}
+% 	\caption[Bar plots showing the Spearman correlation of the \gls{svd}- and TM-derived pathway metagenes across different data sets]{Bar plots showing the Spearman correlations of the \gls{svd}-derived and the transformation matrix-derived \gls{bcat} and \gls{stat3} metagenes across different data sets.
+% 	Only the \gls{bcat} (left) and the \gls{tgfb}  (right) pathway metagenes are shown.
+% 	The scatter plots for the other pathway metagenes are shown in \cref{sec:comparison_of_the_correlation_of_svd_and_tm_generated_pathway_metagenes}.
+% 	}
+% 	\label{fig:allmetacor_bar}
+% \end{figure}
 
-\noindent
+\subsection{Comparison of the obesity and pathway associated signatures}
+\label{sub:comparison_of_the_obesity_and_pathway_associated_signatures}
+
 To visually determine which pathway associated genetic signatures were most similar to the obesity associated genetic signatures, heatmaps were created with the metagenes for all of the genetic signatures.
 The directions of the pathway associated metagenes have already been determined in \cref{sec:pathway_associated_genetic_signatures_from_gatza2010a_study}, and the directions of the obesity associated metagenes were checked in GT data set as described in \cref{sub:metagene_direction}.
-All of the metagenes were created in \gls{rma}-normalised GT data set with \gls{svd}, and the metagene scores were plotted in a heatmap and clustered into groups (\cref{fig:gatza_allmeta}).
+All of the metagenes were created in the GT data set (normalised with \gls{rma}) using \gls{svd}, and the metagene scores were plotted in a heatmap and clustered into groups (\cref{fig:gatza_allmeta}).
 
@@
 	\caption[Heatmap of the Pearson correlation of all the pathway and obesity metagenes with one another in the \acrshort{rma}-normalised GT data]{Heatmap showing the Pearson correlation of all the pathway and obesity metagenes  with one another  in the \gls{rma}-normalised GT data set.
-		High and low correlation were represented as red and blue, respectively, where the colours were matched with the values on the scale shown in the top left histogram.}
+		High and low correlation were represented as red and blue, respectively, where the colours were matched with the values on the scale shown in the top left histogram.
+		Refer to \cref{tab:metagene_direction,tab:mg_abbrev} for the detailed descriptions of the abbreviations used.
+		}
 	\label{fig:gatza_allmeta}
@@
 
-From \cref{fig:gatza_allmeta}, all of the obesity associated metagenes from CR data set clustered together into a group, but none of the pathway associated metagenes associated strongly with the obesity metagenes.
-Though \gls{her2} and p63 pathway metagenes were grouped next to the cluster of obesity metagenes, the similarity was at most about 0.6.
-FM obesity associated metagene did not cluster together with the obesity metagenes from CR data set, nor with any of the other pathway metagenes.
-The fact that FM metagene did not cluster with any of the CR metagenes showed that the nature of FM metagene was different from those metagenes found in CR data set.
-Furthermore, their metagene did not cluster with the Akt pathway in the heatmap, even though \citet{Fuentes-Mattei2014} provided clear evidence that their metagene specifically activated the Akt/\gls{mtor} signalling pathway in a mouse model.
-This may have been due to the lack of consistency of the Akt pathway signature in different data sets (\cref{tab:svd_vs_tm_path}).
-Another likely reason could be that FM genetic signature was not comprised only of the genes from the Akt pathway, but genes from other pathways as well, and so the metagene did not cluster together with the Akt pathway signature in the heatmap.
+In \cref{fig:gatza_allmeta}, all of the obesity associated metagenes from the CR data set cluster together into a group, but none of the pathway associated metagenes correlated strongly with the obesity metagenes.
+Though \gls{her2} and p63 pathway metagenes were grouped next to the cluster of obesity metagenes, the similarity was at most about 0.6  (Pearson correlation; \cref{fig:gatza_allmeta}).
+FM obesity associated metagene did not cluster together with the obesity metagenes from the CR data set, nor with any of the other pathway metagenes.
+The fact that the FM metagene did not cluster with any of the CR metagenes showed that the nature of the FM metagene was different from those metagenes found in the CR data set.
+Furthermore, the FM metagene did not cluster with the Akt pathway in the heatmap, even though \citet{Fuentes-Mattei2014} provided clear evidence that their metagene specifically activated the Akt/\gls{mtor} signalling pathway in a mouse model.
+This may have been due to the lack of consistency of the Akt pathway signature in the different data sets (\cref{tab:svd_vs_tm_path}).
+Another likely reason could be that FM genetic signature was not comprised only of the genes from the Akt pathway, but genes from the other pathways as well, and so the metagene did not cluster together with the Akt pathway signature in the heatmap.
 
 The results in this section suggested that none of the obesity associated genetic signatures were similar or related to any of the pathway associated genetic signatures from the \citet{Gatza2010a} study.
-
-(add heatmaps of other data (Creighton/FM/Cris) as well? or maybe in appendix)
 
@@
 
-% and lastly linear models were constructed based on the pathway metagenes and sample \gls{bmi}/\gls{bmi} status to predict the obesity associated metagenes.
+To confirm that the obesity associated metagenes were not related to any of the biological pathway signatures from the \citet{Gatza2010a} study, linear models were created to predict the obesity metagene scores with the pathway metagene scores.
+If any of the pathway metagenes were significant in the linear model, this would provide evidence that the significant pathways in the linear model were related to the obesity metagene.
+Since most of the obesity metagenes were found in the CR data set, and only the CR and the \gls{nzbc} data sets had \gls{bmi} information for the patients, \gls{nzbc} data set was used as the training data set to create the linear models for obesity metagene score prediction.
 
-To confirm that the obesity associated metagenes were not related to any of the biological pathway signatures from the \citet{Gatza2010a} study, linear models were created to predict the obesity metagene scores with the pathway metagene scores.
-If any of the pathway metagenes were significant in the linear model, this provided evidence that the significant pathways in the linear model were related to the obesity metagene.
-Since most of the obesity metagenes were found in CR data set and only CR and \gls{nzbc} data sets had \gls{bmi}/\gls{bmi} status information for the samples, \gls{nzbc} data set was used as the training data set to create linear models for obesity metagene score prediction.
+\subsection{Linear model prediction in the \gls{nzbc} and CR data sets}
+\label{sub:linear_model_prediction_in_the_nzbc_and_cr_data_sets}
 
-Metagene scores used for the construction of linear models were generated from the application of the transformation matrices derived from GT data (from \cref{sec:pathway_associated_metagenes_and_obesity_associated_metagenes}) to \gls{rma}-normalised \gls{nzbc} data.
+Metagene scores used for the construction of linear models were generated from the application of the transformation matrices derived from the GT data (from \cref{sec:pathway_associated_metagenes_and_obesity_associated_metagenes}) to \gls{rma} normalised \gls{nzbc} data.
 For each of the obesity metagenes identified, seven linear models were created: \gls{bmi}-only, \gls{bmi} status-only, \gls{bmi} and \gls{bmi} status, pathway metagenes-only, and all of the combinations of pathway metagenes with \gls{bmi} and/or \gls{bmi} status.
-The only variables that showed significance in any of the models that predicted the obesity metagenes taken from the CR data set were \gls{pr} metagene and whether samples were obese or not (\cref{tab:lm_sig_var}).
-In the model that predicted the FM metagene scores, the obesity status of the samples and Myc metagene score were significant, but the \gls{pr} metagene score was not (\cref{app:b}).
+The only variables that showed significance in any of the models that predicted the obesity metagenes taken from the CR data set were the \gls{pr} metagene and whether the patients were obese or not (\cref{tab:lm_sig_var}).
+In the model that predicted the FM metagene scores, the obesity status of the patients and Myc metagene score were significant, but the \gls{pr} metagene score was not (\cref{sec:summary_of_the_linear_models_in_nzbc_data}).
 
@@
 
-All of the models constructed from sample \gls{bmi}, \gls{bmi} status and/or GT pathway metagene scores confirmed that none of the pathways were significant, except for the \gls{pr} and Myc pathway metagene scores in the models for CR metagenes and FM metagene, respectively.
-It was not surprising that the \gls{pr} pathway metagene showed significance in the models generated in \gls{nzbc} data set, as \gls{pr} status was very strongly correlated with \gls{nzbc} data set.
-The significance of the Myc metagene in the model for FM metagene was difficult to interpret, as there was no reported association of FM metagene with Myc pathway by \citet{Fuentes-Mattei2014}.
-However, ``sustained cell proliferation'' was one of the cancer hallmarks that \citet{Fuentes-Mattei2014} had shown to be related to their obesity genetic signature, so the Myc pathway metagene may have picked up the proliferation signal in the FM metagene.
+The models constructed from the patient \gls{bmi}, \gls{bmi} status and/or GT pathway metagene scores confirmed that none of the pathways were significant, except for the \gls{pr} and Myc pathway metagene scores in the models for the CR metagenes and the FM metagene, respectively.
+It was not surprising that the \gls{pr} pathway metagene showed significance in the models generated in \gls{nzbc} data set, as majority of the patients in the \gls{nzbc} data set were \gls{pr}$^+$ (\cref{tab:clin_summary}).
+The significance of the Myc metagene in the model for FM metagene was difficult to interpret, as there was no reported association of the FM metagene with the Myc pathway by \citet{Fuentes-Mattei2014}.
+\textit{MYC} is an oncogene that encodes the Myc transcription factor that, when overexpressed, disrupts various pathways that directly promote tumour growth and proliferation, and has a central role in many of the cancer hallmarks \citep{Coller2000,Hanahan2000}.
+``Sustained cell proliferation'' was one of the cancer hallmarks that \citet{Fuentes-Mattei2014} had shown to be related to their obesity associated genetic signature, which could be the reason why the Myc pathway metagene showed up as a significant variable in the linear model.
 
-Since the \gls{pr} pathway metagene showed significant association in many of the linear models, linear models were made with each of the combinations of \gls{bmi}, \gls{bmi} status and \gls{pr} pathway metagene for all of the obesity associated metagenes (\cref{tab:lm_pr_only} and \cref{app:b}).
-As expected, all of the linear models showed significant contribution of the \gls{pr} metagene scores to the model for all obesity metagenes (except FM metagene; \cref{app:b}).
+Since the \gls{pr} pathway metagene showed significant association in many of the linear models, linear models were made with each of the combinations of \gls{bmi}, \gls{bmi} status and \gls{pr} pathway metagene for all of the obesity associated metagenes (\cref{tab:lm_pr_only}; \cref{sec:summary_of_the_linear_models_in_nzbc_data}).
+As expected, all of the linear models showed significant contribution of the \gls{pr} metagene scores to the model for all obesity metagenes (except FM metagene; \cref{sec:summary_of_the_linear_models_in_nzbc_data}).
+% Since the \gls{pr} pathway metagene showed significant association in many of the linear models, linear models were made just with the \gls{pr} metagene scores.
+% For each of the CR and FM obesity metagenes, linear models were constructed using the patient \gls{bmi}, \gls{bmi} status and \gls{pr} pathway metagene as predictor variables (\cref{tab:lm_pr_only}; \cref{sec:summary_of_the_linear_models_in_nzbc_data}).
+% As expected, all of the linear models showed significant contribution of the \gls{pr} metagene scores to the model for all obesity metagenes (except FM metagene; \cref{sec:summary_of_the_linear_models_in_nzbc_data}).
 
@@
 	\centering
-	\caption[Description of the linear models constructed from the \gls{nzbc} data to predict the Cr obesity, using only the sample \gls{bmi}, \gls{bmi} status and the \acrshort{pr} pathway metagene score]{Description of the linear models constructed from the \gls{nzbc} data to predict the Cr obesity, using only the sample \gls{bmi}, \gls{bmi} status and the \gls{pr} pathway metagene score}
+	\caption[Description of the linear models constructed from the \gls{nzbc} data to predict the Cr obesity, using only the patient \gls{bmi}, \gls{bmi} status and the \acrshort{pr} pathway metagene score]{Description of the linear models constructed from the \gls{nzbc} data to predict the Cr obesity, using only the patient \gls{bmi}, \gls{bmi} status and the \gls{pr} pathway metagene score}
 	\label{tab:lm_pr_only}
@@
 To determine whether any of the linear models were able to predict the corresponding obesity metagene scores, the predicted metagene scores were generated from the models and then compared with the original scores.
-The predictions were first made in \gls{nzbc} data, where the metagene scores generated from the application of transformation matrices (from GT data set) were used to make the predictions.
-The predicted metagenes were then compared with the true metagenes that were originally generated from the transformation matrices (\cref{fig:predict_cris_cris}).
+The original obesity metagene scores from the \gls{nzbc} were generated from the application of TM that were derived from the \gls{rma}-normalised GT data set.
+These \gls{nzbc} obesity metagene scores were then compared with the metagene scores that were predicted by the linear models (\cref{fig:predict_cris_cris}).
 
-The predicted metagenes from the \gls{bmi} only, \gls{bmi} status only and \gls{bmi} and \gls{bmi} status linear models were not significantly related to the true metagene in many of the obesity metagenes; though \gls{bmi} and \gls{bmi} status model showed significance in some metagenes (\cref{app:b}).
-On the other hand, all of the predicted metagenes generated from the models that contained any of the pathway metagenes showed significant association with the true metagene scores in all of the obesity metagenes (\cref{fig:predict_cris_cris,app:b}).
+The predicted metagenes from all of the patient \gls{bmi}-related linear models were not significantly related to the original metagene scores in many of the obesity metagenes (\cref{fig:predict_cris_cris}; \cref{sec:summary_of_the_linear_models_in_nzbc_data}).
+On the other hand, all of the predicted metagenes generated from the models that used the pathway metagene scores showed significant association with the original metagene scores (\cref{fig:predict_cris_cris}; \cref{sec:summary_of_the_linear_models_in_nzbc_data}).
 However, the $R^2$ values for all of the predicted metagenes against the true metagenes were very low ($R^2$ \textless{} 0.4).
-Furthermore, all of the plots in \cref{fig:predict_cris_cris,app:b} showed clearly that the data points in the scatter plots were highly variable, suggesting that even though the association between the predicted and the true metagene scores were statistically significant, the variables in the linear models may not be strongly associated to the obesity metagene.
+All of the plots in \cref{fig:predict_cris_cris,sec:summary_of_the_linear_models_in_nzbc_data} showed clearly that the data points in the scatter plots were highly variable, suggesting that even though the association between the predicted and the true metagene scores were statistically significant, the variables in the linear models may not be strongly associated to the obesity metagene.
+Furthermore, these results suggested that most of the variations in the Cr (and other) obesity metagenes were being explained by the \gls{pr} metagene and the patient \gls{bmi} contributed very little to the prediction.
 
@@
 
-To confirm the results shown in \gls{nzbc} data set (training set), the linear models were applied to CR data (test set).
-As shown in \cref{fig:predict_cr_cris}, all of the predicted metagenes generated from the linear models were statistically significantly associated with the true metagenes, although there were some exceptions in the models for other obesity metagenes (see \cref{app:b}).
-Similar to the results from \gls{nzbc} data set, all of the $R^2$ values were low ($R^2$ \textless{} 0.43) and the data points were widely spread out in the plots.
+%TODO: comment on the negative relationship of the predicted and true obesity metagenes in the CR data
+
+To confirm the results shown in the \gls{nzbc} data set (the training set), the linear models were applied to CR data (the test set).
+As shown in \cref{fig:predict_cr_cris}, all of the predicted Cr metagene generated from the linear models were statistically significantly associated with the true Cr metagenes in CR data, but there were some exceptions in the models for the other obesity metagenes (\cref{sec:summary_statistics_of_the_predicted_obesity_metagenes_with_sample_bmi_bmi_status_in_nzbc_and_cr_data}).
+Similar to the results from the \gls{nzbc} data set, all of the $R^2$ values were low ($R^2$ \textless{} 0.43) in the CR data set and the data points were widely spread out in the plots.
 Again, these results suggested that the predicted metagene scores from the linear models may be statistically significant, but may not be relevant to the obesity metagenes which the models predict, as the predicted values were only weakly associated with the true values.
-\\
 
@@
 
-\noindent
+\subsection{Stepwise linear model prediction in \gls{nzbc} and CR data sets}
+\label{sub:stepwise_linear_model_prediction_in_nzbc_and_cr_data_sets}
+
 Lastly, since only the ``consistent'' pathway metagenes were picked from all of the available pathway metagene scores, linear models were constructed with all of the pathway metagenes in a step wise fashion.
-Step wise construction of a linear model is where the variables for a model are selected computationally depending on the contribution of the variable to the resuting linear model, and thus contains only the variables that ``matter'' to the linear model.
-Sample \gls{bmi}, \gls{bmi} status and all of the pathway metagene scores were used to generate the step wise linear model for each of the obesity metagenes.
+Step wise construction of a linear model is where the variables for a model are selected computationally depending on the contribution of the variable to the resulting linear model, and thus contains only the variables that ``matter'' to the final linear model.
+Patient \gls{bmi}, \gls{bmi} status and all of the pathway metagene scores were used to generate the step wise linear model for each of the obesity metagenes.
 The linear models that resulted from the step wise selection of the variables were summarised in \cref{tab:stepwise_model}.
@@
 	\label{tab:stepwise_model}\\
-		Obesity metagene & Variables used in the model & Estimate & {P-value}\\
+		Obesity metagene & Variables retained in the model & Estimate & {P-value}\\
 		\hline
@@
 		Cr         & Akt         & -0.3415   & \bfseries \num{4.43d-8}\\
-				   & \gls{egfr}  & 0.33852   & \bfseries \num{1.94d-05}\\
-				   & \gls{er}    & -0.20237  & \bfseries 0.00419\\
-				   & \gls{ifny}  & 0.12233   & 0.08700\\
-				   & \gls{pr}    & -0.33920  & \bfseries \num{2.09d-5}\\
-				   & Ras         & -0.22611  & \bfseries 0.00193\\
-				   & Src         & -0.11452  & 0.08177\\
-				   & \gls{tgfb}  & 0.55638   & \bfseries \textless{} \num{2d-16}\\
-				   & \gls{tnfa}  & -0.25501  & \bfseries 0.00616\\
+				   & \gls{egfr}  & 0.3385   & \bfseries \num{1.94d-05}\\
+				   & \gls{er}    & -0.2024  & \bfseries 0.004190\\
+				   & \gls{ifny}  & 0.1223   & 0.08700\\
+				   & \gls{pr}    & -0.3392  & \bfseries \num{2.09d-5}\\
+				   & Ras         & -0.2261  & \bfseries 0.001930\\
+				   & Src         & -0.1145  & 0.08177\\
+				   & \gls{tgfb}  & 0.5564   & \bfseries \textless{} \num{2d-16}\\
+				   & \gls{tnfa}  & -0.2550  & \bfseries 0.006160\\
 		\hline
-		CrOl       & Overweight  & 0.040918  & 0.074865\\
-				   & Obese       & -0.006381 & 0.756121\\
-				   & Akt         & -0.356713 & \bfseries \num{1.89d-07}\\
-				   & E2F1        & 0.160229  & \bfseries 0.005411\\
-				   & \gls{egfr}  & 0.148618  & 0.062683\\
-				   & p53         & 0.336552  & \bfseries \num{4.35d-05}\\
-				   & \gls{pr}    & -0.225423 & \bfseries 0.000639\\
-				   & Src         & -0.428758 & \bfseries \num{5.85d-10}\\
-				   & \gls{stat3} & -0.124118 & \bfseries 0.009658\\
+		CrOl       & Overweight  & 0.04092  & 0.07487\\
+				   & Obese       & -0.006381 & 0.7561\\
+				   & Akt         & -0.3567 & \bfseries \num{1.89d-07}\\
+				   & E2F1        & 0.1602  & \bfseries 0.005411\\
+				   & \gls{egfr}  & 0.1486  & 0.06268\\
+				   & p53         & 0.3366  & \bfseries \num{4.35d-05}\\
+				   & \gls{pr}    & -0.2254 & \bfseries 0.0006390\\
+				   & Src         & -0.4288 & \bfseries \num{5.85d-10}\\
+				   & \gls{stat3} & -0.1241 & \bfseries 0.009658\\
 		\hline
-		Res        & Akt         & 0.24613   & \bfseries 0.000289\\
-				   & \gls{egfr}  & -0.46827  & \bfseries \num{2.16d-05}\\
-				   & \gls{er}    & 0.18400   & 0.061446\\
-				   & \gls{her2}  & 0.13835   & 0.118228\\
-				   & \gls{ifny}  & -0.21019  & \bfseries 0.044084\\
-				   & \gls{pr}    & 0.41195   & \bfseries 0.000300\\
-				   & Ras         & 0.26506   & \bfseries 0.010465\\
-				   & \gls{tgfb}  & -0.58075  & \bfseries \num{7.64d-11}\\
-				   & \gls{tnfa}  & 0.42200   & \bfseries 0.001834\\
+		Res        & Akt         & 0.2461   & \bfseries 0.0002890\\
+				   & \gls{egfr}  & -0.4683  & \bfseries \num{2.16d-05}\\
+				   & \gls{er}    & 0.1840   & 0.06145\\
+				   & \gls{her2}  & 0.1384   & 0.1182\\
+				   & \gls{ifny}  & -0.2102  & \bfseries 0.04408\\
+				   & \gls{pr}    & 0.4120   & \bfseries 0.0003000\\
+				   & Ras         & 0.2651   & \bfseries 0.01047\\
+				   & \gls{tgfb}  & -0.5808  & \bfseries \num{7.64d-11}\\
+				   & \gls{tnfa}  & 0.4220   & \bfseries 0.001834\\
 		\hline
-		ResOl      & Akt         & -0.33596  & \bfseries 0.000294\\
-				   & \gls{bcat}  & -0.19866  & 0.138953\\
-				   & E2F1        & 0.36059   & \bfseries 0.009325\\
-				   & \gls{egfr}  & 0.23421   & 0.056865\\
-				   & \gls{her2}  & -0.13652  & 0.121689\\
-				   & p53         & 0.38746   & \bfseries 0.000519\\
-				   & \gls{pi3k}  & -0.21881  & 0.148676\\
-				   & \gls{pr}    & -0.32228  & \bfseries 0.001656\\
-				   & Ras         & -0.28984  & \bfseries 0.033975\\
-				   & Src         & -0.40779  & \bfseries \num{5.26d-06}\\
-				   & \gls{stat3} & -0.11452  & 0.125609\\
-				   & \gls{tnfa}  & -0.11442  & 0.117023\\
+		ResOl      & Akt         & -0.3360  & \bfseries 0.0002940\\
+				   & \gls{bcat}  & -0.1987  & 0.1390\\
+				   & E2F1        & 0.3606   & \bfseries 0.009325\\
+				   & \gls{egfr}  & 0.2342   & 0.05687\\
+				   & \gls{her2}  & -0.1365  & 0.1217\\
+				   & p53         & 0.3875   & \bfseries 0.0005190\\
+				   & \gls{pi3k}  & -0.2188  & 0.1487\\
+				   & \gls{pr}    & -0.3223  & \bfseries 0.001656\\
+				   & Ras         & -0.2898  & \bfseries 0.03398\\
+				   & Src         & -0.4078  & \bfseries \num{5.26d-06}\\
+				   & \gls{stat3} & -0.1145  & 0.1256\\
+				   & \gls{tnfa}  & -0.1144  & 0.1170\\
 		\hline
-		Ca         & Akt         & -0.26243  & \bfseries \num{7.04d-06}\\
-				   & \gls{egfr}  & 0.36305   & \bfseries 0.000139\\
-				   & \gls{er}    & -0.18614  & \bfseries 0.030771\\
-				   & \gls{ifny}  & 0.22118   & \bfseries 0.013034\\
-				   & \gls{pr}    & -0.42911  & \bfseries \num{1.07d-05}\\
-				   & Ras         & -0.28124  & \bfseries 0.000898\\
-				   & \gls{tgfb}  & 0.76139   & \bfseries \textless{} \num{2d-16}\\
-				   & \gls{tnfa}  & -0.33200  & \bfseries 0.003919\\
+		Ca         & Akt         & -0.2624  & \bfseries \num{7.04d-06}\\
+				   & \gls{egfr}  & 0.3631   & \bfseries 0.0001390\\
+				   & \gls{er}    & -0.1861  & \bfseries 0.03077\\
+				   & \gls{ifny}  & 0.2212   & \bfseries 0.01303\\
+				   & \gls{pr}    & -0.4291  & \bfseries \num{1.07d-05}\\
+				   & Ras         & -0.2812  & \bfseries 0.0008980\\
+				   & \gls{tgfb}  & 0.7614   & \bfseries \textless{} \num{2d-16}\\
+				   & \gls{tnfa}  & -0.3320  & \bfseries 0.003919\\
 		\hline
 		CaOl       & Overweight  & 0.07613   & \bfseries 0.006875\\
-				   & Obese       & 0.01408   & 0.564622\\
-				   & Akt         & -0.44727  & \bfseries \num{2.43d-07}\\
-				   & \gls{bcat}  & -0.13463  & 0.140261\\
-				   & E2F1        & 0.33864   & \bfseries 0.001249\\
-				   & p53         & 0.28777   & \bfseries 0.000996\\
-				   & \gls{pr}    & -0.28255  & \bfseries \num{2.18d-05}\\
-				   & Src         & -0.44149  & \bfseries \num{2.02d-08}\\
-				   & \gls{tnfa}  & -0.14215  & \bfseries 0.013897\\
+				   & Obese       & 0.01408   & 0.5646\\
+				   & Akt         & -0.4473  & \bfseries \num{2.43d-07}\\
+				   & \gls{bcat}  & -0.1346  & 0.1403\\
+				   & E2F1        & 0.3386   & \bfseries 0.001249\\
+				   & p53         & 0.2878   & \bfseries 0.0009960\\
+				   & \gls{pr}    & -0.2826  & \bfseries \num{2.18d-05}\\
+				   & Src         & -0.4415  & \bfseries \num{2.02d-08}\\
+				   & \gls{tnfa}  & -0.1422  & \bfseries 0.01390\\
 		\hline
-		CaRes      & \gls{bcat}  & 0.26899   & 0.05238  \\
-				   & E2F1        & -0.22115  & 0.08848  \\
-				   & \gls{egfr}  & 0.44400   & \bfseries \num{6.67d-05 }\\
-				   & \gls{her2}  & -0.20304  & \bfseries 0.04615  \\
-				   & \gls{ifny}  & 0.32874   & \bfseries 0.00574  \\
-				   & \gls{pr}    & -0.28764  & \bfseries 0.00104  \\
-				   & Src         & -0.17989  & 0.05404  \\
-				   & \gls{tgfb}  & 0.52898   & \bfseries \num{2.15d-06 }\\
-				   & \gls{tnfa}  & -0.38677  & \bfseries 0.00564  \\
+		CaRes      & \gls{bcat}  & 0.2690   & 0.05238  \\
+				   & E2F1        & -0.2212  & 0.08848  \\
+				   & \gls{egfr}  & 0.4440   & \bfseries \num{6.67d-05 }\\
+				   & \gls{her2}  & -0.2030  & \bfseries 0.04615  \\
+				   & \gls{ifny}  & 0.3287   & \bfseries 0.005740  \\
+				   & \gls{pr}    & -0.2876  & \bfseries 0.001040  \\
+				   & Src         & -0.1799  & 0.05404  \\
+				   & \gls{tgfb}  & 0.5290   & \bfseries \num{2.15d-06 }\\
+				   & \gls{tnfa}  & -0.3868  & \bfseries 0.005640  \\
 		\hline
-		CaResOl    & \gls{bmi}   & 0.004314  & 0.108546\\
-				   & Overweight  & 0.033086  & 0.306390\\
-				   & Obese       & -0.062257 & 0.175776\\
-				   & Akt         & -0.309706 & \bfseries 0.000339\\
-				   & \gls{bcat}  & -0.198165 & 0.140502\\
-				   & E2F1        & 0.270471  & \bfseries 0.011450\\
-				   & \gls{egfr}  & 0.171219  & 0.113691\\
-				   & Myc         & 0.152007  & 0.056987\\
-				   & p53         & 0.294895  & \bfseries 0.003734\\
-				   & \gls{pr}    & -0.230896 & \bfseries 0.000823\\
-				   & Ras         & -0.270824 & \bfseries 0.044788\\
-				   & Src         & -0.512567 & \bfseries \num{2.72d-08}\\
-				   & \gls{tnfa}  & -0.103302 & 0.083449\\
+		CaResOl    & \gls{bmi}   & 0.004310  & 0.1086\\
+				   & Overweight  & 0.03309  & 0.3064\\
+				   & Obese       & -0.06226 & 0.1758\\
+				   & Akt         & -0.3097 & \bfseries 0.0003390\\
+				   & \gls{bcat}  & -0.1982 & 0.1405\\
+				   & E2F1        & 0.2705  & \bfseries 0.01145\\
+				   & \gls{egfr}  & 0.1712  & 0.1137\\
+				   & Myc         & 0.1520  & 0.05699\\
+				   & p53         & 0.2949  & \bfseries 0.003734\\
+				   & \gls{pr}    & -0.2309 & \bfseries 0.0008230\\
+				   & Ras         & -0.2708 & \bfseries 0.04479\\
+				   & Src         & -0.5126 & \bfseries \num{2.72d-08}\\
+				   & \gls{tnfa}  & -0.1033 & 0.08345\\
 		\hline
-		Or         & \gls{bmi}   & 0.002696  & 0.14407\\
-				   & Overweight  & 0.010522  & 0.62856\\
-				   & Obese       & -0.061020 & 0.05542\\
-				   & Akt         & -0.238412 & \bfseries \num{4.93d-05}\\
-				   & E2F1        & 0.185085  & \bfseries 0.02754\\
-				   & \gls{egfr}  & 0.147219  & 0.07423\\
-				   & \gls{her2}  & -0.086804 & 0.14070\\
-				   & p53         & 0.416414  & \bfseries \num{5.83d-08}\\
-				   & \gls{pi3k}  & -0.216179 & \bfseries 0.03528\\
-				   & \gls{pr}    & -0.199097 & \bfseries 0.00253\\
-				   & Ras         & -0.200544 & \bfseries 0.00898\\
-				   & Src         & -0.543834 & \bfseries \num{2.10d-15}\\
-				   & \gls{stat3} & -0.093363 & \bfseries 0.02749\\
+		Or         & \gls{bmi}   & 0.002696  & 0.1441\\
+				   & Overweight  & 0.01052  & 0.62856\\
+				   & Obese       & -0.06102 & 0.05542\\
+				   & Akt         & -0.2384 & \bfseries \num{4.93d-05}\\
+				   & E2F1        & 0.1851  & \bfseries 0.02754\\
+				   & \gls{egfr}  & 0.1472  & 0.07423\\
+				   & \gls{her2}  & -0.08680 & 0.1407\\
+				   & p53         & 0.4164  & \bfseries \num{5.83d-08}\\
+				   & \gls{pi3k}  & -0.2162 & \bfseries 0.03528\\
+				   & \gls{pr}    & -0.1991 & \bfseries 0.002530\\
+				   & Ras         & -0.2005 & \bfseries 0.008980\\
+				   & Src         & -0.5438 & \bfseries \num{2.10d-15}\\
+				   & \gls{stat3} & -0.0934 & \bfseries 0.02749\\
 		\hline
-		FM         & \gls{bmi}   & 0.011093  & \bfseries 0.036229\\
-				   & Overweight  & -0.153753 & \bfseries 0.014277\\
-				   & Obese       & -0.212483 & \bfseries 0.017378\\
-				   & p53         & 0.460800  & \bfseries 0.002512\\
-				   & \gls{pi3k}  & -0.426246 & \bfseries 0.000118\\
-				   & \gls{pr}    & -0.373119 & \bfseries 0.020737\\
-				   & \gls{stat3} & 0.248745  & \bfseries 0.019073\\
-				   & \gls{tgfb}  & 0.152925  & 0.128933\\
+		FM         & \gls{bmi}   & 0.01109  & \bfseries 0.03623\\
+				   & Overweight  & -0.1538 & \bfseries 0.01428\\
+				   & Obese       & -0.2125 & \bfseries 0.01738\\
+				   & p53         & 0.4608  & \bfseries 0.002512\\
+				   & \gls{pi3k}  & -0.4262 & \bfseries 0.0001180\\
+				   & \gls{pr}    & -0.3731 & \bfseries 0.02074\\
+				   & \gls{stat3} & 0.2487  & \bfseries 0.01907\\
+				   & \gls{tgfb}  & 0.1529  & 0.1289\\
 		\hline
@@
 
-These models were used to predict the obesity metagenes in \gls{nzbc} data to test the performance of the models.
-All of the models were able to make statistically significant prediction of the obesity metagenes in \gls{nzbc} data set (\cref{fig:stepwise_cris}).
-Unlike the previous models, the step wise models showed very high $R^2$ values, where most predictions had $R^2$ \textgreater{} 0.85.
-The statistical significance and the strong association of the predicted metagenes were observed in CR data set as well (\cref{fig:stepwise_cr}).
-Though the $R^2$ values were slightly lower than those from \gls{nzbc} data set, the $R^2$ values were more promising than the non-step wise linear models.
+These models were used to predict the obesity metagenes in the \gls{nzbc} data to test the performance of the models.
+All of the models were able to make statistically significant prediction of the original obesity metagenes in the \gls{nzbc} data set (\cref{fig:stepwise_cris}).
+Unlike the previous models, the stepwise models showed very high $R^2$ values, where most predictions had $R^2$ \textgreater{} 0.85.
+The statistical significance and the strong association of the predicted metagenes were observed in the CR data set as well (\cref{fig:stepwise_cr}).
+Though the $R^2$ values were slightly lower than those from the \gls{nzbc} data set, the $R^2$ values from the CR data prediction were more promising than the predictions made by the non-stepwise linear models.
 
@@
 
-Taken together, these results suggested that the variables included in these step wise models were associated with the obesity metagenes identified in CR and FM data sets.
-However, there were no single common pathway that was present in all of the models; although Akt, \gls{egfr} and \gls{pr} pathway metagenes were frequently included in many of the models.
+Taken together, these results suggested that the variables included in these step wise models were relevant to the biology of the obesity metagenes identified in the CR and FM data sets.
+However, there were no single common pathway that was present in all of the models; although Akt, \gls{egfr} and \gls{pr} pathway metagenes were frequently included in many of the stepwise models.
 The role of the pathways included in the models and the possible biological links of these pathways to the obesity metagenes remain unclear, and should be the focus for future investigations.
+
+% TODO: update the stepwise linear model table and the scatter plots

### 2017-03-23 thesis/discussion/discussion.tex
Finished editing the draft thesis - clean things up

@@
 However, the degree of similarity between these two cancer types are not yet known and further analyses are required for a definitive explanation of the results.
+
+\subsection{Use of \gls{svd} and transformation matrix as a valid method}
+\label{sub:use_of_svd_and_transformation_matrix_as_a_valid_method}
+
+From all of these results so far, one may question how robust or appropriate the use of \gls{svd} and transformation matrices are to generate metagenes in other data sets.
+It appeared as though the CR data-derived obesity associated genetic signatures correlated significantly in the CR data, but as soon as these metagenes were generated in the other data sets via the application of transformation matrices, the correlation became insignificant.
+One possible reason for this was that the obesity metagenes were not actually associated to obesity phenotype in the other data sets.
+Another reason could be that the methodology used were not appropriate for investigating the association between the phenotype and its associated phenotype.
+
+The appropriateness of the use of \gls{svd} and transformation matrices were examined with different genetic signatures: \gls{er} and \gls{pr} pathway associated genetic signatures from the \citet{Gatza2010a} study (covered in \cref{sec:results_for_er_and_pr_metagene_analysis}).
+Three data sets (CR, FM and \gls{nzbc}) were used, as these data sets have both \gls{er} and \gls{pr} status information available for the patients.
+\gls{svd} was first applied to the \gls{rma} normalised CR data to obtain the \gls{er} and \gls{pr} metagenes, along with the transformation matrices for each of the genetic signatures.
+Both the \gls{er} and \gls{pr} metagenes showed good accordance with the gene expressions of the genetic signature in the heatmap, and also showed significant association with the patient \gls{er} and \gls{pr} statuses (\cref{sec:results_for_er_and_pr_metagene_analysis}).
+
+Transformation matrices were applied to the FM and \gls{nzbc} data sets to generate the appropriate pathway metagenes in the data sets.
+Both the \gls{er} and \gls{pr} pathway metagenes showed similar overall gene expression patterning in the heatmap as the CR data (\cref{sec:results_for_er_and_pr_metagene_analysis}).
+Furthermore, both the \gls{er} and \gls{pr} pathway metagenes from these data sets showed significant association with the patient \gls{er} and \gls{pr} statuses, respectively (\cref{sec:results_for_er_and_pr_metagene_analysis}).
+These results clearly showed that the general approach of using \gls{svd} and transformation matrix was appropriate for examining the association of the genetic signatures with the patient phenotype.
 

### 2017-03-23 thesis/misc/appendix.tex
Finished editing the draft thesis - clean things up

@@
 \end{longtable}
+
+\section{Results for \gls{er} and \gls{pr} metagene analysis}
+\label{sec:results_for_er_and_pr_metagene_analysis}
+
+To determine whether the results from all of the obesity metagene analyses were not showing any association with patient \gls{bmi} because of the \gls{svd}/transformation matrix approach used, the same approach was used with \gls{er} and \gls{pr} pathway associated genetic signatures from the \citet{Gatza2010a} study.
+Firstly, both the \gls{er} and \gls{pr} metagenes were generated in the \gls{rma} normalised CR data set.
+The generated metagenes were compared with the gene expressions of the corresponding pathway associated genetic signatures, and the association with the patient \gls{er} and \gls{pr} statuses were also examined.
+As shown in \cref{fig:er_cr,fig:pr_cr}, both the \gls{er} and \gls{pr} metagenes in the CR data showed assocation with the gene expression pattern of the pathway, and also showed significant association with the patient \gls{er} and \gls{pr} statuses.
+
+	\begin{figure}[htp!]
+		\centering
+		\includegraphics[page=8,width=0.45\linewidth]{appendix/er_pr_cr}\\
+		\hfill
+		\includegraphics[page=9,width=0.45\linewidth]{appendix/er_pr_cr}
+		\caption[\acrshort{er} metagene in the CR data]{Heatmap and box plot showing the association of the \gls{er} metagene with the sample gene expression and patient \gls{er} status, respectively, in the CR data.
+		Scales for the heatmap is as described in previou figures.
+		P-values  are calculated relative to the \gls{er}$^-$ group.}
+		\label{fig:er_cr}
+	\end{figure}
+
+	\begin{figure}[htp!]
+		\centering
+		\includegraphics[page=3,width=0.45\linewidth]{appendix/er_pr_cr}\\
+		\hfill
+		\includegraphics[page=4,width=0.45\linewidth]{appendix/er_pr_cr}
+		\caption[\acrshort{pr} metagene in the CR data]{Heatmap and a box plot showing the association of the \gls{pr} metagene with the sample gene expression and patient \gls{pr} status, respectively, in the CR cancer data.
+		Scales and  p-values are as described in previous figures.}
+		\label{fig:pr_cr}
+	\end{figure}
+
+The transformation matrices for the \gls{er} and \gls{pr} metagenes were created in the CR data, and was applied to the \gls{nzbc} and FM data sets (\gls{rma} normalised).
+As clearly shown by \cref{fig:er_cris,fig:pr_cris,fig:er_fm,fig:pr_fm}, both the \gls{er} and \gls{pr} metagenes associated with the gene expressions of the corresponding pathway signature.
+More importantly, both metagenes were significantly associated with the patient \gls{er} and \gls{pr} statuses, even though these metagenes were generated with the transformation matrices from the CR data.
+These results show that the approach taken in this project does not affect the link between the genetic signature and the patient phenotype, if such link exists between them.
+
+	\begin{figure}[htp!]
+		\centering
+		\includegraphics[page=8,width=0.45\linewidth]{appendix/er_pr_cris}\\
+		\hfill
+		\includegraphics[page=9,width=0.45\linewidth]{appendix/er_pr_cris}
+		\caption[\acrshort{er} metagene in the \gls{nzbc} data]{Heatmap and box plot showing the association of the \gls{er} metagene with the sample gene expression and patient \gls{er} status, respectively, in the \gls{nzbc} data.
+		Scales for the heatmap is as described in previou figures.
+		P-values  are calculated relative to the \gls{er}$^-$ group.}
+		\label{fig:er_cris}
+	\end{figure}
+
+	\begin{figure}[htp!]
+		\centering
+		\includegraphics[page=3,width=0.45\linewidth]{appendix/er_pr_cris}\\
+		\hfill
+		\includegraphics[page=4,width=0.45\linewidth]{appendix/er_pr_cris}
+		\caption[\acrshort{pr} metagene in the \gls{nzbc} data]{Heatmap and a box plot showing the association of the \gls{pr} metagene with the sample gene expression and patient \gls{pr} status, respectively, in the \gls{nzbc} cancer data.
+		Scales and  p-values are as described in previous figures.}
+		\label{fig:pr_cris}
+	\end{figure}
+
+	\begin{figure}[htp!]
+		\centering
+		\includegraphics[page=8,width=0.45\linewidth]{appendix/er_pr_fm}\\
+		\hfill
+		\includegraphics[page=9,width=0.45\linewidth]{appendix/er_pr_fm}
+		\caption[\acrshort{er} metagene in the FM data]{Heatmap and box plot showing the association of the \gls{er} metagene with the sample gene expression and patient \gls{er} status, respectively, in the FM data.
+		Scales for the heatmap is as described in previou figures.
+		P-values  are calculated relative to the \gls{er}$^-$ group.}
+		\label{fig:er_fm}
+	\end{figure}
+
+	\begin{figure}[htp!]
+		\centering
+		\includegraphics[page=3,width=0.45\linewidth]{appendix/er_pr_fm}\\
+		\hfill
+		\includegraphics[page=4,width=0.45\linewidth]{appendix/er_pr_fm}
+		\caption[\acrshort{pr} metagene in the FM data]{Heatmap and a box plot showing the association of the \gls{pr} metagene with the sample gene expression and patient \gls{pr} status, respectively, in the FM cancer data.
+		Scales and  p-values are as described in previous figures.}
+		\label{fig:pr_fm}
+	\end{figure}

### 2017-03-23 thesis/discussion/discussion.tex
tmp commit

@@
 The fact that the \gls{er} pathway signature was not significant in the models suggested that the obesity metagenes were related to the \gls{pr} pathway irrespective of the sample \gls{pr} status.
-Furthermore, both the \gls{nzbc} and the CR data sets showed strong association between the \gls{pr} and \gls{er} pathway metagenes with the patient \gls{pr} and \gls{er} status, respectively (\cref{})(add reference to pr and er graph).
+Furthermore, both the \gls{nzbc} and the CR data sets showed strong association between the \gls{pr} and \gls{er} pathway metagenes with the patient \gls{pr} and \gls{er} status, respectively (\cref{sec:results_for_er_and_pr_metagene_analysis}).
 Since both \gls{er} and \gls{pr} signatures were significantly associated with the patient \gls{er} and \gls{pr} satuses, both signatures would have been as equally likely to show up significant in the models, supporting the idea that the \gls{er} and \gls{pr} statuses were independent of the apparent significance of these signatures in the models.
@@
 In the step wise linear models, many variables were identified to be significantly associated with the obesity metagenes.
-With that said, even if these variables were related to the obesity associated genetic signatures, the fact that these results depended on the pathway associated genetic signatures from the \citet{Gatza2010a} study makes the results questionable.
+This suggested that the pathway associated genetic signatures that were signicant in these models may be biologically related to the obesity phenotype, or perhaps the obesity associated genetic signatures were detecting the signals from these pathways rather than the signals from the phenotype.
+In fact, many of the variables included in the linear models were pathway metagenes that negatively correlated with the obesity metagenes in \cref{fig:gatza_allmeta} (\cref{sub:comparison_of_the_obesity_and_pathway_associated_signatures}).
+Therefore, it may be possible that these pathways (Akt, \gls{er}, \gls{egfr}, \gls{pr}, p53, Src and \gls{tgfb}) are related to the underlying biological mechanism that drives breast tumour development, and potentially associated with the patient \gls{bmi} as well.
+
+With that said, even if these pathway signatures were related to the obesity associated genetic signatures, the fact that these results depended on the pathway associated genetic signatures from the \citet{Gatza2010a} study makes the results questionable.
 This was due to the lack of consistency of the metagenes across different cancer data sets, where more than half of the pathway metagenes showed a correlation of less than 0.8 in at least one data set (\cref{sec:pathway_associated_metagenes_and_obesity_associated_metagenes}).
@@
 These results were merely an indication that the pathway signatures may have reflected the pathway activities poorly, and it was not possible to determine whether these signatures were truly related to the corresponding pathways, as the original data was not available to reference the signatures back for validation.
-With that said, few of the pathway signatures correlated well with some clinical variables (such as \gls{er} and \gls{pr}), so not all signatures may have been of poor quality (\cref{}) (add reference to pr and er graph).
+With that said, few of the pathway signatures correlated well with some clinical variables (such as \gls{er} and \gls{pr}), so not all signatures may have been of poor quality (\cref{sec:results_for_er_and_pr_metagene_analysis}).
 
@@
 None of the pathway associated genetic signatures was similar to any of the obesity associated genetic signatures in the heatmap.
+% re-write sentence below:
 The results from the linear model prediction suggested a possible association of the obesity associated genetic signatures with the Akt, \gls{egfr} and \gls{pr} pathways.

### 2017-03-23 thesis/misc/appendix.tex
tmp commit

@@
 
+The complete list of all the pathways significantly enriched using different pathway databases (\gls{fdr} adjusted p-value \textless{} 0.05) are summarised in \cref{tab:sig_path_kegg,tab:sig_path_reactome,tab:sig_path_go}.
+
 \begin{table}[htpb]
@@
 		\begin{footnotesize}
-			\item [1] Not available (no pathway enriched)
+			\item [1] Not available (no pathway enriched).
 		\end{footnotesize}
@@
 		\centering
+		\caption{Significantly enriched \gls{go} pathways in \gls{icgc} cancer data}
 		\label{tab:sig_path_go}\\
 		Cancer type & Pathway\\
+		\hline
+		\hline
 		\endfirsthead
@@
 		\endhead
-	\hline
-	\hline
 		\rule{0pt}{2.25ex}\gls{blca} & Nucleobase-containing small molecule catabolic process                                          \\
@@
 		\hline
-	\gls{coad}                   & NA\tnote{1}                                                                                     \\
+		\gls{coad}                   & NA \tnote{1}                                                                                     \\
 		\hline
@@
 		\centering
+		\includegraphics[page=7,width=0.45\linewidth]{appendix/er_pr_cr}
+		\hfill
 		\includegraphics[page=8,width=0.45\linewidth]{appendix/er_pr_cr}\\
-		\hfill
-		\includegraphics[page=9,width=0.45\linewidth]{appendix/er_pr_cr}
 		\caption[\acrshort{er} metagene in the CR data]{Heatmap and box plot showing the association of the \gls{er} metagene with the sample gene expression and patient \gls{er} status, respectively, in the CR data.
@@
 		\centering
-		\includegraphics[page=3,width=0.45\linewidth]{appendix/er_pr_cr}\\
+		\includegraphics[page=3,width=0.45\linewidth]{appendix/er_pr_cr}
 		\hfill
-		\includegraphics[page=4,width=0.45\linewidth]{appendix/er_pr_cr}
+		\includegraphics[page=4,width=0.45\linewidth]{appendix/er_pr_cr}\\
 		\caption[\acrshort{pr} metagene in the CR data]{Heatmap and a box plot showing the association of the \gls{pr} metagene with the sample gene expression and patient \gls{pr} status, respectively, in the CR cancer data.
@@
 		\centering
+		\includegraphics[page=7,width=0.45\linewidth]{appendix/er_pr_cris}
+		\hfill
 		\includegraphics[page=8,width=0.45\linewidth]{appendix/er_pr_cris}\\
-		\hfill
-		\includegraphics[page=9,width=0.45\linewidth]{appendix/er_pr_cris}
 		\caption[\acrshort{er} metagene in the \gls{nzbc} data]{Heatmap and box plot showing the association of the \gls{er} metagene with the sample gene expression and patient \gls{er} status, respectively, in the \gls{nzbc} data.
@@
 		\centering
-		\includegraphics[page=3,width=0.45\linewidth]{appendix/er_pr_cris}\\
+		\includegraphics[page=3,width=0.45\linewidth]{appendix/er_pr_cris}
 		\hfill
-		\includegraphics[page=4,width=0.45\linewidth]{appendix/er_pr_cris}
+		\includegraphics[page=4,width=0.45\linewidth]{appendix/er_pr_cris}\\
 		\caption[\acrshort{pr} metagene in the \gls{nzbc} data]{Heatmap and a box plot showing the association of the \gls{pr} metagene with the sample gene expression and patient \gls{pr} status, respectively, in the \gls{nzbc} cancer data.
@@
 		\centering
+		\includegraphics[page=7,width=0.45\linewidth]{appendix/er_pr_fm}
+		\hfill
 		\includegraphics[page=8,width=0.45\linewidth]{appendix/er_pr_fm}\\
-		\hfill
-		\includegraphics[page=9,width=0.45\linewidth]{appendix/er_pr_fm}
 		\caption[\acrshort{er} metagene in the FM data]{Heatmap and box plot showing the association of the \gls{er} metagene with the sample gene expression and patient \gls{er} status, respectively, in the FM data.
@@
 		\centering
-		\includegraphics[page=3,width=0.45\linewidth]{appendix/er_pr_fm}\\
+		\includegraphics[page=3,width=0.45\linewidth]{appendix/er_pr_fm}
 		\hfill
-		\includegraphics[page=4,width=0.45\linewidth]{appendix/er_pr_fm}
+		\includegraphics[page=4,width=0.45\linewidth]{appendix/er_pr_fm}\\
 		\caption[\acrshort{pr} metagene in the FM data]{Heatmap and a box plot showing the association of the \gls{pr} metagene with the sample gene expression and patient \gls{pr} status, respectively, in the FM cancer data.

### 2017-03-23 thesis/results/results2.tex
tmp commit

@@
 From these results, it was clear the normalisation methods used had the most significant effect on the resulting metagenes, rather than the transformation matrices themselves (\cref{sec:normalisation_method_for_pathway_associated_genetic_signature_transformation_matrices}).
-This meant that all of the data sets had to be normalised with a single normalisation method, so the \gls{rma} normalisation method was used as this method was more reliable than the \gls{mas} method.
-%TODO: add reference for the above sentence
+This meant that all of the data sets had to be normalised with a single normalisation method, so the \gls{rma} normalisation method was used as this method was more reliable than the \gls{mas} method \citep{Irizarry2003}.
 Therefore, the GT data set was batch corrected first (\cref{sub:batch_correction}), then normalised with the \gls{rma} method and the metagene scores were ranked with fractional ranking (\cref{sub:ranking_of_the_metagene_scores}).
@@
 For example, the \gls{bcat} pathway metagene was highly correlated in all of the data sets (\textgreater{} 0.98), whereas the \gls{stat3} pathway metagene was variable across different data sets, ranging from 0.1902 to 0.9159 (\cref{tab:svd_vs_tm_path}).
-% For example, the \gls{bcat} pathway metagene was highly correlated in all of the data sets (\textgreater{} 0.98), whereas the \gls{stat3} pathway metagene was variable across different data sets, ranging from 0.1902 to 0.9159 (\cref{tab:svd_vs_tm_path,fig:allmetacor_bar}).
 The fact that some pathway associated metagenes were consistent across different data sets suggested that some of these genetic signatures were reliable and did not depend on the data set the transformation matrices were derived from.
@@
 For simplicity, it was decided that the transformation matrices for all the genetic  signatures would be made in the GT data set, since there were some pathway associated signatures that were specific to the GT data, but none of the obesity associated genetic signatures were specific to a data set.
-
-% \begin{figure}[htpb]
-% 	\centering
-% 	\includegraphics[page=2,width=0.45\linewidth]{results2/allmetacor_bar}
-% 	\includegraphics[page=16,width=0.45\linewidth]{results2/allmetacor_bar}
-% 	\caption[Bar plots showing the Spearman correlation of the \gls{svd}- and TM-derived pathway metagenes across different data sets]{Bar plots showing the Spearman correlations of the \gls{svd}-derived and the transformation matrix-derived \gls{bcat} and \gls{stat3} metagenes across different data sets.
-% 	Only the \gls{bcat} (left) and the \gls{tgfb}  (right) pathway metagenes are shown.
-% 	The scatter plots for the other pathway metagenes are shown in \cref{sec:comparison_of_the_correlation_of_svd_and_tm_generated_pathway_metagenes}.
-% 	}
-% 	\label{fig:allmetacor_bar}
-% \end{figure}
 
@@
 
-In \cref{fig:gatza_allmeta}, all of the obesity associated metagenes from the CR data set cluster together into a group, but none of the pathway associated metagenes correlated strongly with the obesity metagenes.
+In \cref{fig:gatza_allmeta}, all of the obesity associated metagenes from the CR data set cluster together into a group, but none of the pathway associated metagenes positively correlated strongly with the obesity metagenes.
 Though \gls{her2} and p63 pathway metagenes were grouped next to the cluster of obesity metagenes, the similarity was at most about 0.6  (Pearson correlation; \cref{fig:gatza_allmeta}).
+However, there seem to be a strong negative correlation with some of the pathway associated genetic signatures, such as the Akt, \gls{er}, \gls{egfr}, \gls{pr}, p53, Src and \gls{tgfb} pathways, and this strong negative correlation was observed in the other data sets as well (\cref{sec:correlation_of_the_pathway_and_obesity_metagenes_in_other_cancer_data_sets}).
+This suggests that the obesity associated genetic signatures from the CR data may be related to these pathway signatures, rather than the obesity phenotype of the patients.
+
 FM obesity associated metagene did not cluster together with the obesity metagenes from the CR data set, nor with any of the other pathway metagenes.
@@
 
-The results in this section suggested that none of the obesity associated genetic signatures were similar or related to any of the pathway associated genetic signatures from the \citet{Gatza2010a} study.
+The results in this section suggested that none of the obesity associated genetic signatures were not positively correlated with to any of the pathway associated genetic signatures from the \citet{Gatza2010a} study.
+However, some pathway signatures were negatively correlated with the obesity associated genetic signatures, which meant that the obesity signatures were likely capturing the signals from these signatures  rather than the biological signals from the obesity phenotype of the patient.
 
-\section{Prediction of obesity associated metagenes with pathway associate metagenes}
+\section{Prediction of obesity associated metagenes with pathway associated metagenes}
 \label{sec:prediction_of_obesity_associated_metagene_with_pathway_associate_metagene}
@@
 Metagene scores used for the construction of linear models were generated from the application of the transformation matrices derived from the GT data (from \cref{sec:pathway_associated_metagenes_and_obesity_associated_metagenes}) to \gls{rma} normalised \gls{nzbc} data.
+From \cref{tab:svd_vs_tm_path}, it was evident that some pathway associated genetic signatures could be variable across different data sets.
+Therefore, only six pathways (\gls{bcat}, \gls{er}, \gls{ifna}, \gls{ifny}, Myc and \gls{pr}) that were ``consistent'' (correlation \textgreater{} 0.9) across all of the data sets were used to construct the linear models.
+
 For each of the obesity metagenes identified, seven linear models were created: \gls{bmi}-only, \gls{bmi} status-only, \gls{bmi} and \gls{bmi} status, pathway metagenes-only, and all of the combinations of pathway metagenes with \gls{bmi} and/or \gls{bmi} status.
@@
 As expected, all of the linear models showed significant contribution of the \gls{pr} metagene scores to the model for all obesity metagenes (except FM metagene; \cref{sec:summary_of_the_linear_models_in_nzbc_data}).
-% Since the \gls{pr} pathway metagene showed significant association in many of the linear models, linear models were made just with the \gls{pr} metagene scores.
-% For each of the CR and FM obesity metagenes, linear models were constructed using the patient \gls{bmi}, \gls{bmi} status and \gls{pr} pathway metagene as predictor variables (\cref{tab:lm_pr_only}; \cref{sec:summary_of_the_linear_models_in_nzbc_data}).
-% As expected, all of the linear models showed significant contribution of the \gls{pr} metagene scores to the model for all obesity metagenes (except FM metagene; \cref{sec:summary_of_the_linear_models_in_nzbc_data}).
 
@@
 All of the plots in \cref{fig:predict_cris_cris,sec:summary_of_the_linear_models_in_nzbc_data} showed clearly that the data points in the scatter plots were highly variable, suggesting that even though the association between the predicted and the true metagene scores were statistically significant, the variables in the linear models may not be strongly associated to the obesity metagene.
-Furthermore, these results suggested that most of the variations in the Cr (and other) obesity metagenes were being explained by the \gls{pr} metagene and the patient \gls{bmi} contributed very little to the prediction.
+Furthermore, the fact that all of the predictions made with the models that contained the \gls{pr} metagene showed greater correlation with the true metagene values suggested that most of the variations in the Cr (and other) obesity metagenes were being explained by the \gls{pr} metagene alone, and the patient \gls{bmi} contributed very little to the prediction.
 
@@
 
-%TODO: comment on the negative relationship of the predicted and true obesity metagenes in the CR data
-
 To confirm the results shown in the \gls{nzbc} data set (the training set), the linear models were applied to CR data (the test set).
@@
 Again, these results suggested that the predicted metagene scores from the linear models may be statistically significant, but may not be relevant to the obesity metagenes which the models predict, as the predicted values were only weakly associated with the true values.
+
+Another thing to note from \cref{fig:predict_cr_cris} was the apparent negative assocation of the predicted metagenes with the original metagenes in the \gls{bmi}-only, the \gls{bmi} status-only and the \gls{bmi} and \gls{bmi} status models.
+One possible explanation for this was that these \gls{bmi}-based models were overfitted in the \gls{nzbc} data set, and was not generalisable in the CR data set, and thus produced negative correlations.
+In fact, it was evident from the metagene analyses in \cref{cha:obesity_genetic_signatures_and_cancer} that these obesity associated genetic signatures were not generalisable with the patient \gls{bmi} in different data sets.
+% An alternative explanation was that the obesity metagenes  were not generalisable in different data sets, and thus the models were predicting the obesity metagenes relatively well in \gls{nzbc} data, but not in the CR data.
 
@@
 
-Lastly, since only the ``consistent'' pathway metagenes were picked from all of the available pathway metagene scores, linear models were constructed with all of the pathway metagenes in a step wise fashion.
-Step wise construction of a linear model is where the variables for a model are selected computationally depending on the contribution of the variable to the resulting linear model, and thus contains only the variables that ``matter'' to the final linear model.
-Patient \gls{bmi}, \gls{bmi} status and all of the pathway metagene scores were used to generate the step wise linear model for each of the obesity metagenes.
-The linear models that resulted from the step wise selection of the variables were summarised in \cref{tab:stepwise_model}.
+Lastly, since only the six most ``consistent'' pathway metagenes were picked from a total of eighteen pathway metagene scores, linear models were constructed with all eighteen pathway metagenes in a stepwise fashion.
+Step wise construction of a linear model is where the variables for a model are selected computationally depending on the contribution of the variable to the resulting linear model, and thus contains only the variables that ``matter'' to the final linear model \cref{sec:obesity_metagene_prediction_with_pathway_metagenes}.
+Patient \gls{bmi}, \gls{bmi} status and all of the pathway metagenes (Akt, \gls{bcat}, E2F1, \gls{egfr}, \gls{er}, \gls{her2}, \gls{ifna}, \gls{ifny}, Myc, p53, p63, \gls{pi3k}, \gls{pr}, Ras, \gls{stat3}, Src, \gls{tgfb} and \gls{tnfa}) were used to generate the stepwise linear model for each of the obesity metagenes.
+The linear models that resulted from the stepwise selection of the variables were summarised in \cref{tab:stepwise_model}.
 
@@
 	\centering
-	\caption[Description of the step wise linear models constructed from the \gls{nzbc} data to predict all of the obesity metagenes]{Description of the step wise linear models constructed from the \gls{nzbc} data to predict all of the obesity metagenes}
+	\caption[Description of the stepwise linear models constructed from the \gls{nzbc} data to predict all of the obesity metagenes]{Description of the stepwise linear models constructed from the \gls{nzbc} data to predict all of the obesity metagenes}
 	\label{tab:stepwise_model}\\
@@
 The statistical significance and the strong association of the predicted metagenes were observed in the CR data set as well (\cref{fig:stepwise_cr}).
-Though the $R^2$ values were slightly lower than those from the \gls{nzbc} data set, the $R^2$ values from the CR data prediction were more promising than the predictions made by the non-stepwise linear models.
+Though the $R^2$ values were slightly lower than those from the \gls{nzbc} data set, the $R^2$ values from the CR data predictions were more promising than the predictions made by the non-stepwise linear models.
+
+With that said, some of the predicted obesity metagenes were negatively correlated with the original metagene values in the CR data (\cref{fig:stepwise_cr}).
+Again, this could have been due to the model being overfitted in the \gls{nzbc} data, and thus made a wrong prediction about the obesity metagenes.
+Nevertheless, the stepwise linear models were able to predict the obesity associated metagenes better than the models constructed from the models made in \cref{sub:linear_model_prediction_in_the_nzbc_and_cr_data_sets}.
 
@@
 	\includegraphics[page=10,width=0.32\linewidth]{results2/stepwise_cris_cris}
-	\caption[Comparison of all the obesity metagene scores predicted from the step wise linear models with the obesity metagene scores from the \gls{nzbc} data]{Scatter plots comparing all of the obesity metagene scores predicted by the step wise linear models constructed from the \gls{nzbc} data, with the corresponding obesity metagene scores from the \gls{nzbc} data.
+	\caption[Comparison of all the obesity metagene scores predicted from the stepwise linear models with the obesity metagene scores from the \gls{nzbc} data]{Scatter plots comparing all of the obesity metagene scores predicted by the stepwise linear models constructed from the \gls{nzbc} data, with the corresponding obesity metagene scores from the \gls{nzbc} data.
 	P-values and $R^2$-value are as described in previous figures.}
@@
 	\includegraphics[page=10,width=0.32\linewidth]{results2/stepwise_cr_cris}
-	\caption[Comparison of all the obesity metagene scores predicted from the step wise linear models with the obesity metagene scores from the CR data]{Scatter plots comparing all of the obesity metagene scores predicted by the step wise linear models constructed from the \gls{nzbc} data, with the corresponding obesity metagene scores from the CR data.
+	\caption[Comparison of all the obesity metagene scores predicted from the stepwise linear models with the obesity metagene scores from the CR data]{Scatter plots comparing all of the obesity metagene scores predicted by the stepwise linear models constructed from the \gls{nzbc} data, with the corresponding obesity metagene scores from the CR data.
 	P-values and $R^2$-value are as described in previous figures.}
@@
 
-Taken together, these results suggested that the variables included in these step wise models were relevant to the biology of the obesity metagenes identified in the CR and FM data sets.
+Taken together, these results suggested that the variables included in these stepwise models were relevant to the biology of the obesity metagenes identified in the CR and FM data sets.
 However, there were no single common pathway that was present in all of the models; although Akt, \gls{egfr} and \gls{pr} pathway metagenes were frequently included in many of the stepwise models.

### 2017-03-23 thesis/intro/intro.tex
tmp commit

@@
 \gls{who} (2014) have used \gls{bmi} measurements to define overweight and obesity in adults as individuals with \gls{bmi} $\geq$ 25 kg/m$^2$ and $\geq$ 30 kg/m$^2$, respectively.
-For children under the age of 5, overweight and obesity were defined as weight-for-height greater than 2 standard deviation and 3 standard deviation above \gls{who} Child Growth Standards median, respectively.
-For children aged between 5 to 19 years, overweight and obesity were defined as BMI-for-age greater than 1 standard deviation and 2 standard deviation above \gls{who} Growth Reference Standards median, respectively.
+For children under the age of 5, overweight and obesity are defined as weight-for-height greater than 2 standard deviation and 3 standard deviation above \gls{who} Child Growth Standards median, respectively.
+For children aged between 5 to 19 years, overweight and obesity are defined as BMI-for-age greater than 1 standard deviation and 2 standard deviation above \gls{who} Growth Reference Standards median, respectively.
 
-However, the use of \gls{bmi} as an accurate measure of body fat composition and/or representation of the individual's metabolic status remains controversial.
-In fact, many studies have suggested the use of other measurements such as waist-to-height and waist-to-hip ratios to classify the patients into different categories, as these measurements were better predictors of the obesity related diseases than \gls{bmi} \citep{Dalton2003,Gelber2008,Lee2008}.
-With that said, \gls{bmi} is still widely used due to the ease of data collection compared to other measurements and, as a result, most of the publicly available clinical data have  height and weight data of the patients from which one can calculate the patients' \gls{bmi}.
+However, the use of \gls{bmi} as an accurate measure of body fat composition and/or representation of the metabolic state of an individual remains controversial.
+In fact, many studies have suggested the use of other measurements such as waist-to-height and waist-to-hip ratios to classify the patients into different categories, as these measurements were better predictors of the diseases related to obesity than \gls{bmi} \citep{Dalton2003,Gelber2008,Lee2008}.
+With that said, \gls{bmi} is still widely used due to the ease of data collection compared to other measurements and, as a result, most of the publicly available clinical data have  height and weight data of the patients from which one can calculate the \gls{bmi} information.
 
@@
 Furthermore, the prevalence of overweight in children under 5 years was estimated to reach from 6.3\% in 2013 to 11\% by 2025 if the current trend continued \citep{WHO2014}.
-The current obesity status of the adult population is bad already, but the fact that more of the younger population who are overweight and obese poses a much greater risk for the health of the population.
+The current obesity status of the adult population is bad already, but the fact that more of the younger population who are overweight and obese poses a much greater risk for the health of the population in the future.
 
@@
 Though there are many hormones that regulate satiety, only leptin and insulin will be covered here.
-Leptin gene was first discovered by \citet{Zhang1994} and the discovery of the leptin receptor gene was followed shortly after by \citet{Tartaglia1995}.
-Together, the identification of the leptin and leptin receptor genes was a major break through in the field of obesity research.
-
+% Leptin gene was first discovered by \citet{Zhang1994} and the discovery of the leptin receptor gene was followed shortly after by \citet{Tartaglia1995}.
+% Together, the identification of the leptin and leptin receptor genes was a major break through in the field of obesity research.
 Leptin is a hormone that signals satiety and reduces food intake, and is secreted mainly by the white adipose tissue \citep{Moustafa2013,Zhang1994}.
@@
 
-The evidence of the involvement of insulin in the satiety control was not particularly clear even though there were strong evidence of obesity with neuron-specific deletion of insulin receptor, and the fact that the insulin level, like leptin, was proportional to the fat mass also suggested the role of insulin in satiety and body weight control \citep{Barsh2002, Bruning2000, Woods1979}.
-This was perhaps due to the fact that leptin and insulin used two different signalling pathways in the cell; leptin used the \gls{jak}/\gls{stat} signalling pathway, whereas insulin used the \gls{pi3k} pathway for signalling \citep{Ghilardi1996}.
+The evidence of the involvement of insulin in the satiety control was not particularly clear even though there were strong evidence of obesity with neuron-specific deletion of insulin receptor.
+% TODO: citation for the sentence above (and check with citation below)
+The fact that the insulin level, like leptin, was proportional to the fat mass also suggested the role of insulin in satiety and body weight control \citep{Barsh2002, Bruning2000, Woods1979}.
+This was perhaps due to the fact that leptin and insulin used two different signalling pathways in the cell: leptin used the \gls{jak}/\gls{stat} signalling pathway, whereas insulin used the \gls{pi3k} pathway for signalling \citep{Ghilardi1996}.
 However, leptin, and subsequently insulin, was shown to activate ATP-dep\-endent K$^+$ channel in glucose-responsive neurons in the hypothalamus \citep{Spanswick1997, Spanswick2000}.
@@
 
-As one can suspect, there are many factors involved with the likelihood of an individual to become overweight and obese.
-
 \subsubsection{Individual-level (micro-level) factors}
@@
 
-Even though environmental factors are probably the most significant contributor to the current global obesity epidemic, there are some individual-level, genetic factors that contribute to the overall increase in the proportion of the population that are obese.
 There is much evidence of obesity being caused by individual-level genetic factors that make some individuals more predisposed to obesity than others.
@@
 Since these mutations are usually directly related to the key constituents in the pathway mentioned earlier (\cref{sub:physiological_mechanism_of_obesity}), the resulting phenotype from the mutation causes severe early-onset obesity.
-However, although these mutations cause a devastating effect on the individual carrying the mutation, these mutations occur very rarely in the population.
+However, although these mutations cause a devastating effect on those who carry the mutation, these mutations occur very rarely in the population.
 
@@
 They further showed that the mutation caused a frameshift mutation in the protein, resulting in an inactive form of leptin that was not able to be secreted \citep{Montague1997}.
-As a result, children with this mutation was constantly hungry due to the lack of anorexogenic signals, and became obese at a very early stage of their lives.
-Though these mutations are very rare and affects only those individuals that have one of these mutations, it had a profound impact in the field of obesity research and contributed significantly to the knowledge of the mechanism of the development of obesity.
+As a result, children with this mutation were constantly hungry due to the lack of anorexogenic signals and became obese at a very early stage of their lives.
+% Though these sort of mutations are very rare and affects only those individuals that have these mutations, it had a profound impact in the field of obesity research and contributed significantly to the knowledge of the mechanism of the development of obesity.
 
-Polygenic obesity is where a combination of variants in multiple genes or multiple regions of the genome, together with the environmental factors, causes obesity in the individual \citep{Moustafa2013}.
+Polygenic obesity is where a combination of variants in multiple genes or multiple regions of the genome (together with the effect from the environmental factors) causes obesity in the individual \citep{Moustafa2013}.
 Unlike in monogenic obesity where a single mutation in a gene causes severe obesity, the variants in some of the genes or genomic regions involved in polygenic obesity have little effect on the obesity status of the individual by itself, but in combination with many other variants and together with environmental factors, it has a serious obesogenic effect on the individual.
@@
 
+These individual-level genetic factors contribute to the overall increase in the proportion of the population that are obese.
+% TODO: citation for the sentence below?
+However, it is likely that the environmental factors are probably the most significant contributor to the current global obesity epidemic,
+
 \subsubsection{Population-level (macro-level) factors}
@@
 
+% TODO: citation for the sentence below?
 It is clear that the individual-level factors cannot be accounted for the global epidemic of obesity that the world currently observes.
@@
 On the other hand, the prices of fruits and vegetables remain expensive due to the lack of support by the government to lower the cost associated with the production \citep{Malik2013}.
-
 As a result, more people are inclined to consume food that are cheap, have little nutritional value and high in energy, than the food that are expensive, nutritional and healthy alternatives.
@@
 As more people take up urban lifestyle, they are exposed to the new environment where there is a better range of food selection, as well as technological and mechanical advancement in transport and other daily chores that may not be available in rural areas.
-
 Although there are many advantages involved with urbanisation, such as access to developed health care systems, education and advanced technologies, the lifestyle change also impose negative health behaviour that ultimately lead to positive energy balance and therefore obesity \citep{Malik2013}.
 Lack of physical activity is one of the obvious components that is linked with the urbanisation of the population.
-In many urban cities, the recommended level of physical exercise is difficult to achieve and maintain due to the shift towards the sedentary behaviours that are encouraged by the developed transport system, automated household chores, and indoor entertainment \citep{Malik2013}.
+In many urban cities, the recommended level of physical exercise is often difficult to achieve and maintain due to the shift towards the sedentary behaviours that are encouraged by the developed transport system, automated household chores and indoor entertainment \citep{Malik2013}.
 
-The last factor to be mentioned is the income and socioeconomic status of the population.
+The last factor to be mentioned here is the income and socioeconomic status of the population.
 As the average income increases, more people are able to buy food and are likely to take up the sedentary lifestyle.
@@
 Access to the developed health care system will allow weight maintenance of the population, but this is likely to be of most benefit for the high-income group in the population.
-The reason for this, quite simply, is the cost associated with the visit to the health care system, as high-income group is more likely to make a consistent and periodic visit than the low or middle-income group of the population \citep{Malik2013}.
+The reason for this, quite simply, is the cost associated with the visit to the health care system, as high-income group is more likely to make a consistent and periodic visit than the low- or middle-income groups of the population \citep{Malik2013}.
 
-So, for the low- and middle-income groups, the rise in average income in these groups allow the people to have access to food and activities that promotes sedentary bahaviours, but the limited access to the health care system, education and recreational activities that promote physical activites ultimately lead to positive energy balance \citep{Malik2013}.
+So, for the low- and middle-income groups, the rise in the average income in these groups allow the people to have access to food and activities that promotes sedentary bahaviours, but the limited access to the health care system, education and recreational activities that promote physical activites ultimately lead to positive energy balance \citep{Malik2013}.
 In contrast, the high-income group of the population have better access to the facilities that promote weight maintenance, thus have less of an impact compared with the low- and middle-income groups of the population.
@@
 More than half of cancer deaths are caused by lung, breast, colorectal, stomach and liver cancers \citep{WHO2014}.
-The leading cause of deaths in high-income countries is lung cancer in both men and women, then colorectal cancer and breast cancer for men and women, respectively \citep{WHO2014}.
+The leading cause of cancer deaths in high-income countries is lung cancer in both men and women, then colorectal cancer and breast cancer for men and women, respectively \citep{WHO2014}.
 Although the levels of mortality vary among different countries, cervical, liver and stomach cancers make up a large proportion of cancer deaths among low- and middle-income countries \citep{WHO2014}.
@@
 \citet{Bray2013} estimated the global prevalence of cancer in adult population in 2008.
-Since there is no clear definition for cancer prevalence, \citet{Bray2013} used \gls{ldp} as an estimate for cancer prevalence, and measured the incidence of cancer during the period between 2004 and 2008 (5 years).
+Since there was no clear definition for cancer prevalence, \citet{Bray2013} used \gls{ldp} as an estimate for cancer prevalence, and measured the incidence of cancer during the period between 2004 and 2008 (5 years).
 From these estimates, Eastern Asia had the greatest 5-year prevalence of cancer, with approximately 7 million people with cancer in 2008, followed by North America and Western Europe with approximately 4.7 and 3 million people with cancer \citep{Bray2013}.
@@
 The rates were 1.2 and 1.4 times greater in the M\=aori population compared to the non-M\=aori population for males and females, respectively \citep{Health2016b}.
-Again, this difference between the M\=aori and non-M\=aori population is unique to \gls{nz}, and improved understanding of the underlying problem is essential for better treatment plans and outcomes.
+Again, this difference between the M\=aori and non-M\=aori population is unique to \gls{nz}, and improved understanding of the underlying problem is essential for better treatment plans and outcomes for the M\=aori population.
 
@@
 
-\begin{figure}[h!]
+\begin{figure}[htb!]
 	\centering
@@
 In short, when \acrshort{dna} damage is sensed, p53 protein transcribes the pro-apoptotic members of the \gls{bcl2} family proteins, such as Bax, Puma and Noxa \citep{Fridman2003,Hanahan2011}.
-The fact that p53 protein is involved not only in the regulation of cell cycle, but also in the apoptotic pathway shows its importance in the control of cancer development, and perhaps that is the reason why more than 50\% of tumours have a mutation in the \textit{TP53} gene \citep{Levine1997}.
+The fact that the p53 protein is involved not only in the regulation of cell cycle, but also in the apoptotic pathway shows its importance in the control of cancer development, and perhaps that is the reason why more than 50\% of tumours have a mutation in the \textit{TP53} gene \citep{Levine1997}.
 
@@
 When acquired by cancer cells, the previous three hallmarks allow the cells to continuously grow in an uncontrolled manner.
-However, this is not the case as the disruption in cell signalling on its own does not trigger the rapid expansion of the tumour; this occurs only when the cells have the capability of replicating infinitely \citep{Hanahan2000, Hanahan2011}.
+However, this is not the case, as the disruption in cell signalling on its own does not trigger the rapid expansion of the tumour; this occurs only when the cells have the capability of replicating infinitely \citep{Hanahan2000, Hanahan2011}.
 
-There are two barriers in achieving the infinite replicative capacity: senescence and crisis \citep{Hanahan2011}.
+There are two barriers in achieving infinite replicative capacity: senescence and crisis \citep{Hanahan2011}.
 When cells are grown in a culture, cells are able to replicate and grow until a certain point, and the cells reach the stage of senescence where the cells stop growing but are still viable \citep{Hanahan2011}.
@@
 \noindent
-Like normal cells, cancer cells also require access to oxygen and nutrient supply from the blood.
+Like normal cells, cancer cells also require access to oxygen and nutrient supplies from the blood.
 However, new blood vessel formation, or angiogenesis, only occur in a handful of occasions, and thus the tumour needs to acquire the ability to stimulate angiogenesis in the microenvironment in which they live in \citep{Hanahan2011}.
@@
 One of the prime reasons that cancers are so difficult to treat and have high mortality rate is due to their ability to metastasise to other parts of the body.
-Once metastasised to different parts of the body, the clinician not only has to consider the primary tumour site, but also the second, third, or even fourth tumour sites, which may or may not have similar tumour biology to the primary tumour, and thus have to consider multiple treatment plans to cure the patient.
+Once metastasised to different parts of the body, the clinician not only has to consider the primary tumour site, but also the second, third, or even fourth tumour sites, which may or may not have similar tumour biology as the primary tumour, and thus have to consider multiple treatment plans to cure the patient.
 
@@
 There are now evidence that show otherwise, whereby a single catastrophic event (termed \gls{chromothripsis}) occurs in the chromosome structure, resulting in major structural rearrangements such as inversions, deletions, duplications and translocations \citep{Leibowitz2015,Stephens2011}.
-Owing to its nature of being a one-off chromosomal disruption, chromothripsis (usually) cause disruptions in tumour suppressor genes rather than duplications of oncogenes \citep{Leibowitz2015}.
+Owing to its nature of being a one-off chromosomal disruption, \gls{chromothripsis} (usually) cause disruptions in tumour suppressor genes rather than duplications of oncogenes \citep{Leibowitz2015}.
 
@@
 Inflammation and immune response, though it may assist in the control of diseases and injuries, have been implicated in the progression of tumour and their acquisition of many of the cancer hallmarks \citep{Hanahan2011}.
-There are two key transcription factors of inflammation that ultimately leads to the prommotion  of many, if not all, of the hallmarks discussed so far: \gls{nfkb} and \gls{stat3} \citep{Mantovani2008}.
+There are two key transcription factors of inflammation that ultimately leads to the promotion  of many, if not all, of the hallmarks discussed so far: \gls{nfkb} and \gls{stat3} \citep{Mantovani2008}.
 
@@
 
-Immunoediting is a process where the immune system successfully eliminates the highly immunogenic cancer cells during surveillance, but fails to eliminate the weakly immunogenic cells and therefore poses a selective advantage on these cells \citep{Hanahan2011,Teng2008}.
-In fact, there have been reports where the cancer cells in the immunodeficient mice were not able to initiate tumour growth in the immunocompetent mice, but the cancer cells from the immunocompetent mice were able to do so in the immunodeficient mice \citep{Hanahan2011}.
+\Gls{immunoediting} is a process where the immune system successfully eliminates the highly immunogenic cancer cells during surveillance, but fails to eliminate the weakly immunogenic cells and therefore poses a selective advantage on these cells \citep{Hanahan2011,Teng2008}.
+In fact, there have been reports where the cancer cells in the immunodeficient mice not being able to initiate tumour growth in the immunocompetent mice, but the cancer cells from the immunocompetent mice were able to do so in the immunodeficient mice \citep{Hanahan2011}.
 One possible reason for this was that the cancer cells from the immunodeficient mice were not immunoedited to the extent in which the cells from immunocompetent mice were, and thus unable to evade the immune system as efficiently as those cells in the immunocompetent mice \citep{Hanahan2011}.
@@
 
-Any one of the ten Hallmarks of Cancers has the potential to initiate cancer growth and progression and every hallmark has their own complex mechanistic pathways.
+Any one of the ten Hallmarks of Cancers has the potential to initiate cancer growth and progression, and every hallmark has their own complex mechanistic pathways.
 Furthermore, every pathway has many proteins associated with its function or role in the cell and, though their impacts may differ from protein to protein, mutations in any one of the proteins involved in a pathway have the potential to cause cancer.
 
-Oncogenes are genes that promote or ``accelerate'' the process of tumorigenesis and/or tumour progression, whereas tumour suppressor genes prevent or ``decelerate'' those effects \citep{Vogelstein2004a}.
-Cancers usually mutate oncogenes in such a way that they become constitutively activated, and mutate the tumour suppressor genes so that these genes are inactivated and/or have reduced expression \citep{Vogelstein2004a}.
-Another thing to note for these genes is that the oncogenes usually only require a ``one-hit'' mutation that affects a single allele, whereas the tumour suppressor genes generally require mutations in both maternal and paternal alleles to have any effect \citep{Stratton2009,Vogelstein2004a}.
+\Glspl{oncogene} are genes that promote or ``accelerate'' the process of tumorigenesis and/or tumour progression, whereas \gls{suppressor} genes prevent or ``decelerate'' those effects \citep{Vogelstein2004a}.
+Cancers usually mutate \glspl{oncogene} in such a way that they become constitutively activated, and mutate the \gls{suppressor} genes so that these genes are inactivated and/or have reduced expression \citep{Vogelstein2004a}.
+Another thing to note for these genes is that the \glspl{oncogene} usually only require a ``one-hit'' mutation that affects a single allele, whereas the \gls{suppressor} genes generally require mutations in both maternal and paternal \glspl{allele} to have any effect \citep{Stratton2009,Vogelstein2004a}.
 
-Stability genes, also known as caretakers, are genes that are involved in, for example,  \gls{mmr}, \gls{ner} and \gls{ber}; in other words, genes that are involved in \acrshort{dna} repair mechanisms \citep{Vogelstein2004a}.
+\Glspl{stability}, also known as caretakers, are genes that are involved in, for example,  \gls{mmr}, \gls{ner} and \gls{ber}; in other words, genes that are involved in \acrshort{dna} repair mechanisms \citep{Vogelstein2004a}.
 These caretaker genes help prevent the cell from developing new mutations during \acrshort{dna} replication and mitosis.
-However, only the oncogenes or the tumour suppressor genes are able to directly affect the biology of the cell, and therefore mutations in stability genes tend to affect the mutation rates of the other classes of genes, ultimately leading to tumorigenesis \citep{Vogelstein2004a}.
+However, only the \glspl{oncogene} or the \gls{suppressor} genes are able to directly affect the biology of the cell, and therefore mutations in \glspl{stability} tend to affect the mutation rates of the other classes of genes, ultimately leading to tumorigenesis \citep{Vogelstein2004a}.
 
-There are two types of mutations that occurs in an individual: germline mutation and somatic mutation.
-Germline mutations are the mutations that are passed on from the parents of an individual, and somatic mutations are the mutations acquired by the cell over the course of the lifetime of an individual.
+There are two types of mutations that occurs in an individual: \gls{germline} and \gls{somatic}.
+\Glspl{germline} are the mutations that are passed on from the parents of an individual, and \glspl{somatic} are the mutations acquired by the cell over the course of the lifetime of an individual.
 
-Unfortunately, there are germline mutations that make the individual more prediposed to cancer \citep{Vogelstein2004a}.
+Unfortunately, there are \glspl{germline} that make the individual more prediposed to cancer \citep{Vogelstein2004a}.
 These individuals have a mutation in one or more of the classes of cancer-related genes, which makes them more likely to develop a tumour than individuals without the mutation \citep{Vogelstein2004a}.
-In contrast, somatic mutations are acquired during the lifetime of an individual, and is not passed down to their children.
-Somatic mutations can either be a driver or a passenger mutation, where a driver mutation helps tumorigenesis and its subsequent progression (in other words, mutations in oncogene, tumour suppressor, or stability gene), and passenger mutation have neither a selective advantage nor contribute to tumour progression \citep{Stratton2009}.
+In contrast, \glspl{somatic} are acquired during the lifetime of an individual, and is not passed down to their children.
+\Glspl{somatic} can either be a \gls{driver} or a \gls{passenger} mutation, where a \gls{driver} mutation helps tumorigenesis and its subsequent progression (in other words, mutations in \gls{oncogene}, \gls{suppressor}, or stability gene), and \gls{passenger} mutation have neither a selective advantage nor contribute to tumour progression \citep{Stratton2009}.
 
-Since cancers require multiple driver mutations, somatic mutations are an essential part of tumour development.
-Perhaps that is one of the reasons why, in some cancers, germline mutation in the stability genes (such as \textit{BRCA1} and \textit{BRCA2} genes) cause greater susceptibility to cancers than mutations in the oncogene or the tumour suppressor gene \citep{Vogelstein2004a}.
+Since cancers require multiple \gls{driver} mutations, \glspl{somatic} are an essential part of tumour development.
+Perhaps that is one of the reasons why, in some cancers, \gls{germline} in the \glspl{stability} (such as \textit{BRCA1} and \textit{BRCA2} genes) cause greater susceptibility to cancers than mutations in \glspl{oncogene} or \gls{suppressor} genes \citep{Vogelstein2004a}.
 
@@
 For current smokers, the relative risk is greater than 1.5 in many cancers, including liver, cervix and kidney, and it can rise up to 8.96 in lung cancer \citep{Gandini2008}.
-Tobacco smoke contain many carcinogens which are processed first by \gls{cyp} enzymes (termed meta\-bolic activation), and the product can either be processed further to be detoxified completely and cause no harm to the cell, or alternatively, the product can covalently bind to the \acrshort{dna}, forming \acrshort{dna} adducts \citep{Hecht1999}.
+Tobacco smoke contain many \gls{carcinogen} which are processed first by \gls{cyp} enzymes (termed meta\-bolic activation), and the product can either be processed further to be detoxified completely and cause no harm to the cell, or alternatively, the product can covalently bind to the \acrshort{dna}, forming \acrshort{dna} adducts \citep{Hecht1999}.
 Unless these adducts are resolved correctly by the \acrshort{dna} repair mechanisms, it can result in a permanent mutation within the genome and lead to oncogenesis \citep{Hecht1999}.
@@
 
-In a more recent meta-analysis study, \citet{Renehan2008} showed that an increase in \gls{bmi} by 5 kg/m$^2$ significantly increased the risks of esophageal, thyroid, colon and renal cancers in men, and endometrial, gallbladder, esophageal and renal cancers in women.
+\citet{Renehan2008} showed in their meta-analysis study  that an increase in \gls{bmi} by 5 kg/m$^2$ significantly increased the risks of esophageal, thyroid, colon and renal cancers in men, and endometrial, gallbladder, esophageal and renal cancers in women.
 The summary of the risks associated with various cancers from this study are shown in \cref{tab:renehan_cancer_risks}.
@@
 \Glspl{adipocyte} synthesises enzymes, such as aromatase, that convert precursor molecules into different forms of androgens and estrogens \citep{Calle2004}.
-Many studies have confirmed the mitogenic effect of sex hormones on certain cell types; mainly breast and endometrial cell types \citep{Roberts2010}.
+Many studies have confirmed the mitogenic effect of sex hormones on certain cell types, mainly breast and endometrial cell types \citep{Roberts2010}.
 
@@
 Adiponectin is a hormone that is synthesised and secreted only by the \glspl{adipocyte} \citep{Kelesidis2006}.
-Unlike leptin, where the concentration of leptin is higher in those who are obese, adiponectin levels are lower in obese people compared to the normal weight people \citep{Kelesidis2006,Renehan2006}.
-Adiponectin have anti-angiogenic and anti-proliferative effects on the cells, and is secreted by mature adipocytes but not by the premature pre-adipocytes \citep{Gilbert2013}.
+Unlike leptin, where the concentration of leptin is higher in those who are obese, adiponectin levels are lower in people that are obese compared to those with normal weight \citep{Kelesidis2006,Renehan2006}.
+Adiponectin have anti-angiogenic and anti-proliferative effects on the cells, and is secreted by mature \glspl{adipocyte} but not by the premature pre-adipocytes \citep{Gilbert2013}.
 During chronic inflammatory state in obese patients, pro-inflammatory proteins such as \gls{tgfb} and \gls{ifny} can block the maturation of the pre-adipocytes, thus decreasing the level of adiponectin and the accompanying beneficial effects \citep{Gilbert2013}.
@@
 
-It is evident that there are many possible mechanisms in which obesity contributes to the progression of cancer.
+It is evident that there are many possible mechanisms in which obesity contributes to the progression of cancer by activating many of the Hallmarks of Cancer mentioned in \cref{sub:mechanisms_of_cancer_development}.
 Furthermore, these mechanisms are related to one another and are part of a larger, more complex physiological network \citep{Renehan2006}.
-Due to this complexity, it is difficult to confidently explain how obesity contributes to the tumour development.
+Due to this complexity, it is difficult to confidently explain how obesity contributes to the tumour development, and therefore the biological link between obesity and cancer is of great interest.
 
@@
 
-One of the difficulties associated with the treatment of cancers is the identification of the underlying mechanisms or hallmarks that drive the cancer progression.
+One of the difficulties associated with the treatment of cancers is the identification of the underlying biological mechanisms or hallmarks that drive the cancer progression.
 As one can imagine, it is difficult for the clinicians to treat a patient if they do not know what is causing the disease.
-The reason why it is so difficult to identify the underlying biological mechanisms is that there are very little reliable biological markers (or ``genetic signatures'') that signify the presence or absence of the causal pathways.
+The reason why it is so difficult to identify the underlyig biological mechanisms is that there are very little reliable biological markers (or ``genetic signatures'') that signify the presence or absence of the causal pathways.
 With that said, significant improvements have been made in recent years owing to the advancement in the sequencing technologies and have allowed the researchers to study the genetic signatures that represent the hallmarks in a much greater detail.
@@
 
-There is a clear evidence of obesity being a major risk factor for many cancer types (\cref{sec:obesity_and_cancer}), but the genes or pathways that directly contribute to tumorigenesis in patients that are obese is yet to be clarified.
+There is a clear epidemiological evidence of obesity being a major risk factor for many cancer types (\cref{sec:obesity_and_cancer}), but the genes or pathways that directly contribute to tumorigenesis in patients that are obese is yet to be clarified.
 Two studies in particular, one by \citet{Creighton2012} and another by \citet{Fuentes-Mattei2014}, have investigated whether there were any difference in genetic composition between the obese and the non-obese patients, specifically in a group of breast cancer patients.
@@
 Of the 103 samples, 35\% were from normal weight, 28\% were from overweight and 37\% were from obese patients.
-It was also noted that African American patients were more likely to be obese than the Caucasian patients.
+It was also noted that African-American patients were more likely to be obese than the Caucasian patients.
 They compared the gene expression between the tumours from obese patients with the tumours from non-obese patients and identified 799 gene probe sets (662 unique genes) with p \textless{} 0.01 and with a \gls{fc} \textgreater{} 1.2.
@@
 One thing to note here is that most of these pathway associated genetic signatures are involved in one or more of the Hallmarks of Cancer (\cref{subsubsec:cancerhallmarks}).
-This means that these pathway associated signatures can be used to provide insights into the underlying biological mechanisms of the samples used in this project and help understand the underlying biology of cancers in obese patients.
+This means that these pathway associated signatures can be used to provide insight into the underlying biological mechanisms of the samples used in this project and help understand the underlying biology of cancers in obese patients.
 

### 2017-03-23 thesis/methods/methods.tex
tmp commit

@@
 The raw microarray data from the \citet{Creighton2012}, \citet{Fuentes-Mattei2014} and \citet{Gatza2010a}  studies were downloaded from the \gls{geo} website (\citealp{Edgar2002}; \citetalias{GEO}).
-In addition to these breast cancer data sets, the raw microarray data set from a cohort of \gls{nz} breast cancer patients used in the \citet{Print2016} study was also downloaded from the \gls{geo} website.
-\gls{rnaseq} and clinical data of multiple different cancer types were downloaded from the \gls{icgc} (\citetalias{ICGC}; \citealp{Zhang2011}) and \gls{tcga} \citepalias{TCGA}, respectively.
+In addition to these breast cancer data sets, the raw microarray data set from a cohort of \gls{nz} breast cancer patients used in the \citet{Print2016} study was downloaded from the \gls{geo} website.
+\gls{rnaseq} and clinical data of multiple different cancer types were downloaded from the \gls{icgc} (\citetalias{ICGC}; \citealp{Zhang2011}) and \gls{tcga} \citepalias{TCGA} websites, respectively.
 
@@
 Clinical data for the samples (age, ethnicity, tumour grade, \gls{er}/\gls{pr}/\gls{her2} statuses, and treatments used) in this study was also downloaded from the \gls{geo} database (same \gls{geo} accession ID).
+Patient height, weight and \gls{bmi} information were not available in this data set.
 130 obesity associated gene probes identified by \citet{Fuentes-Mattei2014} were taken from the Supplementary Table 3 from their paper.
@@
 Since multiple gene probe sets match back to a single gene of interest in a microarray chip, there were conflicting expression data for some of the genes after the conversion of the gene probe sets into gene symbols.
-For the gene symbols that had multiple probe set entries, a single probe set was chosen to represent the gene symbol by using the \texttt{collapseRows} function in the \textit{WGCNA} package in \gls{r}, using the default parameters (which chose the probe set with a maximum mean value) \citep{Langfelder2008}.
+For the gene symbols that had multiple probe set entries, a single probe set was chosen to represent the gene symbol by using the \texttt{collapseRows} function in the \textit{WGCNA} package in \gls{r}, using the default parameters (which chose the probe set with a maximum mean expression value) \citep{Langfelder2008}.
 Likewise, any obesity associated or pathway associated gene probes were converted into gene symbols.
@@
 	\newpage
-	\begin{longtable}{lcc}
-		\caption{Summary of the clinical variables in the Creighton \textit{et al.} and \gls{nzbc} microarray data}
+	\begin{longtable}{lccc}
+		\caption{Summary of the clinical variables in the Creighton \textit{et al.},  Fuentes-Mattei \textit{et al.} and \gls{nzbc} microarray data}
 		\label{tab:clin_summary}\\
-		Clinical variable & Creighton \textit{et al.} & \gls{nzbc} \\
+		Clinical variable & Creighton \textit{et al.} & Fuentes-Mattei \textit{et al.} & \gls{nzbc} \\
+		\hline
+		\hline
 		\endfirsthead
 		\multicolumn{3}{c}{\tablename\ \thetable{}\ (continued)}\\
-		\endhead
 		\hline
 		\hline
+		\endhead
 		\rule{0pt}{2.25ex}Age \\
-		\hspace{1em} Min.                    & 30 & 31 \\
-		\hspace{1em} Max.                    & 72 & 92 \\
-		\hspace{1em} Mean                    & 49 & 59 \\
-		\hspace{1em} Median                  & 48 & 60 \\
+		\hspace{1em} Min.                    & 30 & 26 & 31 \\
+		\hspace{1em} Max.                    & 72 & 79 & 92 \\
+		\hspace{1em} Mean                    & 49 & 52 & 59 \\
+		\hspace{1em} Median                  & 48 & 51 & 60 \\
 		\hline
 		\rule{0pt}{2.25ex}Ethnicity         \\
-		\hspace{1em} Caucasian               & 77 & 71 \\
-		\hspace{1em} African-American        & 16 & 0  \\
-		\hspace{1em} \gls{nz} M\=aori        & 0  & 10 \\
-		\hspace{1em} Pacific Islands         & 0  & 14 \\
-		\hspace{1em} Asian                   & 10 & 4  \\
+		\hspace{1em} Caucasian        & 77 & 176 & 71 \\
+		\hspace{1em} African-American & 16 & 29  & 0  \\
+		\hspace{1em} \gls{nz} M\=aori & 0  & 0   & 10 \\
+		\hspace{1em} Pacific Islands  & 0  & 0   & 14 \\
+		\hspace{1em} Asian            & 10 & 18  & 4  \\
+		\hspace{1em} Other            & 0  & 55  & 0  \\
 		\hline
 		\rule{0pt}{2.25ex}\gls{bmi} status  \\
-		\hspace{1em} Normal weight           & 36 & 28 \\
-		\hspace{1em} Overweight              & 29 & 28 \\
-		\hspace{1em} Obese                   & 38 & 43 \\
+		\hspace{1em} Normal weight           & 36 & NA\tnote{1} & 28 \\
+		\hspace{1em} Overweight              & 29 & NA & 28 \\
+		\hspace{1em} Obese                   & 38 & NA & 43 \\
 		\hline
 		\rule{0pt}{2.25ex}Menopause status  \\
-		\hspace{1em} Premenopause            & 49 & NA\tnote{1}\\
-		\hspace{1em} Perimenopause           & 9  & NA \\
-		\hspace{1em} Postmenopause           & 45 & NA \\
+		\hspace{1em} Premenopause            & 49 & NA & NA\\
+		\hspace{1em} Perimenopause           & 9  & NA & NA \\
+		\hspace{1em} Postmenopause           & 45 & NA & NA \\
 		\hline
 		\rule{0pt}{2.25ex}Tumour grade      \\
-		\hspace{1em} 1                       & 14 & 11 \\
-		\hspace{1em} 2                       & 35 & 39 \\
-		\hspace{1em} 3                       & 54 & 49 \\
+		\hspace{1em} 1                       & 14 & 13 & 11 \\
+		\hspace{1em} 2                       & 35 & 104 & 39 \\
+		\hspace{1em} 3                       & 54 & 150 & 49 \\
 		\hline
 		\rule{0pt}{2.25ex}\gls{er} status   \\
-		\hspace{1em} \gls{er}$^+$            & 58 & 72 \\
-		\hspace{1em} \gls{er}$^-$            & 42 & 27 \\
-		\hspace{1em} Unknown                 & 3  & 0  \\
+		\hspace{1em} \gls{er}$^+$            & 58 & 164 & 72 \\
+		\hspace{1em} \gls{er}$^-$            & 42 & 114 & 27 \\
+		\hspace{1em} Unknown                 & 3  & 0  & 0  \\
 		\hline
 		\rule{0pt}{2.25ex}\gls{pr} status   \\
-		\hspace{1em} \gls{pr}$^+$            & 48 & 62 \\
-		\hspace{1em} \gls{pr}$^-$            & 50 & 37 \\
-		\hspace{1em} Unknown                 & 5  & 0  \\
+		\hspace{1em} \gls{pr}$^+$            & 48 & 121 & 62 \\
+		\hspace{1em} \gls{pr}$^-$            & 50 & 157 & 37 \\
+		\hspace{1em} Unknown                 & 5  & 0  & 0  \\
 		\hline
 		\rule{0pt}{2.25ex}\gls{her2} status \\
-		\hspace{1em} \gls{her2}$^+$          & 12 & NA \\
-		\hspace{1em} \gls{her2}$^-$          & 54 & NA \\
-		\hspace{1em} Unknown                 & 37 & NA \\
+		\hspace{1em} \gls{her2}$^+$          & 12 & 59 & NA \\
+		\hspace{1em} \gls{her2}$^-$          & 54 & 219 & NA \\
+		\hspace{1em} Unknown                 & 37 & 0 & NA \\
 		\hline
 		\rule{0pt}{2.25ex}\gls{ln} status   \\
-		\hspace{1em} \gls{ln}$^+$            & 14 & 56\\
-		\hspace{1em} \gls{ln}$^-$            & 89 & 40\\
-		\hspace{1em} Unknown                 & 0  & 3 \\
+		\hspace{1em} \gls{ln}$^+$            & 14 & NA & 56\\
+		\hspace{1em} \gls{ln}$^-$            & 89 & NA & 40\\
+		\hspace{1em} Unknown                 & 0  & NA & 3 \\
 		\hline
@@
 
-The raw \gls{rnaseq} data from the \gls{icgc} database were formatted in a way that the count data of all the genes were listed for one sample, then the count data for all the genes for the next sample, and so on.
+The raw \gls{rnaseq} data from the \gls{icgc} database were formatted in such a way that the count data of all the genes were listed for one sample, then the count data for all the genes for the next sample, and so on.
 This data format was highly inconvenient for later analyses so the data were reformatted into a ``gene by sample'' matrix format using the \textit{dplyr} package in \gls{r}.
@@
 To associate each sample in the raw \gls{rnaseq} data from the \gls{icgc} database with the correct samples in the clinical data from the \gls{tcga} database, the \gls{icgc} IDs were re-coded so they matched the \gls{tcga} IDs.
-After the IDs were re-coded, all of the samples were checked to see if there were any duplicates in either the \gls{icgc} \gls{rnaseq} data or \gls{tcga} clinical data.
+After the IDs were re-coded, all of the samples were checked to see if there were any duplicates in either the \gls{icgc} \gls{rnaseq} data or the \gls{tcga} clinical data.
 Where there was a duplicate in the sample ID, a single sample with the highest mean expression value was chosen to represent that particular ID by using the \texttt{collapseRows} function in the \textit{WGCNA} package in \gls{r} \citep{Langfelder2008}.
@@
 This ensured that both clinical and \gls{rnaseq} data were available for all of the samples that were included in this project.
-See \cref{tab:samplesize} for a summary of the number of samples included in the analyses for each cancer type.
+See \cref{tab:samplesize} for a summary of the total number of samples included in the analyses for each cancer type.
 
@@
 			UCEC                   & 508 & 482   \\
+			\hline
+			\hline
+		\end{tabular}
+	\end{center}
+\end{table}
+
+\section{\Gls{bmi}}
+\label{sec:bmi}
+
+\subsection{\gls{bmi} Calculation}
+\label{subsec:bmicalc}
+
+Where missing, \cref{eq:bmicalc} was used to calculate the \gls{bmi} of the samples in all of the cancer data with height and weight data available (\cref{sec:data}).
+
+\begin{equation}
+	\label{eq:bmicalc}
+	\gls{bmi} = \frac{Weight (kg)}{Height^2(m^2)}\\
+\end{equation}
+
+\subsection{\gls{bmi} Classification}
+\label{subsec:bmiclassification}
+
+Samples were classified as normal weight (\gls{bmi} \textless{} 25), overweight (25 $\leq$ \gls{bmi} \textless{} 30) or obese (\gls{bmi} $\geq$ 30) based on the \gls{who} definition (\cref{tab:whobmiclass}).
+\cref{eq:bmicalc,tab:whobmiclass} were used to calculate and group the patients from the \gls{icgc} cancer data into appropriate \gls{bmi} classes (\cref{tab:icgc_bmi}).
+
+\begin{table}[htb]
+	\caption{\gls{who} defined \gls{bmi} classification}
+	\label{tab:whobmiclass}
+	\begin{center}
+		\begin{tabular}{lc}
+			Classification & \gls{bmi} Value\\
+			\hline
+			\rule{0pt}{2.25ex}Underweight & \textless{} 20.0\\
+			Normal weight/lean & 20.0$\sim$24.9\\
+			Overweight & 25.0$\sim$29.9\\
+			Obese & $\geq{}$ 30.0\\
 			\hline
@@
 
-\section{\Gls{bmi}}
-\label{sec:bmi}
-
-\subsection{\gls{bmi} Calculation}
-\label{subsec:bmicalc}
-
-Where missing, \cref{eq:bmicalc} was used to calculate the \gls{bmi} of the samples in all of the cancer data with height and weight data available (\cref{sec:data}).
-
-\begin{equation}
-	\label{eq:bmicalc}
-	\gls{bmi} = \frac{Weight (kg)}{Height^2(m^2)}\\
-\end{equation}
-
-\subsection{\gls{bmi} Classification}
-\label{subsec:bmiclassification}
-
-Samples were classified as normal weight (\gls{bmi} \textless{} 25), overweight (25 $\leq$ \gls{bmi} \textless{} 30) or obese (\gls{bmi} $\geq$ 30) based on the \gls{who} definition (\cref{tab:whobmiclass}).
-
-\begin{table}[htb]
-	\caption{\gls{who} defined \gls{bmi} classification}
-	\label{tab:whobmiclass}
-	\begin{center}
-		\begin{tabular}{lc}
-			Classification & \gls{bmi} Value\\
-			\hline
-			\rule{0pt}{2.25ex}Underweight & \textless{} 20.0\\
-			Normal weight/lean & 20.0$\sim$24.9\\
-			Overweight & 25.0$\sim$29.9\\
-			Obese & $\geq{}$ 30.0\\
-			\hline
-			\hline
-		\end{tabular}
-	\end{center}
-\end{table}
-
 \section{Data processing}
@@
 
-All of the data were normalised to remove any experimental bias, errors and noise, so only the true biological signals are considered in the analyses.
 Any experimental procedure is prone to errors due to differences in experimental conditions, machinery used to measure signals and technical procedures in different laboratories, just to name a few.
-In order to remove these experimental noises and focus on the true biological signals in the raw data.
+All of the data were normalised to remove such experimental bias, errors and noise, so only the true biological signals are considered in the analyses.
 
@@
 The reason why the \gls{mas} method was used as well as the \gls{rma} method was becuase the \gls{mas} normalisation method was used in the \citet{Gatza2010a} study.
-Though the normalisation methods were not specified in the \citet{Creighton2012} and \citet{Fuentes-Mattei2014} studies, to allow for data comparison between different datasets and accurate result validation of some of the studies, both \gls{rma} and \gls{mas} methods were used to normalise the microarray data.
+Though the normalisation methods were not specified in the \citet{Creighton2012} and \citet{Fuentes-Mattei2014} studies, to allow for data comparison between different datasets and accurate result validation of some of the studies, both \gls{rma} and \gls{mas} methods were considered for normalising the microarray data.
 
@@
 This means that \gls{rnaseq} data must be processed in a different manner to microarray data.
-However, by analysing \gls{rnaseq} data as a count data, it limits the range of statistical tools and types of analyses that can be done on the data, as many tools are designed for normally distributed data, like the light intensity measures from the microarray data \citep{Law2014}.
+However, by analysing \gls{rnaseq} data as a count data, it limits the range of statistical tools and types of analyses that can be done on the data, as many tools are designed for normally distributed data like the light intensity measures from the microarray data \citep{Law2014}.
 
@@
 All of the simulated results must be derived from a randomised data, and by comparing the observed results with the results that occured by chance, simulation analysis provides statistical evidence for the observed results.
-When a simulation was required, none of the values in the original data were modified but the values in the original data were assigned randomly to different samples; in other words, the samples were shuffled.
+When a simulation was required, none of the values in the original data were modified, but instead the values in the original data were assigned randomly to different samples; in other words, the samples were shuffled.
 This provided a random set of data in which a simulation could be conducted with.
@@
 
-Another important property of \gls{svd} is that the signal used to create the metagene in the first data set can be directly applied to other data set to obtain the metagene scores in that data set.
+Another important property of \gls{svd} is that the signal used to create the metagene in the first data set can be directly applied to other data sets to obtain the metagene scores for the samples in those data set.
 Rearrangement of \cref{eq:svd} for $V^T$ gives the following equation:
@@
 Clearly, by substituting the data matrix $X$ with another data matrix, the transformation matrix allows for the generation of the metagene for that data set.
-The reason why the transformation matrix is used instead of independently applying \gls{svd} to the other data set is because the signature, and therefore the metagene, from a single data set is dependent on the signal within its data set.
+The reason why the transformation matrix is used instead of independently applying \gls{svd} to the other data set is because the genetic signature, and therefore the metagene, from a single data set is dependent on the signal within its data set.
 Therefore, the metagenes created independently from different data sets using \gls{svd} do not have the same weighting as the metagene from the original data set, and consequently does not provide a fair comparison of the metagenes across different data sets.
@@
 Two different approaches were taken in order to rank the metagene scores of the samples.
-The first approach was to rank the metagene scores with the \texttt{rank} function from the \textit{base} package in \gls{r}, then divided the ranked scores with the total number of samples in the data set to obtain a ranked metagene scores between 0 and 1.
-In most analyses, this approach was used to assess the association of the metagene with the sample gene expressions and clinical variables.
+The first approach was to rank the metagene scores with the \texttt{rank} function from the \textit{base} package in \gls{r}, then divided the ranked scores with the total number of samples in the data set to obtain a ranked metagene scores between 0 and 1 (known as fractional ranking).
+In most analyses, fractional ranking was used to assess the association of the metagene with the sample gene expressions and clinical variables.
 
-An alternative ranking method was used in the Gatza \textit{et al.} data to compare the different ranking methods.
-In this method, the sample metagene scores were ranked based on the probit function.
-To do this, the sample metagene scores were scaled and centered using the \texttt{scale} function from the \textit{base} pakage in \gls{r}, then the \texttt{pnorm} function from the \textit{stats} package was used rank the metagene scores.
-Unless stated otherwise, all of the metagenes in this project were ranked based on the number of samples and not with the probit method.
+An alternative approach called probit transformation was used in the Gatza \textit{et al.} data.
+In this transformation, the sample metagene scores were transformed with the probit function.
+To do this, the sample metagene scores were scaled and centered using the \texttt{scale} function from the \textit{base} pakage in \gls{r}, then the \texttt{pnorm} function from the \textit{stats} package was used to transform the metagene scores so that the scores were between 0 and 1.
+Unless stated otherwise, all of the metagenes in this project were fractionally ranked.
 
@@
 	\includegraphics[width=0.45\linewidth,page=5]{methods/gtrmastdrank_corrected}
-	\caption[Example heatmaps showing the direction of the uncorrected and corrected \acrshort{er} pathway metagene with the gene expression of the \acrshort{er} pathway genetic signature in the \acrshort{rma}-normalised Gatza \textit{et al.} data.]{Heatmaps showing the direction of the unflipped (left) and flipped (right) \gls{er} pathway metagene with the gene expression of the \gls{er} pathway genetic signature in the \gls{rma}-normalised Gatza \textit{et al.} data.
+	\caption[Example heatmaps showing the direction of the uncorrected and corrected \acrshort{er} pathway metagene with the gene expression of the \acrshort{er} pathway genetic signature in the \acrshort{rma}-normalised Gatza \textit{et al.} data.]{Heatmaps showing the direction of the uncorrected (left) and corrected (right) \gls{er} pathway metagene with the gene expression of the \gls{er} pathway genetic signature in the \gls{rma}-normalised Gatza \textit{et al.} data.
 	Level of gene expression is represented in the top right histogram, where low and high gene expression are colour-coded with blue and red, respectively.
@@
 In this project, two types of metagenes were created: obesity associated and pathway associated metagenes.
-For obesity associated metagenes, the direction of the metagenes were checked with the obesity status and \gls{bmi} value of the samples from the data in which the metagene was derived from.
-When the data set did not have \gls{bmi}/\gls{bmi} status information for the samples, a set of common genes were taken from all of the obesity associated genetic signatures and their expressions compared with the metagene scores.
-Since all of the genetic signatures are associated with obesity, the genes that are common to all of the obesity associated genetic signatures will be related to the metagenes produced by these signatures, and therefore the expressions of these genes should be in line with the metagene scores.
+For obesity associated metagenes, the direction of the metagenes were checked with the patient obesity status and \gls{bmi} value from the data in which the metagene was derived from.
+When the data set did not have \gls{bmi} or \gls{bmi} status information for the patients (for example the Gatza \textit{et al.} data set), a set of genes that were common in all obesity associated genetic signatures were taken and the expression of these common genes were compared with the metagene scores.
+The genes that are common to all of the obesity associated genetic signatures will be related to the metagenes produced by these signatures, and therefore the expressions of these genes should be in line with the metagene scores.
 
 Checking the direction of the pathway metagenes from \citet{Gatza2010a} study is not as simple as the obesity associated metagenes, as the experimental design that identify which samples were treated to generate the pathway signature was not available.
-Due to this, it was not possible to validate whether the generated pathway metagene was in the ``right'' direction.
+Due to this, it was not possible to validate whether the generated pathway metagene was in the ``correct'' direction.
 As an alternative, the pathway metagenes were visually compared with the gene that represent the pathway using a heatmap, as the expression of this gene would likely be affected the most by the pathway signature.
@@
 
-In addition to this, the correlation of the pathway metagene with its repective gene was checked to further aid the decision on whether the metagene is going in the correct direction.
+In addition to this, the correlation value of the pathway metagene with its respective gene was checked to further aid the decision on whether the metagene is going in the correct direction.
 This was due to the fact that some of the metagenes were visually difficult to determine whether it was in the right direction or not.
@@
 Lastly, the clustering of the pathway metagenes with the other pathway metagenes were also considered.
-It was known from the \citet{Gatza2010a} study that some pathways clustered and grouped together with certain pathways, so the direction of some of the metagenes were altered in order to keep the grouping of the pathways matched to the results in their paper.
+It was known from the \citet{Gatza2010a} study that some pathways clustered and grouped together with certain pathways, so the direction of some of the metagenes were altered in order to keep the grouping of the pathways matched to the results in their paper (\cref{sec:result_from_gatza2010a_study}).
 
@@
 Furthermore, the model that was created based on the significant variables can be used to predict the obesity metagene, which can be used to compare the values with the original obesity metagene.
-If the predicted metagene was similar to the original metagene, then it provides an evidence that the pathway used to construct the model was associated with the obesity metagene.
+If the predicted metagene was similar to the original metagene, then it provides an evidence that the variables used to construct the model was associated with the obesity metagene.
 
+Since the \gls{nzbc} data was the only data set other than the Creighton \textit{et al.} data set that had the patient \gls{bmi} information, all of the linear models were constructed in the \gls{nzbc} data set.
 Eleven linear models were created in the \gls{nzbc} to predict the obesity associated metagene from the \citet{Creighton2012} study.
@@
 When a transformation matrix is applied to the data set in which it was derived from, the resulting metagene is identical to the one created from the direct application of \gls{svd}.
-Therefore, theoretically, if the metagene derived from a transformation matrix is similar to the metagene created from \gls{svd}, then it shows that the genetic signature is consistent and reliable across different data sets.
+Therefore, if the metagene derived from a transformation matrix is similar to the metagene created from \gls{svd}, then it shows that the genetic signature is consistent and reliable across different data sets.
 The above mentioned pathways had high correlation value between the \gls{svd}-derived and transformation matrix-derived pathway metagenes, and so it was used in the construction of the linear model.
@@
 The last four linear models were constructed with the different combinations of the sample \gls{bmi}, \gls{bmi} status and the \gls{pr} pathway metagene.
-The reason the \gls{pr} pathway metagene was used separately was because in the initial analysis, the \gls{pr} pathway metagene was the only pathway associated genetic signature that came up as significant in the other linear models (\cref{sec:prediction_of_obesity_associated_metagene_with_pathway_associate_metagene}).
+The reason that the \gls{pr} pathway metagene was used separately was because in the initial analysis, the \gls{pr} pathway metagene was the only pathway associated genetic signature that came up as significant in the other linear models (\cref{sec:prediction_of_obesity_associated_metagene_with_pathway_associate_metagene}).
 All of the created linear models are summarised in \cref{tab:lm_model_method}.
@@
 
-In addition to these eleven linear models, a linear model was created with the sample \gls{bmi}, \gls{bmi} status and all of the pathway associated metagenes using a stepwise approach.
-Step wise approach allows the construction of a linear model that best fit the data, given a set of variables.
-There are three approaches in creating a linear model in a step wise method: top-down, bottom-up, and both.
+In addition to these eleven linear models, a linear model was created with the sample \gls{bmi}, \gls{bmi} status and all of the pathway associated metagenes (Akt, \gls{bcat}, E2F1, \gls{egfr}, \gls{er}, \gls{her2}, \gls{ifna}, \gls{ifny}, Myc, p53, p63, \gls{pi3k}, \gls{pr}, Ras, \gls{stat3}, Src, \gls{tgfb} and \gls{tnfa}) using a stepwise approach.
+Stepwise approach allows the construction of a linear model that best fit the data, given a set of variables.
+There are three approaches in creating a linear model in a stepwise method: top-down, bottom-up, and both.
 
@@
 Again, this will result in a model with only the variables that are significant in predicting the response variable.
-Alternatively, both methods in combination can be used, where the model begins with all of the variables and at each step, a variable can be added to or removed from the linear model.
-In this project, the combination of top-down and bottom-up approaches was used to create the step wise linear model.
+Alternatively, both methods in combination can be used, where the model begins with all or none of the variables, and at each step a variable can be added to or removed from the linear model.
+In this project, the combination of top-down and bottom-up approaches (beginning with no variables in the model) was used to create the stepwise linear model.
 
@@
 Where appropriate, p-values, \gls{anova} p-values and  $R^2$-values were calculated to assess the relationship between the predicted and the original metagene scores.
-Scatter plots and box plots were also created to visually compare the two metagene scores.
+Scatter plots and box plots were also created to visually compare the two metagene scores (\cref{sub:box_and_scatter_plots}).
 

### 2017-03-23 thesis/misc/glossary.tex
tmp commit
Finished editing methods
Finished editing results1

@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-% TODO: make sure you get a proper definitions for all of these glossary terms
 
@@
 	description={Statistical programming language}
+}
+
+\newglossaryentry{immunoediting}{
+	name=Immunoediting,
+	text=immunoediting,
+	description={Process where the immune system successfully eliminates the highly immunogenic cancer cells during surveillance, but fails to eliminate the weakly immunogenic cells}
 }
@@
 
+\newglossaryentry{carcinogen}{
+	name=Carcinogen,
+	text=carcinogen,
+	plural=carcinogens,
+	description={A substance capable of causing cancer}
+}
+
+\newglossaryentry{passenger}{
+	name=Passanger mutation,
+	text=passanger,
+	description={Mutations that have neither a selective advantage nor contribute to tumour progression}
+}
+
+\newglossaryentry{driver}{
+	name=Driver mutation,
+	text=driver,
+	description={Mutations that helps tumorigenesis and its subsequent progression}
+}
+
+\newglossaryentry{somatic}{
+	name=Somatic mutation,
+	text=somatic mutation,
+	plural=somatic mutations,
+	description={Mutations that are acquried by the cell over the course of the lifetime of an individual}
+}
+
+\newglossaryentry{germline}{
+	name=Germline mutation,
+	text=germline mutation,
+	plural=germline mutations,
+	description={Mutations that are passed on from the parents}
+}
+
+\newglossaryentry{stability}{
+	name=Stability gene,
+	text=stability gene,
+	plural=stability genes,
+	description={Genes that are involved in \acrshort{dna} repair mechanisms}
+}
+
+\newglossaryentry{allele}{
+	name=Allele,
+	text=allele,
+	plural=alleles,
+	description={Variant form of a given gene}
+}
+
+\newglossaryentry{suppressor}{
+	name=Tumour suppressor gene,
+	text=tumour suppressor,
+	description={Genes that prevent or ``decelerate'' the process of tumorigenesis and/or tumour progression}
+}
+
+\newglossaryentry{oncogene}{
+	name=Oncogene,
+	text=oncogene,
+	plural=oncogenes,
+	description={Genes that promote or ``accelerate'' the process of tumorigenesis and/or tumour progression}
+}
+
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 
+\newacronym[longplural={Disability-Adjusted Life Years}]{daly}{DALY}{Disability-Adjusted Life Year}
+\newacronym[longplural={genome wide association studies},\glsshortpluralkey={GWAS}]{gwas}{GWAS}{genome wide association study}
+\newacronym{acth}{ACTH}{adrenocorticotropic hormone}
+\newacronym{agrp}{AGRP}{agouti-related protein}
+\newacronym{aic}{AIC}{Akaike Information Criterion}
+\newacronym{aid}{AID}{activation-induced cytidine deaminase}
+\newacronym{anova}{ANOVA}{analysis of variance}
+\newacronym{apobec}{APOBEC}{apolipoprotein B mRNA editing enzyme, catalytic polypeptide-like}
+\newacronym{atp}{ATP}{adenosine triphosphate}
+\newacronym{bcat}{BCAT}{$\beta$-catenin}
+\newacronym{bcl2}{Bcl-2}{B-cell lymphoma 2}
+\newacronym{ber}{BER}{base-excision repair}
+\newacronym{bh3}{BH3}{Bcl-2 homology 3}
 \newacronym{blca}{BLCA}{bladder urothelial carcinoma}
+\newacronym{bmi}{BMI}{body mass index}
+\newacronym{camera}{CAMERA}{correlation adjusted mean rank gene set test}
+\newacronym{cart}{CART}{cocaine and amphetamine related transcript}
+\newacronym{cdna}{cDNA}{complimentary DNA}
 \newacronym{cesc}{CESC}{cervical squamous cell carcinoma and endocervical adenocarcinoma}
 \newacronym{coad}{COAD}{colon adenocarcinoma}
+\newacronym{cpm}{cpm}{counts per million}
+\newacronym{cyp}{CYP}{cytochrome P450}
+\newacronym{deg}{DEG}{differentially expressed gene}
+\newacronym{dna}{DNA}{deoxyribonucleic acid}
+\newacronym{egfr}{EGFR}{epidermal growth factor receptor}
+\newacronym{emt}{EMT}{endothelial-mesenchymal transition}
+\newacronym{erk}{ERK}{extracellular signal regulated kinase}
+\newacronym{er}{ER}{estrogen receptor}
+\newacronym{fcs}{FCS}{functional class scoring}
+\newacronym{fc}{FC}{fold change}
+\newacronym{fdr}{FDR}{false discovery rate}
+\newacronym{fto}{FTO}{fat mass and obesity associated}
+\newacronym{fto}{FTO}{fat mass and obesity associated}
+\newacronym{fwer}{FWER}{family-wise error rate}
+\newacronym{g1}{G1}{Gap 1}
+\newacronym{gaba}{GABA}{$\gamma$-aminobutyric acid}
+\newacronym{geo}{GEO}{Gene Expression Omnibus}
+\newacronym{ghsr}{GHSR}{growth hormone secretagogue receptor}
+\newacronym{go}{GO}{Gene Ontology}
+\newacronym{gpcr}{GPCR}{G-protein coupled receptor}
+\newacronym{gsea}{GSEA}{Gene Set Enrichment Analysis}
+\newacronym{her2}{HER2}{human epidermal growth factor receptor 2}
+\newacronym{icgc}{ICGC}{International Cancer Genome Consortium}
+\newacronym{ifna}{IFN$\alpha$}{interferon-$\alpha$}
+\newacronym{ifny}{IFN$\gamma$}{interferon-$\gamma$}
+\newacronym{igfbp}{IGFBP}{insulin-like growth factor binding protein}
+\newacronym{igf}{IGF}{insulin-like growth factor}
+\newacronym{il10}{IL-10}{interleukin-10}
+\newacronym{il6}{IL-6}{interleukin-6}
+\newacronym{irs}{IRS}{insulin receptor substrate}
+\newacronym{jak}{JAK}{janus kinase}
+\newacronym{kegg}{KEGG}{Kyoto Encyclopedia of Genes and Genomes}
 \newacronym{kirp}{KIRP}{kidney renal clear cell carcinoma}
 \newacronym{lihc}{LIHC}{liver hepatocellular carcinoma}
-\newacronym{read}{READ}{rectum adenocarcinoma}
-\newacronym{skcm}{SKCM}{skin cutaneous melanoma}
-\newacronym{ucec}{UCEC}{uterine corpus endometrial carcinoma}
-\newacronym{geo}{GEO}{Gene Expression Omnibus}
-\newacronym{tcga}{TCGA}{The Cancer Genome Atlas}
-\newacronym{icgc}{ICGC}{International Cancer Genome Consortium}
-\newacronym{bmi}{BMI}{body mass index}
-\newacronym{who}{WHO}{World Health Organization}
 \newacronym{ln}{LN}{lymph node}
-\newacronym{pr}{PR}{progesterone receptor}
-\newacronym{er}{ER}{estrogen receptor}
-\newacronym{her2}{HER2}{human epidermal growth factor receptor 2}
-\newacronym{svd}{SVD}{singular value decomposition}
-\newacronym{rma}{RMA}{robust multi array}
 \newacronym{mas}{MAS5}{microarray analysis suite 5.0}
-\newacronym{fdr}{FDR}{false discovery rate}
-\newacronym{fwer}{FWER}{family-wise error rate}
-\newacronym{go}{GO}{Gene Ontology}
-\newacronym{kegg}{KEGG}{Kyoto Encyclopedia of Genes and Genomes}
-\newacronym{usa}{USA}{United States of America}
-\newacronym{npy}{NPY}{neuropeptide Y}
-\newacronym{agrp}{AGRP}{agouti-related protein}
-\newacronym{pomc}{POMC}{pro-opiomelanocortin}
-\newacronym{cart}{CART}{cocaine and amphetamine related transcript}
 \newacronym{mc1r}{MC1R}{melanocortin 1 receptor}
 \newacronym{mc4r}{MC4R}{melanocortin 4 receptor}
+\newacronym{met}{MET}{mesenchymal-endothelial transition}
+\newacronym{mmr}{MMR}{mismatch repair}
+\newacronym{mm}{MM}{mismatch}
+\newacronym{mrna}{mRNA}{messenger RNA}
 \newacronym{msh}{MSH}{melanocyte-stimulating hormone}
+\newacronym{mtor}{mTOR}{mammalian target of rapamycin}
+\newacronym{m}{$\mu$}{mean}
+\newacronym{ner}{NER}{nucleotide-excision repair}
+\newacronym{nfkb}{NF-$\kappa$B}{nuclear factor $\kappa$B}
+\newacronym{ngs}{NGS}{next generation sequencing}
+\newacronym{nnk}{NNK}{4-(methylnitrosamino)-1-(3-pyridyl)-1-butanone}
+\newacronym{npy}{NPY}{neuropeptide Y}
 \newacronym{nyr1}{Y1R}{neuropeptide Y1 receptor}
-\newacronym{acth}{ACTH}{adrenocorticotropic hormone}
-\newacronym{ghsr}{GHSR}{growth hormone secretagogue receptor}
-\newacronym{gaba}{GABA}{$\gamma$-aminobutyric acid}
-\newacronym{irs}{IRS}{insulin receptor substrate}
-\newacronym{anova}{ANOVA}{analysis of variance}
-\newacronym{pi3k}{PI3K}{phosphatidylinositol 3-kinase}
-\newacronym{jak}{JAK}{janus kinase}
-\newacronym{pc1}{PC1}{prohormone convertase 1}
-\newacronym{fto}{FTO}{fat mass and obesity associated}
-\newacronym{fto}{FTO}{fat mass and obesity associated}
-\newacronym{snp}{SNP}{single nucleotide polymorphism}
-\newacronym{rb}{RB}{retinoblastoma-associated}
-\newacronym{opa}{OPA}{obesity policy action}
-\newacronym{g1}{G1}{Gap 1}
-\newacronym{bcl2}{Bcl-2}{B-cell lymphoma 2}
-\newacronym{bh3}{BH3}{Bcl-2 homology 3}
-\newacronym{dna}{DNA}{deoxyribonucleic acid}
-\newacronym{vegf}{VEGF}{vascular endothelial growth factor}
-\newacronym{vegfa}{VEGF-A}{vascular endothelial growth factor-A}
-\newacronym{tsp1}{TSP-1}{thrombospondin-1}
-\newacronym{emt}{EMT}{endothelial-mesenchymal transition}
-\newacronym{met}{MET}{mesenchymal-endothelial transition}
-\newacronym{aid}{AID}{activation-induced cytidine deaminase}
-\newacronym{apobec}{APOBEC}{apolipoprotein B mRNA editing enzyme, catalytic polypeptide-like}
-\newacronym{nfkb}{NF-$\kappa$B}{nuclear factor $\kappa$B}
-\newacronym{stat3}{STAT3}{signal transducer and activator of transcription 3}
-\newacronym{il6}{IL-6}{interleukin-6}
-\newacronym{il10}{IL-10}{interleukin-10}
-\newacronym{atp}{ATP}{adenosine triphosphate}
-\newacronym{pep}{PEP}{phosphoenol pyruvate}
-\newacronym{mmr}{MMR}{mismatch repair}
-\newacronym{mtor}{mTOR}{mammalian target of rapamycin}
-\newacronym{ner}{NER}{nucleotide-excision repair}
-\newacronym{ber}{BER}{base-excision repair}
-\newacronym{uv}{UV}{ultraviolet}
-\newacronym{deg}{DEG}{differentially expressed gene}
-\newacronym{pah}{PAH}{polycyclic aromatic hydrocarbon}
-\newacronym{nnk}{NNK}{4-(methylnitrosamino)-1-(3-pyridyl)-1-butanone}
-\newacronym{ros}{ROS}{reactive oxygen species}
-\newacronym{cyp}{CYP}{cytochrome P450}
-\newacronym{igf}{IGF}{insulin-like growth factor}
-\newacronym{igfbp}{IGFBP}{insulin-like growth factor binding protein}
-\newacronym{gpcr}{GPCR}{G-protein coupled receptor}
-\newacronym{pm}{PM}{perfect match}
-\newacronym{mm}{MM}{mismatch}
-\newacronym{tlr}{TLR}{Toll-like receptor}
-\newacronym{cpm}{cpm}{counts per million}
-\newacronym{m}{$\mu$}{mean}
-\newacronym{sd}{$\sigma$}{standard deviation}
-\newacronym{prr}{PRR}{pattern recognition receptor}
-\newacronym{ssdna}{ssDNA}{single-stranded DNA}
-\newacronym{camera}{CAMERA}{correlation adjusted mean rank gene set test}
-\newacronym{ora}{ORA}{over representation analysis}
-\newacronym{fcs}{FCS}{functional class scoring}
-\newacronym{wmw}{WMW}{Wilcoxon-Mann-Whitney}
-\newacronym{vif}{VIF}{variance inflation factor}
-\newacronym{erk}{ERK}{extracellular signal regulated kinase}
-\newacronym{cdna}{cDNA}{complimentary DNA}
-\newacronym{mrna}{mRNA}{messenger RNA}
-\newacronym{ngs}{NGS}{next generation sequencing}
-\newacronym{fc}{FC}{fold change}
-\newacronym{ifny}{IFN$\gamma$}{interferon-$\gamma$}
-\newacronym{ifna}{IFN$\alpha$}{interferon-$\alpha$}
-\newacronym{tgfb}{TGF$\beta$}{transforming growth factor-$\beta$}
-\newacronym{tnfa}{TNF$\alpha$}{tumour necrosis factor-$\alpha$}
-\newacronym{bcat}{BCAT}{$\beta$-catenin}
-\newacronym{egfr}{EGFR}{epidermal growth factor receptor}
-\newacronym[longplural={genome wide association studies},\glsshortpluralkey={GWAS}]{gwas}{GWAS}{genome wide association study}
-\newacronym{stat}{STAT}{signal transducer and activator of transcription}
-\newacronym{pyy}{PYY$_{3-36}$}{peptide YY$_{3-36}$}
 \newacronym{nzbc}{NZBC}{New Zealand Breast Cancer}
 \newacronym{nz}{NZ}{New Zealand}
-\newacronym[longplural={Disability-Adjusted Life Years}]{daly}{DALY}{Disability-Adjusted Life Year}
+\newacronym{opa}{OPA}{obesity policy action}
+\newacronym{ora}{ORA}{over representation analysis}
+\newacronym{pah}{PAH}{polycyclic aromatic hydrocarbon}
+\newacronym{pc1}{PC1}{prohormone convertase 1}
+\newacronym{pep}{PEP}{phosphoenol pyruvate}
+\newacronym{pi3k}{PI3K}{phosphatidylinositol 3-kinase}
+\newacronym{pm}{PM}{perfect match}
+\newacronym{pomc}{POMC}{pro-opiomelanocortin}
+\newacronym{prr}{PRR}{pattern recognition receptor}
+\newacronym{pr}{PR}{progesterone receptor}
+\newacronym{pyy}{PYY$_{3-36}$}{peptide YY$_{3-36}$}
+\newacronym{rb}{RB}{retinoblastoma-associated}
+\newacronym{read}{READ}{rectum adenocarcinoma}
+\newacronym{rma}{RMA}{robust multi array}
+\newacronym{rna}{RNA}{ribonucleic acid}
+\newacronym{ros}{ROS}{reactive oxygen species}
+\newacronym{sd}{$\sigma$}{standard deviation}
+\newacronym{skcm}{SKCM}{skin cutaneous melanoma}
+\newacronym{snp}{SNP}{single nucleotide polymorphism}
+\newacronym{ssdna}{ssDNA}{single-stranded DNA}
+\newacronym{stat3}{STAT3}{signal transducer and activator of transcription 3}
+\newacronym{stat}{STAT}{signal transducer and activator of transcription}
+\newacronym{svd}{SVD}{singular value decomposition}
+\newacronym{tcga}{TCGA}{The Cancer Genome Atlas}
+\newacronym{tgfb}{TGF$\beta$}{transforming growth factor-$\beta$}
+\newacronym{tlr}{TLR}{Toll-like receptor}
+\newacronym{tnfa}{TNF$\alpha$}{tumour necrosis factor-$\alpha$}
+\newacronym{tsp1}{TSP-1}{thrombospondin-1}
+\newacronym{ucec}{UCEC}{uterine corpus endometrial carcinoma}
+\newacronym{usa}{USA}{United States of America}
+\newacronym{uv}{UV}{ultraviolet}
+\newacronym{vegfa}{VEGF-A}{vascular endothelial growth factor-A}
+\newacronym{vegf}{VEGF}{vascular endothelial growth factor}
+\newacronym{vif}{VIF}{variance inflation factor}
+\newacronym{who}{WHO}{World Health Organization}
+\newacronym{wmw}{WMW}{Wilcoxon-Mann-Whitney}

### 2017-03-23 thesis/methods/methods.tex
tmp commit

@@
 Any cancer type with no height and/or weight data for the samples was excluded from the project, as no \gls{bmi} information can be obtained without these data.
-
 Out of these 33 cancer types with clinical data, 14 cancer types had both height and weight data.
@@
 In each of the eight cancer types, the \gls{tcga} sample IDs in the clinical data were cross-checked with the sample IDs in the \gls{rnaseq} data, and vice versa.
-Any samples that did not have either the clinical information or the \gls{rnaseq} data were removed from the analyses.
-This ensured that both clinical and \gls{rnaseq} data were available for all of the samples that were included in this project.
+Any samples that did not have either the patient \gls{bmi} information or the \gls{rnaseq} data were removed from the analyses.
+This ensured that both \gls{bmi} and \gls{rnaseq} data were available for all of the samples that were included in this project.
 See \cref{tab:samplesize} for a summary of the total number of samples included in the analyses for each cancer type.
@@
 For gene expression analyses (\cref{sec:gene_expression_analysis}) and pathway enrichment analyses (\cref{sec:pathway_enrichment_analysis}), voom normalisation from the \textit{limma} package in \gls{r} was used to normalise the data \citep{Ritchie2015}.
-For the purposes of data visualisation (\cref{sec:plot_creation}) or application of the metagene transformation matrices (\cref{sub:svd}) on the \gls{rnaseq} data, the raw data had 1 added (to prevent logging of 0) and then logged to the base of 10.
+For the purposes of data visualisation (\cref{sec:plot_creation}) or the application of the metagene transformation matrices (\cref{sub:svd}) on the \gls{rnaseq} data, the raw data had 1 added (to prevent logging of 0) and then logged to the base of 10.
 
@@
 A linear model was created from the Creighton \textit{et al.} data with age, ethnicity, menopause status, tumour grade, hormone (\gls{er}, \gls{pr}, and \gls{her2}) statuses  and \gls{ln} status included in the model.
-In addition to this model, another linear model was constructed with the Creighton \textit{et al.} data with only the Caucasian patients included in the data (same clinical variables, except, ethnicity were used in the model); this was to completely remove the effect of ethnicity.
-Once a linear model was fitted to the Creighton \textit{et al.} data, the remaining data, or the ``residual data'', represented the data that had been corrected for all of the unwanted variables.
-This residual data was used in the analyses that required focus on certain variables without the other variables affecting the result.
+In addition to this model, another linear model was constructed with the Creighton \textit{et al.} data with only the Caucasian patients included in the data to completely remove the effect of ethnicity (same clinical variables were controlled for, except ethnicity).
+Once a linear model was fitted to the Creighton \textit{et al.} data, the remaining data, or the ``residual data'', represented the data that had been corrected for the effects of the unwanted variables, and was used in the analyses that required focus on certain variables without the other variables affecting the result.
 
@@
 
-Since the data used in \citet{Gatza2010a} study was a combination of multiple microarray data from various studies, the batch effect had to be corrected before any analysis was carried out.
+The data used in \citet{Gatza2010a} study was a combination of multiple microarray data from various studies, so the batch effect had to be corrected before any analysis was carried out.
 Each microarray dataset was normalised with either \gls{rma} or \Gls{mas} separately then combined together into a single dataset, and the batch effect was corrected with the \texttt{ComBat} function (an implementation of the batch correcting method by \citet{Johnson2007}) from the \textit{sva} package in \gls{r} \citep{Leek2012}.
@@
 Two combined data sets were created with the \gls{icgc} data.
-The first combined data set was created by voom normalising each cancer data set separately, combine all of the normalised cancer data sets into one, then applied batch correction to the data set (\cref{sub:batch_correction}).
-The second combined data set was created by first combining the cancer data sets into one, then the combined data was voom normalised, and then the batch effect was corrected to obtain the final data set.
+The first data set was created by voom normalising each cancer data set separately, combine all of the normalised cancer data sets into one, then applied batch correction to the data set (\cref{sub:batch_correction}).
+The second data set was created by first combining the cancer data sets into one, then the combined data was voom normalised, and then the batch effect was corrected to obtain the final data set.
 
@@
 From \cref{eq:fdr}, the adjusted p-value from \gls{fdr} control is the product of the p-value with the number of hypotheses, divided by its rank.
-
 Since \gls{fdr} method controls \gls{type1} in a less conservative manner than the \gls{fwer} methods, the \gls{fdr} method was used to identify the \glspl{deg} for gene expression analyses.
@@
 Therefore, given a list of \glspl{deg}, it is important for any researcher to undertake pathway enrichment analysis in order to obtain useful insights into the underlying biological mechanisms the \glspl{deg} may be involved in.
+Both \acrfull{ora} and \acrfull{fcs} can be used to identify whether certain pathways are enriched, given a list of \glspl{deg}.
 
@@
 \label{sub:ora_and_fcs}
-
-Though the approaches differ between the two, both \acrfull{ora} and \acrfull{fcs} are used to identify whether certain pathways are enriched, given a list of \glspl{deg}.
 
@@
 
-An alternative approach called probit transformation was used in the Gatza \textit{et al.} data.
+An alternative approach called probit transformation was used by \citet{Gatza2010a}.
 In this transformation, the sample metagene scores were transformed with the probit function.
@@
 	Each row of the heatmap represents a gene from the \gls{er} genetic signature, and each column of the heatmap represents a sample from the Gatza \textit{et al.} data.
-	The \textit{ESR1} gene expressions of the samples are shown as a separate row at top of the heatmap with the sample \gls{er} metagene scores below the \textit{ESR1} gene expression, and the tree diagram of the heirarchical clustering of the genes is shown on the left of the heatmap.
+	The \textit{ESR1} gene expressions of the samples are shown as a separate row at the top of the heatmap with the sample \gls{er} metagene scores shown below the \textit{ESR1} gene expression, and the tree diagram of the heirarchical clustering of the genes is shown on the left of the heatmap.
 	}

### 2017-03-23 thesis/discussion/discussion.tex
tmp commit

@@
 In the step wise linear models, many variables were identified to be significantly associated with the obesity metagenes.
-This suggested that the pathway associated genetic signatures that were signicant in these models may be biologically related to the obesity phenotype, or perhaps the obesity associated genetic signatures were detecting the signals from these pathways rather than the signals from the phenotype.
+This suggested that the pathway associated genetic signatures that were signicant in these models may be biologically related to the obesity phenotype, or perhaps the obesity associated genetic signatures were detecting the signals from these pathways rather than the signals from the obesity phenotype.
 In fact, many of the variables included in the linear models were pathway metagenes that negatively correlated with the obesity metagenes in \cref{fig:gatza_allmeta} (\cref{sub:comparison_of_the_obesity_and_pathway_associated_signatures}).
-Therefore, it may be possible that these pathways (Akt, \gls{er}, \gls{egfr}, \gls{pr}, p53, Src and \gls{tgfb}) are related to the underlying biological mechanism that drives breast tumour development, and potentially associated with the patient \gls{bmi} as well.
+Since all of the obesity metagenes failed to show significant association with patient \gls{bmi}, it may be possible that these pathways (Akt, \gls{er}, \gls{egfr}, \gls{pr}, p53, Src and \gls{tgfb}) are related to the underlying biological mechanism that drives the breast tumour development.
 
-With that said, even if these pathway signatures were related to the obesity associated genetic signatures, the fact that these results depended on the pathway associated genetic signatures from the \citet{Gatza2010a} study makes the results questionable.
+With that said, even if these pathway signatures were related to the underlying biology of breast cancer, the fact that these results depended on the pathway associated genetic signatures from the \citet{Gatza2010a} study makes the results questionable.
 This was due to the lack of consistency of the metagenes across different cancer data sets, where more than half of the pathway metagenes showed a correlation of less than 0.8 in at least one data set (\cref{sec:pathway_associated_metagenes_and_obesity_associated_metagenes}).

### 2017-03-23 thesis/thesis.tex
Almost finished editing discussion - tmp commit
Finished editing the draft thesis - clean things up
tmp commit

@@
 \usepackage[T1]{fontenc}                            % package for specifying output font encoding
-\usepackage[acronym]{glossaries}                    % package for making glossaries
+\usepackage[acronym,style=super]{glossaries}                    % package for making glossaries
 \usepackage[flushleft]{threeparttablex}             % package to add notes at the bottom of the table
@@
 % \includeonly{{./intro/intro}}
-\includeonly{{./methods/methods},{./results/results2},{./misc/appendix2}}
+% \includeonly{{./methods/methods}}
+\includeonly{{./results/results2},{./discussion/discussion},{./misc/appendix},{./misc/appendix2}}
 % \includeonly{{./results/results}}
@@
 
+% Define some alias for very long citations:
 \defcitealias{TCGA}{NCI and NHGRI, 2016}
@@
 
-\printglossary[style=altlist]
+\printglossary[style=super]
 
@@
 % Print out the bibliography:
-\bibliography{./MSc,./msc2,./msc3}
+\bibliography{./MSc,./MSc2,./MSc3}
 

### 2017-03-25 thesis/discussion/discussion.tex
finished editing up to the discussion

@@
 This meant that the variables that were significant in the linear models played a role in the obesity associated genetic signatures.
-There was no single pathway signature that was common to all of the linear models, but Akt, \gls{egfr} and \gls{pr} pathway metagenes were included in many of the models, which suggested that these pathway signatures may have an important role in the obesity associated genetic signatures (\cref{sub:stepwise_linear_model_prediction_in_nzbc_and_cr_data_sets}).
+There was no single pathway signature that was common to all of the linear models, but Akt, \gls{egfr} and Src pathway metagenes were included in many of the models, which suggested that these pathway signatures may have an important role in the obesity associated genetic signatures (\cref{sub:stepwise_linear_model_prediction_in_nzbc_and_cr_data_sets}).
 
@@
 The results from \cref{cha:obesity_associated_genetic_signature_and_pathway_signatures} showed that the obesity associated genetic signatures generated from the Creighton \textit{et al.} data set were different from those generated from the Fuentes-Mattei \textit{et al.} data set.
-None of the pathway associated genetic signatures was similar to any of the obesity associated genetic signatures in the heatmap.
-% re-write sentence below:
-The results from the linear model prediction suggested a possible association of the obesity associated genetic signatures with the Akt, \gls{egfr} and \gls{pr} pathways.
+None of the pathway associated genetic signatures showed positive correlation with any of the obesity associated genetic signatures in the heatmap.
+However, strong negative correlation with the Akt, \gls{er}, \gls{egfr}, \gls{pr}, p53, Src and \gls{tgfb} pathways, suggesting that the possibility of the obesity associated genetic signatures were picking up the signals from these pathways, rather than the signals from the obesity phenotype of the patients.
+The results from the linear model prediction suggested that the Akt, \gls{egfr} and Src pathways may be important in breast tumour development and/or progression.
 
 To summarise, there were no common genetic signatures that significantly associated with the obesity status of the patients.
-However, the Akt, \gls{egfr} and \gls{pr} pathways may have a role in promoting the tumour progression in patients that are obese.
+However, the Akt, \gls{egfr} and Src pathways may have a role in promoting the tumour progression in patients that are obese.
 Further investigations with good quality data sets and larger patient cohorts are required to clarify the biological relationship between obesity and cancer.

### 2017-03-25 thesis/misc/appendix.tex
finished editing up to the discussion

@@
 		\label{fig:crob_meta_direction}
+	\end{figure}
+
+	\section{Remainder of the results of the other CR obesity metagenes in the CR data set}
+	\label{sec:rest_of_the_cr_ob_meta_heatmap_results_cr}
+
+	All of the other obesity metagenes identified from the CR data set were able to capture the overall gene expressions of the samples in the CR data set (\cref{fig:appendix/cr_ob_meta_heatmap_cr}).
+	Furthermore, all of the CR obesity metagenes were significantly associated with the patient \gls{bmi}/\gls{bmi} status in the CR data set (\cref{fig:appendix/cr_ob_meta_box_scatter_cr}).
+
+	\begin{figure}[htp!]
+		\centering
+		\includegraphics[width=0.32\linewidth,page=3]{appendix/cr_deg_meta_vs_clin}\\
+		\includegraphics[width=0.32\linewidth,page=8]{appendix/cr_deg_meta_vs_clin}
+		\includegraphics[width=0.32\linewidth,page=13]{appendix/cr_deg_meta_vs_clin}
+		\includegraphics[width=0.32\linewidth,page=18]{appendix/cr_deg_meta_vs_clin}\\
+		\includegraphics[width=0.32\linewidth,page=23]{appendix/cr_deg_meta_vs_clin}
+		\includegraphics[width=0.32\linewidth,page=28]{appendix/cr_deg_meta_vs_clin}
+		\includegraphics[width=0.32\linewidth,page=33]{appendix/cr_deg_meta_vs_clin}\\
+		\caption[Association of the other CR obesity metagenes with the sample gene expressions in the \gls{nzbc} data]{Heatmaps showing the association of the other obesity metagenes identified from the CR data with the sample gene expressions from the \gls{nzbc} data set.
+		Scales are as described in previous figures.}
+		\label{fig:appendix/cr_ob_meta_heatmap_cris}
+	\end{figure}
+
+	\begin{figure}[htpb]
+		\centering
+		\includegraphics[page=4,width=0.42\linewidth]{appendix/cr_deg_meta_vs_clin}
+		\hfill
+		\includegraphics[page=5,width=0.42\linewidth]{appendix/cr_deg_meta_vs_clin}\\
+		\caption[Association of the other CR obesity metagenes with the sample \gls{bmi}/\gls{bmi} status in the \gls{nzbc} data]{Box plots and scatter plots showing the association of the other CR obesity metagenes with the sample \gls{bmi} status  and \gls{bmi} from the \gls{nzbc} data set, respectively.
+		P-values and $R^2$-value are as described in previous figures.}
+		\label{fig:appendix/cr_ob_meta_box_scatter_cris}
+	\end{figure}
+
+	\begin{figure}[htpb]
+		\ContinuedFloat
+		\captionsetup{list=off,format=cont}
+		\centering
+		\includegraphics[page=9,width=0.45\linewidth]{appendix/cr_deg_meta_vs_clin}
+		\hfill
+		\includegraphics[page=10,width=0.45\linewidth]{appendix/cr_deg_meta_vs_clin}\\
+		\includegraphics[page=14,width=0.45\linewidth]{appendix/cr_deg_meta_vs_clin}
+		\hfill
+		\includegraphics[page=15,width=0.45\linewidth]{appendix/cr_deg_meta_vs_clin}\\
+		\includegraphics[page=19,width=0.45\linewidth]{appendix/cr_deg_meta_vs_clin}
+		\hfill
+		\includegraphics[page=20,width=0.45\linewidth]{appendix/cr_deg_meta_vs_clin}\\
+		\caption{Name}
+	\end{figure}
+
+	\begin{figure}[htpb]
+		\ContinuedFloat
+		\captionsetup{list=off,format=cont}
+		\centering
+		\includegraphics[page=24,width=0.45\linewidth]{appendix/cr_deg_meta_vs_clin}
+		\hfill
+		\includegraphics[page=25,width=0.45\linewidth]{appendix/cr_deg_meta_vs_clin}\\
+		\includegraphics[page=29,width=0.45\linewidth]{appendix/cr_deg_meta_vs_clin}
+		\hfill
+		\includegraphics[page=30,width=0.45\linewidth]{appendix/cr_deg_meta_vs_clin}\\
+		\includegraphics[page=34,width=0.45\linewidth]{appendix/cr_deg_meta_vs_clin}
+		\hfill
+		\includegraphics[page=35,width=0.45\linewidth]{appendix/cr_deg_meta_vs_clin}\\
+		\caption{Name}
 	\end{figure}

### 2017-03-25 thesis/results/results.tex
finished editing up to the discussion

@@
 The obesity associated genetic signatures are central to this project, as a means to clarify the relationship between \gls{bmi} and tumour biology.
-In this chapter, the previously identified obesity associated genetic signatures from the studies conducted by \citet{Creighton2012} and \citet{Fuentes-Mattei2014} are examined in turn to judge the agreement of these signatures with the sample \gls{bmi}/\gls{bmi} status, as presented in thier results.
-After this, novel obesity associated genetic signatures are identified in the Creighton \textit{et al.} (CR) data set and was compared with the signatures from the Creighton \textit{et al.} and the Fuentes-Mattei \textit{et al.} studies.
+In this chapter, the previously identified obesity associated genetic signatures from the studies conducted by \citet{Creighton2012} and \citet{Fuentes-Mattei2014} are examined in turn to judge the agreement of these signatures with the patient \gls{bmi} and \gls{bmi} status, as presented in their results.
+After this, novel obesity associated genetic signatures are identified in the Creighton \textit{et al.} (CR) data set and was compared with the obesity associated genetic signatures from the \citet{Creighton2012} and the \citet{Fuentes-Mattei2014}  studies.
 Lastly, the presence of common genes or pathways associated with obesity in multiple types of cancer is explored using the data sets from \gls{icgc} (\citetalias{ICGC}; \citealp{Zhang2011}).
@@
 An obesity metagene was created using the obesity associated genetic signature from the \citet{Creighton2012} study.
-There was no description about the normalisation method used by Creighton \textit{et al.} when they first analysed their data, so the popular \gls{rma} method \citep{Irizarry2003} was used to  normalise the CR data set.
-The obesity metagene was created in the standardised \gls{rma}-normalised CR data set, and  was plotted above a heatmap with the sample gene expression  to check whether the metagene scores were in accordance with the overall gene expression of the samples (\cref{fig:crmetaheat}).
+There was no description about the normalisation method used by Creighton \textit{et al.} when they first analysed their data, so the more popular \gls{rma} method  was used to  normalise the CR data set \citep{Irizarry2003}.
+The obesity metagene was created in the standardised \gls{rma} normalised CR data set, and  was plotted above a heatmap with the sample gene expression  to check whether the metagene scores were in accordance with the overall gene expression of the samples (\cref{fig:crmetaheat}).
 
-\begin{figure}[htb]
+\begin{figure}[tb!]
 	\centering
@@
 As shown in \cref{fig:crmetaheat}, a high obesity associated metagene score  reflects low expression in majority of the genes in the signature, and in contrast, a low obesity associated metagene score reflects high expression in the majority of the genes in the signature.
-This was consistent with the reported property of the obesity associated signature by \citet{Creighton2012} (see \cref{ssub:creighton_study}).
-To provide further evidence that the obesity associated metagenes were in fact associated with both the \gls{bmi} status and the \gls{bmi} value of the samples, a box plot and a scatter plot were created, respectively (\cref{fig:crmetaboxplot}).
+This was consistent with the reported property of the obesity associated genetic signature by \citet{Creighton2012} (see \cref{ssub:creighton_study}).
+To provide further evidence that the obesity metagenes were in fact associated with both the \gls{bmi} status and the \gls{bmi} value of the patients, a box plot and a scatter plot were created, respectively (\cref{fig:crmetaboxplot}).
 
-\begin{figure}[htp!]
+\begin{figure}[tb!]
 	\centering
@@
 
-\cref{fig:crmetaboxplot} clearly showed that the obesity metagene from the \citet{Creighton2012} study significantly associated with the sample \gls{bmi} status, as well as sample \gls{bmi} value (include p-value and r-squared).
-It should be noted here that the obesity associated metagene significantly associated with the samples that were obese, but not with the samples that were overweight.
-This was due to the fact that the obesity associated genetic signatures were originally identified from the comparison of the samples that were obese with the samples that were not obese, and therefore the metagene scores were significant with the obese group, but not with the overweight group.
-In addition to this, although the regression line in the scatter plot showed statistically significant association with sample \gls{bmi}, the sample \gls{bmi} values seem to be randomly dispersed across the metagene scores.
-This suggests that perhaps the obesity metagene from the \citet{Creighton2012} study was not associated with the sample \gls{bmi} as strongly as reported.
+\cref{fig:crmetaboxplot} clearly showed that the obesity metagene from the \citet{Creighton2012} study significantly associated with the obese group of patients, as well as patient \gls{bmi} value.
+It should be noted here that the obesity metagene significantly associated with the samples from the patients that were obese, but not with the sample from the patients that were overweight.
+This was due to the fact that the obesity associated genetic signature was originally identified from the comparison of the patients that were obese with the patients that were not obese, and therefore the obesity metagene scores were significant with the obese group, but not with the overweight group.
+Another thing to note here was that even though  the regression line in the scatter plot showed statistically significant association with the patient \gls{bmi}, patient \gls{bmi} values seem to be randomly dispersed across the obesity metagene scores.
+This suggests that perhaps the obesity metagene from the \citet{Creighton2012} study may not be associated with the patient \gls{bmi} as strongly as reported.
 \\
@@
 \noindent
-Now that the association of the obesity metagene from the \citet{Creighton2012} study was established in CR data set, the obesity associated signature was generated in the \gls{icgc} cancer data.
-The direction of the obesity associated metagene was checked in the CR data first, so that high metagene scores reflected high sample \gls{bmi} and low metagene scores reflected low sample \gls{bmi} (\cref{fig:crmetaboxplot}).
-The transformation matrix was then created in CR data, as described in \cref{sub:svd}.
+Now that the association of the obesity metagene from the \citet{Creighton2012} study was established in the CR data set, the Creighton \textit{et al.} obesity metagene was generated in the \gls{icgc} cancer data.
+The direction of the obesity metagene was checked in the CR data first, so that high metagene scores reflected high patient \gls{bmi} and low metagene scores reflected low patient \gls{bmi} (\cref{fig:crmetaboxplot}).
+The transformation matrix was then created in the \gls{rma} normalised CR data, as described in \cref{sub:svd}.
 
-All of the \gls{icgc} data were normalised as described in \cref{ssub:rna_seq_data}.
-Before the transformation matrix was applied to the log$_{10}$-normalised cancer data, the obesity metagene was created from the standardised data or untouched (non-standardised) \gls{icgc} data and compared with one another to determine which was better for the application of the transformation matrix (see \cref{sec:comp_cr_raw_std_icgc}).
-From these results, the standardised data was found to be the most suitable to use for matrix transformation.
+All of the \gls{icgc} \gls{rnaseq} data were normalised as described in \cref{ssub:rna_seq_data}.
+Before the transformation matrix was applied to the log$_{10}$-normalised cancer data, the obesity metagene was created from the standardised data or untouched (non-standardised) \gls{icgc} data and compared with one another to determine which data format was better for the application of the transformation matrix (\cref{sec:comp_cr_raw_std_icgc}).
+From these results, the standardised data was found to be the most suitable format for the application  of the transformation matrix.
 The transformation matrix was applied to each cancer data set in turn to generate an obesity metagene from each of the data sets.
-Each obesity metagene was plotted in a heatmap with the corresponding data set from which the metagene was taken (\cref{fig:crmetaicgc}; \cref{sec:rest_of_the_cr_icgc_cancer_heatmap_results}).
-These heatmaps confirmed that the obesity associated metagene derived from the CR data was able to capture the overall gene expression pattern, where the metagene scores reflected the expression levels of the majority of the genes in the signature.
-As before, association of the obesity metagene with the sample \gls{bmi} and \gls{bmi} status was checked in their respective cancer data set (\cref{fig:crmetaicgc}; \cref{sec:rest_of_the_cr_icgc_cancer_heatmap_results}).
+Each obesity metagene was plotted in a heatmap with the corresponding \gls{icgc} data set from which the obesity metagene was generated from (\cref{fig:crmetaicgc}; \cref{sec:rest_of_the_cr_icgc_cancer_heatmap_results}).
+These heatmaps confirmed that the obesity metagene was able to capture the overall gene expression pattern in all of the \gls{icgc} cancer data, where the metagene scores reflected the expression levels of the majority of the genes in the signature.
+As before, association of the obesity metagene with the patient \gls{bmi} and \gls{bmi} status was examined in their respective cancer data set (\cref{fig:crmetaicgc}; \cref{sec:rest_of_the_cr_icgc_cancer_heatmap_results}).
 Out of all the cancer types, only the \gls{blca} data set showed a significant association with the obesity metagene,  and only for the overweight group (not the obese group).
-% However, neither the \gls{anova} p-value nor the regression line in the scatter plot were statistically significant, suggesting that this apparent association with the overweight group was not due to the obesity metagene (?).
 
@@
 	\includegraphics[page=5,width=0.45\linewidth]{results1/crtcga_std}
-	\caption[Obesity metagene from the \citet{Creighton2012} study in the \acrshort{icgc} \acrshort{blca} data]{Heatmap, box plot and scatter plot showing the association of the obesity metagene from the \citet{Creighton2012} study with the sample gene expression, \gls{bmi} and \gls{bmi} status, respectively, from the \gls{icgc} \gls{blca} data.
-	The results for other \gls{icgc} cancer types are shown in \cref{sec:rest_of_the_cr_icgc_cancer_heatmap_results}.
+	\caption[Obesity metagene from the \citet{Creighton2012} study in the \acrshort{icgc} \acrshort{blca} data]{Heatmap, scatter plot and box plot showing the association of the obesity metagene from the \citet{Creighton2012} study with the sample gene expression, patient \gls{bmi} and \gls{bmi} status, respectively, from the \gls{icgc} \gls{blca} data.
+	The results for the other \gls{icgc} cancer types are shown in \cref{sec:rest_of_the_cr_icgc_cancer_heatmap_results}.
 	Scales, p-values and $R^2$-value are as described in previous figures.}
@@
 
-There could be several reasons for the apparent lack of association of the metagene with sample \gls{bmi}/\gls{bmi} status.
+There could be several reasons for the apparent lack of association of the obesity metagene with patient \gls{bmi} in most of the \gls{icgc} data.
 First, the transformation matrix was derived from the CR microarray data, but the \gls{icgc} cancer data were generated via \gls{rnaseq}.
 Though the log$_{10}$-normalisation and standardisation of the data was the most appropriate adjustments to be made to the \gls{rnaseq} data, these adjustments were not equivalent to the \gls{rma} normalisation method that was used on the microarray data.
-Secondly, none of the \gls{icgc} cancer originated from breast tumours as in the CR data.
-Since the obesity associated signature was identified in the breast cancer data, the signature may be specific to breast cancer and may not be generalisable to other cancer types.
+Secondly, none of the \gls{icgc} cancer data in this project originated from breast tumours as in the CR data.
+Since the obesity associated genetic signature was identified in the breast cancer data, the signature may be specific to breast cancer and may not be generalisable to other cancer types.
 \\
@@
 \noindent
-To check whether the metagene was specific to breast cancer microarray data, the same transformation matrix was applied to the \gls{nzbc} microarray data from the \citet{Print2016} study (referred to as \gls{nzbc} data hereafter).
-\gls{nzbc} data was normalised with the \gls{rma} method and the transformation matrix was applied to the normalised data to obtain the metagene.
-The metagene was again compared with the gene expression of the samples with a heatmap and the association of the metagene with the sample \gls{bmi}/\gls{bmi} status was examined with box and scatter plots (\cref{fig:crmetaprint}).
+To check whether the signature was specific to breast cancer microarray data, the same transformation matrix used in the \gls{icgc} data was applied to the breast cancer microarray data from the \citet{Print2016} study (referred to as \gls{nzbc} data hereafter).
+\gls{nzbc} data was normalised with the \gls{rma} method and the transformation matrix was applied to the normalised data to obtain the Creighton \textit{et al.} obesity metagene in the \gls{nzbc} data set.
+The generated obesity metagene was again compared with the gene expression of the samples with a heatmap and the association of the metagene with the patient \gls{bmi} and \gls{bmi} status was examined with box and scatter plots, respectively  (\cref{fig:crmetaprint}).
 
-The Creighton \textit{et al.} obesity metagene managed to reflect the overall gene expression of the samples in \gls{nzbc}  data.
-However, as with the \gls{icgc} cancer data, the obesity metagene scores did not significantly associate with the sample \gls{bmi} or \gls{bmi} status (\cref{fig:crmetaprint}).
-These results confirmed that the lack of association of the obesity metagene from the CR data was not due to the technology in which the data was gathered (mcroarray or \gls{rnaseq}), nor the cancer type in which the transformation matrix was applied to.
+The Creighton \textit{et al.} obesity metagene managed to reflect the overall gene expression of the samples in the \gls{nzbc}  data (\cref{fig:crmetaprint}).
+However, as with the \gls{icgc} cancer data, the obesity metagene scores did not significantly associate with patient \gls{bmi} or \gls{bmi} status (\cref{fig:crmetaprint}).
+These results confirmed that the lack of association of the obesity associated genetic signature from the Creighton \textit{et al.} study was not due to the technology in which the data was gathered (microarray or \gls{rnaseq}), nor the cancer type in which the genetic signature was derived from.
 
@@
 	\includegraphics[width=0.45\linewidth,page=5]{results1/cris_cr_trans_meta}
-	\caption[Obesity metagene from \citet{Creighton2012} study in the \gls{nzbc} data]{Heatmap, box plot and scatter plot showing the association of the obesity metagene from the \citet{Creighton2012} study with the sample gene expression, \gls{bmi} and \gls{bmi} status from the \gls{nzbc} data, respectively.
+	\caption[Obesity metagene from the \citet{Creighton2012} study in the \gls{nzbc} data]{Heatmap, scatter plot and box plot showing the association of the obesity metagene from the \citet{Creighton2012} study with the sample gene expression, patient \gls{bmi} and \gls{bmi} status, respectively, from the \gls{nzbc} data.
 	Scales, p-values and $R^2$-value are as described in previous figures.}
@@
 
-Taken together, these results suggest that the obesity metagene identified in the study conducted by \citet{Creighton2012} associated with sample \gls{bmi} and \gls{bmi} status only in that particular data set.
-The lack of association with \gls{bmi} was not due to the type of technology platform in which the data was gathered, as neither the \gls{icgc} \gls{rnaseq} data nor the \gls{nzbc} microarray data showed a significant association with the obesity metagene.
-Furthermore, the obesity metagene was not dependent on the cancer type since the obesity associated genetic signature did not show significant association in either the \gls{icgc} cancer data or in \gls{nzbc} data.
+Taken together, these results suggest that the obesity associated genetic signature identified in the study conducted by \citet{Creighton2012} correlated with patient \gls{bmi} and \gls{bmi} status only in the CR data set.
+The lack of association with patient \gls{bmi} in other cancer data sets was not due to the type of technology platform in which the data was gathered, as neither the \gls{icgc} \gls{rnaseq} data nor the \gls{nzbc} microarray data showed significant association of the obesity metagene with the patient \gls{bmi}.
+Furthermore, the obesity associated genetic signature was not dependent on the cancer type in which it was generated from, since the obesity metagene did not show significant association in either the \gls{icgc} cancer data or in the \gls{nzbc} data.
 
 One possible reason why the obesity metagene from the \citet{Creighton2012} study did not show significant association with other data sets could be because the genetic signature was in fact not an obesity specific signature, but a signature that was detected due to another clinical variable (investigated in \cref{sec:creighton_obesity_metagene_new}).
-Another reason for this apparent lack of association could be that the genetic signature was too specific to the data and was not a broad obesity associated genetic signature, but an obesity associated signature specific to the samples that were profiled in the Creighton \textit{et al.} publication.
+Another reason for this apparent lack of association could be that the obesity associated genetic signature was too specific to the CR data and was not a broad obesity associated genetic signature, but an obesity associated signature that was specific to the patient cohort that was profiled in the \citet{Creighton2012} publication.
+% Another reason for this apparent lack of association could be that the genetic signature was too specific to the data and was not a broad obesity associated genetic signature, but an obesity associated signature that was specific to the samples that were profiled in the Creighton \textit{et al.} publication.
 
@@
 
-Since the obesity associated genetic signature from the \citet{Creighton2012} study was not associated with sample \gls{bmi} or \gls{bmi} status in other data sets, the obesity associated genetic signature from the \citet{Fuentes-Mattei2014} study (FM) was examined to see whether the FM obesity metagene associated with the sample \gls{bmi} or \gls{bmi} status.
-Since the FM data set did not have sample \gls{bmi} information, the FM obesity metagene was not able to be compared with the sample \gls{bmi} or \gls{bmi} status in the original data.
-However, the transformation matrix was still applied to other tumour types to see whether the FM obesity metagene associated with sample \gls{bmi}/\gls{bmi} status in those data sets.
+The obesity associated genetic signature from the \citet{Creighton2012} study was not associated with patient \gls{bmi} or \gls{bmi} status in majority of the cancer data sets, so the obesity associated genetic signature from the \citet{Fuentes-Mattei2014} study (FM) was examined to see whether this obesity metagene was able to significantly associate with the patient \gls{bmi} and \gls{bmi} status across different  data sets.
+Since the FM data set did not have patient \gls{bmi} information, the FM obesity metagene was not able to be compared with the patient \gls{bmi} or \gls{bmi} status in the original FM data.
+Nevertheless, the transformation matrix was still generated in the FM data and applied first to the microarray data (CR and \gls{nzbc} data sets) then to the \gls{icgc} data sets to see whether the FM obesity metagene associated with patient \gls{bmi} and \gls{bmi} status in these data sets.
+% However, the transformation matrix was generated in the FM data and applied to other cancer data sets  to see whether the FM obesity metagene associated with patient \gls{bmi} and \gls{bmi} status in those data sets.
 
-Firstly, FM data was normalised with the \gls{rma} method and \gls{svd} was performed on the normalised FM data to obtain the transformation matrix.
-The transformation matrix was used to transform the \gls{rma}-normalised CR data to extract the FM obesity metagene scores in the CR data.
-The FM obesity metagene scores were compared with gene expression and sample \gls{bmi}/\gls{bmi} status in the CR data, as shown in \cref{fig:fmmetacr}.
-Clearly, as with the CR obesity metagene, the FM obesity metagene was reflective of the overall gene expression of the samples, but did not associate with the sample \gls{bmi} or \gls{bmi} status, suggesting that this signature also does not generalise to other data sets.
-The transformation matrix was applied to \gls{nzbc} data  as well (\cref{fig:fmmetacris}).
-Again, FM obesity metagene scores reflected the gene expression of the samples, did not significantly associate with sample \gls{bmi}/\gls{bmi} status.
-
-Next, the transformation matrix was applied to the \gls{icgc} cancer data and the resulting metagenes were compared with the gene expression and sample \gls{bmi}/\gls{bmi} status.
-As evident in \cref{fig:fmmetaicgc}, the FM obesity metagene scores appeared to reflect the overall gene expression of the FM obesity associated genetic signature.
-As with all the results in this chapter so far, the FM obesity metagene did not significantly associate with any of the \gls{icgc} cancer data, except in the \gls{blca} data set (\cref{fig:fmmetaicgc}).
-The FM obesity metagene significantly associated with the overweight group (but not with the obese group), and also had a significant \gls{anova} p-value.
-On the contrary to the association of the metagene with the sample \gls{bmi} status, the FM obesity metagene was not associated with sample \gls{bmi}.
-These results suggested that the samples that were overweight in the \gls{blca} data set had similar biological properties as the samples that were obese in the FM data set.
-However, due to the fact that the FM metagene lacked association with the sample \gls{bmi} in the \gls{blca} data set and that the metagene did not show any significant association in any other cancer type, it was difficult to determine whether the observed association of the FM obesity metagene with the overweight group of samples was truly reflective of the effect resulted from the FM metagene.
-\\
+The FM data was normalised with the \gls{rma} method and \gls{svd} was performed on the normalised FM data to obtain the transformation matrix.
+The transformation matrix was used to transform the \gls{rma} normalised CR data to extract the FM obesity metagene scores in the CR data.
+The FM obesity metagene scores were compared with sample gene expression levels, patient \gls{bmi} and \gls{bmi} status in the CR data (\cref{fig:fmmetacr}).
+Clearly, as with the CR obesity metagene, the FM obesity metagene was reflective of the overall gene expression of the samples, but did not associate with patient \gls{bmi} or \gls{bmi} status, suggesting that this signature also does not generalise to other data sets.
+The transformation matrix was then applied to the \gls{nzbc} microarray  data.
+Again, FM obesity metagene scores reflected the gene expression of the samples, but did not significantly associate with patient \gls{bmi} or \gls{bmi} status (\cref{fig:fmmetacris}).
 
@@
 	\includegraphics[page=5,width=0.45\linewidth]{results1/cr_fm_meta}
-	\caption[FM obesity metagene in the CR data]{Heatmap, box plot and scatter plot showing the association of the FM obesity associated metagene with the sample gene expression, \gls{bmi} and \gls{bmi} status from the CR  data, respectively.
+	\caption[FM obesity metagene in the CR data]{Heatmap, scatter plot and box plot showing the association of the FM obesity metagene with the sample gene expression, patient \gls{bmi} and \gls{bmi} status, respectively, from the CR  data.
 	Scales, p-values and $R^2$-value are as described in previous figures.}
@@
 	\includegraphics[page=5,width=0.45\linewidth]{results1/cris_fm_meta}
-	\caption[FM metagene in the \gls{nzbc} data]{Heatmap, box plot and scatter plot showing the association of the FM obesity associated metagene with the sample gene expression, \gls{bmi} and \gls{bmi} status from the \gls{nzbc} data, respectively.
+	\caption[FM metagene in the \gls{nzbc} data]{Heatmap, scatter plot and box plot showing the association of the FM obesity associated metagene with the sample gene expression, patient  \gls{bmi} and \gls{bmi} status, respectively, from the \gls{nzbc} data.
 	Scales, p-values and $R^2$-value are as described in previous figures.}
@@
 \end{figure}
+
+Next, the transformation matrix was applied to the \gls{icgc} cancer data and the resulting FM metagene was compared with the sample gene expression, patient \gls{bmi} and \gls{bmi} status in each of the \gls{icgc} cancer data.
+As evident in \cref{fig:fmmetaicgc,sec:rest_of_the_fm_icgc_cancer_heatmap_results}, the FM obesity metagene scores appeared to reflect the overall gene expression of the FM obesity associated genetic signature.
+As with all the results in this chapter so far, the FM obesity metagene did not significantly associate with any of the \gls{icgc} cancer data, except in the \gls{blca} data set (\cref{fig:fmmetaicgc}; \cref{sec:rest_of_the_fm_icgc_cancer_heatmap_results}).
+The FM obesity metagene significantly associated with the overweight group (but not with the obese group), and also had a significant \gls{anova} p-value \cref{fig:fmmetaicgc}.
+On the contrary to the association of the metagene with the patient \gls{bmi} status, the FM obesity metagene was not associated with the patient \gls{bmi}.
+These results suggested that the samples from the patients that were overweight in the \gls{blca} data set had similar biological properties as the samples taken from the patients that were obese in the FM data set.
+However, due to the fact that the FM obesity metagene lacked association with the patient \gls{bmi} in the \gls{blca} data set and that the metagene did not show any significant association in any of the other \gls{icgc} cancer types, it was difficult to determine whether the observed association of the FM obesity metagene with the overweight group was truly reflective of the quality of the FM metagene or observed by chance.
 
@@
 	\includegraphics[page=5,width=0.45\linewidth]{results1/fm_meta_ICGC}
-	\caption[FM obesity metagene in the \acrshort{icgc} \acrshort{blca} data]{Heatmap, box plot and scatter plot showing the association of the FM obesity associated metagene with the sample gene expression, \gls{bmi} and \gls{bmi} status from the \acrshort{icgc} \acrshort{blca} data, respectively.
+	\caption[FM obesity metagene in the \acrshort{icgc} \acrshort{blca} data]{Heatmap, scatter plot and box plot showing the association of the FM obesity metagene with the sample gene expression, patient \gls{bmi} and \gls{bmi} status, respectively, from the \acrshort{icgc} \acrshort{blca} data.
 	The results for other \gls{icgc} cancer types are shown in \cref{sec:rest_of_the_fm_icgc_cancer_heatmap_results}.
@@
 
-\noindent
-These results showed that FM obesity associated metagene was not generalisable to other cancer data sets, similar to the obesity metagene identified by Creighton \textit{et al.}.
-This meant that both the obesity metagenes identified by Creighton \textit{et al.} and Fuentes-Mattei \textit{et al.} may have been too specific to the orignial data set in which the signatures were identified.
+These results showed that the obesity associated genetic signature from the \citet{Fuentes-Mattei2014} study was not generalisable to other cancer data sets, similar to the obesity signature identified by Creighton \textit{et al.}.
+This meant that both the Creighton \textit{et al.} and Fuentes-Mattei \textit{et al.} obesity metagenes may have been too specific to the orignial data set in which the signatures were identified in.
 Furthermore, there was a possibility that these obesity associated metagenes were not related to obesity, but associated with a different clinical variable that may be closely related to \gls{bmi} (\cref{sec:creighton_obesity_metagene_new}).
@@
 
-\subsection{Identification of obesity associated genetic signatures}
+\subsection{Identification of novel obesity associated genetic signatures}
 \label{sub:identification_of_obesity_associated_genetic_signatures}
 
-Both the obesity associated metagenes generated from CR data and FM data were able to capture the overall pattern of gene expression of the genetic signature in the samples, but did not associate with sample \gls{bmi} or \gls{bmi} status.
-One possible reason for this result could be that the obesity associated genetic signatures from the \citet{Creighton2012} and \citet{Fuente-Mattei2014} studies may not have been truly associated with sample \gls{bmi}  status, but with another clinical variable.
-To invevstigate this possibility, obesity associated genetic signatures were identified in CR data after controlling for all the clinical variables in the data set.
-FM data set was not used to get obesity associated genetic signatures, as no \gls{bmi} information was available for the samples in FM data.
+Both the obesity associated genetic signatures identfied from the \citet{Creighton2012} and \citet{Fuentes-Mattei2014} studies were able to capture the overall gene expression patterns  of the genetic signatures of the samples, but did not associate with patient \gls{bmi} or \gls{bmi} status in majority of the cancer data sets.
+One possible reason for this result could be that the obesity associated genetic signatures from the \citet{Creighton2012} and \citet{Fuentes-Mattei2014} studies may not have been truly associated with patient \gls{bmi}, but with another clinical variable.
+To invevstigate this possibility, novel obesity associated genetic signatures were identified in the CR data after controlling for all the clinical variables in the data set.
+The FM data set was not used to get the obesity associated genetic signatures, as no \gls{bmi} information was available for the patients in the FM data.
 
-Firstly, to determine whether the identified obesity associated genetic signature was similar to the signature that Creighton \textit{et al.} had originally found, the identified \glspl{deg} were compared with the original genetic signature identified by Creighton \textit{et al.}.
-\citet{Creighton2012} originally found their obesity associated genetic signature by doing a gene expression analysis between the breast tumour samples from the obese and non-obese patients, and from this list, they selected the 799 significant \glspl{deg} with log$_2$-fold change greater than 1.2.
+Firstly, an attempt was made to replicate the Creighton \textit{et al.} obesity associated genetic signature from the original CR data.
+\citet{Creighton2012} originally found their obesity associated genetic signature by doing a gene expression analysis between the breast tumour samples from the obese and non-obese patients, and from this list, Creighton \textit{et al.} selected the 799 statistically significant \glspl{deg} (p \textless{} 0.01) with log$_2$-fold change greater than 1.2.
+Therefore, gene expression analysis  between the samples from obese and non-obese patients in the \gls{rma} normalised CR data was carried out (described in \cref{sec:gene_expression_analysis}).
+% Therefore, \glspl{deg}  were identified between the samples from obese and non-obese patients in the \gls{rma} normalised CR data, as described in \cref{sec:gene_expression_analysis}.
+Without adjusting the p-value to account for multiple hypothesis testing, 5278 gene probes and 1781 gene probes were significant at p \textless{} 0.05 and p \textless{} 0.01, respectively (\cref{tab:ob_deg_summary}).
+After adjustments were made for multiple hypothesis testing, there were only 9 gene probes significant at p \textless{} 0.05 and no gene was significant at p \textless{} 0.01, which suggest that the the majority of the 799 probe sets reported by Creighton \textit{et al.} may actually be false positives.
 
-Therefore, \glspl{deg}  were identified between the samples from obese and non-obese patients in the \gls{rma}-normalised CR data, as described in \cref{sec:gene_expression_analysis}.
-Without adjusting the p-value to account for multiple hupothesis testing, 5278 gene probes and 1781 gene probes were significant at p \textless{} 0.05 and p \textless{} 0.01, respectively (\cref{tab:ob_deg_summary}).
-After adjustment, there were only 9 gene probes significant at p \textless{} 0.05 and no gene was significant at p \textless{} 0.01, which suggests that the the majority of the 799 probe sets reported by Creighton \textit{et al.} may actually be false positives.
 Furthermore, there were only 61 gene probes that were significantly differentially expressed at p \textless{} 0.05 with a log$_2$-\gls{fc} greater than 1.2.
-From these observations, the log$_2$-fold change of the gene probes were ignored and the threshold p-value was set to 0.01 (without adjustment) for the identification of probe sets.
+From these observations, the log$_2$-fold change of the gene probes were ignored and the threshold p-value was set to 0.01 (unadjusted p-value) for the identification of the significant probe sets.
 Additionally, when there were more than 799 gene probes identified, only the most significant 799 gene probes were taken as the obesity associated genetic signature, as this many genes were originally identified by Creighton \textit{et al.}.
-This was to include as many probe set as possible so the novel genetic signatures identified were able to be compared with the original Creighton \textit{et al.} signature.
+This was to include as many probe set as possible for the novel genetic signatures to be comparable with the original Creighton \textit{et al.} obesity associated genetic signature.
 These criteria were applied for the identification of other genetic signatures as well.
 
-The above analysis was repeated with the residual data (\gls{rma}-normalised CR data that had been controlled for other clinical variables; see \cref{sub:residual_data_creation}).
+% To determine whether the identified obesity associated genetic signature was similar to the signature that Creighton \textit{et al.} had originally found, the identified \glspl{deg} were compared with the original genetic signature identified by Creighton \textit{et al.}.
+
+The above analysis was repeated with the residual data (\gls{rma} normalised CR data that had been controlled for other clinical variables; see \cref{sub:residual_data_creation}).
 The clinical variables controlled were age, ethnicity, menopause status, tumour grade, hormone (\gls{er}, \gls{pr}, and \gls{her2}) statuses  and \gls{ln} status.
-In the residual data, 1104 gene probes were significant with unadjusted p-value (p \textless{} 0.01; additional results shown in \cref{tab:ob_deg_summary}).
-Again, the most significant 799 genes were taken as the obesity associated genetic signature from this data.
+In the residual data, 1104 gene probes were significantly differentially expressed with unadjusted p-value (p \textless{} 0.01; additional results shown in \cref{tab:ob_deg_summary}).
+Again, the most significant 799 gene  probes were taken as the obesity associated genetic signature from this data.
+
+In addition to the above two obesity associated genetic signatures, two more sets of genetic signatures were identified by taking the gene probes that were common between each of the above two signatures with the original obesity associated genetic signatures.
+There were 239 common gene probes between the original Creighton \textit{et al.} obesity associated  genetic signature and the gene probes identified from the unadjusted CR data, and 168 common gene probes between the original signature and the gene probes from the residual CR data.
+The genetic signatures identified were summarised as a Venn diagram, as shown in \cref{fig:venn1}.
+\\
 
@@
 	\centering
-	\caption[Summary of the number of \acrshort{deg}s identified using different unadjusted and \acrshort{fdr}-adjusted p-value threshold in different versions of the \gls{rma}-normalised CR data set]{Summary of the number of \glspl{deg} identified using different unadjusted and \gls{fdr}-adjusted p-value threshold in different versions of the \gls{rma}-normalised CR data set}
+	\caption[Summary of the number of \acrshort{deg}s identified using different unadjusted and \acrshort{fdr}-adjusted p-value threshold in different versions of the \gls{rma} normalised CR data set]{Summary of the number of \glspl{deg} identified using different unadjusted and \gls{fdr}-adjusted p-value thresholds in different versions of the \gls{rma} normalised CR data set}
 	\label{tab:ob_deg_summary}
@@
 
-In addition to the above two obesity associated genetic signatures, two more sets of genetic signatures were identified by taking the common genes between the above two signatures with the original obesity associated genetic signatures.
-There were 239 common gene probes between the original genetic signature and the gene probes identified from the unadjusted CR data, and 168 common gene probes between the original signature and gene probes from residual data in CR data.
-The genetic signatures identified were summarised as a Venn diagram, as shown in \cref{fig:venn1}.
-\\
-
 \begin{figure}[htpb]
@@
 	\includegraphics[width=0.5\linewidth]{results1/deg_venn1}
-	\caption[Venn diagram of the \glspl{deg} identified from the CR data (all patients included)]{Venn diagram showing the common gene probes between the signatures obtained from the unadjusted and the clinical variable-adjusted CR data and the original obesity associated genetic signature.
-}
+	\caption[Venn diagram of the \glspl{deg} identified from the CR data (all patients included)]{Venn diagram showing the common gene probes between the signatures obtained from the unadjusted and the clinical variable-adjusted CR data and the original obesity associated genetic signature from the \citet{Creighton2012} study }
 	\label{fig:venn1}
@@
 \noindent
-There was a possible bias between different ethnic groups, where African American patients were more likely to be obese compared to the Caucasian patients (13 out of 16 were obese in African American group, whereas 22 out of 77 were obese in Caucasian group; see \cref{ssub:creighton_study}).
-Though ethnicity was controlled in the residual data, the effect of ethnicity on the data was completely removed to prevent any possibility of ethnicity influencing the analysis.
-Therefore, the effect of ethnicity was ignored by considering only the Caucasian patients in CR data, which left a total of 77 Caucasian patients in the data set.
-With this Caucasian-only data set, obesity associated genetic signatures were identified as described above; first with the unadjusted data, then with the data with clinical variables adjusted (except ethnicity, as it has already been controlled for by considering Caucasian samples only), and lastly the common genes with the original signature.
+There was a possible bias between the different ethnic groups, where African-American patients were more likely to be obese compared with the Caucasian patients (13 out of 16 were obese in African-American group, whereas 22 out of 77 were obese in Caucasian group; see \cref{ssub:creighton_study,tab:num_sample_microarray} in \cref{sub:breast_cancer_data}).
+Though ethnicity was controlled in the residual CR data, the effect of ethnicity on the CR data was completely removed to prevent any possibility of ethnicity influencing the analysis.
+Therefore, the effect of ethnicity was ignored by considering only the Caucasian patients in the CR data, which left a total of 77 Caucasian patients in the data set.
 
-2129 and 1558 gene probes were identified in the unadjusted and clinical variable-adjusted Caucasian patient data, respectively.
+With the Caucasian-only CR data set, obesity associated genetic signatures were identified as described above.
+% ; first in the unadjusted data, then in the data with clinical variables adjusted (except ethnicity, as it has already been controlled for by considering only the Caucasian samples), and lastly the common genes between each of these two signatures with the original obesity associated genetic signature were identified.
+2129 and 1558 gene probes were significantly differentially expressed (unadjuste p-value \textless{} 0.01) in the unadjusted and clinical variable-adjusted Caucasian patient data, respectively (\cref{tab:ob_deg_summary}).
 As before, the most significant 799 gene probes were selected from these gene probes.
-There were 148 and 92 common gene probes with the original genetic signatures and unadjusted or clinical variable-adjusted Caucasian patient data set, respectively.
+There were 148 and 92 common gene probes with the original Creighton \textit{et al.} obesity associated genetic signatures and unadjusted or clinical variable-adjusted Caucasian patient data set, respectively.
 Again, Venn diagram was used to summarise the genes identified from the Caucasian patient data set (\cref{fig:venn2}).
 
-Since all of these obesity associated genetic signatures were derived based on the discrete values of sample \gls{bmi} staus, a genetic signature was identified by using the correlation of the gene probe expression with the sample \gls{bmi} value (continuous variable).
-However, no significant genes were identified and thus was not used for further analysis (\cref{sec:obesity_associated_genetic_signature_using_sample_bmi_values}).
+\begin{figure}[tb]
+	\centering
+	\includegraphics[width=0.5\linewidth]{results1/deg_venn2}
+	\caption[Venn diagram of the \glspl{deg} identified from the CR data (Caucasian patients only)]{Venn diagram showing the common gene probes between the signatures obtained from the unadjusted and the clinical variable-adjusted CR data (Caucasian patients only) and the original obesity associated genetic signature from the \citet{Creighton2012} study}
+	\label{fig:venn2}
+\end{figure}
+
+Additionally, the correlation of all of the obesity metagenes were examined to see whether these metagenes were similar to one another.
+As clearly shown in \cref{fig:cr_meta_cor}, there were two distinct groups within the eight metagenes: the first group contained the obesity metagenes that were not overlapped with the Or metagene (Cr, Res, Ca and CaRes), while the other group had the overlapped metagenes (CrOl, ResOl, CaOl and CaResOl).
+With that said, all eight metagenes showed high correlation with one another (lowest correlation approximately at 0.85), which suggested that all of these metagenes were detecting similar underlying biological mechanism from the data, as would be expected.
 
@@
 	\centering
-	\includegraphics[width=0.5\linewidth]{results1/deg_venn2}
-	\caption[Venn diagram of the \glspl{deg} identified from the CR data (Caucasian patients only)]{Venn diagram showing the common gene probes between the signatures obtained from the unadjusted and the clinical variable-adjusted CR data (Caucasian patients only) and the original obesity associated genetic signature.}
-	\label{fig:venn2}
+	\includegraphics[width=0.8\linewidth]{results1/cr_meta_cor}
+	\caption[Pearson correlation of all eight obesity metagenes identified in the CR data]{Heatmap showing the Pearson correlation of all eight obesity associated metagenes from the CR data with one another.
+	\gls{svd} was applied to the \gls{rma} normalised CR data to generate each of the eight metagenes.
+	High and low correlation were represented as red and blue, respectively, where the colours were matched with the values on the scale shown in the top right histogram.
+	The lowest correlation was 0.8469, between the CrOl and the CaRes metagenes.
+	}
+	\label{fig:cr_meta_cor}
 \end{figure}
 
-All of these obesity associated genetic signatures (eight in total) were checked to see whether these signatures had significant association with sample \gls{bmi}/\gls{bmi} status in various data set.
+Since all of these obesity associated genetic signatures were derived based on the discrete values of the patient \gls{bmi} status, an obesity associated genetic signature was searched using the correlation of the gene probe expression with the patient \gls{bmi} value (continuous variable) in the CR data with all of the samples included, but no significant genes were identified (\cref{sec:obesity_associated_genetic_signature_using_sample_bmi_values}).
+All of these obesity associated genetic signatures (eight in total) were checked to see whether these signatures had significant association with patient \gls{bmi} or \gls{bmi} status in the \gls{nzbc} and \gls{icgc} cancer data sets (\cref{sub:_novel_obesity_associated_signatures_and_sample_bmi}).
 For simplicity, the abbreviations in \cref{tab:mg_abbrev} will be used to refer to the appropriate genetic signatures.
 
-\begin{table}[htpb]
+\begin{table}[tb]
 	\centering
@@
 
-\subsection{Novel obesity associated signatures and sample \gls{bmi}/\gls{bmi} status}
+\subsection{Novel obesity associated signatures and patient \gls{bmi}/\gls{bmi} status}
 \label{sub:_novel_obesity_associated_signatures_and_sample_bmi}
 
-As with the Or signature, all eight obesity associated genetic signatures were validated in CR data set first, and then compared in other cancer data sets by using the transformation matrix that was generated in \gls{rma}-normalised CR data.
-First, the CR data (unadjusted, all samples included) was normalised with the \gls{rma} method and \gls{svd} was applied to generate the all of the obesity metagenes.
+As with the Or signature, all eight obesity associated genetic signatures were validated in the CR data set first, and then compared in other cancer data sets by using the transformation matrix that was generated in the \gls{rma} normalised CR data.
+The CR data (unadjusted, all samples included) was normalised with the \gls{rma} method and \gls{svd} was applied to generate the all of the obesity metagenes.
 The direction of the metagenes were examined to make sure that all of the metagenes were in line with one another (see \cref{sub:metagene_direction}; \cref{sec:direction_of_all_the_obesity_metagenes_from_the_cr_data}).
-The comparison of the metagenes with the sample gene expression in CR data were displayed as heatmaps (\cref{fig:degmetacr}).
+The comparison of the metagenes with the sample gene expression in the CR data were displayed as heatmaps (\cref{fig:degmetacr}; \cref{sec:rest_of_the_cr_ob_meta_heatmap_results_cr}).
 It was clear from the heatmaps that all of the metagenes reflected the overall expression of the corresponding obesity associated genetic signatures.
-The association of the metagenes with sample \gls{bmi}/\gls{bmi} status was significant for all eight of the metagenes identified (\cref{fig:degmetacr}).
-These results confirmed that all of the obesity associated genetic signatures identified in \cref{sub:identification_of_obesity_associated_genetic_signatures} significantly associated with sample \gls{bmi}/\gls{bmi} status in CR data, where the metagenes were derived from.
-However in \gls{nzbc} data, the higher the metagenes scores of the samples were, the more expressed the genes were (and vice versa), but lacked the association with the sample \gls{bmi}/\gls{bmi} status (\cref{fig:degmetaprint}; \cref{sec:rest_of_the_cr_ob_meta_heatmap_results_cris}).
+The association of the metagenes with patient \gls{bmi} and \gls{bmi} status was significant for all eight of the obesity associated genetic signatures identified (\cref{fig:degmetacr}; \cref{sec:rest_of_the_cr_ob_meta_heatmap_results_cr}).
+These results confirmed that all of the obesity associated genetic signatures identified in \cref{sub:identification_of_obesity_associated_genetic_signatures} significantly associated with patient \gls{bmi} and \gls{bmi} status in the CR data in which the metagenes were derived from.
+In the \gls{nzbc} data, the higher the metagenes scores of the samples were, the more expressed the genes were (and vice versa), but lacked the association with the patient \gls{bmi} or \gls{bmi} status (\cref{fig:degmetaprint}; \cref{sec:rest_of_the_cr_ob_meta_heatmap_results_cris}).
 
@@
 	\includegraphics[page=5,width=0.45\linewidth]{results1/cr_deg_meta_vs_clin}
-	\caption[Cr obesity metagene in the CR data]{Heatmap, box plot and scatter plot showing the association of the Cr obesity associated metagene with the sample gene expression, \gls{bmi} and \gls{bmi} status from the CR data, respectively.
-	The results for the other obesity metagenes are shown in \cref{sec:rest_of_the_cr_ob_meta_heatmap_results_cris}.
+	\caption[Cr obesity metagene in the CR data]{Heatmap, scatter plot and box plot showing the association of the Cr obesity associated metagene with the sample gene expression, patient \gls{bmi} and \gls{bmi} status, respectively, from the CR data.
+	The results for the other obesity metagenes are shown in \cref{sec:rest_of_the_cr_ob_meta_heatmap_results_cr}.
 	Scales, p-values and $R^2$-value are as described in previous figures.}
@@
 \end{figure}
-
-Additionally, the correlation of all of the obesity metagenes were examined to see whether these metagenes were similar to one another.
-As clearly shown in \cref{fig:cr_meta_cor}, there were two distinct groups within the eight metagenes; the first group contained the metagenes that were not overlapped with the Or metagene (Cr, Res, Ca and CaRes), while the other group had the overlapped metagenes (CrOl, ResOl, CaOl and CaResOl).
-With that said, all eight metagenes showed high correlation with one another (lowest correlation approximately at 0.85), which suggested that all of these metagenes were detecting similar underlying biological mechanism from the data, as would be expected.
 
@@
 	\includegraphics[page=5,width=0.45\linewidth]{results1/cris_crdeg_trans_meta}
-	\caption[Cr obesity metagene in the \gls{nzbc} data]{Heatmap, box plot and scatter plot showing the association of the Cr obesity associated metagene with the sample gene expression, \gls{bmi} and \gls{bmi} status from the \gls{nzbc} data, respectively.
+	\caption[Cr obesity metagene in the \gls{nzbc} data]{Heatmap, scatter plot and box plot showing the association of the Cr obesity associated metagene with the sample gene expression, patient \gls{bmi} and \gls{bmi} status, respectively, from the \gls{nzbc} data.
 	The results for the other obesity metagenes are shown in \cref{sec:rest_of_the_cr_ob_meta_heatmap_results_cris}.
@@
 
-\begin{figure}[htpb]
-	\centering
-	\includegraphics[width=0.8\linewidth]{results1/cr_meta_cor}
-	\caption[Pearson correlation of all eight obesity metagenes identified in the CR data]{Heatmap showing the Pearson correlation of all eight obesity associated metagenes from the CR data with one another.
-	\gls{svd} was applied to the \gls{rma}-normalised CR data to generate each of the eight metagenes.
-	High and low correlation were represented as red and blue, respectively, where the colours were matched with the values on the scale shown in the top right histogram.
-	The lowest correlation was 0.8469, between the CrOl and the CaRes metagenes.
-	}
-	\label{fig:cr_meta_cor}
-\end{figure}
+To confirm whether these metagenes showed significant association with patient \gls{bmi} or \gls{bmi} status in other cancer types, the transformation matrix was applied to the \gls{icgc} cancer data.
+In all of the \gls{icgc} cancer data sets, all eight metagene scores were reflective of the sample gene expression of the corresponding obesity associated genetic signatures (\cref{fig:degmetaicgc}; \cref{sec:rest_of_the_cr_icgc_cancer_heatmap_results}).
+In all cancer types but \gls{blca}, none of the obesity metagenes significantly associated with the patient \gls{bmi} or \gls{bmi} status (\cref{sec:rest_of_the_cr_icgc_cancer_heatmap_results}).
+As shown in \cref{fig:degmetaicgc}, the \gls{blca} data set showed significant association with the Cr metagene with the patient \gls{bmi} and \gls{bmi} status.
+Furthermore, Res and Ca metagenes also showed statistical significance in the overweight group, \gls{anova} and the regression line p-values; CaRes and Res metagenes were significant in overweight group and \gls{anova} p-value; CrOl and CaResOl metagenes were significant in overweight group; and ResOl and CaOl metagenes were not significantly associated with the sample \gls{bmi} or \gls{bmi} status in \gls{blca} data set (summarised in \cref{tab:degmetablca}).
 
-To confirm whether these metagenes showed significant association with sample \gls{bmi}/\gls{bmi} status in other cancer types, the transformation matrix was applied to the \gls{icgc} cancer data.
-Firstly in the \gls{icgc} cancer data sets, all eight metagene scores were reflective of the sample gene expression of the corresponding genetic signatures (\cref{fig:degmetaicgc}; \cref{sec:rest_of_the_cr_icgc_cancer_heatmap_results}).
-In all cancer types but \gls{blca}, none of the metagenes significantly associated with the sample \gls{bmi}/\gls{bmi} status (\cref{sec:rest_of_the_cr_icgc_cancer_heatmap_results}).
-As shown in \cref{fig:degmetaicgc}, only the \gls{blca} data set showed significant association of the Cr metagene with the sample \gls{bmi}/\gls{bmi} status.
-The Res and Ca metagenes also showed statistical significance in the overweight group, \gls{anova} and the regression line p-values; CaRes and Res metagenes were significant in overweight group and \gls{anova} p-value; CrOl and CaResOl metagenes were significant in overweight group; and ResOl and CaOl metagenes were not significantly associated with the sample \gls{bmi} or \gls{bmi} status in \gls{blca} data set.
+This was unexpected since all of the genetic signatures resulted from the  gene expression analysis between the non-obese group and the obese group of the samples in the CR data, and yet all of the obesity metagenes showed significant association with the overweight group rather than the obese group in the \gls{blca} data set.
+These results suggested that the samples from the patients that were overweight in the \gls{blca} data were similar in genotype with the samples from the patients that were obese in the CR data.
+Again, due to the fact that all of the metagenes lacked association with the patient \gls{bmi} and  \gls{bmi} status in all of the other \gls{icgc} cancer types, it was difficult to conclude whether the observed association of many of the obesity metagenes with the overweight group of the samples was truly reflective of the effect resulted from these metagenes.
 
-This was unexpected since all of the genetic signatures resulted from the  gene expression analysis between the non-obese group and the obese group of the samples in CR data, and yet the metagenes showed significant association with the overweight group rather than the obese group in the \gls{blca} data set.
-These results suggested that the samples that were overweight in the \gls{blca} data were similar in phenotype with the samples that were obese in CR data.
-Again, due to the fact that all of the metagenes lacked association with the sample \gls{bmi}or \gls{bmi} status in all other \gls{icgc} cancer types, it was difficult to conclude whether the observed association of many of the obesity metagenes with the overweight group of the samples was truly reflective of the effect resulted from these metagenes.
-
-Taken together, these results showed that, even though all of the obesity metagenes significantly associated with the sample \gls{bmi}/\gls{bmi} status in the original data where the genetic signatures were derived from, the metagenes were not generalisable in other cancer data sets.
-Furthermore, these results showed that the lack of association with sample \gls{bmi} and \gls{bmi} status were not due to other clinical variables in the CR data set.
-This raised a question of whether there was any obesity associated genetic signature that was common in multiple types of cancer.
+Taken together, these results showed that, even though all of the obesity metagenes significantly associated with the patient \gls{bmi} and \gls{bmi} status in the original data where the genetic signatures were derived from, none of the metagenes were not generalisable in other cancer data sets.
+Furthermore, these results showed that the lack of association with patient \gls{bmi} and \gls{bmi} status were not due to other clinical variables in the CR data set.
+This raised a question of whether there was any obesity associated genetic signature that was common in multiple types of cancer (\cref{sec:common_genes_across_multiple_cancer_types}).
 
@@
 	\centering
-	\caption{Statistics of all the obesity metagenes with the sample \gls{bmi}/\gls{bmi} status in the \gls{icgc} \gls{blca} cancer data}
+	\caption{Statistics of all the obesity metagenes with the patient \gls{bmi} and \gls{bmi} status in the \gls{icgc} \gls{blca} cancer data}
 	\label{tab:degmetablca}
@@
 	\includegraphics[page=5,width=0.45\linewidth]{results1/rawobsgene_ICGC}
-	\caption[Cr obesity metagene in the \acrshort{icgc} \acrshort{blca} data]{Heatmap, box plot and scatter plot showing the association of the Cr obesity metagene with the sample gene expression, \gls{bmi} and \gls{bmi} status from the \acrshort{icgc} \acrshort{blca} data, respectively.
+	\caption[Cr obesity metagene in the \acrshort{icgc} \acrshort{blca} data]{Heatmap, scatter plot and box plot showing the association of the Cr obesity metagene with the sample gene expression, patient \gls{bmi} and \gls{bmi} status, respectively, from the \acrshort{icgc} \acrshort{blca} data.
 	The results for other other metagenes and other \gls{icgc} cancer types are shown in \cref{sec:rest_of_the_cr_icgc_cancer_heatmap_results}.
@@
 
-It was clear that the obesity associated genetic signatures created from the CR data set were not consistent across different data sets.
+It was clear that the all of the obesity associated genetic signatures created from the CR data set were not significantly associated with patient \gls{bmi} across different \gls{icgc} cancer data sets.
 To determine whether there was any obesity associated genetic signature that was expressed in multiple different cancer types, gene expression analysis was carried out on each of the eight \gls{icgc} cancer types, and common genes were searched from the \glspl{deg} that were identified.
-All of the \gls{icgc} cancer data sets were normalised with voom (\cref{ssub:rna_seq_data}), which were then put through the gene expression analysis pipeline to identify the \glspl{deg} between the obese group of samples and non-obese group of samples (\cref{sec:gene_expression_analysis}).
-\cref{tab:icgcdegnum} summarised the number of genes found from each cancer type with p \textless{} 0.05.
+All of the \gls{icgc} cancer data sets were normalised with voom (\cref{ssub:rna_seq_data}), which were then put through the gene expression analysis pipeline to identify the \glspl{deg} between the obese and non-obese groups of samples (\cref{sec:gene_expression_analysis}).
+\cref{tab:icgcdegnum} summarised the number of \glspl{deg}  (unadjusted p \textless{} 0.05) found from each cancer type.
 
-\begin{table}[htbp]
+\begin{table}[tb]
 	\centering
@@
 
-There were 9695 unique \glspl{deg} across the eight cancer types, and these genes were checked for any commonalities across the different cancer types.
+There were 9695 unique \glspl{deg} across the eight different cancer types, and these genes were checked for any commonalities across the different cancer types.
 There were 7330 genes that were differentially expressed in one cancer type, 2024 genes differentially expressed in any two cancer types, 320 genes expressed in any three cancer types, and 21 genes expressed in any four cancer types.
@@
 Unfortunately, there were no \glspl{deg} expressed in all eight cancer types, which confirmed that there was no common genes that were differentially expressed between the samples that were obese and normal weight.
-When the results from \gls{icgc} gene expression analysis were compared with the simulation results, the number of \glspl{deg} found were statistically significant, as the numbers of genes identified exceeded the 95$^{th}$ percentile values for up to 4 cancer types.
+When the results from the \gls{icgc} gene expression analysis were compared with the simulation results, the number of \glspl{deg} found were statistically significant, as the numbers of genes identified exceeded the 95$^{th}$ percentile values for up to four cancer types.
 This result confirmed that the \glspl{deg} from the \gls{icgc} cancer data were not identified by chance.
-However, this result also showed that many of the genes identified in the gene expression analysis of the cancer types were observed by chance and that the majority of these genes were likely to be false positives.
+However, this also showed that many of the \glspl{deg} identified in the gene expression analysis of the cancer types were observed by chance, and that the majority of these genes were likely to be false positives.
 
-\begin{table}[htbp]
+\begin{table}[tb]
 	\centering
@@
 
-When the pathway enrichment analysis was carried out with the \gls{kegg} pathway data base, only the ``ABC transporter'' was significantly enriched (\gls{fdr}-adjusted p-value \textless{} 0.05) in the \gls{cesc} data, and no other pathways were enriched in any of the other \gls{icgc} cancer types (\cref{sec:pathways_significant_in_each_of_the_cancer_types}).
-With the Reactome data base, ``Phosphatase bond hydrolysis by NUDT proteins'' (in \gls{blca}) and ``Mitochondrial ABC transporters'' (in \gls{cesc}) were significantly enriched (\cref{sec:pathways_significant_in_each_of_the_cancer_types}).
-With the \gls{go} data base, there were 22 pathways significantly enriched in \gls{blca}, 17 pathways in \gls{cesc}, 3 pathways in \gls{kirp}, 21 pathways in \gls{read}, 10 pathways in \gls{skcm}, 14 pathways in \gls{ucec}, and no pathways were significant in the \gls{coad} and \gls{lihc} cancer types (\cref{sec:pathways_significant_in_each_of_the_cancer_types}).
+When the pathway enrichment analysis was carried out with the \gls{kegg} pathway database, only the ``ABC transporter'' was significantly enriched (\gls{fdr}-adjusted p-value \textless{} 0.05) in the \gls{cesc} data, and no other pathways were enriched in any of the other \gls{icgc} cancer types (\cref{sec:pathways_significant_in_each_of_the_cancer_types}).
+With the Reactome database, ``Phosphatase bond hydrolysis by NUDT proteins'' (in \gls{blca}) and ``Mitochondrial ABC transporters'' (in \gls{cesc}) were significantly enriched (\cref{sec:pathways_significant_in_each_of_the_cancer_types}).
+With the \gls{go} database, there were 22 pathways significantly enriched in \gls{blca}, 17 pathways in \gls{cesc}, 3 pathways in \gls{kirp}, 21 pathways in \gls{read}, 10 pathways in \gls{skcm}, 14 pathways in \gls{ucec}, and no pathways were significant in the \gls{coad} and \gls{lihc} cancer types (\cref{sec:pathways_significant_in_each_of_the_cancer_types}).
 
-Though there were many pathways identified as significantly enriched with the \gls{go} data base in each of the cancer types, there were no pathways that were common across the different cancer types.
-Furthermore, these pathways were not similar in terms of the biological activities involved; for example, pathways enriched in the \gls{blca} data suggested possible activation of the \gls{pi3k} pathway, but the pathways enriched in the \gls{read} data were mainly involved in \acrshort{rna} processing (\cref{sec:pathways_significant_in_each_of_the_cancer_types}).
+Though there were many pathways identified as significantly enriched with the \gls{go} database in each of the cancer types, there were no pathways that were common across all of the different cancer types.
+Furthermore, these pathways were not similar in terms of the biological activities involved.
+For example, pathways enriched in the \gls{blca} data suggested possible activation of the \gls{pi3k} pathway, but the pathways enriched in the \gls{read} data were mainly involved in \acrshort{rna} processing (\cref{sec:pathways_significant_in_each_of_the_cancer_types}).
 One interesting result from this enrichment analysis was that the ``ABC transport'' pathway was identified as significantly enriched in the \gls{cesc} data set by all three pathway databases, suggesting that the ``ABC transport'' pathway may be a core component in the \gls{cesc} tumour biology.
@@
 
-Since there were no pathways enriched in any of the cancer types when analysed individually, all of the \gls{icgc} cancer data sets were combined to see if there were any pathways enriched in the combined data set.
-The combined data set was generated in two ways: first combined data set was created by voom normalising each cancer data set separately, combine all of the normalised cancer data sets into one, then applied batch correction to the data set (\cref{sub:batch_correction}); and the second combined data set was created by combining the cancer data sets into one first, then the combined data was voom normalised, and finally batch correction was applied to obtain the final data set.
+Since there were no common pathway enriched in the \gls{icgc} cancer types when analysed individually, all of the \gls{icgc} cancer data sets were combined into a single data set to see if there were any pathways enriched in the combined data set.
+The combined data set was generated in two ways.
+The first combined data set was created by voom normalising each cancer data set separately, combine all of the normalised cancer data sets into one, then applied batch correction to the data set (\cref{sub:batch_correction}).
+The second combined data set was created by combining the cancer data sets into one big data set first, then the combined data was voom normalised, and finally batch correction was applied to obtain the final data set.
 Each of the combined data sets were analysed for enriched pathways, but there were no pathways significantly enriched in either data sets.
-These results suggested that there were no biological pathways associated with obesity in the \gls{icgc} cancer data sets.
+These results suggested that there were no common biological pathways associated with obesity in the \gls{icgc} cancer data sets.

### 2017-03-25 thesis/results/results2.tex
finished editing up to the discussion

@@
 
-In this chapter, the underlying biological mechanism of the obesity associated signatures from CR data were investigated.
+In this chapter, the underlying biological mechanism of the obesity associated signatures from the CR data were investigated.
 In doing so, the pathway genetic signatures from the \citet{Gatza2010a} study (GT) were utilised to determine which biological pathways the obesity associated genetic signatures were most similar to.
@@
 
-Pathway associated genetic signatures from the \citet{Gatza2010a} study were first examined in the GT data to make sure the results from their paper was reproducible.
-
-\subsection{Ranking and normalisation methods}
+\subsection{Ranking and normalisation methods for pathway metagenes}
 \label{sub:ranking_and_normalisation_methods}
 
-In the \citet{Gatza2010a} study, their data comprised of samples from different data sets from other studies, \gls{mas}-normalised and the metagene scores were transformed to a scale between 0 to 1 using a probit function.
+In the \citet{Gatza2010a} study, their data comprised of samples from different studies which were aggregated into a single data set that was \gls{mas} normalised, and the metagene scores were transformed to a scale between 0 to 1 using a probit function.
 However, the analyses so far have used \gls{rma} normalisation method and ranked based on the number of samples present in the data (fractional ranking; \cref{sub:ranking_of_the_metagene_scores}).
@@
 Results shown in \cref{sec:ranking_method_for_the_pathway_associated_genetic_signatures} clarified that there was no significant difference in the ranking approaches used.
-Correlation and scatter plots of the different combinations of transformation matrices (derived from either the \gls{rma}- or \gls{mas}-normalised GT data) with the  GT data normalised with the different methods were calculated and plotted.
-From these results, it was clear the normalisation methods used had the most significant effect on the resulting metagenes, rather than the transformation matrices themselves (\cref{sec:normalisation_method_for_pathway_associated_genetic_signature_transformation_matrices}).
-This meant that all of the data sets had to be normalised with a single normalisation method, so the \gls{rma} normalisation method was used as this method was more reliable than the \gls{mas} method \citep{Irizarry2003}.
-Therefore, the GT data set was batch corrected first (\cref{sub:batch_correction}), then normalised with the \gls{rma} method and the metagene scores were ranked with fractional ranking (\cref{sub:ranking_of_the_metagene_scores}).
 
-One thing noted from these scatter plots was that there were some pathway signatures that were more variable than the others (\cref{sec:normalisation_method_for_pathway_associated_genetic_signature_transformation_matrices}).
+Scatter plots were plotted and the correlations were calculated for each of the GT pathway metagenes derived from either \gls{rma} or \gls{mas} normalised GT data (\cref{sec:normalisation_method_for_pathway_associated_genetic_signature_transformation_matrices}).
+The GT pathway metagenes were generated by using the transformation matrices that were derived from either the \gls{rma} or \gls{mas} normalised GT data.
+From these results, it was clear the normalisation methods used on the data set had the most significant effect on the resulting metagenes, rather than the normalisation methods used to generate the transformation matrices (\cref{sec:normalisation_method_for_pathway_associated_genetic_signature_transformation_matrices}).
+This meant that all of the data sets had to be normalised with a single normalisation method, so the \gls{rma} was chosen as the normalisation method to be used as this method was more reliable than the \gls{mas} method \citep{Irizarry2003}.
+
+Another thing noted from these scatter plots was that there were some pathway signatures that were more variable than the others (\cref{sec:normalisation_method_for_pathway_associated_genetic_signature_transformation_matrices}).
 As an example, the \gls{tgfb} metagenes were significantly more dispersed compared to the \gls{pr} metagenes, even though both of these metagenes were generated in the same data set through similar processes (\cref{fig:gt_rma_vs_mas}).
@@
 This result provided evidence that some of the pathway genetic signatures from the \citet{Gatza2010a} study were less variable across different data sets than the others.
+Taken together, it was decided that the GT data set was batch corrected first (\cref{sub:batch_correction}), then normalised with the \gls{rma} method, and the metagene scores were ranked with fractional ranking (\cref{sub:ranking_of_the_metagene_scores}).
 
@@
 
-Before GT pathway metagenes were compared with the obesity associated metagenes, the direction of the GT pathway metagenes had to be checked to make sure the metagenes were in the ``correct'' direction (\cref{sub:metagene_direction}).
+Before GT pathway metagenes were compared with the obesity associated metagenes, the direction of the GT pathway metagenes had to be checked to make sure that the metagenes were in the ``correct'' direction (\cref{sub:metagene_direction}).
 The correlation of all the pathway metagenes with one another were plotted as a heatmap in \cref{fig:gatza_meta_dir}.
@@
 
-To see whether the directionality of the GT pathway metagenes were being transferred across to other data sets properly, pathway metagenes were generated in other data sets with the transformation matrices and the groupings of the metagenes were examined.
-CR, FM and \gls{nzbc} data were \gls{rma}-normalised and transformation matrices of the pathway genetic signatures (derived from GT data set) were applied to the data sets.
-The metagenes were plotted in a heatmap (shown in \cref{app:b}), which showed similar groupings  as seen in \cref{fig:gatza_meta_dir}.
-This result confirmed that the pathway metagenes were acting similarly in all the data sets in which the transformation matrices have been applied to.
+To see whether the directionality of the GT pathway metagenes were being transferred across to other data sets properly, pathway metagenes were generated in other data sets with the same transformation matrices and the groupings of the metagenes were examined.
+CR, FM and \gls{nzbc} data were \gls{rma} normalised and transformation matrices of the pathway genetic signatures (derived from the GT data set) were applied to the data sets.
+The metagenes were plotted in a heatmap (shown in \cref{sec:correlation_of_the_gt_pathway_metagenes_in_other_cancer_data_sets}), which showed similar groupings  as seen in \cref{fig:gatza_meta_dir}.
+This result confirmed that the pathway metagenes were acting similarly in all of the data sets in which the transformation matrices have been applied to.
 
@@
 As in \cref{sec:pathway_associated_genetic_signatures_from_gatza2010a_study}, all data sets were normalised with the \gls{rma} method and metagenes were ranked with fractional ranking.
-Transformation matrices for the obesity associated genetic signatures were made in the CR data set and the pathway associated genetic signatures were made in the GT data set.
-Metagenes for all of the obesity and pathway associated genetic signatures were generated in all of the data sets with \gls{svd} and transformation matrices.
+TMs for the obesity associated genetic signatures were made in the CR data set and the TMs for pathway associated genetic signatures were made in the GT data set.
+Metagenes for all of the obesity and pathway associated genetic signatures were generated in all of the data sets with \gls{svd} and TMs.
 The Spearman correlation of the \gls{svd}-generated and TM-generated metagene scores were calculated for each genetic signatures (\cref{tab:svd_vs_tm_path,tab:svd_vs_tm_obs}).
@@
 
-In \cref{tab:svd_vs_tm_path,tab:svd_vs_tm_obs}, the obesity and pathway associated genetic signatures had a correlation of 1 in CR and GT data sets respectively, as the transformation matrices were generated in those data sets.
+In \cref{tab:svd_vs_tm_path,tab:svd_vs_tm_obs}, the obesity and pathway associated genetic signatures had a correlation of 1 in CR and GT data sets respectively, as the obesity and pathway TMs were generated in the CR and GT data sets, respectively.
 The correlations of the obesity and pathway metagenes were variable across different data sets, which was as expected since all the data sets were different from one another.
-However, what was unexpected was the fact that there were some genetic signatures that were highly correlated in all of the data sets, whereas the metagene correlation of other signatures were highly variable across different data sets.
-For example, the \gls{bcat} pathway metagene was highly correlated in all of the data sets (\textgreater{} 0.98), whereas the \gls{stat3} pathway metagene was variable across different data sets, ranging from 0.1902 to 0.9159 (\cref{tab:svd_vs_tm_path}).
+However, what was unexpected was the fact that there were some pathway genetic signatures that were highly correlated in all of the data sets, and there were others that varied considerably across different data sets.
+For example, the \gls{bcat} pathway metagene was highly correlated in all of the data sets (\textgreater{} 0.98), whereas the \gls{stat3} pathway metagene was variable across different data sets, ranging from 0.1902 to 0.9159 (\cref{tab:svd_vs_tm_path}; \cref{sec:comparison_of_the_correlation_of_svd_and_tm_generated_pathway_metagenes}).
 The fact that some pathway associated metagenes were consistent across different data sets suggested that some of these genetic signatures were reliable and did not depend on the data set the transformation matrices were derived from.
 On the other hand, the genetic signatures that were not consistent across the data sets were likely to be dependent on the data set in which they were derived from, and therefore the transformation matrices for these signatures must be derived from the GT data set.
-This also showed that the tumour biology may be different for different data sets, likely due to the difference in the patient cohorts that were selected for each of the studies.
+This also showed that the tumour biology may be different across the data sets, likely due to the difference in the patient cohorts that were selected for each of the studies.
 
@@
 This suggested that the obesity associated genetic signatures were relatively consistent across all data sets, and the transformation matrices for these signatures could be made in any of the data sets.
-For simplicity, it was decided that the transformation matrices for all the genetic  signatures would be made in the GT data set, since there were some pathway associated signatures that were specific to the GT data, but none of the obesity associated genetic signatures were specific to a data set.
+For simplicity, it was decided that the transformation matrices for all the genetic  signatures would be made in the GT data set, since there were some pathway associated signatures that were specific to the GT data.
 
@@
 
-In \cref{fig:gatza_allmeta}, all of the obesity associated metagenes from the CR data set cluster together into a group, but none of the pathway associated metagenes positively correlated strongly with the obesity metagenes.
+In \cref{fig:gatza_allmeta}, all of the obesity associated metagenes from the CR data set cluster together into a group, but none of the pathway associated metagenes had  a strong positive correlation with the obesity metagenes.
 Though \gls{her2} and p63 pathway metagenes were grouped next to the cluster of obesity metagenes, the similarity was at most about 0.6  (Pearson correlation; \cref{fig:gatza_allmeta}).
-However, there seem to be a strong negative correlation with some of the pathway associated genetic signatures, such as the Akt, \gls{er}, \gls{egfr}, \gls{pr}, p53, Src and \gls{tgfb} pathways, and this strong negative correlation was observed in the other data sets as well (\cref{sec:correlation_of_the_pathway_and_obesity_metagenes_in_other_cancer_data_sets}).
+With that said, these obesity associated metagenes seem to have a strong negative correlation with some of the pathway associated genetic signatures, such as the Akt, \gls{er}, \gls{egfr}, \gls{pr}, p53, Src and \gls{tgfb} pathways, and this strong negative correlation was observed in the other data sets as well (\cref{sec:correlation_of_the_pathway_and_obesity_metagenes_in_other_cancer_data_sets}).
 This suggests that the obesity associated genetic signatures from the CR data may be related to these pathway signatures, rather than the obesity phenotype of the patients.
@@
 FM obesity associated metagene did not cluster together with the obesity metagenes from the CR data set, nor with any of the other pathway metagenes.
-The fact that the FM metagene did not cluster with any of the CR metagenes showed that the nature of the FM metagene was different from those metagenes found in the CR data set.
+The fact that the FM metagene did not cluster with any of the CR metagenes showed that the signals captures by the FM metagene was different from those metagenes found in the CR data set.
 Furthermore, the FM metagene did not cluster with the Akt pathway in the heatmap, even though \citet{Fuentes-Mattei2014} provided clear evidence that their metagene specifically activated the Akt/\gls{mtor} signalling pathway in a mouse model.
@@
 
-The results in this section suggested that none of the obesity associated genetic signatures were not positively correlated with to any of the pathway associated genetic signatures from the \citet{Gatza2010a} study.
+The results in this section suggested that none of the obesity associated genetic signatures were positively correlated with any of the pathway associated genetic signatures from the \citet{Gatza2010a} study.
 However, some pathway signatures were negatively correlated with the obesity associated genetic signatures, which meant that the obesity signatures were likely capturing the signals from these signatures  rather than the biological signals from the obesity phenotype of the patient.
@@
 
-To confirm that the obesity associated metagenes were not related to any of the biological pathway signatures from the \citet{Gatza2010a} study, linear models were created to predict the obesity metagene scores with the pathway metagene scores.
+To confirm the biological relationship between the obesity associated metagenes with the pathway associated metagenes from the \citet{Gatza2010a} study, linear models were created to predict the obesity metagene scores with the pathway metagene scores.
 If any of the pathway metagenes were significant in the linear model, this would provide evidence that the significant pathways in the linear model were related to the obesity metagene.
@@
 
-For each of the obesity metagenes identified, seven linear models were created: \gls{bmi}-only, \gls{bmi} status-only, \gls{bmi} and \gls{bmi} status, pathway metagenes-only, and all of the combinations of pathway metagenes with \gls{bmi} and/or \gls{bmi} status.
+For each of the obesity metagenes identified, seven linear models were created: \gls{bmi}-only, \gls{bmi} status-only, \gls{bmi} and \gls{bmi} status, pathway metagenes-only, and the combinations of pathway metagenes with \gls{bmi} and/or \gls{bmi} status.
 The only variables that showed significance in any of the models that predicted the obesity metagenes taken from the CR data set were the \gls{pr} metagene and whether the patients were obese or not (\cref{tab:lm_sig_var}).
@@
 On the other hand, all of the predicted metagenes generated from the models that used the pathway metagene scores showed significant association with the original metagene scores (\cref{fig:predict_cris_cris}; \cref{sec:summary_of_the_linear_models_in_nzbc_data}).
-However, the $R^2$ values for all of the predicted metagenes against the true metagenes were very low ($R^2$ \textless{} 0.4).
+However, the $R^2$ values for all of the predicted metagenes against the original metagenes were very low ($R^2$ \textless{} 0.4).
 All of the plots in \cref{fig:predict_cris_cris,sec:summary_of_the_linear_models_in_nzbc_data} showed clearly that the data points in the scatter plots were highly variable, suggesting that even though the association between the predicted and the true metagene scores were statistically significant, the variables in the linear models may not be strongly associated to the obesity metagene.
@@
 	Only the results from the models used to predict the Cr metagene scores are shown.
-	The summary of the statistics for the other obesity metagenes in the \gls{nzbc} data are shown in \cref{sec:summary_statistics_of_the_predicted_obesity_metagenes_with_sample_bmi_bmi_status_in_nzbc_and_cr_data}.
+	The summary of the statistics for the other obesity metagene predictions from the \gls{nzbc} data are shown in \cref{sec:summary_statistics_of_the_predicted_obesity_metagenes_with_sample_bmi_bmi_status_in_nzbc_and_cr_data}.
 	P-values and $R^2$-value are as described in previous figures.
@@
 One possible explanation for this was that these \gls{bmi}-based models were overfitted in the \gls{nzbc} data set, and was not generalisable in the CR data set, and thus produced negative correlations.
-In fact, it was evident from the metagene analyses in \cref{cha:obesity_genetic_signatures_and_cancer} that these obesity associated genetic signatures were not generalisable with the patient \gls{bmi} in different data sets.
-% An alternative explanation was that the obesity metagenes  were not generalisable in different data sets, and thus the models were predicting the obesity metagenes relatively well in \gls{nzbc} data, but not in the CR data.
+In fact, it was evident from the metagene analyses in \cref{cha:obesity_genetic_signatures_and_cancer} that these obesity associated genetic signatures were not generalisable with the patient \gls{bmi} in different data sets, and overfitting of the model in the \gls{nzbc} data was the likely cause of this negative correlation.
 
@@
 	Only the results from the models used to predict the Cr metagene scores are shown.
-	The summary of the statistics for the other obesity metagenes in the CR data are shown in \cref{sec:summary_statistics_of_the_predicted_obesity_metagenes_with_sample_bmi_bmi_status_in_nzbc_and_cr_data}.
+	The summary of the statistics for the other obesity metagene predictions from the CR data are shown in \cref{sec:summary_statistics_of_the_predicted_obesity_metagenes_with_sample_bmi_bmi_status_in_nzbc_and_cr_data}.
 	P-values and $R^2$-value are as described in previous figures.
@@
 Lastly, since only the six most ``consistent'' pathway metagenes were picked from a total of eighteen pathway metagene scores, linear models were constructed with all eighteen pathway metagenes in a stepwise fashion.
-Step wise construction of a linear model is where the variables for a model are selected computationally depending on the contribution of the variable to the resulting linear model, and thus contains only the variables that ``matter'' to the final linear model \cref{sec:obesity_metagene_prediction_with_pathway_metagenes}.
+Step wise construction of a linear model is where the variables for a model are selected computationally depending on the contribution of the variable to the resulting linear model, and thus contain only the variables that ``matter'' to the final linear model \cref{sec:obesity_metagene_prediction_with_pathway_metagenes}.
 Patient \gls{bmi}, \gls{bmi} status and all of the pathway metagenes (Akt, \gls{bcat}, E2F1, \gls{egfr}, \gls{er}, \gls{her2}, \gls{ifna}, \gls{ifny}, Myc, p53, p63, \gls{pi3k}, \gls{pr}, Ras, \gls{stat3}, Src, \gls{tgfb} and \gls{tnfa}) were used to generate the stepwise linear model for each of the obesity metagenes.
@@
 
+These models were used to predict the obesity metagenes in the \gls{nzbc} data to test the performance of the models.
+All of the models were able to make statistically significant prediction of the original obesity metagenes in the \gls{nzbc} data set (\cref{fig:stepwise_cris}).
+Unlike the previous models, the stepwise models showed very high $R^2$ values, where most predictions had $R^2$ \textgreater{} 0.80.
+The statistical significance and the strong association of the predicted metagenes were observed in the CR data set as well (\cref{fig:stepwise_cr}).
+Though the $R^2$ values were slightly lower than those from the \gls{nzbc} data set, the $R^2$ values from the CR data predictions were more promising than the predictions made by the non-stepwise linear models.
+
+\begin{ThreePartTable}
+	\begin{TableNotes}
+		\begin{footnotesize}
+		\item [1] All values in bold are statistically significant (p \textless{} 0.05).
+		\end{footnotesize}
+	\end{TableNotes}
 	\begin{longtable}{llr{\bfseries}S}
@@
 			\hline
-		Cr         & Akt         & -0.3415   & \bfseries \num{4.43d-8}\\
-				   & \gls{egfr}  & 0.3385   & \bfseries \num{1.94d-05}\\
-				   & \gls{er}    & -0.2024  & \bfseries 0.004190\\
-				   & \gls{ifny}  & 0.1223   & 0.08700\\
-				   & \gls{pr}    & -0.3392  & \bfseries \num{2.09d-5}\\
-				   & Ras         & -0.2261  & \bfseries 0.001930\\
-				   & Src         & -0.1145  & 0.08177\\
-				   & \gls{tgfb}  & 0.5564   & \bfseries \textless{} \num{2d-16}\\
-				   & \gls{tnfa}  & -0.2550  & \bfseries 0.006160\\
+			Cr      & Src         & -0.1900  & \bfseries 0.005150\tnote{1}       \\
+					& \gls{egfr}  & 0.1451   & \bfseries 0.002780                \\
+					& \gls{tgfb}  & 0.5105   & \bfseries \num{1.17d-15}          \\
+					& Akt         & -0.2821  & \bfseries \num{7.08d-06}          \\
+					& \gls{ifna}  & -0.06968 & \bfseries 0.02986                 \\
 			\hline
-		CrOl       & Overweight  & 0.04092  & 0.07487\\
-				   & Obese       & -0.006381 & 0.7561\\
-				   & Akt         & -0.3567 & \bfseries \num{1.89d-07}\\
-				   & E2F1        & 0.1602  & \bfseries 0.005411\\
-				   & \gls{egfr}  & 0.1486  & 0.06268\\
-				   & p53         & 0.3366  & \bfseries \num{4.35d-05}\\
-				   & \gls{pr}    & -0.2254 & \bfseries 0.0006390\\
-				   & Src         & -0.4288 & \bfseries \num{5.85d-10}\\
-				   & \gls{stat3} & -0.1241 & \bfseries 0.009658\\
+			CrOl    & Akt         & -0.3134  & \bfseries \num{1.76d-06}          \\
+					& Src         & -0.5386  & \bfseries \textless{} \num{2d-16} \\
+					& p53         & 0.3372   & \bfseries \num{9.18d-07}          \\
+					& \gls{pr}    & -0.1115  & \bfseries 0.02760                 \\
 			\hline
-		Res        & Akt         & 0.2461   & \bfseries 0.0002890\\
-				   & \gls{egfr}  & -0.4683  & \bfseries \num{2.16d-05}\\
-				   & \gls{er}    & 0.1840   & 0.06145\\
-				   & \gls{her2}  & 0.1384   & 0.1182\\
-				   & \gls{ifny}  & -0.2102  & \bfseries 0.04408\\
-				   & \gls{pr}    & 0.4120   & \bfseries 0.0003000\\
-				   & Ras         & 0.2651   & \bfseries 0.01047\\
-				   & \gls{tgfb}  & -0.5808  & \bfseries \num{7.64d-11}\\
-				   & \gls{tnfa}  & 0.4220   & \bfseries 0.001834\\
+			Res     & \gls{tgfb}  & -0.5895  & \bfseries \num{1.53d-15}          \\
+					& \gls{egfr}  & -0.3597  & \bfseries 0.0005390               \\
+					& \gls{tnfa}  & 0.1604   & \bfseries 0.0006970               \\
+					& \gls{bcat}  & -0.2141  & \bfseries 0.01157                 \\
+					& Akt         & 0.1527   & \bfseries 0.02206                 \\
 			\hline
-		ResOl      & Akt         & -0.3360  & \bfseries 0.0002940\\
-				   & \gls{bcat}  & -0.1987  & 0.1390\\
-				   & E2F1        & 0.3606   & \bfseries 0.009325\\
-				   & \gls{egfr}  & 0.2342   & 0.05687\\
-				   & \gls{her2}  & -0.1365  & 0.1217\\
-				   & p53         & 0.3875   & \bfseries 0.0005190\\
-				   & \gls{pi3k}  & -0.2188  & 0.1487\\
-				   & \gls{pr}    & -0.3223  & \bfseries 0.001656\\
-				   & Ras         & -0.2898  & \bfseries 0.03398\\
-				   & Src         & -0.4078  & \bfseries \num{5.26d-06}\\
-				   & \gls{stat3} & -0.1145  & 0.1256\\
-				   & \gls{tnfa}  & -0.1144  & 0.1170\\
+			ResOl   & Akt         & -0.2802  & \bfseries 0.0001420               \\
+					& Src         & -0.5214  & \bfseries \num{2.65d-15}          \\
+					& p53         & 0.2861   & \bfseries \num{5.31d-07}          \\
 			\hline
-		Ca         & Akt         & -0.2624  & \bfseries \num{7.04d-06}\\
-				   & \gls{egfr}  & 0.3631   & \bfseries 0.0001390\\
-				   & \gls{er}    & -0.1861  & \bfseries 0.03077\\
-				   & \gls{ifny}  & 0.2212   & \bfseries 0.01303\\
-				   & \gls{pr}    & -0.4291  & \bfseries \num{1.07d-05}\\
-				   & Ras         & -0.2812  & \bfseries 0.0008980\\
-				   & \gls{tgfb}  & 0.7614   & \bfseries \textless{} \num{2d-16}\\
-				   & \gls{tnfa}  & -0.3320  & \bfseries 0.003919\\
+			Ca      & \gls{tgfb}  & 0.7470   & \bfseries \textless{} \num{2d-16} \\
+					& Akt         & -0.2985  & \bfseries \num{2.83d-10}          \\
 			\hline
-		CaOl       & Overweight  & 0.07613   & \bfseries 0.006875\\
-				   & Obese       & 0.01408   & 0.5646\\
-				   & Akt         & -0.4473  & \bfseries \num{2.43d-07}\\
-				   & \gls{bcat}  & -0.1346  & 0.1403\\
-				   & E2F1        & 0.3386   & \bfseries 0.001249\\
-				   & p53         & 0.2878   & \bfseries 0.0009960\\
-				   & \gls{pr}    & -0.2826  & \bfseries \num{2.18d-05}\\
-				   & Src         & -0.4415  & \bfseries \num{2.02d-08}\\
-				   & \gls{tnfa}  & -0.1422  & \bfseries 0.01390\\
+			CaOl    & Src         & -0.5595  & \bfseries \num{3.88d-16}          \\
+					& Akt         & -0.4331  & \bfseries \num{2.06d-11}          \\
 			\hline
-		CaRes      & \gls{bcat}  & 0.2690   & 0.05238  \\
-				   & E2F1        & -0.2212  & 0.08848  \\
-				   & \gls{egfr}  & 0.4440   & \bfseries \num{6.67d-05 }\\
-				   & \gls{her2}  & -0.2030  & \bfseries 0.04615  \\
-				   & \gls{ifny}  & 0.3287   & \bfseries 0.005740  \\
-				   & \gls{pr}    & -0.2876  & \bfseries 0.001040  \\
-				   & Src         & -0.1799  & 0.05404  \\
-				   & \gls{tgfb}  & 0.5290   & \bfseries \num{2.15d-06 }\\
-				   & \gls{tnfa}  & -0.3868  & \bfseries 0.005640  \\
+			CaRes   & \gls{tgfb}  & 0.7843   & \bfseries \textless{} \num{2d-16} \\
+					& \gls{egfr}  & 0.2418   & \bfseries \num{2.19d-06}          \\
 			\hline
-		CaResOl    & \gls{bmi}   & 0.004310  & 0.1086\\
-				   & Overweight  & 0.03309  & 0.3064\\
-				   & Obese       & -0.06226 & 0.1758\\
-				   & Akt         & -0.3097 & \bfseries 0.0003390\\
-				   & \gls{bcat}  & -0.1982 & 0.1405\\
-				   & E2F1        & 0.2705  & \bfseries 0.01145\\
-				   & \gls{egfr}  & 0.1712  & 0.1137\\
-				   & Myc         & 0.1520  & 0.05699\\
-				   & p53         & 0.2949  & \bfseries 0.003734\\
-				   & \gls{pr}    & -0.2309 & \bfseries 0.0008230\\
-				   & Ras         & -0.2708 & \bfseries 0.04479\\
-				   & Src         & -0.5126 & \bfseries \num{2.72d-08}\\
-				   & \gls{tnfa}  & -0.1033 & 0.08345\\
+			CaResOl & Src         & -0.4553  & \bfseries \num{1.57d-07}          \\
+					& \gls{egfr}  & 0.2776   & \bfseries 0.0009570               \\
+					& Akt         & -0.3001  & \bfseries 0.0001640               \\
+					& E2F1        & 0.2484   & \bfseries 0.001300                \\
+					& Myc         & 0.1584   & \bfseries 0.025220                \\
 			\hline
-		Or         & \gls{bmi}   & 0.002696  & 0.1441\\
-				   & Overweight  & 0.01052  & 0.62856\\
-				   & Obese       & -0.06102 & 0.05542\\
-				   & Akt         & -0.2384 & \bfseries \num{4.93d-05}\\
-				   & E2F1        & 0.1851  & \bfseries 0.02754\\
-				   & \gls{egfr}  & 0.1472  & 0.07423\\
-				   & \gls{her2}  & -0.08680 & 0.1407\\
-				   & p53         & 0.4164  & \bfseries \num{5.83d-08}\\
-				   & \gls{pi3k}  & -0.2162 & \bfseries 0.03528\\
-				   & \gls{pr}    & -0.1991 & \bfseries 0.002530\\
-				   & Ras         & -0.2005 & \bfseries 0.008980\\
-				   & Src         & -0.5438 & \bfseries \num{2.10d-15}\\
-				   & \gls{stat3} & -0.0934 & \bfseries 0.02749\\
+			Or      & Src         & -0.5165  & \bfseries \textless{} \num{2d-16} \\
+					& p53         & 0.3072   & \bfseries \num{2.60d-12}          \\
+					& Akt         & -0.3006  & \bfseries 1.440e-07               \\
+					& E2F1        & 0.09684  & \bfseries 0.01070                 \\
+					& Obese       & -0.02190 & 0.2166                            \\
+					& Overweight  & 0.03099  & 0.1223                            \\
 			\hline
-		FM         & \gls{bmi}   & 0.01109  & \bfseries 0.03623\\
-				   & Overweight  & -0.1538 & \bfseries 0.01428\\
-				   & Obese       & -0.2125 & \bfseries 0.01738\\
-				   & p53         & 0.4608  & \bfseries 0.002512\\
-				   & \gls{pi3k}  & -0.4262 & \bfseries 0.0001180\\
-				   & \gls{pr}    & -0.3731 & \bfseries 0.02074\\
-				   & \gls{stat3} & 0.2487  & \bfseries 0.01907\\
-				   & \gls{tgfb}  & 0.1529  & 0.1289\\
+			FM      & p63         & 0.3507   & \bfseries 0.0006760               \\
+					& \gls{stat3} & 0.2969   & \bfseries 0.004191                \\
+					& p53         & 0.5368   & \bfseries 0.0008240               \\
+					& \gls{pr}    & -0.3462  & \bfseries 0.034070                \\
 			\hline
 			\hline
+			\insertTableNotes
 	\end{longtable}
+\end{ThreePartTable}
 
-These models were used to predict the obesity metagenes in the \gls{nzbc} data to test the performance of the models.
-All of the models were able to make statistically significant prediction of the original obesity metagenes in the \gls{nzbc} data set (\cref{fig:stepwise_cris}).
-Unlike the previous models, the stepwise models showed very high $R^2$ values, where most predictions had $R^2$ \textgreater{} 0.85.
-The statistical significance and the strong association of the predicted metagenes were observed in the CR data set as well (\cref{fig:stepwise_cr}).
-Though the $R^2$ values were slightly lower than those from the \gls{nzbc} data set, the $R^2$ values from the CR data predictions were more promising than the predictions made by the non-stepwise linear models.
-
-With that said, some of the predicted obesity metagenes were negatively correlated with the original metagene values in the CR data (\cref{fig:stepwise_cr}).
-Again, this could have been due to the model being overfitted in the \gls{nzbc} data, and thus made a wrong prediction about the obesity metagenes.
-Nevertheless, the stepwise linear models were able to predict the obesity associated metagenes better than the models constructed from the models made in \cref{sub:linear_model_prediction_in_the_nzbc_and_cr_data_sets}.
+Clearly, the stepwise linear models were able to predict the obesity associated metagenes better than the models created in \cref{sub:linear_model_prediction_in_the_nzbc_and_cr_data_sets}.
+This could be due to the fact that many of the stepwise linear models included many pathways that had a strong negative correlation with the obesity metagenes (Akt, \gls{er}, \gls{egfr}, \gls{pr}, p53, Src and \gls{tgfb} pathways; \cref{fig:gatza_allmeta}).
+These pathway metagenes may have been better predictors of the obesity metagenes than the six pathways that were ``consistent'' across the data sets, and therefore generated better predictions than those linear models.
+With that said, some of the predicted obesity metagenes from the stepwise linear models were negatively correlated with the original metagene values in the CR data (\cref{fig:stepwise_cr}).
+Again, this could have been due to the model being overfitted in the \gls{nzbc} data, and thus made an incorrect prediction about the obesity metagenes in the CR data.
 
@@
 	\centering
-	\includegraphics[page=1,width=0.32\linewidth]{results2/stepwise_cris_cris}
-	\includegraphics[page=2,width=0.32\linewidth]{results2/stepwise_cris_cris}
-	\includegraphics[page=3,width=0.32\linewidth]{results2/stepwise_cris_cris}
-	\includegraphics[page=4,width=0.32\linewidth]{results2/stepwise_cris_cris}
-	\includegraphics[page=5,width=0.32\linewidth]{results2/stepwise_cris_cris}
-	\includegraphics[page=6,width=0.32\linewidth]{results2/stepwise_cris_cris}
-	\includegraphics[page=7,width=0.32\linewidth]{results2/stepwise_cris_cris}
-	\includegraphics[page=8,width=0.32\linewidth]{results2/stepwise_cris_cris}
-	\includegraphics[page=9,width=0.32\linewidth]{results2/stepwise_cris_cris}
-	\includegraphics[page=10,width=0.32\linewidth]{results2/stepwise_cris_cris}
-	\caption[Comparison of all the obesity metagene scores predicted from the stepwise linear models with the obesity metagene scores from the \gls{nzbc} data]{Scatter plots comparing all of the obesity metagene scores predicted by the stepwise linear models constructed from the \gls{nzbc} data, with the corresponding obesity metagene scores from the \gls{nzbc} data.
+	\includegraphics[page=1,width=0.32\linewidth]{results2/stepwise_cris_both_bic}
+	\includegraphics[page=2,width=0.32\linewidth]{results2/stepwise_cris_both_bic}
+	\includegraphics[page=3,width=0.32\linewidth]{results2/stepwise_cris_both_bic}
+	\includegraphics[page=4,width=0.32\linewidth]{results2/stepwise_cris_both_bic}
+	\includegraphics[page=5,width=0.32\linewidth]{results2/stepwise_cris_both_bic}
+	\includegraphics[page=6,width=0.32\linewidth]{results2/stepwise_cris_both_bic}
+	\includegraphics[page=7,width=0.32\linewidth]{results2/stepwise_cris_both_bic}
+	\includegraphics[page=8,width=0.32\linewidth]{results2/stepwise_cris_both_bic}
+	\includegraphics[page=9,width=0.32\linewidth]{results2/stepwise_cris_both_bic}
+	\includegraphics[page=10,width=0.32\linewidth]{results2/stepwise_cris_both_bic}
+	\caption[Comparison of all the obesity metagene scores predicted from the stepwise linear models with the original obesity metagene scores from the \gls{nzbc} data]{Scatter plots comparing all of the obesity metagene scores predicted by the stepwise linear models constructed from the \gls{nzbc} data, with the corresponding obesity metagene scores from the \gls{nzbc} data.
 	P-values and $R^2$-value are as described in previous figures.}
@@
 	\centering
-	\includegraphics[page=1,width=0.32\linewidth]{results2/stepwise_cr_cris}
-	\includegraphics[page=2,width=0.32\linewidth]{results2/stepwise_cr_cris}
-	\includegraphics[page=3,width=0.32\linewidth]{results2/stepwise_cr_cris}
-	\includegraphics[page=4,width=0.32\linewidth]{results2/stepwise_cr_cris}
-	\includegraphics[page=5,width=0.32\linewidth]{results2/stepwise_cr_cris}
-	\includegraphics[page=6,width=0.32\linewidth]{results2/stepwise_cr_cris}
-	\includegraphics[page=7,width=0.32\linewidth]{results2/stepwise_cr_cris}
-	\includegraphics[page=8,width=0.32\linewidth]{results2/stepwise_cr_cris}
-	\includegraphics[page=9,width=0.32\linewidth]{results2/stepwise_cr_cris}
-	\includegraphics[page=10,width=0.32\linewidth]{results2/stepwise_cr_cris}
-	\caption[Comparison of all the obesity metagene scores predicted from the stepwise linear models with the obesity metagene scores from the CR data]{Scatter plots comparing all of the obesity metagene scores predicted by the stepwise linear models constructed from the \gls{nzbc} data, with the corresponding obesity metagene scores from the CR data.
+	\includegraphics[page=1,width=0.32\linewidth]{results2/stepwise_cr_both_bic}
+	\includegraphics[page=2,width=0.32\linewidth]{results2/stepwise_cr_both_bic}
+	\includegraphics[page=3,width=0.32\linewidth]{results2/stepwise_cr_both_bic}
+	\includegraphics[page=4,width=0.32\linewidth]{results2/stepwise_cr_both_bic}
+	\includegraphics[page=5,width=0.32\linewidth]{results2/stepwise_cr_both_bic}
+	\includegraphics[page=6,width=0.32\linewidth]{results2/stepwise_cr_both_bic}
+	\includegraphics[page=7,width=0.32\linewidth]{results2/stepwise_cr_both_bic}
+	\includegraphics[page=8,width=0.32\linewidth]{results2/stepwise_cr_both_bic}
+	\includegraphics[page=9,width=0.32\linewidth]{results2/stepwise_cr_both_bic}
+	\includegraphics[page=10,width=0.32\linewidth]{results2/stepwise_cr_both_bic}
+	\caption[Comparison of all the obesity metagene scores predicted from the stepwise linear models with the original obesity metagene scores from the CR data]{Scatter plots comparing all of the obesity metagene scores predicted by the stepwise linear models constructed from the \gls{nzbc} data, with the corresponding obesity metagene scores from the CR data.
 	P-values and $R^2$-value are as described in previous figures.}
@@
 Taken together, these results suggested that the variables included in these stepwise models were relevant to the biology of the obesity metagenes identified in the CR and FM data sets.
-However, there were no single common pathway that was present in all of the models; although Akt, \gls{egfr} and \gls{pr} pathway metagenes were frequently included in many of the stepwise models.
+However, there were no single common pathway that was present in all of the models; although Akt, \gls{egfr} and Src pathway metagenes were frequently included in many of the stepwise models.
 The role of the pathways included in the models and the possible biological links of these pathways to the obesity metagenes remain unclear, and should be the focus for future investigations.
-
-% TODO: update the stepwise linear model table and the scatter plots

### 2017-03-27 thesis/discussion/discussion.tex
edited in Mik's comments and spell-checked everything

@@
 To conclude this thesis, all of the results were re-examined and are summarised here to highlight the key findings from this project.
-There were many limitations to this project and these limitations are adressed either in the summary in \cref{sec:summary_of_the_results}, or seaprately in \cref{sec:limitations}.
-Lastly, conclusions are drawn from all of the evidenc from this project and suggestions for future aims and experiments are made.
+There were many limitations to this project and these limitations are addressed either in the summary in \cref{sec:summary_of_the_results}, or separately in \cref{sec:limitations}.
+Lastly, conclusions are drawn from all of the evidence from this project and suggestions for future aims and experiments are made.
 
@@
 It appeared as though the CR data-derived obesity associated genetic signatures correlated significantly in the CR data, but as soon as these metagenes were generated in the other data sets via the application of transformation matrices, the correlation became insignificant.
-One possible reason for this was that the obesity metagenes were not actually associated to obesity phenotype in the other data sets.
+One possible reason for this was that the obesity metagenes were not actually associated to the obesity phenotype in the other data sets.
 Another reason could be that the methodology used were not appropriate for investigating the association between the phenotype and its associated phenotype.
@@
 \gls{svd} was first applied to the \gls{rma} normalised CR data to obtain the \gls{er} and \gls{pr} metagenes, along with the transformation matrices for each of the genetic signatures.
-Both the \gls{er} and \gls{pr} metagenes showed good accordance with the gene expressions of the genetic signature in the heatmap, and also showed significant association with the patient \gls{er} and \gls{pr} statuses (\cref{sec:results_for_er_and_pr_metagene_analysis}).
+Both the \gls{er} and \gls{pr} metagenes showed good concordance with the gene expression profiles of the genetic signature in the heatmap, and also showed significant association with the patient \gls{er} and \gls{pr} statuses (\cref{sec:results_for_er_and_pr_metagene_analysis}).
 
@@
 Both the \gls{er} and \gls{pr} pathway metagenes showed similar overall gene expression patterning in the heatmap as the CR data (\cref{sec:results_for_er_and_pr_metagene_analysis}).
-Furthermore, both the \gls{er} and \gls{pr} pathway metagenes from these data sets showed significant association with the patient \gls{er} and \gls{pr} statuses, respectively (\cref{sec:results_for_er_and_pr_metagene_analysis}).
-These results clearly showed that the general approach of using \gls{svd} and transformation matrix was appropriate for examining the association of the genetic signatures with the patient phenotype.
+Furthermore, both the \gls{er} and \gls{pr} pathway metagenes from these data sets show significant association with the patient \gls{er} and \gls{pr} statuses, respectively (\cref{sec:results_for_er_and_pr_metagene_analysis}).
+These results clearly showed that the general approach of using \gls{svd} and transformation matrices was appropriate for examining the association of the genetic signatures with the patient phenotype.
 
@@
 
-(Though how unreliable the Akt pathway metagene scores may be, there is still a possibility that the FM obesity associated genetic signature may be unrelated to the Akt pathway, as Fuentes-Mattei \textit{et al.} validated their human breast tumour-derived genetic signature in a mouse model \citep{Fuentes-Mattei2014}.
-	Mouse models are indeed helpful in studying biological processes \textit{in vivo}, especially when the cause and the outcome of the disease is well established.
-	However, cancer is a complex genetic disorder which may be difficult to model accurately in mice.
-	Furthermore, obesity is a disease that have a variety of biological effects (for example the release of adipokines and the disruption of the hormone levels) that could be acting in a different way in mice compared to humans, which further complicates the causal relationship between obesity and cancer.
-	Nevertheless, the evidence shown by \citet{Fuentes-Mattei2014} was convincing and the role of Akt/\gls{mtor} pathway in obesity and cancer should be explored in greater details.)
-\textit{(include this paragraph or not... tbh, I think this is a weak argument)}
-
 \subsection{Linear models to predict the obesity metagenes}
@@
 Furthermore, both the \gls{nzbc} and the CR data sets showed strong association between the \gls{pr} and \gls{er} pathway metagenes with the patient \gls{pr} and \gls{er} status, respectively (\cref{sec:results_for_er_and_pr_metagene_analysis}).
-Since both \gls{er} and \gls{pr} signatures were significantly associated with the patient \gls{er} and \gls{pr} satuses, both signatures would have been as equally likely to show up significant in the models, supporting the idea that the \gls{er} and \gls{pr} statuses were independent of the apparent significance of these signatures in the models.
+Since both \gls{er} and \gls{pr} signatures were significantly associated with the patient \gls{er} and \gls{pr} statuses, both signatures would have been as equally likely to show up significant in the models, supporting the idea that the \gls{er} and \gls{pr} statuses were independent of the apparent significance of these signatures in the models.
 
 In the step wise linear models, many variables were identified to be significantly associated with the obesity metagenes.
-This suggested that the pathway associated genetic signatures that were signicant in these models may be biologically related to the obesity phenotype, or perhaps the obesity associated genetic signatures were detecting the signals from these pathways rather than the signals from the obesity phenotype.
+This suggested that the pathway associated genetic signatures that were significant in these models may be biologically related to the obesity phenotype, or perhaps the obesity associated genetic signatures were detecting the signals from these pathways rather than the signals from the obesity phenotype.
 In fact, many of the variables included in the linear models were pathway metagenes that negatively correlated with the obesity metagenes in \cref{fig:gatza_allmeta} (\cref{sub:comparison_of_the_obesity_and_pathway_associated_signatures}).
-Since all of the obesity metagenes failed to show significant association with patient \gls{bmi}, it may be possible that these pathways (Akt, \gls{er}, \gls{egfr}, \gls{pr}, p53, Src and \gls{tgfb}) are related to the underlying biological mechanism that drives the breast tumour development.
+Since all of the obesity metagenes failed to show significant association with patient \gls{bmi}, it may be possible that these pathways (Akt, \gls{er}, \gls{egfr}, \gls{pr}, p53, Src and \gls{tgfb}) are actually what was being detected by the ``obesity'' metagenes.
 
@@
 
-There were two main aims to this project: firstly to determine whether there were any obesity specific genetic signatures that could be tranferred across multiple cancer types; and secondly to investigate whether there were any biological pathways dysregulated in the tumour samples from the patients that were obese compared to the samples from non-obese patients.
-These aims were addressed in \cref{cha:obesity_genetic_signatures_and_cancer,cha:obesity_associated_genetic_signature_and_pathway_signatures}, resepectively.
+There were two main aims to this project: firstly to determine whether there were any obesity specific genetic signatures that could be transferred across multiple cancer types; and secondly to investigate whether there were any biological pathways dysregulated in the tumour samples from the patients that were obese compared to the samples from non-obese patients.
+These aims were addressed in \cref{cha:obesity_genetic_signatures_and_cancer,cha:obesity_associated_genetic_signature_and_pathway_signatures}, respectively.
 
-As shown clearly by the results from \cref{cha:obesity_genetic_signatures_and_cancer}, there seemed to be no  genetic signatures that were able to differntiate between the tumour samples from the patients that were obese from those that were not, across different cancer types.
+As shown clearly by the results from \cref{cha:obesity_genetic_signatures_and_cancer}, there seemed to be no  genetic signatures that were able to differentiate between the tumour samples from the patients that were obese from those that were not, across different cancer types.
 Furthermore, there were no genes differentially expressed between obese and non-obese patients that were common across multiple cancer types.
@@
 However, strong negative correlation with the Akt, \gls{er}, \gls{egfr}, \gls{pr}, p53, Src and \gls{tgfb} pathways, suggesting that the possibility of the obesity associated genetic signatures were picking up the signals from these pathways, rather than the signals from the obesity phenotype of the patients.
-The results from the linear model prediction suggested that the Akt, \gls{egfr} and Src pathways may be important in breast tumour development and/or progression.
 
@@
 Although this study failed to identify obesity specific genes that affected tumour biology, it is likely that there is some complex mechanism underlying the relationship between obesity and cancer; otherwise many of the clinical and epidemiological associations between obesity and cancer risk would not make any sense.
-Since Akt, \gls{egfr} and \gls{pr} pathways were signicant in the linear models that predicted the obesity metagenes, more thorough investigation of these pathways may be a good start to clarify the relationship.
+Since Akt, \gls{egfr} and \gls{pr} pathways were significant in the linear models that predicted the obesity metagenes, more thorough investigation of these pathways may be a good start to clarify the relationship.
 Further analyses of \gls{blca} may help clarify the similarity between bladder and breast cancer types and provide insight into the association between the obesity associated genetic signatures and \gls{blca}.

### 2017-03-27 thesis/intro/intro.tex
edited in Mik's comments and spell-checked everything

@@
 For children under the age of 5, overweight and obesity are defined as weight-for-height greater than 2 standard deviation and 3 standard deviation above \gls{who} Child Growth Standards median, respectively.
-For children aged between 5 to 19 years, overweight and obesity are defined as BMI-for-age greater than 1 standard deviation and 2 standard deviation above \gls{who} Growth Reference Standards median, respectively.
+For children aged between 5 to 19 years, overweight and obesity are defined as \gls{bmi}-for-age greater than 1 standard deviation and 2 standard deviation above \gls{who} Growth Reference Standards median, respectively.
 
@@
 
-% TODO: Add western-specific statistics on obesity (wang2007)
-% TODO: Add male/female specific stats
-
 \subsection{Physiological mechanism of obesity}
@@
 The role of hypothalamus is central to the control of satiety and many of the mutations that cause obesity affects the satiety signals to and/or from hypothalamus.
-The hypothalamus receives both nueronal and hormonal inputs from the body and propagates the signal to the downstream neurons that ultimately affect the food intake of the individual \citep{Bell2005, Spiegelman2001}.
+The hypothalamus receives both neuronal and hormonal inputs from the body and propagates the signal to the downstream neurons that ultimately affect the food intake of the individual \citep{Bell2005, Spiegelman2001}.
 
@@
 
-The evidence of the involvement of insulin in the satiety control was not particularly clear even though there were strong evidence of obesity with neuron-specific deletion of insulin receptor.
-% TODO: citation for the sentence above (and check with citation below)
+The exact role of insulin in the satiety control was not particularly clear even though there was evidence of increased food intake with neuron-specific deletion of insulin receptor \citep{Barsh2002,Bruning2000}.
 The fact that the insulin level, like leptin, was proportional to the fat mass also suggested the role of insulin in satiety and body weight control \citep{Barsh2002, Bruning2000, Woods1979}.
@@
 These individual-level genetic factors contribute to the overall increase in the proportion of the population that are obese.
-% TODO: citation for the sentence below?
-However, it is likely that the environmental factors are probably the most significant contributor to the current global obesity epidemic,
+However, it is likely that the environmental factors are the most significant contributor to the current global obesity epidemic \citep{Malik2013}.
 
@@
 
-% TODO: citation for the sentence below?
 It is clear that the individual-level factors cannot be accounted for the global epidemic of obesity that the world currently observes.
@@
 
-So, for the low- and middle-income groups, the rise in the average income in these groups allow the people to have access to food and activities that promotes sedentary bahaviours, but the limited access to the health care system, education and recreational activities that promote physical activites ultimately lead to positive energy balance \citep{Malik2013}.
+So, for the low- and middle-income groups, the rise in the average income in these groups allow the people to have access to food and activities that promotes sedentary behaviours, but the limited access to the health care system, education and recreational activities that promote physical activities ultimately lead to positive energy balance \citep{Malik2013}.
 In contrast, the high-income group of the population have better access to the facilities that promote weight maintenance, thus have less of an impact compared with the low- and middle-income groups of the population.
@@
 
-It is now commonly understood that cancers are caused by dysregulation of various molecular mechanisms or pathways that allow tumour cells to prolilferate, survive and migrate.
+It is now commonly understood that cancers are caused by dysregulation of various molecular mechanisms or pathways that allow tumour cells to proliferate, survive and migrate.
 By taking advantage of these pathways, cancer cells are able to grow and eventually metastasise to other parts of the body.
@@
 
-In addition to this, areas of the chromosome where major structural rearrangements have occured can be prone to a regional hypermutation, termed \gls{kataegis} \citep{Leibowitz2015,Nik-Zainal2012}.
+In addition to this, areas of the chromosome where major structural rearrangements have occurred can be prone to a regional hypermutation, termed \gls{kataegis} \citep{Leibowitz2015,Nik-Zainal2012}.
 Although the full mechanism in which \gls{kataegis} occur is not yet defined, \gls{aid}/\acrshort{apobec} family proteins, a family of proteins involved in gene editing, may play a major role in \gls{kataegis} \citep{Leibowitz2015,Nik-Zainal2012}.
@@
 It seems counterintuitive to limit its full potential of creating \gls{atp} by utilising glycolysis over electron transport chain, as the latter produces almost 18 times more \gls{atp} \citep{Hanahan2011, VanderHeiden2009}.
-However, it is hypothesised  that this apprent inefficiency helps the cell to generate and direct the molecules required for biosynthetic pathways during cell growth and division \citep{Cairns2011,VanderHeiden2009}.
+However, it is hypothesised  that this apparent inefficiency helps the cell to generate and direct the molecules required for biosynthetic pathways during cell growth and division \citep{Cairns2011,VanderHeiden2009}.
 By deliberately relying purely on glycolysis (known as the ``Wardberg effect'') and slowing the conversion of \gls{pep} to pyruvate, the cell is able to direct the carbon molecules into reaction pathways that produce the constituents used to synthesise macromolecules, such as \acrshort{dna} and lipids \citep{Cairns2011,VanderHeiden2009}.
@@
 
-Unfortunately, there are \glspl{germline} that make the individual more prediposed to cancer \citep{Vogelstein2004a}.
+Unfortunately, there are \glspl{germline} that make the individual more predisposed to cancer \citep{Vogelstein2004a}.
 These individuals have a mutation in one or more of the classes of cancer-related genes, which makes them more likely to develop a tumour than individuals without the mutation \citep{Vogelstein2004a}.
@@
 Epidemiological evidence of obesity being a risk factor for many cancers has been well established, yet the mechanisms of its contribution to cancer is still  speculative \citep{Calle2003,Kelesidis2006}.
-Although there are a lot of possible mechcanisms, there is growing evidence of the role of chronic inflammation caused by increased fat mass in the body \citep{Kelesidis2006,Lumeng2011,Hernandez2013}.
+Although there are a lot of possible mechanisms, there is growing evidence of the role of chronic inflammation caused by increased fat mass in the body \citep{Kelesidis2006,Lumeng2011,Hernandez2013}.
 Free fatty acids from excess adipose tissue and/or diet can be recognised by \gls{tlr} (specifically \gls{tlr}4), which can then activate the \gls{nfkb} pathway, leading to the activation of one of the Hallmarks of Cancer (tumour-promoting inflammation) \citep{Lumeng2011}.
@@
 
-Although the mechanisms in which obesity causes cancer are still under debate, the association between obesity and cancer has long been established.
+Although the mechanisms by which obesity causes cancer are still under debate, the association between obesity and cancer has long been established.
 Perhaps the most convincing evidence of the association between obesity and cancer was first presented by \citet{Calle2003}.
 In this study, the link between \gls{bmi} and variety of cancers were investigated in 900,000 adults between 1982 and 1998 \citep{Calle2003}.
-The results from this study showed that morbidly obese men and women (\gls{bmi} \textgreater{} 40) had 52\% and 62\% higher chance of dying from all cancers compared to normal weight adults (18.5 \textgreater{} \gls{bmi} \textgreater{} 25), respectively \citep{Calle2003}.
+The results from this study showed that morbidly obese men and women (\gls{bmi} \textgreater{} 40) had 52\% and 62\% higher chance of dying from all cancers compared to normal weight adults (18.5 \textless{} \gls{bmi} \textless{} 25), respectively \citep{Calle2003}.
 \gls{bmi} was also associated with higher rates of death from esophageal, colorectal, liver, gallbladder, pancreas, kidney, and some sex-specific cancers \citep{Calle2003}.
@@
 
-\subsection{Mechanisms of cancer progression in obese patient}
+\subsection{Mechanisms of cancer progression in obese patients}
 \label{sub:mechanisms_of_cancer_progression_in_obese_patient}
 
-The exact mechanisms in which obesity causes cancer have still not been fully elucidated yet.
-There are several hypotheses that link obesity with cancer: disruption of various hormone levels, chronic and sustained inflammation, and the activation of the insulin/\gls{igf} pathway.
+The exact mechanisms by which obesity causes cancer have still not been fully elucidated yet.
+There are several hypotheses that link obesity with cancer: disruption of various hormone levels, chronic and sustained inflammation, and the activation of the insulin/\gls{igf} pathway \citep{Lumeng2011,Roberts2010}.
 
-Adipose tissue has long been thought to be a group of cells to store excess fats, but it has recently become apparent that adipose tissue is also a metabolically active endocrine organ \citep{Roberts2010}.
+Adipose tissue has long been known to be a group of cells to store excess fats, but it has recently become apparent that adipose tissue is also a metabolically active endocrine organ \citep{Roberts2010}.
 Increased adiposity is known to influence the bioavailability of endogenous sex hormones, including estrogens, androgens and progesterone \citep{Calle2004}.
-\Glspl{adipocyte} synthesises enzymes, such as aromatase, that convert precursor molecules into different forms of androgens and estrogens \citep{Calle2004}.
+\Glspl{adipocyte} synthesise enzymes, such as aromatase, that convert precursor molecules into different forms of androgens and estrogens \citep{Calle2004}.
 Many studies have confirmed the mitogenic effect of sex hormones on certain cell types, mainly breast and endometrial cell types \citep{Roberts2010}.
@@
 One of the difficulties associated with the treatment of cancers is the identification of the underlying biological mechanisms or hallmarks that drive the cancer progression.
-As one can imagine, it is difficult for the clinicians to treat a patient if they do not know what is causing the disease.
-The reason why it is so difficult to identify the underlyig biological mechanisms is that there are very little reliable biological markers (or ``genetic signatures'') that signify the presence or absence of the causal pathways.
+As one can imagine, it is difficult for clinicians to treat a patient if they do not know what is causing the disease.
+The reason it is so difficult to identify the underlying biological mechanisms is that there are very few reliable biological markers (or ``genetic signatures'') that signify the presence or absence of the causal pathways.
 With that said, significant improvements have been made in recent years owing to the advancement in the sequencing technologies and have allowed the researchers to study the genetic signatures that represent the hallmarks in a much greater detail.
@@
 In short, microarray technology uses a glass slide with hundreds of thousands of short \acrshort{dna} sequences attached to it, where fluorescently labelled \glspl{cdna} taken from the sample \glspl{mrna} are hybridised to these immobilised \acrshort{dna} sequences \citep{Schena1995,Schulze2001}.
-The fluorscent probes are then detected to measure the gene expression \citep{Schena1995,Schulze2001}.
+The fluorescent probes are then detected to measure the gene expression \citep{Schena1995,Schulze2001}.
 
@@
 
-\gls{ngs} technology have many advantages over microarray technology.
+\gls{ngs} technology has many advantages over microarray technology.
 Firstly, \gls{ngs} does not require any prior knowledge of the sample like in the microarray technology, as all genetic materials are attached to universal primers when they are sequenced; there is no need to pre-define the sequences to be searched and thus allow the detection of novel genes or transcripts \citep{Hurd2009}.
 This also prevents the introduction of experimental bias, as the sequences are not altered in any way and no specific sequence is searched in a greater detail over another \citep{Hurd2009}.
-Another advantage is that \gls{ngs} produce count data of the genetic material being sequenced instead of relying on the intensity of the fluorescence to detect the presence and amount of certan sequence.
+Another advantage is that \gls{ngs} produce count data of the genetic material being sequenced instead of relying on the intensity of the fluorescence to detect the presence and amount of certain sequence.
 This allows an absolute quantification of the amount of transcripts from the sample, making it possible for a more accurate comparison of the expression levels between samples.
@@
 
-The use of \gls{ngs} technology have significantly improved the workflow of modern genetic studies and its application in variety of research area have made it possible to dig deeper into the causes of complex diseases, such as cancer.
+The use of \gls{ngs} technology has significantly improved the workflow of modern genetic studies and its application in variety of research area have made it possible to dig deeper into the causes of complex diseases, such as cancer.
 With the cost of sequencing declining, personalised genome sequencing and treatment based on the sequence will be a reality in the near future.
@@
 
-There is a clear epidemiological evidence of obesity being a major risk factor for many cancer types (\cref{sec:obesity_and_cancer}), but the genes or pathways that directly contribute to tumorigenesis in patients that are obese is yet to be clarified.
+There is clear epidemiological evidence of obesity being a major risk factor for many cancer types (\cref{sec:obesity_and_cancer}), but the genes or pathways that directly contribute to tumorigenesis in patients that are obese is yet to be clarified.
 Two studies in particular, one by \citet{Creighton2012} and another by \citet{Fuentes-Mattei2014}, have investigated whether there were any difference in genetic composition between the obese and the non-obese patients, specifically in a group of breast cancer patients.
@@
 They compared the gene expression between the tumours from obese patients with the tumours from non-obese patients and identified 799 gene probe sets (662 unique genes) with p \textless{} 0.01 and with a \gls{fc} \textgreater{} 1.2.
-Majority of these genes (725 gene probes) were downregulated and 74 gene probes were upregulated in the breast tumour samples from the obese patients compared with the breast tumour samples from the non-obese patients.
+The majority of these genes (725 gene probes) were downregulated and 74 gene probes were upregulated in the breast tumour samples from the obese patients compared with the breast tumour samples from the non-obese patients.
 
@@
 These pathway specific genes were identified experimentally by altering the expression of one specific gene that is central to the pathway in a model cell culture system.
-For example, all of the genes that have altered its expression in the cells that had  the \textit{MYC} gene induced were noted down as the genes involved in the Myc pathway.
+For example, all of the genes that have altered expression in the cells that had  the \textit{MYC} gene induced were noted as genes involved in the Myc pathway.
 
@@
 
-\citet{Gatza2010a} used 18 pathway signatures to classify human breast tumours into more detailed subtypes.
+Building on the work of \citet{Bild2006}, \citet{Gatza2010a} used 18 pathway signatures to classify human breast tumours into more detailed subtypes.
 The 18 pathways used in this study were: Akt, \gls{bcat}, E2F1, \gls{egfr}, \gls{er}, \gls{her2}, \gls{ifna}, \gls{ifny}, Myc, p53, p63, \gls{pi3k}, \gls{pr}, Ras, \gls{stat3}, Src, \gls{tgfb}, and \gls{tnfa} pathways.
@@
 The results from these analyses are split into two parts, each presented in a separate chapter.
-First part of the results clarify the relationship between the obesity associated genetic signatures with various cancer types.
-Second part of the results consist of the links between the pathway associated genetic signatures and various cancer types, as well as the previously identified obesity associated signatures.
-Lastly, the results will be summarised and the conclusion will be drawn from these results.
+The first part of the results clarify the relationship between the obesity associated genetic signatures with various cancer types.
+The second part of the results consist of the links between the pathway associated genetic signatures and various cancer types, as well as the previously identified obesity associated signatures.
+Lastly, the results will be summarised and conclusions will be drawn from these results.

### 2017-03-27 thesis/methods/methods.tex
finished editing up to the discussion
edited in Mik's comments and spell-checked everything

@@
 See \cref{tab:samplesize} for a summary of the total number of samples included in the analyses for each cancer type.
+Where missing, \cref{eq:bmicalc} was used to calculate the \gls{bmi} of the samples in all of the cancer data with height and weight data available (\cref{sec:data}).
+\\
+
+\begin{equation}
+	\label{eq:bmicalc}
+	\gls{bmi} = \frac{Weight (kg)}{Height^2(m^2)}\\
+\end{equation}
+\\
+
+\noindent
+Samples were classified as normal weight (\gls{bmi} \textless{} 25), overweight (25 $\leq$ \gls{bmi} \textless{} 30) or obese (\gls{bmi} $\geq$ 30) based on the \gls{who} definition (\cref{tab:whobmiclass}).
+\cref{eq:bmicalc,tab:whobmiclass} were used to calculate and group the patients from the \gls{icgc} cancer data into appropriate \gls{bmi} classes (\cref{tab:icgc_bmi}).
 
@@
 \end{table}
-
-\section{\Gls{bmi}}
-\label{sec:bmi}
-
-\subsection{\gls{bmi} Calculation}
-\label{subsec:bmicalc}
-
-Where missing, \cref{eq:bmicalc} was used to calculate the \gls{bmi} of the samples in all of the cancer data with height and weight data available (\cref{sec:data}).
-
-\begin{equation}
-	\label{eq:bmicalc}
-	\gls{bmi} = \frac{Weight (kg)}{Height^2(m^2)}\\
-\end{equation}
-
-\subsection{\gls{bmi} Classification}
-\label{subsec:bmiclassification}
-
-Samples were classified as normal weight (\gls{bmi} \textless{} 25), overweight (25 $\leq$ \gls{bmi} \textless{} 30) or obese (\gls{bmi} $\geq$ 30) based on the \gls{who} definition (\cref{tab:whobmiclass}).
-\cref{eq:bmicalc,tab:whobmiclass} were used to calculate and group the patients from the \gls{icgc} cancer data into appropriate \gls{bmi} classes (\cref{tab:icgc_bmi}).
 
@@
 The raw microarray data were normalised using both \gls{rma} and \gls{mas} methods, each done separately on a copy of the data.
-The reason why the \gls{mas} method was used as well as the \gls{rma} method was becuase the \gls{mas} normalisation method was used in the \citet{Gatza2010a} study.
+The reason why the \gls{mas} method was used as well as the \gls{rma} method was because the \gls{mas} normalisation method was used in the \citet{Gatza2010a} study.
 Though the normalisation methods were not specified in the \citet{Creighton2012} and \citet{Fuentes-Mattei2014} studies, to allow for data comparison between different datasets and accurate result validation of some of the studies, both \gls{rma} and \gls{mas} methods were considered for normalising the microarray data.
@@
 In some cases, an analysis was repeated a number of times to provide statistical evidence that the observed results did not occur by chance.
-All of the simulated results must be derived from a randomised data, and by comparing the observed results with the results that occured by chance, simulation analysis provides statistical evidence for the observed results.
+All of the simulated results must be derived from a randomised data, and by comparing the observed results with the results that occurred by chance, simulation analysis provides statistical evidence for the observed results.
 When a simulation was required, none of the values in the original data were modified, but instead the values in the original data were assigned randomly to different samples; in other words, the samples were shuffled.
@@
 \textit{limma} is a package that contains variety of tools to analyse gene expression data using linear models-based methods, developed by \citet{Ritchie2015}.
-For gene expression analysis, \textit{limma} package fits a linear model in a genewise manner and produce test statistics that allow the assessment of whether the gene is differentially expressed or not between certain groups of samples \citep{Ritchie2015}.
+For gene expression analysis, \textit{limma} package fits a linear model in a gene-wise manner and produce test statistics that allow the assessment of whether the gene is differentially expressed or not between certain groups of samples \citep{Ritchie2015}.
 In this project, the gene expression analysis was carried out to identify \glspl{deg} in the samples that were obese compared to the samples that were non-obese group.
@@
 
-The \texttt{lmFit} function (\textit{limma}) was used to fit a linear model to the normalised data, using the experiental design information from the design matrix.
+The \texttt{lmFit} function (\textit{limma}) was used to fit a linear model to the normalised data, using the experimental design information from the design matrix.
 The output of this function was used in the \texttt{eBayes} function (\textit{limma}) to identify the \glspl{deg} from the data.
@@
 In an experiment where there are multiple hypotheses being tested, the rate or proportion of \gls{type1} appearing from the experiment must be controlled.
-The probability of \gls{type1} occuring for a single hypopthesis is usually controlled at some significance threshold $\alpha$, which is usually set at 0.05 \citep{Shaffer1995}.
+The probability of \gls{type1} occurring for a single hypothesis is usually controlled at some significance threshold $\alpha$, which is usually set at 0.05 \citep{Shaffer1995}.
 In a typical microarray experiment, there are over 20,000 gene probes to be tested for differential expression, and with a significance threshold of $\alpha = 0.05$, this would yield approximately 1,000 \glspl{type1}.
@@
 
-\subsection{\Gls{ora} and \gls{fcs}}
-\label{sub:ora_and_fcs}
-
-\subsubsection{\Gls{ora}}
+\subsection{\Gls{ora}}
 \label{ssub:ora}
@@
 
-\subsubsection{\gls{fcs}}
+\subsection{\Gls{fcs}}
 \label{ssub:fcs}
@@
 There are two major classes of methods when calculating the statistical significance: self-contained or competitive tests \citep{Goeman2007}.
-Self-contained tests are more appropriate in assessing whether a biological process is significantly involved in an expreiment, whereas the competitive tests are better for selecting the most relevant biological processes from those that are not \citep{Wu2012}.
+Self-contained tests are more appropriate in assessing whether a biological process is significantly involved in an experiment, whereas the competitive tests are better for selecting the most relevant biological processes from those that are not \citep{Wu2012}.
 
@@
 These hypotheses show that the self-contained tests care only about the genes defined in the gene set, but the competitive tests examine the genes in the defined set as well as the genes not present in the set \citep{Wu2012}.
-Due to the fact that self-contained tests do not take into account for the genes not defined in the gene set, self-contained tests are more restrictive compared to competitive tests.
+Due to the fact that self-contained tests do not take into account the genes not defined in the gene set, self-contained tests are more restrictive compared to competitive tests.
 The restrictive property of the self-contained tests give it greater power, as they are able to reject the null hypothesis at a higher accuracy for more gene sets than the competitive tests \citep{Goeman2007}.
@@
 By calculating the statistical significance per pathway, the \gls{fcs} approach takes into account of all of the genes involved in the pathways, and not just the genes that are differentially expressed.
-Furthermore, the \gls{fcs} approach also detect small but consistant and coordinated changes in the expression of the genes, unlike in the \gls{ora} where molecular measurements are completely ignored \citep{Khatri2012}.
+Furthermore, the \gls{fcs} approach also detects small but consistent and coordinated changes in the expression of the genes, unlike in the \gls{ora} where molecular measurements are completely ignored \citep{Khatri2012}.
 
 Though the \gls{fcs} approach improves on the \gls{ora} approach, there are still some limitations.
-Since the analyses that use \gls{fcs} compare the pathways independently from one another, it ignores the interactive nature of the biological pathways.
-Another limitation is that the \gls{fcs} methods use the molecular measurements to rank the genes, but they do not consider these changes into further analysis \citep{Khatri2012}.
-For example, if one gene was expressed with a 2-fold change, whereas another gene was expressed with a 20-fold change, the \gls{fcs} method will rank these genes accordingly (first, second, and so on) and disregard the fact that the second gene should be weighted ten times as more than the first gene \citep{Khatri2012}.
+Since the analyses that use \gls{fcs} compare the pathways independently of one another, they ignores the interactive nature of biological pathways.
+Another limitation is that the \gls{fcs} methods use the molecular measurements to rank the genes, but they do not consider these changes in further analysis \citep{Khatri2012}.
+For example, if one gene was expressed with a 2-fold change, whereas another gene was expressed with a 20-fold change, the \gls{fcs} method will rank these genes accordingly (first, second, and so on) and disregard the fact that the second gene is much more highly expressed \citep{Khatri2012}.
 Although both approaches have their own limitations, it is clear that \gls{fcs} approach have more advantages over \gls{ora} methods.
@@
 \Acrfull{camera} is a competitive \gls{fcs}-based method developed by \citet{Wu2012}, implemented as the \texttt{camera} function in the \textit{limma} package \citep{Ritchie2015}.
-Briefly, \gls{camera} fits a linear model in a genewise manner and calculates the genewise test statistics using the log\gls{fc} between the conditions (for example, obese and non-obese samples).
-These genewise test statistics are used in the \gls{wmw} rank sum test to question whether a pathway is significantly enriched in the data \citep{Wu2012}.
+Briefly, \gls{camera} fits a linear model in a gene-wise manner and calculates the gene-wise test statistics using the log\gls{fc} between the conditions (for example, obese and non-obese samples).
+These gene-wise test statistics are used in the \gls{wmw} rank sum test to question whether a pathway is significantly enriched in the data \citep{Wu2012}.
 
@@
 As a consequence, the \gls{type1} rate is increased in these methods.
-However, \Gls{camera} accounts for this correlation by estimating the \gls{vif} from the genewise correlation and the number of genes in the gene set \citep{Wu2012}.
+However, \Gls{camera} accounts for this correlation by estimating the \gls{vif} from the gene-wise correlation and the number of genes in the gene set \citep{Wu2012}.
 
@@
 The \textit{GO.db} package was used to load the \gls{go} database into \gls{r} \citep{Carlson2016}.
-Only the human related \gls{go} terms were used in pathway enrichment analysis, and these terms were organised into a format so that the pathways were able to be queried by using the gene symbol.
+Only the human related \gls{go} terms were used in pathway enrichment analysis, and these terms were organised into a format so that the pathways were able to be queried using the gene symbol.
 
@@
 The \textit{KEGG.db} package was used to load the \gls{kegg} database into \gls{r} \citep{kegg}.
-Only the human related \gls{kegg} terms were used in pathway enrichment analysis, and these terms were organised into a format so that the pathways were able to be queried by using the gene symbol.
+Only the human related \gls{kegg} terms were used in pathway enrichment analysis, and these terms were organised into a format so that the pathways were able to be queried using the gene symbol.
 
@@
 
-The Reactome database is a ``curated, peer-reviewed resource of humana biological processes'' \citep{Joshi2005}.
-Reactome database is constructed base on the reactions that have direct evidence from the literature, which are connected into a higher order pathway structures \citep{Joshi2005}.
+The Reactome database is a ``curated, peer-reviewed resource of human biological processes'' \citep{Joshi2005}.
+The Reactome database is constructed based on the reactions that have direct evidence from the literature, which are connected into higher order pathway structures \citep{Joshi2005}.
 
 The \textit{reactome.db} package was used to load the Reactome database into \gls{r} \citep{reactome}.
-Only the human related Reactome terms were used in pathway enrichment analysis, and these terms were organised into a format so that the pathways were able to be queried by using the gene symbol.
+Only the human related Reactome terms were used in pathway enrichment analysis, and these terms were organised into a format so that the pathways were able to be queried using the gene symbol.
 
@@
 Therefore, the metagenes created independently from different data sets using \gls{svd} do not have the same weighting as the metagene from the original data set, and consequently does not provide a fair comparison of the metagenes across different data sets.
-For this reason, the transformation matrix for a given set of genes was created in the data set in which the genes were first identified in, and then transformed into other data sets to obatain the metagenes.
+For this reason, the transformation matrix for a given set of genes was created in the data set in which the genes were first identified in, and then transformed into other data sets to obtain the metagenes.
 
@@
 In this transformation, the sample metagene scores were transformed with the probit function.
-To do this, the sample metagene scores were scaled and centered using the \texttt{scale} function from the \textit{base} pakage in \gls{r}, then the \texttt{pnorm} function from the \textit{stats} package was used to transform the metagene scores so that the scores were between 0 and 1.
+To do this, the sample metagene scores were scaled and centered using the \texttt{scale} function from the \textit{base} package in \gls{r}, then the \texttt{pnorm} function from the \textit{stats} package was used to transform the metagene scores so that the scores were between 0 and 1.
 Unless stated otherwise, all of the metagenes in this project were fractionally ranked.
@@
 	Each row of the heatmap represents a gene from the \gls{er} genetic signature, and each column of the heatmap represents a sample from the Gatza \textit{et al.} data.
-	The \textit{ESR1} gene expressions of the samples are shown as a separate row at the top of the heatmap with the sample \gls{er} metagene scores shown below the \textit{ESR1} gene expression, and the tree diagram of the heirarchical clustering of the genes is shown on the left of the heatmap.
+	The \textit{ESR1} gene expressions of the samples are shown as a separate row at the top of the heatmap with the sample \gls{er} metagene scores shown below the \textit{ESR1} gene expression, and the tree diagram of the hierarchical clustering of the genes is shown on the left of the heatmap.
 	}
@@
 	\hline
-	\rule{0pt}{2.25ex}\Gls{her2}  & ERBB2  & Also known as ERBB2; involved in cell proliferaion and apoptosis\\
+	\rule{0pt}{2.25ex}\Gls{her2}  & ERBB2  & Also known as ERBB2; involved in cell proliferation and apoptosis\\
 	\hline
@@
 	\hline
-	\rule{0pt}{2.25ex}Myc         & MYC    & Transcription factor involved in cell proliferaion, cell growth and apoptosis\\
+	\rule{0pt}{2.25ex}Myc         & MYC    & Transcription factor involved in cell proliferation, cell growth and apoptosis\\
 	\hline
@@
 	\hline
-	\rule{0pt}{2.25ex}\Gls{pi3k}  & PIK3CA & Involved in cell growth, cell proliferaion, cell differentiation and cell survival\\
+	\rule{0pt}{2.25ex}\Gls{pi3k}  & PIK3CA & Involved in cell growth, cell proliferation, cell differentiation and cell survival\\
 	\hline
@@
 	\hline
-	\rule{0pt}{2.25ex}Ras         & HRAS   & Involved in cell proliferaion, cell differentiation, cell adhesion, apoptosis and cell migration\\
+	\rule{0pt}{2.25ex}Ras         & HRAS   & Involved in cell proliferation, cell differentiation, cell adhesion, apoptosis and cell migration\\
 	\hline
@@
 In addition to these eleven linear models, a linear model was created with the sample \gls{bmi}, \gls{bmi} status and all of the pathway associated metagenes (Akt, \gls{bcat}, E2F1, \gls{egfr}, \gls{er}, \gls{her2}, \gls{ifna}, \gls{ifny}, Myc, p53, p63, \gls{pi3k}, \gls{pr}, Ras, \gls{stat3}, Src, \gls{tgfb} and \gls{tnfa}) using a stepwise approach.
-Stepwise approach allows the construction of a linear model that best fit the data, given a set of variables.
-There are three approaches in creating a linear model in a stepwise method: top-down, bottom-up, and both.
+The stepwise approach allows the construction of a linear model that best fits the data, given a set of variables.
+There are three approaches in creating a linear model in a stepwise method: forward, backward, and both.
 
-In a top-down approach, all of the variables are first included in the linear model.
-At each step, a variable is removed from the model and the \gls{aic} of the model is compared with the original model to determine whether the variable should remain in the model or not.
-This will result in a model with the variables that are most significant to the model.
-In a bottom-up approach, the model begins with no variable and adds one variable at a time to the model.
+In a backward (also known as ``top-down'') approach, all of the variables are first included in the linear model.
+At each step, a variable is removed from the model and the \gls{bic} of the model is compared with the original model to determine whether the variable should remain in the model or not.
+This will result in a model with the variables that are jointly most predictive.
+In a forward, or ``bottom-up'', approach, the model begins with no variable and adds one variable at a time to the model.
 If the variable increased the ability of the model to predict the response variable, then the variable is included in the model.
-Again, this will result in a model with only the variables that are significant in predicting the response variable.
+Again, this will result in a model with only the variables that are contributing to predicting the response variable.
 Alternatively, both methods in combination can be used, where the model begins with all or none of the variables, and at each step a variable can be added to or removed from the linear model.
-In this project, the combination of top-down and bottom-up approaches (beginning with no variables in the model) was used to create the stepwise linear model.
+In this project, the combination of forward and backward approaches (beginning with no variables in the model) was used to create the stepwise linear model.
 
-The obesity metagenes that were predicted from the linear models in the \gls{nzbc} data set was compared with the original obesity metagene scores.
+The obesity metagenes that were predicted from the linear models in the \gls{nzbc} data set were compared with the original obesity metagene scores.
 Where appropriate, p-values, \gls{anova} p-values and  $R^2$-values were calculated to assess the relationship between the predicted and the original metagene scores.
@@
 
-To summarise the number of \glspl{deg} from the gene expression analyses of the Creighton \textit{et al.} data that overlapped with the original Creighton \textit{et al.} obesity associated genetic signature, Venn diagrams were plotted.
+To summarise the number of \glspl{deg} from the gene expression analyses of the Creighton \textit{et al.} data that overlapped with the original Creighton \textit{et al.} obesity associated genetic signature, Venn diagrams were used.
 Venn diagrams were plotted using the \texttt{venn} function from the \textit{gplots} package in \gls{r} \citep{gplots}.
@@
 Some colour palettes such as the colours used by MATLAB are not installed in \gls{r} by default.
+In the \citet{Gatza2010a} study, Gatza \textit{et al.} used MATLAB to generate all of their plots in their paper.
+Since a visual comparison had to be made with one of the plots from the \citet{Gatza2010a} study, some of the heatmaps in this project required MATLAB-like colours for a more accurate comparison with their results.
 For a colour palette used in MATLAB, the \texttt{matlab.like} function from the \textit{colorRamps} package in \gls{r} was used \citep{Keitt2012}.
 For the pastel colour palette used in the bar plots, the \texttt{brewer.pal} function from the \textit{RColorBrewer} package in \gls{r} was used \citep{Neuwirth2014}.
-
-%TODO: add Spearman and Pearson correlation?

### 2017-03-27 thesis/results/results.tex
edited in Mik's comments and spell-checked everything

@@
 	Each row of the heatmap represents a gene from the obesity associated genetic signature, and each column of the heatmap represents a sample from the CR data.
-	The obesity associated metagene scores of the samples are shown as a separate row at top of the heatmap, and the tree diagram of the heirarchical clustering of the genes is shown on the left of the heatmap.
+	The obesity associated metagene scores of the samples are shown as a separate row at top of the heatmap, and the tree diagram of the hierarchical clustering of the genes is shown on the left of the heatmap.
 	For clarity, the sign of the metagene scores were reversed in order to match the results from the \citet{Creighton2012} study.}
@@
 These results showed that the obesity associated genetic signature from the \citet{Fuentes-Mattei2014} study was not generalisable to other cancer data sets, similar to the obesity signature identified by Creighton \textit{et al.}.
-This meant that both the Creighton \textit{et al.} and Fuentes-Mattei \textit{et al.} obesity metagenes may have been too specific to the orignial data set in which the signatures were identified in.
+This meant that both the Creighton \textit{et al.} and Fuentes-Mattei \textit{et al.} obesity metagenes may have been too specific to the original data set in which the signatures were identified in.
 Furthermore, there was a possibility that these obesity associated metagenes were not related to obesity, but associated with a different clinical variable that may be closely related to \gls{bmi} (\cref{sec:creighton_obesity_metagene_new}).
@@
 
-Both the obesity associated genetic signatures identfied from the \citet{Creighton2012} and \citet{Fuentes-Mattei2014} studies were able to capture the overall gene expression patterns  of the genetic signatures of the samples, but did not associate with patient \gls{bmi} or \gls{bmi} status in majority of the cancer data sets.
+Both the obesity associated genetic signatures identified from the \citet{Creighton2012} and \citet{Fuentes-Mattei2014} studies were able to capture the overall gene expression patterns  of the genetic signatures of the samples, but did not associate with patient \gls{bmi} or \gls{bmi} status in majority of the cancer data sets.
 One possible reason for this result could be that the obesity associated genetic signatures from the \citet{Creighton2012} and \citet{Fuentes-Mattei2014} studies may not have been truly associated with patient \gls{bmi}, but with another clinical variable.
-To invevstigate this possibility, novel obesity associated genetic signatures were identified in the CR data after controlling for all the clinical variables in the data set.
+To investigate this possibility, novel obesity associated genetic signatures were identified in the CR data after controlling for all the clinical variables in the data set.
 The FM data set was not used to get the obesity associated genetic signatures, as no \gls{bmi} information was available for the patients in the FM data.
@@
 Without adjusting the p-value to account for multiple hypothesis testing, 5278 gene probes and 1781 gene probes were significant at p \textless{} 0.05 and p \textless{} 0.01, respectively (\cref{tab:ob_deg_summary}).
-After adjustments were made for multiple hypothesis testing, there were only 9 gene probes significant at p \textless{} 0.05 and no gene was significant at p \textless{} 0.01, which suggest that the the majority of the 799 probe sets reported by Creighton \textit{et al.} may actually be false positives.
+After adjustments were made for multiple hypothesis testing, there were only 9 gene probes significant at p \textless{} 0.05 and no gene was significant at p \textless{} 0.01, which suggest that the majority of the 799 probe sets reported by Creighton \textit{et al.} may actually be false positives.
 
@@
 		Caucasian-only CR data          & 2129 & 6029 & 0 & 0 \\
-		Residual Caucasion-only CR data & 1558 & 5427 & 0 & 0 \\
+		Residual Caucasian-only CR data & 1558 & 5427 & 0 & 0 \\
 		\hline
@@
 % ; first in the unadjusted data, then in the data with clinical variables adjusted (except ethnicity, as it has already been controlled for by considering only the Caucasian samples), and lastly the common genes between each of these two signatures with the original obesity associated genetic signature were identified.
-2129 and 1558 gene probes were significantly differentially expressed (unadjuste p-value \textless{} 0.01) in the unadjusted and clinical variable-adjusted Caucasian patient data, respectively (\cref{tab:ob_deg_summary}).
+2129 and 1558 gene probes were significantly differentially expressed (unadjusted p-value \textless{} 0.01) in the unadjusted and clinical variable-adjusted Caucasian patient data, respectively (\cref{tab:ob_deg_summary}).
 As before, the most significant 799 gene probes were selected from these gene probes.
@@
 
-Since all of these obesity associated genetic signatures were derived based on the discrete values of the patient \gls{bmi} status, an obesity associated genetic signature was searched using the correlation of the gene probe expression with the patient \gls{bmi} value (continuous variable) in the CR data with all of the samples included, but no significant genes were identified (\cref{sec:obesity_associated_genetic_signature_using_sample_bmi_values}).
+Since all of these obesity associated genetic signatures were derived based on the discrete values of the patient \gls{bmi} status, an obesity associated genetic signature was searched for using the correlation of the gene probe expression with the patient \gls{bmi} value (continuous variable) in the CR data with all of the samples included, but no significant genes were identified (\cref{sec:obesity_associated_genetic_signature_using_sample_bmi_values}).
 All of these obesity associated genetic signatures (eight in total) were checked to see whether these signatures had significant association with patient \gls{bmi} or \gls{bmi} status in the \gls{nzbc} and \gls{icgc} cancer data sets (\cref{sub:_novel_obesity_associated_signatures_and_sample_bmi}).
@@
 
-When the pathway enrichment analysis was carried out with the \gls{kegg} pathway database, only the ``ABC transporter'' was significantly enriched (\gls{fdr}-adjusted p-value \textless{} 0.05) in the \gls{cesc} data, and no other pathways were enriched in any of the other \gls{icgc} cancer types (\cref{sec:pathways_significant_in_each_of_the_cancer_types}).
+When the pathway enrichment analysis was carried out with the \gls{kegg} pathway database, only the ``ABC transporter'' pathway was significantly enriched (\gls{fdr}-adjusted p-value \textless{} 0.05) in the \gls{cesc} data, and no other pathways were enriched in any of the other \gls{icgc} cancer types (\cref{sec:pathways_significant_in_each_of_the_cancer_types}).
 With the Reactome database, ``Phosphatase bond hydrolysis by NUDT proteins'' (in \gls{blca}) and ``Mitochondrial ABC transporters'' (in \gls{cesc}) were significantly enriched (\cref{sec:pathways_significant_in_each_of_the_cancer_types}).
-With the \gls{go} database, there were 22 pathways significantly enriched in \gls{blca}, 17 pathways in \gls{cesc}, 3 pathways in \gls{kirp}, 21 pathways in \gls{read}, 10 pathways in \gls{skcm}, 14 pathways in \gls{ucec}, and no pathways were significant in the \gls{coad} and \gls{lihc} cancer types (\cref{sec:pathways_significant_in_each_of_the_cancer_types}).
+With the \gls{go} database, there were 22 \gls{go} terms significantly enriched in \gls{blca}, 17 terms in \gls{cesc}, 3 terms in \gls{kirp}, 21 terms in \gls{read}, 10 terms in \gls{skcm}, 14 terms in \gls{ucec}, and no terms were significant in the \gls{coad} and \gls{lihc} cancer types (\cref{sec:pathways_significant_in_each_of_the_cancer_types}).
 
-Though there were many pathways identified as significantly enriched with the \gls{go} database in each of the cancer types, there were no pathways that were common across all of the different cancer types.
-Furthermore, these pathways were not similar in terms of the biological activities involved.
-For example, pathways enriched in the \gls{blca} data suggested possible activation of the \gls{pi3k} pathway, but the pathways enriched in the \gls{read} data were mainly involved in \acrshort{rna} processing (\cref{sec:pathways_significant_in_each_of_the_cancer_types}).
-One interesting result from this enrichment analysis was that the ``ABC transport'' pathway was identified as significantly enriched in the \gls{cesc} data set by all three pathway databases, suggesting that the ``ABC transport'' pathway may be a core component in the \gls{cesc} tumour biology.
-All of these results indicated that the pathways that drive tumour progression were unique to each cancer type, and no one pathway was associated with obesity and tumour biology across multiple cancer types.
+Though there were many \gls{go} terms identified as significantly enriched with the \gls{go} database in each of the cancer types, there were no terms that were common across all of the different cancer types.
+Furthermore, these terms were not similar in terms of the biological activities involved.
+For example, \gls{go} terms enriched in the \gls{blca} data suggested possible activation of the \gls{pi3k} pathway, but those enriched in the \gls{read} data were mainly involved in \acrshort{rna} processing (\cref{sec:pathways_significant_in_each_of_the_cancer_types}).
+One interesting result from this enrichment analysis was that the ``ABC transport'' pathway was identified as significantly enriched in the \gls{cesc} data set by all three databases, suggesting that the ``ABC transport'' pathway may be a core component in the \gls{cesc} tumour biology.
+All of these results indicated that the biological processes that drive tumour progression were unique to each cancer type, and no one pathway was associated with obesity and tumour biology across multiple cancer types.
 
-Since there were no common pathway enriched in the \gls{icgc} cancer types when analysed individually, all of the \gls{icgc} cancer data sets were combined into a single data set to see if there were any pathways enriched in the combined data set.
+Since there were no common pathway enriched in the \gls{icgc} cancer types when analysed individually, all of the \gls{icgc} cancer data sets were combined into a single data set to see any enriched pathways could be identified in a collective analysis.
 The combined data set was generated in two ways.
-The first combined data set was created by voom normalising each cancer data set separately, combine all of the normalised cancer data sets into one, then applied batch correction to the data set (\cref{sub:batch_correction}).
-The second combined data set was created by combining the cancer data sets into one big data set first, then the combined data was voom normalised, and finally batch correction was applied to obtain the final data set.
+The first combined data set was created by normalising (via voom) each cancer data set separately, combine all of the normalised cancer data sets into one, then applying batch correction to the data set (\cref{sub:batch_correction}).
+The second combined data set was created by combining the cancer data sets into one big data set first, then the combined data was normalised, and finally batch correction was applied to obtain the final data set.
 Each of the combined data sets were analysed for enriched pathways, but there were no pathways significantly enriched in either data sets.

### 2017-03-27 thesis/results/results2.tex
edited in Mik's comments and spell-checked everything

@@
 In the \citet{Gatza2010a} study, their data comprised of samples from different studies which were aggregated into a single data set that was \gls{mas} normalised, and the metagene scores were transformed to a scale between 0 to 1 using a probit function.
-However, the analyses so far have used \gls{rma} normalisation method and ranked based on the number of samples present in the data (fractional ranking; \cref{sub:ranking_of_the_metagene_scores}).
+However, the analyses so far have used the \gls{rma} normalisation method and ranked based on the number of samples present in the data (fractional ranking; \cref{sub:ranking_of_the_metagene_scores}).
 To decide which normalisation methods or ranking approaches were suitable for the analysis, the different methods were compared in the GT data (\cref{sec:ranking_method_for_the_pathway_associated_genetic_signatures,sec:normalisation_method_for_pathway_associated_genetic_signature_transformation_matrices}).
-Results shown in \cref{sec:ranking_method_for_the_pathway_associated_genetic_signatures} clarified that there was no significant difference in the ranking approaches used.
+Results shown in \cref{sec:ranking_method_for_the_pathway_associated_genetic_signatures} clarified that there was no substantive difference in the ranking approaches used.
 
-Scatter plots were plotted and the correlations were calculated for each of the GT pathway metagenes derived from either \gls{rma} or \gls{mas} normalised GT data (\cref{sec:normalisation_method_for_pathway_associated_genetic_signature_transformation_matrices}).
-The GT pathway metagenes were generated by using the transformation matrices that were derived from either the \gls{rma} or \gls{mas} normalised GT data.
+Scatter plots were generated and the correlations were calculated for each of the GT pathway metagenes derived from either \gls{rma} or \gls{mas} normalised GT data (\cref{sec:normalisation_method_for_pathway_associated_genetic_signature_transformation_matrices}).
+The GT pathway metagenes were generated using the transformation matrices that were derived from either the \gls{rma} or \gls{mas} normalised GT data.
 From these results, it was clear the normalisation methods used on the data set had the most significant effect on the resulting metagenes, rather than the normalisation methods used to generate the transformation matrices (\cref{sec:normalisation_method_for_pathway_associated_genetic_signature_transformation_matrices}).
-This meant that all of the data sets had to be normalised with a single normalisation method, so the \gls{rma} was chosen as the normalisation method to be used as this method was more reliable than the \gls{mas} method \citep{Irizarry2003}.
+This meant that all of the data sets had to be normalised with a single normalisation method, so \gls{rma} was chosen as the normalisation method to be used as this method is regarded as being more reliable than the \gls{mas} method \citep{Irizarry2003}.
 
@@
 This result provided evidence that some of the pathway genetic signatures from the \citet{Gatza2010a} study were less variable across different data sets than the others.
-Taken together, it was decided that the GT data set was batch corrected first (\cref{sub:batch_correction}), then normalised with the \gls{rma} method, and the metagene scores were ranked with fractional ranking (\cref{sub:ranking_of_the_metagene_scores}).
+Taken together, it was decided that the GT data set should be batch corrected first (\cref{sub:batch_correction}), then normalised with the \gls{rma} method, and the metagene scores were ranked with fractional ranking (\cref{sub:ranking_of_the_metagene_scores}).
 
@@
 In \cref{fig:gatza_allmeta}, all of the obesity associated metagenes from the CR data set cluster together into a group, but none of the pathway associated metagenes had  a strong positive correlation with the obesity metagenes.
-Though \gls{her2} and p63 pathway metagenes were grouped next to the cluster of obesity metagenes, the similarity was at most about 0.6  (Pearson correlation; \cref{fig:gatza_allmeta}).
+Though the  \gls{her2} and p63 pathway metagenes were grouped next to the cluster of obesity metagenes, the similarity was at most about 0.6  (Pearson correlation; \cref{fig:gatza_allmeta}).
 With that said, these obesity associated metagenes seem to have a strong negative correlation with some of the pathway associated genetic signatures, such as the Akt, \gls{er}, \gls{egfr}, \gls{pr}, p53, Src and \gls{tgfb} pathways, and this strong negative correlation was observed in the other data sets as well (\cref{sec:correlation_of_the_pathway_and_obesity_metagenes_in_other_cancer_data_sets}).
+In fact,  high Or metagene scores related to lower levels of gene expression for the genes in the Or signature (\cref{fig:crmetaheat}; \cref{sec:creighton_obesity_metagene}), and therefore it could be possible that many of the genes in the Or signature (and the other obesity associated signatures from the CR data) may be positively correlated with the genes in these pathway signatures.
 This suggests that the obesity associated genetic signatures from the CR data may be related to these pathway signatures, rather than the obesity phenotype of the patients.
 
-FM obesity associated metagene did not cluster together with the obesity metagenes from the CR data set, nor with any of the other pathway metagenes.
+The FM obesity associated metagene did not cluster together with the obesity metagenes from the CR data set, nor with any of the other pathway metagenes.
 The fact that the FM metagene did not cluster with any of the CR metagenes showed that the signals captures by the FM metagene was different from those metagenes found in the CR data set.
@@
 \textit{MYC} is an oncogene that encodes the Myc transcription factor that, when overexpressed, disrupts various pathways that directly promote tumour growth and proliferation, and has a central role in many of the cancer hallmarks \citep{Coller2000,Hanahan2000}.
-``Sustained cell proliferation'' was one of the cancer hallmarks that \citet{Fuentes-Mattei2014} had shown to be related to their obesity associated genetic signature, which could be the reason why the Myc pathway metagene showed up as a significant variable in the linear model.
+``Sustained cell proliferation'' was one of the cancer hallmarks that \citet{Fuentes-Mattei2014} had shown to be related to their obesity associated genetic signature, which could be the reason that the Myc pathway metagene showed up as a significant variable in the linear model.
 
@@
 
-Another thing to note from \cref{fig:predict_cr_cris} was the apparent negative assocation of the predicted metagenes with the original metagenes in the \gls{bmi}-only, the \gls{bmi} status-only and the \gls{bmi} and \gls{bmi} status models.
+Another thing to note from \cref{fig:predict_cr_cris} was the apparent negative association of the predicted metagenes with the original metagenes in the \gls{bmi}-only, the \gls{bmi} status-only and the \gls{bmi} and \gls{bmi} status models.
 One possible explanation for this was that these \gls{bmi}-based models were overfitted in the \gls{nzbc} data set, and was not generalisable in the CR data set, and thus produced negative correlations.
@@
 With that said, some of the predicted obesity metagenes from the stepwise linear models were negatively correlated with the original metagene values in the CR data (\cref{fig:stepwise_cr}).
-Again, this could have been due to the model being overfitted in the \gls{nzbc} data, and thus made an incorrect prediction about the obesity metagenes in the CR data.
+Again, this could have been due to the model being overfit in the \gls{nzbc} data, and thus making an incorrect prediction about the obesity metagenes in the CR data.
 
@@
 
-Taken together, these results suggested that the variables included in these stepwise models were relevant to the biology of the obesity metagenes identified in the CR and FM data sets.
-However, there were no single common pathway that was present in all of the models; although Akt, \gls{egfr} and Src pathway metagenes were frequently included in many of the stepwise models.
+Taken together, these results suggest that the variables included in these stepwise models were relevant to the biology of the obesity metagenes identified in the CR and FM data sets.
+However, there were no common pathways that were present in all of the models; although Akt, \gls{egfr} and Src pathway metagenes were frequently included in many of the stepwise models.
 The role of the pathways included in the models and the possible biological links of these pathways to the obesity metagenes remain unclear, and should be the focus for future investigations.

### 2017-03-28 thesis/intro/intro.tex
tmp commit

@@
 
-Obesity has been one of the major global problems for more than a decade, associated with many noncommunicable diseases such as diabetes, cardiovascular diseases and certain types of cancers \citep{WHO2014}.
-In fact, the risk of comorbidities increases with the increase in \gls{bmi}, where the risk becomes severe as the \gls{bmi} level approaches the obese category \citep{WHO2000}.
-The number of overweight and obese people, in both adults and children, have risen in every country of the world and the trend will likely to continue in the future.
-Estimated to account for 3.4 million deaths per year and 93.6 million \glspl{daly} in 2010, obesity is a serious disease that continues to grow in our society \citep{Lim2012}.
+Obesity has been a major global problem for more than a decade, associated with many noncommunicable diseases such as diabetes, cardiovascular diseases and certain types of cancers \citep{WHO2014}.
+In fact, the risk of comorbidities increases with an increase in \gls{bmi}, where the risk becomes severe as the \gls{bmi} level approaches the obese category \citep{WHO2000}.
+The number of overweight and obese people, in both adults and children, has risen in every country of the world and the trend will likely to continue in the future.
+It is estimated to account for 3.4 million deaths per year and 93.6 million \glspl{daly} in 2010, obesity is a serious disease that continues to grow in our society \citep{Lim2012}.
 
@@
 
-Obesity is defined as an abnormal or excessive fat accumulation that may impair health of that individual \citep{Garrow1988}.
-One common and widely used approach to categorise individuals is by measuring the \gls{bmi} of the individual.
-\gls{bmi} is a measurement based on the weight-to-height ratio of an individual that is often used by clinicians and epidemiologists to classify adults into underweight, normal weight, overweight or obese categories.
-The unit of \gls{bmi} is defined as the weight in kilograms per square of the height in meters (kg/m$^2$).
+Obesity is defined as an abnormal or excessive fat accumulation that may impair the health of an individual \citep{Garrow1988}.
+One common and widely used approach to categorise obesity is to measure the \gls{bmi} of the individual.
+\gls{bmi} is a measurement based on the weight-to-height ratio of an individual.
+It is often used by clinicians and epidemiologists to classify adults into underweight, normal weight, overweight, or obese categories.
+The unit of \gls{bmi} is defined as the weight in kilograms per square of the height in metres (kg/m$^2$).
 
-\gls{who} (2014) have used \gls{bmi} measurements to define overweight and obesity in adults as individuals with \gls{bmi} $\geq$ 25 kg/m$^2$ and $\geq$ 30 kg/m$^2$, respectively.
-For children under the age of 5, overweight and obesity are defined as weight-for-height greater than 2 standard deviation and 3 standard deviation above \gls{who} Child Growth Standards median, respectively.
-For children aged between 5 to 19 years, overweight and obesity are defined as \gls{bmi}-for-age greater than 1 standard deviation and 2 standard deviation above \gls{who} Growth Reference Standards median, respectively.
+The \gls{who} (2014) categorise overweight and obese adults  with a \gls{bmi} of  $\geq$ 25 kg/m$^2$ and $\geq$ 30 kg/m$^2$ respectively.
+For children under the age of 5, overweight and obesity is categorised as a  weight-for-height ratio greater than 2 standard deviation and 3 standard deviation above the \gls{who} Child Growth Standards median respectively.
+For children aged between 5 to 19 years, overweight and obesity is defined as \gls{bmi}-for-age greater than 1 standard deviation and 2 standard deviation above the \gls{who} Growth Reference Standards median respectively.
 
 However, the use of \gls{bmi} as an accurate measure of body fat composition and/or representation of the metabolic state of an individual remains controversial.
-In fact, many studies have suggested the use of other measurements such as waist-to-height and waist-to-hip ratios to classify the patients into different categories, as these measurements were better predictors of the diseases related to obesity than \gls{bmi} \citep{Dalton2003,Gelber2008,Lee2008}.
-With that said, \gls{bmi} is still widely used due to the ease of data collection compared to other measurements and, as a result, most of the publicly available clinical data have  height and weight data of the patients from which one can calculate the \gls{bmi} information.
+In fact, many studies have suggested that the use of other measurements, such as waist-to-height and waist-to-hip ratios, to classify the patients into different categories, as these measurements were better predictors of the diseases related to obesity than \gls{bmi} \citep{Dalton2003,Gelber2008,Lee2008}.
+With that said, \gls{bmi} is still widely used due to the ease of data collection compared to other measurements.
+As a result, most of the publicly available clinical data has height and weight data of the patients from which the \gls{bmi} can be calculated.
+% With that said, \gls{bmi} is still widely used due to the ease of data collection compared to other measurements and, as a result, most of the publicly available clinical data have  height and weight data of the patients from which one can calculate the \gls{bmi} information.
 
@@
 
-From the latest global status report on noncommunicable diseases by \citet{WHO2014}:
+From the latest global status report on noncommunicable diseases by the \citet{WHO2014}:
 \begin{quote}
 	\textit{
-		In 2014, 39\% of adults aged 18 years and older (38\% of men and 40\% of women) were overweight.
+		``In 2014, 39\% of adults aged 18 years and older (38\% of men and 40\% of women) were overweight.
 		The worldwide prevalence of obesity nearly doubled between 1980 and 2014.
 		In 2014, 11\% of men and 15\% of women worldwide were obese.
-		Thus, more than half a billion adults worldwide are classed as obese.
+		Thus, more than half a billion adults worldwide are classed as obese.''
 	}
@@
 \noindent
-This equated to approximately 1 in every 14 people in the world classed as obese in 2014, and therefore, 1 in 14 people around the world were also at severe health risks associated with obesity.
-The highest prevalence of overweight and obesity was in the American regions, where 61\% and 27\% of the population were overweight and obese, respectively \citep{WHO2014}.
+Approximately 1 in every 14 people in the world were classed as obese in 2014; therefore, 1 in every 14 people around the world were also at severe health risks associated with obese people.
+The highest prevalence of overweight and obesity was in the American regions, where 61\% and 27\% of the population were overweight and obese respectively \citep{WHO2014}.
 It was also noted in the report that women were more likely to be obese than men, and that the prevalence of overweight and obesity increased as the income level of the countries increased \citep{WHO2014}.
 
-In addition to this, the worldwide prevalence of childhood overweight and obesity has been increasing steadily since 2000 \citep{WHO2014}.
-Furthermore, the prevalence of overweight in children under 5 years was estimated to reach from 6.3\% in 2013 to 11\% by 2025 if the current trend continued \citep{WHO2014}.
-The current obesity status of the adult population is bad already, but the fact that more of the younger population who are overweight and obese poses a much greater risk for the health of the population in the future.
+In addition to this, the worldwide prevalence of childhood obesity has been increasing steadily since 2000 \citep{WHO2014}.
+The prevalence of overweight children in those under 5 was estimated to rise from 6.3\% in 2013 to 11\% by 2025 if the current trend continued \citep{WHO2014}.
+% The current obesity status of the adult population is high already, but the fact that more of the younger population who are overweight and obese poses a much greater risk for the health of the population in the future.
+The current prevalence of obesity in the younger population poses an even greater risk to the health of the population in the future than the current prevalence of obesity in the adult population.
 
-In \gls{nz}, obesity is a serious problem as approximately one third (31.6\%) of the adults in \gls{nz} were classified as obese in 2015/2016 \citep{Health2016}.
+In \gls{nz}, obesity is a serious problem as approximately one third (31.6\%) of adults in \gls{nz} were classified as obese in 2015/2016 \citep{Health2016}.
 The Pacific and M\=aori populations were affected the most, where 67\% and 47\% of the adults, respectively, were obese \citep{Health2016}.
-\gls{nz} not only have a high prevalence of obesity compared to many of the other countries of the world, but also have a unique health problem that affects the Pacific and M\=aori populations more than the rest of the \gls{nz} population.
+\gls{nz} not only has a high prevalence of obesity compared to many other countries of the world, but also has a unique health problem that affects Pacific and M\=aori populations more than the rest of the \gls{nz} population.
 
@@
 The key to the problem is the control of food intake and energy expenditure, and how they are regulated in the body.
-The role of hypothalamus is central to the control of satiety and many of the mutations that cause obesity affects the satiety signals to and/or from hypothalamus.
-The hypothalamus receives both neuronal and hormonal inputs from the body and propagates the signal to the downstream neurons that ultimately affect the food intake of the individual \citep{Bell2005, Spiegelman2001}.
+The role of the hypothalamus is central to the control of satiety and many of the mutations that cause obesity affects the satiety signals to and/or from hypothalamus.
+The hypothalamus receives both neuronal and hormonal inputs from the body and propagates the signal to the downstream neurons which ultimately affects the food intake of the individual \citep{Bell2005, Spiegelman2001}.
 
@@
 Insulin, perhaps the most well-known hormone for its role in glucose homeostasis, also has an important role in satiety and food intake regulation.
-Briefly, when insulin binds to the insulin receptor, \gls{irs} protein family is recruited and phosphorylated by the  intrinsic tyrosine kinase activity of the insulin receptor \citep{Saltiel2002}.
+In brief terms, when insulin binds to the insulin receptor, \gls{irs} protein family is recruited and phosphorylated by the  intrinsic tyrosine kinase activity of the insulin receptor \citep{Saltiel2002}.
 The phosphorylated \gls{irs} protein then activates \gls{pi3k}, which in turn activates its downstream targets, causing a rapid signal cascade through the cell \citep{Saltiel2002}.
@@
 However, leptin, and subsequently insulin, was shown to activate ATP-dep\-endent K$^+$ channel in glucose-responsive neurons in the hypothalamus \citep{Spanswick1997, Spanswick2000}.
-The activation of K$^+$ channel provided evidence for the two pathways to act similarly on the cell, at least for an acute response, suggesting a convergence in the pathway response by the two hormones, and therefore their role in satiety control.
+The activation of the K$^+$ channel provided evidence for the two pathways to act similarly on the cell, at least for an acute response, suggesting a convergence in the pathway response by the two hormones, and therefore their role in satiety control.
 
 Satiety control is crucial in order to manage  and maintain the body weight in a healthy state.
-As described above, there are many components that regulate satiety and disruption of any part of this satiety controlling pathway can encourage the development of obesity.
+As described above, there are many components that regulate satiety.
+Disruption of any part of the satiety controlling pathway can encourage the development of obesity.
 
@@
 Of those, many of the \glspl{gwas} published in 2007 showed strong evidence of the \gls{fto} gene with \gls{bmi} and obesity  \citep{Dina2007,Frayling2007,Gerken2007,Scuteri2007}.
-Until these studies came out, there has been no common genes or genomic regions that associated with obesity that was reproducible and reliable \citep{Frayling2007}.
+Until these studies came out, there has been no common genes or genomic regions associated with obesity that was reproducible and reliable \citep{Frayling2007}.
 Thus, these studies were the first evidence of common variants that contributed to the polygenic obesity phenotype in the population.
@@
 For this reason, obesity is thought to be caused mainly by the environmental factors that act on a population-level.
-The first key factor that contributes to the global rise in the proportion of the population that is obese is the availability and abundance of certain food in the country.
-With the help from the government trade policies, food and other goods have become easier to trade between countries, helping the economic growth of many countries around the world \citep{Kearney2010}.
-This led to greater abundance of food in general, greater number of choices of food and shifted the overall nutritional status of the country \citep{Malik2013}.
-This shift in the food choice and availability (and the economic benefit associated with the trades) had a positive impact on the country, but the growth in the food market also resulted in the population-wide rise in obesity.
+The first key factor that contributes to the global rise in the proportion of the population that is obese is the availability and abundance of certain foods in the country.
+With the help from government trade policies, food and other goods have become easier to trade between countries, helping the economic growth of many countries around the world \citep{Kearney2010}.
+This led to a greater abundance of food in general, a greater number of food choices and shifted the overall nutritional status of countries \citep{Malik2013}.
+This shift in food choice and availability (and the economic benefit associated with the trades) had a positive impact on countries, but the growth in the food market also resulted in the population-wide rise in obesity.
 
@@
 On the other hand, the prices of fruits and vegetables remain expensive due to the lack of support by the government to lower the cost associated with the production \citep{Malik2013}.
-As a result, more people are inclined to consume food that are cheap, have little nutritional value and high in energy, than the food that are expensive, nutritional and healthy alternatives.
+As a result, more people are inclined to consume foods that are cheap, have little nutritional value and high in energy, than foods that are expensive, nutritional and healthy alternatives.
 Thus, the population is more likely to become obese by making poor food choice.
 
-The second factor to consider is the behavioural changes led by the urbanisation of the population.
-As more people take up urban lifestyle, they are exposed to the new environment where there is a better range of food selection, as well as technological and mechanical advancement in transport and other daily chores that may not be available in rural areas.
-Although there are many advantages involved with urbanisation, such as access to developed health care systems, education and advanced technologies, the lifestyle change also impose negative health behaviour that ultimately lead to positive energy balance and therefore obesity \citep{Malik2013}.
-Lack of physical activity is one of the obvious components that is linked with the urbanisation of the population.
-In many urban cities, the recommended level of physical exercise is often difficult to achieve and maintain due to the shift towards the sedentary behaviours that are encouraged by the developed transport system, automated household chores and indoor entertainment \citep{Malik2013}.
+The second factor to consider is the behavioural changes led by urbanisation.
+As more people take up urban lifestyle, they are exposed to new environments where there is a better range of food selection, as well as technological and mechanical advancement in transport and other daily chores that may not be available in rural areas.
+Although there are many advantages involved with urbanisation, such as access to developed health care systems, education and advanced technologies, the lifestyle change also imposes negative health behaviour that ultimately can lead to positive energy balance and obesity \citep{Malik2013}.
+Lack of physical activity is one of the obvious components linked with urbanisation.
+In many urban cities, the recommended level of physical exercise is often difficult to achieve and maintain due to the shift towards sedentary behaviours that are encouraged by developed transport systems, automated household chores and indoor entertainment \citep{Malik2013}.
 
 The last factor to be mentioned here is the income and socioeconomic status of the population.
-As the average income increases, more people are able to buy food and are likely to take up the sedentary lifestyle.
-From this, one would expect the high-income group to have the largest impact from obesity, but on the contrary to this expectation, the biggest impact from obesity is seen in the low- and middle-income groups of the population \citep{Malik2013}.
-Access to the developed health care system will allow weight maintenance of the population, but this is likely to be of most benefit for the high-income group in the population.
-The reason for this, quite simply, is the cost associated with the visit to the health care system, as high-income group is more likely to make a consistent and periodic visit than the low- or middle-income groups of the population \citep{Malik2013}.
+As the average income increases, more people are able to buy food and are likely to take up a sedentary lifestyle.
+From this, one would expect the high-income group to have the largest impact from obesity, but on the contrary, the biggest impact from obesity is seen in the low- and middle-income groups \citep{Malik2013}.
+Access to a developed health care system will allow weight maintenance, but this is likely to be of most benefit to high-income groups.
+The reason for this is the cost associated with a visit to the health care system, as high-income groups are more likely to make a consistent and periodic visit than a low- or middle-income groups \citep{Malik2013}.
 
-So, for the low- and middle-income groups, the rise in the average income in these groups allow the people to have access to food and activities that promotes sedentary behaviours, but the limited access to the health care system, education and recreational activities that promote physical activities ultimately lead to positive energy balance \citep{Malik2013}.
-In contrast, the high-income group of the population have better access to the facilities that promote weight maintenance, thus have less of an impact compared with the low- and middle-income groups of the population.
+So, for the low- and middle-income groups, the rise in the average income in these groups allows people to have access to food and activities that promotes sedentary behaviours, but the limited access to the health care system, education and recreational activities (that promote physical activity) ultimately leads to positive energy balance \citep{Malik2013}.
+In contrast, the high-income group of the population have better access to facilities that promote weight maintenance, thus have less of an impact compared with the low- and middle-income groups of the population.
 All of these population-level environmental factors in synergy with the individual-level genetic factors have contributed to the global epidemic of obesity.
@@
 
-\citet{Bray2013} estimated the global prevalence of cancer in adult population in 2008.
+\citet{Bray2013} estimated the global prevalence of cancer in the adult population in 2008.
 Since there was no clear definition for cancer prevalence, \citet{Bray2013} used \gls{ldp} as an estimate for cancer prevalence, and measured the incidence of cancer during the period between 2004 and 2008 (5 years).
@@
 Looking at the prevalence of cancer by specific types, breast cancer was the most prevalent type of cancer with 5.2 million cases in 2008 \citep{Bray2013}.
-Colorectal and prostate cancers were second and third most prevalent cancer types, with approximately 3.2 million cases for  both of them \citep{Bray2013}.
-\citet{Bray2013} noted that female-specific cancer types were the most prevalent types of cancer in 75\% of the countries in the world.
-Furthermore, the prevalence of breast, cervix or prostate cancer is the highest in almost 95\% of the countries \citep{Bray2013}.
+Colorectal and prostate cancers were the second and third most prevalent cancer types, with approximately 3.2 million cases \citep{Bray2013}.
+\citet{Bray2013} noted that female-specific cancer types were the most prevalent types of cancer in 75\% of countries in the world.
+Furthermore, the prevalence of breast, cervix or prostate cancer is the highest in almost 95\% of countries \citep{Bray2013}.
 Although these estimates provide a relatively crude measure of cancer prevalence, \citet{Bray2013} provided a useful overview of the prevalence of cancer  around the world.
@@
 The cancer registration rates were higher in females aged 25-54 years than males of the same age group, but the registration rates were significantly higher in males for people who were 60 years or older \citep{Health2016b}.
-Like with obesity, there is a significant difference in the cancer registration rates between the M\=aori and the non-M\=aori populations.
+Like with obesity, there is a significant difference in the cancer registration rates between M\=aori and non-M\=aori.
 The rates were 1.2 and 1.4 times greater in the M\=aori population compared to the non-M\=aori population for males and females, respectively \citep{Health2016b}.
-Again, this difference between the M\=aori and non-M\=aori population is unique to \gls{nz}, and improved understanding of the underlying problem is essential for better treatment plans and outcomes for the M\=aori population.
+Again, this difference between the M\=aori and non-M\=aori is unique to \gls{nz}, and improved understanding of the underlying problem is essential for better treatment plans and outcomes for M\=aori.
 
 The total number of deaths by non-communicable diseases (including cancer) is predicted to increase, reaching 52 million deaths by 2030 \citep{WHO2014}.
-With the prevalence of cancer at such high levels around the world, and with cancer mortality increasing in the future, cancer is a serious burden that the world has to face.
+With the prevalence of cancer at such high levels around the world, and with cancer mortality increasing in the future, cancer is a serious challenge that the world has to face.
 
@@
 Very rarely, some cells are able to survive the crisis phase and acquire the ability to replicate infinitely, and this transition has been termed ``immortalisation'' \citep{Hanahan2011, Wright1989}.
-Without immortalisation, cancer cells are not able to generate a macroscopic tumours that are capable of killing the host organism \citep{Hanahan2000,Hanahan2011}.
+Without immortalisation, cancer cells are not able to generate macroscopic tumours that are capable of killing the host organism \citep{Hanahan2000,Hanahan2011}.
 
@@
 Like normal cells, cancer cells also require access to oxygen and nutrient supplies from the blood.
-However, new blood vessel formation, or angiogenesis, only occur in a handful of occasions, and thus the tumour needs to acquire the ability to stimulate angiogenesis in the microenvironment in which they live in \citep{Hanahan2011}.
+However, new blood vessel formation, or angiogenesis, only occurs in a handful of occasions, and thus the tumour needs to acquire the ability to stimulate angiogenesis in the microenvironment in which they live in \citep{Hanahan2011}.
 
 It is thought that the angiogenic ability of the tumour is gained at an early stage of tumour development by disrupting the ``angiogenic switch'' \citep{Hanahan2011}.
-\Gls{vegfa} is the major growth factor that triggers angiogenesis by signalling through the \acrshort{vegf} receptors (though some receptors have inhibitory effect) \citep{Yancopoulos2000}.
+\Gls{vegfa} is the major growth factor that triggers angiogenesis by signalling through the \acrshort{vegf} receptors (though some receptors have an inhibitory effect) \citep{Yancopoulos2000}.
 Opposing the effect of \Gls{vegfa} is the \gls{tsp1}, a protein that is known to inhibit the process of angiogenesis by affecting the bioavailability of \acrshort{vegf}, as well as promoting apoptosis \citep{Kazerounian2008}.
@@
 \noindent
-One of the prime reasons that cancers are so difficult to treat and have high mortality rate is due to their ability to metastasise to other parts of the body.
-Once metastasised to different parts of the body, the clinician not only has to consider the primary tumour site, but also the second, third, or even fourth tumour sites, which may or may not have similar tumour biology as the primary tumour, and thus have to consider multiple treatment plans to cure the patient.
+One of the prime reasons that cancers are so difficult to treat and have a high mortality rate is due to their ability to metastasise to other parts of the body.
+Once metastasised to different parts of the body, the clinician not only has to consider the primary tumour site, but also the second, third, or even fourth tumour sites, which may or may not have similar tumour biology as the primary tumour, and thus has to consider multiple treatment plans to cure the patient.
 
 Metastasis is a multi-step process that requires the selection and survival of cancer cell(s) that is able to metastasise (``seed''), and is also compatible with the specific target tissue where they prosper (``soil'') \citep{Talmadge2010}.
-Once mobile and invasive, metastatic cancer cell is able to circulate in the vascular system and eventually invade the target organ, where it must undergo \gls{met} and colonise the site \citep{Hanahan2011,Kalluri2009}.
-There are numerous signals, growth factors and pathways that initiate the metastatic phenotype in cancer cell, and it remains an active field of research \citep{Hanahan2011,Kalluri2009}.
+Once mobile and invasive, the metastatic cancer cell is able to circulate in the vascular system and eventually invade the target organ, where it must undergo \gls{met} and colonise the site \citep{Hanahan2011,Kalluri2009}.
+There are numerous signals, growth factors and pathways that initiate the metastatic phenotype in the cancer cell, and it remains an active field of research \citep{Hanahan2011,Kalluri2009}.
 
@@
 
-Until recently, the evolutionary progress of cancers have been thought to be gradual and stepwise, where each driver mutation is acquired over many years, or even decades \citep{Stephens2011}.
-There are now evidence that show otherwise, whereby a single catastrophic event (termed \gls{chromothripsis}) occurs in the chromosome structure, resulting in major structural rearrangements such as inversions, deletions, duplications and translocations \citep{Leibowitz2015,Stephens2011}.
+Until recently, the evolutionary progress of cancers have thought to have been gradual and stepwise, where each driver mutation is acquired over many years, or even decades \citep{Stephens2011}.
+There is now evidence that shows otherwise, whereby a single catastrophic event (termed \gls{chromothripsis}) occurs in the chromosome structure, resulting in major structural rearrangements such as inversions, deletions, duplications and translocations \citep{Leibowitz2015,Stephens2011}.
 Owing to its nature of being a one-off chromosomal disruption, \gls{chromothripsis} (usually) cause disruptions in tumour suppressor genes rather than duplications of oncogenes \citep{Leibowitz2015}.
@@
 In addition to this, areas of the chromosome where major structural rearrangements have occurred can be prone to a regional hypermutation, termed \gls{kataegis} \citep{Leibowitz2015,Nik-Zainal2012}.
-Although the full mechanism in which \gls{kataegis} occur is not yet defined, \gls{aid}/\acrshort{apobec} family proteins, a family of proteins involved in gene editing, may play a major role in \gls{kataegis} \citep{Leibowitz2015,Nik-Zainal2012}.
+Although the full mechanism in which \gls{kataegis} occurs is not yet defined, \gls{aid}/\acrshort{apobec} family proteins, a family of proteins involved in gene editing, may play a major role in \gls{kataegis} \citep{Leibowitz2015,Nik-Zainal2012}.
 It is evident that both \gls{chromothripsis} and \gls{kataegis} contribute to the genomic aberrations and mutations that ultimately encourage the acquisition of many of the tumorigenic phenotypes.
@@
 \Gls{immunoediting} is a process where the immune system successfully eliminates the highly immunogenic cancer cells during surveillance, but fails to eliminate the weakly immunogenic cells and therefore poses a selective advantage on these cells \citep{Hanahan2011,Teng2008}.
-In fact, there have been reports where the cancer cells in the immunodeficient mice not being able to initiate tumour growth in the immunocompetent mice, but the cancer cells from the immunocompetent mice were able to do so in the immunodeficient mice \citep{Hanahan2011}.
+In fact, there have been reports where the cancer cells in the immunodeficient mice were not able to initiate tumour growth in the immunocompetent mice, but the cancer cells from the immunocompetent mice were able to do so in the immunodeficient mice \citep{Hanahan2011}.
 One possible reason for this was that the cancer cells from the immunodeficient mice were not immunoedited to the extent in which the cells from immunocompetent mice were, and thus unable to evade the immune system as efficiently as those cells in the immunocompetent mice \citep{Hanahan2011}.
-Though more experiments must be carried out to confirm its definitive role and mechanisms in which immunoediting contributes to cancer biology, immune system evasion is no doubt a significant factor that contribute to tumour progression.
+Though more experiments must be carried out to confirm its definitive role and mechanisms in which immunoediting contributes to cancer biology, immune system evasion is no doubt a significant factor that contributes to tumour progression.
 \\
@@
 By utilising many, if not all, of these major cancer hallmarks, cancer cells are able to initiate growth, proliferate, invade, metastasise and eventually kill the host if untreated.
-There is an evident need for an efficient way to monitor and treat cancer patients, and these Hallmarks of Cancers will prove to be the best target for effective treatments of cancer in the future.
+There is an evident need for an efficient way to monitor and treat cancer patients, and these Hallmarks of Cancers will prove to be the best target for effective treatment of cancer in the future.
 
@@
 Cancers are developed through many cellular pathways (Hallmarks of Cancers) that are not fully uncovered yet, with either too many or very few causes associated with those pathways.
-Therefore it is difficult to assess how, rather than whether, a given environmental factor causes specific cancer.
+Therefore it is difficult to assess how, rather than whether, a given environmental factor causes a specific cancer.
 With that said, several environmental factors have a strong evidence of causing cancer.
@@
 For current smokers, the relative risk is greater than 1.5 in many cancers, including liver, cervix and kidney, and it can rise up to 8.96 in lung cancer \citep{Gandini2008}.
-Tobacco smoke contain many \gls{carcinogen} which are processed first by \gls{cyp} enzymes (termed meta\-bolic activation), and the product can either be processed further to be detoxified completely and cause no harm to the cell, or alternatively, the product can covalently bind to the \acrshort{dna}, forming \acrshort{dna} adducts \citep{Hecht1999}.
+Tobacco smoke contains many \glspl{carcinogen} which are processed first by \gls{cyp} enzymes (termed meta\-bolic activation), and the product can either be processed further to be detoxified completely and cause no harm to the cell, or alternatively, the product can covalently bind to the \acrshort{dna}, forming \acrshort{dna} adducts \citep{Hecht1999}.
 Unless these adducts are resolved correctly by the \acrshort{dna} repair mechanisms, it can result in a permanent mutation within the genome and lead to oncogenesis \citep{Hecht1999}.
 
-Exposure to \gls{uv} radiation has been significantly associated with the increased risks of skin cancers \citep{Armstrong2001,Gallagher2006}.
+Exposure to \gls{uv} radiation has been significantly associated with an increased risks of skin cancers \citep{Armstrong2001,Gallagher2006}.
 Similar to smoking, the mechanism of how \gls{uv} radiation may cause  skin cancers is related to the \acrshort{dna} repair system.
@@
 These dimers introduce bulky structures in the \acrshort{dna} that prevent the \acrshort{dna} replication and transcription enzymes from functioning properly, and it may cause lasting mutations if not repaired \citep{Friedberg2003,Hoeijmakers2001}.
-These damages caused by \gls{uv} radiation can be relieved by the \gls{ner} system or the photoreactivation process, where a photoreactivating enzyme uses light to monomerize the dimeric pyrimidines \citep{Friedberg2003}.
+The damage caused by \gls{uv} radiation can be relieved by the \gls{ner} system or the photoreactivation process, where a photoreactivating enzyme uses light to monomerize the dimeric pyrimidines \citep{Friedberg2003}.
 
@@
 Traditionally, the treatment of cancer has been surgery, radiation therapy and chemotherapy.
-Though surgery is an effective method to remove the tumour from the patient, there is always a risk of metastasis and recurrence, which is why a combination of surgery with other treatments like chemotherapy is usually considered.
-Earlier chemotherapeutic drugs and radiation therapy had limited effect on tumours due to their resistance against these treatments \citep{Wilhelm2006}.
-However, better understanding of the pathways and mechanisms used by the tumour cells have shed  light on the development of drugs that are more effective against tumours.
+Though surgery is an effective method to remove the tumour from the patient, there is always risk of metastasis and recurrence, which is why a combination of surgery with other treatments like chemotherapy is usually considered.
+Previously, chemotherapeutic drugs and radiation therapy had a limited effect on tumours due to their resistance against these treatments \citep{Wilhelm2006}.
+However, a better understanding of the pathways and mechanisms used by the tumour cells have shed  light on the development of drugs that are more effective against tumours.
 RAF inhibitors are among one of the many chemotherapeutic drugs.
@@
 Development of drugs that target  the downstream RAF and its mutant forms (such as BRAF$^{V600E}$) have been successful, and many RAF inhibitor drugs such as sorafenib and vemurafenib have been approved for clinical use \citep{Samatar2014,Wilhelm2006}.
-Even though the signalling pathways are well understood and the quality of the drugs that are developed based on these knowledge is improving, there is always the possibility of the tumour to acquire resistance to those drugs \citep{Samatar2014}.
+Even though the signalling pathways are well understood and the quality of the drugs that are developed based on this knowledge is improving, there is always the possibility of the tumour to acquire resistance to those drugs \citep{Samatar2014}.
 
-In fact, there have been evidence of pathway cross-talks between the RAS-ERK pathway and the \gls{pi3k}/Akt pathway, which may help the tumour cells achieve resistance to the  drugs that block the RAS-ERK pathway \citep{Moelling2002,Zimmermann1999a}.
+In fact, there is evidence of pathway cross-talks between the RAS-ERK pathway and the \gls{pi3k}/Akt pathway, which may help the tumour cells achieve resistance to the  drugs that block the RAS-ERK pathway \citep{Moelling2002,Zimmermann1999a}.
 A study in a \textit{Drosophila} model showed that the inhibition of just a single pathway does not necessarily stop the effect of the pathway, as the signal can flow through other pathways \citep{Dar2012}.
@@
 
-Further understanding of the tumour biology, its mechanism of resistance to drugs, and the effect of targeting multiple pathways (or targeting effective pathways for a patient) will definitely become important in the near future, both for novel drug development and for better clinical treatment plans for the patients.
+Further understanding of the tumour biology, its mechanism of resistance to drugs, and the effect of targeting multiple pathways (or targeting effective pathways for a patient) will become important in the near future, both for novel drug development and for better clinical treatment plans for the patients.
 
@@
 As mentioned briefly in \cref{sub:causes_and_risk_factors_of_cancer}, obesity is considered a major risk factor for many types of cancer.
-The prevalence of obesity is extremely high around the world and  it is important to establish the link between the apparent risks associated with cancer.
+The prevalence of obesity is extremely high around the world and  it is important to establish the link between the risks associated with cancer.
 
@@
 Although the mechanisms by which obesity causes cancer are still under debate, the association between obesity and cancer has long been established.
-Perhaps the most convincing evidence of the association between obesity and cancer was first presented by \citet{Calle2003}.
-In this study, the link between \gls{bmi} and variety of cancers were investigated in 900,000 adults between 1982 and 1998 \citep{Calle2003}.
-The results from this study showed that morbidly obese men and women (\gls{bmi} \textgreater{} 40) had 52\% and 62\% higher chance of dying from all cancers compared to normal weight adults (18.5 \textless{} \gls{bmi} \textless{} 25), respectively \citep{Calle2003}.
+Perhaps the most convincing evidence was first presented by \citet{Calle2003}.
+In this study, the link between \gls{bmi} and variety of cancers was investigated in 900,000 adults between 1982 and 1998 \citep{Calle2003}.
+The results from this study showed that morbidly obese men and women (\gls{bmi} \textgreater{} 40) had a 52\% and 62\% higher chance of dying from all cancers compared to normal weight adults (18.5 \textless{} \gls{bmi} \textless{} 25) \citep{Calle2003}.
 \gls{bmi} was also associated with higher rates of death from esophageal, colorectal, liver, gallbladder, pancreas, kidney, and some sex-specific cancers \citep{Calle2003}.
@@
 
-The exact mechanisms by which obesity causes cancer have still not been fully elucidated yet.
+The exact mechanisms by which obesity causes cancer has still not been fully elucidated yet.
 There are several hypotheses that link obesity with cancer: disruption of various hormone levels, chronic and sustained inflammation, and the activation of the insulin/\gls{igf} pathway \citep{Lumeng2011,Roberts2010}.
 
-Adipose tissue has long been known to be a group of cells to store excess fats, but it has recently become apparent that adipose tissue is also a metabolically active endocrine organ \citep{Roberts2010}.
+Adipose tissue has long known to have been a group of cells to store excess fats, but it has recently become apparent that adipose tissue is also a metabolically active endocrine organ \citep{Roberts2010}.
 Increased adiposity is known to influence the bioavailability of endogenous sex hormones, including estrogens, androgens and progesterone \citep{Calle2004}.
@@
 Adiponectin is a hormone that is synthesised and secreted only by the \glspl{adipocyte} \citep{Kelesidis2006}.
-Unlike leptin, where the concentration of leptin is higher in those who are obese, adiponectin levels are lower in people that are obese compared to those with normal weight \citep{Kelesidis2006,Renehan2006}.
+Unlike leptin, where the concentration of leptin is higher in those who are obese, adiponectin levels are lower in people that are obese compared to those with a normal weight \citep{Kelesidis2006,Renehan2006}.
 Adiponectin have anti-angiogenic and anti-proliferative effects on the cells, and is secreted by mature \glspl{adipocyte} but not by the premature pre-adipocytes \citep{Gilbert2013}.
@@
 
-Alternatively, or possibly in conjunction with the disruption of hormone levels, inflammation caused by obesity may be associated with tumour-promoting environment.
+Alternatively, or possibly in conjunction with the disruption of hormone levels, inflammation caused by obesity may be associated with a tumour-promoting environment.
 \Gls{tlr}4 is a type of \gls{prr} and, as mentioned earlier, is able to recognise circulating free fatty acids and trigger immune response via \gls{nfkb} pathway \citep{Lumeng2011}.
-In addition to \gls{nfkb} pathway, recent evidence have shown that \gls{tlr}4 could also activate the \gls{jak}/\gls{stat3} pathway \citep{Yu2014}.
+In addition to \gls{nfkb} pathway, recent evidence has shown that \gls{tlr}4 could also activate the \gls{jak}/\gls{stat3} pathway \citep{Yu2014}.
 Activation of these pathways will further enhance the development of the tumour-promoting microenvironment caused by these unwanted inflammatory responses brought upon by obesity.
-Furthermore, since it is likely that the people who are obese have less adiponectin concentration, obesity-induced inflammation may affect the patients that are obese more than those with normal weight.
+Furthermore, since it is likely that people who are obese have less adiponectin concentration, obesity-induced inflammation may affect the patients that are obese more than those with a normal weight.
 
@@
 The reason it is so difficult to identify the underlying biological mechanisms is that there are very few reliable biological markers (or ``genetic signatures'') that signify the presence or absence of the causal pathways.
-With that said, significant improvements have been made in recent years owing to the advancement in the sequencing technologies and have allowed the researchers to study the genetic signatures that represent the hallmarks in a much greater detail.
+With that said, significant improvements have been made in recent years owing to the advancement in the sequencing technologies and have allowed researchers to study the genetic signatures that represent the hallmarks in much greater detail.
 
@@
 One disadvantage of the microarray technology is that it requires previous knowledge of the genes and \glspl{snp} of the samples being investigated \citep{Hurd2009}.
-This is due to the fact that the detection of sequences rely on the hybridisation of the sample sequences with the microarray sequences and therefore the sequences under investigation must be known prior to the investigation to attach the correct sequences to the microarray.
+This is due to the fact that the detection of sequences relies on the hybridisation of the sample sequences with the microarray sequences and therefore the sequences under investigation must be known prior to the investigation to attach the correct sequences to the microarray.
 Another disadvantage of microarrays is the potentially erroneous detection of the sample sequences, both in terms of its sequence and its expression level \citep{Hurd2009}.
@@
 
-\Gls{ngs} technology refers to the high throughput sequencing techniques that allow sequencing of large quantity of genetic material \citep{Metzker2010}.
+\Gls{ngs} technology refers to the high throughput sequencing techniques that allows the sequencing of a large quantity of genetic material \citep{Metzker2010}.
 \gls{ngs} technology has been developed and improved in the last decade at a revolutionary pace, making it possible for researchers to study gene expressions and variations in days to weeks, where the same experiment would have taken years to complete before the technology was developed.
@@
 \gls{ngs} technology has many advantages over microarray technology.
-Firstly, \gls{ngs} does not require any prior knowledge of the sample like in the microarray technology, as all genetic materials are attached to universal primers when they are sequenced; there is no need to pre-define the sequences to be searched and thus allow the detection of novel genes or transcripts \citep{Hurd2009}.
-This also prevents the introduction of experimental bias, as the sequences are not altered in any way and no specific sequence is searched in a greater detail over another \citep{Hurd2009}.
+Firstly, \gls{ngs} does not require any prior knowledge of the sample like in the microarray technology, as all genetic materials are attached to universal primers when they are sequenced; there is no need to pre-define the sequences to be searched and thus allows the detection of novel genes or transcripts \citep{Hurd2009}.
+This also prevents the introduction of experimental bias, as the sequences are not altered in any way and no specific sequence is searched in any greater detail over another \citep{Hurd2009}.
 Another advantage is that \gls{ngs} produce count data of the genetic material being sequenced instead of relying on the intensity of the fluorescence to detect the presence and amount of certain sequence.
 This allows an absolute quantification of the amount of transcripts from the sample, making it possible for a more accurate comparison of the expression levels between samples.
-Lastly, many \gls{ngs} techniques are capable of sequencing multiple samples simultaneously with high throughput, generating tremendous amount of genetic data in a single run.
+Lastly, many \gls{ngs} techniques are capable of sequencing multiple samples simultaneously with high throughput, generating a tremendous amount of genetic data in a single run.
 
-The use of \gls{ngs} technology has significantly improved the workflow of modern genetic studies and its application in variety of research area have made it possible to dig deeper into the causes of complex diseases, such as cancer.
+The use of \gls{ngs} technology has significantly improved the workflow of modern genetic studies and its application in a variety of research areas has made it possible to dig deeper into the causes of complex diseases, such as cancer.
 With the cost of sequencing declining, personalised genome sequencing and treatment based on the sequence will be a reality in the near future.
@@
 There is clear epidemiological evidence of obesity being a major risk factor for many cancer types (\cref{sec:obesity_and_cancer}), but the genes or pathways that directly contribute to tumorigenesis in patients that are obese is yet to be clarified.
-Two studies in particular, one by \citet{Creighton2012} and another by \citet{Fuentes-Mattei2014}, have investigated whether there were any difference in genetic composition between the obese and the non-obese patients, specifically in a group of breast cancer patients.
+Two studies in particular, one by \citet{Creighton2012} and another by \citet{Fuentes-Mattei2014}, have investigated whether there were any differences in the genetic composition between obese and non-obese patients, specifically in a group of breast cancer patients.
 In both studies, the authors were able to identify a set of genes that were up or downregulated in breast tumours from the patients that were obese.

### 2017-03-28 thesis/misc/appendix2.tex
tmp commit
finished editing up to the discussion

@@
-\begin{appendices}
-	\crefalias{chapter}{appcha}
-	\crefalias{section}{appsec}
-	\renewcommand{\thesection}{\Alph{chapter}\arabic{section}}
+% \begin{appendices}
+% 	\crefalias{chapter}{appcha}
+% 	\crefalias{section}{appsec}
+% 	\renewcommand{\thesection}{\Alph{chapter}\arabic{section}}
 
@@
 
-	\section{Comparison of the correlation of \gls{svd} and TM generated pathway metagenes}
+	\section{Visual comparison of the correlation of \gls{svd} and TM generated pathway metagenes}
 	\label{sec:comparison_of_the_correlation_of_svd_and_tm_generated_pathway_metagenes}
@@
 		\begin{threeparttable}
-			\begin{tabular}{llr{\bfseries}S}
+			\begin{tabular}{llrc}
 				Linear Model & Variables & Estimate & P-value\\
@@
 		\begin{threeparttable}
-			\begin{tabular}{llr{\bfseries}S}
+			\begin{tabular}{llrc}
 				Linear Model & Variables & Estimate & {P-value}\\
@@
 		\begin{threeparttable}
-			\begin{tabular}{llr \bfseries S}
+			\begin{tabular}{llrc}
 				Linear Model & Variables & Estimate & {P-value}\\
@@
 
+	\begin{ThreePartTable}
 			\begin{TableNotes}
@@
 
-	\begin{longtable}{llc{\bfseries}S}
+			\begin{longtable}{llcc}
 				\centering
@@
 
+	\end{ThreePartTable}
+
 	\begin{table}[htpb]
@@
 
-	\begin{longtable}{llc{\bfseries}S}
+	\begin{ThreePartTable}
+			\begin{TableNotes}
+				\begin{footnotesize}
+				\item [1] All values in bold are statistically significant (p \textless{} 0.05).
+				\end{footnotesize}
+			\end{TableNotes}
+
+			\begin{longtable}{llcc}
 				\centering
@@
 			\end{longtable}
+	\end{ThreePartTable}
 

### 2017-03-28 thesis/results/results.tex
tmp commit

@@
 	\includegraphics[page=3,width=0.8\linewidth]{results1/creighton_mg_heatmap1}
-	\caption[Obesity metagene from \citet{Creighton2012} study and sample gene expression in CR data] {Heatmap showing the obesity metagene from the \citet{Creighton2012} study with the sample gene expressions of the obesity associated genes in the CR data.
+	\caption[Obesity metagene from the \citet{Creighton2012} study and sample gene expression in CR data] {Heatmap showing the obesity metagene from the \citet{Creighton2012} study with the sample gene expressions of the obesity associated genes in the CR data.
 	Level of expression is represented in the top right histogram, where low and high gene expression are colour-coded with blue and red, respectively.

### 2017-03-28 thesis/results/results2.tex
tmp commit

@@
 	\includegraphics[page=100,width=0.45\linewidth]{results2/allmeta_rma_vs_mas}
-	\caption[Comparison of the pathway metagenes generated in the \acrshort{mas} normalised GT data set from the application of the transformation matrices derived from either the \acrshort{rma} or \acrshort{mas} normalised GT data]{Comparison of the pathway metagenes generated in the \acrshort{mas} normalised GT data set from the application of the transformation matrices derived from either the \acrshort{rma} or \acrshort{mas} normalised GT data.
+	\caption[Comparison of the pathway metagenes generated in the \acrshort{mas} normalised GT data set from the application of TMs, derived from either the \acrshort{rma} or \acrshort{mas} normalised GT data]{Comparison of the pathway metagenes generated in the \acrshort{mas} normalised GT data set from the application of the transformation matrices derived from either the \acrshort{rma} or \acrshort{mas} normalised GT data.
 	Only the \gls{pr} (left) and the \gls{tgfb}  (right) pathway metagenes are shown.
@@
 	\begin{threeparttable}
-			\begin{tabular}{llr{\bfseries}S}
+			\begin{tabular}{llrS}
 				Linear Model & Variables & Estimate & {P-value}\\

### 2017-03-29 thesis/discussion/discussion.tex
pretty much ready to submit - tmp commit

@@
 
-\cref{cha:obesity_genetic_signatures_and_cancer} focussed mainly on establishing the link between the obesity associated genetic signatures from the \citet{Creighton2012} and the \citet{Fuentes-Mattei2014} studies with the sample \gls{bmi} and \gls{bmi} status, as reported by their studies.
+\cref{cha:obesity_genetic_signatures_and_cancer} focussed mainly on establishing the link between the obesity associated genetic signatures from the \citet{Creighton2012} and the \citet{Fuentes-Mattei2014} studies with the patient \gls{bmi} and \gls{bmi} status, as reported by their studies.
 The obesity associated genetic signature from the \citet{Creighton2012} study was examined first.
-The obesity metagene generated from the obesity signature from the \citet{Creighton2012} study was able to ``capture'' the overall genetic expression patterns of the samples, where low and high metagene scores  corresponded to low and high gene expressions, respectively.
+The obesity metagene generated from the Creighton \textit{et al.} obesity signature  was able to ``capture'' the overall genetic expression patterns of the samples, where low and high metagene scores  corresponded to low and high gene expressions, respectively.
 This result was in accordance with the characteristics of the obesity metagene provided by \citet{Creighton2012}.
-Furthermore, the obesity metagene significantly correlated with the sample \gls{bmi} and \gls{bmi} status in the CR data set.
+Furthermore, the obesity metagene significantly correlated with the patient \gls{bmi} and \gls{bmi} status in the CR data set.
 
-To see whether this obesity signature was able to show similar association with the sample \gls{bmi} and \gls{bmi} status in other cancer data sets, metagenes were created in each of the cancer data sets with the transformation matrix generated in the CR  data set.
-Like in the CR  data set, the obesity metagenes were reflective of the gene expression patterns of the genes in the genetic signature.
+To see whether this obesity signature was able to show similar association with the patient \gls{bmi} and \gls{bmi} status in other cancer data sets, metagenes were created in the \gls{nzbc} and \gls{icgc} cancer data sets with the transformation matrix generated in the CR  data set.
+Like in the CR  data set, the obesity metagenes were reflective of the gene expression patterns of the genes in the genetic signature in all of the data sets.
 However, the obesity metagene was not significantly associated with the patient \gls{bmi} or \gls{bmi} status in any of the other cancer data sets, except for  the \gls{blca} data set which showed significant association only with the overweight group (discussed later in \cref{sub:blca_and_obesity_metagenes}).
@@
 Initially, these results were thought to be due to the obesity associated genetic signature being specific only to the CR data set.
-Therefore, a different obesity associated genetic signature from the \citet{Fuentes-Mattei2014} study was used to see whether this signature was able to show significant association with the patient \gls{bmi} in other cancer data sets.
-The results were similar to the obesity metagene from the \citet{Creighton2012} study; FM obesity metagene scores reflected the gene expressions of the signature but the scores were not significantly associated with the patient \gls{bmi} or \gls{bmi} status in any of the cancer data sets (except the \gls{blca} data set).
+Therefore, a different obesity associated genetic signature from the \citet{Fuentes-Mattei2014} study was used to see whether this signature was able to show significant association with the patient \gls{bmi} in the Cr, \gls{nzbc} and \gls{icgc} cancer data sets.
+The results were similar to the obesity metagene from the \citet{Creighton2012} study; FM obesity metagene scores reflected the gene expressions of the signature but the scores were not significantly associated with the patient \gls{bmi} or \gls{bmi} status in any of the cancer data sets investigated (except the \gls{blca} data set).
 Together with the results from the CR obesity metagene, these results suggested that both CR and FM genetic signatures were only significant in the original data set in which the signature was derived in (though this was not confirmed in the FM data set, as no patient \gls{bmi} information was available).
@@
 This question led to the identification of the various versions of obesity associated genetic signatures in \cref{sub:identification_of_obesity_associated_genetic_signatures} that were controlled by clinical variables (sex, age, ethnicity, menopause status, tumour grade, \gls{er}/\gls{her2}/\gls{pr} statuses and \gls{ln} status) in the CR data set.
-All of the metagenes created from these clinical variable-controlled genetic signatures were consistent with the gene expression patterns, but none of the metagenes showed significant association with the sample \gls{bmi}/\gls{bmi} status in the data set other than the CR data set, with the exception of the overweight group in the \gls{blca} data set.
+All of the metagenes created from these clinical variable-controlled genetic signatures were consistent with the gene expression patterns, but none of the metagenes showed significant association with the patient \gls{bmi} and \gls{bmi} status in the data set other than the CR data set, with the exception of the overweight group in the \gls{blca} data set.
 
-It was possible that all of these results were due to the sample quality (and hence due to the quality of the data), or perhaps many of the samples in these data sets had breast cancers due to a cause unrelated to their obesity status
-As many epidemiological studies have pointed out, obesity is a strong risk factor for cancer (covered in \cref{sec:obesity_and_cancer}), but this does not necessarily mean that all cancer in obese patients developed as a result of obesity.
+It was possible that all of these results were due to the sample quality (and hence due to the quality of the data), or perhaps many of the samples in these data sets had breast cancers due to a cause unrelated to their obesity status.
+As many epidemiological studies have pointed out, obesity is a strong risk factor for cancer (covered in \cref{sec:obesity_and_cancer}), but this does not necessarily mean that all cancer in obese patients develop cancer as a result of obesity.
 In other words, the cancer identified in obese patients may have been caused by another, completely different biological mechanism.
@@
 
-In the \gls{blca} data set, many of the obesity metagenes were associated with the overweight group with significant P-value and/or \gls{anova} P-value, but never with the obese group.
+In the \gls{blca} data set, many of the obesity metagenes were associated with the overweight group with significant p-value and/or \gls{anova} p-value, but never with the obese group.
 This was unexpected, as all of the metagenes were generated based on the list of \glspl{deg} between the obese and the non-obese groups in the CR data set, and not the overweight group.
-In addition to the association of the metagenes with the overweight group, some metagenes showed significant association with sample \gls{bmi} in the \gls{blca} data set.
+In addition to the association of the metagenes with the overweight group, some metagenes showed significant association with patient \gls{bmi} in the \gls{blca} data set.
 It was difficult to conclude with confidence that those metagenes that showed significant association with the overweight group in the \gls{blca} data set was truly due to the effect of the metagenes, as these metagenes were generated from the obese group and not the overweight group in the original data sets.
@@
 
-With that said, there is a possibility that the genotypes of the samples that are overweight in the \gls{blca} data set are similar to those samples that are obese in the CR data set.
+With that said, there is a possibility that the genotypes of the patients that are overweight in the \gls{blca} data set are similar to those patients that are obese in the CR data set.
 In fact, \citet{Damrauer2014} have identified ``basal-like'' and ``luminal''  subtypes in bladder cancer that resembled the corresponding subtypes in breast cancer, and suggested that bladder  and breast cancers may have common physiological properties.
@@
 The results so far have provided strong evidence that the obesity associated genetic signatures from the CR and the FM data sets were not associated with the patient \gls{bmi}.
-In the last attempt to identify any obesity associated genes that were common across multiple cancer types, gene expression analyses were carried out on the \gls{icgc} cancer data sets using samples for which \gls{bmi} data were available (\cref{sec:common_genes_across_multiple_cancer_types}).
+In the last attempt to identify any obesity associated genes that were common across multiple cancer types, gene expression analyses were carried out on the \gls{icgc} cancer data sets using samples for which patient \gls{bmi} information were available (\cref{sec:common_genes_across_multiple_cancer_types}).
 There were many \glspl{deg} identified by each of the \gls{icgc} cancer types, but there were no genes in common for all eight cancer types, and this observation was supported by the results from the simulation.
@@
 
-The fact that there were so many false positives in the \glspl{deg} from the \gls{icgc} data sets indicated that the use of the patient \gls{bmi}/\gls{bmi} status as the ``treatment'' conditions (in other words independent variable) for gene expression analysis was prone to \Glspl{type1}.
-This meant that there were very little evidence of true obesity associated genetic signatures in any of the data sets that have been explored in this project, or perhaps the patient \gls{bmi}/\gls{bmi} status was not enough to pinpoint the underlying biological relationship between obesity and cancer (see \cref{sub:discussion_definition_of_obesity} as well).
+The fact that there were so many false positives in the \glspl{deg} from the \gls{icgc} data sets indicated that the use of the patient \gls{bmi} and \gls{bmi} status as the ``treatment'' conditions (in other words independent variable) for gene expression analysis was prone to \Glspl{type1}.
+This meant that there were very little evidence of true obesity associated genetic signatures in any of the data sets that have been explored in this project, or perhaps the patient \gls{bmi} and \gls{bmi} status was not enough to pinpoint the underlying biological relationship between obesity and cancer (see \cref{sub:discussion_definition_of_obesity} as well).
 
@@
 All of the predicted obesity metagene scores significantly correlated with the true obesity metagene scores in both the \gls{nzbc} (training) and CR (testing) data sets.
+% TODO: check r-squared values
 However, the $R^2$-values for all of these predictions were low (highest $R^2$ $\approx$ 0.42; \cref{sec:summary_statistics_of_the_predicted_obesity_metagenes_with_sample_bmi_bmi_status_in_nzbc_and_cr_data}) with highly variable data points in the scatter plots.
@@
 Since linear models were created only with the ``consistent'' pathway metagene scores, a step wise method was used to generate linear models that included the patient \gls{bmi}, \gls{bmi} status and all of the pathway metagenes that allowed the model to predict the true obesity metagene the best.
+% TODO: check r-squared values
 As shown by the results in \cref{sec:prediction_of_obesity_associated_metagene_with_pathway_associate_metagene}, all of the models predicted the true obesity metagene scores with high $R^2$-value in Creighton's data set (highest $R^2$ $\approx$ 0.88) and were statistically significant.
@@
 But was this due to the true biological relationship between the \gls{pr} pathway and the obesity metagenes, or only because the models were created in the \gls{nzbc} data set?
-In fact, the proportion of the samples that were \gls{pr}$^+$ was greater in the \gls{nzbc} data set (62 \gls{pr}$^+$ and 37 \gls{pr}$^-$ samples) compared with the CR data set (48 \gls{pr}$^+$ and 50 \gls{pr}$^-$ samples), which implied that the patients in the \gls{nzbc} data set was more likely to show association with the \gls{pr} pathway metagene.
+In fact, the proportion of the patients that were \gls{pr}$^+$ was greater in the \gls{nzbc} data set (62 \gls{pr}$^+$ and 37 \gls{pr}$^-$ patients) compared with the CR data set (48 \gls{pr}$^+$ and 50 \gls{pr}$^-$ patients), which implied that the patients in the \gls{nzbc} data set was more likely to show association with the \gls{pr} pathway metagene.
 
-However, with this logic, the \gls{er} pathway signature would as likely to be associated, as the  sample  \gls{er} status was similar to the \gls{pr} status in either data sets: 72 \gls{er}$^+$ and 27 \gls{er}$^-$ samples in the \gls{nzbc} data, and 58 \gls{er}$^+$ and  42 \gls{er}$^-$ samples in the  CR data set.
-The fact that the \gls{er} pathway signature was not significant in the models suggested that the obesity metagenes were related to the \gls{pr} pathway irrespective of the sample \gls{pr} status.
+However, with this logic, the \gls{er} pathway signature would as likely to be associated, as the   patient \gls{er} status was similar to the \gls{pr} status in either data sets: 72 \gls{er}$^+$ and 27 \gls{er}$^-$ patient in the \gls{nzbc} data, and 58 \gls{er}$^+$ and  42 \gls{er}$^-$ patients in the  CR data set.
+The fact that the \gls{er} pathway signature was not significant in the models suggested that the obesity metagenes were related to the \gls{pr} pathway irrespective of the patient \gls{pr} status.
 Furthermore, both the \gls{nzbc} and the CR data sets showed strong association between the \gls{pr} and \gls{er} pathway metagenes with the patient \gls{pr} and \gls{er} status, respectively (\cref{sec:results_for_er_and_pr_metagene_analysis}).
@@
 This leads to the second problem: the different technologies used in the studies.
-All of Creighton \textit{et al.}, Fuentes-Mattei \textit{et al.}, Print \textit{et al.} and Gatza \textit{et al.} data sets were analysed using microarray technology, whereas all of the \gls{icgc} cancer data sets were generated with \gls{ngs} technology.
-As mentioned in \cref{sub:data_normalisation}, the two technologies use fundamentally different principals to output sequence information of the sample; microarray uses light intensity and \gls{ngs} uses count data.
+All of Creighton \textit{et al.}, Fuentes-Mattei \textit{et al.}, \gls{nzbc} and Gatza \textit{et al.} data sets were analysed using microarray technology, whereas all of the \gls{icgc} cancer data sets were generated with \gls{ngs} technology.
+As mentioned in \cref{sub:data_normalisation}, the two technologies use fundamentally different principals to output the sequence information of the sample; microarray uses light intensity and \gls{ngs} uses count data.
 
@@
 
-Firstly, the obesity signatures identified by both Creighton \textit{et al.} and Fuentes-Mattei \textit{et al.} relied on the definition of obesity based on the sample \gls{bmi} values.
+Firstly, the obesity signatures identified by both Creighton \textit{et al.} and Fuentes-Mattei \textit{et al.} relied on the definition of obesity based on the patient \gls{bmi} values.
 As discussed in \cref{sub:discussion_definition_of_obesity}, the use of \gls{bmi} may not have been the best measurement for obesity.
@@
 In this project, the lack of association of the obesity associated genetic signatures with any of the cancer types could have been due to the quality of both the raw data sets and the genetic signatures used in this project.
-Therefore, careful selection of the raw data sets with sufficient, and if possible, additional information (such as waist circumference) about the samples and well-validated genetic signatures will be essential for future investigations.
+Therefore, careful selection of the raw data sets with sufficient, and if possible, additional information (such as waist circumference) about the patients and well-validated genetic signatures will be essential for future investigations.

### 2017-03-29 thesis/intro/intro.tex
pretty much ready to submit - tmp commit

@@
 As a result, most of the publicly available clinical data has height and weight data of the patients from which the \gls{bmi} can be calculated.
-% With that said, \gls{bmi} is still widely used due to the ease of data collection compared to other measurements and, as a result, most of the publicly available clinical data have  height and weight data of the patients from which one can calculate the \gls{bmi} information.
 
@@
 The prevalence of overweight children in those under 5 was estimated to rise from 6.3\% in 2013 to 11\% by 2025 if the current trend continued \citep{WHO2014}.
-% The current obesity status of the adult population is high already, but the fact that more of the younger population who are overweight and obese poses a much greater risk for the health of the population in the future.
 The current prevalence of obesity in the younger population poses an even greater risk to the health of the population in the future than the current prevalence of obesity in the adult population.
@@
 The Pacific and M\=aori populations were affected the most, where 67\% and 47\% of the adults, respectively, were obese \citep{Health2016}.
-\gls{nz} not only has a high prevalence of obesity compared to many other countries of the world, but also has a unique health problem that affects Pacific and M\=aori populations more than the rest of the \gls{nz} population.
+\gls{nz} not only has a high prevalence of obesity compared to many other countries, but also has a unique health problem that affects Pacific and M\=aori populations disproportionately than the rest of the \gls{nz} population.
 
@@
 Though there are many hormones that regulate satiety, only leptin and insulin will be covered here.
-% Leptin gene was first discovered by \citet{Zhang1994} and the discovery of the leptin receptor gene was followed shortly after by \citet{Tartaglia1995}.
-% Together, the identification of the leptin and leptin receptor genes was a major break through in the field of obesity research.
 Leptin is a hormone that signals satiety and reduces food intake, and is secreted mainly by the white adipose tissue \citep{Moustafa2013,Zhang1994}.
@@
 Leptin signals through the \gls{jak}/\gls{stat} pathway via the long form of the leptin receptor that is present in high expression in the hypothalamus \citep{Ghilardi1996,Lee1996}.
-Depending on the type of neurons the leptin receptor is expressed, leptin stimulate or inhibit these neurons and ultimately produces an anorexogenic signal to the downstream effector neurons \citep{Bell2005}.
+Depending on the type of neurons the leptin receptor is expressed on, leptin stimulates or inhibits these neurons and ultimately produces an anorexogenic signal to the downstream effector neurons \citep{Bell2005}.
 
@@
 However, leptin, and subsequently insulin, was shown to activate ATP-dep\-endent K$^+$ channel in glucose-responsive neurons in the hypothalamus \citep{Spanswick1997, Spanswick2000}.
-The activation of the K$^+$ channel provided evidence for the two pathways to act similarly on the cell, at least for an acute response, suggesting a convergence in the pathway response by the two hormones, and therefore their role in satiety control.
+The activation of the K$^+$ channel provided evidence for the two pathways to act similarly on the cell, at least for an acute response.
+This suggests a convergence in the pathway response by the two hormones, and therefore their role in satiety control.
 
@@
 As a result, children with this mutation were constantly hungry due to the lack of anorexogenic signals and became obese at a very early stage of their lives.
-% Though these sort of mutations are very rare and affects only those individuals that have these mutations, it had a profound impact in the field of obesity research and contributed significantly to the knowledge of the mechanism of the development of obesity.
 
 Polygenic obesity is where a combination of variants in multiple genes or multiple regions of the genome (together with the effect from the environmental factors) causes obesity in the individual \citep{Moustafa2013}.
-Unlike in monogenic obesity where a single mutation in a gene causes severe obesity, the variants in some of the genes or genomic regions involved in polygenic obesity have little effect on the obesity status of the individual by itself, but in combination with many other variants and together with environmental factors, it has a serious obesogenic effect on the individual.
+Unlike in monogenic obesity, where a single mutation in a gene causes severe obesity, the variants in some of the genes or genomic regions involved in polygenic obesity have little effect on the obesity status of the individual by itself.
+However, in combination with many other variants and together with environmental factors, it has a serious obesogenic effect on the individual.
 
@@
 Of those, many of the \glspl{gwas} published in 2007 showed strong evidence of the \gls{fto} gene with \gls{bmi} and obesity  \citep{Dina2007,Frayling2007,Gerken2007,Scuteri2007}.
-Until these studies came out, there has been no common genes or genomic regions associated with obesity that was reproducible and reliable \citep{Frayling2007}.
+Until these studies came out, there has been no common genes or genomic regions found to be associated with obesity that was reproducible and reliable \citep{Frayling2007}.
 Thus, these studies were the first evidence of common variants that contributed to the polygenic obesity phenotype in the population.
@@
 The first key factor that contributes to the global rise in the proportion of the population that is obese is the availability and abundance of certain foods in the country.
-With the help from government trade policies, food and other goods have become easier to trade between countries, helping the economic growth of many countries around the world \citep{Kearney2010}.
-This led to a greater abundance of food in general, a greater number of food choices and shifted the overall nutritional status of countries \citep{Malik2013}.
-This shift in food choice and availability (and the economic benefit associated with the trades) had a positive impact on countries, but the growth in the food market also resulted in the population-wide rise in obesity.
+With the help from government free trade policies, food and other goods have become easier to trade between countries, helping the economic growth of many countries around the world \citep{Kearney2010}.
+This has led to a greater abundance of food in general, a greater number of food choices and a change in the overall nutritional status of countries \citep{Malik2013}.
+This shift in food choice and availability (and the economic benefit associated with the trade) had a positive impact on countries, but the growth in the food market also resulted in the population-wide rise in obesity.
 
-As an example, in the \gls{usa}, the costs of corn and soy are low due to the fact that they are the raw ingredients for most processed food and beverages, such as high fructose corn syrups used to sweeten many soft drinks in the \gls{usa} \citep{Malik2013}.
-Moreover, corn and soy are also the main food for the livestock, which results in lower prices for meat \citep{Malik2013}.
+As an example, in the \gls{usa}, the cost of corn and soy are low due to the fact that they are the raw ingredients for most processed food and beverages, such as high fructose corn syrups used to sweeten many soft drinks in the \gls{usa} \citep{Malik2013}.
+Moreover, corn and soy are also the main food for livestock, which results in lower price for meat \citep{Malik2013}.
 On the other hand, the prices of fruits and vegetables remain expensive due to the lack of support by the government to lower the cost associated with the production \citep{Malik2013}.
-As a result, more people are inclined to consume foods that are cheap, have little nutritional value and high in energy, than foods that are expensive, nutritional and healthy alternatives.
+As a result, more people are inclined to consume food that are cheap, have little nutritional value and high in energy, than food that are expensive, nutritional and healthy alternatives.
 Thus, the population is more likely to become obese by making poor food choice.
@@
 The second factor to consider is the behavioural changes led by urbanisation.
-As more people take up urban lifestyle, they are exposed to new environments where there is a better range of food selection, as well as technological and mechanical advancement in transport and other daily chores that may not be available in rural areas.
-Although there are many advantages involved with urbanisation, such as access to developed health care systems, education and advanced technologies, the lifestyle change also imposes negative health behaviour that ultimately can lead to positive energy balance and obesity \citep{Malik2013}.
+As more people take up urban lifestyle, they are exposed to new environments with a better range of food selection, as well as technological and mechanical advancement in transport and other daily chores that may not be available in rural areas.
+Although urbanisation has many advantages, such as access to developed health care systems, education and advanced technologies, the lifestyle change also imposes negative health behaviour that ultimately can lead to positive energy balance and obesity \citep{Malik2013}.
 Lack of physical activity is one of the obvious components linked with urbanisation.

### 2017-03-29 thesis/misc/abstract.tex
pretty much ready to submit - tmp commit

--- 2016-09-17 thesis/misc/abstract.tex
@@
 
-It is now commonly understood that cancers are caused by dysregulation of various mechanisms or pathways that allow tumour cells to prolilferate, survive and migrate.
-There have been many reports on the association between obesity and worsened cancer prognosis, suggesting that there may be distinct changes that underlie tumorigenesis and/or cancer progression that are specific to patients with higher \gls{bmi}.
-
-In brief, this research aims to determine whether gene expression signatures exist that are specific to obesity across multiple cancer types, and to investigate whether there are any common pathways being dysregulated in cancers based on these genetic signatures.
-Better understanding of the pathways being dysregulated in cancer cells in obese patients may lead to improved clinical decisions, and could thus contribute towards personalised treatment in the future.
+Obesity has been a major global problem for more than a decade, associated with many noncommunicable diseases such as  cancer.
+The number of obese people, in both adults and children, has risen in every country of the world and the trend will likely to continue in the future.
+Cancers are caused by dysregulation of various molecular pathways that allow tumour cells to proliferate, survive and migrate.
+One of the difficulties associated with the treatment of cancers is the identification of the underlying biological pathways that drive tumorigenesis.
+This research aims to determine whether gene expression signatures exist  that are specific to obesity across multiple cancer types, and to investigate whether there are any common pathways being dysregulated in cancers based on these genetic signatures.
+There were no genetic signatures or genes differentially expressed between obese and non-obese patients that were common across multiple cancer types.
+However, the Akt, \gls{egfr} and Src pathways may have a role in promoting the tumour progression in patients that are obese.
+It is likely that there is some complex mechanism underlying the relationship between obesity and cancer.
+Better understanding of the pathways being dysregulated in cancer cells in obese patients may lead to improved clinical decisions, and contribute towards personalised treatment in the future.
 

### 2017-03-29 thesis/misc/acknowledgement.tex
pretty much ready to submit - tmp commit

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Acknowledgement for my thesis
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\section*{\centering Acknowledgement}
\addcontentsline{toc}{section}{Acknowledgement}

First and foremost, I would like to thank my supervisor Mik, for giving me this project and an
opportunity to learn and develop the basic bioinformatics skills required to complete it.
You have been a great supervisor and I have learnt a lot from you over the past two years, and I have enjoyed being a part of the Black lab.
\\

\noindent
I would like to thank my committee members, Dr.\ Anita Dunbier, Dr.\ Miriam Sharpe, and Dr.\ Peter Mace (temporary member), for being patient during my presentations that were full of statistics and heatmaps.
I am grateful for all the suggestions and inputs you have given me to make this a better project.
\\

\noindent
I would also like to thank both past and present members of the Black lab, for useful tips and tricks to get me through my Masters.
Especially James Boocock, for introducing me to a variety of useful tool sets (by the way, everyone should start using vim); Murray Cadzow, for organising the fortnightly ``Sh*t You Should Know About'' sessions to learn more about R and other useful tools; and Tom Kelly, for being my companion in the Fish-bowl and helping me with some of the codes in my project.
\\

\noindent
I would like to thank all of my friends and family for all their support and fun memories throughout the years that I have been at the University.
Lastly, a special thanks to Gezza and Tessa who have generously agreed to proof-read my introduction.
\\

### 2017-03-29 thesis/misc/appendix2.tex
pretty much ready to submit - tmp commit

@@
                                           & \gls{bmi} and \gls{bmi} status           & 0.02834    & 0.05237                             \\
-		                                  & Pathways only                            & 0.2416     & 1.431e-07 \tnote{1} \\
-		                                  & \gls{bmi} and pathways                   & 0.2466     & 1.031e-07           \\
-		                                  & \gls{bmi} status and pathways            & 0.2719     & 1.883e-08           \\
-		                                  & \gls{bmi}, \gls{bmi} status and pathways & 0.2814     & 9.811e-09           \\
-		                                  & \gls{pr} only                            & 0.2041     & 1.604e-06           \\
-		                                  & \gls{bmi} and \gls{pr}                   & 0.2047     & 1.54e-06            \\
-		                                  & \gls{bmi} status and \gls{pr}            & 0.2249     & 4.237e-07           \\
-		                                  & \gls{bmi}, \gls{bmi} status and \gls{pr} & 0.2372     & 1.908e-07           \\
+                                          & Pathways only                            & 0.2416     & {\bfseries \num{1.431d-07}}\tnote{1} \\
+                                          & \gls{bmi} and pathways                   & 0.2466     & \bfseries \num{1.031d-07}           \\
+                                          & \gls{bmi} status and pathways            & 0.2719     & \bfseries \num{1.883d-08}           \\
+                                          & \gls{bmi}, \gls{bmi} status and pathways & 0.2814     & \bfseries \num{9.811d-09}           \\
+                                          & \gls{pr} only                            & 0.2041     & \bfseries \num{1.604d-06}           \\
+                                          & \gls{bmi} and \gls{pr}                   & 0.2047     & \bfseries \num{1.54d-06}            \\
+                                          & \gls{bmi} status and \gls{pr}            & 0.2249     & \bfseries \num{4.237d-07}           \\
+                                          & \gls{bmi}, \gls{bmi} status and \gls{pr} & 0.2372     & \bfseries \num{1.908d-07}           \\
 				\hline
 				\rule{0pt}{2.25ex}CrOl    & \gls{bmi} only                           & 0.001295   & 0.2910                              \\
-		                                  & \gls{bmi} and \gls{bmi} status           & 0.03509    & 0.03517             \\
-		                                  & Pathways only                            & 0.2041     & 1.604e-06           \\
-		                                  & \gls{bmi} and pathways                   & 0.2173     & 6.941e-07           \\
-		                                  & \gls{bmi} status and pathways            & 0.2411     & 1.48e-07            \\
-		                                  & \gls{bmi}, \gls{bmi} status and pathways & 0.2597     & 4.308e-08           \\
-		                                  & \gls{pr} only                            & 0.2024     & 1.786e-06           \\
-		                                  & \gls{bmi} and \gls{pr}                   & 0.2056     & 1.458e-06           \\
-		                                  & \gls{bmi} status and \gls{pr}            & 0.2212     & 5.398e-07           \\
-		                                  & \gls{bmi}, \gls{bmi} status and \gls{pr} & 0.2336     & 2.417e-07           \\
+                                          & \gls{bmi} and \gls{bmi} status           & 0.03509    & \bfseries 0.03517                   \\
+                                          & Pathways only                            & 0.2041     & \bfseries \num{1.604d-06}           \\
+                                          & \gls{bmi} and pathways                   & 0.2173     & \bfseries \num{6.941d-07}           \\
+                                          & \gls{bmi} status and pathways            & 0.2411     & \bfseries \num{1.48d-07}            \\
+                                          & \gls{bmi}, \gls{bmi} status and pathways & 0.2597     & \bfseries \num{4.308d-08}           \\
+                                          & \gls{pr} only                            & 0.2024     & \bfseries \num{1.786d-06}           \\
+                                          & \gls{bmi} and \gls{pr}                   & 0.2056     & \bfseries \num{1.458d-06}           \\
+                                          & \gls{bmi} status and \gls{pr}            & 0.2212     & \bfseries \num{5.398d-07}           \\
+                                          & \gls{bmi}, \gls{bmi} status and \gls{pr} & 0.2336     & \bfseries \num{2.417d-07}           \\
 				\hline
@@
                                           & \gls{bmi} and \gls{bmi} status           & 0.02452    & 0.06579                             \\
-		                                  & Pathways only                            & 0.2785     & 1.196e-08           \\
-		                                  & \gls{bmi} and pathways                   & 0.2780     & 1.233e-08           \\
-		                                  & \gls{bmi} status and pathways            & 0.2914     & 4.887e-09           \\
-		                                  & \gls{bmi}, \gls{bmi} status and pathways & 0.3195     & 6.576e-10           \\
-		                                  & \gls{pr} only                            & 0.2331     & 2.492e-07           \\
-		                                  & \gls{bmi} and \gls{pr}                   & 0.2349     & 2.215e-07           \\
-		                                  & \gls{bmi} status and \gls{pr}            & 0.2538     & 6.38e-08            \\
-		                                  & \gls{bmi}, \gls{bmi} status and \gls{pr} & 0.2631     & 3.421e-08           \\
+                                          & Pathways only                            & 0.2785     & \bfseries \num{1.196d-08}           \\
+                                          & \gls{bmi} and pathways                   & 0.2780     & \bfseries \num{1.233d-08}           \\
+                                          & \gls{bmi} status and pathways            & 0.2914     & \bfseries \num{4.887d-09}           \\
+                                          & \gls{bmi}, \gls{bmi} status and pathways & 0.3195     & \bfseries \num{6.576d-10}           \\
+                                          & \gls{pr} only                            & 0.2331     & \bfseries \num{2.492d-07}           \\
+                                          & \gls{bmi} and \gls{pr}                   & 0.2349     & \bfseries \num{2.215d-07}           \\
+                                          & \gls{bmi} status and \gls{pr}            & 0.2538     & \bfseries \num{6.38d-08}            \\
+                                          & \gls{bmi}, \gls{bmi} status and \gls{pr} & 0.2631     & \bfseries \num{3.421d-08}           \\
 				\hline
 				\rule{0pt}{2.25ex}ResOl   & \gls{bmi} only                           & -0.002442  & 0.3851                              \\
-		                                  & \gls{bmi} and \gls{bmi} status           & 0.03878    & 0.02835             \\
-		                                  & Pathways only                            & 0.2384     & 1.762e-07           \\
-		                                  & \gls{bmi} and pathways                   & 0.2371     & 1.927e-07           \\
-		                                  & \gls{bmi} status and pathways            & 0.2670     & 2.616e-08           \\
-		                                  & \gls{bmi}, \gls{bmi} status and pathways & 0.2878     & 6.263e-09           \\
-		                                  & \gls{pr} only                            & 0.2251     & 4.198e-07           \\
-		                                  & \gls{bmi} and \gls{pr}                   & 0.2257     & 4.044e-07           \\
-		                                  & \gls{bmi} status and \gls{pr}            & 0.2392     & 1.675e-07           \\
-		                                  & \gls{bmi}, \gls{bmi} status and \gls{pr} & 0.2568     & 5.207e-08           \\
+                                          & \gls{bmi} and \gls{bmi} status           & 0.03878    & \bfseries 0.02835                   \\
+                                          & Pathways only                            & 0.2384     & \bfseries \num{1.762d-07}           \\
+                                          & \gls{bmi} and pathways                   & 0.2371     & \bfseries \num{1.927d-07}           \\
+                                          & \gls{bmi} status and pathways            & 0.2670     & \bfseries \num{2.616d-08}           \\
+                                          & \gls{bmi}, \gls{bmi} status and pathways & 0.2878     & \bfseries \num{6.263d-09}           \\
+                                          & \gls{pr} only                            & 0.2251     & \bfseries \num{4.198d-07}           \\
+                                          & \gls{bmi} and \gls{pr}                   & 0.2257     & \bfseries \num{4.044d-07}           \\
+                                          & \gls{bmi} status and \gls{pr}            & 0.2392     & \bfseries \num{1.675d-07}           \\
+                                          & \gls{bmi}, \gls{bmi} status and \gls{pr} & 0.2568     & \bfseries \num{5.207d-08}           \\
 				\hline
 				\rule{0pt}{2.25ex}Ca      & \gls{bmi} only                           & -0.0004067 & 0.3296                              \\
-		                                  & \gls{bmi} and \gls{bmi} status           & 0.03576    & 0.03383             \\
-		                                  & Pathways only                            & 0.2043     & 1.585e-06           \\
-		                                  & \gls{bmi} and pathways                   & 0.2077     & 1.28e-06            \\
-		                                  & \gls{bmi} status and pathways            & 0.2342     & 2.321e-07           \\
-		                                  & \gls{bmi}, \gls{bmi} status and pathways & 0.2487     & 8.927e-08           \\
-		                                  & \gls{pr} only                            & 0.1362     & 0.0001009           \\
-		                                  & \gls{bmi} and \gls{pr}                   & 0.1364     & 9.944e-05           \\
-		                                  & \gls{bmi} status and \gls{pr}            & 0.1643     & 1.876e-05           \\
-		                                  & \gls{bmi}, \gls{bmi} status and \gls{pr} & 0.1744     & 1.012e-05           \\
+                                          & \gls{bmi} and \gls{bmi} status           & 0.03576    & \bfseries 0.03383                   \\
+                                          & Pathways only                            & 0.2043     & \bfseries \num{1.585d-06}           \\
+                                          & \gls{bmi} and pathways                   & 0.2077     & \bfseries \num{1.28d-06}            \\
+                                          & \gls{bmi} status and pathways            & 0.2342     & \bfseries \num{2.321d-07}           \\
+                                          & \gls{bmi}, \gls{bmi} status and pathways & 0.2487     & \bfseries \num{8.927d-08}           \\
+                                          & \gls{pr} only                            & 0.1362     & \bfseries 0.0001009                 \\
+                                          & \gls{bmi} and \gls{pr}                   & 0.1364     & \bfseries \num{9.944d-05}           \\
+                                          & \gls{bmi} status and \gls{pr}            & 0.1643     & \bfseries \num{1.876d-05}           \\
+                                          & \gls{bmi}, \gls{bmi} status and \gls{pr} & 0.1744     & \bfseries \num{1.012d-05}           \\
 				\hline
 				\rule{0pt}{2.25ex}CaOl    & \gls{bmi} only                           & -0.006593  & 0.5509                              \\
-		                                  & \gls{bmi} and \gls{bmi} status           & 0.0301     & 0.04718             \\
-		                                  & Pathways only                            & 0.1323     & 0.0001273           \\
-		                                  & \gls{bmi} and pathways                   & 0.1347     & 0.0001106           \\
-		                                  & \gls{bmi} status and pathways            & 0.1487     & 4.802e-05           \\
-		                                  & \gls{bmi}, \gls{bmi} status and pathways & 0.1657     & 1.725e-05           \\
-		                                  & \gls{pr} only                            & 0.1103     & 0.0004599           \\
-		                                  & \gls{bmi} and \gls{pr}                   & 0.1106     & 0.0004539           \\
-		                                  & \gls{bmi} status and \gls{pr}            & 0.1235     & 0.0002134           \\
-		                                  & \gls{bmi}, \gls{bmi} status and \gls{pr} & 0.1383     & 8.92e-05            \\
+                                          & \gls{bmi} and \gls{bmi} status           & 0.0301     & \bfseries 0.04718                   \\
+                                          & Pathways only                            & 0.1323     & \bfseries 0.0001273                 \\
+                                          & \gls{bmi} and pathways                   & 0.1347     & \bfseries 0.0001106                 \\
+                                          & \gls{bmi} status and pathways            & 0.1487     & \bfseries \num{4.802d-05}           \\
+                                          & \gls{bmi}, \gls{bmi} status and pathways & 0.1657     & \bfseries \num{1.725d-05}           \\
+                                          & \gls{pr} only                            & 0.1103     & \bfseries 0.0004599                 \\
+                                          & \gls{bmi} and \gls{pr}                   & 0.1106     & \bfseries 0.0004539                 \\
+                                          & \gls{bmi} status and \gls{pr}            & 0.1235     & \bfseries 0.0002134                 \\
+                                          & \gls{bmi}, \gls{bmi} status and \gls{pr} & 0.1383     & \bfseries \num{8.92d-05}            \\
 				\hline
 				\rule{0pt}{2.25ex}CaRes   & \gls{bmi} only                           & 0.001142   & 0.2943                              \\
-		                                  & \gls{bmi} and \gls{bmi} status           & 0.03048    & 0.04612             \\
-		                                  & Pathways only                            & 0.2181     & 6.586e-07           \\
-		                                  & \gls{bmi} and pathways                   & 0.2373     & 1.902e-07           \\
-		                                  & \gls{bmi} status and pathways            & 0.2431     & 1.298e-07           \\
-		                                  & \gls{bmi}, \gls{bmi} status and pathways & 0.2501     & 8.152e-08           \\
-		                                  & \gls{pr} only                            & 0.1566     & 2.979e-05           \\
-		                                  & \gls{bmi} and \gls{pr}                   & 0.1586     & 2.654e-05           \\
-		                                  & \gls{bmi} status and \gls{pr}            & 0.1844     & 5.485e-06           \\
-		                                  & \gls{bmi}, \gls{bmi} status and \gls{pr} & 0.1957     & 2.707e-06           \\
+                                          & \gls{bmi} and \gls{bmi} status           & 0.03048    & \bfseries 0.04612                   \\
+                                          & Pathways only                            & 0.2181     & \bfseries \num{6.586d-07}           \\
+                                          & \gls{bmi} and pathways                   & 0.2373     & \bfseries \num{1.902d-07}           \\
+                                          & \gls{bmi} status and pathways            & 0.2431     & \bfseries \num{1.298d-07}           \\
+                                          & \gls{bmi}, \gls{bmi} status and pathways & 0.2501     & \bfseries \num{8.152d-08}           \\
+                                          & \gls{pr} only                            & 0.1566     & \bfseries \num{2.979d-05}           \\
+                                          & \gls{bmi} and \gls{pr}                   & 0.1586     & \bfseries \num{2.654d-05}           \\
+                                          & \gls{bmi} status and \gls{pr}            & 0.1844     & \bfseries \num{5.485d-06}           \\
+                                          & \gls{bmi}, \gls{bmi} status and \gls{pr} & 0.1957     & \bfseries \num{2.707d-06}           \\
 				\hline
 				\rule{0pt}{2.25ex}CaResOl & \gls{bmi} only                           & -0.006037  & 0.5225                              \\
-		                                  & \gls{bmi} and \gls{bmi} status           & 0.0413     & 0.02449             \\
-		                                  & Pathways only                            & 0.1719     & 1.183e-05           \\
-		                                  & \gls{bmi} and pathways                   & 0.1721     & 1.167e-05           \\
-		                                  & \gls{bmi} status and pathways            & 0.1961     & 2.642e-06           \\
-		                                  & \gls{bmi}, \gls{bmi} status and pathways & 0.2095     & 1.141e-06           \\
-		                                  & \gls{pr} only                            & 0.1571     & 2.894e-05           \\
-		                                  & \gls{bmi} and \gls{pr}                   & 0.1574     & 2.843e-05           \\
-		                                  & \gls{bmi} status and \gls{pr}            & 0.17       & 1.327e-05           \\
-		                                  & \gls{bmi}, \gls{bmi} status and \gls{pr} & 0.1939     & 3.044e-06           \\
+                                          & \gls{bmi} and \gls{bmi} status           & 0.0413     & \bfseries 0.02449                   \\
+                                          & Pathways only                            & 0.1719     & \bfseries \num{1.183d-05}           \\
+                                          & \gls{bmi} and pathways                   & 0.1721     & \bfseries \num{1.167d-05}           \\
+                                          & \gls{bmi} status and pathways            & 0.1961     & \bfseries \num{2.642d-06}           \\
+                                          & \gls{bmi}, \gls{bmi} status and pathways & 0.2095     & \bfseries \num{1.141d-06}           \\
+                                          & \gls{pr} only                            & 0.1571     & \bfseries \num{2.894d-05}           \\
+                                          & \gls{bmi} and \gls{pr}                   & 0.1574     & \bfseries \num{2.843d-05}           \\
+                                          & \gls{bmi} status and \gls{pr}            & 0.17       & \bfseries \num{1.327d-05}           \\
+                                          & \gls{bmi}, \gls{bmi} status and \gls{pr} & 0.1939     & \bfseries \num{3.044d-06}           \\
 				\hline
 				\rule{0pt}{2.25ex}Or      & \gls{bmi} only                           & 0.0008035  & 0.3015                              \\
-		                                  & \gls{bmi} and \gls{bmi} status           & 0.03699    & 0.03147             \\
-		                                  & Pathways only                            & 0.2554     & 5.738e-08           \\
-		                                  & \gls{bmi} and pathways                   & 0.2568     & 5.234e-08           \\
-		                                  & \gls{bmi} status and pathways            & 0.28       & 1.075e-08           \\
-		                                  & \gls{bmi}, \gls{bmi} status and pathways & 0.2974     & 3.194e-09           \\
-		                                  & \gls{pr} only                            & 0.2388     & 1.717e-07           \\
-		                                  & \gls{bmi} and \gls{pr}                   & 0.2382     & 1.79e-07            \\
-		                                  & \gls{bmi} status and \gls{pr}            & 0.2531     & 6.663e-08           \\
-		                                  & \gls{bmi}, \gls{bmi} status and \gls{pr} & 0.2684     & 2.382e-08           \\
+                                          & \gls{bmi} and \gls{bmi} status           & 0.03699    & \bfseries 0.03147                   \\
+                                          & Pathways only                            & 0.2554     & \bfseries \num{5.738d-08}           \\
+                                          & \gls{bmi} and pathways                   & 0.2568     & \bfseries \num{5.234d-08}           \\
+                                          & \gls{bmi} status and pathways            & 0.28       & \bfseries \num{1.075d-08}           \\
+                                          & \gls{bmi}, \gls{bmi} status and pathways & 0.2974     & \bfseries \num{3.194d-09}           \\
+                                          & \gls{pr} only                            & 0.2388     & \bfseries \num{1.717d-07}           \\
+                                          & \gls{bmi} and \gls{pr}                   & 0.2382     & \bfseries \num{1.79d-07}            \\
+                                          & \gls{bmi} status and \gls{pr}            & 0.2531     & \bfseries \num{6.663d-08}           \\
+                                          & \gls{bmi}, \gls{bmi} status and \gls{pr} & 0.2684     & \bfseries \num{2.382d-08}           \\
 				\hline
 				\rule{0pt}{2.25ex}FM      & \gls{bmi} only                           & -0.008238  & 0.6563                              \\
-		                                  & \gls{bmi} and \gls{bmi} status           & 0.07438    & 0.003652            \\
-		                                  & Pathways only                            & 0.268      & 2.449e-08           \\
-		                                  & \gls{bmi} and pathways                   & 0.2645     & 3.106e-08           \\
-		                                  & \gls{bmi} status and pathways            & 0.3067     & 1.654e-09           \\
-		                                  & \gls{bmi}, \gls{bmi} status and pathways & 0.3653     & 2.112e-11           \\
+                                          & \gls{bmi} and \gls{bmi} status           & 0.07438    & \bfseries 0.003652                  \\
+                                          & Pathways only                            & 0.268      & \bfseries \num{2.449d-08}           \\
+                                          & \gls{bmi} and pathways                   & 0.2645     & \bfseries \num{3.106d-08}           \\
+                                          & \gls{bmi} status and pathways            & 0.3067     & \bfseries \num{1.654d-09}           \\
+                                          & \gls{bmi}, \gls{bmi} status and pathways & 0.3653     & \bfseries \num{2.112d-11}           \\
                                           & \gls{pr} only                            & -0.004177  & 0.4434                              \\
                                           & \gls{bmi} and \gls{pr}                   & 0.007149   & 0.1946                              \\
-		                                  & \gls{bmi} status and \gls{pr}            & 0.0394     & 0.02734             \\
-		                                  & \gls{bmi}, \gls{bmi} status and \gls{pr} & 0.1166     & 0.0003183           \\
+                                          & \gls{bmi} status and \gls{pr}            & 0.0394     & \bfseries 0.02734                   \\
+                                          & \gls{bmi}, \gls{bmi} status and \gls{pr} & 0.1166     & \bfseries 0.0003183                 \\
 				\hline
@@
 				\hline
-				\rule{0pt}{2.25ex}Cr      & \gls{bmi} only                           & 0.1239     & 0.000157  \\
-		                                  & \gls{bmi} and \gls{bmi} status           & 0.0471     & 0.01568   \\
-		                                  & Pathways only                            & 0.1756     & 6.307e-06 \\
-		                                  & \gls{bmi} and pathways                   & 0.1407     & 5.596e-05 \\
-		                                  & \gls{bmi} status and pathways            & 0.05509    & 0.009721  \\
-		                                  & \gls{bmi}, \gls{bmi} status and pathways & 0.04186    & 0.02147   \\
-		                                  & \gls{pr} only                            & 0.1327     & 9.191e-05 \\
-		                                  & \gls{bmi} and \gls{pr}                   & 0.1116     & 0.0003317 \\
-		                                  & \gls{bmi} status and \gls{pr}            & 0.03887    & 0.02571   \\
-		                                  & \gls{bmi}, \gls{bmi} status and \gls{pr} & 0.04176    & 0.02161   \\
-				\hline
-				\rule{0pt}{2.25ex}CrOl    & \gls{bmi} only                           & 0.18       & 4.738e-06 \\
-		                                  & \gls{bmi} and \gls{bmi} status           & 0.06523    & 0.005312  \\
-		                                  & Pathways only                            & 0.1296     & 0.000111  \\
-		                                  & \gls{bmi} and pathways                   & 0.09006    & 0.00121   \\
-		                                  & \gls{bmi} status and pathways            & 0.02455    & 0.06179   \\
-		                                  & \gls{bmi}, \gls{bmi} status and pathways & 0.01941    & 0.08532   \\
-		                                  & \gls{pr} only                            & 0.1817     & 4.249e-06 \\
-		                                  & \gls{bmi} and \gls{pr}                   & 0.1558     & 2.188e-05 \\
-		                                  & \gls{bmi} status and \gls{pr}            & 0.07293    & 0.003359  \\
-		                                  & \gls{bmi}, \gls{bmi} status and \gls{pr} & 0.07116    & 0.003732  \\
-				\hline
-				\rule{0pt}{2.25ex}Res     & \gls{bmi} only                           & 0.09072    & 0.001163  \\
-		                                  & \gls{bmi} and \gls{bmi} status           & 0.04082    & 0.02287   \\
-		                                  & Pathways only                            & 0.2263     & 2.275e-07 \\
-		                                  & \gls{bmi} and pathways                   & 0.2012     & 1.206e-06 \\
-		                                  & \gls{bmi} status and pathways            & 0.1184     & 0.0002191 \\
-		                                  & \gls{bmi}, \gls{bmi} status and pathways & 0.1005     & 0.0006487 \\
-		                                  & \gls{pr} only                            & 0.1419     & 5.195e-05 \\
-		                                  & \gls{bmi} and \gls{pr}                   & 0.1224     & 0.0001718 \\
-		                                  & \gls{bmi} status and \gls{pr}            & 0.05342    & 0.01074   \\
-		                                  & \gls{bmi}, \gls{bmi} status and \gls{pr} & 0.05419    & 0.01026   \\
-				\hline
-				\rule{0pt}{2.25ex}ResOl   & \gls{bmi} only                           & 0.1535     & 2.536e-05 \\
-		                                  & \gls{bmi} and \gls{bmi} status           & 0.0466     & 0.01615   \\
-		                                  & Pathways only                            & 0.1581     & 1.899e-05 \\
-		                                  & \gls{bmi} and pathways                   & 0.1329     & 9.083e-05 \\
-		                                  & \gls{bmi} status and pathways            & 0.05085    & 0.01252   \\
-		                                  & \gls{bmi}, \gls{bmi} status and pathways & 0.04523    & 0.01754   \\
-		                                  & \gls{pr} only                            & 0.2067     & 8.421e-07 \\
-		                                  & \gls{bmi} and \gls{pr}                   & 0.1936     & 1.979e-06 \\
-		                                  & \gls{bmi} status and \gls{pr}            & 0.1149     & 0.0002723 \\
-		                                  & \gls{bmi}, \gls{bmi} status and \gls{pr} & 0.1097     & 0.0003713 \\
-				\hline
-				\rule{0pt}{2.25ex}Ca      & \gls{bmi} only                           & 0.07979    & 0.002233  \\
-		                                  & \gls{bmi} and \gls{bmi} status           & 0.05026    & 0.01297   \\
-		                                  & Pathways only                            & 0.2379     & 1.042e-07 \\
-		                                  & \gls{bmi} and pathways                   & 0.2081     & 7.653e-07 \\
-		                                  & \gls{bmi} status and pathways            & 0.09383    & 0.0009655 \\
-		                                  & \gls{bmi}, \gls{bmi} status and pathways & 0.07756    & 0.00255   \\
-		                                  & \gls{pr} only                            & 0.09938    & 0.000692  \\
-		                                  & \gls{bmi} and \gls{pr}                   & 0.08114    & 0.00206   \\
+				\rule{0pt}{2.25ex}Cr      & \gls{bmi} only                           & 0.1239     & {\bfseries 0.000157}\tnote{1}  \\
+		                                  & \gls{bmi} and \gls{bmi} status           & 0.0471     & \bfseries 0.01568   \\
+		                                  & Pathways only                            & 0.1756     & \bfseries \num{6.307d-06} \\
+		                                  & \gls{bmi} and pathways                   & 0.1407     & \bfseries \num{5.596d-05} \\
+		                                  & \gls{bmi} status and pathways            & 0.05509    & \bfseries 0.009721  \\
+		                                  & \gls{bmi}, \gls{bmi} status and pathways & 0.04186    & \bfseries 0.02147   \\
+		                                  & \gls{pr} only                            & 0.1327     & \bfseries \num{9.191d-05} \\
+		                                  & \gls{bmi} and \gls{pr}                   & 0.1116     & \bfseries 0.0003317 \\
+		                                  & \gls{bmi} status and \gls{pr}            & 0.03887    & \bfseries 0.02571   \\
+		                                  & \gls{bmi}, \gls{bmi} status and \gls{pr} & 0.04176    & \bfseries 0.02161   \\
+				\hline
+				\rule{0pt}{2.25ex}CrOl    & \gls{bmi} only                           & 0.18       & \bfseries \num{4.738d-06} \\
+		                                  & \gls{bmi} and \gls{bmi} status           & 0.06523    & \bfseries 0.005312  \\
+		                                  & Pathways only                            & 0.1296     & \bfseries 0.000111  \\
+		                                  & \gls{bmi} and pathways                   & 0.09006    & \bfseries 0.00121   \\
+		                                  & \gls{bmi} status and pathways            & 0.02455    & \bfseries 0.06179   \\
+		                                  & \gls{bmi}, \gls{bmi} status and pathways & 0.01941    & \bfseries 0.08532   \\
+		                                  & \gls{pr} only                            & 0.1817     & \bfseries \num{4.249d-06} \\
+		                                  & \gls{bmi} and \gls{pr}                   & 0.1558     & \bfseries \num{2.188d-05} \\
+		                                  & \gls{bmi} status and \gls{pr}            & 0.07293    & \bfseries 0.003359  \\
+		                                  & \gls{bmi}, \gls{bmi} status and \gls{pr} & 0.07116    & \bfseries 0.003732  \\
+				\hline
+				\rule{0pt}{2.25ex}Res     & \gls{bmi} only                           & 0.09072    & \bfseries 0.001163  \\
+		                                  & \gls{bmi} and \gls{bmi} status           & 0.04082    & \bfseries 0.02287   \\
+		                                  & Pathways only                            & 0.2263     & \bfseries \num{2.275d-07} \\
+		                                  & \gls{bmi} and pathways                   & 0.2012     & \bfseries \num{1.206d-06} \\
+		                                  & \gls{bmi} status and pathways            & 0.1184     & \bfseries 0.0002191 \\
+		                                  & \gls{bmi}, \gls{bmi} status and pathways & 0.1005     & \bfseries 0.0006487 \\
+		                                  & \gls{pr} only                            & 0.1419     & \bfseries \num{5.195d-05} \\
+		                                  & \gls{bmi} and \gls{pr}                   & 0.1224     & \bfseries 0.0001718 \\
+		                                  & \gls{bmi} status and \gls{pr}            & 0.05342    & \bfseries 0.01074   \\
+		                                  & \gls{bmi}, \gls{bmi} status and \gls{pr} & 0.05419    & \bfseries 0.01026   \\
+				\hline
+				\rule{0pt}{2.25ex}ResOl   & \gls{bmi} only                           & 0.1535     & \bfseries \num{2.536d-05} \\
+		                                  & \gls{bmi} and \gls{bmi} status           & 0.0466     & \bfseries 0.01615   \\
+		                                  & Pathways only                            & 0.1581     & \bfseries \num{1.899d-05} \\
+		                                  & \gls{bmi} and pathways                   & 0.1329     & \bfseries \num{9.083d-05} \\
+		                                  & \gls{bmi} status and pathways            & 0.05085    & \bfseries 0.01252   \\
+		                                  & \gls{bmi}, \gls{bmi} status and pathways & 0.04523    & \bfseries 0.01754   \\
+		                                  & \gls{pr} only                            & 0.2067     & \bfseries \num{8.421d-07} \\
+		                                  & \gls{bmi} and \gls{pr}                   & 0.1936     & \bfseries \num{1.979d-06} \\
+		                                  & \gls{bmi} status and \gls{pr}            & 0.1149     & \bfseries 0.0002723 \\
+		                                  & \gls{bmi}, \gls{bmi} status and \gls{pr} & 0.1097     & \bfseries 0.0003713 \\
+				\hline
+				\rule{0pt}{2.25ex}Ca      & \gls{bmi} only                           & 0.07979    & \bfseries 0.002233  \\
+		                                  & \gls{bmi} and \gls{bmi} status           & 0.05026    & \bfseries 0.01297   \\
+		                                  & Pathways only                            & 0.2379     & \bfseries \num{1.042d-07} \\
+		                                  & \gls{bmi} and pathways                   & 0.2081     & \bfseries \num{7.653d-07} \\
+		                                  & \gls{bmi} status and pathways            & 0.09383    & \bfseries 0.0009655 \\
+		                                  & \gls{bmi}, \gls{bmi} status and pathways & 0.07756    & \bfseries 0.00255   \\
+		                                  & \gls{pr} only                            & 0.09938    & \bfseries 0.000692  \\
+		                                  & \gls{bmi} and \gls{pr}                   & 0.08114    & \bfseries 0.00206   \\
 		                                  & \gls{bmi} status and \gls{pr}            & 0.01059    & 0.1512    \\
@@
 				\hline
-				\rule{0pt}{2.25ex}CaOl    & \gls{bmi} only                           & 0.1242     & 0.0001542 \\
-		                                  & \gls{bmi} and \gls{bmi} status           & 0.05609    & 0.009156  \\
-		                                  & Pathways only                            & 0.08453    & 0.001683  \\
-		                                  & \gls{bmi} and pathways                   & 0.07397    & 0.003157  \\
+				\rule{0pt}{2.25ex}CaOl    & \gls{bmi} only                           & 0.1242     & \bfseries 0.0001542 \\
+		                                  & \gls{bmi} and \gls{bmi} status           & 0.05609    & \bfseries 0.009156  \\
+		                                  & Pathways only                            & 0.08453    & \bfseries 0.001683  \\
+		                                  & \gls{bmi} and pathways                   & 0.07397    & \bfseries 0.003157  \\
 		                                  & \gls{bmi} status and pathways            & 0.003647   & 0.244     \\
 		                                  & \gls{bmi}, \gls{bmi} status and pathways & 0.0003557  & 0.3111    \\
-		                                  & \gls{pr} only                            & 0.1253     & 0.0001444 \\
-		                                  & \gls{bmi} and \gls{pr}                   & 0.1243     & 0.0001534 \\
-		                                  & \gls{bmi} status and \gls{pr}            & 0.02838    & 0.04876   \\
+		                                  & \gls{pr} only                            & 0.1253     & \bfseries 0.0001444 \\
+		                                  & \gls{bmi} and \gls{pr}                   & 0.1243     & \bfseries 0.0001534 \\
+		                                  & \gls{bmi} status and \gls{pr}            & 0.02838    & \bfseries 0.04876   \\
 		                                  & \gls{bmi}, \gls{bmi} status and \gls{pr} & 0.02625    & 0.0556    \\
 				\hline
-				\rule{0pt}{2.25ex}CaRes   & \gls{bmi} only                           & 0.05877    & 0.007805  \\
-		                                  & \gls{bmi} and \gls{bmi} status           & 0.0415     & 0.02194   \\
-		                                  & Pathways only                            & 0.2438     & 6.952e-08 \\
-		                                  & \gls{bmi} and pathways                   & 0.2037     & 1.023e-06 \\
-		                                  & \gls{bmi} status and pathways            & 0.1132     & 0.0003018 \\
-		                                  & \gls{bmi}, \gls{bmi} status and pathways & 0.0995     & 0.0006872 \\
-		                                  & \gls{pr} only                            & 0.1096     & 0.0003748 \\
-		                                  & \gls{bmi} and \gls{pr}                   & 0.08751    & 0.001409  \\
+				\rule{0pt}{2.25ex}CaRes   & \gls{bmi} only                           & 0.05877    & \bfseries 0.007805  \\
+		                                  & \gls{bmi} and \gls{bmi} status           & 0.0415     & \bfseries 0.02194   \\
+		                                  & Pathways only                            & 0.2438     & \bfseries \num{6.952d-08} \\
+		                                  & \gls{bmi} and pathways                   & 0.2037     & \bfseries \num{1.023d-06} \\
+		                                  & \gls{bmi} status and pathways            & 0.1132     & \bfseries 0.0003018 \\
+		                                  & \gls{bmi}, \gls{bmi} status and pathways & 0.0995     & \bfseries 0.0006872 \\
+		                                  & \gls{pr} only                            & 0.1096     & \bfseries 0.0003748 \\
+		                                  & \gls{bmi} and \gls{pr}                   & 0.08751    & \bfseries 0.001409  \\
 		                                  & \gls{bmi} status and \gls{pr}            & 0.02671    & 0.05406   \\
@@
 				\hline
-				\rule{0pt}{2.25ex}CaResOl & \gls{bmi} only                           & 0.1167     & 0.0002439 \\
-		                                  & \gls{bmi} and \gls{bmi} status           & 0.04905    & 0.01395   \\
-		                                  & Pathways only                            & 0.07821    & 0.002453  \\
-		                                  & \gls{bmi} and pathways                   & 0.06736    & 0.004681  \\
+				\rule{0pt}{2.25ex}CaResOl & \gls{bmi} only                           & 0.1167     & \bfseries 0.0002439 \\
+		                                  & \gls{bmi} and \gls{bmi} status           & 0.04905    & \bfseries 0.01395   \\
+		                                  & Pathways only                            & 0.07821    & \bfseries 0.002453  \\
+		                                  & \gls{bmi} and pathways                   & 0.06736    & \bfseries 0.004681  \\
 		                                  & \gls{bmi} status and pathways            & -0.0005575 & 0.3338    \\
 		                                  & \gls{bmi}, \gls{bmi} status and pathways & -0.003248  & 0.4151    \\
-		                                  & \gls{pr} only                            & 0.1632     & 1.377e-05 \\
-		                                  & \gls{bmi} and \gls{pr}                   & 0.1628     & 1.412e-05 \\
-		                                  & \gls{bmi} status and \gls{pr}            & 0.05905    & 0.007678  \\
-		                                  & \gls{bmi}, \gls{bmi} status and \gls{pr} & 0.05391    & 0.01043   \\
-				\hline
-				\rule{0pt}{2.25ex}Or      & \gls{bmi} only                           & 0.1246     & 0.0001508 \\
-		                                  & \gls{bmi} and \gls{bmi} status           & 0.03912    & 0.02533   \\
-		                                  & Pathways only                            & 0.1649     & 1.237e-05 \\
-		                                  & \gls{bmi} and pathways                   & 0.1311     & 0.0001014 \\
-		                                  & \gls{bmi} status and pathways            & 0.05614    & 0.009131  \\
-		                                  & \gls{bmi}, \gls{bmi} status and pathways & 0.04897    & 0.01402   \\
-		                                  & \gls{pr} only                            & 0.1805     & 4.604e-06 \\
-		                                  & \gls{bmi} and \gls{pr}                   & 0.1626     & 1.431e-05 \\
-		                                  & \gls{bmi} status and \gls{pr}            & 0.08503    & 0.001634  \\
-		                                  & \gls{bmi}, \gls{bmi} status and \gls{pr} & 0.08819    & 0.001353  \\
+		                                  & \gls{pr} only                            & 0.1632     & \bfseries \num{1.377d-05} \\
+		                                  & \gls{bmi} and \gls{pr}                   & 0.1628     & \bfseries \num{1.412d-05} \\
+		                                  & \gls{bmi} status and \gls{pr}            & 0.05905    & \bfseries 0.007678  \\
+		                                  & \gls{bmi}, \gls{bmi} status and \gls{pr} & 0.05391    & \bfseries 0.01043   \\
+				\hline
+				\rule{0pt}{2.25ex}Or      & \gls{bmi} only                           & 0.1246     & \bfseries 0.0001508 \\
+		                                  & \gls{bmi} and \gls{bmi} status           & 0.03912    & \bfseries 0.02533   \\
+		                                  & Pathways only                            & 0.1649     & \bfseries \num{1.237d-05} \\
+		                                  & \gls{bmi} and pathways                   & 0.1311     & \bfseries 0.0001014 \\
+		                                  & \gls{bmi} status and pathways            & 0.05614    & \bfseries 0.009131  \\
+		                                  & \gls{bmi}, \gls{bmi} status and pathways & 0.04897    & \bfseries 0.01402   \\
+		                                  & \gls{pr} only                            & 0.1805     & \bfseries \num{4.604d-06} \\
+		                                  & \gls{bmi} and \gls{pr}                   & 0.1626     & \bfseries \num{1.431d-05} \\
+		                                  & \gls{bmi} status and \gls{pr}            & 0.08503    & \bfseries 0.001634  \\
+		                                  & \gls{bmi}, \gls{bmi} status and \gls{pr} & 0.08819    & \bfseries 0.001353  \\
 				\hline
@@
 		                                  & \gls{bmi} and \gls{bmi} status           & 0.01947    & 0.08502   \\
-		                                  & Pathways only                            & 0.4202     & 8.085e-14 \\
-		                                  & \gls{bmi} and pathways                   & 0.4213     & 7.317e-14 \\
-		                                  & \gls{bmi} status and pathways            & 0.3386     & 6.885e-11 \\
-		                                  & \gls{bmi}, \gls{bmi} status and pathways & 0.2734     & 8.759e-09 \\
-		                                  & \gls{pr} only                            & 0.09798    & 0.0007529 \\
-		                                  & \gls{bmi} and \gls{pr}                   & 0.09642    & 0.0008267 \\
+		                                  & Pathways only                            & 0.4202     & \bfseries \num{8.085d-14} \\
+		                                  & \gls{bmi} and pathways                   & 0.4213     & \bfseries \num{7.317d-14} \\
+		                                  & \gls{bmi} status and pathways            & 0.3386     & \bfseries \num{6.885d-11} \\
+		                                  & \gls{bmi}, \gls{bmi} status and pathways & 0.2734     & \bfseries \num{8.759d-09} \\
+		                                  & \gls{pr} only                            & 0.09798    & \bfseries 0.0007529 \\
+		                                  & \gls{bmi} and \gls{pr}                   & 0.09642    & \bfseries 0.0008267 \\
 		                                  & \gls{bmi} status and \gls{pr}            & -0.00988   & 0.9634    \\
@@
 				\hline
-				\rule{0pt}{2.25ex}Cr & 3.005e-06 & 0.0002419 & 6.931e-06  \\
-				CrOl                 & 2.924e-07 & 8.516e-06 & 1.467e-07  \\
-				Res                  & 0.2783    & 2.146e-05 & 5.715e-05  \\
-				ResOl                & 2.023e-06 & 3.31e-05  & 1.052e-06  \\
-				Ca                   & 0.00283   & 8.769e-05 & 0.0002354  \\
-				CaOl                 & 0.0001869 & 3.328e-06 & 4.371e-06  \\
-				CaRes                & 0.008582  & 0.0004552 & 0.001334   \\
-				CaResOl              & 0.0004192 & 5.338e-06 & 8.806e-06  \\
-				Or                   & 5.806e-06 & 0.0002258 & 8.092e-06  \\
+				\rule{0pt}{2.25ex}Cr & {\bfseries \num{3.005d-06}} \tnote{1} & \bfseries 0.0002419       & \bfseries \num{6.931d-06} \\
+				CrOl                 & \bfseries \num{2.924d-07}           & \bfseries \num{8.516d-06} & \bfseries \num{1.467d-07} \\
+				Res                  & 0.2783                              & \bfseries \num{2.146d-05} & \bfseries \num{5.715d-05} \\
+				ResOl                & \bfseries \num{2.023d-06}           & \bfseries \num{3.31d-05}  & \bfseries \num{1.052d-06} \\
+				Ca                   & \bfseries 0.00283                   & \bfseries \num{8.769d-05} & \bfseries 0.0002354       \\
+				CaOl                 & \bfseries 0.0001869                 & \bfseries \num{3.328d-06} & \bfseries \num{4.371d-06} \\
+				CaRes                & \bfseries 0.008582                  & \bfseries 0.0004552       & \bfseries 0.001334        \\
+				CaResOl              & \bfseries 0.0004192                 & \bfseries \num{5.338d-06} & \bfseries \num{8.806d-06} \\
+				Or                   & \bfseries \num{5.806d-06}           & \bfseries 0.0002258       & \bfseries \num{8.092d-06} \\
 				FM                   & 0.8573                              & 0.185                     & 0.3568          \\

### 2017-03-29 thesis/misc/title.tex
pretty much ready to submit - tmp commit

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Title page for my thesis - code is pretty self explanatory
% Required packages: setspace
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{titlepage}

\centering

\rule[2.0mm]{\textwidth}{0.5mm}

\begin{doublespace}
    {\Huge \scshape Body Mass Index and Pathway Dysregulation in Cancer}
\end{doublespace}

\rule[2.0mm]{\textwidth}{0.5mm}

\vspace{2.0mm}
{\Large \scshape Riku Takei}

\vfill

\includegraphics[width=0.33\linewidth]{/misc/otago}

\vfill

{\normalsize \scshape A Thesis Submitted For The Degree Of\\}
{\Large \scshape Master Of Science\\}
\vspace{2.0mm}
{\normalsize \scshape At The University Of Otago, Dunedin, New Zealand \\}

\vspace{10.0mm}
{\normalsize \today}

\end{titlepage}

### 2017-03-29 thesis/discussion/discussion.tex
added some pathway stuff into section 5.1.4 - tmp commit

@@
 
-Finally, pathway enrichment analysis was carried out in the \gls{icgc} data sets to investigate whether there were any pathways that significantly associated with obesity (\cref{sec:pathways_enriched_in_icgc_data_sets}).
-There were no pathways enriched in any of the \gls{icgc} data sets, nor the combined \gls{icgc} data set.
-Although there was no significantly enriched  pathway in the \gls{icgc} data sets, it was evident from the obesity metagenes from the CR and FM data sets that the obesity metagenes were associated with the gene expressions of the obesity associated genetic signatures.
+Finally, pathway enrichment analysis was carried out in the \gls{icgc} data sets (first with each cancer type separately, then all cancer types combined) to investigate whether there were any pathways that significantly associated with obesity (\cref{sec:pathways_enriched_in_icgc_data_sets}).
+There were not many pathways identified from the \gls{kegg} and the Reactome pathway databases that were significantly enriched in the \gls{icgc} cancer data sets.
+There was a substantial number of \gls{go} terms that were significantly enriched in the \gls{icgc} data, but there was no pathways that were common across all cancer types.
+One thing noted from these results was that all three databases (\gls{go}, \gls{kegg} and Reactome) identified ``ABC transport'' pathway as significantly enriched in the \gls{cesc} data set.
+This suggests that the ``ABC transport'' pathway may have an important role in \gls{cesc} tumour development in obese patients.
+There was no significantly enriched pathway that was associated with obesity in the combined \gls{icgc} data sets.
+
+Although there was no common pathway that associated with obesity in the \gls{icgc} data sets, it was evident from the CR and FM obesity metagenes that the obesity metagenes were associated with the gene expressions of the obesity signatures.
 This suggested that the metagenes were able to identify some sort of genetic pattern in the data, even though these metagenes were not significantly associated with obesity.

### 2017-03-29 thesis/discussion/discussion.tex
tmp commit

@@
 All of the predicted obesity metagene scores significantly correlated with the true obesity metagene scores in both the \gls{nzbc} (training) and CR (testing) data sets.
-% TODO: check r-squared values
 However, the $R^2$-values for all of these predictions were low (highest $R^2$ $\approx$ 0.42; \cref{sec:summary_statistics_of_the_predicted_obesity_metagenes_with_sample_bmi_bmi_status_in_nzbc_and_cr_data}) with highly variable data points in the scatter plots.
@@
 Since linear models were created only with the ``consistent'' pathway metagene scores, a step wise method was used to generate linear models that included the patient \gls{bmi}, \gls{bmi} status and all of the pathway metagenes that allowed the model to predict the true obesity metagene the best.
-% TODO: check r-squared values
-As shown by the results in \cref{sec:prediction_of_obesity_associated_metagene_with_pathway_associate_metagene}, all of the models predicted the true obesity metagene scores with high $R^2$-value in Creighton's data set (highest $R^2$ $\approx$ 0.88) and were statistically significant.
+As shown by the results in \cref{sec:prediction_of_obesity_associated_metagene_with_pathway_associate_metagene}, all of the models predicted the true obesity metagene scores with high $R^2$-value in the CR data set (highest $R^2$ $\approx$ 0.8869) and were statistically significant.
 This meant that the variables that were significant in the linear models played a role in the obesity associated genetic signatures.
@@
 None of the pathway associated genetic signatures showed positive correlation with any of the obesity associated genetic signatures in the heatmap.
-However, strong negative correlation with the Akt, \gls{er}, \gls{egfr}, \gls{pr}, p53, Src and \gls{tgfb} pathways, suggesting that the possibility of the obesity associated genetic signatures were picking up the signals from these pathways, rather than the signals from the obesity phenotype of the patients.
+However, strong negative correlation with the Akt, \gls{er}, \gls{egfr}, \gls{pr}, p53, Src and \gls{tgfb} pathways, suggesting that the obesity associated genetic signatures were picking up the signals from these pathways, rather than the signals from the obesity phenotype of the patients.
 

### 2017-03-29 thesis/intro/intro.tex
tmp commit

@@
 
+\vspace{-2mm}
+
 \subsubsection{\citet{Gatza2010a} study}
@@
 
+\vspace{-2mm}
+
 \section{Aims of the project}
@@
 This research aims to determine whether gene expression signatures exist  that are specific to obesity across multiple cancer types, and to investigate whether there are any common pathways being dysregulated in cancers based on these genetic signatures.
-Better understanding of the pathways being dysregulated in cancer cells in obese patients may lead to improved clinical decisions, and contribute towards personalised treatment in the future.
+A better understanding of the pathways being dysregulated in cancer cells in obese patients may lead to improved clinical decisions, and contribute towards personalised treatment in the future.
 

### 2017-03-29 thesis/methods/methods.tex
tmp commit

@@
 
-The raw \gls{rnaseq} data from the \gls{icgc} database were formatted in such a way that the count data of all the genes were listed for one sample, then the count data for all the genes for the next sample, and so on.
+The raw \gls{rnaseq} data from the \gls{icgc} database were formatted so that the count data of all the genes were listed for one sample, then the count data for all the genes for the next sample, and so on.
 This data format was highly inconvenient for later analyses so the data were reformatted into a ``gene by sample'' matrix format using the \textit{dplyr} package in \gls{r}.
@@
 Though similar to the \gls{tcga} sample IDs, the \gls{icgc} IDs in the \gls{rnaseq} data had an extra identification code in the sample names.
-To associate each sample in the raw \gls{rnaseq} data from the \gls{icgc} database with the correct samples in the clinical data from the \gls{tcga} database, the \gls{icgc} IDs were re-coded so they matched the \gls{tcga} IDs.
-After the IDs were re-coded, all of the samples were checked to see if there were any duplicates in either the \gls{icgc} \gls{rnaseq} data or the \gls{tcga} clinical data.
+To associate each sample in the raw \gls{rnaseq} data from the \gls{icgc} database with the correct samples in the clinical data from the \gls{tcga} database, the \gls{icgc} IDs were re-coded to match the \gls{tcga} IDs.
+After the IDs were re-coded, samples were checked to see if there were any duplicates in either the \gls{icgc} \gls{rnaseq} data or the \gls{tcga} clinical data.
 Where there was a duplicate in the sample ID, a single sample with the highest mean expression value was chosen to represent that particular ID by using the \texttt{collapseRows} function in the \textit{WGCNA} package in \gls{r} \citep{Langfelder2008}.
@@
 \begin{table}[htpb]
-	\caption{Summary of the sample \gls{bmi} status in the \gls{icgc} cancer data}
+	\caption{Summary of the patient \gls{bmi} status in the \gls{icgc} cancer data}
 	\label{tab:icgc_bmi}
@@
 The \gls{kegg} database also curates other information such as the approved drugs in US and Japan and genes related to diseases, making it a great tool for exploring the biological significance of genes and pathways \citep{Kanehisa2008}.
-
 The \textit{KEGG.db} package was used to load the \gls{kegg} database into \gls{r} \citep{kegg}.
@@
 The Reactome database is constructed based on the reactions that have direct evidence from the literature, which are connected into higher order pathway structures \citep{Joshi2005}.
-
 The \textit{reactome.db} package was used to load the Reactome database into \gls{r} \citep{reactome}.
@@
 
-Since the \gls{nzbc} data was the only data set other than the Creighton \textit{et al.} data set that had the patient \gls{bmi} information, all of the linear models were constructed in the \gls{nzbc} data set.
+Since the \gls{nzbc} data was the only data set (other than the Creighton \textit{et al.} data set) that had the patient \gls{bmi} information, all of the linear models were constructed in the \gls{nzbc} data set.
 Eleven linear models were created in the \gls{nzbc} to predict the obesity associated metagene from the \citet{Creighton2012} study.
@@
 
+\vspace{-2mm}
+
 \subsection{Venn diagrams}
@@
 The \texttt{overLapper} function (from \url{http://faculty.ucr.edu/~tgirke/Documents/R_BioCond/My_R_Scripts/overLapper.R}) was also used to help plot the Venn diagram \citep{Girke2016}.
+
+\vspace{-2mm}
 

### 2017-03-29 thesis/misc/abstract.tex
tmp commit

@@
 It is likely that there is some complex mechanism underlying the relationship between obesity and cancer.
-Better understanding of the pathways being dysregulated in cancer cells in obese patients may lead to improved clinical decisions, and contribute towards personalised treatment in the future.
+A better understanding of the pathways being dysregulated in cancer cells in obese patients may lead to improved clinical decisions, and contribute towards personalised treatment in the future.
 

### 2017-03-29 thesis/misc/acknowledgement.tex
tmp commit

@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+\vspace*{\fill}
 
@@
 First and foremost, I would like to thank my supervisor Mik, for giving me this project and an
-opportunity to learn and develop the basic bioinformatics skills required to complete it.
+opportunity to learn and develop the bioinformatics skills required to complete it.
 You have been a great supervisor and I have learnt a lot from you over the past two years, and I have enjoyed being a part of the Black lab.
@@
 I would also like to thank both past and present members of the Black lab, for useful tips and tricks to get me through my Masters.
-Especially James Boocock, for introducing me to a variety of useful tool sets (by the way, everyone should start using vim); Murray Cadzow, for organising the fortnightly ``Sh*t You Should Know About'' sessions to learn more about R and other useful tools; and Tom Kelly, for being my companion in the Fish-bowl and helping me with some of the codes in my project.
+Especially James Boocock, for introducing me to a variety of useful tool sets (by the way, everyone should start using vim); Murray Cadzow, for organising the fortnightly ``Shit You Should Know About'' (a.k.a.
+SYSKA) sessions to teach us more about R and other useful tools; and Tom Kelly, for being my companion in the Fish-bowl and helping me out with some of the codes in my project.
 \\
@@
 \noindent
-I would like to thank all of my friends and family for all their support and fun memories throughout the years that I have been at the University.
-Lastly, a special thanks to Gezza and Tessa who have generously agreed to proof-read my introduction.
+I would like to thank all of my friends and family, for all their support and fun memories throughout the years that I have been at University.
+I wish I could name everyone, but I do not think it is a good idea to extend this thesis any longer\ldots
+Long story short, my life would not be nearly as interesting or enjoyable without you guys.
 \\
+
+\noindent
+Lastly, a special thanks to Gezza and Tessa who generously proof-read my introduction.
+\\
+
+\vfill

### 2017-03-29 thesis/misc/appendix.tex
tmp commit

@@
 	\renewcommand{\thesection}{\Alph{chapter}\arabic{section}}
+	\renewcommand{\thetable}{S\arabic{table}}
+	\renewcommand{\thefigure}{S\arabic{figure}}
+	\setcounter{figure}{0}
+	\setcounter{table}{0}
 
@@
 
-	\section{Remainder of the results of the other Creighton \textit{et al.} obesity metagenes in the \gls{icgc} cancer data sets}
+	\section{Remainder of the results of the other CR obesity metagenes in the \gls{icgc} cancer data sets}
 	\label{sec:rest_of_the_cr_icgc_cancer_heatmap_results}
@@
 			CaRes   & 0.2327                      & 0.2344  & 0.3738             & -0.006250  & 0.6279              \\
-			CaOl    & 0.3426                      & 0.2775  & 0.4693             & -0.007333  & 0.77469             \\
+			CaOl    & 0.3426                      & 0.2775  & 0.4693             & -0.007333  & 0.7747             \\
 			CaResOl & 0.2643                      & 0.1315  & 0.2410             & -0.001408  & 0.3649              \\
@@
 		\begin{tablenotes}
+			\begin{footnotesize}
 				\item [1] Not available (no pathway enriched)
+			\end{footnotesize}
 		\end{tablenotes}

### 2017-03-29 thesis/misc/appendix2.tex
tmp commit

@@
 
-	\section{Original result from \citet{Gatza2010a} study}
+	\section{Original result from the \citet{Gatza2010a} study}
 	\label{sec:result_from_gatza2010a_study}
@@
 
-	\section{Summary of the remainder of the \gls{pr}-only linear models in \gls{nzbc} data}
+	\section{Summary of the remainder of the \gls{pr} linear models in \gls{nzbc} data}
 	\label{sec:summary_of_the_pr_linear_models_in_nzbc_data}
@@
 
-	\section{Summary statistics of the predicted obesity metagenes with sample \gls{bmi}/\gls{bmi} status in \gls{nzbc} and CR data}
+	\section{Summary statistics of the predicted obesity metagenes with sample \gls{bmi}/\gls{bmi} status in the \gls{nzbc} and CR data}
 	\label{sec:summary_statistics_of_the_predicted_obesity_metagenes_with_sample_bmi_bmi_status_in_nzbc_and_cr_data}
@@
 
+	\section{Code used in this project}
+	\label{sec:code_used_in_this_project}
+
+	All of the \gls{r} source code used in this project is publicly available at \url{https://github.com/rikutakei/bmi-cancer-code.git}.
+	This thesis was written with \LaTeX{}, and the source code is publicly available at \url{https://github.com/rikutakei/mastersDoc.git}.
+
 \end{appendices}

### 2017-03-29 thesis/thesis.tex
finished editing up to the discussion
tmp commit
debugged the blank page before acronyms - tmp commit
pretty much ready to submit - tmp commit

@@
 \usepackage[T1]{fontenc}                                 % package for specifying output font encoding
-\usepackage[acronym,style=super]{glossaries}                    % package for making glossaries
+\usepackage[acronym,style=super,nogroupskip,section=chapter]{glossaries} % package for making glossaries
 \usepackage[flushleft]{threeparttablex}                  % package to add notes at the bottom of the table
@@
 % Change the title of Table of Contents:
-\renewcommand{\thesubfigure}{\Alph{subfigure}}
+% \renewcommand{\thesubfigure}{\Alph{subfigure}}
 
@@
 \titleformat{\paragraph}{\normalfont\itshape}{\theparagraph}{\noindent}{}
+
+% Remove unnecessary spaces between the figures from different chapters in LoF and LoT:
+\newcommand*{\noaddvspace}{\renewcommand*{\addvspace}[1]{}}
+\addtocontents{lof}{\protect\noaddvspace}
+\addtocontents{lot}{\protect\noaddvspace}
 
@@
 % If you want to compile specific files, uncomment the line below and add the filenames in the {} :
+% \includeonly{{./misc/title},{./misc/abstract},{./misc/acknowledgement},{./methods/methods},{./intro/intro},{./results/results},{./results/results2},{./discussion/discussion}}
+\includeonly{{./misc/title},{./misc/abstract},{./misc/acknowledgement}}
 % \includeonly{{./intro/intro}}
 % \includeonly{{./methods/methods}}
-\includeonly{{./results/results2},{./discussion/discussion},{./misc/appendix},{./misc/appendix2}}
+% \includeonly{{./results/results2},{./discussion/discussion},{./misc/appendix},{./misc/appendix2}}
 % \includeonly{{./results/results}}
@@
 
+% Set line spacing to double space:
+\onehalfspacing
+
 \include{./misc/title}
@@
 \pagenumbering{roman} % set page numbering to roman numeral
-
-% Set line spacing to double space:
-\onehalfspacing
 
@@
 
+% Need to set line spacing to single space, otherwise it inserts a blank page for an unknown reason:
+\singlespacing
+
 \clearpage
@@
 \clearpage
+
+% Reset the spacing to 1.5:
+\onehalfspacing
 

### 2017-03-30 thesis/discussion/discussion.tex
tmp commit

@@
 One possible reason for this was that the obesity metagenes were not actually associated to the obesity phenotype in the other data sets.
-Another reason could be that the methodology used were not appropriate for investigating the association between the phenotype and its associated phenotype.
+Another reason could be that the methodology used was not appropriate for investigating the association between the phenotype and its associated phenotype.
 
@@
 \gls{svd} was first applied to the \gls{rma} normalised CR data to obtain the \gls{er} and \gls{pr} metagenes, along with the transformation matrices for each of the genetic signatures.
-Both the \gls{er} and \gls{pr} metagenes showed good concordance with the gene expression profiles of the genetic signature in the heatmap, and also showed significant association with the patient \gls{er} and \gls{pr} statuses (\cref{sec:results_for_er_and_pr_metagene_analysis}).
+Both the \gls{er} and \gls{pr} metagenes showed good concordance with the gene expression profiles of the genetic signature in the heatmap, and also showed significant association with patient \gls{er} and \gls{pr} statuses (\cref{sec:results_for_er_and_pr_metagene_analysis}).
 
@@
 Both the \gls{er} and \gls{pr} pathway metagenes showed similar overall gene expression patterning in the heatmap as the CR data (\cref{sec:results_for_er_and_pr_metagene_analysis}).
-Furthermore, both the \gls{er} and \gls{pr} pathway metagenes from these data sets show significant association with the patient \gls{er} and \gls{pr} statuses, respectively (\cref{sec:results_for_er_and_pr_metagene_analysis}).
+Furthermore, both the \gls{er} and \gls{pr} pathway metagenes from these data sets show significant association with patient \gls{er} and \gls{pr} statuses, respectively (\cref{sec:results_for_er_and_pr_metagene_analysis}).
 These results clearly showed that the general approach of using \gls{svd} and transformation matrices was appropriate for examining the association of the genetic signatures with the patient phenotype.
@@
 
-The results so far have provided strong evidence that the obesity associated genetic signatures from the CR and the FM data sets were not associated with the patient \gls{bmi}.
-In the last attempt to identify any obesity associated genes that were common across multiple cancer types, gene expression analyses were carried out on the \gls{icgc} cancer data sets using samples for which patient \gls{bmi} information were available (\cref{sec:common_genes_across_multiple_cancer_types}).
-There were many \glspl{deg} identified by each of the \gls{icgc} cancer types, but there were no genes in common for all eight cancer types, and this observation was supported by the results from the simulation.
+The results so far have provided strong evidence that the obesity associated genetic signatures from the CR and the FM data sets were not associated with patient \gls{bmi}.
+In a last attempt to identify any obesity associated genes that were common across multiple cancer types, gene expression analyses were carried out on the \gls{icgc} cancer data sets using samples for which patient \gls{bmi} information was available (\cref{sec:common_genes_across_multiple_cancer_types}).
+There were many \glspl{deg} identified in each of the \gls{icgc} cancer types, but there were no genes in common for all eight cancer types, and this observation was supported by the results from the simulation.
 
@@
 There were not many pathways identified from the \gls{kegg} and the Reactome pathway databases that were significantly enriched in the \gls{icgc} cancer data sets.
-There was a substantial number of \gls{go} terms that were significantly enriched in the \gls{icgc} data, but there was no pathways that were common across all cancer types.
-One thing noted from these results was that all three databases (\gls{go}, \gls{kegg} and Reactome) identified ``ABC transport'' pathway as significantly enriched in the \gls{cesc} data set.
+There were a substantial number of \gls{go} terms that were significantly enriched in the \gls{icgc} data, but there were no pathways that were common across all cancer types.
+One thing noted from these results was that all three databases (\gls{go}, \gls{kegg} and Reactome) identified the ``ABC transport'' pathway as significantly enriched in the \gls{cesc} data set.
 This suggests that the ``ABC transport'' pathway may have an important role in \gls{cesc} tumour development in obese patients.
@@
 
-Although there was no common pathway that associated with obesity in the \gls{icgc} data sets, it was evident from the CR and FM obesity metagenes that the obesity metagenes were associated with the gene expressions of the obesity signatures.
+Although there was no common pathway that associated with obesity in the \gls{icgc} data sets, it was evident from the CR and FM obesity metagenes that the obesity metagenes were associated with the gene expression of the obesity signatures.
 This suggested that the metagenes were able to identify some sort of genetic pattern in the data, even though these metagenes were not significantly associated with obesity.
@@
 This meant that the variables that were significant in the linear models played a role in the obesity associated genetic signatures.
-There was no single pathway signature that was common to all of the linear models, but Akt, \gls{egfr} and Src pathway metagenes were included in many of the models, which suggested that these pathway signatures may have an important role in the obesity associated genetic signatures (\cref{sub:stepwise_linear_model_prediction_in_nzbc_and_cr_data_sets}).
+There was no single pathway signature that was common to all of the linear models, but Akt, \gls{egfr}, \gls{tgfb} and Src pathway metagenes were included in many of the models, which suggested that these pathway signatures may have an important role in the obesity associated genetic signatures (\cref{sub:stepwise_linear_model_prediction_in_nzbc_and_cr_data_sets}).
 
@@
 To summarise, there were no common genetic signatures that significantly associated with the obesity status of the patients.
-However, the Akt, \gls{egfr} and Src pathways may have a role in promoting the tumour progression in patients that are obese.
+However, the Akt, \gls{egfr}, \gls{tgfb} and Src pathways may have a role in promoting the tumour progression in patients that are obese.
 Further investigations with good quality data sets and larger patient cohorts are required to clarify the biological relationship between obesity and cancer.
@@
 Although this study failed to identify obesity specific genes that affected tumour biology, it is likely that there is some complex mechanism underlying the relationship between obesity and cancer; otherwise many of the clinical and epidemiological associations between obesity and cancer risk would not make any sense.
-Since Akt, \gls{egfr} and \gls{pr} pathways were significant in the linear models that predicted the obesity metagenes, more thorough investigation of these pathways may be a good start to clarify the relationship.
+Since Akt, \gls{egfr}, \gls{tgfb} and Src pathways were significant in the linear models that predicted the obesity metagenes, more thorough investigation of these pathways may be a good start to clarify the relationship.
 Further analyses of \gls{blca} may help clarify the similarity between bladder and breast cancer types and provide insight into the association between the obesity associated genetic signatures and \gls{blca}.

### 2017-03-30 thesis/intro/intro.tex
tmp commit

@@
 The Pacific and M\=aori populations were affected the most, where 67\% and 47\% of the adults, respectively, were obese \citep{Health2016}.
-\gls{nz} not only has a high prevalence of obesity compared to many other countries, but also has a unique health problem that affects Pacific and M\=aori populations disproportionately than the rest of the \gls{nz} population.
+\gls{nz} not only has a high prevalence of obesity compared to many other countries, but also has a unique health problem that affects Pacific and M\=aori populations disproportionately to the rest of the \gls{nz} population.
 
@@
 The rates were 1.2 and 1.4 times greater in the M\=aori population compared to the non-M\=aori population for males and females, respectively \citep{Health2016b}.
-Again, this difference between the M\=aori and non-M\=aori is unique to \gls{nz}, and improved understanding of the underlying problem is essential for better treatment plans and outcomes for M\=aori.
+Again, this difference between M\=aori and non-M\=aori is unique to \gls{nz}, and improved understanding of the underlying problem is essential for better treatment plans and outcomes for M\=aori.
 

### 2017-03-30 thesis/methods/methods.tex
tmp commit

@@
 The raw microarray data from the \citet{Creighton2012}, \citet{Fuentes-Mattei2014} and \citet{Gatza2010a}  studies were downloaded from the \gls{geo} website (\citealp{Edgar2002}; \citetalias{GEO}).
-In addition to these breast cancer data sets, the raw microarray data set from a cohort of \gls{nz} breast cancer patients used in the \citet{Print2016} study was downloaded from the \gls{geo} website.
+In addition to these breast cancer data sets, the raw microarray data set from a cohort of \gls{nz} breast cancer patients used in the \citet{Print2016} study was downloaded from the \gls{geo} website (abbreviated as \gls{nzbc} data hereafter).
 \gls{rnaseq} and clinical data of multiple different cancer types were downloaded from the \gls{icgc} (\citetalias{ICGC}; \citealp{Zhang2011}) and \gls{tcga} \citepalias{TCGA} websites, respectively.
@@
 
-\subsubsection{\gls{nzbc} data}
+\subsubsection{\acrfull{nzbc} data}
 \label{ssub:nzbc_data}
@@
 
-\begin{table}[htpb]
+\begin{table}[p]
 	\caption{Summary of the total number of samples included in the analyses for each cancer type}
@@
 
-\begin{table}[htb]
+\begin{table}[p]
 	\caption{\gls{who} defined \gls{bmi} classification}
@@
 
-\begin{table}[htpb]
+\begin{table}[p]
 	\caption{Summary of the patient \gls{bmi} status in the \gls{icgc} cancer data}

### 2017-03-30 thesis/misc/abstract.tex
tmp commit

@@
 This research aims to determine whether gene expression signatures exist  that are specific to obesity across multiple cancer types, and to investigate whether there are any common pathways being dysregulated in cancers based on these genetic signatures.
-There were no genetic signatures or genes differentially expressed between obese and non-obese patients that were common across multiple cancer types.
-However, the Akt, \gls{egfr} and Src pathways may have a role in promoting the tumour progression in patients that are obese.
+In this work no genetic signatures or differentially expressed genes were found between obese and non-obese patients that were common across multiple cancer types.
+However, the Akt, \gls{egfr}, \gls{tgfb} and Src pathways may have a role in promoting the tumour progression in patients that are obese.
 It is likely that there is some complex mechanism underlying the relationship between obesity and cancer.

### 2017-03-30 thesis/misc/acknowledgement.tex
tmp commit

@@
 Especially James Boocock, for introducing me to a variety of useful tool sets (by the way, everyone should start using vim); Murray Cadzow, for organising the fortnightly ``Shit You Should Know About'' (a.k.a.
-SYSKA) sessions to teach us more about R and other useful tools; and Tom Kelly, for being my companion in the Fish-bowl and helping me out with some of the codes in my project.
+SYSKA) sessions to teach us more about R and other useful tools; and Tom Kelly, for being my companion in the Fish-bowl and helping me out with some of the code in my project.
 \\
@@
 I would like to thank all of my friends and family, for all their support and fun memories throughout the years that I have been at University.
-I wish I could name everyone, but I do not think it is a good idea to extend this thesis any longer\ldots
+I wish I could name everyone, but I don't think it's a good idea to extend this thesis any longer\ldots{}
 Long story short, my life would not be nearly as interesting or enjoyable without you guys.

### 2017-03-30 thesis/misc/appendix.tex
tmp commit

@@
 
+	\newpage
+
 	\section{Remainder of the results of the other CR obesity metagenes in the CR data set}
@@
 
-	All of the other obesity metagenes identified from the CR data set were able to capture the overall gene expressions of the samples in the CR data set (\cref{fig:appendix/cr_ob_meta_heatmap_cr}).
-	Furthermore, all of the CR obesity metagenes were significantly associated with the patient \gls{bmi}/\gls{bmi} status in the CR data set (\cref{fig:appendix/cr_ob_meta_box_scatter_cr}).
+	All of the other obesity metagenes identified from the CR data set were able to capture the overall gene expressions of the samples in the CR data set (\cref{fig:appendix/cr_ob_meta_heatmap_cris}).
+	Furthermore, all of the CR obesity metagenes were significantly associated with the patient \gls{bmi}/\gls{bmi} status in the CR data set (\cref{fig:appendix/cr_ob_meta_box_scatter_cris}).
 
@@
 
-	\begin{figure}[htpb]
+	\begin{figure}[h]
 		\centering
@@
 
-	All of the other obesity metagenes from the CR data set were able to capture the overall gene expressions of the samples in the \gls{nzbc} data set (\cref{fig:appendix/cr_ob_meta_heatmap_cris}).
-	However, none of the CR obesity metagenes were able to significantly associate with the sample \gls{bmi}/\gls{bmi} status in the \gls{nzbc} data set (\cref{fig:appendix/cr_ob_meta_box_scatter_cris}).
+	All of the other obesity metagenes from the CR data set were able to capture the overall gene expressions of the samples in the \gls{nzbc} data set (\cref{fig:appendix/rest_of_cr_ob_meta_heatmap_cris}).
+	However, none of the CR obesity metagenes were able to significantly associate with the sample \gls{bmi}/\gls{bmi} status in the \gls{nzbc} data set (\cref{fig:appendix/rest_of_cr_ob_meta_box_scatter_cris}).
 
@@
 		Scales are as described in previous figures.}
-		\label{fig:appendix/cr_ob_meta_heatmap_cris}
+		\label{fig:appendix/rest_of_cr_ob_meta_heatmap_cris}
 	\end{figure}
@@
 		P-values and $R^2$-value are as described in previous figures.}
-		\label{fig:appendix/cr_ob_meta_box_scatter_cris}
+		\label{fig:appendix/rest_of_cr_ob_meta_box_scatter_cris}
 	\end{figure}

### 2017-03-30 thesis/results/results.tex
tmp commit

@@
 
-\section{Obesity associated genetic signature from \citet{Creighton2012} study}
+\section{Obesity associated genetic signature from the\\\citet{Creighton2012} study}
 \label{sec:creighton_obesity_metagene}
@@
 Another reason for this apparent lack of association could be that the obesity associated genetic signature was too specific to the CR data and was not a broad obesity associated genetic signature, but an obesity associated signature that was specific to the patient cohort that was profiled in the \citet{Creighton2012} publication.
-% Another reason for this apparent lack of association could be that the genetic signature was too specific to the data and was not a broad obesity associated genetic signature, but an obesity associated signature that was specific to the samples that were profiled in the Creighton \textit{et al.} publication.
 
-\section{Obesity associated genetic signature from \citet{Fuentes-Mattei2014} study}
+\section{Obesity associated genetic signature from the\\\citet{Fuentes-Mattei2014} study}
 \label{sec:fm_obesity_metagene}
@@
 Nevertheless, the transformation matrix was still generated in the FM data and applied first to the microarray data (CR and \gls{nzbc} data sets) then to the \gls{icgc} data sets to see whether the FM obesity metagene associated with patient \gls{bmi} and \gls{bmi} status in these data sets.
-% However, the transformation matrix was generated in the FM data and applied to other cancer data sets  to see whether the FM obesity metagene associated with patient \gls{bmi} and \gls{bmi} status in those data sets.
 
@@
 These results suggested that the samples from the patients that were overweight in the \gls{blca} data set had similar biological properties as the samples taken from the patients that were obese in the FM data set.
-However, due to the fact that the FM obesity metagene lacked association with the patient \gls{bmi} in the \gls{blca} data set and that the metagene did not show any significant association in any of the other \gls{icgc} cancer types, it was difficult to determine whether the observed association of the FM obesity metagene with the overweight group was truly reflective of the quality of the FM metagene or observed by chance.
+However, due to the fact that the FM obesity metagene lacked association with the patient \gls{bmi} in the \gls{blca} data set, and that the metagene did not show any significant association in any of the other \gls{icgc} cancer types, it was difficult to determine whether the observed association of the FM obesity metagene with the overweight group was truly reflective of the quality of the FM metagene or observed by chance.
 
@@
 
-\section{Novel obesity associated genetic signatures from \citet{Creighton2012} data set}
+\section{Novel obesity associated genetic signatures from the \citet{Creighton2012} data set}
 \label{sec:creighton_obesity_metagene_new}
@@
 
-\begin{figure}[htpb]
+\begin{figure}[tpb]
 	\centering
@@
 	\gls{svd} was applied to the \gls{rma} normalised CR data to generate each of the eight metagenes.
-	High and low correlation were represented as red and blue, respectively, where the colours were matched with the values on the scale shown in the top right histogram.
+	High and low correlation were represented as red and yellow, respectively, where the colours were matched with the values on the scale shown in the top right histogram.
 	The lowest correlation was 0.8469, between the CrOl and the CaRes metagenes.

### 2017-03-30 thesis/results/results2.tex
pretty much ready to submit - tmp commit
tmp commit

@@
 
-\section{Pathway associated genetic signatures from \citet{Gatza2010a} study}
+\section{Pathway associated genetic signatures from the\\\citet{Gatza2010a} study}
 \label{sec:pathway_associated_genetic_signatures_from_gatza2010a_study}
@@
 As an example, the \gls{tgfb} metagenes were significantly more dispersed compared to the \gls{pr} metagenes, even though both of these metagenes were generated in the same data set through similar processes (\cref{fig:gt_rma_vs_mas}).
-These differences were seen in other data sets as well (\cref{fig:appendix/gt_meta_rma_mas_other_data}).
+These differences were seen in other data sets as well (\cref{fig:appendix/gt_meta_rma_mas_other_data} in \cref{sec:normalisation_method_for_pathway_associated_genetic_signature_transformation_matrices}).
 This result provided evidence that some of the pathway genetic signatures from the \citet{Gatza2010a} study were less variable across different data sets than the others.
@@
 
-\begin{table}[htpb]
+\begin{table}[tpb]
 	\centering
@@
 Taken together, these results suggest that the variables included in these stepwise models were relevant to the biology of the obesity metagenes identified in the CR and FM data sets.
-However, there were no common pathways that were present in all of the models; although Akt, \gls{egfr} and Src pathway metagenes were frequently included in many of the stepwise models.
+However, there were no common pathways that were present in all of the models; although Akt, \gls{egfr}, \gls{tgfb} and Src pathway metagenes were frequently included in many of the stepwise models.
 The role of the pathways included in the models and the possible biological links of these pathways to the obesity metagenes remain unclear, and should be the focus for future investigations.

### 2017-03-30 thesis/intro/intro.tex
Added Gerry's edit - tmp commit

@@
 Although urbanisation has many advantages, such as access to developed health care systems, education and advanced technologies, the lifestyle change also imposes negative health behaviour that ultimately can lead to positive energy balance and obesity \citep{Malik2013}.
-Lack of physical activity is one of the obvious components linked with urbanisation.
-In many urban cities, the recommended level of physical exercise is often difficult to achieve and maintain due to the shift towards sedentary behaviours that are encouraged by developed transport systems, automated household chores and indoor entertainment \citep{Malik2013}.
+Lack of physical activity is one of the clear components linked with urbanisation.
+In many urban cities, it is often difficult to achieve and maintain the recommended level of physical exercise due to the shift towards sedentary behaviours that are encouraged by developed transport systems, automated household chores and indoor entertainment \citep{Malik2013}.
 
@@
 
-So, for the low- and middle-income groups, the rise in the average income in these groups allows people to have access to food and activities that promotes sedentary behaviours, but the limited access to the health care system, education and recreational activities (that promote physical activity) ultimately leads to positive energy balance \citep{Malik2013}.
+Therefore, for the low- and middle-income groups, the rise in the average income in these groups allows people to have access to food and activities that promotes sedentary behaviours, but the limited access to the health care system, education, and recreational activities (that promote physical activity) ultimately leads to positive energy balance \citep{Malik2013}.
 In contrast, the high-income group of the population have better access to facilities that promote weight maintenance, thus have less of an impact compared with the low- and middle-income groups of the population.
@@
 More than half of cancer deaths are caused by lung, breast, colorectal, stomach and liver cancers \citep{WHO2014}.
-The leading cause of cancer deaths in high-income countries is lung cancer in both men and women, then colorectal cancer and breast cancer for men and women, respectively \citep{WHO2014}.
+The leading cause of cancer deaths in high-income countries is lung cancer for both men and women, then colorectal cancer and breast cancer for men and women, respectively \citep{WHO2014}.
 Although the levels of mortality vary among different countries, cervical, liver and stomach cancers make up a large proportion of cancer deaths among low- and middle-income countries \citep{WHO2014}.
@@
 Extensive research in cancer biology has uncovered many pathways and regulatory mechanisms involved in cancer.
-With more than 100 distinct types and subtypes of cancers, each with its own distinct development mechanisms, it became difficult to focus on the overarching mechanisms of cancer development \citep{Hanahan2000}.
+With more than 100 distinct types and subtypes of cancers (each with its own distinct development mechanisms), it became difficult to focus on the overarching mechanisms of cancer development \citep{Hanahan2000}.
 To resolve this, \citet{Hanahan2000} proposed ``six essential alterations'' that cancer cells usually take advantage of during development and progression.
@@
 Secondly, they can express more receptors for a given growth signal, and thus become ``hyperresponsive'' to the signal \citep{Hanahan2000,Hanahan2011}.
-Thirdly, cancer cells can stimulate the neighbouring normal cells within the tumour microenvironment to produce more growth factors that further assist them to grow \citep{Bhowmick2004, Liotta2001, Wiseman2002}.
+Thirdly, cancer cells can stimulate the neighbouring normal cells within the tumour microenvironment to produce more growth factors that further assist growth \citep{Bhowmick2004, Liotta2001, Wiseman2002}.
 Finally, cancer cells are able to maintain proliferative signalling by mutating the growth factor receptor itself, or the downstream signalling targets of the growth factor receptor pathway \citep{Fuqua1991,SuHuang1997,Satyamoorthy2003}.
@@
 When abundant and active, p53 protein transcribes the p21 protein which, through series of events in the p16-cyclin D$_1$-cdk4-\gls{rb} pathway, triggers cell cycle arrest \citep{Levine1997}.
-In many cancers, p53 protein is mutated and selected against to allow continuous growth of the tumour cells.
+In many cancers, the p53 protein is mutated and selected against to allow continuous growth of the tumour cells.
 
@@
 \noindent
-\Gls{apoptosis}, the natural and controlled death of a cell, is one of many defence mechanisms in the body to prevent the development of cancer.
+\Gls{apoptosis}, the natural and controlled death of a cell and is one of many defence mechanisms in the body to prevent the development of cancer.
 The triggering of \gls{apoptosis} is controlled and maintained by the balance between the pro- and anti-apoptotic signals that affect the downstream apoptotic proteins \citep{Hanahan2011}.
 
-The precise cellular conditions that trigger apoptosis are yet to be uncovered, but there is a clear evidence of p53 tumour suppressor playing a key role in apoptosis initiation.
-In short, when \acrshort{dna} damage is sensed, p53 protein transcribes the pro-apoptotic members of the \gls{bcl2} family proteins, such as Bax, Puma and Noxa \citep{Fridman2003,Hanahan2011}.
-The fact that the p53 protein is involved not only in the regulation of cell cycle, but also in the apoptotic pathway shows its importance in the control of cancer development, and perhaps that is the reason why more than 50\% of tumours have a mutation in the \textit{TP53} gene \citep{Levine1997}.
+The precise cellular conditions that trigger apoptosis are yet to be uncovered, but there is a clear evidence of the p53 tumour suppressor playing a key role in apoptosis initiation.
+In short, when \acrshort{dna} damage is sensed, the p53 protein transcribes the pro-apoptotic members of the \gls{bcl2} family proteins, such as Bax, Puma and Noxa \citep{Fridman2003,Hanahan2011}.
+The fact that the p53 protein is involved not only in the regulation of the cell cycle, but also in the apoptotic pathway shows its importance in the control of cancer development, and perhaps that is the reason why more than 50\% of tumours have a mutation in the \textit{TP53} gene \citep{Levine1997}.
 
@@
 There are two barriers in achieving infinite replicative capacity: senescence and crisis \citep{Hanahan2011}.
-When cells are grown in a culture, cells are able to replicate and grow until a certain point, and the cells reach the stage of senescence where the cells stop growing but are still viable \citep{Hanahan2011}.
+When cells are grown in a culture, the cells are able to replicate and grow until a certain point, and the cells reach the stage of senescence where the cells stop growing but are still viable \citep{Hanahan2011}.
 The cells that manage to bypass this senescence phase by mutating the tumour suppressor proteins such as the p53 protein, undergoes the crisis phase where the majority of the cells in the culture dies \citep{Hanahan2011}.
@@
 Like normal cells, cancer cells also require access to oxygen and nutrient supplies from the blood.
-However, new blood vessel formation, or angiogenesis, only occurs in a handful of occasions, and thus the tumour needs to acquire the ability to stimulate angiogenesis in the microenvironment in which they live in \citep{Hanahan2011}.
+However, new blood vessel formation, or angiogenesis, only occurs in a handful of occasions, and thus the tumour needs to acquire the ability to stimulate angiogenesis in the microenvironment in which they live \citep{Hanahan2011}.
 
@@
 One of the prime reasons that cancers are so difficult to treat and have a high mortality rate is due to their ability to metastasise to other parts of the body.
-Once metastasised to different parts of the body, the clinician not only has to consider the primary tumour site, but also the second, third, or even fourth tumour sites, which may or may not have similar tumour biology as the primary tumour, and thus has to consider multiple treatment plans to cure the patient.
+Once metastasised to different parts of the body, the clinician not only has to consider the primary tumour site, but also the second, third, or even fourth tumour sites.
+These tumours may or may not have similar tumour biology as the primary tumour, and thus has to consider multiple treatment plans to cure the patient.
 
@@
 There is now evidence that shows otherwise, whereby a single catastrophic event (termed \gls{chromothripsis}) occurs in the chromosome structure, resulting in major structural rearrangements such as inversions, deletions, duplications and translocations \citep{Leibowitz2015,Stephens2011}.
-Owing to its nature of being a one-off chromosomal disruption, \gls{chromothripsis} (usually) cause disruptions in tumour suppressor genes rather than duplications of oncogenes \citep{Leibowitz2015}.
+As it is a one-off chromosomal disruption, \gls{chromothripsis} (usually) cause disruptions in tumour suppressor genes rather than duplications of oncogenes \citep{Leibowitz2015}.
 
@@
 \noindent
-Inflammation and immune response, though it may assist in the control of diseases and injuries, have been implicated in the progression of tumour and their acquisition of many of the cancer hallmarks \citep{Hanahan2011}.
+While inflammation and immune response may assist in the control of diseases and injuries, they have been implicated in the progression of tumour and their acquisition of many of the cancer hallmarks \citep{Hanahan2011}.
 There are two key transcription factors of inflammation that ultimately leads to the promotion  of many, if not all, of the hallmarks discussed so far: \gls{nfkb} and \gls{stat3} \citep{Mantovani2008}.
 
-The study by \citep{Pikarsky2004} showed that \gls{nfkb} was an essential component for the progression of hepatocellular carcinoma, providing the first evidence as to how inflammation may affect the tumour biology.
+The study by \citet{Pikarsky2004} showed that \gls{nfkb} was an essential component for the progression of hepatocellular carcinoma, providing the first evidence as to how inflammation may affect the tumour biology.
 \gls{nfkb} regulates the expression of genes that encode the inflammatory cytokines, angiogenic factors, and many others, including anti-apoptotic genes (through the activation of \gls{stat3}) \citep{Elinav2013,Mantovani2008}.
@@
 By increasing the transcription of anti-apoptotic proteins such as \gls{bcl2}, \gls{stat3} promotes tumour survival and proliferation \citep{Yu2007}.
-Furthermore, \gls{stat3} is able to increase the expression of inflammatory cytokines (such as \gls{il6} and \acrshort{il10}) and \acrshort{vegf}, which provides a positive feedback mechanism for further inflammation (and therefore survival and proliferation signals) and advancement of angiogenesis, respectively \citep{Yu2007}.
+Furthermore, \gls{stat3} is able to increase the expression of inflammatory cytokines (such as \gls{il6} and \acrshort{il10}) and \acrshort{vegf} \citep{Yu2007}.
+These cytokines provide a positive feedback mechanism for further inflammation (and therefore survival and proliferation signals) and advancement of angiogenesis, respectively \citep{Yu2007}.
 
@@
 \noindent
-In order to grow and proliferate, cancer cells must adjust their energy metabolism to match their energy requirement \citet{Hanahan2011}.
-\citet{Wardburg1956} found that cancer cells, even with an abundance of oxygen in the cell, utilised the glycolytic pathway instead of the more efficient aerobic respiration, and was later termed ``aerobic glycolysis'' \citep{Hanahan2011}.
-It seems counterintuitive to limit its full potential of creating \gls{atp} by utilising glycolysis over electron transport chain, as the latter produces almost 18 times more \gls{atp} \citep{Hanahan2011, VanderHeiden2009}.
-However, it is hypothesised  that this apparent inefficiency helps the cell to generate and direct the molecules required for biosynthetic pathways during cell growth and division \citep{Cairns2011,VanderHeiden2009}.
-By deliberately relying purely on glycolysis (known as the ``Wardberg effect'') and slowing the conversion of \gls{pep} to pyruvate, the cell is able to direct the carbon molecules into reaction pathways that produce the constituents used to synthesise macromolecules, such as \acrshort{dna} and lipids \citep{Cairns2011,VanderHeiden2009}.
+In order to grow and proliferate, cancer cells must adjust their energy metabolism to match their energy requirement \citep{Hanahan2011}.
+\citet{Wardburg1956} found that even if cancer cells have an abundance of oxygen in the cell, they utilised the glycolytic pathway instead of the more efficient aerobic respiration, and was later termed ``aerobic glycolysis'' \citep{Hanahan2011}.
+It seems counterintuitive to limit its full potential of creating \gls{atp} by utilising glycolysis over \gls{etc}, as the latter produces almost 18 times more \gls{atp} \citep{Hanahan2011, VanderHeiden2009}.
+However, it is hypothesised  that this inefficiency helps the cell to generate and direct the molecules required for biosynthetic pathways during cell growth and division \citep{Cairns2011,VanderHeiden2009}.
+By deliberately relying purely on glycolysis (known as the ``Wardberg effect'') and slowing the conversion of \gls{pep} to pyruvate, the cell is able to direct the carbon molecules into reaction pathways \citep{Cairns2011,VanderHeiden2009}.
+These produce the constituents used to synthesise macromolecules, such as \acrshort{dna} and lipids \citep{Cairns2011,VanderHeiden2009}.
 
@@
 Our immune system is constantly monitoring and maintaining normal cell biology within our body, preventing the emergence of cancer \citep{Hanahan2011}.
-With that said, cancers are frequently observed in clinics, pointing to the fact that the observed cancer cells must have evaded the diligent monitoring by the immune system \citep{Hanahan2011}.
+With that said, cancers are frequently observed in clinics.
+This highlights how the cancer cells can evade the diligent monitoring by the immune system \citep{Hanahan2011}.
+% With that said, cancers are frequently observed in clinics, pointing to the fact that the observed cancer cells must have evaded the diligent monitoring by the immune system \citep{Hanahan2011}.
 
@@
 
-There are two types of mutations that occurs in an individual: \gls{germline} and \gls{somatic}.
+There are two types of mutations that occur in an individual: \gls{germline} and \gls{somatic}.
 \Glspl{germline} are the mutations that are passed on from the parents of an individual, and \glspl{somatic} are the mutations acquired by the cell over the course of the lifetime of an individual.
@@
 
-Smoking is a known risk factor for many cancers; the obvious one being lung cancer \citep{Gandini2008,Hecht1999}.
+Smoking is a known risk factor for many cancers; the main one being lung cancer \citep{Gandini2008,Hecht1999}.
 For current smokers, the relative risk is greater than 1.5 in many cancers, including liver, cervix and kidney, and it can rise up to 8.96 in lung cancer \citep{Gandini2008}.
-Tobacco smoke contains many \glspl{carcinogen} which are processed first by \gls{cyp} enzymes (termed meta\-bolic activation), and the product can either be processed further to be detoxified completely and cause no harm to the cell, or alternatively, the product can covalently bind to the \acrshort{dna}, forming \acrshort{dna} adducts \citep{Hecht1999}.
+Tobacco smoke contains many \glspl{carcinogen} which are processed first by \gls{cyp} enzymes (termed meta\-bolic activation) \citep{Hecht1999}.
+The resulting product from this reaction can either be processed further to be detoxified completely and cause no harm to the cell, or alternatively, the product can covalently bind to the \acrshort{dna}, forming \acrshort{dna} adducts \citep{Hecht1999}.
 Unless these adducts are resolved correctly by the \acrshort{dna} repair mechanisms, it can result in a permanent mutation within the genome and lead to oncogenesis \citep{Hecht1999}.
@@
 
-Diet and obesity are known risk factors for many types of cancers as well \citep{Ames1995,Calle2004}.
+Diet and obesity are also known risk factors for many types of cancers \citep{Ames1995,Calle2004}.
 Epidemiological evidence of obesity being a risk factor for many cancers has been well established, yet the mechanisms of its contribution to cancer is still  speculative \citep{Calle2003,Kelesidis2006}.
@@
 Free fatty acids from excess adipose tissue and/or diet can be recognised by \gls{tlr} (specifically \gls{tlr}4), which can then activate the \gls{nfkb} pathway, leading to the activation of one of the Hallmarks of Cancer (tumour-promoting inflammation) \citep{Lumeng2011}.
-There are many other environmental factors associated with cancers and limiting the exposure to detrimental environmental factors is crucial for the prevention of many cancers.
+There are many other environmental factors associated with cancers, thus limiting the exposure to detrimental environmental factors is crucial for the prevention of many cancers.
 
@@
 Traditionally, the treatment of cancer has been surgery, radiation therapy and chemotherapy.
-Though surgery is an effective method to remove the tumour from the patient, there is always risk of metastasis and recurrence, which is why a combination of surgery with other treatments like chemotherapy is usually considered.
+Though surgery is an effective method to remove the tumour from the patient, there is always risk of metastasis and recurrence.
+This is why a combination of surgery with other treatments like chemotherapy is usually considered.
 Previously, chemotherapeutic drugs and radiation therapy had a limited effect on tumours due to their resistance against these treatments \citep{Wilhelm2006}.
@@
 The RAS-RAF-MEK-ERK signalling pathway is  involved in cell survival, growth and proliferation \citep{Samatar2014,Wilhelm2006}.
-The members of this pathway are often mutated in many types of cancer (approximately a third of cancers contain a mutation in the RAS oncogene), causing an overactivation of this pathway \citep{Samatar2014}.
+The components of this pathway are often mutated in many types of cancer (approximately a third of cancers contain a mutation in the RAS oncogene), causing an overactivation of this pathway \citep{Samatar2014}.
 Development of drugs that target  the downstream RAF and its mutant forms (such as BRAF$^{V600E}$) have been successful, and many RAF inhibitor drugs such as sorafenib and vemurafenib have been approved for clinical use \citep{Samatar2014,Wilhelm2006}.
@@
 
-In fact, there is evidence of pathway cross-talks between the RAS-ERK pathway and the \gls{pi3k}/Akt pathway, which may help the tumour cells achieve resistance to the  drugs that block the RAS-ERK pathway \citep{Moelling2002,Zimmermann1999a}.
+In fact, there has been evidence of pathway cross-talks between the RAS-ERK pathway and the \gls{pi3k}/Akt pathway, which may help the tumour cells achieve resistance to the  drugs that block the RAS-ERK pathway \citep{Moelling2002,Zimmermann1999a}.
 A study in a \textit{Drosophila} model showed that the inhibition of just a single pathway does not necessarily stop the effect of the pathway, as the signal can flow through other pathways \citep{Dar2012}.
@@
 
-Although the mechanisms by which obesity causes cancer are still under debate, the association between obesity and cancer has long been established.
+Although the mechanisms through which obesity causes cancer are still under debate, the association between obesity and cancer has long been established.
 Perhaps the most convincing evidence was first presented by \citet{Calle2003}.
@@
 Leptin is known to activate not only the satiety signalling pathway, but also the \gls{jak}/\gls{stat} and \gls{pi3k}/Akt signalling pathways \citep{Garofalo2006,Renehan2006}.
-Activation of these pathways lead to the activation of the target genes that are involved in cell proliferation and cell survival, as well as pro-angiogenic factors \citep{Garofalo2006}.
+Activation of these pathways leads to the activation of the target genes that are involved in cell proliferation and cell survival, as well as pro-angiogenic factors \citep{Garofalo2006}.
 
-Adiponectin is a hormone that is synthesised and secreted only by the \glspl{adipocyte} \citep{Kelesidis2006}.
+Adiponectin is a hormone that is synthesised and secreted only by \glspl{adipocyte} \citep{Kelesidis2006}.
 Unlike leptin, where the concentration of leptin is higher in those who are obese, adiponectin levels are lower in people that are obese compared to those with a normal weight \citep{Kelesidis2006,Renehan2006}.
@@
 Alternatively, or possibly in conjunction with the disruption of hormone levels, inflammation caused by obesity may be associated with a tumour-promoting environment.
-\Gls{tlr}4 is a type of \gls{prr} and, as mentioned earlier, is able to recognise circulating free fatty acids and trigger immune response via \gls{nfkb} pathway \citep{Lumeng2011}.
+\Gls{tlr}4 is a type of \gls{prr} and, as mentioned earlier, is able to recognise circulating free fatty acids and trigger immune response via the \gls{nfkb} pathway \citep{Lumeng2011}.
 In addition to \gls{nfkb} pathway, recent evidence has shown that \gls{tlr}4 could also activate the \gls{jak}/\gls{stat3} pathway \citep{Yu2014}.
@@
 
-Finally, insulin/\gls{igf} pathway may have a major role in encouraging tumorigenesis in obese patients.
+Finally, the insulin/\gls{igf} pathway may have a major role in encouraging tumorigenesis in obese patients.
 \gls{igf} signalling results in the activation of \gls{pi3k}/Akt and \gls{erk} pathways, resulting in cell proliferation and reduced apoptosis \citep{Roberts2010}.
@@
 As one can imagine, it is difficult for clinicians to treat a patient if they do not know what is causing the disease.
-The reason it is so difficult to identify the underlying biological mechanisms is that there are very few reliable biological markers (or ``genetic signatures'') that signify the presence or absence of the causal pathways.
-With that said, significant improvements have been made in recent years owing to the advancement in the sequencing technologies and have allowed researchers to study the genetic signatures that represent the hallmarks in much greater detail.
+It is difficult to identify the underlying biological mechanisms because there are very few reliable biological markers (or ``genetic signatures'') that signify the presence or absence of the causal pathways.
+With that said, significant improvements have been made in recent years due to the advancement in the sequencing technologies.
+This has allowed researchers to study the genetic signatures that represent the hallmarks in much greater detail.
 
@@
 One disadvantage of the microarray technology is that it requires previous knowledge of the genes and \glspl{snp} of the samples being investigated \citep{Hurd2009}.
-This is due to the fact that the detection of sequences relies on the hybridisation of the sample sequences with the microarray sequences and therefore the sequences under investigation must be known prior to the investigation to attach the correct sequences to the microarray.
+This is due to the fact that the detection of sequences relies on the hybridisation of the sample sequences with the microarray sequences.
+Therefore, the sequences under investigation must be known prior to the investigation to attach the correct sequences to the microarray.
 Another disadvantage of microarrays is the potentially erroneous detection of the sample sequences, both in terms of its sequence and its expression level \citep{Hurd2009}.
 Again, because the microarray slides require pre-defined sequences to be attached to it, there is always a possibility of hybridisation of similar but different sequence to the slide.
-Even if the sequences correctly hybridises to the slide, it may be difficult to measure the exact amount of some sequences in the samples, since the amount of sequences in the sample is measured with the relative strength of the fluorescence of the dyes \citep{Hurd2009}.
+Even if the sequences correctly hybridise to the slide, it may be difficult to measure the exact amount of particular sequences in the samples, since the amount of sequences in the sample is measured with the relative strength of the fluorescence of the dyes \citep{Hurd2009}.
 Therefore, it is difficult to detect the exact level of expressions of the sequences that are either very rare or highly abundant in the sample \citep{Hurd2009}.
@@
 \Gls{ngs} technology refers to the high throughput sequencing techniques that allows the sequencing of a large quantity of genetic material \citep{Metzker2010}.
-\gls{ngs} technology has been developed and improved in the last decade at a revolutionary pace, making it possible for researchers to study gene expressions and variations in days to weeks, where the same experiment would have taken years to complete before the technology was developed.
+\gls{ngs} technology has been developed and improved in the last decade at a revolutionary pace.
+This has made it possible for researchers to study gene expressions and variations in days to weeks, where the same experiment would have taken years to complete before the technology was developed.
 
@@
 This also prevents the introduction of experimental bias, as the sequences are not altered in any way and no specific sequence is searched in any greater detail over another \citep{Hurd2009}.
-Another advantage is that \gls{ngs} produce count data of the genetic material being sequenced instead of relying on the intensity of the fluorescence to detect the presence and amount of certain sequence.
+Another advantage is that \gls{ngs} produces count data of the genetic material being sequenced instead of relying on the intensity of the fluorescence to detect the presence and amount of certain sequence.
 This allows an absolute quantification of the amount of transcripts from the sample, making it possible for a more accurate comparison of the expression levels between samples.
@@
 
-There is clear epidemiological evidence of obesity being a major risk factor for many cancer types (\cref{sec:obesity_and_cancer}), but the genes or pathways that directly contribute to tumorigenesis in patients that are obese is yet to be clarified.
+There is clear epidemiological evidence of obesity being a major risk factor for many cancer types (\cref{sec:obesity_and_cancer}), but the genes or pathways that directly contribute to tumorigenesis in patients that are obese are yet to be clarified.
 Two studies in particular, one by \citet{Creighton2012} and another by \citet{Fuentes-Mattei2014}, have investigated whether there were any differences in the genetic composition between obese and non-obese patients, specifically in a group of breast cancer patients.
@@
 Of the 103 samples, 35\% were from normal weight, 28\% were from overweight and 37\% were from obese patients.
-It was also noted that African-American patients were more likely to be obese than the Caucasian patients.
+It was also noted that African-American patients were more likely to be obese than Caucasian patients.
 They compared the gene expression between the tumours from obese patients with the tumours from non-obese patients and identified 799 gene probe sets (662 unique genes) with p \textless{} 0.01 and with a \gls{fc} \textgreater{} 1.2.
@@
 \citet{Fuentes-Mattei2014} studied 137 breast tumour samples from \gls{er} positive patients and analysed their transcriptome using microarrays.
-Of the 137 patients, 43 patients were obese and 94 patients were not.
+Of the 137 patients, 43 patients were obese and 94 patients were not obese.
 Comparison of the gene expressions of the tumour samples from the obese patients with the tumour samples from the non-obese patients revealed 112 genes that were statistically significantly dysregulated (62 genes upregulated and 50 genes downregulated).
@@
 Like obesity associated genetic signatures, there has also been a focus on the identification of pathways and genes that affect tumour growth.
-About a decade ago, \citet{Bild2006} demonstrated that a collection of genes that are involved in a specific pathway could be used to determine the status of those pathways in tumour samples.
+In 2006, Bild \textit{et al.} demonstrated that a collection of genes that are involved in a specific pathway could be used to determine the status of those pathways in tumour samples \citep{Bild2006}.
 These pathway specific genes were identified experimentally by altering the expression of one specific gene that is central to the pathway in a model cell culture system.
@@
 The 18 pathways used in this study were: Akt, \gls{bcat}, E2F1, \gls{egfr}, \gls{er}, \gls{her2}, \gls{ifna}, \gls{ifny}, Myc, p53, p63, \gls{pi3k}, \gls{pr}, Ras, \gls{stat3}, Src, \gls{tgfb}, and \gls{tnfa} pathways.
-This was a significant improvement from their previous works on pathway signatures \citep{Bild2006,Bild2009}.
-Using more pathways in their analysis have allowed them to identify certain pathways to be coactivated (or cluster) together, suggesting that some of the pathways are related to one another and may have relevance in identifying the biological causes of cancer.
+This was a significant improvement from their previous work on pathway signatures \citep{Bild2006,Bild2009}.
+Using more pathways in their analysis has allowed them to identify certain pathways that are coactivated (or cluster) together, suggesting that some of the pathways are related to one another and may have relevance in identifying the biological causes of cancer.
 

### 2017-03-30 thesis/misc/glossary.tex
edited in Mik's comments and spell-checked everything
Added Gerry's edit - tmp commit

@@
 	name=RNA-seq,
-	description={sequence data using transcriptome}
+	description={Sequence data using the transcriptome}
 }
@@
 	plural=Type I errors,
-	description={Incorrect rejection of a true null hypothesis; False positive}
+	description={Incorrect rejection of a true null hypothesis; false positive}
 }
@@
 	plural=Type II errors,
-	description={Incorrect retaining of a false null hypothesis; False negative}
+	description={Incorrect retaining of a false null hypothesis; false negative}
 }
@@
 	text=power,
-	description={Probability that it will reject a false null hypothesis; It is inversely related to the probability of making a Type II error}
+	description={Probability that it will reject a false null hypothesis; it is inversely related to the probability of making a Type II error}
 }
@@
 
-\newacronym[longplural={Disability-Adjusted Life Years}]{daly}{DALY}{Disability-Adjusted Life Year}
-\newacronym[longplural={genome wide association studies},\glsshortpluralkey={GWAS}]{gwas}{GWAS}{genome wide association study}
-\newacronym{acth}{ACTH}{adrenocorticotropic hormone}
-\newacronym{agrp}{AGRP}{agouti-related protein}
-\newacronym{aic}{AIC}{Akaike Information Criterion}
-\newacronym{aid}{AID}{activation-induced cytidine deaminase}
-\newacronym{anova}{ANOVA}{analysis of variance}
-\newacronym{apobec}{APOBEC}{apolipoprotein B mRNA editing enzyme, catalytic polypeptide-like}
-\newacronym{atp}{ATP}{adenosine triphosphate}
-\newacronym{bcat}{BCAT}{$\beta$-catenin}
-\newacronym{bcl2}{Bcl-2}{B-cell lymphoma 2}
-\newacronym{ber}{BER}{base-excision repair}
-\newacronym{bh3}{BH3}{Bcl-2 homology 3}
-\newacronym{blca}{BLCA}{bladder urothelial carcinoma}
-\newacronym{bmi}{BMI}{body mass index}
-\newacronym{camera}{CAMERA}{correlation adjusted mean rank gene set test}
-\newacronym{cart}{CART}{cocaine and amphetamine related transcript}
-\newacronym{cdna}{cDNA}{complimentary DNA}
-\newacronym{cesc}{CESC}{cervical squamous cell carcinoma and endocervical adenocarcinoma}
-\newacronym{coad}{COAD}{colon adenocarcinoma}
-\newacronym{cpm}{cpm}{counts per million}
-\newacronym{cyp}{CYP}{cytochrome P450}
-\newacronym{deg}{DEG}{differentially expressed gene}
-\newacronym{dna}{DNA}{deoxyribonucleic acid}
-\newacronym{egfr}{EGFR}{epidermal growth factor receptor}
-\newacronym{emt}{EMT}{endothelial-mesenchymal transition}
-\newacronym{erk}{ERK}{extracellular signal regulated kinase}
-\newacronym{er}{ER}{estrogen receptor}
-\newacronym{fcs}{FCS}{functional class scoring}
-\newacronym{fc}{FC}{fold change}
-\newacronym{fdr}{FDR}{false discovery rate}
-\newacronym{fto}{FTO}{fat mass and obesity associated}
-\newacronym{fto}{FTO}{fat mass and obesity associated}
-\newacronym{fwer}{FWER}{family-wise error rate}
-\newacronym{g1}{G1}{Gap 1}
-\newacronym{gaba}{GABA}{$\gamma$-aminobutyric acid}
-\newacronym{geo}{GEO}{Gene Expression Omnibus}
-\newacronym{ghsr}{GHSR}{growth hormone secretagogue receptor}
-\newacronym{go}{GO}{Gene Ontology}
-\newacronym{gpcr}{GPCR}{G-protein coupled receptor}
-\newacronym{gsea}{GSEA}{Gene Set Enrichment Analysis}
-\newacronym{her2}{HER2}{human epidermal growth factor receptor 2}
-\newacronym{icgc}{ICGC}{International Cancer Genome Consortium}
-\newacronym{ifna}{IFN$\alpha$}{interferon-$\alpha$}
-\newacronym{ifny}{IFN$\gamma$}{interferon-$\gamma$}
-\newacronym{igfbp}{IGFBP}{insulin-like growth factor binding protein}
-\newacronym{igf}{IGF}{insulin-like growth factor}
-\newacronym{il10}{IL-10}{interleukin-10}
-\newacronym{il6}{IL-6}{interleukin-6}
-\newacronym{irs}{IRS}{insulin receptor substrate}
-\newacronym{jak}{JAK}{janus kinase}
-\newacronym{kegg}{KEGG}{Kyoto Encyclopedia of Genes and Genomes}
-\newacronym{kirp}{KIRP}{kidney renal clear cell carcinoma}
-\newacronym{lihc}{LIHC}{liver hepatocellular carcinoma}
-\newacronym{ln}{LN}{lymph node}
-\newacronym{mas}{MAS5}{microarray analysis suite 5.0}
-\newacronym{mc1r}{MC1R}{melanocortin 1 receptor}
-\newacronym{mc4r}{MC4R}{melanocortin 4 receptor}
-\newacronym{met}{MET}{mesenchymal-endothelial transition}
-\newacronym{mmr}{MMR}{mismatch repair}
-\newacronym{mm}{MM}{mismatch}
-\newacronym{mrna}{mRNA}{messenger RNA}
-\newacronym{msh}{MSH}{melanocyte-stimulating hormone}
-\newacronym{mtor}{mTOR}{mammalian target of rapamycin}
-\newacronym{m}{$\mu$}{mean}
-\newacronym{ner}{NER}{nucleotide-excision repair}
-\newacronym{nfkb}{NF-$\kappa$B}{nuclear factor $\kappa$B}
-\newacronym{ngs}{NGS}{next generation sequencing}
-\newacronym{nnk}{NNK}{4-(methylnitrosamino)-1-(3-pyridyl)-1-butanone}
-\newacronym{npy}{NPY}{neuropeptide Y}
-\newacronym{nyr1}{Y1R}{neuropeptide Y1 receptor}
-\newacronym{nzbc}{NZBC}{New Zealand Breast Cancer}
-\newacronym{nz}{NZ}{New Zealand}
-\newacronym{opa}{OPA}{obesity policy action}
-\newacronym{ora}{ORA}{over representation analysis}
-\newacronym{pah}{PAH}{polycyclic aromatic hydrocarbon}
-\newacronym{pc1}{PC1}{prohormone convertase 1}
-\newacronym{pep}{PEP}{phosphoenol pyruvate}
-\newacronym{pi3k}{PI3K}{phosphatidylinositol 3-kinase}
-\newacronym{pm}{PM}{perfect match}
-\newacronym{pomc}{POMC}{pro-opiomelanocortin}
-\newacronym{prr}{PRR}{pattern recognition receptor}
-\newacronym{pr}{PR}{progesterone receptor}
-\newacronym{pyy}{PYY$_{3-36}$}{peptide YY$_{3-36}$}
-\newacronym{rb}{RB}{retinoblastoma-associated}
-\newacronym{read}{READ}{rectum adenocarcinoma}
-\newacronym{rma}{RMA}{robust multi array}
-\newacronym{rna}{RNA}{ribonucleic acid}
-\newacronym{ros}{ROS}{reactive oxygen species}
-\newacronym{sd}{$\sigma$}{standard deviation}
-\newacronym{skcm}{SKCM}{skin cutaneous melanoma}
-\newacronym{snp}{SNP}{single nucleotide polymorphism}
-\newacronym{ssdna}{ssDNA}{single-stranded DNA}
-\newacronym{stat3}{STAT3}{signal transducer and activator of transcription 3}
-\newacronym{stat}{STAT}{signal transducer and activator of transcription}
-\newacronym{svd}{SVD}{singular value decomposition}
-\newacronym{tcga}{TCGA}{The Cancer Genome Atlas}
-\newacronym{tgfb}{TGF$\beta$}{transforming growth factor-$\beta$}
-\newacronym{tlr}{TLR}{Toll-like receptor}
-\newacronym{tnfa}{TNF$\alpha$}{tumour necrosis factor-$\alpha$}
-\newacronym{tsp1}{TSP-1}{thrombospondin-1}
-\newacronym{ucec}{UCEC}{uterine corpus endometrial carcinoma}
-\newacronym{usa}{USA}{United States of America}
-\newacronym{uv}{UV}{ultraviolet}
-\newacronym{vegfa}{VEGF-A}{vascular endothelial growth factor-A}
-\newacronym{vegf}{VEGF}{vascular endothelial growth factor}
-\newacronym{vif}{VIF}{variance inflation factor}
-\newacronym{who}{WHO}{World Health Organization}
-\newacronym{wmw}{WMW}{Wilcoxon-Mann-Whitney}
+\newacronym[description={$\beta$-catenin}]{bcat}{BCAT}{$\beta$-catenin}
+\newacronym[description={$\gamma$-aminobutyric acid}]{gaba}{GABA}{$\gamma$-aminobutyric acid}
+\newacronym[description={3-36}]{pyy}{PYY$_{3-36}$}{peptide YY$_{3-36}$}
+\newacronym[description={4-(methylnitrosamino)-1-(3-pyridyl)-1-butanone}]{nnk}{NNK}{4-(methylnitrosamino)-1-(3-pyridyl)-1-butanone}
+\newacronym[description={Activation-induced cytidine deaminase}]{aid}{AID}{activation-induced cytidine deaminase}
+\newacronym[description={Adenosine triphosphate}]{atp}{ATP}{adenosine triphosphate}
+\newacronym[description={Adrenocorticotropic hormone}]{acth}{ACTH}{adrenocorticotropic hormone}
+\newacronym[description={Agouti-related protein}]{agrp}{AGRP}{agouti-related protein}
+\newacronym[description={Analysis of variance}]{anova}{ANOVA}{analysis of variance}
+\newacronym[description={Apolipoprotein B mRNA editing enzyme, catalytic polypeptide-like}]{apobec}{APOBEC}{apolipoprotein B mRNA editing enzyme, catalytic polypeptide-like}
+\newacronym[description={B-cell lymphoma 2}]{bcl2}{Bcl-2}{B-cell lymphoma 2}
+\newacronym[description={Base-excision repair}]{ber}{BER}{base-excision repair}
+\newacronym[description={Bayesian Information Criterion}]{bic}{BIC}{Bayesian Information Criterion}
+\newacronym[description={Bcl-2 homology 3}]{bh3}{BH3}{Bcl-2 homology 3}
+\newacronym[description={Bladder urothelial carcinoma}]{blca}{BLCA}{bladder urothelial carcinoma}
+\newacronym[description={Body mass index}]{bmi}{BMI}{body mass index}
+\newacronym[description={Cervical squamous cell carcinoma and endocervical adenocarcinoma}]{cesc}{CESC}{cervical squamous cell carcinoma and endocervical adenocarcinoma}
+\newacronym[description={Cocaine and amphetamine related transcript}]{cart}{CART}{cocaine and amphetamine related transcript}
+\newacronym[description={Colon adenocarcinoma}]{coad}{COAD}{colon adenocarcinoma}
+\newacronym[description={Complimentary DNA}]{cdna}{cDNA}{complimentary DNA}
+\newacronym[description={Correlation adjusted mean rank gene set test}]{camera}{CAMERA}{correlation adjusted mean rank gene set test}
+\newacronym[description={Counts per million}]{cpm}{cpm}{counts per million}
+\newacronym[description={Cytochrome P450}]{cyp}{CYP}{cytochrome P450}
+\newacronym[description={Deoxyribonucleic acid}]{dna}{DNA}{deoxyribonucleic acid}
+\newacronym[description={Differentially expressed gene}]{deg}{DEG}{differentially expressed gene}
+\newacronym[description={Disability-Adjusted Life Year},longplural={Disability-Adjusted Life Years}]{daly}{DALY}{Disability-Adjusted Life Year}
+\newacronym[description={Electron transport chain}]{etc}{ETC}{electron transport chain}
+\newacronym[description={Endothelial-mesenchymal transition}]{emt}{EMT}{endothelial-mesenchymal transition}
+\newacronym[description={Epidermal growth factor receptor}]{egfr}{EGFR}{epidermal growth factor receptor}
+\newacronym[description={Estrogen receptor}]{er}{ER}{estrogen receptor}
+\newacronym[description={Extracellular signal regulated kinase}]{erk}{ERK}{extracellular signal regulated kinase}
+\newacronym[description={False discovery rate}]{fdr}{FDR}{false discovery rate}
+\newacronym[description={Family-wise error rate}]{fwer}{FWER}{family-wise error rate}
+\newacronym[description={Fat mass and obesity associated}]{fto}{FTO}{fat mass and obesity associated}
+\newacronym[description={Fold change}]{fc}{FC}{fold change}
+\newacronym[description={Functional class scoring}]{fcs}{FCS}{functional class scoring}
+\newacronym[description={G-protein coupled receptor}]{gpcr}{GPCR}{G-protein coupled receptor}
+\newacronym[description={Gap 1}]{g1}{G1}{Gap 1}
+\newacronym[description={Gene Expression Omnibus}]{geo}{GEO}{Gene Expression Omnibus}
+\newacronym[description={Gene Ontology}]{go}{GO}{Gene Ontology}
+\newacronym[description={Gene Set Enrichment Analysis}]{gsea}{GSEA}{Gene Set Enrichment Analysis}
+\newacronym[description={Genome wide association study},longplural={genome wide association studies},\glsshortpluralkey={GWAS}]{gwas}{GWAS}{genome wide association study}
+\newacronym[description={Growth hormone secretagogue receptor}]{ghsr}{GHSR}{growth hormone secretagogue receptor}
+\newacronym[description={Human epidermal growth factor receptor 2}]{her2}{HER2}{human epidermal growth factor receptor 2}
+\newacronym[description={Insulin receptor substrate}]{irs}{IRS}{insulin receptor substrate}
+\newacronym[description={Insulin-like growth factor binding protein}]{igfbp}{IGFBP}{insulin-like growth factor binding protein}
+\newacronym[description={Insulin-like growth factor}]{igf}{IGF}{insulin-like growth factor}
+\newacronym[description={Interferon-$\alpha$}]{ifna}{IFN$\alpha$}{interferon-$\alpha$}
+\newacronym[description={Interferon-$\gamma$}]{ifny}{IFN$\gamma$}{interferon-$\gamma$}
+\newacronym[description={Interleukin-10}]{il10}{IL-10}{interleukin-10}
+\newacronym[description={Interleukin-6}]{il6}{IL-6}{interleukin-6}
+\newacronym[description={International Cancer Genome Consortium}]{icgc}{ICGC}{International Cancer Genome Consortium}
+\newacronym[description={Janus kinase}]{jak}{JAK}{janus kinase}
+\newacronym[description={Kidney renal clear cell carcinoma}]{kirp}{KIRP}{kidney renal clear cell carcinoma}
+\newacronym[description={Kyoto Encyclopedia of Genes and Genomes}]{kegg}{KEGG}{Kyoto Encyclopedia of Genes and Genomes}
+\newacronym[description={Liver hepatocellular carcinoma}]{lihc}{LIHC}{liver hepatocellular carcinoma}
+\newacronym[description={Lymph node}]{ln}{LN}{lymph node}
+\newacronym[description={Mammalian target of rapamycin}]{mtor}{mTOR}{mammalian target of rapamycin}
+\newacronym[description={Mean}]{m}{$\mu$}{mean}
+\newacronym[description={Melanocortin 1 receptor}]{mc1r}{MC1R}{melanocortin 1 receptor}
+\newacronym[description={Melanocortin 4 receptor}]{mc4r}{MC4R}{melanocortin 4 receptor}
+\newacronym[description={Melanocyte-stimulating hormone}]{msh}{MSH}{melanocyte-stimulating hormone}
+\newacronym[description={Mesenchymal-endothelial transition}]{met}{MET}{mesenchymal-endothelial transition}
+\newacronym[description={Messenger RNA}]{mrna}{mRNA}{messenger RNA}
+\newacronym[description={Microarray analysis suite 5.0}]{mas}{MAS5}{microarray analysis suite 5.0}
+\newacronym[description={Mismatch repair}]{mmr}{MMR}{mismatch repair}
+\newacronym[description={Mismatch}]{mm}{MM}{mismatch}
+\newacronym[description={Neuropeptide Y1 receptor}]{nyr1}{Y1R}{neuropeptide Y1 receptor}
+\newacronym[description={Neuropeptide Y}]{npy}{NPY}{neuropeptide Y}
+\newacronym[description={New Zealand Breast Cancer}]{nzbc}{NZBC}{New Zealand Breast Cancer}
+\newacronym[description={New Zealand}]{nz}{NZ}{New Zealand}
+\newacronym[description={Next generation sequencing}]{ngs}{NGS}{next generation sequencing}
+\newacronym[description={Nuclear factor $\kappa$B}]{nfkb}{NF-$\kappa$B}{nuclear factor $\kappa$B}
+\newacronym[description={Nucleotide-excision repair}]{ner}{NER}{nucleotide-excision repair}
+\newacronym[description={Obesity policy action}]{opa}{OPA}{obesity policy action}
+\newacronym[description={Over representation analysis}]{ora}{ORA}{over representation analysis}
+\newacronym[description={Pattern recognition receptor}]{prr}{PRR}{pattern recognition receptor}
+\newacronym[description={Perfect match}]{pm}{PM}{perfect match}
+\newacronym[description={Phosphatidylinositol 3-kinase}]{pi3k}{PI3K}{phosphatidylinositol 3-kinase}
+\newacronym[description={Phosphoenol pyruvate}]{pep}{PEP}{phosphoenol pyruvate}
+\newacronym[description={Polycyclic aromatic hydrocarbon}]{pah}{PAH}{polycyclic aromatic hydrocarbon}
+\newacronym[description={Pro-opiomelanocortin}]{pomc}{POMC}{pro-opiomelanocortin}
+\newacronym[description={Progesterone receptor}]{pr}{PR}{progesterone receptor}
+\newacronym[description={Prohormone convertase 1}]{pc1}{PC1}{prohormone convertase 1}
+\newacronym[description={Reactive oxygen species}]{ros}{ROS}{reactive oxygen species}
+\newacronym[description={Rectum adenocarcinoma}]{read}{READ}{rectum adenocarcinoma}
+\newacronym[description={Retinoblastoma-associated}]{rb}{RB}{retinoblastoma-associated}
+\newacronym[description={Ribonucleic acid}]{rna}{RNA}{ribonucleic acid}
+\newacronym[description={Robust multi array}]{rma}{RMA}{robust multi array}
+\newacronym[description={Signal transducer and activator of transcription 3}]{stat3}{STAT3}{signal transducer and activator of transcription 3}
+\newacronym[description={Signal transducer and activator of transcription}]{stat}{STAT}{signal transducer and activator of transcription}
+\newacronym[description={Single nucleotide polymorphism}]{snp}{SNP}{single nucleotide polymorphism}
+\newacronym[description={Single-stranded DNA}]{ssdna}{ssDNA}{single-stranded DNA}
+\newacronym[description={Singular value decomposition}]{svd}{SVD}{singular value decomposition}
+\newacronym[description={Skin cutaneous melanoma}]{skcm}{SKCM}{skin cutaneous melanoma}
+\newacronym[description={Standard deviation}]{sd}{$\sigma$}{standard deviation}
+\newacronym[description={The Cancer Genome Atlas}]{tcga}{TCGA}{The Cancer Genome Atlas}
+\newacronym[description={Thrombospondin-1}]{tsp1}{TSP-1}{thrombospondin-1}
+\newacronym[description={Toll-like receptor}]{tlr}{TLR}{Toll-like receptor}
+\newacronym[description={Transforming growth factor-$\beta$}]{tgfb}{TGF$\beta$}{transforming growth factor-$\beta$}
+\newacronym[description={Tumour necrosis factor-$\alpha$}]{tnfa}{TNF$\alpha$}{tumour necrosis factor-$\alpha$}
+\newacronym[description={Ultraviolet}]{uv}{UV}{ultraviolet}
+\newacronym[description={United States of America}]{usa}{USA}{United States of America}
+\newacronym[description={Uterine corpus endometrial carcinoma}]{ucec}{UCEC}{uterine corpus endometrial carcinoma}
+\newacronym[description={Variance inflation factor}]{vif}{VIF}{variance inflation factor}
+\newacronym[description={Vascular endothelial growth factor-A}]{vegfa}{VEGF-A}{vascular endothelial growth factor-A}
+\newacronym[description={Vascular endothelial growth factor}]{vegf}{VEGF}{vascular endothelial growth factor}
+\newacronym[description={Wilcoxon-Mann-Whitney}]{wmw}{WMW}{Wilcoxon-Mann-Whitney}
+\newacronym[description={World Health Organization}]{who}{WHO}{World Health Organization}

### 2017-03-30 thesis/discussion/discussion.tex
Final commit before submission of this thesis

@@
 Initially, these results were thought to be due to the obesity associated genetic signature being specific only to the CR data set.
-Therefore, a different obesity associated genetic signature from the \citet{Fuentes-Mattei2014} study was used to see whether this signature was able to show significant association with the patient \gls{bmi} in the Cr, \gls{nzbc} and \gls{icgc} cancer data sets.
+Therefore, a different obesity associated genetic signature from the \citet{Fuentes-Mattei2014} study was used to see whether this signature was able to show significant association with the patient \gls{bmi} in the CR, \gls{nzbc} and \gls{icgc} cancer data sets.
 The results were similar to the obesity metagene from the \citet{Creighton2012} study; FM obesity metagene scores reflected the gene expressions of the signature but the scores were not significantly associated with the patient \gls{bmi} or \gls{bmi} status in any of the cancer data sets investigated (except the \gls{blca} data set).
-Together with the results from the CR obesity metagene, these results suggested that both CR and FM genetic signatures were only significant in the original data set in which the signature was derived in (though this was not confirmed in the FM data set, as no patient \gls{bmi} information was available).
+Together, these results suggested that both CR and FM genetic signatures were only significant in the original data set in which the signature was derived in (though this was not confirmed in the FM data set, as no patient \gls{bmi} information was available).
 
 The fact that both CR and FM obesity associated genetic signatures showed no sign of significant association with the patient \gls{bmi} in majority of the cancer data sets raised a question of whether these signatures were actually related to obesity, or a different clinical variable within the data set.
-This question led to the identification of the various versions of obesity associated genetic signatures in \cref{sub:identification_of_obesity_associated_genetic_signatures} that were controlled by clinical variables (sex, age, ethnicity, menopause status, tumour grade, \gls{er}/\gls{her2}/\gls{pr} statuses and \gls{ln} status) in the CR data set.
-All of the metagenes created from these clinical variable-controlled genetic signatures were consistent with the gene expression patterns, but none of the metagenes showed significant association with the patient \gls{bmi} and \gls{bmi} status in the data set other than the CR data set, with the exception of the overweight group in the \gls{blca} data set.
+This question led to the identification of various versions of obesity associated genetic signatures in \cref{sub:identification_of_obesity_associated_genetic_signatures}.
+The CR data set was controlled for all the clinical variables (sex, age, ethnicity, menopause status, tumour grade, \gls{er}/\gls{her2}/\gls{pr} statuses and \gls{ln} status) to generate some of these signatures.
+All of the metagenes created from these novel obesity associated genetic signatures were consistent with the gene expression patterns, but none of the metagenes showed significant association with the patient \gls{bmi} or \gls{bmi} status in the data set other than the CR data set, with the exception of the overweight group in the \gls{blca} data set.
 
@@
 Thus the tumour samples taken from obese and overweight patients will be a mixture of samples that are driven by obesity-related mechanisms and mechanisms that are not related to obesity.
-These different subgroups of tumours (obesity-driven or not) cannot be easily identified and isolated, which may affect the detection of the truly obesity associated genes, as some of the tumour samples from the patients will have nothing to do with obesity and cancer, and hence misclassified into wrong groups.
+These different subgroups of tumours (obesity-driven or not) cannot be easily identified and isolated.
+This may affect the detection of the truly obesity associated genes, as some of the tumour samples from the patients will have nothing to do with obesity and cancer, and hence misclassified into wrong groups.
 
@@
 
-Even if the patient \gls{bmi} and/or \gls{bmi} status was enough to identify the genes that were truly associated with obesity, it would not be a trivial task to identify the truly obesity associated genes from those that were not.
+Even if the patient \gls{bmi} and/or \gls{bmi} status was enough to identify the genes that were truly associated with obesity, it would not be a trivial task to identify the truly obesity associated genes from those that were not, given a list of \glspl{deg}.
 This raises an important question of whether the patient \gls{bmi} was an appropriate clinical variable to uncover the true association between obesity and cancer (more in \cref{sec:limitations,sec:future_directions}).
@@
 Since linear models were created only with the ``consistent'' pathway metagene scores, a step wise method was used to generate linear models that included the patient \gls{bmi}, \gls{bmi} status and all of the pathway metagenes that allowed the model to predict the true obesity metagene the best.
-As shown by the results in \cref{sec:prediction_of_obesity_associated_metagene_with_pathway_associate_metagene}, all of the models predicted the true obesity metagene scores with high $R^2$-value in the CR data set (highest $R^2$ $\approx$ 0.8869) and were statistically significant.
+As shown by the results in \cref{sec:prediction_of_obesity_associated_metagene_with_pathway_associate_metagene}, all of the models predicted the true obesity metagene scores with high $R^2$-value in the CR data set (highest $R^2$ $\approx$ 0.89) and were statistically significant.
 This meant that the variables that were significant in the linear models played a role in the obesity associated genetic signatures.
@@
 None of the pathway associated genetic signatures showed positive correlation with any of the obesity associated genetic signatures in the heatmap.
-However, strong negative correlation with the Akt, \gls{er}, \gls{egfr}, \gls{pr}, p53, Src and \gls{tgfb} pathways, suggesting that the obesity associated genetic signatures were picking up the signals from these pathways, rather than the signals from the obesity phenotype of the patients.
+However, all of the obesity metagenes derived from the Creighton \textit{et al.} showed strong negative correlation with the Akt, \gls{er}, \gls{egfr}, \gls{pr}, p53, Src and \gls{tgfb} pathways.
+This suggested that the obesity associated genetic signatures were picking up the signals from these pathways, rather than the signals from the obesity phenotype of the patients.
 

### 2017-03-30 thesis/misc/acknowledgement.tex
Final commit before submission of this thesis

@@
 
-First and foremost, I would like to thank my supervisor Mik, for giving me this project and an
-opportunity to learn and develop the bioinformatics skills required to complete it.
-You have been a great supervisor and I have learnt a lot from you over the past two years, and I have enjoyed being a part of the Black lab.
+First and foremost, I would like to thank my supervisor Mik, for giving me this project and the opportunity to learn and develop the bioinformatics skills required to complete it.
+You have been a great supervisor; I have learnt a lot from you over the past two years, and I have enjoyed being a part of the Black lab.
 \\
@@
 I would like to thank my committee members, Dr.\ Anita Dunbier, Dr.\ Miriam Sharpe, and Dr.\ Peter Mace (temporary member), for being patient during my presentations that were full of statistics and heatmaps.
-I am grateful for all the suggestions and inputs you have given me to make this a better project.
+I am grateful for all the suggestions and input you have given me to make this a better project.
 \\
@@
 I would like to thank all of my friends and family, for all their support and fun memories throughout the years that I have been at University.
-I wish I could name everyone, but I don't think it's a good idea to extend this thesis any longer\ldots{}
+I wish I could name everyone, but extending this thesis any longer is probably not a good idea\ldots{}
 Long story short, my life would not be nearly as interesting or enjoyable without you guys.

### 2017-03-30 thesis/results/results2.tex
Final commit before submission of this thesis

@@
 First, the direction of GT pathway associated genetic signatures were resolved, then the pathway associated metagenes were compared with the obesity associated metagenes, and lastly linear models were constructed based on the pathway metagenes and patient \gls{bmi} to predict the obesity associated metagenes.
+
+\vspace{-3mm}
 
@@
 
-\begin{figure}[htpb]
+\begin{figure}[htb]
 	\centering
@@
 Metagenes generated from the application of \gls{svd} and metagenes generated from the transformation matrix are exactly the same if the transformation matrix is derived from the same data set.
-For example, an obesity associated metagene generated from the CR data with \gls{svd} would have the same values as the metagenes generated in the CR data with the transformation matrix (where the transformation matrix was derived from CR data).
+For example, an obesity associated metagene generated from the CR data with \gls{svd} would have the same values as the metagenes generated in the CR data with the transformation matrix (since the transformation matrix was derived from the CR data).
 Furthermore, if the \gls{svd}-derived metagenes and transformation matrix-derived metagenes were the same (or at least similar) in a different data set, this suggests that the metagene-related biology contained in the two data sets is highly similar.
@@
 
-\begin{figure}[htpb]
+\begin{figure}[tb]
 	\centering
@@
 However, some pathway signatures were negatively correlated with the obesity associated genetic signatures, which meant that the obesity signatures were likely capturing the signals from these signatures  rather than the biological signals from the obesity phenotype of the patient.
+
+\vspace{-3mm}
 
@@
 ``Sustained cell proliferation'' was one of the cancer hallmarks that \citet{Fuentes-Mattei2014} had shown to be related to their obesity associated genetic signature, which could be the reason that the Myc pathway metagene showed up as a significant variable in the linear model.
-
 Since the \gls{pr} pathway metagene showed significant association in many of the linear models, linear models were made with each of the combinations of \gls{bmi}, \gls{bmi} status and \gls{pr} pathway metagene for all of the obesity associated metagenes (\cref{tab:lm_pr_only}; \cref{sec:summary_of_the_linear_models_in_nzbc_data}).

### 2017-08-22 marking_response/response.tex
Added marking_response to the thesis directory temporarily

%--------------------------------------------------------------------------------
%     File Name           :     response.tex
%     Created By          :     rikutakei
%     Creation Date       :     [2017-08-15 15:29]
%     Last Modified       :     [2017-08-22 17:44]
%     Description         :     Response to the markers' comments for my MSc thesis.
%--------------------------------------------------------------------------------

\documentclass[a4paper, 12pt]{article}

\usepackage{xcolor}

\begin{document}

\section*{Response to the comments made by Examiner ``1136''}
\label{sec:response_to_comments_from_examiner_1136}

\noindent
\textcolor{blue}{
	\textit{A reasonably comprehensive literature review was provided, however of concern, over 8.5 pages of the introduction is effectively a regurgitation of the `Hallmarks of cancer' review by Weinberg and Hanahan.
	This could be substantially cropped.
	}
}\\

\noindent
\\

\noindent
\textcolor{blue}{
	\textit{Given this thesis is focused on a known environmental influence on cancer (obesity), it was a surprise the section on environmental factors associated with cancer incidence (1.2.3) is underdeveloped and superficial.
	}
}\\

\noindent
I understand the Examiner's concern about the relatively brief description of the known environmental factors that affect cancer incidence.
As you may know, the focus of this project was the relationship between cancer and obesity, and how obesity affects cancer (which is covered more in depth in Section 1.3.2).

Section 1.2 was written as a section introducing cancer in general, and was not specifically focused on any single cancer type, nor environmental factor.
So it follows that Section 1.2.3 describes, in very general terms, how \textit{some} (not \textit{all}) environmental factors can influence the cancer incidence, and I think the amount of information placed in this subsection is sufficient for this purpose.
Additionally, the environmental factors chosen to be included in this section was based on either (1) (very) strong evidence of their role in cancer onset/progression in the literature (in terms of both epidemiological and biological evidence/association), or (2) some sort of relationship with obesity.

Perhaps I could emphasise these points within Section 1.2.3 and guide the readers to Section 1.3.2 for specific details on the effect of obesity on cancer.
\\

\noindent
\textcolor{blue}{
	\textit{In Section 1.2.4, Provide a citation for a recent list of molecular targets in cancer therapy, and their companion approved targeted agents (or present it in a table in appendix).
	This first paragraph is superficial and na\"ive.
	What is the rationale for only mentioning the RAS/RAF/MEK/ERK and PI3K/AKT pathways?
	Is it because these are the most relevant to those impacted by obesity?
	It would follow that those pathways be more in alignment with inflammatory signalling pathways rather than those associated with cell proliferation (as mentioned later in section 1.3.2).
	The rationale for cherry picking select cell signalling information for this section of the introduction is unclear.
	}
}\\

\noindent
\\

\noindent
\textcolor{blue}{
\textit{What is the rationale for selection of one TCGA cancer dataset over another?
Not all cancers are known to have obesity as a risk factor (as mentioned at the beginning of this report).
It would therefore follow that not all cancers would be expected to have a signature consistent with a high BMI score metagene.
A more targeted approach would perhaps have been better to rationalise which datasets to examine based on known risk of association with obesity, then look within any of those relevant cancer types for datasets for which height and weight data were available.
This may have aided the pathway enrichment analysis.
}
}\\

\noindent
The only reason why the eight TCGA data sets were chosen for the analysis was because these cancer types were the only data sets that had both height and weight information of the samples.
I agree that there may have been other cancer types more suitable than the eight that were chosen for this project, but those cancer types did not have the crucial clinical information (i.e. height or weight).
In addition, it was of interest in this project to identify novel genes/pathways that associated with obesity across multiple cancer types, and hence including as many cancer types was important to get as many DEGs that were potentially obesity-associated in the first instance.

But, as you mentioned in your comment, perhaps for the purpose of pathway enrichment analysis, using only the cancer types that were known to be associated with obesity may have been more useful than using all eight cancer types.
Since this analysis will take a considerable amount of time and addition of information to the thesis, I would like to dismiss this idea for the current project, but I will add this to the discussion.
\\

\noindent
\textcolor{blue}{
\textit{It was not clear why it was necessary to run the simulation of the number of differentially expressed genes between 2 or more ICGC datasets when the cell of origin of these datasets were so highly variable (p83).
}
}\\

\noindent
As mentioned above, it was of interest to find novel obesity-associated pathways or genes that were common across multiple cancer types, so DEG analysis was carried out on all available cancer types, regardless of the tissue of origin.
The purpose of the simulation was to make sure, statistically, that the number of DEGs identified between multiple cancer types were significant, and not by chance.
However, as you may know from the report, majority of these DEGs would have been false positives.

Again, using only the cancer types that were known to be associated with obesity may have aided the discovery of significant pathways and/or genes.
\\

\noindent
\textcolor{blue}{
\textit{On page 110, it is suggested that as the obesity-associated genetic signatures were derived from breast cancer datasets, that it made sense that these signatures were significant in the bladder cancer data sets as `basal-like' and `luminal-like' subtypes had been identified in bladder cancer.
As adenocarcinomas (epithelial-cell derived, and glandular in origin) all have these characteristics due to cell of origin context, I am not convinced by interpretations made in section 5.1.2, and feel the statements made are ungrounded.
}
}\\

\noindent
\\

\subsection*{Minor corrections}
\label{sub:minor_corrections}

\noindent
\textcolor{blue}{
\textit{Acronym list are not all acronyms -- some are definitions, or abbreviations.
}
}

\noindent
Changed the heading to ``Acronyms, Abbreviations, and Definitions'' from ``Acronyms''.
\\

\noindent
\textcolor{blue}{
\textit{Endothelial should be epithelial (p16)
}
}\\

\noindent
Changed `endothelial' to `epithelial'.
\\

\noindent
\textcolor{blue}{
\textit{``\ldots only logical to induce\ldots'' should be ``\ldots only logical to deduce\ldots'' (p16)
}
}\\

\noindent
Changed `induce' to `deduce'.
\\

\noindent
\textcolor{blue}{
\textit{Define NER (p20)
}
}\\

\noindent
NER is mentioned in the previous page (p19) as ``nucleotide-excision repair'' and also in the Acronyms list.
\\

\noindent
\textcolor{blue}{
\textit{Use full first and last names when referencing personal communication
}
}\\

\noindent
Changed ``Assoc. Prof. Mik Black'' to ``Assoc. Prof. Michael Black''.
\\

\noindent
\textcolor{blue}{
\textit{Throughout the thesis, the use of hyphens are inconsistent (e.g. obesity-associated, RMA-normalised, etc.)
}
}\\

\noindent
These errors have been corrected.
\\

\noindent
\textcolor{blue}{
\textit{``Hormone'' status is incorrect -- it should be ``receptor'' status (p73)
}
}\\

\noindent
Changed ``hormone'' to ``receptor''.
\\

\noindent
\textcolor{blue}{
\textit{``\ldots none of the metagenes were not generalisable in other cancer datasets.'' does not make grammatical sense (p83)
}
}\\

\noindent
Changed ``\ldots none of the metagenes were not generalisable in other cancer datasets'' to ``\ldots none of the metagenes were generalisable in other cancer datasets''
\\

\noindent
\textcolor{blue}{
\textit{The font used in the plots for figures 19-22 are exceedingly small (p100, 106)
}
}\\

\noindent
I agree with the Examiner that the fonts are indeed small, but due to the nature of the plot arrangement for these figures the fonts cannot be made any larger (at least not trivially).
Also, the font size was chosen so that the statistical values and letters did not obscure the points and boxes of the plots, and I would like to avoid this as much as possible, so I have left the font size unchanged.
\\

\newpage

\section*{Response to the comments made by Examiner ``1137''}
\label{sec:response_to_comments_from_examiner_1137}

\noindent
\textcolor{blue}{
	\textit{Individually published papers could have been critically appraised in the Introduction when it was introduced.
	}
}\\

\noindent
\\

\noindent
\textcolor{blue}{
	\textit{Potential bias and confounding in previously published studies due to the patient selection and the use of BMI as a marker of obesity are highly pertinent to thesis and could have been explored in more detail.
	}
}\\

\noindent
\\

\noindent
\textcolor{blue}{
	\textit{The effect of obesity on treatment response in cancer, and the influence of obesity on the likelihood of initial diagnosis, may also have implications for obesity-gene expression associations found in the literature.
	}
}\\

\noindent
\\

\noindent
\textcolor{blue}{
	\textit{Potential effects on gene expression in tumours by lifestyle factors associated with obesity, but not recorded in most data sets was not discussed in great detail, however this is relevant to the concepts explored later in Section 3.3.
	}
}\\

\noindent
\\

\noindent
\textcolor{blue}{
	\textit{It would be interesting to read here about the growth inhibiting mechanisms in cancer cells of anti-obesity drugs like orlistat.
	}
}\\

\noindent
\\

\noindent
\textcolor{blue}{
	\textit{The interesting literature from \textit{in vitro} studies about the effect on cancer cells of factors produced by adipocytes in culture was not explored.
	}
}\\

\noindent
\\

\noindent
\textcolor{blue}{
	\textit{Potential mechanisms of obesity-altering hormone receptor expression could have been explored in more depth \textit{at the level of cancer cells}.
	The effects of endocrine factors like leptin and the IGFs were discussed in general but not the cancer cell type-specific gene expression changes they cause.
	}
}\\

\noindent
\\

\noindent
\textcolor{blue}{
	\textit{A general in-depth exploration of the `meaning' of the different data types used in the thesis and how they are affected by normalisation and statistical test choice would have been appropriate in the Introduction.
	}
}\\

\noindent
\\

\noindent
\textcolor{blue}{
	\textit{How does the overlap between the genes that constitute each metagene influence the statistical analysis?
	}
}\\

\noindent
\\

\noindent
\textcolor{blue}{
	\textit{Biologically important multiple splice variants of many RNAs are available in both the Affymetrix microarray and TCGA/ICGC RNAseq data.
	Could the use of all probe sets/isoform variants have provided useful additional information?
	}
}\\

\noindent
\\

\noindent
\textcolor{blue}{
	\textit{Although reasoning behind exclusion of some patients due to missing clinical data is discussed, I am not completely clear about this after reading the description, and the numbers in Table 2 do not concord with what I know of these data sets from my own research.
	I am sure that patients will only have been excluded with good reason and it seems unlikely these exclusions will have introduced bias, however I would have liked a clearer summary of what patients had been excluded and why.
	Perhaps I have missed this?
	}
}\\

\noindent
From Table 2, the NZBC and the Gatza \textit{et al.} data sets were the only data set where the samples were excluded.
For NZBC data set, only the samples that had BMI or obesity status  information was included in the analysis (not stated in the thesis --  will be included after revision).
For Gatza \textit{et al.} data set, only the samples that were genotyped with Affymetrix HGU133A platform was used in the analysis, as all of the other three microarray data sets used this platform.

These information were stated in Section 2.2.1 under the corresponding data sets, but perhaps it might have been better if it was stated separately in a section of its own.
\\

\noindent
\textcolor{blue}{
	\textit{What basic quality control was possible on each data set?
	}
}\\

\noindent
\\

\noindent
\textcolor{blue}{
	\textit{Why was the BRCA RNAseq data set in TCGA not used?
	Surely this could have been informative?
	}
}\\

\noindent
I completely agree with the Examiner and the BRCA data set would have aided greatly in the exploration of obesity-associated signatures from both Creighton \textit{et al.} and Fuentes-Mattei \textit{et al.} data sets.
However, at the time of data retrieval (March 31st 2015), there were no height, weight or BMI information for the BRCA data set, and could not use the data set for the analyses.
\\

\noindent
\textcolor{blue}{
	\textit{On page 93 you say ``The fact that some pathway associated metagenes were consistent across different data sets suggested that some of these genetic signatures were reliable and did not depend on the data set the transformation matrices were derived from.''
	However, could the pathway expression signatures that were highly correlated in all of the  data sets instead reflect a technical issue, perhaps an artefact affecting a dominant probe set?
	}
}\\

\noindent
This is an interesting point.
All of the pathway-associated signatures were derived from the Gatza \textit{et al.} data set and was then transformed into other data sets.
For the pathway signatures to highly correlate between the data sets, all of the probe sets included in the pathway signature have to have similar expression in the other data sets as well.
With this in mind, I doubt that there was a similar technical error in all of these data sets, for a particular set of gene probe sets, given that each one of these data sets was from an independent study.
\\

\noindent
\textcolor{blue}{
	\textit{The justification presented about the appropriateness of using SVD and transformation matrices is convincing but I wonder about the effects of small numbers of probe sets/RNAs with specific biological connotations dominating some metagenes.
	}
}\\

\noindent
I agree that there will be some probe sets/RNAs that would have influenced the metagene scores more than the others, and the identification of these influencial probes/RNAs would be of great interest.
However, I feel like an investigation  of the most informative probes/RNAs for each of the obesity- or pathway-associated signatures (or any genetic signature in question) would require an effort of a whole new project.
\\

\noindent
\textcolor{blue}{
	\textit{It would be interesting to explore whether the differences in risk ratio between cancer types was due to technical issues, or represented real biology.
	}
}\\

\noindent
\\

\noindent
\textcolor{blue}{
	\textit{Cancer patients can have vast differences in BMI at different points in their treatment.
	Was the clinical timing of weight measures standardised in some way to reduce the impact of this problem, e.g. using weight at initial diagnosis?
	}
}\\

\noindent
This is a good point, and none of the papers define at which stage of the disease the patients' clinical data was taken (check).
\\

\noindent
\textcolor{blue}{
	\textit{In the Introduction, add a few sentences of additional explanation about potential bias and confounding introduced into previously published studies of obesity-cancer gene expression relationships, including potential confounding by lifestyle factors associated with obesity.
	}
}\\

\noindent
\\

\noindent
\textcolor{blue}{
	\textit{In the Introduction, refer to the published in vitro studies of the effect on cancer cells of factors produced by adipocytes in culture.
	}
}\\

\noindent
\\

\noindent
\textcolor{blue}{
	\textit{In the Introduction, add a few additional sentences to more fully introduce potential mechanisms at the level of cancer cells for obesity to alter biology through changes in hormone receptor expression and transcriptional effects of endocrine factors like leptin and the IGFs.
	}
}\\

\noindent
\\

\noindent
\textcolor{blue}{
	\textit{Add a table supplementing existing tables to more clearly show what patients from each data set were included, or excluded, and why.
	}
}\\

\noindent
This will be considered after discussion with Mik and HOD.
\\

\noindent
\textcolor{blue}{
	\textit{Add additional description about what basic quality control was possible for each data set (perhaps in section 2.3), and the results of the QC.
	The discussion in section 5.2.2 is hard to interpret without this.
	}
}\\

\noindent
\\

\noindent
\textcolor{blue}{
	\textit{Discuss whether the biology underpinning differences between cancer types in terms of obesity risk ratios in Table 1 could be explored statistically using genomic data.
	}
}\\

\noindent
\\

\noindent
\textcolor{blue}{
	\textit{When discussing future work, suggest specifically how the limitations of using BMI as an indicative biomarker could be addressed.
	What additional clinical data fields would be helpful to include as factors in the models?
	}
}\\

\noindent
I have added an extra sentence in the last paragraph addressing the concerns about the use of BMI as a measure to classify obesity.
\\

\end{document}

### 2017-08-22 thesis/discussion/discussion.tex
Added marking_response to the thesis directory temporarily

@@
 
-\subsection{Obesity associated genetic signatures}
+\subsection{Obesity-associated genetic signatures}
 \label{sub:obesity_associated_genetic_signatures}
 
-\cref{cha:obesity_genetic_signatures_and_cancer} focussed mainly on establishing the link between the obesity associated genetic signatures from the \citet{Creighton2012} and the \citet{Fuentes-Mattei2014} studies with the patient \gls{bmi} and \gls{bmi} status, as reported by their studies.
-The obesity associated genetic signature from the \citet{Creighton2012} study was examined first.
+\cref{cha:obesity_genetic_signatures_and_cancer} focussed mainly on establishing the link between the obesity-associated genetic signatures from the \citet{Creighton2012} and the \citet{Fuentes-Mattei2014} studies with the patient \gls{bmi} and \gls{bmi} status, as reported by their studies.
+The obesity-associated genetic signature from the \citet{Creighton2012} study was examined first.
 The obesity metagene generated from the Creighton \textit{et al.} obesity signature  was able to ``capture'' the overall genetic expression patterns of the samples, where low and high metagene scores  corresponded to low and high gene expressions, respectively.
@@
 
-Initially, these results were thought to be due to the obesity associated genetic signature being specific only to the CR data set.
-Therefore, a different obesity associated genetic signature from the \citet{Fuentes-Mattei2014} study was used to see whether this signature was able to show significant association with the patient \gls{bmi} in the CR, \gls{nzbc} and \gls{icgc} cancer data sets.
+Initially, these results were thought to be due to the obesity-associated genetic signature being specific only to the CR data set.
+Therefore, a different obesity-associated genetic signature from the \citet{Fuentes-Mattei2014} study was used to see whether this signature was able to show significant association with the patient \gls{bmi} in the CR, \gls{nzbc} and \gls{icgc} cancer data sets.
 The results were similar to the obesity metagene from the \citet{Creighton2012} study; FM obesity metagene scores reflected the gene expressions of the signature but the scores were not significantly associated with the patient \gls{bmi} or \gls{bmi} status in any of the cancer data sets investigated (except the \gls{blca} data set).
@@
 
-The fact that both CR and FM obesity associated genetic signatures showed no sign of significant association with the patient \gls{bmi} in majority of the cancer data sets raised a question of whether these signatures were actually related to obesity, or a different clinical variable within the data set.
-This question led to the identification of various versions of obesity associated genetic signatures in \cref{sub:identification_of_obesity_associated_genetic_signatures}.
+The fact that both CR and FM obesity-associated genetic signatures showed no sign of significant association with the patient \gls{bmi} in majority of the cancer data sets raised a question of whether these signatures were actually related to obesity, or a different clinical variable within the data set.
+This question led to the identification of various versions of obesity-associated genetic signatures in \cref{sub:identification_of_obesity_associated_genetic_signatures}.
 The CR data set was controlled for all the clinical variables (sex, age, ethnicity, menopause status, tumour grade, \gls{er}/\gls{her2}/\gls{pr} statuses and \gls{ln} status) to generate some of these signatures.
-All of the metagenes created from these novel obesity associated genetic signatures were consistent with the gene expression patterns, but none of the metagenes showed significant association with the patient \gls{bmi} or \gls{bmi} status in the data set other than the CR data set, with the exception of the overweight group in the \gls{blca} data set.
+All of the metagenes created from these novel obesity-associated genetic signatures were consistent with the gene expression patterns, but none of the metagenes showed significant association with the patient \gls{bmi} or \gls{bmi} status in the data set other than the CR data set, with the exception of the overweight group in the \gls{blca} data set.
 
@@
 These different subgroups of tumours (obesity-driven or not) cannot be easily identified and isolated.
-This may affect the detection of the truly obesity associated genes, as some of the tumour samples from the patients will have nothing to do with obesity and cancer, and hence misclassified into wrong groups.
+This may affect the detection of the truly obesity-associated genes, as some of the tumour samples from the patients will have nothing to do with obesity and cancer, and hence misclassified into wrong groups.
 
@@
 In fact, \citet{Damrauer2014} have identified ``basal-like'' and ``luminal''  subtypes in bladder cancer that resembled the corresponding subtypes in breast cancer, and suggested that bladder  and breast cancers may have common physiological properties.
-Since all of the obesity associated genetic signatures were derived from breast cancer data sets, it made sense that these signatures were significant in the \gls{blca} data set, as bladder cancer and breast cancer could have been biologically similar to one another.
+Since all of the obesity-associated genetic signatures were derived from breast cancer data sets, it made sense that these signatures were significant in the \gls{blca} data set, as bladder cancer and breast cancer could have been biologically similar to one another.
 However, the degree of similarity between these two cancer types are not yet known and further analyses are required for a definitive explanation of the results.
@@
 From all of these results so far, one may question how robust or appropriate the use of \gls{svd} and transformation matrices are to generate metagenes in other data sets.
-It appeared as though the CR data-derived obesity associated genetic signatures correlated significantly in the CR data, but as soon as these metagenes were generated in the other data sets via the application of transformation matrices, the correlation became insignificant.
+It appeared as though the CR data-derived obesity-associated genetic signatures correlated significantly in the CR data, but as soon as these metagenes were generated in the other data sets via the application of transformation matrices, the correlation became insignificant.
 One possible reason for this was that the obesity metagenes were not actually associated to the obesity phenotype in the other data sets.
@@
 
-The appropriateness of the use of \gls{svd} and transformation matrices were examined with different genetic signatures: \gls{er} and \gls{pr} pathway associated genetic signatures from the \citet{Gatza2010a} study (covered in \cref{sec:results_for_er_and_pr_metagene_analysis}).
+The appropriateness of the use of \gls{svd} and transformation matrices were examined with different genetic signatures: \gls{er} and \gls{pr} pathway-associated genetic signatures from the \citet{Gatza2010a} study (covered in \cref{sec:results_for_er_and_pr_metagene_analysis}).
 Three data sets (CR, FM and \gls{nzbc}) were used, as these data sets have both \gls{er} and \gls{pr} status information available for the patients.
-\gls{svd} was first applied to the \gls{rma} normalised CR data to obtain the \gls{er} and \gls{pr} metagenes, along with the transformation matrices for each of the genetic signatures.
+\gls{svd} was first applied to the \gls{rma}-normalised CR data to obtain the \gls{er} and \gls{pr} metagenes, along with the transformation matrices for each of the genetic signatures.
 Both the \gls{er} and \gls{pr} metagenes showed good concordance with the gene expression profiles of the genetic signature in the heatmap, and also showed significant association with patient \gls{er} and \gls{pr} statuses (\cref{sec:results_for_er_and_pr_metagene_analysis}).
@@
 
-\subsection{Common obesity associated genes and pathways across multiple cancer types}
+\subsection{Common obesity-associated genes and pathways across multiple cancer types}
 \label{sub:common_obesity_associated_genes_and_pathways_across_multiple_cancer_types}
 
-The results so far have provided strong evidence that the obesity associated genetic signatures from the CR and the FM data sets were not associated with patient \gls{bmi}.
-In a last attempt to identify any obesity associated genes that were common across multiple cancer types, gene expression analyses were carried out on the \gls{icgc} cancer data sets using samples for which patient \gls{bmi} information was available (\cref{sec:common_genes_across_multiple_cancer_types}).
+The results so far have provided strong evidence that the obesity-associated genetic signatures from the CR and the FM data sets were not associated with patient \gls{bmi}.
+In a last attempt to identify any obesity-associated genes that were common across multiple cancer types, gene expression analyses were carried out on the \gls{icgc} cancer data sets using samples for which patient \gls{bmi} information was available (\cref{sec:common_genes_across_multiple_cancer_types}).
 There were many \glspl{deg} identified in each of the \gls{icgc} cancer types, but there were no genes in common for all eight cancer types, and this observation was supported by the results from the simulation.
@@
 However, as observed by the results from the simulation, there were many genes identified by chance alone, which suggested that majority of the \glspl{deg} found in the eight cancer types were false positives.
-This apparently high level of \Glspl{type1} was not only observed in the \gls{icgc} data sets, but also in the CR data set when gene expression analyses were carried out to generate various obesity associated genetic signatures in \cref{sub:identification_of_obesity_associated_genetic_signatures} (discussed in more detail in \cref{sub:false_positives_in_gene_expression_analyses}).
+This apparently high level of \Glspl{type1} was not only observed in the \gls{icgc} data sets, but also in the CR data set when gene expression analyses were carried out to generate various obesity-associated genetic signatures in \cref{sub:identification_of_obesity_associated_genetic_signatures} (discussed in more detail in \cref{sub:false_positives_in_gene_expression_analyses}).
 
@@
 The fact that there were so many false positives in the \glspl{deg} from the \gls{icgc} data sets indicated that the use of the patient \gls{bmi} and \gls{bmi} status as the ``treatment'' conditions (in other words independent variable) for gene expression analysis was prone to \Glspl{type1}.
-This meant that there were very little evidence of true obesity associated genetic signatures in any of the data sets that have been explored in this project, or perhaps the patient \gls{bmi} and \gls{bmi} status was not enough to pinpoint the underlying biological relationship between obesity and cancer (see \cref{sub:discussion_definition_of_obesity} as well).
+This meant that there were very little evidence of true obesity-associated genetic signatures in any of the data sets that have been explored in this project, or perhaps the patient \gls{bmi} and \gls{bmi} status was not enough to pinpoint the underlying biological relationship between obesity and cancer (see \cref{sub:discussion_definition_of_obesity} as well).
 
-Evidence of large number of false positives was also seen in \cref{sub:identification_of_obesity_associated_genetic_signatures}, where there were many \glspl{deg} identified from the gene expression analyses when obesity associated genetic signatures were sought for in the CR data set.
+Evidence of large number of false positives was also seen in \cref{sub:identification_of_obesity_associated_genetic_signatures}, where there were many \glspl{deg} identified from the gene expression analyses when obesity-associated genetic signatures were sought for in the CR data set.
 This suggested that there was a significant number of genes that were not related to obesity (false positives) in the genetic signatures.
@@
 
-Even if the patient \gls{bmi} and/or \gls{bmi} status was enough to identify the genes that were truly associated with obesity, it would not be a trivial task to identify the truly obesity associated genes from those that were not, given a list of \glspl{deg}.
+Even if the patient \gls{bmi} and/or \gls{bmi} status was enough to identify the genes that were truly associated with obesity, it would not be a trivial task to identify the truly obesity-associated genes from those that were not, given a list of \glspl{deg}.
 This raises an important question of whether the patient \gls{bmi} was an appropriate clinical variable to uncover the true association between obesity and cancer (more in \cref{sec:limitations,sec:future_directions}).
@@
 In \cref{cha:obesity_associated_genetic_signature_and_pathway_signatures}, analyses were carried out to determine what the genetic pattern the obesity metagenes from \cref{cha:obesity_genetic_signatures_and_cancer} were associating with.
-The pathway associated genetic signatures from the \citet{Gatza2010a} study were used to establish the biological connection with the obesity associated genetic signatures.
-In order to make the obesity and pathway associated genetic signatures comparable with one another, the orientations of the metagenes were examined so that all of the metagenes were in the ``correct'' orientation (see \cref{sec:pathway_associated_genetic_signatures_from_gatza2010a_study}).
+The pathway-associated genetic signatures from the \citet{Gatza2010a} study were used to establish the biological connection with the obesity-associated genetic signatures.
+In order to make the obesity- and pathway-associated genetic signatures comparable with one another, the orientations of the metagenes were examined so that all of the metagenes were in the ``correct'' orientation (see \cref{sec:pathway_associated_genetic_signatures_from_gatza2010a_study}).
 
@@
 The clustering of the obesity metagene was not surprising, as the results from \cref{sub:_novel_obesity_associated_signatures_and_sample_bmi} already showed that the metagenes were highly correlated with one another.
-This result showed that all of the obesity metagenes created from the CR data associated with an unknown genetic signature that was not related to any of the pathway associated genetic signatures from the \citet{Gatza2010a} study.
+This result showed that all of the obesity metagenes created from the CR data associated with an unknown genetic signature that was not related to any of the pathway-associated genetic signatures from the \citet{Gatza2010a} study.
 
@@
 
-To further investigate whether there was any evidence of a pathway signature being related to the obesity associated genetic signatures, linear models were created in the \gls{nzbc} data set with the pathway metagene scores.
+To further investigate whether there was any evidence of a pathway signature being related to the obesity-associated genetic signatures, linear models were created in the \gls{nzbc} data set with the pathway metagene scores.
 First, linear models were created with the patient \gls{bmi}, \gls{bmi} status and a selection of the most ``consistent'' pathway metagene scores (based on the consistency across different data sets).
-From these linear models, \gls{pr} pathway metagene scores showed significant contribution to the model in predicting the obesity associated metagene scores in all of the linear models, with the exception of the FM obesity metagene (where Myc was the significant variable; \cref{sec:summary_of_the_linear_models_in_nzbc_data}).
+From these linear models, \gls{pr} pathway metagene scores showed significant contribution to the model in predicting the obesity-associated metagene scores in all of the linear models, with the exception of the FM obesity metagene (where Myc was the significant variable; \cref{sec:summary_of_the_linear_models_in_nzbc_data}).
 
@@
 However, the $R^2$-values for all of these predictions were low (highest $R^2$ $\approx$ 0.42; \cref{sec:summary_statistics_of_the_predicted_obesity_metagenes_with_sample_bmi_bmi_status_in_nzbc_and_cr_data}) with highly variable data points in the scatter plots.
-It was unclear from these predictions and plots whether the models explained the true underlying biological mechanisms of the obesity associated genetic signatures.
+It was unclear from these predictions and plots whether the models explained the true underlying biological mechanisms of the obesity-associated genetic signatures.
 Nevertheless, the models did explain some variations in the obesity metagenes, but there may be some other unknown biological processes that provide more information about the signatures.
@@
 As shown by the results in \cref{sec:prediction_of_obesity_associated_metagene_with_pathway_associate_metagene}, all of the models predicted the true obesity metagene scores with high $R^2$-value in the CR data set (highest $R^2$ $\approx$ 0.89) and were statistically significant.
-This meant that the variables that were significant in the linear models played a role in the obesity associated genetic signatures.
-There was no single pathway signature that was common to all of the linear models, but Akt, \gls{egfr}, \gls{tgfb} and Src pathway metagenes were included in many of the models, which suggested that these pathway signatures may have an important role in the obesity associated genetic signatures (\cref{sub:stepwise_linear_model_prediction_in_nzbc_and_cr_data_sets}).
+This meant that the variables that were significant in the linear models played a role in the obesity-associated genetic signatures.
+There was no single pathway signature that was common to all of the linear models, but Akt, \gls{egfr}, \gls{tgfb} and Src pathway metagenes were included in many of the models, which suggested that these pathway signatures may have an important role in the obesity-associated genetic signatures (\cref{sub:stepwise_linear_model_prediction_in_nzbc_and_cr_data_sets}).
 
-\subsection{Significance of the pathway associated signatures with the obesity associated signatures}
+\subsection{Significance of the pathway-associated signatures with the obesity-associated signatures}
 \label{sub:significance_of_pr_pathway}
@@
 But was this due to the true biological relationship between the \gls{pr} pathway and the obesity metagenes, or only because the models were created in the \gls{nzbc} data set?
-In fact, the proportion of the patients that were \gls{pr}$^+$ was greater in the \gls{nzbc} data set (62 \gls{pr}$^+$ and 37 \gls{pr}$^-$ patients) compared with the CR data set (48 \gls{pr}$^+$ and 50 \gls{pr}$^-$ patients), which implied that the patients in the \gls{nzbc} data set was more likely to show association with the \gls{pr} pathway metagene.
+In fact, the proportion of the patients that were \gls{pr}$^+$ was greater in the \gls{nzbc} data set (62 \gls{pr}$^+$ and 37 \gls{pr}$^-$ patients) compared with the CR data set (48 \gls{pr}$^+$ and 50 \gls{pr}$^-$ patients), which implied that the patients in the \gls{nzbc} data set were more likely to show association with the \gls{pr} pathway metagene.
 
@@
 In the step wise linear models, many variables were identified to be significantly associated with the obesity metagenes.
-This suggested that the pathway associated genetic signatures that were significant in these models may be biologically related to the obesity phenotype, or perhaps the obesity associated genetic signatures were detecting the signals from these pathways rather than the signals from the obesity phenotype.
+This suggested that the pathway-associated genetic signatures that were significant in these models may be biologically related to the obesity phenotype, or perhaps the obesity-associated genetic signatures were detecting the signals from these pathways rather than the signals from the obesity phenotype.
 In fact, many of the variables included in the linear models were pathway metagenes that negatively correlated with the obesity metagenes in \cref{fig:gatza_allmeta} (\cref{sub:comparison_of_the_obesity_and_pathway_associated_signatures}).
@@
 
-With that said, even if these pathway signatures were related to the underlying biology of breast cancer, the fact that these results depended on the pathway associated genetic signatures from the \citet{Gatza2010a} study makes the results questionable.
+With that said, even if these pathway signatures were related to the underlying biology of breast cancer, the fact that these results depended on the pathway-associated genetic signatures from the \citet{Gatza2010a} study makes the results questionable.
 This was due to the lack of consistency of the metagenes across different cancer data sets, where more than half of the pathway metagenes showed a correlation of less than 0.8 in at least one data set (\cref{sec:pathway_associated_metagenes_and_obesity_associated_metagenes}).
 This meant that at least half of the signatures were specific to the GT data set and could be highly variable in the other data sets, reflecting the potential differences in the cohorts (at the molecular level).
-Thus, without further validation of the quality of the pathway associated genetic signatures and verification of the roles of these signatures in the biological relationship between obesity and cancer, these results should be interpreted with care.
+Thus, without further validation of the quality of the pathway-associated genetic signatures and verification of the roles of these signatures in the biological relationship between obesity and cancer, these results should be interpreted with care.
 
@@
 Obesity is a term given to those individuals with excessive amount of fat and can be defined in various ways.
-It was important to define the obese phenotype accurately, as all of the obesity associated genetic signatures, gene expression analyses and pathway enrichment analysis were dependent on the patient \gls{bmi} status.
+It was important to define the obese phenotype accurately, as all of the obesity-associated genetic signatures, gene expression analyses and pathway enrichment analysis were dependent on the patient \gls{bmi} status.
 \Gls{bmi} is perhaps the most popular measurement to estimate obesity owing to the relative ease of measurement in clinical settings, and hence used in various clinical trials and large epidemiological studies as an indication of the patients' metabolic state.
@@
 
-Secondly, the method in which the pathway associated genetic signatures was defined in the \citet{Gatza2010a} study may have been questionable.
-In their studies, Gatza \textit{et al.} used model cell lines to identify the pathway associated genetic signatures by altering the expression of the representative gene for the pathway and assigned all of the genes affected by the change in representative gene expression to be associated with that pathway \citep{Bild2006,Gatza2010a}.
-However, some of these pathway associated genetic signatures were derived in different cell lines to one another, which meant that some pathway signatures may not reflect the true pathway activity \textit{in vivo}.
+Secondly, the method in which the pathway-associated genetic signatures was defined in the \citet{Gatza2010a} study may have been questionable.
+In their studies, Gatza \textit{et al.} used model cell lines to identify the pathway-associated genetic signatures by altering the expression of the representative gene for the pathway and assigned all of the genes affected by the change in representative gene expression to be associated with that pathway \citep{Bild2006,Gatza2010a}.
+However, some of these pathway-associated genetic signatures were derived in different cell lines to one another, which meant that some pathway signatures may not reflect the true pathway activity \textit{in vivo}.
 In fact, as seen in \cref{sec:pathway_associated_genetic_signatures_from_gatza2010a_study,sec:pathway_associated_metagenes_and_obesity_associated_metagenes}, some pathway signatures produced more consistent metagenes from the data sets than the other signatures.
@@
 These results have suggested that there were no obesity specific genetic signatures that were common across different cancer types.
-The results from \cref{cha:obesity_associated_genetic_signature_and_pathway_signatures} showed that the obesity associated genetic signatures generated from the Creighton \textit{et al.} data set were different from those generated from the Fuentes-Mattei \textit{et al.} data set.
-None of the pathway associated genetic signatures showed positive correlation with any of the obesity associated genetic signatures in the heatmap.
+The results from \cref{cha:obesity_associated_genetic_signature_and_pathway_signatures} showed that the obesity-associated genetic signatures generated from the Creighton \textit{et al.} data set were different from those generated from the Fuentes-Mattei \textit{et al.} data set.
+None of the pathway-associated genetic signatures showed positive correlation with any of the obesity-associated genetic signatures in the heatmap.
 However, all of the obesity metagenes derived from the Creighton \textit{et al.} showed strong negative correlation with the Akt, \gls{er}, \gls{egfr}, \gls{pr}, p53, Src and \gls{tgfb} pathways.
-This suggested that the obesity associated genetic signatures were picking up the signals from these pathways, rather than the signals from the obesity phenotype of the patients.
+This suggested that the obesity-associated genetic signatures were picking up the signals from these pathways, rather than the signals from the obesity phenotype of the patients.
 
@@
 Since Akt, \gls{egfr}, \gls{tgfb} and Src pathways were significant in the linear models that predicted the obesity metagenes, more thorough investigation of these pathways may be a good start to clarify the relationship.
-Further analyses of \gls{blca} may help clarify the similarity between bladder and breast cancer types and provide insight into the association between the obesity associated genetic signatures and \gls{blca}.
+Further analyses of \gls{blca} may help clarify the similarity between bladder and breast cancer types and provide insight into the association between the obesity-associated genetic signatures and \gls{blca}.
 
-In this project, the lack of association of the obesity associated genetic signatures with any of the cancer types could have been due to the quality of both the raw data sets and the genetic signatures used in this project.
+In this project, the lack of association of the obesity-associated genetic signatures with any of the cancer types could have been due to the quality of both the raw data sets and the genetic signatures used in this project.
+Furthermore, the use of \gls{bmi} may have been insufficient to describe the patients' obesity status, and would likely require more informative clinical measurements, such as waist circumference, to aid the classification.
 Therefore, careful selection of the raw data sets with sufficient, and if possible, additional information (such as waist circumference) about the patients and well-validated genetic signatures will be essential for future investigations.

### 2017-08-22 thesis/intro/intro.tex
Added marking_response to the thesis directory temporarily

@@
 The number of overweight and obese people, in both adults and children, has risen in every country of the world and the trend will likely to continue in the future.
-It is estimated to account for 3.4 million deaths per year and 93.6 million \glspl{daly} in 2010, obesity is a serious disease that continues to grow in our society \citep{Lim2012}.
+Estimated to account for 3.4 million deaths per year and 93.6 million \glspl{daly} in 2010, obesity is a serious disease that continues to grow in our society \citep{Lim2012}.
 
@@
 \noindent
-Approximately 1 in every 14 people in the world were classed as obese in 2014; therefore, 1 in every 14 people around the world were also at severe health risks associated with obese people.
+Approximately 1 in every 14 people in the world were classed as obese in 2014; therefore, 1 in every 14 people around the world were also at severe health risks associated with obesity.
 The highest prevalence of overweight and obesity was in the American regions, where 61\% and 27\% of the population were overweight and obese respectively \citep{WHO2014}.
@@
 Insulin, perhaps the most well-known hormone for its role in glucose homeostasis, also has an important role in satiety and food intake regulation.
-In brief terms, when insulin binds to the insulin receptor, \gls{irs} protein family is recruited and phosphorylated by the  intrinsic tyrosine kinase activity of the insulin receptor \citep{Saltiel2002}.
+In brief terms, when insulin binds to the insulin receptor, \gls{irs} protein family members are recruited and phosphorylated by the  intrinsic tyrosine kinase activity of the insulin receptor \citep{Saltiel2002}.
 The phosphorylated \gls{irs} protein then activates \gls{pi3k}, which in turn activates its downstream targets, causing a rapid signal cascade through the cell \citep{Saltiel2002}.
@@
 Extensive research in cancer biology has uncovered many pathways and regulatory mechanisms involved in cancer.
-With more than 100 distinct types and subtypes of cancers (each with its own distinct development mechanisms), it became difficult to focus on the overarching mechanisms of cancer development \citep{Hanahan2000}.
+With more than 100 distinct types and subtypes of cancers (each with its own distinct developmental mechanisms), it became difficult to focus on the overarching mechanisms of cancer development \citep{Hanahan2000}.
 To resolve this, \citet{Hanahan2000} proposed ``six essential alterations'' that cancer cells usually take advantage of during development and progression.
@@
 One of the prime reasons that cancers are so difficult to treat and have a high mortality rate is due to their ability to metastasise to other parts of the body.
-Once metastasised to different parts of the body, the clinician not only has to consider the primary tumour site, but also the second, third, or even fourth tumour sites.
+Once metastasised, the clinician not only has to consider the primary tumour site, but also the second, third, or even fourth tumour sites.
 These tumours may or may not have similar tumour biology as the primary tumour, and thus has to consider multiple treatment plans to cure the patient.
@@
 Many of the hallmarks are acquired by the cancer cells during tumour growth, through many genomic changes and mutations of the essential genes that allow cancer cells to survive, proliferate, invade and metastasise.
-Thus, it is only logical to induce that many, if not all, cancers are more prone to genetic mutations and restructuring, which allows the cancers to subsequently acquire the already described Hallmarks of Cancer \citep{Hanahan2011}.
+Thus, it is only logical to deduce that many, if not all, cancers are more prone to genetic mutations and restructuring, which allows the cancers to subsequently acquire the already described Hallmarks of Cancer \citep{Hanahan2011}.
 
@@
 \noindent
-While inflammation and immune response may assist in the control of diseases and injuries, they have been implicated in the progression of tumour and their acquisition of many of the cancer hallmarks \citep{Hanahan2011}.
+While inflammation and immune response may assist in the control of diseases and injuries, they have been implicated in the progression of tumours and their acquisition of many of the cancer hallmarks \citep{Hanahan2011}.
 There are two key transcription factors of inflammation that ultimately leads to the promotion  of many, if not all, of the hallmarks discussed so far: \gls{nfkb} and \gls{stat3} \citep{Mantovani2008}.
@@
 
-\subsection{Obesity associated genetic signatures}
+\subsection{Obesity-associated genetic signatures}
 \label{sub:obesity_associated_genetic_signatures}
@@
 
-\subsection{Pathway associated genetic signatures}
+\subsection{Pathway-associated genetic signatures}
 \label{sub:pathway_associated_genetic_signatures}
 
-Like obesity associated genetic signatures, there has also been a focus on the identification of pathways and genes that affect tumour growth.
+Like obesity-associated genetic signatures, there has also been a focus on the identification of pathways and genes that affect tumour growth.
 In 2006, Bild \textit{et al.} demonstrated that a collection of genes that are involved in a specific pathway could be used to determine the status of those pathways in tumour samples \citep{Bild2006}.
@@
 
-One thing to note here is that most of these pathway associated genetic signatures are involved in one or more of the Hallmarks of Cancer (\cref{subsubsec:cancerhallmarks}).
-This means that these pathway associated signatures can be used to provide insight into the underlying biological mechanisms of the samples used in this project and help understand the underlying biology of cancers in obese patients.
+One thing to note here is that most of these pathway-associated genetic signatures are involved in one or more of the Hallmarks of Cancer (\cref{subsubsec:cancerhallmarks}).
+This means that these pathway-associated signatures can be used to provide insight into the underlying biological mechanisms of the samples used in this project and help understand the underlying biology of cancers in obese patients.
 
@@
 The results from these analyses are split into two parts, each presented in a separate chapter.
-The first part of the results clarify the relationship between the obesity associated genetic signatures with various cancer types.
-The second part of the results consist of the links between the pathway associated genetic signatures and various cancer types, as well as the previously identified obesity associated signatures.
+The first part of the results clarify the relationship between the obesity-associated genetic signatures with various cancer types.
+The second part of the results consist of the links between the pathway-associated genetic signatures and various cancer types, as well as the previously identified obesity-associated signatures.
 Lastly, the results will be summarised and conclusions will be drawn from these results.

### 2017-08-22 thesis/methods/methods.tex
Added marking_response to the thesis directory temporarily

@@
 Clinical data of the samples (age, ethnicity, tumour grade, menopause status, \gls{bmi}, \gls{er} status, \gls{pr} status, \gls{her2} status, and \gls{ln} status) were obtained from the Supplementary Table 1 from the \citet{Creighton2012} study.
-799 obesity associated gene probes identified in the \citet{Creighton2012} study were obtained from the Supplementary Data File 1 from the \citet{Creighton2012} study.
+799 obesity-associated gene probes identified in the \citet{Creighton2012} study were obtained from the Supplementary Data File 1 from the \citet{Creighton2012} study.
 
@@
 Patient height, weight and \gls{bmi} information were not available in this data set.
-130 obesity associated gene probes identified by \citet{Fuentes-Mattei2014} were taken from the Supplementary Table 3 from their paper.
+130 obesity-associated gene probes identified by \citet{Fuentes-Mattei2014} were taken from the Supplementary Table 3 from their paper.
 
@@
 
-To provide an insight into the relevance of the obesity associated genetic signatures in \gls{nz} patient cohort, the \gls{nzbc} data set used by \citet{Print2016} were used.
+To provide an insight into the relevance of the obesity-associated genetic signatures in \gls{nz} patient cohort, the \gls{nzbc} data set used by \citet{Print2016} were used.
 The raw Affymetrix HGU133A microarray gene expression data files from the \citet{Print2016} study were downloaded from the \gls{geo} database (\gls{geo} accession ID: GSE3\-6771).
@@
 Clinical data for the samples in \citet{Gatza2010a} study was not available, as these samples were a combination of many different datasets.
-The Akt, \gls{bcat}, E2F1, \gls{egfr}, \gls{er}, \gls{her2}, \gls{ifna}, \gls{ifny}, Myc, p53, p63, \gls{pi3k}, \gls{pr}, Ras, \gls{stat3}, Src, \gls{tgfb}, and \gls{tnfa} pathway associated genetic signatures were used by Gatza \textit{et al.}, and the gene probes of all the pathway associated genetic signatures were obtained from the Supplemental Table 10 from their paper.
+The Akt, \gls{bcat}, E2F1, \gls{egfr}, \gls{er}, \gls{her2}, \gls{ifna}, \gls{ifny}, Myc, p53, p63, \gls{pi3k}, \gls{pr}, Ras, \gls{stat3}, Src, \gls{tgfb}, and \gls{tnfa} pathway-associated genetic signatures were used by Gatza \textit{et al.}, and the gene probes of all the pathway-associated genetic signatures were obtained from the Supplemental Table 10 from their paper.
 
@@
 For the gene symbols that had multiple probe set entries, a single probe set was chosen to represent the gene symbol by using the \texttt{collapseRows} function in the \textit{WGCNA} package in \gls{r}, using the default parameters (which chose the probe set with a maximum mean expression value) \citep{Langfelder2008}.
-Likewise, any obesity associated or pathway associated gene probes were converted into gene symbols.
+Likewise, any obesity-associated or pathway-associated gene probes were converted into gene symbols.
 
@@
 			\hline
+			\hline
 			\rule{0pt}{2.25ex}Underweight & \textless{} 20.0\\
@@
 
-\Gls{mas} and \gls{rma} normalisation methods are both part of the \textit{affy} package in \gls{r} \citep{Gautier2004}.
+\Gls{mas} and \gls{rma}-normalisation methods are both part of the \textit{affy} package in \gls{r} \citep{Gautier2004}.
 \Gls{mas} uses a difference-based method where it subtracts a value derived from \gls{mm} from \gls{pm}, but this approach may introduce additional noise and/or errors in some cases \citep{Irizarry2003}.
@@
 The raw microarray data were normalised using both \gls{rma} and \gls{mas} methods, each done separately on a copy of the data.
-The reason why the \gls{mas} method was used as well as the \gls{rma} method was because the \gls{mas} normalisation method was used in the \citet{Gatza2010a} study.
+The reason why the \gls{mas} method was used as well as the \gls{rma} method was because the \gls{mas}-normalisation method was used in the \citet{Gatza2010a} study.
 Though the normalisation methods were not specified in the \citet{Creighton2012} and \citet{Fuentes-Mattei2014} studies, to allow for data comparison between different datasets and accurate result validation of some of the studies, both \gls{rma} and \gls{mas} methods were considered for normalising the microarray data.
@@
 
-With respect to this project, \gls{svd} was used to assess whether the metagenes created from various obesity or pathway associated genes were significantly associated with certain clinical variables, such as \gls{bmi}.
-To do this, the normalised original expression data was reduced in size so that it only included the genes of interest (either obesity or pathway associated) and the samples, then \texttt{svd} function from the \textit{base} package in \gls{r} was used to apply \gls{svd} to the matrix.
+With respect to this project, \gls{svd} was used to assess whether the metagenes created from various obesity or pathway-associated genes were significantly associated with certain clinical variables, such as \gls{bmi}.
+To do this, the normalised original expression data was reduced in size so that it only included the genes of interest (either obesity or pathway-associated) and the samples, then \texttt{svd} function from the \textit{base} package in \gls{r} was used to apply \gls{svd} to the matrix.
 The first principal component was taken from the $V$ matrix and was used as the metagene scores for each samples in the data set.
@@
 
-In this project, two types of metagenes were created: obesity associated and pathway associated metagenes.
-For obesity associated metagenes, the direction of the metagenes were checked with the patient obesity status and \gls{bmi} value from the data in which the metagene was derived from.
-When the data set did not have \gls{bmi} or \gls{bmi} status information for the patients (for example the Gatza \textit{et al.} data set), a set of genes that were common in all obesity associated genetic signatures were taken and the expression of these common genes were compared with the metagene scores.
-The genes that are common to all of the obesity associated genetic signatures will be related to the metagenes produced by these signatures, and therefore the expressions of these genes should be in line with the metagene scores.
+In this project, two types of metagenes were created: obesity-associated and pathway-associated metagenes.
+For obesity-associated metagenes, the direction of the metagenes were checked with the patient obesity status and \gls{bmi} value from the data in which the metagene was derived from.
+When the data set did not have \gls{bmi} or \gls{bmi} status information for the patients (for example the Gatza \textit{et al.} data set), a set of genes that were common in all obesity-associated genetic signatures were taken and the expression of these common genes were compared with the metagene scores.
+The genes that are common to all of the obesity-associated genetic signatures will be related to the metagenes produced by these signatures, and therefore the expressions of these genes should be in line with the metagene scores.
 
-Checking the direction of the pathway metagenes from \citet{Gatza2010a} study is not as simple as the obesity associated metagenes, as the experimental design that identify which samples were treated to generate the pathway signature was not available.
+Checking the direction of the pathway metagenes from \citet{Gatza2010a} study is not as simple as the obesity-associated metagenes, as the experimental design that identify which samples were treated to generate the pathway signature was not available.
 Due to this, it was not possible to validate whether the generated pathway metagene was in the ``correct'' direction.
@@
 Since the \gls{nzbc} data was the only data set (other than the Creighton \textit{et al.} data set) that had the patient \gls{bmi} information, all of the linear models were constructed in the \gls{nzbc} data set.
-Eleven linear models were created in the \gls{nzbc} to predict the obesity associated metagene from the \citet{Creighton2012} study.
-First seven models created were combinations of the sample \gls{bmi}, \gls{bmi} status and a selection of pathway associated metagene scores.
+Eleven linear models were created in the \gls{nzbc} to predict the obesity-associated metagene from the \citet{Creighton2012} study.
+First seven models created were combinations of the sample \gls{bmi}, \gls{bmi} status and a selection of pathway-associated metagene scores.
 The selected pathways used in the linear model construction were \gls{bcat}, \gls{er}, \gls{ifna}, \gls{ifny}, Myc and \gls{pr}.
@@
 The last four linear models were constructed with the different combinations of the sample \gls{bmi}, \gls{bmi} status and the \gls{pr} pathway metagene.
-The reason that the \gls{pr} pathway metagene was used separately was because in the initial analysis, the \gls{pr} pathway metagene was the only pathway associated genetic signature that came up as significant in the other linear models (\cref{sec:prediction_of_obesity_associated_metagene_with_pathway_associate_metagene}).
+The reason that the \gls{pr} pathway metagene was used separately was because in the initial analysis, the \gls{pr} pathway metagene was the only pathway-associated genetic signature that came up as significant in the other linear models (\cref{sec:prediction_of_obesity_associated_metagene_with_pathway_associate_metagene}).
 All of the created linear models are summarised in \cref{tab:lm_model_method}.
@@
 
-In addition to these eleven linear models, a linear model was created with the sample \gls{bmi}, \gls{bmi} status and all of the pathway associated metagenes (Akt, \gls{bcat}, E2F1, \gls{egfr}, \gls{er}, \gls{her2}, \gls{ifna}, \gls{ifny}, Myc, p53, p63, \gls{pi3k}, \gls{pr}, Ras, \gls{stat3}, Src, \gls{tgfb} and \gls{tnfa}) using a stepwise approach.
+In addition to these eleven linear models, a linear model was created with the sample \gls{bmi}, \gls{bmi} status and all of the pathway-associated metagenes (Akt, \gls{bcat}, E2F1, \gls{egfr}, \gls{er}, \gls{her2}, \gls{ifna}, \gls{ifny}, Myc, p53, p63, \gls{pi3k}, \gls{pr}, Ras, \gls{stat3}, Src, \gls{tgfb} and \gls{tnfa}) using a stepwise approach.
 The stepwise approach allows the construction of a linear model that best fits the data, given a set of variables.
@@
 
-To summarise the number of \glspl{deg} from the gene expression analyses of the Creighton \textit{et al.} data that overlapped with the original Creighton \textit{et al.} obesity associated genetic signature, Venn diagrams were used.
+To summarise the number of \glspl{deg} from the gene expression analyses of the Creighton \textit{et al.} data that overlapped with the original Creighton \textit{et al.} obesity-associated genetic signature, Venn diagrams were used.
 Venn diagrams were plotted using the \texttt{venn} function from the \textit{gplots} package in \gls{r} \citep{gplots}.

### 2017-08-22 thesis/misc/acknowledgement.tex
Added marking_response to the thesis directory temporarily

@@
 I would like to thank my committee members, Dr.\ Anita Dunbier, Dr.\ Miriam Sharpe, and Dr.\ Peter Mace (temporary member), for being patient during my presentations that were full of statistics and heatmaps.
-I am grateful for all the suggestions and input you have given me to make this a better project.
+I am grateful for all the suggestions and inputs you have given me to make this a better project.
 \\
@@
 \noindent
-Lastly, a special thanks to Gezza and Tessa who generously proof-read my introduction.
+Lastly, a special thanks to Gerry and Tessa who generously proof-read my introduction.
 \\

### 2017-08-22 thesis/misc/appendix.tex
Added marking_response to the thesis directory temporarily

@@
 
-	\cref{fig:appendix/fm_ob_meta_heatmap_icgc} shows the association of the FM obesity metagene with the gene expression pattern of the FM obesity associated genetic signature in the other \gls{icgc} cancer data sets.
+	\cref{fig:appendix/fm_ob_meta_heatmap_icgc} shows the association of the FM obesity metagene with the gene expression pattern of the FM obesity-associated genetic signature in the other \gls{icgc} cancer data sets.
 	\cref{fig:appendix/fm_ob_meta_box_scatter_icgc} shows the association of the FM obesity metagene scores with the patient \gls{bmi} and \gls{bmi} status.
@@
 	The direction of the obesity metagenes identified from the CR data were examined and corrected for (\cref{fig:crob_meta_direction}).
-	Gene probes that were common to all obesity associated genetic signatures were used to examine whether the metagene was in the correct direction, relative to the other obesity metagenes
+	Gene probes that were common to all obesity-associated genetic signatures were used to examine whether the metagene was in the correct direction, relative to the other obesity metagenes
 
@@
 		\includegraphics[width=0.32\linewidth,page=24]{appendix/cr_meta_direction_raw}\\
-		\caption[Directionality of the obesity metagenes in the CR data]{Heatmaps showing the gene expression of the obesity associated genetic signatures with the obesity metagene in \gls{rma}-normalised CR data set.
-		The gene probes that were common to all of the obesity associated genetic signatures were used to determine the direction of the obesity metagenes.
+		\caption[Directionality of the obesity metagenes in the CR data]{Heatmaps showing the gene expression of the obesity-associated genetic signatures with the obesity metagene in \gls{rma}-normalised CR data set.
+		The gene probes that were common to all of the obesity-associated genetic signatures were used to determine the direction of the obesity metagenes.
 		The scales and axes are the same as the previous heatmaps.
@@
 
-	As shown in \cref{fig:appendix/cr_icgc_heatmap}, the original obesity metagene from the Creighton \textit{et al.} study was reflective of the sample gene expression levels of the obesity associated genetic signature in all of the other \gls{icgc} cancer data sets.
+	As shown in \cref{fig:appendix/cr_icgc_heatmap}, the original obesity metagene from the Creighton \textit{et al.} study was reflective of the sample gene expression levels of the obesity-associated genetic signature in all of the other \gls{icgc} cancer data sets.
 	However, none of the \gls{icgc} cancer data sets (other than the \gls{blca} data set; \cref{fig:crmetaicgc} in \cref{sec:creighton_obesity_metagene}) showed any significant association of the sample \gls{bmi} or \gls{bmi} status with the Or obesity metagene (\cref{fig:appendix/cr_icgc_box_scatter}).
@@
 		\caption[Obesity metagene from the \citet{Creighton2012} study and the sample gene expressions in the other \gls{icgc} cancer types]{Heatmaps showing the association of obesity metagene from the \citet{Creighton2012} study with sample gene expression from the other \gls{icgc} data sets.
-		Level of expression is represented in the top right histogram, where low and high gene expression were colour-coded with blue and red, respectively.
-		Each row of the heatmap represents a gene from the obesity associated genetic signature, and each column of the heatmap represents a sample from the \gls{icgc} data.
-		The obesity associated metagene scores of the samples are shown in a separate row at top of the heatmap, and the tree diagram of the heirarchical clustering of the genes is shown to the left of the heatmap.
+		Level of expression is represented in the top left histogram, where low and high gene expression were colour-coded with blue and red, respectively.
+		Each row of the heatmap represents a gene from the obesity-associated genetic signature, and each column of the heatmap represents a sample from the \gls{icgc} data.
+		The obesity-associated metagene scores of the samples are shown in a separate row at top of the heatmap, and the tree diagram of the heirarchical clustering of the genes is shown to the left of the heatmap.
 }
@@
 
-	\section{Obesity associated genetic signature using sample \gls{bmi} values}
+	\section{Obesity-associated genetic signature using sample \gls{bmi} values}
 	\label{sec:obesity_associated_genetic_signature_using_sample_bmi_values}
 
-	Since all of these obesity associated genetic signatures were derived based on the discrete values of sample \gls{bmi} staus, a genetic signature was identified using the sample \gls{bmi} value (continuous variable).
+	Since all of these obesity-associated genetic signatures were derived based on the discrete values of sample \gls{bmi} staus, a genetic signature was identified using the sample \gls{bmi} value (continuous variable).
 	Gene probe expressions were correlated with the sample \gls{bmi} values and was corrected for multiple hypothesis testing with \gls{fdr}.
@@
 	With that said, the top 799 gene probes that correlated the most with the sample \gls{bmi} were examined to see if there was any association.
-	The results were similar to the other obesity associated genetic signatures: the continuous \gls{bmi} signature significantly associated with sample \gls{bmi} and \gls{bmi} status in the CR data, but did not show any significance in any of the other cancer data.
+	The results were similar to the other obesity-associated genetic signatures: the continuous \gls{bmi} signature significantly associated with sample \gls{bmi} and \gls{bmi} status in the CR data, but did not show any significance in any of the other cancer data.
 
@@
 		\centering
-		\caption{Significantly enriched \gls{go} pathways in \gls{icgc} cancer data}
+		\caption[Significantly enriched \gls{go} pathways in \gls{icgc} cancer data]{Significantly enriched \gls{go} pathways in \gls{icgc} cancer data}
 		\label{tab:sig_path_go}\\
 		Cancer type & Pathway\\
-		\hline
-		\hline
 		\endfirsthead
@@
 		\endhead
+		\hline
+		\hline
 		\rule{0pt}{2.25ex}\gls{blca} & Nucleobase-containing small molecule catabolic process                                          \\
@@
 
-To determine whether the results from all of the obesity metagene analyses were not showing any association with patient \gls{bmi} because of the \gls{svd}/transformation matrix approach used, the same approach was used with \gls{er} and \gls{pr} pathway associated genetic signatures from the \citet{Gatza2010a} study.
+To determine whether the results from all of the obesity metagene analyses were not showing any association with patient \gls{bmi} because of the \gls{svd}/transformation matrix approach used, the same approach was used with \gls{er} and \gls{pr} pathway-associated genetic signatures from the \citet{Gatza2010a} study.
 Firstly, both the \gls{er} and \gls{pr} metagenes were generated in the \gls{rma} normalised CR data set.
-The generated metagenes were compared with the gene expressions of the corresponding pathway associated genetic signatures, and the association with the patient \gls{er} and \gls{pr} statuses were also examined.
+The generated metagenes were compared with the gene expressions of the corresponding pathway-associated genetic signatures, and the association with the patient \gls{er} and \gls{pr} statuses were also examined.
 As shown in \cref{fig:er_cr,fig:pr_cr}, both the \gls{er} and \gls{pr} metagenes in the CR data showed assocation with the gene expression pattern of the pathway, and also showed significant association with the patient \gls{er} and \gls{pr} statuses.

### 2017-08-22 thesis/misc/appendix2.tex
tmp commit
Added marking_response to the thesis directory temporarily

@@
 
-	\section{Ranking method for the pathway associated genetic signatures}
+	\section{Ranking method for the pathway-associated genetic signatures}
 	\label{sec:ranking_method_for_the_pathway_associated_genetic_signatures}
@@
 	Since the results from the \gls{rma} and \gls{mas} normalised GT data were so similar to one another, only the results from the \gls{rma} normalised data are presented (\cref{fig:appendix/rank_scatter}).
-	\cref{fig:appendix/rank_scatter} showed that the metagene scores for all of the GT pathway associated genetic signatures were approximately the same in either of the ranking methods used.
+	\cref{fig:appendix/rank_scatter} showed that the metagene scores for all of the GT pathway-associated genetic signatures were approximately the same in either of the ranking methods used.
 
@@
 
-	\section{Normalisation method for the pathway associated genetic signature transformation matrices}
+	\section{Normalisation method for the pathway-associated genetic signature transformation matrices}
 	\label{sec:normalisation_method_for_pathway_associated_genetic_signature_transformation_matrices}
 
-	To investigate whether the normalisation methods were important in the creation of the pathway associated genetic signature transformation matrices, and whether the normalisation methods of the data affected the resulting metagenes, transformation matrices were generated in \gls{rma} or \gls{mas} normalised GT data set and then applied to the \gls{rma} or \gls{mas} normalised GT data set.
+	To investigate whether the normalisation methods were important in the creation of the pathway-associated genetic signature transformation matrices, and whether the normalisation methods of the data affected the resulting metagenes, transformation matrices were generated in \gls{rma} or \gls{mas} normalised GT data set and then applied to the \gls{rma} or \gls{mas} normalised GT data set.
 	The results in \cref{fig:appendix/gt_meta_rma_mas} showed less variability of the metagene scores when the metagenes were derived from the data with the same normalisation method, regardless of the transformation matrix used to produce the metagenes.
-	This was most clearly demonstrated by the \gls{bcat}, \gls{er}, \gls{ifna}, \gls{ifny}, p53 and \gls{pr} pathway associated genetic signatures.
+	This was most clearly demonstrated by the \gls{bcat}, \gls{er}, \gls{ifna}, \gls{ifny}, p53 and \gls{pr} pathway-associated genetic signatures.
 
@@
 
-	The directionality of the pathway metagenes were determined by comparing the metagene values with the gene expression of the representative gene for the pathway associated genetic signature, as well as the gene expression pattern of the pathway signature (\cref{fig:appendix/gt_pathmeta_rank}).
+	The directionality of the pathway metagenes were determined by comparing the metagene values with the gene expression of the representative gene for the pathway-associated genetic signature, as well as the gene expression pattern of the pathway signature (\cref{fig:appendix/gt_pathmeta_rank}).
 	Together with the original results presented by \citet{Gatza2010a} (\cref{sec:result_from_gatza2010a_study}), the final directions of the pathway metagenes were decided.
@@
 		\includegraphics[width=0.32\linewidth,page=6]{appendix/gtrmastdrank}\\
-		\caption[Directionality of the pathway metagenes in the GT data]{Heatmaps showing the gene expression of the pathway associated genetic signatures, pathway metagenes, and the expression of the representative pathway gene in \gls{rma}-normalised GT data set.
-		For each pathway associated genetic signature, the gene expression pattern of the signature was plotted in the heatmap.
+		\caption[Directionality of the pathway metagenes in the GT data]{Heatmaps showing the gene expression of the pathway-associated genetic signatures, pathway metagenes, and the expression of the representative pathway gene in \gls{rma}-normalised GT data set.
+		For each pathway-associated genetic signature, the gene expression pattern of the signature was plotted in the heatmap.
 		The expression of the representative gene for the pathway signature (see \cref{tab:metagene_direction} in \cref{sub:metagene_direction}) was plotted seaparately as a bar above the pathway metagene scores.
@@
 
-	The original result from the \citet{Gatza2010a} study were used to determine the directions of the GT pathway associated genetic signatures in \cref{sec:pathway_associated_genetic_signatures_from_gatza2010a_study} (\cref{fig:gatza_paper_res}).
+	The original result from the \citet{Gatza2010a} study were used to determine the directions of the GT pathway-associated genetic signatures in \cref{sec:pathway_associated_genetic_signatures_from_gatza2010a_study} (\cref{fig:gatza_paper_res}).
 
@@
 		\includegraphics[width=0.7\linewidth]{results2/gatzares1}
-		\caption[Original result presented in the \citet{Gatza2010a} study]{Heatmap showing the correlation of all of the GT pathway associated genetic signatures with one another.
+		\caption[Original result presented in the \citet{Gatza2010a} study]{Heatmap showing the correlation of all of the GT pathway-associated genetic signatures with one another.
 		Figure adapted from the \citet{Gatza2010a} study.}
@@
 
-	The correlation of the pathway associated genetic signatures were plotted as heatmaps for the CR, FM and \gls{nzbc} data sets to confirm that the pathway transformation matrices from the GT data set were generating metagenes that were similar to those in the original \gls{rma} normalised GT data.
+	The correlation of the pathway-associated genetic signatures were plotted as heatmaps for the CR, FM and \gls{nzbc} data sets to confirm that the pathway transformation matrices from the GT data set were generating metagenes that were similar to those in the original \gls{rma} normalised GT data.
 	Most importantly, the pathway metagenes generated from the transformation matrices had to cluster in a similar pattern as the pathway metagenes in the GT data set.
@@
 
-	The directionality of the CR obesity associated genetic signatures were determined in the \gls{rma} normalised GT data set to make sure these signatures were in the ``correct'' direction, and therefore able to be compared with the pathway associated genetic signatures appropriately.
+	The directionality of the CR obesity-associated genetic signatures were determined in the \gls{rma} normalised GT data set to make sure these signatures were in the ``correct'' direction, and therefore able to be compared with the pathway-associated genetic signatures appropriately.
 	As clearly shown in \cref{fig:appendix/ob_meta_dir_gt}, the Res metagene was flipped as the Res metagene was in the oppsite direction compared with the other obesity metagenes.
@@
 		\includegraphics[width=0.32\linewidth,page=9]{appendix/gtobrma}\\
-		\caption[Heatmaps of all the obesity and the pathway metagenes in the CR, \gls{nzbc} and FM data]{Heatmaps showing the obesity metagenes from the CR data set with the gene expressions of the obesity associated genetic signatures in the \gls{rma} normalised GT data set.
+		\caption[Heatmaps of all the obesity and the pathway metagenes in the CR, \gls{nzbc} and FM data]{Heatmaps showing the obesity metagenes from the CR data set with the gene expressions of the obesity-associated genetic signatures in the \gls{rma} normalised GT data set.
 		High and low correlation were represented as red and blue, respectively, where the colours were matched with the values on the scale shown in the top left histogram.}
@@
 
-	Correlation of the pathway and obesity associated genetic signatures were plotted in a heatmap for the GT data in \cref{sec:pathway_associated_metagenes_and_obesity_associated_metagenes}.
+	Correlation of the pathway and obesity-associated genetic signatures were plotted in a heatmap for the GT data in \cref{sec:pathway_associated_metagenes_and_obesity_associated_metagenes}.
 	In this section, the correlation of these genetic signatures were plotted as heatmaps for the CR, FM and \gls{nzbc} data sets.
@@
 
-	The \gls{pr}-only linear models for the other obesity associated genetic signatures that were generated in the \gls{nzbc} data were summarised in \cref{tab:lm_pr_only_crol,tab:lm_pr_only_res,tab:lm_pr_only_resol,tab:lm_pr_only_ca,tab:lm_pr_only_caol,tab:lm_pr_only_cares,tab:lm_pr_only_caresol,tab:lm_pr_only_or,tab:lm_pr_only_fm}.
+	The \gls{pr}-only linear models for the other obesity-associated genetic signatures that were generated in the \gls{nzbc} data were summarised in \cref{tab:lm_pr_only_crol,tab:lm_pr_only_res,tab:lm_pr_only_resol,tab:lm_pr_only_ca,tab:lm_pr_only_caol,tab:lm_pr_only_cares,tab:lm_pr_only_caresol,tab:lm_pr_only_or,tab:lm_pr_only_fm}.
 	The \gls{pr} pathway metagene was significant in all of the linear models that predicted the obesity metagenes derived from the CR data set \cref{tab:lm_pr_only_crol,tab:lm_pr_only_res,tab:lm_pr_only_resol,tab:lm_pr_only_ca,tab:lm_pr_only_caol,tab:lm_pr_only_cares,tab:lm_pr_only_caresol,tab:lm_pr_only_or}.
@@
 
-	All of the \gls{r} source code used in this project is publicly available at \url{https://github.com/rikutakei/bmi-cancer-code.git}.
+	All of the \gls{r} source code used in this project is publicly available at: \url{https://github.com/rikutakei/bmi-cancer-code.git}.
 	This thesis was written with \LaTeX{}, and the source code is publicly available at \url{https://github.com/rikutakei/mastersDoc.git}.

### 2017-08-22 thesis/misc/glossary.tex
Added marking_response to the thesis directory temporarily

@@
 \newacronym[description={Electron transport chain}]{etc}{ETC}{electron transport chain}
-\newacronym[description={Endothelial-mesenchymal transition}]{emt}{EMT}{endothelial-mesenchymal transition}
+\newacronym[description={Epithelial-mesenchymal transition}]{emt}{EMT}{epithelial-mesenchymal transition}
 \newacronym[description={Epidermal growth factor receptor}]{egfr}{EGFR}{epidermal growth factor receptor}
@@
 \newacronym[description={Melanocyte-stimulating hormone}]{msh}{MSH}{melanocyte-stimulating hormone}
-\newacronym[description={Mesenchymal-endothelial transition}]{met}{MET}{mesenchymal-endothelial transition}
+\newacronym[description={Mesenchymal-epithelial transition}]{met}{MET}{mesenchymal-epithelial transition}
 \newacronym[description={Messenger RNA}]{mrna}{mRNA}{messenger RNA}

### 2017-08-22 thesis/results/results.tex
Final commit before submission of this thesis
Added marking_response to the thesis directory temporarily

@@
-\chapter{Obesity associated genetic signatures and cancer}
+\chapter{Obesity-associated genetic signatures and cancer}
 \label{cha:obesity_genetic_signatures_and_cancer}
 
-The obesity associated genetic signatures are central to this project, as a means to clarify the relationship between \gls{bmi} and tumour biology.
-In this chapter, the previously identified obesity associated genetic signatures from the studies conducted by \citet{Creighton2012} and \citet{Fuentes-Mattei2014} are examined in turn to judge the agreement of these signatures with the patient \gls{bmi} and \gls{bmi} status, as presented in their results.
-After this, novel obesity associated genetic signatures are identified in the Creighton \textit{et al.} (CR) data set and was compared with the obesity associated genetic signatures from the \citet{Creighton2012} and the \citet{Fuentes-Mattei2014}  studies.
+The obesity-associated genetic signatures are central to this project, as a means to clarify the relationship between \gls{bmi} and tumour biology.
+In this chapter, the previously identified obesity-associated genetic signatures from the studies conducted by \citet{Creighton2012} and \citet{Fuentes-Mattei2014} are examined in turn to judge the agreement of these signatures with the patient \gls{bmi} and \gls{bmi} status, as presented in their results.
+After this, novel obesity-associated genetic signatures were identified in the Creighton \textit{et al.} (CR) data set and was compared with the obesity-associated genetic signatures from the \citet{Creighton2012} and the \citet{Fuentes-Mattei2014}  studies.
 Lastly, the presence of common genes or pathways associated with obesity in multiple types of cancer is explored using the data sets from \gls{icgc} (\citetalias{ICGC}; \citealp{Zhang2011}).
 
-\section{Obesity associated genetic signature from the\\\citet{Creighton2012} study}
+\section{Obesity-associated genetic signature from the\\\citet{Creighton2012} study}
 \label{sec:creighton_obesity_metagene}
 
-An obesity metagene was created using the obesity associated genetic signature from the \citet{Creighton2012} study.
+An obesity metagene was created using the obesity-associated genetic signature from the \citet{Creighton2012} study.
 There was no description about the normalisation method used by Creighton \textit{et al.} when they first analysed their data, so the more popular \gls{rma} method  was used to  normalise the CR data set \citep{Irizarry2003}.
@@
 	\includegraphics[page=3,width=0.8\linewidth]{results1/creighton_mg_heatmap1}
-	\caption[Obesity metagene from the \citet{Creighton2012} study and sample gene expression in CR data] {Heatmap showing the obesity metagene from the \citet{Creighton2012} study with the sample gene expressions of the obesity associated genes in the CR data.
-	Level of expression is represented in the top right histogram, where low and high gene expression are colour-coded with blue and red, respectively.
-	Each row of the heatmap represents a gene from the obesity associated genetic signature, and each column of the heatmap represents a sample from the CR data.
-	The obesity associated metagene scores of the samples are shown as a separate row at top of the heatmap, and the tree diagram of the hierarchical clustering of the genes is shown on the left of the heatmap.
+	\caption[Obesity metagene from the \citet{Creighton2012} study and sample gene expression in CR data] {Heatmap showing the obesity metagene from the \citet{Creighton2012} study with the sample gene expressions of the obesity-associated genes in the CR data.
+	Level of expression is represented in the top left histogram, where low and high gene expression are colour-coded with blue and red, respectively.
+	Each row of the heatmap represents a gene from the obesity-associated genetic signature, and each column of the heatmap represents a sample from the CR data.
+	The obesity-associated metagene scores of the samples are shown as a separate row at top of the heatmap, and the tree diagram of the hierarchical clustering of the genes is shown on the left of the heatmap.
 	For clarity, the sign of the metagene scores were reversed in order to match the results from the \citet{Creighton2012} study.}
@@
 
-As shown in \cref{fig:crmetaheat}, a high obesity associated metagene score  reflects low expression in majority of the genes in the signature, and in contrast, a low obesity associated metagene score reflects high expression in the majority of the genes in the signature.
-This was consistent with the reported property of the obesity associated genetic signature by \citet{Creighton2012} (see \cref{ssub:creighton_study}).
+As shown in \cref{fig:crmetaheat}, a high obesity-associated metagene score  reflects low expression in majority of the genes in the signature, and in contrast, a low obesity-associated metagene score reflects high expression in the majority of the genes in the signature.
+This was consistent with the reported property of the obesity-associated genetic signature by \citet{Creighton2012} (see \cref{ssub:creighton_study}).
 To provide further evidence that the obesity metagenes were in fact associated with both the \gls{bmi} status and the \gls{bmi} value of the patients, a box plot and a scatter plot were created, respectively (\cref{fig:crmetaboxplot}).
@@
 It should be noted here that the obesity metagene significantly associated with the samples from the patients that were obese, but not with the sample from the patients that were overweight.
-This was due to the fact that the obesity associated genetic signature was originally identified from the comparison of the patients that were obese with the patients that were not obese, and therefore the obesity metagene scores were significant with the obese group, but not with the overweight group.
+This was due to the fact that the obesity-associated genetic signature was originally identified from the comparison of the patients that were obese with the patients that were not obese, and therefore the obesity metagene scores were significant with the obese group, but not with the overweight group.
 Another thing to note here was that even though  the regression line in the scatter plot showed statistically significant association with the patient \gls{bmi}, patient \gls{bmi} values seem to be randomly dispersed across the obesity metagene scores.
@@
 First, the transformation matrix was derived from the CR microarray data, but the \gls{icgc} cancer data were generated via \gls{rnaseq}.
-Though the log$_{10}$-normalisation and standardisation of the data was the most appropriate adjustments to be made to the \gls{rnaseq} data, these adjustments were not equivalent to the \gls{rma} normalisation method that was used on the microarray data.
+Though the log$_{10}$-normalisation and standardisation of the data were the most appropriate adjustments to be made to the \gls{rnaseq} data, these adjustments were not equivalent to the \gls{rma}-normalisation method that was used on the microarray data.
 Secondly, none of the \gls{icgc} cancer data in this project originated from breast tumours as in the CR data.
-Since the obesity associated genetic signature was identified in the breast cancer data, the signature may be specific to breast cancer and may not be generalisable to other cancer types.
+Since the obesity-associated genetic signature was identified in the breast cancer data, the signature may be specific to breast cancer and may not be generalisable to other cancer types.
 \\
@@
 However, as with the \gls{icgc} cancer data, the obesity metagene scores did not significantly associate with patient \gls{bmi} or \gls{bmi} status (\cref{fig:crmetaprint}).
-These results confirmed that the lack of association of the obesity associated genetic signature from the Creighton \textit{et al.} study was not due to the technology in which the data was gathered (microarray or \gls{rnaseq}), nor the cancer type in which the genetic signature was derived from.
+These results confirmed that the lack of association of the obesity-associated genetic signature from the Creighton \textit{et al.} study was not due to the technology in which the data was gathered (microarray or \gls{rnaseq}), nor the cancer type in which the genetic signature was derived from.
 
@@
 
-Taken together, these results suggest that the obesity associated genetic signature identified in the study conducted by \citet{Creighton2012} correlated with patient \gls{bmi} and \gls{bmi} status only in the CR data set.
+Taken together, these results suggest that the obesity-associated genetic signature identified in the study conducted by \citet{Creighton2012} correlated with patient \gls{bmi} and \gls{bmi} status only in the CR data set.
 The lack of association with patient \gls{bmi} in other cancer data sets was not due to the type of technology platform in which the data was gathered, as neither the \gls{icgc} \gls{rnaseq} data nor the \gls{nzbc} microarray data showed significant association of the obesity metagene with the patient \gls{bmi}.
-Furthermore, the obesity associated genetic signature was not dependent on the cancer type in which it was generated from, since the obesity metagene did not show significant association in either the \gls{icgc} cancer data or in the \gls{nzbc} data.
+Furthermore, the obesity-associated genetic signature was not dependent on the cancer type in which it was generated from, since the obesity metagene did not show significant association in either the \gls{icgc} cancer data or in the \gls{nzbc} data.
 
 One possible reason why the obesity metagene from the \citet{Creighton2012} study did not show significant association with other data sets could be because the genetic signature was in fact not an obesity specific signature, but a signature that was detected due to another clinical variable (investigated in \cref{sec:creighton_obesity_metagene_new}).
-Another reason for this apparent lack of association could be that the obesity associated genetic signature was too specific to the CR data and was not a broad obesity associated genetic signature, but an obesity associated signature that was specific to the patient cohort that was profiled in the \citet{Creighton2012} publication.
+Another reason for this apparent lack of association could be that the obesity-associated genetic signature was too specific to the CR data and was not a broad obesity-associated genetic signature, but an obesity-associated signature that was specific to the patient cohort that was profiled in the \citet{Creighton2012} publication.
 
-\section{Obesity associated genetic signature from the\\\citet{Fuentes-Mattei2014} study}
+\vspace{-3mm}
+
+\section{Obesity-associated genetic signature from the\\\citet{Fuentes-Mattei2014} study}
 \label{sec:fm_obesity_metagene}
 
-The obesity associated genetic signature from the \citet{Creighton2012} study was not associated with patient \gls{bmi} or \gls{bmi} status in majority of the cancer data sets, so the obesity associated genetic signature from the \citet{Fuentes-Mattei2014} study (FM) was examined to see whether this obesity metagene was able to significantly associate with the patient \gls{bmi} and \gls{bmi} status across different  data sets.
+The obesity-associated genetic signature from the \citet{Creighton2012} study was not associated with patient \gls{bmi} or \gls{bmi} status in majority of the cancer data sets, so the obesity-associated genetic signature from the \citet{Fuentes-Mattei2014} study (FM) was examined to see whether this obesity metagene was able to significantly associate with the patient \gls{bmi} and \gls{bmi} status across different  data sets.
 Since the FM data set did not have patient \gls{bmi} information, the FM obesity metagene was not able to be compared with the patient \gls{bmi} or \gls{bmi} status in the original FM data.
@@
 	\includegraphics[page=5,width=0.45\linewidth]{results1/cris_fm_meta}
-	\caption[FM metagene in the \gls{nzbc} data]{Heatmap, scatter plot and box plot showing the association of the FM obesity associated metagene with the sample gene expression, patient  \gls{bmi} and \gls{bmi} status, respectively, from the \gls{nzbc} data.
+	\caption[FM metagene in the \gls{nzbc} data]{Heatmap, scatter plot and box plot showing the association of the FM obesity-associated metagene with the sample gene expression, patient  \gls{bmi} and \gls{bmi} status, respectively, from the \gls{nzbc} data.
 	Scales, p-values and $R^2$-value are as described in previous figures.}
@@
 Next, the transformation matrix was applied to the \gls{icgc} cancer data and the resulting FM metagene was compared with the sample gene expression, patient \gls{bmi} and \gls{bmi} status in each of the \gls{icgc} cancer data.
-As evident in \cref{fig:fmmetaicgc,sec:rest_of_the_fm_icgc_cancer_heatmap_results}, the FM obesity metagene scores appeared to reflect the overall gene expression of the FM obesity associated genetic signature.
+As evident in \cref{fig:fmmetaicgc,sec:rest_of_the_fm_icgc_cancer_heatmap_results}, the FM obesity metagene scores appeared to reflect the overall gene expression of the FM obesity-associated genetic signature.
 As with all the results in this chapter so far, the FM obesity metagene did not significantly associate with any of the \gls{icgc} cancer data, except in the \gls{blca} data set (\cref{fig:fmmetaicgc}; \cref{sec:rest_of_the_fm_icgc_cancer_heatmap_results}).
@@
 
-These results showed that the obesity associated genetic signature from the \citet{Fuentes-Mattei2014} study was not generalisable to other cancer data sets, similar to the obesity signature identified by Creighton \textit{et al.}.
+These results showed that the obesity-associated genetic signature from the \citet{Fuentes-Mattei2014} study was not generalisable to other cancer data sets, similar to the obesity signature identified by Creighton \textit{et al.}.
 This meant that both the Creighton \textit{et al.} and Fuentes-Mattei \textit{et al.} obesity metagenes may have been too specific to the original data set in which the signatures were identified in.
-Furthermore, there was a possibility that these obesity associated metagenes were not related to obesity, but associated with a different clinical variable that may be closely related to \gls{bmi} (\cref{sec:creighton_obesity_metagene_new}).
+Furthermore, there was a possibility that these obesity-associated metagenes were not related to obesity, but associated with a different clinical variable that may be closely related to \gls{bmi} (\cref{sec:creighton_obesity_metagene_new}).
 
-\section{Novel obesity associated genetic signatures from the \citet{Creighton2012} data set}
+\section{Novel obesity-associated genetic signatures from the \citet{Creighton2012} data set}
 \label{sec:creighton_obesity_metagene_new}
 
-\subsection{Identification of novel obesity associated genetic signatures}
+\subsection{Identification of novel obesity-associated genetic signatures}
 \label{sub:identification_of_obesity_associated_genetic_signatures}
 
-Both the obesity associated genetic signatures identified from the \citet{Creighton2012} and \citet{Fuentes-Mattei2014} studies were able to capture the overall gene expression patterns  of the genetic signatures of the samples, but did not associate with patient \gls{bmi} or \gls{bmi} status in majority of the cancer data sets.
-One possible reason for this result could be that the obesity associated genetic signatures from the \citet{Creighton2012} and \citet{Fuentes-Mattei2014} studies may not have been truly associated with patient \gls{bmi}, but with another clinical variable.
-To investigate this possibility, novel obesity associated genetic signatures were identified in the CR data after controlling for all the clinical variables in the data set.
-The FM data set was not used to get the obesity associated genetic signatures, as no \gls{bmi} information was available for the patients in the FM data.
+Both the obesity-associated genetic signatures identified from the \citet{Creighton2012} and \citet{Fuentes-Mattei2014} studies were able to capture the overall gene expression patterns  of the genetic signatures of the samples, but did not associate with patient \gls{bmi} or \gls{bmi} status in majority of the cancer data sets.
+One possible reason for this result could be that the obesity-associated genetic signatures from the \citet{Creighton2012} and \citet{Fuentes-Mattei2014} studies may not have been truly associated with patient \gls{bmi}, but with another clinical variable.
+To investigate this possibility, novel obesity-associated genetic signatures were identified in the CR data after controlling for all the clinical variables in the data set.
+The FM data set was not used to get the obesity-associated genetic signatures, as no \gls{bmi} information was available for the patients in the FM data.
 
-Firstly, an attempt was made to replicate the Creighton \textit{et al.} obesity associated genetic signature from the original CR data.
-\citet{Creighton2012} originally found their obesity associated genetic signature by doing a gene expression analysis between the breast tumour samples from the obese and non-obese patients, and from this list, Creighton \textit{et al.} selected the 799 statistically significant \glspl{deg} (p \textless{} 0.01) with log$_2$-fold change greater than 1.2.
+Firstly, an attempt was made to replicate the Creighton \textit{et al.} obesity-associated genetic signature from the original CR data.
+\citet{Creighton2012} originally found their obesity-associated genetic signature by doing a gene expression analysis between the breast tumour samples from the obese and non-obese patients, and from this list, Creighton \textit{et al.} selected the 799 statistically significant \glspl{deg} (p \textless{} 0.01) with log$_2$-fold change greater than 1.2.
 Therefore, gene expression analysis  between the samples from obese and non-obese patients in the \gls{rma} normalised CR data was carried out (described in \cref{sec:gene_expression_analysis}).
@@
 From these observations, the log$_2$-fold change of the gene probes were ignored and the threshold p-value was set to 0.01 (unadjusted p-value) for the identification of the significant probe sets.
-Additionally, when there were more than 799 gene probes identified, only the most significant 799 gene probes were taken as the obesity associated genetic signature, as this many genes were originally identified by Creighton \textit{et al.}.
-This was to include as many probe set as possible for the novel genetic signatures to be comparable with the original Creighton \textit{et al.} obesity associated genetic signature.
+Additionally, when there were more than 799 gene probes identified, only the most significant 799 gene probes were taken as the obesity-associated genetic signature, as this many genes were originally identified by Creighton \textit{et al.}.
+This was to include as many probe set as possible for the novel genetic signatures to be comparable with the original Creighton \textit{et al.} obesity-associated genetic signature.
 These criteria were applied for the identification of other genetic signatures as well.
 
-% To determine whether the identified obesity associated genetic signature was similar to the signature that Creighton \textit{et al.} had originally found, the identified \glspl{deg} were compared with the original genetic signature identified by Creighton \textit{et al.}.
+% To determine whether the identified obesity-associated genetic signature was similar to the signature that Creighton \textit{et al.} had originally found, the identified \glspl{deg} were compared with the original genetic signature identified by Creighton \textit{et al.}.
 
 The above analysis was repeated with the residual data (\gls{rma} normalised CR data that had been controlled for other clinical variables; see \cref{sub:residual_data_creation}).
-The clinical variables controlled were age, ethnicity, menopause status, tumour grade, hormone (\gls{er}, \gls{pr}, and \gls{her2}) statuses  and \gls{ln} status.
+The clinical variables controlled were age, ethnicity, menopause status, tumour grade, receptor (\gls{er}, \gls{pr}, and \gls{her2}) statuses  and \gls{ln} status.
 In the residual data, 1104 gene probes were significantly differentially expressed with unadjusted p-value (p \textless{} 0.01; additional results shown in \cref{tab:ob_deg_summary}).
-Again, the most significant 799 gene  probes were taken as the obesity associated genetic signature from this data.
+Again, the most significant 799 gene  probes were taken as the obesity-associated genetic signature from this data.
 
-In addition to the above two obesity associated genetic signatures, two more sets of genetic signatures were identified by taking the gene probes that were common between each of the above two signatures with the original obesity associated genetic signatures.
-There were 239 common gene probes between the original Creighton \textit{et al.} obesity associated  genetic signature and the gene probes identified from the unadjusted CR data, and 168 common gene probes between the original signature and the gene probes from the residual CR data.
+In addition to the above two obesity-associated genetic signatures, two more sets of genetic signatures were identified by taking the gene probes that were common between each of the above two signatures with the original obesity-associated genetic signatures.
+There were 239 common gene probes between the original Creighton \textit{et al.} obesity-associated  genetic signature and the gene probes identified from the unadjusted CR data, and 168 common gene probes between the original signature and the gene probes from the residual CR data.
 The genetic signatures identified were summarised as a Venn diagram, as shown in \cref{fig:venn1}.
@@
 	\includegraphics[width=0.5\linewidth]{results1/deg_venn1}
-	\caption[Venn diagram of the \glspl{deg} identified from the CR data (all patients included)]{Venn diagram showing the common gene probes between the signatures obtained from the unadjusted and the clinical variable-adjusted CR data and the original obesity associated genetic signature from the \citet{Creighton2012} study }
+	\caption[Venn diagram of the \glspl{deg} identified from the CR data (all patients included)]{Venn diagram showing the common gene probes between the signatures obtained from the unadjusted and the clinical variable-adjusted CR data and the original obesity-associated genetic signature from the \citet{Creighton2012} study }
 	\label{fig:venn1}
@@
 
-With the Caucasian-only CR data set, obesity associated genetic signatures were identified as described above.
-% ; first in the unadjusted data, then in the data with clinical variables adjusted (except ethnicity, as it has already been controlled for by considering only the Caucasian samples), and lastly the common genes between each of these two signatures with the original obesity associated genetic signature were identified.
+With the Caucasian-only CR data set, obesity-associated genetic signatures were identified as described above.
+% ; first in the unadjusted data, then in the data with clinical variables adjusted (except ethnicity, as it has already been controlled for by considering only the Caucasian samples), and lastly the common genes between each of these two signatures with the original obesity-associated genetic signature were identified.
 2129 and 1558 gene probes were significantly differentially expressed (unadjusted p-value \textless{} 0.01) in the unadjusted and clinical variable-adjusted Caucasian patient data, respectively (\cref{tab:ob_deg_summary}).
 As before, the most significant 799 gene probes were selected from these gene probes.
-There were 148 and 92 common gene probes with the original Creighton \textit{et al.} obesity associated genetic signatures and unadjusted or clinical variable-adjusted Caucasian patient data set, respectively.
+There were 148 and 92 common gene probes with the original Creighton \textit{et al.} obesity-associated genetic signatures and unadjusted or clinical variable-adjusted Caucasian patient data set, respectively.
 Again, Venn diagram was used to summarise the genes identified from the Caucasian patient data set (\cref{fig:venn2}).
@@
 	\includegraphics[width=0.5\linewidth]{results1/deg_venn2}
-	\caption[Venn diagram of the \glspl{deg} identified from the CR data (Caucasian patients only)]{Venn diagram showing the common gene probes between the signatures obtained from the unadjusted and the clinical variable-adjusted CR data (Caucasian patients only) and the original obesity associated genetic signature from the \citet{Creighton2012} study}
+	\caption[Venn diagram of the \glspl{deg} identified from the CR data (Caucasian patients only)]{Venn diagram showing the common gene probes between the signatures obtained from the unadjusted and the clinical variable-adjusted CR data (Caucasian patients only) and the original obesity-associated genetic signature from the \citet{Creighton2012} study}
 	\label{fig:venn2}
@@
 	\includegraphics[width=0.8\linewidth]{results1/cr_meta_cor}
-	\caption[Pearson correlation of all eight obesity metagenes identified in the CR data]{Heatmap showing the Pearson correlation of all eight obesity associated metagenes from the CR data with one another.
+	\caption[Pearson correlation of all eight obesity metagenes identified in the CR data]{Heatmap showing the Pearson correlation of all eight obesity-associated metagenes from the CR data with one another.
 	\gls{svd} was applied to the \gls{rma} normalised CR data to generate each of the eight metagenes.
-	High and low correlation were represented as red and yellow, respectively, where the colours were matched with the values on the scale shown in the top right histogram.
+	High and low correlation were represented as red and yellow, respectively, where the colours were matched with the values on the scale shown in the top left histogram.
 	The lowest correlation was 0.8469, between the CrOl and the CaRes metagenes.
@@
 
-Since all of these obesity associated genetic signatures were derived based on the discrete values of the patient \gls{bmi} status, an obesity associated genetic signature was searched for using the correlation of the gene probe expression with the patient \gls{bmi} value (continuous variable) in the CR data with all of the samples included, but no significant genes were identified (\cref{sec:obesity_associated_genetic_signature_using_sample_bmi_values}).
-All of these obesity associated genetic signatures (eight in total) were checked to see whether these signatures had significant association with patient \gls{bmi} or \gls{bmi} status in the \gls{nzbc} and \gls{icgc} cancer data sets (\cref{sub:_novel_obesity_associated_signatures_and_sample_bmi}).
+Since all of these obesity-associated genetic signatures were derived based on the discrete values of the patient \gls{bmi} status, an obesity-associated genetic signature was searched for using the correlation of the gene probe expression with the patient \gls{bmi} value (continuous variable) in the CR data with all of the samples included, but no significant genes were identified (\cref{sec:obesity_associated_genetic_signature_using_sample_bmi_values}).
+All of these obesity-associated genetic signatures (eight in total) were checked to see whether these signatures had significant association with patient \gls{bmi} or \gls{bmi} status in the \gls{nzbc} and \gls{icgc} cancer data sets (\cref{sub:_novel_obesity_associated_signatures_and_sample_bmi}).
 For simplicity, the abbreviations in \cref{tab:mg_abbrev} will be used to refer to the appropriate genetic signatures.
@@
 	\centering
-	\caption{Summary of the abbreviations used to refer to the different obesity associated genetic signatures identified in the CR data}
+	\caption{Summary of the abbreviations used to refer to the different obesity-associated genetic signatures identified in the CR data}
 	\label{tab:mg_abbrev}
@@
 		\hline
-		\rule{0pt}{2.25ex}Or      & Original obesity associated genetic signature identified by \citet{Creighton2012}                       & 799\\
-		\rule{0pt}{2.25ex}Cr      & Obesity associated genetic signature identified from unadjusted CR data                                & 799 \\
+		\rule{0pt}{2.25ex}Or      & Original obesity-associated genetic signature identified by \citet{Creighton2012}                       & 799\\
+		\rule{0pt}{2.25ex}Cr      & Obesity-associated genetic signature identified from unadjusted CR data                                & 799 \\
 		\rule{0pt}{2.25ex}CrOl    & Genes common between Or and Cr genetic signatures                                                       & 239\\
-		\rule{0pt}{2.25ex}Res     & Obesity associated genetic signature identified from clinical variable-adjusted CR data                & 799\\
+		\rule{0pt}{2.25ex}Res     & Obesity-associated genetic signature identified from clinical variable-adjusted CR data                & 799\\
 		\rule{0pt}{2.25ex}ResOl   & Genes common between Or and Res genetic signatures                                                      & 168\\
-		\rule{0pt}{2.25ex}Ca      & Obesity associated genetic signature identified from unadjusted Caucasian-only CR data                 & 799\\
+		\rule{0pt}{2.25ex}Ca      & Obesity-associated genetic signature identified from unadjusted Caucasian-only CR data                 & 799\\
 		\rule{0pt}{2.25ex}CaOl    & Genes common between Or and Ca genetic signatures                                                       & 148\\
-		\rule{0pt}{2.25ex}CaRes   & Obesity associated genetic signature identified from clinical variable-adjusted Caucasian-only CR data & 799\\
+		\rule{0pt}{2.25ex}CaRes   & Obesity-associated genetic signature identified from clinical variable-adjusted Caucasian-only CR data & 799\\
 		\rule{0pt}{2.25ex}CaResOl & Genes common between Or and CaRes genetic signatures                                                    & 92\\
@@
 
-\subsection{Novel obesity associated signatures and patient \gls{bmi}/\gls{bmi} status}
+\subsection{Novel obesity-associated signatures and patient \gls{bmi}/\gls{bmi} status}
 \label{sub:_novel_obesity_associated_signatures_and_sample_bmi}
 
-As with the Or signature, all eight obesity associated genetic signatures were validated in the CR data set first, and then compared in other cancer data sets by using the transformation matrix that was generated in the \gls{rma} normalised CR data.
+As with the Or signature, all eight obesity-associated genetic signatures were validated in the CR data set first, and then compared in other cancer data sets by using the transformation matrix that was generated in the \gls{rma} normalised CR data.
 The CR data (unadjusted, all samples included) was normalised with the \gls{rma} method and \gls{svd} was applied to generate the all of the obesity metagenes.
@@
 The comparison of the metagenes with the sample gene expression in the CR data were displayed as heatmaps (\cref{fig:degmetacr}; \cref{sec:rest_of_the_cr_ob_meta_heatmap_results_cr}).
-It was clear from the heatmaps that all of the metagenes reflected the overall expression of the corresponding obesity associated genetic signatures.
-The association of the metagenes with patient \gls{bmi} and \gls{bmi} status was significant for all eight of the obesity associated genetic signatures identified (\cref{fig:degmetacr}; \cref{sec:rest_of_the_cr_ob_meta_heatmap_results_cr}).
-These results confirmed that all of the obesity associated genetic signatures identified in \cref{sub:identification_of_obesity_associated_genetic_signatures} significantly associated with patient \gls{bmi} and \gls{bmi} status in the CR data in which the metagenes were derived from.
+It was clear from the heatmaps that all of the metagenes reflected the overall expression of the corresponding obesity-associated genetic signatures.
+The association of the metagenes with patient \gls{bmi} and \gls{bmi} status was significant for all eight of the obesity-associated genetic signatures identified (\cref{fig:degmetacr}; \cref{sec:rest_of_the_cr_ob_meta_heatmap_results_cr}).
+These results confirmed that all of the obesity-associated genetic signatures identified in \cref{sub:identification_of_obesity_associated_genetic_signatures} significantly associated with patient \gls{bmi} and \gls{bmi} status in the CR data in which the metagenes were derived from.
 In the \gls{nzbc} data, the higher the metagenes scores of the samples were, the more expressed the genes were (and vice versa), but lacked the association with the patient \gls{bmi} or \gls{bmi} status (\cref{fig:degmetaprint}; \cref{sec:rest_of_the_cr_ob_meta_heatmap_results_cris}).
@@
 	\includegraphics[page=5,width=0.45\linewidth]{results1/cr_deg_meta_vs_clin}
-	\caption[Cr obesity metagene in the CR data]{Heatmap, scatter plot and box plot showing the association of the Cr obesity associated metagene with the sample gene expression, patient \gls{bmi} and \gls{bmi} status, respectively, from the CR data.
+	\caption[Cr obesity metagene in the CR data]{Heatmap, scatter plot and box plot showing the association of the Cr obesity-associated metagene with the sample gene expression, patient \gls{bmi} and \gls{bmi} status, respectively, from the CR data.
 	The results for the other obesity metagenes are shown in \cref{sec:rest_of_the_cr_ob_meta_heatmap_results_cr}.
@@
 	\includegraphics[page=5,width=0.45\linewidth]{results1/cris_crdeg_trans_meta}
-	\caption[Cr obesity metagene in the \gls{nzbc} data]{Heatmap, scatter plot and box plot showing the association of the Cr obesity associated metagene with the sample gene expression, patient \gls{bmi} and \gls{bmi} status, respectively, from the \gls{nzbc} data.
+	\caption[Cr obesity metagene in the \gls{nzbc} data]{Heatmap, scatter plot and box plot showing the association of the Cr obesity-associated metagene with the sample gene expression, patient \gls{bmi} and \gls{bmi} status, respectively, from the \gls{nzbc} data.
 	The results for the other obesity metagenes are shown in \cref{sec:rest_of_the_cr_ob_meta_heatmap_results_cris}.
@@
 To confirm whether these metagenes showed significant association with patient \gls{bmi} or \gls{bmi} status in other cancer types, the transformation matrix was applied to the \gls{icgc} cancer data.
-In all of the \gls{icgc} cancer data sets, all eight metagene scores were reflective of the sample gene expression of the corresponding obesity associated genetic signatures (\cref{fig:degmetaicgc}; \cref{sec:rest_of_the_cr_icgc_cancer_heatmap_results}).
+In all of the \gls{icgc} cancer data sets, all eight metagene scores were reflective of the sample gene expression of the corresponding obesity-associated genetic signatures (\cref{fig:degmetaicgc}; \cref{sec:rest_of_the_cr_icgc_cancer_heatmap_results}).
 In all cancer types but \gls{blca}, none of the obesity metagenes significantly associated with the patient \gls{bmi} or \gls{bmi} status (\cref{sec:rest_of_the_cr_icgc_cancer_heatmap_results}).
@@
 
-Taken together, these results showed that, even though all of the obesity metagenes significantly associated with the patient \gls{bmi} and \gls{bmi} status in the original data where the genetic signatures were derived from, none of the metagenes were not generalisable in other cancer data sets.
+Taken together, these results showed that, even though all of the obesity metagenes significantly associated with the patient \gls{bmi} and \gls{bmi} status in the original data where the genetic signatures were derived from, none of the metagenes were generalisable in other cancer data sets.
 Furthermore, these results showed that the lack of association with patient \gls{bmi} and \gls{bmi} status were not due to other clinical variables in the CR data set.
-This raised a question of whether there was any obesity associated genetic signature that was common in multiple types of cancer (\cref{sec:common_genes_across_multiple_cancer_types}).
+This raised a question of whether there was any obesity-associated genetic signature that was common in multiple types of cancer (\cref{sec:common_genes_across_multiple_cancer_types}).
 
-\begin{table}[htpb]
+\begin{table}[p]
 	\centering
@@
 
-\begin{figure}[htp!]
+\begin{figure}[p]
 	\centering
@@
 
-It was clear that the all of the obesity associated genetic signatures created from the CR data set were not significantly associated with patient \gls{bmi} across different \gls{icgc} cancer data sets.
-To determine whether there was any obesity associated genetic signature that was expressed in multiple different cancer types, gene expression analysis was carried out on each of the eight \gls{icgc} cancer types, and common genes were searched from the \glspl{deg} that were identified.
+It was clear that all of the obesity-associated genetic signatures created from the CR data set were not significantly associated with patient \gls{bmi} across different \gls{icgc} cancer data sets.
+To determine whether there was any obesity-associated genetic signature that was expressed in multiple different cancer types, gene expression analysis was carried out on each of the eight \gls{icgc} cancer types, and common genes were searched from the \glspl{deg} that were identified.
 All of the \gls{icgc} cancer data sets were normalised with voom (\cref{ssub:rna_seq_data}), which were then put through the gene expression analysis pipeline to identify the \glspl{deg} between the obese and non-obese groups of samples (\cref{sec:gene_expression_analysis}).
@@
 
-\begin{table}[tb]
+\begin{table}[p]
 	\centering

### 2017-08-22 thesis/results/results2.tex
Added marking_response to the thesis directory temporarily

@@
-\chapter{Obesity associated genetic signatures and pathway signatures}
+\chapter{Obesity-associated genetic signatures and pathway signatures}
 \label{cha:obesity_associated_genetic_signature_and_pathway_signatures}
 
-In this chapter, the underlying biological mechanism of the obesity associated signatures from the CR data were investigated.
-In doing so, the pathway genetic signatures from the \citet{Gatza2010a} study (GT) were utilised to determine which biological pathways the obesity associated genetic signatures were most similar to.
-First, the direction of GT pathway associated genetic signatures were resolved, then the pathway associated metagenes were compared with the obesity associated metagenes, and lastly linear models were constructed based on the pathway metagenes and patient \gls{bmi} to predict the obesity associated metagenes.
+In this chapter, the underlying biological mechanism of the obesity-associated signatures from the CR data were investigated.
+In doing so, the pathway genetic signatures from the \citet{Gatza2010a} study (GT) were utilised to determine which biological pathways the obesity-associated genetic signatures were most similar to.
+First, the direction of GT pathway-associated genetic signatures were resolved, then the pathway-associated metagenes were compared with the obesity-associated metagenes, and lastly linear models were constructed based on the pathway metagenes and patient \gls{bmi} to predict the obesity-associated metagenes.
 
@@
 
-\section{Pathway associated genetic signatures from the\\\citet{Gatza2010a} study}
+\section{Pathway-associated genetic signatures from the\\\citet{Gatza2010a} study}
 \label{sec:pathway_associated_genetic_signatures_from_gatza2010a_study}
@@
 In the \citet{Gatza2010a} study, their data comprised of samples from different studies which were aggregated into a single data set that was \gls{mas} normalised, and the metagene scores were transformed to a scale between 0 to 1 using a probit function.
-However, the analyses so far have used the \gls{rma} normalisation method and ranked based on the number of samples present in the data (fractional ranking; \cref{sub:ranking_of_the_metagene_scores}).
+However, the analyses so far have used the \gls{rma}-normalisation method and ranked based on the number of samples present in the data (fractional ranking; \cref{sub:ranking_of_the_metagene_scores}).
 To decide which normalisation methods or ranking approaches were suitable for the analysis, the different methods were compared in the GT data (\cref{sec:ranking_method_for_the_pathway_associated_genetic_signatures,sec:normalisation_method_for_pathway_associated_genetic_signature_transformation_matrices}).
@@
 
-Before GT pathway metagenes were compared with the obesity associated metagenes, the direction of the GT pathway metagenes had to be checked to make sure that the metagenes were in the ``correct'' direction (\cref{sub:metagene_direction}).
+Before GT pathway metagenes were compared with the obesity-associated metagenes, the direction of the GT pathway metagenes had to be checked to make sure that the metagenes were in the ``correct'' direction (\cref{sub:metagene_direction}).
 The correlation of all the pathway metagenes with one another were plotted as a heatmap in \cref{fig:gatza_meta_dir}.
-The most prominent group had five pathways (E2F1, \gls{pi3k}, Myc, \gls{bcat} and Ras) that clustered at the top right hand corner of the heatmap.
+The most prominent group had five pathways (E2F1, \gls{pi3k}, Myc, \gls{bcat} and Ras) that clustered at the top left hand corner of the heatmap.
 Other groups included \gls{ifna}/\gls{ifny}/\gls{tnfa} pathways, \gls{er}/\gls{pr}/p53 pathways and p63/\gls{her2} pathways.
@@
 
-\section{Pathway associated metagenes and obesity associated metagenes}
+\section{Pathway-associated metagenes and obesity-associated metagenes}
 \label{sec:pathway_associated_metagenes_and_obesity_associated_metagenes}
@@
 Results from the previous section confirmed that the pathway metagenes from the \citet{Gatza2010a} study were in the correct directions, and that the pathway metagenes were behaving as expected in the different data sets.
-Now that the directions of the pathway metagenes were established, these metagenes were ready for comparison with the obesity associated metagenes.
-However, all of the pathway metagenes were derived from the GT data, whereas the majority of the obesity associated metagenes were derived from the CR data, which presents a problem in terms of deciding which data set the transformation matrices should be generated in.
+Now that the directions of the pathway metagenes were established, these metagenes were ready for comparison with the obesity-associated metagenes.
+However, all of the pathway metagenes were derived from the GT data, whereas the majority of the obesity-associated metagenes were derived from the CR data, which presents a problem in terms of deciding which data set the transformation matrices should be generated in.
 
@@
 Metagenes generated from the application of \gls{svd} and metagenes generated from the transformation matrix are exactly the same if the transformation matrix is derived from the same data set.
-For example, an obesity associated metagene generated from the CR data with \gls{svd} would have the same values as the metagenes generated in the CR data with the transformation matrix (since the transformation matrix was derived from the CR data).
+For example, an obesity-associated metagene generated from the CR data with \gls{svd} would have the same values as the metagenes generated in the CR data with the transformation matrix (since the transformation matrix was derived from the CR data).
 Furthermore, if the \gls{svd}-derived metagenes and transformation matrix-derived metagenes were the same (or at least similar) in a different data set, this suggests that the metagene-related biology contained in the two data sets is highly similar.
@@
 
-To decide which data set the transformation matrices for obesity and pathway associated genetic signatures should be generated in, the \gls{svd}- and transformation matrix(TM)-generated metagene scores were compared in all of the data sets.
+To decide which data set the transformation matrices for obesity and pathway-associated genetic signatures should be generated in, the \gls{svd}- and transformation matrix(TM)-generated metagene scores were compared in all of the data sets.
 As in \cref{sec:pathway_associated_genetic_signatures_from_gatza2010a_study}, all data sets were normalised with the \gls{rma} method and metagenes were ranked with fractional ranking.
-TMs for the obesity associated genetic signatures were made in the CR data set and the TMs for pathway associated genetic signatures were made in the GT data set.
-Metagenes for all of the obesity and pathway associated genetic signatures were generated in all of the data sets with \gls{svd} and TMs.
+TMs for the obesity-associated genetic signatures were made in the CR data set and the TMs for pathway-associated genetic signatures were made in the GT data set.
+Metagenes for all of the obesity and pathway-associated genetic signatures were generated in all of the data sets with \gls{svd} and TMs.
 The Spearman correlation of the \gls{svd}-generated and TM-generated metagene scores were calculated for each genetic signatures (\cref{tab:svd_vs_tm_path,tab:svd_vs_tm_obs}).
@@
 
-In \cref{tab:svd_vs_tm_path,tab:svd_vs_tm_obs}, the obesity and pathway associated genetic signatures had a correlation of 1 in CR and GT data sets respectively, as the obesity and pathway TMs were generated in the CR and GT data sets, respectively.
+In \cref{tab:svd_vs_tm_path,tab:svd_vs_tm_obs}, the obesity and pathway-associated genetic signatures had a correlation of 1 in CR and GT data sets respectively, as the obesity and pathway TMs were generated in the CR and GT data sets, respectively.
 The correlations of the obesity and pathway metagenes were variable across different data sets, which was as expected since all the data sets were different from one another.
@@
 For example, the \gls{bcat} pathway metagene was highly correlated in all of the data sets (\textgreater{} 0.98), whereas the \gls{stat3} pathway metagene was variable across different data sets, ranging from 0.1902 to 0.9159 (\cref{tab:svd_vs_tm_path}; \cref{sec:comparison_of_the_correlation_of_svd_and_tm_generated_pathway_metagenes}).
-The fact that some pathway associated metagenes were consistent across different data sets suggested that some of these genetic signatures were reliable and did not depend on the data set the transformation matrices were derived from.
+The fact that some pathway-associated metagenes were consistent across different data sets suggested that some of these genetic signatures were reliable and did not depend on the data set the transformation matrices were derived from.
 On the other hand, the genetic signatures that were not consistent across the data sets were likely to be dependent on the data set in which they were derived from, and therefore the transformation matrices for these signatures must be derived from the GT data set.
@@
 
-In contrast to the pathway associated genetic signatures, all of the obesity associated genetic signatures showed high correlation for the \gls{svd}- and transformation matrix-generated metagenes across the different data sets.
-This suggested that the obesity associated genetic signatures were relatively consistent across all data sets, and the transformation matrices for these signatures could be made in any of the data sets.
-For simplicity, it was decided that the transformation matrices for all the genetic  signatures would be made in the GT data set, since there were some pathway associated signatures that were specific to the GT data.
+In contrast to the pathway-associated genetic signatures, all of the obesity-associated genetic signatures showed high correlation for the \gls{svd}- and transformation matrix-generated metagenes across the different data sets.
+This suggested that the obesity-associated genetic signatures were relatively consistent across all data sets, and the transformation matrices for these signatures could be made in any of the data sets.
+For simplicity, it was decided that the transformation matrices for all the genetic  signatures would be made in the GT data set, since there were some pathway-associated signatures that were specific to the GT data.
 
-\subsection{Comparison of the obesity and pathway associated signatures}
+\subsection{Comparison of the obesity and pathway-associated signatures}
 \label{sub:comparison_of_the_obesity_and_pathway_associated_signatures}
 
-To visually determine which pathway associated genetic signatures were most similar to the obesity associated genetic signatures, heatmaps were created with the metagenes for all of the genetic signatures.
-The directions of the pathway associated metagenes have already been determined in \cref{sec:pathway_associated_genetic_signatures_from_gatza2010a_study}, and the directions of the obesity associated metagenes were checked in GT data set as described in \cref{sub:metagene_direction}.
+To visually determine which pathway-associated genetic signatures were most similar to the obesity-associated genetic signatures, heatmaps were created with the metagenes for all of the genetic signatures.
+The directions of the pathway-associated metagenes have already been determined in \cref{sec:pathway_associated_genetic_signatures_from_gatza2010a_study}, and the directions of the obesity-associated metagenes were checked in GT data set as described in \cref{sub:metagene_direction}.
 All of the metagenes were created in the GT data set (normalised with \gls{rma}) using \gls{svd}, and the metagene scores were plotted in a heatmap and clustered into groups (\cref{fig:gatza_allmeta}).
@@
 
-In \cref{fig:gatza_allmeta}, all of the obesity associated metagenes from the CR data set cluster together into a group, but none of the pathway associated metagenes had  a strong positive correlation with the obesity metagenes.
+In \cref{fig:gatza_allmeta}, all of the obesity-associated metagenes from the CR data set cluster together into a group, but none of the pathway-associated metagenes had  a strong positive correlation with the obesity metagenes.
 Though the  \gls{her2} and p63 pathway metagenes were grouped next to the cluster of obesity metagenes, the similarity was at most about 0.6  (Pearson correlation; \cref{fig:gatza_allmeta}).
-With that said, these obesity associated metagenes seem to have a strong negative correlation with some of the pathway associated genetic signatures, such as the Akt, \gls{er}, \gls{egfr}, \gls{pr}, p53, Src and \gls{tgfb} pathways, and this strong negative correlation was observed in the other data sets as well (\cref{sec:correlation_of_the_pathway_and_obesity_metagenes_in_other_cancer_data_sets}).
-In fact,  high Or metagene scores related to lower levels of gene expression for the genes in the Or signature (\cref{fig:crmetaheat}; \cref{sec:creighton_obesity_metagene}), and therefore it could be possible that many of the genes in the Or signature (and the other obesity associated signatures from the CR data) may be positively correlated with the genes in these pathway signatures.
-This suggests that the obesity associated genetic signatures from the CR data may be related to these pathway signatures, rather than the obesity phenotype of the patients.
+With that said, these obesity-associated metagenes seem to have a strong negative correlation with some of the pathway-associated genetic signatures, such as the Akt, \gls{er}, \gls{egfr}, \gls{pr}, p53, Src and \gls{tgfb} pathways, and this strong negative correlation was observed in the other data sets as well (\cref{sec:correlation_of_the_pathway_and_obesity_metagenes_in_other_cancer_data_sets}).
+In fact,  high Or metagene scores related to lower levels of gene expression for the genes in the Or signature (\cref{fig:crmetaheat}; \cref{sec:creighton_obesity_metagene}), and therefore it could be possible that many of the genes in the Or signature (and the other obesity-associated signatures from the CR data) may be positively correlated with the genes in these pathway signatures.
+This suggests that the obesity-associated genetic signatures from the CR data may be related to these pathway signatures, rather than the obesity phenotype of the patients.
 
-The FM obesity associated metagene did not cluster together with the obesity metagenes from the CR data set, nor with any of the other pathway metagenes.
-The fact that the FM metagene did not cluster with any of the CR metagenes showed that the signals captures by the FM metagene was different from those metagenes found in the CR data set.
+The FM obesity-associated metagene did not cluster together with the obesity metagenes from the CR data set, nor with any of the other pathway metagenes.
+The fact that the FM metagene did not cluster with any of the CR metagenes showed that the signals captured by the FM metagene was different from those metagenes found in the CR data set.
 Furthermore, the FM metagene did not cluster with the Akt pathway in the heatmap, even though \citet{Fuentes-Mattei2014} provided clear evidence that their metagene specifically activated the Akt/\gls{mtor} signalling pathway in a mouse model.
@@
 
-The results in this section suggested that none of the obesity associated genetic signatures were positively correlated with any of the pathway associated genetic signatures from the \citet{Gatza2010a} study.
-However, some pathway signatures were negatively correlated with the obesity associated genetic signatures, which meant that the obesity signatures were likely capturing the signals from these signatures  rather than the biological signals from the obesity phenotype of the patient.
+The results in this section suggested that none of the obesity-associated genetic signatures were positively correlated with any of the pathway-associated genetic signatures from the \citet{Gatza2010a} study.
+However, some pathway signatures were negatively correlated with the obesity-associated genetic signatures, which meant that the obesity signatures were likely capturing the signals from these signatures  rather than the biological signals from the obesity phenotype of the patient.
 
@@
 
-\section{Prediction of obesity associated metagenes with pathway associated metagenes}
+\section{Prediction of obesity-associated metagenes with pathway-associated metagenes}
 \label{sec:prediction_of_obesity_associated_metagene_with_pathway_associate_metagene}
 
-To confirm the biological relationship between the obesity associated metagenes with the pathway associated metagenes from the \citet{Gatza2010a} study, linear models were created to predict the obesity metagene scores with the pathway metagene scores.
+To confirm the biological relationship between the obesity-associated metagenes with the pathway-associated metagenes from the \citet{Gatza2010a} study, linear models were created to predict the obesity metagene scores with the pathway metagene scores.
 If any of the pathway metagenes were significant in the linear model, this would provide evidence that the significant pathways in the linear model were related to the obesity metagene.
@@
 Metagene scores used for the construction of linear models were generated from the application of the transformation matrices derived from the GT data (from \cref{sec:pathway_associated_metagenes_and_obesity_associated_metagenes}) to \gls{rma} normalised \gls{nzbc} data.
-From \cref{tab:svd_vs_tm_path}, it was evident that some pathway associated genetic signatures could be variable across different data sets.
+From \cref{tab:svd_vs_tm_path}, it was evident that some pathway-associated genetic signatures could be variable across different data sets.
 Therefore, only six pathways (\gls{bcat}, \gls{er}, \gls{ifna}, \gls{ifny}, Myc and \gls{pr}) that were ``consistent'' (correlation \textgreater{} 0.9) across all of the data sets were used to construct the linear models.
@@
 \textit{MYC} is an oncogene that encodes the Myc transcription factor that, when overexpressed, disrupts various pathways that directly promote tumour growth and proliferation, and has a central role in many of the cancer hallmarks \citep{Coller2000,Hanahan2000}.
-``Sustained cell proliferation'' was one of the cancer hallmarks that \citet{Fuentes-Mattei2014} had shown to be related to their obesity associated genetic signature, which could be the reason that the Myc pathway metagene showed up as a significant variable in the linear model.
-Since the \gls{pr} pathway metagene showed significant association in many of the linear models, linear models were made with each of the combinations of \gls{bmi}, \gls{bmi} status and \gls{pr} pathway metagene for all of the obesity associated metagenes (\cref{tab:lm_pr_only}; \cref{sec:summary_of_the_linear_models_in_nzbc_data}).
+``Sustained cell proliferation'' was one of the cancer hallmarks that \citet{Fuentes-Mattei2014} had shown to be related to their obesity-associated genetic signature, which could be the reason that the Myc pathway metagene showed up as a significant variable in the linear model.
+Since the \gls{pr} pathway metagene showed significant association in many of the linear models, linear models were made with each of the combinations of \gls{bmi}, \gls{bmi} status and \gls{pr} pathway metagene for all of the obesity-associated metagenes (\cref{tab:lm_pr_only}; \cref{sec:summary_of_the_linear_models_in_nzbc_data}).
 As expected, all of the linear models showed significant contribution of the \gls{pr} metagene scores to the model for all obesity metagenes (except FM metagene; \cref{sec:summary_of_the_linear_models_in_nzbc_data}).
@@
 One possible explanation for this was that these \gls{bmi}-based models were overfitted in the \gls{nzbc} data set, and was not generalisable in the CR data set, and thus produced negative correlations.
-In fact, it was evident from the metagene analyses in \cref{cha:obesity_genetic_signatures_and_cancer} that these obesity associated genetic signatures were not generalisable with the patient \gls{bmi} in different data sets, and overfitting of the model in the \gls{nzbc} data was the likely cause of this negative correlation.
+In fact, it was evident from the metagene analyses in \cref{cha:obesity_genetic_signatures_and_cancer} that these obesity-associated genetic signatures were not generalisable with the patient \gls{bmi} in different data sets, and overfitting of the model in the \gls{nzbc} data was the likely cause of this negative correlation.
 
@@
 
-Clearly, the stepwise linear models were able to predict the obesity associated metagenes better than the models created in \cref{sub:linear_model_prediction_in_the_nzbc_and_cr_data_sets}.
+Clearly, the stepwise linear models were able to predict the obesity-associated metagenes better than the models created in \cref{sub:linear_model_prediction_in_the_nzbc_and_cr_data_sets}.
 This could be due to the fact that many of the stepwise linear models included many pathways that had a strong negative correlation with the obesity metagenes (Akt, \gls{er}, \gls{egfr}, \gls{pr}, p53, Src and \gls{tgfb} pathways; \cref{fig:gatza_allmeta}).

### 2017-08-28 marking_response/response.tex
tmp commit

@@
 %     Creation Date       :     [2017-08-15 15:29]
-%     Last Modified       :     [2017-08-22 17:44]
+%     Last Modified       :     [2017-08-28 21:55]
 %     Description         :     Response to the markers' comments for my MSc thesis.
@@
 
-Perhaps I could emphasise these points within Section 1.2.3 and guide the readers to Section 1.3.2 for specific details on the effect of obesity on cancer.
+Perhaps, on request, I could emphasise these points within Section 1.2.3 and guide the readers to Section 1.3.2 for specific details on the effect of obesity on cancer.
 \\
@@
 \noindent
+I would like to point out that Damrauer \textit{et al.} not only identified luminal and basal-like subtypes in bladder cancer, but they also showed that the molecular signatures (both gene expression pattern and gene set enrichment) between the breast and bladder cancer subtypes had striking similarity.
+This suggests that the underlying biology of the subtypes between the two cancer types are very similar.
+I am not aware of any other cancer type showing this sort of similarity.
+
+It could be possible that many other adenocarcinoma may have these characteristics due to their cell of origin context.
+However, based  on my very limited knowledge of the broad field of cancer biology, breast cancer-like subtypes  have not yet been identified in any other cancer types, let alone compared with the breast cancer subtypes like in the Damrauer \textit{et al.} paper.
+
+Based on the current knowledge of adenocarcinoma subtypes and the unexpected analysis results from this thesis, I personally think that the interpretation made on page 110  was the best explanation one could make out from these evidence.
 \\
@@
 \noindent
+%TODO: "critically appraise" the three papers.
+Creighton \textit{et al.}, Fuentes-Mattei \textit{et al.} and Gatza \textit{et al.} papers have been more thoroughly introduced in Section 1.4.
 \\
@@
 \noindent
+Adjusting the linear models with other clincial variables such as smoking would have been interesting to see.
+Extra paragraph has been added to Section 1.2.3 and a sentence has also been added towards the end of Section 5.1.1 to address this point.
 \\
@@
 \textcolor{blue}{
-	\textit{Potential mechanisms of obesity-altering hormone receptor expression could have been explored in more depth \textit{at the level of cancer cells}.
+	\textit{Potential mechanisms of obesity-altering hormone receptor expression could have been explored in more depth \textbf{at the level of cancer cells}.
 	The effects of endocrine factors like leptin and the IGFs were discussed in general but not the cancer cell type-specific gene expression changes they cause.
@@
 \textcolor{blue}{
-	\textit{It would be interesting to explore whether the differences in risk ratio between cancer types was due to technical issues, or represented real biology.
-	}
-}\\
-
-\noindent
-\\
-
-\noindent
-\textcolor{blue}{
 	\textit{Cancer patients can have vast differences in BMI at different points in their treatment.
@@
 \noindent
-This is a good point, and none of the papers define at which stage of the disease the patients' clinical data was taken (check).
+This is a good point, and only Creighton \textit{et al.} have specifically stated in their paper that the patient BMI was measured at the time of diagnosis.
+Fuentes-Mattei \textit{et al.} paper did not describe when the patient BMI was measured.
+Although I am not aware of how and when clinicians record patient data, I would suspect they would have measured BMI at the time of diagnosis to keep the data consistent with other variables (e.g.
+tumour grade).
 \\
@@
 \textcolor{blue}{
-	\textit{Discuss whether the biology underpinning differences between cancer types in terms of obesity risk ratios in Table 1 could be explored statistically using genomic data.
+	\textit{It would be interesting to explore whether the differences in risk ratio between cancer types was due to technical issues, or represented real biology.
+	Discuss whether the biology underpinning differences between cancer types in terms of obesity risk ratios in Table 1 could be explored statistically using genomic data.
 	}
@@
 \noindent
+Though there would be some technical issues in the individual studies include in the meta-analysis study by Renehan \textit{et al.}, I suspect that there is an effect of obesity on certain types of cancer.
+I am not sure whether there are any genomic data that could clarify the Examiner's question, as it is possible that obesity affects many different biological functions/pathways of a cell.
+However, I wonder if it would be possible to use SVD, or some other weighting scores, that take these risk ratios into account and essentially ``normalise'' the genetic signals (e.g.
+gene expression and/or methylation) across different cancer types, and thereby focus on the underlying biology of each cancer type.
 \\

### 2017-08-28 thesis/discussion/discussion.tex
tmp commit

@@
 As many epidemiological studies have pointed out, obesity is a strong risk factor for cancer (covered in \cref{sec:obesity_and_cancer}), but this does not necessarily mean that all cancer in obese patients develop cancer as a result of obesity.
+As briefly mentioned in \cref{ssub:environmental_factors}, there are many other environmental factors that can cause cancer and patients in these studies may have been exposed to these factors, but not recorded or measured the exposure in the clinical data for adjustment in the analysis.
+
 In other words, the cancer identified in obese patients may have been caused by another, completely different biological mechanism.
@@
 In this project, the lack of association of the obesity-associated genetic signatures with any of the cancer types could have been due to the quality of both the raw data sets and the genetic signatures used in this project.
-Furthermore, the use of \gls{bmi} may have been insufficient to describe the patients' obesity status, and would likely require more informative clinical measurements, such as waist circumference, to aid the classification.
+Furthermore, the use of \gls{bmi} may have been insufficient to describe the patients' obesity status, and would likely require more informative clinical measurements (such as waist circumference) to aid the classification.
 Therefore, careful selection of the raw data sets with sufficient, and if possible, additional information (such as waist circumference) about the patients and well-validated genetic signatures will be essential for future investigations.

### 2017-08-28 thesis/intro/intro.tex
tmp commit

@@
 Free fatty acids from excess adipose tissue and/or diet can be recognised by \gls{tlr} (specifically \gls{tlr}4), which can then activate the \gls{nfkb} pathway, leading to the activation of one of the Hallmarks of Cancer (tumour-promoting inflammation) \citep{Lumeng2011}.
-There are many other environmental factors associated with cancers, thus limiting the exposure to detrimental environmental factors is crucial for the prevention of many cancers.
+
+In addition to the mechanisms described above, there is a growing body of evidence where some environmental factors may affect the transcriptional regulation of a set of genes in certain cancer type.
+A relatively recent review by \citet{Perez-Solis2016} describes potential mechanism in which obesity, alcoholism and smoking may regulate gene expression in breast cancer cells.
+By affecting the so-called ``Master'' transcription factors, environmental factors are able to affect the expression of the genes that are downstream to the Master transcription factor, and as a result, lead to the development of cancerous cells \citep{Perez-Solis2016}.
+It is not surprising that different environmental factors affect different Master transcriptional factor, and therefore have different biological mechanism of gene regulation.
+
+It is also worth noting here that multiple  environmental factors may be exposed to an individual at any given point of time, and perhaps a combination of multiple environmental factors may have more significant effect on the individual.
+There are many other environmental factors associated with cancers that have not been mentioned here.
+Nevertheless, it is evident that limiting the exposure to detrimental environmental factors is crucial for the prevention of many cancers.
 

### 2017-09-19 marking_response/response.tex
Final edit

@@
 %     Creation Date       :     [2017-08-15 15:29]
-%     Last Modified       :     [2017-08-28 21:55]
+%     Last Modified       :     [2017-08-29 13:48]
 %     Description         :     Response to the markers' comments for my MSc thesis.
@@
 \noindent
-I understand the Examiner's concern about the relatively brief description of the known environmental factors that affect cancer incidence.
+I understand your concern about the relatively brief description of the known environmental factors that affect cancer incidence.
 As you may know, the focus of this project was the relationship between cancer and obesity, and how obesity affects cancer (which is covered more in depth in Section 1.3.2).
@@
 \noindent
+As mentioned above, Section 1.2 was intended as a general overview of cancer, and not specific to any particular environmental factor or cancer type.
+RAS-ERK pathway was chosen as an example of the target for cancer treatment, as this signalling pathway was arguably the most well-studied and understood signalling pathway in cancer.
+PI3K/Akt pathway was included very briefly as an example of how complex the biology of a tumour could be, and the associated difficulties in treating cancer.
 \\
@@
 \noindent
-I agree with the Examiner that the fonts are indeed small, but due to the nature of the plot arrangement for these figures the fonts cannot be made any larger (at least not trivially).
+I agree that the fonts are indeed small, but due to the nature of the plot arrangement for these figures the fonts cannot be made any larger (at least not trivially).
 Also, the font size was chosen so that the statistical values and letters did not obscure the points and boxes of the plots, and I would like to avoid this as much as possible, so I have left the font size unchanged.
@@
 \noindent
-%TODO: "critically appraise" the three papers.
-Creighton \textit{et al.}, Fuentes-Mattei \textit{et al.} and Gatza \textit{et al.} papers have been more thoroughly introduced in Section 1.4.
+Slightly more details on Creighton \textit{et al.} and Fuentes-Mattei \textit{et al.} papers have been added in Section 1.4.
 \\
@@
 \textcolor{blue}{
-	\textit{The interesting literature from \textit{in vitro} studies about the effect on cancer cells of factors produced by adipocytes in culture was not explored.
-	}
-}\\
-
-\noindent
-\\
-
-\noindent
-\textcolor{blue}{
-	\textit{Potential mechanisms of obesity-altering hormone receptor expression could have been explored in more depth \textbf{at the level of cancer cells}.
-	The effects of endocrine factors like leptin and the IGFs were discussed in general but not the cancer cell type-specific gene expression changes they cause.
-	}
-}\\
-
-\noindent
-\\
-
-\noindent
-\textcolor{blue}{
 	\textit{A general in-depth exploration of the `meaning' of the different data types used in the thesis and how they are affected by normalisation and statistical test choice would have been appropriate in the Introduction.
@@
 \noindent
+I have not looked into the overlap between the genes/probes between different genetic signatures in great details.
+The overlap of the genes between the different signatures will make it more/less similar to one another (depending on the overlap).
+Since there were no direct comparison between any two genetic signatures, I doubt that there was any influence on the statistical analysis carried out in this project.
+Note that multiple genetic signatures were used in the linear models as covariates for controlling their influence on the models that were constructed, and therefore it would have had no negative influence on the analyses.
 \\
@@
 \textcolor{blue}{
-	\textit{What basic quality control was possible on each data set?
-	}
-}\\
-
-\noindent
-\\
-
-\noindent
-\textcolor{blue}{
 	\textit{Why was the BRCA RNAseq data set in TCGA not used?
@@
 \noindent
-I completely agree with the Examiner and the BRCA data set would have aided greatly in the exploration of obesity-associated signatures from both Creighton \textit{et al.} and Fuentes-Mattei \textit{et al.} data sets.
+I agree with the Examiner and the BRCA data set would have aided greatly in the exploration of obesity-associated signatures from both Creighton \textit{et al.} and Fuentes-Mattei \textit{et al.} data sets.
 However, at the time of data retrieval (March 31st 2015), there were no height, weight or BMI information for the BRCA data set, and could not use the data set for the analyses.
@@
 \noindent
+This has been included in the added paragraph in Section 1.2.3.
 \\
@@
 \noindent
+Added a couple of sentences at the end of second paragraph in Section 1.3.2.
 \\
@@
 \noindent
-Though there would be some technical issues in the individual studies include in the meta-analysis study by Renehan \textit{et al.}, I suspect that there is an effect of obesity on certain types of cancer.
+Though there would have been some technical issues in the individual studies included in the meta-analysis study by Renehan \textit{et al.}, I suspect that there is an effect of obesity on certain types of cancer.
 I am not sure whether there are any genomic data that could clarify the Examiner's question, as it is possible that obesity affects many different biological functions/pathways of a cell.

### 2017-09-19 thesis/intro/intro.tex
Final edit

@@
 In addition to the mechanisms described above, there is a growing body of evidence where some environmental factors may affect the transcriptional regulation of a set of genes in certain cancer type.
-A relatively recent review by \citet{Perez-Solis2016} describes potential mechanism in which obesity, alcoholism and smoking may regulate gene expression in breast cancer cells.
+A relatively recent review by \citet{Perez-Solis2016} describes potential mechanism in which obesity, alcoholism and smoking (all a lifestyle factor related to obesity) may regulate gene expression in breast cancer cells.
 By affecting the so-called ``Master'' transcription factors, environmental factors are able to affect the expression of the genes that are downstream to the Master transcription factor, and as a result, lead to the development of cancerous cells \citep{Perez-Solis2016}.
@@
 Many studies have confirmed the mitogenic effect of sex hormones on certain cell types, mainly breast and endometrial cell types \citep{Roberts2010}.
+In addition to this, there has also been evidence of serum from obese patient influencing the gene expression of sex hormone receptor (estrogen receptor $\beta$) in a breast cancer cell line, suggesting there may be a gene regulatory effect caused by the obese phenotype \citep{Bowers2015}.
+Through the effects of dysregulated levels of hormones and hormone receptors, obesity may have significant impact on the tumour biology through changes in gene expression level.
 
-On top of the effects that \glspl{adipocyte} have on circulating concentration of sex hormones, \glspl{adipocyte} also produce adipose tissue-specific hormones called adipokines \citep{Roberts2010}.
+On top of the effects that \glspl{adipocyte} have on circulating concentration of sex hormones as well as gene expression of sex hormone receptors, \glspl{adipocyte} also produce adipose tissue-specific hormones called adipokines \citep{Roberts2010}.
 Of the many adipokines identified so far, leptin and adipo\-nectin have been studied the most in terms of cancer development \citep{Renehan2006,Roberts2010}.
@@
 \citet{Creighton2012} used 103 breast tumour samples and analysed these samples using microarrays.
+The tumour samples were trimmed upon collection, to ensure all of the samples had $\geq$ 70\% tumour cells before \acrshort{rna} was extracted from the samples.
 Of the 103 samples, 35\% were from normal weight, 28\% were from overweight and 37\% were from obese patients.
@@
 The majority of these genes (725 gene probes) were downregulated and 74 gene probes were upregulated in the breast tumour samples from the obese patients compared with the breast tumour samples from the non-obese patients.
+
+Major concern with this study was the quality of the \acrshort{rna} that was collected from the samples.
+Ensuring that all of the samples had more than 70\% of tumour cells meant that the \acrshort{rna} was likely from the tumour cells.
+However, this did not control for the quality of the \acrshort{rna} that has been extracted, as some samples may have had significantly less tumour volume compared to another, but still have $\geq$ 70\% tumour cells in both samples.
+Another thing to note from this paper was the predictive ability of their obesity signatures in other data sets.
+\citet{Creighton2012} used six independent breast cancer  data sets to test whether their signature was able to replicate their findings.
+Although they were able to replicate their results in one of the data sets, they were not able to find any significant association in the other data sets.
 
@@
 Furthermore, functional transcriptomic analysis suggested that the insulin signalling and inflammation affected human \gls{er}$^+$ breast cancer.
+One minor thing to point out in this paper was that the normalisation procedure carried out before the differential expression analysis was not clear.
 
-On top of this, \citet{Fuentes-Mattei2014} validated their results using a mouse model, where they compared the expression level in the breast tumours from the obese mice with those from the lean mice.
+On top of the bioinformatic analysis that was carried out, \citet{Fuentes-Mattei2014} validated their results using a mouse model, where they compared the expression level in the breast tumours from the obese mice with those from the lean mice.
 Results from this animal model provided strong evidence that obesity significantly accelerated tumour growth through Akt/\gls{mtor} signalling pathway.

### 2017-09-19 thesis/methods/methods.tex
Final edit

@@
 	\label{tab:num_sample_microarray}
-	\begin{tabular}{lc}
-		Data set & Number of samples\\
+	\begin{threeparttable}
+		\begin{tabular}{lcc}
+			Data set & Number of samples before exclusion\tnote{1} & Number of samples\\
 			\hline
 			\hline
-		\rule{0pt}{2.25ex}Creighton \textit{et al.} & 103\\
-		Fuentes-Mattei \textit{et al.}              & 278\\
-		\gls{nzbc}                                  & 99 \\
-		Gatza \textit{et al.}                       & 1060\\
+			\rule{0pt}{2.25ex}Creighton \textit{et al.} & 103 & 103\\
+			Fuentes-Mattei \textit{et al.}              & 278 & 278\\
+			\gls{nzbc}                                  & 107 & 99 \\
+			Gatza \textit{et al.}                       & 1965 & 1060\\
 			\hline
@@
 		\end{tabular}
+		\begin{tablenotes}
+			\begin{footnotesize}
+			\item [1] See \cref{sub:breast_cancer_data} for exclusion criteria.
+			\end{footnotesize}
+		\end{tablenotes}
+	\end{threeparttable}
 \end{table}

### 2017-09-19 thesis/misc/appendix2.tex
Final edit

@@
 
+	\section{Overlap of the genes between the pathway-associated signatures}
+	\label{sec:overlap_of_the_genes_between_the_pathway_associated_signatures}
+
+	\begin{figure}[htpb]
+		\centering
+		\includegraphics[width=0.8\linewidth,page=1]{results2/overlap}
+		\caption[Heatmap showing the overlap of genes between the different pathway-associated signatures]{Heatmap showing the overlap of genes between the different pathway-associated signatures.
+For each pairs of pathway-associated signatures, the overlap of genes were counted and divided by the total number of unique genes from the two signatures to get a ratio.
+This ratio was then scaled and plotted as shown}
+		\label{fig:overlap_pathway_genetic_signature_extra}
+	\end{figure}
+
 	\section{Correlation of the pathway and obesity metagenes in other cancer data sets}

### 2017-09-19 thesis/results/results2.tex
Final edit

@@
 The fact that some pathway-associated metagenes were consistent across different data sets suggested that some of these genetic signatures were reliable and did not depend on the data set the transformation matrices were derived from.
+Adding to this, there were very little overlap in the genes between the different pathway-associated genetic signatures, suggesting that the consistency across the data sets were not due to an artefact that was affecting the dominant probe set in the signatures (\cref{sec:overlap_of_the_genes_between_the_pathway_associated_signatures}).
 On the other hand, the genetic signatures that were not consistent across the data sets were likely to be dependent on the data set in which they were derived from, and therefore the transformation matrices for these signatures must be derived from the GT data set.

### 2017-10-12 thesis/discussion/discussion.tex
Final revision completed

@@
 
-The fact that both CR and FM obesity-associated genetic signatures showed no sign of significant association with the patient \gls{bmi} in majority of the cancer data sets raised a question of whether these signatures were actually related to obesity, or a different clinical variable within the data set.
+The fact that neither CR nor FM obesity-associated genetic signatures showed any sign of significant association with the patient \gls{bmi} in the majority of the cancer data sets raised a question of whether these signatures were actually related to obesity, or to a different clinical variable within the data set.
 This question led to the identification of various versions of obesity-associated genetic signatures in \cref{sub:identification_of_obesity_associated_genetic_signatures}.
@@
 As many epidemiological studies have pointed out, obesity is a strong risk factor for cancer (covered in \cref{sec:obesity_and_cancer}), but this does not necessarily mean that all cancer in obese patients develop cancer as a result of obesity.
-As briefly mentioned in \cref{ssub:environmental_factors}, there are many other environmental factors that can cause cancer and patients in these studies may have been exposed to these factors, but not recorded or measured the exposure in the clinical data for adjustment in the analysis.
+As briefly mentioned in \cref{ssub:environmental_factors}, there are many other environmental factors that can cause cancer, and patients in these studies may have been exposed to these factors, but this exposure information is not recorded in the clinical data, so cannot be accounted for in the analysis.
 
@@
 These different subgroups of tumours (obesity-driven or not) cannot be easily identified and isolated.
-This may affect the detection of the truly obesity-associated genes, as some of the tumour samples from the patients will have nothing to do with obesity and cancer, and hence misclassified into wrong groups.
+This may affect the detection of the truly obesity-associated genes, as some of the tumour samples from the patients will have nothing to do with obesity and cancer, and hence will be misclassified into the wrong groups.
 
@@
 In fact, \citet{Damrauer2014} have identified ``basal-like'' and ``luminal''  subtypes in bladder cancer that resembled the corresponding subtypes in breast cancer, and suggested that bladder  and breast cancers may have common physiological properties.
-Since all of the obesity-associated genetic signatures were derived from breast cancer data sets, it made sense that these signatures were significant in the \gls{blca} data set, as bladder cancer and breast cancer could have been biologically similar to one another.
+Since all of the obesity-associated genetic signatures were derived from breast cancer data sets, it makes some sense that these signatures were significant in the \gls{blca} data set, as there is evidence that bladder cancer and breast cancer are biologically similar to one another \citep{Damrauer2014}.
 However, the degree of similarity between these two cancer types are not yet known and further analyses are required for a definitive explanation of the results.
@@
 From all of these results so far, one may question how robust or appropriate the use of \gls{svd} and transformation matrices are to generate metagenes in other data sets.
-It appeared as though the CR data-derived obesity-associated genetic signatures correlated significantly in the CR data, but as soon as these metagenes were generated in the other data sets via the application of transformation matrices, the correlation became insignificant.
+The CR data-derived obesity-associated genetic signatures correlated significantly in the CR data, but as soon as these metagenes were generated in the other data sets via the application of transformation matrices, the correlation became non-significant.
 One possible reason for this was that the obesity metagenes were not actually associated to the obesity phenotype in the other data sets.
@@
 The fact that there were so many false positives in the \glspl{deg} from the \gls{icgc} data sets indicated that the use of the patient \gls{bmi} and \gls{bmi} status as the ``treatment'' conditions (in other words independent variable) for gene expression analysis was prone to \Glspl{type1}.
-This meant that there were very little evidence of true obesity-associated genetic signatures in any of the data sets that have been explored in this project, or perhaps the patient \gls{bmi} and \gls{bmi} status was not enough to pinpoint the underlying biological relationship between obesity and cancer (see \cref{sub:discussion_definition_of_obesity} as well).
+This meant that there was very little evidence of true obesity-associated genetic signatures in any of the data sets that were  explored in this project, or perhaps the patient \gls{bmi} and \gls{bmi} status was not enough to pinpoint the underlying biological relationship between obesity and cancer (also see \cref{sub:discussion_definition_of_obesity}).
 
-Evidence of large number of false positives was also seen in \cref{sub:identification_of_obesity_associated_genetic_signatures}, where there were many \glspl{deg} identified from the gene expression analyses when obesity-associated genetic signatures were sought for in the CR data set.
+Evidence of a large number of false positives was also seen in \cref{sub:identification_of_obesity_associated_genetic_signatures}, where there were many \glspl{deg} identified from the gene expression analyses when obesity-associated genetic signatures were sought for in the CR data set.
 This suggested that there was a significant number of genes that were not related to obesity (false positives) in the genetic signatures.
@@
 The pathway-associated genetic signatures from the \citet{Gatza2010a} study were used to establish the biological connection with the obesity-associated genetic signatures.
-In order to make the obesity- and pathway-associated genetic signatures comparable with one another, the orientations of the metagenes were examined so that all of the metagenes were in the ``correct'' orientation (see \cref{sec:pathway_associated_genetic_signatures_from_gatza2010a_study}).
+In order to make the obesity- and pathway-associated genetic signatures comparable with one another, the orientations of the metagenes were examined so that all of the metagenes were in a consistent orientation (see \cref{sec:pathway_associated_genetic_signatures_from_gatza2010a_study}).
 
@@
 First, linear models were created with the patient \gls{bmi}, \gls{bmi} status and a selection of the most ``consistent'' pathway metagene scores (based on the consistency across different data sets).
-From these linear models, \gls{pr} pathway metagene scores showed significant contribution to the model in predicting the obesity-associated metagene scores in all of the linear models, with the exception of the FM obesity metagene (where Myc was the significant variable; \cref{sec:summary_of_the_linear_models_in_nzbc_data}).
+From these linear models, \gls{pr} pathway metagene scores showed a significant contribution to the model in predicting the obesity-associated metagene scores in all of the linear models, with the exception of the FM obesity metagene (where Myc was the significant variable; \cref{sec:summary_of_the_linear_models_in_nzbc_data}).
 
@@
 But was this due to the true biological relationship between the \gls{pr} pathway and the obesity metagenes, or only because the models were created in the \gls{nzbc} data set?
-In fact, the proportion of the patients that were \gls{pr}$^+$ was greater in the \gls{nzbc} data set (62 \gls{pr}$^+$ and 37 \gls{pr}$^-$ patients) compared with the CR data set (48 \gls{pr}$^+$ and 50 \gls{pr}$^-$ patients), which implied that the patients in the \gls{nzbc} data set were more likely to show association with the \gls{pr} pathway metagene.
+In fact, the proportion of patients that were \gls{pr}$^+$ was greater in the \gls{nzbc} data set (62 \gls{pr}$^+$ and 37 \gls{pr}$^-$ patients) compared with the CR data set (48 \gls{pr}$^+$ and 50 \gls{pr}$^-$ patients), which implied that the patients in the \gls{nzbc} data set were more likely to show association with the \gls{pr} pathway metagene.
 
@@
 In the step wise linear models, many variables were identified to be significantly associated with the obesity metagenes.
-This suggested that the pathway-associated genetic signatures that were significant in these models may be biologically related to the obesity phenotype, or perhaps the obesity-associated genetic signatures were detecting the signals from these pathways rather than the signals from the obesity phenotype.
+This suggested that the pathway-associated genetic signatures that were significant in these models may be biologically related to the obesity phenotype, or perhaps the obesity-associated genetic signatures were detecting signals from these pathways rather than the signals from the obesity phenotype.
 In fact, many of the variables included in the linear models were pathway metagenes that negatively correlated with the obesity metagenes in \cref{fig:gatza_allmeta} (\cref{sub:comparison_of_the_obesity_and_pathway_associated_signatures}).

### 2017-10-12 thesis/intro/intro.tex
Final revision completed

@@
 \noindent
-Approximately 1 in every 14 people in the world were classed as obese in 2014; therefore, 1 in every 14 people around the world were also at severe health risks associated with obesity.
+Approximately 1 in every 14 people in the world were classed as obese in 2014; therefore, 1 in every 14 people around the world were also at risk of the severe health consequences associated with obesity.
 The highest prevalence of overweight and obesity was in the American regions, where 61\% and 27\% of the population were overweight and obese respectively \citep{WHO2014}.
@@
 In addition to the mechanisms described above, there is a growing body of evidence where some environmental factors may affect the transcriptional regulation of a set of genes in certain cancer type.
-A relatively recent review by \citet{Perez-Solis2016} describes potential mechanism in which obesity, alcoholism and smoking (all a lifestyle factor related to obesity) may regulate gene expression in breast cancer cells.
+A relatively recent review by \citet{Perez-Solis2016} describes potential mechanisms in which obesity, alcoholism and smoking (all lifestyle factors related to obesity) may regulate gene expression in breast cancer cells.
 By affecting the so-called ``Master'' transcription factors, environmental factors are able to affect the expression of the genes that are downstream to the Master transcription factor, and as a result, lead to the development of cancerous cells \citep{Perez-Solis2016}.
-It is not surprising that different environmental factors affect different Master transcriptional factor, and therefore have different biological mechanism of gene regulation.
+It is not surprising that different environmental factors affect different Master transcriptional factors, and therefore have different biological mechanisms of gene regulation.
 
-It is also worth noting here that multiple  environmental factors may be exposed to an individual at any given point of time, and perhaps a combination of multiple environmental factors may have more significant effect on the individual.
+It is also worth noting here that an individual may be exposed to  multiple  environmental factors at any given point of time, and perhaps a combination of multiple environmental factors may have more significant effect on the individual.
 There are many other environmental factors associated with cancers that have not been mentioned here.
@@
 Many studies have confirmed the mitogenic effect of sex hormones on certain cell types, mainly breast and endometrial cell types \citep{Roberts2010}.
-In addition to this, there has also been evidence of serum from obese patient influencing the gene expression of sex hormone receptor (estrogen receptor $\beta$) in a breast cancer cell line, suggesting there may be a gene regulatory effect caused by the obese phenotype \citep{Bowers2015}.
+In addition to this, there has also been evidence of serum from obese patients influencing the gene expression of sex hormone receptor (estrogen receptor $\beta$) in a breast cancer cell line, suggesting there may be a gene regulatory effect caused by the obese phenotype \citep{Bowers2015}.
 Through the effects of dysregulated levels of hormones and hormone receptors, obesity may have significant impact on the tumour biology through changes in gene expression level.
@@
 
-Major concern with this study was the quality of the \acrshort{rna} that was collected from the samples.
+One major concern with this study was the quality of the \acrshort{rna} that was collected from the samples.
 Ensuring that all of the samples had more than 70\% of tumour cells meant that the \acrshort{rna} was likely from the tumour cells.
-However, this did not control for the quality of the \acrshort{rna} that has been extracted, as some samples may have had significantly less tumour volume compared to another, but still have $\geq$ 70\% tumour cells in both samples.
+However, this did not control for the quality of the \acrshort{rna} that was extracted, as some samples may have had significantly less tumour volume compared to another, but still have $\geq$ 70\% tumour cells in both samples.
 Another thing to note from this paper was the predictive ability of their obesity signatures in other data sets.

### 2017-10-12 thesis/methods/methods.tex
Final revision completed

@@
 The raw Affymetrix HGU\-133A microarray gene expression data files from the \citet{Creighton2012} study were downloaded from the \gls{geo} database (\gls{geo} accession ID: GSE24185).
-Clinical data of the samples (age, ethnicity, tumour grade, menopause status, \gls{bmi}, \gls{er} status, \gls{pr} status, \gls{her2} status, and \gls{ln} status) were obtained from the Supplementary Table 1 from the \citet{Creighton2012} study.
-799 obesity-associated gene probes identified in the \citet{Creighton2012} study were obtained from the Supplementary Data File 1 from the \citet{Creighton2012} study.
+Clinical data of the samples (age, ethnicity, tumour grade, menopause status, \gls{bmi}, \gls{er} status, \gls{pr} status, \gls{her2} status, and \gls{ln} status) were obtained from Supplementary Table 1 of the \citet{Creighton2012} study.
+799 obesity-associated gene probes identified in the \citet{Creighton2012} study were obtained from Supplementary Data File 1 from the \citet{Creighton2012} study.
 
@@
 Patient height, weight and \gls{bmi} information were not available in this data set.
-130 obesity-associated gene probes identified by \citet{Fuentes-Mattei2014} were taken from the Supplementary Table 3 from their paper.
+130 obesity-associated gene probes identified by \citet{Fuentes-Mattei2014} were taken from Supplementary Table 3 of their paper.
 
@@
 Clinical data for the samples in \citet{Gatza2010a} study was not available, as these samples were a combination of many different datasets.
-The Akt, \gls{bcat}, E2F1, \gls{egfr}, \gls{er}, \gls{her2}, \gls{ifna}, \gls{ifny}, Myc, p53, p63, \gls{pi3k}, \gls{pr}, Ras, \gls{stat3}, Src, \gls{tgfb}, and \gls{tnfa} pathway-associated genetic signatures were used by Gatza \textit{et al.}, and the gene probes of all the pathway-associated genetic signatures were obtained from the Supplemental Table 10 from their paper.
+The Akt, \gls{bcat}, E2F1, \gls{egfr}, \gls{er}, \gls{her2}, \gls{ifna}, \gls{ifny}, Myc, p53, p63, \gls{pi3k}, \gls{pr}, Ras, \gls{stat3}, Src, \gls{tgfb}, and \gls{tnfa} pathway-associated genetic signatures were used by Gatza \textit{et al.}, and the gene probes of all the pathway-associated genetic signatures were obtained from Supplemental Table 10 of their paper.
 
@@
 The raw microarray data were normalised using both \gls{rma} and \gls{mas} methods, each done separately on a copy of the data.
-The reason why the \gls{mas} method was used as well as the \gls{rma} method was because the \gls{mas}-normalisation method was used in the \citet{Gatza2010a} study.
+The reason that the \gls{mas} method was used as well as the \gls{rma} method was because the \gls{mas}-normalisation method was used in the \citet{Gatza2010a} study.
 Though the normalisation methods were not specified in the \citet{Creighton2012} and \citet{Fuentes-Mattei2014} studies, to allow for data comparison between different datasets and accurate result validation of some of the studies, both \gls{rma} and \gls{mas} methods were considered for normalising the microarray data.
@@
 In this project, two types of metagenes were created: obesity-associated and pathway-associated metagenes.
-For obesity-associated metagenes, the direction of the metagenes were checked with the patient obesity status and \gls{bmi} value from the data in which the metagene was derived from.
+For obesity-associated metagenes, the direction of each metagene was checked with the patient obesity status and \gls{bmi} value from the data in which the metagene was derived from.
 When the data set did not have \gls{bmi} or \gls{bmi} status information for the patients (for example the Gatza \textit{et al.} data set), a set of genes that were common in all obesity-associated genetic signatures were taken and the expression of these common genes were compared with the metagene scores.
-The genes that are common to all of the obesity-associated genetic signatures will be related to the metagenes produced by these signatures, and therefore the expressions of these genes should be in line with the metagene scores.
+The genes that are common to all of the obesity-associated genetic signatures will be related to the metagenes produced by these signatures, and therefore the expression patterns of these genes should be in line with the metagene scores.
 
-Checking the direction of the pathway metagenes from \citet{Gatza2010a} study is not as simple as the obesity-associated metagenes, as the experimental design that identify which samples were treated to generate the pathway signature was not available.
+Checking the direction of the pathway metagenes from \citet{Gatza2010a} study is not as simple as the obesity-associated metagenes, as the gene lists for each pathway were derived from a separate study and per-gene directionality information was not provided.
 Due to this, it was not possible to validate whether the generated pathway metagene was in the ``correct'' direction.
@@
 Eleven linear models were created in the \gls{nzbc} to predict the obesity-associated metagene from the \citet{Creighton2012} study.
-First seven models created were combinations of the sample \gls{bmi}, \gls{bmi} status and a selection of pathway-associated metagene scores.
+The first seven models created were combinations of the sample \gls{bmi}, \gls{bmi} status and a selection of pathway-associated metagene scores.
 The selected pathways used in the linear model construction were \gls{bcat}, \gls{er}, \gls{ifna}, \gls{ifny}, Myc and \gls{pr}.

### 2017-10-12 thesis/misc/appendix.tex
Final revision completed

@@
 	Since all of these obesity-associated genetic signatures were derived based on the discrete values of sample \gls{bmi} staus, a genetic signature was identified using the sample \gls{bmi} value (continuous variable).
-	Gene probe expressions were correlated with the sample \gls{bmi} values and was corrected for multiple hypothesis testing with \gls{fdr}.
+	Gene probe expressions were correlated with the sample \gls{bmi} values and were corrected for multiple hypothesis testing with \gls{fdr}.
 	There were no significant gene that correlated with the sample \gls{bmi} (p \textless{} 0.05).
@@
 
-To determine whether the results from all of the obesity metagene analyses were not showing any association with patient \gls{bmi} because of the \gls{svd}/transformation matrix approach used, the same approach was used with \gls{er} and \gls{pr} pathway-associated genetic signatures from the \citet{Gatza2010a} study.
+To determine whether the results from all of the obesity metagene analyses were not showing association with patient \gls{bmi} because of the \gls{svd}/transformation matrix approach used, the same approach was used with \gls{er} and \gls{pr} pathway-associated genetic signatures from the \citet{Gatza2010a} study.
 Firstly, both the \gls{er} and \gls{pr} metagenes were generated in the \gls{rma} normalised CR data set.

### 2017-10-12 thesis/misc/appendix2.tex
Final revision completed

@@
 
-	The original result from the \citet{Gatza2010a} study were used to determine the directions of the GT pathway-associated genetic signatures in \cref{sec:pathway_associated_genetic_signatures_from_gatza2010a_study} (\cref{fig:gatza_paper_res}).
+	The original result from the \citet{Gatza2010a} study was used to determine the directions of the GT pathway-associated genetic signatures in \cref{sec:pathway_associated_genetic_signatures_from_gatza2010a_study} (\cref{fig:gatza_paper_res}).
 
@@
 		\includegraphics[width=0.32\linewidth,page=9]{appendix/gtobrma}\\
-		\caption[Heatmaps of all the obesity and the pathway metagenes in the CR, \gls{nzbc} and FM data]{Heatmaps showing the obesity metagenes from the CR data set with the gene expressions of the obesity-associated genetic signatures in the \gls{rma} normalised GT data set.
+		\caption[Heatmaps of all the obesity and pathway metagenes in the CR, \gls{nzbc} and FM data]{Heatmaps showing the obesity metagenes from the CR data set with the gene expression of the obesity-associated genetic signatures in the \gls{rma} normalised GT data set.
 		High and low correlation were represented as red and blue, respectively, where the colours were matched with the values on the scale shown in the top left histogram.}
@@
 		\caption[Heatmap showing the overlap of genes between the different pathway-associated signatures]{Heatmap showing the overlap of genes between the different pathway-associated signatures.
-For each pairs of pathway-associated signatures, the overlap of genes were counted and divided by the total number of unique genes from the two signatures to get a ratio.
+For each pair of pathway-associated signatures, the overlap of genes was counted and divided by the total number of unique genes from the two signatures to get a ratio.
 This ratio was then scaled and plotted as shown}
